### 2017-12-06 thesis_main.tex
Only scheleton

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% yale_thesis.tex
% Elena Gramellini
% Started: 2017/12/05
%
% A bare, sample template for a Yale PhD thesis using yalephd.cls
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[letterpaper,12pt,draft]{yalephd}
% remove draft option for final printing.
% font size must be between 10pt-12pt.

\usepackage{geometry} % you need this for yalephd.cls to work.
\usepackage{graphicx} % you probably want the rest of these.
\usepackage{dcolumn}
\usepackage{bm}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{appendix}
\usepackage{comment}
\usepackage{cite}
\bibliographystyle{plain}

\newenvironment{dedication}
  {%\clearpage           % we want a new page          %% I commented this
   \thispagestyle{empty}% no header and footer
   \vspace*{\stretch{1}}% some space at the top
   \itshape             % the text is in italics
   \raggedleft          % flush to the right margin
  }
  {\par % end the paragraph
   \vspace{\stretch{3}} % space at bottom is three times that at the top
   \clearpage           % finish off the page
  }

\begin{document}

% Need to define title before the abstract.
\title{Measurement of the positive kaon-argon total hadronic differential cross section in the LArIAT experiment}
\author{Elena Gramellini}
\advisor{Bonnie T. Fleming}
\date{Date you'll receive your degree} % usually not \today.

% All the stuff at the front of your thesis.
\frontmatter

\begin{abstract}
Abstract goes here. Limit 750 words.
\end{abstract}

\maketitle
\makecopyright{2017}

%  \begin{dedication}
%A mia mamma e mio babbo,\\
%grazie per le radici e grazie per le ali.
%    \par   %% or a blank line
%    \vspace{2\baselineskip}
%To my mom and dad,\\
%thank you for the roots and thank you for the wings.
%    \vspace{\baselineskip}
%  \end{dedication}
\tableofcontents
\listoffigures % remove this if you have no figures.
\listoftables % remove this if you have no tables.

\chapter{Acknowledgements} % this needs to be before \mainmatter.
A lot of people are awesome. But not you, who are reading my thesis before it's done.

% Starts proper arabic numbering of pages and chapters.
\mainmatter

\input{0-Introduction}
\input{1-LArTPCs}
\input{2-TheLArIATExperiment}
\input{3-KaonInteractionsInLAr}
\input{4-DataCollection}
\input{5-EnergyCalibration}
\input{6-TrackingOptimization}
\input{7-KaonCrossSectionAnalysis}

% Add additional \chapter{}s as necessary.

% use \cite{} to cite a reference in your bibliography file.
% use \ref{} to reference a \label{} from an equation, figure, or table.

% for sets of equations use align:
%\begin{align}
%\end{align}

% for figures:
%\begin{figure}[ht]
%\centering
%\includegraphics[width=.45\textwidth]{name_of_figure.eps}
%\caption{A caption! \label{a_figure}}
%\end{figure}

% for tables:
%\begin{table}
%\begin{tabular}{c|c|c}
% 1 & 2 & 3 \\
%\hline
%\end{tabular}
%\caption{Another caption! \label{a_table}}
%\end{table}

% Only call appendix once, here.
\appendix
\input{A-MicroBooNECRT}

% Any chapters such as End Notes go after this.
\backmatter

\bibliography{bib.bib}
\bibliographystyle{plain}
% for your own sake, use a bibtex file, so all of the numbering of references will be done
% automatically.

\end{document}

### 2018-01-30 Chapter-0/0-Introduction.tex
it is been a while

\chapter{Introduction}

\section{The Standard Model}
The Standard Model (SM) of particle physics the most accurate theoretical description of the subatomic world and, more generically, one of the most precisely tested theories in the history of physics.  The SM describes the strong, electromagnetic and weak interactions among  elementary particles in the framework of quantum field theory, accounting for the unification of electromagnetic and weak interactions for energies above the  vacuum expectation value of the Higgs field. The SM does not describe gravity or general relativity.

The Standard Model is a gauge theory based on the local group of symmetry
\begin{equation}
G_{SM} = SU(3)_C  \otimes SU(2)_T \otimes U(1)_Y
\end{equation}

where the subscripts indicate the conserved charges: the strong charge, or color C, the weak isospin T (or rather its third component T3) and the hypercharge Y. These quantities can be related to the electric charge Q through the Gell-Mann-Nishijima relation:
\begin{equation}
Q = \frac{Y}{2} + T_3.
\end{equation}

In the quantum field framework, the elementary particles correspond to the irreducible representations of the G$_{SM}$ symmetry group. In particular, the particles are divided in two categories, fermions and bosons, according to their spin-statistics. Described by the Fermi-Dirac statistics, Fermions have half-integer spin and are sometimes called ``matter-particles". Bosons or ``force carriers" have integer spin, follow the Bose-Einstein statistics and mediate the interaction between fermions. The fundamental fermions and their quantum numbers are listed in Tab \ref{tab:SMParticles}.

\begin{table}[]
\centering
\begin{tabular}{|cccc|c|c|c|}\hline
Generation               & I                   & II                  & III                 & T                       & Y                   & Q                   \\\hline

\multirow{6}{*}{Leptons}                          &                     &                     &                     &                         &                     &                     \\
& $\begin{pmatrix}\ \nu_e\\ e \end{pmatrix}_L$ & $\begin{pmatrix}\ \nu_\mu\\ \mu \end{pmatrix}_L$ & $\begin{pmatrix}\ \nu_\tau\\ \tau \end{pmatrix}_L$ & $\begin{matrix}\ 1/2\\ -1/2 \end{matrix}$ & $\begin{matrix}\ -1\\ -1 \end{matrix}$ & $\begin{matrix}\ 0\\ -1 \end{matrix}$ \\
                         &                     &                     &                     &                         &                     &                     \\
                         & $e_R$         & $\mu_R$      & $\tau_R$                  & 0                      & -2                  & 1    \\
                         &                     &                     &                     &                         &                     &                     \\\hline
\multirow{7}{*}{Quarks} 
                         &                     &                     &                     &                         &                     &                     \\
& $\begin{pmatrix}\ u\\ d' \end{pmatrix}_L$ & $\begin{pmatrix}\ c\\ s' \end{pmatrix}_L$ & $\begin{pmatrix}\ t\\ b' \end{pmatrix}_L$ & $\begin{matrix}\ 1/2\\ -1/2 \end{matrix}$ & $\begin{matrix}\ 1/3\\ 1/3 \end{matrix}$ & $\begin{matrix}\ 2/3\\ -1/3 \end{matrix}$ \\
                         &                     &                     &                     &                         &                     &                     \\
& $\begin{matrix}\ u_R\\ d'_R \end{matrix}$ & $\begin{matrix}\ c_R\\ s'_R \end{matrix}$ & $\begin{matrix}\ t_R\\ b'_R \end{matrix}$ & $\begin{matrix}\ 0\\ 0 \end{matrix}$ & $\begin{matrix}\ 4/3\\ -2/3 \end{matrix}$ & $\begin{matrix}\ 2/3\\ -1/3 \end{matrix}$ \\
                         &                     &                     &                     &                         &                     &                    \\\hline
\end{tabular}
\caption{SM elementary fermions. The subscripts L and R indicate respectively the negative helicity (left-handed) and the positive helicity (right-handed).}
\label{tab:SMParticles}
\end{table}

Quarks can interact via all three the fundamental forces; they are triplets of SU(3)$_C$, that is they can exist in three different colors: C = R, G, B. If one chooses a base where $u$, $c$ and $t$ quarks are simultaneously eigenstates of both the strong and the weak interactions, the remaining eigenstates are usually written as $d$, $s$ and $b$ for the strong interaction and $d'$, $s'$ and $b'$ for the weak interaction, because the latter ones are the result of a Cabibbo rotation on the first ones.
Charged leptons interact via the weak and the electromagnetic forces, while neutrinos only interact via the weak force. 
The gauge group univocally determines the number of gauge bosons that carry the interaction; the gauge bosons correspond to the generators of the group: eight gluons (g) for the strong interaction, one photon ($\gamma$) and three bosons (W$^\pm$, Z$^0$) for the electroweak interaction.
A gauge theory by itself can not provide a description of massive particles, but it is experimentally well know that most of the elementary particles have non-zero masses. The introduction of massive fields in the Standard Model lagrangian would make the theory non-renormalizable, and - so far - mathematically impossible to handle. This problem is solved in the Standard Model by the introduction of a scalar iso-doublet $\Phi(x)$, the Higgs field, which gives mass to W$^\pm$ and Z$^0$ gauge bosons through the electroweak symmetry breaking and to the fermions through Yukawa coupling \cite{Higgs1964,Higgs19642}.

\section{Neutrinos in the Standard Model}
The neutrino was introduced in the SM as a left-handed massless Weyl spinor.
The Dirac equation of motion
\begin{equation}
(i\gamma^ \mu \partial_\mu - m) \psi = 0
\end{equation}
for a fermionic field 
\begin{equation}
 \psi =  \psi_L +  \psi_R
\end{equation}
is equivalent to the equaitons
\begin{equation}
i\gamma^ \mu \partial_\mu  \psi_L = m \psi_R
\end{equation}
\begin{equation}
i\gamma^ \mu \partial_\mu  \psi_R = m \psi_L
\end{equation}

for the chiral fields $\psi_R$ and $\psi_L$, whose evolution in space and time is coupled through the mass $m$.
If the fermion is massless, the chiral fields decouple and the fermion can be described by a single Weyl spinor with two independent components. Pauli initially rejected the description of a physical particle through a single Wyle spinor because of its implication of parity violation. In fact, since the spatial inversion operator throws $\psi_R \leftrightarrow \psi_L$, parity is conserved only if the both the chiral components exist at the same time. 
\textcolor{red}{ADD CITATIONS}.  For the neutrino introduction in the SM, experiments came in help of the theoretical description:  the constraint of parity conservation weakened after Wu's experiment \textcolor{red}{ADD CITATIONS AND DATES}, there was no experimental indication for massive neutrinos and neutrinos likely interacted only via the left-handed component.

The symmetry group $SU(2)_T \otimes U(1)_Y$ is the only group relevant for neutrino interactions. The SM electroweak lagrangian is the most general renormalizable lagrangian invariant under the local symmetry group $SU(2)_T \otimes U(1)_Y$. The lagrangian couples the weak isotopic spin doublets and singlets described in \ref{tab:SMParticles} with the gauge bosons  $A^{\mu}_{a}$ ($a$ $=$ 1,2,3) and $B^{\mu}$, and Higgs doublet $\Phi(x)$:

\begin{eqnarray}
\lefteqn{\mathcal{L} = i\sum_{\alpha=e,\mu,\tau} \bar{L}'_{\alpha L}  \slashed D L'_{\alpha L} + 
 i\sum_{\alpha=1,2,3} \bar{Q}'_{\alpha L}  \slashed D Q'_{\alpha L} {}}
 \nonumber\\
 & & {} + i\sum_{\alpha=e,\mu,\tau} \bar{l}'_{\alpha R}  \slashed D l'_{\alpha R} + i\sum_{\alpha=d,s,b} \bar{q}'^D_{\alpha R}  \slashed D q'^D_{\alpha R} + i\sum_{\alpha=u,c,t} \bar{q}'^U_{\alpha R}  \slashed D q'^U_{\alpha R}
 \nonumber\\
 & & {} -\frac{1}{4}A_{\mu \nu}A^{\mu \nu} - \frac{1}{4}B_{\mu \nu}B^{\mu \nu}
 \nonumber\\
 & & {} +(D_{\rho}\Phi)^\dagger(D^{\rho}\Phi) - \mu^2\Phi^\dagger\Phi - \lambda(\Phi^\dagger\Phi)^2 
 \nonumber\\
 & & {} -\sum_{\alpha,\beta=e,\mu,\tau} \Big(Y'^l_{\alpha\beta}\bar{L}'_{\alpha L}  \Phi l'_{\beta R} + Y'^{l*}_{\alpha\beta}\bar{l}'_{\beta R}  \Phi^\dagger L'_{\alpha L}\Big)
  \nonumber\\
 & & {} -\sum_{\alpha=1,2,3} \sum_{\beta=d,s,b} \Big(Y'^D_{\alpha\beta}\bar{Q}'_{\alpha L}  \Phi q'^D_{\beta R} + Y'^{D*}_{\alpha\beta}\bar{q}'^D_{\beta R}  \Phi^\dagger Q'_{\alpha L}\Big)
  \nonumber\\
 & & {} -\sum_{\alpha=1,2,3} \sum_{\beta=u,c,t} \Big(Y'^U_{\alpha\beta}\bar{Q}'_{\alpha L}   \widetilde{\Phi} q'^U_{\beta R} + Y'^{U*}_{\alpha\beta}\bar{q}'^U_{\beta R} \widetilde{\Phi}^\dagger Q'_{\alpha L}\Big).
\end{eqnarray}

The first two lines of the lagrangian summarize the kinetic terms for the fermionic fields and their coupling to the gauge bosons $A^{\mu\nu}_a$, $B^{\mu\nu}$ \footnote{In gauge theories the ordinary derivative $\partial_\mu$  is substitued with èthe covariant derivative $D_\mu$. Here $D_\mu = \partial_\mu + igA_\mu \cdot I + ig'B_\mu\frac{Y}{2}$, where I and Y are the SU(2)$_L$ and U(1)$_Y$ generators, respectively.}.
The third line describes the kinetic terms and the self-coupling terms of the gauge bosons. The forth line is the Higgs lagrangian, which results in the spontaneous symmetry breaking. The last three lines describe the Yukawa coupling between fermions and the Higgs field, origin of the fermion's mass.

The coupling between left-handed and right-handed field generates the mass term for fermions. The SM assumes only left-handed components for neutrinos, thus implying zero neutrino mass. Since any linear combination of massless fields results in a massless field, the flavor eigenstates are identical to the mass eigenstates in the SM.

\section{Beyond the Standard Model}
The discovery of neutrino oscillation and its implication of non-zero neutrino mass marks  the beginning of a new, exciting era in neutrino physics: the era of physics Beyond the Standard Model (BSM) in the neutrino sector.
We are currently searching for new, deeper theories that can accommodate neutrinos with non-zero mass, while remaining consistent with the rest of the Standard Model. %More excitingly, we need to probe these theories experimentally. 
\subsection{Neutrino Oscillations}
\subsection{Proton Decay}

### 2018-01-30 Chapters/0-Introduction.tex
it is been a while

--- 2018-01-30 Chapter-0/0-Introduction.tex
@@
 \section{The Standard Model}
-The Standard Model (SM) of particle physics the most accurate theoretical description of the subatomic world and, more generically, one of the most precisely tested theories in the history of physics.
-The SM describes the strong, electromagnetic and weak interactions among  elementary particles in the framework of quantum field theory, accounting for the unification of electromagnetic and weak interactions for energies above the  vacuum expectation value of the Higgs field.
-The SM does not describe gravity or general relativity.
-
+The Standard Model of particle physics the most accurate theoretical description of the subatomic world.
+The Standard Model describes the strong, electromagnetic and weak interactions among  elementary particles in the framework of quantum field theory.
+The weak and the electromagnetic interactions are unified into the electroweak interaction.
+The Standard Model does not describe gravity or general relativity.
 The Standard Model is a gauge theory based on the local group of symmetry
@@
 \begin{equation}
-Q = \frac{Y}{2} + T_3.
+Q = \frac{Y}{2} + T_3
 \end{equation}
@@
 Bosons or ``force carriers" have integer spin, follow the Bose-Einstein statistics and mediate the interaction between fermions.
-The fundamental fermions and their quantum numbers are listed in Tab \ref{tab:SMParticles}.
+The fundamental fermions and their quantum numbers are listed in Tab \ref{tab:SMParticles}\textcolor{red}{INSERT TABLE}
+\begin{table}[!htbp]
 
-\begin{table}[]
-\centering
-\begin{tabular}{|cccc|c|c|c|}\hline
-Generation               & I                   & II                  & III                 & T                       & Y                   & Q                   \\\hline
-
-\multirow{6}{*}{Leptons}                          &                     &                     &                     &                         &                     &                     \\
-& $\begin{pmatrix}\ \nu_e\\ e \end{pmatrix}_L$ & $\begin{pmatrix}\ \nu_\mu\\ \mu \end{pmatrix}_L$ & $\begin{pmatrix}\ \nu_\tau\\ \tau \end{pmatrix}_L$ & $\begin{matrix}\ 1/2\\ -1/2 \end{matrix}$ & $\begin{matrix}\ -1\\ -1 \end{matrix}$ & $\begin{matrix}\ 0\\ -1 \end{matrix}$ \\
-                         &                     &                     &                     &                         &                     &                     \\
-                         & $e_R$         & $\mu_R$      & $\tau_R$                  & 0                      & -2                  & 1    \\
-                         &                     &                     &                     &                         &                     &                     \\\hline
-\multirow{7}{*}{Quarks} 
-                         &                     &                     &                     &                         &                     &                     \\
-& $\begin{pmatrix}\ u\\ d' \end{pmatrix}_L$ & $\begin{pmatrix}\ c\\ s' \end{pmatrix}_L$ & $\begin{pmatrix}\ t\\ b' \end{pmatrix}_L$ & $\begin{matrix}\ 1/2\\ -1/2 \end{matrix}$ & $\begin{matrix}\ 1/3\\ 1/3 \end{matrix}$ & $\begin{matrix}\ 2/3\\ -1/3 \end{matrix}$ \\
-                         &                     &                     &                     &                         &                     &                     \\
-& $\begin{matrix}\ u_R\\ d'_R \end{matrix}$ & $\begin{matrix}\ c_R\\ s'_R \end{matrix}$ & $\begin{matrix}\ t_R\\ b'_R \end{matrix}$ & $\begin{matrix}\ 0\\ 0 \end{matrix}$ & $\begin{matrix}\ 4/3\\ -2/3 \end{matrix}$ & $\begin{matrix}\ 2/3\\ -1/3 \end{matrix}$ \\
-                         &                     &                     &                     &                         &                     &                    \\\hline
+\begin{tabular}{|l ccc|c|c|c|}
+\hline
+Generation: & I & II & III & T$_3$ & Y & Q \\
+\hline
+Leptonic doubles: & e$_L$ & e$_L$ & e$_L$ & 0 & -2 & -2 \\
+Leptonic singlets: & e$_R$ & e$_R$ & e$_R$ & 0 & -2 & -2 \\
+\hline
+Quark doubles: & I & II & III & T$_3$ & Y & Q \\
+Quark singlets: & I & II & III & T$_3$ & Y & Q \\
+\hline
 \end{tabular}
-\caption{SM elementary fermions. The subscripts L and R indicate respectively the negative helicity (left-handed) and the positive helicity (right-handed).}
+\caption{SM elementary fermions. The subscripts L and R indicate respectively the negative helicity (left-handed) and the positive helicity (right- handed).}
 \label{tab:SMParticles}
@@
 Quarks can interact via all three the fundamental forces; they are triplets of SU(3)$_C$, that is they can exist in three different colors: C = R, G, B.
-If one chooses a base where $u$, $c$ and $t$ quarks are simultaneously eigenstates of both the strong and the weak interactions, the remaining eigenstates are usually written as $d$, $s$ and $b$ for the strong interaction and $d'$, $s'$ and $b'$ for the weak interaction, because the latter ones are the result of a Cabibbo rotation on the first ones.
+If one chooses a base where u, c and t quarks are simultaneously eigenstates of both the strong and the weak interactions, the remaining eigenstates are usually written as d, s and b for the strong interaction and d', s' and b' for the weak interaction, because the latter ones are the result of a Cabibbo rotation on the first ones.
 Charged leptons interact via the weak and the electromagnetic forces, while neutrinos only interact via the weak force. 
@@
 The introduction of massive fields in the Standard Model lagrangian would make the theory non-renormalizable, and - so far - mathematically impossible to handle.
-This problem is solved in the Standard Model by the introduction of a scalar iso-doublet $\Phi(x)$, the Higgs field, which gives mass to W$^\pm$ and Z$^0$ gauge bosons through the electroweak symmetry breaking and to the fermions through Yukawa coupling \cite{Higgs1964,Higgs19642}.
+This problem is solved in the Standard Model by the introduction of a scalar iso-doublet $\phi$(x), the Higgs field, see Tab \textcolor{red}{INSERT TABLE}, which gives masse to W$^\pm$ and Z$^0$ gauge bosons and to the fermions [1, 2].
 
 \section{Neutrinos in the Standard Model}
-The neutrino was introduced in the SM as a left-handed massless Weyl spinor.
-The Dirac equation of motion
-\begin{equation}
-(i\gamma^ \mu \partial_\mu - m) \psi = 0
-\end{equation}
-for a fermionic field 
-\begin{equation}
- \psi =  \psi_L +  \psi_R
-\end{equation}
-is equivalent to the equaitons
-\begin{equation}
-i\gamma^ \mu \partial_\mu  \psi_L = m \psi_R
-\end{equation}
-\begin{equation}
-i\gamma^ \mu \partial_\mu  \psi_R = m \psi_L
-\end{equation}
-
-for the chiral fields $\psi_R$ and $\psi_L$, whose evolution in space and time is coupled through the mass $m$.
-If the fermion is massless, the chiral fields decouple and the fermion can be described by a single Weyl spinor with two independent components.
-Pauli initially rejected the description of a physical particle through a single Wyle spinor because of its implication of parity violation.
-In fact, since the spatial inversion operator throws $\psi_R \leftrightarrow \psi_L$, parity is conserved only if the both the chiral components exist at the same time.
-\textcolor{red}{ADD CITATIONS}.
-For the neutrino introduction in the SM, experiments came in help of the theoretical description:  the constraint of parity conservation weakened after Wu's experiment \textcolor{red}{ADD CITATIONS AND DATES}, there was no experimental indication for massive neutrinos and neutrinos likely interacted only via the left-handed component.
-
-The symmetry group $SU(2)_T \otimes U(1)_Y$ is the only group relevant for neutrino interactions.
-The SM electroweak lagrangian is the most general renormalizable lagrangian invariant under the local symmetry group $SU(2)_T \otimes U(1)_Y$.
-The lagrangian couples the weak isotopic spin doublets and singlets described in \ref{tab:SMParticles} with the gauge bosons  $A^{\mu}_{a}$ ($a$ $=$ 1,2,3) and $B^{\mu}$, and Higgs doublet $\Phi(x)$:
-
-\begin{eqnarray}
-\lefteqn{\mathcal{L} = i\sum_{\alpha=e,\mu,\tau} \bar{L}'_{\alpha L}  \slashed D L'_{\alpha L} + 
- i\sum_{\alpha=1,2,3} \bar{Q}'_{\alpha L}  \slashed D Q'_{\alpha L} {}}
- \nonumber\\
- & & {} + i\sum_{\alpha=e,\mu,\tau} \bar{l}'_{\alpha R}  \slashed D l'_{\alpha R} + i\sum_{\alpha=d,s,b} \bar{q}'^D_{\alpha R}  \slashed D q'^D_{\alpha R} + i\sum_{\alpha=u,c,t} \bar{q}'^U_{\alpha R}  \slashed D q'^U_{\alpha R}
- \nonumber\\
- & & {} -\frac{1}{4}A_{\mu \nu}A^{\mu \nu} - \frac{1}{4}B_{\mu \nu}B^{\mu \nu}
- \nonumber\\
- & & {} +(D_{\rho}\Phi)^\dagger(D^{\rho}\Phi) - \mu^2\Phi^\dagger\Phi - \lambda(\Phi^\dagger\Phi)^2 
- \nonumber\\
- & & {} -\sum_{\alpha,\beta=e,\mu,\tau} \Big(Y'^l_{\alpha\beta}\bar{L}'_{\alpha L}  \Phi l'_{\beta R} + Y'^{l*}_{\alpha\beta}\bar{l}'_{\beta R}  \Phi^\dagger L'_{\alpha L}\Big)
-  \nonumber\\
- & & {} -\sum_{\alpha=1,2,3} \sum_{\beta=d,s,b} \Big(Y'^D_{\alpha\beta}\bar{Q}'_{\alpha L}  \Phi q'^D_{\beta R} + Y'^{D*}_{\alpha\beta}\bar{q}'^D_{\beta R}  \Phi^\dagger Q'_{\alpha L}\Big)
-  \nonumber\\
- & & {} -\sum_{\alpha=1,2,3} \sum_{\beta=u,c,t} \Big(Y'^U_{\alpha\beta}\bar{Q}'_{\alpha L}   \widetilde{\Phi} q'^U_{\beta R} + Y'^{U*}_{\alpha\beta}\bar{q}'^U_{\beta R} \widetilde{\Phi}^\dagger Q'_{\alpha L}\Big).
-\end{eqnarray}
-
-The first two lines of the lagrangian summarize the kinetic terms for the fermionic fields and their coupling to the gauge bosons $A^{\mu\nu}_a$, $B^{\mu\nu}$ \footnote{In gauge theories the ordinary derivative $\partial_\mu$  is substitued with èthe covariant derivative $D_\mu$.
-Here $D_\mu = \partial_\mu + igA_\mu \cdot I + ig'B_\mu\frac{Y}{2}$, where I and Y are the SU(2)$_L$ and U(1)$_Y$ generators, respectively.}.
-The third line describes the kinetic terms and the self-coupling terms of the gauge bosons.
-The forth line is the Higgs lagrangian, which results in the spontaneous symmetry breaking.
-The last three lines describe the Yukawa coupling between fermions and the Higgs field, origin of the fermion's mass.
-
-The coupling between left-handed and right-handed field generates the mass term for fermions.
-The SM assumes only left-handed components for neutrinos, thus implying zero neutrino mass.
-Since any linear combination of massless fields results in a massless field, the flavor eigenstates are identical to the mass eigenstates in the SM.
-
 \section{Beyond the Standard Model}
-The discovery of neutrino oscillation and its implication of non-zero neutrino mass marks  the beginning of a new, exciting era in neutrino physics: the era of physics Beyond the Standard Model (BSM) in the neutrino sector.
+The discovery of neutrino oscillation marks  the beginning of a new, exciting era in neutrino physics: the era of physics Beyond the Standard Model (BSM) in the neutrino sector.
 We are currently searching for new, deeper theories that can accommodate neutrinos with non-zero mass, while remaining consistent with the rest of the Standard Model.
-%More excitingly, we need to probe these theories experimentally.
+More excitingly, we need to probe these theories experimentally.
 \subsection{Neutrino Oscillations}

### 2018-01-30 Chapters/1-LArTPCs.tex
it is been a while

\chapter{Liquid Argon Detectors at the Intensity Frontier}
Following the recommendation of the latest Particle Physics Project Prioritization Panel  \cite{ParticlePhysicsProjectPrioritizationPanel(P5):2014pwa}, the US is dedicating substantial resources to the development of a short- and long- baseline neutrino program to address the fundamental questions in neutrino physics today. A large portion of this program pivots on the Liquid Argon Time Projection Chamber (LArTPC) detector technology. In the next few years, LArTPC experiments -- such as the Short-Baseline Neutrino program (SBN) and DUNE -- will be major players in the intensity frontier field.

\section{The SBN Program}
\subsection{SBN Goals}
\subsection{Neutrino Interactions and Detection }

\section{DUNE}
\subsection{DUNE Non-Accelerator Physics Program}
\subsection{Rare Decay Searches: Experimental Limit}
\subsection{Nucleon Decay Detection in LAr}

\section{Liquid Argon Time Projection Chambers at the Intensity Frontier}

%Bubble-chamber experiments played a key role in probing the properties of ?-interactions. The Liquid Argon Time Projection Chambers (LArTPC) technology,  first proposed by C.Rubbia in 1977 with ICARUS project [14], is considered the modern evolution of bubble-camber concept, with the additional features of three-dimensional event reconstruction, high-resolution calorimetry, active mass coincident with detector sensitive mass and can intrinsically supply a trigger signal (self-triggering) by means of the scintillation light produced in the liquid noble gas. This technology is ideal to perform $\nu$-studies in a broad energy range, from MeV up to few GeV, with high event reconstruction efficiency, thanks to the capability of particle identifcation and detailed reconstruction of different interaction topologies. In Figure 1.4 is shown a neutrino interaction event, producing a proton, a pion and a muon, as seen in a bubble chamber and in a LArTPC.

\subsection{Time Projection Chamber}
\subsection{Ionization Detectors with Noble Liquids}
\subsection{LArTPC: Principles of Operation}
\subsection{Liquid Argon Ionization Charge Detection}
\subsubsection{Electron Life Time \& purity}
\subsubsection{Space Charge Effect}
\subsubsection{Recombination Effect}
\subsection{Liquid Argon scintillation Light Detection}
\subsubsection{LAr Scintillation Process}
\subsubsection{Wavelength Shifting of LAr Scintillation Light}

### 2018-01-30 Chapters/6-TrackingOptimization.tex
it is been a while

\chapter{Tracking Optimization}
Your first chapter is probably an introduction. But who knows.

### 2018-02-04 Chapter-0/0-Introduction.tex
s

@@
 G_{SM} = SU(3)_C  \otimes SU(2)_T \otimes U(1)_Y
+\label{eq:SMGroup}
 \end{equation}
@@
 
-\section{Neutrinos in the Standard Model}
-The neutrino was introduced in the SM as a left-handed massless Weyl spinor.
+\section{Neutrinos: in and beyond the Standard Model}
+\subsection{Neutrino in the SM}
+Neutrino were introduced in the SM as left-handed massless Weyl spinors.
 The Dirac equation of motion
@@
 
+\subsection{Neutrino Oscillations}
+The determination of the flavor of a neutrino dynamically arises from the corresponding charged lepton associated in a CC interaction; for example, a $\nu_e$ is a neutrino which produces an $e^-$, a $\bar\nu_\mu$ is a neutrino which produces a $\mu^+$, $etc$.
+The neutrino flavor eigenstates $\ket{\nu_\alpha}$,  with $\alpha = e,\mu,\tau$, are orthogonal to each other and form a base for the the weak interaction matrix.
+
+Overwhelming experimental data show neutrinos change flavor during their propagation.
+This phenomenon, predicted by Bruno Pontecorvo in 1957, is  called ``neutrino oscillations" and it is possible only if the flavor eigenstate are not identical to the mass eigenstates.
+A minimal extension of the Standard Model introduces three mass eigenstates, $\ket{\nu_i}$ ($i = 1,2, 3$), whose mass $m_i$ is well defined.
+The unitary Pontecorvo-Maki-Nakagawa-Sakata matrix transforms the spinor wave functions ($\psi$) of each component  between flavor and mass bases as follows
+
+\begin{equation} 
+\sum \psi_\alpha \ket{\nu_\alpha} =  \sum \psi_i \ket{\nu_i}, \rightarrow \psi_\alpha =  U_{PMNS} \psi_i, 
+\end{equation}
+
+with
+
+\begin{equation}
+U_{PMNS}=\\
+\left[ 
+\begin{array}{ccc}
+c_{12} & s_{12} & 0\\
+-s_{12} & c_{12} & 0\\
+0 & 0 & 1\\
+\end{array}
+\right]
+\left[ 
+\begin{array}{ccc}
+c_{13} & 0 &s_{13}e^{-i\delta} \\
+0 & 1 & 0\\
+-s_{13}e^{-i\delta}& 0 & c_{13}\\
+\end{array}
+\right]
+\left[
+\begin{array}{ccc}
+1 & 0 & 0\\
+0 & c_{23} & s_{23} \\
+0& -s_{23} & c_{23} \\
+\end{array}
+\right]
+\left[ 
+\begin{array}{ccc}
+e^{i\alpha_1} & 0 & 0\\
+0& e^{i\alpha_2} & 0\\
+0 & 0 & 1\\
+\end{array}
+\right]
+\end{equation}
+
+where $c$ e $s$ stand respectively for cosine and sine of the corresponding  mixing angles ($\theta_{12}$, $\theta_{23}$ and $\theta_{13}$), $\delta$ è is the Dirac CP violation phase, $\alpha_1$ and $\alpha_{2}$   è is the eventual Majorana CP violation phases.
+
+\textcolor{red}{Experimental results summary}
+
 \section{Beyond the Standard Model}
-The discovery of neutrino oscillation and its implication of non-zero neutrino mass marks  the beginning of a new, exciting era in neutrino physics: the era of physics Beyond the Standard Model (BSM) in the neutrino sector.
+The discovery of neutrino oscillation and its implication of non-zero neutrino mass mark  the beginning of a new, exciting era in neutrino physics: the era of physics Beyond the Standard Model (BSM) in the neutrino sector.
 We are currently searching for new, deeper theories that can accommodate neutrinos with non-zero mass, while remaining consistent with the rest of the Standard Model.
 %More excitingly, we need to probe these theories experimentally.
-\subsection{Neutrino Oscillations}
-\subsection{Proton Decay}
+
+\subsubsection{Open Questions in Neutrino Physics}
+On one hand, the last X decades of experiments in neutrino oscillations brought spectacular advancements in the understanding of the oscillations pattern, measuring the neutrino mixing angles and mass splitting with a precision of less than 10\%.
+On the other, it opened the field for a series of questions needing experimental answers.
+Critical challenges in the next decade will entail an even deeper understanding of the neutrino mixing pattern, investigation of the neutrino mass origin, neutrino number and nature, and assessing CP violation in the lepton sector.
+
+Following the recommendation of the latest Particle Physics Project Prioritization Panel  \cite{ParticlePhysicsProjectPrioritizationPanel(P5):2014pwa}, the US  is dedicating substantial resources to the development of a short- and long- baseline neutrino program to address many of these fundamental questions.
+This program pivots on the Liquid Argon Time Projection Chamber (LArTPC) detector technology which will be described in \ref{ch:1}.
+
+The main goals of these research programs include:
+\begin{itemize}
+\item[-] Assessment of the existence of right-handed sterile neutrinos. 
+\item[-] Determination of the sign of $\Delta m^2_{13}$ (or $\Delta m^2_{23}$), i.e., the ``neutrino mass ordering".
+\item[-] Determination of the octant, i.e.  whether $\theta_{23}$ is maximal.
+\item[-] Determination the status of CP symmetry in the lepton sector.
+\end{itemize}
+
+\textcolor{red}{Smallness of neutrino masses, link to GUTs}
+
+\subsubsection{Towards a more fundamental theory}
+Despite its highly predictive power, a number of conceptual issues arise in the SM which disfavor it to be a good candidate for a fundamental theory.
+
+The SM rather complex group structure, where a gauge group is formed with the direct product of other three groups as shown in eq.
+\ref{eq:SMGroup},  is unexplained.
+Also, the SM fails to include a suitable dark matter candidate and a mechanisms that accounts for the baryon asymmetry of the universe.
+Within the SM, a total of 25 parameters remain seemingly arbitrary and need to be fitted to data: 3 gauge couplings, 9 charged fermion masses, 3 mixing angles and one CP phase in the CKM matrix, the Higgs mass and quartic coupling, $\theta_{QCD}$, 3 neutrino masses, 3 neutrino mixing angles, 1 Dirac phase and, eventually,  2 Majorana phases.

### 2018-02-04 Chapters/1-LArTPCs.tex
s

@@
-\chapter{Liquid Argon Detectors at the Intensity Frontier}
+\chapter{Liquid Argon Detectors at the Intensity Frontier}\label{ch:1}
 Following the recommendation of the latest Particle Physics Project Prioritization Panel  \cite{ParticlePhysicsProjectPrioritizationPanel(P5):2014pwa}, the US is dedicating substantial resources to the development of a short- and long- baseline neutrino program to address the fundamental questions in neutrino physics today.

### 2018-02-04 Chapters/6-TrackingOptimization.tex
s

@@
 \chapter{Tracking Optimization}
-Your first chapter is probably an introduction. But who knows.
+Understanding how kaons and pions are tracked inside the TPC and optimizing the reconstruction algorithms to maximize the correct identification of the interaction point is a fundamental step of the analysis.
+
+\section{MC sample and WC2TPC match}
+The optimization is performed on a MC sample of \textcolor{red}{HOW MANY?} kaons and \textcolor{red}{HOW MANY?} pions produced with the DDMC technique. 
+DDMC particles are shot from the WC4 location into the TPC following the beam profile.
+We mimic the matching between the WC and the TPC track on Monte Carlo by constructing a fake WC track using truth information at the true initial point.
+We then apply the same WC to TPC matching algorithm as in data.
+
+Plots I want in this section:
+\begin{enumerate}
+\item WC2TPC MC DeltaX, DeltaY and $\alpha$
+\item Delta L, reco - true
+\item Delta L, reco - true Elastic, Delta L, reco - true Inelastic, other
+\item Length Quality cut
+\item Efficiency as a function of true KE and Angle
+\end{enumerate}

### 2018-02-10 Chapters/6-TrackingOptimization.tex
s

@@
 \section{MC sample and WC2TPC match}
-The optimization is performed on a MC sample of \textcolor{red}{HOW MANY?} kaons and \textcolor{red}{HOW MANY?} pions produced with the DDMC technique. 
+The optimization is performed on a MC sample of \textcolor{red}{HOW MANY?} kaons and 359000 pions produced with the DDMC technique.
 DDMC particles are shot from the WC4 location into the TPC following the beam profile.
-We mimic the matching between the WC and the TPC track on Monte Carlo by constructing a fake WC track using truth information at the true initial point.
+We mimic the matching between the WC and the TPC track on Monte Carlo by constructing a fake WC track using truth information at wire chamber four.
 We then apply the same WC to TPC matching algorithm as in data.
@@
 \end{enumerate}
+
+\section{Wire chamber-to-TPC match}
+\subsection{Wire chamber-to-TPC match: importance and definition}
+\textcolor{red}{something something about why this match is important}
+
+In data, we attempt to uniquely match one WC-Track to one and only one reconstructed TPC track.
+This match is done by using in the $X$ and $Y$ coordinate of the extrapolated WC-Track to the upstream most point of the reconstructed TPC Track and by using the angle between the incoming track angle and the reconstructed TPC.
+We define $\Delta$X as the difference between the $x$ position of the most upstream point of the TPC track and the $x$ position of the WC track as projected to the TPC front face.
+$\Delta$Y is defined analogously.
+We define  $\Delta$R as $ \Delta \text{R} =  \sqrt{ \Delta \text{X}^2 +  \Delta \text{Y}^2}  $.
+The angle between the incident WC Track and the TPC track in the plane that contains them defines $\alpha$.
+
+We define a match between WC-track and TPC reconstructed track if  $\Delta \text{R} < r_{T}$, $\alpha < \alpha_{T}$ and the Z position of the first reconstructed point of the TPC track is within 2 cm from the TPC front face.
+The determination of the best $r_{T}$ and $\alpha_{T}$ is the scope of the following section.
+
+In MC, we mimic the matching between the WC and the TPC track on Monte Carlo by constructing a fake WC track using truth information at wire chamber four.
+We then apply the same WC to TPC matching algorithm as in data.
+
+\subsection{Matching optimization}
+Scope of this optimization study is assessing the goodness of the wire chamber-to-TPC match on Monte Carlo and porting the optimized selection cuts to data.
+A word of caution is necessary here.
+With this study, we want to minimize pathologies associated with the presence of the primary kaon itself, like the incorrect association between the beamline kaon its decay products inside the TPC.
+Assessing the contamination from pile-up\footnote{We remind the reader that the DDMC is a single particle Monte Carlo, where the beam pile up is not simulated.}, albeit related, is beyond the scope of this study .
+
+In MC, we are able to define a correct WC2TPC match using the Geant4 truth information.
+We are thus able to count how many times the WC tracks get associated incorrectly with the wrong TPC reconstructed track.
+
+%We define two populations: TPC tracks correctly matched and one of all the other match.
+We define a correct match if the all following conditions are met:
+\begin{itemize}
+\item[-] the length of the true primary Geant4 track in the TPC is greater than 2 cm,  
+\item[-] the length of the reconstructed track length is greater than 2 cm,
+\item[-] the Z position of the first reconstructed point is within 2 cm from the TPC front face
+\item[-] the distance between the reconstructed track and the true entering point is the minimum compared with all the other reconstructed tracks.
+\end{itemize}
+
+In order to count the wrong matches, we consider all the reconstructed tracks whose Z position of the first reconstructed point lies within 2 cm from the TPC front face.
+Events with true length in TPC < 2 cm included.
+Since kaons are shot 100 cm upstream from the TPC front face, the following two scenarios are possible from a truth standpoint: 
+\begin{itemize}
+\item[[$Ta$]] the primary kaon decays or interact strongly before getting to the TPC,
+\item[[$Tb$]] the primary kaon enters the TPC.
+\end{itemize}
+
+Once we choose the selection cuts to determine a reconstructed wire chamber-to-TPC match, the following four scenarios are possible : 
+\begin{itemize}
+\item[1)] no reconstructed tracks are matched
+\item[2)] the correct track and one (or more) wrong tracks are matched
+\item[3)] only the correct track is matched
+\item[4)] one (or more) wrong track is matched.
+\end{itemize}
+
+If we keep only events with one and only one match, we discard cases 1) and 2) from the cross section measurement.
+For each set of $\alpha$ and $\Delta R$ selection value, we define purity and efficiency of the selection as follows:
+\begin{equation}
+\text{Efficiency} = \frac{\text{events with correct track matched}}{\text{events with primary in TPC}}
+\end{equation}
+
+\begin{equation}
+\text{Purity} = \frac{\text{Events with correct track matched}}{\text{Events with correct track matched + Events with wrong track matched}}
+\end{equation}

### 2018-02-13 Chapter-6/6-TrackingOptimization.tex
a bit of tracking

--- 2018-02-10 Chapters/6-TrackingOptimization.tex
@@
 \section{MC sample and WC2TPC match}
-The optimization is performed on a MC sample of \textcolor{red}{HOW MANY?} kaons and 359000 pions produced with the DDMC technique.
-DDMC particles are shot from the WC4 location into the TPC following the beam profile.
+The optimization is performed on a MC sample of 191000 kaons and 359000 pions produced with the DDMC technique. DDMC particles are shot from the WC4 location into the TPC following the beam profile.
 We mimic the matching between the WC and the TPC track on Monte Carlo by constructing a fake WC track using truth information at wire chamber four.
@@
 With this study, we want to minimize pathologies associated with the presence of the primary kaon itself, like the incorrect association between the beamline kaon its decay products inside the TPC.
-Assessing the contamination from pile-up\footnote{We remind the reader that the DDMC is a single particle Monte Carlo, where the beam pile up is not simulated.}, albeit related, is beyond the scope of this study .
+Assessing the contamination from pile-up\footnote{We remind the reader that the DDMC is a single particle Monte Carlo, where the beam pile up is not simulated.}, albeit related, is beyond the scope of this study.
 
-In MC, we are able to define a correct WC2TPC match using the Geant4 truth information.
-We are thus able to count how many times the WC tracks get associated incorrectly with the wrong TPC reconstructed track.
+In MC, we are able to define a correct WC2TPC match using the Geant4 truth information. We are thus able to count how many times the WC tracks is associated with the wrong TPC reconstructed track.
 
@@
 In order to count the wrong matches, we consider all the reconstructed tracks whose Z position of the first reconstructed point lies within 2 cm from the TPC front face.
-Events with true length in TPC < 2 cm included.
+Events with true length in TPC $<$ 2 cm are included.
 Since kaons are shot 100 cm upstream from the TPC front face, the following two scenarios are possible from a truth standpoint: 
@@
 
-Once we choose the selection cuts to determine a reconstructed wire chamber-to-TPC match, the following four scenarios are possible : 
+Once we choose the selection cuts to determine a reconstructed wire chamber-to-TPC match $r_{T}$ and $\alpha_{T}$, the following four scenarios are possible : 
 \begin{itemize}
@@
 If we keep only events with one and only one match, we discard cases 1) and 2) from the cross section measurement.
-For each set of $\alpha$ and $\Delta R$ selection value, we define purity and efficiency of the selection as follows:
+For each set of $r_{T}$ and $\alpha_{T}$ selection value, we define purity and efficiency of the selection as follows:
 \begin{equation}
-\text{Efficiency} = \frac{\text{events with correct track matched}}{\text{events with primary in TPC}}
+\text{Efficiency} = \frac{\text{Number of events correctly matched}}{\text{ Number of events with primary in TPC}}
 \end{equation}
@@
 \begin{equation}
-\text{Purity} = \frac{\text{Events with correct track matched}}{\text{Events with correct track matched + Events with wrong track matched}}
+\text{Purity} = \frac{\text{Number of events correctly matched}}{\text{Total number of matched events}}.
 \end{equation}
+
+Figure \ref{fig:EffPurityK} shows the efficiency (left) and purity (right) for wire chamber-to-TPC match as a function of the radius, $r_{T}$, and angle, $\alpha_{T}$, selection value.
+It is apparent how both efficiency and purity are fairly flat as a function of the radius selection value at a given angle.
+This is not surprising, given the fact that the wrong matches can occur  in a single particle gun MC  only for mis-tracking of the primary or for association with decay products, which are generally different in angle, but might be more similar in $x$ and $y$ projection.
+The radius cut would play a key role in removing pile up events.
+
+For LArIAT cross section measurements, we generally prefer purity over efficiency, since a sample of particle of a pure species will lead to a better measurement.
+In the case of kaons however, efficiency needs not to drop too low, given the smallness of the kaon sample.
+We choose $(\alpha_{T}$, $r_{T}) = (8 \text{ deg}, 4 \text{ cm} )$ and get a MC 85\% efficiency and 98\% purity.
+
+\begin{figure}[hpbt]
+\centering
+\includegraphics[width=15cm]{Chapter-6/Images/KEffPurity.png}
+\caption{Efficiency (left) and purity (right) for wire chamber-to-TPC match as a function of the radius and angle selections.}
+\label{fig:EffPurityK}
+\end{figure}
+
+\subsection{Porting optimization to data}

### 2018-03-11 AppentixA-EField/A-LArIATEField.tex
s

\chapter{Measurement of LArIAT Electric Field}
The electric field of a LArTPC in the drift volume is a fundamental quantity for the proper functionality of this technology, as it affects almost every reconstructed quantity such as the position of hits or their collected charge. Given its importance, we calculate the electric field for LArIAT with a single line diagram from our HV circuit and we cross check the obtained value with a measurement relying only on TPC data.

Before getting into the details of the measurement procedures, it is important to explicit the relationship between some  quantities in play. The electric field and the drift velocity ($v_{drift}$) are related as follows 
\begin{equation} v_{drift} = \mu(E_{field},T) E_{field}, \label{eq:vd}
\end{equation}
where $\mu$ is the electron mobility, which depends on the electric field and on the temperature (T). The empirical formula for this dependency is described in ~\cite{WWW} and shown in Figure \ref{fig:EV} for several argon temperatures.

\begin{figure}[htb]
\centering
\includegraphics[scale=0.45]{./AppentixA-EField/Images/Walkowiak.png}\\
\caption{Drift velocity dependence on electric field for several temperatures. The slope of the line at any one point represents the electron mobility for that given temperature and electric field.}
\label{fig:EV}
\end{figure}

The relationship between the drift time ($t_{drift}$) and the drift velocity is trivially given by
\begin{equation}
t_{drift} = \Delta x/v_{drift}, \label{eq:drifttime}
\end{equation}
where $\Delta x$ is the distance between the edges of the drift region.
Table \ref{tab:Efields} reports the values of the electric field, drift velocity, and drift times for the smaller drift volumes.

\begin{table}[]
\centering
\caption{Electric field and drift velocities in LArIAT smaller drift volumes}
\label{tab:Efields}
\begin{tabular}{|l|l|l|}
\hline
& Shield-Induction & Induction-Collection \\ \hline
E$_{filed}$ &                 700.625 V/cm        &                892.5  V/cm             \\ \hline
v$_{drift}$ &                   1.73  mm/$\mu$s   &                  1.90 mm/$\mu$s        \\ \hline
t$_{drift}$ &                   2.31  $\mu$s      &                   2.11 $\mu$s          \\ \hline

\end{tabular}
\end{table}

With these basic parameters established, we can now move on to calculating the electric field in the main drift region (between the cathode and the shield plane).

\subsection*{Single line diagram method}
The electric field strength in the LArIAT main drift volume can be determined knowing the voltage applied to the cathode, the voltage applied at the shield plane, and the distance between them. We assume the distance between the cathode and the shield plane to be 470 mm and any length contraction due to the liquid argon is negligibly small ($\sim$2~mm).

The voltage applied to the cathode can be calculated using Ohm's law and the single line diagram shown in Figure \ref{fig:circuit}.  A set of two of filter pots for emergency power dissipation are positioned between the Glassman power supply and the cathode, one at each end of the feeder cable, each with an internal resistance of 40~M$\Omega$.  The output current of the Glassman power supply is then used to determine the electric field strength.  Figure \ref{fig:currentMeasurement} shows an average current of 0.004172 mA from the Glassman power supply.

\begin{figure}[h]
\centering
\begin{minipage}{0.45\textwidth}
\centering
\includegraphics[width=3in]{AppentixA-EField/Images/CircuitLArIAT.png}
\caption{\textcolor{red}{get rid of current line} LArIAT HV simple schematics.}
\label{fig:circuit}
\end{minipage}\hfill
\begin{minipage}{0.45\textwidth}
\centering
\includegraphics[width=3in]{AppentixA-EField/Images/glassman_current_20160525-30.png}
\caption{\textcolor{red}{the axis is wrong!!} Current reading from the Glassman between May 25th and May 30th, 2016 (typical Run-II conditions).}
\label{fig:currentMeasurement}
\end{minipage}
\end{figure}

Using this current, the voltage at the cathode is calculated as
\begin{equation} \label{eq:VBC}
V_{BC}=V_{PS} - (I \times R_{eq}) = -23.5\text{ kV} + ( 0.00417\text{ mA} \times 80\text{ M}\Omega ) = -23.17\text{ kV}, 
\end{equation}
where $I$ is the current and $R_{eq}$ is the equivalent resistor representing the two filter pots. The electric field, drift voltage, and drift time are then calculated to be
\begin{equation}E_{\text{field}} = \frac{V_{BC} - V_{\text{shield}}}{\Delta x} = 486.54\text{ V/cm}
\end{equation}
%\begin{equation}v_{drift} = \mu E_{field} = 1.5097 \textit{ mm/$\mu$s}
%\end{equation}
%\begin{equation}t_{drift} = \frac{\Delta x}{v_{drift}} = 311.316 \textit{ $\mu$s.}
%\end{equation}

\subsection*{E field using cathode-anode piercing tracks}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[b]
\centering
\begin{minipage}{0.45\textwidth}
\centering
\includegraphics[width=3in]{AppentixA-EField/Images/TPCCrossSectionView.png}
\caption{Pictorial representation of the YX view of the TPC. The distance within the anode planes and between the shield plane and the cathode is purposely out of proportion to illustrate the time difference between hits on collection and induciton. A ACP track is shown as an example.}
\label{fig:Scheme}
\end{minipage}\hfill
\begin{minipage}{0.45\textwidth}
\centering
\includegraphics[width=3in]{AppentixA-EField/Images/AngleDef.png}
\caption{Angle definition in the context of LArIAT coordinates system.}
\label{fig:AngleDef}
\end{minipage}
\end{figure}
We devise an independent method to measure the drift time (and consequently drift velocity and electric field) using TPC cathode to anode piercing tracks. We use this method as a cross check to the single line method.
The basic idea is simple:
\begin{itemize}
\item[0.] Select cosmic ray events with only 1 reconstructed track 
\item[1.] Reduce the events to the one contaoning tracks that cross both anode and cathode
\item[2.] Identify the first and last hit of the track
\item[3.] Measure the time difference between these two hits ($\Delta t$).
\end{itemize}
This method works under the assumptions that the time it takes for a cosmic particle to cross the chamber ($\sim$ns) is small compared to the charge drift time ($\sim$ hundreds of $\mu$s).

We choose cosmic events to allow for a high number of anode to cathode piercing tracks (ACP tracks), rejecting  beam  events where the particles travel almost perpendicularly to drift direction. We select events with only one reconstructed track to  maximize the chance of selecting a single crossing muon (no-michel electron). We utilize ACP tracks because their hits span the full drift length of the TPC, see figure \ref{fig:Scheme}, allowing us to define where the first and last hit of the tracks are located in space regardless of our assumption of the electric field. %The definition of the last hit is easy: it is the hit closest to the cathode. The definition of the first hit is a bit more complicated. A track that crosses the anode planes will deposit charge in the small drift volumes (S-I and I-C). The drift time in S-I and I-C region was already calculated in Table \ref{tab:Efields}. This is to say that the position of the first hit matters when calculating the drift time at the microsecond precision. Single hits on the collection plan will not form a 3D object. This means that we can safely exclude that the reconstruction of ACP tracks starts at the cathode plane (black point in figure \ref{fig:Scheme}). %Understanding if the first hit of a track in on the induction or on the shield plan is more complicated. For now, we'll take an uncertainty hit of 2.3 $\mu$s.

One of the main features of this method is that it doesn't rely on the measurement of the trigger time. Since $\Delta t$ is the time difference between the first and last hit of a track and we assume the charge started drifting at the same time for both hits, the measurement of the absolute beginning of drift time $t_0$ is unnecessary. We boost the presence of ACP tracks in the cosmic sample by imposing the following requirements on tracks:

\begin{itemize}
\item vertical position (Y) of first and last hits within $\pm$ 18 cm from TPC center (avoid Top-Bottom tracks) 
\item horizontal position (Z) of first and last hits within 2 and 86 cm from TPC front face (avoid through going tracks) 
\item track length greater than 48 cm (more likely to be crossing)
\item angle from the drift direction (phi in figure \ref{fig:AngleDef}) smaller than 50 deg (more reliable tracking)
\item angle from the beam direction (theta in figure \ref{fig:AngleDef}) grater than 50 deg (more reliable tracking)
\end{itemize}

Tracks passing all these selection requirements are used for the $\Delta t$ calculation.

For each track passing our selection, we loop through the associated hits in order to retrieve the timing information. The analysis is performed separately on hits on the collection plane and induction plane, but lead to consistent results. As an example of the time difference, figures \ref{fig:Run2PosColFit} and \ref{fig:Run2PosIndFit} represent the difference in time between the last and first hit of the selected tracks for Run-II Positive Polarity sample on the collection and induction plane respectively.  We fit with a Gaussian to the peak of the $\Delta t$ distributions to extract the mean drift time and the uncertainty associated with it. The long tail at low $\Delta t$ represent contamination of non-ACP tracks in the track selection.  We apply the same procedure to Run-I and Run-II, positive and negative polarity alike.

\begin{figure}[h!]
\begin{minipage}{0.40\textwidth}
\centering
\includegraphics[width=3in]{AppentixA-EField/Images/RunIIPosCol.png}
\caption{Collection plane $\Delta$t fit for Run II positive polarity ACP data  selected tracks.}
\label{fig:Run2PosColFit}
\end{minipage}\hfill
\begin{minipage}{0.40\textwidth}
\centering
\includegraphics[width=3in]{AppentixA-EField/Images/RunIIPosInd.png}
\caption{Induction plane $\Delta$t fit for Run II positive polarity ACP data  selected tracks.}
\label{fig:Run2PosIndFit}
\end{minipage}
\end{figure}

To convert $\Delta t$ recorded for the hits on the induction plane to the drift time we utilize the formula
\begin{equation}
t_{drift} = \Delta t - t_{S-I}
\end{equation}
where $t_{drift}$ is the time the charge takes to drift in the main volume between the cathode and the shield plane and $t_{S-I}$ is the time it takes for the charge to drift from the shield plane to the induction plane. In Table \ref{tab:Efields} we calculated the drift velocity in the S-I region, thus we can calculate $t_{S-I}$ as 
\begin{equation}
t_{S-I} = \frac{l_{S-I}}{v_{S-I}} = \frac{4 mm}{1.745 mm/ \mu s}
\end{equation}
where $\l_{S-I}$ is the distance between the shield and induction plane and $v_{S-I}$ is the drift velocity in the same region. A completely analogous procedure is followed for the hits on the collection plane, taking into account the time the charge spent in drifting from shield to induction as well as between the induction and collection plane
The value for $\Delta t_{drift}$ , the calculated drift velocity ($v_{drift}$), and corresponding drift electric field for the various run periods is given in Table \ref{tab:deltaTACP} and are consistent with the electric field value calculated with the single line diagram method.

\begin{center}
\begin{table}[htb]
  \begin{center}
    %\resizebox{0.45\textwidth}{!}{%
    \begin{tabular}{|c|c|c|c|}
      \multicolumn{4}{c}{\textbf{Delta t$_{drift}$, drift v and E field with ACP tracks}} \\
      \hline \hline
       Data Period  & $\Delta t_{Drift}$ [$\mu s$] & Drift velocity [mm/$\mu$s] & E field [V/cm] \\
       \hline
       RunI Positive Polarity Induction &  311.1 $\pm$ 2.4   &1.51 $\pm$ 0.01  & 486.6 $\pm$ 21\\
       \hline
       RunI Positive Polarity Collection &  310.9 $\pm$ 2.6 & 1.51 $\pm$ 0.01  &  487.2 $\pm$ 21\\
       \hline
       RunII Positive Polarity Induction &   315.7 $\pm$ 2.8 & 1.49 $\pm$ 0.01 &  467.9 $\pm$ 21\\
       \hline
       RunII Positive Polarity Collection &  315.7 $\pm$ 2.7 & 1.49 $\pm$ 0.01 &  467.9 $\pm$ 21\\
       \hline
       RunII Negative Polarity Induction &   315.9 $\pm$ 2.6 & 1.49 $\pm$ 0.01  & 467.1 $\pm$ 21 \\
       \hline
       RunII Negative Polarity Collection &  315.1 $\pm$ 2.8 & 1.49 $\pm$ 0.01  & 470.3 $\pm$ 21  \\
       \hline
       \hline
       Average Values & 314.1 & 1.50 $\pm$ 0.01 & 474.3 $\pm$ 21 \\
       \hline
       \end{tabular}
    \caption{$\Delta t$ for the different data samples used for the Anode-Cathode Piercing tracks study. }
    \label{tab:deltaTACP}
    \end{center}
\end{table}
\end{center}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2018-03-11 Chapter-4/3-KaonInteractionsInLAr.tex
s

\chapter{Kaons Interactions in Argon: Cross Section}\label{ch:Interactions}
\section{Literature Review}
\section{How to Measure Hadron Cross Section in LArIAT}\label{ch:methodology}
\subsection{Particle Selection}\label{ch:ParticleSelectionMethod}
\subsection{Beamline to TPC Handshake}\label{ch:WC2TPCMatchMethod}
\subsection{The Thin Slice Method}
\subsubsection{Cross Sections on Thin Target}
Cross section measurements on a thin target have been the bread and butter of nuclear and particle experimentalists since the Rutherford experiments \textcolor{red}{NEED CITATION}. At their core, this type of experiments consists in shooting a beam of particles with a known flux on a thin target and recording the outgoing flux.

In general, the target is not a single particle, but rather a slab of material containing many diffusion centers. The so-called  ``thin target" approximation assumes that the target centers are uniformly distributed in the material and the target is thin compared to the interaction length so that no center of interaction sits in front of another. In this approximation, the ratio between the number of particles interacting in the target $N_{Interacting}$ and number of incident particles $N_{incident}$ determines the interaction probability $P_{Interacting}$, which is the complementary to one of the survival probability $P_{Survival}$. 
Equation \ref{eq:thinTargetXS} 
\begin{equation}
P_{Survival} = 1- P_{Interacting} = 1 - \frac{N_{Interacting}}{N_{incident}} = e^{-\sigma_{TOT} n \delta X}
\label{eq:thinTargetXS}
\end{equation}
describes the probability for a particle to survive the thin target. This formula relates the total cross section $\sigma_{TOT}$, the density of the target centers  $n$  and  the thickness of the target  along the incident hadron direction $\delta X$, to the interaction probability\footnote{The scattering center density in the target, {\emph{n}},  relates to the argon density $\rho$, the Avogadro number  $ N_{A} $ and the argon molar mass $m_A$ as $n=\frac{\rho N_{A} }{m_A}$.}. If the target is thin compared to the interaction length of the process considered, we can Taylor expand the exponential function in equation \ref{eq:thinTargetXS} and find a simple proportionality relationship between the number of incident and interacting particles, and the cross section, as shown in equation \ref{eq:thinTargetXSTaylor}:
\begin{equation}
1 - \frac{N_{Interacting}}{N_{incident}} =  1 -\sigma_{TOT} n \delta X + O(\delta X^2).
\label{eq:thinTargetXSTaylor}
\end{equation}

Solving for the cross section, we find:
\begin{equation}
 \sigma_{TOT}  = \frac{1}{n \delta X}\frac{N_{Interacting}}{N_{incident}}.
\label{eq:thinTargetXSSolved}
\end{equation}

\subsubsection{Not-so-Thin Target: Slicing the Argon}
The LArIAT TPC, with its 90 cm of length, is not a thin target. \textcolor{red}{Find expected interaction length for hadrons and kaons}. However, the fine-grained tracking of the LArIAT LArTPC allows us to treat the argon volume as a sequence of many adjacent thin targets.

As described in section \ref{sec:experimentDescription}, LArIAT wire planes count 240 wires each. The wires are oriented at +/- $60^{\circ}$ from the vertical direction at 4 mm spacing, while the beam direction is oriented 3 degrees off the $z$ axis in the $XZ$ plane. \textcolor{red}{review this math} The wires collect signals proportional to the energy loss of the hadron along its path in a  $\delta${\emph{X}} = 4 mm/sin($60^{\circ}$) $\approx$ 4.7~mm slab of liquid argon. Thus, one can think to slice the TPC into many thin targets of $\delta${\emph{X}} = 4.7~mm thickness along the direction of the incident particle.

Considering each slice {\emph{j}}  a ``thin target",  we can apply the cross section calculation from Eq.~\ref{eq:thinTargetXSSolved} iteratively, evaluating the kinetic energy of the hadron as it enters each slice, $E_{j}^{kin}$.  For each WC-to-TPC matched particle, the energy of the hadron entering the TPC is known thanks to the momentum and mass determination by the tertiary beamline,

\begin{equation}
 E^{kin}_{Front Face}  = \sqrt{p^2_{Beam} - m^2_{Beam}} - E_{loss},
\label{eq:thinTargetXSSolved}
\end{equation}
where $E_{loss}$ is a correction for the energy loss in the dead material between the beamline and the TPC front face (more on \ref{sec:Eloss}). The  energy of the hadron at the each slab is determined by subtracting the energy released by the particle in the previous slabs. For example, at the $j^{th}$ point of a track, the kinetic energy will be

\begin{equation}
 E_{j}^{kin} =  E^{kin}_{Front Face} - \sum_{i < j} \Delta E_i,
\label{eq:KEj}
\end{equation}
where $\Delta E_i$ is the energy deposited at each argon slice before the $j^{th}$ point as measured by the calorimetry associated with the tracking.

If the particle enters a slice, it contributes to $N_{incident}( E^{kin})$ in the energy bin corresponding to its kinetic energy in that slice. If it interacts in the slice, it then also contributes to $N_{interacting}(E^{kin})$ in the appropriate energy bin. The cross section as a function of kinetic energy, $\sigma_{TOT}( E^{kin})$ will then be proportional to the ratio $\frac{N_{interacting}( E^{kin})}{N_{incident}( E^{kin})}$ .

The statistical uncertainty for each energy bin is calculated by error propagation from the statistical  uncertainty on $N_{incident}$ and $N_{interacting}$. 
Since the number of incident hadrons in each energy bin is given by a simple counting, we assume that $N_{incident}$ is distributed as a poissonian with mean and $\sigma^2$ equal to $N_{incident}$ in each bin.  
On the other hand, $N_{interacting}$ follows a binomial distribution: a particle in a given energy bin might or might not interact.  The square of the variance for the binomial is given by  
\begin{equation}
\sigma^2 = \mathcal{N}P_{Interacting}(1-P_{Interacting});
\label{eq:binVar}
\end{equation}

since the interaction probability $P_{Interacting}$ is $\frac{ N_{interacting}}{N_{incident}}$ and the number of tries $\mathcal{N}$ is $N_{incident}$, equation \ref{eq:binVar} translates into
\begin{equation}
\sigma^2 = N_{incident}\frac{ N_{interacting}}{N_{incident}} (1-\frac{ N_{interacting}}{N_{incident}}) = N_{interacting}(1-\frac{ N_{interacting}}{N_{incident}}).
\end{equation}

$N_{incident}$ and $N_{interacting}$ are not independent.
The uncertainty on the cross section is thus calculated as 
\begin{equation}
\delta\sigma_{tot}(E) = \sigma_{tot}(E) \Big(\frac{\delta N_{interacting}}{N_{interacting}}+\frac{\delta N_{incident}}{N_{incident}}\Big) 
\end{equation}
where:
\begin{eqnarray}
\delta N_{incident} = \sqrt[]{N_{incident}} \\
\delta N_{interacting} = \sqrt[]{N_{interacting}(1-\frac{ N_{interacting}}{N_{incident}})}.
\end{eqnarray}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Procedure testing with truth quantities}
We describe here a closure test done on Monte Carlo quantities to prove that the methodology of slicing the TPC works within statistical error under the working assumption of perfect reconstruction.

With this test, we compare the  by comparing the results of this method with the prediction of the total hadronic interaction cross section ($\pi^{\pm}$, Ar) for thin-target simulations from two Monte Carlo generators (Geant 4.10.1 with Bertini Cascade model~\cite{geant4, g4bert} and Genie v2.8.2 with intranuke-hA model). The thick target simulation used a simple stand-alone Geant4 simulation  (i.e., no other detector features were taken into account except the geometry of the thin slices in the LAr volume). Fig.~\ref{fig:xsplot} shows the resulting total ${\pi^-}$ cross section extracted by the sliced TPC technique; it agrees well with the Geant 4 thin-target cross section. The Genie thin-target cross section for $^{40}$Ar shown in this figure is significantly different than that of Geant at low kinetic energies due to the fact that Geant4 models are tuned on $^{12}$C while Genie ones are tuned on a much heavier $^{56}$Fe target. The extrapolated cross section predicted for $^{40}$Ar can be then very different, especially in the resonance region where the model is strongly target-dependent from one generator to another.

%The comparison of the Geant4 thin- and thick-target cross section results demonstrates the power of the "sliced TPC" method for the measurement of the ($\pi^{\pm}$, Ar) cross section in LArIAT TPC geometry.

%Having validated the ``Sliced TPC'' technique and shown that the this technique recovers the simulated cross-section, we now move to performing this measurement utilizing the complete LArIAT simulation within the LArSoft \cite{} based simulation.  Furthermore, we move from utilizing particle level MC-truth information to performing fully automated reconstruction of the charged hadron events within the LArTPC.

### 2018-03-11 Chapter-4/3-KaonInteractionsInLAr.tex
s

@@
 \section{Literature Review}
+
+%the prediction of the total hadronic interaction cross section ($\pi^{\pm}$, Ar) for thin-target simulations from two Monte Carlo generators (Geant 4.10.1 with Bertini Cascade model~\cite{geant4, g4bert} and Genie v2.8.2 with intranuke-hA model).
+The thick target simulation used a simple stand-alone Geant4 simulation  (i.e., no other detector features were taken into account except the geometry of the thin slices in the LAr volume).
+Fig.~\ref{fig:xsplot} shows the resulting total ${\pi^-}$ cross section extracted by the sliced TPC technique; it agrees well with the Geant 4 thin-target cross section.
+The Genie thin-target cross section for $^{40}$Ar shown in this figure is significantly different than that of Geant at low kinetic energies due to the fact that Geant4 models are tuned on $^{12}$C while Genie ones are tuned on a much heavier $^{56}$Fe target.
+The extrapolated cross section predicted for $^{40}$Ar can be then very different, especially in the resonance region where the model is strongly target-dependent from one generator to another.
+
+%The comparison of the Geant4 thin- and thick-target cross section results demonstrates the power of the "sliced TPC" method for the measurement of the ($\pi^{\pm}$, Ar) cross section in LArIAT TPC geometry.
+
+%Having validated the ``Sliced TPC'' technique and shown that the this technique recovers the simulated cross-section, we now move to performing this measurement utilizing the complete LArIAT simulation within the LArSoft \cite{} based simulation.
+Furthermore, we move from utilizing particle level MC-truth information to performing fully automated reconstruction of the charged hadron events within the LArTPC.
+
+%Since the kaon cross section in argon has never been measured before, the Geant4 Monte Carlo tunes kaon transportation in argon by extrapolation from lighter and heavier nuclei.
+As shown in the previous section,  kaon data on carbon are available and  can be used as a metric to evaluate the Geant4 prediction performances.
+Figure \ref{fig:TrueCarbon} shows the total hadronic cross section for carbon implemented in Geant4 10.01.p3 overlaid with the Bugg and Friedman data.
+Unfortunately, the current version of Geant4 does not reproduce the data for carbon closely.
+On one hand, this evidence makes us even more wary when using the Monte Carlo in simulating the kaon-argon interactions.
+On the other, it further highlights the importance of kaon measurements.
+
 \section{How to Measure Hadron Cross Section in LArIAT}\label{ch:methodology}
@@
 \subsubsection{Cross Sections on Thin Target}
-Cross section measurements on a thin target have been the bread and butter of nuclear and particle experimentalists since the Rutherford experiments \textcolor{red}{NEED CITATION}.
+Cross section measurements on a thin target have been the bread and butter of nuclear and particle {Iexperimentalists since the Rutherford experiments \textcolor{red}{NEED CITATION}.
 At their core, this type of experiments consists in shooting a beam of particles with a known flux on a thin target and recording the outgoing flux.
@@
 In general, the target is not a single particle, but rather a slab of material containing many diffusion centers.
-The so-called  ``thin target" approximation assumes that the target centers are uniformly distributed in the material and the target is thin compared to the interaction length so that no center of interaction sits in front of another.
-In this approximation, the ratio between the number of particles interacting in the target $N_{Interacting}$ and number of incident particles $N_{incident}$ determines the interaction probability $P_{Interacting}$, which is the complementary to one of the survival probability $P_{Survival}$.
+The so-called  ``thin target" approximation assumes that the target centers are uniformly distributed in the material and that the target is thin compared to the interaction length so that no center of interaction sits in front of another.
+In this approximation, the ratio between the number of particles interacting in the target $N_{Interacting}$ and number of incident particles $N_{Incident}$ determines the interaction probability $P_{Interacting}$, which is the complementary to one of the survival probability $P_{Survival}$.
 Equation \ref{eq:thinTargetXS} 
 \begin{equation}
-P_{Survival} = 1- P_{Interacting} = 1 - \frac{N_{Interacting}}{N_{incident}} = e^{-\sigma_{TOT} n \delta X}
+P_{Survival} = 1- P_{Interacting} = 1 - \frac{N_{Interacting}}{N_{Incident}} = e^{-\sigma_{TOT} n \delta X}
 \label{eq:thinTargetXS}
@@
 \begin{equation}
-1 - \frac{N_{Interacting}}{N_{incident}} =  1 -\sigma_{TOT} n \delta X + O(\delta X^2).
+1 - \frac{N_{Interacting}}{N_{Incident}} =  1 -\sigma_{TOT} n \delta X + O(\delta X^2).
 \label{eq:thinTargetXSTaylor}
@@
 \begin{equation}
- \sigma_{TOT}  = \frac{1}{n \delta X}\frac{N_{Interacting}}{N_{incident}}.
+ \sigma_{TOT}  = \frac{1}{n \delta X}\frac{N_{Interacting}}{N_{Incident}}.
 \label{eq:thinTargetXSSolved}
@@
  E^{kin}_{Front Face}  = \sqrt{p^2_{Beam} - m^2_{Beam}} - E_{loss},
-\label{eq:thinTargetXSSolved}
+\label{eq:enFF}
 \end{equation}
@@
 
-If the particle enters a slice, it contributes to $N_{incident}( E^{kin})$ in the energy bin corresponding to its kinetic energy in that slice.
-If it interacts in the slice, it then also contributes to $N_{interacting}(E^{kin})$ in the appropriate energy bin.
-The cross section as a function of kinetic energy, $\sigma_{TOT}( E^{kin})$ will then be proportional to the ratio $\frac{N_{interacting}( E^{kin})}{N_{incident}( E^{kin})}$ .
+If the particle enters a slice, it contributes to $N_{Incident}( E^{kin})$ in the energy bin corresponding to its kinetic energy in that slice.
+If it interacts in the slice, it then also contributes to $N_{Interacting}(E^{kin})$ in the appropriate energy bin.
+The cross section as a function of kinetic energy, $\sigma_{TOT}( E^{kin})$ will then be proportional to the ratio $\frac{N_{Interacting}( E^{kin})}{N_{Incident}( E^{kin})}$ .
 
-The statistical uncertainty for each energy bin is calculated by error propagation from the statistical  uncertainty on $N_{incident}$ and $N_{interacting}$. 
-Since the number of incident hadrons in each energy bin is given by a simple counting, we assume that $N_{incident}$ is distributed as a poissonian with mean and $\sigma^2$ equal to $N_{incident}$ in each bin.
-On the other hand, $N_{interacting}$ follows a binomial distribution: a particle in a given energy bin might or might not interact.  The square of the variance for the binomial is given by  
+The statistical uncertainty for each energy bin is calculated by error propagation from the statistical  uncertainty on $N_{Incident}$ and $N_{Interacting}$. 
+Since the number of incident hadrons in each energy bin is given by a simple counting, we assume that $N_{Incident}$ is distributed as a poissonian with mean and $\sigma^2$ equal to $N_{Incident}$ in each bin.
+On the other hand, $N_{Interacting}$ follows a binomial distribution: a particle in a given energy bin might or might not interact.  The square of the variance for the binomial is given by  
 \begin{equation}
@@
 
-since the interaction probability $P_{Interacting}$ is $\frac{ N_{interacting}}{N_{incident}}$ and the number of tries $\mathcal{N}$ is $N_{incident}$, equation \ref{eq:binVar} translates into
+since the interaction probability $P_{Interacting}$ is $\frac{ N_{Interacting}}{N_{Incident}}$ and the number of tries $\mathcal{N}$ is $N_{Incident}$, equation \ref{eq:binVar} translates into
 \begin{equation}
-\sigma^2 = N_{incident}\frac{ N_{interacting}}{N_{incident}} (1-\frac{ N_{interacting}}{N_{incident}}) = N_{interacting}(1-\frac{ N_{interacting}}{N_{incident}}).
+\sigma^2 = N_{Incident}\frac{ N_{Interacting}}{N_{Incident}} (1-\frac{ N_{Interacting}}{N_{Incident}}) = N_{Interacting}(1-\frac{ N_{Interacting}}{N_{Incident}}).
 \end{equation}
 
-$N_{incident}$ and $N_{interacting}$ are not independent.
+$N_{Incident}$ and $N_{Interacting}$ are not independent.
 The uncertainty on the cross section is thus calculated as 
 \begin{equation}
-\delta\sigma_{tot}(E) = \sigma_{tot}(E) \Big(\frac{\delta N_{interacting}}{N_{interacting}}+\frac{\delta N_{incident}}{N_{incident}}\Big) 
+\delta\sigma_{tot}(E) = \sigma_{tot}(E) \Big(\frac{\delta N_{Interacting}}{N_{Interacting}}+\frac{\delta N_{Incident}}{N_{Incident}}\Big) 
 \end{equation}
@@
 \begin{eqnarray}
-\delta N_{incident} = \sqrt[]{N_{incident}} \\
-\delta N_{interacting} = \sqrt[]{N_{interacting}(1-\frac{ N_{interacting}}{N_{incident}})}.
+\delta N_{Incident} = \sqrt[]{N_{Incident}} \\
+\delta N_{Interacting} = \sqrt[]{N_{Interacting}(1-\frac{ N_{Interacting}}{N_{Incident}})}.
 \end{eqnarray}
@@
 \subsection{Procedure testing with truth quantities}
-We describe here a closure test done on Monte Carlo quantities to prove that the methodology of slicing the TPC works within statistical error under the working assumption of perfect reconstruction.
+The $\pi^{-}$-Ar and K$^{+}$-Ar total hadronic cross section implemented in Geant4 can be used as a tool to validate the measurement methodology.
+We describe here a closure test done on Monte Carlo to prove that the methodology of slicing the TPC retrieves the underlying cross section distribution implemented in Geant4 within the statistical error.
+%under the working assumption of perfect reconstruction.
 
-With this test, we compare the  by comparing the results of this method with the prediction of the total hadronic interaction cross section ($\pi^{\pm}$, Ar) for thin-target simulations from two Monte Carlo generators (Geant 4.10.1 with Bertini Cascade model~\cite{geant4, g4bert} and Genie v2.8.2 with intranuke-hA model).
-The thick target simulation used a simple stand-alone Geant4 simulation  (i.e., no other detector features were taken into account except the geometry of the thin slices in the LAr volume).
-Fig.~\ref{fig:xsplot} shows the resulting total ${\pi^-}$ cross section extracted by the sliced TPC technique; it agrees well with the Geant 4 thin-target cross section.
-The Genie thin-target cross section for $^{40}$Ar shown in this figure is significantly different than that of Geant at low kinetic energies due to the fact that Geant4 models are tuned on $^{12}$C while Genie ones are tuned on a much heavier $^{56}$Fe target.
-The extrapolated cross section predicted for $^{40}$Ar can be then very different, especially in the resonance region where the model is strongly target-dependent from one generator to another.
+For pions in the considered energy range, \textcolor{red}{the Geant4 inelastic model adopted to is ``BertiniCascade", while the elastic model ``hElasticLHEP".}
+For kaons, the Geant4 inelastic model adopted to is ``BertiniCascade", while the elastic model ``hElasticLHEP".
 
-%The comparison of the Geant4 thin- and thick-target cross section results demonstrates the power of the "sliced TPC" method for the measurement of the ($\pi^{\pm}$, Ar) cross section in LArIAT TPC geometry.
+For the validation test, we fire about 390000 pions and 140000 kaons inside the LArIAT TPC active volume using the DDMC (see sec \ref{ch:DDMC}).
+We apply  the thin-sliced method on using true quantities to calculate the hadron kinetic energy at each slab in order to decouple reconstruction effects to eventual issues with the methodology.
+For each slab of 4.7 mm length on the path of the hadron, we integrate the true energy deposition as given by the Geant4 transportation model.
+Then, we recursively subtracted it from the hadron kinetic energy at the TPC front face to evaluate the kinetic energy at each slab until the true interaction point is reached.
+Doing so, we obtain the true interacting and incident distributions for the considered hadron and we obtain the true MC cross section as a function of the hadron true kinetic energy.
 
-%Having validated the ``Sliced TPC'' technique and shown that the this technique recovers the simulated cross-section, we now move to performing this measurement utilizing the complete LArIAT simulation within the LArSoft \cite{} based simulation.
-Furthermore, we move from utilizing particle level MC-truth information to performing fully automated reconstruction of the charged hadron events within the LArTPC.
+Figure \ref{fig:TrueMCXS} shows the total hadronic cross section for argon implemented in Geant4 10.01.p3 (solid lines) overlaid with the true MC cross section as obtained with the sliced TPC method (markers) for pions on the left and kaons on the right; the total cross section is shown in green,  the elastic cross section in blue and the inelastic cross section in red.
+The nice agreement with the Geant4 distribution and the cross section  obtained with the sliced TPC method gives us confidence in the  validity of the methodology.
+
+%\begin{comment}     
+\begin{figure}
+%\captionsetup{justification=raggedright}  
+\begin{minipage}[b]{.53\textwidth}  
+  \centering  
+\includegraphics[width=3in]{Chapter-4/Images/PionTrueXS.png}
+\end{minipage}%  
+\begin{minipage}[b]{0.53\textwidth}  
+  \centering  
+\includegraphics[width=3in]{Chapter-4/Images/KaonTrueXS.png}
+\end{minipage}
+\label{fig:TrueMCXS}
+\caption{Hadronic cross sections for $\pi^-$-Ar (left) and K$^+$-Ar (right) implemented in Geant4 10.01.p3 (solid lines) overlaid the true MC cross section as obtained with the sliced TPC method (markers).
+The total cross section is shown in green,  the elastic cross section in blue and the inelastic cross section in red.}
+%\par
+%\begin{minipage}[t]{.53\textwidth}
+%\caption{total hadronic cross section for carbon implemented in Geant4  10.01.p3  with overlaid with the Bugg %and Frideman data.}
+%\label{fig:TrueCarbon}
+%\end{minipage}%
+%\begin{minipage}[t]{.5\textwidth}  
+%\caption{Hadronic cross sections for argon implemented in Geant4 10.01.p3 (solid lines) overlaid the true MC cross section as obtained with the sliced TPC method (markers).
+The total cross section is shown in green,  the elastic cross section in blue and the inelastic cross section in red.}
+%\label{fig:TrueArgon}
+%\end{minipage}  
+\end{figure}
+%\end{comment}

### 2018-03-12 Chapter-2/2-TheLArIATExperiment.tex
s

\chapter{LArIAT: Liquid Argon In A Testbeam}\label{sec:experimentDescription}
\section{LArIAT \& the Intensity Frontier}
\section{The Particles Path to LArIAT}

LArIAT's home at Fermilab is the Fermilab Test Beam Facility (FTBF), where the experiment 
characterizes a beam of charge particles downstream from the Meson Center beam line.

LArIAT's particles history begins in the Fermilab accelerator complex with a beam of protons. 
The process of protons acceleration develops in gradual stages, see picture \ref{fig:Accelerator}: gaseous hydrogen is ionized in order to form H$^{-}$ ions; these ions are boosted to 750 keV by a Cockroft-Walton accelerator and injected to the Linac linear accelerator that increases their energy up to 400 MeV; then, H$^{-}$ ions pass through a carbon foil and lose the two electrons; the resulting protons are then injected into a rapid cycling synchrotron, called Booster; at this stage protons reach 8 GeV of energy and are compacted into bunches; the next stage of acceleration is the Main Injector, a synchrotron which accelerates the bunches up to 120 GeV; in the Main Injector, several bunches are merged into one and used for the injection in the last stage.

The Fermilab accelerator complex works in supercycles of roughly 60 seconds in duration. The beam is then split by electrostatic septa and delivered at different experimental halls all over the lab. A 120~GeV$/c$ primary proton beam with variable intensity is extracted in four-second ``spills" and sent to the Meson Center beam line. Here, this primary beam is focused onto a tungsten target to create LArIAT's secondary beam. The composition of the secondary particle beam is mainly positive pions. For the LArIAT data considered in this work the secondary beam peak momentum was fixed at 64~GeV$/c$, although the beam is tunable in momentum between 8-80\,GeV$/c$; this was deemed a secondary beamline configuration which allowed a stable beam operation of the FTBF.
The beam of pions impinges then on a copper target within a steel collimator inside the LArIAT experimental hall (MC7) to create LArIAT tertiary beam, whose geometry in MC7 has been optimized for LArIAT (shown in  Fig.~\ref{fig:tert-layout}).   The steel collimator selects particles produced with a $13^\circ$ production angle at the target down the beamline.  The particles are then bend by  $~10^\circ$  through a pair of dipole magnets.  Tuning the magnets field intensity results in a range of particle momenta from 0.2 to 1.4~GeV/c.
The tertiary beam composition counts mostly pions and protons with a small fraction of electrons, muons, and kaons present as well. It is the job of the LArIAT beamline detectors to select the particles polarity,  to perform particle identification (beamPID) and to measure the momentum of the tertiary beam particles before they get to the LArTPC. The LArIAT detectors are described in the following paragraphs.

%\begin{comment}     
\begin{figure}
  \centering  	
\includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{Chapter-2/Images/AcceleratorFNAL.png}
\label{fig:Accelerator}
\caption{Layout of Fermilab Acellerator complex.}
\end{figure}

%\begin{comment}     
\begin{figure}
  \centering  	
\includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{Chapter-2/Images/Tertiary.png}
\label{fig:tert-layout}
\caption{Birds eye view of the LArIAT tertiary beamline. In grey: upstream and downstream collimators; in yellow: bending magnets in yellow; in red: wire chambers; in blue: time of flight; in green: liquid argon TPC volume; in maroon: muon range statck.}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{LArIAT Tertiary Beam Instrumentation}\label{sec:Instrumentation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The instrumentation of  LArIAT tertiary beam and the TPC components have changed several times during the three years of LArIAT data taking. The following paragraphs describe the components operational during the data taking period relevant to the hadron cross section measurements (a.k.a. Run II).

The components of the tertiary beamline instrumentation key for the hadron cross section analyses are the target and collimators system, the two bending magnets (in a similar configuration used for the  MINERvA T-977 test beam calibration~\cite{MinervaTestbeam}) a set of four wire chambers (WCs) and two time-of-flight scintillating paddles (TOF) and, of course, the LArTPC.  The magnets determine the polarity of the particles in the tertiary beam; the combination of magnets and wire chambers determine the particles momentum, which is used in conjunction with the TOF to determine the particle species.
A muon range stack downstream from the TPC and two sets of cosmic paddles configured as a telescope surrounding the TPC are also used for calibration purposes.

\section{In the Cryostat}

\subsection{TPC: Charge Collection}\label{sec:TPC}
\subsection{TPC: Light Collection System}
\subsection{Cryogenics and Purity Control}
\subsection{TPC: Electric Field Measurement}
\section{Trigger and DAQ}
\section{Control Systems}

\begin{comment}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Time-of-Flight System}\label{sec:TOF}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The LArIAT time-of-flight (TOF) detector system consists broadly consists of two scintillator paddles, which bracket the beamline and are shown in blue in Figure~\ref{fig:tert-layout}. The exact details of the of the TOF system were changed between Run-II and Run-III to allow for better performance and timing resolution.

\paragraph{\textbf{TOF System for Run-I and Run-II:}}
The upstream paddle, shown in Figure \ref{}, has a relatively small active area (10~cm x 6~cm ) and was chosen to be a relatively thin piece of scintillator to minimize any impact on the momentum of the particles coming from the target. Light guides were mounted on all four edges which lead to two \textit{insert PMT type} PMTs mounted on the beam left side. The downstream paddle, shown in Figure \ref{fig:TOFSystemRunIandII}, was chosen to have a slightly larger area (14~cm x 14~cm) and had two \textit{insert PMT type} PMTs which were read out on opposite ends of the paddle.

\begin{figure}[!h]
\begin{centering}
\vspace{-0.3cm}
%\includegraphics[height=2.3in]{figures/tofdelay.png}
\caption{
{\scriptsize \sf Pictures of the TOF system as was deployed during Run-I and Run-II data taking. The left image is of the upstream TOF paddle and the right image is of the downstream TOF paddle }
}
\label{fig:TOFSystemRunIandII}
\end{centering}
\end{figure}

\paragraph{\textbf{TOF System for Run-III:}}
Two modifications were made to the downstream TOF paddle for Run-III to improve the efficiency and timing of the system. The large area scintillator paddle was replaced with a slightly smaller paddle more similar to the upstream and was now readout with four new PMT's (2" Hamamatsu PMTs (R-1828) inherited from the muon g-2 experiment(E821) at BNL) located on the diagonals of the counter, as shown in Figure \ref{fig:TOFSystemRunIII}.

\begin{figure}[!h]
\begin{centering}
\vspace{-0.3cm}
%\includegraphics[height=2.3in]{figures/TOFRunIII.png}
\caption{
{\scriptsize \sf Image of the modified downstream TOF system deployed during Run-III data taking }
}
\label{fig:TOFSystemRunIII}
\end{centering}
\end{figure}

%The active area of the upstream paddle is relatively small ( 10~cm x 6~cm ) and that of the downstream paddle is somewhat larger ( 14~cmx 14~cm ). Lightguides are mounted on all four edges of each paddle. In the first two LARIAT production runs, each of the paddles was read out by two (xxkxkxkxk) photomultiplier tubes (PMTs), The long axis of the downstream paddle was directed horizontally and read out at either end. The upstream paddle was rotated by 45 degrees with respect to the horizontal and its two PMTs were mounted to the left side. In run 3, four more PMTs were added to the system. The upstream paddle was read out with the original set of PMTs and the downstream paddle with four 2" Hamamatsu PMTs (R-1828) inherited from the muon g-2 experiment(E821) at BNL.

During data-taking cycles, signals from the TOF PMTs were sampled at 1 GHz
with a CAEN 1751 digitizer and 12 bit samples were stored in a circular memory buffer. In response to an experimental trigger, a 28.7 $\mu$ second window of samples, starting approximately 8.4 $\mu$sec before the trigger, was written to the
output. The amplitude of the TOF signals was typically 200 mV in the upstream TOF PMTs but only 50 mV in those of the downstream counter. The signals have a rise time (10-90\%) of 4 ns and a full width, half-maximum of 9 ns. The rate in the upstream counter was typically 15 kHz and much less, 400 Hz in the downstream counter.

Because the shape of signals from each TOF PMT was highly uniform, the time of the pulses was determined using an oversampled template derived from the data itself. The DC offset (pulse pedestal) is taken from samples far from the pulse, The template is stretched vertically to match the pedestal-subtracted pulse amplitude and moved horizontally to match the time. The pulse time-pickoff resolution is better than 100 ps and relative amplitude resolution is better than x\%. Since the average rate of pulses incident on the ToF system is modest, pulse pile up is not a significant problem. Given the uniform width of the pulses produced by any given PMT (sigma of 400 ps), the pulse width can be used to flag events where two pulses overlap closely in time. Simulations indicate that for pulses as close as 4 ns, a well-chosen pulse width cut is 90\% efficient.

To determine the particles' time of arrival, pulse times from the multiple PMTs mounted on each paddle had to be combined. During Run-I and II, with two PMTs mounted on each paddle, a simple average was used. In the case of the downstream paddle, with PMTs mounted at opposite ends, the average corrected for the fact that transit time to the two PMTs depended on the particle's point of impact on the scintillator. A plot of the time of arrival {\it difference} between the two downstream PMTs vs the longitudinal impact point, made with the help of track stubs from the wire chamber system, is shown in Fig.~\ref{fig:tofdelay}. As expected, the time difference
changes systematically across the face of the scintillator.

\begin{figure}[!h]
\begin{centering}
\vspace{-0.3cm}
%\includegraphics[height=2.3in]{figures/tofdelay.png}
\caption{
{\scriptsize \sf $\Delta-t$ between times registered on two downstream paddles.}
}
\label{fig:tofdelay}
\end{centering}
\end{figure}

Taking the average minimizes ToF errors arising from optical path differences in the scintillator, However, even for a set of particles which pass through a single small area of the paddle, the times of pulses registered in the two PMTs are still spread by approximately 300~ps, an uncertainty which is probably caused by transit time jitter in the PMTs themselves. This jitter is evident in both the upstream and downstream detectors.

Although climate-control in the LArIAT experiment hall is fairly primitive, there is little sign of systematic timing drift over longer periods. The average time differences between pairs of PMTs reading out the same scintillator varied by no more than 150~ps over the 3-4 months of a data-taking period.

Finally, the amplitude of the ToF pulses (peak minus pedestal) reflects the energy deposited in the scintillator, largely through ionization.  At a given momentum, more massive particles such deuterons, protons, and kaons, will deposit more energy than less massive particles, such as electrons, muons, and pions. As described below, ToF and momentum information from the wire chambers can be combined to determine the particle's mass and hence its identity. Although the energy deposited in the scintillator isn't used to distinguish among the various particle types, it's clear from Figure~\ref{fig:tofamplitude} that the spectra of energy deposition for
electrons, pions, and protons are very different.

\begin{figure}[!h]
\begin{centering}
\vspace{-0.3cm}
%\includegraphics[height=2.3in]{figures/tofamplitude.png}
\caption{
{\scriptsize \sf Average upstream ToF pulse amplitude for protons and pions/positrons.}
}
\label{fig:tofamplitude}
\end{centering}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Bending Magnets}\label{sec:Magnets}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The pair of spectrometer magnets in the LArIAT tertiary beam are Fermilab type "NDB" electromagnets, originally used in the anti-proton ring. These magnets provide the necessary bend in the particle trajectory to allow for a momentum measurement to be performed.  The magnets have a gap height of 14.224~cm, a gap width of 31.75~cm, and have an iron length of 46.67~cm.  The aperture presented by the magnets is larger than that of the wire chambers ($\sim$12.5cm) and thus only the central part of the magnet is ever utilized.  Over this limited aperture there is negligible variation of the field integral and thus a very uniform field is provided. The field intensity in one of the pair of magnets was measured using two Hall probes, both calibrated with NMR, to obtain the excitation curve shown in Figure~\ref{fig:magnet_excitation}). The second magnet, having been made to the same standard and having the same history, is assumed to have a very similar response.
\begin{figure}[!h]
\begin{centering}
\vspace{-0.3cm}
%\includegraphics[height=3.0in]{figures/ExcitationCurves.png}
\caption{
{\scriptsize \sf Excitation Curve, B/I vs. I, for one NDB magnet, using two Hall probes (blue and green). Parameters from a cubic fit (black curve) to average of measurements (red) given in legend.}
}
\label{fig:magnet_excitation}
\end{centering}
\end{figure}

These magnets were air- and water-cooled during operations and their temperatures monitored. Due to concerns of overheating, the maximum usable field at the center of the magnets is $\sim$0.35 Tesla with 100 Amps of current being passed through the magnets.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Multi-Wire Proportional Chambers}\label{sec:MWPC}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The wire chambers are based on the Fenker Chambers~\cite{Fenker} long in use at Fermilab.  The chambers have been upgraded by adding additional grounding to improve the signal to noise in the electronic readout.  The chambers, shown in Figure~\ref{fig:wirechamber}, have an effective aperture of 128~mm in both horizontal and vertical distances.  The wires are spaced at 1~mm, with 128 wires in each view.  In a test beam, the chambers have typical efficiencies of 98\% to 99\%. The gas used is 85\% Argon + 15\% isobutane.  The wire chambers typically operate between 2400 and 2500 volts.

\begin{figure}[!h]
\begin{centering}
\vspace{-0.3cm}
%\includegraphics[height=2.3in]{figures/WireChamber.png}
\caption{
{\scriptsize \sf Image of one of the wire chambers used in the LArIAT tertiary beamline.}
}
\label{fig:wirechamber}
\end{centering}
\end{figure}

The front end amplifier/discriminator uses the ASDQ chip~\cite{ASDQchip}. Two of the ASDQ chips are mounted on a mother board that is plugged into the chamber.  Short flat cables are used to connect these ASDQ mother board to a new, multi-hit TDC~\cite{Sten}. This TDC provides a (maskable) fast OR output that can be used in a first level trigger.

This TDC can accept multiple hits per wire with a time resolution of 1.18~ns/bin.  The TDCs can accept 2 edges per 9~ns.  The maximum event rate that can be accepted by the chamber system is of the order of one megahertz.  These limitations are clearly not an issue for the LArIAT beamline where the rates and timing resolution required are much less stringent than this.  There is enough memory on the TDC board to buffer a full spill of data (up to about 100 events/spill).  As in this application the spills occur only once per minute, there is ample time to read out the whole spill through a controller specially designed to support these TDCs.  Power and data flow to and from the controller to the TDCs and from the controller to the rest of the DAQ is via LVDS.  The time window for acceptance for hits, time offsets, front end threshold, and pulse shaping parameters may be programmed through the controller via a USB link from a PC or through an Ethernet connection.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Aerogel Cherenkov Detectors}\label{sec:Aerogel}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Aerogel is an ultra-light material in which the liquid component of the gel has been replaced with a gas. The result is a solid with extremely low density, low thermal conductivity, and low index of refraction. The goal of having an aerogel threshold Cherenkov detector in the LArIAT beam line is to separate muons and pions in the momentum range where muons emit Cherenkov radiation while pions do not. This technique was demonstrated by the MICE and Belle II experiments ~\cite{MICE-aerogel,BelleII-aerogel} and in LArIAT we use two aerogel threshold Cherenkov detector with index of refraction of 1.057 and 1.103. Having different indices of aerogel allows LArIAT to do this separation in two different momentum ranges of interest.

\begin{figure}[htb]
\centering
%\includegraphics[height=1.85in]{figures/Aerogel1103Figure.png}
\hspace{1cm}
%\includegraphics[height=1.85in]{figures/Aerogel1057Figure.png}
\caption{Number of photoelectrons per cm versus momentum for the two indices of refraction that are employed by the LArIAT aerogel detectors. The photoelectron yield includes the response of a bialkali photocathode following the recipe in ~\cite{bib8}. }
\end{figure}

%%% KEK couter
The Aerogel Threshold Cherenkov Detector with the refractive index of 1.103 is placed just behind the second collimator on the tertiary beam line, between the two downstream wire chambers.
Figure~\ref{fig:kek_aerogel} is a picture of this aerogel detector.
It consists of 7 aerogel tiles (totaling 110 mm long in the beam direction), and the cross section of the detector is 108 mm $\times$ 108 mm.
The array of the aerogel tiles is wrapped with a diffusion sheet and viewed by two PMTs (HAMAMATSU R329-02) from the sides.
Figure~\ref{fig:kek_aerogel_ly} shows the observed light yield distribution for 550-800~MeV pions hitting the middle of the detector.
The obtained light yield is equivalent to 22 photoelectrons for $\beta=1$.
The fake rate for particles with velocity below the Cherenkov threshold is measured to be around 0.5\% with the threshold of 1.5 photoelectrons.

\begin{figure}[htb]
\centering
%\includegraphics[height=1.85in]{figures/kek_aerogel.jpg}
\caption{The Aerogel Threshold Cherenkov Detector with the refractive index of 1.103.}
\label{fig:kek_aerogel}
\end{figure}

\begin{figure}[h]
\centering
%\includegraphics[height=1.85in]{figures/kek_aerogel_ly.png}
\caption{The observed light yield distribution for 550-800 MeV pions hitting the middle of the detector.
The obtained light yield is equivalent to 22 photoelectrons for $\beta=1$.}
\label{fig:kek_aerogel_ly}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Punch-Through and Muon Range Stack Instruments}\label{sec:MuRS}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{muon-Punch.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{LArIAT Cosmic Ray Paddle Detectors}\label{sec:CosmicRayPaddle}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

LArIAT's system to trigger on cosmic rays is based on two so-called "cosmic towers" which stand  upstream and downstream of the cryostat, one on beam's right and one on beam's left, framing the cryostat.  Each cosmic tower is composed of two paddle assemblies, upper and lower.  The paddle assemblies each consist of four paddles, a matched pair which stand upright and a second matched pair lying across the top  of the assembly, to act as a veto for downward-going cosmic ray air showers.  Unless vetoes by the horizontal paddles, signals from paddle assemblies along the body diagonals of the TPC are combined in a logical ``AND'' to select cosmic muons crossing the TPC along one of its diagonals. A high proportion of events triggered this way contain cosmic ray tracks crossing both anode and cathode.  Such tracks provide a sample of LAr ionization with effectively uniform linear ionization density, but experience the entire range of charge attenuation available in the TPC before they drift to the anode. These tracks are used to calculate and monitor the level of electronegative contaminants in the liquid argon and provide a calibration sample for calorimetry and electric field studies outlined in Section \ref{sec:DetectorPerformance}.

The paddles have trapezoidal shape and are each enclosed in an aluminum case. An example is shown in Fig. \ref{pic:cosmicpaddle}. They come in two different sizes: a smaller version, with bases $32.2~cm$ and $26.7~cm$, $61.0~cm$ height and $3.02~cm$ thickness, and a bigger version, with bases $33.2~cm$ and $27.0~cm$, $70.8~cm$ height and $3.02~cm$ thickness. Each paddle is equipped with wavelength-shifting optical fibers running along one of the long sides and optically coupled to a low voltage, Zener-diode Hamamatsu H5783 PMT.  Signals from the PMT are amplified and discriminated by a custom-made PMT Amplifier and Discrimination (PAD) circuit mounted at one end of the paddle, then sent through a CAT5 cable to a Control and Concentrator Unit (CCU) which both power the PMT, controlling voltage and threshold, and output the PMT signal as logic ECL pulse.

\begin{figure}[h!]
 \centering
% \includegraphics[angle=90,width=0.7\textwidth]{figures/Cosmic_Paddle.jpg}
\caption{
Photograph of one of the scintillation counters used in the cosmic towers. 
} 
\label{pic:cosmicpaddle}
\end{figure}

The paddles were selected from a pool of over 300 scintillating counters collected during the decommissioning of the CDF detector at Fermilab. 
For each counter, both the efficiency $\varepsilon_P$ and the noise $\eta_P$ as a function of the voltage were determined. The measurement was performed sandwiching the given paddle among 4 sample counters, placed two above and two below the paddle under test. The efficiency 
$\varepsilon_P$ was defined as the ratio between the 5-fold coincidence and the 4-fold coincidence of the sample counters. The accidental rate $\eta_P$ was instead defined as the number of 5-fold coincidence observed during ten minutes of data acquisition, when the signal of the paddle under test was delayed by $5 \mu s$. Each paddle with $\varepsilon_P \geq 95\%$ and $\eta_P = 0$ at working voltage was identified as a candidate for the Cosmic Ray system. The ones with the highest efficiency and lowest single count rate were then selected to realize the system itself. 
The plot of $\varepsilon_P$ as a function of the PMT voltage for one example paddle is shown in Fig. \ref{pic:CR_Effplot}.

\begin{figure}[h!]
 \centering
% \includegraphics[width=0.7\textwidth]{figures/TSU29_Efficiency.png}
\caption{
Plot of the counter efficiency $\varepsilon_P$ as a function of the PMT voltage for one of the paddles composing the Cosmic Ray system. 
} 
\label{pic:CR_Effplot}
\end{figure}

%$\varepsilon_P$ at a given voltage is defined as:

%$$
%\varepsilon_P = \frac{R_{1-2} \land R_{3-4} \land R_P}{R_{1-2} \land R_{3-4}}
%$$

%where $R_P$ is the rate of the paddle $P$ to be tested, $R_{1-2}$ is the coincidence rate of two sample paddles positioned above $P$, while $R_{3-4}$ is the coincidence rate of two sample paddles positioned below $P$.
%$\eta_P$ at a given voltage is instead defined as:

%$$
%\eta_P = \frac{R_{1-2} \land R_{3-4} \land R_{P^*}}{R_{1-2} \land R_{3-4}}
%$$

%where $R_{P^*}$ is the Rate of the paddle $P$

The trigger rate $R$ of the whole system is $R=0.032Hz$, corresponding to $\sim 1.9~\mu/minute$.
\end{comment}

### 2018-03-12 Chapter-4/3-KaonInteractionsInLAr.tex
s

@@
 \section{How to Measure Hadron Cross Section in LArIAT}\label{ch:methodology}
+We use both the LArIAT  beamline detectors and the LArTPC information in order to measure hadronic cross sections in argon.
+Albeit with small differences, both the  $\pi^{-}$-Ar and K$^{+}$-Ar total hadronic cross section measurements rely on the same procedure described in details in the following paragraphs: we select the particle of interest using a combination of beamline detectors and TPC information (paragraph \ref{ch:ParticleSelectionMethod}), we perform a handshake between the beamline information and the TPC tracking to assure we are selecting the right TPC track (paragraph \ref{ch:WC2TPCMatchMethod}), and we apply the ``thin slice" method to get to the final result (paragraph \ref{ch:ThinSliceMethod}).
 \subsection{Particle Selection}\label{ch:ParticleSelectionMethod}
-\subsection{Beamline to TPC Handshake}\label{ch:WC2TPCMatchMethod}
-\subsection{The Thin Slice Method}
+\subsection{Wire Chamber to TPC Match}\label{ch:WC2TPCMatchMethod}
+After an event passes the selection described in the previous paragraph, we need to identify the track inside the TPC corresponding to the particle which triggered the beamline detectors, a procedure we refer to as ``WC to TPC match".
+In general, the tracking algorithm will reconstruct more than one track in the event, partially due to the fact that hadrons interact in the chamber, as shown in figure , and partially because of pile up events where the beam of charge particle is too intense compared to the drift time, as shown in figure.
+\textcolor{red}{EVENT DISPLAYS}
+
+\subsection{The Thin Slice Method}\label{ch:ThinSliceMethod}
 \subsubsection{Cross Sections on Thin Target}

### 2018-03-12 Drafting.tex
s

\section{Non-LArTPC Instruments}\label{sec:FTBF}
The Fermilab Test Beam Facility (FTBF) receives a 120~GeV$/c$ primary proton beam with variable intensity, split by electrostatic septa from beam circulating in the Main Injector. Each supercycle of the Fermilab accelerator complex, roughly 60 seconds in duration, includes a four-second "spill" of primary beam to FTBF. Primary beam is sent to either or both of two experimental halls, named Meson Test and Meson Center.

The LArIAT experiment utilized the Meson Center beam line and had this primary beam focused onto a tungsten target. The resulting secondary charged particle beam is selected to have peak momentum between 8-80\,GeV$/c$ and is transported to the LArIAT experimental hall, designated MC7, where it is focused onto a copper target. This target, and the steel collimator hosting it, form the beginning of the tertiary particle beam and the LArIAT apparatus shown in  Fig.~\ref{fig:tert-layout}.  For most of the LArIAT data taking this secondary beam peak momentum was fixed at 64~GeV$/c$, although some lower secondary beam momentum settings were explored.  With such an energetic secondary beam on it, the copper target produces a tertiary beam which is comprised of mostly pions and protons with a small fraction of electrons, muons, and kaons present as well.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{LArIAT Tertiary Beam}\label{sec:TertiaryBeam}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The tertiary beam consists of a target and collimator system and two bending magnets in a similar configuration used for the  MINERvA T-977 test beam calibration~\cite{MinervaTestbeam}. A set of four wire chambers and two time-of-flight scintillating paddles, also shown in Fig.~\ref{fig:tert-layout}, provide tracking, momentum determination, and particle identification (PID).

The geometry of the tertiary beam line in MC7 has been optimized for LArIAT.  A $13^\circ$ production angle at the target and its enclosing collimator, as well as a $10^\circ$ bend through a pair of dipole magnets provide particle momenta spectra tunable in the range from 0.2 to 2.0~GeV/c, as a function of the field intensity in the magnets.

\begin{figure}[!h]
\begin{centering}
\vspace{-0.3cm}
\includegraphics[height=5.3in]{figures/LArIATBeamline.png}
\caption{
{\scriptsize \sf The tertiary beam line layout. Upstream and downstream collimators are in gray, bending magnets in yellow, wire chambers are in red, and the liquid argon TPC volume in green.}
}
\label{fig:tert-layout}
\end{centering}
\end{figure}

In the subsequent sections, we will detail the various subsystems installed along the beamline

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Time-of-Flight System}\label{sec:TOF}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The LArIAT time-of-flight (TOF) detector system consists broadly consists of two scintillator paddles, which bracket the beamline and are shown in blue in Figure~\ref{fig:tert-layout}. The exact details of the of the TOF system were changed between Run-II and Run-III to allow for better performance and timing resolution.

\paragraph{\textbf{TOF System for Run-I and Run-II:}}
The upstream paddle, shown in Figure \ref{}, has a relatively small active area (10~cm x 6~cm ) and was chosen to be a relatively thin piece of scintillator to minimize any impact on the momentum of the particles coming from the target. Light guides were mounted on all four edges which lead to two \textit{insert PMT type} PMTs mounted on the beam left side. The downstream paddle, shown in Figure \ref{fig:TOFSystemRunIandII}, was chosen to have a slightly larger area (14~cm x 14~cm) and had two \textit{insert PMT type} PMTs which were read out on opposite ends of the paddle.

\begin{figure}[!h]
\begin{centering}
\vspace{-0.3cm}
%\includegraphics[height=2.3in]{figures/tofdelay.png}
\caption{
{\scriptsize \sf Pictures of the TOF system as was deployed during Run-I and Run-II data taking. The left image is of the upstream TOF paddle and the right image is of the downstream TOF paddle }
}
\label{fig:TOFSystemRunIandII}
\end{centering}
\end{figure}

\paragraph{\textbf{TOF System for Run-III:}}
Two modifications were made to the downstream TOF paddle for Run-III to improve the efficiency and timing of the system. The large area scintillator paddle was replaced with a slightly smaller paddle more similar to the upstream and was now readout with four new PMT's (2" Hamamatsu PMTs (R-1828) inherited from the muon g-2 experiment(E821) at BNL) located on the diagonals of the counter, as shown in Figure \ref{fig:TOFSystemRunIII}.

\begin{figure}[!h]
\begin{centering}
\vspace{-0.3cm}
\includegraphics[height=2.3in]{figures/TOFRunIII.png}
\caption{
{\scriptsize \sf Image of the modified downstream TOF system deployed during Run-III data taking }
}
\label{fig:TOFSystemRunIII}
\end{centering}
\end{figure}

%The active area of the upstream paddle is relatively small ( 10~cm x 6~cm ) and that of the downstream paddle is somewhat larger ( 14~cmx 14~cm ). Lightguides are mounted on all four edges of each paddle. In the first two LARIAT production runs, each of the paddles was read out by two (xxkxkxkxk) photomultiplier tubes (PMTs), The long axis of the downstream paddle was directed horizontally and read out at either end. The upstream paddle was rotated by 45 degrees with respect to the horizontal and its two PMTs were mounted to the left side. In run 3, four more PMTs were added to the system. The upstream paddle was read out with the original set of PMTs and the downstream paddle with four 2" Hamamatsu PMTs (R-1828) inherited from the muon g-2 experiment(E821) at BNL.

During data-taking cycles, signals from the TOF PMTs were sampled at 1 GHz
with a CAEN 1751 digitizer and 12 bit samples were stored in a circular memory buffer. In response to an experimental trigger, a 28.7 $\mu$ second window of samples, starting approximately 8.4 $\mu$sec before the trigger, was written to the
output. The amplitude of the TOF signals was typically 200 mV in the upstream TOF PMTs but only 50 mV in those of the downstream counter. The signals have a rise time (10-90\%) of 4 ns and a full width, half-maximum of 9 ns. The rate in the upstream counter was typically 15 kHz and much less, 400 Hz in the downstream counter.

Because the shape of signals from each TOF PMT was highly uniform, the time of the pulses was determined using an oversampled template derived from the data itself. The DC offset (pulse pedestal) is taken from samples far from the pulse, The template is stretched vertically to match the pedestal-subtracted pulse amplitude and moved horizontally to match the time. The pulse time-pickoff resolution is better than 100 ps and relative amplitude resolution is better than x\%. Since the average rate of pulses incident on the ToF system is modest, pulse pile up is not a significant problem. Given the uniform width of the pulses produced by any given PMT (sigma of 400 ps), the pulse width can be used to flag events where two pulses overlap closely in time. Simulations indicate that for pulses as close as 4 ns, a well-chosen pulse width cut is 90\% efficient.

To determine the particles' time of arrival, pulse times from the multiple PMTs mounted on each paddle had to be combined. During Run-I and II, with two PMTs mounted on each paddle, a simple average was used. In the case of the downstream paddle, with PMTs mounted at opposite ends, the average corrected for the fact that transit time to the two PMTs depended on the particle's point of impact on the scintillator. A plot of the time of arrival {\it difference} between the two downstream PMTs vs the longitudinal impact point, made with the help of track stubs from the wire chamber system, is shown in Fig.~\ref{fig:tofdelay}. As expected, the time difference
changes systematically across the face of the scintillator.

\begin{figure}[!h]
\begin{centering}
\vspace{-0.3cm}
%\includegraphics[height=2.3in]{figures/tofdelay.png}
\caption{
{\scriptsize \sf $\Delta-t$ between times registered on two downstream paddles.}
}
\label{fig:tofdelay}
\end{centering}
\end{figure}

Taking the average minimizes ToF errors arising from optical path differences in the scintillator, However, even for a set of particles which pass through a single small area of the paddle, the times of pulses registered in the two PMTs are still spread by approximately 300~ps, an uncertainty which is probably caused by transit time jitter in the PMTs themselves. This jitter is evident in both the upstream and downstream detectors.

Although climate-control in the LArIAT experiment hall is fairly primitive, there is little sign of systematic timing drift over longer periods. The average time differences between pairs of PMTs reading out the same scintillator varied by no more than 150~ps over the 3-4 months of a data-taking period.

Finally, the amplitude of the ToF pulses (peak minus pedestal) reflects the energy deposited in the scintillator, largely through ionization.  At a given momentum, more massive particles such deuterons, protons, and kaons, will deposit more energy than less massive particles, such as electrons, muons, and pions. As described below, ToF and momentum information from the wire chambers can be combined to determine the particle's mass and hence its identity. Although the energy deposited in the scintillator isn't used to distinguish among the various particle types, it's clear from Figure~\ref{fig:tofamplitude} that the spectra of energy deposition for
electrons, pions, and protons are very different.

\begin{figure}[!h]
\begin{centering}
\vspace{-0.3cm}
%\includegraphics[height=2.3in]{figures/tofamplitude.png}
\caption{
{\scriptsize \sf Average upstream ToF pulse amplitude for protons and pions/positrons.}
}
\label{fig:tofamplitude}
\end{centering}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Bending Magnets}\label{sec:Magnets}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The pair of spectrometer magnets in the LArIAT tertiary beam are Fermilab type "NDB" electromagnets, originally used in the anti-proton ring. These magnets provide the necessary bend in the particle trajectory to allow for a momentum measurement to be performed.  The magnets have a gap height of 14.224~cm, a gap width of 31.75~cm, and have an iron length of 46.67~cm.  The aperture presented by the magnets is larger than that of the wire chambers ($\sim$12.5cm) and thus only the central part of the magnet is ever utilized.  Over this limited aperture there is negligible variation of the field integral and thus a very uniform field is provided. The field intensity in one of the pair of magnets was measured using two Hall probes, both calibrated with NMR, to obtain the excitation curve shown in Figure~\ref{fig:magnet_excitation}). The second magnet, having been made to the same standard and having the same history, is assumed to have a very similar response.
\begin{figure}[!h]
\begin{centering}
\vspace{-0.3cm}
\includegraphics[height=3.0in]{figures/ExcitationCurves.png}
\caption{
{\scriptsize \sf Excitation Curve, B/I vs. I, for one NDB magnet, using two Hall probes (blue and green). Parameters from a cubic fit (black curve) to average of measurements (red) given in legend.}
}
\label{fig:magnet_excitation}
\end{centering}
\end{figure}

These magnets were air- and water-cooled during operations and their temperatures monitored. Due to concerns of overheating, the maximum usable field at the center of the magnets is $\sim$0.35 Tesla with 100 Amps of current being passed through the magnets.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Multi-Wire Proportional Chambers}\label{sec:MWPC}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The wire chambers are based on the Fenker Chambers~\cite{Fenker} long in use at Fermilab.  The chambers have been upgraded by adding additional grounding to improve the signal to noise in the electronic readout.  The chambers, shown in Figure~\ref{fig:wirechamber}, have an effective aperture of 128~mm in both horizontal and vertical distances.  The wires are spaced at 1~mm, with 128 wires in each view.  In a test beam, the chambers have typical efficiencies of 98\% to 99\%. The gas used is 85\% Argon + 15\% isobutane.  The wire chambers typically operate between 2400 and 2500 volts.

\begin{figure}[!h]
\begin{centering}
\vspace{-0.3cm}
\includegraphics[height=2.3in]{figures/WireChamber.png}
\caption{
{\scriptsize \sf Image of one of the wire chambers used in the LArIAT tertiary beamline.}
}
\label{fig:wirechamber}
\end{centering}
\end{figure}

The front end amplifier/discriminator uses the ASDQ chip~\cite{ASDQchip}. Two of the ASDQ chips are mounted on a mother board that is plugged into the chamber.  Short flat cables are used to connect these ASDQ mother board to a new, multi-hit TDC~\cite{Sten}. This TDC provides a (maskable) fast OR output that can be used in a first level trigger.

This TDC can accept multiple hits per wire with a time resolution of 1.18~ns/bin.  The TDCs can accept 2 edges per 9~ns.  The maximum event rate that can be accepted by the chamber system is of the order of one megahertz.  These limitations are clearly not an issue for the LArIAT beamline where the rates and timing resolution required are much less stringent than this.  There is enough memory on the TDC board to buffer a full spill of data (up to about 100 events/spill).  As in this application the spills occur only once per minute, there is ample time to read out the whole spill through a controller specially designed to support these TDCs.  Power and data flow to and from the controller to the TDCs and from the controller to the rest of the DAQ is via LVDS.  The time window for acceptance for hits, time offsets, front end threshold, and pulse shaping parameters may be programmed through the controller via a USB link from a PC or through an Ethernet connection.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Aerogel Cherenkov Detectors}\label{sec:Aerogel}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Aerogel is an ultra-light material in which the liquid component of the gel has been replaced with a gas. The result is a solid with extremely low density, low thermal conductivity, and low index of refraction. The goal of having an aerogel threshold Cherenkov detector in the LArIAT beam line is to separate muons and pions in the momentum range where muons emit Cherenkov radiation while pions do not. This technique was demonstrated by the MICE and Belle II experiments ~\cite{MICE-aerogel,BelleII-aerogel} and in LArIAT we use two aerogel threshold Cherenkov detector with index of refraction of 1.057 and 1.103. Having different indices of aerogel allows LArIAT to do this separation in two different momentum ranges of interest.

\begin{figure}[htb]
\centering
\includegraphics[height=1.85in]{figures/Aerogel1103Figure.png}
\hspace{1cm}
\includegraphics[height=1.85in]{figures/Aerogel1057Figure.png}
\caption{Number of photoelectrons per cm versus momentum for the two indices of refraction that are employed by the LArIAT aerogel detectors. The photoelectron yield includes the response of a bialkali photocathode following the recipe in ~\cite{bib8}. }
\end{figure}

%%% KEK couter
The Aerogel Threshold Cherenkov Detector with the refractive index of 1.103 is placed just behind the second collimator on the tertiary beam line, between the two downstream wire chambers.
Figure~\ref{fig:kek_aerogel} is a picture of this aerogel detector.
It consists of 7 aerogel tiles (totaling 110 mm long in the beam direction), and the cross section of the detector is 108 mm $\times$ 108 mm.
The array of the aerogel tiles is wrapped with a diffusion sheet and viewed by two PMTs (HAMAMATSU R329-02) from the sides.
Figure~\ref{fig:kek_aerogel_ly} shows the observed light yield distribution for 550-800~MeV pions hitting the middle of the detector.
The obtained light yield is equivalent to 22 photoelectrons for $\beta=1$.
The fake rate for particles with velocity below the Cherenkov threshold is measured to be around 0.5\% with the threshold of 1.5 photoelectrons.

\begin{figure}[htb]
\centering
\includegraphics[height=1.85in]{figures/kek_aerogel.jpg}
\caption{The Aerogel Threshold Cherenkov Detector with the refractive index of 1.103.}
\label{fig:kek_aerogel}
\end{figure}

\begin{figure}[h]
\centering
\includegraphics[height=1.85in]{figures/kek_aerogel_ly.png}
\caption{The observed light yield distribution for 550-800 MeV pions hitting the middle of the detector.
The obtained light yield is equivalent to 22 photoelectrons for $\beta=1$.}
\label{fig:kek_aerogel_ly}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Punch-Through and Muon Range Stack Instruments}\label{sec:MuRS}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{muon-Punch.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{LArIAT Cosmic Ray Paddle Detectors}\label{sec:CosmicRayPaddle}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

LArIAT's system to trigger on cosmic rays is based on two so-called "cosmic towers" which stand  upstream and downstream of the cryostat, one on beam's right and one on beam's left, framing the cryostat.  Each cosmic tower is composed of two paddle assemblies, upper and lower.  The paddle assemblies each consist of four paddles, a matched pair which stand upright and a second matched pair lying across the top  of the assembly, to act as a veto for downward-going cosmic ray air showers.  Unless vetoes by the horizontal paddles, signals from paddle assemblies along the body diagonals of the TPC are combined in a logical ``AND'' to select cosmic muons crossing the TPC along one of its diagonals. A high proportion of events triggered this way contain cosmic ray tracks crossing both anode and cathode.  Such tracks provide a sample of LAr ionization with effectively uniform linear ionization density, but experience the entire range of charge attenuation available in the TPC before they drift to the anode. These tracks are used to calculate and monitor the level of electronegative contaminants in the liquid argon and provide a calibration sample for calorimetry and electric field studies outlined in Section \ref{sec:DetectorPerformance}.

The paddles have trapezoidal shape and are each enclosed in an aluminum case. An example is shown in Fig. \ref{pic:cosmicpaddle}. They come in two different sizes: a smaller version, with bases $32.2~cm$ and $26.7~cm$, $61.0~cm$ height and $3.02~cm$ thickness, and a bigger version, with bases $33.2~cm$ and $27.0~cm$, $70.8~cm$ height and $3.02~cm$ thickness. Each paddle is equipped with wavelength-shifting optical fibers running along one of the long sides and optically coupled to a low voltage, Zener-diode Hamamatsu H5783 PMT.  Signals from the PMT are amplified and discriminated by a custom-made PMT Amplifier and Discrimination (PAD) circuit mounted at one end of the paddle, then sent through a CAT5 cable to a Control and Concentrator Unit (CCU) which both power the PMT, controlling voltage and threshold, and output the PMT signal as logic ECL pulse.

\begin{figure}[h!]
 \centering
 \includegraphics[angle=90,width=0.7\textwidth]{figures/Cosmic_Paddle.jpg}
\caption{
Photograph of one of the scintillation counters used in the cosmic towers. 
} 
\label{pic:cosmicpaddle}
\end{figure}

The paddles were selected from a pool of over 300 scintillating counters collected during the decommissioning of the CDF detector at Fermilab. 
For each counter, both the efficiency $\varepsilon_P$ and the noise $\eta_P$ as a function of the voltage were determined. The measurement was performed sandwiching the given paddle among 4 sample counters, placed two above and two below the paddle under test. The efficiency 
$\varepsilon_P$ was defined as the ratio between the 5-fold coincidence and the 4-fold coincidence of the sample counters. The accidental rate $\eta_P$ was instead defined as the number of 5-fold coincidence observed during ten minutes of data acquisition, when the signal of the paddle under test was delayed by $5 \mu s$. Each paddle with $\varepsilon_P \geq 95\%$ and $\eta_P = 0$ at working voltage was identified as a candidate for the Cosmic Ray system. The ones with the highest efficiency and lowest single count rate were then selected to realize the system itself. 
The plot of $\varepsilon_P$ as a function of the PMT voltage for one example paddle is shown in Fig. \ref{pic:CR_Effplot}.

\begin{figure}[h!]
 \centering
 \includegraphics[width=0.7\textwidth]{figures/TSU29_Efficiency.png}
\caption{
Plot of the counter efficiency $\varepsilon_P$ as a function of the PMT voltage for one of the paddles composing the Cosmic Ray system. 
} 
\label{pic:CR_Effplot}
\end{figure}

%$\varepsilon_P$ at a given voltage is defined as:

%$$
%\varepsilon_P = \frac{R_{1-2} \land R_{3-4} \land R_P}{R_{1-2} \land R_{3-4}}
%$$

%where $R_P$ is the rate of the paddle $P$ to be tested, $R_{1-2}$ is the coincidence rate of two sample paddles positioned above $P$, while $R_{3-4}$ is the coincidence rate of two sample paddles positioned below $P$.
%$\eta_P$ at a given voltage is instead defined as:

%$$
%\eta_P = \frac{R_{1-2} \land R_{3-4} \land R_{P^*}}{R_{1-2} \land R_{3-4}}
%$$

%where $R_{P^*}$ is the Rate of the paddle $P$

The trigger rate $R$ of the whole system is $R=0.032Hz$, corresponding to $\sim 1.9~\mu/minute$.

\section{The LArIAT Detector}\label{sec:LArIATDetector}

An Intro paragraph, as Will F. suggested, would be good here at the start of section 3. I might also rename this section to ?The LArIAT LArTPC Detector?, since beam line stuff is also ?The LArIAT Detector?.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Cryogenics}\label{sec:Cryo}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{CryoDetails.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Time Projection Chamber}\label{sec:TPC}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{TPCDetails.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{TPC Electronics}\label{sec:MSUElectronics}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{TPCelectronics.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Digitization}\label{sec:Digitization}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{Digitization.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Trigger and Readout System}\label{sec:Trigger}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{Trigger.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Photon Detection System}\label{sec:PhotonSystem}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{LightDetectionSystem.tex}

The LArIAT cryogenic system consists of the repurposed ArgoNeuT cryostat, the new process piping, and a liquid argon filtration system installed at Fermilab's Test Beam Facility.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{subsubsection}{Cryostat}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The LArIAT cryostat, shown in Figure \ref{fig:LArIATCryoStat}, was repurposed from the ArgoNeuT experiment with a series of modifications to allow for operations in a charged particle test beam. The cryostat \cite{ArgoNeuT} consists of an inner volume containing the purified liquid argon and an outer volume serving as a vacuum jacket with layers of aluminized mylar superinsulation. Both the inner and outer cryostat vessels are cylindrical in shape with convex end caps.  Access to the internal volume is possible by opening the upstream end caps of the inner and outer vessels. The main axis of the cryostat is horizontal and oriented parallel to the beam. The inner vessel is 76.2~cm in diameter and 130~cm in length, corresponding to a liquid argon volume of about 550~L, or a mass of 0.76 t. The cryostat has a wide neck, or ``chimney'', protruding from its top at mid-length which serves as an access path for signal cables from the LArTPC and the internal instrumentation, as well as for the high voltage feedthrough.

\begin{figure}[htb]
\centering
\includegraphics[scale=0.18]{./figures/Cryostat1.jpg}
\includegraphics[scale=0.07]{./figures/Cryostat2.jpg}
\caption{\emph{(left)} The LArIAT cryostat open with the TPC placed in the inner volume. \emph{(right)} The LArIAT cryostat fully sealed during initial commissioning prior to installation at Fermilab Testbeam Facility.}
\label{fig:LArIATCryoStat}
\end{figure}

Several modifications to the cryostat were made to meet the requirements of operating in a charged particle test beam and to facilitate the new liquid argon purification system. These modifications, shown in Figure \ref{fig:LArIATCryoMods}, include:
\begin{enumerate}[label=\roman*),topsep=0pt,itemsep=-1ex,partopsep=1ex,parsep=1ex]
\item A ``beam window'' on the front outer end cap and an ``excluder'' on the inner endcap.  The purpose of these modifications is to minimize the amount of material upstream of the TPC's active volume along the beam line.  On the original ArgoNeuT cryostat the stainless steel end caps of the cryostat's inner and outer vessels were 3/16'' (4.8 mm) thick. In addition, upstream of the TPC active volume there was a region of uninstrumented liquid argon approximately 15~cm in length. This corresponded to $\sim$ 1.6 radiation lengths ($X_{0}$) of dead material between the outside of the cryostat and the TPC, which would have impeded charged test beam particles prior to their entrance into the active volume.  To mitigate this effect, the 9'' (22.9~cm) diameter beam window located at the center of the outer vessel's front flange was installed along with a blanked 0.4~mm titanium sheet. On the inner vessel's end cap, a hollow concave volume (the ``excluder'') was installed to reduce the amount of liquid argon charged particles would traverse before reaching the active volume of the TPC. With these modificaitons in place, the total material thickness seen by charged particles entering the front face of the LArIAT cryostat is now less than 0.3 $X_{0}$. 
\item A connection for a liquid argon transfer line to the new argon cooling and purification system. This modification consists of an outlet at the bottom of the cryostat with Conflat and ISO flange sealing. 
\item The addition of side port feedthroughs to accomodate signal and high-voltage bias connections for the scintillation light detection system (described in more detail in Section \ref{sec:PhotonSystem}). Both the inner and outer cryostat vessels had their existing side ports modified with CF-flanged apertures to accomplish this.
\end{enumerate}

%$i)$ A ``beam window'' on the front outer end cap and an ``excluder'' on the inner endcap to minimize the amount of material upstream of the TPC's active volume along the beam line; $ii)$ a connection at the base of the cryostat for the argon cooling and purification system; and $iii)$ the addition of a side port feedthrough to accomodate bias and signal connections for the scintillation light detectors.

%On the original ArgoNeuT cryostat the stainless steel end caps of the cryostat's inner and outer vessels were 3/16'' (4.8 mm) thick. In addition, upstream of the TPC active volume there was a region of uninstrumented liquid argon approximately 15~cm in length. This corresponded to $\sim$ 1.6 radiation lengths ($X_{0}$) of dead material between the outside of the cryostat and the TPC, which would have impeded charged test beam particles prior to their entrance into the active volume.  To minimize the effect of this dead region, a 9'' (22.9~cm) diameter beam window located at the center of the outer vessel's front flange was installed along with a blanked 0.4~mm titanium sheet. On the inner vessel's end cap, a hollow concave volume (the ``excluder'') was installed to reduce the amount of liquid argon charged particles would traverse before reaching the active volume of the TPC. With these modificaitons in place, the total material thickness seen by charged particles entering the front face of the LArIAT cryostat is now less than 0.3 $X_{0}$.

%The cryostat vessels were modified to connect a liquid argon transfer line to the new cooling and purification system. This modification consists of an outlet at the bottom of the cryostat with Conflat and ISO flange sealing.

%Finally, both the inner and outer cryostat vessels had their existing side ports modified with CF-flanged apertures for the signal and high-voltage bias feedthroughs for the scintillation light system, described in more detail in Section \ref{sec:PhotonSystem}.

\begin{figure}[htb]
\centering
\includegraphics[scale=0.35]{./figures/CryoMods.png}
\caption{Pictures of the modified components of the cryostat. $1)$ The addition of an outlet to the bottom of the cryostat to allow connections to the purification system; $2)$ The ``beam-window'' on the outer endcap and the concave inner surface of the inner endcap (referred to as the excluder) to reduce the amount of material through which beam particles must travel before entering the TPC; $3)$ The modified side port for the LArIAT light collection system.}
\label{fig:LArIATCryoMods}
\end{figure}

\end{subsubsection}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{subsubsection}{Liquid Argon System}\label{sec:LArCryoSystem}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
LArIAT is supplied with liquid argon from a commercial dewar positioned outside of the experimental enclosure.  Liquid argon is delivered by a vendor with a specified maximum contamination of 2 parts per million (ppm) oxygen, 3.5~ppm water, and 10~ppm nitrogen. Oxygen and water contamination is detrimental to the ability of the ionization electrons to drift through the volume, as free electrons readily attach to these molecules. Nitrogen contamination is detrimental to the collection of scintillation light.  The level of these contaminants in the delivered argon is determined using a suite of commercial gas analyzers.  In practice, the delivered argon has much less than 1~ppm of all these contaminants.  Even such apparently low levels of contamination are too large to effectively drift ionization electrons, so filtration of the argon is required to achieve oxygen and water contamination at the level of 100 parts per trillion (ppt) or better. %{\bf 100 ppt is sufficient for a 3 ms lifetime, I think LArIAT has less than that}.

The argon is delivered from the commercial dewar to the cryostat through 2.54~cm diameter schedule 10 stainless steel piping.  The piping was insulated with 20.32~cm of polyurethane foam by the manufacturer.  The piping was cleaned to remove oil and grease before being welded into the system. The argon then passes through LArIAT's argon filtration system which is based on the Liquid Argon Purity Demonstrator (LAPD)~\cite{LAPD} design. The purification system consists of a single 77~liter filter which is filled halfway with a 4A molecular sieve supplied by Sigma-Aldrich~\cite{sigma-aldrich} that primarily removes water contamination but can also remove small amounts of nitrogen and oxygen. The remaining volume of the filter contains BASF~CU-0226~S, a highly dispersed copper oxide impregnated on a high surface area alumina, to remove oxygen~\cite{basf} and to a lesser extent, water.  The filter is insulated with a vacuum jacket and aluminum radiation shields.  The filter media are regenerated in place using heated gas, as was done for LAPD. The filter media are very efficient at removing oxygen and water and the argon is pure enough after a single pass through the media to allow several millisecond electron drift lifetimes in the TPC.

From the filter the argon is directed into the inner cryostat via a liquid feedthrough on the top of the cryostat. The argon level, temperatures, and pressures are continuously monitored during operation both in the commercial dewar supplying the argon as well as the levels inside the cryostat, as shown on the top of Figure \ref{fig:LArIATCryoMonitor}. The argon in the cryostat is allowed to boil and vent to the atmosphere during operation and the argon level is monitored via the level probes. Flow of argon into the cryostat is enabled whenever the liquid level goes below a predetermined value to ensure that the TPC high voltage feedthrough and cold electronics are always submerged. During normal operations, the liquid level inside the cryostat is replenished several times per day, as can be seen by the cycle of the liquid level and valve positions shown at the bottom of Figure \ref{fig:LArIATCryoMonitor}

\begin{figure}[htb]
\centering
\includegraphics[scale=0.35]{./figures/CryoMonitor.png}
\includegraphics[scale=0.34]{./figures/CryoMonitor2.png}
\caption{\emph{(top)} Screenshot of the LArIAT cyrosystem monitoring page showing the levels of the argon both inside the cryostat and in the supply dewar as well as the monitored levels plotted over a twenty-four hour period. \emph{(bottom)} Plots showing the liquid argon level inside the cryostat as well as the corresponding liquid valve which allows argon to flow into the cryostat and the level drops due to boiling. The frequency of the typical fill/vent cycle can be seen with the typical time between fills being slightly more than 3 hours.}
\label{fig:LArIATCryoMonitor}
\end{figure}

The temperature of the argon inside the cryostat can be determined through direct use of temperature probes deployed at the bottom, middle, and top of the cryostat (listed as TE213A, TE314A, and TE212A in Figure \ref{fig:LArIATCryoMonitor} respectively). This can be crossed check using the pressure in the cryostat gas volume and extrapoloting this to the center of the liquid volume. The pressure at the surface of the liquid argon is maintained by the cryosystem at $\sim$20~psi $\pm$.0.4 psi.

In order to determine the temperature inside the TPC, we measure the pressure in the cryostat gas volume and calculate the pressure at the center of the TPC. From this pressure and the boiling point curve of argon, we calculate the temperature. The pressure at the surface of the liquid argon is maintained by the cryosystem at $\sim$20~psi as shown in Figure \ref{fig:pressure} for a period of 13 days with an average pressure of 20.0 $\pm$.0.4 psi. If we consider approximately 40 cm of argon from the ulage surface to the center of the TPC this would give the pressure there of $\sim$20.8 psi, which according to the Reference \cite{} corresponds to a temperature of 90.7 K. This is consistent with the readings given of the temperature probes, shown in Table \ref{tab:temp}. This cross-check both provides a basis for the electron mobility calculation used when calibrating the drift speed, as is discussed in Section \ref{}, as well as providing confidence in the monitoring tools.

\begin{table}[h!]
\centering
\caption{Average temperatures measured at the top, middle and bottom in the LArIAT cryostat.}
\label{tab:temp}
\begin{tabular}{|c|c|c|}
\hline
Top Probe Temp (TE212A) & Middle Probe Temp (TE214A)   & Bottom Probe Temp (TE213A)  \\ \hline
91.7 $\pm$ 0.7 K &  91.8 $\pm$ 0.7 K                   & 93.3 $\pm$ 2.7 K       \\ \hline
\end{tabular}
\end{table}

\end{subsubsection}
The Liquid Argon Time Projection Chamber (LArTPC) can be broken down into three major subcomponents: $1)$ The high voltage system which provides the drift field voltage; $2)$ the cathode and field cage which steps down the high voltage through a network of voltage-dividing resistors to form a uniform electric field for charges to drift within; and $3)$ the wire planes which provide the charge sensitive readout for the detector. Here we describe each of the subcomponents which make up the LArIAT LArTPC.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{subsubsection}{High Voltage}\label{sec:DriftVoltage}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The drift high voltage system for the LArIAT detector, shown pictorially in Figure~\ref{fig:HVScheme}, is designed to allow ionization electrons from the interaction of charged particles in the liquid argon to drift to the wireplanes. The high voltage system consists of a Glassman LX125N16 power supply~\cite{GlassmanPS} capable of generating -125~kV and 16~mA of current. This power supply was typically operated much below this limit during normal data taking operations. The voltage from the power supply is transmitted through high voltage cables to a series of filter pots before finally reaching the high-voltage feedthrough on the top of the LArIAT cryostat. This feedthrough brings the voltage into the liquid argon volume to be transmitted to the TPC cathode.

\begin{figure}[htb]
\centering
\includegraphics[scale=0.35]{./figures/HVSchematic.png}\\
\caption{Schematic of the LArIAT high voltage system.}
\label{fig:HVScheme}
\end{figure}%See DocDB 1472

The first high voltage cable which connects the Glassman power supply to the first filter pot is a 1/2-inch HV cable, Glassman cable model 2121, supplied by the Glassman manufacturer. The second cable is a $\sim$100-ft long 1/2-inch Type-2134 Dielectric Sciences cable which connects the first filter pot to the second filter pot. Finally, the same type of cable is used to connect the second filter pot to the high voltage feedthrough at the cryostat.

The two filter pots that are inline with the high voltage system serve to: $a)$ limit the current draw on the power supply; $b)$  provide a low-pass filter to help reduce the voltage ripple on the cathode; and $c)$ partition the energy stored in the system in the event of a high voltage trip. Each filter pot, as shown in Figure~\ref{fig:FilterPot}, is a cylindrical aluminum pot 18.5 inches tall by 20 inches in diameter and 3/16 inches thick, with welded tops each having an opening that allows for a flange and o-ring to receive the high voltage cables. Within each pot, the flanges are connected to four 10~M$\Omega$ resistors connected in series. This assembly is submerged in $\sim$16 gallons of Diala transformer oil~\cite{ShellDiala} to aid in the suppression of corona and discharges.

\begin{figure}[htb]
\centering
\includegraphics[scale=0.35]{./figures/FilterPot.png}\\
\caption{Filter pots used in the LArIAT high voltage system.}
\label{fig:FilterPot}
\end{figure}%See DocDB 1472

The final piece of the high voltage system is the high voltage feedthrough which delivers the drift voltage to the cathode of the TPC. This feedthrough is based on a a custom design that was used by ICARUS \cite{IcarusDetector}. The feedthrough consists of a stainless steel inner conductor surrounded by a tube of ultra high molecular weight polyethylene further surrounded by a stainless steel outer ground tube. The high-voltage feedthrough enters the cryostat from a dedicated 4-5/8 inch Conflat flange at the top of the cryostat. The end of the inner conductor is finally attached to the cathode by a flexible conductor bolted at both ends. A technical drawing of the LArIAt high voltage feedthrough can be found in Figure \ref{fig:HVFT}.

\begin{figure}[htb]
\centering
\includegraphics[scale=0.35]{./figures/lariatFeedthrough.pdf}\\
\caption{Technical drawing of the LArIAT high voltage feedthrough. Various liquid argon levels are shown in cyan to indicate the levels to ensure the feedthrough is immersed in argon during operations}
\label{fig:HVFT}
\end{figure}%See DocDB 1472

During usual operations, the operating voltage from the Glassman power supply was -23.5~kV corresponding to a drift field of $\sim$500 V/cm. During special runs, the high voltage system was operated in a range from 0~kV to 32.8~kV (corresponding to drift fields ranging from 0~V/cm to 700~V/cm) and the feedthrough was tested up to 60~kV with no electrical breakdown observed.

\end{subsubsection}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{subsubsection}{Cathode and Field Cage}\label{sec:FieldCage}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\textcolor{red}{Description of the cathode and field cage is missing}

As shown in Figure \ref{fig:driftregions}, the LArIAT TPC has three drift volumes, each of which has its own electric field. The main drift volume is defined as the region between the cathode plane and the shield plane (C-S). Note that the shield plane wires are \emph{not} read out. The other two drift regions are those between the shield plane and the induction plane (S-I), and between the induction plane and the collection plane (I-C). The electric field in these regions is chosen to satisfy the charge transparency condition to allow for 100$\%$ transmission of the drifting electrons through the shield and then the induction planes. We provide a detailed description of the electric field measurement in each region in \ref{sec:EFieldMeasurements}.

\begin{figure}[htb]
\centering
\includegraphics[scale=0.35]{./images/DriftRegions.png}\\
\caption{Schematic of the three drift regions inside the LArIAT TPC: the main drift volume between the cathode and the shield plane (C-S) in green, the region between the shield plane and the induction plane (S-I) in purple, and the region between the induction plane and the collection plane (I-C) in pink.}
\label{fig:driftregions}
\end{figure}

Table \ref{tab:voltages} provides the default voltages applied to the cathode and the shield, induction, and collection plane for all the runs. These voltages were tuned to maintain the transparency condition when the (I-C) and (S-I) spacing changed. For data taken in Run-I, Run-II, and Run-IIIb the spacing was a constant 4~mm; the spacing was 5~mm for Run-IIIa.

\begin{table}[htpb]
\centering
\caption{Anode planes default voltages}
\label{tab:voltages}
\begin{tabular}{lllll}
\hline
\multicolumn{1}{|l|}{V Cathode} & 
\multicolumn{1}{|l|}{V Shield} & \multicolumn{1}{l|}{V Induction} & \multicolumn{1}{l|}{V Collection} & \multicolumn{1}{|l|}{Run} \\ \hline
\multicolumn{1}{|l|}{-23.17 kV} &
\multicolumn{1}{|l|}{-298.8 V} & \multicolumn{1}{l|}{-18.5 V}      & \multicolumn{1}{l|}{338.5 V} & \multicolumn{1}{|l|}{Run-I, Run-II \& Run-IIIb}      \\ \hline
\multicolumn{1}{|l|}{-23.17 kV} & 
\multicolumn{1}{|l|}{-325.8 V} & \multicolumn{1}{l|}{-0.6 V}      & \multicolumn{1}{l|}{421.9 V} & \multicolumn{1}{|l|}{Run-IIIa}      \\ \hline

                              &                                 &                                 
\end{tabular}
\end{table}

\end{subsubsection}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{subsubsection}{Wire Plane Assembly}\label{sec:WirePlanes}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{subsubsection}*{Wire Winding}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Wires for the LArIAT wire planes are wound to the correct spacing and tension on an independent winding machine before being transferred to the printed circuit boards on which they are installed in the TPC.

\begin{figure}[h!]
\centering
\includegraphics[height=2in]{figures/MamaBear.jpg}
\caption{The wire-winding apparatus in action. Both frames are attached and a single wire has been wound.}
% CAN PROBABLY FIND A BETTER PICTURE!
\label{pic:mamabear}
\end{figure}

Two rectangular steel frames are bolted back-to-back on an apparatus that rotates at a constant speed, shown in Figure \ref{pic:mamabear}. As the frames rotate, wire is drawn from a spool and wrapped around the two frames at constant tension (monitored by a tensiometer in the device paying out the spool). Once the requisite number of wires have been wound, the wires are epoxied onto the steel frames with five-minute epoxy and cut, giving two separate wire planes (one on each frame).
\end{subsubsection}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{subsubsection}*{Wire Transfer}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Once the wire planes have been wound, they must be transferred from the winding frames onto the G10 wire-carrier boards that go into the LArIAT TPC. To achieve this, the frames are unbolted from the winding apparatus and laid over the board onto which the wires are to be transferred, which itself is bolted to a rigid aluminium frame to prevent it from flexing under tension. Careful manual alignment is performed, using guide holes drilled in the board for reference. This alignment is shown in Figure \ref{pic:wiretransfer}.

Once the wires are aligned to the solder pads on the board, they are glued to the board behind the solder pads with a temporary epoxy strip (using five-minute epoxy) to preserve the alignment. This is then followed by a permanent epoxy strip on the inside of the solder pads, using a slow-curing epoxy that is safe for cryogenic conditions. This epoxy is allowed to cure for 12 to 24 hours before the wires are cut behind the temporary strip and the winding frame can be removed.

\begin{figure}[h!]
\centering
\includegraphics[height=2in]{figures/WireTransfer.jpg}
\caption{One of the \textcolor{red}{This can be 4mm or 5 mm wire plane} 3~mm wire planes during transfer, with the wires not yet cut from the winding frame. A steel bar was used for this plane to ensure all wires made good contact with the board.}
\label{pic:wiretransfer}
\end{figure}

\end{subsubsection}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{subsubsection}*{Soldering of Electrical Connections}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

When the wire transfer is complete, the wires are soldered to their pads using [SOLDER DETAILS HERE], as are the resistor and capacitor components for the U and V planes \textcolor{red}{Which are the values of capacitances and resistances?}. Mylar sheeting is used to protect the active region of the wire planes from evaporating solder flux during this process. Once all wires have been soldered in place, the wires are broken off behind the solder pads and the temporary epoxy strip is removed. The entire board is cleaned with ethyl alcohol to remove any residue of solder flux before the finished wire planes ahead of installation in the TPC.
\end{subsubsection}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{subsubsection}*{Wire Plane Testing}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\textbf{Electrical Testing}\\
\textcolor{red}{Both the test on the RC and on the wire tension are very bad described. According to me it is better to delete the section if a better description can not be provided.}
Every channel on each completed plane is probed with a multimeter and a signal generator to check for electrical continuity, and in the case of the U and V planes, for the correct RC time constant from the components.\\
\textbf{Tension Testing}\\
A laser tensiometer is mounted on a rigid aluminium arm over each completed frame and used to verify the tension on a representative sample of the wires. [QUANTITATIVE RESULTS?]

\end{subsubsection}

\end{subsubsection}

\textcolor{red}{The title says "Cold" but all the electronic, cold and warm, is described}
The LArIAT TPC front-end electronics comprises a 480-channel analog signal path from the TPC wireplanes to the signal digitizers. The front-end system also includes a digital control system for the TPC-mounted electronics, a power supply, and a distribution system.
A block diagram of the overall system is shown in Fig. \ref{pic:FEelectronics}.

\begin{figure}[htbp]
 \centering
 \includegraphics[width=1.0\textwidth]{figures/LArIAT_FE_Electronics.png}
\caption{
Overview of LArIAT Front End electronics. 
} 
\label{pic:FEelectronics}
\end{figure}

The electrical signals on the TPC readout wires are typically quite small, being a direct readout of the ionization of the liquid argon. To achieve a good Signal-to-Noise Ratio (SNR) for such signals, the LArIAT TPC is instrumented with cold amplifiers developed by Brookhaven National Lab (BNL) and mounted directly to the TPC frame inside the liquid argon cryostat. The BNL amplifiers are built as custom Application-Specific Integrated Circuits (ASICs) [ADD reference]. The BNL ASICs adopted in LAriAT are designated as LArASIC, version 4-star.

The LArASICs are mixed signal devices, where the analog amplifier characteristics are controlled by digitally-set parameters. The maximum gain setting of the LArASIC is 25 mV/fc. At this gain a 3.5~fC charge deposition - corresponding to the expected charge arriving at an individual wire from a passing MIP energy deposition - on a readout wire will generate an output with 90~mV amplitude. At the other end of the readout chain, the CAEN V1740 [\textbf{ADD}: CAEN refernece] digitizers have a 12 bit resolution and a maximum input range of 2~VDC, yielding about 0.5~mV LSB. With the analog signal path from LArASIC to the digitizers being designed to provide unity gain, a passing MIP in the TPC is expected to generate a digital signal with amplitude of about 180 ADC counts.  [\textbf{UPDATE}: Dean found gain to be less than unity, add details; what was the measured MIP amplitude?] \textcolor{red}{what is the sampling rate of the board?}

The readout racks containing the signal digitizers are positioned at a distance of $\sim$8~mt from the cryostat and are referenced to a different ground respect to the TPC electronics ground.  The distance minimizes the presence of secondary, tertiary and halo charged particles passing through the digitial electronics.  To prevent problems caused by differing grounds and noise pick-up along the signal transmission line, the single-ended analog signals from the ASICs are transmitted from the cryostat to the readout racks as differential analog pairs, each pair individually shielded.  This is achieved using four types of interconnected cards.  The first type of card in the system is a 48-channel cold motherboard (CMB-48) that mounts directly to the TPC and houses three 16-channel LArASICs per card.  The next card in the signal path is a single cryostat feed through card (FT) that carries the 480 signals, along with power and control lines, across the cryostat boundary.  A set of warm receiver and driver (WRD-48) cards plug directly into the cryostat feed-through and amplify the single-ended TPC signals as differential analog. The differential analog signals are then driven through 8~mt high quality pleated foil cables to a set of differential-to-single-ended (D2S-64) cards, converting the differential signals into single-ended ones, with enough drive capability to allow direct input into an array of CAEN V1740 digitizers.

At the receiving end of the differential path the differential-mode analog signal arrives with common-mode noise cancelled and converted to single-ended analog for the short distance transmission to the CAEN V1740 digitizers.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Cryostat feed through}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The feedthrough (FT) card developed for LArIAT is sandwiched between an ASA flange and cap, and sealed with o-rings. A stiff mechanical structure and backplane-style connectors support the FT card and allow for the WRD-48 cards to be plugged directly into it. Such an arrangement facilitates the assembly and potential repair work while reducing the amount of cables and connectors needed.

\begin{figure}[htbp]
 \centering
 \includegraphics[width=0.5\textwidth]{figures/FT_drawings.png}
\caption{
Drawing of LArIAT feed through and Warm Receiver Driver crate. 
} 
\label{pic:FeedthroughElectronics}
\end{figure}

The FT card is built as an 8-layer PCB with dimensions 254~mm $\times$ 432~mm. The outer layers of the card are mostly uninterrupted ground planes to increase noise immunity. The internal traces are built with relatively large 12~mil design rules to enhance manufacturing yield and physical robustness. Signal on adjacent layers are staggered to reduce capacitive cross-talk couplings, and ground fills are provided whenever possible on all copper layers.

Supporting the FT card is a mechanical structure that is rooted on the cryostat flange. The overall assembly of the card, the ten WRD-48 cards, and the WRD card cage is light enough to be safely cantilevered from the cryostat flange. The backplate of the structure is a solid copper sheet which is electrically connected with the cryostat flange though lock washers and also connected to the FT card through 40 conductive standoffs. The copper plate and the cryostat flange are defined as the electrical ground of the TPC readout electronics.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection*{CMB-48 card details}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Each CMB-48 card holds 3 LArASICs. The card provides electrical connection to the LArIAT TPC wires and mechanical connection to the TPC structure. The CMB-48 board is located inside the cryostat and is thus inaccessible during data-taking periods. The board is designed to minimize the source of electrical noise, the risk of failure due to thermal stress, and the risk of component failure.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection*{WRD-48 card details}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The WRD-48 cards are designed to be paired one-to-one with the CMB-48 cards inside the cryostat.The WRD-48 cards hold 48 channels of single-ended to differential analog amplifiers to transmit the TPC signals to the distant digitizers. These cards also provide low-noise power regulation, digital signal conditioning, front-panel diagnostic LEDs and voltage monitor ports, and an on-board test-pulse generator.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{D2S-64 card details}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The D2S-64 cards convert the differential analog signals from the WRD-48 cards to single-ended signals and drive these signals into the $50 \Omega$ input impedance of the CAEN V1740 digitizers.

The LArIAT data acquisition system is triggered to read out the digitizing buffers by a CAEN V1495 module. The V1495 is a powerful, easily configurable coincidence module featuring a user-programmable FPGA.  Sixteen logical inputs (upgraded to 32 in Run-III) are sampled on a 10~ns clock, checking for matches to any of the sixteen user-defined patterns in the trigger menu.  If a given trigger pattern is satisfied for two consecutive clock ticks, that pattern fires.

NIM-standard logic pulse inputs to the trigger decision come from any of the instruments in the experiment, in principle; trigger inputs are derived from beam instruments, from the cosmic ray taggers, and from the cryostat's scintillation detectors. LArIAT takes full advantage of the trigger card's great flexibility.  As conditions change and innovations are gradually made to LArIAT's data-taking process, improvements are made to the trigger input designs and the combinations of inputs which define trigger patterns for the V1495.  The configuration of the trigger inputs is automatically logged in a database at the start of each run as part of the DAQ configuration, along with the rest of the XML configuration file for the V1495.

\subsubsection{Inputs to the Trigger System}

Two primary inputs to the trigger card are from the time of flight (see Sec.~\ref{sec:TOF}) and the wire chamber (see Sec.~\ref{sec:MWPC}) systems in the tertiary beamline.  Coincidence of activity in both of these systems strongly suggests that a charged particle has made the journey down the tertiary beam line from the copper target to the LArTPC in the cryostat, and that we will have a measure of its momentum and its velocity. PMT pulses from the time of flight system each pass through a 
%FIXME WHAT MODEL linear fanout?
linear fan-out, one output of which is threshold discriminated by a
%FIXME WHAT MODEL
discriminator to produce a NIM logic pulse for use in trigger logic.  On each of the upstream or downstream TOF paddles, we form a coincidence to within 20~ns of pulses from all the PMTs observing that same block of scintillator.  The upstream paddle coincidence signal is delayed 20~ns to allow any approximately lightspeed particles to travel 6.5~m to the downstream paddle.  The same upstream paddle coincidence signal is also widened to 100~ns to allow for slow-traveling (high-mass) particles.  Coincidences of these upstream and downstream TOF trigger input signals can be made in the V1495 trigger card.

Each wire chamber is read out by four multi-hit TDCs (time to digital converters), with sixty-four wires per TDC and two TDCs per plane (one horizontal and one vertical).  The TDCs each provide a logical ``fast'' OR of their inputs, indicating that one or more of their sixty-four wires went over the settable threshold.  Using NIM logic units, the OR of the horizontal wires and the OR of the vertical wires are input to a coincidence unit for each wire chamber, providing a single logical pulse for each of the four wire chambers, indicating that at least one horizontal wire and one vertical wire saw significant signal.  The single culminating logical pulses from each of the four wire chambers make up the first four inputs to the V1495 trigger card.  Within the user-programmable FPGA, the V1495 looks for the coincidence within 20~ns of at least three of these four specially-treated logical inputs.

Another primary input to the trigger card is from the cosmic towers (see Section~\ref{sec:CosmicRayPaddle}). To capture cosmic ray events in which a minimally ionizing cosmic ray muon crossed the TPC along the body diagonal, NIM modules form the logical coincidences from the two cosmic towers, one upper and one lower paddle assembly, in each combination.  The OR of these is provided as an input to the V1495.

Three important logic pulses are derived from the timing of the beam.  These include a pulse in a brief window before the beam, a pulse indicating that the beam is on, and a pulse which defines the beam-free period which may be used for collecting cosmic-ray events.  An adjustable pulser is a fourth trigger input which does not depend on any particular activity in the experiment hall,  useful for collecting background events with zero bias.

The PMTs observing liquid argon scintillation light (see Section \ref{sec:PhotonSystem}) produce pulses which form the foundation of several interesting trigger inputs.  Thresholding a copy of each PMT pulse (after amplification), and requiring a coincidence of pulses within $\sim$20~ns, creates simple trigger inputs indicating ionizing radiation was produced in the TPC.  This scintillation logic pulse is used to initiate a gate which spans the length of the TPC drift time, creating a logic signal which is remains ``on'' while significant drift charge may still be present in the TPC.  In addition, requiring a delayed coincidence of two subsequent scintillation logic pulses, separated by a variable length of time ranging from 300~ns to 7~$\mu$s, is used to create a trigger input to select events where a cosmic muon stops and decays to a Michel electron in the TPC.  A few different versions of this light-based trigger were implemented throughout the course of LArIAT's run time to allow reconstruction and calorimetric studies of Michel electrons. Figure~\ref{michel_logic} shows a schematic diagram of the logic comprising the Michel electron trigger.

\begin{figure}
\includegraphics[width=\textwidth]{figures/trigger_michellogic.png}
\caption{\label{michel_logic}A schematic diagram of the trigger logic used to select Michel electron events during the cosmic readout window of the LArIAT supercycle.  The two PMT signals refer to the Hamamatsu (``HMM'') and ETL PMTs described in Section~\ref{sec:PhotonSystem}.  For some data-taking periods in Run-II, un-amplified pulses were discriminated at 180 mV to act as a veto on events that may saturate the dynamic range of the V1751 digitizer.  The discriminator thresholds used on the amplified (x10) PMT signal copies (\emph{ThA}, \emph{ThB}) as well as the Gate Delay period, were adjusted between run periods while experimenting with different configurations.}
\end{figure}

Further trigger inputs come from the beam line instrumentation behind the LArTPC cryostat, the PMTs of the Punch-Through scintillator paddles and those of the scintillator paddles instrumenting the Muon Range Stack.  The PMT pulses of all four of the broad-faced Punch Through paddles are discriminated to form logic pulses.  A single logic pulse is formed from these, indicating activity in at least two overlapping paddles at the rear of cryostat, before the steel block of the range stack.  PMT pulses from the Muon Range Stack are amplified and threshold discriminated.  These MuRS paddle pulses are then combined as in the Punch Through, creating single-bit indicators for each of the four instrumented layers that at least one pair of overlapping scintillator paddles sent signals within a 20~ns coincidence window.

\subsubsection{Trigger Decision and Issuance}

The V1495 may be configured to have up to sixteen trigger patterns and sixteen veto patterns, based on the trigger input signals.  A trigger pattern is defined as the AND of one or more defined inputs, and may include the NOT of the AND of further inputs.  Veto patterns are independently defined in the same way, but they have a very different effect.  When any of the trigger patterns fire, a ``fast trigger'' signal is issued and an adjustable countdown is initiated.  If the countdown completes without a veto pattern firing, the ``slow trigger'' signal is also issued and on a distinct hardware channel. Otherwise, if a veto pattern fires during the countdown, the slow trigger signal is vetoed.

The fast trigger signal prompts readout of all the `short' data buffers, which include the V1751 modules, the V1495 itself, and the MWPC controller.  The V1751 buffers typically contain digitized PMT signals from the time of flight and cryogenic light collection detectors. Readout of the TPC wire signals, which are much longer and more numerous, is only prompted at the issuance of the slow trigger.

In addition to the drift electrons, LArIAT also collects liquid argon scintillation photons.  This is achieved using several cryogenic photodetectors held in place by a PEEK support structure which screws into a side access flange as shown in Figure~\ref{lightsys_pmts}.  Two PMTs, a 3-inch diameter Hamamatsu R-11065 and 2-inch diameter ETL D757KFL~\cite{lightsys-pmttests}, contribute to most of the system's light collection capacity.  During Run-I and Run-II, several silicon photomultiplier (SiPM) arrays were also included as part of the photon detection system.  These SiPMs consisted of two Hamamatsu S11828-3344M 4x4 arrays (each with 12x12mm$^2$ total active area) as well as one single-channel SensL MicroFB-60035 (6x6mm$^2$ active area) on custom preamplifier readout boards which were mounted along the edges of the PMT holder.  The windows of all devices are held parallel to the wireplanes with approximately 5~cm of clearance.

%------------------------------------------
\begin{figure}
\centering
\includegraphics[height=2.2in]{figures/lightsys_pmts.png}
\hspace{1cm}
\includegraphics[height=2.2in]{figures/lightsys_wls.png}
\caption{LArIAT's photodetector system for observing LAr scintillation light inside the TPC (left), and a simplified schematic of VUV light being wavelength-shifting along the TPB-coated reflecting foils (right).}
\label{lightsys_pmts}
\end{figure}
\begin{figure}
\centering
\includegraphics[height=0.25\textheight]{figures/lightsys_etlbase.jpeg}
\hspace{0.5cm}
\includegraphics[height=0.25\textheight]{figures/lightsys_hmmbase.jpg}
\caption{\label{voltagedividers}Photos of the voltage divider bases for the ETL PMT (left) and the Hamamatsu PMT (right) used in Run-II.  The cable connections to the bases seen here were used for powering and testing prior to installation.  The yellow through-hole signal coupling capacitors seen on both bases are 18~nF (X7R) and are rated to 2~kV.}
\end{figure}
%------------------------------------------

Liquid argon scintillation light is produced in the vacuum-ultraviolet (VUV) range and must be shifted to visible before it can be detected by most photosensors.  This wavelength-shifting occurs within a thin layer of tetraphenyl-butadiene (TPB) which coats highly-reflective VIKUITY dielectric substrate foils lining four walls of TPC.  A photo of these foils mounted inside the TPC is shown in Figure~\ref{lightsys_foils}.  A VUV photon interacts with the TPB to induce emission of one or more visible photons which are then emitted or reflected back into the active volume, increasing the probability of their detection by the photosystem~\cite{lightsys-szelc}. This technique also improves light yield uniformity across the TPC's active volume, a feature relevant to light-augmented calorimetry.

%------------------------------------------
\begin{figure}
\centering
\includegraphics[scale=1.3]{figures/lightsys_foils.png}
\caption{\label{lightsys_foils}The TPB-coated reflector foils mounted to the TPC field cage walls as viewed through the front cryostat opening. {\textcolor{red}{Better pictures needed.}}}
\end{figure}
%------------------------------------------

A TPB coating\footnote{The TPB coating was made using a solution of 100~mL toluene, 2~g polystyrene, and 800~mg of TPB prior to Run-II. The coating was re-applied to the ETL midway through Run-II using the same solution but with less TPB (600~mg).} was added to the windows of the ETL PMT and SensL SiPM prior to Run-II and reapplied about midway through the run.  This allowed some of the VUV scintillation light to down-convert directly at their faces instead of at the coated reflector foils, rendering the photodetectors directionally sensitive to this light emanating from the point of energy deposition.  These coatings were removed prior to Run-III.

For Run-I, the voltage dividers for the PMTs were configured for negative bias with a DC-coupled anode, and were switched to positive bias mode (AC-coupled anode with grounded photocathode) for Run-II to minimize induced noise on wires near the charged photocathodes and outer PMT chasis.  To configure each base for positive polarity before the start of Run-II, 18 nF (X7R) capacitors rated to 2kV were soldered onto the circuits to couple each PMT's anode to the outgoing signal line.  At liquid argon temperature and PMT operating voltages, these capacitances reduce to $\sim$1/3 of their nominal value~\cite{lightsys-capacitors,lightsys-ubsplitting}.

For Run-III, a prototype ARAPUCA~\cite{lightsys-arapuca} detector was installed, replacing the SiPM boards. The main purpose of the prototype is research and development of new alternatives for photosensors for liquid argon-based detectors.
The central concept of ARAPUCA is to capture photons within a box with highly reflective internal surfaces (reflectivity~>~98~\% ), resulting in high photon detection efficiency even for limited active coverage, which is defined as the total area of photosensors viewing the internal volume. The key to the ARAPUCA's photon-trapping mechanism is the use of dichroic short-pass optical filters which are highly transparent to wavelengths below a particular cutoff and highly reflective above it. The filter is externally coated by a wavelength-shifting (WLS) film. The WLS film on the outer side of the filter converts the VUV scintillation light from the liquid argon to a wavelength below the cutoff, so the filter is transparent to the incident photons, allowing them to enter into the box. A second WLS coating covers the inner side of the filter or optionally the internal walls of the box. The internal WLS layer enables a photon to convert a second time to a wavelength greater than the cutoff condition for which the filter is highly reflective, trapping the photons inside the box. The interior of the ARAPUCA is viewed by light sensors which capture the trapped photons after few reflections.
%covered by one or more SiPMs. The number of SiPMs can be increased and read out using an active ganging circuit. Figure \ref{arapuca-principle} illustrates the working principle of the device.

The prototype ARAPUCA device installed in LArIAT, shown in Figure~\ref{fig:arapuca_photo}, is 4.5$\times$5.5 cm$^2$, with a total exposed filter area of 3.5$\times$4.5 cm$^2$.  The WLS film on the outer surface of the filter is P-Terphenyl ($\lambda \sim  350$~nm), while the inner surface is coated in TPB ($\lambda \sim 430$ nm). It is equipped with a single SensL MicroFC 60035 SiPM biased at +24~V in liquid argon.  Its signals are sent through a commercial preamplifier before being sent to the DAQ.

%------------------------------------------
\begin{figure}
\centering
\includegraphics[height=5cm]{figures/lightsys_arapuca2.jpg}
\hspace{0.5cm}
\includegraphics[height=5cm]{figures/lightsys_arapucaPmts.jpg}
\caption{The prototype ARAPUCA device without its dichroic filter (left) as well as mounted to the PMT support structure after the installation of the filter (right).}
\label{fig:arapuca_photo}

\end{figure}
%------------------------------------------

\subsection*{Signal Digitization}

Signals from all photodetectors are routed from the side-flange of the cryostat to NIM fan-outs to provide the 50-Ohm termination needed to minimize reflections along the cables. Copies of the PMT signals are made to aid in several light-based triggers, and copies of all signals are then recored into data using CAEN V1751 digitizer boards with 1~GHz sampling rate.

%------------------------------------------
\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{figures/lightsys_signals.png}
\caption{Some example signals, in ADC units, from LArIAT photodetectors for a triggered cosmic Michel electron event.}
\label{lightsys_signals}
\end{figure}
\begin{figure}
\centering
\includegraphics[width=0.6\textwidth]{figures/lightsys_avewfm_hmm.png}
\caption{\label{wfm_hmmpmt}An example of an averaged inverted waveform of the Hamamatsu PMT after event-by-event noise removal and overshoot correction.  Data was taken from a small subset of events from Run-II.  In this example, a falling exponential is fit to the signal region from 300~ns to 3~$\mu$s following the onset of the pulse to extract the late-light component lifetime $\tau$ = 1117~ns.}
\end{figure}
%------------------------------------------

\subsection*{Nitrogen Contamination Measurement Using Light}

The light detection system can also be used to derive information about the nitrogen (N2) contamination in liquid argon. The influence of N2 on liquid argon scintillation light emission has been observed in the WArP experiment~\cite{WARP-nitrogen}. With a higher N2 concentration, the observed decay time of the liquid argon slow component decreases significantly, from over 1~$\mu$s to hundreds of nanoseconds. In this measurement, waveforms recorded by the ETL PMT in Run-I, selected to cover periods of varying N2 concentrations (as measured by the external sensor in Figure~\ref{light_nitro}), were analyzed. An exponential fit was performed to estimate the decay time of the slow light component. Results, shown in Figure~\ref{light_nitro}, were compared with the model from the WArP paper~\cite{WARP-nitrogen}.

%------------------------------------------
\begin{figure}
\includegraphics[height=0.2\textheight]{figures/nitrosensor.jpeg}
\hspace{0.5cm}
\includegraphics[height=0.2\textheight]{figures/nitro_qt.jpeg}
\label{light_nitro}
\caption{Concentration of N2 in LArIAT as measured by an external sensor in Run-I (left), and results of fitting exponential models to waveforms recorded by the ETL PMT compared with the model from WArP~\cite{WARP-nitrogen} (red line).}
\end{figure}
%------------------------------------------

### 2018-03-12 Chapter-2/2-TheLArIATExperiment.tex
s

@@
 The components of the tertiary beamline instrumentation key for the hadron cross section analyses are the target and collimators system, the two bending magnets (in a similar configuration used for the  MINERvA T-977 test beam calibration~\cite{MinervaTestbeam}) a set of four wire chambers (WCs) and two time-of-flight scintillating paddles (TOF) and, of course, the LArTPC.
-The magnets determine the polarity of the particles in the tertiary beam; the combination of magnets and wire chambers determine the particles momentum, which is used in conjunction with the TOF to determine the particle species.
+The magnets determine the polarity of the particles in the tertiary beam; the combination of magnets and wire chambers determine the particles momentum, which is used to determine the particle species in conjunction with the TOF.
 A muon range stack downstream from the TPC and two sets of cosmic paddles configured as a telescope surrounding the TPC are also used for calibration purposes.
+
+\subsection{Bending Magnets}\label{sec:Magnets}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+LArIAT uses a pair of identical Fermilab type ``NDB" electromagnets, recycled from the Tevatron's anti-proton ring \textcolor{red}{CITE CDF?}. 
+The magnets are a fundamental piece of the LArIAT beamline as they are used in all the three tasks of the LArIAT beamline: the sign of the current in the magnet provide the selection of either positively or negatively charged particles, the value of the magnetic field is used in the momentum determination and the subsequent particle identification.
+
+We describe here the characteristics and response of one magnet.
+We expect the second one have a similar response, being identical in shape and with a similar history.
+The magnet aperture measures the gap dimensions to be 14.224~cm of height, 31.75~cm width, and  46.67~cm length.
+The wire chambers aperture ($\sim$12.5~cm) is smaller than the magnet aperture, thus, only the central part of the magnet gap is utilized.
+The field is extremely uniform over this limited aperture and was measured with two Hall probes, both calibrated with nuclear magnetic resonance probes.
+The probes measured the excitation curve shown in Figure~\ref{fig:magnet_excitation}.
+
+\begin{figure}[!h]
+\begin{centering}
+\vspace{-0.3cm}
+\includegraphics[height=3.0in]{Chapter-2/Images/ExcitationCurves.png}
+\caption{
+{ Magnetic field over current as a function of the current, for one NDB magnet (excitation curve).
+The data was collected using two Hall probes (blue and green).
+We fit the readings with a cubic function (black) to average of measurements (red) given in the legend.}
+}
+\label{fig:magnet_excitation}
+\end{centering}
+\end{figure}
+
+The current being passed through the magnets at a given time is identical in both magnets.
+For Run II period, the current settings explored were 60A (B $\sim$0.21 T) and 100A (B $\sim$0.35 T) in both polarities.
+Albeit advantageous to enrich the tertiary beam composition with high mass particles such as kaons, we never pushed the magnets current over 100 A, not to incur in overheating.
+During operation, we operated a air and water cooling system on the magnets and we remotely monitored the magnets temperature.
 
@@
 
-The LArIAT time-of-flight (TOF) detector system consists broadly consists of two scintillator paddles, which bracket the beamline and are shown in blue in Figure~\ref{fig:tert-layout}.
+The LArIAT time-of-flight (TOF) detector system consists broadly of two scintillator paddles, which bracket the beamline and are shown in blue in Figure~\ref{fig:tert-layout}.
 The exact details of the of the TOF system were changed between Run-II and Run-III to allow for better performance and timing resolution.
@@
 \label{fig:TOFSystemRunIandII}
-\end{centering}
-\end{figure}
-
-\paragraph{\textbf{TOF System for Run-III:}}
-Two modifications were made to the downstream TOF paddle for Run-III to improve the efficiency and timing of the system.
-The large area scintillator paddle was replaced with a slightly smaller paddle more similar to the upstream and was now readout with four new PMT's (2" Hamamatsu PMTs (R-1828) inherited from the muon g-2 experiment(E821) at BNL) located on the diagonals of the counter, as shown in Figure \ref{fig:TOFSystemRunIII}.
-
-\begin{figure}[!h]
-\begin{centering}
-\vspace{-0.3cm}
-%\includegraphics[height=2.3in]{figures/TOFRunIII.png}
-\caption{
-{\scriptsize \sf Image of the modified downstream TOF system deployed during Run-III data taking }
-}
-\label{fig:TOFSystemRunIII}
 \end{centering}
@@
 
-Finally, the amplitude of the ToF pulses (peak minus pedestal) reflects the energy deposited in the scintillator, largely through ionization.
-At a given momentum, more massive particles such deuterons, protons, and kaons, will deposit more energy than less massive particles, such as electrons, muons, and pions.
-As described below, ToF and momentum information from the wire chambers can be combined to determine the particle's mass and hence its identity.
-Although the energy deposited in the scintillator isn't used to distinguish among the various particle types, it's clear from Figure~\ref{fig:tofamplitude} that the spectra of energy deposition for
-electrons, pions, and protons are very different.
-
-\begin{figure}[!h]
-\begin{centering}
-\vspace{-0.3cm}
-%\includegraphics[height=2.3in]{figures/tofamplitude.png}
-\caption{
-{\scriptsize \sf Average upstream ToF pulse amplitude for protons and pions/positrons.}
-}
-\label{fig:tofamplitude}
-\end{centering}
-\end{figure}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Bending Magnets}\label{sec:Magnets}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-The pair of spectrometer magnets in the LArIAT tertiary beam are Fermilab type "NDB" electromagnets, originally used in the anti-proton ring.
-These magnets provide the necessary bend in the particle trajectory to allow for a momentum measurement to be performed.
-The magnets have a gap height of 14.224~cm, a gap width of 31.75~cm, and have an iron length of 46.67~cm.
-The aperture presented by the magnets is larger than that of the wire chambers ($\sim$12.5cm) and thus only the central part of the magnet is ever utilized.
-Over this limited aperture there is negligible variation of the field integral and thus a very uniform field is provided.
-The field intensity in one of the pair of magnets was measured using two Hall probes, both calibrated with NMR, to obtain the excitation curve shown in Figure~\ref{fig:magnet_excitation}).
-The second magnet, having been made to the same standard and having the same history, is assumed to have a very similar response.
-\begin{figure}[!h]
-\begin{centering}
-\vspace{-0.3cm}
-%\includegraphics[height=3.0in]{figures/ExcitationCurves.png}
-\caption{
-{\scriptsize \sf Excitation Curve, B/I vs. I, for one NDB magnet, using two Hall probes (blue and green). Parameters from a cubic fit (black curve) to average of measurements (red) given in legend.}
-}
-\label{fig:magnet_excitation}
-\end{centering}
-\end{figure}
-
-These magnets were air- and water-cooled during operations and their temperatures monitored.
-Due to concerns of overheating, the maximum usable field at the center of the magnets is $\sim$0.35 Tesla with 100 Amps of current being passed through the magnets.
 

### 2018-03-12 Chapter-2/2-TheLArIATExperiment.tex
s

@@
 
+\subsection{Multi-Wire Proportional Chambers}\label{sec:MWPC}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure}[!h]
+\begin{centering}
+\vspace{-0.3cm}
+\includegraphics[height=2.3in]{Chapter-2/Images/WireChamber.png}
+\caption{
+{One of the four Multi Wire Proportional Chambers (WC) used in the LArIAT tertiary beamline.}
+}
+\label{fig:wirechamber}
+\end{centering}
+\end{figure}
+
+LArIAT uses four Multi-Wire proportional chambers, or wire chambers (WC) for short, two upstream and two downstream from the bending magnets.
+The geometry of one chamber is shown in Figure~\ref{fig:wirechamber}: the WC effective aperture is a square of  12.8~cm perpendicular to the beam direction.
+Inside the chamber, the 128 horizontal and 128 vertical wires hang at a distance of 1~mm from each other in a mixture of 85\% Argon and 15\% isobutane gas.
+The WC operating voltage is between 2400 V and 2500 V.
+LArIAT wire chambers are an upgraded version of the Fenker Chambers~\cite{Fenker}, where an extra grounding improves the signal to noise ratio of the electronic readout.
+
+Two ASDQ chips~\cite{ASDQchip} mounted on a mother board plugged into the chamber serve as front end amplifier/discriminator.
+The chips are connected to a multi-hit TDC~\cite{Sten} which provides a fast OR output used as first level trigger.
+The TDC time resolution is 1.18~ns/bin and can accept 2 edges per 9~ns.
+The maximum event rate acceptable by the chamber system is of 1 MHz: this rate is not a limiting factor considering that the rate of the tertiary particle beam at the first wire chamber is estimated to be less than 15 kHz.
+A full spill of data occurring once per supercycle is stored on the TDC board memory at once and read out by a controller specially designed controller.
+We use LVDS cables to carry both power and data between the controller and the TDCs and from the controller to the rest of the DAQ.
+%It is possible to program the time window for acceptance for hits, time offsets, front end threshold, and pulse shaping parameters through the controller via a USB from a PC or through an Ethernet connection.
+
+\subsubsection{Multi-Wire Proportional Chambers functionality}\label{sec:MWPCfunc}
+
+\subsection{Time-of-Flight System}\label{sec:TOF}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+Two scintillator paddles, one upstream to the first set of WCs and one downstream to the second set of WCs  form LArIAT  time-of-flight (TOF) detector system.
+
+ \textcolor{red}{find exact dimentsions}
+The upstream paddle is made of a 10 x 6 x \textcolor{red}{?}~cm scintillator piece, read out by two \textcolor{red}{Hamamatsu 2 inch} PMTs mounted on the beam left side which collect the light from light guides mounted on all four edges of the scintillator.
+The downstream paddle dimensions are  14 x 14 x \textcolor{red}{?}~cm and it is read out by two \textcolor{red}{Hamamatsu 2 inch} PMTs on the opposite ends of the scintillator.
+The relatively thin width on the beamline direction minimizes energy loss of the particles coming from the target in the scintillatior material.
+
+\begin{figure}[!h]
+\begin{centering}
+\vspace{-0.3cm}
+%\includegraphics[height=2.3in]{figures/tofdelay.png}
+\caption{
+{\scriptsize \sf Pictures of the TOF system as was deployed during Run-I and Run-II data taking. The left image is of the upstream TOF paddle and the right image is of the downstream TOF paddle }
+}
+\label{fig:TOFSystemRunIandII}
+\end{centering}
+\end{figure}
+
+The CAEN 1751 digitizer is used to digitize the TOF PMTs signals at a sampling rate of 1 GHz.
+The 12 bit samples are stored in a circular memory buffer.
+At trigger time, data from the TOF PMTs are recorded to output in a 28.7 $\mu$ second windows starting  approximately 8.4 $\mu$sec before the trigger time.
+
+\subsubsection{TOF functionality}\label{sec:MWPCfunc}
+
+The TOF signals rise time (10-90\%) is 4 ns and a full width, half-maximum of 9 ns consistent in time.
+The signal amplitudes from the upstream TOF and  downstream TOF are slightly different:  200 mV for the upstream PMTs but only 50 mV for downstream PMTs.
+
+The time of the pulses was calculated utilizing an oversampled template derived from the data itself.
+The pulse pedestal was taken from samples far from the pulse and subtracted to the pulse amplitude.
+We then stretch vertically a template to match the pedestal-subtracted pulse amplitude and we move it horizontally to find the time.
+With this technique, we find a pulse time-pickoff resolution better than 100 ps.
+The pulse pile up is not a significant problem given the TOF timing resolution and the rate of the particle beam.
+Leveraging on the pulses width uniformity of any given PMT (sigma of 400 ps),  we flag events where two pulses overlap as closely in time as 4 ns with an 90\% efficiency according to simulation.
+
+We combine the pulses from the two PMTs on each paddle to determine the particles' arrival time by averaging the time measured from the single PMT, so to minimize errors due to optical path differences in the scintillator.
+However, a time spread of approximately 300~ps is present in both the upstream and downstream detectors, likely due to transit time jitter in the PMTs themselves.
+There is no evidence of systematic timing drift over long periods such as 3-4 months of a data-taking: the maximum variation of the average time differences between pairs of PMTs reading out the same scintillator is of the order of 150~ps.
+
 \section{In the Cryostat}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Time-of-Flight System}\label{sec:TOF}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-The LArIAT time-of-flight (TOF) detector system consists broadly of two scintillator paddles, which bracket the beamline and are shown in blue in Figure~\ref{fig:tert-layout}.
-The exact details of the of the TOF system were changed between Run-II and Run-III to allow for better performance and timing resolution.
-
-\paragraph{\textbf{TOF System for Run-I and Run-II:}}
-The upstream paddle, shown in Figure \ref{}, has a relatively small active area (10~cm x 6~cm ) and was chosen to be a relatively thin piece of scintillator to minimize any impact on the momentum of the particles coming from the target.
-Light guides were mounted on all four edges which lead to two \textit{insert PMT type} PMTs mounted on the beam left side.
-The downstream paddle, shown in Figure \ref{fig:TOFSystemRunIandII}, was chosen to have a slightly larger area (14~cm x 14~cm) and had two \textit{insert PMT type} PMTs which were read out on opposite ends of the paddle.
-
-\begin{figure}[!h]
-\begin{centering}
-\vspace{-0.3cm}
-%\includegraphics[height=2.3in]{figures/tofdelay.png}
-\caption{
-{\scriptsize \sf Pictures of the TOF system as was deployed during Run-I and Run-II data taking. The left image is of the upstream TOF paddle and the right image is of the downstream TOF paddle }
-}
-\label{fig:TOFSystemRunIandII}
-\end{centering}
-\end{figure}
-
-%The active area of the upstream paddle is relatively small ( 10~cm x 6~cm ) and that of the downstream paddle is somewhat larger ( 14~cmx 14~cm ).
-Lightguides are mounted on all four edges of each paddle.
-In the first two LARIAT production runs, each of the paddles was read out by two (xxkxkxkxk) photomultiplier tubes (PMTs), The long axis of the downstream paddle was directed horizontally and read out at either end.
-The upstream paddle was rotated by 45 degrees with respect to the horizontal and its two PMTs were mounted to the left side.
-In run 3, four more PMTs were added to the system.
-The upstream paddle was read out with the original set of PMTs and the downstream paddle with four 2" Hamamatsu PMTs (R-1828) inherited from the muon g-2 experiment(E821) at BNL.
-
-During data-taking cycles, signals from the TOF PMTs were sampled at 1 GHz
-with a CAEN 1751 digitizer and 12 bit samples were stored in a circular memory buffer.
-In response to an experimental trigger, a 28.7 $\mu$ second window of samples, starting approximately 8.4 $\mu$sec before the trigger, was written to the
-output.
-The amplitude of the TOF signals was typically 200 mV in the upstream TOF PMTs but only 50 mV in those of the downstream counter.
-The signals have a rise time (10-90\%) of 4 ns and a full width, half-maximum of 9 ns.
-The rate in the upstream counter was typically 15 kHz and much less, 400 Hz in the downstream counter.
-
-Because the shape of signals from each TOF PMT was highly uniform, the time of the pulses was determined using an oversampled template derived from the data itself.
-The DC offset (pulse pedestal) is taken from samples far from the pulse, The template is stretched vertically to match the pedestal-subtracted pulse amplitude and moved horizontally to match the time.
-The pulse time-pickoff resolution is better than 100 ps and relative amplitude resolution is better than x\%.
-Since the average rate of pulses incident on the ToF system is modest, pulse pile up is not a significant problem.
-Given the uniform width of the pulses produced by any given PMT (sigma of 400 ps), the pulse width can be used to flag events where two pulses overlap closely in time.
-Simulations indicate that for pulses as close as 4 ns, a well-chosen pulse width cut is 90\% efficient.
-
-To determine the particles' time of arrival, pulse times from the multiple PMTs mounted on each paddle had to be combined.
-During Run-I and II, with two PMTs mounted on each paddle, a simple average was used.
-In the case of the downstream paddle, with PMTs mounted at opposite ends, the average corrected for the fact that transit time to the two PMTs depended on the particle's point of impact on the scintillator.
-A plot of the time of arrival {\it difference} between the two downstream PMTs vs the longitudinal impact point, made with the help of track stubs from the wire chamber system, is shown in Fig.~\ref{fig:tofdelay}.
-As expected, the time difference
-changes systematically across the face of the scintillator.
-
-\begin{figure}[!h]
-\begin{centering}
-\vspace{-0.3cm}
-%\includegraphics[height=2.3in]{figures/tofdelay.png}
-\caption{
-{\scriptsize \sf $\Delta-t$ between times registered on two downstream paddles.}
-}
-\label{fig:tofdelay}
-\end{centering}
-\end{figure}
-
-Taking the average minimizes ToF errors arising from optical path differences in the scintillator, However, even for a set of particles which pass through a single small area of the paddle, the times of pulses registered in the two PMTs are still spread by approximately 300~ps, an uncertainty which is probably caused by transit time jitter in the PMTs themselves.
-This jitter is evident in both the upstream and downstream detectors.
-
-Although climate-control in the LArIAT experiment hall is fairly primitive, there is little sign of systematic timing drift over longer periods.
-The average time differences between pairs of PMTs reading out the same scintillator varied by no more than 150~ps over the 3-4 months of a data-taking period.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Multi-Wire Proportional Chambers}\label{sec:MWPC}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-The wire chambers are based on the Fenker Chambers~\cite{Fenker} long in use at Fermilab.
-The chambers have been upgraded by adding additional grounding to improve the signal to noise in the electronic readout.
-The chambers, shown in Figure~\ref{fig:wirechamber}, have an effective aperture of 128~mm in both horizontal and vertical distances.
-The wires are spaced at 1~mm, with 128 wires in each view.
-In a test beam, the chambers have typical efficiencies of 98\% to 99\%.
-The gas used is 85\% Argon + 15\% isobutane.
-The wire chambers typically operate between 2400 and 2500 volts.
-
-\begin{figure}[!h]
-\begin{centering}
-\vspace{-0.3cm}
-%\includegraphics[height=2.3in]{figures/WireChamber.png}
-\caption{
-{\scriptsize \sf Image of one of the wire chambers used in the LArIAT tertiary beamline.}
-}
-\label{fig:wirechamber}
-\end{centering}
-\end{figure}
-
-The front end amplifier/discriminator uses the ASDQ chip~\cite{ASDQchip}.
-Two of the ASDQ chips are mounted on a mother board that is plugged into the chamber.
-Short flat cables are used to connect these ASDQ mother board to a new, multi-hit TDC~\cite{Sten}.
-This TDC provides a (maskable) fast OR output that can be used in a first level trigger.
-
-This TDC can accept multiple hits per wire with a time resolution of 1.18~ns/bin.
-The TDCs can accept 2 edges per 9~ns.
-The maximum event rate that can be accepted by the chamber system is of the order of one megahertz.
-These limitations are clearly not an issue for the LArIAT beamline where the rates and timing resolution required are much less stringent than this.
-There is enough memory on the TDC board to buffer a full spill of data (up to about 100 events/spill).
-As in this application the spills occur only once per minute, there is ample time to read out the whole spill through a controller specially designed to support these TDCs.
-Power and data flow to and from the controller to the TDCs and from the controller to the rest of the DAQ is via LVDS.
-The time window for acceptance for hits, time offsets, front end threshold, and pulse shaping parameters may be programmed through the controller via a USB link from a PC or through an Ethernet connection.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Aerogel Cherenkov Detectors}\label{sec:Aerogel}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-Aerogel is an ultra-light material in which the liquid component of the gel has been replaced with a gas.
-The result is a solid with extremely low density, low thermal conductivity, and low index of refraction.
-The goal of having an aerogel threshold Cherenkov detector in the LArIAT beam line is to separate muons and pions in the momentum range where muons emit Cherenkov radiation while pions do not.
-This technique was demonstrated by the MICE and Belle II experiments ~\cite{MICE-aerogel,BelleII-aerogel} and in LArIAT we use two aerogel threshold Cherenkov detector with index of refraction of 1.057 and 1.103.
-Having different indices of aerogel allows LArIAT to do this separation in two different momentum ranges of interest.
-
-\begin{figure}[htb]
-\centering
-%\includegraphics[height=1.85in]{figures/Aerogel1103Figure.png}
-\hspace{1cm}
-%\includegraphics[height=1.85in]{figures/Aerogel1057Figure.png}
-\caption{Number of photoelectrons per cm versus momentum for the two indices of refraction that are employed by the LArIAT aerogel detectors.
-The photoelectron yield includes the response of a bialkali photocathode following the recipe in ~\cite{bib8}.
-}
-\end{figure}
-
-%%% KEK couter
-The Aerogel Threshold Cherenkov Detector with the refractive index of 1.103 is placed just behind the second collimator on the tertiary beam line, between the two downstream wire chambers.
-Figure~\ref{fig:kek_aerogel} is a picture of this aerogel detector.
-It consists of 7 aerogel tiles (totaling 110 mm long in the beam direction), and the cross section of the detector is 108 mm $\times$ 108 mm.
-The array of the aerogel tiles is wrapped with a diffusion sheet and viewed by two PMTs (HAMAMATSU R329-02) from the sides.
-Figure~\ref{fig:kek_aerogel_ly} shows the observed light yield distribution for 550-800~MeV pions hitting the middle of the detector.
-The obtained light yield is equivalent to 22 photoelectrons for $\beta=1$.
-The fake rate for particles with velocity below the Cherenkov threshold is measured to be around 0.5\% with the threshold of 1.5 photoelectrons.
-
-\begin{figure}[htb]
-\centering
-%\includegraphics[height=1.85in]{figures/kek_aerogel.jpg}
-\caption{The Aerogel Threshold Cherenkov Detector with the refractive index of 1.103.}
-\label{fig:kek_aerogel}
-\end{figure}
-
-\begin{figure}[h]
-\centering
-%\includegraphics[height=1.85in]{figures/kek_aerogel_ly.png}
-\caption{The observed light yield distribution for 550-800 MeV pions hitting the middle of the detector.
-The obtained light yield is equivalent to 22 photoelectrons for $\beta=1$.}
-\label{fig:kek_aerogel_ly}
-\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsubsection{Punch-Through and Muon Range Stack Instruments}\label{sec:MuRS}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\input{muon-Punch.tex}
 

### 2018-03-12 Chapter-2/2-TheLArIATExperiment.tex
s

@@
 The signal amplitudes from the upstream TOF and  downstream TOF are slightly different:  200 mV for the upstream PMTs but only 50 mV for downstream PMTs.
-
 The time of the pulses was calculated utilizing an oversampled template derived from the data itself.
-The pulse pedestal was taken from samples far from the pulse and subtracted to the pulse amplitude.
+We take the pulse pedestal from samples far from the pulse and subtract it to the pulse amplitude.
 We then stretch vertically a template to match the pedestal-subtracted pulse amplitude and we move it horizontally to find the time.
@@
 However, a time spread of approximately 300~ps is present in both the upstream and downstream detectors, likely due to transit time jitter in the PMTs themselves.
-There is no evidence of systematic timing drift over long periods such as 3-4 months of a data-taking: the maximum variation of the average time differences between pairs of PMTs reading out the same scintillator is of the order of 150~ps.
+There is no evidence of systematic timing drift over long data-taking periods such as 3-4 months: the maximum variation of the average time differences between pairs of PMTs reading out the same scintillator is of the order of 150~ps.
 
-\section{In the Cryostat}
-
-\subsection{TPC: Charge Collection}\label{sec:TPC}
-\subsection{TPC: Light Collection System}
-\subsection{Cryogenics and Purity Control}
-\subsection{TPC: Electric Field Measurement}
-\section{Trigger and DAQ}
-\section{Control Systems}
-
-\begin{comment}
+\textcolor{red}{calculated TOF with error}
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Punch-Through and Muon Range Stack Instruments}\label{sec:MuRS}
+\subsection{Punch-Through and Muon Range Stack Instruments}\label{sec:MuRS}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+The punch-thorough and the muon range stack (MuRS) detectors are located downstream of the TPC.
+These detectors provide a sample of  TPC crossing tracks without relying on TPC information and can be used to improve particle ID for  muons and pions with momentum higher than  450 MeV/c.
+
+The punch-thorough is simply a \textcolor{red}{? x ? x ?}~cm scintillator piece, read out by \textcolor{red}{ two? Hamamatsu? 2? inch} PMTs. 
+The MuRS is a segmented block of steel with four slots instrumented with scintillation bars.
+The four steel layers in front of each instrumented slot are 2 cm, 2 cm, 14 cm and 16 cm wide in the beam direction.
+Each instrumented slot is equipped with four scintillation bars each, positioned vertically in the direction orthogonal to the beam.
+Each scintillator bar measures  \textcolor{red}{?
+x ?
+x 2}~cm and it is read out by  \textcolor{red}{ two?
+Hamamatsu?
+2?
+inch} PMTs.
+
+The signals from both the punch-thorough and the MuRS PMTs are digitized in the CAEN V1740, same as the TPC; the details of this discriminator are laid out in~\ref{sec:TPCCharge}.
+It is worth noticing that the sampling time of the CAEN V1740 is slow (of the order of 128 ns), so pulse shape information from the PMT is lost.
+Punch-thorough and MuRS hits are formed utilizing the OR between the PMTs digital discriminator signals under threshold at a given time, where we obtain the threshold for each PMT directly on data distributions.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+Besides on beam data, LArIAT also triggers on cosmic rays events by using two sets of cosmic ray paddle detectors (a.k.a.
+``cosmic towers".) The cosmic towers frame the LArIAT cryostat, as one sits in the downstream left corner and the other sits in the upstream right corner of the cryostat.
+Two paddle sets of four scintillators pieces each, an upper and a lower set, make up each cosmic tower.
+Of the four paddles, a couple of two matched paddles stands upright while the a second matched pair lies across the top of the assembly in the top sets (or across the bottom of the assembly in the bottom sets).
+The horizontal couple is used as a veto for particles traveling from the TPC out.
+The four signals  from the vertical paddles along one of the body diagonals of the TPC are combined in a logical ``AND''.
+This allows to select cosmic muons crossing the TPC along one of its diagonals.
+Cosmic ray tracks crossing both anode and cathode populate the events triggered this way.
+This particularly useful sample of tracks (which we can safely assume to be associated with ~5 GeV muons MIPs) can be used for many tasks; for example, we use anode-cathode piercing tracks to cross check the TPC electric field on data (see \ref{ch:AppendixC}), to calibrate the charge response of the TPC wires for the full TPC volume and to measure the electron lifetime in the chamber \textcolor{red}{ ADD reference to different chapters}.
 
-LArIAT's system to trigger on cosmic rays is based on two so-called "cosmic towers" which stand  upstream and downstream of the cryostat, one on beam's right and one on beam's left, framing the cryostat.
-Each cosmic tower is composed of two paddle assemblies, upper and lower.
-The paddle assemblies each consist of four paddles, a matched pair which stand upright and a second matched pair lying across the top  of the assembly, to act as a veto for downward-going cosmic ray air showers.
-Unless vetoes by the horizontal paddles, signals from paddle assemblies along the body diagonals of the TPC are combined in a logical ``AND'' to select cosmic muons crossing the TPC along one of its diagonals.
-A high proportion of events triggered this way contain cosmic ray tracks crossing both anode and cathode.
-Such tracks provide a sample of LAr ionization with effectively uniform linear ionization density, but experience the entire range of charge attenuation available in the TPC before they drift to the anode.
-These tracks are used to calculate and monitor the level of electronegative contaminants in the liquid argon and provide a calibration sample for calorimetry and electric field studies outlined in Section \ref{sec:DetectorPerformance}.
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 The paddles have trapezoidal shape and are each enclosed in an aluminum case.
@@
 
-%$\varepsilon_P$ at a given voltage is defined as:
+The trigger rate $R$ of the whole system is $R=0.032Hz$, corresponding to $\sim 1.9~\mu/minute$.
 
-%$$
-%\varepsilon_P = \frac{R_{1-2} \land R_{3-4} \land R_P}{R_{1-2} \land R_{3-4}}
-%$$
-
-%where $R_P$ is the rate of the paddle $P$ to be tested, $R_{1-2}$ is the coincidence rate of two sample paddles positioned above $P$, while $R_{3-4}$ is the coincidence rate of two sample paddles positioned below $P$.
-%$\eta_P$ at a given voltage is instead defined as:
-
-%$$
-%\eta_P = \frac{R_{1-2} \land R_{3-4} \land R_{P^*}}{R_{1-2} \land R_{3-4}}
-%$$
-
-%where $R_{P^*}$ is the Rate of the paddle $P$
-
-The trigger rate $R$ of the whole system is $R=0.032Hz$, corresponding to $\sim 1.9~\mu/minute$.
-\end{comment}
+\section{In the Cryostat}
+\subsection{TPC: Charge Collection}\label{sec:TPCCharge}
+\subsection{TPC: Light Collection System}\label{sec:TPCLight}
+\subsection{Cryogenics and Purity Control}
+%\subsection{TPC: Electric Field Measurement}
+\section{Trigger and DAQ}
+\section{Control Systems}

### 2018-03-13 Chapter-2/2-TheLArIATExperiment.tex
s

@@
 The downstream paddle dimensions are  14 x 14 x \textcolor{red}{?}~cm and it is read out by two \textcolor{red}{Hamamatsu 2 inch} PMTs on the opposite ends of the scintillator.
-The relatively thin width on the beamline direction minimizes energy loss of the particles coming from the target in the scintillatior material.
+The relatively thin width on the beamline direction minimizes energy loss of the particles coming from the target in the scintillator material.
 
-\begin{figure}[!h]
-\begin{centering}
-\vspace{-0.3cm}
-%\includegraphics[height=2.3in]{figures/tofdelay.png}
-\caption{
-{\scriptsize \sf Pictures of the TOF system as was deployed during Run-I and Run-II data taking. The left image is of the upstream TOF paddle and the right image is of the downstream TOF paddle }
-}
-\label{fig:TOFSystemRunIandII}
-\end{centering}
-\end{figure}
+%\begin{figure}[!h]
+%\begin{centering}
+%\vspace{-0.3cm}
+%\includegraphics[height=2.3in]{Chapter-2/Images/tofdelay.png}
+%\caption{
+%{\scriptsize \sf Pictures of the TOF system as was deployed during Run-I and Run-II data taking. The left image is of the upstream TOF paddle and the right image is of the downstream TOF paddle }
+%}
+%\label{fig:TOFSystemRunIandII}
+%\end{centering}
+%\end{figure}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-The paddles have trapezoidal shape and are each enclosed in an aluminum case.
-An example is shown in Fig.
-\ref{pic:cosmicpaddle}.
-They come in two different sizes: a smaller version, with bases $32.2~cm$ and $26.7~cm$, $61.0~cm$ height and $3.02~cm$ thickness, and a bigger version, with bases $33.2~cm$ and $27.0~cm$, $70.8~cm$ height and $3.02~cm$ thickness.
-Each paddle is equipped with wavelength-shifting optical fibers running along one of the long sides and optically coupled to a low voltage, Zener-diode Hamamatsu H5783 PMT.
-Signals from the PMT are amplified and discriminated by a custom-made PMT Amplifier and Discrimination (PAD) circuit mounted at one end of the paddle, then sent through a CAT5 cable to a Control and Concentrator Unit (CCU) which both power the PMT, controlling voltage and threshold, and output the PMT signal as logic ECL pulse.
+All the paddles are $3.02~cm$ thick and are  trapezoidal in shape.
+The paddles come in two sizes: the smaller version has bases $32.2~cm$ and $26.7~cm$, and $61.0~cm$ height, while the bigger version has bases $33.2~cm$ and $27.0~cm$, and $70.8~cm$ height.
+A Zener-diode Hamamatsu H5783 PMT collects the light from a wavelength-shifting optical fiber which runs along one of the long sides of each paddle.
+A custom-made PMT Amplifier and Discrimination (PAD) circuit mounted at one end of the paddle collects signals from the PMTs and sends them to the Control and Concentrator Unit (CCU).
+We use the same connection to  power the PMT, control voltage and threshold, and output the PMT signal as logic ECL pulse.
+We retrieved the scintillation paddles from the decommissioning of the CDF detector at Fermilab and we used only the paddles with a counting efficiency greater than 95\% and low noise at working voltage.
+The measured trigger rate of the whole system is $0.032Hz$, corresponding to $\sim 2$ muons per minute.
 
@@
 \centering
-% \includegraphics[angle=90,width=0.7\textwidth]{figures/Cosmic_Paddle.jpg}
-\caption{
-Photograph of one of the scintillation counters used in the cosmic towers. 
-} 
+ \includegraphics[angle=90,width=0.7\textwidth]{Chapter-2/Images/Cosmic_Paddle.jpg}
+\caption{Photograph of one of the scintillation counters used in the cosmic towers. } 
 \label{pic:cosmicpaddle}
 \end{figure}
-
-The paddles were selected from a pool of over 300 scintillating counters collected during the decommissioning of the CDF detector at Fermilab. 
-For each counter, both the efficiency $\varepsilon_P$ and the noise $\eta_P$ as a function of the voltage were determined.
-The measurement was performed sandwiching the given paddle among 4 sample counters, placed two above and two below the paddle under test.
-The efficiency
-$\varepsilon_P$ was defined as the ratio between the 5-fold coincidence and the 4-fold coincidence of the sample counters.
-The accidental rate $\eta_P$ was instead defined as the number of 5-fold coincidence observed during ten minutes of data acquisition, when the signal of the paddle under test was delayed by $5 \mu s$.
-Each paddle with $\varepsilon_P \geq 95\%$ and $\eta_P = 0$ at working voltage was identified as a candidate for the Cosmic Ray system.
-The ones with the highest efficiency and lowest single count rate were then selected to realize the system itself.
-The plot of $\varepsilon_P$ as a function of the PMT voltage for one example paddle is shown in Fig. \ref{pic:CR_Effplot}.
-
-\begin{figure}[h!]
- \centering
-% \includegraphics[width=0.7\textwidth]{figures/TSU29_Efficiency.png}
-\caption{
-Plot of the counter efficiency $\varepsilon_P$ as a function of the PMT voltage for one of the paddles composing the Cosmic Ray system. 
-} 
-\label{pic:CR_Effplot}
-\end{figure}
-
-The trigger rate $R$ of the whole system is $R=0.032Hz$, corresponding to $\sim 1.9~\mu/minute$.
 
@@
 \subsection{TPC: Light Collection System}\label{sec:TPCLight}
+The mechanism of particle detection in argon other than drift electrons is the collection of scintillation photons.
+Over the course of LArIAT three years of data taking, the light collection system changed several times.
+We describe here the light collection system for Run II.
+Two PMTs, a 3-inch diameter Hamamatsu R-11065 and 2-inch diameter ETL D757KFL~\cite{lightsys-pmttests}, as well as three SiPMs arrays (two Hamamatsu S11828-3344M 4x4 arrays and one single-channel SensL MicroFB-60035 ) are mounted on the PEEK support structure.
+PEEK screws into an access flange as shown in Figure~\ref{lightsys_pmts}, on the anode side, leaving  approximately 5~cm of clearance from the collection plane.
+
+%------------------------------------------
+\begin{figure}
+\centering
+\includegraphics[height=2.2in]{Chapter-2/Images/lightsys_pmts.png}
+\hspace{1cm}
+\includegraphics[height=2.2in]{Chapter-2/Images/lightsys_wls.png}
+\caption{LArIAT's photodetector system for observing LAr scintillation light inside the TPC (left), and a simplified schematic of VUV light being wavelength-shifting along the TPB-coated reflecting foils (right).}
+\label{lightsys_pmts}
+\end{figure}
+\begin{figure}
+\centering
+\includegraphics[height=0.25\textheight]{Chapter-2/Images/lightsys_etlbase.jpeg}
+\hspace{0.5cm}
+\includegraphics[height=0.25\textheight]{Chapter-2/Images/lightsys_hmmbase.jpg}
+\caption{\label{voltagedividers}Photos of the voltage divider bases for the ETL PMT (left) and the Hamamatsu PMT (right) used in Run-II.
+The cable connections to the bases seen here were used for powering and testing prior to installation.
+The yellow through-hole signal coupling capacitors seen on both bases are 18~nF (X7R) and are rated to 2~kV.}
+\end{figure}
+%------------------------------------------
+Liquid argon scintillates in vacuum-ultraviolet (VUV) range at 128 nm; since cryogenic PMTs are not sensitive to VUV wavelengths, we need to shift the light in a region visible to the PMTs.
+In LArIAT, the wavelength shifting is achieved by installing on the four walls of the TPC highly-reflective VIKUITY dielectric substrate foils coated with a thin layer of tetraphenyl-butadiene (TPB) (see Figure ~\ref{lightsys_foils}).
+One or more visible photons  are emitted and reflected into the chamber during the interaction of a VUV photon interacts with the TPB.
+Thus, the light yield is increased and made more uniform across the TPC active volume, allowing the possibility of light-based calorimetry (under study).
+
+%------------------------------------------
+%\begin{figure}
+%\centering
+%\includegraphics[scale=1.3]{Chapter-2/Images/lightsys_foils.png}
+%\caption{\label{lightsys_foils}The TPB-coated reflector foils mounted to the TPC field cage walls as viewed through the front cryostat opening. {\textcolor{red}{Better pictures needed.}}}
+%\end{figure}
+%------------------------------------------
+
+For Run II, we coated both  the windows of the ETL PMT and SensL SiPM  with a thin layer of TPB.
+In doing so, some of the VUV scintillation light converts into visible right at the sensor faces, keeping information on the direction of the light source.
+Information about the light directionality is lost for light reflected on foils, as the reflection is uniform in angle.
+For Run-II, the voltage dividers for the PMTs were configured for positive bias with a DC-coupled anode (AC-coupled anode with grounded photocathode) to minimize induced noise on the TPC wires and modify the PMT bases accordingly.
+
 \subsection{Cryogenics and Purity Control}

### 2018-03-13 Chapter-2/2-TheLArIATExperiment.tex
s

@@
 \chapter{LArIAT: Liquid Argon In A Testbeam}\label{sec:experimentDescription}
-\section{LArIAT \& the Intensity Frontier}
+
+LArIAT, a small Liquid Argon Time Projection Chamber (LArTPC) in a test beam,  is designed to perform an extensive physics campaign centered on charged particle cross section measurements while characterizing the detector performance for future LArTPCs.
+LArTPC represents one of the most advanced experimental technologies for physics at the Intensity Frontier due to its full 3D-imaging, excellent particle identification and precise calorimetric energy reconstruction.
+This complex technology however needs a thorough calibration and dedicated measurements of some key quantities to achieve the precision required for the next generation of discoveries at the Intensity Frontier which LArIAT can provide.
+
+The LArIAT LArTPC is deployed in a dedicated calibration test beamline at Fermilab.
+We use the LArIAT beamline to characterize the charge particles before they enter the TPC: the particle type and initial momentum is known from beamline information.
+The precise calorimetric energy reconstruction of the LArTPC technology enables the measurement of the total differential cross section for  tagged hadrons.
+The Pion-Nucleus and Kaon-Nucleus total hadronic interaction cross section have never been measured before in argon and they are a fundamental step to shed light on light meson interaction in nuclei.
+Additionally, these measures provides a key input to neutrino physics and proton decay studies in future LArTPC experiments like SBN and DUNE.
+
+%\section{LArIAT \& the Intensity Frontier}
 \section{The Particles Path to LArIAT}
 
-LArIAT's home at Fermilab is the Fermilab Test Beam Facility (FTBF), where the experiment 
-characterizes a beam of charge particles downstream from the Meson Center beam line.
+LArIAT's home at Fermilab is the Fermilab Test Beam Facility (FTBF), where the experiment characterizes a beam of charge particles downstream from the Meson Center beam line.
 
@@
 
-\subsection{Cryogenics and Purity Control}
-%\subsection{TPC: Electric Field Measurement}
+\subsection{Cryogenics and Purity Control}\label{ch:Cryo}
+LArIAT repurposed the ArgoNeuT cryostat \cite{argoneut} in order to use it in a beam of charge particles. We also added a new process piping and a new liquid argon filtration system in FTBF.
+Inside the LArIAT experimental hall, the cryostat sits on the beam of charge particles with its horizontal main axis oriented parallel to the beam.
+
+Two volumes make up LArIAT cryostat, shown in Figure \ref{fig:LArIATCryoStat}: purified liquid argon fills the inner vessel, while the outer volume insulates it with a vacuum jacket equipped with layers of aluminized mylar superinsulation.
+The inner vessel is a cylinder of 130~cm length and 6.2~cm diameter, containing about 550~L of LAr, corresponding to a mass of 0.76 ton.
+We run the signal cables for the LArTPC and the high voltage feedthrough through a ``chimney'' at the top and mid-length of the cryostat.
+
+\begin{figure}[htb]
+\centering
+\includegraphics[scale=0.18]{Chapter-2/Images/Cryostat1.jpg}
+\includegraphics[scale=0.07]{Chapter-2/Images/Cryostat2.jpg}
+\caption{\emph{(left)} The LArIAT cryostat open with the TPC placed in the inner volume.
+\emph{(right)} The LArIAT cryostat fully sealed during initial commissioning prior to installation at Fermilab Testbeam Facility.Access to the internal volume is possible by opening the upstream end caps of the inner and outer vessels.
+}
+\label{fig:LArIATCryoStat}
+\end{figure}
+
+Given the different scopes of the ArgoNeuT and LArIAT detectors, we made several modification to the ArgoNeuT cryostat in order to use it in LArIAT.
+In particular, the modification  shown in Figure \ref{fig:LArIATCryoMods} were necessary to account for the beam of charged particles entering the TPC and to employ the new FTBT liquid argon purification system.
+We added a ``beam window'' on the front outer end cap and an ``excluder'' on the inner endcap, with the scope of minimizing the amount of dead material upstream of the TPC's active volume.
+Doing so, we reduced the amount of uninstrumented material before the TPC from $\sim$ 1.6 radiation lengths ($X_{0}$) (ArgoNeuT) to less than 0.3 $X_{0}$ (LArIAT).
+To allow studies of the scintillation light, we added a side port feedthrough which enables the mounting of the light collection system, as well as the connections for the corresponding signal and high-voltage cables (see Section \ref{sec:TPCLight}).
+We modified the bottom of the cryostat adding Conflat and ISO flange sealing to connect the liquid argon transfer line to the new argon cooling and purification system.
+
+\begin{figure}[htb]
+\centering
+\includegraphics[scale=0.35]{Chapter-2/Images/CryoMods.png}
+\caption{Pictures of the modified components of the cryostat.
+$1)$ The addition of an outlet to the bottom of the cryostat to allow connections to the purification system; $2)$ The ``beam-window'' on the outer endcap and the concave inner surface of the inner endcap (referred to as the excluder) to reduce the amount of material through which beam particles must travel before entering the TPC; $3)$ The modified side port for the LArIAT light collection system.}
+\label{fig:LArIATCryoMods}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%
+As in any other LArTPC, argon purity is a crucial parameter for LArIAT.
+Indeed, the presence of contaminants effects both the basic working principles of a LArTPC: electronegative contaminants such as oxygen and water decrease the number of ionization electrons collected on the wires after drifting through the volume, while contaminants such as Nitrogen decrease the light yield from scintillation light, especially in its slow component.
+
+The argon path
+
+LArIAT argon is less than  1~ppm of oxygen, water and nitrogen contaminants.
+monitor with the use of  commercial gas analyzer.
+LArIAT makes use of LAr supplies from a vendor
+
+LArIAT is supplied with liquid argon from a commercial dewar positioned outside of the experimental enclosure.  
+Liquid argon is delivered by a vendor with a specified maximum contamination of 2 parts per million (ppm) oxygen, 3.5~ppm water, and 10~ppm nitrogen which we
+
+In practice, the delivered argon has much less than 1~ppm of all these contaminants.
+Even such apparently low levels of contamination are too large to effectively drift ionization electrons, so filtration of the argon is required to achieve oxygen and water contamination at the level of 100 parts per trillion (ppt) or better.
+
+The argon is delivered from the commercial dewar to the cryostat through 2.54~cm diameter schedule 10 stainless steel piping.
+The piping was insulated with 20.32~cm of polyurethane foam by the manufacturer.
+The piping was cleaned to remove oil and grease before being welded into the system.
+The argon then passes through LArIAT's argon filtration system which is based on the Liquid Argon Purity Demonstrator (LAPD)~\cite{LAPD} design.
+The purification system consists of a single 77~liter filter which is filled halfway with a 4A molecular sieve supplied by Sigma-Aldrich~\cite{sigma-aldrich} that primarily removes water contamination but can also remove small amounts of nitrogen and oxygen.
+The remaining volume of the filter contains BASF~CU-0226~S, a highly dispersed copper oxide impregnated on a high surface area alumina, to remove oxygen~\cite{basf} and to a lesser extent, water.
+The filter is insulated with a vacuum jacket and aluminum radiation shields.
+The filter media are regenerated in place using heated gas, as was done for LAPD.
+The filter media are very efficient at removing oxygen and water and the argon is pure enough after a single pass through the media to allow several millisecond electron drift lifetimes in the TPC.
+
+From the filter the argon is directed into the inner cryostat via a liquid feedthrough on the top of the cryostat.
+The argon level, temperatures, and pressures are continuously monitored during operation both in the commercial dewar supplying the argon as well as the levels inside the cryostat, as shown on the top of Figure \ref{fig:LArIATCryoMonitor}.
+The argon in the cryostat is allowed to boil and vent to the atmosphere during operation and the argon level is monitored via the level probes.
+Flow of argon into the cryostat is enabled whenever the liquid level goes below a predetermined value to ensure that the TPC high voltage feedthrough and cold electronics are always submerged.
+During normal operations, the liquid level inside the cryostat is replenished several times per day, as can be seen by the cycle of the liquid level and valve positions shown at the bottom of Figure \ref{fig:LArIATCryoMonitor}
+
+\begin{figure}[htb]
+\centering
+\includegraphics[scale=0.35]{Chapter-2/Images/CryoMonitor.png}
+\includegraphics[scale=0.34]{.Chapter-2/Images/CryoMonitor2.png}
+\caption{\emph{(top)} Screenshot of the LArIAT cyrosystem monitoring page showing the levels of the argon both inside the cryostat and in the supply dewar as well as the monitored levels plotted over a twenty-four hour period.
+\emph{(bottom)} Plots showing the liquid argon level inside the cryostat as well as the corresponding liquid valve which allows argon to flow into the cryostat and the level drops due to boiling.
+The frequency of the typical fill/vent cycle can be seen with the typical time between fills being slightly more than 3 hours.}
+\label{fig:LArIATCryoMonitor}
+\end{figure}
+
 \section{Trigger and DAQ}

### 2018-03-13 Drafting.tex
s

@@
-\section{Non-LArTPC Instruments}\label{sec:FTBF}
-The Fermilab Test Beam Facility (FTBF) receives a 120~GeV$/c$ primary proton beam with variable intensity, split by electrostatic septa from beam circulating in the Main Injector.
-Each supercycle of the Fermilab accelerator complex, roughly 60 seconds in duration, includes a four-second "spill" of primary beam to FTBF.
-Primary beam is sent to either or both of two experimental halls, named Meson Test and Meson Center.
-
-The LArIAT experiment utilized the Meson Center beam line and had this primary beam focused onto a tungsten target.
-The resulting secondary charged particle beam is selected to have peak momentum between 8-80\,GeV$/c$ and is transported to the LArIAT experimental hall, designated MC7, where it is focused onto a copper target.
-This target, and the steel collimator hosting it, form the beginning of the tertiary particle beam and the LArIAT apparatus shown in  Fig.~\ref{fig:tert-layout}.
-For most of the LArIAT data taking this secondary beam peak momentum was fixed at 64~GeV$/c$, although some lower secondary beam momentum settings were explored.
-With such an energetic secondary beam on it, the copper target produces a tertiary beam which is comprised of mostly pions and protons with a small fraction of electrons, muons, and kaons present as well.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{LArIAT Tertiary Beam}\label{sec:TertiaryBeam}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-The tertiary beam consists of a target and collimator system and two bending magnets in a similar configuration used for the  MINERvA T-977 test beam calibration~\cite{MinervaTestbeam}.
-A set of four wire chambers and two time-of-flight scintillating paddles, also shown in Fig.~\ref{fig:tert-layout}, provide tracking, momentum determination, and particle identification (PID).
-
-The geometry of the tertiary beam line in MC7 has been optimized for LArIAT.
-A $13^\circ$ production angle at the target and its enclosing collimator, as well as a $10^\circ$ bend through a pair of dipole magnets provide particle momenta spectra tunable in the range from 0.2 to 2.0~GeV/c, as a function of the field intensity in the magnets.
-
-\begin{figure}[!h]
-\begin{centering}
-\vspace{-0.3cm}
-\includegraphics[height=5.3in]{figures/LArIATBeamline.png}
-\caption{
-{\scriptsize \sf The tertiary beam line layout. Upstream and downstream collimators are in gray, bending magnets in yellow, wire chambers are in red, and the liquid argon TPC volume in green.}
-}
-\label{fig:tert-layout}
-\end{centering}
-\end{figure}
-
-In the subsequent sections, we will detail the various subsystems installed along the beamline
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Time-of-Flight System}\label{sec:TOF}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-The LArIAT time-of-flight (TOF) detector system consists broadly consists of two scintillator paddles, which bracket the beamline and are shown in blue in Figure~\ref{fig:tert-layout}.
-The exact details of the of the TOF system were changed between Run-II and Run-III to allow for better performance and timing resolution.
-
-\paragraph{\textbf{TOF System for Run-I and Run-II:}}
-The upstream paddle, shown in Figure \ref{}, has a relatively small active area (10~cm x 6~cm ) and was chosen to be a relatively thin piece of scintillator to minimize any impact on the momentum of the particles coming from the target.
-Light guides were mounted on all four edges which lead to two \textit{insert PMT type} PMTs mounted on the beam left side.
-The downstream paddle, shown in Figure \ref{fig:TOFSystemRunIandII}, was chosen to have a slightly larger area (14~cm x 14~cm) and had two \textit{insert PMT type} PMTs which were read out on opposite ends of the paddle.
-
-\begin{figure}[!h]
-\begin{centering}
-\vspace{-0.3cm}
-%\includegraphics[height=2.3in]{figures/tofdelay.png}
-\caption{
-{\scriptsize \sf Pictures of the TOF system as was deployed during Run-I and Run-II data taking. The left image is of the upstream TOF paddle and the right image is of the downstream TOF paddle }
-}
-\label{fig:TOFSystemRunIandII}
-\end{centering}
-\end{figure}
-
-\paragraph{\textbf{TOF System for Run-III:}}
-Two modifications were made to the downstream TOF paddle for Run-III to improve the efficiency and timing of the system.
-The large area scintillator paddle was replaced with a slightly smaller paddle more similar to the upstream and was now readout with four new PMT's (2" Hamamatsu PMTs (R-1828) inherited from the muon g-2 experiment(E821) at BNL) located on the diagonals of the counter, as shown in Figure \ref{fig:TOFSystemRunIII}.
-
-\begin{figure}[!h]
-\begin{centering}
-\vspace{-0.3cm}
-\includegraphics[height=2.3in]{figures/TOFRunIII.png}
-\caption{
-{\scriptsize \sf Image of the modified downstream TOF system deployed during Run-III data taking }
-}
-\label{fig:TOFSystemRunIII}
-\end{centering}
-\end{figure}
-
-%The active area of the upstream paddle is relatively small ( 10~cm x 6~cm ) and that of the downstream paddle is somewhat larger ( 14~cmx 14~cm ).
-Lightguides are mounted on all four edges of each paddle.
-In the first two LARIAT production runs, each of the paddles was read out by two (xxkxkxkxk) photomultiplier tubes (PMTs), The long axis of the downstream paddle was directed horizontally and read out at either end.
-The upstream paddle was rotated by 45 degrees with respect to the horizontal and its two PMTs were mounted to the left side.
-In run 3, four more PMTs were added to the system.
-The upstream paddle was read out with the original set of PMTs and the downstream paddle with four 2" Hamamatsu PMTs (R-1828) inherited from the muon g-2 experiment(E821) at BNL.
-
-During data-taking cycles, signals from the TOF PMTs were sampled at 1 GHz
-with a CAEN 1751 digitizer and 12 bit samples were stored in a circular memory buffer.
-In response to an experimental trigger, a 28.7 $\mu$ second window of samples, starting approximately 8.4 $\mu$sec before the trigger, was written to the
-output.
-The amplitude of the TOF signals was typically 200 mV in the upstream TOF PMTs but only 50 mV in those of the downstream counter.
-The signals have a rise time (10-90\%) of 4 ns and a full width, half-maximum of 9 ns.
-The rate in the upstream counter was typically 15 kHz and much less, 400 Hz in the downstream counter.
-
-Because the shape of signals from each TOF PMT was highly uniform, the time of the pulses was determined using an oversampled template derived from the data itself.
-The DC offset (pulse pedestal) is taken from samples far from the pulse, The template is stretched vertically to match the pedestal-subtracted pulse amplitude and moved horizontally to match the time.
-The pulse time-pickoff resolution is better than 100 ps and relative amplitude resolution is better than x\%.
-Since the average rate of pulses incident on the ToF system is modest, pulse pile up is not a significant problem.
-Given the uniform width of the pulses produced by any given PMT (sigma of 400 ps), the pulse width can be used to flag events where two pulses overlap closely in time.
-Simulations indicate that for pulses as close as 4 ns, a well-chosen pulse width cut is 90\% efficient.
-
-To determine the particles' time of arrival, pulse times from the multiple PMTs mounted on each paddle had to be combined.
-During Run-I and II, with two PMTs mounted on each paddle, a simple average was used.
-In the case of the downstream paddle, with PMTs mounted at opposite ends, the average corrected for the fact that transit time to the two PMTs depended on the particle's point of impact on the scintillator.
-A plot of the time of arrival {\it difference} between the two downstream PMTs vs the longitudinal impact point, made with the help of track stubs from the wire chamber system, is shown in Fig.~\ref{fig:tofdelay}.
-As expected, the time difference
-changes systematically across the face of the scintillator.
-
-\begin{figure}[!h]
-\begin{centering}
-\vspace{-0.3cm}
-%\includegraphics[height=2.3in]{figures/tofdelay.png}
-\caption{
-{\scriptsize \sf $\Delta-t$ between times registered on two downstream paddles.}
-}
-\label{fig:tofdelay}
-\end{centering}
-\end{figure}
-
-Taking the average minimizes ToF errors arising from optical path differences in the scintillator, However, even for a set of particles which pass through a single small area of the paddle, the times of pulses registered in the two PMTs are still spread by approximately 300~ps, an uncertainty which is probably caused by transit time jitter in the PMTs themselves.
-This jitter is evident in both the upstream and downstream detectors.
-
-Although climate-control in the LArIAT experiment hall is fairly primitive, there is little sign of systematic timing drift over longer periods.
-The average time differences between pairs of PMTs reading out the same scintillator varied by no more than 150~ps over the 3-4 months of a data-taking period.
-
-Finally, the amplitude of the ToF pulses (peak minus pedestal) reflects the energy deposited in the scintillator, largely through ionization.
-At a given momentum, more massive particles such deuterons, protons, and kaons, will deposit more energy than less massive particles, such as electrons, muons, and pions.
-As described below, ToF and momentum information from the wire chambers can be combined to determine the particle's mass and hence its identity.
-Although the energy deposited in the scintillator isn't used to distinguish among the various particle types, it's clear from Figure~\ref{fig:tofamplitude} that the spectra of energy deposition for
-electrons, pions, and protons are very different.
-
-\begin{figure}[!h]
-\begin{centering}
-\vspace{-0.3cm}
-%\includegraphics[height=2.3in]{figures/tofamplitude.png}
-\caption{
-{\scriptsize \sf Average upstream ToF pulse amplitude for protons and pions/positrons.}
-}
-\label{fig:tofamplitude}
-\end{centering}
-\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Bending Magnets}\label{sec:Magnets}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-The pair of spectrometer magnets in the LArIAT tertiary beam are Fermilab type "NDB" electromagnets, originally used in the anti-proton ring.
-These magnets provide the necessary bend in the particle trajectory to allow for a momentum measurement to be performed.
-The magnets have a gap height of 14.224~cm, a gap width of 31.75~cm, and have an iron length of 46.67~cm.
-The aperture presented by the magnets is larger than that of the wire chambers ($\sim$12.5cm) and thus only the central part of the magnet is ever utilized.
-Over this limited aperture there is negligible variation of the field integral and thus a very uniform field is provided.
-The field intensity in one of the pair of magnets was measured using two Hall probes, both calibrated with NMR, to obtain the excitation curve shown in Figure~\ref{fig:magnet_excitation}).
-The second magnet, having been made to the same standard and having the same history, is assumed to have a very similar response.
-\begin{figure}[!h]
-\begin{centering}
-\vspace{-0.3cm}
-\includegraphics[height=3.0in]{figures/ExcitationCurves.png}
-\caption{
-{\scriptsize \sf Excitation Curve, B/I vs. I, for one NDB magnet, using two Hall probes (blue and green). Parameters from a cubic fit (black curve) to average of measurements (red) given in legend.}
-}
-\label{fig:magnet_excitation}
-\end{centering}
-\end{figure}
-
-These magnets were air- and water-cooled during operations and their temperatures monitored.
-Due to concerns of overheating, the maximum usable field at the center of the magnets is $\sim$0.35 Tesla with 100 Amps of current being passed through the magnets.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Multi-Wire Proportional Chambers}\label{sec:MWPC}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-The wire chambers are based on the Fenker Chambers~\cite{Fenker} long in use at Fermilab.
-The chambers have been upgraded by adding additional grounding to improve the signal to noise in the electronic readout.
-The chambers, shown in Figure~\ref{fig:wirechamber}, have an effective aperture of 128~mm in both horizontal and vertical distances.
-The wires are spaced at 1~mm, with 128 wires in each view.
-In a test beam, the chambers have typical efficiencies of 98\% to 99\%.
-The gas used is 85\% Argon + 15\% isobutane.
-The wire chambers typically operate between 2400 and 2500 volts.
-
-\begin{figure}[!h]
-\begin{centering}
-\vspace{-0.3cm}
-\includegraphics[height=2.3in]{figures/WireChamber.png}
-\caption{
-{\scriptsize \sf Image of one of the wire chambers used in the LArIAT tertiary beamline.}
-}
-\label{fig:wirechamber}
-\end{centering}
-\end{figure}
-
-The front end amplifier/discriminator uses the ASDQ chip~\cite{ASDQchip}.
-Two of the ASDQ chips are mounted on a mother board that is plugged into the chamber.
-Short flat cables are used to connect these ASDQ mother board to a new, multi-hit TDC~\cite{Sten}.
-This TDC provides a (maskable) fast OR output that can be used in a first level trigger.
-
-This TDC can accept multiple hits per wire with a time resolution of 1.18~ns/bin.
-The TDCs can accept 2 edges per 9~ns.
-The maximum event rate that can be accepted by the chamber system is of the order of one megahertz.
-These limitations are clearly not an issue for the LArIAT beamline where the rates and timing resolution required are much less stringent than this.
-There is enough memory on the TDC board to buffer a full spill of data (up to about 100 events/spill).
-As in this application the spills occur only once per minute, there is ample time to read out the whole spill through a controller specially designed to support these TDCs.
-Power and data flow to and from the controller to the TDCs and from the controller to the rest of the DAQ is via LVDS.
-The time window for acceptance for hits, time offsets, front end threshold, and pulse shaping parameters may be programmed through the controller via a USB link from a PC or through an Ethernet connection.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Aerogel Cherenkov Detectors}\label{sec:Aerogel}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-Aerogel is an ultra-light material in which the liquid component of the gel has been replaced with a gas.
-The result is a solid with extremely low density, low thermal conductivity, and low index of refraction.
-The goal of having an aerogel threshold Cherenkov detector in the LArIAT beam line is to separate muons and pions in the momentum range where muons emit Cherenkov radiation while pions do not.
-This technique was demonstrated by the MICE and Belle II experiments ~\cite{MICE-aerogel,BelleII-aerogel} and in LArIAT we use two aerogel threshold Cherenkov detector with index of refraction of 1.057 and 1.103.
-Having different indices of aerogel allows LArIAT to do this separation in two different momentum ranges of interest.
-
-\begin{figure}[htb]
-\centering
-\includegraphics[height=1.85in]{figures/Aerogel1103Figure.png}
-\hspace{1cm}
-\includegraphics[height=1.85in]{figures/Aerogel1057Figure.png}
-\caption{Number of photoelectrons per cm versus momentum for the two indices of refraction that are employed by the LArIAT aerogel detectors.
-The photoelectron yield includes the response of a bialkali photocathode following the recipe in ~\cite{bib8}.
-}
-\end{figure}
-
-%%% KEK couter
-The Aerogel Threshold Cherenkov Detector with the refractive index of 1.103 is placed just behind the second collimator on the tertiary beam line, between the two downstream wire chambers.
-Figure~\ref{fig:kek_aerogel} is a picture of this aerogel detector.
-It consists of 7 aerogel tiles (totaling 110 mm long in the beam direction), and the cross section of the detector is 108 mm $\times$ 108 mm.
-The array of the aerogel tiles is wrapped with a diffusion sheet and viewed by two PMTs (HAMAMATSU R329-02) from the sides.
-Figure~\ref{fig:kek_aerogel_ly} shows the observed light yield distribution for 550-800~MeV pions hitting the middle of the detector.
-The obtained light yield is equivalent to 22 photoelectrons for $\beta=1$.
-The fake rate for particles with velocity below the Cherenkov threshold is measured to be around 0.5\% with the threshold of 1.5 photoelectrons.
-
-\begin{figure}[htb]
-\centering
-\includegraphics[height=1.85in]{figures/kek_aerogel.jpg}
-\caption{The Aerogel Threshold Cherenkov Detector with the refractive index of 1.103.}
-\label{fig:kek_aerogel}
-\end{figure}
-
-\begin{figure}[h]
-\centering
-\includegraphics[height=1.85in]{figures/kek_aerogel_ly.png}
-\caption{The observed light yield distribution for 550-800 MeV pions hitting the middle of the detector.
-The obtained light yield is equivalent to 22 photoelectrons for $\beta=1$.}
-\label{fig:kek_aerogel_ly}
-\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Punch-Through and Muon Range Stack Instruments}\label{sec:MuRS}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\input{muon-Punch.tex}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{LArIAT Cosmic Ray Paddle Detectors}\label{sec:CosmicRayPaddle}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-LArIAT's system to trigger on cosmic rays is based on two so-called "cosmic towers" which stand  upstream and downstream of the cryostat, one on beam's right and one on beam's left, framing the cryostat.
-Each cosmic tower is composed of two paddle assemblies, upper and lower.
-The paddle assemblies each consist of four paddles, a matched pair which stand upright and a second matched pair lying across the top  of the assembly, to act as a veto for downward-going cosmic ray air showers.
-Unless vetoes by the horizontal paddles, signals from paddle assemblies along the body diagonals of the TPC are combined in a logical ``AND'' to select cosmic muons crossing the TPC along one of its diagonals.
-A high proportion of events triggered this way contain cosmic ray tracks crossing both anode and cathode.
-Such tracks provide a sample of LAr ionization with effectively uniform linear ionization density, but experience the entire range of charge attenuation available in the TPC before they drift to the anode.
-These tracks are used to calculate and monitor the level of electronegative contaminants in the liquid argon and provide a calibration sample for calorimetry and electric field studies outlined in Section \ref{sec:DetectorPerformance}.
-
-The paddles have trapezoidal shape and are each enclosed in an aluminum case.
-An example is shown in Fig.
-\ref{pic:cosmicpaddle}.
-They come in two different sizes: a smaller version, with bases $32.2~cm$ and $26.7~cm$, $61.0~cm$ height and $3.02~cm$ thickness, and a bigger version, with bases $33.2~cm$ and $27.0~cm$, $70.8~cm$ height and $3.02~cm$ thickness.
-Each paddle is equipped with wavelength-shifting optical fibers running along one of the long sides and optically coupled to a low voltage, Zener-diode Hamamatsu H5783 PMT.
-Signals from the PMT are amplified and discriminated by a custom-made PMT Amplifier and Discrimination (PAD) circuit mounted at one end of the paddle, then sent through a CAT5 cable to a Control and Concentrator Unit (CCU) which both power the PMT, controlling voltage and threshold, and output the PMT signal as logic ECL pulse.
-
-\begin{figure}[h!]
- \centering
- \includegraphics[angle=90,width=0.7\textwidth]{figures/Cosmic_Paddle.jpg}
-\caption{
-Photograph of one of the scintillation counters used in the cosmic towers. 
-} 
-\label{pic:cosmicpaddle}
-\end{figure}
-
-The paddles were selected from a pool of over 300 scintillating counters collected during the decommissioning of the CDF detector at Fermilab. 
-For each counter, both the efficiency $\varepsilon_P$ and the noise $\eta_P$ as a function of the voltage were determined.
-The measurement was performed sandwiching the given paddle among 4 sample counters, placed two above and two below the paddle under test.
-The efficiency
-$\varepsilon_P$ was defined as the ratio between the 5-fold coincidence and the 4-fold coincidence of the sample counters.
-The accidental rate $\eta_P$ was instead defined as the number of 5-fold coincidence observed during ten minutes of data acquisition, when the signal of the paddle under test was delayed by $5 \mu s$.
-Each paddle with $\varepsilon_P \geq 95\%$ and $\eta_P = 0$ at working voltage was identified as a candidate for the Cosmic Ray system.
-The ones with the highest efficiency and lowest single count rate were then selected to realize the system itself.
-The plot of $\varepsilon_P$ as a function of the PMT voltage for one example paddle is shown in Fig. \ref{pic:CR_Effplot}.
-
-\begin{figure}[h!]
- \centering
- \includegraphics[width=0.7\textwidth]{figures/TSU29_Efficiency.png}
-\caption{
-Plot of the counter efficiency $\varepsilon_P$ as a function of the PMT voltage for one of the paddles composing the Cosmic Ray system. 
-} 
-\label{pic:CR_Effplot}
-\end{figure}
-
-%$\varepsilon_P$ at a given voltage is defined as:
-
-%$$
-%\varepsilon_P = \frac{R_{1-2} \land R_{3-4} \land R_P}{R_{1-2} \land R_{3-4}}
-%$$
-
-%where $R_P$ is the rate of the paddle $P$ to be tested, $R_{1-2}$ is the coincidence rate of two sample paddles positioned above $P$, while $R_{3-4}$ is the coincidence rate of two sample paddles positioned below $P$.
-%$\eta_P$ at a given voltage is instead defined as:
-
-%$$
-%\eta_P = \frac{R_{1-2} \land R_{3-4} \land R_{P^*}}{R_{1-2} \land R_{3-4}}
-%$$
-
-%where $R_{P^*}$ is the Rate of the paddle $P$
-
-The trigger rate $R$ of the whole system is $R=0.032Hz$, corresponding to $\sim 1.9~\mu/minute$.
-
-\section{The LArIAT Detector}\label{sec:LArIATDetector}
-
-An Intro paragraph, as Will F.
-suggested, would be good here at the start of section 3.
-I might also rename this section to ?The LArIAT LArTPC Detector?, since beam line stuff is also ?The LArIAT Detector?.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Cryogenics}\label{sec:Cryo}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\input{CryoDetails.tex}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Time Projection Chamber}\label{sec:TPC}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\input{TPCDetails.tex}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{TPC Electronics}\label{sec:MSUElectronics}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\input{TPCelectronics.tex}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Digitization}\label{sec:Digitization}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\input{Digitization.tex}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Trigger and Readout System}\label{sec:Trigger}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\input{Trigger.tex}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Photon Detection System}\label{sec:PhotonSystem}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\input{LightDetectionSystem.tex}
-
-The LArIAT cryogenic system consists of the repurposed ArgoNeuT cryostat, the new process piping, and a liquid argon filtration system installed at Fermilab's Test Beam Facility.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{subsubsection}{Cryostat}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-The LArIAT cryostat, shown in Figure \ref{fig:LArIATCryoStat}, was repurposed from the ArgoNeuT experiment with a series of modifications to allow for operations in a charged particle test beam.
-The cryostat \cite{ArgoNeuT} consists of an inner volume containing the purified liquid argon and an outer volume serving as a vacuum jacket with layers of aluminized mylar superinsulation.
-Both the inner and outer cryostat vessels are cylindrical in shape with convex end caps.
-Access to the internal volume is possible by opening the upstream end caps of the inner and outer vessels.
-The main axis of the cryostat is horizontal and oriented parallel to the beam.
-The inner vessel is 76.2~cm in diameter and 130~cm in length, corresponding to a liquid argon volume of about 550~L, or a mass of 0.76 t.
-The cryostat has a wide neck, or ``chimney'', protruding from its top at mid-length which serves as an access path for signal cables from the LArTPC and the internal instrumentation, as well as for the high voltage feedthrough.
-
-\begin{figure}[htb]
-\centering
-\includegraphics[scale=0.18]{./figures/Cryostat1.jpg}
-\includegraphics[scale=0.07]{./figures/Cryostat2.jpg}
-\caption{\emph{(left)} The LArIAT cryostat open with the TPC placed in the inner volume.
-\emph{(right)} The LArIAT cryostat fully sealed during initial commissioning prior to installation at Fermilab Testbeam Facility.}
-\label{fig:LArIATCryoStat}
-\end{figure}
-
-Several modifications to the cryostat were made to meet the requirements of operating in a charged particle test beam and to facilitate the new liquid argon purification system.
-These modifications, shown in Figure \ref{fig:LArIATCryoMods}, include:
-\begin{enumerate}[label=\roman*),topsep=0pt,itemsep=-1ex,partopsep=1ex,parsep=1ex]
-\item A ``beam window'' on the front outer end cap and an ``excluder'' on the inner endcap.
-The purpose of these modifications is to minimize the amount of material upstream of the TPC's active volume along the beam line.
-On the original ArgoNeuT cryostat the stainless steel end caps of the cryostat's inner and outer vessels were 3/16'' (4.8 mm) thick.
-In addition, upstream of the TPC active volume there was a region of uninstrumented liquid argon approximately 15~cm in length.
-This corresponded to $\sim$ 1.6 radiation lengths ($X_{0}$) of dead material between the outside of the cryostat and the TPC, which would have impeded charged test beam particles prior to their entrance into the active volume.
-To mitigate this effect, the 9'' (22.9~cm) diameter beam window located at the center of the outer vessel's front flange was installed along with a blanked 0.4~mm titanium sheet.
-On the inner vessel's end cap, a hollow concave volume (the ``excluder'') was installed to reduce the amount of liquid argon charged particles would traverse before reaching the active volume of the TPC.
-With these modificaitons in place, the total material thickness seen by charged particles entering the front face of the LArIAT cryostat is now less than 0.3 $X_{0}$.
-\item A connection for a liquid argon transfer line to the new argon cooling and purification system.
-This modification consists of an outlet at the bottom of the cryostat with Conflat and ISO flange sealing.
-\item The addition of side port feedthroughs to accomodate signal and high-voltage bias connections for the scintillation light detection system (described in more detail in Section \ref{sec:PhotonSystem}).
-Both the inner and outer cryostat vessels had their existing side ports modified with CF-flanged apertures to accomplish this.
-\end{enumerate}
-
-%$i)$ A ``beam window'' on the front outer end cap and an ``excluder'' on the inner endcap to minimize the amount of material upstream of the TPC's active volume along the beam line; $ii)$ a connection at the base of the cryostat for the argon cooling and purification system; and $iii)$ the addition of a side port feedthrough to accomodate bias and signal connections for the scintillation light detectors.
-
-%On the original ArgoNeuT cryostat the stainless steel end caps of the cryostat's inner and outer vessels were 3/16'' (4.8 mm) thick.
-In addition, upstream of the TPC active volume there was a region of uninstrumented liquid argon approximately 15~cm in length.
-This corresponded to $\sim$ 1.6 radiation lengths ($X_{0}$) of dead material between the outside of the cryostat and the TPC, which would have impeded charged test beam particles prior to their entrance into the active volume.
-To minimize the effect of this dead region, a 9'' (22.9~cm) diameter beam window located at the center of the outer vessel's front flange was installed along with a blanked 0.4~mm titanium sheet.
-On the inner vessel's end cap, a hollow concave volume (the ``excluder'') was installed to reduce the amount of liquid argon charged particles would traverse before reaching the active volume of the TPC.
-With these modificaitons in place, the total material thickness seen by charged particles entering the front face of the LArIAT cryostat is now less than 0.3 $X_{0}$.
-
-%The cryostat vessels were modified to connect a liquid argon transfer line to the new cooling and purification system.
-This modification consists of an outlet at the bottom of the cryostat with Conflat and ISO flange sealing.
-
-%Finally, both the inner and outer cryostat vessels had their existing side ports modified with CF-flanged apertures for the signal and high-voltage bias feedthroughs for the scintillation light system, described in more detail in Section \ref{sec:PhotonSystem}.
-
-\begin{figure}[htb]
-\centering
-\includegraphics[scale=0.35]{./figures/CryoMods.png}
-\caption{Pictures of the modified components of the cryostat.
-$1)$ The addition of an outlet to the bottom of the cryostat to allow connections to the purification system; $2)$ The ``beam-window'' on the outer endcap and the concave inner surface of the inner endcap (referred to as the excluder) to reduce the amount of material through which beam particles must travel before entering the TPC; $3)$ The modified side port for the LArIAT light collection system.}
-\label{fig:LArIATCryoMods}
-\end{figure}
-
-\end{subsubsection}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{subsubsection}{Liquid Argon System}\label{sec:LArCryoSystem}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-LArIAT is supplied with liquid argon from a commercial dewar positioned outside of the experimental enclosure.
-Liquid argon is delivered by a vendor with a specified maximum contamination of 2 parts per million (ppm) oxygen, 3.5~ppm water, and 10~ppm nitrogen.
-Oxygen and water contamination is detrimental to the ability of the ionization electrons to drift through the volume, as free electrons readily attach to these molecules.
-Nitrogen contamination is detrimental to the collection of scintillation light.
-The level of these contaminants in the delivered argon is determined using a suite of commercial gas analyzers.
-In practice, the delivered argon has much less than 1~ppm of all these contaminants.
-Even such apparently low levels of contamination are too large to effectively drift ionization electrons, so filtration of the argon is required to achieve oxygen and water contamination at the level of 100 parts per trillion (ppt) or better.
-%{\bf 100 ppt is sufficient for a 3 ms lifetime, I think LArIAT has less than that}.
-
-The argon is delivered from the commercial dewar to the cryostat through 2.54~cm diameter schedule 10 stainless steel piping.
-The piping was insulated with 20.32~cm of polyurethane foam by the manufacturer.
-The piping was cleaned to remove oil and grease before being welded into the system.
-The argon then passes through LArIAT's argon filtration system which is based on the Liquid Argon Purity Demonstrator (LAPD)~\cite{LAPD} design.
-The purification system consists of a single 77~liter filter which is filled halfway with a 4A molecular sieve supplied by Sigma-Aldrich~\cite{sigma-aldrich} that primarily removes water contamination but can also remove small amounts of nitrogen and oxygen.
-The remaining volume of the filter contains BASF~CU-0226~S, a highly dispersed copper oxide impregnated on a high surface area alumina, to remove oxygen~\cite{basf} and to a lesser extent, water.
-The filter is insulated with a vacuum jacket and aluminum radiation shields.
-The filter media are regenerated in place using heated gas, as was done for LAPD.
-The filter media are very efficient at removing oxygen and water and the argon is pure enough after a single pass through the media to allow several millisecond electron drift lifetimes in the TPC.
-
-From the filter the argon is directed into the inner cryostat via a liquid feedthrough on the top of the cryostat.
-The argon level, temperatures, and pressures are continuously monitored during operation both in the commercial dewar supplying the argon as well as the levels inside the cryostat, as shown on the top of Figure \ref{fig:LArIATCryoMonitor}.
-The argon in the cryostat is allowed to boil and vent to the atmosphere during operation and the argon level is monitored via the level probes.
-Flow of argon into the cryostat is enabled whenever the liquid level goes below a predetermined value to ensure that the TPC high voltage feedthrough and cold electronics are always submerged.
-During normal operations, the liquid level inside the cryostat is replenished several times per day, as can be seen by the cycle of the liquid level and valve positions shown at the bottom of Figure \ref{fig:LArIATCryoMonitor}
-
-\begin{figure}[htb]
-\centering
-\includegraphics[scale=0.35]{./figures/CryoMonitor.png}
-\includegraphics[scale=0.34]{./figures/CryoMonitor2.png}
-\caption{\emph{(top)} Screenshot of the LArIAT cyrosystem monitoring page showing the levels of the argon both inside the cryostat and in the supply dewar as well as the monitored levels plotted over a twenty-four hour period.
-\emph{(bottom)} Plots showing the liquid argon level inside the cryostat as well as the corresponding liquid valve which allows argon to flow into the cryostat and the level drops due to boiling.
-The frequency of the typical fill/vent cycle can be seen with the typical time between fills being slightly more than 3 hours.}
-\label{fig:LArIATCryoMonitor}
-\end{figure}
-
-The temperature of the argon inside the cryostat can be determined through direct use of temperature probes deployed at the bottom, middle, and top of the cryostat (listed as TE213A, TE314A, and TE212A in Figure \ref{fig:LArIATCryoMonitor} respectively).
-This can be crossed check using the pressure in the cryostat gas volume and extrapoloting this to the center of the liquid volume.
-The pressure at the surface of the liquid argon is maintained by the cryosystem at $\sim$20~psi $\pm$.0.4 psi.
-
-In order to determine the temperature inside the TPC, we measure the pressure in the cryostat gas volume and calculate the pressure at the center of the TPC.
-From this pressure and the boiling point curve of argon, we calculate the temperature.
-The pressure at the surface of the liquid argon is maintained by the cryosystem at $\sim$20~psi as shown in Figure \ref{fig:pressure} for a period of 13 days with an average pressure of 20.0 $\pm$.0.4 psi.
-If we consider approximately 40 cm of argon from the ulage surface to the center of the TPC this would give the pressure there of $\sim$20.8 psi, which according to the Reference \cite{} corresponds to a temperature of 90.7 K.
-This is consistent with the readings given of the temperature probes, shown in Table \ref{tab:temp}.
-This cross-check both provides a basis for the electron mobility calculation used when calibrating the drift speed, as is discussed in Section \ref{}, as well as providing confidence in the monitoring tools.
-
-\begin{table}[h!]
-\centering
-\caption{Average temperatures measured at the top, middle and bottom in the LArIAT cryostat.}
-\label{tab:temp}
-\begin{tabular}{|c|c|c|}
-\hline
-Top Probe Temp (TE212A) & Middle Probe Temp (TE214A)   & Bottom Probe Temp (TE213A)  \\ \hline
-91.7 $\pm$ 0.7 K &  91.8 $\pm$ 0.7 K                   & 93.3 $\pm$ 2.7 K       \\ \hline
-\end{tabular}
-\end{table}
-
-\end{subsubsection}
 The Liquid Argon Time Projection Chamber (LArTPC) can be broken down into three major subcomponents: $1)$ The high voltage system which provides the drift field voltage; $2)$ the cathode and field cage which steps down the high voltage through a network of voltage-dividing resistors to form a uniform electric field for charges to drift within; and $3)$ the wire planes which provide the charge sensitive readout for the detector.
@@
 
-In addition to the drift electrons, LArIAT also collects liquid argon scintillation photons.
-This is achieved using several cryogenic photodetectors held in place by a PEEK support structure which screws into a side access flange as shown in Figure~\ref{lightsys_pmts}.
-Two PMTs, a 3-inch diameter Hamamatsu R-11065 and 2-inch diameter ETL D757KFL~\cite{lightsys-pmttests}, contribute to most of the system's light collection capacity.
-During Run-I and Run-II, several silicon photomultiplier (SiPM) arrays were also included as part of the photon detection system.
-These SiPMs consisted of two Hamamatsu S11828-3344M 4x4 arrays (each with 12x12mm$^2$ total active area) as well as one single-channel SensL MicroFB-60035 (6x6mm$^2$ active area) on custom preamplifier readout boards which were mounted along the edges of the PMT holder.
-The windows of all devices are held parallel to the wireplanes with approximately 5~cm of clearance.
-
-%------------------------------------------
-\begin{figure}
-\centering
-\includegraphics[height=2.2in]{figures/lightsys_pmts.png}
-\hspace{1cm}
-\includegraphics[height=2.2in]{figures/lightsys_wls.png}
-\caption{LArIAT's photodetector system for observing LAr scintillation light inside the TPC (left), and a simplified schematic of VUV light being wavelength-shifting along the TPB-coated reflecting foils (right).}
-\label{lightsys_pmts}
-\end{figure}
-\begin{figure}
-\centering
-\includegraphics[height=0.25\textheight]{figures/lightsys_etlbase.jpeg}
-\hspace{0.5cm}
-\includegraphics[height=0.25\textheight]{figures/lightsys_hmmbase.jpg}
-\caption{\label{voltagedividers}Photos of the voltage divider bases for the ETL PMT (left) and the Hamamatsu PMT (right) used in Run-II.
-The cable connections to the bases seen here were used for powering and testing prior to installation.
-The yellow through-hole signal coupling capacitors seen on both bases are 18~nF (X7R) and are rated to 2~kV.}
-\end{figure}
-%------------------------------------------
-
-Liquid argon scintillation light is produced in the vacuum-ultraviolet (VUV) range and must be shifted to visible before it can be detected by most photosensors.
-This wavelength-shifting occurs within a thin layer of tetraphenyl-butadiene (TPB) which coats highly-reflective VIKUITY dielectric substrate foils lining four walls of TPC.
-A photo of these foils mounted inside the TPC is shown in Figure~\ref{lightsys_foils}.
-A VUV photon interacts with the TPB to induce emission of one or more visible photons which are then emitted or reflected back into the active volume, increasing the probability of their detection by the photosystem~\cite{lightsys-szelc}.
-This technique also improves light yield uniformity across the TPC's active volume, a feature relevant to light-augmented calorimetry.
-
-%------------------------------------------
-\begin{figure}
-\centering
-\includegraphics[scale=1.3]{figures/lightsys_foils.png}
-\caption{\label{lightsys_foils}The TPB-coated reflector foils mounted to the TPC field cage walls as viewed through the front cryostat opening. {\textcolor{red}{Better pictures needed.}}}
-\end{figure}
-%------------------------------------------
-
-A TPB coating\footnote{The TPB coating was made using a solution of 100~mL toluene, 2~g polystyrene, and 800~mg of TPB prior to Run-II.
-The coating was re-applied to the ETL midway through Run-II using the same solution but with less TPB (600~mg).} was added to the windows of the ETL PMT and SensL SiPM prior to Run-II and reapplied about midway through the run.
-This allowed some of the VUV scintillation light to down-convert directly at their faces instead of at the coated reflector foils, rendering the photodetectors directionally sensitive to this light emanating from the point of energy deposition.
-These coatings were removed prior to Run-III.
-
-For Run-I, the voltage dividers for the PMTs were configured for negative bias with a DC-coupled anode, and were switched to positive bias mode (AC-coupled anode with grounded photocathode) for Run-II to minimize induced noise on wires near the charged photocathodes and outer PMT chasis.
-To configure each base for positive polarity before the start of Run-II, 18 nF (X7R) capacitors rated to 2kV were soldered onto the circuits to couple each PMT's anode to the outgoing signal line.
-At liquid argon temperature and PMT operating voltages, these capacitances reduce to $\sim$1/3 of their nominal value~\cite{lightsys-capacitors,lightsys-ubsplitting}.
-
-For Run-III, a prototype ARAPUCA~\cite{lightsys-arapuca} detector was installed, replacing the SiPM boards.
-The main purpose of the prototype is research and development of new alternatives for photosensors for liquid argon-based detectors.
-The central concept of ARAPUCA is to capture photons within a box with highly reflective internal surfaces (reflectivity~>~98~\% ), resulting in high photon detection efficiency even for limited active coverage, which is defined as the total area of photosensors viewing the internal volume.
-The key to the ARAPUCA's photon-trapping mechanism is the use of dichroic short-pass optical filters which are highly transparent to wavelengths below a particular cutoff and highly reflective above it.
-The filter is externally coated by a wavelength-shifting (WLS) film.
-The WLS film on the outer side of the filter converts the VUV scintillation light from the liquid argon to a wavelength below the cutoff, so the filter is transparent to the incident photons, allowing them to enter into the box.
-A second WLS coating covers the inner side of the filter or optionally the internal walls of the box.
-The internal WLS layer enables a photon to convert a second time to a wavelength greater than the cutoff condition for which the filter is highly reflective, trapping the photons inside the box.
-The interior of the ARAPUCA is viewed by light sensors which capture the trapped photons after few reflections.
-%covered by one or more SiPMs. The number of SiPMs can be increased and read out using an active ganging circuit. Figure \ref{arapuca-principle} illustrates the working principle of the device.
-
-The prototype ARAPUCA device installed in LArIAT, shown in Figure~\ref{fig:arapuca_photo}, is 4.5$\times$5.5 cm$^2$, with a total exposed filter area of 3.5$\times$4.5 cm$^2$.
-The WLS film on the outer surface of the filter is P-Terphenyl ($\lambda \sim  350$~nm), while the inner surface is coated in TPB ($\lambda \sim 430$ nm).
-It is equipped with a single SensL MicroFC 60035 SiPM biased at +24~V in liquid argon.
-Its signals are sent through a commercial preamplifier before being sent to the DAQ.
-
-%------------------------------------------
-\begin{figure}
-\centering
-\includegraphics[height=5cm]{figures/lightsys_arapuca2.jpg}
-\hspace{0.5cm}
-\includegraphics[height=5cm]{figures/lightsys_arapucaPmts.jpg}
-\caption{The prototype ARAPUCA device without its dichroic filter (left) as well as mounted to the PMT support structure after the installation of the filter (right).}
-\label{fig:arapuca_photo}
-
-\end{figure}
-%------------------------------------------
-
 \subsection*{Signal Digitization}

### 2018-03-13 Chapter-2/2-TheLArIATExperiment.tex
s

@@
 
-\subsection{Cryogenics and Purity Control}\label{ch:Cryo}
+\subsection{Cryogenics and Argon Purity}\label{ch:Cryo}
 LArIAT repurposed the ArgoNeuT cryostat \cite{argoneut} in order to use it in a beam of charge particles. We also added a new process piping and a new liquid argon filtration system in FTBF.
@@
 Indeed, the presence of contaminants effects both the basic working principles of a LArTPC: electronegative contaminants such as oxygen and water decrease the number of ionization electrons collected on the wires after drifting through the volume, while contaminants such as Nitrogen decrease the light yield from scintillation light, especially in its slow component.
-
-The argon path
-
-LArIAT argon is less than  1~ppm of oxygen, water and nitrogen contaminants.
-monitor with the use of  commercial gas analyzer.
-LArIAT makes use of LAr supplies from a vendor
-
-LArIAT is supplied with liquid argon from a commercial dewar positioned outside of the experimental enclosure.  
-Liquid argon is delivered by a vendor with a specified maximum contamination of 2 parts per million (ppm) oxygen, 3.5~ppm water, and 10~ppm nitrogen which we
-
-In practice, the delivered argon has much less than 1~ppm of all these contaminants.
-Even such apparently low levels of contamination are too large to effectively drift ionization electrons, so filtration of the argon is required to achieve oxygen and water contamination at the level of 100 parts per trillion (ppt) or better.
-
-The argon is delivered from the commercial dewar to the cryostat through 2.54~cm diameter schedule 10 stainless steel piping.
+In LArIAT, contaminations should not exceed the level of 100 parts per trillion (ppt).
+We achieve this level of purity in several stages.
+The specifics required for the commercial argon bought for LArIAT are 2 parts per million (ppm) oxygen, 3.5~ppm water, and 10~ppm nitrogen.
+This argon is monitored with the use of commercial gas analyzer.
+Argon is stored in a dewar external to LArIAT hall and filtered before filling the TPC.
+%The argon is delivered from the commercial dewar to the cryostat through 2.54~cm diameter schedule 10 stainless steel piping.
 The piping was insulated with 20.32~cm of polyurethane foam by the manufacturer.
 The piping was cleaned to remove oil and grease before being welded into the system.
-The argon then passes through LArIAT's argon filtration system which is based on the Liquid Argon Purity Demonstrator (LAPD)~\cite{LAPD} design.
-The purification system consists of a single 77~liter filter which is filled halfway with a 4A molecular sieve supplied by Sigma-Aldrich~\cite{sigma-aldrich} that primarily removes water contamination but can also remove small amounts of nitrogen and oxygen.
-The remaining volume of the filter contains BASF~CU-0226~S, a highly dispersed copper oxide impregnated on a high surface area alumina, to remove oxygen~\cite{basf} and to a lesser extent, water.
-The filter is insulated with a vacuum jacket and aluminum radiation shields.
-The filter media are regenerated in place using heated gas, as was done for LAPD.
-The filter media are very efficient at removing oxygen and water and the argon is pure enough after a single pass through the media to allow several millisecond electron drift lifetimes in the TPC.
-
-From the filter the argon is directed into the inner cryostat via a liquid feedthrough on the top of the cryostat.
-The argon level, temperatures, and pressures are continuously monitored during operation both in the commercial dewar supplying the argon as well as the levels inside the cryostat, as shown on the top of Figure \ref{fig:LArIATCryoMonitor}.
-The argon in the cryostat is allowed to boil and vent to the atmosphere during operation and the argon level is monitored via the level probes.
-Flow of argon into the cryostat is enabled whenever the liquid level goes below a predetermined value to ensure that the TPC high voltage feedthrough and cold electronics are always submerged.
-During normal operations, the liquid level inside the cryostat is replenished several times per day, as can be seen by the cycle of the liquid level and valve positions shown at the bottom of Figure \ref{fig:LArIATCryoMonitor}
-
-\begin{figure}[htb]
-\centering
-\includegraphics[scale=0.35]{Chapter-2/Images/CryoMonitor.png}
-\includegraphics[scale=0.34]{.Chapter-2/Images/CryoMonitor2.png}
-\caption{\emph{(top)} Screenshot of the LArIAT cyrosystem monitoring page showing the levels of the argon both inside the cryostat and in the supply dewar as well as the monitored levels plotted over a twenty-four hour period.
-\emph{(bottom)} Plots showing the liquid argon level inside the cryostat as well as the corresponding liquid valve which allows argon to flow into the cryostat and the level drops due to boiling.
-The frequency of the typical fill/vent cycle can be seen with the typical time between fills being slightly more than 3 hours.}
-\label{fig:LArIATCryoMonitor}
-\end{figure}
+LArIAT uses a filtration system designed for the Liquid Argon Purity Demonstrator (LAPD)~\cite{LAPD}: half of a 77~liter filter contains a 4A molecular sieve (Sigma-Aldrich~\cite{sigma-aldrich}) apt to remove mainly water, while the other half contains BASF~CU-0226~S, a highly dispersed copper oxide impregnated on a high surface area alumina, apt to remove mainly oxygen~\cite{basf}.
+A single pass of argon in the filter is sufficient to achieve the necessary purity, unless the filter is saturated.
+In case the filter saturates, the media needs to be regenerated by using heated gas; this happened twice during the Run II period.
+We deemed the filter regeneration necessary every time the electron lifetime dropped under 100 \textmu s.
+The filtered argon reaches the inner vessel via a liquid feedthrough on the top of the cryostat.
+Argon is not recirculated in the system, but it rather boils off and vent to the atmosphere.
+During data taking, we replenish the argon in the cryostat several times per day to keep the TPC high voltage feedthrough and cold electronics always submerged.
+In fact, we need to monitor the level, temperatures, and pressures  of the argon both in the commercial dewar and inside the cryostat.
 

### 2018-03-14 AppentixA-EField/A-LArIATEField.tex
s

@@
-\chapter{Measurement of LArIAT Electric Field}
+\chapter{Measurement of LArIAT Electric Field}{ch:AppendixA}
 The electric field of a LArTPC in the drift volume is a fundamental quantity for the proper functionality of this technology, as it affects almost every reconstructed quantity such as the position of hits or their collected charge.

### 2018-03-14 Chapter-2/2-TheLArIATExperiment.tex
s

@@
 The 12 bit samples are stored in a circular memory buffer.
-At trigger time, data from the TOF PMTs are recorded to output in a 28.7 $\mu$ second windows starting  approximately 8.4 $\mu$sec before the trigger time.
+At trigger time, data from the TOF PMTs are recorded to output in a 28.7 \textmu s windows starting  approximately 8.4 \textmu s before the trigger time.
 
@@
 \section{In the Cryostat}
-\subsection{TPC: Charge Collection}\label{sec:TPCCharge}
-\subsection{TPC: Light Collection System}\label{sec:TPCLight}
+\subsection{LArTPC: Charge Collection}\label{sec:TPCCharge}
+The LArIAT Liquid Argon Time Projection Chamber is a rectangular box of dimensions 47 cm (width) x 40 cm (height) x 90 cm (length), containing 170 liters of Liquid Argon.
+The LArTPC three major subcomponents are 
+\begin{itemize} 
+\item[1)] the cathode and field cage,
+\item [2)] the wire planes, 
+\item [3)] the read-out electronics. %
+\end{itemize}
+
+\subsubsection{Cathode and field cage}
+A G10 plain sheet with copper metallization on one of the 40x90~cm inner surfaces forms the cathode. 
+A high-voltage feedthrough on the top of the LArIAT cryostat delivers the high voltage to the cathode; scope of the high voltage system (Figure~\ref{fig:HVScheme}) is to drift ionization electrons from the interaction of charged particles in the liquid argon to the wire planes.
+The power supply used in this system is a Glassman LX125N16 ~\cite{GlassmanPS} capable of generating up to -125~kV and 16~mA of current, but operated at -23.5kV during LArIAT Run-II.
+The power supply is connected via high voltage cables to a series of filter pots before finally reaching the cathode.
+
+\begin{figure}[htb]
+\centering
+\includegraphics[scale=0.35]{Chapter-2/Images//HVSchematic.png}\\
+\caption{Schematic of the LArIAT high voltage system.}
+\label{fig:HVScheme}
+\end{figure}%See DocDB 1472
+
+The field cage is made of \textcolor{red}{13 parallel copper rings} framing the inner walls of the G10 TPC structure.
+A network of voltage-dividing resistors connected to the field cage rings steps down the high voltage from the cathode to form a uniform electric field.
+The electric field over the entire TPC drift volume is  486 V/cm (see \ref{ch:AppendixA}).
+The  maximum drift length, i.e.
+the distance between cathode and anode planes, is 47 cm.
+
+\subsubsection{Wire planes}
+The wire planes measure the charge deposited in the TPC active volume. The drifting charge induces a current on the wire of the inner planes and it is collected on the collection plane wires.
+LArIAT counts three wire planes separated by 4 mm spaces: in order of distance from the cathode, they are the shield, the induction and the collection plane.
+The distance between two consecutive wires in each given plane (aka wire pitch) is 4 mm.
+The shield plane counts 225 parallel wires of equal length oriented vertically.
+This plane is not connected with the read-out electronics; rather it shields the outer planes from extremely long induction signals due to the ionization chamber in the whole drift volume.
+As the shield plane acts almost like a Faraday cage, the shape of signals in the first instrumented plane (induction)  results easier to reconstruct.
+Both the induction and collection planes count 240 parallel wires of different length oriented at 60$^\circ$ from the vertical with opposite signs.
+Moving electrons moving past the induction plane will induce a bipolar pulse on its wires and will form a unipolar pulse when collected on the last plane wires.
+
+The three wire planes and the cathode form three drift volumes, as shown in Figure \ref{fig:driftregions}. 
+The main drift volume is defined as the region between the cathode plane and the shield plane (C-S).
+The other two drift regions are those between the shield plane and the induction plane (S-I), and between the induction plane and the collection plane (I-C).
+The electric field in these regions is chosen to satisfy the charge transparency condition to allow for 100$\%$ transmission of the drifting electrons through the shield and then the induction planes.
+
+\begin{figure}[htb]
+\centering
+\includegraphics[scale=0.35]{Chapter-2/Images/DriftRegions.png}\\
+\caption{Schematic of the three drift regions inside the LArIAT TPC: the main drift volume between the cathode and the shield plane (C-S) in green, the region between the shield plane and the induction plane (S-I) in purple, and the region between the induction plane and the collection plane (I-C) in pink.}
+\label{fig:driftregions}
+\end{figure}
+
+Table \ref{tab:voltages} provides the default voltages applied to the cathode and the shield, induction, and collection plane.
+
+\begin{table}[htpb]
+\centering
+\caption{Cathode and anode planes default voltages}
+\label{tab:voltages}
+\begin{tabular}{llll}
+\hline
+\multicolumn{1}{|l|}{ Cathode} & 
+\multicolumn{1}{|l|}{ Shield} & \multicolumn{1}{l|}{ Induction} & \multicolumn{1}{l|}{ Collection}  \\ \hline
+\multicolumn{1}{|l|}{-23.17 kV} &
+\multicolumn{1}{|l|}{-298.8 V} & \multicolumn{1}{l|}{-18.5 V}      & \multicolumn{1}{l|}{338.5 V} \\ \hline
+\end{tabular}
+\end{table}
+
+\subsubsection{Electronics}
+
+Dedicated electronics read the induction and collection plane wires, for a total of  480-channel analog signal path from the TPC wires to the signal digitizers.
+A digital control system for the TPC-mounted electronics, a power supply, and a distribution system complete the front-end system.
+Figure \ref{pic:FEelectronics} shows a block diagram of the overall system.
+The direct readout of the ionization electrons in liquid argon forms typically small signals on the wires, which needs to be amplified in oder to be processed.
+LArIAT  performs the amplification stage directly in cold with  amplifiers developed by Brookhaven National Lab (BNL) and mounted on the TPC frame inside the liquid argon, achieving a remarkable Signal-to-Noise ratio.
+%The BNL ASICs adopted in LAriAT are designated as LArASIC, version 4-star.%The ASIC signals for each wire are then driven out of the vessel  to DAQ boards that act as waveform recorders.
+The signal from the ASICs are driven to the other end of the readout chain, to the CAEN V1740 digitizers.
+The CAEN V1740 has a 12 bit resolution and a maximum input range of 2~VDC, resulting in about 180 ADC count for a crossing MIP.
+
+\begin{figure}[htbp]
+ \centering
+ \includegraphics[width=1.0\textwidth]{Chapter-2/Images/LArIAT_FE_Electronics.png}
+\caption{Overview of LArIAT Front End electronics. } 
+\label{pic:FEelectronics}
+\end{figure}
+
+\subsection{LArTPC: Light Collection System}\label{sec:TPCLight}
 The mechanism of particle detection in argon other than drift electrons is the collection of scintillation photons.
@@
 A single pass of argon in the filter is sufficient to achieve the necessary purity, unless the filter is saturated.
-In case the filter saturates, the media needs to be regenerated by using heated gas; this happened twice during the Run II period.
-We deemed the filter regeneration necessary every time the electron lifetime dropped under 100 \textmu s.
+In case the filter saturates, the media needs to be regenerated by using heated gas; this happened twice during the Run II period\footnote{We deemed the filter regeneration necessary every time the electron lifetime dropped under 100 \textmu s}.
 The filtered argon reaches the inner vessel via a liquid feedthrough on the top of the cryostat.
@@
 \section{Trigger and DAQ}
+%The read-out of a LArIAT event is triggerred by a CAEN V1495 module.
+
+%The LArIAT data acquisition system is triggered to read out the digitizing buffers by a CAEN V1495 module.
+The V1495 is a powerful, easily configurable coincidence module featuring a user-programmable FPGA.
+Sixteen logical inputs  are sampled on a 10~ns clock, checking for matches to any of the sixteen user-defined patterns in the trigger menu.
+If a given trigger pattern is satisfied for two consecutive clock ticks, that pattern fires.
+
+%Scope of the LArIAT DAQ and trigger system is to start the read out of the detector, to convert signals from analog to digital, to interface and synchronize with the accelerator system, to store raw data on disk, transfer data to archival storage and conduct a first pass data quality monitoring.
+
+%A CAEN V1495 governs LArIAT DAQ system.
+
+%The system is made by commercial VME electronics from CAEN, a VME time to digital converter for the beamline time of flight system, and a readout system for beamline wire chambers.
+
+%A CAEN V1495 board is used to perform the trigger of the whole detectors system.
+It is a VME board, which can be directly customized by the user through two programmable FPGA.
+For LArIAT experiment 16 trigger input signals are used, see Fig.
+2.14.
+Through the FPGA configuration file  V1495.config  it is possible to define different trigger conditions - trigger output paths, with several logic combinations of the input signals.
+%The former trigger output path is the one that selects particles from the beam that cross the whole beamline and get to the TPC.
+%Moreover it's possible to add a delay to the trigger and to set some combinations of signals as a VETO.
+%Waveforms from beamline detectors and optical devices (PMTs and SiPMs) inside the cryostat are collected using two CAEN V1751 boards (For v1751 boards inputs, see Table 2.2.).
+They are both fed by the V1495 fast trigger.
+Each has a unit boardId and produces CAENFragment raw data objects.
+%The CAEN V1751 Waveform Digitizer [40] is VME module housing 8 Channel 10 bit 1 GS/s Waveform Digitizer with 1 Vpp input dynamic range; the range that has been set for our purposes is 204 mV, so 1 ADC count corresponds to 0.2 mV.
+The input channel provides a programmable voltage o set in the ± 0.5 V range too; we used this feature to adjust baseline o sets on our signals.
+The time window can also be adjusted by user maintaining the same sample %frequency (1 GS/s); we've started with a 7.168 ?s
+%time window, but after preliminary studies on the signals collected by PMTs and SiPMs we've moved to a 14.336 ?s window that could allow us to sample over all the slow LAr scintillation component (?S) signal tails, to study single photoelectron response of the optical devices.
+
+%The LArIAT data acquisition system is triggered to read out the digitizing buffers by a CAEN V1495 module.
+The V1495 is a powerful, easily configurable coincidence module featuring a user-programmable FPGA.
+Sixteen logical inputs  are sampled on a 10~ns clock, checking for matches to any of the sixteen user-defined patterns in the trigger menu.
+If a given trigger pattern is satisfied for two consecutive clock ticks, that pattern fires.
+
+%NIM-standard logic pulse inputs to the trigger decision come from any of the instruments in the experiment, in principle; trigger inputs are derived from beam instruments, from the cosmic ray taggers, and from the cryostat's scintillation detectors.
+LArIAT takes full advantage of the trigger card's great flexibility.
+As conditions change and innovations are gradually made to LArIAT's data-taking process, improvements are made to the trigger input designs and the combinations of inputs which define trigger patterns for the V1495.
+The configuration of the trigger inputs is automatically logged in a database at the start of each run as part of the DAQ configuration, along with the rest of the XML configuration file for the V1495.
+
+%\subsubsection{Inputs to the Trigger System}
+
+%Two primary inputs to the trigger card are from the time of flight (see Sec.~\ref{sec:TOF}) and the wire chamber (see Sec.~\ref{sec:MWPC}) systems in the tertiary beamline.
+Coincidence of activity in both of these systems strongly suggests that a charged particle has made the journey down the tertiary beam line from the copper target to the LArTPC in the cryostat, and that we will have a measure of its momentum and its velocity.
+PMT pulses from the time of flight system each pass through a linear fan-out, one output of which is threshold discriminated by a
+%discriminator to produce a NIM logic pulse for use in trigger logic.
+On each of the upstream or downstream TOF paddles, we form a coincidence to within 20~ns of pulses from all the PMTs observing that same block of scintillator.
+The upstream paddle coincidence signal is delayed 20~ns to allow any approximately lightspeed particles to travel 6.5~m to the downstream paddle.
+The same upstream paddle coincidence signal is also widened to 100~ns to allow for slow-traveling (high-mass) particles.
+Coincidences of these upstream and downstream TOF trigger input signals can be made in the V1495 trigger card.
+
+%Each wire chamber is read out by four multi-hit TDCs (time to digital converters), with sixty-four wires per TDC and two TDCs per plane (one horizontal and one vertical).
+The TDCs each provide a logical ``fast'' OR of their inputs, indicating that one or more of their sixty-four wires went over the settable threshold.
+Using NIM logic units, the OR of the horizontal wires and the OR of the vertical wires are input to a coincidence unit for each wire chamber, providing a single logical pulse for each of the four wire chambers, indicating that at least one horizontal wire and one vertical wire saw significant signal.
+The single culminating logical pulses from each of the four wire chambers make up the first four inputs to the V1495 trigger card.
+Within the user-programmable FPGA, the V1495 looks for the coincidence within 20~ns of at least three of these four specially-treated logical inputs.
+
+%Another primary input to the trigger card is from the cosmic towers (see Section~\ref{sec:CosmicRayPaddle}).
+To capture cosmic ray events in which a minimally ionizing cosmic ray muon crossed the TPC along the body diagonal, NIM modules form the logical coincidences from the two cosmic towers, one upper and one lower paddle assembly, in each combination.
+The OR of these is provided as an input to the V1495.
+
+%Three important logic pulses are derived from the timing of the beam.
+These include a pulse in a brief window before the beam, a pulse indicating that the beam is on, and a pulse which defines the beam-free period which may be used for collecting cosmic-ray events.
+An adjustable pulser is a fourth trigger input which does not depend on any particular activity in the experiment hall,  useful for collecting background events with zero bias.
+
+%The PMTs observing liquid argon scintillation light (see Section \ref{sec:PhotonSystem}) produce pulses which form the foundation of several interesting trigger inputs.
+Thresholding a copy of each PMT pulse (after amplification), and requiring a coincidence of pulses within $\sim$20~ns, creates simple trigger inputs indicating ionizing radiation was produced in the TPC.
+This scintillation logic pulse is used to initiate a gate which spans the length of the TPC drift time, creating a logic signal which is remains ``on'' while significant drift charge may still be present in the TPC.
+In addition, requiring a delayed coincidence of two subsequent scintillation logic pulses, separated by a variable length of time ranging from 300~ns to 7~$\mu$s, is used to create a trigger input to select events where a cosmic muon stops and decays to a Michel electron in the TPC.
+A few different versions of this light-based trigger were implemented throughout the course of LArIAT's run time to allow reconstruction and calorimetric studies of Michel electrons.
+Figure~\ref{michel_logic} shows a schematic diagram of the logic comprising the Michel electron trigger.
+
+%\begin{figure}
+%\includegraphics[width=\textwidth]{figures/trigger_michellogic.png}
+%\caption{\label{michel_logic}A schematic diagram of the trigger logic used to select Michel electron events during the cosmic readout window of the LArIAT supercycle.
+The two PMT signals refer to the Hamamatsu (``HMM'') and ETL PMTs described in Section~\ref{sec:PhotonSystem}.
+For some data-taking periods in Run-II, un-amplified pulses were discriminated at 180 mV to act as a veto on events that may saturate the dynamic range of the V1751 digitizer.
+The discriminator thresholds used on the amplified (x10) PMT signal copies (\emph{ThA}, \emph{ThB}) as well as the Gate Delay period, were adjusted between run periods while experimenting with different configurations.}
+%\end{figure}
+
+%Further trigger inputs come from the beam line instrumentation behind the LArTPC cryostat, the PMTs of the Punch-Through scintillator paddles and those of the scintillator paddles instrumenting the Muon Range Stack.
+The PMT pulses of all four of the broad-faced Punch Through paddles are discriminated to form logic pulses.
+A single logic pulse is formed from these, indicating activity in at least two overlapping paddles at the rear of cryostat, before the steel block of the range stack.
+PMT pulses from the Muon Range Stack are amplified and threshold discriminated.
+These MuRS paddle pulses are then combined as in the Punch Through, creating single-bit indicators for each of the four instrumented layers that at least one pair of overlapping scintillator paddles sent signals within a 20~ns coincidence window.
+
+%\subsubsection{Trigger Decision and Issuance}
+
+%The V1495 may be configured to have up to sixteen trigger patterns and sixteen veto patterns, based on the trigger input signals.
+A trigger pattern is defined as the AND of one or more defined inputs, and may include the NOT of the AND of further inputs.
+Veto patterns are independently defined in the same way, but they have a very different effect.
+When any of the trigger patterns fire, a ``fast trigger'' signal is issued and an adjustable countdown is initiated.
+If the countdown completes without a veto pattern firing, the ``slow trigger'' signal is also issued and on a distinct hardware channel.
+Otherwise, if a veto pattern fires during the countdown, the slow trigger signal is vetoed.
+
+%The fast trigger signal prompts readout of all the `short' data buffers, which include the V1751 modules, the V1495 itself, and the MWPC controller.
+The V1751 buffers typically contain digitized PMT signals from the time of flight and cryogenic light collection detectors.
+Readout of the TPC wire signals, which are much longer and more numerous, is only prompted at the issuance of the slow trigger.
+
 \section{Control Systems}
+LArIAT is a complex ensemble of systems which needed to be monitored at once during data taking.
+We performed the monitoring of the systems operations with a slow control system, a DAQ monitoring system and a low level data quality monitoring described in the following sections.
+
+\subsubsection{Slow Control}
+We used the Synoptic Java Web Start framework as a real-time display of subsystem conditions. Its simple 
+Graphical User Interface allowed us to change the operating parameters and to graph the trends of several variables of interest for all the tertiary beam detectors.
+Among the most important quantities monitored by Synoptic there are the level of argon in both the inner vessel and the external dewar, the operating voltages of cathode and wire planes, of the PMTs and SiPMs, and of the four wire chambers, as well as the magnets temperature.
+Figure \ref{fig:synoptics} shows an example of the monitoring system.
+LArIAT uses the Accelerator Control NETwork system (ACNET) to monitor the beam conditions of the MCenter beamline.
+For example, the X and Y position of the beam at the first two wire chambers (WC1 and WC2) are shown in \ref{fig:ACNET}.
+
+\begin{figure}[htb]
+\centering
+\includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{Chapter-2/Images/BeamOverview.png}
+\caption{Interface of the Synoptic slow control system}
+\label{fig:synoptics}
+\end{figure}
+
+\begin{figure}[htb]
+\centering
+\includegraphics[scale=0.5]{Chapter-2/Images/BeamPosition.png}
+\caption{Beam position at the upstream wire chambers monitored with ACNET.}
+\label{fig:ACNET}
+\end{figure}
+
+\subsubsection{DAQ Monitoring}
+
+We monitor the data taking and the run time evolution with the Run Status Webpage (\href{http://lariat-wbm.fnal.gov/lariat/run.html}{http://lariat-wbm.fnal.gov/lariat/run.html}), a  webpage updated in real-time.
+The page displays, among other information, the total number of triggers in the event per CAEN digitizer board, the total number of detectors triggered during a beam spill,  the trigger patterns (the number of times a particular trigger pattern was satisfied  during a beam spill) and current time relative to the Fermilab accelerator complex supercycle.
+A screen shot of the page is show in figure \ref{fig:runcond}.
+
+\begin{figure}[htb]
+\centering
+\includegraphics[scale=0.6]{Chapter-2/Images/RunConditions.png}
+\caption{Run Status page at LArIAT downtime.
+At the top the yellow bar displays the current position in the Fermilab supercycle.
+Interesting information to be monitored by the shifter were the run number and number of spills, time elapsed from data taking (here in red), the energy of the secondary beam and the trigger paths.}
+\label{fig:runcond}
+\end{figure}
+
+\subsubsection{Data Quality Monitoring}
+We employ two systems to ensure the quality of out data during data taking: the Near-real-time Data Quality Monitoring and the Event Viewer.
+
+\href{http://lariat-daq01.fnal.gov:5000/}{The Near-real-time Data Quality Monitoring} (DQM).
+This webpage receives updates from all the VME boards in the trigger system and displays the results of a quick analysis of the DAQ stream of raw data on a spill-by-spill basis.
+The DQM allows the shifter to monitor almost in real time (typically with a 2-minute delay)  a series of low level-quantities and compare them to past collections of beam spills.
+Some of the variables monitored in the DQM are  the pedestal mean and RMS on CAEN digitizer boards
+of the TPC wires and PMTs of the beamline detectors, the hit occupancy and timing plots on the multi-wire chambers, and number of data fragments recorded that are used to build a TPC event.
+Abnormal values for  low-level quantity in the data  activate a series of alarms in the DQM; this quick feedback on the DAQ and beam conditions is fundamental to assure a fast debugging of the detector and a very efficient data taking during beam uptime.
+
+The online Event Viewer displays a two dimensional representation of LArIAT TPC events on both the Induction and the Collection planes in near real time.
+The raw pulses collected by the DAQ on each wire are plotted as a function of drift time, resulting in an image of the TPC event easily readable by the shifter.
+This tool guarantees a particularly good  check of the TPC operation which activate an immediate feedback for troubleshooting a number of issues.
+For example,  it easy for the shifter to spot high occupancy events and request a reduction of the primary beam intensity, or to spot a decrease of the argon purity which requires the regeneration of filters, or to catch the presence of electronic noise and reboot the ASICs.
+An example of high occupancy event is shown in \ref{fig:highOcc}.
+
+\begin{figure}[htb]
+\centering
+%\includegraphics[scale=0.6]{Chapter-2/Images/RunConditions.png}
+\caption{High occupancy event display.}
+\label{fig:highOcc}
+\end{figure}

### 2018-03-14 Drafting.tex
s

@@
-The Liquid Argon Time Projection Chamber (LArTPC) can be broken down into three major subcomponents: $1)$ The high voltage system which provides the drift field voltage; $2)$ the cathode and field cage which steps down the high voltage through a network of voltage-dividing resistors to form a uniform electric field for charges to drift within; and $3)$ the wire planes which provide the charge sensitive readout for the detector.
-Here we describe each of the subcomponents which make up the LArIAT LArTPC.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{subsubsection}{High Voltage}\label{sec:DriftVoltage}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-The drift high voltage system for the LArIAT detector, shown pictorially in Figure~\ref{fig:HVScheme}, is designed to allow ionization electrons from the interaction of charged particles in the liquid argon to drift to the wireplanes.
-The high voltage system consists of a Glassman LX125N16 power supply~\cite{GlassmanPS} capable of generating -125~kV and 16~mA of current.
-This power supply was typically operated much below this limit during normal data taking operations.
-The voltage from the power supply is transmitted through high voltage cables to a series of filter pots before finally reaching the high-voltage feedthrough on the top of the LArIAT cryostat.
-This feedthrough brings the voltage into the liquid argon volume to be transmitted to the TPC cathode.
-
-\begin{figure}[htb]
-\centering
-\includegraphics[scale=0.35]{./figures/HVSchematic.png}\\
-\caption{Schematic of the LArIAT high voltage system.}
-\label{fig:HVScheme}
-\end{figure}%See DocDB 1472
-
-The first high voltage cable which connects the Glassman power supply to the first filter pot is a 1/2-inch HV cable, Glassman cable model 2121, supplied by the Glassman manufacturer.
-The second cable is a $\sim$100-ft long 1/2-inch Type-2134 Dielectric Sciences cable which connects the first filter pot to the second filter pot.
-Finally, the same type of cable is used to connect the second filter pot to the high voltage feedthrough at the cryostat.
-
-The two filter pots that are inline with the high voltage system serve to: $a)$ limit the current draw on the power supply; $b)$  provide a low-pass filter to help reduce the voltage ripple on the cathode; and $c)$ partition the energy stored in the system in the event of a high voltage trip.
-Each filter pot, as shown in Figure~\ref{fig:FilterPot}, is a cylindrical aluminum pot 18.5 inches tall by 20 inches in diameter and 3/16 inches thick, with welded tops each having an opening that allows for a flange and o-ring to receive the high voltage cables.
-Within each pot, the flanges are connected to four 10~M$\Omega$ resistors connected in series.
-This assembly is submerged in $\sim$16 gallons of Diala transformer oil~\cite{ShellDiala} to aid in the suppression of corona and discharges.
-
-\begin{figure}[htb]
-\centering
-\includegraphics[scale=0.35]{./figures/FilterPot.png}\\
-\caption{Filter pots used in the LArIAT high voltage system.}
-\label{fig:FilterPot}
-\end{figure}%See DocDB 1472
-
-The final piece of the high voltage system is the high voltage feedthrough which delivers the drift voltage to the cathode of the TPC.
-This feedthrough is based on a a custom design that was used by ICARUS \cite{IcarusDetector}.
-The feedthrough consists of a stainless steel inner conductor surrounded by a tube of ultra high molecular weight polyethylene further surrounded by a stainless steel outer ground tube.
-The high-voltage feedthrough enters the cryostat from a dedicated 4-5/8 inch Conflat flange at the top of the cryostat.
-The end of the inner conductor is finally attached to the cathode by a flexible conductor bolted at both ends.
-A technical drawing of the LArIAt high voltage feedthrough can be found in Figure \ref{fig:HVFT}.
-
-\begin{figure}[htb]
-\centering
-\includegraphics[scale=0.35]{./figures/lariatFeedthrough.pdf}\\
-\caption{Technical drawing of the LArIAT high voltage feedthrough.
-Various liquid argon levels are shown in cyan to indicate the levels to ensure the feedthrough is immersed in argon during operations}
-\label{fig:HVFT}
-\end{figure}%See DocDB 1472
-
-During usual operations, the operating voltage from the Glassman power supply was -23.5~kV corresponding to a drift field of $\sim$500 V/cm.
-During special runs, the high voltage system was operated in a range from 0~kV to 32.8~kV (corresponding to drift fields ranging from 0~V/cm to 700~V/cm) and the feedthrough was tested up to 60~kV with no electrical breakdown observed.
-
-\end{subsubsection}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{subsubsection}{Cathode and Field Cage}\label{sec:FieldCage}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\textcolor{red}{Description of the cathode and field cage is missing}
-
-As shown in Figure \ref{fig:driftregions}, the LArIAT TPC has three drift volumes, each of which has its own electric field.
-The main drift volume is defined as the region between the cathode plane and the shield plane (C-S).
-Note that the shield plane wires are \emph{not} read out.
-The other two drift regions are those between the shield plane and the induction plane (S-I), and between the induction plane and the collection plane (I-C).
-The electric field in these regions is chosen to satisfy the charge transparency condition to allow for 100$\%$ transmission of the drifting electrons through the shield and then the induction planes.
-We provide a detailed description of the electric field measurement in each region in \ref{sec:EFieldMeasurements}.
-
-\begin{figure}[htb]
-\centering
-\includegraphics[scale=0.35]{./images/DriftRegions.png}\\
-\caption{Schematic of the three drift regions inside the LArIAT TPC: the main drift volume between the cathode and the shield plane (C-S) in green, the region between the shield plane and the induction plane (S-I) in purple, and the region between the induction plane and the collection plane (I-C) in pink.}
-\label{fig:driftregions}
-\end{figure}
-
-Table \ref{tab:voltages} provides the default voltages applied to the cathode and the shield, induction, and collection plane for all the runs.
-These voltages were tuned to maintain the transparency condition when the (I-C) and (S-I) spacing changed.
-For data taken in Run-I, Run-II, and Run-IIIb the spacing was a constant 4~mm; the spacing was 5~mm for Run-IIIa.
-
-\begin{table}[htpb]
-\centering
-\caption{Anode planes default voltages}
-\label{tab:voltages}
-\begin{tabular}{lllll}
-\hline
-\multicolumn{1}{|l|}{V Cathode} & 
-\multicolumn{1}{|l|}{V Shield} & \multicolumn{1}{l|}{V Induction} & \multicolumn{1}{l|}{V Collection} & \multicolumn{1}{|l|}{Run} \\ \hline
-\multicolumn{1}{|l|}{-23.17 kV} &
-\multicolumn{1}{|l|}{-298.8 V} & \multicolumn{1}{l|}{-18.5 V}      & \multicolumn{1}{l|}{338.5 V} & \multicolumn{1}{|l|}{Run-I, Run-II \& Run-IIIb}      \\ \hline
-\multicolumn{1}{|l|}{-23.17 kV} & 
-\multicolumn{1}{|l|}{-325.8 V} & \multicolumn{1}{l|}{-0.6 V}      & \multicolumn{1}{l|}{421.9 V} & \multicolumn{1}{|l|}{Run-IIIa}      \\ \hline
-
-                              &                                 &                                 
-\end{tabular}
-\end{table}
-
-\end{subsubsection}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{subsubsection}{Wire Plane Assembly}\label{sec:WirePlanes}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{subsubsection}*{Wire Winding}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-Wires for the LArIAT wire planes are wound to the correct spacing and tension on an independent winding machine before being transferred to the printed circuit boards on which they are installed in the TPC.
-
-\begin{figure}[h!]
-\centering
-\includegraphics[height=2in]{figures/MamaBear.jpg}
-\caption{The wire-winding apparatus in action. Both frames are attached and a single wire has been wound.}
-% CAN PROBABLY FIND A BETTER PICTURE!
-\label{pic:mamabear}
-\end{figure}
-
-Two rectangular steel frames are bolted back-to-back on an apparatus that rotates at a constant speed, shown in Figure \ref{pic:mamabear}.
-As the frames rotate, wire is drawn from a spool and wrapped around the two frames at constant tension (monitored by a tensiometer in the device paying out the spool).
-Once the requisite number of wires have been wound, the wires are epoxied onto the steel frames with five-minute epoxy and cut, giving two separate wire planes (one on each frame).
-\end{subsubsection}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{subsubsection}*{Wire Transfer}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-Once the wire planes have been wound, they must be transferred from the winding frames onto the G10 wire-carrier boards that go into the LArIAT TPC.
-To achieve this, the frames are unbolted from the winding apparatus and laid over the board onto which the wires are to be transferred, which itself is bolted to a rigid aluminium frame to prevent it from flexing under tension.
-Careful manual alignment is performed, using guide holes drilled in the board for reference.
-This alignment is shown in Figure \ref{pic:wiretransfer}.
-
-Once the wires are aligned to the solder pads on the board, they are glued to the board behind the solder pads with a temporary epoxy strip (using five-minute epoxy) to preserve the alignment.
-This is then followed by a permanent epoxy strip on the inside of the solder pads, using a slow-curing epoxy that is safe for cryogenic conditions.
-This epoxy is allowed to cure for 12 to 24 hours before the wires are cut behind the temporary strip and the winding frame can be removed.
-
-\begin{figure}[h!]
-\centering
-\includegraphics[height=2in]{figures/WireTransfer.jpg}
-\caption{One of the \textcolor{red}{This can be 4mm or 5 mm wire plane} 3~mm wire planes during transfer, with the wires not yet cut from the winding frame.
-A steel bar was used for this plane to ensure all wires made good contact with the board.}
-\label{pic:wiretransfer}
-\end{figure}
-
-\end{subsubsection}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{subsubsection}*{Soldering of Electrical Connections}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-When the wire transfer is complete, the wires are soldered to their pads using [SOLDER DETAILS HERE], as are the resistor and capacitor components for the U and V planes \textcolor{red}{Which are the values of capacitances and resistances?}.
-Mylar sheeting is used to protect the active region of the wire planes from evaporating solder flux during this process.
-Once all wires have been soldered in place, the wires are broken off behind the solder pads and the temporary epoxy strip is removed.
-The entire board is cleaned with ethyl alcohol to remove any residue of solder flux before the finished wire planes ahead of installation in the TPC.
-\end{subsubsection}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{subsubsection}*{Wire Plane Testing}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\textbf{Electrical Testing}\\
-\textcolor{red}{Both the test on the RC and on the wire tension are very bad described. According to me it is better to delete the section if a better description can not be provided.}
-Every channel on each completed plane is probed with a multimeter and a signal generator to check for electrical continuity, and in the case of the U and V planes, for the correct RC time constant from the components.\\
-\textbf{Tension Testing}\\
-A laser tensiometer is mounted on a rigid aluminium arm over each completed frame and used to verify the tension on a representative sample of the wires. [QUANTITATIVE RESULTS?]
-
-\end{subsubsection}
-
-\end{subsubsection}
-
 \textcolor{red}{The title says "Cold" but all the electronic, cold and warm, is described}
-The LArIAT TPC front-end electronics comprises a 480-channel analog signal path from the TPC wireplanes to the signal digitizers.
-The front-end system also includes a digital control system for the TPC-mounted electronics, a power supply, and a distribution system.
-A block diagram of the overall system is shown in Fig. \ref{pic:FEelectronics}.
-
-\begin{figure}[htbp]
- \centering
- \includegraphics[width=1.0\textwidth]{figures/LArIAT_FE_Electronics.png}
-\caption{
-Overview of LArIAT Front End electronics. 
-} 
-\label{pic:FEelectronics}
-\end{figure}
-
-The electrical signals on the TPC readout wires are typically quite small, being a direct readout of the ionization of the liquid argon.
-To achieve a good Signal-to-Noise Ratio (SNR) for such signals, the LArIAT TPC is instrumented with cold amplifiers developed by Brookhaven National Lab (BNL) and mounted directly to the TPC frame inside the liquid argon cryostat.
-The BNL amplifiers are built as custom Application-Specific Integrated Circuits (ASICs) [ADD reference].
-The BNL ASICs adopted in LAriAT are designated as LArASIC, version 4-star.
-
-The LArASICs are mixed signal devices, where the analog amplifier characteristics are controlled by digitally-set parameters.
-The maximum gain setting of the LArASIC is 25 mV/fc.
-At this gain a 3.5~fC charge deposition - corresponding to the expected charge arriving at an individual wire from a passing MIP energy deposition - on a readout wire will generate an output with 90~mV amplitude.
-At the other end of the readout chain, the CAEN V1740 [\textbf{ADD}: CAEN refernece] digitizers have a 12 bit resolution and a maximum input range of 2~VDC, yielding about 0.5~mV LSB.
-With the analog signal path from LArASIC to the digitizers being designed to provide unity gain, a passing MIP in the TPC is expected to generate a digital signal with amplitude of about 180 ADC counts.
-[\textbf{UPDATE}: Dean found gain to be less than unity, add details; what was the measured MIP amplitude?] \textcolor{red}{what is the sampling rate of the board?}
-
-The readout racks containing the signal digitizers are positioned at a distance of $\sim$8~mt from the cryostat and are referenced to a different ground respect to the TPC electronics ground.
-The distance minimizes the presence of secondary, tertiary and halo charged particles passing through the digitial electronics.
-To prevent problems caused by differing grounds and noise pick-up along the signal transmission line, the single-ended analog signals from the ASICs are transmitted from the cryostat to the readout racks as differential analog pairs, each pair individually shielded.
-This is achieved using four types of interconnected cards.
-The first type of card in the system is a 48-channel cold motherboard (CMB-48) that mounts directly to the TPC and houses three 16-channel LArASICs per card.
-The next card in the signal path is a single cryostat feed through card (FT) that carries the 480 signals, along with power and control lines, across the cryostat boundary.
-A set of warm receiver and driver (WRD-48) cards plug directly into the cryostat feed-through and amplify the single-ended TPC signals as differential analog.
-The differential analog signals are then driven through 8~mt high quality pleated foil cables to a set of differential-to-single-ended (D2S-64) cards, converting the differential signals into single-ended ones, with enough drive capability to allow direct input into an array of CAEN V1740 digitizers.
-
-At the receiving end of the differential path the differential-mode analog signal arrives with common-mode noise cancelled and converted to single-ended analog for the short distance transmission to the CAEN V1740 digitizers.
 

### 2018-03-14 Drafting.tex
s

@@
-\textcolor{red}{The title says "Cold" but all the electronic, cold and warm, is described}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection*{Cryostat feed through}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-The feedthrough (FT) card developed for LArIAT is sandwiched between an ASA flange and cap, and sealed with o-rings.
-A stiff mechanical structure and backplane-style connectors support the FT card and allow for the WRD-48 cards to be plugged directly into it.
-Such an arrangement facilitates the assembly and potential repair work while reducing the amount of cables and connectors needed.
-
-\begin{figure}[htbp]
- \centering
- \includegraphics[width=0.5\textwidth]{figures/FT_drawings.png}
-\caption{
-Drawing of LArIAT feed through and Warm Receiver Driver crate. 
-} 
-\label{pic:FeedthroughElectronics}
-\end{figure}
-
-The FT card is built as an 8-layer PCB with dimensions 254~mm $\times$ 432~mm.
-The outer layers of the card are mostly uninterrupted ground planes to increase noise immunity.
-The internal traces are built with relatively large 12~mil design rules to enhance manufacturing yield and physical robustness.
-Signal on adjacent layers are staggered to reduce capacitive cross-talk couplings, and ground fills are provided whenever possible on all copper layers.
-
-Supporting the FT card is a mechanical structure that is rooted on the cryostat flange.
-The overall assembly of the card, the ten WRD-48 cards, and the WRD card cage is light enough to be safely cantilevered from the cryostat flange.
-The backplate of the structure is a solid copper sheet which is electrically connected with the cryostat flange though lock washers and also connected to the FT card through 40 conductive standoffs.
-The copper plate and the cryostat flange are defined as the electrical ground of the TPC readout electronics.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection*{CMB-48 card details}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-Each CMB-48 card holds 3 LArASICs.
-The card provides electrical connection to the LArIAT TPC wires and mechanical connection to the TPC structure.
-The CMB-48 board is located inside the cryostat and is thus inaccessible during data-taking periods.
-The board is designed to minimize the source of electrical noise, the risk of failure due to thermal stress, and the risk of component failure.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection*{WRD-48 card details}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-The WRD-48 cards are designed to be paired one-to-one with the CMB-48 cards inside the cryostat.The WRD-48 cards hold 48 channels of single-ended to differential analog amplifiers to transmit the TPC signals to the distant digitizers.
-These cards also provide low-noise power regulation, digital signal conditioning, front-panel diagnostic LEDs and voltage monitor ports, and an on-board test-pulse generator.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection*{D2S-64 card details}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-The D2S-64 cards convert the differential analog signals from the WRD-48 cards to single-ended signals and drive these signals into the $50 \Omega$ input impedance of the CAEN V1740 digitizers.
-
-The LArIAT data acquisition system is triggered to read out the digitizing buffers by a CAEN V1495 module.
-The V1495 is a powerful, easily configurable coincidence module featuring a user-programmable FPGA.
-Sixteen logical inputs (upgraded to 32 in Run-III) are sampled on a 10~ns clock, checking for matches to any of the sixteen user-defined patterns in the trigger menu.
-If a given trigger pattern is satisfied for two consecutive clock ticks, that pattern fires.
-
-NIM-standard logic pulse inputs to the trigger decision come from any of the instruments in the experiment, in principle; trigger inputs are derived from beam instruments, from the cosmic ray taggers, and from the cryostat's scintillation detectors.
-LArIAT takes full advantage of the trigger card's great flexibility.
-As conditions change and innovations are gradually made to LArIAT's data-taking process, improvements are made to the trigger input designs and the combinations of inputs which define trigger patterns for the V1495.
-The configuration of the trigger inputs is automatically logged in a database at the start of each run as part of the DAQ configuration, along with the rest of the XML configuration file for the V1495.
-
 \subsubsection{Inputs to the Trigger System}
@@
 Readout of the TPC wire signals, which are much longer and more numerous, is only prompted at the issuance of the slow trigger.
-
-\subsection*{Signal Digitization}
-
-Signals from all photodetectors are routed from the side-flange of the cryostat to NIM fan-outs to provide the 50-Ohm termination needed to minimize reflections along the cables.
-Copies of the PMT signals are made to aid in several light-based triggers, and copies of all signals are then recored into data using CAEN V1751 digitizer boards with 1~GHz sampling rate.
-
-%------------------------------------------
-\begin{figure}
-\centering
-\includegraphics[width=0.8\textwidth]{figures/lightsys_signals.png}
-\caption{Some example signals, in ADC units, from LArIAT photodetectors for a triggered cosmic Michel electron event.}
-\label{lightsys_signals}
-\end{figure}
-\begin{figure}
-\centering
-\includegraphics[width=0.6\textwidth]{figures/lightsys_avewfm_hmm.png}
-\caption{\label{wfm_hmmpmt}An example of an averaged inverted waveform of the Hamamatsu PMT after event-by-event noise removal and overshoot correction.
-Data was taken from a small subset of events from Run-II.
-In this example, a falling exponential is fit to the signal region from 300~ns to 3~$\mu$s following the onset of the pulse to extract the late-light component lifetime $\tau$ = 1117~ns.}
-\end{figure}
-%------------------------------------------
-
-\subsection*{Nitrogen Contamination Measurement Using Light}
-
-The light detection system can also be used to derive information about the nitrogen (N2) contamination in liquid argon.
-The influence of N2 on liquid argon scintillation light emission has been observed in the WArP experiment~\cite{WARP-nitrogen}.
-With a higher N2 concentration, the observed decay time of the liquid argon slow component decreases significantly, from over 1~$\mu$s to hundreds of nanoseconds.
-In this measurement, waveforms recorded by the ETL PMT in Run-I, selected to cover periods of varying N2 concentrations (as measured by the external sensor in Figure~\ref{light_nitro}), were analyzed.
-An exponential fit was performed to estimate the decay time of the slow light component.
-Results, shown in Figure~\ref{light_nitro}, were compared with the model from the WArP paper~\cite{WARP-nitrogen}.
-
-%------------------------------------------
-\begin{figure}
-\includegraphics[height=0.2\textheight]{figures/nitrosensor.jpeg}
-\hspace{0.5cm}
-\includegraphics[height=0.2\textheight]{figures/nitro_qt.jpeg}
-\label{light_nitro}
-\caption{Concentration of N2 in LArIAT as measured by an external sensor in Run-I (left), and results of fitting exponential models to waveforms recorded by the ETL PMT compared with the model from WArP~\cite{WARP-nitrogen} (red line).}
-\end{figure}
-%------------------------------------------

### 2018-03-16 Chapter-2/2-TheLArIATExperiment.tex
s

@@
 \subsubsection{Multi-Wire Proportional Chambers functionality}\label{sec:MWPCfunc}
+We use the wire chamber system together with the bending magnets to measure the particles momentum.
+
+In the simplest scenario, only one hit on each and every of the four wire chambers is recorder during a single readout of the detector systems.
+Thus, we use the hits position in the two wire chambers upstream to the magnets to form a trajectory before the bent and the hits position in the two wire chambers downstream to the magnets to form a trajectory after the bent.
+We use the angles in the XZ plane between the upstream and downstream trajectories  to calculate the z component of the momentum as follows:
+
+\begin{equation}
+P_z=\frac{B_{eff}L_{eff}}{3.3(sin(\theta_{DS})-sin(\theta_{US}))},
+\label{eq:momformula}
+\end{equation}
+
+where $B_{eff}$ is the effective maximum field in a square field approximation,  $L_{eff}$ is the effective length of both magnets (twice the effective length of one magnet) and  3.3~$c^{-1}$ the conversion factor from from T*m to MeV/c.
+By using the hit positions on WC3 and WC4, we estimate the azimutal and polar angles of the particle trajectory, and we are able to calculate the other components of the momentum.
+
+This simple scenario can get complicated in two ways:  the presence of multiple hits in a single wire chamber or by the absence of hits in one (or more) wire chamber.
+The first complication is due to beam pile up, while the latter is due to wire chamber inefficiency.
+Even in the case of multiple hits, at most one wire chamber track is reconstructed per event; since the magnets bend particles only in the X direction, we assume the particle trajectory to be roughly constant in the YZ plane, thus we keep the combination of hits most coplanar on YZ plane.
+It is possible to reconstruct the particle momentum even in case of wire chamber inefficiency, if the information is missing in either of the two middle wire chambers (WC2 or WC3).
+%Assuming identical magnets, the upstream and downstream trajectories projected without bending should intersect at some point in the space between the magnets.
+Over a collection of many tracks, these points of intersection would trace a plane, therefore the term "midplane" is used.
+The completed half of the wire chamber track is projected to the midplane and the point of intersection is used with the incomplete half to finish the track.
+
+\textcolor{red}{BETTER EXPLAINATION OF HIGH YIELD}
+\textcolor{red}{UNCERTAINTY ON MOMENTUM MEASUREMENT}
+Events satisfying the simplest scenario of one single hit in each of the four wire chambers form the so called ``Picky track" sample.
+We construct an other, higher statistics sample where we loose the requirements on single hit and wire chamber efficiency called "High Yield".
+For LArIAT Run II, the "High Yield" sample is
+\textcolor{red}{ HIGH YIELD increase in stat} the "Picky Tracks" sample.
+We use Picky tracks for calibration purposes: \textcolor{red}{ I DONT UNDERSTAND THIS Plotting the fractional error of the corrected three-point momentum compared to the known four-point momentum shows an uncertainty of approximately 2\%.}
 
@@
 \section{In the Cryostat}
+\subsection{Cryogenics and Argon Purity}\label{ch:Cryo}
+LArIAT repurposed the ArgoNeuT cryostat \cite{argoneut} in order to use it in a beam of charge particles. We also added a new process piping and a new liquid argon filtration system in FTBF.
+Inside the LArIAT experimental hall, the cryostat sits on the beam of charge particles with its horizontal main axis oriented parallel to the beam.
+
+Two volumes make up LArIAT cryostat, shown in Figure \ref{fig:LArIATCryoStat}: purified liquid argon fills the inner vessel, while the outer volume insulates it with a vacuum jacket equipped with layers of aluminized mylar superinsulation.
+The inner vessel is a cylinder of 130~cm length and 6.2~cm diameter, containing about 550~L of LAr, corresponding to a mass of 0.76 ton.
+We run the signal cables for the LArTPC and the high voltage feedthrough through a ``chimney'' at the top and mid-length of the cryostat.
+
+\begin{figure}[htb]
+\centering
+\includegraphics[scale=0.18]{Chapter-2/Images/Cryostat1.jpg}
+\includegraphics[scale=0.07]{Chapter-2/Images/Cryostat2.jpg}
+\caption{\emph{(left)} The LArIAT cryostat open with the TPC placed in the inner volume.
+\emph{(right)} The LArIAT cryostat fully sealed during initial commissioning prior to installation at Fermilab Testbeam Facility.Access to the internal volume is possible by opening the upstream end caps of the inner and outer vessels.
+}
+\label{fig:LArIATCryoStat}
+\end{figure}
+
+Given the different scopes of the ArgoNeuT and LArIAT detectors, we made several modification to the ArgoNeuT cryostat in order to use it in LArIAT.
+In particular, the modification  shown in Figure \ref{fig:LArIATCryoMods} were necessary to account for the beam of charged particles entering the TPC and to employ the new FTBT liquid argon purification system.
+We added a ``beam window'' on the front outer end cap and an ``excluder'' on the inner endcap, with the scope of minimizing the amount of dead material upstream of the TPC's active volume.
+Doing so, we reduced the amount of uninstrumented material before the TPC from $\sim$ 1.6 radiation lengths ($X_{0}$) (ArgoNeuT) to less than 0.3 $X_{0}$ (LArIAT).
+To allow studies of the scintillation light, we added a side port feedthrough which enables the mounting of the light collection system, as well as the connections for the corresponding signal and high-voltage cables (see Section \ref{sec:TPCLight}).
+We modified the bottom of the cryostat adding Conflat and ISO flange sealing to connect the liquid argon transfer line to the new argon cooling and purification system.
+
+\begin{figure}[htb]
+\centering
+\includegraphics[scale=0.35]{Chapter-2/Images/CryoMods.png}
+\caption{Pictures of the modified components of the cryostat.
+$1)$ The addition of an outlet to the bottom of the cryostat to allow connections to the purification system; $2)$ The ``beam-window'' on the outer endcap and the concave inner surface of the inner endcap (referred to as the excluder) to reduce the amount of material through which beam particles must travel before entering the TPC; $3)$ The modified side port for the LArIAT light collection system.}
+\label{fig:LArIATCryoMods}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%
+As in any other LArTPC, argon purity is a crucial parameter for LArIAT.
+Indeed, the presence of contaminants effects both the basic working principles of a LArTPC: electronegative contaminants such as oxygen and water decrease the number of ionization electrons collected on the wires after drifting through the volume, while contaminants such as Nitrogen decrease the light yield from scintillation light, especially in its slow component.
+In LArIAT, contaminations should not exceed the level of 100 parts per trillion (ppt).
+We achieve this level of purity in several stages.
+The specifics required for the commercial argon bought for LArIAT are 2 parts per million (ppm) oxygen, 3.5~ppm water, and 10~ppm nitrogen.
+This argon is monitored with the use of commercial gas analyzer.
+Argon is stored in a dewar external to LArIAT hall and filtered before filling the TPC.
+%The argon is delivered from the commercial dewar to the cryostat through 2.54~cm diameter schedule 10 stainless steel piping.
+The piping was insulated with 20.32~cm of polyurethane foam by the manufacturer.
+The piping was cleaned to remove oil and grease before being welded into the system.
+LArIAT uses a filtration system designed for the Liquid Argon Purity Demonstrator (LAPD)~\cite{LAPD}: half of a 77~liter filter contains a 4A molecular sieve (Sigma-Aldrich~\cite{sigma-aldrich}) apt to remove mainly water, while the other half contains BASF~CU-0226~S, a highly dispersed copper oxide impregnated on a high surface area alumina, apt to remove mainly oxygen~\cite{basf}.
+A single pass of argon in the filter is sufficient to achieve the necessary purity, unless the filter is saturated.
+In case the filter saturates, the media needs to be regenerated by using heated gas; this happened twice during the Run II period\footnote{We deemed the filter regeneration necessary every time the electron lifetime dropped under 100 \textmu s}.
+The filtered argon reaches the inner vessel via a liquid feedthrough on the top of the cryostat.
+Argon is not recirculated in the system, but it rather boils off and vent to the atmosphere.
+During data taking, we replenish the argon in the cryostat several times per day to keep the TPC high voltage feedthrough and cold electronics always submerged.
+In fact, we need to monitor the level, temperatures, and pressures  of the argon both in the commercial dewar and inside the cryostat.
 \subsection{LArTPC: Charge Collection}\label{sec:TPCCharge}
@@
 
-\subsection{Cryogenics and Argon Purity}\label{ch:Cryo}
-LArIAT repurposed the ArgoNeuT cryostat \cite{argoneut} in order to use it in a beam of charge particles. We also added a new process piping and a new liquid argon filtration system in FTBF.
-Inside the LArIAT experimental hall, the cryostat sits on the beam of charge particles with its horizontal main axis oriented parallel to the beam.
+\section{Trigger and DAQ}
+The LArIAT DAQ and trigger system governs the read out of all the many detectors forming LArIAT. 
+The CAEN V1495 module featuring a user-programmable FPGA  is the core of this system.
+Every 10~ns, this module checks for matches between sixteen logical inputs and user-defined patterns in the trigger menu; if it finds a match for two consecutive clock ticks, that trigger fires.
 
-Two volumes make up LArIAT cryostat, shown in Figure \ref{fig:LArIATCryoStat}: purified liquid argon fills the inner vessel, while the outer volume insulates it with a vacuum jacket equipped with layers of aluminized mylar superinsulation.
-The inner vessel is a cylinder of 130~cm length and 6.2~cm diameter, containing about 550~L of LAr, corresponding to a mass of 0.76 ton.
-We run the signal cables for the LArTPC and the high voltage feedthrough through a ``chimney'' at the top and mid-length of the cryostat.
-
-\begin{figure}[htb]
-\centering
-\includegraphics[scale=0.18]{Chapter-2/Images/Cryostat1.jpg}
-\includegraphics[scale=0.07]{Chapter-2/Images/Cryostat2.jpg}
-\caption{\emph{(left)} The LArIAT cryostat open with the TPC placed in the inner volume.
-\emph{(right)} The LArIAT cryostat fully sealed during initial commissioning prior to installation at Fermilab Testbeam Facility.Access to the internal volume is possible by opening the upstream end caps of the inner and outer vessels.
-}
-\label{fig:LArIATCryoStat}
-\end{figure}
-
-Given the different scopes of the ArgoNeuT and LArIAT detectors, we made several modification to the ArgoNeuT cryostat in order to use it in LArIAT.
-In particular, the modification  shown in Figure \ref{fig:LArIATCryoMods} were necessary to account for the beam of charged particles entering the TPC and to employ the new FTBT liquid argon purification system.
-We added a ``beam window'' on the front outer end cap and an ``excluder'' on the inner endcap, with the scope of minimizing the amount of dead material upstream of the TPC's active volume.
-Doing so, we reduced the amount of uninstrumented material before the TPC from $\sim$ 1.6 radiation lengths ($X_{0}$) (ArgoNeuT) to less than 0.3 $X_{0}$ (LArIAT).
-To allow studies of the scintillation light, we added a side port feedthrough which enables the mounting of the light collection system, as well as the connections for the corresponding signal and high-voltage cables (see Section \ref{sec:TPCLight}).
-We modified the bottom of the cryostat adding Conflat and ISO flange sealing to connect the liquid argon transfer line to the new argon cooling and purification system.
-
-\begin{figure}[htb]
-\centering
-\includegraphics[scale=0.35]{Chapter-2/Images/CryoMods.png}
-\caption{Pictures of the modified components of the cryostat.
-$1)$ The addition of an outlet to the bottom of the cryostat to allow connections to the purification system; $2)$ The ``beam-window'' on the outer endcap and the concave inner surface of the inner endcap (referred to as the excluder) to reduce the amount of material through which beam particles must travel before entering the TPC; $3)$ The modified side port for the LArIAT light collection system.}
-\label{fig:LArIATCryoMods}
-\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%
-As in any other LArTPC, argon purity is a crucial parameter for LArIAT.
-Indeed, the presence of contaminants effects both the basic working principles of a LArTPC: electronegative contaminants such as oxygen and water decrease the number of ionization electrons collected on the wires after drifting through the volume, while contaminants such as Nitrogen decrease the light yield from scintillation light, especially in its slow component.
-In LArIAT, contaminations should not exceed the level of 100 parts per trillion (ppt).
-We achieve this level of purity in several stages.
-The specifics required for the commercial argon bought for LArIAT are 2 parts per million (ppm) oxygen, 3.5~ppm water, and 10~ppm nitrogen.
-This argon is monitored with the use of commercial gas analyzer.
-Argon is stored in a dewar external to LArIAT hall and filtered before filling the TPC.
-%The argon is delivered from the commercial dewar to the cryostat through 2.54~cm diameter schedule 10 stainless steel piping.
-The piping was insulated with 20.32~cm of polyurethane foam by the manufacturer.
-The piping was cleaned to remove oil and grease before being welded into the system.
-LArIAT uses a filtration system designed for the Liquid Argon Purity Demonstrator (LAPD)~\cite{LAPD}: half of a 77~liter filter contains a 4A molecular sieve (Sigma-Aldrich~\cite{sigma-aldrich}) apt to remove mainly water, while the other half contains BASF~CU-0226~S, a highly dispersed copper oxide impregnated on a high surface area alumina, apt to remove mainly oxygen~\cite{basf}.
-A single pass of argon in the filter is sufficient to achieve the necessary purity, unless the filter is saturated.
-In case the filter saturates, the media needs to be regenerated by using heated gas; this happened twice during the Run II period\footnote{We deemed the filter regeneration necessary every time the electron lifetime dropped under 100 \textmu s}.
-The filtered argon reaches the inner vessel via a liquid feedthrough on the top of the cryostat.
-Argon is not recirculated in the system, but it rather boils off and vent to the atmosphere.
-During data taking, we replenish the argon in the cryostat several times per day to keep the TPC high voltage feedthrough and cold electronics always submerged.
-In fact, we need to monitor the level, temperatures, and pressures  of the argon both in the commercial dewar and inside the cryostat.
-
-\section{Trigger and DAQ}
-%The read-out of a LArIAT event is triggerred by a CAEN V1495 module.
-
-%The LArIAT data acquisition system is triggered to read out the digitizing buffers by a CAEN V1495 module.
-The V1495 is a powerful, easily configurable coincidence module featuring a user-programmable FPGA.
-Sixteen logical inputs  are sampled on a 10~ns clock, checking for matches to any of the sixteen user-defined patterns in the trigger menu.
-If a given trigger pattern is satisfied for two consecutive clock ticks, that pattern fires.
-
-%Scope of the LArIAT DAQ and trigger system is to start the read out of the detector, to convert signals from analog to digital, to interface and synchronize with the accelerator system, to store raw data on disk, transfer data to archival storage and conduct a first pass data quality monitoring.
-
-%A CAEN V1495 governs LArIAT DAQ system.
-
-%The system is made by commercial VME electronics from CAEN, a VME time to digital converter for the beamline time of flight system, and a readout system for beamline wire chambers.
+% converts signals from analog to digital, interfaces and synchronizes with the accelerator system, stores raw data on disk, transfers data to archival storage and conduct a first pass data quality monitoring.
 

### 2018-03-16 Chapter-4/3-KaonInteractionsInLAr.tex
s

@@
 %\end{comment}
+
+\subsection{Uncertainty budget}
+Measuring an hadronic cross section  in LArIAT translates into counting how many hadrons impinged on a slab of argon at a given energy and how many of those hadrons interacted at said energy.
+So, the key questions here are:
+\begin{itemize}
+\item[]a) how well do we know the kinetic energy at each point of the tracking? %(Incident Kinetic Energy bins)
+\item[]b) how well do we know when the tracking stops? %(Interacting Kinetic Energy bin)
+\item[]c) are there any systematic shifts?
+\end{itemize}
+
+In order to answer this question, will discuss first a simple scenario  were our beam is 100\% made of pions which arrive as primaries in the TPC (no decay in the beam and no inelastic interaction before the TPC front face).
+We will then add a layer of complexity by discussing how we handle beamline contamination.
+\subsubsection{Pure beam of pions}
+\subsubsection{Handling beamline contamination}
+What is the beamline contamination? We define beamline contamination every TPC track matched to the WC track which is not a primary pion. There are 4 different types of beamline contaminations:
+\begin{itemize}
+\item[]1) electrons,
+\item[]2) muons,
+\item[]3) secondary pions,
+\item[]4) matched pile up events.
+\end{itemize}
+So, how do we handle this contamination?
+The first step is understanding what percentage of events used in the data cross section calculation is not a primary pion.
+
+The number of electrons and muons in the pion sample is estimated via the beamline MC. We simulate the beamline composition for each of the magnet settings present on data (-60A and -100A). 
+and we weight for the number of events which pass the mass cut in both the configurations. 
+For each beamline
+
+We estimate the matched pile up events to be a negligible fraction, because of the definition of the WC2TPC match: we deem the probability of a single match\footnote{ Events with multiple matches are rejected.} with a pile up event in the absence of a real particle  within a four centimeter radius from the WC track projection extremely small.

### 2018-03-16 Chapter-4/3-KaonInteractionsInLAr.tex
s

@@
 
-The number of electrons and muons in the pion sample is estimated via the beamline MC. We simulate the beamline composition for each of the magnet settings present on data (-60A and -100A). 
-and we weight for the number of events which pass the mass cut in both the configurations. 
-For each beamline
+The number of electrons and muons in the pion sample is estimated via the beamline MC.
+We simulate the beamline composition for each of the magnet settings present on data (-60A and -100A).
+This composition is weighted according to the number of event in each magnet configuration after the mass cut in order to obtain the right electron to pion and muon to pion ratio of the whole sample.
+We simulate electron, muons and pions with the DDMC and we subject the three samples to the same selection chain (WC2TPC match, shower filter, pile up filter, etc...).
+The percentage of electrons and muons surviving the cut chain is the number of electrons and muons in the pion sample.
+The number of secondary
 

### 2018-03-18 Chapter-4/3-KaonInteractionsInLAr.tex
s

--- 2018-03-16 Chapter-4/3-KaonInteractionsInLAr.tex
@@
 \item[]2) muons,
-\item[]3) secondary pions,
+\item[]3) secondaries from pion events,
 \item[]4) matched pile up events.
 \end{itemize}
+
 So, how do we handle this contamination?
-The first step is understanding what percentage of events used in the data cross section calculation is not a primary pion.
 
-The number of electrons and muons in the pion sample is estimated via the beamline MC. We simulate the beamline composition for each of the magnet settings present on data (-60A and -100A). 
-and we weight for the number of events which pass the mass cut in both the configurations. 
-For each beamline
+The first step is to estimate what percentage of events used in the cross section calculation is not a primary pion.  
+We estimate the percentage of electrons and muons in the beam via the beamline MC\footnote{Since the beamline composition is a function of the magnet settings, we simulate separately events for magnet current of -60A and -100A.
+We calculate the electron to pion and muon to pion ratio on the whole sample as the weighted sum of the corresponding ratio in the two current settings, 
+\begin{equation}
+\frac{N_e}{N_\pi}_{Data} = w_{60A}\frac{N_e}{N_\pi}_{60A}  + w_{100A}\frac{N_e}{N_\pi}_{100A},
+\end{equation}
+\begin{equation}
+\frac{N_\mu}{N_\pi}_{Data} = w_{60A}\frac{N_\mu}{N_\pi}_{60A}  + w_{100A}\frac{N_\mu}{N_\pi}_{100A},
+\end{equation}
+where the weights $w_{60A}$ and $w_{100A}$ are the percentage of events in the corresponding magnet configuration passing the mass selection in data. }.
+Once the beam composition is know,  we simulate the electrons, muons and pions with the DDMC and we subject the three samples to the same selection chain (WC2TPC match, shower filter, pile up filter, etc...).
+The percentage of electrons and muons surviving the selection chain is the  electron and muon contamination in the pion cross section sample.
+The percentage of secondaries is given in the MC by the number of matched WC2TPC tracks which are not flagged as primary by Geant4.
+We estimate the last type of contamination, the ``matched pile up" events, to be a negligible fraction, because of the definition of the WC2TPC match: we deem the probability of a single match with a halo particle in the absence of a beamline particle\footnote{ Events with multiple WC2TPC matches are always rejected.} extremely small.
 
-We estimate the matched pile up events to be a negligible fraction, because of the definition of the WC2TPC match: we deem the probability of a single match\footnote{ Events with multiple matches are rejected.} with a pile up event in the absence of a real particle  within a four centimeter radius from the WC track projection extremely small.
+Once we estimate the contaminants to primary pion ratio, the next step is subtracting their contribution from data for each type of contaminant independently.
+The contaminant samples are reconstructed and the corresponding interacting and incident histograms are produced.
+We then perform a bin by bin subtraction in the data interacting and incident histograms separately.
+A graphical rendering of this procedure is shown in Fig \ref{fig:backgroundSubtraction}
+Once the data is background subtracted, we apply the correction laid out in the previous section.
+
+\begin{figure}
+\includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{Chapter-4/Images/FakePlot.jpg}
+\label{fig:backgroundSubtraction}
+\caption{A graphical rendering of the beamline contamination background subtraction.
+The contribution of the contaminants is shown in green for the secondaries, in orange for the muons and in pink for electrons.
+The colored plots are coming from the MC and are staggered.
+The percentages shown in the legend are the percentages of contaminants over the total number of events  passing the selection chain.
+We actually expect way less contamination.}
+\end{figure}

### 2018-03-18 Chapter-4/3-KaonInteractionsInLAr.tex
s

@@
 \begin{equation}
- E^{kin}_{Front Face}  = \sqrt{p^2_{Beam} - m^2_{Beam}} - E_{loss},
+ E^{kin}_{Front Face}  = \sqrt{p^2_{Beam} - m^2_{Beam}} - m_{Beam} - E_{loss},
 \label{eq:enFF}
@@
 We will then add a layer of complexity by discussing how we handle beamline contamination.
+
 \subsubsection{Pure beam of pions}
+Assuming a beam of pure pions gets to the TPC, let us explicit some of the variables in the kinetic energy equation \ref{eq:KEj}  to point out the important quantities in the error budget,
+
+\begin{align}
+ E_{j}^{kin} &=  E_{Beam}^{kin}  - E_{loss} - \sum_{i < j} \frac {dE_i}{dx_i}*dx_i\\
+                  &=  \sqrt{p^2_{Beam} - m^2_{Beam}} - m_{Beam} - E_{loss} - \sum_{i < j} \frac {dE_i}{dx_i}*dx_i.
+\end{align}
+
+Let us start by discussing the uncertainty on $E_{Beam}^{kin}$.
+Since we are assuming a beam of pions, the uncertainty on the value of mass of the pion ($m_{Beam}$) as given by the pdg is irrelevant compared to the momentum uncertainties, thus $\delta E_{Beam}^{kin} = \delta p_{Beam}^{kin}$.
+\textcolor{blue}{Ideas on how to estimate the momentum uncertainty.
+Base line:  4-points track.
+In case of 3-points track, add an additional 2\%.
+}
+
+KEWC = (pWC2 - m2)?   => ?m very small, ?KEWC = ?pWC
+ELOSS                        => ?ELOSS MC truth spread
+dE/dxi                  => ?dE/dxi MC Reco - Data comparison
+dxi                        => ?dxi      MC Reco - Data comparison
+?TOT, ?i<j                  => Efficiency correction
+
 \subsubsection{Handling beamline contamination}
@@
 Once the data is background subtracted, we apply the correction laid out in the previous section.
+\textcolor{blue}{How do we account for the error in the contamination subtraction? We change the electron/pion and muon/pion ratio and we see how much difference we get?}
 

### 2018-03-18 Chapter-4/3-KaonInteractionsInLAr.tex
s

@@
 Since we are assuming a beam of pions, the uncertainty on the value of mass of the pion ($m_{Beam}$) as given by the pdg is irrelevant compared to the momentum uncertainties, thus $\delta E_{Beam}^{kin} = \delta p_{Beam}^{kin}$.
-\textcolor{blue}{Ideas on how to estimate the momentum uncertainty.
-Base line:  4-points track.
-In case of 3-points track, add an additional 2\%.
+We estimate the momentum uncertainty as follows.
+
+\textcolor{blue}{  
+We estimate the uncertainty on a 4-point track. In case of 3-points track, we add an additional 2\% coming from Greg's study. 
+Uncertainty on a 4-point track:
+\begin{itemize}
+\item[-]  Alignment surveys. 1mm misalignment translates to 3\% in overall
+\item[-] Doug study dp/p = ~2\% based on field map (docdb 1710)
+\item[-] Minerva test beam paper
+\end{itemize}
 }
 
-KEWC = (pWC2 - m2)?   => ?m very small, ?KEWC = ?pWC
-ELOSS                        => ?ELOSS MC truth spread
+We estimate the uncertainty on the energy loss between the beamline momentum measurement and the TPC, $E_{loss}$, using the DDMC pion sample.
+We shoot pions from WC4 with the same momentum distribution as in the beamline data and plot the true $E_{loss}$ for that sample.
+The width of the $E_{loss}$ distribution is the $\delta E_{loss}$.
+
+\textcolor{blue}{  TO DO HERE: make sure we have the geometry right, cause otherwise this correction is meaningless.}
+
+\begin{comment}
 dE/dxi                  => ?dE/dxi MC Reco - Data comparison
@@
 ?TOT, ?i<j                  => Efficiency correction
+\end{comment}
 

### 2018-03-18 Chapter-4/3-KaonInteractionsInLAr.tex
s

@@
 
-\subsection{Uncertainty budget}
+\chapter{Uncertainty budget}
 Measuring an hadronic cross section  in LArIAT translates into counting how many hadrons impinged on a slab of argon at a given energy and how many of those hadrons interacted at said energy.
@@
 
-\subsubsection{Pure beam of pions}
-Assuming a beam of pure pions gets to the TPC, let us explicit some of the variables in the kinetic energy equation \ref{eq:KEj}  to point out the important quantities in the error budget,
+\section{Pure beam of pions}
+Assuming a beam of pure pions gets to the TPC, let us explicit some of the variables in the kinetic energy equation \ref{eq:KEj}  to point out the important quantities in the uncertainty budget,
 
@@
 
+\subsection{Uncertainty on $E_{Beam}^{kin}$}
 Let us start by discussing the uncertainty on $E_{Beam}^{kin}$.
@@
 
+\subsection{Uncertainty on $E_{loss}$}
 We estimate the uncertainty on the energy loss between the beamline momentum measurement and the TPC, $E_{loss}$, using the DDMC pion sample.
@@
 
-\textcolor{blue}{  TO DO HERE: make sure we have the geometry right, cause otherwise this correction is meaningless.}
+\textcolor{blue}{ TO DO HERE: make sure we have the geometry right, cause otherwise this correction is meaningless.  With this method, so far we get a mean ~40 MeV, but uncertainty ~7MeV. 
+The trajectory method does not improve uncertainty, why? It's a mystery I don't think we should solve before June :) .
+Back of the envelope material budget calculation:}
+\begin{table}[h!]
+\centering
+\caption{Back of the envelope calculation}
+\label{my-label}
+\begin{tabular}{|l|l|l|l|}
+\hline
+dEdx for MIP, MPV {[}MeV cm$^2$/gr{]} & density {[}g/cm$^3${]} & width {[}cm{]} & E$_{loss}$ {[}MeV{]} \\ \hline
+1.6                                                 & 1.7 (G10)                            & 1.3            & 3.5                   \\
+1.6                                                 & 1.4 (LAr)                            & 1.77           & 4.0                   \\
+1.6                                                 & 7.7 (S.S.)                           & 0.23           & 2.8                   \\
+1.6                                                 & 4.5 (Ti)                             & 0.04           & 0.3                   \\ 
+1.6                                                 & 1.03 (Plastic Sci)                   & 1.1            & 1.8                   \\ \hline
+Total                                               &                                      &                & 12.4                  \\ \hline
+\end{tabular}
+\end{table}
 
-\begin{comment}
-dE/dxi                  => ?dE/dxi MC Reco - Data comparison
-dxi                        => ?dxi      MC Reco - Data comparison
-?TOT, ?i<j                  => Efficiency correction
-\end{comment}
+\textcolor{blue}{Event taking into account a 3 degree bent, we get 12.41 MeV, which is quite far from 40 MeV... something smells here ;)}
 
-\subsubsection{Handling beamline contamination}
+\subsection{Uncertainty on dE/dx and pitch}
+We obtain the uncertainty on dE/dx and track pitch by comparing the dE/dx and pitch distributions in data and MC.
+\textcolor{blue}{ Currently, MPV MC = 1.70 and MPV DATA = 1.72 MeV/cm (~3\% higher).
+TO DO HERE: calculate Argon density from mid-RTD temperature. Compare this  density with MC Argon density. 
+Density change  affects dE/dx (in MeV/cm!). Try changing MC density up to ``real one" and see if dEdX agrees between DATA and MC}
+
+\subsection{Uncertainty on track end, aka efficiency correction}
+From the MC, we obtain an efficiency correction on the interacting and incident distributions separately.
+This is done by comparing the MC reconstructed with the true MC deposition on an event by event basis.
+This correction is applied bin by bin on the data interacting and incident distributions.
+The better our tracking, the smaller this efficiency correction will be. So, step number one is improving the tracking.
+\textcolor{blue}{Need to talk to Bruce about this.}
+\textcolor{blue}{ I don't understand the angle cut that Dave Schmitz and Jon Paley were so vocal about.}
+
+Now, the key question remains: does the tracking behave in the same way in data and MC? 
+We can compare some key plots between reconstructed data and MC which gives us confidence this is true: the track pitch, the tracks straightness and the goodness of fit in data and MC.
+\textcolor{blue}{ Does such a variable as ``goodness of fit" exists in the tracking?
+We should ask Bruce.}
+
+\section{Handling beamline contamination}
 What is the beamline contamination? We define beamline contamination every TPC track matched to the WC track which is not a primary pion. There are 4 different types of beamline contaminations:

### 2018-03-19 Chapter-4/3-KaonInteractionsInLAr.tex
s

@@
-\chapter{Kaons Interactions in Argon: Cross Section}\label{ch:Interactions}
+\chapter{Hadron Interactions in Argon: Cross Section}\label{ch:Interactions}
 \section{Literature Review}
@@
 
-\section{How to Measure Hadron Cross Section in LArIAT}\label{ch:methodology}
+\section{How to Measure a Hadron Cross Section in LArIAT}\label{ch:methodology}
 We use both the LArIAT  beamline detectors and the LArTPC information in order to measure hadronic cross sections in argon.
-Albeit with small differences, both the  $\pi^{-}$-Ar and K$^{+}$-Ar total hadronic cross section measurements rely on the same procedure described in details in the following paragraphs: we select the particle of interest using a combination of beamline detectors and TPC information (paragraph \ref{ch:ParticleSelectionMethod}), we perform a handshake between the beamline information and the TPC tracking to assure we are selecting the right TPC track (paragraph \ref{ch:WC2TPCMatchMethod}), and we apply the ``thin slice" method to get to the final result (paragraph \ref{ch:ThinSliceMethod}).
-\subsection{Particle Selection}\label{ch:ParticleSelectionMethod}
+Albeit with small differences, both the  $\pi^{-}$~-~Ar and K$^{+}$~-~Ar total hadronic cross section measurements rely on the same procedure described in details in the following paragraphs: we select the particle of interest using a combination of beamline detectors and TPC information (paragraph \ref{ch:ParticleSelectionMethod}), we perform a handshake between the beamline information and the TPC tracking to assure we are selecting the right TPC track (paragraph \ref{ch:WC2TPCMatchMethod}), and we apply the ``thin slice" method to get to the final result (paragraph \ref{ch:ThinSliceMethod}).
+We show a cross check of this method in paragraph \ref{ch:procedureTesting}.
+
+\subsection{Event Selection}\label{ch:ParticleSelectionMethod}
+
+\subsubsection{Particle Identification in the beamline}
+In data, the main tool used to identify the hadron of interest for the cross section calculation is the LArIAT tertiary beamline, in its function of mass spectrometer.
+We combine the measurement of the time of flight, $TOF$, and the beamline momentum, $p_{Beam}$, to reconstruct the invariant mass of the particles in the beamline, $m_{Beam}$, as follows
+\begin{equation}
+m_{Beam} = \frac{p_{Beam}}{c}\sqrt{\biggl(\frac{TOF*c}{l}\biggr)^2 -1},
+\end{equation}
+ where $c$ is the speed of light and $l$ is the length of the particle trajectory between the time of flight paddels.
+
+Figure \ref{fig:mass} shows the mass distribution for the Run II negative polarity runs on the left and positive polarity runs on the right.
+We perform the classification of events into the different samples as follows:
+
+\begin{itemize}
+\item \underline{$\pi, \mu, e$:} 0~MeV $<$ mass $<$ 350~MeV
+
+\item \underline{kaon:} 350~MeV $<$ mass $<$ 650~MeV
+
+\item \underline{proton:} 650~MeV $<$ mass $<$ 3000~MeV.
+
+\end{itemize}
+
+%\begin{comment}     
+\begin{figure}
+%\captionsetup{justification=raggedright}  
+\begin{minipage}[b]{.53\textwidth}  
+  \centering  
+\includegraphics[width=3in]{Chapter-4/Images/massRunIIPos.png}
+\end{minipage}%  
+\begin{minipage}[b]{0.53\textwidth}  
+  \centering  
+\includegraphics[width=3in]{Chapter-4/Images/massRunIINeg.png}
+\end{minipage}
+\caption{The mass plotted for a sample of Run-II events reconstructed in the beamline, negative polarity runs on the left and positive polarity runs on the right.
+The classification of the events into $\pi, \mu, e$, kaon, or proton is based on this distribution.\textcolor{red}{CHANGE PLOTS}}
+\label{fig:mass}
+\end{figure}
+
+\subsubsection{Additional Particle Identification technique}\label{ch:electrons}
+For the $\pi^-$-Ar cross section, the resolution of  beamline mass spectrometer is not sufficient to select a beam of pure pions.
+In fact, muons and electrons survive the selection on the beamline mass value.
+It is important to notice that the composition of the negative polarity beam is mostly pions, as discussed in \ref{ch:beamlineComposition}.
+Anyhow, we devise a selection on the TPC information to mitigate the presence of electrons in the sample used for the pion cross section.
+The selection relies on the different topologies of a pion and an electron event in the argon: while the former will trace a track inside the TPC active volume, the latter will tend to ``shower", i.e.
+interact with the medium, produce bremsstrahlung photons which pair convert into several short tracks.
+We provide details of this selection in section \ref{ch:electronCuts}.
+
+\subsubsection{Pile up mitigation }\label{ch:pileUp}
+The secondary beam impinging on LArIAT secondary target produces a pletora of particles.
+The number of particle which trace down the LArIAT beamline is greatly abated by the presence of upstream and downstream collimators.
+%because of pile up particles during the triggered TPC drift time
+
+We adjust the primary beam intensity between LArIAT Run I and Run II to minimize the presence of pile up events in the data sample.
+We remove from the cross section sample events with more than 4 tracks in the first 14 cm upstream portion of the TPC.
+
 \subsection{Wire Chamber to TPC Match}\label{ch:WC2TPCMatchMethod}
-After an event passes the selection described in the previous paragraph, we need to identify the track inside the TPC corresponding to the particle which triggered the beamline detectors, a procedure we refer to as ``WC to TPC match".
-In general, the tracking algorithm will reconstruct more than one track in the event, partially due to the fact that hadrons interact in the chamber, as shown in figure , and partially because of pile up events where the beam of charge particle is too intense compared to the drift time, as shown in figure.
-\textcolor{red}{EVENT DISPLAYS}
+If an event passes the selection on its beamline information, we need to identify the track inside the TPC corresponding to the particle which triggered the beamline detectors, a procedure we refer to as ``WC to TPC match" (WC2TPC for short).
+In general, the TPC tracking algorithm will reconstruct more than one track in the event, partially due to the fact that hadrons interact in the chamber, as shown in figure \ref{fig:pionInteraction}, and partially because of pile up particles during the triggered TPC drift time, as shown in figure \ref{fig:pionPileUp}.
+\textcolor{red}{ADD EVENT DISPLAYS}
+
+In data, we attempt to uniquely match one wire chamber track to one and only one reconstructed TPC track.
+This match leverages on a geometrical selection on both the position of the wire chamber and TPC tracks and the angle between them.
+We consider only TPC tracks whose first point is in the first 2 cm upstream portion of the TPC for the match.
+We project the wire chamber track to the TPC front face where we define the $x_{FF}$ and $y_{FF}$ coordinates used for evaluating the match.
+We define $\Delta$X as the difference between the $x$ position of the most upstream point of the TPC track and $x_{FF}$.
+$\Delta$Y is defined analogously.
+We define the radius difference, $\Delta$R, as $ \Delta \text{R} =  \sqrt{ \Delta \text{X}^2 +  \Delta \text{Y}^2}  $.
+The angle between the incident WC track and the TPC track in the plane that contains them defines $\alpha$.
+If  $\Delta \text{R} < 4 $~cm, $\alpha < 8^\circ $,  a match between WC-track and TPC reconstructed track is found.
+We describe  how we determinate the best value for the radius and angular selection in sec \ref{ch:WC2TPCMatchOptimization}.
+In MC, we mimic the matching between the WC and the TPC track by constructing a fake WC track using truth information at wire chamber four.
+We then apply the same WC to TPC matching algorithm as in data.
+We discard events with multiple WC2TPC matches. We use only TPC track matched to WC tracks in the cross section calculation.
 
@@
 \subsubsection{Cross Sections on Thin Target}
-Cross section measurements on a thin target have been the bread and butter of nuclear and particle {Iexperimentalists since the Rutherford experiments \textcolor{red}{NEED CITATION}.
+Cross section measurements on a thin target have been the bread and butter of nuclear and particle experimentalists since the Rutherford experiments \textcolor{red}{NEED CITATION}.
 At their core, this type of experiments consists in shooting a beam of particles with a known flux on a thin target and recording the outgoing flux.
@@
 \delta N_{Incident} = \sqrt[]{N_{Incident}} \\
-\delta N_{Interacting} = \sqrt[]{N_{Interacting}(1-\frac{ N_{Interacting}}{N_{Incident}})}.
+\delta N_{Interacting} = \sqrt[]{N_{Interacting}\Big(1-\frac{ N_{Interacting}}{N_{Incident}}\Big)}.
 \end{eqnarray}
@@
 
-\subsection{Procedure testing with truth quantities}
+\subsection{Procedure testing with truth quantities}\label{ch:procedureTesting}
 The $\pi^{-}$-Ar and K$^{+}$-Ar total hadronic cross section implemented in Geant4 can be used as a tool to validate the measurement methodology.
@@
 %\end{comment}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 

### 2018-03-19 Chapter-6/6-TrackingOptimization.tex
s

@@
 
-In data, we attempt to uniquely match one WC-Track to one and only one reconstructed TPC track.
+%In data, we attempt to uniquely match one WC-Track to one and only one reconstructed TPC track.
 This match is done by using in the $X$ and $Y$ coordinate of the extrapolated WC-Track to the upstream most point of the reconstructed TPC Track and by using the angle between the incoming track angle and the reconstructed TPC.
@@
 
-We define a match between WC-track and TPC reconstructed track if  $\Delta \text{R} < r_{T}$, $\alpha < \alpha_{T}$ and the Z position of the first reconstructed point of the TPC track is within 2 cm from the TPC front face.
+%We define a match between WC-track and TPC reconstructed track if  $\Delta \text{R} < r_{T}$, $\alpha < \alpha_{T}$ and the Z position of the first reconstructed point of the TPC track is within 2 cm from the TPC front face.
 The determination of the best $r_{T}$ and $\alpha_{T}$ is the scope of the following section.
 
-In MC, we mimic the matching between the WC and the TPC track on Monte Carlo by constructing a fake WC track using truth information at wire chamber four.
+%In MC, we mimic the matching between the WC and the TPC track on Monte Carlo by constructing a fake WC track using truth information at wire chamber four.
 We then apply the same WC to TPC matching algorithm as in data.
 
-\subsection{Matching optimization}
+\subsection{Matching optimization}\label{ch:WC2TPCMatchOptimization}
 Scope of this optimization study is assessing the goodness of the wire chamber-to-TPC match on Monte Carlo and porting the optimized selection cuts to data.

### 2018-03-19 Chapter-4/3-KaonInteractionsInLAr.tex
s

@@
 \section{How to Measure a Hadron Cross Section in LArIAT}\label{ch:methodology}
-We use both the LArIAT  beamline detectors and the LArTPC information in order to measure hadronic cross sections in argon.
+We use both the LArIAT  beamline detectors and the LArTPC information to measure hadronic cross sections in argon.
 Albeit with small differences, both the  $\pi^{-}$~-~Ar and K$^{+}$~-~Ar total hadronic cross section measurements rely on the same procedure described in details in the following paragraphs: we select the particle of interest using a combination of beamline detectors and TPC information (paragraph \ref{ch:ParticleSelectionMethod}), we perform a handshake between the beamline information and the TPC tracking to assure we are selecting the right TPC track (paragraph \ref{ch:WC2TPCMatchMethod}), and we apply the ``thin slice" method to get to the final result (paragraph \ref{ch:ThinSliceMethod}).
@@
 \subsection{Event Selection}\label{ch:ParticleSelectionMethod}
+\subsubsection{Beamline events}\label{ch:beamlineDetectorsData}
+As will be clear in paragraph \ref{ch:ThinSliceMethod}, beamline particle identification and momentum measurement before entering the TPC are fundamental information for the hadronic cross sections measurements in LArIAT.
+Thus, we scan the LArIAT data to keep only events whose wire chamber and time of flight information is registered.
+Additionally, we perform a check of the plausibility of the trajectory inside the beamline detectors: given the position of the hits in the four wire chambers, we make sure the particle trajectory does not cross any impenetrable material such as the collimator and the magnets steel.
 
 \subsubsection{Particle Identification in the beamline}
-In data, the main tool used to identify the hadron of interest for the cross section calculation is the LArIAT tertiary beamline, in its function of mass spectrometer.
+In data, the main tool to establish the identity of the hadron of interest is the LArIAT tertiary beamline, in its function of mass spectrometer.
 We combine the measurement of the time of flight, $TOF$, and the beamline momentum, $p_{Beam}$, to reconstruct the invariant mass of the particles in the beamline, $m_{Beam}$, as follows
@@
 \subsubsection{Additional Particle Identification technique}\label{ch:electrons}
-For the $\pi^-$-Ar cross section, the resolution of  beamline mass spectrometer is not sufficient to select a beam of pure pions.
+In the case of the $\pi^-$-Ar cross section, the resolution of  beamline mass spectrometer is not sufficient to select a beam of pure pions.
 In fact, muons and electrons survive the selection on the beamline mass value.
@@
 \subsubsection{Pile up mitigation }\label{ch:pileUp}
-The secondary beam impinging on LArIAT secondary target produces a pletora of particles.
-The number of particle which trace down the LArIAT beamline is greatly abated by the presence of upstream and downstream collimators.
-%because of pile up particles during the triggered TPC drift time
-
-We adjust the primary beam intensity between LArIAT Run I and Run II to minimize the presence of pile up events in the data sample.
-We remove from the cross section sample events with more than 4 tracks in the first 14 cm upstream portion of the TPC.
+The secondary beam impinging on LArIAT secondary target produces a plethora  of particles.
+The presence of upstream and downstream collimators greatly abates the number of particles tracing down the LArIAT beamline.
+However, more than one beamline particles, or particles produced from the beam interaction with the beamline detectors, may sneak into the LArTPC during its readout time.
+The TPC readout is triggered by the actual particle  firing the beamline detectors; we call ``pile up" the additional traces in the TPC.
+We adjusted the primary beam intensity between LArIAT Run I and Run II to minimize the presence of events with high pile up particles in the data sample.
+For the cross section analyses, we remove events with more than 4 tracks in the first 14 cm upstream portion of the TPC from the sample.
+\textcolor{red}{probably need to do a better job explaining pile up}
 
 \subsection{Wire Chamber to TPC Match}\label{ch:WC2TPCMatchMethod}
-If an event passes the selection on its beamline information, we need to identify the track inside the TPC corresponding to the particle which triggered the beamline detectors, a procedure we refer to as ``WC to TPC match" (WC2TPC for short).
+For each event passing the selection on its beamline information we need to identify the track inside the TPC corresponding to the particle which triggered the beamline detectors, a procedure we refer to as ``WC to TPC match" (WC2TPC for short).
 In general, the TPC tracking algorithm will reconstruct more than one track in the event, partially due to the fact that hadrons interact in the chamber, as shown in figure \ref{fig:pionInteraction}, and partially because of pile up particles during the triggered TPC drift time, as shown in figure \ref{fig:pionPileUp}.
@@
 
-In data, we attempt to uniquely match one wire chamber track to one and only one reconstructed TPC track.
-This match leverages on a geometrical selection on both the position of the wire chamber and TPC tracks and the angle between them.
+We attempt to uniquely match one wire chamber track to one and only one reconstructed TPC track.
+In data, this match leverages on a geometrical selection exploiting both the position of the wire chamber and TPC tracks, and the angle between them.
 We consider only TPC tracks whose first point is in the first 2 cm upstream portion of the TPC for the match.
@@
 %\end{comment}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\chapter{Uncertainty budget}
-Measuring an hadronic cross section  in LArIAT translates into counting how many hadrons impinged on a slab of argon at a given energy and how many of those hadrons interacted at said energy.
-So, the key questions here are:
-\begin{itemize}
-\item[]a) how well do we know the kinetic energy at each point of the tracking? %(Incident Kinetic Energy bins)
-\item[]b) how well do we know when the tracking stops? %(Interacting Kinetic Energy bin)
-\item[]c) are there any systematic shifts?
-\end{itemize}
-
-In order to answer this question, will discuss first a simple scenario  were our beam is 100\% made of pions which arrive as primaries in the TPC (no decay in the beam and no inelastic interaction before the TPC front face).
-We will then add a layer of complexity by discussing how we handle beamline contamination.
-
-\section{Pure beam of pions}
-Assuming a beam of pure pions gets to the TPC, let us explicit some of the variables in the kinetic energy equation \ref{eq:KEj}  to point out the important quantities in the uncertainty budget,
-
-\begin{align}
- E_{j}^{kin} &=  E_{Beam}^{kin}  - E_{loss} - \sum_{i < j} \frac {dE_i}{dx_i}*dx_i\\
-                  &=  \sqrt{p^2_{Beam} - m^2_{Beam}} - m_{Beam} - E_{loss} - \sum_{i < j} \frac {dE_i}{dx_i}*dx_i.
-\end{align}
-
-\subsection{Uncertainty on $E_{Beam}^{kin}$}
-Let us start by discussing the uncertainty on $E_{Beam}^{kin}$.
-Since we are assuming a beam of pions, the uncertainty on the value of mass of the pion ($m_{Beam}$) as given by the pdg is irrelevant compared to the momentum uncertainties, thus $\delta E_{Beam}^{kin} = \delta p_{Beam}^{kin}$.
-We estimate the momentum uncertainty as follows.
-
-\textcolor{blue}{  
-We estimate the uncertainty on a 4-point track. In case of 3-points track, we add an additional 2\% coming from Greg's study. 
-Uncertainty on a 4-point track:
-\begin{itemize}
-\item[-]  Alignment surveys. 1mm misalignment translates to 3\% in overall
-\item[-] Doug study dp/p = ~2\% based on field map (docdb 1710)
-\item[-] Minerva test beam paper
-\end{itemize}
-}
-
-\subsection{Uncertainty on $E_{loss}$}
-We estimate the uncertainty on the energy loss between the beamline momentum measurement and the TPC, $E_{loss}$, using the DDMC pion sample.
-We shoot pions from WC4 with the same momentum distribution as in the beamline data and plot the true $E_{loss}$ for that sample.
-The width of the $E_{loss}$ distribution is the $\delta E_{loss}$.
-
-\textcolor{blue}{ TO DO HERE: make sure we have the geometry right, cause otherwise this correction is meaningless.  With this method, so far we get a mean ~40 MeV, but uncertainty ~7MeV. 
-The trajectory method does not improve uncertainty, why? It's a mystery I don't think we should solve before June :) .
-Back of the envelope material budget calculation:}
-\begin{table}[h!]
-\centering
-\caption{Back of the envelope calculation}
-\label{my-label}
-\begin{tabular}{|l|l|l|l|}
-\hline
-dEdx for MIP, MPV {[}MeV cm$^2$/gr{]} & density {[}g/cm$^3${]} & width {[}cm{]} & E$_{loss}$ {[}MeV{]} \\ \hline
-1.6                                                 & 1.7 (G10)                            & 1.3            & 3.5                   \\
-1.6                                                 & 1.4 (LAr)                            & 1.77           & 4.0                   \\
-1.6                                                 & 7.7 (S.S.)                           & 0.23           & 2.8                   \\
-1.6                                                 & 4.5 (Ti)                             & 0.04           & 0.3                   \\ 
-1.6                                                 & 1.03 (Plastic Sci)                   & 1.1            & 1.8                   \\ \hline
-Total                                               &                                      &                & 12.4                  \\ \hline
-\end{tabular}
-\end{table}
-
-\textcolor{blue}{Event taking into account a 3 degree bent, we get 12.41 MeV, which is quite far from 40 MeV... something smells here ;)}
-
-\subsection{Uncertainty on dE/dx and pitch}
-We obtain the uncertainty on dE/dx and track pitch by comparing the dE/dx and pitch distributions in data and MC.
-\textcolor{blue}{ Currently, MPV MC = 1.70 and MPV DATA = 1.72 MeV/cm (~3\% higher).
-TO DO HERE: calculate Argon density from mid-RTD temperature. Compare this  density with MC Argon density. 
-Density change  affects dE/dx (in MeV/cm!). Try changing MC density up to ``real one" and see if dEdX agrees between DATA and MC}
-
-\subsection{Uncertainty on track end, aka efficiency correction}
-From the MC, we obtain an efficiency correction on the interacting and incident distributions separately.
-This is done by comparing the MC reconstructed with the true MC deposition on an event by event basis.
-This correction is applied bin by bin on the data interacting and incident distributions.
-The better our tracking, the smaller this efficiency correction will be. So, step number one is improving the tracking.
-\textcolor{blue}{Need to talk to Bruce about this.}
-\textcolor{blue}{ I don't understand the angle cut that Dave Schmitz and Jon Paley were so vocal about.}
-
-Now, the key question remains: does the tracking behave in the same way in data and MC? 
-We can compare some key plots between reconstructed data and MC which gives us confidence this is true: the track pitch, the tracks straightness and the goodness of fit in data and MC.
-\textcolor{blue}{ Does such a variable as ``goodness of fit" exists in the tracking?
-We should ask Bruce.}
-
-\section{Handling beamline contamination}
-What is the beamline contamination? We define beamline contamination every TPC track matched to the WC track which is not a primary pion. There are 4 different types of beamline contaminations:
-\begin{itemize}
-\item[]1) electrons,
-\item[]2) muons,
-\item[]3) secondaries from pion events,
-\item[]4) matched pile up events.
-\end{itemize}
-
-So, how do we handle this contamination?
-
-The first step is to estimate what percentage of events used in the cross section calculation is not a primary pion.  
-We estimate the percentage of electrons and muons in the beam via the beamline MC\footnote{Since the beamline composition is a function of the magnet settings, we simulate separately events for magnet current of -60A and -100A.
-We calculate the electron to pion and muon to pion ratio on the whole sample as the weighted sum of the corresponding ratio in the two current settings, 
-\begin{equation}
-\frac{N_e}{N_\pi}_{Data} = w_{60A}\frac{N_e}{N_\pi}_{60A}  + w_{100A}\frac{N_e}{N_\pi}_{100A},
-\end{equation}
-\begin{equation}
-\frac{N_\mu}{N_\pi}_{Data} = w_{60A}\frac{N_\mu}{N_\pi}_{60A}  + w_{100A}\frac{N_\mu}{N_\pi}_{100A},
-\end{equation}
-where the weights $w_{60A}$ and $w_{100A}$ are the percentage of events in the corresponding magnet configuration passing the mass selection in data. }.
-Once the beam composition is know,  we simulate the electrons, muons and pions with the DDMC and we subject the three samples to the same selection chain (WC2TPC match, shower filter, pile up filter, etc...).
-The percentage of electrons and muons surviving the selection chain is the  electron and muon contamination in the pion cross section sample.
-The percentage of secondaries is given in the MC by the number of matched WC2TPC tracks which are not flagged as primary by Geant4.
-We estimate the last type of contamination, the ``matched pile up" events, to be a negligible fraction, because of the definition of the WC2TPC match: we deem the probability of a single match with a halo particle in the absence of a beamline particle\footnote{ Events with multiple WC2TPC matches are always rejected.} extremely small.
-
-Once we estimate the contaminants to primary pion ratio, the next step is subtracting their contribution from data for each type of contaminant independently.
-The contaminant samples are reconstructed and the corresponding interacting and incident histograms are produced.
-We then perform a bin by bin subtraction in the data interacting and incident histograms separately.
-A graphical rendering of this procedure is shown in Fig \ref{fig:backgroundSubtraction}
-Once the data is background subtracted, we apply the correction laid out in the previous section.
-\textcolor{blue}{How do we account for the error in the contamination subtraction? We change the electron/pion and muon/pion ratio and we see how much difference we get?}
-
-\begin{figure}
-\includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{Chapter-4/Images/FakePlot.jpg}
-\label{fig:backgroundSubtraction}
-\caption{A graphical rendering of the beamline contamination background subtraction.
-The contribution of the contaminants is shown in green for the secondaries, in orange for the muons and in pink for electrons.
-The colored plots are coming from the MC and are staggered.
-The percentages shown in the legend are the percentages of contaminants over the total number of events  passing the selection chain.
-We actually expect way less contamination.}
-\end{figure}

### 2018-03-19 Chapter-8/8-Uncertainty.tex
s

\chapter{Uncertainty budget}\label{ch:Uncertainty}
Measuring an hadronic cross section  in LArIAT translates into counting how many hadrons impinged on a slab of argon at a given energy and how many of those hadrons interacted at said energy. So, the key questions here are:
\begin{itemize}
\item[]a) how well do we know the kinetic energy at each point of the tracking? %(Incident Kinetic Energy bins)
\item[]b) how well do we know when the tracking stops? %(Interacting Kinetic Energy bin)
\item[]c) are there any systematic shifts?
\end{itemize}

In order to answer this question, will discuss first a simple scenario  were our beam is 100\% made of pions which arrive as primaries in the TPC (no decay in the beam and no inelastic interaction before the TPC front face). We will then add a layer of complexity by discussing how we handle beamline contamination.

\section{Pure beam of pions}
Assuming a beam of pure pions gets to the TPC, let us explicit some of the variables in the kinetic energy equation \ref{eq:KEj}  to point out the important quantities in the uncertainty budget,

\begin{align}
 E_{j}^{kin} &=  E_{Beam}^{kin}  - E_{loss} - \sum_{i < j} \frac {dE_i}{dx_i}*dx_i\\
                  &=  \sqrt{p^2_{Beam} - m^2_{Beam}} - m_{Beam} - E_{loss} - \sum_{i < j} \frac {dE_i}{dx_i}*dx_i.
\end{align}

\subsection{Uncertainty on $E_{Beam}^{kin}$}
Let us start by discussing the uncertainty on $E_{Beam}^{kin}$. Since we are assuming a beam of pions, the uncertainty on the value of mass of the pion ($m_{Beam}$) as given by the pdg is irrelevant compared to the momentum uncertainties, thus $\delta E_{Beam}^{kin} = \delta p_{Beam}^{kin}$. 
We estimate the momentum uncertainty as follows.

\textcolor{blue}{  
We estimate the uncertainty on a 4-point track. In case of 3-points track, we add an additional 2\% coming from Greg's study. 
Uncertainty on a 4-point track:
\begin{itemize}
\item[-]  Alignment surveys. 1mm misalignment translates to 3\% in overall
\item[-] Doug study dp/p = ~2\% based on field map (docdb 1710)
\item[-] Minerva test beam paper
\end{itemize}
}

\subsection{Uncertainty on $E_{loss}$}
We estimate the uncertainty on the energy loss between the beamline momentum measurement and the TPC, $E_{loss}$, using the DDMC pion sample. We shoot pions from WC4 with the same momentum distribution as in the beamline data and plot the true $E_{loss}$ for that sample. The width of the $E_{loss}$ distribution is the $\delta E_{loss}$.

\textcolor{blue}{ TO DO HERE: make sure we have the geometry right, cause otherwise this correction is meaningless.  With this method, so far we get a mean ~40 MeV, but uncertainty ~7MeV. 
The trajectory method does not improve uncertainty, why? It's a mystery I don't think we should solve before June :) .
Back of the envelope material budget calculation:}
\begin{table}[h!]
\centering
\caption{Back of the envelope calculation}
\label{my-label}
\begin{tabular}{|l|l|l|l|}
\hline
dEdx for MIP, MPV {[}MeV cm$^2$/gr{]} & density {[}g/cm$^3${]} & width {[}cm{]} & E$_{loss}$ {[}MeV{]} \\ \hline
1.6                                                 & 1.7 (G10)                            & 1.3            & 3.5                   \\
1.6                                                 & 1.4 (LAr)                            & 1.77           & 4.0                   \\
1.6                                                 & 7.7 (S.S.)                           & 0.23           & 2.8                   \\
1.6                                                 & 4.5 (Ti)                             & 0.04           & 0.3                   \\ 
1.6                                                 & 1.03 (Plastic Sci)                   & 1.1            & 1.8                   \\ \hline
Total                                               &                                      &                & 12.4                  \\ \hline
\end{tabular}
\end{table}

\textcolor{blue}{Event taking into account a 3 degree bent, we get 12.41 MeV, which is quite far from 40 MeV... something smells here ;)}

\subsection{Uncertainty on dE/dx and pitch}
We obtain the uncertainty on dE/dx and track pitch by comparing the dE/dx and pitch distributions in data and MC.
\textcolor{blue}{ Currently, MPV MC = 1.70 and MPV DATA = 1.72 MeV/cm (~3\% higher).
TO DO HERE: calculate Argon density from mid-RTD temperature. Compare this  density with MC Argon density. 
Density change  affects dE/dx (in MeV/cm!). Try changing MC density up to ``real one" and see if dEdX agrees between DATA and MC}

\subsection{Uncertainty on track end, aka efficiency correction}
From the MC, we obtain an efficiency correction on the interacting and incident distributions separately. This is done by comparing the MC reconstructed with the true MC deposition on an event by event basis.
This correction is applied bin by bin on the data interacting and incident distributions.
The better our tracking, the smaller this efficiency correction will be. So, step number one is improving the tracking.
\textcolor{blue}{Need to talk to Bruce about this.}
\textcolor{blue}{ I don't understand the angle cut that Dave Schmitz and Jon Paley were so vocal about.}

Now, the key question remains: does the tracking behave in the same way in data and MC? 
We can compare some key plots between reconstructed data and MC which gives us confidence this is true: the track pitch, the tracks straightness and the goodness of fit in data and MC. \textcolor{blue}{ Does such a variable as ``goodness of fit" exists in the tracking? We should ask Bruce.}

\section{Handling beamline contamination}
What is the beamline contamination? We define beamline contamination every TPC track matched to the WC track which is not a primary pion. There are 4 different types of beamline contaminations:
\begin{itemize}
\item[]1) electrons,
\item[]2) muons,
\item[]3) secondaries from pion events,
\item[]4) matched pile up events.
\end{itemize}

So, how do we handle this contamination?

The first step is to estimate what percentage of events used in the cross section calculation is not a primary pion.  
We estimate the percentage of electrons and muons in the beam via the beamline MC\footnote{Since the beamline composition is a function of the magnet settings, we simulate separately events for magnet current of -60A and -100A. 
We calculate the electron to pion and muon to pion ratio on the whole sample as the weighted sum of the corresponding ratio in the two current settings, 
\begin{equation}
\frac{N_e}{N_\pi}_{Data} = w_{60A}\frac{N_e}{N_\pi}_{60A}  + w_{100A}\frac{N_e}{N_\pi}_{100A},
\end{equation}
\begin{equation}
\frac{N_\mu}{N_\pi}_{Data} = w_{60A}\frac{N_\mu}{N_\pi}_{60A}  + w_{100A}\frac{N_\mu}{N_\pi}_{100A},
\end{equation}
where the weights $w_{60A}$ and $w_{100A}$ are the percentage of events in the corresponding magnet configuration passing the mass selection in data. }.
Once the beam composition is know,  we simulate the electrons, muons and pions with the DDMC and we subject the three samples to the same selection chain (WC2TPC match, shower filter, pile up filter, etc...). The percentage of electrons and muons surviving the selection chain is the  electron and muon contamination in the pion cross section sample.
The percentage of secondaries is given in the MC by the number of matched WC2TPC tracks which are not flagged as primary by Geant4.
We estimate the last type of contamination, the ``matched pile up" events, to be a negligible fraction, because of the definition of the WC2TPC match: we deem the probability of a single match with a halo particle in the absence of a beamline particle\footnote{ Events with multiple WC2TPC matches are always rejected.} extremely small.

Once we estimate the contaminants to primary pion ratio, the next step is subtracting their contribution from data for each type of contaminant independently. The contaminant samples are reconstructed and the corresponding interacting and incident histograms are produced. We then perform a bin by bin subtraction in the data interacting and incident histograms separately. A graphical rendering of this procedure is shown in Fig \ref{fig:backgroundSubtraction}
Once the data is background subtracted, we apply the correction laid out in the previous section.
\textcolor{blue}{How do we account for the error in the contamination subtraction? We change the electron/pion and muon/pion ratio and we see how much difference we get?}

\begin{figure}
\includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{Chapter-8/Images/FakePlot.jpg}
\label{fig:backgroundSubtraction}
\caption{A graphical rendering of the beamline contamination background subtraction. The contribution of the contaminants is shown in green for the secondaries, in orange for the muons and in pink for electrons. The colored plots are coming from the MC and are staggered. The percentages shown in the legend are the percentages of contaminants over the total number of events  passing the selection chain. We actually expect way less contamination.}
\end{figure}

### 2018-03-19 AppendixB-EField/B-LArIATEField.tex
Changing organization structure

--- 2018-03-11 AppentixA-EField/A-LArIATEField.tex
@@
-\chapter{Measurement of LArIAT Electric Field}
+\chapter{Measurement of LArIAT Electric Field}{ch:AppendixB}
 The electric field of a LArTPC in the drift volume is a fundamental quantity for the proper functionality of this technology, as it affects almost every reconstructed quantity such as the position of hits or their collected charge.
@@
 \centering
-\includegraphics[scale=0.45]{./AppentixA-EField/Images/Walkowiak.png}\\
+\includegraphics[scale=0.45]{./AppendixB-EField/Images/Walkowiak.png}\\
 \caption{Drift velocity dependence on electric field for several temperatures. The slope of the line at any one point represents the electron mobility for that given temperature and electric field.}
@@
 \centering
-\includegraphics[width=3in]{AppentixA-EField/Images/CircuitLArIAT.png}
+\includegraphics[width=3in]{AppendixB-EField/Images/CircuitLArIAT.png}
 \caption{\textcolor{red}{get rid of current line} LArIAT HV simple schematics.}
@@
 \centering
-\includegraphics[width=3in]{AppentixA-EField/Images/glassman_current_20160525-30.png}
+\includegraphics[width=3in]{AppendixB-EField/Images/glassman_current_20160525-30.png}
 \caption{\textcolor{red}{the axis is wrong!!} Current reading from the Glassman between May 25th and May 30th, 2016 (typical Run-II conditions).}
@@
 \centering
-\includegraphics[width=3in]{AppentixA-EField/Images/TPCCrossSectionView.png}
+\includegraphics[width=3in]{AppendixB-EField/Images/TPCCrossSectionView.png}
 \caption{Pictorial representation of the YX view of the TPC.
@@
 \centering
-\includegraphics[width=3in]{AppentixA-EField/Images/AngleDef.png}
+\includegraphics[width=3in]{AppendixB-EField/Images/AngleDef.png}
 \caption{Angle definition in the context of LArIAT coordinates system.}
@@
 \centering
-\includegraphics[width=3in]{AppentixA-EField/Images/RunIIPosCol.png}
+\includegraphics[width=3in]{AppendixB-EField/Images/RunIIPosCol.png}
 \caption{Collection plane $\Delta$t fit for Run II positive polarity ACP data  selected tracks.}
@@
 \centering
-\includegraphics[width=3in]{AppentixA-EField/Images/RunIIPosInd.png}
+\includegraphics[width=3in]{AppendixB-EField/Images/RunIIPosInd.png}
 \caption{Induction plane $\Delta$t fit for Run II positive polarity ACP data  selected tracks.}

### 2018-03-19 Chapter-2/2-LArTPCs.tex
Changing organization structure

--- 2018-01-30 Chapters/1-LArTPCs.tex
@@
-\chapter{Liquid Argon Detectors at the Intensity Frontier}
-Following the recommendation of the latest Particle Physics Project Prioritization Panel  \cite{ParticlePhysicsProjectPrioritizationPanel(P5):2014pwa}, the US is dedicating substantial resources to the development of a short- and long- baseline neutrino program to address the fundamental questions in neutrino physics today.
-A large portion of this program pivots on the Liquid Argon Time Projection Chamber (LArTPC) detector technology.
+\chapter{Liquid Argon Detectors at the Intensity Frontier}\label{ch:1}
+
 In the next few years, LArTPC experiments -- such as the Short-Baseline Neutrino program (SBN) and DUNE -- will be major players in the intensity frontier field.
-
-\section{The SBN Program}
-\subsection{SBN Goals}
-\subsection{Neutrino Interactions and Detection }
-
-\section{DUNE}
-\subsection{DUNE Non-Accelerator Physics Program}
-\subsection{Rare Decay Searches: Experimental Limit}
-\subsection{Nucleon Decay Detection in LAr}
 
@@
 \subsubsection{Wavelength Shifting of LAr Scintillation Light}
+\section{The SBN Program: Neutrino Interaction and Detection}
+%\subsection{SBN Goals}
+%\subsection{Neutrino Interactions and Detection }
+\section{DUNE: Rare Decay Searches}
+%\subsection{Non-Accelerator Physics Program}
+%\subsection{Rare Decay Searches: Experimental Limit}
+%\subsection{Nucleon Decay Detection in LAr}
+\section{Enabling the next generation of discoveries: LArIAT}

### 2018-03-19 Chapter-3/3-TheLArIATExperiment.tex
Changing organization structure

--- 2018-03-16 Chapter-2/2-TheLArIATExperiment.tex
@@
   \centering  	
-\includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{Chapter-2/Images/AcceleratorFNAL.png}
+\includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{Chapter-3/Images/AcceleratorFNAL.png}
 \label{fig:Accelerator}
@@
   \centering  	
-\includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{Chapter-2/Images/Tertiary.png}
+\includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{Chapter-3/Images/Tertiary.png}
 \label{fig:tert-layout}
@@
 
+\subsection{Collimators}\label{sec:Collimators}
+
 \subsection{Bending Magnets}\label{sec:Magnets}
@@
 \vspace{-0.3cm}
-\includegraphics[height=3.0in]{Chapter-2/Images/ExcitationCurves.png}
+\includegraphics[height=3.0in]{Chapter-3/Images/ExcitationCurves.png}
 \caption{
@@
 \vspace{-0.3cm}
-\includegraphics[height=2.3in]{Chapter-2/Images/WireChamber.png}
+\includegraphics[height=2.3in]{Chapter-3/Images/WireChamber.png}
 \caption{
@@
 %\vspace{-0.3cm}
-%\includegraphics[height=2.3in]{Chapter-2/Images/tofdelay.png}
+%\includegraphics[height=2.3in]{Chapter-3/Images/tofdelay.png}
 %\caption{
@@
 \centering
- \includegraphics[angle=90,width=0.7\textwidth]{Chapter-2/Images/Cosmic_Paddle.jpg}
+ \includegraphics[angle=90,width=0.7\textwidth]{Chapter-3/Images/Cosmic_Paddle.jpg}
 \caption{Photograph of one of the scintillation counters used in the cosmic towers. } 
@@
 \centering
-\includegraphics[scale=0.18]{Chapter-2/Images/Cryostat1.jpg}
-\includegraphics[scale=0.07]{Chapter-2/Images/Cryostat2.jpg}
+\includegraphics[scale=0.18]{Chapter-3/Images/Cryostat1.jpg}
+\includegraphics[scale=0.07]{Chapter-3/Images/Cryostat2.jpg}
 \caption{\emph{(left)} The LArIAT cryostat open with the TPC placed in the inner volume.
@@
 \centering
-\includegraphics[scale=0.35]{Chapter-2/Images/CryoMods.png}
+\includegraphics[scale=0.35]{Chapter-3/Images/CryoMods.png}
 \caption{Pictures of the modified components of the cryostat.
@@
 \centering
-\includegraphics[scale=0.35]{Chapter-2/Images//HVSchematic.png}\\
+\includegraphics[scale=0.35]{Chapter-3/Images//HVSchematic.png}\\
 \caption{Schematic of the LArIAT high voltage system.}
@@
 \centering
-\includegraphics[scale=0.35]{Chapter-2/Images/DriftRegions.png}\\
+\includegraphics[scale=0.35]{Chapter-3/Images/DriftRegions.png}\\
 \caption{Schematic of the three drift regions inside the LArIAT TPC: the main drift volume between the cathode and the shield plane (C-S) in green, the region between the shield plane and the induction plane (S-I) in purple, and the region between the induction plane and the collection plane (I-C) in pink.}
@@
  \centering
- \includegraphics[width=1.0\textwidth]{Chapter-2/Images/LArIAT_FE_Electronics.png}
+ \includegraphics[width=1.0\textwidth]{Chapter-3/Images/LArIAT_FE_Electronics.png}
 \caption{Overview of LArIAT Front End electronics. } 
@@
 \centering
-\includegraphics[height=2.2in]{Chapter-2/Images/lightsys_pmts.png}
+\includegraphics[height=2.2in]{Chapter-3/Images/lightsys_pmts.png}
 \hspace{1cm}
-\includegraphics[height=2.2in]{Chapter-2/Images/lightsys_wls.png}
+\includegraphics[height=2.2in]{Chapter-3/Images/lightsys_wls.png}
 \caption{LArIAT's photodetector system for observing LAr scintillation light inside the TPC (left), and a simplified schematic of VUV light being wavelength-shifting along the TPB-coated reflecting foils (right).}
@@
 \centering
-\includegraphics[height=0.25\textheight]{Chapter-2/Images/lightsys_etlbase.jpeg}
+\includegraphics[height=0.25\textheight]{Chapter-3/Images/lightsys_etlbase.jpeg}
 \hspace{0.5cm}
-\includegraphics[height=0.25\textheight]{Chapter-2/Images/lightsys_hmmbase.jpg}
+\includegraphics[height=0.25\textheight]{Chapter-3/Images/lightsys_hmmbase.jpg}
 \caption{\label{voltagedividers}Photos of the voltage divider bases for the ETL PMT (left) and the Hamamatsu PMT (right) used in Run-II.
@@
 %\centering
-%\includegraphics[scale=1.3]{Chapter-2/Images/lightsys_foils.png}
+%\includegraphics[scale=1.3]{Chapter-3/Images/lightsys_foils.png}
 %\caption{\label{lightsys_foils}The TPB-coated reflector foils mounted to the TPC field cage walls as viewed through the front cryostat opening. {\textcolor{red}{Better pictures needed.}}}
@@
 \centering
-\includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{Chapter-2/Images/BeamOverview.png}
+\includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{Chapter-3/Images/BeamOverview.png}
 \caption{Interface of the Synoptic slow control system}
@@
 \centering
-\includegraphics[scale=0.5]{Chapter-2/Images/BeamPosition.png}
+\includegraphics[scale=0.5]{Chapter-3/Images/BeamPosition.png}
 \caption{Beam position at the upstream wire chambers monitored with ACNET.}
@@
 \centering
-\includegraphics[scale=0.6]{Chapter-2/Images/RunConditions.png}
+\includegraphics[scale=0.6]{Chapter-3/Images/RunConditions.png}
 \caption{Run Status page at LArIAT downtime.
@@
 \centering
-%\includegraphics[scale=0.6]{Chapter-2/Images/RunConditions.png}
+%\includegraphics[scale=0.6]{Chapter-3/Images/RunConditions.png}
 \caption{High occupancy event display.}

### 2018-03-19 Chapter-7/7-BackgroundSub.tex
Changing organization structure

--- 2018-03-19 Chapter-8/8-Uncertainty.tex
@@
-\chapter{Uncertainty budget}\label{ch:Uncertainty}
-Measuring an hadronic cross section  in LArIAT translates into counting how many hadrons impinged on a slab of argon at a given energy and how many of those hadrons interacted at said energy.
-So, the key questions here are:
-\begin{itemize}
-\item[]a) how well do we know the kinetic energy at each point of the tracking? %(Incident Kinetic Energy bins)
-\item[]b) how well do we know when the tracking stops? %(Interacting Kinetic Energy bin)
-\item[]c) are there any systematic shifts?
-\end{itemize}
-
-In order to answer this question, will discuss first a simple scenario  were our beam is 100\% made of pions which arrive as primaries in the TPC (no decay in the beam and no inelastic interaction before the TPC front face).
-We will then add a layer of complexity by discussing how we handle beamline contamination.
-
-\section{Pure beam of pions}
-Assuming a beam of pure pions gets to the TPC, let us explicit some of the variables in the kinetic energy equation \ref{eq:KEj}  to point out the important quantities in the uncertainty budget,
-
-\begin{align}
- E_{j}^{kin} &=  E_{Beam}^{kin}  - E_{loss} - \sum_{i < j} \frac {dE_i}{dx_i}*dx_i\\
-                  &=  \sqrt{p^2_{Beam} - m^2_{Beam}} - m_{Beam} - E_{loss} - \sum_{i < j} \frac {dE_i}{dx_i}*dx_i.
-\end{align}
-
-\subsection{Uncertainty on $E_{Beam}^{kin}$}
-Let us start by discussing the uncertainty on $E_{Beam}^{kin}$.
-Since we are assuming a beam of pions, the uncertainty on the value of mass of the pion ($m_{Beam}$) as given by the pdg is irrelevant compared to the momentum uncertainties, thus $\delta E_{Beam}^{kin} = \delta p_{Beam}^{kin}$.
-We estimate the momentum uncertainty as follows.
-
-\textcolor{blue}{  
-We estimate the uncertainty on a 4-point track. In case of 3-points track, we add an additional 2\% coming from Greg's study. 
-Uncertainty on a 4-point track:
-\begin{itemize}
-\item[-]  Alignment surveys. 1mm misalignment translates to 3\% in overall
-\item[-] Doug study dp/p = ~2\% based on field map (docdb 1710)
-\item[-] Minerva test beam paper
-\end{itemize}
-}
-
-\subsection{Uncertainty on $E_{loss}$}
-We estimate the uncertainty on the energy loss between the beamline momentum measurement and the TPC, $E_{loss}$, using the DDMC pion sample.
-We shoot pions from WC4 with the same momentum distribution as in the beamline data and plot the true $E_{loss}$ for that sample.
-The width of the $E_{loss}$ distribution is the $\delta E_{loss}$.
-
-\textcolor{blue}{ TO DO HERE: make sure we have the geometry right, cause otherwise this correction is meaningless.  With this method, so far we get a mean ~40 MeV, but uncertainty ~7MeV. 
-The trajectory method does not improve uncertainty, why? It's a mystery I don't think we should solve before June :) .
-Back of the envelope material budget calculation:}
-\begin{table}[h!]
-\centering
-\caption{Back of the envelope calculation}
-\label{my-label}
-\begin{tabular}{|l|l|l|l|}
-\hline
-dEdx for MIP, MPV {[}MeV cm$^2$/gr{]} & density {[}g/cm$^3${]} & width {[}cm{]} & E$_{loss}$ {[}MeV{]} \\ \hline
-1.6                                                 & 1.7 (G10)                            & 1.3            & 3.5                   \\
-1.6                                                 & 1.4 (LAr)                            & 1.77           & 4.0                   \\
-1.6                                                 & 7.7 (S.S.)                           & 0.23           & 2.8                   \\
-1.6                                                 & 4.5 (Ti)                             & 0.04           & 0.3                   \\ 
-1.6                                                 & 1.03 (Plastic Sci)                   & 1.1            & 1.8                   \\ \hline
-Total                                               &                                      &                & 12.4                  \\ \hline
-\end{tabular}
-\end{table}
-
-\textcolor{blue}{Event taking into account a 3 degree bent, we get 12.41 MeV, which is quite far from 40 MeV... something smells here ;)}
-
-\subsection{Uncertainty on dE/dx and pitch}
-We obtain the uncertainty on dE/dx and track pitch by comparing the dE/dx and pitch distributions in data and MC.
-\textcolor{blue}{ Currently, MPV MC = 1.70 and MPV DATA = 1.72 MeV/cm (~3\% higher).
-TO DO HERE: calculate Argon density from mid-RTD temperature. Compare this  density with MC Argon density. 
-Density change  affects dE/dx (in MeV/cm!). Try changing MC density up to ``real one" and see if dEdX agrees between DATA and MC}
-
-\subsection{Uncertainty on track end, aka efficiency correction}
-From the MC, we obtain an efficiency correction on the interacting and incident distributions separately.
-This is done by comparing the MC reconstructed with the true MC deposition on an event by event basis.
-This correction is applied bin by bin on the data interacting and incident distributions.
-The better our tracking, the smaller this efficiency correction will be. So, step number one is improving the tracking.
-\textcolor{blue}{Need to talk to Bruce about this.}
-\textcolor{blue}{ I don't understand the angle cut that Dave Schmitz and Jon Paley were so vocal about.}
-
-Now, the key question remains: does the tracking behave in the same way in data and MC? 
-We can compare some key plots between reconstructed data and MC which gives us confidence this is true: the track pitch, the tracks straightness and the goodness of fit in data and MC.
-\textcolor{blue}{ Does such a variable as ``goodness of fit" exists in the tracking?
-We should ask Bruce.}
-
-\section{Handling beamline contamination}
+\chapter{Background subtraction}
+\section{Assessing Beamline Contamination}
 What is the beamline contamination? We define beamline contamination every TPC track matched to the WC track which is not a primary pion. There are 4 different types of beamline contaminations:
@@
 
+\section{Subtraction}
 Once we estimate the contaminants to primary pion ratio, the next step is subtracting their contribution from data for each type of contaminant independently.
@@
 \begin{figure}
-\includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{Chapter-8/Images/FakePlot.jpg}
+\includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{Chapter-9/Images/FakePlot.jpg}
 \label{fig:backgroundSubtraction}

### 2018-03-19 Chapter-8/8-TrackingOptimization.tex
Changing organization structure

--- 2018-03-19 Chapter-6/6-TrackingOptimization.tex
@@
 \chapter{Tracking Optimization}
-Understanding how kaons and pions are tracked inside the TPC and optimizing the reconstruction algorithms to maximize the correct identification of the interaction point is a fundamental step of the analysis.
+In this chapter, we describe how we optimize the reconstruction algorithms to maximize the correct identification of the interaction point for hadrons inside the TPC and
 
@@
 We mimic the matching between the WC and the TPC track on Monte Carlo by constructing a fake WC track using truth information at wire chamber four.
-We then apply the same WC to TPC matching algorithm as in data.
+We then apply the same WC to TPC ma-8tching algorithm as in data.
 
@@
 
-\section{Wire chamber-to-TPC match}
-\subsection{Wire chamber-to-TPC match: importance and definition}
+\section{Optimizing the Wire Chamber to TPC match}\label{ch:WC2TPCMatchOptimization}
 \textcolor{red}{something something about why this match is important}
@@
 
-\subsection{Matching optimization}\label{ch:WC2TPCMatchOptimization}
+\subsection{Matching optimization}
 Scope of this optimization study is assessing the goodness of the wire chamber-to-TPC match on Monte Carlo and porting the optimized selection cuts to data.
@@
 \centering
-\includegraphics[width=15cm]{Chapter-6/Images/KEffPurity.png}
+\includegraphics[width=15cm]{Chapter-8/Images/KEffPurity.png}
 \caption{Efficiency (left) and purity (right) for wire chamber-to-TPC match as a function of the radius and angle selections.}

### 2018-03-19 Chapter-9/9-Uncertainty.tex
Changing organization structure

--- 2018-03-19 Chapter-8/8-Uncertainty.tex
@@
 We should ask Bruce.}
-
-\section{Handling beamline contamination}
-What is the beamline contamination? We define beamline contamination every TPC track matched to the WC track which is not a primary pion. There are 4 different types of beamline contaminations:
-\begin{itemize}
-\item[]1) electrons,
-\item[]2) muons,
-\item[]3) secondaries from pion events,
-\item[]4) matched pile up events.
-\end{itemize}
-
-So, how do we handle this contamination?
-
-The first step is to estimate what percentage of events used in the cross section calculation is not a primary pion.  
-We estimate the percentage of electrons and muons in the beam via the beamline MC\footnote{Since the beamline composition is a function of the magnet settings, we simulate separately events for magnet current of -60A and -100A.
-We calculate the electron to pion and muon to pion ratio on the whole sample as the weighted sum of the corresponding ratio in the two current settings, 
-\begin{equation}
-\frac{N_e}{N_\pi}_{Data} = w_{60A}\frac{N_e}{N_\pi}_{60A}  + w_{100A}\frac{N_e}{N_\pi}_{100A},
-\end{equation}
-\begin{equation}
-\frac{N_\mu}{N_\pi}_{Data} = w_{60A}\frac{N_\mu}{N_\pi}_{60A}  + w_{100A}\frac{N_\mu}{N_\pi}_{100A},
-\end{equation}
-where the weights $w_{60A}$ and $w_{100A}$ are the percentage of events in the corresponding magnet configuration passing the mass selection in data. }.
-Once the beam composition is know,  we simulate the electrons, muons and pions with the DDMC and we subject the three samples to the same selection chain (WC2TPC match, shower filter, pile up filter, etc...).
-The percentage of electrons and muons surviving the selection chain is the  electron and muon contamination in the pion cross section sample.
-The percentage of secondaries is given in the MC by the number of matched WC2TPC tracks which are not flagged as primary by Geant4.
-We estimate the last type of contamination, the ``matched pile up" events, to be a negligible fraction, because of the definition of the WC2TPC match: we deem the probability of a single match with a halo particle in the absence of a beamline particle\footnote{ Events with multiple WC2TPC matches are always rejected.} extremely small.
-
-Once we estimate the contaminants to primary pion ratio, the next step is subtracting their contribution from data for each type of contaminant independently.
-The contaminant samples are reconstructed and the corresponding interacting and incident histograms are produced.
-We then perform a bin by bin subtraction in the data interacting and incident histograms separately.
-A graphical rendering of this procedure is shown in Fig \ref{fig:backgroundSubtraction}
-Once the data is background subtracted, we apply the correction laid out in the previous section.
-\textcolor{blue}{How do we account for the error in the contamination subtraction? We change the electron/pion and muon/pion ratio and we see how much difference we get?}
-
-\begin{figure}
-\includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{Chapter-8/Images/FakePlot.jpg}
-\label{fig:backgroundSubtraction}
-\caption{A graphical rendering of the beamline contamination background subtraction.
-The contribution of the contaminants is shown in green for the secondaries, in orange for the muons and in pink for electrons.
-The colored plots are coming from the MC and are staggered.
-The percentages shown in the legend are the percentages of contaminants over the total number of events  passing the selection chain.
-We actually expect way less contamination.}
-\end{figure}

### 2018-03-19 Chapters/AppendixC_Pion.tex
Changing organization structure

\chapter{Pion Analysis}

\subsection{Data Sample}

We decided to use only the data from the  -60 A, -100 A RunII configurations, because the beam composition for these 2 beamline configurations is available in G4Beamline. Run II data is better than Run I in terms of calorimetry and understanding of the detector.

The used SamWeb definition is PionAna\_RunII\_60A\_100A\_lovely1\_elenag\_00  =

Defined as ``defname: Lovely1 and data\_tier digits and lariat\_mid\_f\_mc7anb < 0 and create\_date < '2017-06-02' and run\_number != 9276 and run\_number != 9277 and run\_number != 8836 and run\_number != 9083 and run\_number != 9122 and run\_number != 8977 and run\_number != 8981 and run\_number != 8982 and run\_number != 8983 and run\_number != 8984 and run\_number != 8985 and run\_number != 8991 and run\_number != 8994 and run\_number != 8996 and run\_number >= 8767 and
run\_number <= 9635"

\begin{table}[]
\centering
\caption{My caption}
\label{my-label}
\begin{tabular}{|c|c|c|c|}
\hline
Stage                                          & -100A +64 GeV & -60A +64 GeV & Number of Evt \\
\hline
SamWebDefinition                     & 32.7\%        & 67.3\% & 3569206      \\
WCQuality                                  & 37.8\%        & 63.2\% &  553486      \\
P$_{WC4} > $ 420 MeV/c               & 50.8\%        & 49.2\% &  423294      \\
TOF Cut                                     & 32.7\%        & 67.3\% &       \\
\hline
                                               &               &             &\\
\hline
\end{tabular}
\end{table}

\subsection{Capture and Decay}
Our goal is to measure the total hadronic cross section for negative pions in argon. Since pion capture can be classified as an electromagnetic process and pion decay is a week process,  capture and decay represent unwanted interactions in our sample. We present here a study of capture and decay in Monte Carlo and the solution we adopted to mitigate their present in our sample.

For this MC study, we use a sample of 359000 MC pions generated according to the beam profile with the DDMC described in \ref{sec:DDMC}. Unlike decay -- which may occur both in flight and at rest -- capture occurs predominantly
at rest. Thus, we can highly mitigate capture and decay at rest by removing pions whose energy is low enough to stop in the TPC. This translates into a momentum selection, where we keep only events whose WC momentum is above a certain threshold. 
Figure \ref{fig:CaptureMom} shows the true momentum distribution for the primary\footnote{We use here the Geant4 denomination "primary" to indicate that the pion considered does not undergo interactions modifying its energy before getting to the TPC. In fact, not every pion shot from wire chamber four will arrive to the TPC as primary,  some will decay or interact before the TPC, as explained in \textcolor{red}{reference to WC 2 TPC chapter}} pions that arrive to the TPC (pink), that capture (green) or decay (blue) inside the TPC, on a linear and log scale vertical axis.

\begin{figure}[hpbt]
\centering
\includegraphics[width=15cm]{Chapters/C-PionImages/CDAsMomentumFunct.png}
\caption{True momentum distribution at wire chamber 4 for every simulated pion arriving in the TPC (pink), ending its life in capture (green) or in decay (blue) in the TPC, linear vertical axis on the left, logarithmic on the right. }
\label{fig:CaptureMom}
\end{figure}

In order to choose the selection value for the wire chamber momentum, it is beneficial to estimate the ratio of events which capture or decay that survive the selection in MC as a function of the momentum threshold, and compare it with the survival ratio for all events. This is done in figure \ref{fig:survRatio}. We define the survival ratio simply  as the number of events surviving the true momentum cut divided by the number of events of that given category. The survival ratio calculated separately for the three event categories explained above: total (pink), capture (green) and decay (blue).
Selecting pions with momentum greater than 420 MeV/c reduces the capture events by ~99\% while maintaining about 80\% of the total data sample. 
Figure \ref{fig:evtRatio} shows the ratio of events which end their life in capture (green) or decay (blue) over the total number of events as a as a function of the true momentum at wire chamber four. This ratio is slightly dependent on the inelastic cross section implemented in Geant4, as we are able to register a pion capture (or decay) only if it doesn't interact inelastically in the TPC. For momenta greater than 420 MeV/c, the percentage of capture events drops below 1\% and the percentage of decays is never above 2\%.

\begin{figure}[h!]
\centering
\begin{minipage}[t]{0.45\textwidth}
\centering
\includegraphics[width=7.5cm]{Chapters/C-PionImages/CDThreshold.png}
\caption{Survival ratio as a function of selection threshold on true momentum at wire chamber four for for every simulated pion arriving in the TPC (pink), capture (green) or in decay (blue).   }
\label{fig:survRatio}
\end{minipage}\hfill
\begin{minipage}[t]{0.45\textwidth}
\centering
\includegraphics[width=7.5cm]{Chapters/C-PionImages/CDRatio.png}
\caption{Ratio between the capture (green) and decay (blue) events over the total number of events as a as a function of the true momentum at wire chamber four.}
\label{fig:evtRatio}
\end{minipage}
\end{figure}

### 2018-03-20 Chapter-2/2-LArTPCs.tex
s

@@
 \section{Enabling the next generation of discoveries: LArIAT}
+LArIAT, a small Liquid Argon Time Projection Chamber (LArTPC) in a test beam,  is designed to perform an extensive physics campaign centered on charged particle cross section measurements while characterizing the detector performance for future LArTPCs.
+LArTPC represents one of the most advanced experimental technologies for physics at the Intensity Frontier due to its full 3D-imaging, excellent particle identification and precise calorimetric energy reconstruction.
+This complex technology however needs a thorough calibration and dedicated measurements of some key quantities to achieve the precision required for the next generation of discoveries at the Intensity Frontier which LArIAT can provide.
+
+The LArIAT LArTPC is deployed in a dedicated calibration test beamline at Fermilab.
+We use the LArIAT beamline to characterize the charge particles before they enter the TPC: the particle type and initial momentum is known from beamline information.
+The precise calorimetric energy reconstruction of the LArTPC technology enables the measurement of the total differential cross section for  tagged hadrons.
+The Pion-Nucleus and Kaon-Nucleus total hadronic interaction cross section have never been measured before in argon and they are a fundamental step to shed light on light meson interaction in nuclei.
+Additionally, these measures provides a key input to neutrino physics and proton decay studies in future LArTPC experiments like SBN and DUNE.
+\textcolor{red}{add paragraph on all wonderful things lariat can do... some event displays would be nice!}

### 2018-03-20 Chapter-3/3-TheLArIATExperiment.tex
s

@@
 \chapter{LArIAT: Liquid Argon In A Testbeam}\label{sec:experimentDescription}
-
-LArIAT, a small Liquid Argon Time Projection Chamber (LArTPC) in a test beam,  is designed to perform an extensive physics campaign centered on charged particle cross section measurements while characterizing the detector performance for future LArTPCs.
-LArTPC represents one of the most advanced experimental technologies for physics at the Intensity Frontier due to its full 3D-imaging, excellent particle identification and precise calorimetric energy reconstruction.
-This complex technology however needs a thorough calibration and dedicated measurements of some key quantities to achieve the precision required for the next generation of discoveries at the Intensity Frontier which LArIAT can provide.
-
-The LArIAT LArTPC is deployed in a dedicated calibration test beamline at Fermilab.
-We use the LArIAT beamline to characterize the charge particles before they enter the TPC: the particle type and initial momentum is known from beamline information.
-The precise calorimetric energy reconstruction of the LArTPC technology enables the measurement of the total differential cross section for  tagged hadrons.
-The Pion-Nucleus and Kaon-Nucleus total hadronic interaction cross section have never been measured before in argon and they are a fundamental step to shed light on light meson interaction in nuclei.
-Additionally, these measures provides a key input to neutrino physics and proton decay studies in future LArTPC experiments like SBN and DUNE.
+In this chapter, we describe the LArIAT experimental setup.
+We start by illustrating the journey of the charge particles in the Fermilab accelerator complex, from gaseous thermal hydrogen in \textcolor{red}{find place where hydrogen is} to the delivery of the LArIAT tertiary beam to our experimental hall.
+We  then describe the LArIAT beamline detectors, the LArTPC, the DAQ and the monitoring system.
 
@@
 The particles are then bend by  $~10^\circ$  through a pair of dipole magnets.
-Tuning the magnets field intensity results in a range of particle momenta from 0.2 to 1.4~GeV/c.
+Tuning the magnets field intensity results in a range of particle's momenta from 0.2 to 1.4~GeV/c.
 The tertiary beam composition counts mostly pions and protons with a small fraction of electrons, muons, and kaons present as well.
@@
 The TDC time resolution is 1.18~ns/bin and can accept 2 edges per 9~ns.
-The maximum event rate acceptable by the chamber system is of 1 MHz: this rate is not a limiting factor considering that the rate of the tertiary particle beam at the first wire chamber is estimated to be less than 15 kHz.
-A full spill of data occurring once per supercycle is stored on the TDC board memory at once and read out by a controller specially designed controller.
+The maximum event rate acceptable by the chamber system is of 1 MHz: this rate is not a limiting factor considering that \textcolor{red}{the rate of the tertiary particle beam at the first wire chamber is estimated to be less than 15 kHz}.
+A full spill of data occurring once per supercycle is stored on the TDC board memory at once and read out by a specially designed controller.
 We use LVDS cables to carry both power and data between the controller and the TDCs and from the controller to the rest of the DAQ.
@@
 \subsubsection{Multi-Wire Proportional Chambers functionality}\label{sec:MWPCfunc}
-We use the wire chamber system together with the bending magnets to measure the particles momentum.
+We use the wire chamber system together with the bending magnets to measure the particle's momentum.
 
-In the simplest scenario, only one hit on each and every of the four wire chambers is recorder during a single readout of the detector systems.
-Thus, we use the hits position in the two wire chambers upstream to the magnets to form a trajectory before the bent and the hits position in the two wire chambers downstream to the magnets to form a trajectory after the bent.
+In the simplest scenario, only one hit on each and every of the four wire chambers is recorded during a single readout of the detector systems.
+Thus, we use the hit positions in the two wire chambers upstream of the magnets to form a trajectory before the bend, and the hit positions in the two wire chambers downstream of the magnets to form a trajectory after the bend.
 We use the angles in the XZ plane between the upstream and downstream trajectories  to calculate the z component of the momentum as follows:
@@
 
-where $B_{eff}$ is the effective maximum field in a square field approximation,  $L_{eff}$ is the effective length of both magnets (twice the effective length of one magnet) and  3.3~$c^{-1}$ the conversion factor from from T*m to MeV/c.
+where $B_{eff}$ is the effective maximum field in a square field approximation,  $L_{eff}$ is the effective length of both magnets (twice the effective length of one magnet), $\theta_{US}$ is the angle off the $z$ axis of the upstream trajectory, $\theta_{DS}$ is the angle off the $z$ axis of the downstream trajectory  and  3.3~$c^{-1}$ is the conversion factor from [T$\cdot$m] to [MeV/c].
 By using the hit positions on WC3 and WC4, we estimate the azimutal and polar angles of the particle trajectory, and we are able to calculate the other components of the momentum.
 
-This simple scenario can get complicated in two ways:  the presence of multiple hits in a single wire chamber or by the absence of hits in one (or more) wire chamber.
+This simple scenario can get complicated in two ways:  the presence of multiple hits in a single wire chamber or by the absence of hits in one (or more) wire chambers.
 The first complication is due to beam pile up, while the latter is due to wire chamber inefficiency.
-Even in the case of multiple hits, at most one wire chamber track is reconstructed per event; since the magnets bend particles only in the X direction, we assume the particle trajectory to be roughly constant in the YZ plane, thus we keep the combination of hits most coplanar on YZ plane.
-It is possible to reconstruct the particle momentum even in case of wire chamber inefficiency, if the information is missing in either of the two middle wire chambers (WC2 or WC3).
-%Assuming identical magnets, the upstream and downstream trajectories projected without bending should intersect at some point in the space between the magnets.
-Over a collection of many tracks, these points of intersection would trace a plane, therefore the term "midplane" is used.
-The completed half of the wire chamber track is projected to the midplane and the point of intersection is used with the incomplete half to finish the track.
+Even in the case of multiple hits, at most one wire chamber track is reconstructed per event.
+Since the magnets bend particles only in the X direction, we assume the particle trajectory to be roughly constant in the YZ plane, thus we keep the combination of hits which fit best with a straight line.
+Even if the information is missing in either of the two middle wire chambers (WC2 or WC3), it is still possible to reconstruct the particle's momentum.
+Under the assumption of identical magnets, we define a plane centered in the middle of the two magnets (called ``midplane") that the physical particles need to cross.
+We project the completed half of the wire chamber track to the midplane, assuming no bending, to find the point of intersection.
+We use this point to complete the other half of the wire chamber track and to calculate the reconstructed momentum  with these four points.
+To account for the lack of bending in our reconstructed wire chamber track, we apply to the calculated momentum  a correction obtained with a sample of 4-point, single hit tracks.
 
-\textcolor{red}{BETTER EXPLAINATION OF HIGH YIELD}
-\textcolor{red}{UNCERTAINTY ON MOMENTUM MEASUREMENT}
-Events satisfying the simplest scenario of one single hit in each of the four wire chambers form the so called ``Picky track" sample.
-We construct an other, higher statistics sample where we loose the requirements on single hit and wire chamber efficiency called "High Yield".
-For LArIAT Run II, the "High Yield" sample is
-\textcolor{red}{ HIGH YIELD increase in stat} the "Picky Tracks" sample.
-We use Picky tracks for calibration purposes: \textcolor{red}{ I DONT UNDERSTAND THIS Plotting the fractional error of the corrected three-point momentum compared to the known four-point momentum shows an uncertainty of approximately 2\%.}
+Events satisfying the simplest scenario of one single hit in each of the four wire chambers form the ``Picky track" sample.
+We construct another, higher statistics sample, where we loosen the requirements on single hit and wire chamber efficiency: the ``High Yield" sample.
+For LArIAT Run II, the High Yield sample is about three times the Picky Tracks statistics.
+We use Picky Tracks for calibration purposes, in particular to obtain the momentum correction for the 3-point tracks; this correction adds an uncertainty of approximately 2\% to the momentum calculation of the High Yield sample.
+
+\textcolor{red}{Four point track momentum uncertainty}
 
@@
 The LArIAT DAQ and trigger system governs the read out of all the many detectors forming LArIAT. 
-The CAEN V1495 module featuring a user-programmable FPGA  is the core of this system.
+The CAEN V1495 module and its user-programmable FPGA  are the core of this system.
 Every 10~ns, this module checks for matches between sixteen logical inputs and user-defined patterns in the trigger menu; if it finds a match for two consecutive clock ticks, that trigger fires.
 
-% converts signals from analog to digital, interfaces and synchronizes with the accelerator system, stores raw data on disk, transfers data to archival storage and conduct a first pass data quality monitoring.
+The beam instruments,  the cosmic ray taggers, and the light collection system provide NIM-standard logic pulse inputs to the trigger decision.
+We automatically log the trigger inputs configuration with the rest of the DAQ configuration at the beginning of each run.
 
-%A CAEN V1495 board is used to perform the trigger of the whole detectors system.
-It is a VME board, which can be directly customized by the user through two programmable FPGA.
-For LArIAT experiment 16 trigger input signals are used, see Fig.
-2.14.
-Through the FPGA configuration file  V1495.config  it is possible to define different trigger conditions - trigger output paths, with several logic combinations of the input signals.
-%The former trigger output path is the one that selects particles from the beam that cross the whole beamline and get to the TPC.
-%Moreover it's possible to add a delay to the trigger and to set some combinations of signals as a VETO.
-%Waveforms from beamline detectors and optical devices (PMTs and SiPMs) inside the cryostat are collected using two CAEN V1751 boards (For v1751 boards inputs, see Table 2.2.).
-They are both fed by the V1495 fast trigger.
-Each has a unit boardId and produces CAENFragment raw data objects.
-%The CAEN V1751 Waveform Digitizer [40] is VME module housing 8 Channel 10 bit 1 GS/s Waveform Digitizer with 1 Vpp input dynamic range; the range that has been set for our purposes is 204 mV, so 1 ADC count corresponds to 0.2 mV.
-The input channel provides a programmable voltage o set in the ± 0.5 V range too; we used this feature to adjust baseline o sets on our signals.
-The time window can also be adjusted by user maintaining the same sample %frequency (1 GS/s); we've started with a 7.168 ?s
-%time window, but after preliminary studies on the signals collected by PMTs and SiPMs we've moved to a 14.336 ?s window that could allow us to sample over all the slow LAr scintillation component (?S) signal tails, to study single photoelectron response of the optical devices.
+%%%%%%%%%%%%%%%%%%%%%%%%%
 
-%The LArIAT data acquisition system is triggered to read out the digitizing buffers by a CAEN V1495 module.
-The V1495 is a powerful, easily configurable coincidence module featuring a user-programmable FPGA.
-Sixteen logical inputs  are sampled on a 10~ns clock, checking for matches to any of the sixteen user-defined patterns in the trigger menu.
-If a given trigger pattern is satisfied for two consecutive clock ticks, that pattern fires.
-
-%NIM-standard logic pulse inputs to the trigger decision come from any of the instruments in the experiment, in principle; trigger inputs are derived from beam instruments, from the cosmic ray taggers, and from the cryostat's scintillation detectors.
-LArIAT takes full advantage of the trigger card's great flexibility.
-As conditions change and innovations are gradually made to LArIAT's data-taking process, improvements are made to the trigger input designs and the combinations of inputs which define trigger patterns for the V1495.
-The configuration of the trigger inputs is automatically logged in a database at the start of each run as part of the DAQ configuration, along with the rest of the XML configuration file for the V1495.
-
+%
 %\subsubsection{Inputs to the Trigger System}
@@
 An adjustable pulser is a fourth trigger input which does not depend on any particular activity in the experiment hall,  useful for collecting background events with zero bias.
+
+%%%%%%%%%%%%%%%%%%%%%%
 

### 2018-03-20 Chapter-7/7-BackgroundSub.tex
s

@@
 \end{figure}
+
+\section{Capture and decay}\label{ch:CaptureAndDecay}
+Our goal is to measure the total hadronic cross section for negative pions in argon.
+Since pion capture can be classified as an electromagnetic process and pion decay is a week process,  capture and decay represent unwanted interactions.
+We present here a study of capture and decay in Monte Carlo and the solution we adopted to mitigate their present in the data sample.
+
+For this MC study, we use a sample of 359000 MC pions generated according to the beam profile with the DDMC described in \ref{sec:DDMC}.
+It is important to notice that capture occurs predominantly at rest, while decay may occur both in flight and at rest.
+Thus, we can highly mitigate capture and decay at rest by removing pions which would release all their energy in the TPC and stop.
+This translates into a momentum selection, where we keep only events whose WC momentum is above a certain threshold.
+Figure \ref{fig:CaptureMom} shows the true momentum distribution for the primary\footnote{We use here the Geant4 denomination ``primary" to indicate that the pion considered does not undergo interactions modifying its energy before getting to the TPC.
+In fact, not every pion shot from wire chamber four will arrive to the TPC as primary,  some will decay or interact before the TPC.} pions that arrive to the TPC (pink), that capture (green) or decay (blue) inside the TPC, on a linear and log scale vertical axis.
+
+\begin{figure}[hpbt]
+\centering
+\includegraphics[width=15cm]{Chapter-7/Images/CDAsMomentumFunct.png}
+\caption{True momentum distribution at wire chamber 4 for every simulated pion arriving in the TPC (pink), ending its life in capture (green) or in decay (blue) in the TPC, linear vertical axis on the left, logarithmic on the right.
+}
+\label{fig:CaptureMom}
+\end{figure}
+
+In order to choose the selection value for the wire chamber momentum, it is beneficial to estimate the ratio of events which capture or decay that survive the selection in MC as a function of the momentum threshold, and compare it with the survival ratio for all events.
+This is done in figure \ref{fig:survRatio}.
+We define the survival ratio simply  as the number of events surviving the true momentum selection divided by the number of events of that category.
+We calculate the survival ratio separately for the three event categories explained above: total (pink), capture (green) and decay (blue).
+Selecting pions with momentum greater than 420 MeV/c reduces the capture events by ~99\% while maintaining about 80\% of the total data sample. 
+Figure \ref{fig:evtRatio} shows the ratio of events which end their life in capture (green) or decay (blue) over the total number of events as a as a function of the true momentum at wire chamber four.
+This ratio is slightly dependent on the inelastic cross section implemented in Geant4, as we are able to register a pion capture (or decay) only if it did not interact inelastically in the TPC.
+We choose a momentum threshold of 420~MeV/c because the percentage of capture events drops below 1\% and the percentage of decays is never above 2\% for momenta greater than 420~MeV/c.
+
+\begin{figure}[h!]
+\centering
+\begin{minipage}[t]{0.45\textwidth}
+\centering
+\includegraphics[width=7.5cm]{Chapter-7/Images/CDThreshold.png}
+\caption{Survival ratio as a function of selection threshold on true momentum at wire chamber four for for every simulated pion arriving in the TPC (pink), capture (green) or in decay (blue).   }
+\label{fig:survRatio}
+\end{minipage}\hfill
+\begin{minipage}[t]{0.45\textwidth}
+\centering
+\includegraphics[width=7.5cm]{Chapter-7/Images/CDRatio.png}
+\caption{Ratio between the capture (green) and decay (blue) events over the total number of events as a as a function of the true momentum at wire chamber four.}
+\label{fig:evtRatio}
+\end{minipage}
+\end{figure}

### 2018-03-20 Chapters/AppendixC_Pion.tex
s

@@
 \centering
-\includegraphics[width=15cm]{Chapters/C-PionImages/CDAsMomentumFunct.png}
+\includegraphics[width=15cm]{Chapter-7/Images/CDAsMomentumFunct.png}
 \caption{True momentum distribution at wire chamber 4 for every simulated pion arriving in the TPC (pink), ending its life in capture (green) or in decay (blue) in the TPC, linear vertical axis on the left, logarithmic on the right.
@@
 \centering
-\includegraphics[width=7.5cm]{Chapters/C-PionImages/CDThreshold.png}
+\includegraphics[width=7.5cm]{Chapter-7/Images/CDThreshold.png}
 \caption{Survival ratio as a function of selection threshold on true momentum at wire chamber four for for every simulated pion arriving in the TPC (pink), capture (green) or in decay (blue).   }
@@
 \centering
-\includegraphics[width=7.5cm]{Chapters/C-PionImages/CDRatio.png}
+\includegraphics[width=7.5cm]{Chapter-7/Images/CDRatio.png}
 \caption{Ratio between the capture (green) and decay (blue) events over the total number of events as a as a function of the true momentum at wire chamber four.}

### 2018-03-28 Chapter-7/7-BackgroundSub.tex
s

@@
 So, how do we handle this contamination?
+The first step is to estimate what percentage of events used in the cross section calculation is not a primary pion.
 
-The first step is to estimate what percentage of events used in the cross section calculation is not a primary pion.  
-We estimate the percentage of electrons and muons in the beam via the beamline MC\footnote{Since the beamline composition is a function of the magnet settings, we simulate separately events for magnet current of -60A and -100A.
+\subsection{Electron and Muon contamination}
+We estimate the percentage of electrons and muons in the beam via the beamline MC. 
+Since the beamline composition is a function of the magnet settings, we simulate separately events for magnet current of -60A and -100A.
+
+\begin{figure}
+\includegraphics[width=0.5\textwidth,height=\textheight,keepaspectratio]{Chapter-7/Images/Beam60A.png}
+\includegraphics[width=0.5\textwidth,height=\textheight,keepaspectratio]{Chapter-7/Images/Beam100A.png}
+\label{fig:BeamComposition}
+\caption{Beam composition for the -60A runs (left) and -100A runs (right).
+The solid blue plot represents the pion content, the yellow plot represents the muon content and the grey plot represents the electron content.
+The plots are area normalized to the number of data events, shown in red.
+}
+\end{figure}
+
+Table \ref{tab:beamline} shows the beam composition per magnet setting after the mass selection according to the G4simulation, 
+\begin{table}[]
+\centering
+\label{tab:beamline}
+\begin{tabular}{|l|c|c|}
+\hline
+                     & I = -60 A           & I = -100 A \\ \hline
+G4Pions       &   68.8 \%           &      87.4 \%        \\ \hline
+G4Muons     &     4.6 \%           &        3.7 \%         \\ \hline
+G4Electrons &   26.6 \%           &        8.9 \%        \\ \hline
+\end{tabular}
+\caption{Beamline composition per magnet settings}
+\end{table}
+
+\begin{table}[]
+\centering
+\label{tab:databreakdown}
+\begin{tabular}{|l|c|c|c|c|c|}
+\hline
+                                                     & I = -60 A          & I = -100 A   & Total     &  w$_{60A}$ &  w$_{100A}$\\ \hline
+Data events after Mass Cut         &     70192          &  76056       & 146248 & 0.52 & 0.48\\ \hline
+Data events for Cross Section     &                         &                   &              & & \\ \hline
+\end{tabular}
+\caption{Data events per magnet settings}
+\end{table}
+
 We calculate the electron to pion and muon to pion ratio on the whole sample as the weighted sum of the corresponding ratio in the two current settings, 
@@
 \end{equation}
-where the weights $w_{60A}$ and $w_{100A}$ are the percentage of events in the corresponding magnet configuration passing the mass selection in data. }.
+where the weights $w_{60A}$ and $w_{100A}$ are the percentage of events in the corresponding magnet configuration passing the mass selection in data, as shown in \ref{tab:databreakdown}.
+
 Once the beam composition is know,  we simulate the electrons, muons and pions with the DDMC and we subject the three samples to the same selection chain (WC2TPC match, shower filter, pile up filter, etc...).
 The percentage of electrons and muons surviving the selection chain is the  electron and muon contamination in the pion cross section sample.
+
+\subsection{Contamination from secondaries}
 The percentage of secondaries is given in the MC by the number of matched WC2TPC tracks which are not flagged as primary by Geant4.

### 2018-03-28 Chapter-8/8-TrackingOptimization.tex
s

@@
 \subsection{Porting optimization to data}
+
+\section{Tracking spatial and angular resolution}
+Scope of this study is understanding and comparing the tracking spatial and angular resolution on data and MC.
+We start by selecting all the WC2TPC matched tracks. 
+We fit a line on all the space points of the track and calculate the $\chi^2$. The $\chi^2$ distribution for data and MC is shown in Figure \ref{fig:Chi2AllPts}.
+
+For the spatial and angular resolution study, we reject tracks with less than 14 space points.
+For each track, we order the space points according to their Z position and we split them in two sets: the first set counts all the points belonging to the first half of the track and the second set counts all the points belonging to the second half of the track.
+We remove the last 5 points in the first set and the first 5 points in the second set, so to have a gap in the middle of the original track.
+We fit the first and the second set of points with a line separately.
+We reject the event entirely if the  $\chi^2$ for the fit of either of the halves is greater than four.
+We define a track middle plane as the plane perpendicular to the original track fit, positioned in the middle of its length.
+We project the tracks on the middle plane and calculate the impact parameter, $d$, i.e.
+the distance between the projected points.
+We also calculate the angle between the original track direction and the fit of the first and second half, called $\alpha_1$ and $\alpha_2$ respectively.
+The spatial resolution of the track will be $\sigma_S = \frac{d}{\sqrt 2}$ while the angular resolution of the tracks will be  $\sigma_\alpha = \alpha_1 - \alpha_2$.
+The distributions for data and MC for $\sigma_\alpha$ and $\sigma_S$ are given in \ref{fig:trackingResolution}.

### 2018-03-29 Chapter-3/3-TheLArIATExperiment.tex
s

@@
 \section{Trigger and DAQ}
-The LArIAT DAQ and trigger system governs the read out of all the many detectors forming LArIAT. 
+The LArIAT DAQ and trigger system governs the read out of all the many subsystems forming LArIAT. 
 The CAEN V1495 module and its user-programmable FPGA  are the core of this system.
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%
-%\subsubsection{Inputs to the Trigger System}
-
-%Two primary inputs to the trigger card are from the time of flight (see Sec.~\ref{sec:TOF}) and the wire chamber (see Sec.~\ref{sec:MWPC}) systems in the tertiary beamline.
-Coincidence of activity in both of these systems strongly suggests that a charged particle has made the journey down the tertiary beam line from the copper target to the LArTPC in the cryostat, and that we will have a measure of its momentum and its velocity.
-PMT pulses from the time of flight system each pass through a linear fan-out, one output of which is threshold discriminated by a
-%discriminator to produce a NIM logic pulse for use in trigger logic.
-On each of the upstream or downstream TOF paddles, we form a coincidence to within 20~ns of pulses from all the PMTs observing that same block of scintillator.
-The upstream paddle coincidence signal is delayed 20~ns to allow any approximately lightspeed particles to travel 6.5~m to the downstream paddle.
-The same upstream paddle coincidence signal is also widened to 100~ns to allow for slow-traveling (high-mass) particles.
-Coincidences of these upstream and downstream TOF trigger input signals can be made in the V1495 trigger card.
-
-%Each wire chamber is read out by four multi-hit TDCs (time to digital converters), with sixty-four wires per TDC and two TDCs per plane (one horizontal and one vertical).
-The TDCs each provide a logical ``fast'' OR of their inputs, indicating that one or more of their sixty-four wires went over the settable threshold.
-Using NIM logic units, the OR of the horizontal wires and the OR of the vertical wires are input to a coincidence unit for each wire chamber, providing a single logical pulse for each of the four wire chambers, indicating that at least one horizontal wire and one vertical wire saw significant signal.
-The single culminating logical pulses from each of the four wire chambers make up the first four inputs to the V1495 trigger card.
-Within the user-programmable FPGA, the V1495 looks for the coincidence within 20~ns of at least three of these four specially-treated logical inputs.
+Fundamental inputs to the trigger card come from the TOF (see Sec.~\ref{sec:TOF}) and the wire chambers (see Sec.~\ref{sec:MWPC}), as activity in these systems points to the presence of a charged particle in tertiary beam line.
+In particular, the discriminated pulses from the TOF PMTs form a NIM logic pulse for the trigger logic.
+We ask for a coincidence within a 20~ns window for all the pulses from the PMTs looking at the same scintillator block and use the coincidence int the upstream and downstream paddle to inform the trigger decision.
+In order to form a coincidence between the upstream and downstream paddles, we delay the upstream paddle coincidence by 20~ns and widen it by 100~ns.
+The delay and widening are necessary to account for both  lightspeed particles and slower particles (high-mass) to travel the 6.5~m between the upstream and the downstream paddles.
+Four multi-hit TDCs read out each wire chamber: two TDC per plane (horizontal and vertical), sixty-four wires per TDC.
+In each TDC, we keep the logical ``OR" for any signal over threshold from the sixty-four wires.
+We then require a coincidence between the ``OR" for the horizontal TDCs and the ``OR" for the vertical TDCs: with this logic we make sure that at least one horizontal wire and one vertical wire saw significant signal in one wire chamber.
+The single logical pulse from each of the four wire chambers feeds into the first four inputs to the V1495 trigger card.
+We require a coincidence within 20~ns of at least three logical inputs to form a trigger.
 

### 2018-03-29 Chapter-5/5-SamplesPreparation.tex
s

--- 2018-03-28 Chapter-8/8-TrackingOptimization.tex
@@
-\chapter{Tracking Optimization}
-In this chapter, we describe how we optimize the reconstruction algorithms to maximize the correct identification of the interaction point for hadrons inside the TPC and
+\chapter{Samples Preparation}\label{ch:samples}
 
-\section{MC sample and WC2TPC match}
-The optimization is performed on a MC sample of 191000 kaons and 359000 pions produced with the DDMC technique. DDMC particles are shot from the WC4 location into the TPC following the beam profile.
+\section{LArIAT Data}
+\section{LArIAT Monte Carlo}
+\subsection{G4Beamline}\label{beamlineComposition}
+\subsection{Data Driven MC}\label{sec:DDMC}
+
+\section{Energy Calibration}
+\section{Tracking Optimization}
+In this section, we describe two studies.
+The first is a description of the selection criteria for the beamline handshake with the TPC information.
+We perform this study to maximize  the correct identification of the particles in the TPC associated with the beamline information, while maintaining sufficient statistics for the cross section measurement.
+The second study is an optimization of the tracking algorithm, with the scope of maximizing the identification of the hadronic interaction point inside the TPC.
+
+%\section{MC sample and WC2TPC match}
+We perform the following studies on a MC sample of 191000 kaons and 359000 pions produced with the DDMC technique.
+DDMC particles are shot from the WC4 location into the TPC following the beam profile.
 We mimic the matching between the WC and the TPC track on Monte Carlo by constructing a fake WC track using truth information at wire chamber four.
-We then apply the same WC to TPC ma-8tching algorithm as in data.
+We then apply the same WC to TPC matching algorithm as in data described in\ref{sec:WC2TPC}.
 
@@
 \end{enumerate}
-
-\section{Optimizing the Wire Chamber to TPC match}\label{ch:WC2TPCMatchOptimization}
-\textcolor{red}{something something about why this match is important}
 
@@
 
-\subsection{Matching optimization}
-Scope of this optimization study is assessing the goodness of the wire chamber-to-TPC match on Monte Carlo and porting the optimized selection cuts to data.
+\subsection{Selection Study for the Wire Chamber to TPC Match}\label{ch:WC2TPCMatchOptimization}
+Scope of this study is assessing the goodness of the wire chamber to TPC match on Monte Carlo and porting the optimized selections to data.
 A word of caution is necessary here.
-With this study, we want to minimize pathologies associated with the presence of the primary kaon itself, like the incorrect association between the beamline kaon its decay products inside the TPC.
+With this study, we want to minimize pathologies associated with the presence of the primary hadron itself, e.g.
+the incorrect association between the beamline kaon its decay products inside the TPC.
 Assessing the contamination from pile-up\footnote{We remind the reader that the DDMC is a single particle Monte Carlo, where the beam pile up is not simulated.}, albeit related, is beyond the scope of this study.
@@
 
-%We define two populations: TPC tracks correctly matched and one of all the other match.
 We define a correct match if the all following conditions are met:
@@
 \centering
-\includegraphics[width=15cm]{Chapter-8/Images/KEffPurity.png}
+\includegraphics[width=15cm]{Chapter-5/Images/KEffPurity.png}
 \caption{Efficiency (left) and purity (right) for wire chamber-to-TPC match as a function of the radius and angle selections.}
@@
 
-\subsection{Porting optimization to data}
+\subsection{Interaction Point Optimization}\label{ch:TrackingOptimization}
 
-\section{Tracking spatial and angular resolution}
+\subsection{Tracking spatial and angular resolution}
 Scope of this study is understanding and comparing the tracking spatial and angular resolution on data and MC.

### 2018-03-30 Chapter-3/3-TheLArIATExperiment.tex
s

@@
+% This chapter to do:
+% Re-write the red parts
+% write DAQ part
+% write collimators description
+% put references
+% re-read
+
 \chapter{LArIAT: Liquid Argon In A Testbeam}\label{sec:experimentDescription}
 In this chapter, we describe the LArIAT experimental setup.
-We start by illustrating the journey of the charge particles in the Fermilab accelerator complex, from gaseous thermal hydrogen in \textcolor{red}{find place where hydrogen is} to the delivery of the LArIAT tertiary beam to our experimental hall.
+We start by illustrating the journey of the charge particles in the Fermilab accelerator complex, from the gaseous thermal hydrogen at the ion source to the delivery of the LArIAT tertiary beam at MC7.
 We  then describe the LArIAT beamline detectors, the LArTPC, the DAQ and the monitoring system.
@@
 
-LArIAT's home at Fermilab is the Fermilab Test Beam Facility (FTBF), where the experiment characterizes a beam of charge particles downstream from the Meson Center beam line.
-
 LArIAT's particles history begins in the Fermilab accelerator complex with a beam of protons.
-The process of protons acceleration develops in gradual stages, see picture \ref{fig:Accelerator}: gaseous hydrogen is ionized in order to form H$^{-}$ ions; these ions are boosted to 750 keV by a Cockroft-Walton accelerator and injected to the Linac linear accelerator that increases their energy up to 400 MeV; then, H$^{-}$ ions pass through a carbon foil and lose the two electrons; the resulting protons are then injected into a rapid cycling synchrotron, called Booster; at this stage protons reach 8 GeV of energy and are compacted into bunches; the next stage of acceleration is the Main Injector, a synchrotron which accelerates the bunches up to 120 GeV; in the Main Injector, several bunches are merged into one and used for the injection in the last stage.
+The process of protons acceleration develops in gradual stages (see picture \ref{fig:Accelerator}): gaseous hydrogen is ionized in order to form H$^{-}$ ions; these ions are boosted to 750 keV by a Cockroft-Walton accelerator and injected to the Linac linear accelerator that increases their energy up to 400 MeV; then, H$^{-}$ ions pass through a carbon foil and lose the two electrons; the resulting protons are then injected into a rapid cycling synchrotron, called Booster; at this stage protons reach 8 GeV of energy and are compacted into bunches; the next stage of acceleration is the Main Injector, a synchrotron which accelerates the bunches up to 120 GeV; in the Main Injector, several bunches are merged into one and used for the injection in the last stage.
 
 The Fermilab accelerator complex works in supercycles of roughly 60 seconds in duration.
-The beam is then split by electrostatic septa and delivered at different experimental halls all over the lab.
+The beam is split by electrostatic septa and delivered at different experimental halls all over the lab.
 A 120~GeV$/c$ primary proton beam with variable intensity is extracted in four-second ``spills" and sent to the Meson Center beam line.
-Here, this primary beam is focused onto a tungsten target to create LArIAT's secondary beam.
+
+LArIAT's home at Fermilab is the Fermilab Test Beam Facility (FTBF), where the experiment characterizes a beam of charge particles downstream from the Meson Center beam line. 
+Here, the primary beam is focused onto a tungsten target to create LArIAT's secondary beam.
 The composition of the secondary particle beam is mainly positive pions.
-For the LArIAT data considered in this work the secondary beam peak momentum was fixed at 64~GeV$/c$, although the beam is tunable in momentum between 8-80\,GeV$/c$; this was deemed a secondary beamline configuration which allowed a stable beam operation of the FTBF.
-The beam of pions impinges then on a copper target within a steel collimator inside the LArIAT experimental hall (MC7) to create LArIAT tertiary beam, whose geometry in MC7 has been optimized for LArIAT (shown in  Fig.~\ref{fig:tert-layout}).
+For the LArIAT data considered in this work the secondary beam peak momentum was fixed at 64~GeV$/c$, although the beam is tunable in momentum between 8-80\,GeV$/c$; this was deemed a secondary beamline configuration which allowed a stable beam operation for LArIAT.
+The secondary beam of pions impinges then on a copper target within a steel collimator inside the LArIAT experimental hall (MC7) to create the LArIAT tertiary beam, whose geometry in MC7 has been optimized for LArIAT (shown in  Fig.~\ref{fig:tert-layout}).
 The steel collimator selects particles produced with a $13^\circ$ production angle at the target down the beamline.
-The particles are then bend by  $~10^\circ$  through a pair of dipole magnets.
-Tuning the magnets field intensity results in a range of particle's momenta from 0.2 to 1.4~GeV/c.
-The tertiary beam composition counts mostly pions and protons with a small fraction of electrons, muons, and kaons present as well.
+The particles are then bent by  $~10^\circ$  through a pair of dipole magnets.
+The configuration of the magnets' field intensity allows the particles of LArIAT's tertiary beam to span a momentum range from 0.2 to 1.4~GeV/c.
+The polarity of the magnet is also configurable and determines the sign of the beamline particles which are focused on the LArTPC.
+In the positive magnets polarity, the tertiary beam composition counts mostly pions and protons with a small fraction of electrons, muons, and kaons.
 It is the job of the LArIAT beamline detectors to select the particles polarity,  to perform particle identification (beamPID) and to measure the momentum of the tertiary beam particles before they get to the LArTPC.
@@
 
-The components of the tertiary beamline instrumentation key for the hadron cross section analyses are the target and collimators system, the two bending magnets (in a similar configuration used for the  MINERvA T-977 test beam calibration~\cite{MinervaTestbeam}) a set of four wire chambers (WCs) and two time-of-flight scintillating paddles (TOF) and, of course, the LArTPC.
-The magnets determine the polarity of the particles in the tertiary beam; the combination of magnets and wire chambers determine the particles momentum, which is used to determine the particle species in conjunction with the TOF.
+The key components of the tertiary beamline instrumentation for the hadron cross section analyses are the two bending magnets (in a similar configuration used for the  MINERvA T-977 test beam calibration~\cite{MinervaTestbeam}), a set of four wire chambers (WCs) and two time-of-flight scintillating paddles (TOF) and, of course, the LArTPC.
+The magnets determine the polarity of the particles in the tertiary beam; the combination of magnets and wire chambers determines the particles' momentum, which is used to determine the particle species in conjunction with the TOF.
 A muon range stack downstream from the TPC and two sets of cosmic paddles configured as a telescope surrounding the TPC are also used for calibration purposes.
-
-\subsection{Collimators}\label{sec:Collimators}
 

### 2018-03-30 Chapter-7/7-BackgroundSub.tex
s

@@
                                                      & I = -60 A          & I = -100 A   & Total     &  w$_{60A}$ &  w$_{100A}$\\ \hline
-Data events after Mass Cut         &     70192          &  76056       & 146248 & 0.52 & 0.48\\ \hline
+Data events after Mass Cut         &     70192          &  76056       & 146248 & 0.48 & 0.52\\ \hline
 Data events for Cross Section     &                         &                   &              & & \\ \hline
+Estimated e number                     &   18671.1         &    6769.0  &              & & \\ \hline
+Estimated $\mu$ number             &     3228.8         &     2814.1     &              & & \\ \hline
+Estimated $\pi$ number               &    48292.1        &   66472.9   &              & & \\ \hline
 \end{tabular}

### 2018-04-02 Chapter-3/3-TheLArIATExperiment.tex
s

@@
 In this chapter, we describe the LArIAT experimental setup.
-We start by illustrating the journey of the charge particles in the Fermilab accelerator complex, from the gaseous thermal hydrogen at the ion source to the delivery of the LArIAT tertiary beam at MC7.
+We start by illustrating the journey of the charge particles in the Fermilab accelerator complex, from the gaseous thermal hydrogen at the Fermilab ion source to the delivery of the LArIAT tertiary beam at MC7.
 We  then describe the LArIAT beamline detectors, the LArTPC, the DAQ and the monitoring system.
@@
 LArIAT's particles history begins in the Fermilab accelerator complex with a beam of protons.
-The process of protons acceleration develops in gradual stages (see picture \ref{fig:Accelerator}): gaseous hydrogen is ionized in order to form H$^{-}$ ions; these ions are boosted to 750 keV by a Cockroft-Walton accelerator and injected to the Linac linear accelerator that increases their energy up to 400 MeV; then, H$^{-}$ ions pass through a carbon foil and lose the two electrons; the resulting protons are then injected into a rapid cycling synchrotron, called Booster; at this stage protons reach 8 GeV of energy and are compacted into bunches; the next stage of acceleration is the Main Injector, a synchrotron which accelerates the bunches up to 120 GeV; in the Main Injector, several bunches are merged into one and used for the injection in the last stage.
+The process of protons acceleration develops in gradual stages (see picture \ref{fig:Accelerator}): gaseous hydrogen is ionized in order to form H$^{-}$ ions; these ions are boosted to 750 keV by a Cockroft-Walton accelerator and injected to the Linac linear accelerator that increases their energy up to 400 MeV; then, H$^{-}$ ions pass through a carbon foil and lose the two electrons; the resulting protons are then injected into a rapid cycling synchrotron, called Booster; at this stage, protons reach 8 GeV of energy and are compacted into bunches; the next stage of acceleration is the Main Injector, a synchrotron which accelerates the bunches up to 120 GeV; in the Main Injector, several bunches are merged into one and used for the injection in the last stage.
 
@@
 The composition of the secondary particle beam is mainly positive pions.
-For the LArIAT data considered in this work the secondary beam peak momentum was fixed at 64~GeV$/c$, although the beam is tunable in momentum between 8-80\,GeV$/c$; this was deemed a secondary beamline configuration which allowed a stable beam operation for LArIAT.
-The secondary beam of pions impinges then on a copper target within a steel collimator inside the LArIAT experimental hall (MC7) to create the LArIAT tertiary beam, whose geometry in MC7 has been optimized for LArIAT (shown in  Fig.~\ref{fig:tert-layout}).
+The momentum peak of the secondary beam was fixed at 64~GeV/c for the LArIAT data considered in this work, although the beam is tunable in momentum between 8-80\,GeV/c; this configuration of the secondary beamline assured a stable beam delivery at the LArIAT experimental hall.
+
+The secondary beam impinges then on a copper target within a steel collimator inside the LArIAT experimental hall (MC7) to create the LArIAT tertiary beam, (shown in  Fig.~\ref{fig:tert-layout}).
 The steel collimator selects particles produced with a $13^\circ$ production angle at the target down the beamline.
 The particles are then bent by  $~10^\circ$  through a pair of dipole magnets.
-The configuration of the magnets' field intensity allows the particles of LArIAT's tertiary beam to span a momentum range from 0.2 to 1.4~GeV/c.
+By configuring the field intensity of the magnets we allow the particles of LArIAT's tertiary beam to span a momentum range from 0.2 to 1.4~GeV/c.
 The polarity of the magnet is also configurable and determines the sign of the beamline particles which are focused on the LArTPC.
-In the positive magnets polarity, the tertiary beam composition counts mostly pions and protons with a small fraction of electrons, muons, and kaons.
+If the magnets polarity is positive the tertiary beam composition counts mostly pions and protons with a small fraction of electrons, muons, and kaons.
 It is the job of the LArIAT beamline detectors to select the particles polarity,  to perform particle identification (beamPID) and to measure the momentum of the tertiary beam particles before they get to the LArTPC.
@@
 \includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{Chapter-3/Images/AcceleratorFNAL.png}
+\caption{Layout of Fermilab Acellerator complex.}
 \label{fig:Accelerator}
-\caption{Layout of Fermilab Acellerator complex.}
 \end{figure}
@@
 \includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{Chapter-3/Images/Tertiary.png}
+\caption{Bird's eye view of the LArIAT tertiary beamline.
+In grey: upstream and downstream collimators; in yellow: bending magnets; in red: wire chambers; in blue: time of flight; in green: liquid argon TPC volume; in maroon: muon range statck.}
 \label{fig:tert-layout}
-\caption{Birds eye view of the LArIAT tertiary beamline.
-In grey: upstream and downstream collimators; in yellow: bending magnets in yellow; in red: wire chambers; in blue: time of flight; in green: liquid argon TPC volume; in maroon: muon range statck.}
 \end{figure}
@@
 The instrumentation of  LArIAT tertiary beam and the TPC components have changed several times during the three years of LArIAT data taking.
-The following paragraphs describe the components operational during the data taking period relevant to the hadron cross section measurements (a.k.a.
-Run II).
+The following paragraphs describe the components operational during ``Run II", the data taking period relevant to the hadron cross section measurements.
 
-The key components of the tertiary beamline instrumentation for the hadron cross section analyses are the two bending magnets (in a similar configuration used for the  MINERvA T-977 test beam calibration~\cite{MinervaTestbeam}), a set of four wire chambers (WCs) and two time-of-flight scintillating paddles (TOF) and, of course, the LArTPC.
+The key components of the tertiary beamline instrumentation for the hadron cross section analyses are the two bending magnets, a set of four wire chambers (WCs) and two time-of-flight scintillating paddles (TOF) and, of course, the LArTPC.
 The magnets determine the polarity of the particles in the tertiary beam; the combination of magnets and wire chambers determines the particles' momentum, which is used to determine the particle species in conjunction with the TOF.
@@
 \subsection{Bending Magnets}\label{sec:Magnets}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-LArIAT uses a pair of identical Fermilab type ``NDB" electromagnets, recycled from the Tevatron's anti-proton ring \textcolor{red}{CITE CDF?}. 
-The magnets are a fundamental piece of the LArIAT beamline as they are used in all the three tasks of the LArIAT beamline: the sign of the current in the magnet provide the selection of either positively or negatively charged particles, the value of the magnetic field is used in the momentum determination and the subsequent particle identification.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-We describe here the characteristics and response of one magnet.
-We expect the second one have a similar response, being identical in shape and with a similar history.
-The magnet aperture measures the gap dimensions to be 14.224~cm of height, 31.75~cm width, and  46.67~cm length.
-The wire chambers aperture ($\sim$12.5~cm) is smaller than the magnet aperture, thus, only the central part of the magnet gap is utilized.
-The field is extremely uniform over this limited aperture and was measured with two Hall probes, both calibrated with nuclear magnetic resonance probes.
+LArIAT uses a pair of identical Fermilab type ``NDB" electromagnets, recycled from the Tevatron's anti-proton ring, in a similar configuration used for the  MINERvA T-977 test beam calibration~\cite{MinervaTestbeam}).
+The magnets are a fundamental piece of the LArIAT beamline equipment, as they are used for both particle identification and momentum measurement before the LArTPC.
+The sign of the current in the magnets allows us to select either positively or negatively charged particles; the value of the magnetic field is used in the momentum determination and in the subsequent particle identification.
+
+We describe here the characteristics and response of one magnet, as the second one has a similar response, given its identical shape and history.
+Each magnet is a box with a rectangular aperture gap in the center to allow for the particle passage.
+The magnet aperture measures 14.224~cm in height, 31.75~cm in width, and  46.67~cm in length.
+Since the wire chambers aperture ($\sim$12.8~cm$^2$) is smaller than the magnet aperture, only the central part of the magnet gap is utilized.
+The field is extremely uniform over this limited aperture and was measured with two hall probes, both calibrated with nuclear magnetic resonance probes.
 The probes measured the excitation curve shown in Figure~\ref{fig:magnet_excitation}.
@@
 
-The current being passed through the magnets at a given time is identical in both magnets.
-For Run II period, the current settings explored were 60A (B $\sim$0.21 T) and 100A (B $\sim$0.35 T) in both polarities.
+The current through the magnets at a given time is identical in both magnets.
+For the Run II data taking period, the current settings explored were 60A (B $\sim$0.21 T) and 100A (B $\sim$0.35 T) in both polarities.
 Albeit advantageous to enrich the tertiary beam composition with high mass particles such as kaons, we never pushed the magnets current over 100 A, not to incur in overheating.
@@
 
-LArIAT uses four Multi-Wire proportional chambers, or wire chambers (WC) for short, two upstream and two downstream from the bending magnets.
+LArIAT uses four multi-wire proportional chambers, or wire chambers (WC) for short, two upstream and two downstream from the bending magnets.
 The geometry of one chamber is shown in Figure~\ref{fig:wirechamber}: the WC effective aperture is a square of  12.8~cm perpendicular to the beam direction.
 Inside the chamber, the 128 horizontal and 128 vertical wires hang at a distance of 1~mm from each other in a mixture of 85\% Argon and 15\% isobutane gas.
-The WC operating voltage is between 2400 V and 2500 V.
-LArIAT wire chambers are an upgraded version of the Fenker Chambers~\cite{Fenker}, where an extra grounding improves the signal to noise ratio of the electronic readout.
+The WC operating voltage is between 2400~V and 2500~V.
+The LArIAT wire chambers are an upgraded version of the Fenker Chambers~\cite{Fenker}, where an extra grounding improves the signal to noise ratio of the electronic readout.
 
@@
 Thus, we use the hit positions in the two wire chambers upstream of the magnets to form a trajectory before the bend, and the hit positions in the two wire chambers downstream of the magnets to form a trajectory after the bend.
-We use the angles in the XZ plane between the upstream and downstream trajectories  to calculate the z component of the momentum as follows:
+We use the angles in the XZ plane between the upstream and downstream trajectories  to calculate the $Z$ component of the momentum as follows:
 
@@
 where $B_{eff}$ is the effective maximum field in a square field approximation,  $L_{eff}$ is the effective length of both magnets (twice the effective length of one magnet), $\theta_{US}$ is the angle off the $z$ axis of the upstream trajectory, $\theta_{DS}$ is the angle off the $z$ axis of the downstream trajectory  and  3.3~$c^{-1}$ is the conversion factor from [T$\cdot$m] to [MeV/c].
-By using the hit positions on WC3 and WC4, we estimate the azimutal and polar angles of the particle trajectory, and we are able to calculate the other components of the momentum.
+By using the hit positions on the third and fourth wire chamber, we estimate the azimutal and polar angles of the particle trajectory, and we are able to calculate the other components of the momentum.
 
-This simple scenario can get complicated in two ways:  the presence of multiple hits in a single wire chamber or by the absence of hits in one (or more) wire chambers.
+The presence of multiple hits in a single wire chamber or the absence of hits in one (or more) wire chambers can complicate this simple scenario.
 The first complication is due to beam pile up, while the latter is due to wire chamber inefficiency.
-Even in the case of multiple hits, at most one wire chamber track is reconstructed per event.
+In the case of multiple hits on a single WC, at most one wire chamber track is reconstructed per event.
 Since the magnets bend particles only in the X direction, we assume the particle trajectory to be roughly constant in the YZ plane, thus we keep the combination of hits which fit best with a straight line.
-Even if the information is missing in either of the two middle wire chambers (WC2 or WC3), it is still possible to reconstruct the particle's momentum.
-Under the assumption of identical magnets, we define a plane centered in the middle of the two magnets (called ``midplane") that the physical particles need to cross.
+It is still possible to reconstruct the particle's momentum  even if the information is missing in either of the two middle wire chambers (WC2 or WC3), by constraining the particle trajectory to cross the plane in between the magnets.
+%Under the assumption of identical magnets, we define a plane centered in the middle of the two magnets (called ``midplane") that the physical particles need to cross.
 We project the completed half of the wire chamber track to the midplane, assuming no bending, to find the point of intersection.
@@
 
-Events satisfying the simplest scenario of one single hit in each of the four wire chambers form the ``Picky track" sample.
+Events satisfying the simplest scenario of one single hit in each of the four wire chambers form the ``Picky Track" sample.
 We construct another, higher statistics sample, where we loosen the requirements on single hit and wire chamber efficiency: the ``High Yield" sample.
 For LArIAT Run II, the High Yield sample is about three times the Picky Tracks statistics.
-We use Picky Tracks for calibration purposes, in particular to obtain the momentum correction for the 3-point tracks; this correction adds an uncertainty of approximately 2\% to the momentum calculation of the High Yield sample.
+For the first measurements of the LArIAT hadronic cross section, we use the Picky Tracks sample because the uncertainty on the momentum is smaller and the comparison with the beamline MC results is straightforward compared with the High Yield sample;  a possible future update and cross check of these analysis would be the use of the High Yield sample.
+
+%We use Picky Tracks to calibrate the momentum measurement for the High Yield sample, in particular to obtain a momentum correction for  tracks missing information from the central WCs;  this correction adds an uncertainty of approximately 2\% to the momentum calculation of the High Yield sample.
 
@@
 
- \textcolor{red}{find exact dimentsions}
-The upstream paddle is made of a 10 x 6 x \textcolor{red}{?}~cm scintillator piece, read out by two \textcolor{red}{Hamamatsu 2 inch} PMTs mounted on the beam left side which collect the light from light guides mounted on all four edges of the scintillator.
-The downstream paddle dimensions are  14 x 14 x \textcolor{red}{?}~cm and it is read out by two \textcolor{red}{Hamamatsu 2 inch} PMTs on the opposite ends of the scintillator.
+The upstream paddle is made of a 10 x 6 x 1~cm scintillator piece, read out by two PMTs mounted on the beam left side which collect the light from light guides mounted on all four edges of the scintillator.
+The downstream paddle is a   14 x 14 x 1~cm scintillator piece read out by two PMTs on the opposite ends of the scintillator.
 The relatively thin width on the beamline direction minimizes energy loss of the particles coming from the target in the scintillator material.

### 2018-04-02 Chapter-5/5-SamplesPreparation.tex
s

@@
 \section{Energy Calibration}
-\section{Tracking Optimization}
-In this section, we describe two studies.
-The first is a description of the selection criteria for the beamline handshake with the TPC information.
-We perform this study to maximize  the correct identification of the particles in the TPC associated with the beamline information, while maintaining sufficient statistics for the cross section measurement.
+\section{Tracking Studies}
+In this section, we describe three studies.
+The first is a justification of the selection criteria for the beamline handshake with the TPC information.
+We perform this study to boost  the correct identification of the particles in the TPC associated with the beamline information, while maintaining sufficient statistics for the cross section measurement.
 The second study is an optimization of the tracking algorithm, with the scope of maximizing the identification of the hadronic interaction point inside the TPC.
+These two studies are related, since the optimization of the tracking is performed on TPC tracks which have been matched to the wire chamber track; in turn, the tracking algorithm for TPC tracks determine the number of reconstructed tracks in each event used to try the matching with the wire chamber track.
+Starting with a sensible tracking reconstruction, we perform the WC2TPC matching optimization first, then the tracking optimization.
+The WC2TPC match purity and efficiency  are then calculated again with the optimized tracking.
 
@@
 We mimic the matching between the WC and the TPC track on Monte Carlo by constructing a fake WC track using truth information at wire chamber four.
-We then apply the same WC to TPC matching algorithm as in data described in\ref{sec:WC2TPC}.
-
-Plots I want in this section:
-\begin{enumerate}
-\item WC2TPC MC DeltaX, DeltaY and $\alpha$
-\item Delta L, reco - true
-\item Delta L, reco - true Elastic, Delta L, reco - true Inelastic, other
-\item Length Quality cut
-\item Efficiency as a function of true KE and Angle
-\end{enumerate}
+We then apply the same WC to TPC matching algorithm as in data described in \ref{ch:WC2TPCMatchMethod}.
 
@@
 \subsection{Selection Study for the Wire Chamber to TPC Match}\label{ch:WC2TPCMatchOptimization}
-Scope of this study is assessing the goodness of the wire chamber to TPC match on Monte Carlo and porting the optimized selections to data.
+Plots I want in this section:
+\begin{enumerate}
+\item WC2TPC MC DeltaX, DeltaY and $\alpha$
+\end{enumerate}
+
+Scope of this study is assessing the goodness of the wire chamber to TPC match on Monte Carlo and decide the selection values we will use on data.
 A word of caution is necessary here.
 With this study, we want to minimize pathologies associated with the presence of the primary hadron itself, e.g.
-the incorrect association between the beamline kaon its decay products inside the TPC.
+the incorrect association between the beamline hadron and its decay products inside the TPC.
 Assessing the contamination from pile-up\footnote{We remind the reader that the DDMC is a single particle Monte Carlo, where the beam pile up is not simulated.}, albeit related, is beyond the scope of this study.
@@
 Events with true length in TPC $<$ 2 cm are included.
-Since kaons are shot 100 cm upstream from the TPC front face, the following two scenarios are possible from a truth standpoint: 
+Since hadrons are shot 100 cm upstream from the TPC front face, the following two scenarios are possible from a truth standpoint: 
 \begin{itemize}
-\item[[$Ta$]] the primary kaon decays or interact strongly before getting to the TPC,
-\item[[$Tb$]] the primary kaon enters the TPC.
+\item[[$Ta$]] the primary hadron decays or interact strongly before getting to the TPC,
+\item[[$Tb$]] the primary hadron enters the TPC.
 \end{itemize}
 
-Once we choose the selection cuts to determine a reconstructed wire chamber-to-TPC match $r_{T}$ and $\alpha_{T}$, the following four scenarios are possible : 
+Once we choose the selection cuts to determine a reconstructed wire chamber-to-TPC match $r_{T}$ and $\alpha_{T}$, the following five scenarios are possible in the truth to reconstruction interplay : 
 \begin{itemize}
-\item[1)] no reconstructed tracks are matched
-\item[2)] the correct track and one (or more) wrong tracks are matched
-\item[3)] only the correct track is matched
-\item[4)] one (or more) wrong track is matched.
+\item[1)] only the correct track is matched
+\item[2)] only one wrong track is matched 
+\item[3)] the correct track and one (or more) wrong tracks are matched
+\item[4)] multiple wrong tracks  matched.
+\item[5)] no reconstructed tracks are matched
 \end{itemize}
 
-If we keep only events with one and only one match, we discard cases 1) and 2) from the cross section measurement.
+Since we keep only events with one and only one match, we discard cases 3), 4) and 5) from the events used in the cross section measurement.
 For each set of $r_{T}$ and $\alpha_{T}$ selection value, we define purity and efficiency of the selection as follows:
@@
 It is apparent how both efficiency and purity are fairly flat as a function of the radius selection value at a given angle.
-This is not surprising, given the fact that the wrong matches can occur  in a single particle gun MC  only for mis-tracking of the primary or for association with decay products, which are generally different in angle, but might be more similar in $x$ and $y$ projection.
+This is not surprising.
+Since we are studying a single particle gun Monte Carlo sample, the wrong matches can occur only for mis-tracking of the primary or for association with decay products;  decay products will tend to be produced at large angles compared to the primary, but could be fairly close to the in $x$ and $y$ projection of the primary.
 The radius cut would play a key role in removing pile up events.
 
-For LArIAT cross section measurements, we generally prefer purity over efficiency, since a sample of particle of a pure species will lead to a better measurement.
-In the case of kaons however, efficiency needs not to drop too low, given the smallness of the kaon sample.
-We choose $(\alpha_{T}$, $r_{T}) = (8 \text{ deg}, 4 \text{ cm} )$ and get a MC 85\% efficiency and 98\% purity.
+For LArIAT cross section measurements, we generally prefer purity over efficiency, since a sample of particles of a pure species will lead to a better measurement.
+Obviously, purity should be balanced with a sensible efficiency to avoid rejecting the whole sample.
+
+We choose $(\alpha_{T}$, $r_{T}) = (8 \text{ deg}, 4 \text{ cm} )$ and get a MC 85\% efficiency and 98\% purity for the kaon sample and a MC \textcolor{red}{BOH}\% efficiency and 98\% purity for the \textcolor{red}{BOH} sample.
 
@@
 \subsection{Interaction Point Optimization}\label{ch:TrackingOptimization}
+Scheme of this subsection
+\subsubsection{Brief Explanation of the reconstruction chain}
+\subsubsection{Explanation of clustering parameters}
+\subsubsection{Figure of merit and  spanning of cluster}
+\subsubsection{Important numbers out of this optimization}
+
+Plots I want in this section:
+\begin{enumerate}
+\item Delta L, reco - true
+\item Delta L, reco - true Elastic, Delta L, reco - true Inelastic, other
+\item Length Quality cut
+\item Efficiency as a function of true KE and Angle
+\end{enumerate}
 

### 2018-04-03 AppendixB-EField/B-LArIATEField.tex
s

@@
-\chapter{Measurement of LArIAT Electric Field}{ch:AppendixB}
+\chapter{Measurement of LArIAT Electric Field}\label{ch:AppendixB}
 The electric field of a LArTPC in the drift volume is a fundamental quantity for the proper functionality of this technology, as it affects almost every reconstructed quantity such as the position of hits or their collected charge.

### 2018-04-03 Chapter-1/1-Theory.tex
Changing organization structure
s

--- 2018-02-04 Chapter-0/0-Introduction.tex
@@
-\chapter{Introduction}
+\chapter{The theoretical framework}
 
@@
 \section{Neutrinos: in and beyond the Standard Model}
-\subsection{Neutrino in the SM}
+\subsection{Neutrinos in the SM}
 Neutrino were introduced in the SM as left-handed massless Weyl spinors.
@@
 
-The first two lines of the lagrangian summarize the kinetic terms for the fermionic fields and their coupling to the gauge bosons $A^{\mu\nu}_a$, $B^{\mu\nu}$ \footnote{In gauge theories the ordinary derivative $\partial_\mu$  is substitued with èthe covariant derivative $D_\mu$.
+The first two lines of the lagrangian summarize the kinetic terms for the fermionic fields and their coupling to the gauge bosons $A^{\mu\nu}_a$, $B^{\mu\nu}$ \footnote{In gauge theories the ordinary derivative $\partial_\mu$  is substitued with the covariant derivative $D_\mu$.
 Here $D_\mu = \partial_\mu + igA_\mu \cdot I + ig'B_\mu\frac{Y}{2}$, where I and Y are the SU(2)$_L$ and U(1)$_Y$ generators, respectively.}.
@@
 
-where $c$ e $s$ stand respectively for cosine and sine of the corresponding  mixing angles ($\theta_{12}$, $\theta_{23}$ and $\theta_{13}$), $\delta$ è is the Dirac CP violation phase, $\alpha_1$ and $\alpha_{2}$   è is the eventual Majorana CP violation phases.
+where $c$ e $s$ stand respectively for cosine and sine of the corresponding  mixing angles ($\theta_{12}$, $\theta_{23}$ and $\theta_{13}$), $\delta$  is the Dirac CP violation phase, $\alpha_1$ and $\alpha_{2}$    is the eventual Majorana CP violation phases.
 
@@
 
-Following the recommendation of the latest Particle Physics Project Prioritization Panel  \cite{ParticlePhysicsProjectPrioritizationPanel(P5):2014pwa}, the US  is dedicating substantial resources to the development of a short- and long- baseline neutrino program to address many of these fundamental questions.
+Following the recommendation of the latest Particle Physics Project Prioritization Panel  \cite{P5}, the US  is dedicating substantial resources to the development of a short- and long- baseline neutrino program to address many of these fundamental questions.
 This program pivots on the Liquid Argon Time Projection Chamber (LArTPC) detector technology which will be described in \ref{ch:1}.
@@
 Within the SM, a total of 25 parameters remain seemingly arbitrary and need to be fitted to data: 3 gauge couplings, 9 charged fermion masses, 3 mixing angles and one CP phase in the CKM matrix, the Higgs mass and quartic coupling, $\theta_{QCD}$, 3 neutrino masses, 3 neutrino mixing angles, 1 Dirac phase and, eventually,  2 Majorana phases.
+
+\section{Motivations for Hadronic Cross Sections in Argon}

### 2018-04-03 Chapter-3/3-TheLArIATExperiment.tex
s

@@
 % write DAQ part
-% write collimators description
+% re-write captions
 % put references
@@
 If the magnets polarity is positive the tertiary beam composition counts mostly pions and protons with a small fraction of electrons, muons, and kaons.
-It is the job of the LArIAT beamline detectors to select the particles polarity,  to perform particle identification (beamPID) and to measure the momentum of the tertiary beam particles before they get to the LArTPC.
+It is the job of the LArIAT beamline detectors to select the particles polarity,  to perform particle identification in the beamline and to measure the momentum of the tertiary beam particles before they get to the LArTPC.
 The LArIAT detectors are described in the following paragraphs.
@@
 { Magnetic field over current as a function of the current, for one NDB magnet (excitation curve).
-The data was collected using two Hall probes (blue and green).
+The data was collected using two hall probes (blue and green).
 We fit the readings with a cubic function (black) to average of measurements (red) given in the legend.}
@@
 \caption{
-{One of the four Multi Wire Proportional Chambers (WC) used in the LArIAT tertiary beamline.}
+{One of the four Multi Wire Proportional Chambers (WC) used in the LArIAT tertiary beamline and relative read-out electronics.}
 }
@@
 
-%\begin{figure}[!h]
-%\begin{centering}
-%\vspace{-0.3cm}
-%\includegraphics[height=2.3in]{Chapter-3/Images/tofdelay.png}
-%\caption{
-%{\scriptsize \sf Pictures of the TOF system as was deployed during Run-I and Run-II data taking. The left image is of the upstream TOF paddle and the right image is of the downstream TOF paddle }
-%}
-%\label{fig:TOFSystemRunIandII}
-%\end{centering}
-%\end{figure}
-
 The CAEN 1751 digitizer is used to digitize the TOF PMTs signals at a sampling rate of 1 GHz.
@@
 The pulse pile up is not a significant problem given the TOF timing resolution and the rate of the particle beam.
-Leveraging on the pulses width uniformity of any given PMT (sigma of 400 ps),  we flag events where two pulses overlap as closely in time as 4 ns with an 90\% efficiency according to simulation.
+Leveraging on the pulses width uniformity of any given PMT,  we flag events where two pulses overlap as closely in time as 4 ns with an 90\% efficiency according to simulation.
 
@@
 
-\textcolor{red}{calculated TOF with error}
+\begin{figure}[h!]
+\centering
+\includegraphics[width=0.5\textwidth]{Chapter-3/Images/DSTOF.jpg}
+\caption{Image of the down stream time of flight paddle, PMTs and relative support structure before mounting. } 
+\label{pic:cosmicpaddle}
+\end{figure}
+
+%\textcolor{red}{calculated TOF with error}
 
@@
 
-The punch-thorough is simply a \textcolor{red}{? x ? x ?}~cm scintillator piece, read out by \textcolor{red}{ two? Hamamatsu? 2? inch} PMTs. 
+The punch-thorough is simple sheet of scintillator material, read out by two PMTs. 
 The MuRS is a segmented block of steel with four slots instrumented with scintillation bars.
 The four steel layers in front of each instrumented slot are 2 cm, 2 cm, 14 cm and 16 cm wide in the beam direction.
-Each instrumented slot is equipped with four scintillation bars each, positioned vertically in the direction orthogonal to the beam.
+Each instrumented slot is equipped with four scintillation bars each, positioned horizontally in the direction orthogonal to the beam.
 Each scintillator bar measures  \textcolor{red}{?
 x ?
-x 2}~cm and it is read out by  \textcolor{red}{ two?
-Hamamatsu?
-2?
-inch} PMTs.
+x 2}~cm and it is read out by one PMT.
 
@@
 It is worth noticing that the sampling time of the CAEN V1740 is slow (of the order of 128 ns), so pulse shape information from the PMT is lost.
-Punch-thorough and MuRS hits are formed utilizing the OR between the PMTs digital discriminator signals under threshold at a given time, where we obtain the threshold for each PMT directly on data distributions.
+Punch-thorough and MuRS hits are formed utilizing the digital discriminator signals under threshold at a given time, where we obtain the threshold for each PMT directly on data distributions.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-Besides on beam data, LArIAT also triggers on cosmic rays events by using two sets of cosmic ray paddle detectors (a.k.a.
+LArIAT triggers both on beam events and on cosmic rays events.
+We perform this latter trigger by using two sets of cosmic ray paddle detectors (a.k.a.
 ``cosmic towers".) The cosmic towers frame the LArIAT cryostat, as one sits in the downstream left corner and the other sits in the upstream right corner of the cryostat.
-Two paddle sets of four scintillators pieces each, an upper and a lower set, make up each cosmic tower.
+Two paddle sets of four scintillators pieces each make up each cosmic tower, an upper set and a lower set per tower.
 Of the four paddles, a couple of two matched paddles stands upright while the a second matched pair lies across the top of the assembly in the top sets (or across the bottom of the assembly in the bottom sets).
-The horizontal couple is used as a veto for particles traveling from the TPC out.
+The horizontal couple is used as a veto for particles traveling from inside the TPC out.
 The four signals  from the vertical paddles along one of the body diagonals of the TPC are combined in a logical ``AND''.
@@
 Cosmic ray tracks crossing both anode and cathode populate the events triggered this way.
-This particularly useful sample of tracks (which we can safely assume to be associated with ~5 GeV muons MIPs) can be used for many tasks; for example, we use anode-cathode piercing tracks to cross check the TPC electric field on data (see \ref{ch:AppendixC}), to calibrate the charge response of the TPC wires for the full TPC volume and to measure the electron lifetime in the chamber \textcolor{red}{ ADD reference to different chapters}.
+This particularly useful sample of tracks (which we can safely assume to be associated with ~5 GeV muons) can be used for many tasks; for example, we use anode-cathode piercing tracks to cross check the TPC electric field on data (see Appendix \ref{ch:AppendixB}), to calibrate the charge response of the TPC wires for the full TPC volume and to measure the electron lifetime in the chamber (see section \ref{ch:EnCalibration})\textcolor{red}{ ADD pictures}.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-All the paddles are $3.02~cm$ thick and are  trapezoidal in shape.
-The paddles come in two sizes: the smaller version has bases $32.2~cm$ and $26.7~cm$, and $61.0~cm$ height, while the bigger version has bases $33.2~cm$ and $27.0~cm$, and $70.8~cm$ height.
+%All the paddles are 3.02~cm thick and are  trapezoidal in shape.
+The paddles come in two sizes: the smaller version has bases 32.2~cm and 26.7~cm, and 61.0~cm height, while the bigger version has bases 33.2~cm and 27.0~cm, and $70.8~cm$ height.
  A Zener-diode Hamamatsu H5783 PMT collects the light from a wavelength-shifting optical fiber which runs along one of the long sides of each paddle.
@@
 We retrieved the scintillation paddles from the decommissioning of the CDF detector at Fermilab and we used only the paddles with a counting efficiency greater than 95\% and low noise at working voltage.
-The measured trigger rate of the whole system is $0.032Hz$, corresponding to $\sim 2$ muons per minute.
+The measured trigger rate of the whole system is 0.032~Hz, corresponding to $\sim 2$ muons per minute.
 
@@
 \subsection{Cryogenics and Argon Purity}\label{ch:Cryo}
-LArIAT repurposed the ArgoNeuT cryostat \cite{argoneut} in order to use it in a beam of charge particles. We also added a new process piping and a new liquid argon filtration system in FTBF.
+LArIAT repurposed the ArgoNeuT cryostat \cite{ArgoNeuT-det} in order to use it in a beam of charge particles, and added a new process piping and a new liquid argon filtration system in FTBF.
+%The main modifications to the ArgoNeuT cryostat are the addition of a beam flange and a port for the light collection system,
 Inside the LArIAT experimental hall, the cryostat sits on the beam of charge particles with its horizontal main axis oriented parallel to the beam.
 
-Two volumes make up LArIAT cryostat, shown in Figure \ref{fig:LArIATCryoStat}: purified liquid argon fills the inner vessel, while the outer volume insulates it with a vacuum jacket equipped with layers of aluminized mylar superinsulation.
+Two volumes make up LArIAT cryostat, shown in Figure \ref{fig:LArIATCryoStat}:  the inner vessel and the outer vessel.
+Purified liquid argon fills the inner vessel, while the outer volume provides insulation through a vacuum jacket equipped with layers of aluminized mylar superinsulation.
 The inner vessel is a cylinder of 130~cm length and 6.2~cm diameter, containing about 550~L of LAr, corresponding to a mass of 0.76 ton.
@@
 \includegraphics[scale=0.07]{Chapter-3/Images/Cryostat2.jpg}
-\caption{\emph{(left)} The LArIAT cryostat open with the TPC placed in the inner volume.
-\emph{(right)} The LArIAT cryostat fully sealed during initial commissioning prior to installation at Fermilab Testbeam Facility.Access to the internal volume is possible by opening the upstream end caps of the inner and outer vessels.
-}
+\caption{ Left: the LArIAT TPC in the inner volume of the open cryostat. Right: cryostat fully sealed ready to be transported to FTBF. }
 \label{fig:LArIATCryoStat}
@@
 \includegraphics[scale=0.35]{Chapter-3/Images/CryoMods.png}
-\caption{Pictures of the modified components of the cryostat.
-$1)$ The addition of an outlet to the bottom of the cryostat to allow connections to the purification system; $2)$ The ``beam-window'' on the outer endcap and the concave inner surface of the inner endcap (referred to as the excluder) to reduce the amount of material through which beam particles must travel before entering the TPC; $3)$ The modified side port for the LArIAT light collection system.}
+\caption{Main modifications to the ArgoNeuT cryostat: $1)$ outlet for connection to the purification system at the bottom of the cryostat; $2)$ the ``beam-window'' on the outer endcap and ``excluder"  which reduce the amount of non-instrumented material before the TPC; $3)$ the side port to host  the  light collection system.}
 \label{fig:LArIATCryoMods}
@@
 A single pass of argon in the filter is sufficient to achieve the necessary purity, unless the filter is saturated.
-In case the filter saturates, the media needs to be regenerated by using heated gas; this happened twice during the Run II period\footnote{We deemed the filter regeneration necessary every time the electron lifetime dropped under 100 \textmu s}.
+In case the filter saturates, the media needs to be regenerated by using heated gas; this happened twice during the Run II period\footnote{We deemed the filter regeneration necessary every time the electron lifetime dropped under 100 \textmu s.}.
 The filtered argon reaches the inner vessel via a liquid feedthrough on the top of the cryostat.
-Argon is not recirculated in the system, but it rather boils off and vent to the atmosphere.
+Argon is not recirculated in the system, rather it boils off and vent to the atmosphere.
 During data taking, we replenish the argon in the cryostat several times per day to keep the TPC high voltage feedthrough and cold electronics always submerged.
-In fact, we need to monitor the level, temperatures, and pressures  of the argon both in the commercial dewar and inside the cryostat.
+In fact, we constantly monitor the level, temperature, and pressure of the argon both in the commercial dewar and inside the cryostat during data taking.
 \subsection{LArTPC: Charge Collection}\label{sec:TPCCharge}
@@
 \subsubsection{Cathode and field cage}
-A G10 plain sheet with copper metallization on one of the 40x90~cm inner surfaces forms the cathode. 
+A G10 plain sheet with copper metallization on one of the 40 x 90~cm inner surfaces forms the cathode. 
 A high-voltage feedthrough on the top of the LArIAT cryostat delivers the high voltage to the cathode; scope of the high voltage system (Figure~\ref{fig:HVScheme}) is to drift ionization electrons from the interaction of charged particles in the liquid argon to the wire planes.
@@
 
-The field cage is made of \textcolor{red}{13 parallel copper rings} framing the inner walls of the G10 TPC structure.
+The field cage is made of twenty-three parallel copper rings framing the inner walls of the G10 TPC structure.
 A network of voltage-dividing resistors connected to the field cage rings steps down the high voltage from the cathode to form a uniform electric field.
-The electric field over the entire TPC drift volume is  486 V/cm (see \ref{ch:AppendixA}).
+The electric field over the entire TPC drift volume is  486 V/cm (see \ref{ch:AppendixB}).
 The  maximum drift length, i.e.
@@
 The wire planes measure the charge deposited in the TPC active volume. The drifting charge induces a current on the wire of the inner planes and it is collected on the collection plane wires.
-LArIAT counts three wire planes separated by 4 mm spaces: in order of distance from the cathode, they are the shield, the induction and the collection plane.
-The distance between two consecutive wires in each given plane (aka wire pitch) is 4 mm.
+LArIAT counts three wire planes separated by 4 mm spaces: in order of increasing distance from the cathode, they are the shield, the induction and the collection plane.
+The ``wire pitch", i.e., the distance between two consecutive wires in a given plane, is 4 mm.
 The shield plane counts 225 parallel wires of equal length oriented vertically.
@@
 Both the induction and collection planes count 240 parallel wires of different length oriented at 60$^\circ$ from the vertical with opposite signs.
-Moving electrons moving past the induction plane will induce a bipolar pulse on its wires and will form a unipolar pulse when collected on the last plane wires.
+Electrons moving past the induction plane will induce a bipolar pulse on its wires and will form a unipolar pulse when collected on the wires of the collection plane.
 
@@
 The other two drift regions are those between the shield plane and the induction plane (S-I), and between the induction plane and the collection plane (I-C).
-The electric field in these regions is chosen to satisfy the charge transparency condition to allow for 100$\%$ transmission of the drifting electrons through the shield and then the induction planes.
+The electric field in these regions is chosen to satisfy the charge transparency condition and allow for 100$\%$ transmission of the drifting electrons through the shield and the induction planes.
 

### 2018-04-03 Chapter-3/3-TheLArIATExperiment.tex
s

@@
 Cosmic ray tracks crossing both anode and cathode populate the events triggered this way.
-This particularly useful sample of tracks (which we can safely assume to be associated with ~5 GeV muons) can be used for many tasks; for example, we use anode-cathode piercing tracks to cross check the TPC electric field on data (see Appendix \ref{ch:AppendixB}), to calibrate the charge response of the TPC wires for the full TPC volume and to measure the electron lifetime in the chamber (see section \ref{ch:EnCalibration})\textcolor{red}{ ADD pictures}.
+This particularly useful sample of tracks (which we can safely assume to be associated with ~5 GeV muons) can be used for many tasks; for example, we use anode-cathode piercing tracks to cross check the TPC electric field on data (see Appendix \ref{ch:AppendixB}), to calibrate the charge response of the TPC wires for the full TPC volume and to measure the electron lifetime in the chamber (see section \ref{ch:EnCalibration}).
 
@@
 Both the induction and collection planes count 240 parallel wires of different length oriented at 60$^\circ$ from the vertical with opposite signs.
-Electrons moving past the induction plane will induce a bipolar pulse on its wires and will form a unipolar pulse when collected on the wires of the collection plane.
+Electrons moving past the induction plane will induce a bipolar pulse on its wires; the drifting electrons will be then collected on the collection plane's wires, forming a unipolar pulse.
 
@@
 Figure \ref{pic:FEelectronics} shows a block diagram of the overall system.
-The direct readout of the ionization electrons in liquid argon forms typically small signals on the wires, which needs to be amplified in oder to be processed.
-LArIAT  performs the amplification stage directly in cold with  amplifiers developed by Brookhaven National Lab (BNL) and mounted on the TPC frame inside the liquid argon, achieving a remarkable Signal-to-Noise ratio.
+The direct readout of the ionization electrons in liquid argon forms typically small signals on the wires, which need amplification in oder to be processed.
+LArIAT  performs the amplification stage directly in cold with  amplifiers %developed by the Brookhaven National Lab (BNL) and
+mounted on the TPC frame inside the liquid argon, achieving a remarkable Signal-to-Noise ratio.
 %The BNL ASICs adopted in LAriAT are designated as LArASIC, version 4-star.%The ASIC signals for each wire are then driven out of the vessel  to DAQ boards that act as waveform recorders.
@@
 \subsection{LArTPC: Light Collection System}\label{sec:TPCLight}
-The mechanism of particle detection in argon other than drift electrons is the collection of scintillation photons.
-Over the course of LArIAT three years of data taking, the light collection system changed several times.
+The collection of scintillation photons is the second mechanism of particle detection in argon other than the ionization electrons.
+Over the course of LArIAT's three years of data taking, the light collection system changed several times.
 We describe here the light collection system for Run II.
-Two PMTs, a 3-inch diameter Hamamatsu R-11065 and 2-inch diameter ETL D757KFL~\cite{lightsys-pmttests}, as well as three SiPMs arrays (two Hamamatsu S11828-3344M 4x4 arrays and one single-channel SensL MicroFB-60035 ) are mounted on the PEEK support structure.
+Two PMTs, a 3-inch diameter Hamamatsu R-11065 and 2-inch diameter ETL D757KFL~\cite{lightsys-pmttests}, as well as three SiPMs arrays (two Hamamatsu S11828-3344M 4x4 arrays and one single-channel SensL MicroFB-60035) are mounted on the PEEK support structure.
 PEEK screws into an access flange as shown in Figure~\ref{lightsys_pmts}, on the anode side, leaving  approximately 5~cm of clearance from the collection plane.
@@
 \end{figure}
-%------------------------------------------
+
 Liquid argon scintillates in vacuum-ultraviolet (VUV) range at 128 nm; since cryogenic PMTs are not sensitive to VUV wavelengths, we need to shift the light in a region visible to the PMTs.
-In LArIAT, the wavelength shifting is achieved by installing on the four walls of the TPC highly-reflective VIKUITY dielectric substrate foils coated with a thin layer of tetraphenyl-butadiene (TPB) (see Figure ~\ref{lightsys_foils}).
-One or more visible photons  are emitted and reflected into the chamber during the interaction of a VUV photon interacts with the TPB.
-Thus, the light yield is increased and made more uniform across the TPC active volume, allowing the possibility of light-based calorimetry (under study).
-
-%------------------------------------------
-%\begin{figure}
-%\centering
-%\includegraphics[scale=1.3]{Chapter-3/Images/lightsys_foils.png}
-%\caption{\label{lightsys_foils}The TPB-coated reflector foils mounted to the TPC field cage walls as viewed through the front cryostat opening. {\textcolor{red}{Better pictures needed.}}}
-%\end{figure}
-%------------------------------------------
+In LArIAT, the wavelength shifting is achieved by installing on the four walls of the TPC highly-reflective VIKUITY dielectric substrate foils coated with a thin layer of tetraphenyl-butadiene (TPB).
+The scintillation light interaction with the TPC emits one or more visible photons, we are then reflected into the chamber.
+Thus, the light yield increases and results more uniform across the TPC active volume, allowing the possibility of light-based calorimetry, currently under study.
 
@@
 Information about the light directionality is lost for light reflected on foils, as the reflection is uniform in angle.
-For Run-II, the voltage dividers for the PMTs were configured for positive bias with a DC-coupled anode (AC-coupled anode with grounded photocathode) to minimize induced noise on the TPC wires and modify the PMT bases accordingly.
+%For Run-II, the voltage dividers for the PMTs were configured for positive bias with a DC-coupled anode (AC-coupled anode with grounded photocathode) to minimize induced noise on the TPC wires and modify the PMT bases accordingly.
 
@@
 
-The beam instruments,  the cosmic ray taggers, and the light collection system provide NIM-standard logic pulse inputs to the trigger decision.
+LArIAT receives three logic from the Fermilab accelerator complex related to the beam timing which we use a input triggers: a pulse just before before the beam, a pulse indicating beam-on, and a beam-off pulse.
+
+The beam instruments,  the cosmic ray taggers, and the light collection system provide the other NIM-standard logic pulse inputs to the trigger decision.
 We automatically log the trigger inputs configuration with the rest of the DAQ configuration at the beginning of each run.
@@
 In particular, the discriminated pulses from the TOF PMTs form a NIM logic pulse for the trigger logic.
-We ask for a coincidence within a 20~ns window for all the pulses from the PMTs looking at the same scintillator block and use the coincidence int the upstream and downstream paddle to inform the trigger decision.
+We ask for a coincidence within a 20~ns window for all the pulses from the PMTs looking at the same scintillator block and use the coincidence between the upstream and downstream paddle to inform the trigger decision.
 In order to form a coincidence between the upstream and downstream paddles, we delay the upstream paddle coincidence by 20~ns and widen it by 100~ns.
 The delay and widening are necessary to account for both  lightspeed particles and slower particles (high-mass) to travel the 6.5~m between the upstream and the downstream paddles.
-Four multi-hit TDCs read out each wire chamber: two TDC per plane (horizontal and vertical), sixty-four wires per TDC.
+For the read out of the wire chambers, we use a total of sixteen multi-hit TDCs, four per chamber: two TDC per plane (horizontal and vertical), sixty-four wires per TDC.
 In each TDC, we keep the logical ``OR" for any signal over threshold from the sixty-four wires.
@@
 
-%Another primary input to the trigger card is from the cosmic towers (see Section~\ref{sec:CosmicRayPaddle}).
-To capture cosmic ray events in which a minimally ionizing cosmic ray muon crossed the TPC along the body diagonal, NIM modules form the logical coincidences from the two cosmic towers, one upper and one lower paddle assembly, in each combination.
-The OR of these is provided as an input to the V1495.
+The cosmic towers (see Section~\ref{sec:CosmicRayPaddle}) provide another primary input to the trigger, in order to capture long tracks from cosmic muons crossing the TPC.
+We use NIM modules to require coincidences between one upper and one lower paddle set of any opposite cosmic towers.
+The OR all the opposite towers' coincidences is fed as an input to the trigger card.
 
-%Three important logic pulses are derived from the timing of the beam.
-These include a pulse in a brief window before the beam, a pulse indicating that the beam is on, and a pulse which defines the beam-free period which may be used for collecting cosmic-ray events.
-An adjustable pulser is a fourth trigger input which does not depend on any particular activity in the experiment hall,  useful for collecting background events with zero bias.
-
-%%%%%%%%%%%%%%%%%%%%%%
-
-%The PMTs observing liquid argon scintillation light (see Section \ref{sec:PhotonSystem}) produce pulses which form the foundation of several interesting trigger inputs.
-Thresholding a copy of each PMT pulse (after amplification), and requiring a coincidence of pulses within $\sim$20~ns, creates simple trigger inputs indicating ionizing radiation was produced in the TPC.
-This scintillation logic pulse is used to initiate a gate which spans the length of the TPC drift time, creating a logic signal which is remains ``on'' while significant drift charge may still be present in the TPC.
-In addition, requiring a delayed coincidence of two subsequent scintillation logic pulses, separated by a variable length of time ranging from 300~ns to 7~$\mu$s, is used to create a trigger input to select events where a cosmic muon stops and decays to a Michel electron in the TPC.
-A few different versions of this light-based trigger were implemented throughout the course of LArIAT's run time to allow reconstruction and calorimetric studies of Michel electrons.
-Figure~\ref{michel_logic} shows a schematic diagram of the logic comprising the Michel electron trigger.
-
-%\begin{figure}
-%\includegraphics[width=\textwidth]{figures/trigger_michellogic.png}
-%\caption{\label{michel_logic}A schematic diagram of the trigger logic used to select Michel electron events during the cosmic readout window of the LArIAT supercycle.
-The two PMT signals refer to the Hamamatsu (``HMM'') and ETL PMTs described in Section~\ref{sec:PhotonSystem}.
-For some data-taking periods in Run-II, un-amplified pulses were discriminated at 180 mV to act as a veto on events that may saturate the dynamic range of the V1751 digitizer.
-The discriminator thresholds used on the amplified (x10) PMT signal copies (\emph{ThA}, \emph{ThB}) as well as the Gate Delay period, were adjusted between run periods while experimenting with different configurations.}
-%\end{figure}
-
-%Further trigger inputs come from the beam line instrumentation behind the LArTPC cryostat, the PMTs of the Punch-Through scintillator paddles and those of the scintillator paddles instrumenting the Muon Range Stack.
-The PMT pulses of all four of the broad-faced Punch Through paddles are discriminated to form logic pulses.
-A single logic pulse is formed from these, indicating activity in at least two overlapping paddles at the rear of cryostat, before the steel block of the range stack.
-PMT pulses from the Muon Range Stack are amplified and threshold discriminated.
-These MuRS paddle pulses are then combined as in the Punch Through, creating single-bit indicators for each of the four instrumented layers that at least one pair of overlapping scintillator paddles sent signals within a 20~ns coincidence window.
-
-%\subsubsection{Trigger Decision and Issuance}
-
-%The V1495 may be configured to have up to sixteen trigger patterns and sixteen veto patterns, based on the trigger input signals.
-A trigger pattern is defined as the AND of one or more defined inputs, and may include the NOT of the AND of further inputs.
-Veto patterns are independently defined in the same way, but they have a very different effect.
-When any of the trigger patterns fire, a ``fast trigger'' signal is issued and an adjustable countdown is initiated.
-If the countdown completes without a veto pattern firing, the ``slow trigger'' signal is also issued and on a distinct hardware channel.
-Otherwise, if a veto pattern fires during the countdown, the slow trigger signal is vetoed.
-
-%The fast trigger signal prompts readout of all the `short' data buffers, which include the V1751 modules, the V1495 itself, and the MWPC controller.
-The V1751 buffers typically contain digitized PMT signals from the time of flight and cryogenic light collection detectors.
-Readout of the TPC wire signals, which are much longer and more numerous, is only prompted at the issuance of the slow trigger.
+We use the signal from the cryogenic PMTs (see Section \ref{sec:PhotonSystem}) to form several interesting triggers.
+The coincidence of signal from all the PMT pulses within $\sim$20~ns is an indication of ionizing radiation in the TPC and forms a trigger input.
+The coincidence of two subsequent scintillation logic pulses delayed by a maximum of  7~$\mu$s forms the Michel electron trigger.
 

### 2018-04-03 Chapter-3/3-TheLArIATExperiment.tex
s

@@
 LArIAT uses the Accelerator Control NETwork system (ACNET) to monitor the beam conditions of the MCenter beamline.
-For example, the X and Y position of the beam at the first two wire chambers (WC1 and WC2) are shown in \ref{fig:ACNET}.
+For example, the horizontal and vertical position of the beam at the first two wire chambers (WC1 and WC2) are shown in \ref{fig:ACNET} as seen by the shifter during data taking.
 
@@
 We monitor the data taking and the run time evolution with the Run Status Webpage (\href{http://lariat-wbm.fnal.gov/lariat/run.html}{http://lariat-wbm.fnal.gov/lariat/run.html}), a  webpage updated in real-time.
-The page displays, among other information, the total number of triggers in the event per CAEN digitizer board, the total number of detectors triggered during a beam spill,  the trigger patterns (the number of times a particular trigger pattern was satisfied  during a beam spill) and current time relative to the Fermilab accelerator complex supercycle.
+The page displays, among other information, the total number of triggers in the event, the total number of detectors triggered during a beam spill,  the trigger patterns, the number of times a particular trigger pattern was satisfied  during a beam spill, and the current time relative to the Fermilab accelerator complex supercycle.
 A screen shot of the page is show in figure \ref{fig:runcond}.
@@
 \subsubsection{Data Quality Monitoring}
-We employ two systems to ensure the quality of out data during data taking: the Near-real-time Data Quality Monitoring and the Event Viewer.
+We employ two systems to ensure the quality of out data during data taking: the Near-Real-Time Data Quality Monitoring and the Event Viewer.
 
-\href{http://lariat-daq01.fnal.gov:5000/}{The Near-real-time Data Quality Monitoring} (DQM).
-This webpage receives updates from all the VME boards in the trigger system and displays the results of a quick analysis of the DAQ stream of raw data on a spill-by-spill basis.
+\href{http://lariat-daq01.fnal.gov:5000/}{The Near-Real-Time Data Quality Monitoring} (DQM) is a  webpage which receives updates from all the VME boards in the trigger system and displays the results of a quick analysis of the DAQ stream of raw data on a spill-by-spill basis.
 The DQM allows the shifter to monitor almost in real time (typically with a 2-minute delay)  a series of low level-quantities and compare them to past collections of beam spills.
 Some of the variables monitored in the DQM are  the pedestal mean and RMS on CAEN digitizer boards
-of the TPC wires and PMTs of the beamline detectors, the hit occupancy and timing plots on the multi-wire chambers, and number of data fragments recorded that are used to build a TPC event.
+of the TPC wires and PMTs of the beamline detectors, the hit occupancy and timing plots on the wire chambers, and number of data fragments recorded that are used to build a TPC event.
 Abnormal values for  low-level quantity in the data  activate a series of alarms in the DQM; this quick feedback on the DAQ and beam conditions is fundamental to assure a fast debugging of the detector and a very efficient data taking during beam uptime.
@@
 \centering
-%\includegraphics[scale=0.6]{Chapter-3/Images/RunConditions.png}
-\caption{High occupancy event display.}
+\includegraphics[scale=0.2]{Chapter-3/Images/highOccupancy.png}
+\caption{High occupancy event display: induction plane (top) and collection plane (bottom).}
 \label{fig:highOcc}

### 2018-04-03 Chapter-1/1-Theory.tex
s

@@
+% Have someone read this chapter
+% Add citations
+  % Higgs discovery papers CMS+Atlas
+  % Wu experiment (parity violation)
+  %Super Kamiokande, MINOS
+  %No$\nu$a, MACRO 
+  %SNO,Gallex,
+  %SAGE,  KamLAND 
+  % DAYA  Bay,   RENO
+% Red stuff!
+
 \chapter{The theoretical framework}
@@
 \section{The Standard Model}
-The Standard Model (SM) of particle physics the most accurate theoretical description of the subatomic world and, more generically, one of the most precisely tested theories in the history of physics.
+The Standard Model (SM) of particle physics is the most accurate theoretical description of the subatomic world and, in general, one of the most precisely tested theories in the history of physics.
 The SM describes the strong, electromagnetic and weak interactions among  elementary particles in the framework of quantum field theory, accounting for the unification of electromagnetic and weak interactions for energies above the  vacuum expectation value of the Higgs field.
@@
 The introduction of massive fields in the Standard Model lagrangian would make the theory non-renormalizable, and - so far - mathematically impossible to handle.
-This problem is solved in the Standard Model by the introduction of a scalar iso-doublet $\Phi(x)$, the Higgs field, which gives mass to W$^\pm$ and Z$^0$ gauge bosons through the electroweak symmetry breaking and to the fermions through Yukawa coupling \cite{Higgs1964,Higgs19642}.
+This problem is solved in the Standard Model by the introduction of a scalar iso-doublet $\Phi(x)$, the Higgs field, which gives mass to W$^\pm$ and Z$^0$ gauge bosons through the electroweak symmetry breaking and to the fermions through Yukawa coupling \cite{Higgs1964,Higgs19642}, discovered in 2012 at the LHC experiments \textcolor{red}{\cite{CMS,ATLAS}}.
 
@@
 
-The first two lines of the lagrangian summarize the kinetic terms for the fermionic fields and their coupling to the gauge bosons $A^{\mu\nu}_a$, $B^{\mu\nu}$ \footnote{In gauge theories the ordinary derivative $\partial_\mu$  is substitued with the covariant derivative $D_\mu$.
+The first two lines of the lagrangian summarize the kinetic terms for the fermionic fields and their coupling to the gauge bosons $A^{\mu\nu}_a$, $B^{\mu\nu}$ \footnote{In gauge theories the ordinary derivative $\partial_\mu$  is substitued with ¬èthe covariant derivative $D_\mu$.
 Here $D_\mu = \partial_\mu + igA_\mu \cdot I + ig'B_\mu\frac{Y}{2}$, where I and Y are the SU(2)$_L$ and U(1)$_Y$ generators, respectively.}.
@@
 
-where $c$ e $s$ stand respectively for cosine and sine of the corresponding  mixing angles ($\theta_{12}$, $\theta_{23}$ and $\theta_{13}$), $\delta$  is the Dirac CP violation phase, $\alpha_1$ and $\alpha_{2}$    is the eventual Majorana CP violation phases.
+where $c$ e $s$ stand respectively for cosine and sine of the corresponding  mixing angles ($\theta_{12}$, $\theta_{23}$ and $\theta_{13}$), $\delta$ èis the Dirac CP violation phase, $\alpha_1$ and $\alpha_{2}$   is the eventual Majorana CP violation phases.
+Experimental results on neutrino oscillations are generally reported in terms of the mixing angles and of the squared mass splitting $\Delta m^2_{ab} = | m^2_{a} - m^2_{b}|$, where $a$ and $b$ represent the mass eigenstates.
+A summary of the current status of experimental results, albeit partial, is given in table \ref{tab:nuosc}.
 
-\textcolor{red}{Experimental results summary}
+\begin{table}[htpb]
+\centering
+\caption{Summary of experimental results on neutrino oscillation parameters. \textcolor{red}{ADD CITATIONS}}
+\label{tab:nuosc}
+\begin{tabular}{|r|c|c|c|}
+\hline
+        & Value   & Precision & Experiment \\
+\hline
+$\theta_{23}$            & 45$^\circ$                      & 9.0\%   & Super Kamiokande, MINOS,\\
+$\Delta m^2_{23}$    & 2.5 $10^{-3}$ eV$^2$   & 1.8\% &  No$\nu$a, MACRO \\
+\hline
+$\theta_{12}$            & 34$^\circ$                     & 5.8\% &  SNO,Gallex,\\
+$\Delta m^2_{12}$    & 7.4 $10^{-5}$ eV$^2$   & 2.8\% &  SAGE,  KamLAND \\
+\hline
+$\theta_{13}$            &   9$^\circ$                    & 4.7\%  & DAYA  Bay, \\
+$\Delta m^2_{13}$    & 2.5 $10^{-3}$ eV$^2$   & 1.8\% &  RENO\\
+\hline
+\end{tabular}
+\end{table}
 
@@
 \subsubsection{Open Questions in Neutrino Physics}
-On one hand, the last X decades of experiments in neutrino oscillations brought spectacular advancements in the understanding of the oscillations pattern, measuring the neutrino mixing angles and mass splitting with a precision of less than 10\%.
+On one hand, the last \textcolor{red}{X} decades of experiments in neutrino oscillations brought spectacular advancements in the understanding of the oscillations pattern, measuring the neutrino mixing angles and mass splitting with a precision of less than 10\%.
 On the other, it opened the field for a series of questions needing experimental answers.

### 2018-04-03 Chapter-3/3-TheLArIATExperiment.tex
s

@@
 % This chapter to do:
-% Re-write the red parts
-% write DAQ part
-% re-write captions
-% put references
-% re-read
+% change figure with cryostat  modifications
+% Add TOF and Momentum Plots?
 
@@
 \end{figure}
-\begin{figure}
-\centering
-\includegraphics[height=0.25\textheight]{Chapter-3/Images/lightsys_etlbase.jpeg}
-\hspace{0.5cm}
-\includegraphics[height=0.25\textheight]{Chapter-3/Images/lightsys_hmmbase.jpg}
-\caption{\label{voltagedividers}Photos of the voltage divider bases for the ETL PMT (left) and the Hamamatsu PMT (right) used in Run-II.
-The cable connections to the bases seen here were used for powering and testing prior to installation.
-The yellow through-hole signal coupling capacitors seen on both bases are 18~nF (X7R) and are rated to 2~kV.}
-\end{figure}
 

### 2018-04-03 thesis_main.tex
it is been a while
s
a bit of tracking
Changing organization structure

@@
 
-\documentclass[letterpaper,12pt,draft]{yalephd}
+\documentclass[letterpaper,12pt]{yalephd}
 % remove draft option for final printing.
@@
 \usepackage{bm}
+\usepackage{hyperref}
+
+\usepackage{mathtools}
+\DeclarePairedDelimiter\bra{\langle}{\rvert}
+\DeclarePairedDelimiter\ket{\lvert}{\rangle}
+\DeclarePairedDelimiterX\braket[2]{\langle}{\rangle}{#1 \delimsize\vert #2}
+\usepackage{textgreek}
+
+\usepackage{multirow}
 \usepackage{amsmath}
@@
 \usepackage{cite}
+\usepackage{color}
 \bibliographystyle{plain}
+\usepackage{slashed}
+
+\setcounter{chapter}{-1}
 
@@
 % Need to define title before the abstract.
-\title{Measurement of the positive kaon-argon total hadronic differential cross section in the LArIAT experiment}
+\title{Measurement of total hadronic differential cross sections in the LArIAT experiment}
 \author{Elena Gramellini}
@@
 \chapter{Acknowledgements} % this needs to be before \mainmatter.
-A lot of people are awesome. But not you, who are reading my thesis before it's done.
+A lot of people are awesome, especially you, since you probably agreed to read this when it was a draft. 
+%But you, who is reading my thesis before it's done without my permission, you're the worst.
+%The truth is, younger Elena was a theorists at heart and didn't know.
+She absolutely loved QFT.
+But, she hated every single person of her class who self-identified as a theorist.
+She swore never to become one of them and buried that love deep in her heart, never to be touched by those jerks who traded in Lagrangians and measurements of each other's dicks.
+Was that love not strong enough to overcome the filth of the everyday shade?
+Was she not armored enough to shield from other people?s low self-esteem?
+Who?s fault was that?
+Not the theorist?s?
+she always thought it was hers, but was it?
+I don?t know.
+But sure, younger Elena had to find a way to get by.
+And she did.
+Painfully, with a lot of bruises and an ocean across.
+But she did.
 
+% A big thank you to my academic sisters, a bunch of extraordinary women who have taught me so much on the value
+%from the youngest to the oldest (again academically speaking).
+Supraja, Brooke, the depth of your knowledge, your drive and love for our field is a constant inspiration;  Ariana, thanks for keeping me sane for the last four year, love and hugs to you dudette; and last but not least Xiao, I don't think Fermilab has given me a bigger gift than your friendship.
 % Starts proper arabic numbering of pages and chapters.
 \mainmatter
+\input{Chapters/0-Introduction}
+\input{Chapter-1/1-Theory}
+\input{Chapter-2/2-LArTPCs}
+\input{Chapter-3/3-TheLArIATExperiment}
+\input{Chapter-4/4-InteractionsInLAr}
+\input{Chapter-5/5-SamplesPreparation}
+\input{Chapter-7/7-BackgroundSub}
+\input{Chapters/8-CrossSectionAnalysis}
+\input{Chapter-9/9-Uncertainty}
 
-\input{0-Introduction}
-\input{1-LArTPCs}
-\input{2-TheLArIATExperiment}
-\input{3-KaonInteractionsInLAr}
-\input{4-DataCollection}
-\input{5-EnergyCalibration}
-\input{6-TrackingOptimization}
-\input{7-KaonCrossSectionAnalysis}
+% Only call appendix once, here.
+\appendix
+%\input{Chapters/A-KaonAna}
+\input{AppendixB-EField/B-LArIATEField}
+%\input{Chapters/C-MicroBooNECRT}
+
+% Old chapters configuration %
+%\input{Chapters/4-DataCollection}
+%\input{Chapters/4-MonteCarlo}
+%\input{Chapters/5-EnergyCalibration}
+%\input{Chapter-8/8-TrackingOptimization}
 
@@
 
-% Only call appendix once, here.
-\appendix
-\input{A-MicroBooNECRT}
-
 % Any chapters such as End Notes go after this.
@@
 \bibliography{bib.bib}
-\bibliographystyle{plain}
+%\bibliographystyle{plain}
 % for your own sake, use a bibtex file, so all of the numbering of references will be done

### 2018-04-05 Chapter-1/1-Theory.tex
s

@@
 
-The first two lines of the lagrangian summarize the kinetic terms for the fermionic fields and their coupling to the gauge bosons $A^{\mu\nu}_a$, $B^{\mu\nu}$ \footnote{In gauge theories the ordinary derivative $\partial_\mu$  is substitued with ¬èthe covariant derivative $D_\mu$.
+The first two lines of the lagrangian summarize the kinetic terms for the fermionic fields and their coupling to the gauge bosons $A^{\mu\nu}_a$, $B^{\mu\nu}$ \footnote{In gauge theories the ordinary derivative $\partial_\mu$  is substitued with Âthe covariant derivative $D_\mu$.
 Here $D_\mu = \partial_\mu + igA_\mu \cdot I + ig'B_\mu\frac{Y}{2}$, where I and Y are the SU(2)$_L$ and U(1)$_Y$ generators, respectively.}.
@@
 
-where $c$ e $s$ stand respectively for cosine and sine of the corresponding  mixing angles ($\theta_{12}$, $\theta_{23}$ and $\theta_{13}$), $\delta$ èis the Dirac CP violation phase, $\alpha_1$ and $\alpha_{2}$   is the eventual Majorana CP violation phases.
+where $c$ e $s$ stand respectively for cosine and sine of the corresponding  mixing angles ($\theta_{12}$, $\theta_{23}$ and $\theta_{13}$), $\delta$ is the Dirac CP violation phase, $\alpha_1$ and $\alpha_{2}$   is the eventual Majorana CP violation phases.
 Experimental results on neutrino oscillations are generally reported in terms of the mixing angles and of the squared mass splitting $\Delta m^2_{ab} = | m^2_{a} - m^2_{b}|$, where $a$ and $b$ represent the mass eigenstates.
@@
 On the other, it opened the field for a series of questions needing experimental answers.
-Critical challenges in the next decade will entail an even deeper understanding of the neutrino mixing pattern, investigation of the neutrino mass origin, neutrino number and nature, and assessing CP violation in the lepton sector.
+Critical challenges in the next decade will entail an even deeper understanding of the neutrino mixing pattern, investigation of the neutrino mass origin, the determination of the number of neutrinos and their nature, and the measurement of CP violation in the lepton sector.
 
@@
 
+The smallness of neutrino masses remains a theoretical puzzle.
 \textcolor{red}{Smallness of neutrino masses, link to GUTs}
@@
 
+\subsubsection{Nucleon decay}\label{sec:theoryPDK}
+Baryon number is accidentally conserved in the Standard Model.
+Even though no baryon number violation has been experimentally observed thus far, no underlying symmetry in line with the Noether paradigm \cite{Noether1971} explains its conservation.
+Almost all Grand Unified Theories predict at some level baryon number violation in the form of nucleon decay on long time-scales.
+Given the impossibility to reach grand unification energy scales with collider experiments ($\sqrt{s} > 10^{15}$ GeV),  an indirect proof of GUT is needed.
+The experimental observation of nucleon decay may be the only viable way to explore these theories and it is therefore a subject of great interest \cite{Adams:LBNE}.
+%Both experiments and theory indeed suggest the energy scale for convergence of the running coupling constants of the Standard Model to be over $10^{15}$ GeV.
+This energy scale seems impossible to access by any foreseeable accelerator experiment, leaving baryon number violation  to be the only testable process.
+
 \section{Motivations for Hadronic Cross Sections in Argon}
+
+\subsection{Pion-Argon Total Hadronic Cross Section}
+\textcolor{red}{Here we put the link to previous sections: why is this important in the current context?}
+\subsubsection{Previous measurements}
+\subsubsection{Signal Signatures}
+\subsection{Kaon-Argon Total Hadronic Cross Section}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%% ---------------------------------------- Hadronic Cross sections ---------------------------------------- %%%%%
+
+\subsubsection{K$^{+}$Ar Cross section in the Context of Nucleon Decay Searches}
+If nucleon decay was experimentally found, additional information about the GUT type may be derived from the dominant decay mode.
+Supersymmetric GUTs \cite{Dimopoulos:1981dw,Bajc20161} prefer the presence of kaons in the products of the decay, e.g.
+$p\rightarrow K^+\bar{\nu}$  (see fig \ref{fig:MandatoryFeynmannDiagrams}, left).
+Gauge mediation GUTs, in which new gauge bosons are introduced that allow for the transformation of quarks into leptons, and vice versa, prefer the mode $p\rightarrow e^+\pi^0$ (see fig \ref{fig:MandatoryFeynmannDiagrams}, right).
+%%%%%%%%%%%%%%%% Find theory papers!!!!!! %%%%%%%%%%
+
+\begin{figure}[hbpt]
+\centering
+\includegraphics[width=6.5in]{Chapter-1/Images/MandatoryFeynmannDiagrams.png}
+\caption{Feynman diagrams for proton decay ``golden modes": $p \rightarrow K^+ \bar{\nu}$ for supersymmetric GUTs on the left and  $p \rightarrow e^+ \pi^0$ for gauge-mediation GUTs  on the right.}
+\label{fig:MandatoryFeynmannDiagrams}
+\end{figure}
+
+LArIAT tiny active volume makes it impossible for the experiment to place competitive limits on nucleon decay searches.
+However,  LArIAT provides excellent data to characterize kaons in LAr for the ``LAr golden mode", $p \rightarrow K^+ \bar{\nu}$.
+The result of these studies will affect future proton decay searches in LArTPCs.
+Previous work has been done to assess the potential identification efficiency for different decay modes in a LArTPC \cite{Bueno2007}, but, as the time of this  writing, no study of kaon selection efficiency in LArTPCs has been performed on data.
+The K$^+$-Ar interaction cross section has never been measured before and can affect the possibility of detecting and measuring kaons when produced in a proton decay event. 
+Kaon interactions with argon can distort the kaon energy spectrum as well as change the topology of single kaon events.
+In a LArTPC, non-interacting kaons appear as straight tracks with a high ionization depositions at the end (Bragg peak).
+The topology of interacting kaons can be quite different.
+In case of elastic scattering, a distinct kink will be present in the track.
+In case of inelastic scattering the Bragg peak will not be present and additional tracks will populate the event.
+Performing the total hadronic K$^+$-Ar cross section measurement on data serves the double purpose of identifying the rate of ``unusual" topologies (kinks and additional tracks) and of developing tools for kaon tracking in LAr.
+
+\subsubsection{K$^{+}$Ar Cross section in the Context of Light Mesons Interaction with Nuclei}
+\label{sec:theoryStrangeMeson}
+The intrinsic value of this measurement is that kaon interactions complement the measurements of $\pi$ interactions as a probe of  hadron interaction inside the nucleus in the strange sector.  
+\textcolor{red}{CHIEDI REFERENZE A FLAVIO}
+
+\subsubsection{Signal Signatures}\label{sec:KSignalSignature}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+The interaction of a mildly relativistic charged kaon with an argon nucleus is determined largely by the strong force.
+The total hadronic K$^{+}$-Ar interaction cross section is defined as the one related to the single (hadronic) process driven only by the strong interaction.
+In this case, ``total" indicates all strong interactions regardless of the final state.
+This condition purposefully includes both elastic and inelastic (reaction) channels.
+Indeed, the total cross section section can be then decomposed into
+$$\sigma_{Tot} = \sigma_{Elastic}+ \sigma_{Reaction}.$$
+
+For this analysis, kaons are selected from the LArIAT beamline in the momentum range between \textcolor{red}{500} MeV/c and \textcolor{red}{1000} MeV/c (see Fig \ref{fig:TOFK}).
+
+\begin{figure}[hpbt]
+\centering
+%\includegraphics[width=5in]{Chapter-1/Images/KaonTOF}
+\caption{Time of flight versus momentum distributions as produced by the LArIAT TOF and Wire Chambers systems.
+The Kaon population lies between the proton and the muon/pion populations, allowing PID of Kaons in the beam line.
+}
+\label{fig:TOFK}
+\end{figure}
+
+In this energy range, the relevant K-Nucleon interactions are according to \cite{fesbach1992theoretical}:
+\begin{equation}
+K^{+} + N \rightarrow K^{+} + N\textit{ (elastic)}
+\end{equation}
+\begin{equation}K^{+} + n \rightarrow K^{0} + p\textit{ (elastic)}\end{equation}
+\begin{equation}K^{+} + N \rightarrow K + N + \pi \textit{ (inelastic)}\end{equation}
+\begin{equation}K^{+} + N \rightarrow K^{*} + N\textit{ (inelastic)}.\end{equation}
+
+\subsubsection{Previous Measurements: Lighter and Heavier Nuclei}
+In general, data on kaon cross sections are  extremely scarce.
+The measurement of kaon interaction cross section would bring the additional benefit of reducing the uncertainties associated  with hadron interaction models adopted in MC simulations for argon targets.
+
+The data on nuclear effects for specific hadronic final states are extremely scarce, resulting in big uncertainties in modeling final state interactions \cite{Drakoulakos:2004gn}.
+Figure \ref{fig:Friedmann} shows a 1997 measurement on several elements as performed by  Friedmann et al.
+\cite{Friedman:1997eq}.
+As a reference, this paper measures a $\sigma_{Tot}$ for Si of  366.5  $\pm$  4.8 mb and a $\sigma_{Tot}$ for Ca of 494.6  $\pm$ 7.7 mb at 488 MeV/c.
+The cross section for argon is expected to lie in between these two measurements.
+Additional data on the kaon cross section are provided by Bugg et al. \cite{PhysRev.168.1466}. Bugg performs a measurement of the total 
+K$^+$ and K$^-$ cross sections on protons and deuterons over the range of 0.6-2.65 GeV/c, as well as a measurement of the total K$^+$ and K$^-$  cross sections on carbon for a number of momenta; the results of this paper on carbon are reported in Figure \ref{fig:Bugg}.
+
+\begin{figure}
+\begin{minipage}[b]{.5\textwidth}  
+  \centering  
+\includegraphics[width=3in]{Chapter-1/Images/Friedmann.png}
+\end{minipage}%  
+\begin{minipage}[b]{0.5\textwidth}  
+  \centering  
+\includegraphics[width=3in]{Chapter-1/Images/Bugg.png}
+\end{minipage}
+\par
+\begin{minipage}[t]{.53\textwidth}
+\caption{Ratios between experimental and calculated cross sections as from \cite{Friedman:1997eq}. Top: Total cross sections. Bottom: reaction cross sections.}
+\label{fig:Friedmann}
+\end{minipage}%
+\begin{minipage}[t]{.5\textwidth}  
+\caption{Total K$^+$  and K$^-$ cross sections on carbon as from \cite{PhysRev.168.1466}.}
+\label{fig:Bugg}
+\end{minipage}  
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% PRETTY EVENT DISPLAY WITH TEXT, NOT SURE IF USEFUL
+%\begin{figure}[h!]
+%\centering
+%\includegraphics[width=6.5in]{Chapter-1/Images/KLariat.png}
+%\caption{LArIAT Data $K^+$ candidate.
+$K^+$ enters TPC, undergoes a hadronic scatter, and then decays into $\pi^+$ and $\pi^0$.
+The the $\pi^0$ decays into 2 photons while the $\pi^+$ stops quickly in the TPC.
+Collection plane view.}
+%\label{Fig:KLariat}
+%\end{figure}
+
+%Fig \ref{Fig:KLariat} shows a $K^+$ candidate event in the LArIAT TPC.
+Following the kaon candidate track from left to right, two important elements are visible by eye: a change in the K momentum due to hadronic scatter and a Bragg peak by the end of the track due to an augment of ionization as the kaon slows down in the TPC.
+The track "kink" is only visible thanks to the millimetric spacial resolution of the TPC, while the Bragg shows the calorimetric power of this technology.
+The kaon in this event decays hadronically into $\pi^+$ and $\pi^0$.
+The the $\pi^0$ decays into 2 photons while the $\pi^+$ stops quickly in the TPC.
+The ability to distinguish the topology of this decay from the most frequent one, i.e.
+$K^+\rightarrow\mu^+\nu$, remarks the versatility of the LArTPC technology.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+ \subsubsection{Kaon Interaction Cross Section for thin target in Geant4}
+Since the kaon cross section in argon has never been measured before, the Geant4 Monte Carlo tunes kaon transportation in argon by extrapolation from lighter and heavier nuclei.
+As shown in the previous section,  kaon data on carbon are available and  can be used as a metric to evaluate the Geant4 prediction performances.
+Figure \ref{fig:TrueCarbon} shows the total hadronic cross section for carbon implemented in Geant4 10.01.p3 overlaid with the Bugg and Friedman data.
+Unfortunately, the current version of Geant4 does not reproduce the data for carbon closely.
+On one hand, this evidence makes us even more wary when using the Monte Carlo in simulating the kaon-argon interactions.
+On the other, it further highlights the importance of kaon measurements.

### 2018-04-05 Chapter-2/2-LArTPCs.tex
s

@@
 \section{DUNE: Rare Decay Searches}
+The key elements for a rare decay experiment are: massive active volume, long exposure, high identification efficiency and low background. 
+%The limit to proton lifetime in case of absence of signal and backgrounds is set by calculating
+%$$\tau/B > M\times \epsilon\times T \times 10^{32},$$ 
+%where M is the detector mass in kton, $\epsilon$ the signal detection efficiency after cuts to suppress backgrounds (dependent on the considered decay mode), T is the exposure in years, B the assumed branching fraction for the considered mode and  $10^{32}$ is a factor accounting for the number of nucleons in a kton of material \cite{Bueno2007}.
+Figure \ref{fig:PDKExperimentalLImit} shows the current best experimental limits on nucleon decay lifetime over branching ratio (dots).
+Historically, the dominant technology used in these searches has been water Cherenkov detectors: all the best experimental limits on every decay mode are indeed set by Super-Kamiokande \cite{PhysRevD.90.072005,PhysRevLett.115.121803}.
+It is particularly important to notice that the kaon energy for the proton decay mode $p \rightarrow K^+ \bar{\nu}$ is under Cherenkov threshold.
+Super-Kamiokande set the limit on the lifetime for the $p \rightarrow K^+ \bar{\nu}$ mode by  relying exclusively on photons from nuclear de-excitation.
+For this reason, an attractive alternative approach to identifying nucleon decay is the use of a Liquid Argon Time Projection Chamber (LArTPC).
+
+LArTPCs can complement nucleon decay searches in modes where water Cherenkov detectors are less sensitive, especially $p\rightarrow K^+\bar{\nu}$.
+According to \cite{Acciarri:Dune}, DUNE will have an active volume large enough, have sufficient shielding from the surface, and will run for lengths of time sufficient to compete with Hyper-K, opening up the opportunity for the discovery of nucleon decay.
+
+\begin{figure}[hbpt]
+\centering
+\includegraphics[width=6.5in]{Chapter-2/Images/PDKExperimentalLImit.png}
+\caption{Proton decay lifetime limits from passed and future experiments.}
+\label{fig:PDKExperimentalLImit}
+\end{figure}
+
 %\subsection{Non-Accelerator Physics Program}

### 2018-04-08 Chapter-1/1-Theory.tex
s

@@
 \section{Motivations for Hadronic Cross Sections in Argon}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%% ---------------------------------------- Hadronic Cross sections ---------------------------------------- %%%%%
+%%%%% ----------------------------------------     Pion Cross Section     ---------------------------------------- %%%%%
 
 \subsection{Pion-Argon Total Hadronic Cross Section}
-\textcolor{red}{Here we put the link to previous sections: why is this important in the current context?}
-\subsubsection{Previous measurements}
+This section outlines the importance of the pion-argon total hadronic cross section.
+We start by discussing the measurement in the context of neutrino interaction searches and of light mesons interaction with nuclei studies.
+We then describe the signal signature and historical measurements of pion-nucleus cross section, as well as the implementation of this cross sections in the current version of the simulation package used by LArIAT.
+\subsubsection{$\pi^{-}$Ar Cross section in the Context of Neutrino Searches}
+\subsubsection{$\pi^{-}$Ar Cross Section in the Context of Light Mesons Interaction with Nuclei}
 \subsubsection{Signal Signatures}
-\subsection{Kaon-Argon Total Hadronic Cross Section}
+\subsubsection{Previous measurements: Lighter and Heavier Nuclei}
+\subsubsection{Pion Interaction Cross Section for thin target in Geant4}
 
@@
 %%%%% ---------------------------------------- Hadronic Cross sections ---------------------------------------- %%%%%
+%%%%% ----------------------------------------     Kaon Cross Section     ---------------------------------------- %%%%%
+
+\subsection{Kaon-Argon Total Hadronic Cross Section}
+This section outlines the importance of the kaon-argon total hadronic cross section.
+We start by discussing the measurement in the context of nucleon decay searches and of light mesons interaction with nuclei studies.
+We then describe the signal signature and historical measurements of kaon-nucleus cross section, as well as the implementation of this cross sections in the current version of the simulation package used by LArIAT.
 
 \subsubsection{K$^{+}$Ar Cross section in the Context of Nucleon Decay Searches}
-If nucleon decay was experimentally found, additional information about the GUT type may be derived from the dominant decay mode.
+In case of nucleon decay discovery, the dominant decay mode may uncover additional information about the GUT type.
 Supersymmetric GUTs \cite{Dimopoulos:1981dw,Bajc20161} prefer the presence of kaons in the products of the decay, e.g.
 $p\rightarrow K^+\bar{\nu}$  (see fig \ref{fig:MandatoryFeynmannDiagrams}, left).
-Gauge mediation GUTs, in which new gauge bosons are introduced that allow for the transformation of quarks into leptons, and vice versa, prefer the mode $p\rightarrow e^+\pi^0$ (see fig \ref{fig:MandatoryFeynmannDiagrams}, right).
-%%%%%%%%%%%%%%%% Find theory papers!!!!!! %%%%%%%%%%
+Gauge mediated GUTs, in which new gauge bosons are introduced that allow for the transformation of quarks into leptons, and vice versa, prefer the mode $p\rightarrow e^+\pi^0$ (see fig \ref{fig:MandatoryFeynmannDiagrams}, right).
 
@@
 \includegraphics[width=6.5in]{Chapter-1/Images/MandatoryFeynmannDiagrams.png}
-\caption{Feynman diagrams for proton decay ``golden modes": $p \rightarrow K^+ \bar{\nu}$ for supersymmetric GUTs on the left and  $p \rightarrow e^+ \pi^0$ for gauge-mediation GUTs  on the right.}
+\caption{Feynman diagrams for proton decay ``golden modes": $p \rightarrow K^+ \bar{\nu}$ for supersymmetric GUTs on the left and  $p \rightarrow e^+ \pi^0$ for gauge-mediated GUTs  on the right.}
 \label{fig:MandatoryFeynmannDiagrams}
@@
 LArIAT tiny active volume makes it impossible for the experiment to place competitive limits on nucleon decay searches.
-However,  LArIAT provides excellent data to characterize kaons in LAr for the ``LAr golden mode", $p \rightarrow K^+ \bar{\nu}$.
+However,  LArIAT provides excellent data to characterize kaons in liquid argon for the ``LAr golden mode", $p \rightarrow K^+ \bar{\nu}$.
 The result of these studies will affect future proton decay searches in LArTPCs.
@@
 \label{sec:theoryStrangeMeson}
-The intrinsic value of this measurement is that kaon interactions complement the measurements of $\pi$ interactions as a probe of  hadron interaction inside the nucleus in the strange sector.  
-\textcolor{red}{CHIEDI REFERENZE A FLAVIO}
+The intrinsic value of the total hadronic K$^{+}$-Ar cross section measurement is that kaon interactions complement the measurements of $\pi$ interactions as a probe of  hadron interaction inside the nucleus in the strange sector.
+\textcolor{red}{High theoretical interest in probing constituent quark model of nuclear structure with
+KAON-NUCLEON INTERACTIONS CHIEDI REFERENZE A FLAVIO}
+
+%Total cross sections for the interaction of mildly relativistic kaons with several nuclei were derived from transmission experiments performed at the alternating-gradient synchrotron in Brookhaven National Laboratory.
+The high precision of these cross sections ~about 1\% led to analyses of the data in terms of KN nucleus potentials, based on the expectation that the KN nucleus interaction is simply related to the KN interaction.
+In particular, in this energy range the KN interaction does not vary strongly with energy and together with the relative weakness of the interaction, one expects that optical potentials close to the ??tr??
+approximation ~see below will be capable of describing the data.
+However, all such analyses showed disagreement between calculation and experiment at the level of 5?15\%, which caused speculations about modifications in the nuclear medium of the KN interaction @5?8#.
 
@@
 \subsubsection{Previous Measurements: Lighter and Heavier Nuclei}
-In general, data on kaon cross sections are  extremely scarce.
-The measurement of kaon interaction cross section would bring the additional benefit of reducing the uncertainties associated  with hadron interaction models adopted in MC simulations for argon targets.
+In general, measurements on kaon cross sections are  extremely scarce.
+The measurement of the kaon interaction cross section would bring the additional benefit of reducing the uncertainties associated  with hadron interaction models adopted in MC simulations for argon targets, beneficial for both proton decay studies and kaon production from neutrino interaction studies, where the  uncertainties for final state interaction models are big \cite{Drakoulakos:2004gn}.
 
-The data on nuclear effects for specific hadronic final states are extremely scarce, resulting in big uncertainties in modeling final state interactions \cite{Drakoulakos:2004gn}.
 Figure \ref{fig:Friedmann} shows a 1997 measurement on several elements as performed by  Friedmann et al.
@@
 \begin{figure}
-\begin{minipage}[b]{.5\textwidth}  
+\captionsetup{justification=raggedright}  
+	\begin{minipage}[t]{.53\textwidth}  
 	  \centering  
 	   \includegraphics[width=3in]{Chapter-1/Images/Friedmann.png}
+	   	        \caption{Ratios between experimental and calculated cross sections as from \cite{Friedman:1997eq}. Top: Total cross sections. \\Bottom: reaction cross sections.}
+        \label{fig:Friedmann}
 	\end{minipage}%  
-\begin{minipage}[b]{0.5\textwidth}  
+	\begin{minipage}[t]{0.53\textwidth}  
 	  \centering  
 	\includegraphics[width=3in]{Chapter-1/Images/Bugg.png}
-\end{minipage}
-\par
-\begin{minipage}[t]{.53\textwidth}
-\caption{Ratios between experimental and calculated cross sections as from \cite{Friedman:1997eq}. Top: Total cross sections. Bottom: reaction cross sections.}
-\label{fig:Friedmann}
-\end{minipage}%
-\begin{minipage}[t]{.5\textwidth}  
         \caption{Total K$^+$  and K$^-$ cross sections on carbon as from \cite{PhysRev.168.1466}.}
@@
 	\end{minipage}
+	\par
 \end{figure}
@@
  \subsubsection{Kaon Interaction Cross Section for thin target in Geant4}
-Since the kaon cross section in argon has never been measured before, the Geant4 Monte Carlo tunes kaon transportation in argon by extrapolation from lighter and heavier nuclei.
-As shown in the previous section,  kaon data on carbon are available and  can be used as a metric to evaluate the Geant4 prediction performances.
+Since the kaon cross section in argon has never been measured before, simulation packages tune kaon transportation in argon by extrapolation from lighter and heavier nuclei.
+LArIAT uses the Geant4 suite for particle transportation.
+Since kaon data on carbon are available, we used it as a metric to evaluate the Geant4 prediction performances.
 Figure \ref{fig:TrueCarbon} shows the total hadronic cross section for carbon implemented in Geant4 10.01.p3 overlaid with the Bugg and Friedman data.
@@
 On one hand, this evidence makes us even more wary when using the Monte Carlo in simulating the kaon-argon interactions.
-On the other, it further highlights the importance of kaon measurements.
+On the other, it further highlights the importance of the kaon measurement.
+
+\begin{figure}
+\captionsetup{justification=raggedright}  
+  \centering  
+\includegraphics[width=3in]{Chapter-1/Images/CarbonG4.png}
+\caption{total hadronic cross section for carbon implemented in Geant4  10.01.p3  with overlaid with the Bugg and Frideman data.}
+\label{fig:TrueCarbon}
+\end{figure}

### 2018-04-08 Chapter-1/1-Theory.tex
s

@@
 
-For this analysis, kaons are selected from the LArIAT beamline in the momentum range between \textcolor{red}{500} MeV/c and \textcolor{red}{1000} MeV/c (see Fig \ref{fig:TOFK}).
+%For this analysis, kaons are selected from the LArIAT beamline in the momentum range between \textcolor{red}{500} MeV/c and \textcolor{red}{1000} MeV/c (see Fig \ref{fig:TOFK}).
 
-\begin{figure}[hpbt]
-\centering
+%\begin{figure}[hpbt]
+%\centering
 %\includegraphics[width=5in]{Chapter-1/Images/KaonTOF}
-\caption{Time of flight versus momentum distributions as produced by the LArIAT TOF and Wire Chambers systems.
+%\caption{Time of flight versus momentum distributions as produced by the LArIAT TOF and Wire Chambers systems.
 The Kaon population lies between the proton and the muon/pion populations, allowing PID of Kaons in the beam line.
 }
-\label{fig:TOFK}
-\end{figure}
+%\label{fig:TOFK}
+%\end{figure}
 
+For the LArIAT cross section analysis, the kaons considered span a momentum inside the TPC from 800 MeV/c and 100 MeV/c.
 In this energy range, the relevant K-Nucleon interactions are according to \cite{fesbach1992theoretical}:
-\begin{equation}
-K^{+} + N \rightarrow K^{+} + N\textit{ (elastic)}
-\end{equation}
-\begin{equation}K^{+} + n \rightarrow K^{0} + p\textit{ (elastic)}\end{equation}
-\begin{equation}K^{+} + N \rightarrow K + N + \pi \textit{ (inelastic)}\end{equation}
-\begin{equation}K^{+} + N \rightarrow K^{*} + N\textit{ (inelastic)}.\end{equation}
+
+\begin{align}
+K^{+} + N &\rightarrow K^{+} + N\textit{ (elastic)}\\
+K^{+} + n &\rightarrow K^{0} + p\textit{ (elastic)}\\
+K^{+} + N &\rightarrow K + N + \pi \textit{ (inelastic)}\\
+K^{+} + N &\rightarrow K^{*} + N\textit{ (inelastic)}.
+\end{align}
 

### 2018-04-08 Chapter-2/2-LArTPCs.tex
s

@@
 
+\begin{figure}[hbpt]
+\centering
+\includegraphics[width=3.5in]{Chapter-2/Images/pdkGenie.png}
+\caption{Momentum of the kaon outgoing a proton decay event as simulated by the Genie 2.8.10 event generator in argon.
+The red line represent the kaon momentum distribution before undergoing the simulated final state interaction inside the argon nucleus, while the blue line represents the momentum distribution after FSI.
+}
+\label{fig:PDKGENIE}
+\end{figure}
+
 %\subsection{Non-Accelerator Physics Program}
@@
 \textcolor{red}{add paragraph on all wonderful things lariat can do... some event displays would be nice!}
+
+\textcolor{red}{ADD genie proton decay kaon distribution and lariat beamline overlaied}
+The signature of a proton decay event in the ``LAr golden mode" is the presence of a single kaon of about 400 MeV in the detector.

### 2018-04-08 Chapter-1/1-Theory.tex
s

@@
 % Add citations
-  % Higgs discovery papers CMS+Atlas
   % Wu experiment (parity violation)
@@
 In particular, the particles are divided in two categories, fermions and bosons, according to their spin-statistics.
-Described by the Fermi-Dirac statistics, Fermions have half-integer spin and are sometimes called ``matter-particles".
+Described by the Fermi-Dirac statistics, fermions have half-integer spin and are sometimes called ``matter-particles".
 Bosons or ``force carriers" have integer spin, follow the Bose-Einstein statistics and mediate the interaction between fermions.
@@
 The introduction of massive fields in the Standard Model lagrangian would make the theory non-renormalizable, and - so far - mathematically impossible to handle.
-This problem is solved in the Standard Model by the introduction of a scalar iso-doublet $\Phi(x)$, the Higgs field, which gives mass to W$^\pm$ and Z$^0$ gauge bosons through the electroweak symmetry breaking and to the fermions through Yukawa coupling \cite{Higgs1964,Higgs19642}, discovered in 2012 at the LHC experiments \textcolor{red}{\cite{CMS,ATLAS}}.
+This problem is solved in the Standard Model by the introduction of a scalar iso-doublet $\Phi(x)$, the Higgs field, which gives mass to W$^\pm$ and Z$^0$ gauge bosons through the electroweak symmetry breaking and to the fermions through Yukawa coupling \cite{Higgs1964,Higgs19642}.
+The discovery of the Higgs boson in 2012 by the LHC experiments \cite{201230,Aad2012} marked the ultimate confirmation  of a long history of Standard Model successful predictions.
 
-\section{Neutrinos: in and beyond the Standard Model}
-\subsection{Neutrinos in the SM}
+\section{Neutrinos:  tiny cracks in the Standard Model}
+\subsection{Neutrinos in the Standard Model}
 Neutrino were introduced in the SM as left-handed massless Weyl spinors.
@@
 for the chiral fields $\psi_R$ and $\psi_L$, whose evolution in space and time is coupled through the mass $m$.
-If the fermion is massless, the chiral fields decouple and the fermion can be described by a single Weyl spinor with two independent components.
-Pauli initially rejected the description of a physical particle through a single Wyle spinor because of its implication of parity violation.
+If the fermion is massless, the chiral fields decouple and the fermion can be described by a single Weyl spinor with two independent components~\cite{Weyl:10.2307}.
+Pauli initially rejected the description of a physical particle through a single Weyl spinor because of its implication of parity violation.
 In fact, since the spatial inversion operator throws $\psi_R \leftrightarrow \psi_L$, parity is conserved only if the both the chiral components exist at the same time.
-\textcolor{red}{ADD CITATIONS}.
-For the neutrino introduction in the SM, experiments came in help of the theoretical description:  the constraint of parity conservation weakened after Wu's experiment \textcolor{red}{ADD CITATIONS AND DATES}, there was no experimental indication for massive neutrinos and neutrinos likely interacted only via the left-handed component.
+For the neutrino introduction in the SM, experiments came in help of the theoretical description.
+The constraint of parity conservation weakened after Wu's experiment in 1957 \cite{PhysRev.105.1413}.
+Additionally,  there was no experimental indication for massive neutrinos nor evidence of interaction via the neutrino right-handed component.% neutrinos likely interacted only via the left-handed component.
 
@@
 \subsection{Neutrino Oscillations}
-The determination of the flavor of a neutrino dynamically arises from the corresponding charged lepton associated in a CC interaction; for example, a $\nu_e$ is a neutrino which produces an $e^-$, a $\bar\nu_\mu$ is a neutrino which produces a $\mu^+$, $etc$.
+The determination of the flavor of a neutrino dynamically arises from the corresponding charged lepton associated in a change current interaction; for example, a $\nu_e$ is a neutrino which produces an $e^-$, a $\bar\nu_\mu$ is a neutrino which produces a $\mu^+$, $etc$.
 The neutrino flavor eigenstates $\ket{\nu_\alpha}$,  with $\alpha = e,\mu,\tau$, are orthogonal to each other and form a base for the the weak interaction matrix.
 
-Overwhelming experimental data show neutrinos change flavor during their propagation.
-This phenomenon, predicted by Bruno Pontecorvo in 1957, is  called ``neutrino oscillations" and it is possible only if the flavor eigenstate are not identical to the mass eigenstates.
+Overwhelming experimental data show neutrinos change flavor during their propagation~\cite{Patrignani:2016xqp}.
+This phenomenon, called ``neutrino oscillations",  was predicted first by Bruno Pontecorvo in 1957 ~\cite{Pontecorvo:1967fh}.
+Neutrino oscillations are possible only if the neutrino flavor eigenstate are not identical to the mass eigenstates, thus resulting in the first evidence of physics beyond the Standard Model.
 A minimal extension of the Standard Model introduces three mass eigenstates, $\ket{\nu_i}$ ($i = 1,2, 3$), whose mass $m_i$ is well defined.
@@
 
+\subsection{Why so sneaky? Neutrino Cross Section Challenge}
+
 \section{Beyond the Standard Model}
 The discovery of neutrino oscillation and its implication of non-zero neutrino mass mark  the beginning of a new, exciting era in neutrino physics: the era of physics Beyond the Standard Model (BSM) in the neutrino sector.
-We are currently searching for new, deeper theories that can accommodate neutrinos with non-zero mass, while remaining consistent with the rest of the Standard Model.
-%More excitingly, we need to probe these theories experimentally.
+We are currently searching for new, deeper theories that can accommodate neutrinos with tiny but non-zero mass, while remaining consistent with the rest of the Standard Model.
 
-\subsubsection{Open Questions in Neutrino Physics}
-On one hand, the last \textcolor{red}{X} decades of experiments in neutrino oscillations brought spectacular advancements in the understanding of the oscillations pattern, measuring the neutrino mixing angles and mass splitting with a precision of less than 10\%.
+\subsection{Open Questions in Neutrino Physics}
+On one hand, the last three decades of experiments in neutrino oscillations brought spectacular advancements in the understanding of the oscillations pattern, measuring the neutrino mixing angles and mass splitting with a precision of less than 10\%.
 On the other, it opened the field for a series of questions needing experimental answers.
-Critical challenges in the next decade will entail an even deeper understanding of the neutrino mixing pattern, investigation of the neutrino mass origin, the determination of the number of neutrinos and their nature, and the measurement of CP violation in the lepton sector.
 
+%%%%%%%%%%
+Hints to the existence of at least one additional neutrino, in the form of various anomalies, have been puzzling physicists almost from the beginning of neutrino oscillation searches. 
+Originally designed to look for evidence of neutrino oscillation, the Liquid Scintillator Neutrino Detector (LSND) \cite{Athanassopoulos1997} provided a first conflicting result with the Standard Model expectation of only three neutrino flavors.
+A second conflicting result has also been provided by the MiniBooNE experiment \cite{AGUILARAREVALO200928}.
+The  LSND and MiniBooNE $\nu_e$ and $\bar\nu_e$ appearance results, known as the ``LSND and MiniBooNE anomalies" \cite{Aguilar:2001ty, Athanassopoulos:1997pv, Aguilar-Arevalo:2013pmq}, may be interpreted under the assumption of a new neutrino.
+The additional neutrino needs to be ``sterile", i.e needs not to couple with the electroweak force carriers, in order to be consistent with the measurement of the width of the Z boson~\cite{2006257}.
+The new sterile neutrino would mainly be composed of a heavy neutrino $\nu_4$ with mass $m_4$ such that  $m_1, m_2, m_3 \ll m_4$ and  $\Delta m^2= \Delta m^2_{14} \sim [0.1 - 10]$ eV$^2$.
+The introduction of sterile neutrinos is an appealing line of thinking, since this renormalizable generalization of the Standard Model has the potential to impact long standing questions in high energy physics and cosmology: light sterile neutrinos are candidates for dark matter particles and there are ideas that the theory could be adjusted to explain the baryon asymmetry of the Universe via leptogenesis \cite{1063-7869-57-5-503}.
+
+%%%%%%%%%%
+\begin{enumerate}
+\item CP violation in the neutrino sector, and mass ordering and determination of the octant => hidden symmetries and . 
+\item Neutrino nature Majorana or Dirac?
+\item Neutrino masses?
+\end{enumerate}
+
+Critical challenges await  the next decade of experimental neutrino physics.
+% will entail an even deeper understanding of the neutrino mixing pattern, investigation of the neutrino mass origin, the determination of the number of neutrinos and their nature, and the measurement of CP violation in the lepton sector.
+%%% Plug to LArTPC %%%%%%%%%%%%
 Following the recommendation of the latest Particle Physics Project Prioritization Panel  \cite{P5}, the US  is dedicating substantial resources to the development of a short- and long- baseline neutrino program to address many of these fundamental questions.
@@
 
-The smallness of neutrino masses remains a theoretical puzzle.
-\textcolor{red}{Smallness of neutrino masses, link to GUTs}
-
-\subsubsection{Towards a more fundamental theory}
+\subsection{Towards a more fundamental theory: GUTs}
 Despite its highly predictive power, a number of conceptual issues arise in the SM which disfavor it to be a good candidate for a fundamental theory.

### 2018-04-09 Chapter-1/1-Theory.tex
s

@@
 \right]
+\label{eq:PMNS}
 \end{equation}
@@
 A second conflicting result has also been provided by the MiniBooNE experiment \cite{AGUILARAREVALO200928}.
-The  LSND and MiniBooNE $\nu_e$ and $\bar\nu_e$ appearance results, known as the ``LSND and MiniBooNE anomalies" \cite{Aguilar:2001ty, Athanassopoulos:1997pv, Aguilar-Arevalo:2013pmq}, may be interpreted under the assumption of a new neutrino.
+The  LSND and MiniBooNE $\nu_e$ and $\bar\nu_e$ appearance results, known as the ``LSND and MiniBooNE anomalies" \cite{Aguilar:2001ty, Athanassopoulos:1997pv, Aguilar-Arevalo:2013pmq}, may be interpreted under the assumption of a new right-handed neutrino.
 The additional neutrino needs to be ``sterile", i.e needs not to couple with the electroweak force carriers, in order to be consistent with the measurement of the width of the Z boson~\cite{2006257}.
@@
 %%%%%%%%%%
+The measurement of non-zero value for the oscillation parameter $\theta_{13}$ unleashes the exploration of low-energy CP violation in the neutrino sector.
+Using the parameterization of the PMNS matrix shown in Equation \ref{eq:PMNS},  the difference in the probability of $\nu_e$ to $\nu_\mu$ oscillation and the probability of $\bar\nu_e$ to $\bar\nu_\mu$ oscillation can be parametrized as follows,
+\begin{equation}
+P_{\nu_e\rightarrow\nu_\mu} - P_{\bar\nu_e\rightarrow\bar\nu_\mu} = J \cos\Big(\pm\delta - \frac{\Delta_{31} L}{2}\Big) \sin\Big(\frac{\Delta_{21} L}{2}\Big)\sin\Big(\frac{\Delta_{31} L}{2}\Big)
+\end{equation}
+\textcolor{red}{asymmetry formula}
+where. 
+  The magnitude of the CP- violating terms in the oscillation depends  directly on the size of the Jarlskog Invariant [56], but it aslo related to the octant and the the mass hiararchy.
+
+Exciting theoretical results tie low-energy CP violation in the neutrino sector directly to the generation of the baryon asymmetry of the Universe at the Grand Unified Theory scale \textcolor{red}{a couple of cit would be nice}.
+According to theoretical model described in \cite{PASCOLI20071}, for example, leptogenesis can be achieved if $|\sin\theta_{13} \sin \delta| > 0.11$, i.e.
+$\sin \delta > 0.7$.
+
 \begin{enumerate}
+
 \item CP violation in the neutrino sector, and mass ordering and determination of the octant => hidden symmetries and .
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \item Neutrino nature Majorana or Dirac?

### 2018-04-09 Chapter-1/1-Theory.tex
s

--- 2018-04-08 Chapter-1/1-Theory.tex
@@
 where $c$ e $s$ stand respectively for cosine and sine of the corresponding  mixing angles ($\theta_{12}$, $\theta_{23}$ and $\theta_{13}$), $\delta$ is the Dirac CP violation phase, $\alpha_1$ and $\alpha_{2}$   is the eventual Majorana CP violation phases.
-Experimental results on neutrino oscillations are generally reported in terms of the mixing angles and of the squared mass splitting $\Delta m^2_{ab} = | m^2_{a} - m^2_{b}|$, where $a$ and $b$ represent the mass eigenstates.
+Experimental results on neutrino oscillations are generally reported in terms of the mixing angles and of the squared mass splitting $\Delta m^2_{ab} = m^2_{a} - m^2_{b}$, where $a$ and $b$ represent the mass eigenstates.
 A summary of the current status of experimental results, albeit partial, is given in table \ref{tab:nuosc}.
@@
 %%%%%%%%%%
-Hints to the existence of at least one additional neutrino, in the form of various anomalies, have been puzzling physicists almost from the beginning of neutrino oscillation searches. 
+\textbf{Sterile neutrinos.} Hints to the existence of at least one additional neutrino, in the form of various anomalies, have been puzzling physicists almost from the beginning of neutrino oscillation searches.
 Originally designed to look for evidence of neutrino oscillation, the Liquid Scintillator Neutrino Detector (LSND) \cite{Athanassopoulos1997} provided a first conflicting result with the Standard Model expectation of only three neutrino flavors.
 A second conflicting result has also been provided by the MiniBooNE experiment \cite{AGUILARAREVALO200928}.
-The  LSND and MiniBooNE $\nu_e$ and $\bar\nu_e$ appearance results, known as the ``LSND and MiniBooNE anomalies" \cite{Aguilar:2001ty, Athanassopoulos:1997pv, Aguilar-Arevalo:2013pmq}, may be interpreted under the assumption of a new neutrino.
+The  LSND and MiniBooNE $\nu_e$ and $\bar\nu_e$ appearance results, known as the ``LSND and MiniBooNE anomalies" \cite{Aguilar:2001ty, Athanassopoulos:1997pv, Aguilar-Arevalo:2013pmq}, may be interpreted under the assumption of a new right-handed neutrino.
 The additional neutrino needs to be ``sterile", i.e needs not to couple with the electroweak force carriers, in order to be consistent with the measurement of the width of the Z boson~\cite{2006257}.
@@
 %%%%%%%%%%
+\textbf{CP Violation In Lepton Sector.} The measurement of non-zero value for the oscillation parameter $\theta_{13}$ allows the exploration of low-energy CP violation in the lepton sector at neutrino oscillation experiments, enabling the possibility to measure the Dirac CP-violating phase $\delta$.
+Exciting theoretical results tie $\delta$ directly to the generation of the baryon asymmetry of the Universe at the Grand Unified Theory scale \textcolor{red}{a couple of cit would be nice}.
+According to theoretical model described in \cite{PASCOLI20071}, for example, leptogenesis can be achieved if $|\sin\theta_{13} \sin \delta| > 0.11$, i.e.
+$\sin \delta > 0.7$.\\
+The asymmetry in the oscillation probability of neutrinos and antineutrinos is the observable sensitive to the Dirac CP-violating phase $\delta$ leveraged in neutrino oscillation experiments.
+Using the parameterization of the PMNS matrix shown in Equation \ref{eq:PMNS},  the difference in the probability of $\nu_e \rightarrow \nu_\mu$ oscillation and the probability of $\bar\nu_e \rightarrow \bar\nu_\mu$ oscillation can be parametrized as follows \cite{Cervera2000},
+\begin{equation}
+P_{\nu_e\rightarrow\nu_\mu} - P_{\bar\nu_e\rightarrow\bar\nu_\mu} = J \cos\Big(\pm\delta - \frac{\Delta_{31}L}{2}\Big) \sin\Big(\frac{\Delta_{21}L}{2}\Big)\sin\Big(\frac{\Delta_{31}L }{2}\Big)
+\end{equation}
+where 
+\begin{equation}
+J = \cos\theta_{13}\sin2\theta_{13}\sin2\theta_{12}\sin2\theta_{23}
+\end{equation}
+is the Jarlskog invariant \cite{Jarlskog1985}, $L$ the neutrino baseline and $\Delta_{ab}$ a factor proportional to the sign and magnitude of the mass splitting. 
+From these equations, it is clear how the relative  large value of $\theta_{13}$ is a happy accident necessary not to completely suppress the sensitivity to CP violation.
+The equations also show how the sensitivity to $\delta$ is tied to the measurement of the least precisely measured mixing angle,  $\theta_{23}$ (via the $\sin2\theta_{23}$ term) and to an other unknown quantity, the neutrino ``mass hierarchy" (via the $\Delta_{ab}$ terms).
+The precise determination of $\theta_{23}$ is often referred as to ``the octant problem".
+Current experimental results \textcolor{red}{cite NOVA T2K} are consistent with  $\theta_{23}=45^\circ$, which would imply maximal mixing between $\nu_\mu$ - $\nu_\tau$, hinting to an intriguing new symmetry.
+Therefore, a precise measurement of $\theta_{23}$  is of great interest for theoretical models of quark-lepton universality [59,84,85,86,87,88] , whose  quark and lepton mixing matrices are proportional to the deviation of $\theta_{23}$ from $45^\circ$.
+The  ``mass hierarchy" problem refers to the unknown ordering of the value of absolute mass of the neutrino mass eigenstates.
+Current oscillation experiments are sensitive only to the magnitude of the mass splitting, but not to its sign.
+Therefore, in a framework where the lightest neutrino mass (arbitrarily) corresponds to the first eigenstate $m_1$, we do not know whether $m_2 - m_1 < m_3 - m_1$ (Normal Hierarchy) or $m_2 - m_1 > m_3 - m_1$ (Inverted Hierarchy).
+
 \begin{enumerate}
+
 \item CP violation in the neutrino sector, and mass ordering and determination of the octant => hidden symmetries and .
+
 \item Neutrino nature Majorana or Dirac?

### 2018-04-09 Chapter-1/1-Theory.tex
s

@@
 i\gamma^ \mu \partial_\mu  \psi_L = m \psi_R
+\label{eq:15}
 \end{equation}
@@
 i\gamma^ \mu \partial_\mu  \psi_R = m \psi_L
+\label{eq:16}
 \end{equation}
@@
 \section{Beyond the Standard Model}
-The discovery of neutrino oscillation and its implication of non-zero neutrino mass mark  the beginning of a new, exciting era in neutrino physics: the era of physics Beyond the Standard Model (BSM) in the neutrino sector.
-We are currently searching for new, deeper theories that can accommodate neutrinos with tiny but non-zero mass, while remaining consistent with the rest of the Standard Model.
+The discovery of neutrino oscillation and its implication of non-zero neutrino mass mark  the beginning of a new, exciting era in neutrino physics: the era of physics Beyond the Standard Model (BSM) at the intensity frontier.
+We are currently searching for new, deeper theories that can accommodate neutrinos with tiny but non-zero masses, while remaining consistent with the rest of the Standard Model.
 
@@
 %%%%%%%%%%
-\textbf{CP Violation In Lepton Sector.} The measurement of non-zero value for the oscillation parameter $\theta_{13}$ allows the exploration of low-energy CP violation in the lepton sector at neutrino oscillation experiments, enabling the possibility to measure the Dirac CP-violating phase $\delta$.
+\textbf{CP Violation In Lepton Sector.} The measurement of non-zero value for the oscillation parameter $\theta_{13}$ allows the exploration of low-energy CP violation in the lepton sector at neutrino long baseline oscillation experiments, enabling the possibility to measure the Dirac CP-violating phase $\delta$.
 Exciting theoretical results tie $\delta$ directly to the generation of the baryon asymmetry of the Universe at the Grand Unified Theory scale \textcolor{red}{a couple of cit would be nice}.
@@
 Therefore, a precise measurement of $\theta_{23}$  is of great interest for theoretical models of quark-lepton universality [59,84,85,86,87,88] , whose  quark and lepton mixing matrices are proportional to the deviation of $\theta_{23}$ from $45^\circ$.
-The  ``mass hierarchy" problem refers to the unknown ordering of the value of absolute mass of the neutrino mass eigenstates.
+
+\textbf{Neutrino mass hierarchy.} The  ``mass hierarchy" problem refers to the unknown ordering of the value of absolute mass of the neutrino mass eigenstates.
 Current oscillation experiments are sensitive only to the magnitude of the mass splitting, but not to its sign.
-Therefore, in a framework where the lightest neutrino mass (arbitrarily) corresponds to the first eigenstate $m_1$, we do not know whether $m_2 - m_1 < m_3 - m_1$ (Normal Hierarchy) or $m_2 - m_1 > m_3 - m_1$ (Inverted Hierarchy).
+\textcolor{red}{cite hints} Therefore, in a framework where the lightest neutrino mass (arbitrarily) corresponds to the first eigenstate $m_1$, it is unknown whether $m_2 - m_1 < m_3 - m_1$ (Normal Hierarchy) or $m_2 - m_1 > m_3 - m_1$ (Inverted Hierarchy).
+The mass hierarchy affects not only the sensitivity to CP violation searches in long baseline oscillation experiments, but also the sensitivity to determine whether neutrinos are Majorana particles in neutrinoless double beta decay experiments.
+
+\textbf{Majorana or Dirac?}
+Evidence of neutrino oscillations demand the introduction in the theory of a mechanism which can give mass to the neutrinos.
+This mechanism should possibly also explain why neutrino masses are at least six orders of magnitude lower than all the other SM fermions' masses.
+In a description of neutrinos as Dirac 4-component spinors, the neutrino field acquires mass  via the Higgs mechanism as any other fermion of the SM. In this case, the neutrino mass is given by 
+\begin{equation} m_a = \frac{y^\nu_a v}{\sqrt2} \end{equation} 
+where $v$ is the Higgs VEV and   $y^\nu_a$ is the Yukawa coupling between the Higgs and the neutrino.
+The smallness of neutrino masses can only be pinned on a tiny Yukawa coupling which is not justified by the theory.\\
+In 1937, Majorana demonstrated that the introduction of a two components spinor is sufficient to describe a neutral massive fermion, under the assumption that the chiral components $\psi_R$ and $\psi_L$ are correlated as follows
+$$\psi_R = \mathcal{C}\bar\psi_L$$ where $\mathcal{C}$ is the charge conjugation matrix.
+Neutrinos are the only neutral elementary particles in the SM, thus the only possible Majorana particle candidate.
+\textcolor{red}{ a few words on see-saw}
+
+From a kinematic point of view, Dirac and Majorana neutrinos satisfy the same energy-momentum dispersion relationship.
+Thus, it is impossible to discern the neutrino nature through kinematic effects such as neutrino oscillations.
+
+Il modo più promettente per capire se il neutrino è una particella di Majorana è cercare sperimentalmente il doppio decadimento ? senza neutrini.
+Il doppio decadimento. E? chiaro inoltre che tale processo viola il numero leptonico di due unità
+
+The puzzle of the neutrino mass hints to the fascinating idea of considering the SM as an effective theory at low energy, obtained as the limit of a more complete theory.
 
 \begin{enumerate}
-
-\item CP violation in the neutrino sector, and mass ordering and determination of the octant => hidden symmetries and .
-
 \item Neutrino nature Majorana or Dirac?
@@
 \item[-] Assessment of the existence of right-handed sterile neutrinos. 
-\item[-] Determination of the sign of $\Delta m^2_{13}$ (or $\Delta m^2_{23}$), i.e., the ``neutrino mass ordering".
+\item[-] Determination of the sign of $\Delta m^2_{13}$ (or $\Delta m^2_{23}$), i.e., the neutrino mass hierarchy.
 \item[-] Determination of the octant, i.e.  whether $\theta_{23}$ is maximal.

### 2018-04-09 Chapter-1/1-Theory.tex
s

@@
 The  LSND and MiniBooNE $\nu_e$ and $\bar\nu_e$ appearance results, known as the ``LSND and MiniBooNE anomalies" \cite{Aguilar:2001ty, Athanassopoulos:1997pv, Aguilar-Arevalo:2013pmq}, may be interpreted under the assumption of a new right-handed neutrino.
-The additional neutrino needs to be ``sterile", i.e needs not to couple with the electroweak force carriers, in order to be consistent with the measurement of the width of the Z boson~\cite{2006257}.
+The additional neutrino needs to be ``sterile", i.e needs not to couple with the electroweak force carriers, in order to meet the constraint imposed by  the measurement of the width of the Z boson~\cite{2006257}.
 The new sterile neutrino would mainly be composed of a heavy neutrino $\nu_4$ with mass $m_4$ such that  $m_1, m_2, m_3 \ll m_4$ and  $\Delta m^2= \Delta m^2_{14} \sim [0.1 - 10]$ eV$^2$.
@@
 Exciting theoretical results tie $\delta$ directly to the generation of the baryon asymmetry of the Universe at the Grand Unified Theory scale \textcolor{red}{a couple of cit would be nice}.
-According to theoretical model described in \cite{PASCOLI20071}, for example, leptogenesis can be achieved if $|\sin\theta_{13} \sin \delta| > 0.11$, i.e.
+According to the theoretical model described in \cite{PASCOLI20071}, for example, leptogenesis can be achieved if $|\sin\theta_{13} \sin \delta| > 0.11$, i.e.
 $\sin \delta > 0.7$.\\
@@
 Current oscillation experiments are sensitive only to the magnitude of the mass splitting, but not to its sign.
-\textcolor{red}{cite hints} Therefore, in a framework where the lightest neutrino mass (arbitrarily) corresponds to the first eigenstate $m_1$, it is unknown whether $m_2 - m_1 < m_3 - m_1$ (Normal Hierarchy) or $m_2 - m_1 > m_3 - m_1$ (Inverted Hierarchy).
+\textcolor{red}{cite hints} In a framework where the lightest neutrino mass (arbitrarily) corresponds to the first eigenstate $m_1$, it is unknown whether $m_2 - m_1 < m_3 - m_1$ (Normal Hierarchy) or $m_2 - m_1 > m_3 - m_1$ (Inverted Hierarchy).
 The mass hierarchy affects not only the sensitivity to CP violation searches in long baseline oscillation experiments, but also the sensitivity to determine whether neutrinos are Majorana particles in neutrinoless double beta decay experiments.
@@
 Evidence of neutrino oscillations demand the introduction in the theory of a mechanism which can give mass to the neutrinos.
-This mechanism should possibly also explain why neutrino masses are at least six orders of magnitude lower than all the other SM fermions' masses.
+This mechanism should possibly also explain why neutrino masses are at least six orders of magnitude lower than the electron mass (the second lightest SM fermion).
 In a description of neutrinos as Dirac 4-component spinors, the neutrino field acquires mass  via the Higgs mechanism as any other fermion of the SM. In this case, the neutrino mass is given by 
-\begin{equation} m_a = \frac{y^\nu_a v}{\sqrt2} \end{equation} 
+%\begin{equation} 
+$m_a = \frac{y^\nu_a v}{\sqrt2}$, %\end{equation} 
 where $v$ is the Higgs VEV and   $y^\nu_a$ is the Yukawa coupling between the Higgs and the neutrino.
 The smallness of neutrino masses can only be pinned on a tiny Yukawa coupling which is not justified by the theory.\\
-In 1937, Majorana demonstrated that the introduction of a two components spinor is sufficient to describe a neutral massive fermion, under the assumption that the chiral components $\psi_R$ and $\psi_L$ are correlated as follows
-$$\psi_R = \mathcal{C}\bar\psi_L$$ where $\mathcal{C}$ is the charge conjugation matrix.
-Neutrinos are the only neutral elementary particles in the SM, thus the only possible Majorana particle candidate.
-\textcolor{red}{ a few words on see-saw}
+In 1937, Majorana demonstrated that the introduction of a two components spinor is sufficient to describe a massive fermion \cite{Majorana1937}.
+The Dirac equations of motion for the chiral fields (equations \ref{eq:15} and \ref{eq:16}) hold true under the assumption that the chiral components $\psi_R$ and $\psi_L$ are correlated through the charge conjugation matrix $\mathcal{C}$, $\psi_R = \mathcal{C}\bar\psi_L$, thus the theory is applicable only to neutral fermions.
+Neutrinos are the only neutral elementary particles in the SM -- the only possible Majorana particle candidate.
+This theory constructs a  neutrino Majorana mass term $\mathcal{L}_5 $ of the following form in the Higgs unitary gauge
+\begin{equation}
+\mathcal{L}_5 = \frac{1}{2}\frac{gv^2}{\mathcal{M}}\nu^T_L\mathcal{C^\dagger}\nu_L,
+ \end{equation}
+ where $g$ is the coupling coefficient, $v$ the Higgs VEV and $\mathcal{M}$ a constant with the dimension of the mass proportional to the scale of new physics.
+The largrangian term $\mathcal{L}_5$ would introduce nonrenormalizable term, since it has dimensions of energy to the fifth power.
+ %%%%%%%%%%%%%%%%%
+ This is not allowed in the SM lagrangian; however, if the SM is seen as an effective manifestation at low energy 
+ an effective theory at low energy 
+manifestation of symmetry breaking of a unified theory 
+Se si vede il MS come una teoria efficace in una scala di basse energie, manifestazione della rottura di simmetria di una teoria unificata ad alte energie, e non come una teoria definitiva, diventa plausibile l?esistenza di termini non rinormalizzabili nella lagrangiana effettiva; tali termini devono comunque rispettare le simmetrie del MS perché la teoria ad alta energia da cui sono generati deve includere le simmetrie di gauge del MS, in quanto essa si riduce a basse energie al MS stesso.
+%%%%%%%%%%%
 
+ In this context, the neutrino mass is given by $m = \frac{g v^2}{\mathcal{M}}$: the coupling coefficient can be of the order of any other coupling coefficient in the SM, since a big value of the new physics mass scale implies small neutrino masses.
+This vanilla formulation is the conceptual basis for many flavors of theories based on the \emph{see-saw mechanism}, which we will not discuss here in any detail.
+However, it is fascinating how the puzzle of the neutrino mass hints to the idea of considering the SM as an effective theory at low energy, obtained as the limit of a more complete theory.\\
 From a kinematic point of view, Dirac and Majorana neutrinos satisfy the same energy-momentum dispersion relationship.
 Thus, it is impossible to discern the neutrino nature through kinematic effects such as neutrino oscillations.
-
-Il modo più promettente per capire se il neutrino è una particella di Majorana è cercare sperimentalmente il doppio decadimento ? senza neutrini.
-Il doppio decadimento. E? chiaro inoltre che tale processo viola il numero leptonico di due unità
-
-The puzzle of the neutrino mass hints to the fascinating idea of considering the SM as an effective theory at low energy, obtained as the limit of a more complete theory.
-
-\begin{enumerate}
-\item Neutrino nature Majorana or Dirac?
-\item Neutrino masses?
-\end{enumerate}
+Neutrinoless double beta decay searches are the most promising way to discern the nature of the neutrino and are therefore subject of great theoretical and experimental interest.
+Observation of the lepton number violating process  $0\nu\beta\beta$  would imply neutrinos have a Majorana component.
+Depending on the mass hierarchy, the theory also predicts  $0\nu\beta\beta$  exclusion regions and confirmation of the sole Dirac component for neutrinos\textcolor{red}{find CIT}.\\
 

### 2018-04-09 Chapter-1/1-Theory.tex
s

@@
 
-\subsection{Why so sneaky? Neutrino Cross Section Challenge}
+\subsection{Make up of Neutrino Interactions}
+Any neutrino experiment involving the detection of neutrinos is concerned with neutrino interactions (and neutrino cross sections) on nuclei. 
+Given the invisible nature of neutrinos, characterizing the products of their interaction is the only method to retrieve their flavor and eventually energy.
+In real life experiments, this is complicated
 
@@
 In 1937, Majorana demonstrated that the introduction of a two components spinor is sufficient to describe a massive fermion \cite{Majorana1937}.
-The Dirac equations of motion for the chiral fields (equations \ref{eq:15} and \ref{eq:16}) hold true under the assumption that the chiral components $\psi_R$ and $\psi_L$ are correlated through the charge conjugation matrix $\mathcal{C}$, $\psi_R = \mathcal{C}\bar\psi_L$, thus the theory is applicable only to neutral fermions.
+The Dirac equations of motion for the chiral fields (equations \ref{eq:15} and \ref{eq:16}) hold true in the case of two components spinor under the assumption that the chiral components $\psi_R$ and $\psi_L$ are correlated through the charge conjugation matrix $\mathcal{C}$, $\psi_R = \mathcal{C}\bar\psi_L$, thus the theory is applicable only to neutral fermions.
 Neutrinos are the only neutral elementary particles in the SM -- the only possible Majorana particle candidate.
@@
 \mathcal{L}_5 = \frac{1}{2}\frac{gv^2}{\mathcal{M}}\nu^T_L\mathcal{C^\dagger}\nu_L,
+\label{eq:majoMass}
  \end{equation}
  where $g$ is the coupling coefficient, $v$ the Higgs VEV and $\mathcal{M}$ a constant with the dimension of the mass proportional to the scale of new physics.
-The largrangian term $\mathcal{L}_5$ would introduce nonrenormalizable term, since it has dimensions of energy to the fifth power.
- %%%%%%%%%%%%%%%%%
- This is not allowed in the SM lagrangian; however, if the SM is seen as an effective manifestation at low energy 
- an effective theory at low energy 
-manifestation of symmetry breaking of a unified theory 
-Se si vede il MS come una teoria efficace in una scala di basse energie, manifestazione della rottura di simmetria di una teoria unificata ad alte energie, e non come una teoria definitiva, diventa plausibile l?esistenza di termini non rinormalizzabili nella lagrangiana effettiva; tali termini devono comunque rispettare le simmetrie del MS perché la teoria ad alta energia da cui sono generati deve includere le simmetrie di gauge del MS, in quanto essa si riduce a basse energie al MS stesso.
-%%%%%%%%%%%
-
- In this context, the neutrino mass is given by $m = \frac{g v^2}{\mathcal{M}}$: the coupling coefficient can be of the order of any other coupling coefficient in the SM, since a big value of the new physics mass scale implies small neutrino masses.
-This vanilla formulation is the conceptual basis for many flavors of theories based on the \emph{see-saw mechanism}, which we will not discuss here in any detail.
-However, it is fascinating how the puzzle of the neutrino mass hints to the idea of considering the SM as an effective theory at low energy, obtained as the limit of a more complete theory.\\
+The  $\mathcal{L}_5$ term would introduce a non-renormalizable term in the lagrangian, since it has dimensions of energy to the fifth power.
+This is not allowed in the SM term; however, the existence of such terms is plausible if the we consider the SM an effective theory at low energy, manifestation of the symmetry breaking of a Grand Unified Theory (GUT) at higher energy, and not the definitive theory.
+The mass term in eq \ref{eq:majoMass} implies the neutrino mass to be  $m = \frac{g v^2}{\mathcal{M}}$.
+The coupling coefficient can be of the order of any other fermion's coupling coefficient, since the smallness of neutrino masses is achieved by the big value of the new physics mass scale alone.
+This vanilla formulation is the conceptual basis for many flavors of GUT-based \emph{seesaw mechanism} ~\cite{Yanagida1980}, which we will not discuss here in any detail.
+However, it is fascinating how the puzzle of the neutrino mass hints to the existence of a deeper and more complete theory.\\
 From a kinematic point of view, Dirac and Majorana neutrinos satisfy the same energy-momentum dispersion relationship.
 Thus, it is impossible to discern the neutrino nature through kinematic effects such as neutrino oscillations.
-Neutrinoless double beta decay searches are the most promising way to discern the nature of the neutrino and are therefore subject of great theoretical and experimental interest.
+Neutrinoless double beta decay searches are the most promising way to understand the nature of the neutrino and are therefore subject of great theoretical and experimental interest.
 Observation of the lepton number violating process  $0\nu\beta\beta$  would imply neutrinos have a Majorana component.

### 2018-04-09 Chapter-1/1-Theory.tex
s

@@
 \subsection{Make up of Neutrino Interactions}
-Any neutrino experiment involving the detection of neutrinos is concerned with neutrino interactions (and neutrino cross sections) on nuclei. 
-Given the invisible nature of neutrinos, characterizing the products of their interaction is the only method to retrieve their flavor and eventually energy.
-In real life experiments, this is complicated
+All neutrino experiments involving the detection of single neutrinos are concerned with neutrino interactions (and neutrino cross sections) on nuclei. 
+Given the invisible nature of neutrinos, characterizing the products of their interaction is the only method to assess their flavor and eventually reconstruct their energy.
+In real life experiments, this task is complicated by the structure of the nucleus.
+Nuclear effects modify the kinematic and final state of the neutrino interaction products and are particularly relevant in modern neutrino experiments that use fairly heavy nuclear targets to increase their event yields.
+Data on neutrino scattering off many different nuclei are available for both charged current (CC) and neutral current (NC) channels, as summarized here \cite{RevModPhys.84.1307}.
+For current and forthcoming oscillation experiments, the interesting neutrino energy region is around 1 GeV.
+
+%%%%%%%%%%%%%%%%%%%%%%%%  
+measurements performed in the ?
+0.1 ?
+300 GeV range with an emphasis on inclusive, quasi-elastic, and pion production processes, areas where we have the most experimental input at present (Table 49.1 and Table 49.2).
+For a more comprehensive discussion of neutrino cross sections, including neutrino-electron elastic scattering and lower energy neutrino measurements, the reader is directed to a recent review of this subject [1].
+Here, we survey existing experimental data on neutrino interactions and do not attempt to provide a census of the associated theoretical calculations, which are both important and plentiful.
+
+Inclusive Scattering
+Over the years, many experiments have measured the total inclusive cross section
+for neutrino (?? N ? ?? X) and antineutrino (
+?? N ? ?+ X) scattering off nucleons
+covering a broad range of neutrino energies.
+As can be seen in Fig.
+49.1, the inclusive cross section approaches a linear dependence on neutrino energy.
+Such behavior is expected for point-like scattering of neutrinos from quarks, an assumption which breaks down at lower energies.
+To provide a more complete picture, differential cross sections for such inclusive scattering processes have been reported ?
+these include measurements on iron from NuTeV [42] and, more recently, at lower neutrino energies on argon from ArgoNeuT [2,3] and on carbon from T2K [37].
+MINERvA has also provided new measurements of the ratios of the muon neutrino CC inclusive scattering cross section on a variety of nuclear targets such as lead, iron, and carbon [11].
+At high energy, the inclusive cross section is dominated by deep inelastic scattering (DIS).
+Several high energy neutrino experiments have measured the DIS cross sections for specific final states, for example opposite-sign dimuon production.
+The most recent dimuon cross section measurements include those from CHORUS [43], NOMAD [44], and NuTeV [45].
+At lower neutrino energies, the inclusive cross section is an additionally complex combination of quasi-elastic scattering
+
+Quasi-elastic scattering
+Quasi-elastic (QE) scattering is the dominant neutrino interaction for neutrino energies less than ?
+1 GeV and represents a large fraction of the signal samples in many neutrino oscillation experiments.
+Historically, neutrino (antineutrino) quasi-elastic scattering refers to the process, ??
+n ?
+??
+p (??
+p ?
+?+ n), where a charged lepton and single nucleon are ejected in the elastic interaction of a neutrino (or antineutrino) with a nucleon in the target material.
+This is the final state one would strictly observe, for example, in scattering off of a free nucleon target.
+Fig.
+49.2 displays the current status of existing measurements of ??
+and ??
+QE scattering cross sections as a function of neutrino energy.
+In this plot, and all others in this review, the prediction from a representative neutrino event generator (NUANCE) [46] provides a theoretical comparator.
+Other generators and more sophisticated calculations exist which can yield significantly different predictions [47].
+Note that modern experiments have recently opted to report QE cross sections as a function of final state muon or proton kinematics [17,18,48].
+Such distributions are more difficult to compare between experiments but are much less model-dependent and provide more stringent tests of the theory than cross sections as a function of neutrino energy (E?) or 4-momentum transfer (Q2).
+
+In many of these initial measurements of the neutrino QE cross section, bubble chamber experiments employed light targets (H2 or D2) and required both the detection of the final state muon and single nucleon?; thus the final state was clear and elastic kinematic conditions could be verified.
+The situation is more complicated, of course, for heavier nuclear targets.
+In this case, nuclear effects can impact the size and shape of the cross section as well as the final state kinematics and topology.
+Due to intranuclear hadron rescattering and the possible effects of correlations between target nucleons, additional nucleons may be ejected in the final state; hence, a QE interaction on a nuclear target does not always imply the ejection of a single nucleon.
+One therefore needs to take some care in defining what one means by neutrino QE scattering when scattering off targets heavier than H2 or D2.
+Adding to the complexity, recent MiniBooNE measurements of the ??
+and ??
+QE scattering cross sections on carbon near 1 GeV have revealed a significantly larger cross section than originally anticipated [17,18].
+Such an enhancement was observed many years prior in electron-nucleus scattering [57] and is believed to be due to the presence of correlations between target nucleons in the nucleus.
+As a result, the impact of such nuclear effects on neutrino QE scattering has recently been the subject of intense experimental and theoretical scrutiny with potential implications on event rates, nucleon emission, neutrino energy reconstruction, and neutrino/antineutrino ratios.
+The reader is referred to a recent review of the situation in [58,59].
+Additional measurements are clearly needed before a complete understanding is achieved.
+To help drive further progress, neutrino-nucleus QE cross sections have been reported for the first time in the form of double-differential distributions in muon kinematics, d2?/dT?dcos??, by MiniBooNE [17,18] thus reducing the model-dependence of the reported data and allowing a more rigorous two-dimensional test of the underlying nuclear theory.
+Experiments such as ArgoNeuT have begun to provide the first measurements of proton multiplicities in neutrino-argon QE scattering [4,48], a critical ingredient in understanding the hadronic side of these interactions and final state effects.
+MINOS, NOvA, and T2K have started to study QE interactions in their near detectors with sizable statistics [29,39,60].
+Most recently, MINERvA has measured the differential cross section, d?/dQ2QE, ?e QE scattering [61], single proton emission [14], and vertex energy in both ??
+and ??
+QE interactions in hydrocarbon [12,13].
+With the MiniBooNE results having revealed additional complexities in the QE channel, measurements from other neutrino experiments are crucial for getting a better handle on the underlying nuclear physics impacting neutrino-nucleus interactions.
+What we once thought was ?simple?
+QE scattering is in fact not so simple.
+In addition to such charged current investigations, measurements of the neutral current counterpart of this channel have also been performed.
+The most recent NC elastic scattering cross section measurements include those from BNL E734 [62], MiniBooNE [20,21,22], and T2K [40].
+A number of measurements of the Cabibbo- suppressed antineutrino QE hyperon production cross section have additionally been reported [56,63], although not in recent years.
+
+Pion Production
+In addition to such elastic processes, neutrinos can also inelastically scatter producing a nucleon excited state (?, N?).
+Such baryonic resonances quickly decay, most often to a nucleon and single-pion final state.
+Historically, experiments have measured various exclusive final states associated with these reactions, the majority of which have been on hydrogen and deuterium targets [1].
+In addition to such resonance production processes, neutrinos can also coherently scatter off of the entire nucleus and produce a distinctly forward-scattered single pion final state.
+BothCC(??A???A?+,??A??+A??
+)andNC(??A???A?0,??A???A?0) processes are possible in this case.
+Even though the level of coherent pion production is small compared to resonant processes, observations exist across a broad energy range andon multiple nuclear targets [64].
+More recently, several modern neutrino experiments have measured coherent pion production cross sections including ArgoNeuT [5], K2K [8], MINERvA [15], MiniBooNE [27], MINOS [65], NOMAD [32], and SciBooNE [34,36].
+As with QE scattering, a new appreciation for the significance of nuclear effects has surfaced in pion production channels, again due to the use of heavy nuclear targets in modern neutrino experiments.
+Many experiments have been careful to report cross sections for various detected final states, thereby not correcting for large and uncertain nuclear effects (e.g., pion rescattering, charge exchange, and absorption) which can introduce unwanted sources of uncertainty and model dependence.
+Recent measurements of single-pion cross sections, as published by K2K [9,10], MiniBooNE [24], and SciBooNE [35], take the form of ratios with respect to QE or CC inclusive scattering samples.
+Providing the most comprehensive survey of neutrino single-pion production to date, MiniBooNE has recently published a total of 16 single- and double-differential cross sections for both the final state muon (in the case of CC scattering) and pion in these interactions; thus, providing the first measurements of these distributions (Fig.
+49.3) [23?26].
+MINERvA has recently produced similar kinematic measurements at higher neutrino energies (Fig.
+49.4) [16,66].
+Regardless of the interaction channel, such differential cross section measurements (in terms of observed final state particle kinematics) are now preferred for their reduced model dependence and for the additional kinematic information they provide.
+Such a new direction has been the focus of modern measurements as opposed to the reporting of more model-dependent, historical cross sections as a function of E?
+or Q2.
+Together with similar results for other interaction channels, a better understanding and modeling of nuclear effects will be possible moving forward.It should be noted that baryonic resonances can also decay to multi-pion, other mesonic (K, ?, ?, etc.), and even photon final states.
+Experimental results for these channels are typically sparse or non-existent [1]; however, photon production processes can be an important background for ??
+?
+?e appearance searches and thus have become the focus of some recent experimental investigations; for example, in NOMAD [67].
+
+%%%%%%%%%%%%%%%%%%%%%%%%
 

### 2018-04-10 Chapter-1/1-Theory.tex
s

@@
 All neutrino experiments involving the detection of single neutrinos are concerned with neutrino interactions (and neutrino cross sections) on nuclei. 
-Given the invisible nature of neutrinos, characterizing the products of their interaction is the only method to assess their flavor and eventually reconstruct their energy.
-In real life experiments, this task is complicated by the structure of the nucleus.
-Nuclear effects modify the kinematic and final state of the neutrino interaction products and are particularly relevant in modern neutrino experiments that use fairly heavy nuclear targets to increase their event yields.
+Given the invisible nature of the neutrino, characterizing the products of its interaction is the only method to a) assess the neutrino existence, b) detect its flavor in case of a charge current interaction and c) eventually reconstruct its energy.
+The complexity of the nuclear structure complicate this task.
+Nuclear effects modify the kinematic and final state of the neutrino interaction products and play a key role in modern neutrino experiments that use fairly heavy nuclear targets.
 Data on neutrino scattering off many different nuclei are available for both charged current (CC) and neutral current (NC) channels, as summarized here \cite{RevModPhys.84.1307}.
-For current and forthcoming oscillation experiments, the interesting neutrino energy region is around 1 GeV.
+
+Historically, neutrino interactions with the nucleus are divided into three categories as a function of increasing neutrino energy: quasi elastic, resonant and deep inelastic scattering.
+All current and forthcoming oscillation experiments live in
+the 0.1-10 GeV transition region, which encompasses the energy where the quasi-elastic neutrino-nucleus interaction transitions into resonance scattering and the energy  resonance scattering  transitions in to deep inelastic scattering.
+\textcolor{red}{PLOT}
 
 %%%%%%%%%%%%%%%%%%%%%%%%  
-measurements performed in the ?
-0.1 ?
-300 GeV range with an emphasis on inclusive, quasi-elastic, and pion production processes, areas where we have the most experimental input at present (Table 49.1 and Table 49.2).
-For a more comprehensive discussion of neutrino cross sections, including neutrino-electron elastic scattering and lower energy neutrino measurements, the reader is directed to a recent review of this subject [1].
-Here, we survey existing experimental data on neutrino interactions and do not attempt to provide a census of the associated theoretical calculations, which are both important and plentiful.
-
-Inclusive Scattering
-Over the years, many experiments have measured the total inclusive cross section
-for neutrino (?? N ? ?? X) and antineutrino (
-?? N ? ?+ X) scattering off nucleons
-covering a broad range of neutrino energies.
-As can be seen in Fig.
-49.1, the inclusive cross section approaches a linear dependence on neutrino energy.
-Such behavior is expected for point-like scattering of neutrinos from quarks, an assumption which breaks down at lower energies.
-To provide a more complete picture, differential cross sections for such inclusive scattering processes have been reported ?
-these include measurements on iron from NuTeV [42] and, more recently, at lower neutrino energies on argon from ArgoNeuT [2,3] and on carbon from T2K [37].
-MINERvA has also provided new measurements of the ratios of the muon neutrino CC inclusive scattering cross section on a variety of nuclear targets such as lead, iron, and carbon [11].
-At high energy, the inclusive cross section is dominated by deep inelastic scattering (DIS).
-Several high energy neutrino experiments have measured the DIS cross sections for specific final states, for example opposite-sign dimuon production.
-The most recent dimuon cross section measurements include those from CHORUS [43], NOMAD [44], and NuTeV [45].
-At lower neutrino energies, the inclusive cross section is an additionally complex combination of quasi-elastic scattering
-
-Quasi-elastic scattering
+\textbf{Quasi-elastic scattering}
 Quasi-elastic (QE) scattering is the dominant neutrino interaction for neutrino energies less than ?
 1 GeV and represents a large fraction of the signal samples in many neutrino oscillation experiments.
-Historically, neutrino (antineutrino) quasi-elastic scattering refers to the process, ??
-n ?
-??
-p (??
-p ?
-?+ n), where a charged lepton and single nucleon are ejected in the elastic interaction of a neutrino (or antineutrino) with a nucleon in the target material.
+Historically, neutrino (antineutrino) quasi-elastic scattering refers to the process $\nu_\mu n \rightarrow \mu^- p$ and ($\bar\nu_\mu p \rightarrow \mu^+ n$) where a charged lepton and single nucleon are ejected in the elastic interaction of a neutrino (or antineutrino) with a nucleon in the target material.
 This is the final state one would strictly observe, for example, in scattering off of a free nucleon target.
+%% ADD DIFFERENCES WITH HEAVIER TARGETS
+In many of these initial measurements of the neutrino QE cross section, bubble chamber experiments employed light targets (H2 or D2) and required both the detection of the final state muon and single nucleon; thus the final state was clear and elastic kinematic conditions could be verified.
+The situation is more complicated, of course, for heavier nuclear targets.
+In this case, nuclear effects can impact the size and shape of the cross section as well as the final state kinematics and topology.
+Due to intra-nuclear hadron rescattering and the possible effects of correlations between target nucleons, additional nucleons may be ejected in the final state; hence, a QE interaction on a nuclear target does not always imply the ejection of a single nucleon.
+One therefore needs to take some care in defining what one means by neutrino QE scattering when scattering off targets heavier than H2 or D2.
+
 Fig.
-49.2 displays the current status of existing measurements of ??
-and ??
-QE scattering cross sections as a function of neutrino energy.
+49.2 displays the current status of existing measurements of $\nu_\mu$ and $\bar\nu_\mu$ QE scattering cross sections as a function of neutrino energy.
 In this plot, and all others in this review, the prediction from a representative neutrino event generator (NUANCE) [46] provides a theoretical comparator.
-Other generators and more sophisticated calculations exist which can yield significantly different predictions [47].
-Note that modern experiments have recently opted to report QE cross sections as a function of final state muon or proton kinematics [17,18,48].
+MiniBooNE XS enanchement Such an enhancement is believed to be due to the presence of correlations between target nucleons in the nucleus.
+As a result, the impact of such nuclear effects on neutrino QE scattering has recently been the subject of intense experimental and theoretical scrutiny with potential implications on event rates, nucleon emission, neutrino energy reconstruction, and neutrino/antineutrino ratios.
+Experiments such as ArgoNeuT have begun to provide the first measurements of proton multiplicities in neutrino-argon QE scattering [4,48], a critical ingredient in understanding the hadronic side of these interactions and final state effects.
+
+%Note that modern experiments have recently opted to report QE cross sections as a function of final state muon or proton kinematics [17,18,48].
 Such distributions are more difficult to compare between experiments but are much less model-dependent and provide more stringent tests of the theory than cross sections as a function of neutrino energy (E?) or 4-momentum transfer (Q2).
 
-In many of these initial measurements of the neutrino QE cross section, bubble chamber experiments employed light targets (H2 or D2) and required both the detection of the final state muon and single nucleon?; thus the final state was clear and elastic kinematic conditions could be verified.
-The situation is more complicated, of course, for heavier nuclear targets.
-In this case, nuclear effects can impact the size and shape of the cross section as well as the final state kinematics and topology.
-Due to intranuclear hadron rescattering and the possible effects of correlations between target nucleons, additional nucleons may be ejected in the final state; hence, a QE interaction on a nuclear target does not always imply the ejection of a single nucleon.
-One therefore needs to take some care in defining what one means by neutrino QE scattering when scattering off targets heavier than H2 or D2.
-Adding to the complexity, recent MiniBooNE measurements of the ??
-and ??
-QE scattering cross sections on carbon near 1 GeV have revealed a significantly larger cross section than originally anticipated [17,18].
-Such an enhancement was observed many years prior in electron-nucleus scattering [57] and is believed to be due to the presence of correlations between target nucleons in the nucleus.
-As a result, the impact of such nuclear effects on neutrino QE scattering has recently been the subject of intense experimental and theoretical scrutiny with potential implications on event rates, nucleon emission, neutrino energy reconstruction, and neutrino/antineutrino ratios.
-The reader is referred to a recent review of the situation in [58,59].
-Additional measurements are clearly needed before a complete understanding is achieved.
-To help drive further progress, neutrino-nucleus QE cross sections have been reported for the first time in the form of double-differential distributions in muon kinematics, d2?/dT?dcos??, by MiniBooNE [17,18] thus reducing the model-dependence of the reported data and allowing a more rigorous two-dimensional test of the underlying nuclear theory.
-Experiments such as ArgoNeuT have begun to provide the first measurements of proton multiplicities in neutrino-argon QE scattering [4,48], a critical ingredient in understanding the hadronic side of these interactions and final state effects.
-MINOS, NOvA, and T2K have started to study QE interactions in their near detectors with sizable statistics [29,39,60].
-Most recently, MINERvA has measured the differential cross section, d?/dQ2QE, ?e QE scattering [61], single proton emission [14], and vertex energy in both ??
-and ??
-QE interactions in hydrocarbon [12,13].
-With the MiniBooNE results having revealed additional complexities in the QE channel, measurements from other neutrino experiments are crucial for getting a better handle on the underlying nuclear physics impacting neutrino-nucleus interactions.
-What we once thought was ?simple?
-QE scattering is in fact not so simple.
 In addition to such charged current investigations, measurements of the neutral current counterpart of this channel have also been performed.
 The most recent NC elastic scattering cross section measurements include those from BNL E734 [62], MiniBooNE [20,21,22], and T2K [40].
-A number of measurements of the Cabibbo- suppressed antineutrino QE hyperon production cross section have additionally been reported [56,63], although not in recent years.
 
-Pion Production
-In addition to such elastic processes, neutrinos can also inelastically scatter producing a nucleon excited state (?, N?).
+\textbf{Pion Production}
+In addition to such elastic processes, neutrinos can also inelastically scatter producing a nucleon excited state ($\Delta$,N*).
 Such baryonic resonances quickly decay, most often to a nucleon and single-pion final state.
@@
 In addition to such resonance production processes, neutrinos can also coherently scatter off of the entire nucleus and produce a distinctly forward-scattered single pion final state.
-BothCC(??A???A?+,??A??+A??
-)andNC(??A???A?0,??A???A?0) processes are possible in this case.
-Even though the level of coherent pion production is small compared to resonant processes, observations exist across a broad energy range andon multiple nuclear targets [64].
+Both CC %(??A???A?+,??A??+A??
+)
+and NC %(??A???A?0,??A???A?0) 
+processes are possible in this case.
+Even though the level of coherent pion production is small compared to resonant processes, observations exist across a broad energy range and on multiple nuclear targets [64].
 More recently, several modern neutrino experiments have measured coherent pion production cross sections including ArgoNeuT [5], K2K [8], MINERvA [15], MiniBooNE [27], MINOS [65], NOMAD [32], and SciBooNE [34,36].
 As with QE scattering, a new appreciation for the significance of nuclear effects has surfaced in pion production channels, again due to the use of heavy nuclear targets in modern neutrino experiments.
-Many experiments have been careful to report cross sections for various detected final states, thereby not correcting for large and uncertain nuclear effects (e.g., pion rescattering, charge exchange, and absorption) which can introduce unwanted sources of uncertainty and model dependence.
+Many experiments have been careful to report cross sections for various detected final states, thereby not correcting for large and uncertain nuclear effects (e.g., pion re-scattering, charge exchange, and absorption) which can introduce unwanted sources of uncertainty and model dependence.
 Recent measurements of single-pion cross sections, as published by K2K [9,10], MiniBooNE [24], and SciBooNE [35], take the form of ratios with respect to QE or CC inclusive scattering samples.
@@
 ?e appearance searches and thus have become the focus of some recent experimental investigations; for example, in NOMAD [67].
+
+\textbf{Inclusive Scattering -- DIS}
+Over the years, many experiments have measured the total inclusive cross section ($\nu_\mu N \rightarrow \mu^- X$) and antineutrino ($\bar\nu_\mu N \rightarrow \mu^+ X$) scattering off nucleons covering a broad range of neutrino energies.
+As can be seen in Fig.
+49.1, the inclusive cross section approaches a linear dependence on neutrino energy.
+Such behavior is expected for point-like scattering of neutrinos from quarks, an assumption which breaks down at lower energies.
+To provide a more complete picture, differential cross sections for such inclusive scattering processes have been reported ?
+these include measurements on iron from NuTeV [42] and, more recently, at lower neutrino energies on argon from ArgoNeuT [2,3] and on carbon from T2K [37].
+MINERvA has also provided new measurements of the ratios of the muon neutrino CC inclusive scattering cross section on a variety of nuclear targets such as lead, iron, and carbon [11].
+At high energy, the inclusive cross section is dominated by deep inelastic scattering (DIS).
+Several high energy neutrino experiments have measured the DIS cross sections for specific final states, for example opposite-sign dimuon production.
+The most recent dimuon cross section measurements include those from CHORUS [43], NOMAD [44], and NuTeV [45].
+At lower neutrino energies, the inclusive cross section is an additionally complex combination of quasi-elastic scattering
 

### 2018-04-10 Chapter-1/1-Theory.tex
s

@@
 
-Historically, neutrino interactions with the nucleus are divided into three categories as a function of increasing neutrino energy: quasi elastic, resonant and deep inelastic scattering.
+Historically, neutrino interactions with the nucleus in the few GeV region are divided into three categories as a function of increasing neutrino energy: quasi elastic, resonant and deep inelastic scattering.
+In this simple framework, neutrino (antineutrino) quasi-elastic scattering refers to the process $\nu_\mu n \rightarrow \mu^- p$ and ($\bar\nu_\mu p \rightarrow \mu^+ n$) where a charged lepton and single nucleon are ejected in the elastic interaction of a neutrino (or antineutrino) with a nucleon in the target material.
+Resonant scattering refers to an inelastic collision producing a nucleon excited state ($\Delta$,N*).
+The excited resonance quickly decays, most often to a nucleon and single-pion final state.
+Deep inelastic refers to the head on collision between the neutrino and a parton inside the nucleon, producing a subsequent hadronization.
+This simple model
+
 All current and forthcoming oscillation experiments live in
-the 0.1-10 GeV transition region, which encompasses the energy where the quasi-elastic neutrino-nucleus interaction transitions into resonance scattering and the energy  resonance scattering  transitions in to deep inelastic scattering.
+the 0.1-10 GeV transition region, which encompasses the energy where the quasi-elastic neutrino-nucleus interaction transitions into resonance scattering and the energy  where resonance scattering  transitions in to deep inelastic scattering.
+
 \textcolor{red}{PLOT}
+more schematic: what 
+\begin{enumerate}
+\item QE
+\item Resonance
+\item DIS
+\end{enumerate}
 
+Complications
 %%%%%%%%%%%%%%%%%%%%%%%%
+
 \textbf{Quasi-elastic scattering}
-Quasi-elastic (QE) scattering is the dominant neutrino interaction for neutrino energies less than ?
-1 GeV and represents a large fraction of the signal samples in many neutrino oscillation experiments.
+Quasi-elastic (QE) scattering is the dominant neutrino interaction for neutrino energies less than $\sim1$ GeV and represents a large fraction of the signal samples in many neutrino oscillation experiments.
 Historically, neutrino (antineutrino) quasi-elastic scattering refers to the process $\nu_\mu n \rightarrow \mu^- p$ and ($\bar\nu_\mu p \rightarrow \mu^+ n$) where a charged lepton and single nucleon are ejected in the elastic interaction of a neutrino (or antineutrino) with a nucleon in the target material.
-This is the final state one would strictly observe, for example, in scattering off of a free nucleon target.
-%% ADD DIFFERENCES WITH HEAVIER TARGETS
-In many of these initial measurements of the neutrino QE cross section, bubble chamber experiments employed light targets (H2 or D2) and required both the detection of the final state muon and single nucleon; thus the final state was clear and elastic kinematic conditions could be verified.
+
+This is the final state one would strictly observe, for example, in scattering off of a free nucleon target. %% ADD DIFFERENCES WITH HEAVIER TARGETS
 The situation is more complicated, of course, for heavier nuclear targets.
@@
 One therefore needs to take some care in defining what one means by neutrino QE scattering when scattering off targets heavier than H2 or D2. 
-
 Fig.

### 2018-04-10 Chapter-1/1-Theory.tex
s

--- 2018-04-10 Chapter-1/1-Theory.tex
@@
 All neutrino experiments involving the detection of single neutrinos are concerned with neutrino interactions (and neutrino cross sections) on nuclei. 
-Given the invisible nature of the neutrino, characterizing the products of its interaction is the only method to a) assess the neutrino existence, b) detect its flavor in case of a charge current interaction and c) eventually reconstruct its energy.
-The complexity of the nuclear structure complicate this task.
-Nuclear effects modify the kinematic and final state of the neutrino interaction products and play a key role in modern neutrino experiments that use fairly heavy nuclear targets.
+Given the invisible nature of the neutrino, characterizing the products of its interaction is the only method to a) assess the neutrino presence, b) detect its flavor in case of a charge current interaction and c) eventually reconstruct its energy.
+
+Historically, neutrino interactions with the nucleus in the GeV region are divided into three categories as a function of increasing neutrino energy: quasi elastic (QE), resonant and deep inelastic scattering (DIS).
+All current and forthcoming oscillation experiments live in the 0.1-10 GeV transition region, which encompasses the energy where the QE neutrino-nucleus interaction transitions into resonant scattering and the energy  where resonance scattering  transitions in to DIS.
+ Neutrino and antineutrino QE charge current scattering refers to the process $\nu_\mu n \rightarrow \mu^- p$ and $\bar\nu_\mu p \rightarrow \mu^+ n$ where a charged lepton and single nucleon are ejected in the elastic interaction and the target nucleus remains at ground state.
+%of a neutrino (or antineutrino) with a nucleon in the target material.
+ Resonant scattering refers to an inelastic collision producing a nucleon excited state ($\Delta$,N*); the excited resonance quickly decays, most often to a nucleon and single-pion final state.
+DIS refers to the head-on collision between the neutrino and a parton inside the nucleon, producing hadronization and subsequent  abundant production of mesons and nucleons.
+In addition to such interactions between the neutrino and a single component of the nucleus, neutrinos can also interact with the nucleus as a whole, albeit more rarely, a well documented process called coherent meson production scattering \cite{Vilain1993}; the signature of such process is the production of a distinctly forward-scattered single meson final state, most often a pion.
+This simple picture of neutrino interactions works rather well for scattering
+off of light targets, such as the H$_2$ and D$_2$ of bubble chamber experiments \cite{RevModPhys.84.1307}, but the complexity of the nuclear structure for heavier nuclei such as argon complicates this model.
+
+As we will discuss in Chapter \ref{ch:2}, the properties of argon make it a good candidate for interacting medium in neutrino experiments; in particular the density of its interaction centers  augments the yield of neutrino interactions and allows for relatively compact detectors.
+Though, the choice of a relatively heavy nuclear target comes at the cost of enhancing  nuclear effects which modify the kinematic and final state of the neutrino interaction products.
+
+Nuclear effects can potentially affect the neutrino event rates, nucleon emission, neutrino energy reconstruction, and neutrino/antineutrino ratios, carrying deep implications for oscillation experiments.
+Even in the case of ``simple" QE scattering  nuclear effects can impact the size and shape of the cross section as well as the final state kinematics and topology.
+Intra-nuclear hadron rescattering and correlation effects between the target nucleons can cause the ejection of additional nucleons in the final state.
+In case of resonant and DIS scattering, the hadronic interactions of meson and nucleons produced in the decay of the resonance or during hadronization  complicate this picture even more.
+A large source of uncertainty in modeling  nuclear effects in neutrino interactions come from mesons interactions (and re-interactions) in the nucleus, e.g., pion re-scattering, charge exchange, and absorption.
+
+A renewed interest for neutrino cross section measurements surged in recent year, along with lively discussion on the data reporting; the historical method of reporting the neutrino cross section as a function of the neutrino energy or momentum transferred shakes under the weight of its dependency to the chosen nuclear model.
+On one hand, correcting for nuclear effects in neutrino interaction can introduce unwanted sources of uncertainty and model dependence due to the mis-modeling of the meson interactions.
+On the other, avoiding this correction however makes a comparison between neutrino interactions on different target nuclei extremely difficult.
+
 Data on neutrino scattering off many different nuclei are available for both charged current (CC) and neutral current (NC) channels, as summarized here \cite{RevModPhys.84.1307}.
-
-Historically, neutrino interactions with the nucleus are divided into three categories as a function of increasing neutrino energy: quasi elastic, resonant and deep inelastic scattering.
-All current and forthcoming oscillation experiments live in
-the 0.1-10 GeV transition region, which encompasses the energy where the quasi-elastic neutrino-nucleus interaction transitions into resonance scattering and the energy  resonance scattering  transitions in to deep inelastic scattering.
-\textcolor{red}{PLOT}
-
-%%%%%%%%%%%%%%%%%%%%%%%%  
-\textbf{Quasi-elastic scattering}
-Quasi-elastic (QE) scattering is the dominant neutrino interaction for neutrino energies less than ?
-1 GeV and represents a large fraction of the signal samples in many neutrino oscillation experiments.
-Historically, neutrino (antineutrino) quasi-elastic scattering refers to the process $\nu_\mu n \rightarrow \mu^- p$ and ($\bar\nu_\mu p \rightarrow \mu^+ n$) where a charged lepton and single nucleon are ejected in the elastic interaction of a neutrino (or antineutrino) with a nucleon in the target material.
-This is the final state one would strictly observe, for example, in scattering off of a free nucleon target.
-%% ADD DIFFERENCES WITH HEAVIER TARGETS
-In many of these initial measurements of the neutrino QE cross section, bubble chamber experiments employed light targets (H2 or D2) and required both the detection of the final state muon and single nucleon; thus the final state was clear and elastic kinematic conditions could be verified.
-The situation is more complicated, of course, for heavier nuclear targets.
-In this case, nuclear effects can impact the size and shape of the cross section as well as the final state kinematics and topology.
-Due to intra-nuclear hadron rescattering and the possible effects of correlations between target nucleons, additional nucleons may be ejected in the final state; hence, a QE interaction on a nuclear target does not always imply the ejection of a single nucleon.
-One therefore needs to take some care in defining what one means by neutrino QE scattering when scattering off targets heavier than H2 or D2.
-
-Fig.
-49.2 displays the current status of existing measurements of $\nu_\mu$ and $\bar\nu_\mu$ QE scattering cross sections as a function of neutrino energy.
-In this plot, and all others in this review, the prediction from a representative neutrino event generator (NUANCE) [46] provides a theoretical comparator.
-MiniBooNE XS enanchement Such an enhancement is believed to be due to the presence of correlations between target nucleons in the nucleus.
-As a result, the impact of such nuclear effects on neutrino QE scattering has recently been the subject of intense experimental and theoretical scrutiny with potential implications on event rates, nucleon emission, neutrino energy reconstruction, and neutrino/antineutrino ratios.
-Experiments such as ArgoNeuT have begun to provide the first measurements of proton multiplicities in neutrino-argon QE scattering [4,48], a critical ingredient in understanding the hadronic side of these interactions and final state effects.
-
-%Note that modern experiments have recently opted to report QE cross sections as a function of final state muon or proton kinematics [17,18,48].
-Such distributions are more difficult to compare between experiments but are much less model-dependent and provide more stringent tests of the theory than cross sections as a function of neutrino energy (E?) or 4-momentum transfer (Q2).
-
-In addition to such charged current investigations, measurements of the neutral current counterpart of this channel have also been performed.
-The most recent NC elastic scattering cross section measurements include those from BNL E734 [62], MiniBooNE [20,21,22], and T2K [40].
-
-\textbf{Pion Production}
-In addition to such elastic processes, neutrinos can also inelastically scatter producing a nucleon excited state ($\Delta$,N*).
-Such baryonic resonances quickly decay, most often to a nucleon and single-pion final state.
-Historically, experiments have measured various exclusive final states associated with these reactions, the majority of which have been on hydrogen and deuterium targets [1].
-In addition to such resonance production processes, neutrinos can also coherently scatter off of the entire nucleus and produce a distinctly forward-scattered single pion final state.
-Both CC %(??A???A?+,??A??+A??
-)
-and NC %(??A???A?0,??A???A?0) 
-processes are possible in this case.
-Even though the level of coherent pion production is small compared to resonant processes, observations exist across a broad energy range and on multiple nuclear targets [64].
-More recently, several modern neutrino experiments have measured coherent pion production cross sections including ArgoNeuT [5], K2K [8], MINERvA [15], MiniBooNE [27], MINOS [65], NOMAD [32], and SciBooNE [34,36].
-As with QE scattering, a new appreciation for the significance of nuclear effects has surfaced in pion production channels, again due to the use of heavy nuclear targets in modern neutrino experiments.
-Many experiments have been careful to report cross sections for various detected final states, thereby not correcting for large and uncertain nuclear effects (e.g., pion re-scattering, charge exchange, and absorption) which can introduce unwanted sources of uncertainty and model dependence.
-Recent measurements of single-pion cross sections, as published by K2K [9,10], MiniBooNE [24], and SciBooNE [35], take the form of ratios with respect to QE or CC inclusive scattering samples.
-Providing the most comprehensive survey of neutrino single-pion production to date, MiniBooNE has recently published a total of 16 single- and double-differential cross sections for both the final state muon (in the case of CC scattering) and pion in these interactions; thus, providing the first measurements of these distributions (Fig.
-49.3) [23?26].
-MINERvA has recently produced similar kinematic measurements at higher neutrino energies (Fig.
-49.4) [16,66].
-Regardless of the interaction channel, such differential cross section measurements (in terms of observed final state particle kinematics) are now preferred for their reduced model dependence and for the additional kinematic information they provide.
-Such a new direction has been the focus of modern measurements as opposed to the reporting of more model-dependent, historical cross sections as a function of E?
-or Q2.
-Together with similar results for other interaction channels, a better understanding and modeling of nuclear effects will be possible moving forward.It should be noted that baryonic resonances can also decay to multi-pion, other mesonic (K, ?, ?, etc.), and even photon final states.
-Experimental results for these channels are typically sparse or non-existent [1]; however, photon production processes can be an important background for ??
-?
-?e appearance searches and thus have become the focus of some recent experimental investigations; for example, in NOMAD [67].
-
-\textbf{Inclusive Scattering -- DIS}
-Over the years, many experiments have measured the total inclusive cross section ($\nu_\mu N \rightarrow \mu^- X$) and antineutrino ($\bar\nu_\mu N \rightarrow \mu^+ X$) scattering off nucleons covering a broad range of neutrino energies.
-As can be seen in Fig.
-49.1, the inclusive cross section approaches a linear dependence on neutrino energy.
-Such behavior is expected for point-like scattering of neutrinos from quarks, an assumption which breaks down at lower energies.
-To provide a more complete picture, differential cross sections for such inclusive scattering processes have been reported ?
-these include measurements on iron from NuTeV [42] and, more recently, at lower neutrino energies on argon from ArgoNeuT [2,3] and on carbon from T2K [37].
-MINERvA has also provided new measurements of the ratios of the muon neutrino CC inclusive scattering cross section on a variety of nuclear targets such as lead, iron, and carbon [11].
-At high energy, the inclusive cross section is dominated by deep inelastic scattering (DIS).
-Several high energy neutrino experiments have measured the DIS cross sections for specific final states, for example opposite-sign dimuon production.
-The most recent dimuon cross section measurements include those from CHORUS [43], NOMAD [44], and NuTeV [45].
-At lower neutrino energies, the inclusive cross section is an additionally complex combination of quasi-elastic scattering
-
-%%%%%%%%%%%%%%%%%%%%%%%%
+A summary of is reported in Figure \textcolor{red}{PLOT} where the (NUANCE) [46] event generator provides a theoretical comparator.
 

### 2018-04-10 Chapter-1/1-Theory.tex
s

@@
 
-Historically, neutrino interactions with the nucleus in the GeV region are divided into three categories as a function of increasing neutrino energy: quasi elastic (QE), resonant and deep inelastic scattering (DIS).
-All current and forthcoming oscillation experiments live in the 0.1-10 GeV transition region, which encompasses the energy where the QE neutrino-nucleus interaction transitions into resonant scattering and the energy  where resonance scattering  transitions in to DIS.
+Historically, neutrino interactions with the nucleus in the GeV region are divided into three categories as a function of increasing neutrino energy: quasi elastic (QE), resonant, and deep inelastic (DIS) scattering.
+All current and forthcoming oscillation experiments live in the 0.1-10 GeV transition region, which encompasses the energy where the QE neutrino-nucleus interaction transitions into resonant scattering and the energy  where resonance scattering  transitions into DIS.
  Neutrino and antineutrino QE charge current scattering refers to the process $\nu_\mu n \rightarrow \mu^- p$ and $\bar\nu_\mu p \rightarrow \mu^+ n$ where a charged lepton and single nucleon are ejected in the elastic interaction and the target nucleus remains at ground state.
 %of a neutrino (or antineutrino) with a nucleon in the target material.
- Resonant scattering refers to an inelastic collision producing a nucleon excited state ($\Delta$,N*); the excited resonance quickly decays, most often to a nucleon and single-pion final state.
+ Resonant scattering refers to an inelastic collision producing a nucleon excited state ($\Delta$,N*) -- the resonance, indeed -- which then quickly decays, most often to a nucleon and single-pion final state.
 DIS refers to the head-on collision between the neutrino and a parton inside the nucleon, producing hadronization and subsequent  abundant production of mesons and nucleons.
@@
 Nuclear effects can potentially affect the neutrino event rates, nucleon emission, neutrino energy reconstruction, and neutrino/antineutrino ratios, carrying deep implications for oscillation experiments.
-Even in the case of ``simple" QE scattering  nuclear effects can impact the size and shape of the cross section as well as the final state kinematics and topology.
-Intra-nuclear hadron rescattering and correlation effects between the target nucleons can cause the ejection of additional nucleons in the final state.
+Even in the case of ``simple" QE scattering, intra-nuclear hadron rescattering and correlation effects between the target nucleons can cause the ejection of additional nucleons in the final state, modifying the final state kinematics and topology.
 In case of resonant and DIS scattering, the hadronic interactions of meson and nucleons produced in the decay of the resonance or during hadronization  complicate this picture even more.
@@
 
-A renewed interest for neutrino cross section measurements surged in recent year, along with lively discussion on the data reporting; the historical method of reporting the neutrino cross section as a function of the neutrino energy or momentum transferred shakes under the weight of its dependency to the chosen nuclear model.
-On one hand, correcting for nuclear effects in neutrino interaction can introduce unwanted sources of uncertainty and model dependence due to the mis-modeling of the meson interactions.
-On the other, avoiding this correction however makes a comparison between neutrino interactions on different target nuclei extremely difficult.
+A renewed interest for neutrino cross section measurements surged in recent year, along with lively discussion on the data reporting; the historical method of reporting the neutrino cross section as a function of the neutrino energy or momentum transferred shakes under the weight of its dependency on the chosen nuclear model.
+On one hand, correcting for nuclear effects in neutrino interaction can introduce unwanted sources of uncertainty and model dependence especially due to the mis-modeling of the meson interactions.
+On the other, avoiding this correction makes a comparison between neutrino interactions on different target nuclei extremely difficult.
 
 Data on neutrino scattering off many different nuclei are available for both charged current (CC) and neutral current (NC) channels, as summarized here \cite{RevModPhys.84.1307}.
-A summary of is reported in Figure \textcolor{red}{PLOT} where the (NUANCE) [46] event generator provides a theoretical comparator.
+A summary of the results on QE, resonant and DIS scattering for neutrinos and antineutrinos from accelerators on different target is reported in Figure \ref{fig:nuXS}, where the (NUANCE) \cite{Casper2002} event generator is used as comparison with the theory.
+\begin{figure}[hbpt]
+\centering
+\includegraphics[scale=0.6]{Chapter-1/Images/NuCrossSec.png}
+\includegraphics[scale=0.95]{Chapter-1/Images/AntiNuCrossSec.png}
+\caption{Total neutrino (top) and antineutrino (bottom) CC cross sections per nucleon divided by neutrino energy as a function of energy as reported in \cite{RevModPhys.84.1307}.
+Predictions for the total (black), the QE (red), resonant (blue) and DIS (green) are provided by the NUANCE generator.
+The quasi-elastic scattering data and predictions have been averaged over neutron and proton targets (isoscalar target).}
+\label{fig:nuXS}
+\end{figure}
 

### 2018-04-10 Chapter-1/1-Theory.tex
s

@@
 
-\subsection{Make up of Neutrino Interactions}
+\subsection{Make up of Neutrino Interactions}\label{ch:NuInt}
 All neutrino experiments involving the detection of single neutrinos are concerned with neutrino interactions (and neutrino cross sections) on nuclei. 
@@
 All current and forthcoming oscillation experiments live in the 0.1-10 GeV transition region, which encompasses the energy where the QE neutrino-nucleus interaction transitions into resonant scattering and the energy  where resonance scattering  transitions into DIS.
- Neutrino and antineutrino QE charge current scattering refers to the process $\nu_\mu n \rightarrow \mu^- p$ and $\bar\nu_\mu p \rightarrow \mu^+ n$ where a charged lepton and single nucleon are ejected in the elastic interaction and the target nucleus remains at ground state.
+ Neutrino and antineutrino QE charge current scattering refers to the process $\nu_l n \rightarrow l^- p$ and $\bar\nu_l p \rightarrow l^+ n$ where a charged lepton and single nucleon are ejected in the elastic interaction and the target nucleus remains at ground state.
 %of a neutrino (or antineutrino) with a nucleon in the target material.
@@
 We then describe the signal signature and historical measurements of pion-nucleus cross section, as well as the implementation of this cross sections in the current version of the simulation package used by LArIAT.
+
 \subsubsection{$\pi^{-}$Ar Cross section in the Context of Neutrino Searches}
+As outlined in \ref{ch:NuInt}, neutrino experiments use the products of neutrino interactions to identify the energy and flavor of the incoming neutrino.
+In particular,  assumptions on the nuclear models and on the interaction of hadrons in the nucleus bridge the measurement of the products of a neutrino interaction to the reconstruction of the neutrino energy and flavor.
+Understanding pion hadronic interactions with the nucleus is particularly important to model resonant, DIS and coherent pion production in neutrino interactions, where the presence of pions in the nucleus is abundant.
+For example, in case of resonant scattering
+\begin{equation}
+\nu_l + N \rightarrow l + \Delta/N^* \rightarrow l+ \pi + N' 
+\end{equation}
+the $\Delta$ and $N^*$ and excited states will decay inside the nucleus producing pions. The decay modes for the lower mass $\Delta$ (1232) and $N^*$(1440)  are listed in table \ref{tab:decaysdelta}.
+
+\begin{table}[b]
+\centering
+\begin{tabular}{|l|l|}
+\hline
+$\Delta$ (1232) 3/2$^+$                 & $N^*$ (1440) 1/2$^+$    \\ \hline
+$\Delta^{++} \text{(uuu)} \rightarrow p \pi^+$ & $N^* \rightarrow N \pi$ \\
+$\Delta^{+} \text{(uud)} \rightarrow n \pi^+$  &  $N^* \rightarrow N \pi\pi$\\ 
+$\Delta^{+} \text{(uud)} \rightarrow p \pi^0$  &                         \\ 
+$\Delta^{0} \text{(udd)} \rightarrow n \pi^0$  &                         \\ 
+$\Delta^{0} \text{(udd)} \rightarrow p \pi^-$  &                         \\ 
+$\Delta^{-} \text{(ddd)} \rightarrow n \pi^-$  &                         \\ \hline
+\end{tabular}
+\caption{Main decay modes of the lightest Delta resonance and Nucleon excited state.}
+\label{tab:decaysdelta}
+\end{table}
+
 \subsubsection{$\pi^{-}$Ar Cross Section in the Context of Light Mesons Interaction with Nuclei}

### 2018-04-10 Chapter-1/1-Theory.tex
s

@@
 As outlined in \ref{ch:NuInt}, neutrino experiments use the products of neutrino interactions to identify the energy and flavor of the incoming neutrino.
-In particular,  assumptions on the nuclear models and on the interaction of hadrons in the nucleus bridge the measurement of the products of a neutrino interaction to the reconstruction of the neutrino energy and flavor.
-Understanding pion hadronic interactions with the nucleus is particularly important to model resonant, DIS and coherent pion production in neutrino interactions, where the presence of pions in the nucleus is abundant.
+In particular,  assumptions on the nuclear models and on the interaction of hadrons in the nucleus performed at the level of the neutrino event generator bridge the measurement of the products of a neutrino interaction to the reconstruction of the neutrino energy and flavor.
+Understanding pion hadronic interactions with the nucleus is particularly important to model correctly resonant, DIS and coherent pion production in neutrino interactions, where the presence of pions in the nucleus is abundant.
 For example, in case of resonant scattering
 \begin{equation}
-\nu_l + N \rightarrow l + \Delta/N^* \rightarrow l+ \pi + N' 
+\nu_l + N \rightarrow l + \Delta/N^* \rightarrow l+ \pi + N', 
 \end{equation}
-the $\Delta$ and $N^*$ and excited states will decay inside the nucleus producing pions. The decay modes for the lower mass $\Delta$ (1232) and $N^*$(1440)  are listed in table \ref{tab:decaysdelta}.
+the $\Delta$ and $N^*$ and excited states will decay hadronically in matters of $\sim10^{-24}$~s inside the nucleus producing pions which will bounce within the nuclear medium.
+The decay modes for the lower mass $\Delta$ (1232) and $N^*$(1440)  are listed in table \ref{tab:decaysdelta}.
 
@@
 \end{table}
+
+The key elements of a neutrino event generators for resonance and DIS events are the nuclear model and the hadron treatment (both production and transportation).
+As one of the most popular event generators for LAr experiments, we take GENIE~\cite{Andreopoulos:2009rq} here as an example.
+The nuclear model used by GENIE for all processes is a relativistic Fermi gas (RFG) modified to incorporate short range nucleon-nucleon correlations \cite{PhysRevD.24.1400}.
+This means that the initial momentum and binding energy of the struck nucleon is determined by assuming nucleons inside the nucleus are quasi-free, acting independently in the mean field of the nucleus.
+For A $ >$ 20 like argon, the 2-parameter Woods-Saxon shell model for density function is used.
+The GENIE module INTRANUKE \cite{Juszczak2006} is used to simulate the final-state interactions (FSI) which is the hadron re-interaction inside the nucleus.
+This module places the outgoing particles in the nucleus and propagates them.
+In the INRANUKE model, hadrons cannot undergo more than one FSI per event.
+When possible, external hadron-nucleus scattering data are used to tune INTRANUKE.
+Since no data is available for Argon, an interpolation of data from heavier and lighter nuclei is used for  the pion-argon cross section.
 

### 2018-04-10 Chapter-1/1-Theory.tex
s

@@
 
-Critical challenges await  the next decade of experimental neutrino physics.
-% will entail an even deeper understanding of the neutrino mixing pattern, investigation of the neutrino mass origin, the determination of the number of neutrinos and their nature, and the measurement of CP violation in the lepton sector.
-%%% Plug to LArTPC %%%%%%%%%%%%
-Following the recommendation of the latest Particle Physics Project Prioritization Panel  \cite{P5}, the US  is dedicating substantial resources to the development of a short- and long- baseline neutrino program to address many of these fundamental questions.
-This program pivots on the Liquid Argon Time Projection Chamber (LArTPC) detector technology which will be described in \ref{ch:1}.
-
-The main goals of these research programs include:
-\begin{itemize}
-\item[-] Assessment of the existence of right-handed sterile neutrinos. 
-\item[-] Determination of the sign of $\Delta m^2_{13}$ (or $\Delta m^2_{23}$), i.e., the neutrino mass hierarchy.
-\item[-] Determination of the octant, i.e.  whether $\theta_{23}$ is maximal.
-\item[-] Determination the status of CP symmetry in the lepton sector.
-\end{itemize}
-
 \subsection{Towards a more fundamental theory: GUTs}
@@
 \section{Motivations for Hadronic Cross Sections in Argon}
+Critical challenges await  the next decade of experimental physics at the intensity frontier.
+% will entail an even deeper understanding of the neutrino mixing pattern, investigation of the neutrino mass origin, the determination of the number of neutrinos and their nature, and the measurement of CP violation in the lepton sector.
+%%% Plug to LArTPC %%%%%%%%%%%%
+Following the recommendation of the latest Particle Physics Project Prioritization Panel  \cite{P5}, the US  is dedicating substantial resources to the development of a short- and long- baseline neutrino program to address many of open questions in neutrino physics today.
+This program pivots on the Liquid Argon Time Projection Chamber (LArTPC) detector technology which will be described in \ref{ch:1}.
+
+The main goals of these research programs include:
+\begin{itemize}
+\item[-] Assessment of the existence of right-handed sterile neutrinos. 
+\item[-] Determination of the sign of $\Delta m^2_{13}$ (or $\Delta m^2_{23}$), i.e., the neutrino mass hierarchy.
+\item[-] Determination of the octant, i.e.  whether $\theta_{23}$ is maximal.
+\item[-] Determination the status of CP symmetry in the lepton sector.
+\item[-] \textcolor{red}{something about GUTs}.
+\end{itemize}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 As outlined in \ref{ch:NuInt}, neutrino experiments use the products of neutrino interactions to identify the energy and flavor of the incoming neutrino.
-In particular,  assumptions on the nuclear models and on the interaction of hadrons in the nucleus performed at the level of the neutrino event generator bridge the measurement of the products of a neutrino interaction to the reconstruction of the neutrino energy and flavor.
-Understanding pion hadronic interactions with the nucleus is particularly important to model correctly resonant, DIS and coherent pion production in neutrino interactions, where the presence of pions in the nucleus is abundant.
+Pions are a common product of neutrino interaction, especially in resonant scattering, DIS and coherent pion production.
+For neutrino experiments in argon, there are two main reasons why understanding pion hadronic interactions with argon is important: to model the behavior of the pion inside the nucleus struck by the neutrino and to model the behavior of the pion during its propagation inside the detector medium.
+
+Assumptions on the nuclear models and on the interaction of hadrons in the nucleus performed at the level of the neutrino event generator bridge the measurement of the products of a neutrino interaction to the reconstruction of the neutrino energy and flavor.
+Thus, understanding pion hadronic interactions with the nucleus is particularly important to model correctly resonant, DIS and coherent pion production in neutrino interactions, where the presence of pions in the nucleus is abundant.
 For example, in case of resonant scattering
@@
 The decay modes for the lower mass $\Delta$ (1232) and $N^*$(1440)  are listed in table \ref{tab:decaysdelta}.
-
 \begin{table}[b]
@@
 \end{table}
-
 The key elements of a neutrino event generators for resonance and DIS events are the nuclear model and the hadron treatment (both production and transportation).
@@
 When possible, external hadron-nucleus scattering data are used to tune INTRANUKE.
-Since no data is available for Argon, an interpolation of data from heavier and lighter nuclei is used for  the pion-argon cross section.
+Since no data is available for Argon, GENIE uses an interpolation of data from heavier and lighter nuclei for the pion-argon cross section leading to big uncertainties in the INTRANUKE module.
 

### 2018-04-11 Chapter-1/1-Theory.tex
s

@@
 Current experimental results \textcolor{red}{cite NOVA T2K} are consistent with  $\theta_{23}=45^\circ$, which would imply maximal mixing between $\nu_\mu$ - $\nu_\tau$, hinting to an intriguing new symmetry.
-Therefore, a precise measurement of $\theta_{23}$  is of great interest for theoretical models of quark-lepton universality [59,84,85,86,87,88] , whose  quark and lepton mixing matrices are proportional to the deviation of $\theta_{23}$ from $45^\circ$.
+Therefore, a precise measurement of $\theta_{23}$  is of great interest for theoretical models of quark-lepton universality \textcolor{red}{change cit [59,84,85,86,87,88]}, whose  quark and lepton mixing matrices are proportional to the deviation of $\theta_{23}$ from $45^\circ$.
 
@@
 
-The SM rather complex group structure, where a gauge group is formed with the direct product of other three groups as shown in eq.
-\ref{eq:SMGroup},  is unexplained.
-Also, the SM fails to include a suitable dark matter candidate and a mechanisms that accounts for the baryon asymmetry of the universe.
-Within the SM, a total of 25 parameters remain seemingly arbitrary and need to be fitted to data: 3 gauge couplings, 9 charged fermion masses, 3 mixing angles and one CP phase in the CKM matrix, the Higgs mass and quartic coupling, $\theta_{QCD}$, 3 neutrino masses, 3 neutrino mixing angles, 1 Dirac phase and, eventually,  2 Majorana phases.
+The SM does not include a suitable dark matter candidate and a mechanisms that accounts for the baryon asymmetry of the universe.
+Additionally, a total of \textcolor{red}{25} parameters remain seemingly arbitrary and need to be fitted to data: 3 gauge couplings, 9 charged fermion masses, 3 mixing angles and one CP phase in the CKM matrix, the Higgs mass and quartic coupling, $\theta_{QCD}$, 3 neutrino masses, 3 neutrino mixing angles, 1 Dirac phase and, eventually,  2 Majorana phases.
+
+From a group theory perspective, the SM has a rather complex group structure, where a gauge group is formed with the direct product of other three groups as shown in eq.
+\ref{eq:SMGroup}.
+Drawing a parallel with the electroweak symmetry breaking mechanism, where the $ SU(2)_T \otimes U(1)_Y$ is recovered from $U(1)_{EM}$, an interesting line of simplification for the SM group structure would be to think about a similar mechanism where $SU(3)_C  \otimes SU(2)_T \otimes U(1)_Y$ is recovered from an hypothetical larger group.
+Just as the electroweak unification becomes evident at energies higher than the Higgs VEV,  a direct manifestation of Grand Unification Theories (GUTs) would occur at even higher energies.
+
+As the smallness of neutrino masses suggests to the existence of a higher mass scale \textcolor{red}{IS THIS TRUE?}, an other, even stronger, hint to Grand Unification comes from the slope of running of the coupling constants.
+The coupling constants for the electromagnetic, weak and strong interactions in the SM vary as a function of the interaction energy as shown in figure \ref{fig:constants}; they do not exactly meet, but their trend is interesting enough to push theorists to construct theories where perfect unification is achieved with the addition of new particles.
+
+%the running is parametrized as follows
+%where $b_i$ are coefficients for each interaction, and  the constants assume the same value $\alpha_{GUT}$ at energy $E=M_{GUT}$
 
@@
 Following the recommendation of the latest Particle Physics Project Prioritization Panel  \cite{P5}, the US  is dedicating substantial resources to the development of a short- and long- baseline neutrino program to address many of open questions in neutrino physics today.
-This program pivots on the Liquid Argon Time Projection Chamber (LArTPC) detector technology which will be described in \ref{ch:1}.
+This program pivots on the Liquid Argon Time Projection Chamber (LArTPC) detector technology which will be described in Chapter \ref{ch:2}.
 
@@
 
-Assumptions on the nuclear models and on the interaction of hadrons in the nucleus performed at the level of the neutrino event generator bridge the measurement of the products of a neutrino interaction to the reconstruction of the neutrino energy and flavor.
+Assumptions on the nuclear models and on the interaction of hadrons inside the nucleus performed at the level of the neutrino event generator bridge the measurement of the products of a neutrino interaction to the reconstruction of the neutrino energy and flavor.
 Thus, understanding pion hadronic interactions with the nucleus is particularly important to model correctly resonant, DIS and coherent pion production in neutrino interactions, where the presence of pions in the nucleus is abundant.
@@
 Since no data is available for Argon, GENIE uses an interpolation of data from heavier and lighter nuclei for the pion-argon cross section leading to big uncertainties in the INTRANUKE module.
+
+Once the pion has left the target nucleus, the pion-argon hadronic cross section plays an important role in the pion transportation inside the argon medium: processes like pion absorption with emission of protons or pion charge exchange can greatly modify the topology of a neutrino interactions in the detector and lead to errors in the event classification.
+Being able to reconstruct the details of pions inside the detector is an imperative for modern liquid argon neutrino experiment to achieve the design resolution for their key physics measurements.
 

### 2018-04-11 Chapter-1/1-Theory.tex
s

@@
 \ref{eq:SMGroup}.
-Drawing a parallel with the electroweak symmetry breaking mechanism, where the $ SU(2)_T \otimes U(1)_Y$ is recovered from $U(1)_{EM}$, an interesting line of simplification for the SM group structure would be to think about a similar mechanism where $SU(3)_C  \otimes SU(2)_T \otimes U(1)_Y$ is recovered from an hypothetical larger group.
+Drawing a parallel with the electroweak symmetry breaking mechanism, where the $ SU(2)_T \otimes U(1)_Y$ is recovered from $U(1)_{EM}$, an interesting line of simplification for the SM group structure would be to devise a similar mechanism where $SU(3)_C  \otimes SU(2)_T \otimes U(1)_Y$ is recovered from an hypothetical larger group.
 Just as the electroweak unification becomes evident at energies higher than the Higgs VEV,  a direct manifestation of Grand Unification Theories (GUTs) would occur at even higher energies.
 
-As the smallness of neutrino masses suggests to the existence of a higher mass scale \textcolor{red}{IS THIS TRUE?}, an other, even stronger, hint to Grand Unification comes from the slope of running of the coupling constants.
-The coupling constants for the electromagnetic, weak and strong interactions in the SM vary as a function of the interaction energy as shown in figure \ref{fig:constants}; they do not exactly meet, but their trend is interesting enough to push theorists to construct theories where perfect unification is achieved with the addition of new particles.
+As the smallness of neutrino masses suggests the existence of a higher mass scale, an other, even stronger, hint to Grand Unification comes from the slope of running of the coupling constants.
+The coupling constants for the electromagnetic, weak and strong interactions in the SM vary as a function of the interaction energy as shown in figure \ref{fig:constants};  they cannot exactly meet under the current experimental constraints, but their trend is interesting enough to push theorists to construct theories where perfect unification is achieved with the addition of new particles.
+Both experiments and theory indeed suggest the energy scale for convergence of the running coupling constants of the Standard Model to be over $10^{15}$ GeV.
+This energy scale seems impossible to access by any foreseeable accelerator experiment.%, leaving baryon number violation  to be the only testable process.
+
+%Models I want to mention: SUSY and SU(5) kaon modes and p -> epi0
+
+%The $SU(5)$ group is the smallest simple group containing $SU(3)_C  \otimes SU(2)_T \otimes U(1)_Y$. Georgi and Glashow proposed the ``minimal SU(5)"
 
@@
 The experimental observation of nucleon decay may be the only viable way to explore these theories and it is therefore a subject of great interest \cite{Adams:LBNE}.
-%Both experiments and theory indeed suggest the energy scale for convergence of the running coupling constants of the Standard Model to be over $10^{15}$ GeV.
-This energy scale seems impossible to access by any foreseeable accelerator experiment, leaving baryon number violation  to be the only testable process.
 

### 2018-04-11 Chapter-1/1-Theory.tex
s

@@
 As the smallness of neutrino masses suggests the existence of a higher mass scale, an other, even stronger, hint to Grand Unification comes from the slope of running of the coupling constants.
-The coupling constants for the electromagnetic, weak and strong interactions in the SM vary as a function of the interaction energy as shown in figure \ref{fig:constants};  they cannot exactly meet under the current experimental constraints, but their trend is interesting enough to push theorists to construct theories where perfect unification is achieved with the addition of new particles.
+The coupling constants for the electromagnetic, weak and strong interactions in the SM vary as a function of the interaction energy as shown in figure \ref{fig:constants} \textcolor{red}{ find reference to coupling constant picture};  they cannot exactly meet under the current experimental constraints, but their trend is interesting enough to push theorists to construct theories where perfect unification is achieved with the addition of new particles.
 Both experiments and theory indeed suggest the energy scale for convergence of the running coupling constants of the Standard Model to be over $10^{15}$ GeV.
@@
 \subsubsection{Signal Signatures}
+The strong force determines most of the interactions between an argon nucleus and a charged pion in the [100 -1200] MeV/c momentum range.
+The total hadronic $\pi^{-}$-Ar interaction cross section is defined as the one related to the single (hadronic) process driven only by the strong interaction.
+In measuring the ``total" cross section, we include both the elastic and  reaction channels, regardless of the final state,
+\begin{equation}
+\sigma_{Tot} = \sigma_{Elastic}+ \sigma_{Reaction};
+\end{equation}
+the reaction channel is further characterizes by several exclusive channels with defined topologies,
+\begin{equation}
+\sigma_{Reaction} = \sigma_{Inelastic} + \sigma_{abs} + \sigma_{chex}+ \sigma_{\pi prod}.
+\end{equation}
+
+A summary of the pion final states in order of pion multiplicity is given in table
+
+%\ref{tab:pionFinal}
+\Rotatebox{90}{%
+\begin{tabular}{llll}
+\hline
+N $\pi$ in final state & Channel Name & Channel description & Reaction \\
+\hline
+                       &              &                     &          \\
+                       &              &                     &          \\
+                       &              &                     &          \\
+                       &              &                     &          \\
+                       &              &                     &         \\
+\hline
+\end{tabular}
+}
+
+Pion capture on the argon orbital and pion decay at rest dominate the cross section under 100 MeV/c.
+Since pion capture can be considered an electromagnetic process and pion decay is a weak process, this energy region is purposely excluded from the hadronic cross section measurement.
+
 \subsubsection{Previous measurements: Lighter and Heavier Nuclei}

### 2018-04-11 Chapter-1/1-Theory.tex
s

@@
 This section outlines the importance of the pion-argon total hadronic cross section.
-We start by discussing the measurement in the context of neutrino interaction searches and of light mesons interaction with nuclei studies.
+We start by discussing the measurement in the context of neutrino interaction searches.
+% and of light mesons interaction with nuclei studies.
  We then describe the signal signature and historical measurements of pion-nucleus cross section, as well as the implementation of this cross sections in the current version of the simulation package used by LArIAT.
 
-\subsubsection{$\pi^{-}$Ar Cross section in the Context of Neutrino Searches}
+\subsubsection{$\pi^{-}$Ar Cross Section in the Context of Neutrino Searches}
 As outlined in \ref{ch:NuInt}, neutrino experiments use the products of neutrino interactions to identify the energy and flavor of the incoming neutrino.
@@
 
-\subsubsection{$\pi^{-}$Ar Cross Section in the Context of Light Mesons Interaction with Nuclei}
+%\subsubsection{$\pi^{-}$Ar Cross Section in the Context of Light Mesons Interaction with Nuclei}
 \subsubsection{Signal Signatures}
-The strong force determines most of the interactions between an argon nucleus and a charged pion in the [100 -1200] MeV/c momentum range.
+The strong force determines most of the interactions between an argon nucleus and a charged pion in the [100 -1200] MeV energy range.
 The total hadronic $\pi^{-}$-Ar interaction cross section is defined as the one related to the single (hadronic) process driven only by the strong interaction.
@@
 
-A summary of the pion final states in order of pion multiplicity is given in table
+A summary of the pion final states in order of pion multiplicity is given in table \ref{tab:pionFinal}.
+Pion capture on the argon orbital and pion decay at rest dominate the cross section under 100 MeV.
+Since pion capture can be considered an electromagnetic process and pion decay is a weak process, this energy region is purposely excluded from the hadronic cross section measurement.
 
-%\ref{tab:pionFinal}
-\Rotatebox{90}{%
-\begin{tabular}{llll}
-\hline
-N $\pi$ in final state & Channel Name & Channel description & Reaction \\
+\begin{sidewaystable}
+%\Rotatebox{90}{%
+\begin{tabular}{| c | l | m{7cm} | m{8cm} |}
 \hline
                     &              &                     &           \\
+N $\pi$ in FS & Channel Name & Reaction &  Notes \\
                     &              &                     &           \\
+\hline
                     &              &                     &           \\
+ 0  & \textbf{Pion Absorption}, $\sigma_{abs}$ &  \begin{tabular}{@{}l@{}} $\pi^- (np) \rightarrow nn$ (2-body abs) \\ $\pi^- (nnp) \rightarrow nnn$ (3-body abs)  \\ $\pi^- (npp) \rightarrow pnn$ (3-body abs)\\  $\pi^- (nnpp) \rightarrow pnn$ (Multi-body abs)  \end{tabular}  &Suppressed on single nucleon by energy conservation: the process occurs on at least two nucleons system.
+\\
                     &              &                     &           \\
+ \hline
+                     &              &                     &           \\
+ 1 &  \textbf{Elastic Scattering}, $\sigma_{el}$ & $\pi^- + N \rightarrow \pi^- + N$   & Scattering on nucleon or nucleus, the target is left in ground state         \\
+                    &              &                     &           \\
+ 1 &  \textbf{Charge Exchange}, $\sigma_{chex}$ & \begin{tabular}{@{}l@{}} $\pi^- + p \rightarrow \Delta^0 \rightarrow \pi^0+n$ \\ $\pi^- + N \rightarrow \pi^+ + $ nucleons  \end{tabular} &\begin{tabular}{@{}m{8cm}@{}}Single charge exchange: charged pion converts into neutral pion \\  \\Double charge exchange: charged pion converts into opposite charge pion  \end{tabular}   \\
+                    &              &                     &           \\
+  1 &  \textbf{Inelastic Scattering}, $\sigma_{inel}$ &    \begin{tabular}{@{}l@{}} $\pi^- + p \rightarrow \Delta^0 \rightarrow \pi^-+p$ (knock-out)  \\  $\pi^- + n \rightarrow \Delta^- \rightarrow \pi^-+n$ (knock-out)   \end{tabular}  &\begin{tabular}{@{}m{8cm}@{}} Other possible reactions: \\ Pure  Inelastic scattering:  population of low energy bound excited states \\ Nuclear break-up with nucleons or fragments knock-out    \end{tabular}   \\
+                    &              &                     &           \\
+  \hline
+                      &              &                     &           \\
+ 2+ & \textbf{Pion Production}, $\sigma_{\pi prod}$    &           $\pi^- + N \rightarrow \ge2 \pi $ + nucleons            &  Possible if pion K.E $\ge$ 500 Mev/c       \\
                     &              &                     &           \\
@@
 \end{tabular}
-}
-
-Pion capture on the argon orbital and pion decay at rest dominate the cross section under 100 MeV/c.
-Since pion capture can be considered an electromagnetic process and pion decay is a weak process, this energy region is purposely excluded from the hadronic cross section measurement.
+%}
+\caption{Summary of negative pion hadronic interactions as a function of the pion multiplicity in the final state in the energy range [100-1200] MeV. }
+\label{tab:pionFinal}
+\end{sidewaystable}
 
@@
 This section outlines the importance of the kaon-argon total hadronic cross section.
-We start by discussing the measurement in the context of nucleon decay searches and of light mesons interaction with nuclei studies.
+We start by discussing the measurement in the context of nucleon decay searches.
+% and of light mesons interaction with nuclei studies.
  We then describe the signal signature and historical measurements of kaon-nucleus cross section, as well as the implementation of this cross sections in the current version of the simulation package used by LArIAT.
@@
 
-\subsubsection{K$^{+}$Ar Cross section in the Context of Light Mesons Interaction with Nuclei}
-\label{sec:theoryStrangeMeson}
-The intrinsic value of the total hadronic K$^{+}$-Ar cross section measurement is that kaon interactions complement the measurements of $\pi$ interactions as a probe of  hadron interaction inside the nucleus in the strange sector.
-\textcolor{red}{High theoretical interest in probing constituent quark model of nuclear structure with
-KAON-NUCLEON INTERACTIONS CHIEDI REFERENZE A FLAVIO}
+%\subsubsection{K$^{+}$Ar Cross section in the Context of Light Mesons Interaction with Nuclei}
+%\label{sec:theoryStrangeMeson}
+%The intrinsic value of the total hadronic K$^{+}$-Ar cross section measurement is that kaon interactions complement the measurements of $\pi$ interactions as a probe of  hadron interaction inside the nucleus in the strange sector.
+%\textcolor{red}{High theoretical interest in probing constituent quark model of nuclear structure with
+%KAON-NUCLEON INTERACTIONS CHIEDI REFERENZE A FLAVIO}
 
@@
 $$\sigma_{Tot} = \sigma_{Elastic}+ \sigma_{Reaction}.$$
-
-%For this analysis, kaons are selected from the LArIAT beamline in the momentum range between \textcolor{red}{500} MeV/c and \textcolor{red}{1000} MeV/c (see Fig \ref{fig:TOFK}).
-
-%\begin{figure}[hpbt]
-%\centering
-%\includegraphics[width=5in]{Chapter-1/Images/KaonTOF}
-%\caption{Time of flight versus momentum distributions as produced by the LArIAT TOF and Wire Chambers systems.
-The Kaon population lies between the proton and the muon/pion populations, allowing PID of Kaons in the beam line.
-}
-%\label{fig:TOFK}
-%\end{figure}
 

### 2018-04-11 Chapter-1/1-Theory.tex
s

@@
 \item[-] Determination the status of CP symmetry in the lepton sector.
-\item[-] \textcolor{red}{something about GUTs}.
+\item[-] Search for observables predicted by GUTs, such as proton decay.
 \end{itemize}
@@
 \subsubsection{Signal Signatures}
-The strong force determines most of the interactions between an argon nucleus and a charged pion in the [100 -1200] MeV energy range.
-The total hadronic $\pi^{-}$-Ar interaction cross section is defined as the one related to the single (hadronic) process driven only by the strong interaction.
+Strong hadronic interaction models \cite{9780198520085,9780471779957} predict the pion interaction processes with argon in the [100 -1200] MeV energy range. 
+The total hadronic $\pi^{-}$-Ar interaction cross section is defined as the one related to the single process driven only by the strong interaction which is dominant in the considered energy range. 
 In measuring the ``total" cross section, we include both the elastic and  reaction channels, regardless of the final state, 
@@
 Pion capture on the argon orbital and pion decay at rest dominate the cross section under 100 MeV.
+The pion decay life time is $\tau_\pi = 2.6 \times 10^-8$ s and the main decay mode is  $\pi^-\rightarrow \mu^- + \bar\nu_\mu$ (BR 99.98\%); for pion decay at rest, the muon is emitted with an energy of $\sim5$ MeV and it decays into $\mu^-\rightarrow e^- + \bar\nu_e + \nu_\mu$ in $\tau_\mu = 2.3 $ $\mu$s .
 Since pion capture can be considered an electromagnetic process and pion decay is a weak process, this energy region is purposely excluded from the hadronic cross section measurement.
@@
 \subsubsection{Previous measurements: Lighter and Heavier Nuclei}
+Many experiments with pion beams have studied the hadronic interaction of pions on light and heavy materials, such as  He, Li, C, Fe, Pb\textcolor{red}{\cite{s}}.
+However, data on Argon are rare: the total differential hadronic cross section has never been measured before on argon.
+Simulation packages like Geant base their pion transportation for argon on data from lighter and heavier nuclei: the goal of LArIAT's dedicated measurement on argon is to bridge this gap in data, thus reducing the uncertainties related to pions interaction in argon in both neutrino event generators and in simulation packages of pion transportation.
+
 \subsubsection{Pion Interaction Cross Section for thin target in Geant4}

### 2018-04-11 Chapter-1/1-Theory.tex
s

@@
 A summary of the pion final states in order of pion multiplicity is given in table \ref{tab:pionFinal}.
-Pion capture on the argon orbital and pion decay at rest dominate the cross section under 100 MeV.
-The pion decay life time is $\tau_\pi = 2.6 \times 10^-8$ s and the main decay mode is  $\pi^-\rightarrow \mu^- + \bar\nu_\mu$ (BR 99.98\%); for pion decay at rest, the muon is emitted with an energy of $\sim5$ MeV and it decays into $\mu^-\rightarrow e^- + \bar\nu_e + \nu_\mu$ in $\tau_\mu = 2.3 $ $\mu$s .
+Pion capture and pion decay at rest dominate the cross section under 100 MeV.
+We define pion capture as the process determining the formation of a pionic atom and the subsequent pion's end of life.
+Stopping negative pions can form pionic argon, where the negative pion plays the role of an orbital electron.
+Since the pion mass is two orders of magnitude greater than the electron mass, the spatial wave function of the pion will overlap way more with the nucleus compared to the electron case.
+After the electromagnetic formation of the pionic atom, the pion will quickly absorbed by the nucleus, which is put in an excited state.
+The nucleus then de-excite with the emission of low energy nucleons and photons.
+Pion capture is predominant compared to pion decay, the other important process for very low energy pions.
+The decay of a pion is governed by the weak force; the pion decay life time is $\tau_\pi = 2.6 \times 10^{-8}$ s and the main decay mode is  $\pi^-\rightarrow \mu^- + \bar\nu_\mu$ (BR 99.98\%).
+%; for pion decay at rest, the muon is emitted with an energy of $\sim5$ MeV and it decays into $\mu^-\rightarrow e^- + \bar\nu_e + \nu_\mu$ in $\tau_\mu = 2.3 $ $\mu$s.
  Since pion capture can be considered an electromagnetic process and pion decay is a weak process, this energy region is purposely excluded from the hadronic cross section measurement.

### 2018-04-11 Chapter-1/1-Theory.tex
s

@@
 \centering
-\begin{tabular}{|l|l|}
+\begin{tabular}{|l|l|l|}
 \hline
-$\Delta$ (1232) 3/2$^+$                 & $N^*$ (1440) 1/2$^+$    \\ \hline
-$\Delta^{++} \text{(uuu)} \rightarrow p \pi^+$ & $N^* \rightarrow N \pi$ \\
-$\Delta^{+} \text{(uud)} \rightarrow n \pi^+$  &  $N^* \rightarrow N \pi\pi$\\ 
-$\Delta^{+} \text{(uud)} \rightarrow p \pi^0$  &                         \\ 
-$\Delta^{0} \text{(udd)} \rightarrow n \pi^0$  &                         \\ 
-$\Delta^{0} \text{(udd)} \rightarrow p \pi^-$  &                         \\ 
-$\Delta^{-} \text{(ddd)} \rightarrow n \pi^-$  &                         \\ \hline
+Resonance  & Decay Mode & Lifetime (s)\\
+\hline
+    &  &  \\
+$\Delta$ (1232) 3/2$^+$  & \begin{tabular}{@{}l@{}} $\Delta^{++} \text{(uuu)} \rightarrow p \pi^+$ \\$\Delta^{+} \text{(uud)} \rightarrow n \pi^+$  \\ $\Delta^{+} \text{(uud)} \rightarrow p \pi^0$\\$\Delta^{0} \text{(udd)} \rightarrow n \pi^0$ \\ $\Delta^{0} \text{(udd)} \rightarrow p \pi^-$ \\$\Delta^{-} \text{(ddd)} \rightarrow n \pi^-$  \end{tabular}
+ & $\sim5.6 \times10^{-24}$\\
+     &  &  \\
+\hline
+    &  &  \\
+$N^*$ (1440) 1/2$^+$     & \begin{tabular}{@{}l@{}}$N^* \rightarrow N \pi$ \\$N^* \rightarrow N \pi\pi$\end{tabular} &   $\sim2.2 \times10^{-24}$\\
+    &  &  \\
+\hline
 \end{tabular}
@@
 \end{table}
+
 The key elements of a neutrino event generators for resonance and DIS events are the nuclear model and the hadron treatment (both production and transportation).
@@
 Many experiments with pion beams have studied the hadronic interaction of pions on light and heavy materials, such as  He, Li, C, Fe, Pb\textcolor{red}{\cite{s}}.
-However, data on Argon are rare: the total differential hadronic cross section has never been measured before on argon.
+However, data on argon are rare: the total differential hadronic cross section has never been measured before on argon.
 Simulation packages like Geant base their pion transportation for argon on data from lighter and heavier nuclei: the goal of LArIAT's dedicated measurement on argon is to bridge this gap in data, thus reducing the uncertainties related to pions interaction in argon in both neutrino event generators and in simulation packages of pion transportation.
 
+The shape of the pion-nucleus interaction cross section in the energy range considered shows the distinct features that indicate the presence of a resonance.
+The mean free path of a pion of kinetic energy between 100 and 400 MeV is much shorter than the average distance between nucleons (which is of the order of 1 fm).
+Therefore, the pion interacts with surface nucleons.
+A $\Delta$ resonance is often produced in the interaction, which subsequently decays inside the nucleus.
+
+Experimental results on several nuclei are reported in Figure as from \cite{}.
+
 \subsubsection{Pion Interaction Cross Section for thin target in Geant4}
+
+\clearpage
 

### 2018-04-11 Chapter-1/1-Theory.tex
s

@@
 
-Experimental results on several nuclei are reported in Figure as from \cite{}.
+Experimental results on several nuclei are reported in Figure as from \cite{PhysRevC.14.635}.
+
+\begin{figure}[hbpt]
+\centering
+\includegraphics[]{Chapter-1/Images/PionsOnNuclei.png}
+\caption{Pion-nucleus total cross sections:  $\sigma_{\pi^+}$ for positive pions (right) and $\sigma_{AV}$ (left) for the average between positive and negative pions $\sigma_{AV} = \frac{\sigma_{\pi^+} + \sigma_{\pi^-}}{2}$  in  the $\Delta$ resonance region.
+The error bars include estimates of systematic uncertainties.
+The curves are the results of fits to the data assuming a Breit-Wigner shape.
+This summary plot  is reported in \cite{PhysRevC.14.635} and uses data from \cite{Clough1974, Wilkin:1973xd}.
+}
+\end{figure}
 

### 2018-04-11 Chapter-1/1-Theory.tex
s

@@
 
-\begin{sidewaystable}
+\begin{sidewaystable}[hp]
 %\Rotatebox{90}{%
@@
 However, data on argon are rare: the total differential hadronic cross section has never been measured before on argon.
-Simulation packages like Geant base their pion transportation for argon on data from lighter and heavier nuclei: the goal of LArIAT's dedicated measurement on argon is to bridge this gap in data, thus reducing the uncertainties related to pions interaction in argon in both neutrino event generators and in simulation packages of pion transportation.
+Simulation packages like Geant4 base their pion transportation for argon on data from lighter and heavier nuclei: the goal of LArIAT's dedicated measurement on argon is to bridge this gap in data, thus reducing the uncertainties related to pions interaction in argon in both neutrino event generators and in simulation packages of pion transportation.
 
 The shape of the pion-nucleus interaction cross section in the energy range considered shows the distinct features that indicate the presence of a resonance.
-The mean free path of a pion of kinetic energy between 100 and 400 MeV is much shorter than the average distance between nucleons (which is of the order of 1 fm).
+In fact, the mean free path of a pion of kinetic energy between 100 and 400 MeV is much shorter than the average distance between nucleons (which is of the order of 1 fm).
 Therefore, the pion interacts with surface nucleons.
 A $\Delta$ resonance is often produced in the interaction, which subsequently decays inside the nucleus.
-
-Experimental results on several nuclei are reported in Figure as from \cite{PhysRevC.14.635}.
+Experimental results on several nuclei as reported in \cite{PhysRevC.14.635} are shown in Figure \ref{fig:pi1}; it is interesting to notice here how the shape fo the $\Delta$ resonance become less pronounced as a function of the mass number of the target nucleus.
+Pion interactions with heavier nuclei also shift the peak of the resonance at lower energy; this effect is due to kinematic considerations and to the difference in propagation of the $\Delta$ inside the nucleus.
+Multiple scattering effect modify the resonance width, which is larger than the natural-decay width.
+As an example of a fairly well studied target, Figure \ref{fig:pi2} reports the negative pion cross section on Carbon for the elastic and reaction\footnote{What we refer as to ``reaction channel" is called ``inelastic" in the paper.} channels, and their sum \cite{BINON1970168}.
 
@@
 \centering
-\includegraphics[]{Chapter-1/Images/PionsOnNuclei.png}
+\includegraphics[scale=0.7]{Chapter-1/Images/PionsOnNuclei.png}
 \caption{Pion-nucleus total cross sections:  $\sigma_{\pi^+}$ for positive pions (right) and $\sigma_{AV}$ (left) for the average between positive and negative pions $\sigma_{AV} = \frac{\sigma_{\pi^+} + \sigma_{\pi^-}}{2}$  in  the $\Delta$ resonance region.
@@
 }
+\label{fig:pi1}
 \end{figure}
 
-\subsubsection{Pion Interaction Cross Section for thin target in Geant4}
+\begin{figure}[hbpt]
+\centering
+\includegraphics[scale=0.7]{Chapter-1/Images/PionCarbon.png}
+\caption{Negative pion nucleus total, elastic and reaction cross sections on $^{12}$C as from \cite{BINON1970168}.  }
+\label{fig:pi2}
+\end{figure}
+
+\subsubsection{Pion Interaction Cross Section in generators}
+A broad review of the current status of negative and positive pion simulation in Geant4 and GENIE for $^{12}$C, $^{56}$Fe, and $^{40}$Ca can be found in \cite{Irene} \textcolor{red}{Irene}.
+From that work, we report the results for $^{12}$C (left) in Figure \ref{fig:pionC} and  $^{40}$Ca (right) in Figure \ref{fig:pionCa}, since the argon cross section should lie in between.
+
+\begin{figure}
+\captionsetup{justification=raggedright}  
+	\begin{minipage}[t]{.53\textwidth}  
+	  \centering  
+	   \includegraphics[width=3in]{Chapter-1/Images/IreneCarbon.png}
+	   	        \caption{ Total, elastic and reaction cross section for $\pi^-$ on $^{12}$C.
+Comparison between results from Geant4 simulation (Bertini cascade model) and experimental data \cite{Clough1974, PhysRevC.23.2173, ALLARDYCE19731,osti_4488432}.}
+        \label{fig:pionC}
+	\end{minipage}%  
+	\begin{minipage}[t]{0.53\textwidth}  
+	  \centering  
+	\includegraphics[width=3in]{Chapter-1/Images/IreneCalcium.png}
+        \caption{Total, elastic and reaction cross section for $\pi^-$ on $^{40}$Ca.
+Comparison between results from Geant4 simulation (Bertini cascade model),  Genie simulation (hA model), and experimental data \cite{ALLARDYCE19731}.}
+        \label{fig:pionCa}
+	\end{minipage}
+	\par
+\end{figure}
 
@@
 
-\begin{figure}
+\begin{figure}[b!]
 \captionsetup{justification=raggedright}  

### 2018-04-11 Chapter-1/1-Theory.tex
s

@@
 
+\textcolor{red}{explain geant4 }
+\textcolor{red}{explain Genie }
+Figure \ref{fig:PionTrueArgon} shows the Geant4 prediction for the total, elastic and reaction cross section for $\pi^-$ on argon.
+
 \begin{figure}
@@
 	\par
+\end{figure}
+
+\begin{figure}[hbpt]
+\centering
+\includegraphics[width=3in]{Chapter-1/Images/cPionsG4.png}
+\caption{Total, elastic and reaction hadronic cross section for $\pi^-$-argon implemented in Geant4  10.01.p3.}
+\label{fig:PionTrueArgon}
 \end{figure}
@@
 On the other, it further highlights the importance of the kaon measurement.
+Figure \ref{fig:TrueArgon} shows the Geant4 prediction for the total, elastic and reaction cross section for $K^+$ on argon.
 
-\begin{figure}[b!]
+\begin{figure}
 \captionsetup{justification=raggedright}  
+	\begin{minipage}[t]{.48\textwidth}  
 	  \centering  
 \includegraphics[width=3in]{Chapter-1/Images/CarbonG4.png}
-\caption{total hadronic cross section for carbon implemented in Geant4  10.01.p3  with overlaid with the Bugg and Frideman data.}
+\caption{Total hadronic cross section for carbon implemented in Geant4  10.01.p3  with overlaid with the Bugg and Frideman data.}
 \label{fig:TrueCarbon}
+	\end{minipage}%  
+	\begin{minipage}[t]{0.48\textwidth}  
+	  \centering  
+\includegraphics[width=3in]{Chapter-1/Images/cKaonsG4.png}
+\caption{Total, elastic and reaction hadronic cross section for $K^+$-argon implemented in Geant4  10.01.p3.}
+\label{fig:TrueArgon}
+	\end{minipage}
+	\par
 \end{figure}

### 2018-04-11 Chapter-1/1-Theory.tex
s

@@
 The GENIE module INTRANUKE \cite{Juszczak2006} is used to simulate the final-state interactions (FSI) which is the hadron re-interaction inside the nucleus.
-This module places the outgoing particles in the nucleus and propagates them.
-In the INRANUKE model, hadrons cannot undergo more than one FSI per event.
+This module places the outgoing particles in the nucleus and propagates them using the ``hA model".
+In the INRANUKE hA model, hadrons can undergo at most one FSI per event.
 When possible, external hadron-nucleus scattering data are used to tune INTRANUKE.
@@
 
-\textcolor{red}{explain geant4 }
-\textcolor{red}{explain Genie }
+LArIAT uses Geant4 as the default simulation package.
+In particular, pions (and kaons) transportation is achieved through the Geant4 FTFP\_BERT physics list.
+In this physics list, Geant4 uses the Bertini cascade model  \cite{Wright:2015xia}  to simulate the products of the pion-nucleus interaction as well as the hadrons re-interactions inside the target nucleus (intra-nuclear cascade).
+REWRITE THE FOLLOWING
+%The target nucleus is treated as an average nuclear medium, in which the nucleons are assumed to have a Fermi gas momentum distribution.
+The final state of each collision is sampled according to free-particle cross section data.
+Clusters of nucleons are considered only in case of pion absorption which requires at least dinucleons.
+Nuclear effects, i.e.
+secondaries which can interact with other protons and neutrons inside the nucleus and so on until the last secondaries es-cape the nucleus or are absorbed, are also taken in account.
+This model reproduces detailed cross section data for nucleons, pions and kaons in the region below 1 GeV and is expected to do reasonably well in the multi-GeV region.
+
+%Genie
+The relevance of the GENIE generator for neutrino physics and its basic working principles have been outlined earlier in this section.
+Information on hadron-nucleus interactions can be extracted from the  INTRANUKE module and compared against the Geant4 predictions, as shown in Figure    \ref{fig:pionCa}.
+%Calcium explained
+%Carbon explained
+
 Figure \ref{fig:PionTrueArgon} shows the Geant4 prediction for the total, elastic and reaction cross section for $\pi^-$ on argon.

### 2018-04-12 Chapter-1/1-Theory.tex
s

--- 2018-04-11 Chapter-1/1-Theory.tex
@@
 \centering
-\includegraphics[scale=0.7]{Chapter-1/Images/PionsOnNuclei.png}
+\includegraphics[scale=0.9]{Chapter-1/Images/PionsOnNuclei.png}
 \caption{Pion-nucleus total cross sections:  $\sigma_{\pi^+}$ for positive pions (right) and $\sigma_{AV}$ (left) for the average between positive and negative pions $\sigma_{AV} = \frac{\sigma_{\pi^+} + \sigma_{\pi^-}}{2}$  in  the $\Delta$ resonance region.
@@
 \centering
-\includegraphics[scale=0.7]{Chapter-1/Images/PionCarbon.png}
+\includegraphics[scale=0.9]{Chapter-1/Images/PionCarbon.png}
 \caption{Negative pion nucleus total, elastic and reaction cross sections on $^{12}$C as from \cite{BINON1970168}.  }
@@
 \subsubsection{Pion Interaction Cross Section in generators}
+LArIAT uses Geant4 as the default simulation package.
+In particular, pions (and kaons) transportation is achieved through the Geant4 FTFP\_BERT physics list.
+In this physics list, Geant4 uses the Bertini cascade model  \cite{Wright:2015xia}  to simulate the products of the pion-nucleus interaction as well as the secondary hadrons re-interactions inside the target nucleus (intra-nuclear cascade).
+The target nucleus is represented as a continuous gas where the nuclear potential follows concentrical shells whose depths approximate the Woods-Saxon shape.
+The CERN-HERA compilations \cite{Compilation1,Compilation2} of hadron-nucleon interaction data is the data base used for the decision making process after the cascade is invoked.
+The cross section model determines if the pion interacts, the eventual type of interaction and the interaction multiplicity.
+For  hadron projectiles with energy less than 20 GeV, Geant4 reports  the uncertainty on the cross section model to be about the size of the error bars on the data used, or about 10\%, increasing to 20-30\% in energy regions where data is sparse.
+
+The relevance of the GENIE generator for neutrino physics and its basic working principles have been outlined earlier in this section.
+Given GENIE's modularity, information on hadron-nucleus interactions can be extracted from the  INTRANUKE module and directly compared against the Geant4 predictions.%, as shown in Figure    \ref{fig:pionCa}.
+
 A broad review of the current status of negative and positive pion simulation in Geant4 and GENIE for $^{12}$C, $^{56}$Fe, and $^{40}$Ca can be found in \cite{Irene} \textcolor{red}{Irene}.
-From that work, we report the results for $^{12}$C (left) in Figure \ref{fig:pionC} and  $^{40}$Ca (right) in Figure \ref{fig:pionCa}, since the argon cross section should lie in between.
+From that work, we report the results for $^{12}$C in Figure \ref{fig:pionC} as it allows a direct comparison between Geant4, GENIE and data.
+Geant4 predictions for $\pi^-$  on Carbon are in good agreement with data over all the spectrum, while GENIE predictions seem to show some features at around 500 MeV and 900 MeV maybe due ti higher resonances in the hA model.
+From the same work, we also report the negative pion cross section on $^{40}$Ca in Figure \ref{fig:pionCa}, since the this is the nuclear medium closest argon with some available data.
+The predictions from both Geant4 and GENIE agree with data in the high energy region.
+The predictions diverge in the resonance region, where data is not available.
 
-\textcolor{red}{explain geant4 }
-\textcolor{red}{explain Genie }
-Figure \ref{fig:PionTrueArgon} shows the Geant4 prediction for the total, elastic and reaction cross section for $\pi^-$ on argon.
+For the $\pi^-$ argon total cross section measurement, we use the Geant4 Bertini Cascade model, whose  predictions for the total, elastic and reaction cross sections are show in Figure \ref{fig:PionTrueArgon}.
 
@@
 	   	        \caption{ Total, elastic and reaction cross section for $\pi^-$ on $^{12}$C.
-Comparison between results from Geant4 simulation (Bertini cascade model) and experimental data \cite{Clough1974, PhysRevC.23.2173, ALLARDYCE19731,osti_4488432}.}
+Comparison between results from Geant4 simulation (Bertini cascade model), Genie simulation (hA model), and experimental data \cite{Clough1974, PhysRevC.23.2173, ALLARDYCE19731,osti_4488432}.}
         \label{fig:pionC}

### 2018-04-12 Chapter-1/1-Theory.tex
s

@@
 The Standard Model (SM) of particle physics is the most accurate theoretical description of the subatomic world and, in general, one of the most precisely tested theories in the history of physics.
-The SM describes the strong, electromagnetic and weak interactions among  elementary particles in the framework of quantum field theory, accounting for the unification of electromagnetic and weak interactions for energies above the  vacuum expectation value of the Higgs field.
+The SM describes the strong, electromagnetic and weak interactions among  elementary particles in the framework of quantum field theory, accounting for the unification of electromagnetic and weak interactions for energies above the  vacuum expectation value (VEV) of the Higgs field.
 The SM does not describe gravity or general relativity.
@@
 The introduction of massive fields in the Standard Model lagrangian would make the theory non-renormalizable, and - so far - mathematically impossible to handle.
-This problem is solved in the Standard Model by the introduction of a scalar iso-doublet $\Phi(x)$, the Higgs field, which gives mass to W$^\pm$ and Z$^0$ gauge bosons through the electroweak symmetry breaking and to the fermions through Yukawa coupling \cite{Higgs1964,Higgs19642}.
-The discovery of the Higgs boson in 2012 by the LHC experiments \cite{201230,Aad2012} marked the ultimate confirmation  of a long history of Standard Model successful predictions.
+This problem is solved in the SM by the introduction of a scalar iso-doublet $\Phi(x)$, the Higgs field, which gives mass to W$^\pm$ and Z$^0$ gauge bosons through the electroweak symmetry breaking mechanism and to the fermions through Yukawa coupling \cite{Higgs1964,Higgs19642}.
+The discovery of the Higgs boson in 2012 by the LHC experiments \cite{201230,Aad2012} marked the ultimate confirmation  of a long history of successful predictions by the SM.
 
@@
 The constraint of parity conservation weakened after Wu's experiment in 1957 \cite{PhysRev.105.1413}.
-Additionally,  there was no experimental indication for massive neutrinos nor evidence of interaction via the neutrino right-handed component.% neutrinos likely interacted only via the left-handed component.
+Additionally,  there was no experimental indication for massive neutrinos, nor evidence of interaction via the neutrino right-handed component.% neutrinos likely interacted only via the left-handed component.
 
@@
 The SM electroweak lagrangian is the most general renormalizable lagrangian invariant under the local symmetry group $SU(2)_T \otimes U(1)_Y$.
-The lagrangian couples the weak isotopic spin doublets and singlets described in \ref{tab:SMParticles} with the gauge bosons  $A^{\mu}_{a}$ ($a$ $=$ 1,2,3) and $B^{\mu}$, and Higgs doublet $\Phi(x)$:
+The lagrangian couples the weak isotopic spin doublets and singlets described in Table \ref{tab:SMParticles} with the gauge bosons  $A^{\mu}_{a}$ ($a$ $=$ 1,2,3) and $B^{\mu}$, and Higgs doublet $\Phi(x)$:
 
@@
 
-The first two lines of the lagrangian summarize the kinetic terms for the fermionic fields and their coupling to the gauge bosons $A^{\mu\nu}_a$, $B^{\mu\nu}$ \footnote{In gauge theories the ordinary derivative $\partial_\mu$  is substitued with Âthe covariant derivative $D_\mu$.
+The first two lines of the lagrangian summarize the kinetic terms for the fermionic fields and their coupling to the gauge bosons $A^{\mu\nu}_a$, $B^{\mu\nu}$ \footnote{In gauge theories the ordinary derivative $\partial_\mu$  is substitued with the covariant derivative $D_\mu$.
 Here $D_\mu = \partial_\mu + igA_\mu \cdot I + ig'B_\mu\frac{Y}{2}$, where I and Y are the SU(2)$_L$ and U(1)$_Y$ generators, respectively.}.
@@
 The forth line is the Higgs lagrangian, which results in the spontaneous symmetry breaking.
-The last three lines describe the Yukawa coupling between fermions and the Higgs field, origin of the fermion's mass.
+The last three lines describe the Yukawa coupling between fermions and the Higgs field, origin of the fermions' mass.
 
@@
 Neutrino oscillations are possible only if the neutrino flavor eigenstate are not identical to the mass eigenstates, thus resulting in the first evidence of physics beyond the Standard Model.
-A minimal extension of the Standard Model introduces three mass eigenstates, $\ket{\nu_i}$ ($i = 1,2, 3$), whose mass $m_i$ is well defined.
-The unitary Pontecorvo-Maki-Nakagawa-Sakata matrix transforms the spinor wave functions ($\psi$) of each component  between flavor and mass bases as follows
+A minimal extension of the SM introduces three mass eigenstates, $\ket{\nu_i}$ ($i = 1,2, 3$), whose mass $m_i$ is well defined.
+The unitary Pontecorvo-Maki-Nakagawa-Sakata matrix transforms the spinor wave functions ($\psi$) of each component  between the flavor and mass bases as follows
 
 \begin{equation} 
-\sum \psi_\alpha \ket{\nu_\alpha} =  \sum \psi_i \ket{\nu_i}, \rightarrow \psi_\alpha =  U_{PMNS} \psi_i, 
+\sum_\alpha \psi_\alpha \ket{\nu_\alpha} =  \sum_i \psi_i \ket{\nu_i}, \rightarrow \psi_\alpha =  U_{PMNS} \psi_i, 
 \end{equation}
@@
 All current and forthcoming oscillation experiments live in the 0.1-10 GeV transition region, which encompasses the energy where the QE neutrino-nucleus interaction transitions into resonant scattering and the energy  where resonance scattering  transitions into DIS.
- Neutrino and antineutrino QE charge current scattering refers to the process $\nu_l n \rightarrow l^- p$ and $\bar\nu_l p \rightarrow l^+ n$ where a charged lepton and single nucleon are ejected in the elastic interaction and the target nucleus remains at ground state.
+ Schematically, neutrino and antineutrino QE charge current scattering refers to the process $\nu_l n \rightarrow l^- p$ and $\bar\nu_l p \rightarrow l^+ n$ where a charged lepton and single nucleon are ejected in the elastic interaction, leaving the target nucleus in its ground state.
 %of a neutrino (or antineutrino) with a nucleon in the target material.
@@
 This simple picture of neutrino interactions works rather well for scattering
-off of light targets, such as the H$_2$ and D$_2$ of bubble chamber experiments \cite{RevModPhys.84.1307}, but the complexity of the nuclear structure for heavier nuclei such as argon complicates this model.
+off of light nuclear targets, such as the H$_2$ and D$_2$ of bubble chamber experiments \cite{RevModPhys.84.1307}, but the complexity of the nuclear structure for heavier nuclei such as argon complicates this model.
 
@@
 
-Nuclear effects can potentially affect the neutrino event rates, nucleon emission, neutrino energy reconstruction, and neutrino/antineutrino ratios, carrying deep implications for oscillation experiments.
+Nuclear effects can potentially affect the neutrino event rates, nucleon emission, neutrino energy reconstruction, and the neutrino/antineutrino ratios, carrying deep implications for oscillation experiments.
 Even in the case of ``simple" QE scattering, intra-nuclear hadron rescattering and correlation effects between the target nucleons can cause the ejection of additional nucleons in the final state, modifying the final state kinematics and topology.
@@
 
-A renewed interest for neutrino cross section measurements surged in recent year, along with lively discussion on the data reporting; the historical method of reporting the neutrino cross section as a function of the neutrino energy or momentum transferred shakes under the weight of its dependency on the chosen nuclear model.
-On one hand, correcting for nuclear effects in neutrino interaction can introduce unwanted sources of uncertainty and model dependence especially due to the mis-modeling of the meson interactions.
+A renewed interest for neutrino cross section measurements surged in recent years, along with a lively discussion on the data reporting; the historical method of reporting the neutrino cross section as a function of the neutrino energy or momentum transferred shakes under the weight of its dependency on the chosen nuclear model.
+On one hand, correcting for nuclear effects in neutrino interaction can introduce unwanted sources of uncertainty and model dependency especially due to the mis-modeling of the meson interactions.
 On the other, avoiding this correction makes a comparison between neutrino interactions on different target nuclei extremely difficult.
 
-Data on neutrino scattering off many different nuclei are available for both charged current (CC) and neutral current (NC) channels, as summarized here \cite{RevModPhys.84.1307}.
+Data on neutrino scattering off many different nuclei are available for both charged current (CC) and neutral current (NC) channels, as summarized in \cite{RevModPhys.84.1307}.
 A summary of the results on QE, resonant and DIS scattering for neutrinos and antineutrinos from accelerators on different target is reported in Figure \ref{fig:nuXS}, where the (NUANCE) \cite{Casper2002} event generator is used as comparison with the theory.
@@
 \textbf{Sterile neutrinos.} Hints to the existence of at least one additional neutrino, in the form of various anomalies, have been puzzling physicists almost from the beginning of neutrino oscillation searches.
-Originally designed to look for evidence of neutrino oscillation, the Liquid Scintillator Neutrino Detector (LSND) \cite{Athanassopoulos1997} provided a first conflicting result with the Standard Model expectation of only three neutrino flavors.
+Originally designed to look for evidence of neutrino oscillation, the Liquid Scintillator Neutrino Detector (LSND) \cite{Athanassopoulos1997} provided a first conflicting result with the Standard Model expectation of only three neutrinos.
 A second conflicting result has also been provided by the MiniBooNE experiment \cite{AGUILARAREVALO200928}.
@@
 The new sterile neutrino would mainly be composed of a heavy neutrino $\nu_4$ with mass $m_4$ such that  $m_1, m_2, m_3 \ll m_4$ and  $\Delta m^2= \Delta m^2_{14} \sim [0.1 - 10]$ eV$^2$.
-The introduction of sterile neutrinos is an appealing line of thinking, since this renormalizable generalization of the Standard Model has the potential to impact long standing questions in high energy physics and cosmology: light sterile neutrinos are candidates for dark matter particles and there are ideas that the theory could be adjusted to explain the baryon asymmetry of the Universe via leptogenesis \cite{1063-7869-57-5-503}.
+The introduction of sterile neutrinos is an appealing line of thinking, since this renormalizable generalization of the SM has the potential to impact long standing questions in high energy physics and cosmology: light sterile neutrinos are candidates for dark matter particles and there are ideas that the theory could be adjusted to explain the baryon asymmetry of the Universe via leptogenesis \cite{1063-7869-57-5-503}.
 
@@
 The asymmetry in the oscillation probability of neutrinos and antineutrinos is the observable sensitive to the Dirac CP-violating phase $\delta$ leveraged in neutrino oscillation experiments.
-Using the parameterization of the PMNS matrix shown in Equation \ref{eq:PMNS},  the difference in the probability of $\nu_e \rightarrow \nu_\mu$ oscillation and the probability of $\bar\nu_e \rightarrow \bar\nu_\mu$ oscillation can be parametrized as follows \cite{Cervera2000},
+Using the parameterization of the PMNS matrix shown in Equation \ref{eq:PMNS},  the difference between the probability of $\nu_e \rightarrow \nu_\mu$ oscillation and the probability of $\bar\nu_e \rightarrow \bar\nu_\mu$ oscillation can be parametrized as follows \cite{Cervera2000},
 \begin{equation}
@@
 \textbf{Majorana or Dirac?}
-Evidence of neutrino oscillations demand the introduction in the theory of a mechanism which can give mass to the neutrinos.
+Evidence of neutrino oscillations demands the introduction of a mechanism which can give mass to the neutrinos.
 This mechanism should possibly also explain why neutrino masses are at least six orders of magnitude lower than the electron mass (the second lightest SM fermion).
@@
 In 1937, Majorana demonstrated that the introduction of a two components spinor is sufficient to describe a massive fermion \cite{Majorana1937}.
-The Dirac equations of motion for the chiral fields (equations \ref{eq:15} and \ref{eq:16}) hold true in the case of two components spinor under the assumption that the chiral components $\psi_R$ and $\psi_L$ are correlated through the charge conjugation matrix $\mathcal{C}$, $\psi_R = \mathcal{C}\bar\psi_L$, thus the theory is applicable only to neutral fermions.
+The Dirac equations of motion for the chiral fields (equations \ref{eq:15} and \ref{eq:16}) hold true in the case of two components spinor under the assumption that the chiral components $\psi_R$ and $\psi_L$ are correlated through the charge conjugation matrix $\mathcal{C}$, $\psi_R = \mathcal{C}\bar\psi_L$.
+Therefore the theory is applicable only to neutral fermions.
 Neutrinos are the only neutral elementary particles in the SM -- the only possible Majorana particle candidate.
@@
 The  $\mathcal{L}_5$ term would introduce a non-renormalizable term in the lagrangian, since it has dimensions of energy to the fifth power.
-This is not allowed in the SM term; however, the existence of such terms is plausible if the we consider the SM an effective theory at low energy, manifestation of the symmetry breaking of a Grand Unified Theory (GUT) at higher energy, and not the definitive theory.
+This is not allowed in the SM lagrangian; however, the existence of such terms is plausible if the we consider the SM as an effective theory at low energy, manifestation of the symmetry breaking of a Grand Unified Theory (GUT) at higher energy, and not the definitive theory.
 The mass term in eq \ref{eq:majoMass} implies the neutrino mass to be  $m = \frac{g v^2}{\mathcal{M}}$.
 The coupling coefficient can be of the order of any other fermion's coupling coefficient, since the smallness of neutrino masses is achieved by the big value of the new physics mass scale alone.
-This vanilla formulation is the conceptual basis for many flavors of GUT-based \emph{seesaw mechanism} ~\cite{Yanagida1980}, which we will not discuss here in any detail.
+This vanilla formulation is the conceptual basis for many flavors of \emph{seesaw mechanism} ~\cite{Yanagida1980}, which we will not discuss here in any detail.
 However, it is fascinating how the puzzle of the neutrino mass hints to the existence of a deeper and more complete theory.\\
@@
 The SM does not include a suitable dark matter candidate and a mechanisms that accounts for the baryon asymmetry of the universe.
-Additionally, a total of \textcolor{red}{25} parameters remain seemingly arbitrary and need to be fitted to data: 3 gauge couplings, 9 charged fermion masses, 3 mixing angles and one CP phase in the CKM matrix, the Higgs mass and quartic coupling, $\theta_{QCD}$, 3 neutrino masses, 3 neutrino mixing angles, 1 Dirac phase and, eventually,  2 Majorana phases.
+Additionally, up to a total of 25 parameters remain seemingly arbitrary and need to be fitted to data: 3 gauge couplings, 9 charged fermion masses, 3 mixing angles and one CP phase in the CKM matrix, the Higgs mass and quartic coupling, $\theta_{QCD}$, 3 neutrino mixing angles, 1 Dirac phase and, eventually,  2 Majorana phases.
 
@@
 \ref{eq:SMGroup}.
-Drawing a parallel with the electroweak symmetry breaking mechanism, where the $ SU(2)_T \otimes U(1)_Y$ is recovered from $U(1)_{EM}$, an interesting line of simplification for the SM group structure would be to devise a similar mechanism where $SU(3)_C  \otimes SU(2)_T \otimes U(1)_Y$ is recovered from an hypothetical larger group.
-Just as the electroweak unification becomes evident at energies higher than the Higgs VEV,  a direct manifestation of Grand Unification Theories (GUTs) would occur at even higher energies.
+\textcolor{red}{Drawing a parallel with the electroweak symmetry breaking mechanism, where the $ SU(2)_T \otimes U(1)_Y$ is recovered from $U(1)_{EM}$, an interesting line of simplification for the SM group structure would be to devise a similar mechanism where $SU(3)_C  \otimes SU(2)_T \otimes U(1)_Y$ is recovered from an hypothetical larger group.
+IS THIS CORRECT?} Just as the electroweak unification becomes evident at energies higher than the Higgs VEV,  a direct manifestation of Grand Unification Theories (GUTs) would occur at even higher energies.
 
 As the smallness of neutrino masses suggests the existence of a higher mass scale, an other, even stronger, hint to Grand Unification comes from the slope of running of the coupling constants.
-The coupling constants for the electromagnetic, weak and strong interactions in the SM vary as a function of the interaction energy as shown in figure \ref{fig:constants} \textcolor{red}{ find reference to coupling constant picture};  they cannot exactly meet under the current experimental constraints, but their trend is interesting enough to push theorists to construct theories where perfect unification is achieved with the addition of new particles.
-Both experiments and theory indeed suggest the energy scale for convergence of the running coupling constants of the Standard Model to be over $10^{15}$ GeV.
+The coupling constants for the electromagnetic, weak and strong interactions in the SM vary as a function of the interaction energy as shown in figure \ref{fig:constants};  they cannot exactly meet under the current experimental constraints, but their trend is interesting enough to push for the construction of theories where perfect unification is achieved through the addition of new particles.
+
+\begin{figure}[hbpt]
+\centering
+\includegraphics[scale=0.5]{Chapter-1/Images/RunninConstant.png}
+\caption{Evolution of the inverse of the three coupling constants in the Standard Model as a function of the momentum transferred, \cite{Kazakov:2000ra}.}
+\label{fig:constants}
+\end{figure}
+
+\textbf{SU(5).} The smallest simple group containing $SU(3)_C  \otimes SU(2)_T \otimes U(1)_Y$ is SU(5), as shown first by Georgi and Glashow in~\cite{PhysRevLett.32.438}.
+Quarks and leptons in this group fit the $\bar 5$ and 10 representations.
+The representation for left-handed fermions are the following
+\begin{align}
+\bar 5 &= (\nu_e, e^-)_L + \bar d_L\\
+10 &=  e^+_L  +  \bar u_L  + (u, d)_L ,
+\end{align}
+while the boson structure gains a new couple of super heavy bosons (X,Y)
+\begin{equation}
+24 = \underbrace{(8,1)}_{\text{gluons}} + \underbrace{ (1,3) + (1,1)}_{\text{W$^\pm$, Z, $\gamma$}}.. +  \underbrace{(3,2) + (\bar 3,2)}_{\text{X,Y bosons}}. 
+\end{equation}
+
+Nice features such as charge quantization and the identity between the positron and proton charge value come directly from the group structure. 
+The new super heavy bosons are colored and form a weak doublet.
+Their are the mediator of the interaction that turns quarks into leptons, leading to predict the existence of processes that violate baryon number, such as $p \rightarrow \pi^0  + e^+$  (see fig~ \ref{fig:MandatoryFeynmannDiagrams}, right).
+The prediction for proton decay lifetime, $\tau_p \sim \frac{M^4_X}{m^5_p}\sim10^{30\pm1.5}$~years, is unfortunately experimentally disproved by IMB and Super-K \cite{s}.
+
+\textbf{SO(10).} More complicated group structures, such as SO(10) are still viable candidates for GUT.
+SO(10) includes the same type of $X$ and $Y$ bosons as SU(5).
+Right-handed massive neutrinos are embedded in the construction of the irreducible representation of SO(10).
+Different patterns of SO(10) symmetry breaking to recover the SM are possible and lead to different predictions for the proton decay lifetime; some of these predictions are not excluded by the experiments \cite{PhysRevD.51.229}.
+
+\textbf{SUSY GUTs.} Supersymmetry theories allow for another family of GUTs\cite{supersymmetry}.
+In SUSY, every fundamental particle in the SM has a ``superpartner", identical in each quantum number except for the spin-statistics: the fermion supersymmetric partners are bosons and vice versa.
+Collider experiments (mainly LHC)  constrain the mass of the supersymmetric partners to be very heavy\cite{LHC}.
+The SU(5), SU(10) groups with a SUSY twist are the basic groups for SUSY GUTs.
+From the phenomenology point of view, SUSY models tend to push the proton decay life time higher by a factor of four, they solve the ``hierarchy problem", and they also predict new channels for the proton decay.
+In particular they predict the presence of kaons in the final product, with a dominant mode of $p\rightarrow\bar\nu K^+$%, since particles from the first generation (e.g.
+$u$ and $d$) need to decay into second and third generation particles (such as $s$).
+Predictions on the proton decay lifetime depend on the chosen SUSY model; again, some of the predictions are not excluded by the experiments \cite{}.
+
+%Both experiments and theory indeed suggest the energy scale for convergence of the running coupling constants of the Standard Model to be over $10^{15}$ GeV.
 This energy scale seems impossible to access by any foreseeable accelerator experiment.%, leaving baryon number violation  to be the only testable process.
@@
 %where $b_i$ are coefficients for each interaction, and  the constants assume the same value $\alpha_{GUT}$ at energy $E=M_{GUT}$
-
-\subsubsection{Nucleon decay}\label{sec:theoryPDK}
-Baryon number is accidentally conserved in the Standard Model.
-Even though no baryon number violation has been experimentally observed thus far, no underlying symmetry in line with the Noether paradigm \cite{Noether1971} explains its conservation.
-Almost all Grand Unified Theories predict at some level baryon number violation in the form of nucleon decay on long time-scales.
-Given the impossibility to reach grand unification energy scales with collider experiments ($\sqrt{s} > 10^{15}$ GeV),  an indirect proof of GUT is needed.
-The experimental observation of nucleon decay may be the only viable way to explore these theories and it is therefore a subject of great interest \cite{Adams:LBNE}.
 
@@
 \subsubsection{K$^{+}$Ar Cross section in the Context of Nucleon Decay Searches}
+\subsubsection{Nucleon decay}\label{sec:theoryPDK}
+Baryon number is accidentally conserved in the Standard Model.
+Even though no baryon number violation has been experimentally observed thus far, no underlying symmetry in line with the Noether paradigm \cite{Noether1971} explains its conservation.
+Almost all Grand Unified Theories predict at some level baryon number violation in the form of nucleon decay on long time-scales.
+Given the impossibility to reach grand unification energy scales with collider experiments ($\sqrt{s} > 10^{15}$ GeV),  an indirect proof of GUT is needed.
+The experimental observation of nucleon decay may be the only viable way to explore these theories and it is therefore a subject of great interest.
+
 In case of nucleon decay discovery, the dominant decay mode may uncover additional information about the GUT type.

### 2018-04-12 Chapter-1/1-Theory.tex
s

@@
 
-\subsection{Towards a more fundamental theory: GUTs}
+\subsection{Towards a more fundamental theory: GUTs}\label{ch:GUTsTheory}
 Despite its highly predictive power, a number of conceptual issues arise in the SM which disfavor it to be a good candidate for a fundamental theory.
@@
 Thus, understanding pion hadronic interactions with the nucleus is particularly important to model correctly resonant, DIS and coherent pion production in neutrino interactions, where the presence of pions in the nucleus is abundant.
-For example, in case of resonant scattering
+For example, in case of resonant scattering,
 \begin{equation}
@@
 
-Once the pion has left the target nucleus, the pion-argon hadronic cross section plays an important role in the pion transportation inside the argon medium: processes like pion absorption with emission of protons or pion charge exchange can greatly modify the topology of a neutrino interactions in the detector and lead to errors in the event classification.
+Once the pion has left the target nucleus, the pion-argon hadronic cross section plays an important role in the pion transportation inside the argon medium: processes like pion absorption with emission of nucleons or pion charge exchange can greatly modify the topology of a neutrino interactions in the detector and lead to errors in the event classification.
 Being able to reconstruct the details of pions inside the detector is an imperative for modern liquid argon neutrino experiment to achieve the design resolution for their key physics measurements.
@@
 \subsubsection{Signal Signatures}
-Strong hadronic interaction models \cite{9780198520085,9780471779957} predict the pion interaction processes with argon in the [100 -1200] MeV energy range. 
+Strong hadronic interaction models \cite{9780198520085,9780471779957} predict the pion interaction processes with argon in the [100 - 1200] MeV energy range. 
 The total hadronic $\pi^{-}$-Ar interaction cross section is defined as the one related to the single process driven only by the strong interaction which is dominant in the considered energy range. 
@@
 Stopping negative pions can form pionic argon, where the negative pion plays the role of an orbital electron.
-Since the pion mass is two orders of magnitude greater than the electron mass, the spatial wave function of the pion will overlap way more with the nucleus compared to the electron case.
+Since the pion mass is two orders of magnitude greater than the electron mass, the spatial wave form of the pion will overlap way more with the nucleus compared to the electron case.
 After the electromagnetic formation of the pionic atom, the pion will quickly absorbed by the nucleus, which is put in an excited state.
-The nucleus then de-excite with the emission of low energy nucleons and photons.
+The nucleus then de-excites with the emission of low energy nucleons and photons.
 Pion capture is predominant compared to pion decay, the other important process for very low energy pions.
@@
 A $\Delta$ resonance is often produced in the interaction, which subsequently decays inside the nucleus.
-Experimental results on several nuclei as reported in \cite{PhysRevC.14.635} are shown in Figure \ref{fig:pi1}; it is interesting to notice here how the shape fo the $\Delta$ resonance become less pronounced as a function of the mass number of the target nucleus.
+Experimental results on several nuclei as reported in \cite{PhysRevC.14.635} are shown in Figure \ref{fig:pi1}; it is interesting to notice here how the shape of the $\Delta$ resonance become less pronounced as a function of the mass number of the target nucleus.
 Pion interactions with heavier nuclei also shift the peak of the resonance at lower energy; this effect is due to kinematic considerations and to the difference in propagation of the $\Delta$ inside the nucleus.
@@
 
-\subsubsection{Pion Interaction Cross Section in generators}
+\subsubsection{Negative Pion Interaction Cross Section in Generators}
 LArIAT uses Geant4 as the default simulation package.
@@
 
-A broad review of the current status of negative and positive pion simulation in Geant4 and GENIE for $^{12}$C, $^{56}$Fe, and $^{40}$Ca can be found in \cite{Irene} \textcolor{red}{Irene}.
+The work in \cite{Irene} \textcolor{red}{Irene} reviews  the current status of negative and positive pion simulation in Geant4 and GENIE for $^{12}$C, $^{56}$Fe, and $^{40}$Ca.
 From that work, we report the results for $^{12}$C in Figure \ref{fig:pionC} as it allows a direct comparison between Geant4, GENIE and data.
-Geant4 predictions for $\pi^-$  on Carbon are in good agreement with data over all the spectrum, while GENIE predictions seem to show some features at around 500 MeV and 900 MeV maybe due ti higher resonances in the hA model.
+Geant4 predictions for $\pi^-$  on Carbon are in good agreement with data over all the spectrum, while GENIE predictions seem to show some features at around 500 MeV and 900 MeV, maybe due ti higher resonances in the hA model.
 From the same work, we also report the negative pion cross section on $^{40}$Ca in Figure \ref{fig:pionCa}, since the this is the nuclear medium closest argon with some available data.
@@
 Even though no baryon number violation has been experimentally observed thus far, no underlying symmetry in line with the Noether paradigm \cite{Noether1971} explains its conservation.
-Almost all Grand Unified Theories predict at some level baryon number violation in the form of nucleon decay on long time-scales.
-Given the impossibility to reach grand unification energy scales with collider experiments ($\sqrt{s} > 10^{15}$ GeV),  an indirect proof of GUT is needed.
-The experimental observation of nucleon decay may be the only viable way to explore these theories and it is therefore a subject of great interest.
+As shown in section \ref{ch:GUTsTheory}, almost all Grand Unified Theories predict at some level baryon number violation in the form of nucleon decay on long time-scales.
+Given the impossibility to reach grand unification energy scales with collider experiments ($\sqrt{s} > 10^{15}$ GeV),  an indirect proof of GUTs is needed.
+The experimental observation of nucleon decay may be the only viable way to explore these theories.
 
@@
 
-For the LArIAT cross section analysis, the kaons considered span a momentum inside the TPC from 800 MeV/c and 100 MeV/c.
+For the LArIAT cross section analysis, the kaons considered span a momentum inside the TPC from 100 MeV/c to 800 MeV/c.
 In this energy range, the relevant K-Nucleon interactions are according to \cite{fesbach1992theoretical}:

### 2018-04-15 Chapter-2/2-LArTPCs.tex
s

@@
-\chapter{Liquid Argon Detectors at the Intensity Frontier}\label{ch:1}
+\chapter{Liquid Argon Detectors at the Intensity Frontier}\label{ch:2}
 
@@
 \section{Liquid Argon Time Projection Chambers at the Intensity Frontier}
-
-%Bubble-chamber experiments played a key role in probing the properties of ?-interactions.
-The Liquid Argon Time Projection Chambers (LArTPC) technology,  first proposed by C.Rubbia in 1977 with ICARUS project [14], is considered the modern evolution of bubble-camber concept, with the additional features of three-dimensional event reconstruction, high-resolution calorimetry, active mass coincident with detector sensitive mass and can intrinsically supply a trigger signal (self-triggering) by means of the scintillation light produced in the liquid noble gas.
-This technology is ideal to perform $\nu$-studies in a broad energy range, from MeV up to few GeV, with high event reconstruction efficiency, thanks to the capability of particle identifcation and detailed reconstruction of different interaction topologies.
-In Figure 1.4 is shown a neutrino interaction event, producing a proton, a pion and a muon, as seen in a bubble chamber and in a LArTPC.
 
@@
 \subsubsection{LAr Scintillation Process}
+Liquid argon emits scintillation light at the passage of charge particle.
+The light yield depends on the argon purity, the electric field, the dE/dx and particle type, averaging at the tens of thousands of photons per MeV.
+Light emission peaks in the ultraviolet at a 128 nm, shown in comparison to Xenon and Kypton in Figure  [183].
+
+The de-excitation of Rydberg dimers in the argon is responsible for the scintillation light.
+Rydberg dimers exist in two states:  a singlet and a triplet.
+The time constant for the singlet radiative decay is 6 ns, resulting in a prompt component for the scintillation light.
+The decay of the triplet is delayed by intersystem crossing, producing a slow component with a time constant of $\sim$~1500 ns.
+``Self-trapped exciton luminescence" and  ``recombination luminescence" are the two processes responsible for the creation of the Rydberg dimers.
+In the first process, a charged particle excites an argon atom which becomes self-trapped in the surrounding bulk of argon,  forming a dimer; the dimer is in the singlet state 65\% of the times and in the triplet state 35\% of the times.
+In case of recombination luminescence, the charged particle transfers enough energy to ionize the argon.
+The argon form a charged argon dimer state, which quickly recombines with the thermalized free electron cloud.
+Excimer states are produced in the recombination, roughly half in the singlet and half in the triplet state.
+The light yield dependency on the electric field, on the dEdx and particle type derives from the role of free charge in the recombination luminescence process.
+The separation between the argon ions and the free electron cloud depend on the electric field.
+On one hand, a strong electric field diminishes the recombination probability, leading to a smaller light yield; on the other, it increases the free charge drifting towards the anode plane.
+Hence, charge and light as a function of the electric field are anti-correlated.
+
+%The recombination rate is proportional to the local densities of both free electrons and
+%positive ions, so the square of the local ionization density. Recombination is therefore
+%enhanced for highly ionizing particles [102]. This results in a characteristic dE/dx
+%dependence in the time profile of scintillation light [106], which can be used for pulseshape-based particle identification [186], [187].
+
+%Although attempts to measure the absorption length of pure argon at 128 nm
+%have been hindered by the presence of impurities [188], it is expected to be much
+%longer than the length scale of existing or proposed LArTPCs.
+
+The production mechanism through emission from bound excimer states implies that argon is  transparent  to its own scintillation light: the emitted photons are not energetic enough to re-excite the argon bulk.
+%Scattering processes, on the other hand, do occur on length scales comparable to
+%the size of existing detectors. In particular, Rayleigh scattering of 128 nm light
+%from thermal density fluctuations in the liquid has a predicted scattering length of
+%approximately 90 cm [190]. Scattering causes the light propagating between a source
+%and optical detector to take a longer path than the direct line of sight, and affects the
+%arrival times and directions of detected photons. This is particularly important for ??0
+%determination in a LArPC. A direct measurement of the Rayleigh scattering length
+%has been made [191] and a value 66 cm was obtained, shorter than the theoretical
+%prediction.
+
 \subsubsection{Wavelength Shifting of LAr Scintillation Light}
+\subsection{Signal processing}
 \section{The SBN Program: Neutrino Interaction and Detection}

### 2018-04-15 Chapter-7/7-BackgroundSub.tex
s

--- 2018-03-28 Chapter-7/7-BackgroundSub.tex
@@
 \section{Assessing Beamline Contamination}
-What is the beamline contamination? We define beamline contamination every TPC track matched to the WC track which is not a primary pion. There are 4 different types of beamline contaminations:
+Even if pions are by far the biggest beam component in negative polarity runs, the LArIAT beam is not a pure pion beam.
+While useful to discriminate between pions, kaons, and protons, the beamline detectors are not sensitive enough to  discriminate among the lighter particles in the beam: electrons, muons and pions fall under the same mass hypothesis.
+Thus, we need to assess the contamination from beamline particles other than pions in the event selections used for the pion cross section analysis and correct for its effects.
+
+We define beamline contamination every TPC track matched to the WC track which is not a primary pion. Potentially, there are 4 different types of beamline contaminations:
 \begin{itemize}
@@
 The first step is to estimate what percentage of events used in the cross section calculation is not a primary pion.
+The next two sections will illustrate this estimate for the electrons, muons and secondaries from pion event.
+We estimate the last type of contamination, the ``matched pile up" events, to be a negligible fraction, because of the definition of the WC2TPC match: we deem the probability of a single match with a halo particle in the absence of a beamline particle\footnote{ Events with multiple WC2TPC matches are always rejected.} negligibly small.
 
@@
 \includegraphics[width=0.5\textwidth,height=\textheight,keepaspectratio]{Chapter-7/Images/Beam100A.png}
-\label{fig:BeamComposition}
 \caption{Beam composition for the -60A runs (left) and -100A runs (right).
-The solid blue plot represents the pion content, the yellow plot represents the muon content and the grey plot represents the electron content.
+The solid blue plot represents the simulated pion content, the yellow plot represents the simulated muon content and the grey plot represents the simulated electron content.
 The plots are area normalized to the number of data events, shown in red.
 }
+\label{fig:BeamComposition}
 \end{figure}
 
-Table \ref{tab:beamline} shows the beam composition per magnet setting after the mass selection according to the G4simulation, 
+Table \ref{tab:beamline} shows the beam composition per magnet setting after the mass selection according to the G4Beamline simulation.
 \begin{table}[]
 \centering
-\label{tab:beamline}
 \begin{tabular}{|l|c|c|}
@@
 \end{tabular}
-\caption{Beamline composition per magnet settings}
+\caption{Simulated beamline composition per magnet settings}
+\label{tab:beamline}
 \end{table}
@@
 \centering
-\label{tab:databreakdown}
 \begin{tabular}{|l|c|c|c|c|c|}
@@
                                                      & I = -60 A          & I = -100 A   & Total     &  w$_{60A}$ &  w$_{100A}$\\ \hline
-Data events after Mass Cut         &     70192          &  76056       & 146248 & 0.52 & 0.48\\ \hline
-Data events for Cross Section     &                         &                   &              & & \\ \hline
+N Data Events after Mass Selection     &     70192          &  76056       & 146248 & 0.48 & 0.52\\ \hline
+%Data events for Cross Section          &                         &                   &              & & \\ \hline
+%Estimated e in the beamline             &   18671.1         &    6769.0  &              & & \\ \hline
+%Estimated $\mu$ in the beamline     &     3228.8         &     2814.1     &              & & \\ \hline
+%Estimated $\pi$ in the beamline       &    48292.1        &   66472.9   &              & & \\ \hline
 \end{tabular}
-\caption{Data events per magnet settings}
+\caption{Number of data events which fit the pion mass hypothesis as a function of magnet settings. The last two columns represent the faction of the data in the given magnet setting.}
+\label{tab:databreakdown}
 \end{table}
 
-We calculate the electron to pion and muon to pion ratio on the whole sample as the weighted sum of the corresponding ratio in the two current settings, 
+We calculate the electron to pion, as well as the muon to pion ratio on the whole sample as the weighted sum of the corresponding ratio in the two current settings, 
 \begin{equation}
@@
 \end{equation}
-where the weights $w_{60A}$ and $w_{100A}$ are the percentage of events in the corresponding magnet configuration passing the mass selection in data, as shown in \ref{tab:databreakdown}.
+where the weights $w_{60A}$ and $w_{100A}$ are the percentage of events in the corresponding magnet configuration passing the mass selection in data, as shown in table \ref{tab:databreakdown}.
+Figure \ref{fig:BeamComposition} shows the mometum predictions from G4Beamline overlaid with data for the 60A runs (left) and for the 100A runs (right).
+The predictions for electrons, muons and pions have been staggered and their sum is area normalized to data.
+Albeit not perfect, these plots show a reasonable agreement between the momentum shapes in data and MC.
+We attribute  the difference in shape to the lack of simulation of the WC efficiency in the MC which is momentum dependent and leads to enhance the number events in the center of the momentum distribution.
 
-Once the beam composition is know,  we simulate the electrons, muons and pions with the DDMC and we subject the three samples to the same selection chain (WC2TPC match, shower filter, pile up filter, etc...).
-The percentage of electrons and muons surviving the selection chain is the  electron and muon contamination in the pion cross section sample.
+Once the beam composition is know,  we simulate the electrons, muons and pions with the DDMC and we subject the three samples to the same selection chain (WC2TPC match, shower filter, pile up filter).
+The percentage of electrons and muons surviving the selection chain weighted by the beam composition is the  electron and muon contamination in the pion cross section sample, as shown in Table \ref{tab:MCafterCutContaminants}.
 
 \subsection{Contamination from secondaries}
-The percentage of secondaries is given in the MC by the number of matched WC2TPC tracks which are not flagged as primary by Geant4.
-We estimate the last type of contamination, the ``matched pile up" events, to be a negligible fraction, because of the definition of the WC2TPC match: we deem the probability of a single match with a halo particle in the absence of a beamline particle\footnote{ Events with multiple WC2TPC matches are always rejected.} extremely small.
+Pions can travel the length of the LArIAT beamline and interact hadronically in the steel or in the non-instrumented argon upstream to the TPC front face.
+One of these products can leak into the TPC and be matched with the WC track, contribuiting to the pool of events used for the cross section calculation.
+We call this type of particles ``secondaries" from pion events, with a terminology inspired by Geant4.
+We estimate the number of secondaries using the DDMC pion sample.
+The percentage of secondaries is given by the number of matched WC2TPC tracks whose corresponding particle is not flagged as primary by Geant4 and is not a muon, to avoid double counting with the G4Beamline estimate.
+The secondary to pion ratio is $X$\% in the 60A sample and $Y$\% in the 100A sample.
 
-\section{Subtraction}
-Once we estimate the contaminants to primary pion ratio, the next step is subtracting their contribution from data for each type of contaminant independently.
-The contaminant samples are reconstructed and the corresponding interacting and incident histograms are produced.
-We then perform a bin by bin subtraction in the data interacting and incident histograms separately.
-A graphical rendering of this procedure is shown in Fig \ref{fig:backgroundSubtraction}
-Once the data is background subtracted, we apply the correction laid out in the previous section.
-\textcolor{blue}{How do we account for the error in the contamination subtraction? We change the electron/pion and muon/pion ratio and we see how much difference we get?}
+\section{Beamline Background Subtraction}
+Once we estimate the contaminants to primary pion ratio, the next step is subtracting their collective contribution from data.
+To do so, we simulate the same number of electrons, muons and pions with the DDMC separately for the two magnet settings, and we apply the same selection filters on the three samples.
+The number of events per particle species surviving this selection is shown on table \ref{tab:MCafterCutContaminants}.
+
+\begin{table}[]
+\centering
+\begin{tabular}{| l | l | l | l | l | l | l | l | }
+\hline
+                                                  &  $\pi^-$  60A  & $\mu^-$ 60A & $e^-$  60A & $\pi^-$ 100A  & $\mu^-$ 100A & $e^-$  100A  \\
+\hline
+Total Initial events                     & 334500   & 334500  & 334500  & & & \\
+After Multiplicity Rejection & 331313   & 322436   &  186261& & &\\
+After WC2TPC: Selection & 201458   &  285686  &  79109  & & & \\
+Evts After Shower Rejection    & 191655   &  277914 &  17477    & & &\\
+  \hline
+     &     &   &     & & &\\
+Survival rate               &57\%  &  83\%   & 5\%  &     &  &\\
+     &     &   &     & & &\\
+Beam Composition               &     &   &     & & &\\ 
+After Selection                      &     &   &     & & &\\
+                                             & 88.5\%  & 8.5\%  &  3\%   & & &\\
+\hline
+%         &                              &             &                      &  \\ \hline
+\end{tabular}
+\caption{MC selection flow per particle species.}
+\label{tab:MCafterCutContaminants}
+\end{table}
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=\textwidth]{Chapter-7/Images/Staggered60A.png}
+\caption{Left: staggered contributions to the interacting kinetic energy distribution for electron (grey), muons (yellow) and pion (blue) in the 60A simulation sample.
+Right: staggered contributions to the incident kinetic energy distribution for electron (grey), muons (yellow) and pion (blue) in the 60A simulation sample.
+}
+\label{fig:stag60A}
+\end{figure}
+
+We then produce the interacting and incident histograms for the events surviving the selection for both the pions and the contaminants, weighted by the estimated beam composition.
 
@@
 \end{figure}
+
+We then evaluate the relative contribution of the contaminants bin by bin in the interacting and incident histograms separately.
+In data, we subtract this estimated relative contaminants contribution on the interacting and incident histograms  bin by bin.
+
+We estimate the systematic uncertainty on the cross section from this subtraction procedure by varying the electron to pion and muon to pion ratio in a suitable range of values. Figure
 

### 2018-04-17 Chapter-5/5-SamplesPreparation.tex
s

@@
 \chapter{Samples Preparation}\label{ch:samples}
-
+This chapter describes the data and Monte Carlo samples used for the cross section analyses, % as well as the work necessary to prepare the samples in a usable state for the physics analyses.
 \section{LArIAT Data}
@@
 
-\section{Energy Calibration}
+\section{Energy Calibration}\label{ch:EnCalibration}
 \section{Tracking Studies}
@@
 The distributions for data and MC for $\sigma_\alpha$ and $\sigma_S$ are given in \ref{fig:trackingResolution}.
+
+\subsection{Estimate of $E_{loss}$ before the TPC}
+The beamline particles travel a path from when their momentum is measured by the beamline detector, until they are tracked inside the TPC.
+In the current LArIAT geometry, a particle leaving the fourth wire chamber will encounter the materials listed in Table \ref{tab:budget} before being registered again.
+The energy lost by the particle in this non instrumented material modifies the particle's kinetic energy and directly affects the cross section measurement, as shown in equation \ref{eq:enFF}.
+
+\begin{table}[h!]
+\centering
+\begin{tabular}{|l|l|l|}
+\hline
+Material  & density {[}g/cm$^3${]} & width {[}cm{]}    \\ \hline
+Fiberglass laminate (G10)      & 1.7                             & 1.28                              \\
+Liquid Argon                           & 1.4                             & 3.72                             \\
+Stainless Steel                        & 7.7                            & 0.23                             \\
+Titanium                                  & 4.5                            & 0.04                             \\ 
+Air                                            &  1.2 $\cdot10^{-3}$  & 89.43                              \\
+Plastic Scintillator                    & 1.03                          & 5.30                              \\ \hline
+\end{tabular}
+\caption{LArIAT material budget from WC4 to the TPC Front Face.}
+\label{tab:budget}
+\end{table}
+
+We estimate the uncertainty on the energy loss between the beamline momentum measurement and the TPC, $E_{loss}$, using the DDMC pion sample.
+We shoot pions from WC4 with the same momentum distribution as in the beamline data and plot the true $E_{loss}$ for that sample.
+The distribution for $E_{loss}$ for the pion sample is shown in Figure \ref{fig:Eloss}.
+We estimate the energy loss for pions to be $E_{loss} = 37 \pm 7 $~MeV where we use the average energy lost as the central value and the standard deviation of the distribution as the uncertainty.
+
+\begin{figure}[hpbt]
+\centering
+\includegraphics[scale=0.4]{Chapter-5/Images/cELossPions.png}
+\caption{Energy loss by simulated negative pions downstream from WC4 and upstream from the TPC.}
+\label{fig:Eloss}
+\end{figure}

### 2018-04-17 Chapter-2/2-LArTPCs.tex
s

--- 2018-04-08 Chapter-2/2-LArTPCs.tex
@@
 
-%Bubble-chamber experiments played a key role in probing the properties of ?-interactions.
-The Liquid Argon Time Projection Chambers (LArTPC) technology,  first proposed by C.Rubbia in 1977 with ICARUS project [14], is considered the modern evolution of bubble-camber concept, with the additional features of three-dimensional event reconstruction, high-resolution calorimetry, active mass coincident with detector sensitive mass and can intrinsically supply a trigger signal (self-triggering) by means of the scintillation light produced in the liquid noble gas.
-This technology is ideal to perform $\nu$-studies in a broad energy range, from MeV up to few GeV, with high event reconstruction efficiency, thanks to the capability of particle identifcation and detailed reconstruction of different interaction topologies.
-In Figure 1.4 is shown a neutrino interaction event, producing a proton, a pion and a muon, as seen in a bubble chamber and in a LArTPC.
-
 \subsection{Time Projection Chamber}
 \subsection{Ionization Detectors with Noble Liquids}
-\subsection{LArTPC: Principles of Operation}
+\subsection{LArTPC: Principles of Operation}\label{sec:LArTPCWorkingPrinciple}
 \subsection{Liquid Argon Ionization Charge Detection}
@@
 \subsubsection{Recombination Effect}
-\subsection{Liquid Argon scintillation Light Detection}
-\subsubsection{LAr Scintillation Process}
+\subsection{Liquid Argon Scintillation Light }
+Liquid argon emits scintillation light at the passage of charged particles.
+LArTPCs leverage this property to determine when the ionization charge begins to drift towards the anode plane.
+%Scintillation light can also be used for particle identification, as discussed in the next sections.
+
+\subsubsection{Scintillation Process}
+Scintillation light in argon peaks in the ultraviolet at a 128 nm, shown in comparison to Xenon and Kypton in Figure  [183].
+The light yield collected by the optical detector depends on the argon purity, the electric field, the dE/dx and particle type, averaging at the tens of thousands of photons per MeV.
+The de-excitation of Rydberg dimers in the argon is responsible for the scintillation light. 
+Rydberg dimers exist in two states:  singlets and a triplets.
+The time constant for the singlet radiative decay is 6 ns, resulting in a prompt component for the scintillation light.
+The decay of the triplet is delayed by intersystem crossing, producing a slow component with a time constant of $\sim$~1500 ns.
+``Self-trapped exciton luminescence" and  ``recombination luminescence" are the two processes responsible for the creation of the Rydberg dimers.
+In the first process, a charged particle excites an argon atom which becomes self-trapped in the surrounding bulk of argon,  forming a dimer; the dimer is in the singlet state 65\% of the times and in the triplet state 35\% of the times.
+In case of recombination luminescence, the charged particle transfers enough energy to ionize the argon.
+The argon ion forms a charged argon dimer state, which quickly recombines with the thermalized free electron cloud.
+Excimer states are produced in the recombination, roughly half in the singlet and half in the triplet state.
+The light yield dependency on the electric field, on the dE/dx and particle type derives from the role of free charge in the recombination luminescence process.
+The spacial separation between the argon ions and the free electron cloud depends on the electric field.
+On one hand, a strong electric field diminishes the recombination probability, leading to a smaller light yield; on the other, it increases the free charge drifting towards the anode plane.
+Hence, the amount of  measurable charge and light anti-correlates as a function of the electric field.
+Ionizing particles in the argon modify the local density of both free electrons and ions depending on their dE/dx.
+Since the recombination rate is proportional to the square of the local ionization density, highly ionizing particles boost recombination and the subsequent light yield compared to MIPs.
+The possibility to leverage this dependency for pulseshape-based particle identification has been shown in [186], [187].
+
+\subsubsection{Attenuation of Scintillation Light}
+The production mechanism through emission from bound excimer states implies that argon is  transparent to its own scintillation light.
+The photons emitted from these metastable states are not energetic enough to re-excite the argon bulk, greatly suppressing absorption mechanisms.
+In a LArTPC however, several processes modify the light yield in between the production location and the optical detector.
+The most important processes are  Rayleigh scattering in the argon and light quenching due to impurities.
+Rayleigh scattering changes the path of light propagation in argon, prolonging the time between light production and detection.
+The scattering length has been measured to be 66 cm [191] , shorter than the theoretical prediction of $\sim$~90 cm [190]; this value is short enough to be relevant for the current size of LArTPCs detectors.
+In fact, Rayleigh scattering worsen the resolution on $t_0$, the start time for charge drifting, and  alters the light directionality, complicating the matching between light and charge coming from the same object in case of multiple particles in the detector.
+
 \subsubsection{Wavelength Shifting of LAr Scintillation Light}
+\subsection{Signal processing}
 \section{The SBN Program: Neutrino Interaction and Detection}

### 2018-04-17 Chapter-3/3-TheLArIATExperiment.tex
s

@@
 As in any other LArTPC, argon purity is a crucial parameter for LArIAT.
-Indeed, the presence of contaminants effects both the basic working principles of a LArTPC: electronegative contaminants such as oxygen and water decrease the number of ionization electrons collected on the wires after drifting through the volume, while contaminants such as Nitrogen decrease the light yield from scintillation light, especially in its slow component.
+Indeed, the presence of contaminants effects both the basic working principles of a LArTPC, as shown in section \ref{sec:LArTPCWorkingPrinciple}: electronegative contaminants such as oxygen and water decrease the number of ionization electrons collected on the wires after drifting through the volume.
+In addition, contaminants such as Nitrogen decrease the light yield from scintillation light, especially in its slow component.
 In LArIAT, contaminations should not exceed the level of 100 parts per trillion (ppt).

### 2018-04-19 Chapter-2/2-LArTPCs.tex
s

@@
 \subsection{LArTPC: Principles of Operation}\label{sec:LArTPCWorkingPrinciple}
-\subsection{Liquid Argon Ionization Charge Detection}
+Details on the production and detection of ionization charge and scintillation light are provided in \ref{sec:light} and \ref{sec:light} respectively.
+\subsection{Liquid Argon Ionization Charge Detection}\label{sec:charge}
 \subsubsection{Electron Life Time \& purity}
+LArTPCs use  hermetically sealed and leak-checked vessels to abate the leakage and diffusion of contaminants into the system.
+The liquid argon filling of the volume occurs after the vessel is evacuated or purged with gaseous argon [205]  to reduce remaining gases in the volume.
+Even so, the construction of a pure tank of argon is unviable, as several sources of impurity remain.
+In particular, impurities can come from the raw argon supply, the argon filtration system and from the outgassing from internal surfaces.
+Outgassing is a continuous diffusive process  producing contaminants, especially water, even after the vessel is sealed, particularly from materials in the ullage region\footnote{.
+While the liquid argon low temperature reduces outgassing in the liquid, this process remains significant for absorptive material (such as plastic) above the surface of the liquid phase}.
+Since research-grade argon comes from the industrial distillation of air, the impurities with the highest concentration are nitrogen, oxygen and water, generally maintained under the 1 part per million level by the vendor.
+Even so, a higher level of purity is necessary to achieve a free electron life time usable in meter scale detectors.
+Thus, argon  is constantly  filtered in the cryogenic system, which reduce the oxygen and water contamination to less than 100 parts per trillion.
+The filtration system depends on the size and drift distance of the experiment and, for experiments on several meters scale, it includes an argon recirculation system.
+
 \subsubsection{Space Charge Effect}
 \subsubsection{Recombination Effect}
-\subsection{Liquid Argon Scintillation Light }
+\subsection{Liquid Argon Scintillation Light }\label{sec:light}
 Liquid argon emits scintillation light at the passage of charged particles.
@@
 
-\subsubsection{Attenuation of Scintillation Light}
+\subsubsection{Effects modifying the  Light Yield}
 The production mechanism through emission from bound excimer states implies that argon is  transparent to its own scintillation light.
@@
 In a LArTPC however, several processes modify the light yield in between the production location and the optical detector.
-The most important processes are  Rayleigh scattering in the argon and light quenching due to impurities.
+In a hypothetical pure tank of argon, Rayleigh scattering would be the most important processes modifying the light yield.
 Rayleigh scattering changes the path of light propagation in argon, prolonging the time between light production and detection.
@@
 In fact, Rayleigh scattering worsen the resolution on $t_0$, the start time for charge drifting, and  alters the light directionality, complicating the matching between light and charge coming from the same object in case of multiple particles in the detector.
+
+Traces of impurities such as oxygen, water and nitrogen  in liquid argon can effect the light yield via two mechanisms: absorption and quenching. 
+Absorption occurs as the interaction of a 128 nm photon directly with the impurity dissolved in the liquid argon.
+Differently, quenching occurs as the interaction of an argon excimer and an impurity, where the excimer transfers its excitation to the impurity and  dissociates non-radiatively.
+Given this mechanism, it is evident how quenching is both a function of the impurity concentrations and the excimer lifetime.
+Since the triplet states live much longer than the singlet states,  quenching occurs mainly on triplet states, affecting primarily the slow component of the light,  reducing the scintillation yield and a shortening of the scintillation time constants.
+
+The stringent constraints for the electron life time limit the presence of oxygen and water to such a low level that both absorption and quenching on these impurity is not expected to be significant [210].
+Contrarily, the nitrogen level is not bound by the electron life time constraints -- nitrogen being an inert gas, expensive to filter.
+Thus, nitrogen is often present at the level provided by the vendor.
+For a nitrogen  concentration of 2 parts per million,  typical of the current generation of LArTPC, the attenuation length due to nitrogen has been measured to be $\sim$30 meters \cite{1748-0221-8-07-P07011}.
+
+\textcolor{red}{ADD quenching effects}
 

### 2018-04-19 Chapter-2/2-LArTPCs.tex
s

@@
 
-\subsubsection{Effects modifying the  Light Yield}
+\subsubsection{Effects Modifying the  Light Yield}
 The production mechanism through emission from bound excimer states implies that argon is  transparent to its own scintillation light.
-The photons emitted from these metastable states are not energetic enough to re-excite the argon bulk, greatly suppressing absorption mechanisms.
-In a LArTPC however, several processes modify the light yield in between the production location and the optical detector.
+In fact, the photons emitted from these metastable states are not energetic enough to re-excite the argon bulk, greatly suppressing absorption mechanisms.
+In a LArTPC however, several processes modify the light yield in between the location where light is produced and the optical detector.
 In a hypothetical pure tank of argon, Rayleigh scattering would be the most important processes modifying the light yield.
@@
 The scattering length has been measured to be 66 cm [191] , shorter than the theoretical prediction of $\sim$~90 cm [190]; this value is short enough to be relevant for the current size of LArTPCs detectors.
-In fact, Rayleigh scattering worsen the resolution on $t_0$, the start time for charge drifting, and  alters the light directionality, complicating the matching between light and charge coming from the same object in case of multiple particles in the detector.
+In fact, Rayleigh scattering worsen the resolution on $t_0$, the start time for charge drifting, and  alters the light directionality, complicating the matching between light and charge coming from the same object in case of multiple charged particles in the detector.
 
-Traces of impurities such as oxygen, water and nitrogen  in liquid argon can effect the light yield via two mechanisms: absorption and quenching. 
+Traces of impurities in argon such as oxygen, water and nitrogen  also affect the light yield, mainly  via absorption and quenching mechanisms. 
 Absorption occurs as the interaction of a 128 nm photon directly with the impurity dissolved in the liquid argon.

### 2018-04-19 Chapter-2/2-LArTPCs.tex
s

@@
 \subsection{LArTPC: Principles of Operation}\label{sec:LArTPCWorkingPrinciple}
+
+%The time projection chamber (TPC) was invented by Nygren in 1974 [91].
+The proposal to implement a liquid argon TPC (LArTPC) for neutrino physics was made by Rubbia in 1977 [92], and the concept was implemented by the ICARUS collaboration [93].
+
+To the bare bones, a LArTPC is a bulk of liquid argon sandwiched in a flat capacitor, equipped with a light collection system.
+A uniform electric field of the order of ~500 V/cm is maintained constant between the faces of the capacitor.
+Two processes occurs when an ionizing radiation traverses the argon active volume: the argon ionize and scintillates along the charge particle's trajectory.
 Details on the production and detection of ionization charge and scintillation light are provided in \ref{sec:light} and \ref{sec:light} respectively.
+The optical part of the detector collects the argon scintillation light in matters of nanoseconds.
+This flash of light flags the starting time of an event in the chamber, $t_0$.
+The uniform electric field drifts the electrons from ionization towards the anode plane and the ionized argon towards the cathode.
+
 \subsection{Liquid Argon Ionization Charge Detection}\label{sec:charge}
@@
 \subsubsection{Recombination Effect}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%                     Light Detection             %%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Liquid Argon Scintillation Light }\label{sec:light}
@@
 Thus, nitrogen is often present at the level provided by the vendor.
+The effects of nitrogen on argon scintillation light have been studied in the WArP R\&D program and at several test stands.
+The quenching process induced by nitrogen in liquid Ar has been measured to be proportional to the nitrogen concentration, with a rate constant of $\sim  0.11$ $\mu s^{-1}$ ppm$^{-1}$; appreciable decreasing in lifetime and relative amplitude of the slow component have been shown for contamination as high as a few ppm of nitrogen \cite{1748-0221-5-06-P06003}.
 For a nitrogen  concentration of 2 parts per million,  typical of the current generation of LArTPC, the attenuation length due to nitrogen has been measured to be $\sim$30 meters \cite{1748-0221-8-07-P07011}.
 
-\textcolor{red}{ADD quenching effects}
+\subsubsection{Wavelength Shifting of LAr Scintillation Light}
+Liquid argon scintillation light is invisible for most optical detectors deployed in LArTPC, such as cryogenic PMTs and SiPMs, since a wavelength of 128 nm is  generally too short to be absorbed from most in glasses, polymers and semiconductor materials.
+Research on prototype SiPMs absorbing directly VUV light and their deployment in noble gasses experiment is ongoing but not mature \cite{1748-0221-8-01-C01003}.
+Thus, experiments need to shift the wavelength of scintillation light to be able to detect it.
+Albeit deployed in different ways, neutrinos and dark matter experiments commonly use  1,1,4,4-tetraphenyl-butadiene (TPB) to shift the scintillation light.
+TPB, whose chemical structure is shown in figure , absorbs the vacuum ultraviolet (VUV) light and emits in the visible at $\sim$~425 nm [231], with a ratio of visible photon emitted per VUV photon absorbed of $\sim$1.2:1 \cite{GEHMAN2011116}.
 
-\subsubsection{Wavelength Shifting of LAr Scintillation Light}
+Neutrino experiments typically coat their optical detector system evaporating a layer of TPB either directly on the PMTs glass surface or on acrylic plates mounted in front of the PMTs \textcolor{red}{cite microboone}; this technique allows the fast detection light coming directly from the neutrino interaction.
+Dark matter experiments typically evaporate TPB on reflective foils mounted on the inside walls of the sensitive volume and detect the light after it has been reflected; this technique leads to a higher and more uniform light yield, though scattering effects for both the visible and VUV light augment the propagation time and hinder directionality information\textcolor{red}{some DM detector} .
+In order to take advantage of both these techniques, hybrid systems with PMT coating and foils are being considered for the next generation of large neutrino detectors \textcolor{red}{cite SBND?}.
+
 \subsection{Signal processing}

### 2018-04-19 Chapter-2/2-LArTPCs.tex
s

@@
 To the bare bones, a LArTPC is a bulk of liquid argon sandwiched in a flat capacitor, equipped with a light collection system.
-A uniform electric field of the order of ~500 V/cm is maintained constant between the faces of the capacitor.
+A uniform electric field of the order of 500 V/cm is maintained constant between the faces of the capacitor.
+The anode is sensitive to ionization charge and it is usually made of two or more planes segmented into multiple parallel sense wires; different geometries for the anode segmentation are under study \cite{1748-0221-8-07-P07002}.
 Two processes occurs when an ionizing radiation traverses the argon active volume: the argon ionize and scintillates along the charge particle's trajectory.
@@
 The optical part of the detector collects the argon scintillation light in matters of nanoseconds.
-This flash of light flags the starting time of an event in the chamber, $t_0$.
-The uniform electric field drifts the electrons from ionization towards the anode plane and the ionized argon towards the cathode.
+This flash of light determines the start time of an event in the chamber, $t_0$.
+The uniform electric field drifts the electrons from ionization towards the anode plane where the anode sense wires see either an induced current by the drifting charge (on induction planes) or the injection of the ionization charge (collection plane) [94] \footnote{The ionized argon also drifts,  in the opposite directions compared to the electrons and with a much longer drift time, given that the drift time is proportional to the particle mass.
+Ionized argon is collected on the cathode which is not instrumented, so it is not used to infer information about the interactions in the chamber.}.
+The arrival time of the charge on the anode sense wires is used to measure the position of the original ionizing radiation in the drift direction.
+In fact, since the constant electric field implies that the drift velocity is also constant in the chamber, the position of the original ionization is simply given by the multiplication of the drift velocity by the drift time, where we define as ``drift time" the difference between $t_0$ and the charge arrival time on the wire planes.
+Combining the spatial information on the different wire planes results in the mapping of a bi-dimensional projection of the interaction pattern in the plane perpendicular to the drift direction.
+
+Calorimetry
 

### 2018-04-19 Chapter-2/2-LArTPCs.tex
s

@@
 \chapter{Liquid Argon Detectors at the Intensity Frontier}\label{ch:2}
+In the next few years, LArTPC experiments -- such as the Short-Baseline Neutrino program (SBN) and DUNE -- will  be the tools to answer some of the burning questions in neutrino physics today.
+This section illustrates the operational principles of this detector technology, as well as the scope of SBN, DUNE and LArIAT in the current intensity frontier panorama.
 
-In the next few years, LArTPC experiments -- such as the Short-Baseline Neutrino program (SBN) and DUNE -- will be major players in the intensity frontier field.
-
-\section{Liquid Argon Time Projection Chambers at the Intensity Frontier}
+\section{The Liquid Argon Time Projection Chamber Technology}
 
 \subsection{Time Projection Chamber}
-\subsection{Ionization Detectors with Noble Liquids}
 \subsection{LArTPC: Principles of Operation}\label{sec:LArTPCWorkingPrinciple}
@@
 A uniform electric field of the order of 500 V/cm is maintained constant between the faces of the capacitor.
-The anode is sensitive to ionization charge and it is usually made of two or more planes segmented into multiple parallel sense wires; different geometries for the anode segmentation are under study \cite{1748-0221-8-07-P07002}.
-Two processes occurs when an ionizing radiation traverses the argon active volume: the argon ionize and scintillates along the charge particle's trajectory.
-Details on the production and detection of ionization charge and scintillation light are provided in \ref{sec:light} and \ref{sec:light} respectively.
+The anode is sensitive to ionization charge and it is usually made of two or more planes segmented into several hundreds parallel sense wires a few millimeters apart; different geometries for the anode segmentation are under study \cite{1748-0221-8-07-P07002}.
+
+Argon ionization and scintillation are the processes leveraged to detect particles in the LArTPC active volume.
+When a ionizing radiation traverses the argon active volume it leaves a trail of ionization electrons along its trajectory and it excites the argon, leading to the production of scintillation light -- details on the production and detection of ionization charge and scintillation light are provided in \ref{sec:light} and \ref{sec:light} respectively.
 The optical part of the detector collects the argon scintillation light in matters of nanoseconds.
 This flash of light determines the start time of an event in the chamber, $t_0$.
-The uniform electric field drifts the electrons from ionization towards the anode plane where the anode sense wires see either an induced current by the drifting charge (on induction planes) or the injection of the ionization charge (collection plane) [94] \footnote{The ionized argon also drifts,  in the opposite directions compared to the electrons and with a much longer drift time, given that the drift time is proportional to the particle mass.
+The uniform electric field drifts the ionization  electrons from the production point towards the anode plane in order of hundreds of microseconds or more depending on the chamber dimensions\footnote{The ionized argon also drifts, but  in the opposite directions compared to the electrons.
+Since the drift time is proportional to the particle mass,  the ions drift time is much longer than the electrons'.
 Ionized argon is collected on the cathode which is not instrumented, so it is not used to infer information about the interactions in the chamber.}.
+The anode sense wires see either an induced current by the drifting charge (on induction planes) or an injection of the ionization charge (collection plane) [94].
+An appropriate choice of the voltage bias on each wire plane assures ideal charge transparency, so that all the ionization charge is collected on the collection plane and none on the induction planes [95].
+
 The arrival time of the charge on the anode sense wires is used to measure the position of the original ionizing radiation in the drift direction.
 In fact, since the constant electric field implies that the drift velocity is also constant in the chamber, the position of the original ionization is simply given by the multiplication of the drift velocity by the drift time, where we define as ``drift time" the difference between $t_0$ and the charge arrival time on the wire planes.
-Combining the spatial information on the different wire planes results in the mapping of a bi-dimensional projection of the interaction pattern in the plane perpendicular to the drift direction.
+The spacial resolution on this dimension is limited by the time resolution of the electronics or by longitudinal diffusion of the electrons.
+The spatial information on the different wire planes maps a bi-dimensional projection of the interaction pattern in the plane perpendicular to the drift direction.
+The spacial resolution on this dimension is limited by the transverse electron diffusion in argon and by the grain of the anode segmentation, i.e.
+the spacing between the wires in the sense planes [100].
+The off-line combination of the 2-D information on the wire planes with the timing information allows for the 3D reconstruction of the event in the chamber.
 
@@
 
-\subsection{Liquid Argon Ionization Charge Detection}\label{sec:charge}
+\subsection{Liquid Argon: Ionization Charge}\label{sec:charge}
 \subsubsection{Electron Life Time \& purity}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Liquid Argon Scintillation Light }\label{sec:light}
+\subsection{Liquid Argon: Scintillation Light }\label{sec:light}
 Liquid argon emits scintillation light at the passage of charged particles.
@@
 \subsection{Signal processing}
-\section{The SBN Program: Neutrino Interaction and Detection}
+
+\section{The Intensity Frontier Program}
+\subsection{SBN: Neutrino Interaction and Detection}
 %\subsection{SBN Goals}
 %\subsection{Neutrino Interactions and Detection }
-\section{DUNE: Rare Decay Searches}
+\subsection{DUNE: Rare Decay Searches}
 The key elements for a rare decay experiment are: massive active volume, long exposure, high identification efficiency and low background. 
@@
 %\subsection{Nucleon Decay Detection in LAr}
-\section{Enabling the next generation of discoveries: LArIAT}
+\subsection{Enabling the next generation of discoveries: LArIAT}
 LArIAT, a small Liquid Argon Time Projection Chamber (LArTPC) in a test beam,  is designed to perform an extensive physics campaign centered on charged particle cross section measurements while characterizing the detector performance for future LArTPCs.

### 2018-04-20 Chapter-2/2-LArTPCs.tex
s

@@
 \chapter{Liquid Argon Detectors at the Intensity Frontier}\label{ch:2}
-In the next few years, LArTPC experiments -- such as the Short-Baseline Neutrino program (SBN) and DUNE -- will  be the tools to answer some of the burning questions in neutrino physics today.
-This section illustrates the operational principles of this detector technology, as well as the scope of SBN, DUNE and LArIAT in the current intensity frontier panorama.
+In the next few years, LArTPCs will  be the tools to answer some of the burning questions in neutrino physics today.
+This section illustrates the operational principles of this detector technology, as well as the scope of the key detectors in the US liquid argon program -- SBN, DUNE and LArIAT.
 
@@
 
-\subsection{Time Projection Chamber}
+\subsection{Argon, Particle Detectors and TPCs}
+
+%% Noble elements in particle phyisics
+
+%% Table with argon properties
+
+%Ionization detectors have been frequently used in Experimental Particle Physics for the possibility to reconstruct particle trajectories and to have calorimetric information with a non-destructive technique.
+
+%Among them the most used have been the proportional counters, as Multi Wire Proportional Chambers (MWPC) and Drift Chambers, in which the charge collected after an ionizing process is directly proportional to the energy deposit of the primary particle that crosses the active volume or interacts there.
+
+%% Why liquid?
+Using nobles elements in the liquid form for neutrino detectors is advantageous with respect to noble gases for several reasons.
+The density of liquids is $\sim$1000 times greater than gases, increasing the number of target centers for neutrino's interaction in the same volume.
+Since the energy loss of charged particle is proportional to the target material density, as shown in the Bethe-Block equation in \ref{eq:BB}, the increased density reflects into a proportionally higher energy loss per particle, enhancing the calorimetry capability of detectors with a liquid active medium.
+Additionally, the ionization energy of liquids is smaller than gasses by the order of tens of eV.
+Thus, at the passage of charged particles, liquid generally produce more ionization electrons than gas for the same deposited energy and force the particles to deposit more energy in a shorter range.
+The downside of using noble liquid gasses in experiments is that they require a cryogenic system to cool the gas until it transitions to its the liquid form.
+
+%%% TPCs
+David Nygren designed the first Time Projection Chamber (TPC) in the late 1970s [15] for the  PEP-4, a detector  apt to study electron-positron collisions at the PEP storage ring at the SLAC National Accelerator Laboratory.
+From the original design  in the seventies -- a cylindrical chamber filled with methane gas -- the TPC detector concept has seen many incarnations, the employment of several different active media and a variety of different particle physics applications, including, but not limited to the study of electron/positron storage rings (e.g.
+PEP4, TOPAZ, ALEPH and DELPHI), heavy ions collisions in fixed target and collider experiments (e.g.
+EOS/HISS and ALICE ), dark matter (ArDM), rare decays and capture (e.g.
+TRIUMP, MuCap),  neutrino detectors and nucleon decay (ICARUS, SBN, DUNE), and neutrino less double beta decay (Next).
+A nice review of the history of TPCs and working principles is provided in \cite{0034-4885-73-11-116201}.
+
+The many useful characteristics of TPC make this technology a very versatile tool for particle detection and explain such a wide popularity.
+TPCs are the only electronically read detector which deliver simultaneous  three-dimensional track information and a measurement of the particle energy loss.
+Leveraging on both tracking and calorimetry,  particle identification (PID) capabilities are enhanced  over a wide momentum range.
+
+Carlo Rubbia first proposed the use of a Liquid Argon TPC for a neutrino experiment, ICARUS [14], in 1977.
+LArTPCs are some times referred as to ``electronic" bubble-camber, for the similar tracking and energy resolution.
+Compared to these historic detectors however, LArTPC bestow tridimensional tracking and a self triggering mechanism provided by the scintillation light in the noble gas.
+
 \subsection{LArTPC: Principles of Operation}\label{sec:LArTPCWorkingPrinciple}
@@
 \subsection{Liquid Argon: Ionization Charge}\label{sec:charge}
+
+\begin{eqnarray*}
+			- \frac{dE}{dx} = K z^2 \frac{Z}{A} \varrho \frac{1}{\beta^2} \left[ \frac{1}{2} \ln{\frac{2 m_e c^2 \beta^2 \gamma^2 T_{max}}{I^2}} - \beta^2 - \frac{\delta}{2}\right] 
+			\label{eq:BB}
+		\end{eqnarray*}
+
 \subsubsection{Electron Life Time \& purity}

### 2018-04-20 Chapter-2/2-LArTPCs.tex
s

@@
 
-\subsection{Argon, Particle Detectors and TPCs}
-
-%% Noble elements in particle phyisics
-
-%% Table with argon properties
-
-%Ionization detectors have been frequently used in Experimental Particle Physics for the possibility to reconstruct particle trajectories and to have calorimetric information with a non-destructive technique.
-
-%Among them the most used have been the proportional counters, as Multi Wire Proportional Chambers (MWPC) and Drift Chambers, in which the charge collected after an ionizing process is directly proportional to the energy deposit of the primary particle that crosses the active volume or interacts there.
-
-%% Why liquid?
-Using nobles elements in the liquid form for neutrino detectors is advantageous with respect to noble gases for several reasons.
-The density of liquids is $\sim$1000 times greater than gases, increasing the number of target centers for neutrino's interaction in the same volume.
-Since the energy loss of charged particle is proportional to the target material density, as shown in the Bethe-Block equation in \ref{eq:BB}, the increased density reflects into a proportionally higher energy loss per particle, enhancing the calorimetry capability of detectors with a liquid active medium.
-Additionally, the ionization energy of liquids is smaller than gasses by the order of tens of eV.
-Thus, at the passage of charged particles, liquid generally produce more ionization electrons than gas for the same deposited energy and force the particles to deposit more energy in a shorter range.
-The downside of using noble liquid gasses in experiments is that they require a cryogenic system to cool the gas until it transitions to its the liquid form.
-
-%%% TPCs
-David Nygren designed the first Time Projection Chamber (TPC) in the late 1970s [15] for the  PEP-4, a detector  apt to study electron-positron collisions at the PEP storage ring at the SLAC National Accelerator Laboratory.
+\subsection{TPCs, Neutrinos \& Argon}
+David Nygren designed the first Time Projection Chamber (TPC) in the late 1970s \textcolor{red}{[15]} for the  PEP-4 experiment, a detector  apt to study electron-positron collisions at the PEP storage ring at the SLAC National Accelerator Laboratory.
 From the original design  in the seventies -- a cylindrical chamber filled with methane gas -- the TPC detector concept has seen many incarnations, the employment of several different active media and a variety of different particle physics applications, including, but not limited to the study of electron/positron storage rings (e.g.
@@
 
-The many useful characteristics of TPC make this technology a very versatile tool for particle detection and explain such a wide popularity.
+Several features of the TPC technology make these detectors a more versatile tool compared to other ionization detectors and explain such a wide popularity.
 TPCs are the only electronically read detector which deliver simultaneous  three-dimensional track information and a measurement of the particle energy loss.
@@
 
-Carlo Rubbia first proposed the use of a Liquid Argon TPC for a neutrino experiment, ICARUS [14], in 1977.
-LArTPCs are some times referred as to ``electronic" bubble-camber, for the similar tracking and energy resolution.
+Carlo Rubbia first proposed the use of a Liquid Argon TPC for a neutrino experiment, ICARUS \textcolor{red}{[14]}, in 1977.
+The active medium in ionization detector has historically been in the gaseous form.
+Nevertheless, using nobles elements in the liquid form for neutrino detectors is advantageous for several reasons.
+The density of liquids is $\sim$1000 times greater than gases, increasing the number of target centers for neutrino's interaction in the same volume.
+Since the energy loss of charged particle is proportional to the target material density, as shown in the Bethe-Block equation in \ref{eq:BB}, the increased density reflects into a proportionally higher energy loss, enhancing the calorimetry capability of detectors with a liquid active medium.
+Additionally, the ionization energy of liquids is smaller than gasses by the order of tens of eV.
+Thus, at the passage of charged particles, liquid generally produce more ionization electrons than gas for the same deposited energy and force the particles to deposit more energy in a shorter range.
+The downside of using noble liquid gasses in experiments is that they require a cryogenic system to cool the gas until it transitions to its the liquid form.
+The properties of liquid argon in comparison liquid xenon -- a popular choice for dark matter and neutrinoless double beta decay detectors -- are summarized in table \ref{tab:properties}.
+Albeit xenon would be more desirable than argon given some superior properties such as lower ionization energy and higher density and light yield, argon relative abundance abates the cost of argon compared to xenon, making argon a more viable choice for the construction of big neutrino detectors.
+
+\begin{table}[]
+\centering
+\begin{tabular}{|l|c|c|}\hline
+Element & LAr & LXe \\
+\hline
+\hline
+Atomic Number &  18 &54 \\
+Atomic weight A & 40  & 131\\
+Boiling Point Tb at 1 atm & 87.3 K & 165.0 K\\
+Density  & 1.4 g/cm$^3$& 3.0 g/cm$^3$\\
+Radiation length  & 14.0 cm& 2.8 cm \\
+Moliere Radius  &10.0 cm& 5.7 cm\\
+Work function  & 23.6 eV&15.6 eV\\
+Electron Mobility at $E_{field} =10^4$ V/m &0.047 m$^2$/Vs& 0.22 m$^2$/Vs\\
+Average dE/dx MIP  & 2.1 MeV/cm&3.8 MeV/cm\\
+Average Scintillation Light Yield & 40000 $\gamma$/MeV&42000 $\gamma$/MeV\\
+Scintillation $\lambda$  &128 nm&175 nm\\
+\hline
+\end{tabular}
+\caption{LAr, LXe summary of properties relevant for neutrino detectors.}
+\label{tab:properties}
+\end{table}
+
+LArTPCs are some times referred as to ``electronic" bubble-chambers, for the similarity in the tracking and energy resolution which is coupled with an electronic readout of the imaging information in LArTPCs.
 Compared to these historic detectors however, LArTPC bestow tridimensional tracking and a self triggering mechanism provided by the scintillation light in the noble gas.
+A comparison between a neutrino interaction in the ArgoNeut LArTPC and in the bubble chamber is shown in picture \textcolor{red}{EVENT DISPLAY}.
 
@@
 Ionized argon is collected on the cathode which is not instrumented, so it is not used to infer information about the interactions in the chamber.}.
-The anode sense wires see either an induced current by the drifting charge (on induction planes) or an injection of the ionization charge (collection plane) [94].
-An appropriate choice of the voltage bias on each wire plane assures ideal charge transparency, so that all the ionization charge is collected on the collection plane and none on the induction planes [95].
+The anode sense wires see either an induced current by the drifting charge (on induction planes) or an injection of the ionization charge (collection plane) \textcolor{red}{[94]}.
+An appropriate choice of the voltage bias on each wire plane assures ideal charge transparency, so that all the ionization charge is collected on the collection plane and none on the induction planes \textcolor{red}{[95]}.
 
@@
 The spacial resolution on this dimension is limited by the transverse electron diffusion in argon and by the grain of the anode segmentation, i.e.
-the spacing between the wires in the sense planes [100].
+the spacing between the wires in the sense planes \textcolor{red}{[100]}.
 The off-line combination of the 2-D information on the wire planes with the timing information allows for the 3D reconstruction of the event in the chamber.
 
-Calorimetry
+Since the charge deposited by the ionizing radiation in the chamber is proportional to the deposited energy and the charge collected on the sense plane is a function of the deposited charge, LArTPC allow the measurement of the energy deposit in the active volume.
+Effects due to the presence of free charge and impurities in the active volume, such as a finite electron lifetime, recombination and space charge, complicate the relationship between deposited and collected charge affecting the measurement of the particle's energy, as described in the next section.
 
@@
 
-\subsubsection{Electron Life Time \& purity}
+\subsubsection{Purity \& Electron Life Time }
+The presence of electronegative contaminants in liquid argon, such as oxygen and water, is particularly
+pernicious, since these molecules quench the charge produced by the ionizing radiation.
+Thus, amount of charge per unit of length $dQ/dx$ collected on the collection plane depends on the charge's production point in the detector: a ionization produced  close to the cathode will see more impurities along its journey to the collection plane than ionization produced close to the anode, resulting in greater attenuation.
+As a result,  the amount of charge collected on the sense wires as a function of the traveled distance follows an exponential decay trend.
+The traveled distance is generally measured in terms of drift time and the  characteristic time constant of the exponential decay is called electron lifetime $\tau_e$.
+Figure \ref{fig:Elifetime} shows the typical life time for LArIAT data \textcolor{red}{cite technote}.
+LArIAT small drift distance (47 cm) allows for a relatively short electron life time.
+The life time for bigger detectors such as MicroBooNE, whose drift distance is 2.5 m, needs to be of the order of  to allow charge collection usable for physics analyses.
+Energy reconstruction in LArTPC applies a correction for the finite lifetime to calibrate the detector calorimetric response; details for LArIAT are provided in Section s\ref{ch:energyCalibration}.
+
 LArTPCs use  hermetically sealed and leak-checked vessels to abate the leakage and diffusion of contaminants into the system.
-The liquid argon filling of the volume occurs after the vessel is evacuated or purged with gaseous argon [205]  to reduce remaining gases in the volume.
+The liquid argon filling of the volume occurs after the vessel is evacuated or purged with gaseous argon \textcolor{red}{[205]} to reduce remaining gases in the volume.
 Even so, the construction of a pure tank of argon is unviable, as several sources of impurity remain.
@@
 The filtration system depends on the size and drift distance of the experiment and, for experiments on several meters scale, it includes an argon recirculation system.
+
+%% Life time value for LArIAT
 
@@
 \subsubsection{Scintillation Process}
-Scintillation light in argon peaks in the ultraviolet at a 128 nm, shown in comparison to Xenon and Kypton in Figure  [183].
+Scintillation light in argon peaks in the ultraviolet at a 128 nm, shown in comparison to Xenon and Kypton in Figure  \textcolor{red}{[183]}.
 The light yield collected by the optical detector depends on the argon purity, the electric field, the dE/dx and particle type, averaging at the tens of thousands of photons per MeV.
@@
 Since the recombination rate is proportional to the square of the local ionization density, highly ionizing particles boost recombination and the subsequent light yield compared to MIPs.
-The possibility to leverage this dependency for pulseshape-based particle identification has been shown in [186], [187].
+The possibility to leverage this dependency for pulseshape-based particle identification has been shown in \textcolor{red}{[186], [187]}.
 
@@
 Rayleigh scattering changes the path of light propagation in argon, prolonging the time between light production and detection.
-The scattering length has been measured to be 66 cm [191] , shorter than the theoretical prediction of $\sim$~90 cm [190]; this value is short enough to be relevant for the current size of LArTPCs detectors.
+The scattering length has been measured to be 66 cm \textcolor{red}{[191]} , shorter than the theoretical prediction of $\sim$~90 cm \textcolor{red}{[190]}; this value is short enough to be relevant for the current size of LArTPCs detectors.
 In fact, Rayleigh scattering worsen the resolution on $t_0$, the start time for charge drifting, and  alters the light directionality, complicating the matching between light and charge coming from the same object in case of multiple charged particles in the detector.
@@
 
-The stringent constraints for the electron life time limit the presence of oxygen and water to such a low level that both absorption and quenching on these impurity is not expected to be significant [210].
+The stringent constraints for the electron life time limit the presence of oxygen and water to such a low level that both absorption and quenching on these impurity is not expected to be significant \textcolor{red}{[210]}.
 Contrarily, the nitrogen level is not bound by the electron life time constraints -- nitrogen being an inert gas, expensive to filter.
@@
 Albeit deployed in different ways, neutrinos and dark matter experiments commonly use  1,1,4,4-tetraphenyl-butadiene (TPB) to shift the scintillation light.
-TPB, whose chemical structure is shown in figure , absorbs the vacuum ultraviolet (VUV) light and emits in the visible at $\sim$~425 nm [231], with a ratio of visible photon emitted per VUV photon absorbed of $\sim$1.2:1 \cite{GEHMAN2011116}.
+TPB, whose chemical structure is shown in figure , absorbs the vacuum ultraviolet (VUV) light and emits in the visible at $\sim$~425 nm \textcolor{red}{[231]}, with a ratio of visible photon emitted per VUV photon absorbed of $\sim$1.2:1 \cite{GEHMAN2011116}.
 

### 2018-04-21 Chapter-2/2-LArTPCs.tex
s

@@
 The life time for bigger detectors such as MicroBooNE, whose drift distance is 2.5 m, needs to be of the order of  to allow charge collection usable for physics analyses.
-Energy reconstruction in LArTPC applies a correction for the finite lifetime to calibrate the detector calorimetric response; details for LArIAT are provided in Section s\ref{ch:energyCalibration}.
+Energy reconstruction in LArTPC applies a correction for the finite lifetime to calibrate the detector calorimetric response; details for LArIAT are provided in Section \ref{ch:energyCalibration}.
 
@@
 
+\subsubsection{Recombination Effect}
+After production, ionization electrons thermalize with the surrounding medium and may recombine with nearby ions.
+Recombination might occur either between the electron and the parent ion through Coulomb attraction (Onsager geminate theory  [2] ) or thanks to the collective charge density of electrons and ions from multiple ionizations in a cylindrical volume surrounding the particle trajectory (Jaffe[3]  columnar model ).
+Consideration on the  average electron-ion distance and the average ion-ion distance for argon show that the probability of geminate recombination is low; thus recombination in argon is mainly due to collective effects\cite{1748-0221-8-08-P08005}.
+Since protons, kaons and stopping particles present a higher ionization compared to MIPs, recombination effects are more prominent when considering the reconstruction of energy deposited by these particles.
+
+Models for a theoretical descriptions of recombination base on are provided in (Birks model [6]) and in (Box model [7]).
+The Birks model assumes a gaussian spatial distribution around the particle trajectory during the entire recombination phase and identical charge mobility for ions and electrons.
+The Box model also assumes that electron diffusion and ion mobility are negligible in liquid argon during recombination
+In these models, the fraction of ionization electrons surviving recombination is a function of the number of ion-electron pairs per unit length, the electric field, the average ion-electron separation distance after thermalization and the angle of the particle with respect to the direction of the electric field -- plus the diffusion coefficient in the Birks model.
+Given the stringent assumptions, it is perhaps  not surprising that these models are in accordance to data only in specific regimes: the Birks model is generally used to describe recombination for low dE/dx, the Box model for high dEdX.
+In LArTPC, the ICARUS and ArgoNeut have measured recombination in [8] and \cite{1748-0221-8-08-P08005} respetively.
+Since LArIAT uses the refurbished ArgoNeut TPC and cryostat,  LArIAT currently corrects for recombination using the ArgoNeut measurement, shown in figure \textcolor{red}{find figure}.
+
 \subsubsection{Space Charge Effect}
-\subsubsection{Recombination Effect}
+Slow-moving positive argon ions created during ionization can build-up in LArTPC, causing the distortion of the
+electric field within the detector.
+This effect, called  ``space charge effect" leads to a displacement in the reconstructed position of the signal ionization electrons.
+In surface LArTPCs the space charge effect is primarily due to the rate of ionization produced by cosmic rays which is slowly drifting in the chamber at all times.
+Surface LArTPC of the size of several meters are expected to be modestly impacted from the space charge effect, where charge build-up create anisotropy of the electric field magnitude of the order of  5\% at a drift field of 500 V/cm.
+The smallness of the LArIAT drift volume is such that  effect of  space charge on the electric field is expected to b even smaller.
+\textcolor{red}{CHIEDI A FLAVIO}
 

### 2018-04-22 Chapter-2/2-LArTPCs.tex
s

@@
 
-\subsection{Signal processing}
+\subsection{Signal Processing and Event Reconstruction}
+In this section we illustrate the processing and reconstruction chain of the TPC signals, from the pulses on the sense wire to the construction of three dimensional objects with associated calorimetry.\\
 
+\textbf{Deconvolution.} Induction and collection planes have different field responses, given the different nature of the signals on these planes: the wires on the induction planes see inductive signal of the drifting charge, while the wires on the collection planes see the current derived from the charge entering the conductor.
+Thus, signals on the induction plane are bi-polar pulse and signal on the collection plane are unipolar pulses, see Figure \textcolor{red}{ADD PULSES FIGURE}.
+The first step in signal processing is deconvolution, that is a series of off-line algorithms geared towards undoing the detector effects.
+The result of the deconvolution step is  the production of  a comparable set waveforms on all planes presenting unipolar, approximately gaussian-like pulses.
+Signal from all planes are treated on equal footage beyond this point.
+Some LArTPC apply noise filtering in the frequency domain just after the deconvolution to clean up wire cross talk.
+Since signals from the LArIAT TPC are extremely clean, noise filtering is not necessary.\\
+
+\textbf{Hit Reconstruction.} The second stage of the signal processing is the reconstruction of hits, indicating an energy deposition in the detector.
+A peak finder scans the deconvolved TPC waveforms for each wire on the whole readout time looking for spikes  above  the waveform's baseline.
+It then fits these peaks with gaussian shapes and stores the fit parameters such as the quality of the fit, the peak time, height and area under the gaussian fit.
+The information resulting fron this process a single spike form a single reconstructed ``hit".
+The next steps in the event reconstruction chain will then decide if rejecting hits with poor fits.
+It is important to notice how the height and width of the hit depend on the topology of the event: for example, a particle running  parallel to the wire planes will leave a series of sharp hits on many consecutive wires, while a particle traveling towards the planes will leave a long, wide hit on very few wires.
+The height of the hits and their integral is proportional to the charge collected on the wire, so it depends on the particle type.
+
+The event reconstruction chain uses collection of hits to form more complex objects associated with the particles in the detector.
+The development of different approaches to accomplish this task is an extremely hot topic in LArTPC event reconstruction which spans from more traditional approaches such as line-clustering  \cite{MicroBooNE} \textcolor{red}{argoneut paper} to the use of machine learning tools \cite{MicroBooNE} \textcolor{red}{KAZU'S paper}.
+Generally speaking, the scope of hit clustering and event reconstruction to provide shower-like or track like-objects with an associated energy reconstruction.
+This is because different particles have different topology in the detector -- electrons and photon create electromagnetic showers,  resulting in shower-like topologies, while muons and hadrons  leave track-like signals.
+For the scope of these thesis, we will describe only LArIAT's approach to track reconstruction even if we recognize the breath of LArTPC event reconstruction is much wider.
+We are interested in the reconstruction of pions and kaons in the active volume, whose topology is track-like.
+
+\textbf{Track Reconstruction} 
 \section{The Intensity Frontier Program}
@@
 \subsection{DUNE: Rare Decay Searches}
+DUNE will achieve a wide non-accelerator physics program.
 The key elements for a rare decay experiment are: massive active volume, long exposure, high identification efficiency and low background. 

### 2018-04-22 Chapter-2/2-LArTPCs.tex
s

@@
 For the scope of these thesis, we will describe only LArIAT's approach to track reconstruction even if we recognize the breath of LArTPC event reconstruction is much wider.
-We are interested in the reconstruction of pions and kaons in the active volume, whose topology is track-like.
+We are interested in the reconstruction of pions and kaons in the active volume, whose topology is track-like.\\
 
-\textbf{Track Reconstruction} 
+\textbf{2D Clustering Reconstruction.} 
+The LArIAT reconstruction of track-like objects starts by clustering hits on the collection and induction planes separately with the use of the TrajCluster clustering package\cite{Bruce} \textcolor{red}{Bruce}.
+TrajCluster looks for a collection of hits in the wire-time 2D space which can be described with a line-like 2D trajectory.
+TrajCluster reconstructs trajectories by adding trajectory points to the leading edge of the trajectory while stepping through the 2D space of hits.
+Several factors determine whether a hit is added to the trajectory, including but not limited to
+\begin{enumerate}
+\item the goodness of the fit of the single hit,
+\item the charge of the hit compared to the average charge and RMS of the hits already forming the trajectory,
+\item the goodness of trajectory fit with and without the hit addition,
+\item the angle between the two lines formed by the collection of hits before and after the considered hit in the trajectory.
+\end{enumerate}
+
+\textbf{TrackReconstruction.}
+
 \section{The Intensity Frontier Program}

### 2018-04-23 Chapter-2/2-LArTPCs.tex
s

@@
 \subsection{Liquid Argon: Ionization Charge}\label{sec:charge}
-
+The mean rate of energy loss by moderately relativistic elementary charge particles heavier than electrons in the region $0.1 < \frac{\beta}{\sqrt{1-\beta^2}} < 1000$ is well described by the modified Bethe-Bloch \cite{BB} equation
 \begin{eqnarray*}
-			- \frac{dE}{dx} = K z^2 \frac{Z}{A} \varrho \frac{1}{\beta^2} \left[ \frac{1}{2} \ln{\frac{2 m_e c^2 \beta^2 \gamma^2 T_{max}}{I^2}} - \beta^2 - \frac{\delta}{2}\right] 
+			- \frac{dE}{dx} = K z^2 \frac{Z}{A} \varrho \frac{1}{\beta^2} \left[ \frac{1}{2} \ln{\frac{2 m_e c^2 \beta^2 \gamma^2 T_{max}}{I^2}} - \beta^2 - \frac{\delta}{2}\right] ,
 			\label{eq:BB}
 \end{eqnarray*}
+where $K = 0.307 075 \text{ MeV g}^1\text{ cm}^2$ is a numerical conversion, $z$ is the number of unit charge of the ionizing radiation, $Z$, $A$  and $\varrho$ are the atomic number, mass number and density of the medium,  $m_e$  is the electron mass, $\gamma = \frac{\beta}{\sqrt{1-\beta^2}} $ is the Lorentz factor of the ionizing radiation,  $T_{max}$ is the maximum kinetic energy which can be
+imparted to a free electron in a single collision, $I$ is the mean excitation energy on eV and  $\delta$ is the  density correction.
+The Bethe-Bloch treats as an uniform and continuous process the energy loss by an ionizing radiation via quantum-mechanical collisions producing ionization or an excitation in the medium.
+The density correction  terms becomes relevant for incident particle with high energy, where screening effects due to the polarization of the medium by high energy particles occur.
+
+Excitation and ionization of the detector medium occur in similar amounts.
+Since the ionizing collisions occur randomly, we can parametrize their number $k$ in a segment of length $s$ along the track  with a Poissonian function
+\begin{equation}
+P(k) = \frac{s^k}{k! \lambda^k} e^{-s/ \lambda }, 
+\end{equation}
+where $\lambda = 1/N_{e}\sigma_i$, with $N_{e}$ being the electron density of $\sigma_i$ the ionization cross-section per electron.
+About 66\% of the ionizing collisions in Argon produce only a single electron/ion pair \cite{0034-4885-73-11-116201}; in the other cases, the transferred kinetic energy is enough for the primary electron to liberate one or more secondary electrons, which usually stay close to the original pair.
+Occasionally, electrons can receive enough energy to be ejected with high energy, forming so-called ``$\delta$-ray": a detectable ionization short track off the particle trajectory, as shown in figure \textcolor{red}{ADD figure}.
+The average number of $\delta$-ray  with energy E$>$E$_0$ per cm follows the empirical form
+\begin{equation}
+P(E>E_0) \sim \frac{y}{\beta^2 E_0},
+\end{equation}
+where $y$ is an empirical factor depending on the medium (0.114 for  gaseous Ar), and $\beta$ is $v/c$.
 
@@
 
-\textbf{TrackReconstruction.}
+\textbf{3D TrackReconstruction.} The 3D tracking set of algorithms uses clusters close in time on the induction and collection planes as starting point to form a 3D track.
+Firstly, it construct a tentative 3D trajectory using the edges of the clusters.
+Then, it  projected back the tentative trajectory on to the planes and adjusts the parameters of the 3D track fit such that they minimize the distance between the fit projections and the track hits in all wire planes simultaneously.
+The track algorithm can use multiple clusters in one plane, but it can never break them in smaller groups of hits.
+This algorithm was first developed for the ICARUS collaboration\cite{Antonello2013}.
 
@@
 It is particularly important to notice that the kaon energy for the proton decay mode $p \rightarrow K^+ \bar{\nu}$ is under Cherenkov threshold.
-Super-Kamiokande set the limit on the lifetime for the $p \rightarrow K^+ \bar{\nu}$ mode by  relying exclusively on photons from nuclear de-excitation.
+Super-Kamiokande set the limit on the lifetime for the $p \rightarrow K^+ \bar{\nu}$ mode by  relying  on photons from nuclear de-excitation and on the muon tagging in the kaon decay leptonic mode.
 For this reason, an attractive alternative approach to identifying nucleon decay is the use of a Liquid Argon Time Projection Chamber (LArTPC).

### 2018-04-23 Chapter-1/1-Theory.tex
s

@@
 The determination of the flavor of a neutrino dynamically arises from the corresponding charged lepton associated in a change current interaction; for example, a $\nu_e$ is a neutrino which produces an $e^-$, a $\bar\nu_\mu$ is a neutrino which produces a $\mu^+$, $etc$.
-The neutrino flavor eigenstates $\ket{\nu_\alpha}$,  with $\alpha = e,\mu,\tau$, are orthogonal to each other and form a base for the the weak interaction matrix.
+The neutrino flavor eigenstates $\ket{\nu_\alpha}$,  with $\alpha = e,\mu,\tau$, are orthogonal to each other and form a base for the weak interaction matrix.
 
@@
 \end{equation}
-is the Jarlskog invariant \cite{Jarlskog1985}, $L$ the neutrino baseline and $\Delta_{ab}$ a factor proportional to the sign and magnitude of the mass splitting. 
+is the Jarlskog invariant \cite{Jarlskog1985}, $L$ the neutrino baseline, i.e.
+the distance between the neutrino production and detection points, and $\Delta_{ab}$ a factor proportional to the sign and magnitude of the mass splitting.
 From these equations, it is clear how the relative  large value of $\theta_{13}$ is a happy accident necessary not to completely suppress the sensitivity to CP violation.
@@
 The precise determination of $\theta_{23}$ is often referred as to ``the octant problem".
-Current experimental results \textcolor{red}{cite NOVA T2K} are consistent with  $\theta_{23}=45^\circ$, which would imply maximal mixing between $\nu_\mu$ - $\nu_\tau$, hinting to an intriguing new symmetry.
-Therefore, a precise measurement of $\theta_{23}$  is of great interest for theoretical models of quark-lepton universality \textcolor{red}{change cit [59,84,85,86,87,88]}, whose  quark and lepton mixing matrices are proportional to the deviation of $\theta_{23}$ from $45^\circ$.
+Current experimental results \cite{PhysRevLett.118.151801,PhysRevLett.118.231801} are consistent with  $\theta_{23}=45^\circ$, which would imply maximal mixing between $\nu_\mu$ - $\nu_\tau$, hinting to an intriguing new symmetry.
+Therefore, a precise measurement of $\theta_{23}$  is of great interest for theoretical models of quark-lepton universality \cite{PhysRevLett.93.161801,PhysRevD.70.073009,0295-5075-103-2-21001}, whose  quark and lepton mixing matrices are proportional to the deviation of $\theta_{23}$ from $45^\circ$.
 
 \textbf{Neutrino mass hierarchy.} The  ``mass hierarchy" problem refers to the unknown ordering of the value of absolute mass of the neutrino mass eigenstates.
-Current oscillation experiments are sensitive only to the magnitude of the mass splitting, but not to its sign.
-\textcolor{red}{cite hints} In a framework where the lightest neutrino mass (arbitrarily) corresponds to the first eigenstate $m_1$, it is unknown whether $m_2 - m_1 < m_3 - m_1$ (Normal Hierarchy) or $m_2 - m_1 > m_3 - m_1$ (Inverted Hierarchy).
+Current oscillation experiments are sensitive only to the magnitude of the mass splitting, and not directly to its sign.
+In a framework where the lightest neutrino mass (arbitrarily) corresponds to the first eigenstate $m_1$, it is unknown whether $m_2 - m_1 < m_3 - m_1$ (Normal Hierarchy) or $m_2 - m_1 > m_3 - m_1$ (Inverted Hierarchy).
 The mass hierarchy affects not only the sensitivity to CP violation searches in long baseline oscillation experiments, but also the sensitivity to determine whether neutrinos are Majorana particles in neutrinoless double beta decay experiments.
@@
 The  $\mathcal{L}_5$ term would introduce a non-renormalizable term in the lagrangian, since it has dimensions of energy to the fifth power.
-This is not allowed in the SM lagrangian; however, the existence of such terms is plausible if the we consider the SM as an effective theory at low energy, manifestation of the symmetry breaking of a Grand Unified Theory (GUT) at higher energy, and not the definitive theory.
+This is not allowed in the SM lagrangian; however, the existence of such terms is plausible if the we consider the SM as an effective theory at low energy, manifestation of the symmetry breaking of a more general theory at higher energy, e.g.
+a Grand Unified Theory (GUT), and not the definitive theory.
 The mass term in eq \ref{eq:majoMass} implies the neutrino mass to be  $m = \frac{g v^2}{\mathcal{M}}$.
@@
 Observation of the lepton number violating process  $0\nu\beta\beta$  would imply neutrinos have a Majorana component.
-Depending on the mass hierarchy, the theory also predicts  $0\nu\beta\beta$  exclusion regions and confirmation of the sole Dirac component for neutrinos\textcolor{red}{find CIT}.\\
+Depending on the mass hierarchy, the theory also predicts  $0\nu\beta\beta$  exclusion regions and confirmation of the sole Dirac component for neutrinos \cite{DellOro2016}.\\
 
@@
 As the smallness of neutrino masses suggests the existence of a higher mass scale, an other, even stronger, hint to Grand Unification comes from the slope of running of the coupling constants.
-The coupling constants for the electromagnetic, weak and strong interactions in the SM vary as a function of the interaction energy as shown in figure \ref{fig:constants};  they cannot exactly meet under the current experimental constraints, but their trend is interesting enough to push for the construction of theories where perfect unification is achieved through the addition of new particles.
+The coupling constants for the electromagnetic, weak and strong interactions in the SM vary as a function of the interaction energy as shown in figure \ref{fig:constants};  they do not exactly meet under the current experimental constraints, but their trend is interesting enough to push for the construction of theories where perfect unification is achieved through the addition of new particles.
 
@@
 From the phenomenology point of view, SUSY models tend to push the proton decay life time higher by a factor of four, they solve the ``hierarchy problem", and they also predict new channels for the proton decay.
-In particular they predict the presence of kaons in the final product, with a dominant mode of $p\rightarrow\bar\nu K^+$%, since particles from the first generation (e.g.
+In particular they predict the presence of kaons in the final product, with a dominant mode of $p\rightarrow\bar\nu K^+$.
+%, since particles from the first generation (e.g.
 $u$ and $d$) need to decay into second and third generation particles (such as $s$).
-Predictions on the proton decay lifetime depend on the chosen SUSY model; again, some of the predictions are not excluded by the experiments \cite{}.
+Predictions on the proton decay lifetime depend on the chosen SUSY model; again, some of the predictions are not excluded by the experiments \cite{PhysRevD.55.6986, PhysRevD.67.075007,Lopez:1991eca}.
 
@@
 The main goals of these research programs include:
+
 \begin{itemize}
-\item[-] Assessment of the existence of right-handed sterile neutrinos. 
-\item[-] Determination of the sign of $\Delta m^2_{13}$ (or $\Delta m^2_{23}$), i.e., the neutrino mass hierarchy.
-\item[-] Determination of the octant, i.e.  whether $\theta_{23}$ is maximal.
-\item[-] Determination the status of CP symmetry in the lepton sector.
-\item[-] Search for observables predicted by GUTs, such as proton decay.
+\item[-] the assessment of the existence of right-handed sterile neutrinos via the study of accelerator neutrinos on a short baseline (SBN), 
+\item[-] the determination of the sign of $\Delta m^2_{13}$ (or $\Delta m^2_{23}$), i.e., the neutrino mass hierarchy via the study of accelerator neutrinos on a long baseline (DUNE),
+\item[-] the determination of the octant, i.e.  whether $\theta_{23}$ is maximal, via the study of accelerator neutrinos on a long baseline (DUNE),
+\item[-] the determination the status of CP symmetry in the lepton sector, via the study of accelerator neutrinos on a long baseline (DUNE),
+\item[-] the search for observables predicted by GUTs, such as proton decay via the study of non accelerator physics in massive underground detectors (DUNE).
 \end{itemize}
@@
 \subsection{Pion-Argon Total Hadronic Cross Section}
-This section outlines the importance of the pion-argon total hadronic cross section.
-We start by discussing the measurement in the context of neutrino interaction searches.
+This section outlines the importance of the pion-argon total hadronic cross section in the context of the current and upcoming liquid argon neutrino experiments, SBN and DUNE.
+%We start by discussing the measurement in the context of neutrino interaction searches.
 % and of light mesons interaction with nuclei studies.
- We then describe the signal signature and historical measurements of pion-nucleus cross section, as well as the implementation of this cross sections in the current version of the simulation package used by LArIAT.
+ We describe the signal signature and historical measurements of pion-nucleus cross section, as well as the implementation of this cross sections in the current version of the simulation package used by LArIAT.
 
@@
 
+We illustrate here the basis of the GENIE Neutrino Generator~\cite{Andreopoulos:2009rq}, since it is one the most popular event generators for LAr experiments.
+
 The key elements of a neutrino event generators for resonance and DIS events are the nuclear model and the hadron treatment (both production and transportation).
-As one of the most popular event generators for LAr experiments, we take GENIE~\cite{Andreopoulos:2009rq} here as an example.
-The nuclear model used by GENIE for all processes is a relativistic Fermi gas (RFG) modified to incorporate short range nucleon-nucleon correlations \cite{PhysRevD.24.1400}.
+We illustrate here the conceptual basis of the GENIE Neutrino Generator~\cite{Andreopoulos:2009rq} as an example, since GENIE is one the most popular event generators for LAr experiments.
+For example, the nuclear model used by GENIE for all processes is a relativistic Fermi gas (RFG) modified to incorporate short range nucleon-nucleon correlations \cite{PhysRevD.24.1400}.
 This means that the initial momentum and binding energy of the struck nucleon is determined by assuming nucleons inside the nucleus are quasi-free, acting independently in the mean field of the nucleus.
@@
 
-%\subsubsection{$\pi^{-}$Ar Cross Section in the Context of Light Mesons Interaction with Nuclei}
-\subsubsection{Signal Signatures}
+\subsubsection{$\pi^{-}$Ar Hadronic Interaction: Signal Signatures}
 Strong hadronic interaction models \cite{9780198520085,9780471779957} predict the pion interaction processes with argon in the [100 - 1200] MeV energy range. 
@@
 \subsubsection{Previous measurements: Lighter and Heavier Nuclei}
-Many experiments with pion beams have studied the hadronic interaction of pions on light and heavy materials, such as  He, Li, C, Fe, Pb\textcolor{red}{\cite{s}}.
+Many experiments with pion beams have studied the hadronic interaction of pions on light and heavy materials, such as  He, Li, C, Fe, Pb \cite{PhysRevC.14.635}.
 However, data on argon are rare: the total differential hadronic cross section has never been measured before on argon.
@@
 Multiple scattering effect modify the resonance width, which is larger than the natural-decay width.
-As an example of a fairly well studied target, Figure \ref{fig:pi2} reports the negative pion cross section on Carbon for the elastic and reaction\footnote{What we refer as to ``reaction channel" is called ``inelastic" in the paper.} channels, and their sum \cite{BINON1970168}.
+As an example of a fairly well studied target, Figure \ref{fig:pi2} reports the negative pion cross section on Carbon for the elastic and reaction\footnote{ This paper calls  ``inelastic interaction" what we refer as to ``reaction channel".} channels, and their sum \cite{BINON1970168}.
 
@@
 
-\subsubsection{Negative Pion Interaction Cross Section in Generators}
+\subsubsection{Negative Pion Interaction Cross Section in Simulation Packages}
 LArIAT uses Geant4 as the default simulation package.
@@
 
-The work in \cite{Irene} \textcolor{red}{Irene} reviews  the current status of negative and positive pion simulation in Geant4 and GENIE for $^{12}$C, $^{56}$Fe, and $^{40}$Ca.
+The work in \cite{Nutini2015}  reviews  the current status of negative and positive pion simulation in Geant4 and GENIE for $^{12}$C, $^{56}$Fe, and $^{40}$Ca.
 From that work, we report the results for $^{12}$C in Figure \ref{fig:pionC} as it allows a direct comparison between Geant4, GENIE and data.
@@
 As shown in section \ref{ch:GUTsTheory}, almost all Grand Unified Theories predict at some level baryon number violation in the form of nucleon decay on long time-scales.
-Given the impossibility to reach grand unification energy scales with collider experiments ($\sqrt{s} > 10^{15}$ GeV),  an indirect proof of GUTs is needed.
+Given the impossibility to reach grand unification energy scales with collider experiments (Energy Scale $> 10^{15}$ GeV),  an indirect proof of GUTs is needed.
 The experimental observation of nucleon decay may be the only viable way to explore these theories.
@@
 
-\subsubsection{Signal Signatures}\label{sec:KSignalSignature}
+\subsubsection{K$^{+}$Ar Hadronic Interaction: Signal Signatures}\label{sec:KSignalSignature}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% PRETTY EVENT DISPLAY WITH TEXT, NOT SURE IF USEFUL
-%\begin{figure}[h!]
-%\centering
-%\includegraphics[width=6.5in]{Chapter-1/Images/KLariat.png}
-%\caption{LArIAT Data $K^+$ candidate.
-$K^+$ enters TPC, undergoes a hadronic scatter, and then decays into $\pi^+$ and $\pi^0$.
-The the $\pi^0$ decays into 2 photons while the $\pi^+$ stops quickly in the TPC.
-Collection plane view.}
-%\label{Fig:KLariat}
-%\end{figure}
-
-%Fig \ref{Fig:KLariat} shows a $K^+$ candidate event in the LArIAT TPC.
-Following the kaon candidate track from left to right, two important elements are visible by eye: a change in the K momentum due to hadronic scatter and a Bragg peak by the end of the track due to an augment of ionization as the kaon slows down in the TPC.
-The track "kink" is only visible thanks to the millimetric spacial resolution of the TPC, while the Bragg shows the calorimetric power of this technology.
-The kaon in this event decays hadronically into $\pi^+$ and $\pi^0$.
-The the $\pi^0$ decays into 2 photons while the $\pi^+$ stops quickly in the TPC.
-The ability to distinguish the topology of this decay from the most frequent one, i.e.
-$K^+\rightarrow\mu^+\nu$, remarks the versatility of the LArTPC technology.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
  \subsubsection{Kaon Interaction Cross Section for thin target in Geant4}

### 2018-04-23 Chapter-1/1-Theory.tex
s

@@
 
-A summary of the pion final states in order of pion multiplicity is given in table \ref{tab:pionFinal}.
+A summary of the pion final states in order of pion multiplicity for the reaction channel is given in table \ref{tab:pionFinal}.
 Pion capture and pion decay at rest dominate the cross section under 100 MeV.
@@
 %}
-\caption{Summary of negative pion hadronic interactions as a function of the pion multiplicity in the final state in the energy range [100-1200] MeV. }
+\caption{Summary of negative pion hadronic interactions of the reaction channel as a function of the pion multiplicity in the final state in the energy range [100-1200] MeV. }
 \label{tab:pionFinal}
@@
 From that work, we report the results for $^{12}$C in Figure \ref{fig:pionC} as it allows a direct comparison between Geant4, GENIE and data.
-Geant4 predictions for $\pi^-$  on Carbon are in good agreement with data over all the spectrum, while GENIE predictions seem to show some features at around 500 MeV and 900 MeV, maybe due ti higher resonances in the hA model.
-From the same work, we also report the negative pion cross section on $^{40}$Ca in Figure \ref{fig:pionCa}, since the this is the nuclear medium closest argon with some available data.
-The predictions from both Geant4 and GENIE agree with data in the high energy region.
-The predictions diverge in the resonance region, where data is not available.
+Geant4 predictions for $\pi^-$  on Carbon are in good agreement with data over all the spectrum, while GENIE predictions seem to show some features at around 500 MeV and 900 MeV, maybe due to higher resonances in the hA model.
+From the same work, we also report the negative pion cross section on $^{40}$Ca in Figure \ref{fig:pionCa}, since this is the nuclear medium closest argon with some available data.
+The predictions from both Geant4 and GENIE agree with data in the high energy region; the Geant4 and GENIE predictions diverge in the resonance region, where data is not available.
+These few examples highlight how cross section data for the specific nucleus considered in the neutrino experiments is fundamental to inform the Monte Carlo simulation.
 

### 2018-04-23 Chapter-2/2-LArTPCs.tex
s

@@
 \subsection{TPCs, Neutrinos \& Argon}
-David Nygren designed the first Time Projection Chamber (TPC) in the late 1970s \textcolor{red}{[15]} for the  PEP-4 experiment, a detector  apt to study electron-positron collisions at the PEP storage ring at the SLAC National Accelerator Laboratory.
+David Nygren designed the first Time Projection Chamber (TPC) in the late 1970s~\cite{FirstTPC} for the  PEP-4 experiment, a detector  apt to study electron-positron collisions at the PEP storage ring at the SLAC National Accelerator Laboratory.
 From the original design  in the seventies -- a cylindrical chamber filled with methane gas -- the TPC detector concept has seen many incarnations, the employment of several different active media and a variety of different particle physics applications, including, but not limited to the study of electron/positron storage rings (e.g.
@@
 
-Carlo Rubbia first proposed the use of a Liquid Argon TPC for a neutrino experiment, ICARUS \textcolor{red}{[14]}, in 1977.
-The active medium in ionization detector has historically been in the gaseous form.
-Nevertheless, using nobles elements in the liquid form for neutrino detectors is advantageous for several reasons.
+Historically, the active medium in ionization detectors has been in the gaseous form.
+Carlo Rubbia first proposed the use of a Liquid Argon TPC for a neutrino experiment, ICARUS \cite{Rubbia:1977zz}, in 1977.
+Using nobles elements in the liquid form for neutrino detectors is advantageous for several reasons.
 The density of liquids is $\sim$1000 times greater than gases, increasing the number of target centers for neutrino's interaction in the same volume.
@@
 Thus, at the passage of charged particles, liquid generally produce more ionization electrons than gas for the same deposited energy and force the particles to deposit more energy in a shorter range.
-The downside of using noble liquid gasses in experiments is that they require a cryogenic system to cool the gas until it transitions to its the liquid form.
+The downside of using noble liquid gasses in experiments is that they require expensive cryogenic systems to cool the gas until it transitions to its the liquid form.
 The properties of liquid argon in comparison liquid xenon -- a popular choice for dark matter and neutrinoless double beta decay detectors -- are summarized in table \ref{tab:properties}.
-Albeit xenon would be more desirable than argon given some superior properties such as lower ionization energy and higher density and light yield, argon relative abundance abates the cost of argon compared to xenon, making argon a more viable choice for the construction of big neutrino detectors.
+Albeit xenon would be more desirable than argon given some superior properties such as lower ionization energy and higher density and light yield, argon relative abundance abates the cost of argon compared to xenon, making argon a more viable choice for the construction of kilo-ton scale neutrino detectors.
 
@@
 Compared to these historic detectors however, LArTPC bestow tridimensional tracking and a self triggering mechanism provided by the scintillation light in the noble gas.
-A comparison between a neutrino interaction in the ArgoNeut LArTPC and in the bubble chamber is shown in picture \textcolor{red}{EVENT DISPLAY}.
+A comparison between a neutrino interaction in the ArgoNeut LArTPC and in the bubble chamber is shown in picture \ref{fig:NuEvd}.
+\begin{figure}[hbpt]
+\centering
+\includegraphics[width=6.5in]{Chapter-2/Images/run3469_subrun1064_event53223_col.pdf}
+\caption{Proton decay lifetime limits from passed and future experiments.}
+\label{fig:NuEvd}
+\end{figure}
 

### 2018-04-25 Chapter-2/2-LArTPCs.tex
s

@@
 The density of liquids is $\sim$1000 times greater than gases, increasing the number of target centers for neutrino's interaction in the same volume.
-Since the energy loss of charged particle is proportional to the target material density, as shown in the Bethe-Block equation in \ref{eq:BB}, the increased density reflects into a proportionally higher energy loss, enhancing the calorimetry capability of detectors with a liquid active medium.
+Since the energy loss of charged particle is proportional to the target material density, as shown in the Bethe-Block equation, eq.
+\ref{eq:BB}, the increased density reflects into a proportionally higher energy loss, enhancing the calorimetry capability of detectors with a liquid active medium.
 Additionally, the ionization energy of liquids is smaller than gasses by the order of tens of eV.
@@
 Compared to these historic detectors however, LArTPC bestow tridimensional tracking and a self triggering mechanism provided by the scintillation light in the noble gas.
-A comparison between a neutrino interaction in the ArgoNeut LArTPC and in the bubble chamber is shown in picture \ref{fig:NuEvd}.
+An event display of a $\nu_\mu$ CC interaction candidate in the MicroBooNE detector is shown in picture \ref{fig:NuEvd} to display the level of spatial details these detectors a capable of; the color scale of the image is proportional to the energy deposited, hinting to the calorimetry capabilities of the detectors.
 \begin{figure}[hbpt]
 \centering
-\includegraphics[width=6.5in]{Chapter-2/Images/run3469_subrun1064_event53223_col.pdf}
-\caption{Proton decay lifetime limits from passed and future experiments.}
+\includegraphics[width=\textwidth]{Chapter-2/Images/run3469_subrun1064_event53223_col.pdf}
+\caption{Event display of a $\nu_\mu$ CC interaction candidate in the MicroBooNE detector.}
 \label{fig:NuEvd}
@@
 Argon ionization and scintillation are the processes leveraged to detect particles in the LArTPC active volume.
-When a ionizing radiation traverses the argon active volume it leaves a trail of ionization electrons along its trajectory and it excites the argon, leading to the production of scintillation light -- details on the production and detection of ionization charge and scintillation light are provided in \ref{sec:light} and \ref{sec:light} respectively.
-The optical part of the detector collects the argon scintillation light in matters of nanoseconds.
+When a ionizing radiation traverses the argon active volume it leaves a trail of ionization electrons along its trajectory and it excites the argon producing of scintillation light -- details on the production and detection of ionization charge and scintillation light are provided in \ref{sec:light} and \ref{sec:light} respectively.
+The optical the detector sees the argon scintillation light in matters of nanoseconds.
 This flash of light determines the start time of an event in the chamber, $t_0$.
-The uniform electric field drifts the ionization  electrons from the production point towards the anode plane in order of hundreds of microseconds or more depending on the chamber dimensions\footnote{The ionized argon also drifts, but  in the opposite directions compared to the electrons.
-Since the drift time is proportional to the particle mass,  the ions drift time is much longer than the electrons'.
+The uniform electric field drifts the ionization  electrons from the production point towards the anode in order of hundreds of microseconds or more depending on the chamber dimensions\footnote{The ionized argon also drifts, but  in the opposite directions compared to the electrons.
+Since the drift time is proportional to the particle mass,  the ions' drift time is much longer than the electrons'.
 Ionized argon is collected on the cathode which is not instrumented, so it is not used to infer information about the interactions in the chamber.}.
-The anode sense wires see either an induced current by the drifting charge (on induction planes) or an injection of the ionization charge (collection plane) \textcolor{red}{[94]}.
-An appropriate choice of the voltage bias on each wire plane assures ideal charge transparency, so that all the ionization charge is collected on the collection plane and none on the induction planes \textcolor{red}{[95]}.
+The anode sense wires see either an induced current by the drifting charge (on induction planes) or an injection of the ionization charge (collection plane).
+An appropriate choice of the voltage bias on each wire plane assures ideal charge transparency, so that all the ionization charge is collected on the collection plane and none on the induction planes.
 
 The arrival time of the charge on the anode sense wires is used to measure the position of the original ionizing radiation in the drift direction.
-In fact, since the constant electric field implies that the drift velocity is also constant in the chamber, the position of the original ionization is simply given by the multiplication of the drift velocity by the drift time, where we define as ``drift time" the difference between $t_0$ and the charge arrival time on the wire planes.
+In fact, since the constant electric field implies that the drift velocity is also constant, the position of the original ionization is simply given by the multiplication of the drift velocity by the drift time, where the ``drift time" is the difference between $t_0$ and the charge arrival time on the wire planes.
 The spacial resolution on this dimension is limited by the time resolution of the electronics or by longitudinal diffusion of the electrons.
@@
 The spacial resolution on this dimension is limited by the transverse electron diffusion in argon and by the grain of the anode segmentation, i.e.
-the spacing between the wires in the sense planes \textcolor{red}{[100]}.
+the spacing between the wires in the sense planes \cite{DERENZO1974319}.
 The off-line combination of the 2-D information on the wire planes with the timing information allows for the 3D reconstruction of the event in the chamber.
 
-Since the charge deposited by the ionizing radiation in the chamber is proportional to the deposited energy and the charge collected on the sense plane is a function of the deposited charge, LArTPC allow the measurement of the energy deposit in the active volume.
+Since the charge deposited by the ionizing radiation is proportional to the deposited energy and the charge collected on the sense plane is a function of the deposited charge, LArTPC allow the measurement of the energy deposit in the active volume.
 Effects due to the presence of free charge and impurities in the active volume, such as a finite electron lifetime, recombination and space charge, complicate the relationship between deposited and collected charge affecting the measurement of the particle's energy, as described in the next section.
@@
 \subsection{Liquid Argon: Ionization Charge}\label{sec:charge}
-The mean rate of energy loss by moderately relativistic elementary charge particles heavier than electrons in the region $0.1 < \frac{\beta}{\sqrt{1-\beta^2}} < 1000$ is well described by the modified Bethe-Bloch \cite{BB} equation
-\begin{eqnarray*}
+The mean rate of energy loss by moderately relativistic elementary charge particles heavier than electrons is well described by the modified Bethe-Bloch \cite{Patrignani:2016xqp} equation
+\begin{eqnarray}
 			- \frac{dE}{dx} = K z^2 \frac{Z}{A} \varrho \frac{1}{\beta^2} \left[ \frac{1}{2} \ln{\frac{2 m_e c^2 \beta^2 \gamma^2 T_{max}}{I^2}} - \beta^2 - \frac{\delta}{2}\right] ,
 			\label{eq:BB}
-\end{eqnarray*}
-where $K = 0.307 075 \text{ MeV g}^1\text{ cm}^2$ is a numerical conversion, $z$ is the number of unit charge of the ionizing radiation, $Z$, $A$  and $\varrho$ are the atomic number, mass number and density of the medium,  $m_e$  is the electron mass, $\gamma = \frac{\beta}{\sqrt{1-\beta^2}} $ is the Lorentz factor of the ionizing radiation,  $T_{max}$ is the maximum kinetic energy which can be
-imparted to a free electron in a single collision, $I$ is the mean excitation energy on eV and  $\delta$ is the  density correction.
+\end{eqnarray}
+where  $z$ is the number of unit charge of the ionizing radiation, $Z$, $A$  and $\varrho$ are the atomic number, mass number and density of the medium,  $m_e$  is the electron mass, $\gamma = \frac{\beta}{\sqrt{1-\beta^2}} $ is the Lorentz factor of the ionizing radiation,  $T_{max}$ is the maximum kinetic energy which can be
+imparted to a free electron in a single collision, $I$ is the mean excitation energy on eV,  $\delta$ is the  density correction and $K = 0.307 075 \text{ MeV g}^{-1}\text{ cm}^2$ is a numerical conversion factor.
 The Bethe-Bloch treats as an uniform and continuous process the energy loss by an ionizing radiation via quantum-mechanical collisions producing ionization or an excitation in the medium.

### 2018-04-26 Chapter-2/2-LArTPCs.tex
s

@@
 About 66\% of the ionizing collisions in Argon produce only a single electron/ion pair \cite{0034-4885-73-11-116201}; in the other cases, the transferred kinetic energy is enough for the primary electron to liberate one or more secondary electrons, which usually stay close to the original pair.
-Occasionally, electrons can receive enough energy to be ejected with high energy, forming so-called ``$\delta$-ray": a detectable ionization short track off the particle trajectory, as shown in figure \textcolor{red}{ADD figure}.
+Occasionally, electrons can receive enough energy to be ejected with high energy, forming so-called ``$\delta$-ray": a detectable ionization short track off the particle trajectory, as shown in figure \ref{fig:delta}.
 The average number of $\delta$-ray  with energy E$>$E$_0$ per cm follows the empirical form
@@
 where $y$ is an empirical factor depending on the medium (0.114 for  gaseous Ar), and $\beta$ is $v/c$.
+
+\begin{figure}[hbpt]
+\centering
+\includegraphics[width=\textwidth]{Chapter-2/Images/Delta.png}
+\caption{Events display for a LArIAT pion absorption candidate on the induction plane, with highlighted delta ray.}
+\label{fig:delta}
+\end{figure}
 
@@
 The traveled distance is generally measured in terms of drift time and the  characteristic time constant of the exponential decay is called electron lifetime $\tau_e$.
-Figure \ref{fig:Elifetime} shows the typical life time for LArIAT data \textcolor{red}{cite technote}.
+Figure \ref{fig:Elifetime} shows the typical life time for LArIAT data.
+The procedure to measure the electron lifetime in LArIAT is outlined in \cite{LArIATLifeTime}.
 LArIAT small drift distance (47 cm) allows for a relatively short electron life time.

### 2018-04-27 Chapter-2/2-LArTPCs.tex
s

@@
 imparted to a free electron in a single collision, $I$ is the mean excitation energy on eV,  $\delta$ is the  density correction and $K = 0.307 075 \text{ MeV g}^{-1}\text{ cm}^2$ is a numerical conversion factor.
-The Bethe-Bloch treats as an uniform and continuous process the energy loss by an ionizing radiation via quantum-mechanical collisions producing ionization or an excitation in the medium.
+The Bethe-Bloch treats the energy loss by an ionizing radiation via quantum-mechanical collisions producing ionization or an excitation in the medium as an uniform and continuous process.
 The density correction  terms becomes relevant for incident particle with high energy, where screening effects due to the polarization of the medium by high energy particles occur.
@@
 pernicious, since these molecules quench the charge produced by the ionizing radiation.
-Thus, amount of charge per unit of length $dQ/dx$ collected on the collection plane depends on the charge's production point in the detector: a ionization produced  close to the cathode will see more impurities along its journey to the collection plane than ionization produced close to the anode, resulting in greater attenuation.
+Thus, amount of charge per unit of length $dQ/dx$ collected on the collection plane depends on the charge's production point in the detector: ionization produced  close to the cathode will see more impurities along its journey to the collection plane than ionization produced close to the anode, resulting in greater attenuation of its charge.
 As a result,  the amount of charge collected on the sense wires as a function of the traveled distance follows an exponential decay trend.
@@
 Energy reconstruction in LArTPC applies a correction for the finite lifetime to calibrate the detector calorimetric response; details for LArIAT are provided in Section \ref{ch:energyCalibration}.
+
+\begin{figure}[hbpt]
+\centering
+\includegraphics[width=\textwidth]{Chapter-2/Images/ELifetime.png}
+\caption{Electron lifetime during the LArIAT run period \cite{detectorPaper}.}
+\label{fig:Elifetime}
+\end{figure}
 

### 2018-04-27 Chapter-2/2-LArTPCs.tex
s

@@
 LArTPCs use  hermetically sealed and leak-checked vessels to abate the leakage and diffusion of contaminants into the system.
-The liquid argon filling of the volume occurs after the vessel is evacuated or purged with gaseous argon \textcolor{red}{[205]} to reduce remaining gases in the volume.
+The liquid argon filling of the volume occurs after the vessel is evacuated or purged with gaseous argon \cite{1748-0221-9-07-P07005} to reduce remaining gases in the volume.
 Even so, the construction of a pure tank of argon is unviable, as several sources of impurity remain.
 In particular, impurities can come from the raw argon supply, the argon filtration system and from the outgassing from internal surfaces.
-Outgassing is a continuous diffusive process  producing contaminants, especially water, even after the vessel is sealed, particularly from materials in the ullage region\footnote{.
-While the liquid argon low temperature reduces outgassing in the liquid, this process remains significant for absorptive material (such as plastic) above the surface of the liquid phase}.
+Outgassing is a continuous diffusive process  producing contaminants, especially water, even after the vessel is sealed, particularly from materials in the ullage region\footnote{While the liquid argon low temperature reduces outgassing in the liquid, this process remains significant for absorptive material (such as plastic) above the surface of the liquid phase.}.
 Since research-grade argon comes from the industrial distillation of air, the impurities with the highest concentration are nitrogen, oxygen and water, generally maintained under the 1 part per million level by the vendor.
@@
 After production, ionization electrons thermalize with the surrounding medium and may recombine with nearby ions.
-Recombination might occur either between the electron and the parent ion through Coulomb attraction (Onsager geminate theory  [2] ) or thanks to the collective charge density of electrons and ions from multiple ionizations in a cylindrical volume surrounding the particle trajectory (Jaffe[3]  columnar model ).
+Recombination might occur either between the electron and the parent ion through Coulomb attraction, according to Onsager geminate theory  [2], or thanks to the collective charge density of electrons and ions from multiple ionizations in a cylindrical volume surrounding the particle trajectory, according to Jaffe[3]  columnar model.
 Consideration on the  average electron-ion distance and the average ion-ion distance for argon show that the probability of geminate recombination is low; thus recombination in argon is mainly due to collective effects\cite{1748-0221-8-08-P08005}.
@@
 
-Models for a theoretical descriptions of recombination base on are provided in (Birks model [6]) and in (Box model [7]).
+Models for a theoretical descriptions of recombination based on the Birks model and the Box model are provided in [6] and  [7] , respectively.
 The Birks model assumes a gaussian spatial distribution around the particle trajectory during the entire recombination phase and identical charge mobility for ions and electrons.
-The Box model also assumes that electron diffusion and ion mobility are negligible in liquid argon during recombination
+The Box model also assumes that electron diffusion and ion mobility are negligible in liquid argon during recombination.
 In these models, the fraction of ionization electrons surviving recombination is a function of the number of ion-electron pairs per unit length, the electric field, the average ion-electron separation distance after thermalization and the angle of the particle with respect to the direction of the electric field -- plus the diffusion coefficient in the Birks model.

### 2018-04-27 Chapter-2/2-LArTPCs.tex
s

@@
 LArIAT small drift distance (47 cm) allows for a relatively short electron life time.
-The life time for bigger detectors such as MicroBooNE, whose drift distance is 2.5 m, needs to be of the order of  to allow charge collection usable for physics analyses.
+The life time for bigger detectors such as MicroBooNE, whose drift distance is 2.5 m, needs to be of the order of tens of milliseconds to allow charge collection usable for physics analyses.
 Energy reconstruction in LArTPC applies a correction for the finite lifetime to calibrate the detector calorimetric response; details for LArIAT are provided in Section \ref{ch:energyCalibration}.
@@
 After production, ionization electrons thermalize with the surrounding medium and may recombine with nearby ions.
-Recombination might occur either between the electron and the parent ion through Coulomb attraction, according to Onsager geminate theory  [2], or thanks to the collective charge density of electrons and ions from multiple ionizations in a cylindrical volume surrounding the particle trajectory, according to Jaffe[3]  columnar model.
+Recombination might occur either between the electron and the parent ion through Coulomb attraction, as described in the geminate theory  \cite{PhysRev.54.554}, or thanks to the collective charge density of electrons and ions from multiple ionizations in a cylindrical volume surrounding the particle trajectory, as described in the  columnar model \cite{Jaff1913}.
 Consideration on the  average electron-ion distance and the average ion-ion distance for argon show that the probability of geminate recombination is low; thus recombination in argon is mainly due to collective effects\cite{1748-0221-8-08-P08005}.
@@
 
-Models for a theoretical descriptions of recombination based on the Birks model and the Box model are provided in [6] and  [7] , respectively.
+Models for a theoretical descriptions of recombination based on the Birks model and the Box model are provided in \cite{0370-1298-64-10-303} and  \cite{PhysRevA.36.614}, respectively.
 The Birks model assumes a gaussian spatial distribution around the particle trajectory during the entire recombination phase and identical charge mobility for ions and electrons.
@@
 Given the stringent assumptions, it is perhaps  not surprising that these models are in accordance to data only in specific regimes: the Birks model is generally used to describe recombination for low dE/dx, the Box model for high dEdX.
-In LArTPC, the ICARUS and ArgoNeut have measured recombination in [8] and \cite{1748-0221-8-08-P08005} respetively.
-Since LArIAT uses the refurbished ArgoNeut TPC and cryostat,  LArIAT currently corrects for recombination using the ArgoNeut measurement, shown in figure \textcolor{red}{find figure}.
+In LArTPC, the ICARUS and ArgoNeut have measured recombination in \cite{Amoruso:2004dy} and \cite{1748-0221-8-08-P08005} respectively.
+Since LArIAT uses the refurbished ArgoNeut TPC and cryostat at the same electric field,  LArIAT currently corrects for recombination using the ArgoNeut measured recombination parameters in \cite{1748-0221-8-08-P08005}.
 
@@
 In surface LArTPCs the space charge effect is primarily due to the rate of ionization produced by cosmic rays which is slowly drifting in the chamber at all times.
-Surface LArTPC of the size of several meters are expected to be modestly impacted from the space charge effect, where charge build-up create anisotropy of the electric field magnitude of the order of  5\% at a drift field of 500 V/cm.
-The smallness of the LArIAT drift volume is such that  effect of  space charge on the electric field is expected to b even smaller.
+Surface LArTPC of the size of several meters are expected to be modestly impacted from the space charge effect, where charge build-up create anisotropy of the electric field magnitude of the order of  5\% at a drift field of 500 V/cm \cite{SpaceCharge}.
+The smallness of the LArIAT drift volume is such that  effect of  space charge on the electric field is expected to be even smaller.
 \textcolor{red}{CHIEDI A FLAVIO}
@@
 \subsubsection{Scintillation Process}
-Scintillation light in argon peaks in the ultraviolet at a 128 nm, shown in comparison to Xenon and Kypton in Figure  \textcolor{red}{[183]}.
+Scintillation light in argon peaks in the ultraviolet at a 128 nm, shown in comparison to Xenon and Kypton in Figure  \ref{fig:ArLight}, from \cite{Morikawa1989}.
 The light yield collected by the optical detector depends on the argon purity, the electric field, the dE/dx and particle type, averaging at the tens of thousands of photons per MeV.
+
+\begin{figure}[hbpt]
+\centering
+\includegraphics[scale=0.5]{Chapter-2/Images/Light.png}
+\caption{Emission spectra of the fast and slow emission components in Xenon, Kypton and Argon according to \cite{Morikawa1989}. The dotted lines correspond to the Gaussian fits. }
+\label{fig:ArLight}
+\end{figure}
+
 The de-excitation of Rydberg dimers in the argon is responsible for the scintillation light. 
@@
 The decay of the triplet is delayed by intersystem crossing, producing a slow component with a time constant of $\sim$~1500 ns.
-``Self-trapped exciton luminescence" and  ``recombination luminescence" are the two processes responsible for the creation of the Rydberg dimers.
+``Self-trapped exciton luminescence" and  ``recombination luminescence" are the two processes responsible for the creation of the Rydberg dimers \cite{Jones:2015bya}.
 In the first process, a charged particle excites an argon atom which becomes self-trapped in the surrounding bulk of argon,  forming a dimer; the dimer is in the singlet state 65\% of the times and in the triplet state 35\% of the times.
@@
 Since the recombination rate is proportional to the square of the local ionization density, highly ionizing particles boost recombination and the subsequent light yield compared to MIPs.
-The possibility to leverage this dependency for pulseshape-based particle identification has been shown in \textcolor{red}{[186], [187]}.
+The possibility to leverage this dependency for pulseshape-based particle identification has been shown in \cite{Boulay:2004dk, PhysRevC.78.035801}.
 

### 2018-04-27 Chapter-9/9-Uncertainty.tex
s

@@
 
-\subsection{Uncertainty on $E_{loss}$}
-We estimate the uncertainty on the energy loss between the beamline momentum measurement and the TPC, $E_{loss}$, using the DDMC pion sample.
+\subsection{$E_{loss}$}
+We derive an estimate of the energy loss between the beamline momentum measurement and the TPC ($E_{loss}$) from the Monte Carlo using the DDMC pion sample, since this quantity is not  measurable directly on data.
 We shoot pions from WC4 with the same momentum distribution as in the beamline data and plot the true $E_{loss}$ for that sample.
+The $E_{loss}$ distribution for the 60A  and 100A pion sample is shown in figure \ref{fig:ELoss60A}, left and right respectively.
+A clear double peaked structure is visible, which is due to the particles either missing or hitting the HALO paddle: a schematic rendering of this occurrence is  shown in figure \ref{fig:Halo}.
+The kinematic at WC4 determine the trajectory of a particle and whether or not it will hit the halo paddle.
+Thus, to
+
+\begin{figure}[hbpt]
+\centering
+\includegraphics[width=0.45\textwidth]{Chapter-9/Images/E_loss60A.png}
+\includegraphics[width=0.45\textwidth]{Chapter-9/Images/E_loss100A.png}
+\caption{True energy loss between WC4 and the TPC front face according to the MC simulation of the 60A runs (left) and of the 100A runs (right).}
+\label{fig:ELoss60A}
+\end{figure}
+
+\begin{figure}[hbpt]
+\centering
+\includegraphics[scale=0.5]{Chapter-9/Images/Halo.png}
+\caption{Schematic rendering of the particle path between WC4 and the TPC front face.
+The paddle with the hollow central circle represents the Halo paddle.
+We illustrate two possible trajectories: in black, a trajectory that miss the paddle and goes through the hole in the Halo, in blue a trajectory that hits the Halo paddle and goes through the scintillation material.}
+\label{fig:Halo}
+\end{figure}
+
 The width of the $E_{loss}$ distribution is the $\delta E_{loss}$.

### 2018-04-27 Chapter-9/9-Uncertainty.tex
s

@@
 A clear double peaked structure is visible, which is due to the particles either missing or hitting the HALO paddle: a schematic rendering of this occurrence is  shown in figure \ref{fig:Halo}.
-The kinematic at WC4 determine the trajectory of a particle and whether or not it will hit the halo paddle.
-Thus, to
+The kinematic at WC4 determines the trajectory of a particle and whether or not it will hit the halo paddle.
+In figure \ref{fig:PxVsXTrue} , we plot the true  $X$ component of the momentum versus the true $X$ position at WC4 for pions missing the halo paddle (left) and for pions hitting the halo paddle (right) for the 60A MC simulation runs -- analogous plots are obtained with the 100A simulation.
+These distributions can be separated drawing a line in this position-momentum space.
+We use a logistic regression  \cite{agresti2013categorical}  as a classifier to find the best separating line, shown in both plots as the red line.
+We classify as ``hitting the halo paddle" all pions whose $P_x$ and $X$ are such that $$P_x +0.02* X - 0.4 < 0 $$ and as ``missing the halo  paddle" all pions whose $P_x$ and $X$ are such that $$P_x +0.02*X - 0.4 > 0, $$ where the coefficients of the line are empirically found by the logistic regression estimation.
+Overall, this simple classifier classifies in the right category (hit or miss) about 86\% of the pion events.
+We apply the same classifier on data.
+We assign  $E_{loss} = 32 \pm 4 $~MeV for events classified as ``hitting the halo paddle"; we assign  $E_{loss} = 24 \pm 3 $~MeV for events classified as ``missing the halo paddle".
+
+%We use the separation of these two distributions to decide what the energy loss for each event on data. 
+%Thus,  we assign the value for energy loss is used in the data.
 
@@
 \includegraphics[width=0.45\textwidth]{Chapter-9/Images/E_loss100A.png}
-\caption{True energy loss between WC4 and the TPC front face according to the MC simulation of the 60A runs (left) and of the 100A runs (right).}
+\caption{True energy loss between WC4 and the TPC front face according to the MC simulation of the 60A runs (left) and of the 100A runs (right).
+The distribution for the whole data sample is shown in blue, the distribution for the pions missing the halo is shown in red, and the distribution for the pions hitting the halo is shown in green.
+}
 \label{fig:ELoss60A}
@@
 
-The width of the $E_{loss}$ distribution is the $\delta E_{loss}$.
+\begin{figure}[hbpt]
+\centering
+\includegraphics[width=\textwidth]{Chapter-9/Images/PXVsX60A.png}
+\caption{Horizontal component of the true momentum vs the horizontal position at WC4 for MC simulated pions of the 60A runs.
+The plot on the left shows the distribution for pion that miss the halo paddle and the plot on the right shows the distributions for pions that hit the halo.
+The form of the classifier is overlaid to both plots (red line).}
+\label{fig:PxVsXTrue}
+\end{figure}
 
-\textcolor{blue}{ TO DO HERE: make sure we have the geometry right, cause otherwise this correction is meaningless.  With this method, so far we get a mean ~40 MeV, but uncertainty ~7MeV. 
-The trajectory method does not improve uncertainty, why? It's a mystery I don't think we should solve before June :) .
-Back of the envelope material budget calculation:}
-\begin{table}[h!]
-\centering
-\caption{Back of the envelope calculation}
-\label{my-label}
-\begin{tabular}{|l|l|l|l|}
-\hline
-dEdx for MIP, MPV {[}MeV cm$^2$/gr{]} & density {[}g/cm$^3${]} & width {[}cm{]} & E$_{loss}$ {[}MeV{]} \\ \hline
-1.6                                                 & 1.7 (G10)                            & 1.3            & 3.5                   \\
-1.6                                                 & 1.4 (LAr)                            & 1.77           & 4.0                   \\
-1.6                                                 & 7.7 (S.S.)                           & 0.23           & 2.8                   \\
-1.6                                                 & 4.5 (Ti)                             & 0.04           & 0.3                   \\ 
-1.6                                                 & 1.03 (Plastic Sci)                   & 1.1            & 1.8                   \\ \hline
-Total                                               &                                      &                & 12.4                  \\ \hline
-\end{tabular}
-\end{table}
-
-\textcolor{blue}{Event taking into account a 3 degree bent, we get 12.41 MeV, which is quite far from 40 MeV... something smells here ;)}
+%\textcolor{blue}{ TO DO HERE: make sure we have the geometry right, cause otherwise this correction is meaningless.  With this method, so far we get a mean ~40 MeV, but uncertainty ~7MeV. 
+%The trajectory method does not improve uncertainty, why? It's a mystery I don't think we should solve before June :) .
+%Back of the envelope material budget calculation:}
+%\begin{table}[h!]
+%\centering
+%\caption{Back of the envelope calculation}
+%\label{my-label}
+%\begin{tabular}{|l|l|l|l|}
+%\hline
+%dEdx for MIP, MPV {[}MeV cm$^2$/gr{]} & density {[}g/cm$^3${]} & width {[}cm{]} & E$_{loss}$ {[}MeV{]} \\ \hline
+%1.6                                                 & 1.7 (G10)                            & 1.3            & 3.5                   \\
+%1.6                                                 & 1.4 (LAr)                            & 1.77           & 4.0                   \\
+%1.6                                                 & 7.7 (S.S.)                           & 0.23           & 2.8                   \\
+%1.6                                                 & 4.5 (Ti)                             & 0.04           & 0.3                   \\ 
+%1.6                                                 & 1.03 (Plastic Sci)                   & 1.1            & 1.8                   \\ \hline
+%Total                                               &                                      &                & 12.4                  \\ \hline
+%\end{tabular}
+%\end{table}
 

### 2018-04-27 Chapter-9/9-Uncertainty.tex
s

@@
 
+\textbf{Systematics}
+Discrepancies between the real TPC geometry and the simulated geometry can lead to a systematic in the $E_{loss}$ calculation.
+In particular, we found a difference in the depth of the un-instrumented argon upstream to the TPC front face, the MC geometry reporting $~\sim 3.3$ cm more un-instrumented argon than the TPC survey.
+For a pion MIP, this depth corresponds to 7.4 MeV which we account for as a double sided systematic in the determination of the pion kinetic energy.
+
 %\textcolor{blue}{ TO DO HERE: make sure we have the geometry right, cause otherwise this correction is meaningless.  With this method, so far we get a mean ~40 MeV, but uncertainty ~7MeV. 

### 2018-04-27 Chapter-2/2-LArTPCs.tex
s

@@
 Rayleigh scattering changes the path of light propagation in argon, prolonging the time between light production and detection.
-The scattering length has been measured to be 66 cm \textcolor{red}{[191]} , shorter than the theoretical prediction of $\sim$~90 cm \textcolor{red}{[190]}; this value is short enough to be relevant for the current size of LArTPCs detectors.
+The scattering length has been measured to be 66 cm \cite{Ishida1997} , shorter than the theoretical prediction of $\sim$~90 cm \cite{Teague1968}; this value is short enough to be relevant for the current size of LArTPCs detectors.
 In fact, Rayleigh scattering worsen the resolution on $t_0$, the start time for charge drifting, and  alters the light directionality, complicating the matching between light and charge coming from the same object in case of multiple charged particles in the detector.
@@
 
-The stringent constraints for the electron life time limit the presence of oxygen and water to such a low level that both absorption and quenching on these impurity is not expected to be significant \textcolor{red}{[210]}.
+The stringent constraints for the electron life time limit the presence of oxygen and water to such a low level that both absorption and quenching on these impurity is not expected to be significant.
 Contrarily, the nitrogen level is not bound by the electron life time constraints -- nitrogen being an inert gas, expensive to filter.
@@
 \subsubsection{Wavelength Shifting of LAr Scintillation Light}
-Liquid argon scintillation light is invisible for most optical detectors deployed in LArTPC, such as cryogenic PMTs and SiPMs, since a wavelength of 128 nm is  generally too short to be absorbed from most in glasses, polymers and semiconductor materials.
+Liquid argon scintillation light is invisible for most optical detectors deployed in a LArTPC, such as cryogenic PMTs and SiPMs, since a wavelength of 128 nm is  generally too short to be absorbed from most in glasses, polymers and semiconductor materials.
 Research on prototype SiPMs absorbing directly VUV light and their deployment in noble gasses experiment is ongoing but not mature \cite{1748-0221-8-01-C01003}.
@@
 Albeit deployed in different ways, neutrinos and dark matter experiments commonly use  1,1,4,4-tetraphenyl-butadiene (TPB) to shift the scintillation light.
-TPB, whose chemical structure is shown in figure , absorbs the vacuum ultraviolet (VUV) light and emits in the visible at $\sim$~425 nm \textcolor{red}{[231]}, with a ratio of visible photon emitted per VUV photon absorbed of $\sim$1.2:1 \cite{GEHMAN2011116}.
+TPB  absorbs the vacuum ultraviolet (VUV) light and emits in the visible at $\sim$~425 nm \cite{Burton1973}, with a ratio of visible photon emitted per VUV photon absorbed of $\sim$1.2:1 \cite{GEHMAN2011116}.
 
-Neutrino experiments typically coat their optical detector system evaporating a layer of TPB either directly on the PMTs glass surface or on acrylic plates mounted in front of the PMTs \textcolor{red}{cite microboone}; this technique allows the fast detection light coming directly from the neutrino interaction.
-Dark matter experiments typically evaporate TPB on reflective foils mounted on the inside walls of the sensitive volume and detect the light after it has been reflected; this technique leads to a higher and more uniform light yield, though scattering effects for both the visible and VUV light augment the propagation time and hinder directionality information\textcolor{red}{some DM detector} .
-In order to take advantage of both these techniques, hybrid systems with PMT coating and foils are being considered for the next generation of large neutrino detectors \textcolor{red}{cite SBND?}.
+Neutrino experiments typically coat their optical detector system evaporating a layer of TPB either directly on the PMTs glass surface or on acrylic plates mounted in front of the PMTs \cite{Acciarri2017}; this technique allows the fast detection light coming directly from the neutrino interaction.
+Dark matter experiments typically evaporate TPB on reflective foils mounted on the inside walls of the sensitive volume and detect the light after it has been reflected; this technique leads to a higher and more uniform light yield, though scattering effects for both the visible and VUV light augment the propagation time and hinder directionality information\cite{Aalseth2018}.
+In order to take advantage of both these techniques, hybrid systems with PMT coating and foils are being considered for the next generation of large neutrino detectors.
 

### 2018-04-28 Chapter-2/2-LArTPCs.tex
s

@@
 The event reconstruction chain uses collection of hits to form more complex objects associated with the particles in the detector.
-The development of different approaches to accomplish this task is an extremely hot topic in LArTPC event reconstruction which spans from more traditional approaches such as line-clustering  \cite{MicroBooNE} \textcolor{red}{argoneut paper} to the use of machine learning tools \cite{MicroBooNE} \textcolor{red}{KAZU'S paper}.
+The development of different approaches to accomplish this task is an extremely hot topic in LArTPC event reconstruction which spans from more traditional approaches such as line-clustering  \cite{MicroBooNE} \textcolor{red}{argoneut paper} to the use of machine learning tools \cite{1748-0221-12-03-P03011}.
 Generally speaking, the scope of hit clustering and event reconstruction to provide shower-like or track like-objects with an associated energy reconstruction.

### 2018-04-28 Chapter-2/2-LArTPCs.tex
s

@@
 \subsection{Signal Processing and Event Reconstruction}
-In this section we illustrate the processing and reconstruction chain of the TPC signals, from the pulses on the sense wire to the construction of three dimensional objects with associated calorimetry.\\
+In this section we illustrate the processing and reconstruction chain of the TPC signals, from the pulses on the sense wire to the construction of three dimensional objects with associated calorimetry.
+Different experiments can chose different software packages for their off line signal processing and event reconstruction, but  a popular choice for  US based  LArTPCs is LArSoft \cite{EricFChurck}.
+Based on the Art framework \cite{Green:2012gv}, LArSoft is an event-based toolkit to perform simulation, analysis and reconstruction of LArTPCs events.\\
+
+LArTPC signal processing develops in several consecutive stages that we summarize here in the following categories: \emph{Deconvolution}, \emph{Hit Reconstruction}, \emph{2D Clustering}, \emph{3D Tracking}, \emph{Calorimetry Reconstruction}.
+A visualization of the signal processing workflow is shown in figure \ref{fig:SignalProc}.\\
+
+\begin{figure}[hbpt]
+\centering
+\includegraphics[width=\textwidth]{Chapter-2/Images/SignalProc.jpg}
+\caption{A scheme of a typical signal processing workflow in LArSoft.}
+\label{fig:SignalProc}
+\end{figure}
 
 \textbf{Deconvolution.} Induction and collection planes have different field responses, given the different nature of the signals on these planes: the wires on the induction planes see inductive signal of the drifting charge, while the wires on the collection planes see the current derived from the charge entering the conductor.
-Thus, signals on the induction plane are bi-polar pulse and signal on the collection plane are unipolar pulses, see Figure \textcolor{red}{ADD PULSES FIGURE}.
+Thus, signals on the induction plane are bi-polar pulse and signal on the collection plane are unipolar pulses, see Figure \ref{fig:SignalProc} panel a.
 The first step in signal processing is deconvolution, that is a series of off-line algorithms geared towards undoing the detector effects.
-The result of the deconvolution step is  the production of  a comparable set waveforms on all planes presenting unipolar, approximately gaussian-like pulses.
+The result of the deconvolution step is  the production of  a comparable set waveforms on all planes presenting unipolar, approximately gaussian-like pulses (Figure \ref{fig:SignalProc} panel b).
 Signal from all planes are treated on equal footage beyond this point.
@@
 It is important to notice how the height and width of the hit depend on the topology of the event: for example, a particle running  parallel to the wire planes will leave a series of sharp hits on many consecutive wires, while a particle traveling towards the planes will leave a long, wide hit on very few wires.
-The height of the hits and their integral is proportional to the charge collected on the wire, so it depends on the particle type.
+The height of the hits and their integral is proportional to the charge collected on the wire, so it depends on the particle type.\\
 
 The event reconstruction chain uses collection of hits to form more complex objects associated with the particles in the detector.
-The development of different approaches to accomplish this task is an extremely hot topic in LArTPC event reconstruction which spans from more traditional approaches such as line-clustering  \cite{MicroBooNE} \textcolor{red}{argoneut paper} to the use of machine learning tools \cite{1748-0221-12-03-P03011}.
+The development of different approaches to accomplish this task is an extremely hot topic in LArTPC event reconstruction which spans from more traditional approaches such as line-clustering  \cite{Barker2011} to the use of machine learning tools \cite{1748-0221-12-03-P03011}.
 Generally speaking, the scope of hit clustering and event reconstruction to provide shower-like or track like-objects with an associated energy reconstruction.
@@
 \textbf{2D Clustering Reconstruction.} 
-The LArIAT reconstruction of track-like objects starts by clustering hits on the collection and induction planes separately with the use of the TrajCluster clustering package\cite{Bruce} \textcolor{red}{Bruce}.
+The LArIAT reconstruction of track-like objects starts by clustering hits on the collection and induction planes separately with the use of the TrajCluster clustering package\cite{Baller2016}. 
 TrajCluster looks for a collection of hits in the wire-time 2D space which can be described with a line-like 2D trajectory.
@@
 
-\textbf{3D TrackReconstruction.} The 3D tracking set of algorithms uses clusters close in time on the induction and collection planes as starting point to form a 3D track.
+\textbf{3D Tracking.} The 3D tracking set of algorithms uses clusters close in time on the induction and collection planes as starting point to form a 3D track.
 Firstly, it construct a tentative 3D trajectory using the edges of the clusters.
@@
 The track algorithm can use multiple clusters in one plane, but it can never break them in smaller groups of hits.
-This algorithm was first developed for the ICARUS collaboration\cite{Antonello2013}.
+This algorithm was first developed for the ICARUS collaboration\cite{Antonello2013}.\\
+
+\textbf{Calorimetry.} The last step in the event reconstruction chain is to assign calorimetric information to the track (or shower) objects.
+Calorimetry is performed separately on the different planes.
+A multi-step procedure is needed to retrieve the energy deposited in the TPC  from the charge seen by the wires.
+For each hit associated with the track object, the calorimetry algorithms calculate the charge seen on every wire using the area underneath the gaussian fit; then they correct this raw charge by the electron life time, the electronic noise on the considered wire and the recombination effect.
+Lastly an overall calibration of the energy, explained in detail in section \ref{ch:energyCalibration}, is applied and the calorimetric information for the given track is assigned.
+Even if calorimetry is done in 2D, it benefits from the 3D tracking information; typical information available after the calorimetric reconstruction are the total energy deposited by the particle and its stopping power $dE/dx$ at each ``track pitch", i.e.
+at each 2D projection on the wire plane of the 3D trajectory.
 
@@
 \centering
-\includegraphics[width=6.5in]{Chapter-2/Images/PDKExperimentalLImit.png}
+\includegraphics[width=\textwidth]{Chapter-2/Images/PDKExperimentalLImit.png}
 \caption{Proton decay lifetime limits from passed and future experiments.}

### 2018-04-28 Chapter-2/2-LArTPCs.tex
s

@@
 \section{The Intensity Frontier Program}
-\subsection{SBN: Neutrino Interaction and Detection}
+\subsection{Prospects for LArTPCs in Neutrino Physics: SBN and DUNE}
 %\subsection{SBN Goals}
 %\subsection{Neutrino Interactions and Detection }
-\subsection{DUNE: Rare Decay Searches}
-DUNE will achieve a wide non-accelerator physics program.
+\subsection{Prospects for LArTPCs in GUT Physics: DUNE}
+The experimental exploration of a manifestation of Grand Unified Theory is possible in DUNE thanks to its sheer huge mass.
+In particular, proton decay searches are a capital topic of DUNE's wide non-accelerator physics program.
 The key elements for a rare decay experiment are: massive active volume, long exposure, high identification efficiency and low background. 
@@
 Historically, the dominant technology used in these searches has been water Cherenkov detectors: all the best experimental limits on every decay mode are indeed set by Super-Kamiokande \cite{PhysRevD.90.072005,PhysRevLett.115.121803}.
-It is particularly important to notice that the kaon energy for the proton decay mode $p \rightarrow K^+ \bar{\nu}$ is under Cherenkov threshold.
-Super-Kamiokande set the limit on the lifetime for the $p \rightarrow K^+ \bar{\nu}$ mode by  relying  on photons from nuclear de-excitation and on the muon tagging in the kaon decay leptonic mode.
-For this reason, an attractive alternative approach to identifying nucleon decay is the use of a Liquid Argon Time Projection Chamber (LArTPC).
-
-LArTPCs can complement nucleon decay searches in modes where water Cherenkov detectors are less sensitive, especially $p\rightarrow K^+\bar{\nu}$.
-According to \cite{Acciarri:Dune}, DUNE will have an active volume large enough, have sufficient shielding from the surface, and will run for lengths of time sufficient to compete with Hyper-K, opening up the opportunity for the discovery of nucleon decay.
+As shown in section \ref{ch:GUTsTheory}, different family of GUTs predict the proton to decay in different modes.
+In particular, SUSY flavored GUTs prefer the presence of kaons in the decay products, e.g.
+$p \rightarrow K^+ \bar{\nu}$.
+It is particularly important to notice that the kaon energy for the proton decay mode $p \rightarrow K^+ \bar{\nu}$ is under Cherenkov threshold in water.
+Thus, Super-Kamiokande set the limit on the lifetime for the $p \rightarrow K^+ \bar{\nu}$ mode by  relying  on photons from nuclear de-excitation and on the muon tagging in the kaon decay leptonic mode.
+For this reason, an attractive alternative approach to identifying nucleon decay is the use of a LArTPCs, where the kaon is directly visible in the detector.
+According to \cite{Adams:2013qkq}, DUNE will have an active volume large enough, have sufficient shielding from the surface, and will run for lengths of time sufficient to compete with Hyper-K, opening up the opportunity for the discovery of nucleon decay.
 
@@
 \caption{Momentum of the kaon outgoing a proton decay event as simulated by the Genie 2.8.10 event generator in argon.
-The red line represent the kaon momentum distribution before undergoing the simulated final state interaction inside the argon nucleus, while the blue line represents the momentum distribution after FSI.
+The red line represents the kaon momentum distribution before undergoing the simulated final state interaction inside the argon nucleus, while the blue line represents the momentum distribution after FSI.
 }
@@
 \subsection{Enabling the next generation of discoveries: LArIAT}
-LArIAT, a small Liquid Argon Time Projection Chamber (LArTPC) in a test beam,  is designed to perform an extensive physics campaign centered on charged particle cross section measurements while characterizing the detector performance for future LArTPCs.
-LArTPC represents one of the most advanced experimental technologies for physics at the Intensity Frontier due to its full 3D-imaging, excellent particle identification and precise calorimetric energy reconstruction.
-This complex technology however needs a thorough calibration and dedicated measurements of some key quantities to achieve the precision required for the next generation of discoveries at the Intensity Frontier which LArIAT can provide.
-
+LArIAT, a small LArTPC in a test beam,  is designed to perform an extensive physics campaign centered on charged particle cross section measurements while characterizing the detector performance for future LArTPCs.
+Since LArTPCs represent the most advanced experiments for physics at the Intensity Frontier, their complex technology needs a thorough calibration and dedicated measurements of some key quantities to achieve the precision required for the next generation of discoveries.
+LArIAT's goal is to provide such calibration and dedicated measurements.
 The LArIAT LArTPC is deployed in a dedicated calibration test beamline at Fermilab.
@@
 The precise calorimetric energy reconstruction of the LArTPC technology enables the measurement of the total differential cross section for  tagged hadrons.
-The Pion-Nucleus and Kaon-Nucleus total hadronic interaction cross section have never been measured before in argon and they are a fundamental step to shed light on light meson interaction in nuclei.
-Additionally, these measures provides a key input to neutrino physics and proton decay studies in future LArTPC experiments like SBN and DUNE.
-\textcolor{red}{add paragraph on all wonderful things lariat can do... some event displays would be nice!}
+The Pion-Nucleus and Kaon-Nucleus total hadronic interaction cross section have never been measured before in argon and they are a fundamental step to shed light on light meson interaction in nuclei per se, while providing a key input to neutrino physics and proton decay studies in future LArTPC experiments like SBN and DUNE.
+
+\textcolor{red}{Add Kaons and Pions spectrum}
 

### 2018-04-29 Chapter-1/1-Theory.tex
s

@@
 Their are the mediator of the interaction that turns quarks into leptons, leading to predict the existence of processes that violate baryon number, such as $p \rightarrow \pi^0  + e^+$  (see fig~ \ref{fig:MandatoryFeynmannDiagrams}, right).
-The prediction for proton decay lifetime, $\tau_p \sim \frac{M^4_X}{m^5_p}\sim10^{30\pm1.5}$~years, is unfortunately experimentally disproved by IMB and Super-K \cite{s}.
+The prediction for proton decay lifetime, $\tau_p \sim \frac{M^4_X}{m^5_p}\sim10^{30\pm1.5}$~years, is unfortunately experimentally disproved by IMB and Super-Kamiokande \cite{s}.
 

### 2018-04-29 Chapter-2/2-LArTPCs.tex
s

@@
 \subsection{Prospects for LArTPCs in Neutrino Physics: SBN and DUNE}
+The ArgoNeut  experiment \cite{ArgoNeuT-det} initiated the US LArTPC neutrino program.
+Following the success of this small TPC on the NuMI beam, a wide program of LArTPCs on neutrino beams has flourished.
+The construction of  LArTPCs as near and far detectors at different baseline allows for the exploration of different fundamental questions in neutrino physics today.
+
+The Short-Baseline Neutrino (SBN) \cite{Antonello:2015lea} program at Fermilab  is tasked with conclusively addressing the ``LSND and MiniBooNE anomalies" \cite{Aguilar:2001ty, Athanassopoulos:1997pv, Aguilar-Arevalo:2013pmq}, whose $\nu_e$ and $\bar\nu_e$ appearance results may be interpreted under the assumption of a new sterile neutrino.
+The SBN program entails three surface LArTPCs positioned on the Booster Neutrino Beam at different distances from the neutrino production in oder to fully exploit  the L/E dependence of the oscillation pattern:  SBND (100 m from the decay pipe), MicroBooNE (450 m), and ICARUS (600 m).
+Within the oscillation context, the choice of the LArTPC technology for the SBN detectors changes the set of systematics with respect to LSND and MiniBooNE, whose detection techniques were both based on Cherenkov light.
+In particular, LArTPCs provide excellent electron/photon separation \cite{Acciarri:2016sli} lacking in Cherenkov detectors which can be leveraged to abate the photon background from neutral current interactions  in $\nu_e$ searches.
+MicroBooNE\cite{MicroBooNE-det}, the first detector of the SBN program to be fully operational, started its first neutrino run in October 2015.
+MicroBooNE is a 89 ton active volume LArTPC, single drift chamber with TPC dimensions of 2.6 m (drift) x 2.3 m (heigh) x 10.4 m (depth).
+MicroBooNE is positioned at a very similar L/E on the Booster neutrino beam as MiniBooNE has the scope to directly cross check the MiniBooNE oscillation measurement.
+In case MicroBooNE confirms the presence of the ``low energy excess" anomaly, SBND and ICARUS will provide the full measurement of the oscillation parameters.
+SBND and ICARUS are both dual drift chambers, whose active volume is respectively 112 ton and 600 ton.
+ICARUS is scheduled to become operational by the end of 2018 and SBND shortly after.
+Besides the oscillation analysis, the second main goals of SBN is to perform an extensive campaign of neutrino cross section measurements in argon.
+Given the importance of nuclear effects in (relatively) heavy materials, both the oscillation analysis of the SBN program and the measurements of neutrino properties in DUNE will benefit from such a campaign.
+
+On a different neutrino beam and baseline, the DUNE  experiment, n\'ee LBNE\cite{Adams:2013qkq},  is the flagship experiment on the medium-long term of US-based neutrino physics, scheduled to start data taking in 2026.
+Shooting neutrinos from Fermilab for 800 miles to the SURF laboratory in South Dakota, DUNE is tasked with preforming conclusive measurements of CP violation in the lepton sector,  the neutrino mass ordering and the $\theta_{23}$ octant.
+The DUNE far detector will count four 10 kton LArTPCs, roughly of dimensions of  19 m (horizontally) x 18 m (vertically) x 66 m (depth).
+
 %\subsection{SBN Goals}
@@
 \subsection{Prospects for LArTPCs in GUT Physics: DUNE}
-The experimental exploration of a manifestation of Grand Unified Theory is possible in DUNE thanks to its sheer huge mass.
+The experimental exploration of a manifestation of Grand Unified Theory is possible in DUNE thanks to its sheer mass.
 In particular, proton decay searches are a capital topic of DUNE's wide non-accelerator physics program.
@@
 
-\begin{figure}[hbpt]
-\centering
-\includegraphics[width=3.5in]{Chapter-2/Images/pdkGenie.png}
-\caption{Momentum of the kaon outgoing a proton decay event as simulated by the Genie 2.8.10 event generator in argon.
-The red line represents the kaon momentum distribution before undergoing the simulated final state interaction inside the argon nucleus, while the blue line represents the momentum distribution after FSI.
-}
-\label{fig:PDKGENIE}
-\end{figure}
-
 %\subsection{Non-Accelerator Physics Program}
@@
 
-\textcolor{red}{Add Kaons and Pions spectrum}
+In order to showcase LArIAT's utility to SBN and DUNE, we illustrate briefly two comparisons as examples: one regarding neutrino interactions and the second regarding proton decay studies.\\
+The left side of figure \ref{fig:NuSimulation} shows the distribution of products in momentum spectrum and particle type as simulated in a $\nu_e$ CC interaction in DUNE (according to \cite{LeiguideOliveira:1953730}) as  a comparative example to the momentum distribution of particles in the LArIAT beamline -- shown on the right side of figure \ref{fig:NuSimulation}.
+The momentum spectrum in the LArIAT beamline for electrons, muons and pions -- the most abundant particles produced in  a $\nu_e$ CC interaction -- covers a wide range of the expected momentum distribution in a neutrino event.
 
-\textcolor{red}{ADD genie proton decay kaon distribution and lariat beamline overlaied}
-The signature of a proton decay event in the ``LAr golden mode" is the presence of a single kaon of about 400 MeV in the detector.
+The signature of a proton decay event in the ``LAr golden mode" is the presence of a single kaon of about 400 MeV in the detector; the momentum spectrum of the kaon pre and post FSI in such an event is shown on the left side of figure \ref{fig:PDKGENIE}.
+The right side of figure \ref{fig:PDKGENIE} shows the momentum spectrum of kaons in the LArIAT beamline.
+Kaons arriving to the LArIAT TPC are ideal for proton decay studies, since their momentum in the beamline is just above the typical momentum for kaons in a proton decay event: the majority of LArIAT kaons slow down in the TPC enough to enter the desired momentum window.
+
+\begin{figure}[hbpt]
+\centering
+\includegraphics[width=0.45\textwidth]{Chapter-2/Images/NueCCSim.png}	
+\includegraphics[width=0.45\textwidth]{Chapter-2/Images/MomentumKaonsDatabeamline.png}
+\caption{$Left$. Simulation of the products of a $\nu_e$ CC interaction in DUNE, both in particles type and momentum. \\
+$Right$. Momentum spectrum for low mass particles ($e,\mu,\pi$) in the LArIAT beamline,  negative tune. Data is shown in black, while the beamline simulation is shown in solid colors. }
+\label{fig:NuSimulation}
+\end{figure}
+
+\begin{figure}[hbpt]
+\centering
+\includegraphics[width=0.45\textwidth]{Chapter-2/Images/pdkGenie.png}	
+\includegraphics[width=0.45\textwidth]{Chapter-2/Images/MomentumKaonsDatabeamline.png}
+\caption{$Left$.
+Momentum of the kaon outgoing a proton decay $p\rightarrow K^+\bar\nu$ event as simulated by the Genie 2.8.10 event generator in argon.
+The red line represents the kaon momentum distribution before undergoing the simulated final state interaction inside the argon nucleus, while the blue line represents the momentum distribution after FSI.\\
+$Right$. Positive Kaon momentum spectrum in the LArIAT beamline. }
+\label{fig:PDKGENIE}
+\end{figure}

### 2018-04-29 Chapter-2/2-LArTPCs.tex
s

@@
 \caption{$Left$. Simulation of the products of a $\nu_e$ CC interaction in DUNE, both in particles type and momentum. \\
-$Right$. Momentum spectrum for low mass particles ($e,\mu,\pi$) in the LArIAT beamline,  negative tune. Data is shown in black, while the beamline simulation is shown in solid colors. }
+$Right$. Momentum spectrum for low mass particles ($e,\mu,\pi$) in the LArIAT beamline,  negative tune, Run II, High Yield see section \ref{sec:MWPCfunc}. }
 \label{fig:NuSimulation}
@@
 The red line represents the kaon momentum distribution before undergoing the simulated final state interaction inside the argon nucleus, while the blue line represents the momentum distribution after FSI.\\
-$Right$. Positive Kaon momentum spectrum in the LArIAT beamline. }
+$Right$. Positive Kaon momentum spectrum in the LArIAT beamline,  positive tune, Run II, Picky Tracks see section \ref{sec:MWPCfunc}. }
 \label{fig:PDKGENIE}

### 2018-04-30 Chapter-5/5-SamplesPreparation.tex
s

@@
 \chapter{Samples Preparation}\label{ch:samples}
-This chapter describes the data and Monte Carlo samples used for the cross section analyses, % as well as the work necessary to prepare the samples in a usable state for the physics analyses.
+This chapter describes the preparation of the data and Monte Carlo samples used for the cross section analyses. This entails:
+\begin{enumerate}
+\item the beamline event selection on data,
+\item the MC production,
+\item the energy calibration of the detector both in data and MC,
+\item the optimization of the tracking algorithm for the total cross section analyses.
+\end{enumerate}
+
 \section{LArIAT Data}
 \section{LArIAT Monte Carlo}
+For the simulation of LArIAT events and their particle make up, we use a combination of two MC generators: the G4Beamline Monte Carlo and the Data Driven single particle Monte Carlo (DDMC).
+We use the G4Beamline MC to simulate the particle transportation in the beamline and calculate the particle composition of the beam just past the fourth Wire Chamber (WC4).
+In order to simulate the beam line particles after WC4 and in the TPC, we use the DDMC.
 \subsection{G4Beamline}\label{beamlineComposition}
+G4Beamline simulates the beam collision with the LArIAT secondary target, the energy deposited by the particles in the LArIAT beamline detectors and the action of the LArIAT magnets, effectively accounting for particle transportation through the beam line from the LArIAT target until ``Big Disk", a fictional, void detector located just before the cryostat.
+ At the moment of this writing, G4Beamline does not simulated the responses of the beam line detectors.
+It is possible to interrogate the truth level information of the simulated particles in several points of the geometry.
+In order to ease the handshake between G4Beamline and the DDMC, we ask for the beam composition just after WC4.
+Since LArIAT data are taken under different beam conditions, G4Beamline simulates separately the beam composition according to the magnets' settings and the secondary beam intensity.
+For the pion cross section analysis the relevant beam conditions are  secondary beam energy of 64 GeV, negative polarity magnet with current of 100 A and 60 A.
+For the kaon cross section analysis the relevant beam conditions is a secondary beam energy of 64 GeV, positive polarity magnet with current of 100 A.
+
+\textcolor{red}{DECIDE IF YOU WANT THE BEAM COMPOSITION HERE}
+%Figure 16 shows the tertiary beam spectra for the 64 GeV and 100 A.
+
+%In Table 3, the beam composition is given in terms of percentage of different particle species per spill for positive polarity.
+The values reported are the weighted average on the two beam conditions considered.
+The weights are calculated according to the fourth column of Table 2.
+
 \subsection{Data Driven MC}\label{sec:DDMC}
+The Data Driven single particle Monte Carlo (DDMC) is a single particle MC gun which simulates the particle transportation from WC4 into the TPC leveraging on the beamline data information.
+The DDMC uses the data momentum and position at WC4 to derive its initial conditions: a general sketch of the DDMC workflow is shown in Figure \ref{fig:DDMCSketch}.
 
-\section{Energy Calibration}\label{ch:EnCalibration}
+\begin{figure}[hpbt]
+\centering
+\includegraphics[width=\textwidth]{Chapter-5/Images/DDMCQuantities.png}
+\caption{Workflow for Data Driven single particle Monte Carlo production.}
+\label{fig:DDMCSketch}
+\end{figure}
+
+When producing a DDMC sample, beam line data from a particular running period and/or running condition are selected first.
+Figure 18 schematically shows the data quantities of interest: the measured momentum ($P_x, P_y, P_z$) and position ($X, Y$) at WC4.
+The momentum and position of the selected data is sampled thousand of times through a 5-dimensional hit-or-miss sampling procedure.
+This produces MC distributions with the same momentum and position distributions as data, with the additional benefit of accounting for the correlations between the considered variables.
+A LArSoft simulation module then launches single particle MC from z = -100 cm (the location of the WC4) using the sampled momentum and position distributions as a template.
+An illustration of this procedure is shown in Figure 17 with the results of the DDMC generation compared to a sample of wire chamber track data.
+Using this technique ensures the MC and data have very similar momentum, position and angular distributions at Wire-Chamber 4 and allow us to calibrate the energy loss upstream of the TPC as precisely as possible.
+The DDMC is a single particle Monte Carlo: the beam pile-up is not simulated.
+A sample of \textcolor{red}{NUMBERS} pions and \textcolor{red}{NUMBERS} kaons have been generated with the DDMC and used for the MC cross section study.
+
+\begin{figure}[hpbt]
+\centering
+\includegraphics[width=\textwidth]{Chapter-5/Images/DDMCScheme.png}
+\caption{Workflow for Data Driven single particle Monte Carlo production.}
+\label{fig:DDMCQuantities}
+\end{figure}
+
+\section{Energy Calibration}\label{ch:energyCalibration}
 \section{Tracking Studies}

### 2018-04-30 Chapter-5/5-SamplesPreparation.tex
s

@@
 
-\begin{figure}[hpbt]
-\centering
-\includegraphics[width=\textwidth]{Chapter-5/Images/DDMCQuantities.png}
-\caption{Workflow for Data Driven single particle Monte Carlo production.}
-\label{fig:DDMCSketch}
-\end{figure}
-
 When producing a DDMC sample, beam line data from a particular running period and/or running condition are selected first.
-Figure 18 schematically shows the data quantities of interest: the measured momentum ($P_x, P_y, P_z$) and position ($X, Y$) at WC4.
+Figure \ref{fig:DDMCQuantities}  schematically shows the data quantities of interest leveraged from data: the momentum ($P_x, P_y, P_z$) and position ($X, Y$) at WC4.
+For each data event, we obtain the particle position ($X, Y$) at WC4 directly from the data measurement.
+On the contrary, we calculate the components of the momentum using the beamline measurement of the momentum magnitude (see section \ref{sec:MWPCfunc}) in conjunction with the hits on WC3 and WC4 to determine the direction of the momentum vector.
 The momentum and position of the selected data is sampled thousand of times through a 5-dimensional hit-or-miss sampling procedure.
@@
 A LArSoft simulation module then launches single particle MC from z = -100 cm (the location of the WC4) using the sampled momentum and position distributions as a template.
-An illustration of this procedure is shown in Figure 17 with the results of the DDMC generation compared to a sample of wire chamber track data.
-Using this technique ensures the MC and data have very similar momentum, position and angular distributions at Wire-Chamber 4 and allow us to calibrate the energy loss upstream of the TPC as precisely as possible.
-The DDMC is a single particle Monte Carlo: the beam pile-up is not simulated.
-A sample of \textcolor{red}{NUMBERS} pions and \textcolor{red}{NUMBERS} kaons have been generated with the DDMC and used for the MC cross section study.
+As an example, the results of the DDMC generation compared to data for the pion 60A sample are shown in figure \ref{DDMCComparison}; as expected, MC and data agree within the statistical uncertainty by construction.
+Using this technique ensures the MC and data particles have very similar momentum, position and angular distributions at WC4 and allow us to us the MC sample in several occasions, for example to calibrate the energy loss upstream of the TPC or account for the WC2TPC match inefficiency.
+A small caveat is in order here: the DDMC is a single particle Monte Carlo, which means that the beam pile-up is not simulated.
+Three sample of \textcolor{red}{NUMBERS} pions, muons and electrons, as well as  a sample of \textcolor{red}{NUMBERS} kaons have been generated with the DDMC and are used for the MC cross section study.
 
@@
 \caption{Workflow for Data Driven single particle Monte Carlo production.}
+\label{fig:DDMCSketch}
+\end{figure}
+
+\begin{figure}[hpbt]
+\centering
+\includegraphics[width=\textwidth]{Chapter-5/Images/DDMCQuantities.png}
+\caption{Scheme of the quantities of interest for the DDMC event generation: $P_x, P_y, P_z, X, Y$ at WC4.}
 \label{fig:DDMCQuantities}
+\end{figure}
+
+\begin{figure}[hpbt]
+\centering
+\includegraphics[width=\textwidth]{Chapter-5/Images/.png}
+\caption{Comparison between generated quantities and data distributions for the 60A pion sample: Z component of the momentum (top left), X position at Wire Chamber 4 (top right), Y position at Wire Chamber 4 (bottom).}
+\label{fig:DDMCComparison}
 \end{figure}

### 2018-04-30 Chapter-1/1-Theory.tex
s

@@
 The gauge group univocally determines the number of gauge bosons that carry the interaction; the gauge bosons correspond to the generators of the group: eight gluons (g) for the strong interaction, one photon ($\gamma$) and three bosons (W$^\pm$, Z$^0$) for the electroweak interaction.
-A gauge theory by itself can not provide a description of massive particles, but it is experimentally well know that most of the elementary particles have non-zero masses.
+A gauge theory by itself cannot provide a description of massive particles, but it is experimentally well know that most of the elementary particles have non-zero masses.
 The introduction of massive fields in the Standard Model lagrangian would make the theory non-renormalizable, and - so far - mathematically impossible to handle.

### 2018-04-30 Chapter-2/2-LArTPCs.tex
s

@@
 
-\subsection{Signal Processing and Event Reconstruction}
+\subsection{Signal Processing and Event Reconstruction}\label{sec:SignalProc}
 In this section we illustrate the processing and reconstruction chain of the TPC signals, from the pulses on the sense wire to the construction of three dimensional objects with associated calorimetry.
@@
 \includegraphics[width=0.45\textwidth]{Chapter-2/Images/NueCCSim.png}	
-\includegraphics[width=0.45\textwidth]{Chapter-2/Images/MomentumKaonsDatabeamline.png}
+\includegraphics[width=0.45\textwidth]{Chapter-2/Images/momentumPiMuE.png}
 \caption{$Left$. Simulation of the products of a $\nu_e$ CC interaction in DUNE, both in particles type and momentum. \\
-$Right$. Momentum spectrum for low mass particles ($e,\mu,\pi$) in the LArIAT beamline,  negative tune, Run II, High Yield see section \ref{sec:MWPCfunc}. }
+$Right$. Momentum spectrum for low mass particles ($e,\mu,\pi$) in the LArIAT beamline,  negative tune, Run II, Picky Tracks see section \ref{sec:MWPCfunc}. }
 \label{fig:NuSimulation}

### 2018-04-30 Chapter-5/5-SamplesPreparation.tex
s

@@
 \centering
-\includegraphics[width=\textwidth]{Chapter-5/Images/.png}
+\includegraphics[width=\textwidth]{Chapter-5/Images/Pz.png}
 \caption{Comparison between generated quantities and data distributions for the 60A pion sample: Z component of the momentum (top left), X position at Wire Chamber 4 (top right), Y position at Wire Chamber 4 (bottom).}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%         Energy Calibration         %%%%%%%%%%%%%%         
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Energy Calibration}\label{ch:energyCalibration}
+Scope of the energy calibration is to identify the factors which convert the charge collected (dQ) to energy deposited in the chamber(dE).
+As described in section \ref{sec:SignalProc}, this is a multi-step procedure.
+In LArIAT, we first correct the raw charge by the electronic noise on the considered wire \cite{technote}, then by the electron life time \cite{technote},  and then by the recombination effect using the ArgoNeut recombination values.
+Lastly, we apply overall calibration of the energy, i.e.
+we determine the ``calorimetry constants" using the procedure described in this paragraph.
+
+The general idea is to use a parametrization of the energy deposited per unit length (dE/dX) as a function of momentum to independently determine the calorimetry constants for the data and Monte Carlo in the LArIAT Run-II data samples.
+This is done by comparing the stopping power measured on reconstructed quantities against the Bethe-Bloch theoretical prediction for various particle species.
+
+This is best represented by Figure \ref{fig:PDGEnergyLoss}, taken from the Particle Data Group \cite{PDG}.
+Using the tables provided by the PDG for liquid argon (\cite{PDG-Argon}), we calculate the theoretical values for the most probably value for the stopping power of pions ($\pi$), muons ($\mu$), kaons ($K$), and protons ($p$) in the momentum range most relevant for LArIAT, shown in Figure \ref{fig:PDGEnergyLossArgon}.
+
+\begin{figure}[htb]
+\centering
+%\includegraphics[width=0.50\textwidth]{images/PDGdEdX.png}
+\caption{Mean energy loss in various materials over a range of particle momenta as produced in Reference \cite{PDG}.}
+\label{fig:PDGEnergyLoss}
+\end{figure}
+
+\begin{figure}[htb]
+\centering
+%\includegraphics[width=0.50\textwidth]{images/dEdXvsMomentumTemplate}
+\caption{Mean energy loss for pions, muons, and protons in liquid argon over the momentum range most relvant for LArIAT.}
+\label{fig:PDGEnergyLossArgon}
+\end{figure}
+
+Using the predictions in Figure \ref{fig:PDGEnergyLossArgon}, allows us to tune the calorimetry constants.
+The goal is to have the data and the Bethe-Bloch prediction agree across the broad range of momentum.
+
+The basic idea of this calibration technique is to utilize a portion of a track within the LArTPC that has a well known momentum and particle species to measure the stopping power (or  energy deposited per unit length, $dE/dX$) as recorded inside the TPC.
+Once a sample of particles $dE/dX$ has been measured at various momenta, we then tune to calorimetry constants within the reconstruction software to align these measured values to match the theoretical ones found in Figure \ref{fig:PDGEnergyLossArgon}.
+
+The calibration procedure follows the following steps:
+\begin{itemize}
+\item \textbf{Species and Momentum Determination} We first select a sample of beamline events that correspond to either a sample of $\pi, \mu, e$, kaons or protons.
+This is done by selecting based on the invariant beamline mass as shown in section \ref{}.
+The beamline particle is matched to the TPC information.
+In order to determine the particle momentum at the TPC front face, we use the momentum measured by the LArIAT spectrometer and we apply a flat correction for the energy loss between the beamline and the TPC.
+For the Monte Carlo, no such beamline identification is done and instead we generate DDMC for each particle species.
+
+\item \textbf{dE/dX sampling:} With the track within the TPC identified and the momentum for that track measured, we require the track to be of a minimum length of 10~cm long (to ensure we are away from any interaction point where the track may be broken into subsequent tracks).
+We then take the first twelve spacepoints of the track (excluding the first point to avoid edge effects near the field cage) and sample the reconstructed dE/dX for each point along the track.
+On average, this samples 5~cm of the track (shown in Section \ref{sec:Results}).
+These dE/dX measurements are then put into a histogram that corresponds to measured momentum of the track.
+The dE/dX histograms are sampled every 50 MeV in momentum (e.g.
+150~MeV $< P <$ 200~MeV, 200~MeV $< P <$ 250~MeV, etc...).
+On average, pions and muons only lose $\sim$10 MeV in this 5~cm section of the track and protons lose $\sim$20 MeV.
+Thus choosing 50 MeV size bins for our histograms covers the energy spread within those bins due to energy loss from ionization.
+This process of selecting, sampling, and recording the dE/dX for various momentum bins is now repeated over the entire sample of events, allowing us to collect sufficient statistic in most of the momentum bins between 150~MeV and 1100~MeV.
+
+\item \textbf{Fit, tune, repeat:} Each 50 MeV momentum binned dE/dX histogram is now fit with a simple Landau function.
+The most probable value (MPV) and the associated error on the MPV from the fit are extracted and plotted on Figure \ref{fig:PDGEnergyLossArgon}.
+Depending on the outcome of the fit, the calorimetry constants are either tuned up or down.
+The values are tuned for both the collection and induction plane to try to achieve the best possible match to the theoretical curve.
+The exact match is left as a qualitative exercise and is not quantitatively evaluated.
+\end{itemize}
+
+\begin{figure}[htb]
+\centering
+%\includegraphics[width=0.50\textwidth]{images/CalibrationExample.png}
+\caption{Illustration of the calibration technique.
+Here we depict a 325 MeV wire chamber track (shown in green) which enters the TPC (taking into account the energy loss from the upstream material) and we sample the first 12 spacepoints (shown in teal) to extract the dE/dX distribution which is fit with a Landau.}
+\label{fig:CalibrationExample}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%         Energy Calibration         %%%%%%%%%%%%%%         
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \section{Tracking Studies}
@@
 The distributions for data and MC for $\sigma_\alpha$ and $\sigma_S$ are given in \ref{fig:trackingResolution}.
-
-\subsection{Estimate of $E_{loss}$ before the TPC}
-The beamline particles travel a path from when their momentum is measured by the beamline detector, until they are tracked inside the TPC.
-In the current LArIAT geometry, a particle leaving the fourth wire chamber will encounter the materials listed in Table \ref{tab:budget} before being registered again.
-The energy lost by the particle in this non instrumented material modifies the particle's kinetic energy and directly affects the cross section measurement, as shown in equation \ref{eq:enFF}.
-
-\begin{table}[h!]
-\centering
-\begin{tabular}{|l|l|l|}
-\hline
-Material  & density {[}g/cm$^3${]} & width {[}cm{]}    \\ \hline
-Fiberglass laminate (G10)      & 1.7                             & 1.28                              \\
-Liquid Argon                           & 1.4                             & 3.72                             \\
-Stainless Steel                        & 7.7                            & 0.23                             \\
-Titanium                                  & 4.5                            & 0.04                             \\ 
-Air                                            &  1.2 $\cdot10^{-3}$  & 89.43                              \\
-Plastic Scintillator                    & 1.03                          & 5.30                              \\ \hline
-\end{tabular}
-\caption{LArIAT material budget from WC4 to the TPC Front Face.}
-\label{tab:budget}
-\end{table}
-
-We estimate the uncertainty on the energy loss between the beamline momentum measurement and the TPC, $E_{loss}$, using the DDMC pion sample.
-We shoot pions from WC4 with the same momentum distribution as in the beamline data and plot the true $E_{loss}$ for that sample.
-The distribution for $E_{loss}$ for the pion sample is shown in Figure \ref{fig:Eloss}.
-We estimate the energy loss for pions to be $E_{loss} = 37 \pm 7 $~MeV where we use the average energy lost as the central value and the standard deviation of the distribution as the uncertainty.
-
-\begin{figure}[hpbt]
-\centering
-\includegraphics[scale=0.4]{Chapter-5/Images/cELossPions.png}
-\caption{Energy loss by simulated negative pions downstream from WC4 and upstream from the TPC.}
-\label{fig:Eloss}
-\end{figure}

### 2018-04-30 Chapter-8/8-CrossSectionAnalysis.tex
s

--- 2018-04-27 Chapter-9/9-Uncertainty.tex
@@
-\chapter{Uncertainty budget}\label{ch:Uncertainty}
-Measuring an hadronic cross section  in LArIAT translates into counting how many hadrons impinged on a slab of argon at a given energy and how many of those hadrons interacted at said energy.
-So, the key questions here are:
-\begin{itemize}
-\item[]a) how well do we know the kinetic energy at each point of the tracking? %(Incident Kinetic Energy bins)
-\item[]b) how well do we know when the tracking stops? %(Interacting Kinetic Energy bin)
-\item[]c) are there any systematic shifts?
-\end{itemize}
+\chapter{Negative Pion Cross Section Measurement}
 
-In order to answer this question, will discuss first a simple scenario  were our beam is 100\% made of pions which arrive as primaries in the TPC (no decay in the beam and no inelastic interaction before the TPC front face).
-We will then add a layer of complexity by discussing how we handle beamline contamination.
+\section{Estimate of $E_{loss}$ before the TPC}
+The beamline particles travel a path from when their momentum is measured by the beamline detector, until they are tracked again inside the TPC.
+In the current LArIAT geometry, a particle leaving the fourth wire chamber will encounter the materials listed in Table \ref{tab:budget} before being registered again.
+The energy lost by the particle in this non-instrumented material modifies the particle's kinetic energy and directly affects the cross section measurement, as shown in equation \ref{eq:enFF}.
 
-\section{Pure beam of pions}
-Assuming a beam of pure pions gets to the TPC, let us explicit some of the variables in the kinetic energy equation \ref{eq:KEj}  to point out the important quantities in the uncertainty budget,
+\begin{table}[h!]
+\centering
+\begin{tabular}{|l|l|l|}
+\hline
+Material  & density {[}g/cm$^3${]} & width {[}cm{]}    \\ \hline
+Fiberglass laminate (G10)      & 1.7                             & 1.28                              \\
+Liquid Argon                           & 1.4                             & 3.20                             \\
+Stainless Steel                        & 7.7                            & 0.23                             \\
+Titanium                                  & 4.5                            & 0.04                             \\ 
+Air                                            &  1.2 $\cdot10^{-3}$  & 89.43                              \\
+Plastic Scintillator                    & 1.03                          & 1.20 (+ 1.30)                             \\ \hline
+\end{tabular}
+\caption{LArIAT material budget from WC4 to the TPC Front Face.}
+\label{tab:budget}
+\end{table}
 
-\begin{align}
- E_{j}^{kin} &=  E_{Beam}^{kin}  - E_{loss} - \sum_{i < j} \frac {dE_i}{dx_i}*dx_i\\
-                  &=  \sqrt{p^2_{Beam} - m^2_{Beam}} - m_{Beam} - E_{loss} - \sum_{i < j} \frac {dE_i}{dx_i}*dx_i.
-\end{align}
-
-\subsection{Uncertainty on $E_{Beam}^{kin}$}
-Let us start by discussing the uncertainty on $E_{Beam}^{kin}$.
-Since we are assuming a beam of pions, the uncertainty on the value of mass of the pion ($m_{Beam}$) as given by the pdg is irrelevant compared to the momentum uncertainties, thus $\delta E_{Beam}^{kin} = \delta p_{Beam}^{kin}$.
-We estimate the momentum uncertainty as follows.
-
-\textcolor{blue}{  
-We estimate the uncertainty on a 4-point track. In case of 3-points track, we add an additional 2\% coming from Greg's study. 
-Uncertainty on a 4-point track:
-\begin{itemize}
-\item[-]  Alignment surveys. 1mm misalignment translates to 3\% in overall
-\item[-] Doug study dp/p = ~2\% based on field map (docdb 1710)
-\item[-] Minerva test beam paper
-\end{itemize}
-}
-
-\subsection{$E_{loss}$}
-We derive an estimate of the energy loss between the beamline momentum measurement and the TPC ($E_{loss}$) from the Monte Carlo using the DDMC pion sample, since this quantity is not  measurable directly on data.
-We shoot pions from WC4 with the same momentum distribution as in the beamline data and plot the true $E_{loss}$ for that sample.
+We derive an estimate of the energy loss between the beamline momentum measurement and the TPC ($E_{loss}$) from the Data Driven Monte Carlo using the pion and kaon samples separately, since this quantity is not  measurable directly on data.
 The $E_{loss}$ distribution for the 60A  and 100A pion sample is shown in figure \ref{fig:ELoss60A}, left and right respectively.
@@
 \centering
-\includegraphics[width=0.45\textwidth]{Chapter-9/Images/E_loss60A.png}
-\includegraphics[width=0.45\textwidth]{Chapter-9/Images/E_loss100A.png}
+\includegraphics[width=0.45\textwidth]{Chapter-5/Images/E_loss60A.png}
+\includegraphics[width=0.45\textwidth]{Chapter-5/Images/E_loss100A.png}
 \caption{True energy loss between WC4 and the TPC front face according to the MC simulation of the 60A runs (left) and of the 100A runs (right).
@@
 \centering
-\includegraphics[scale=0.5]{Chapter-9/Images/Halo.png}
+\includegraphics[scale=0.5]{Chapter-5/Images/Halo.png}
 \caption{Schematic rendering of the particle path between WC4 and the TPC front face.
@@
 \centering
-\includegraphics[width=\textwidth]{Chapter-9/Images/PXVsX60A.png}
+\includegraphics[width=\textwidth]{Chapter-5/Images/PXVsX60A.png}
 \caption{Horizontal component of the true momentum vs the horizontal position at WC4 for MC simulated pions of the 60A runs.
@@
 
-%\textcolor{blue}{ TO DO HERE: make sure we have the geometry right, cause otherwise this correction is meaningless.  With this method, so far we get a mean ~40 MeV, but uncertainty ~7MeV. 
-%The trajectory method does not improve uncertainty, why? It's a mystery I don't think we should solve before June :) .
-%Back of the envelope material budget calculation:}
-%\begin{table}[h!]
-%\centering
-%\caption{Back of the envelope calculation}
-%\label{my-label}
-%\begin{tabular}{|l|l|l|l|}
-%\hline
-%dEdx for MIP, MPV {[}MeV cm$^2$/gr{]} & density {[}g/cm$^3${]} & width {[}cm{]} & E$_{loss}$ {[}MeV{]} \\ \hline
-%1.6                                                 & 1.7 (G10)                            & 1.3            & 3.5                   \\
-%1.6                                                 & 1.4 (LAr)                            & 1.77           & 4.0                   \\
-%1.6                                                 & 7.7 (S.S.)                           & 0.23           & 2.8                   \\
-%1.6                                                 & 4.5 (Ti)                             & 0.04           & 0.3                   \\ 
-%1.6                                                 & 1.03 (Plastic Sci)                   & 1.1            & 1.8                   \\ \hline
-%Total                                               &                                      &                & 12.4                  \\ \hline
-%\end{tabular}
-%\end{table}
+\section{Interacting and Incident Distributions}
 
-\subsection{Uncertainty on dE/dx and pitch}
-We obtain the uncertainty on dE/dx and track pitch by comparing the dE/dx and pitch distributions in data and MC.
-\textcolor{blue}{ Currently, MPV MC = 1.70 and MPV DATA = 1.72 MeV/cm (~3\% higher).
-TO DO HERE: calculate Argon density from mid-RTD temperature. Compare this  density with MC Argon density. 
-Density change  affects dE/dx (in MeV/cm!). Try changing MC density up to ``real one" and see if dEdX agrees between DATA and MC}
-
-\subsection{Uncertainty on track end, aka efficiency correction}
-From the MC, we obtain an efficiency correction on the interacting and incident distributions separately.
-This is done by comparing the MC reconstructed with the true MC deposition on an event by event basis.
-This correction is applied bin by bin on the data interacting and incident distributions.
-The better our tracking, the smaller this efficiency correction will be. So, step number one is improving the tracking.
-\textcolor{blue}{Need to talk to Bruce about this.}
-\textcolor{blue}{ I don't understand the angle cut that Dave Schmitz and Jon Paley were so vocal about.}
-
-Now, the key question remains: does the tracking behave in the same way in data and MC? 
-We can compare some key plots between reconstructed data and MC which gives us confidence this is true: the track pitch, the tracks straightness and the goodness of fit in data and MC.
-\textcolor{blue}{ Does such a variable as ``goodness of fit" exists in the tracking?
-We should ask Bruce.}
+\section{Total Hadronic Negative Pion-Argon Differential Cross Section}

### 2018-04-30 Chapter-9/9-Uncertainty.tex
s

@@
 
-\subsection{$E_{loss}$}
-We derive an estimate of the energy loss between the beamline momentum measurement and the TPC ($E_{loss}$) from the Monte Carlo using the DDMC pion sample, since this quantity is not  measurable directly on data.
-We shoot pions from WC4 with the same momentum distribution as in the beamline data and plot the true $E_{loss}$ for that sample.
-The $E_{loss}$ distribution for the 60A  and 100A pion sample is shown in figure \ref{fig:ELoss60A}, left and right respectively.
-A clear double peaked structure is visible, which is due to the particles either missing or hitting the HALO paddle: a schematic rendering of this occurrence is  shown in figure \ref{fig:Halo}.
-The kinematic at WC4 determines the trajectory of a particle and whether or not it will hit the halo paddle.
-In figure \ref{fig:PxVsXTrue} , we plot the true  $X$ component of the momentum versus the true $X$ position at WC4 for pions missing the halo paddle (left) and for pions hitting the halo paddle (right) for the 60A MC simulation runs -- analogous plots are obtained with the 100A simulation.
-These distributions can be separated drawing a line in this position-momentum space.
-We use a logistic regression  \cite{agresti2013categorical}  as a classifier to find the best separating line, shown in both plots as the red line.
-We classify as ``hitting the halo paddle" all pions whose $P_x$ and $X$ are such that $$P_x +0.02* X - 0.4 < 0 $$ and as ``missing the halo  paddle" all pions whose $P_x$ and $X$ are such that $$P_x +0.02*X - 0.4 > 0, $$ where the coefficients of the line are empirically found by the logistic regression estimation.
-Overall, this simple classifier classifies in the right category (hit or miss) about 86\% of the pion events.
-We apply the same classifier on data.
-We assign  $E_{loss} = 32 \pm 4 $~MeV for events classified as ``hitting the halo paddle"; we assign  $E_{loss} = 24 \pm 3 $~MeV for events classified as ``missing the halo paddle".
-
-%We use the separation of these two distributions to decide what the energy loss for each event on data. 
-%Thus,  we assign the value for energy loss is used in the data.
-
-\begin{figure}[hbpt]
-\centering
-\includegraphics[width=0.45\textwidth]{Chapter-9/Images/E_loss60A.png}
-\includegraphics[width=0.45\textwidth]{Chapter-9/Images/E_loss100A.png}
-\caption{True energy loss between WC4 and the TPC front face according to the MC simulation of the 60A runs (left) and of the 100A runs (right).
-The distribution for the whole data sample is shown in blue, the distribution for the pions missing the halo is shown in red, and the distribution for the pions hitting the halo is shown in green.
-}
-\label{fig:ELoss60A}
-\end{figure}
-
-\begin{figure}[hbpt]
-\centering
-\includegraphics[scale=0.5]{Chapter-9/Images/Halo.png}
-\caption{Schematic rendering of the particle path between WC4 and the TPC front face.
-The paddle with the hollow central circle represents the Halo paddle.
-We illustrate two possible trajectories: in black, a trajectory that miss the paddle and goes through the hole in the Halo, in blue a trajectory that hits the Halo paddle and goes through the scintillation material.}
-\label{fig:Halo}
-\end{figure}
-
-\begin{figure}[hbpt]
-\centering
-\includegraphics[width=\textwidth]{Chapter-9/Images/PXVsX60A.png}
-\caption{Horizontal component of the true momentum vs the horizontal position at WC4 for MC simulated pions of the 60A runs.
-The plot on the left shows the distribution for pion that miss the halo paddle and the plot on the right shows the distributions for pions that hit the halo.
-The form of the classifier is overlaid to both plots (red line).}
-\label{fig:PxVsXTrue}
-\end{figure}
+\subsection{Systematics on $E_{loss}$}
 

### 2018-04-30 Chapter-1/1-Theory.tex
s

@@
 
-Overwhelming experimental data show neutrinos change flavor during their propagation~\cite{Patrignani:2016xqp}.
+Overwhelming experimental data show that neutrinos change flavor during their propagation~\cite{Patrignani:2016xqp}.
 This phenomenon, called ``neutrino oscillations",  was predicted first by Bruno Pontecorvo in 1957 ~\cite{Pontecorvo:1967fh}.
@@
 
-where $c$ e $s$ stand respectively for cosine and sine of the corresponding  mixing angles ($\theta_{12}$, $\theta_{23}$ and $\theta_{13}$), $\delta$ is the Dirac CP violation phase, $\alpha_1$ and $\alpha_{2}$   is the eventual Majorana CP violation phases.
+where $c$ e $s$ stand respectively for cosine and sine of the corresponding  mixing angles ($\theta_{12}$, $\theta_{23}$ and $\theta_{13}$), $\delta$ is the Dirac CP violation phase, $\alpha_1$ and $\alpha_{2}$  are the eventual Majorana CP violation phases.
 Experimental results on neutrino oscillations are generally reported in terms of the mixing angles and of the squared mass splitting $\Delta m^2_{ab} = m^2_{a} - m^2_{b}$, where $a$ and $b$ represent the mass eigenstates.
@@
 
-Historically, neutrino interactions with the nucleus in the GeV region are divided into three categories as a function of increasing neutrino energy: quasi elastic (QE), resonant, and deep inelastic (DIS) scattering.
-All current and forthcoming oscillation experiments live in the 0.1-10 GeV transition region, which encompasses the energy where the QE neutrino-nucleus interaction transitions into resonant scattering and the energy  where resonance scattering  transitions into DIS.
+Historically, neutrino interactions with the nucleus in the GeV region are divided into three categories as a function of increasing neutrino energy: quasi elastic (QE), resonant (RES), and deep inelastic (DIS) scattering.
+All current and forthcoming oscillation experiments live in the 0.1-10 GeV transition region, which encompasses the energy where the QE neutrino-nucleus interaction transitions into RES and the energy  where RES  transitions into DIS.
  Schematically, neutrino and antineutrino QE charge current scattering refers to the process $\nu_l n \rightarrow l^- p$ and $\bar\nu_l p \rightarrow l^+ n$ where a charged lepton and single nucleon are ejected in the elastic interaction, leaving the target nucleus in its ground state.
@@
 On one hand, the last three decades of experiments in neutrino oscillations brought spectacular advancements in the understanding of the oscillations pattern, measuring the neutrino mixing angles and mass splitting with a precision of less than 10\%.
-On the other, it opened the field for a series of questions needing experimental answers.
+On the other, they opened the field for a series of questions needing experimental answers.
 
@@
 \begin{equation}
-24 = \underbrace{(8,1)}_{\text{gluons}} + \underbrace{ (1,3) + (1,1)}_{\text{W$^\pm$, Z, $\gamma$}}.. +  \underbrace{(3,2) + (\bar 3,2)}_{\text{X,Y bosons}}. 
+24 = \underbrace{(8,1)}_{\text{gluons}} + \underbrace{ (1,3) + (1,1)}_{\text{W$^\pm$, Z, $\gamma$}} +  \underbrace{(3,2) + (\bar 3,2)}_{\text{X,Y bosons}}. 
 \end{equation}
@@
 Their are the mediator of the interaction that turns quarks into leptons, leading to predict the existence of processes that violate baryon number, such as $p \rightarrow \pi^0  + e^+$  (see fig~ \ref{fig:MandatoryFeynmannDiagrams}, right).
-The prediction for proton decay lifetime, $\tau_p \sim \frac{M^4_X}{m^5_p}\sim10^{30\pm1.5}$~years, is unfortunately experimentally disproved by IMB and Super-Kamiokande \cite{s}.
+The prediction for proton decay lifetime, $\tau_p \sim \frac{M^4_X}{m^5_p}\sim10^{30\pm1.5}$~years, is unfortunately experimentally disproved by IMB and Super-Kamiokande \cite{PhysRevD.42.2974,PhysRevD.95.012004}.
 
@@
 
-\textbf{SUSY GUTs.} Supersymmetry theories allow for another family of GUTs\cite{supersymmetry}.
+\textbf{SUSY GUTs.} Supersymmetry theories allow for another family of GUTs.
 In SUSY, every fundamental particle in the SM has a ``superpartner", identical in each quantum number except for the spin-statistics: the fermion supersymmetric partners are bosons and vice versa.
@@
 From the phenomenology point of view, SUSY models tend to push the proton decay life time higher by a factor of four, they solve the ``hierarchy problem", and they also predict new channels for the proton decay.
-In particular they predict the presence of kaons in the final product, with a dominant mode of $p\rightarrow\bar\nu K^+$.
+In particular they predict the presence of kaons in the final product, with a dominant mode of $p\rightarrow K^+ \bar\nu$.
 %, since particles from the first generation (e.g.
@@
 \section{Motivations for Hadronic Cross Sections in Argon}
-Critical challenges await  the next decade of experimental physics at the intensity frontier.
+Critical challenges await  the next decade of high energy physics at the intensity frontier.
 % will entail an even deeper understanding of the neutrino mixing pattern, investigation of the neutrino mass origin, the determination of the number of neutrinos and their nature, and the measurement of CP violation in the lepton sector.
@@
 % and of light mesons interaction with nuclei studies.
- We describe the signal signature and historical measurements of pion-nucleus cross section, as well as the implementation of this cross sections in the current version of the simulation package used by LArIAT.
+ We describe the signal signature and historic measurements of pion-nucleus cross section, as well as the implementation of this cross sections in the current version of the simulation package used by LArIAT.
 
@@
 
-We illustrate here the basis of the GENIE Neutrino Generator~\cite{Andreopoulos:2009rq}, since it is one the most popular event generators for LAr experiments.
+%We illustrate here the basis of the GENIE Neutrino Generator~\cite{Andreopoulos:2009rq}, since it is one the most popular event generators for LAr experiments.
 
 The key elements of a neutrino event generators for resonance and DIS events are the nuclear model and the hadron treatment (both production and transportation).
-We illustrate here the conceptual basis of the GENIE Neutrino Generator~\cite{Andreopoulos:2009rq} as an example, since GENIE is one the most popular event generators for LAr experiments.
+We illustrate here the conceptual basis of the GENIE Neutrino Generator~\cite{Andreopoulos:2009rq} as an example, since GENIE is one the most popular event generators for liquid argon experiments.
 For example, the nuclear model used by GENIE for all processes is a relativistic Fermi gas (RFG) modified to incorporate short range nucleon-nucleon correlations \cite{PhysRevD.24.1400}.
@@
 Stopping negative pions can form pionic argon, where the negative pion plays the role of an orbital electron.
-Since the pion mass is two orders of magnitude greater than the electron mass, the spatial wave form of the pion will overlap way more with the nucleus compared to the electron case.
-After the electromagnetic formation of the pionic atom, the pion will quickly absorbed by the nucleus, which is put in an excited state.
+Since the pion mass is two orders of magnitude greater than the electron mass, the spatial wave form of the pion will overlap more with the nucleus compared to the electron case.
+After the electromagnetic formation of the pionic atom, the pion will be quickly absorbed by the nucleus, which is put in an excited state.
 The nucleus then de-excites with the emission of low energy nucleons and photons.
@@
 
-The shape of the pion-nucleus interaction cross section in the energy range considered shows the distinct features that indicate the presence of a resonance.
+The shape of the pion-nucleus interaction cross section in the energy range considered shows the distinct features indicating the presence of a resonance.
 In fact, the mean free path of a pion of kinetic energy between 100 and 400 MeV is much shorter than the average distance between nucleons (which is of the order of 1 fm).
@@
 A $\Delta$ resonance is often produced in the interaction, which subsequently decays inside the nucleus.
-Experimental results on several nuclei as reported in \cite{PhysRevC.14.635} are shown in Figure \ref{fig:pi1}; it is interesting to notice here how the shape of the $\Delta$ resonance become less pronounced as a function of the mass number of the target nucleus.
+Experimental results on several nuclei as reported in \cite{PhysRevC.14.635} are shown in Figure \ref{fig:pi1}; it is interesting to notice here how the shape of the $\Delta$ resonance becomes less pronounced as a function of the mass number of the target nucleus.
 Pion interactions with heavier nuclei also shift the peak of the resonance at lower energy; this effect is due to kinematic considerations and to the difference in propagation of the $\Delta$ inside the nucleus.
@@
 The relevance of the GENIE generator for neutrino physics and its basic working principles have been outlined earlier in this section.
-Given GENIE's modularity, information on hadron-nucleus interactions can be extracted from the  INTRANUKE module and directly compared against the Geant4 predictions.%, as shown in Figure    \ref{fig:pionCa}.
-
+Given GENIE's modularity, information on hadron-nucleus interactions can be extracted from the  INTRANUKE module and directly compared against the Geant4 predictions.
 The work in \cite{Nutini2015}  reviews  the current status of negative and positive pion simulation in Geant4 and GENIE for $^{12}$C, $^{56}$Fe, and $^{40}$Ca.
@@
 
-For the $\pi^-$ argon total cross section measurement, we use the Geant4 Bertini Cascade model, whose  predictions for the total, elastic and reaction cross sections are show in Figure \ref{fig:PionTrueArgon}.
+For the LArIAT simulation of the MC sample used in the $\pi^-$ argon total hadronic cross section measurement we use the Geant4 Bertini Cascade model, whose  predictions for the total, elastic and reaction hadronic cross sections are show in Figure \ref{fig:PionTrueArgon}.
 
@@
 
-\subsubsection{K$^{+}$Ar Cross section in the Context of Nucleon Decay Searches}
-\subsubsection{Nucleon decay}\label{sec:theoryPDK}
+\subsubsection{K$^{+}$Ar Cross section in the Context of Nucleon Decay Searches}\label{sec:theoryPDK}
 Baryon number is accidentally conserved in the Standard Model.
@@
 On the other, it further highlights the importance of the kaon measurement.
-Figure \ref{fig:TrueArgon} shows the Geant4 prediction for the total, elastic and reaction cross section for $K^+$ on argon.
+For the LArIAT simulation of the MC sample used in the K$^+$ argon total hadronic cross section measurement we use the Geant4 Bertini Cascade model, whose  predictions for the total, elastic and reaction hadronic cross sections are show in Figure \ref{fig:TrueArgon}.
 

### 2018-04-30 Chapter-2/2-LArTPCs.tex
s

@@
 \section{The Liquid Argon Time Projection Chamber Technology}
-
 \subsection{TPCs, Neutrinos \& Argon}
@@
 Using nobles elements in the liquid form for neutrino detectors is advantageous for several reasons.
-The density of liquids is $\sim$1000 times greater than gases, increasing the number of target centers for neutrino's interaction in the same volume.
-Since the energy loss of charged particle is proportional to the target material density, as shown in the Bethe-Block equation, eq.
-\ref{eq:BB}, the increased density reflects into a proportionally higher energy loss, enhancing the calorimetry capability of detectors with a liquid active medium.
+The density of liquids is $\sim$1000 times greater than gases, augmenting the number of targets for neutrino's interaction in the same volume, in a effort to balance the smallness of neutrino cross section.
+Since the energy loss of charged particle is proportional to the target material density, as shown in the Bethe-Block equation (eq.
+\ref{eq:BB}), the increased density reflects into a proportionally higher energy loss, enhancing the calorimetry capability of detectors with a liquid active medium.
 Additionally, the ionization energy of liquids is smaller than gasses by the order of tens of eV.
 Thus, at the passage of charged particles, liquid generally produce more ionization electrons than gas for the same deposited energy and force the particles to deposit more energy in a shorter range.
-The downside of using noble liquid gasses in experiments is that they require expensive cryogenic systems to cool the gas until it transitions to its the liquid form.
+The downside of using noble liquid elements in experiments is that they require expensive cryogenic systems to cool the gas until it transitions to its the liquid form.
 The properties of liquid argon in comparison liquid xenon -- a popular choice for dark matter and neutrinoless double beta decay detectors -- are summarized in table \ref{tab:properties}.

### 2018-04-30 Chapter-2/2-LArTPCs.tex
s

@@
 Additionally, the ionization energy of liquids is smaller than gasses by the order of tens of eV.
-Thus, at the passage of charged particles, liquid generally produce more ionization electrons than gas for the same deposited energy and force the particles to deposit more energy in a shorter range.
+Thus, at the passage of charged particles, liquids generally produce more ionization electrons than gases for the same deposited energy, forcing the particles to deposit more energy in a shorter range.
 The downside of using noble liquid elements in experiments is that they require expensive cryogenic systems to cool the gas until it transitions to its the liquid form.
 The properties of liquid argon in comparison liquid xenon -- a popular choice for dark matter and neutrinoless double beta decay detectors -- are summarized in table \ref{tab:properties}.
-Albeit xenon would be more desirable than argon given some superior properties such as lower ionization energy and higher density and light yield, argon relative abundance abates the cost of argon compared to xenon, making argon a more viable choice for the construction of kilo-ton scale neutrino detectors.
+Albeit xenon would be more desirable than argon given some superior properties such as lower ionization energy and higher density and light yield, argon relative abundance abates the cost of argon compared to xenon, making argon a more viable choice for the construction of ton  (and kilo-ton) scale neutrino detectors.
 
@@
 LArTPCs are some times referred as to ``electronic" bubble-chambers, for the similarity in the tracking and energy resolution which is coupled with an electronic readout of the imaging information in LArTPCs.
-Compared to these historic detectors however, LArTPC bestow tridimensional tracking and a self triggering mechanism provided by the scintillation light in the noble gas.
-An event display of a $\nu_\mu$ CC interaction candidate in the MicroBooNE detector is shown in picture \ref{fig:NuEvd} to display the level of spatial details these detectors a capable of; the color scale of the image is proportional to the energy deposited, hinting to the calorimetry capabilities of the detectors.
+Compared to these historic detectors however, LArTPC bestow tridimensional tracking and a self triggering mechanism provided by the scintillation light in the liquid argon.
+An event display of a $\nu_\mu$ CC interaction candidate in the MicroBooNE detector is shown in picture \ref{fig:NuEvd} to display the level of spatial details these detectors are capable of; the color scale of the image is proportional to the energy deposited, hinting to these calorimetry capabilities of the detectors.
 \begin{figure}[hbpt]
@@
 
-%The time projection chamber (TPC) was invented by Nygren in 1974 [91].
-The proposal to implement a liquid argon TPC (LArTPC) for neutrino physics was made by Rubbia in 1977 [92], and the concept was implemented by the ICARUS collaboration [93].
+\begin{figure}[hbpt]
+\centering
+\includegraphics[width=\textwidth]{Chapter-2/Images/Cartoon.png}
+\caption{A cartoonish sketch of a LArTPC.}
+\label{fig:cartoon}
+\end{figure}
 
-To the bare bones, a LArTPC is a bulk of liquid argon sandwiched in a flat capacitor, equipped with a light collection system.
+To the bare bones, a LArTPC is a bulk of liquid argon sandwiched in a flat capacitor, equipped with a light collection system, as the cartoon in \ref{fig:cartoon} shows.
 A uniform electric field of the order of 500 V/cm is maintained constant between the faces of the capacitor.
@@
 Argon ionization and scintillation are the processes leveraged to detect particles in the LArTPC active volume.
-When a ionizing radiation traverses the argon active volume it leaves a trail of ionization electrons along its trajectory and it excites the argon producing of scintillation light -- details on the production and detection of ionization charge and scintillation light are provided in \ref{sec:light} and \ref{sec:light} respectively.
-The optical the detector sees the argon scintillation light in matters of nanoseconds.
+When a ionizing radiation traverses the argon active volume it leaves a trail of ionization electrons along its trajectory and it excites the argon producing scintillation light -- details on the production and detection of ionization charge and scintillation light are provided in \ref{sec:light} and \ref{sec:light} respectively.
+The optical detector sees the argon scintillation light in matters of nanoseconds.
 This flash of light determines the start time of an event in the chamber, $t_0$.
@@
 Ionized argon is collected on the cathode which is not instrumented, so it is not used to infer information about the interactions in the chamber.}.
-The anode sense wires see either an induced current by the drifting charge (on induction planes) or an injection of the ionization charge (collection plane).
+The anode sense wires see either an induced current by the drifting ionization charge (on induction planes) or an injection of such charge (collection plane).
 An appropriate choice of the voltage bias on each wire plane assures ideal charge transparency, so that all the ionization charge is collected on the collection plane and none on the induction planes.
@@
 
-Since the charge deposited by the ionizing radiation is proportional to the deposited energy and the charge collected on the sense plane is a function of the deposited charge, LArTPC allow the measurement of the energy deposit in the active volume.
+Since the charge deposited by the ionizing radiation is proportional to the deposited energy and the charge collected on the sense plane is a function of the deposited charge, LArTPCs allow the measurement of the energy deposit in the active volume.
 Effects due to the presence of free charge and impurities in the active volume, such as a finite electron lifetime, recombination and space charge, complicate the relationship between deposited and collected charge affecting the measurement of the particle's energy, as described in the next section.
@@
 About 66\% of the ionizing collisions in Argon produce only a single electron/ion pair \cite{0034-4885-73-11-116201}; in the other cases, the transferred kinetic energy is enough for the primary electron to liberate one or more secondary electrons, which usually stay close to the original pair.
-Occasionally, electrons can receive enough energy to be ejected with high energy, forming so-called ``$\delta$-ray": a detectable ionization short track off the particle trajectory, as shown in figure \ref{fig:delta}.
+Occasionally, electrons can receive enough energy to be ejected with high energy, forming a so-called ``$\delta$-ray": a detectable  short track off the particle trajectory, as shown in figure \ref{fig:delta}.
 The average number of $\delta$-ray  with energy E$>$E$_0$ per cm follows the empirical form
@@
 \subsubsection{Purity \& Electron Life Time }
-The presence of electronegative contaminants in liquid argon, such as oxygen and water, is particularly
+The presence of electronegative contaminants in liquid argon, such as oxygen $O_2$ and water $H_2O$, is particularly
 pernicious, since these molecules quench the charge produced by the ionizing radiation.
@@
 LArIAT small drift distance (47 cm) allows for a relatively short electron life time.
-The life time for bigger detectors such as MicroBooNE, whose drift distance is 2.5 m, needs to be of the order of tens of milliseconds to allow charge collection usable for physics analyses.
+The life time for bigger detectors such as MicroBooNE, whose drift distance is 2.6 m, needs to be of the order of tens of milliseconds to allow a charge collection usable for physics analyses.
 Energy reconstruction in LArTPC applies a correction for the finite lifetime to calibrate the detector calorimetric response; details for LArIAT are provided in Section \ref{ch:energyCalibration}.
@@
 
-Models for a theoretical descriptions of recombination based on the Birks model and the Box model are provided in \cite{0370-1298-64-10-303} and  \cite{PhysRevA.36.614}, respectively.
+Theoretical descriptions of recombination based on the Birks model and the Box model are provided in \cite{0370-1298-64-10-303} and  \cite{PhysRevA.36.614}, respectively.
 The Birks model assumes a gaussian spatial distribution around the particle trajectory during the entire recombination phase and identical charge mobility for ions and electrons.
@@
 In these models, the fraction of ionization electrons surviving recombination is a function of the number of ion-electron pairs per unit length, the electric field, the average ion-electron separation distance after thermalization and the angle of the particle with respect to the direction of the electric field -- plus the diffusion coefficient in the Birks model.
-Given the stringent assumptions, it is perhaps  not surprising that these models are in accordance to data only in specific regimes: the Birks model is generally used to describe recombination for low dE/dx, the Box model for high dEdX.
-In LArTPC, the ICARUS and ArgoNeut have measured recombination in \cite{Amoruso:2004dy} and \cite{1748-0221-8-08-P08005} respectively.
+Given the stringent assumptions, it is perhaps  not surprising that these models are in accordance to data only in specific regimes: the Birks model is generally used to describe recombination for low dE/dx, the Box model for high dE/dX.
+In LArTPC, the ICARUS and ArgoNeut experiments have measured recombination in \cite{Amoruso:2004dy} and \cite{1748-0221-8-08-P08005} respectively.
 Since LArIAT uses the refurbished ArgoNeut TPC and cryostat at the same electric field,  LArIAT currently corrects for recombination using the ArgoNeut measured recombination parameters in \cite{1748-0221-8-08-P08005}.
@@
 Surface LArTPC of the size of several meters are expected to be modestly impacted from the space charge effect, where charge build-up create anisotropy of the electric field magnitude of the order of  5\% at a drift field of 500 V/cm \cite{SpaceCharge}.
-The smallness of the LArIAT drift volume is such that  effect of  space charge on the electric field is expected to be even smaller.
+The smallness of the LArIAT drift volume is such that  effect of  space charge on the electric field is expected to be negligible.
 \textcolor{red}{CHIEDI A FLAVIO}
@@
 
-\textbf{Deconvolution.} Induction and collection planes have different field responses, given the different nature of the signals on these planes: the wires on the induction planes see inductive signal of the drifting charge, while the wires on the collection planes see the current derived from the charge entering the conductor.
+\textbf{Deconvolution.} Induction and collection planes have different field responses, given the different nature of the signals on these planes: the wires on the induction planes see the inductive signal of the drifting charge, while the wires on the collection planes see the current derived from the charge entering the conductor.
 Thus, signals on the induction plane are bi-polar pulse and signal on the collection plane are unipolar pulses, see Figure \ref{fig:SignalProc} panel a.

### 2018-04-30 Chapter-1/1-Theory.tex
s

@@
 
-\subsection{Open Questions in Neutrino Physics}
+\subsection{Open Questions in Neutrino Physics}\label{ch:questions}
 On one hand, the last three decades of experiments in neutrino oscillations brought spectacular advancements in the understanding of the oscillations pattern, measuring the neutrino mixing angles and mass splitting with a precision of less than 10\%.

### 2018-04-30 Chapter-2/2-LArTPCs.tex
s

@@
 \textbf{Deconvolution.} Induction and collection planes have different field responses, given the different nature of the signals on these planes: the wires on the induction planes see the inductive signal of the drifting charge, while the wires on the collection planes see the current derived from the charge entering the conductor.
-Thus, signals on the induction plane are bi-polar pulse and signal on the collection plane are unipolar pulses, see Figure \ref{fig:SignalProc} panel a.
+Thus, signals on the induction plane are bi-polar pulse and signal on the collection plane are unipolar pulses, see Figure \ref{fig:SignalProc} panel a).
 The first step in signal processing is deconvolution, that is a series of off-line algorithms geared towards undoing the detector effects.
@@
 It then fits these peaks with gaussian shapes and stores the fit parameters such as the quality of the fit, the peak time, height and area under the gaussian fit.
-The information resulting fron this process a single spike form a single reconstructed ``hit".
+The information resulting from this process on a single spike form a single reconstructed ``hit", see Figure \ref{fig:SignalProc} panel c).
 The next steps in the event reconstruction chain will then decide if rejecting hits with poor fits.
@@
 \end{enumerate}
+The final product of this reconstruction stage is the collection of bidimensional clusters on each wire plane, see Figure \ref{fig:SignalProc} panel d).
 
@@
 Then, it  projected back the tentative trajectory on to the planes and adjusts the parameters of the 3D track fit such that they minimize the distance between the fit projections and the track hits in all wire planes simultaneously.
-The track algorithm can use multiple clusters in one plane, but it can never break them in smaller groups of hits.
-This algorithm was first developed for the ICARUS collaboration\cite{Antonello2013}.\\
+Tridimensional tracking can use multiple clusters in one plane, but it can never break them in smaller groups of hits.
+This algorithm was first developed for the ICARUS collaboration\cite{Antonello2013}.
+The final product of this reconstruction stage is the formation of  tridimensional objects in the TPC active volume, see Figure \ref{fig:SignalProc} panel e).\\
 
@@
 A multi-step procedure is needed to retrieve the energy deposited in the TPC  from the charge seen by the wires.
-For each hit associated with the track object, the calorimetry algorithms calculate the charge seen on every wire using the area underneath the gaussian fit; then they correct this raw charge by the electron life time, the electronic noise on the considered wire and the recombination effect.
+For each hit associated with the track object, the calorimetry algorithms calculate the charge seen on every wire using the area underneath the gaussian fit; then, they correct this raw charge by the electron life time, the electronic noise on the considered wire and the recombination effect.
 Lastly an overall calibration of the energy, explained in detail in section \ref{ch:energyCalibration}, is applied and the calorimetric information for the given track is assigned.
@@
 Following the success of this small TPC on the NuMI beam, a wide program of LArTPCs on neutrino beams has flourished.
-The construction of  LArTPCs as near and far detectors at different baseline allows for the exploration of different fundamental questions in neutrino physics today.
+The construction of  LArTPCs as near and far detectors at different baseline allows for the exploration of some of the fundamental questions in neutrino physics today illustrated in section \ref{ch:questions}.
 
-The Short-Baseline Neutrino (SBN) \cite{Antonello:2015lea} program at Fermilab  is tasked with conclusively addressing the ``LSND and MiniBooNE anomalies" \cite{Aguilar:2001ty, Athanassopoulos:1997pv, Aguilar-Arevalo:2013pmq}, whose $\nu_e$ and $\bar\nu_e$ appearance results may be interpreted under the assumption of a new sterile neutrino.
+The Short-Baseline Neutrino (SBN) \cite{Antonello:2015lea} program at Fermilab  is tasked with conclusively assess the nature of the ``LSND and MiniBooNE anomalies" \cite{Aguilar:2001ty, Athanassopoulos:1997pv, Aguilar-Arevalo:2013pmq}, resolving the mystery of  sterile neutrinos at the eV$^2$ scale.
 The SBN program entails three surface LArTPCs positioned on the Booster Neutrino Beam at different distances from the neutrino production in oder to fully exploit  the L/E dependence of the oscillation pattern:  SBND (100 m from the decay pipe), MicroBooNE (450 m), and ICARUS (600 m).
@@
 Besides the oscillation analysis, the second main goals of SBN is to perform an extensive campaign of neutrino cross section measurements in argon.
-Given the importance of nuclear effects in (relatively) heavy materials, both the oscillation analysis of the SBN program and the measurements of neutrino properties in DUNE will benefit from such a campaign.
+Given the importance of nuclear effects in (relatively) heavy materials, as discussed in section \ref{ch:NuInt}, both the oscillation analysis of the SBN program and the measurements of neutrino properties in DUNE will benefit from such a campaign.
 
@@
 
-%\subsection{SBN Goals}
-%\subsection{Neutrino Interactions and Detection }
 \subsection{Prospects for LArTPCs in GUT Physics: DUNE}
@@
 In order to showcase LArIAT's utility to SBN and DUNE, we illustrate briefly two comparisons as examples: one regarding neutrino interactions and the second regarding proton decay studies.\\
-The left side of figure \ref{fig:NuSimulation} shows the distribution of products in momentum spectrum and particle type as simulated in a $\nu_e$ CC interaction in DUNE (according to \cite{LeiguideOliveira:1953730}) as  a comparative example to the momentum distribution of particles in the LArIAT beamline -- shown on the right side of figure \ref{fig:NuSimulation}.
+The left side of figure \ref{fig:NuSimulation} shows the distribution of products in momentum spectrum and particle type as simulated in a $\nu_e$ CC interaction in DUNE (according to \cite{LeiguideOliveira:1953730}); the range of these distribution is to compare with  the momentum distribution of light particles in the LArIAT beamline -- shown on the right side of figure \ref{fig:NuSimulation}.
 The momentum spectrum in the LArIAT beamline for electrons, muons and pions -- the most abundant particles produced in  a $\nu_e$ CC interaction -- covers a wide range of the expected momentum distribution in a neutrino event.
 
-The signature of a proton decay event in the ``LAr golden mode" is the presence of a single kaon of about 400 MeV in the detector; the momentum spectrum of the kaon pre and post FSI in such an event is shown on the left side of figure \ref{fig:PDKGENIE}.
+The signature of a proton decay event in the ``LAr golden mode" is the presence of a single kaon of about 400 MeV in the detector; the momentum spectrum of the kaon pre and post FSI in such an event as simulated by GENIE is shown on the left side of figure \ref{fig:PDKGENIE}.
 The right side of figure \ref{fig:PDKGENIE} shows the momentum spectrum of kaons in the LArIAT beamline.

### 2018-04-30 Chapter-3/3-TheLArIATExperiment.tex
s

@@
 LArIAT's particles history begins in the Fermilab accelerator complex with a beam of protons.
-The process of protons acceleration develops in gradual stages (see picture \ref{fig:Accelerator}): gaseous hydrogen is ionized in order to form H$^{-}$ ions; these ions are boosted to 750 keV by a Cockroft-Walton accelerator and injected to the Linac linear accelerator that increases their energy up to 400 MeV; then, H$^{-}$ ions pass through a carbon foil and lose the two electrons; the resulting protons are then injected into a rapid cycling synchrotron, called Booster; at this stage, protons reach 8 GeV of energy and are compacted into bunches; the next stage of acceleration is the Main Injector, a synchrotron which accelerates the bunches up to 120 GeV; in the Main Injector, several bunches are merged into one and used for the injection in the last stage.
+The process of protons acceleration develops in gradual stages (see picture \ref{fig:Accelerator}): gaseous hydrogen is ionized in order to form H$^{-}$ ions; these ions are boosted to 750 keV by a Cockroft-Walton accelerator and injected to the Linac linear accelerator that increases their energy up to 400 MeV; then, H$^{-}$ ions pass through a carbon foil and lose the two electrons; the resulting protons are then injected into a rapid cycling synchrotron, called Booster; at this stage, protons reach 8 GeV of energy and are compacted into bunches; the next stage of acceleration is the Main Injector, a synchrotron which accelerates the bunches up to 120 GeV; in the Main Injector, several bunches are merged into one and are ready for delivery.
 
@@
 The beam is split by electrostatic septa and delivered at different experimental halls all over the lab.
-A 120~GeV$/c$ primary proton beam with variable intensity is extracted in four-second ``spills" and sent to the Meson Center beam line.
+A 120~GeV primary proton beam with variable intensity is extracted in four-second ``spills" and sent to the Meson Center beam line.
 
@@
 The steel collimator selects particles produced with a $13^\circ$ production angle at the target down the beamline.
-The particles are then bent by  $~10^\circ$  through a pair of dipole magnets.
+The particles are then bent by roughly $~10^\circ$  through a pair of dipole magnets.
 By configuring the field intensity of the magnets we allow the particles of LArIAT's tertiary beam to span a momentum range from 0.2 to 1.4~GeV/c.
@@
 The instrumentation of  LArIAT tertiary beam and the TPC components have changed several times during the three years of LArIAT data taking.
-The following paragraphs describe the components operational during ``Run II", the data taking period relevant to the hadron cross section measurements.
+The following paragraphs describe the components operational during ``Run II", the data taking period relevant to the hadron cross section measurements considered in this thesis.
 
@@
 
-LArIAT uses a pair of identical Fermilab type ``NDB" electromagnets, recycled from the Tevatron's anti-proton ring, in a similar configuration used for the  MINERvA T-977 test beam calibration~\cite{MinervaTestbeam}).
+LArIAT uses a pair of identical Fermilab type ``NDB" electromagnets, recycled from the Tevatron's anti-proton ring, in a similar configuration used for the  MINERvA T-977 test beam calibration~\cite{MinervaTestbeam}.
 The magnets are a fundamental piece of the LArIAT beamline equipment, as they are used for both particle identification and momentum measurement before the LArTPC.
@@
 The data was collected using two hall probes (blue and green).
-We fit the readings with a cubic function (black) to average of measurements (red) given in the legend.}
+We fit the readings with a cubic function (black) to average of measurements (red) given in the legend \cite{detectorPaper}.}
 }
@@
 Cosmic ray tracks crossing both anode and cathode populate the events triggered this way.
-This particularly useful sample of tracks (which we can safely assume to be associated with ~5 GeV muons) can be used for many tasks; for example, we use anode-cathode piercing tracks to cross check the TPC electric field on data (see Appendix \ref{ch:AppendixB}), to calibrate the charge response of the TPC wires for the full TPC volume and to measure the electron lifetime in the chamber (see section \ref{ch:EnCalibration}).
+This particularly useful sample of tracks (which we can safely assume to be associated with ~5 GeV muons) can be used for many tasks; for example, we use anode-cathode piercing tracks to cross check the TPC electric field on data (see Appendix \ref{ch:AppendixB}), to calibrate the charge response of the TPC wires for the full TPC volume and to measure the electron lifetime in the chamber \cite{LArIATLifeTime}.
 
@@
 The paddles come in two sizes: the smaller version has bases 32.2~cm and 26.7~cm, and 61.0~cm height, while the bigger version has bases 33.2~cm and 27.0~cm, and $70.8~cm$ height.
- A Zener-diode Hamamatsu H5783 PMT collects the light from a wavelength-shifting optical fiber which runs along one of the long sides of each paddle.
+% A Zener-diode Hamamatsu H5783 PMT collects the light from a wavelength-shifting optical fiber which runs along one of the long sides of each paddle.
 A custom-made PMT Amplifier and Discrimination (PAD) circuit mounted at one end of the paddle collects signals from the PMTs and sends them to the Control and Concentrator Unit (CCU).
@@
 In particular, the modification  shown in Figure \ref{fig:LArIATCryoMods} were necessary to account for the beam of charged particles entering the TPC and to employ the new FTBT liquid argon purification system.
-We added a ``beam window'' on the front outer end cap and an ``excluder'' on the inner endcap, with the scope of minimizing the amount of dead material upstream of the TPC's active volume.
-Doing so, we reduced the amount of uninstrumented material before the TPC from $\sim$ 1.6 radiation lengths ($X_{0}$) (ArgoNeuT) to less than 0.3 $X_{0}$ (LArIAT).
+We added a ``beam window'' on the front outer end cap and an ``excluder'' on the inner endcap, with the scope of minimizing the amount of non-instrumented material upstream of the TPC's active volume.
+The amount of non-instrumented material in front of the TPC for LArIAT corresponds to $\sim$0.3 electron radiation lengths ($X_{0}$), to compare against the  $\sim 1.6 X_{0}$ of ArgoNeuT.
 To allow studies of the scintillation light, we added a side port feedthrough which enables the mounting of the light collection system, as well as the connections for the corresponding signal and high-voltage cables (see Section \ref{sec:TPCLight}).
@@
 The filtered argon reaches the inner vessel via a liquid feedthrough on the top of the cryostat.
-Argon is not recirculated in the system, rather it boils off and vent to the atmosphere.
+Argon is not recirculated in the system; rather, it boils off and vent to the atmosphere.
 During data taking, we replenish the argon in the cryostat several times per day to keep the TPC high voltage feedthrough and cold electronics always submerged.
@@
 \subsection{LArTPC: Charge Collection}\label{sec:TPCCharge}
-The LArIAT Liquid Argon Time Projection Chamber is a rectangular box of dimensions 47 cm (width) x 40 cm (height) x 90 cm (length), containing 170 liters of Liquid Argon.
+The LArIAT Liquid Argon Time Projection Chamber is a rectangular box of dimensions 47 cm (drift) x 40 cm (height) x 90 cm (length), containing 170 liters of Liquid Argon.
 The LArTPC three major subcomponents are 
@@
 A network of voltage-dividing resistors connected to the field cage rings steps down the high voltage from the cathode to form a uniform electric field.
-The electric field over the entire TPC drift volume is  486 V/cm (see \ref{ch:AppendixB}).
+The electric field over the entire TPC drift volume is  486 V/cm, as measured in appendix \ref{ch:AppendixB}.
 The  maximum drift length, i.e.
@@
 \subsubsection{Wire planes}
-The wire planes measure the charge deposited in the TPC active volume. The drifting charge induces a current on the wire of the inner planes and it is collected on the collection plane wires.
 LArIAT counts three wire planes separated by 4 mm spaces: in order of increasing distance from the cathode, they are the shield, the induction and the collection plane.
@@
 Liquid argon scintillates in vacuum-ultraviolet (VUV) range at 128 nm; since cryogenic PMTs are not sensitive to VUV wavelengths, we need to shift the light in a region visible to the PMTs.
-In LArIAT, the wavelength shifting is achieved by installing on the four walls of the TPC highly-reflective VIKUITY dielectric substrate foils coated with a thin layer of tetraphenyl-butadiene (TPB).
+In LArIAT, the wavelength shifting is achieved by installing highly-reflective VIKUITY dielectric substrate foils coated with a thin layer of tetraphenyl-butadiene (TPB)  on the four unbiased walls of the TPB .
 The scintillation light interaction with the TPC emits one or more visible photons, we are then reflected into the chamber.
@@
 In doing so, some of the VUV scintillation light converts into visible right at the sensor faces, keeping information on the direction of the light source.
-Information about the light directionality is lost for light reflected on foils, as the reflection is uniform in angle.
+Information about the light directionality is hindered for the light reflected on foils, as the reflection is uniform in angle.
 %For Run-II, the voltage dividers for the PMTs were configured for positive bias with a DC-coupled anode (AC-coupled anode with grounded photocathode) to minimize induced noise on the TPC wires and modify the PMT bases accordingly.
@@
 
-LArIAT receives three logic from the Fermilab accelerator complex related to the beam timing which we use a input triggers: a pulse just before before the beam, a pulse indicating beam-on, and a beam-off pulse.
+LArIAT receives three logic signals from the Fermilab accelerator complex related to the beam timing which we use as input triggers: a pulse just  before the beam, a pulse indicating beam-on, and a beam-off pulse.
 
@@
 
-Fundamental inputs to the trigger card come from the TOF (see Sec.~\ref{sec:TOF}) and the wire chambers (see Sec.~\ref{sec:MWPC}), as activity in these systems points to the presence of a charged particle in tertiary beam line.
+Fundamental inputs to the trigger card come from the TOF (see section~\ref{sec:TOF}) and the wire chambers (see section~\ref{sec:MWPC}), as activity in these systems points to the presence of a charged particle in tertiary beam line.
 In particular, the discriminated pulses from the TOF PMTs form a NIM logic pulse for the trigger logic.
@@
 
-We use the signal from the cryogenic PMTs (see Section \ref{sec:PhotonSystem}) to form several interesting triggers.
+We use the signal from the cryogenic PMTs (see Section \ref{sec:TPCLight}) to form several interesting triggers.
 The coincidence of signal from all the PMT pulses within $\sim$20~ns is an indication of ionizing radiation in the TPC and forms a trigger input.
@@
 
-The online Event Viewer displays a two dimensional representation of LArIAT TPC events on both the Induction and the Collection planes in near real time.
+The online Event Viewer displays a two dimensional representation (Wire vs Time) of LArIAT TPC events  on both the Induction and the Collection planes in near real time.
 The raw pulses collected by the DAQ on each wire are plotted as a function of drift time, resulting in an image of the TPC event easily readable by the shifter.

### 2018-04-30 Chapter-5/5-SamplesPreparation.tex
s

@@
 For each data event, we obtain the particle position ($X, Y$) at WC4 directly from the data measurement.
-On the contrary, we calculate the components of the momentum using the beamline measurement of the momentum magnitude (see section \ref{sec:MWPCfunc}) in conjunction with the hits on WC3 and WC4 to determine the direction of the momentum vector.
+On the contrary, we calculate the components of the momentum using the beamline measurement of the momentum magnitude (see section \ref{sec:MWPCfunc}) in conjunction with the hits on WC3 and WC4 to determine the direction of the momentum vector, as described in \ref{sec:MWPCfunc}.
 The momentum and position of the selected data is sampled thousand of times through a 5-dimensional hit-or-miss sampling procedure.

### 2018-05-01 Chapter-3/3-TheLArIATExperiment.tex
s

@@
 
-\subsubsection{TOF functionality}\label{sec:MWPCfunc}
+\subsubsection{TOF functionality}\label{sec:TOFfunc}
 
@@
 Punch-thorough and MuRS hits are formed utilizing the digital discriminator signals under threshold at a given time, where we obtain the threshold for each PMT directly on data distributions.
+
+It is worth mentioning here the presence of an additional scintillation paddle between WC4 and the downstream paddle of the TOF system, called halo.
+The halo is a 39x38x1 cm$^3$ paddle with a 6.5 cm radius hole in the center, whose original function was to reject  beam particles slightly offset from the beamline center.
+Data from this paddle turned out to be unusable, so our data events include both particle going through the halo scintillation material or through the halo hole.
 

### 2018-05-01 Chapter-4/4-InteractionsInLAr.tex
s

--- 2018-03-19 Chapter-4/3-KaonInteractionsInLAr.tex
@@
-\chapter{Hadron Interactions in Argon: Cross Section}\label{ch:Interactions}
-\section{Literature Review}
+\chapter{Total Hadronic Cross Section Measurement Methodology}\label{ch:Interactions}
 
@@
 \section{How to Measure a Hadron Cross Section in LArIAT}\label{ch:methodology}
-We use both the LArIAT  beamline detectors and the LArTPC information to measure hadronic cross sections in argon.
+We use both the LArIAT  beamline detectors and the LArTPC information to measure the total hadronic cross sections in argon.
 Albeit with small differences, both the  $\pi^{-}$~-~Ar and K$^{+}$~-~Ar total hadronic cross section measurements rely on the same procedure described in details in the following paragraphs: we select the particle of interest using a combination of beamline detectors and TPC information (paragraph \ref{ch:ParticleSelectionMethod}), we perform a handshake between the beamline information and the TPC tracking to assure we are selecting the right TPC track (paragraph \ref{ch:WC2TPCMatchMethod}), and we apply the ``thin slice" method to get to the final result (paragraph \ref{ch:ThinSliceMethod}).
-We show a cross check of this method in paragraph \ref{ch:procedureTesting}.
+At the end of this chapter, we show a cross check of this method against MC truth information (paragraph \ref{ch:procedureTesting}).
 
@@
   \centering  
-\includegraphics[width=3in]{Chapter-4/Images/massRunIINeg.png}
+\includegraphics[width=3in]{Chapter-4/Images/face-smile-big.png}
 \end{minipage}
@@
 In fact, muons and electrons survive the selection on the beamline mass value.
-It is important to notice that the composition of the negative polarity beam is mostly pions, as discussed in \ref{ch:beamlineComposition}.
+It is important to notice that the composition of the negative polarity beam is mostly pions, as discussed in section \ref{ch:beamlineComposition}.
 Anyhow, we devise a selection on the TPC information to mitigate the presence of electrons in the sample used for the pion cross section.
@@
 interact with the medium, produce bremsstrahlung photons which pair convert into several short tracks.
-We provide details of this selection in section \ref{ch:electronCuts}.
+%We provide details of this selection in section \ref{ch:electronCuts}.
 
@@
 For the cross section analyses, we remove events with more than 4 tracks in the first 14 cm upstream portion of the TPC from the sample.
-\textcolor{red}{probably need to do a better job explaining pile up}
 
@@
 For each event passing the selection on its beamline information we need to identify the track inside the TPC corresponding to the particle which triggered the beamline detectors, a procedure we refer to as ``WC to TPC match" (WC2TPC for short).
-In general, the TPC tracking algorithm will reconstruct more than one track in the event, partially due to the fact that hadrons interact in the chamber, as shown in figure \ref{fig:pionInteraction}, and partially because of pile up particles during the triggered TPC drift time, as shown in figure \ref{fig:pionPileUp}.
+In general, the TPC tracking algorithm will reconstruct more than one track in the event, partially due to the fact that hadrons interact in the chamber and partially because of pile up particles during the triggered TPC drift time, as shown in figure \ref{fig:kaonInteraction}.
 \textcolor{red}{ADD EVENT DISPLAYS}
+
+\begin{figure}
+  \centering  
+\includegraphics[width=3in]{Chapter-4/Images/face-smile-big.png}
+\caption{Kaon candidate event: on the right, event display showing raw quantities; on the left, event display showing reconstructed tracks.
+In the reconstructed event display, different colors represent different track objects.
+A kink is visible in the kaon ionization, signature of a hadronic interaction: the tracking correctly stops at the kink position and two tracks are formed.
+An additional pile-up track is so present in the event (top track).}
+\label{fig:kaonInteraction}
+\end{figure}
 
@@
 \subsubsection{Cross Sections on Thin Target}
-Cross section measurements on a thin target have been the bread and butter of nuclear and particle experimentalists since the Rutherford experiments \textcolor{red}{NEED CITATION}.
+Cross section measurements on a thin target have been the bread and butter of nuclear and particle experimentalists since the Geiger-Marsden experiments \cite{Geiger1909}.
 At their core, this type of experiments consists in shooting a beam of particles with a known flux on a thin target and recording the outgoing flux.
@@
 \subsubsection{Not-so-Thin Target: Slicing the Argon}
-The LArIAT TPC, with its 90 cm of length, is not a thin target.
-\textcolor{red}{Find expected interaction length for hadrons and kaons}.
-However, the fine-grained tracking of the LArIAT LArTPC allows us to treat the argon volume as a sequence of many adjacent thin targets.
+The LArIAT TPC, with its 90 cm of length, is not a thin target. However, the fine-grained tracking of the LArIAT LArTPC allows us to treat the argon volume as a sequence of many adjacent thin targets.
 
-As described in section \ref{sec:experimentDescription}, LArIAT wire planes count 240 wires each.
+As described in Chapter \ref{sec:experimentDescription}, LArIAT wire planes count 240 wires each.
 The wires are oriented at +/- $60^{\circ}$ from the vertical direction at 4 mm spacing, while the beam direction is oriented 3 degrees off the $z$ axis in the $XZ$ plane.
-\textcolor{red}{review this math} The wires collect signals proportional to the energy loss of the hadron along its path in a  $\delta${\emph{X}} = 4 mm/sin($60^{\circ}$) $\approx$ 4.7~mm slab of liquid argon.
+The wires collect signals proportional to the energy loss of the hadron along its path in a  $\delta${\emph{X}} = 4 mm/sin($60^{\circ}$) $\approx$ 4.7~mm slab of liquid argon.
 Thus, one can think to slice the TPC into many thin targets of $\delta${\emph{X}} = 4.7~mm thickness along the direction of the incident particle.
@@
 \end{equation}
-where $E_{loss}$ is a correction for the energy loss in the dead material between the beamline and the TPC front face (more on \ref{sec:Eloss}).
+where $E_{loss}$ is a correction for the energy loss in the dead material between the beamline and the TPC front face.
 The  energy of the hadron at the each slab is determined by subtracting the energy released by the particle in the previous slabs.
@@
 The $\pi^{-}$-Ar and K$^{+}$-Ar total hadronic cross section implemented in Geant4 can be used as a tool to validate the measurement methodology.
-We describe here a closure test done on Monte Carlo to prove that the methodology of slicing the TPC retrieves the underlying cross section distribution implemented in Geant4 within the statistical error.
+We describe here a closure test done on Monte Carlo to prove that the methodology of slicing the TPC retrieves the underlying cross section distribution implemented in Geant4 within the statistical uncertainty.
 %under the working assumption of perfect reconstruction.
@@
 
-For the validation test, we fire about 390000 pions and 140000 kaons inside the LArIAT TPC active volume using the DDMC (see sec \ref{ch:DDMC}).
-We apply  the thin-sliced method on using true quantities to calculate the hadron kinetic energy at each slab in order to decouple reconstruction effects to eventual issues with the methodology.
+For the validation test, we fire about 390000 pions and 140000 kaons inside the LArIAT TPC active volume using the Data Driven Monte Carlo (see section \ref{sec:DDMC}).
+We apply  the thin-sliced method using only true quantities to calculate the hadron kinetic energy at each slab in order to decouple reconstruction effects to eventual issues with the methodology.
 For each slab of 4.7 mm length on the path of the hadron, we integrate the true energy deposition as given by the Geant4 transportation model.
@@
 
-Figure \ref{fig:TrueMCXS} shows the total hadronic cross section for argon implemented in Geant4 10.01.p3 (solid lines) overlaid with the true MC cross section as obtained with the sliced TPC method (markers) for pions on the left and kaons on the right; the total cross section is shown in green,  the elastic cross section in blue and the inelastic cross section in red.
+Figure \ref{fig:TrueMCXS2} shows the total hadronic cross section for argon implemented in Geant4 10.01.p3 (solid lines) overlaid with the true MC cross section as obtained with the sliced TPC method (markers) for pions on the left and kaons on the right; the total cross section is shown in green,  the elastic cross section in blue and the inelastic cross section in red.
 The nice agreement with the Geant4 distribution and the cross section  obtained with the sliced TPC method gives us confidence in the  validity of the methodology.
@@
 \end{minipage}
-\label{fig:TrueMCXS}
 \caption{Hadronic cross sections for $\pi^-$-Ar (left) and K$^+$-Ar (right) implemented in Geant4 10.01.p3 (solid lines) overlaid the true MC cross section as obtained with the sliced TPC method (markers).
 The total cross section is shown in green,  the elastic cross section in blue and the inelastic cross section in red.}
-%\par
-%\begin{minipage}[t]{.53\textwidth}
-%\caption{total hadronic cross section for carbon implemented in Geant4  10.01.p3  with overlaid with the Bugg %and Frideman data.}
-%\label{fig:TrueCarbon}
-%\end{minipage}%
-%\begin{minipage}[t]{.5\textwidth}  
-%\caption{Hadronic cross sections for argon implemented in Geant4 10.01.p3 (solid lines) overlaid the true MC cross section as obtained with the sliced TPC method (markers).
-The total cross section is shown in green,  the elastic cross section in blue and the inelastic cross section in red.}
-%\label{fig:TrueArgon}
-%\end{minipage}  
+\label{fig:TrueMCXS2}
 \end{figure}

### 2018-05-01 Chapter-3/3-TheLArIATExperiment.tex
s

@@
 In this chapter, we describe the LArIAT experimental setup.
-We start by illustrating the journey of the charge particles in the Fermilab accelerator complex, from the gaseous thermal hydrogen at the Fermilab ion source to the delivery of the LArIAT tertiary beam at MC7.
+We start by illustrating the journey of the charged particles in the Fermilab accelerator complex, from the gaseous thermal hydrogen at the Fermilab ion source to the delivery of the LArIAT tertiary beam at MC7.
 We  then describe the LArIAT beamline detectors, the LArTPC, the DAQ and the monitoring system.
@@
 %\section{LArIAT \& the Intensity Frontier}
-\section{The Particles Path to LArIAT}
+\section{The Particles' Path to LArIAT}
 
-LArIAT's particles history begins in the Fermilab accelerator complex with a beam of protons.
-The process of protons acceleration develops in gradual stages (see picture \ref{fig:Accelerator}): gaseous hydrogen is ionized in order to form H$^{-}$ ions; these ions are boosted to 750 keV by a Cockroft-Walton accelerator and injected to the Linac linear accelerator that increases their energy up to 400 MeV; then, H$^{-}$ ions pass through a carbon foil and lose the two electrons; the resulting protons are then injected into a rapid cycling synchrotron, called Booster; at this stage, protons reach 8 GeV of energy and are compacted into bunches; the next stage of acceleration is the Main Injector, a synchrotron which accelerates the bunches up to 120 GeV; in the Main Injector, several bunches are merged into one and are ready for delivery.
+LArIAT's particle history begins in the Fermilab accelerator complex with a beam of protons.
+The process of proton acceleration develops in gradual stages (see picture \ref{fig:Accelerator}): gaseous hydrogen is ionized in order to form H$^{-}$ ions; these ions are boosted to 750 keV by a Cockroft-Walton accelerator and injected into the linear accelerator (Linac) that increases their energy up to 400 MeV; then, H$^{-}$ ions pass through a carbon foil and lose the two electrons; the resulting protons are then injected into a rapid cycling synchrotron, called the Booster; at this stage, protons reach 8 GeV of energy and are compacted into bunches; the next stage of acceleration is the Main Injector, a synchrotron which accelerates the bunches up to 120 GeV; in the Main Injector, several bunches are merged into one and are ready for delivery.
 
-The Fermilab accelerator complex works in supercycles of roughly 60 seconds in duration.
-The beam is split by electrostatic septa and delivered at different experimental halls all over the lab.
+The Fermilab accelerator complex works in supercycles of  60 seconds in duration.
 A 120~GeV primary proton beam with variable intensity is extracted in four-second ``spills" and sent to the Meson Center beam line.
+%The beam is split by electrostatic septa and delivered to different experimental halls all over the lab.
 
-LArIAT's home at Fermilab is the Fermilab Test Beam Facility (FTBF), where the experiment characterizes a beam of charge particles downstream from the Meson Center beam line. 
-Here, the primary beam is focused onto a tungsten target to create LArIAT's secondary beam.
-The composition of the secondary particle beam is mainly positive pions.
+LArIAT's home at Fermilab is the Fermilab Test Beam Facility (FTBF), where the experiment characterizes a beam of charged particles in the Meson Center beam line. 
+At FTBF, the primary beam is focused onto a tungsten target to create LArIAT's secondary beam.
+The secondary beamline is set such that the composition of the secondary particle beam is mainly positive pions.
 The momentum peak of the secondary beam was fixed at 64~GeV/c for the LArIAT data considered in this work, although the beam is tunable in momentum between 8-80\,GeV/c; this configuration of the secondary beamline assured a stable beam delivery at the LArIAT experimental hall.
@@
 The secondary beam impinges then on a copper target within a steel collimator inside the LArIAT experimental hall (MC7) to create the LArIAT tertiary beam, (shown in  Fig.~\ref{fig:tert-layout}).
-The steel collimator selects particles produced with a $13^\circ$ production angle at the target down the beamline.
+The steel collimator selects particles produced with a $13^\circ$ production angle.
 The particles are then bent by roughly $~10^\circ$  through a pair of dipole magnets.
@@
 The polarity of the magnet is also configurable and determines the sign of the beamline particles which are focused on the LArTPC.
-If the magnets polarity is positive the tertiary beam composition counts mostly pions and protons with a small fraction of electrons, muons, and kaons.
-It is the job of the LArIAT beamline detectors to select the particles polarity,  to perform particle identification in the beamline and to measure the momentum of the tertiary beam particles before they get to the LArTPC.
+If the magnet polarity is positive the tertiary beam composition is mostly pions and protons with a small fraction of electrons, muons, and kaons.
+It is the job of the LArIAT beamline equipment to select the particles polarity,  to perform particle identification in the beamline and to measure the momentum of the tertiary beam particles before they get to the LArTPC.
 The LArIAT detectors are described in the following paragraphs.
@@
 \includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{Chapter-3/Images/AcceleratorFNAL.png}
-\caption{Layout of Fermilab Acellerator complex.}
+\caption{Layout of Fermilab Accelerator complex.}
 \label{fig:Accelerator}
@@
 \caption{Bird's eye view of the LArIAT tertiary beamline.
-In grey: upstream and downstream collimators; in yellow: bending magnets; in red: wire chambers; in blue: time of flight; in green: liquid argon TPC volume; in maroon: muon range statck.}
+In grey: upstream and downstream collimators; in yellow: bending magnets; in red: multi wire proportional chambers; in blue: time of flight; in green: liquid argon TPC volume; in maroon: muon range statck.}
 \label{fig:tert-layout}
@@
 The key components of the tertiary beamline instrumentation for the hadron cross section analyses are the two bending magnets, a set of four wire chambers (WCs) and two time-of-flight scintillating paddles (TOF) and, of course, the LArTPC.
-The magnets determine the polarity of the particles in the tertiary beam; the combination of magnets and wire chambers determines the particles' momentum, which is used to determine the particle species in conjunction with the TOF.
+The magnets determine the polarity of the particles in the tertiary beam; the combination of magnets and wire chambers determines the particles' momenta, which is used to determine the particle species in conjunction with the TOF.
 A muon range stack downstream from the TPC and two sets of cosmic paddles configured as a telescope surrounding the TPC are also used for calibration purposes.
@@
 LArIAT uses a pair of identical Fermilab type ``NDB" electromagnets, recycled from the Tevatron's anti-proton ring, in a similar configuration used for the  MINERvA T-977 test beam calibration~\cite{MinervaTestbeam}.
-The magnets are a fundamental piece of the LArIAT beamline equipment, as they are used for both particle identification and momentum measurement before the LArTPC.
+The magnets are a fundamental piece of the LArIAT beamline equipment, as they are used for the selection of the particle polarity and for the momentum measurement before the LArTPC.
 The sign of the current in the magnets allows us to select either positively or negatively charged particles; the value of the magnetic field is used in the momentum determination and in the subsequent particle identification.
@@
 Each magnet is a box with a rectangular aperture gap in the center to allow for the particle passage.
-The magnet aperture measures 14.224~cm in height, 31.75~cm in width, and  46.67~cm in length.
+The magnet aperture measures 14.22~cm in height, 31.75~cm in width, and  46.67~cm in length.
 Since the wire chambers aperture ($\sim$12.8~cm$^2$) is smaller than the magnet aperture, only the central part of the magnet gap is utilized.
@@
 Albeit advantageous to enrich the tertiary beam composition with high mass particles such as kaons, we never pushed the magnets current over 100 A, not to incur in overheating.
-During operation, we operated a air and water cooling system on the magnets and we remotely monitored the magnets temperature.
+During operation, we operated an air and water cooling system on the magnets and we remotely monitored the magnet temperatures.
 
@@
 \caption{
-{One of the four Multi Wire Proportional Chambers (WC) used in the LArIAT tertiary beamline and relative read-out electronics.}
+{One of the four Multi Wire Proportional Chambers (WC) used in the LArIAT tertiary beamline and related read-out electronics.}
 }
@@
 The geometry of one chamber is shown in Figure~\ref{fig:wirechamber}: the WC effective aperture is a square of  12.8~cm perpendicular to the beam direction.
-Inside the chamber, the 128 horizontal and 128 vertical wires hang at a distance of 1~mm from each other in a mixture of 85\% Argon and 15\% isobutane gas.
+Inside the chamber, the 128 horizontal and 128 vertical wires strung at a distance of 1~mm from each other in a mixture of 85\% Argon and 15\% isobutane gas.
 The WC operating voltage is between 2400~V and 2500~V.
@@
 The TDC time resolution is 1.18~ns/bin and can accept 2 edges per 9~ns.
-The maximum event rate acceptable by the chamber system is of 1 MHz: this rate is not a limiting factor considering that \textcolor{red}{the rate of the tertiary particle beam at the first wire chamber is estimated to be less than 15 kHz}.
+The maximum event rate acceptable by the chamber system is 1 MHz: this rate is not a limiting factor considering that \textcolor{red}{the rate of the tertiary particle beam at the first wire chamber is estimated to be less than 15 kHz}.
 A full spill of data occurring once per supercycle is stored on the TDC board memory at once and read out by a specially designed controller.
@@
 where $B_{eff}$ is the effective maximum field in a square field approximation,  $L_{eff}$ is the effective length of both magnets (twice the effective length of one magnet), $\theta_{US}$ is the angle off the $z$ axis of the upstream trajectory, $\theta_{DS}$ is the angle off the $z$ axis of the downstream trajectory  and  3.3~$c^{-1}$ is the conversion factor from [T$\cdot$m] to [MeV/c].
-By using the hit positions on the third and fourth wire chamber, we estimate the azimutal and polar angles of the particle trajectory, and we are able to calculate the other components of the momentum.
+By using the hit positions on the third and fourth wire chamber, we estimate the azimuthal and polar angles of the particle trajectory, and we are able to calculate the other components of the momentum.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-Two scintillator paddles, one upstream to the first set of WCs and one downstream to the second set of WCs  form LArIAT  time-of-flight (TOF) detector system.
+Two scintillator paddles, one upstream of the first set of WCs and one downstream of the second set of WCs  form LArIAT  time-of-flight (TOF) detector system.
 
 The upstream paddle is made of a 10 x 6 x 1~cm scintillator piece, read out by two PMTs mounted on the beam left side which collect the light from light guides mounted on all four edges of the scintillator.
-The downstream paddle is a   14 x 14 x 1~cm scintillator piece read out by two PMTs on the opposite ends of the scintillator.
-The relatively thin width on the beamline direction minimizes energy loss of the particles coming from the target in the scintillator material.
+The downstream paddle is a   14 x 14 x 1~cm scintillator piece read out by two PMTs on the opposite ends of the scintillator, as shown in figure \ref{fig:TOF}.
+The relatively thin width in the beamline direction minimizes energy loss of beam particles traveling through the scintillator material.
 
@@
 The time of the pulses was calculated utilizing an oversampled template derived from the data itself.
-We take the pulse pedestal from samples far from the pulse and subtract it to the pulse amplitude.
-We then stretch vertically a template to match the pedestal-subtracted pulse amplitude and we move it horizontally to find the time.
+We take the pulse pedestal from samples far from the pulse and subtract it from the pulse amplitude.
+We then vertically stretch  a template to match the pedestal-subtracted pulse amplitude and we move it horizontally to find the time.
 With this technique, we find a pulse time-pickoff resolution better than 100 ps.
 The pulse pile up is not a significant problem given the TOF timing resolution and the rate of the particle beam.
-Leveraging on the pulses width uniformity of any given PMT,  we flag events where two pulses overlap as closely in time as 4 ns with an 90\% efficiency according to simulation.
+Leveraging on the pulses width uniformity of any given PMT,  we flag events where two pulses overlap as closely in time as 4 ns with a 90\% efficiency according to simulation.
 
@@
 However, a time spread of approximately 300~ps is present in both the upstream and downstream detectors, likely due to transit time jitter in the PMTs themselves.
-There is no evidence of systematic timing drift over long data-taking periods such as 3-4 months: the maximum variation of the average time differences between pairs of PMTs reading out the same scintillator is of the order of 150~ps.
+%There is no evidence of systematic timing drift over long data-taking periods such as 3-4 months: the maximum variation of the average time differences between pairs of PMTs reading out the same scintillator is of the order of 150~ps.
 
@@
 \caption{Image of the down stream time of flight paddle, PMTs and relative support structure before mounting. } 
-\label{pic:cosmicpaddle}
+\label{fig:TOF}
 \end{figure}
@@
 
-The punch-thorough and the muon range stack (MuRS) detectors are located downstream of the TPC.
+The punch-through and the muon range stack (MuRS) detectors are located downstream of the TPC.
 These detectors provide a sample of  TPC crossing tracks without relying on TPC information and can be used to improve particle ID for  muons and pions with momentum higher than  450 MeV/c.
 
-The punch-thorough is simple sheet of scintillator material, read out by two PMTs. 
+The punch-through is simple sheet of scintillator material, read out by two PMTs. 
 The MuRS is a segmented block of steel with four slots instrumented with scintillation bars.
-The four steel layers in front of each instrumented slot are 2 cm, 2 cm, 14 cm and 16 cm wide in the beam direction.
+The four steel layers in front of each instrumented slot are 2 cm, 2 cm, 14 cm and 16 cm deep in the beam direction.
 Each instrumented slot is equipped with four scintillation bars each, positioned horizontally in the direction orthogonal to the beam.
@@
 
-The signals from both the punch-thorough and the MuRS PMTs are digitized in the CAEN V1740, same as the TPC; the details of this discriminator are laid out in~\ref{sec:TPCCharge}.
-It is worth noticing that the sampling time of the CAEN V1740 is slow (of the order of 128 ns), so pulse shape information from the PMT is lost.
-Punch-thorough and MuRS hits are formed utilizing the digital discriminator signals under threshold at a given time, where we obtain the threshold for each PMT directly on data distributions.
+The signals from both the punch-thorough and the MuRS PMTs  are sent to a NIM discriminator.
+If the signal crosses the discriminator threshold, it is digitized in the CAEN V1740, same as the TPC.
+The sampling time of the CAEN V1740 is slow (of the order of 128 ns) and that the pulse shape information from the PMT is lost.
+A Punch-thorough and MuRS signal will then be simply a ``hit" at a given time in the beamline event.
 
@@
 \subsection{Cryogenics and Argon Purity}\label{ch:Cryo}
-LArIAT repurposed the ArgoNeuT cryostat \cite{ArgoNeuT-det} in order to use it in a beam of charge particles, and added a new process piping and a new liquid argon filtration system in FTBF.
+LArIAT repurposed the ArgoNeuT cryostat \cite{ArgoNeuT-det} in order to use it in a beam of charged particles, and added a new process piping and a new liquid argon filtration system in FTBF.
 %The main modifications to the ArgoNeuT cryostat are the addition of a beam flange and a port for the light collection system,
-Inside the LArIAT experimental hall, the cryostat sits on the beam of charge particles with its horizontal main axis oriented parallel to the beam.
+Inside the LArIAT experimental hall, the cryostat sits on the beam of charged particles with its horizontal main axis oriented parallel to the beam.
 

### 2018-05-01 Chapter-3/3-TheLArIATExperiment.tex
s

@@
 A muon range stack downstream from the TPC and two sets of cosmic paddles configured as a telescope surrounding the TPC are also used for calibration purposes.
+A couple of Aerogel Cherenkov counters, which we will not describe here as they are not used in the hadron cross section measurements,  completes the beamline instrumentation.
 
@@
 The four signals  from the vertical paddles along one of the body diagonals of the TPC are combined in a logical ``AND''.
-This allows to select cosmic muons crossing the TPC along one of its diagonals.
-Cosmic ray tracks crossing both anode and cathode populate the events triggered this way.
-This particularly useful sample of tracks (which we can safely assume to be associated with ~5 GeV muons) can be used for many tasks; for example, we use anode-cathode piercing tracks to cross check the TPC electric field on data (see Appendix \ref{ch:AppendixB}), to calibrate the charge response of the TPC wires for the full TPC volume and to measure the electron lifetime in the chamber \cite{LArIATLifeTime}.
+This allows to select track due to cosmic muons at the ground level crossing the TPC along one of its diagonals.
+Cosmic ray muons whose average energy is in the few GeV range crossing both anode and cathode populate the events triggered this way.
+This particularly useful sample of tracks is associated can be used for many tasks; for example, we use anode-cathode piercing tracks to cross check the TPC electric field on data (see Appendix \ref{ch:AppendixB}), to calibrate the charge response of the TPC wires for the full TPC volume and to measure the electron lifetime in the chamber \cite{LArIATLifeTime}.
 
@@
 %The main modifications to the ArgoNeuT cryostat are the addition of a beam flange and a port for the light collection system,
-Inside the LArIAT experimental hall, the cryostat sits on the beam of charged particles with its horizontal main axis oriented parallel to the beam.
+Inside the LArIAT experimental hall, the cryostat sits in the beam of charged particles with its horizontal main axis oriented parallel to the secondary beam, ~3$^\circ$ off axis from the tertiary beam
 
@@
 Purified liquid argon fills the inner vessel, while the outer volume provides insulation through a vacuum jacket equipped with layers of aluminized mylar superinsulation.
-The inner vessel is a cylinder of 130~cm length and 6.2~cm diameter, containing about 550~L of LAr, corresponding to a mass of 0.76 ton.
+The inner vessel is a cylinder of 130~cm length and 76.2~cm diameter, containing about 550~L of LAr, corresponding to a mass of 0.77 ton.
 We run the signal cables for the LArTPC and the high voltage feedthrough through a ``chimney'' at the top and mid-length of the cryostat.
@@
 
-Given the different scopes of the ArgoNeuT and LArIAT detectors, we made several modification to the ArgoNeuT cryostat in order to use it in LArIAT.
-In particular, the modification  shown in Figure \ref{fig:LArIATCryoMods} were necessary to account for the beam of charged particles entering the TPC and to employ the new FTBT liquid argon purification system.
-We added a ``beam window'' on the front outer end cap and an ``excluder'' on the inner endcap, with the scope of minimizing the amount of non-instrumented material upstream of the TPC's active volume.
+Given the different scopes of the ArgoNeuT and LArIAT detectors, we made several modifications to the ArgoNeuT cryostat in order to use it in LArIAT.
+In particular, the modifications  shown in Figure \ref{fig:LArIATCryoMods} were necessary to account for the beam of charged particles entering the TPC and to employ the new FTBF liquid argon purification system.
+We added a ``beam window'' on the front outer end cap and an ``excluder'' on the inner endcap, with the purpose of minimizing the amount of non-instrumented material upstream of the TPC's active volume.
 The amount of non-instrumented material in front of the TPC for LArIAT corresponds to $\sim$0.3 electron radiation lengths ($X_{0}$), to compare against the  $\sim 1.6 X_{0}$ of ArgoNeuT.
@@
 \includegraphics[scale=0.35]{Chapter-3/Images/CryoMods.png}
-\caption{Main modifications to the ArgoNeuT cryostat: $1)$ outlet for connection to the purification system at the bottom of the cryostat; $2)$ the ``beam-window'' on the outer endcap and ``excluder"  which reduce the amount of non-instrumented material before the TPC; $3)$ the side port to host  the  light collection system.}
+\caption{Main modifications to the ArgoNeuT cryostat: $1)$ outlet for connection to the purification system at the bottom of the cryostat; $2)$ the ``beam-window'' on the outer endcap and ``excluder"  which reduces the amount of non-instrumented material before the TPC; $3)$ the side port to host  the  light collection system.}
 \label{fig:LArIATCryoMods}
@@
 As in any other LArTPC, argon purity is a crucial parameter for LArIAT.
-Indeed, the presence of contaminants effects both the basic working principles of a LArTPC, as shown in section \ref{sec:LArTPCWorkingPrinciple}: electronegative contaminants such as oxygen and water decrease the number of ionization electrons collected on the wires after drifting through the volume.
+Indeed, the presence of contaminants affects both the basic working principles of a LArTPC, as shown in section \ref{sec:LArTPCWorkingPrinciple}: electronegative contaminants such as oxygen and water decrease the number of ionization electrons collected on the wires after drifting through the volume.
 In addition, contaminants such as Nitrogen decrease the light yield from scintillation light, especially in its slow component.
-In LArIAT, contaminations should not exceed the level of 100 parts per trillion (ppt).
+In LArIAT, contaminations should not exceed the level of 0.2 parts per billion (ppb).
 We achieve this level of purity in several stages.
@@
 The piping was cleaned to remove oil and grease before being welded into the system.
-LArIAT uses a filtration system designed for the Liquid Argon Purity Demonstrator (LAPD)~\cite{LAPD}: half of a 77~liter filter contains a 4A molecular sieve (Sigma-Aldrich~\cite{sigma-aldrich}) apt to remove mainly water, while the other half contains BASF~CU-0226~S, a highly dispersed copper oxide impregnated on a high surface area alumina, apt to remove mainly oxygen~\cite{basf}.
+LArIAT uses a filtration system designed for the Liquid Argon Purity Demonstrator (LAPD)~\cite{LAPD}: half of a 77~liter filter contains a 4A molecular sieve (Sigma-Aldrich~\cite{sigma-aldrich}) able to remove mainly water, while the other half contains BASF~CU-0226~S, a highly dispersed copper oxide impregnated on a high surface area alumina, apt to remove mainly oxygen~\cite{basf}.
 A single pass of argon in the filter is sufficient to achieve the necessary purity, unless the filter is saturated.
 In case the filter saturates, the media needs to be regenerated by using heated gas; this happened twice during the Run II period\footnote{We deemed the filter regeneration necessary every time the electron lifetime dropped under 100 \textmu s.}.
-The filtered argon reaches the inner vessel via a liquid feedthrough on the top of the cryostat.
-Argon is not recirculated in the system; rather, it boils off and vent to the atmosphere.
-During data taking, we replenish the argon in the cryostat several times per day to keep the TPC high voltage feedthrough and cold electronics always submerged.
+The electron lifetime during the full LArIAT data taking are shown in Figure \ref{fig:Elifetime}.
+The filtered argon reaches the inner vessel via a liquid feedthrough which is routed to the bottom of the cryostat.
+Argon is not recirculated in the system; rather, it boils off and vents to the atmosphere.
+During data taking, we replenish the argon in the cryostat every 6 hours to keep the TPC high voltage feedthrough and cold electronics always submerged.
 In fact, we constantly monitor the level, temperature, and pressure of the argon both in the commercial dewar and inside the cryostat during data taking.
@@
 A G10 plain sheet with copper metallization on one of the 40 x 90~cm inner surfaces forms the cathode. 
-A high-voltage feedthrough on the top of the LArIAT cryostat delivers the high voltage to the cathode; scope of the high voltage system (Figure~\ref{fig:HVScheme}) is to drift ionization electrons from the interaction of charged particles in the liquid argon to the wire planes.
+A high-voltage feedthrough on the top of the LArIAT cryostat delivers the high voltage to the cathode; the purpose of the high voltage system (Figure~\ref{fig:HVScheme}) is to drift ionization electrons from the interaction of charged particles in the liquid argon to the wire planes.
 The power supply used in this system is a Glassman LX125N16 ~\cite{GlassmanPS} capable of generating up to -125~kV and 16~mA of current, but operated at -23.5kV during LArIAT Run-II.
@@
 \subsubsection{Wire planes}
-LArIAT counts three wire planes separated by 4 mm spaces: in order of increasing distance from the cathode, they are the shield, the induction and the collection plane.
-The ``wire pitch", i.e., the distance between two consecutive wires in a given plane, is 4 mm.
+LArIAT Run-II has three wire planes separated by 4 mm spaces: in order of increasing distance from the cathode, they are the shield, the induction and the collection plane.
+The ``wire pitch", i.e., the distance between two adjacent wires in a given plane, is 4 mm.
 The shield plane counts 225 parallel wires of equal length oriented vertically.
-This plane is not connected with the read-out electronics; rather it shields the outer planes from extremely long induction signals due to the ionization chamber in the whole drift volume.
-As the shield plane acts almost like a Faraday cage, the shape of signals in the first instrumented plane (induction)  results easier to reconstruct.
+This plane is not connected with the read-out electronics; rather it shields the outer planes from extremely long induction signals due to the ionization in the whole drift volume.
+As the shield plane acts almost like a Faraday cage, the resulting shape of signals in the first instrumented plane (induction) is easier to reconstruct.
 Both the induction and collection planes count 240 parallel wires of different length oriented at 60$^\circ$ from the vertical with opposite signs.
@@
 LArIAT  performs the amplification stage directly in cold with  amplifiers %developed by the Brookhaven National Lab (BNL) and
-mounted on the TPC frame inside the liquid argon, achieving a remarkable Signal-to-Noise ratio.
-%The BNL ASICs adopted in LAriAT are designated as LArASIC, version 4-star.%The ASIC signals for each wire are then driven out of the vessel  to DAQ boards that act as waveform recorders.
-The signal from the ASICs are driven to the other end of the readout chain, to the CAEN V1740 digitizers.
+mounted on the TPC frame inside the liquid argon. The BNL ASICs adopted in LAriAT are designated as LArASIC, version 4-star and are the same used by the MicroBooNE experiment \cite{Acciarri2017}.
+%The ASIC signals for each wire are then driven out of the vessel  to DAQ boards that act as waveform recorders.
+The signal from the ASICs are driven to the other end of the readout chain, to the CAEN V1740 digitizers \cite{CAENV1740}.
 The CAEN V1740 has a 12 bit resolution and a maximum input range of 2~VDC, resulting in about 180 ADC count for a crossing MIP.
@@
 
-Liquid argon scintillates in vacuum-ultraviolet (VUV) range at 128 nm; since cryogenic PMTs are not sensitive to VUV wavelengths, we need to shift the light in a region visible to the PMTs.
-In LArIAT, the wavelength shifting is achieved by installing highly-reflective VIKUITY dielectric substrate foils coated with a thin layer of tetraphenyl-butadiene (TPB)  on the four unbiased walls of the TPB .
-The scintillation light interaction with the TPC emits one or more visible photons, we are then reflected into the chamber.
-Thus, the light yield increases and results more uniform across the TPC active volume, allowing the possibility of light-based calorimetry, currently under study.
+Liquid argon scintillates in vacuum-ultraviolet (VUV) range at 128 nm; since cryogenic PMTs are not sensitive to VUV wavelengths, we need to shift the light to a range that is visible to the PMTs.
+In LArIAT, the wavelength shifting is achieved by installing highly-reflective 3M VIKUITI dielectric substrate foils coated with a thin layer of tetraphenyl-butadiene (TPB) on the four unbiased walls of the TPC.
+The scintillation light interaction with the TPB emits one or more visible photons, which are then reflected into the chamber.
+Thus, the light yield increases and results in higher uniformity of light across the TPC active volume, allowing the possibility of light-based calorimetry, currently under study.
 
-For Run II, we coated both  the windows of the ETL PMT and SensL SiPM  with a thin layer of TPB.
+For Run II, we coated   the windows of the ETL PMT and the SensL SiPM  with a thin layer of TPB.
 In doing so, some of the VUV scintillation light converts into visible right at the sensor faces, keeping information on the direction of the light source.
@@
 The LArIAT DAQ and trigger system governs the read out of all the many subsystems forming LArIAT. 
-The CAEN V1495 module and its user-programmable FPGA  are the core of this system.
+The CAEN V1495 module \cite{CAEN1495} and its user-programmable FPGA  are the core of this system.
 Every 10~ns, this module checks for matches between sixteen logical inputs and user-defined patterns in the trigger menu; if it finds a match for two consecutive clock ticks, that trigger fires.
@@
 In particular, the discriminated pulses from the TOF PMTs form a NIM logic pulse for the trigger logic.
-We ask for a coincidence within a 20~ns window for all the pulses from the PMTs looking at the same scintillator block and use the coincidence between the upstream and downstream paddle to inform the trigger decision.
+We ask for a coincidence within a 20~ns window for all the pulses from the PMTs looking at the same scintillator block and use a delayed coincidence between the upstream and downstream paddle to inform the trigger decision.
 In order to form a coincidence between the upstream and downstream paddles, we delay the upstream paddle coincidence by 20~ns and widen it by 100~ns.
 The delay and widening are necessary to account for both  lightspeed particles and slower particles (high-mass) to travel the 6.5~m between the upstream and the downstream paddles.
-For the read out of the wire chambers, we use a total of sixteen multi-hit TDCs, four per chamber: two TDC per plane (horizontal and vertical), sixty-four wires per TDC.
+For the read out of the wire chambers, we use a total of sixteen multi-hit TDCs\cite{Sten}, four per chamber: two TDC per plane (horizontal and vertical), sixty-four wires per TDC.
 In each TDC, we keep the logical ``OR" for any signal over threshold from the sixty-four wires.
@@
 We use the signal from the cryogenic PMTs (see Section \ref{sec:TPCLight}) to form several interesting triggers.
-The coincidence of signal from all the PMT pulses within $\sim$20~ns is an indication of ionizing radiation in the TPC and forms a trigger input.
+The coincidence of signals from all the PMT pulses within $\sim$20~ns is an indication of ionizing radiation in the TPC and forms a trigger input.
 The coincidence of two subsequent scintillation logic pulses delayed by a maximum of  7~$\mu$s forms the Michel electron trigger.
@@
 \section{Control Systems}
-LArIAT is a complex ensemble of systems which needed to be monitored at once during data taking.
+LArIAT is a complex ensemble of systems which needed to be monitored simultaneously during data taking.
 We performed the monitoring of the systems operations with a slow control system, a DAQ monitoring system and a low level data quality monitoring described in the following sections.
@@
 \subsubsection{Slow Control}
-We used the Synoptic Java Web Start framework as a real-time display of subsystem conditions. Its simple 
-Graphical User Interface allowed us to change the operating parameters and to graph the trends of several variables of interest for all the tertiary beam detectors.
-Among the most important quantities monitored by Synoptic there are the level of argon in both the inner vessel and the external dewar, the operating voltages of cathode and wire planes, of the PMTs and SiPMs, and of the four wire chambers, as well as the magnets temperature.
+We used the Synoptic Java Web Start framework \cite{Synoptic} as a real-time display of subsystem conditions.
+Synoptic provides a Graphical User Interface that talks to the Fermilab Accelerator Control System via the ACNET protocol.
+Its simple GUI allowed us to change the operating parameters and to graph the trends of several variables of interest for all of the tertiary beam detectors.
+Among the most important quantities monitored by Synoptic there are the level of argon in both the inner vessel and the external dewar, the operating voltages of cathode and wire planes, of the PMTs and SiPMs, and of the four wire chambers, as well as the magnet temperatures.
 Figure \ref{fig:synoptics} shows an example of the monitoring system.
@@
 \subsubsection{Data Quality Monitoring}
-We employ two systems to ensure the quality of out data during data taking: the Near-Real-Time Data Quality Monitoring and the Event Viewer.
+We employ two systems to ensure the quality of our data during data taking: the Near-Real-Time Data Quality Monitoring and the Event Viewer.
 
@@
 of the TPC wires and PMTs of the beamline detectors, the hit occupancy and timing plots on the wire chambers, and number of data fragments recorded that are used to build a TPC event.
-Abnormal values for  low-level quantity in the data  activate a series of alarms in the DQM; this quick feedback on the DAQ and beam conditions is fundamental to assure a fast debugging of the detector and a very efficient data taking during beam uptime.
+Abnormal values for  low-level quantity in the data  activates a series of alarms in the DQM; this quick feedback on the DAQ and beam conditions is fundamental to assure a fast debugging of the detector and a very efficient data taking during beam uptime.
 
@@
 This tool guarantees a particularly good  check of the TPC operation which activate an immediate feedback for troubleshooting a number of issues.
-For example,  it easy for the shifter to spot high occupancy events and request a reduction of the primary beam intensity, or to spot a decrease of the argon purity which requires the regeneration of filters, or to catch the presence of electronic noise and reboot the ASICs.
+For example,  it is easy for the shifter to spot high occupancy events and request a reduction of the primary beam intensity, or to spot a decrease of the argon purity which requires the regeneration of filters, or to catch the presence of electronic noise and reboot the ASICs.
 An example of high occupancy event is shown in \ref{fig:highOcc}.

### 2018-05-01 Chapter-4/4-InteractionsInLAr.tex
s

@@
 %The comparison of the Geant4 thin- and thick-target cross section results demonstrates the power of the "sliced TPC" method for the measurement of the ($\pi^{\pm}$, Ar) cross section in LArIAT TPC geometry.
+section
 
@@
 We use both the LArIAT  beamline detectors and the LArTPC information to measure the total hadronic cross sections in argon.
-Albeit with small differences, both the  $\pi^{-}$~-~Ar and K$^{+}$~-~Ar total hadronic cross section measurements rely on the same procedure described in details in the following paragraphs: we select the particle of interest using a combination of beamline detectors and TPC information (paragraph \ref{ch:ParticleSelectionMethod}), we perform a handshake between the beamline information and the TPC tracking to assure we are selecting the right TPC track (paragraph \ref{ch:WC2TPCMatchMethod}), and we apply the ``thin slice" method to get to the final result (paragraph \ref{ch:ThinSliceMethod}).
-At the end of this chapter, we show a cross check of this method against MC truth information (paragraph \ref{ch:procedureTesting}).
+Albeit with small differences, both the  $\pi^{-}$~-~Ar and K$^{+}$~-~Ar total hadronic cross section measurements rely on the same procedure described in details in the following sections: we select the particle of interest using a combination of beamline detectors and TPC information (section \ref{ch:ParticleSelectionMethod}), we perform a handshake between the beamline information and the TPC tracking to assure we are selecting the right TPC track (section \ref{ch:WC2TPCMatchMethod}), and we apply the ``thin slice" method to get to the final result (section \ref{ch:ThinSliceMethod}).
+At the end of this chapter, we show a cross check of this method against MC truth information (section \ref{ch:procedureTesting}).
 
@@
 \subsubsection{Beamline events}\label{ch:beamlineDetectorsData}
-As will be clear in paragraph \ref{ch:ThinSliceMethod}, beamline particle identification and momentum measurement before entering the TPC are fundamental information for the hadronic cross sections measurements in LArIAT.
-Thus, we scan the LArIAT data to keep only events whose wire chamber and time of flight information is registered.
+As will be clear in section \ref{ch:ThinSliceMethod}, beamline particle identification and momentum measurement before entering the TPC are fundamental information for the hadronic cross sections measurements in LArIAT.
+Thus, we apply selections to the LArIAT data to keep only events whose wire chamber and time of flight information is registered.
 Additionally, we perform a check of the plausibility of the trajectory inside the beamline detectors: given the position of the hits in the four wire chambers, we make sure the particle trajectory does not cross any impenetrable material such as the collimator and the magnets steel.
@@
 We then apply the same WC to TPC matching algorithm as in data.
-We discard events with multiple WC2TPC matches. We use only TPC track matched to WC tracks in the cross section calculation.
+We discard events with multiple WC2TPC matches. We use only those TPC tracks that are matched to WC tracks in the cross section calculation.
 
@@
 Cross section measurements on a thin target have been the bread and butter of nuclear and particle experimentalists since the Geiger-Marsden experiments \cite{Geiger1909}.
-At their core, this type of experiments consists in shooting a beam of particles with a known flux on a thin target and recording the outgoing flux.
+At their core, these types of experiments consist in shooting a beam of particles with a known flux on a thin target and recording the outgoing flux.
 
@@
 
-As described in Chapter \ref{sec:experimentDescription}, LArIAT wire planes count 240 wires each.
+As described in Chapter \ref{sec:experimentDescription}, LArIAT wire planes consist of 240 wires each.
 The wires are oriented at +/- $60^{\circ}$ from the vertical direction at 4 mm spacing, while the beam direction is oriented 3 degrees off the $z$ axis in the $XZ$ plane.
 The wires collect signals proportional to the energy loss of the hadron along its path in a  $\delta${\emph{X}} = 4 mm/sin($60^{\circ}$) $\approx$ 4.7~mm slab of liquid argon.
-Thus, one can think to slice the TPC into many thin targets of $\delta${\emph{X}} = 4.7~mm thickness along the direction of the incident particle.
+Thus, one can think to slice the TPC into many thin targets of $\delta${\emph{X}} = 4.7~mm thickness along the direction of the incident particle, making a measurement at each wire along the path.
 
@@
 where $E_{loss}$ is a correction for the energy loss in the dead material between the beamline and the TPC front face.
-The  energy of the hadron at the each slab is determined by subtracting the energy released by the particle in the previous slabs.
+The  energy of the hadron at each slab is determined by subtracting the energy released by the particle in the previous slabs.
 For example, at the $j^{th}$ point of a track, the kinetic energy will be
@@
 For the validation test, we fire about 390000 pions and 140000 kaons inside the LArIAT TPC active volume using the Data Driven Monte Carlo (see section \ref{sec:DDMC}).
-We apply  the thin-sliced method using only true quantities to calculate the hadron kinetic energy at each slab in order to decouple reconstruction effects to eventual issues with the methodology.
-For each slab of 4.7 mm length on the path of the hadron, we integrate the true energy deposition as given by the Geant4 transportation model.
+We apply  the thin-sliced method using only true quantities to calculate the hadron kinetic energy at each slab in order to decouple reconstruction effects from issues with the methodology.
+For each slab of 4.7 mm length along the path of the hadron, we integrate the true energy deposition as given by the Geant4 transportation model.
 Then, we recursively subtracted it from the hadron kinetic energy at the TPC front face to evaluate the kinetic energy at each slab until the true interaction point is reached.

### 2018-05-01 AppendixB-EField/B-LArIATEField.tex
s

@@
 & Shield-Induction & Induction-Collection \\ \hline
-E$_{filed}$ &                 700.625 V/cm        &                892.5  V/cm             \\ \hline
+E$_{field}$ &                 700.63 V/cm        &                892.5  V/cm             \\ \hline
 v$_{drift}$ &                   1.73  mm/$\mu$s   &                  1.90 mm/$\mu$s        \\ \hline
@@
 \includegraphics[width=3in]{AppendixB-EField/Images/CircuitLArIAT.png}
-\caption{\textcolor{red}{get rid of current line} LArIAT HV simple schematics.}
+\caption{LArIAT HV simple schematics.}
 \label{fig:circuit}
@@
 \includegraphics[width=3in]{AppendixB-EField/Images/glassman_current_20160525-30.png}
-\caption{\textcolor{red}{the axis is wrong!!} Current reading from the Glassman between May 25th and May 30th, 2016 (typical Run-II conditions).}
+\caption{Current reading from the Glassman between May 25th and May 30th, 2016 (typical Run-II conditions).}
 \label{fig:currentMeasurement}

### 2018-05-01 AppendixB-EField/B-LArIATEField.tex
s

@@
 A set of two of filter pots for emergency power dissipation are positioned between the Glassman power supply and the cathode, one at each end of the feeder cable, each with an internal resistance of 40~M$\Omega$.
-The output current of the Glassman power supply is then used to determine the electric field strength.
-Figure \ref{fig:currentMeasurement} shows an average current of 0.004172 mA from the Glassman power supply.
+
+Given the TPC resistor chain, the total TPC impedance is ~6 G$\Omega$. Since the total resistance on the circuit is driven by the TPC impedance, we expect the resulting current to be 
+\begin{equation}
+I=V_{PS} /R_{tot} = -23.5\text{ kV}/6 \text{ G}\Omega  \sim 4 \text{ $\mu$A}, 
+\end{equation}
+
+which we measure with the Glassman power supply, shown in  Figure \ref{fig:currentMeasurement}.
 
@@
 \end{equation}
-where $I$ is the current and $R_{eq}$ is the equivalent resistor representing the two filter pots. The electric field, drift voltage, and drift time are then calculated to be
-\begin{equation}E_{\text{field}} = \frac{V_{BC} - V_{\text{shield}}}{\Delta x} = 486.54\text{ V/cm}
+where $I$ is the current and $R_{eq}$ is the equivalent resistor representing the two filter pots. The electric field is then calculated to be
+\begin{equation}E_{\text{field}} = \frac{V_{BC} - V_{\text{shield}}}{\Delta x} = 486.54\text{ V/cm}.
 \end{equation}
@@
 \caption{Pictorial representation of the YX view of the TPC.
-The distance within the anode planes and between the shield plane and the cathode is purposely out of proportion to illustrate the time difference between hits on collection and induciton.
-A ACP track is shown as an example.}
+The distance within the anode planes and between the shield plane and the cathode is purposely out of proportion to illustrate the time difference between hits on collection and induction.
+An ACP track is shown as an example.}
 \label{fig:Scheme}
@@
 \includegraphics[width=3in]{AppendixB-EField/Images/AngleDef.png}
-\caption{Angle definition in the context of LArIAT coordinates system.}
+\caption{Angle definition in the context of LArIAT coordinate system.}
 \label{fig:AngleDef}
@@
 \item[0.] Select cosmic ray events with only 1 reconstructed track 
-\item[1.] Reduce the events to the one contaoning tracks that cross both anode and cathode
+\item[1.] Reduce the events to the one containing tracks that cross both anode and cathode
 \item[2.] Identify the first and last hit of the track
@@
 \item angle from the drift direction (phi in figure \ref{fig:AngleDef}) smaller than 50 deg (more reliable tracking)
-\item angle from the beam direction (theta in figure \ref{fig:AngleDef}) grater than 50 deg (more reliable tracking)
+\item angle from the beam direction (theta in figure \ref{fig:AngleDef}) greater than 50 deg (more reliable tracking)
 \end{itemize}
@@
 
-For each track passing our selection, we loop through the associated hits in order to retrieve the timing information.
+For each track passing our selection, we loop through the associated hits to retrieve the timing information.
 The analysis is performed separately on hits on the collection plane and induction plane, but lead to consistent results.
@@
 We fit with a Gaussian to the peak of the $\Delta t$ distributions to extract the mean drift time and the uncertainty associated with it.
-The long tail at low $\Delta t$ represent contamination of non-ACP tracks in the track selection.
+The long tail at low $\Delta t$ represents contamination of non-ACP tracks in the track selection.
 We apply the same procedure to Run-I and Run-II, positive and negative polarity alike.
@@
 
-To convert $\Delta t$ recorded for the hits on the induction plane to the drift time we utilize the formula
+To convert $\Delta t$ recorded for the hits on the induction plane to the drift time we employ the formula
 \begin{equation}
@@
 \begin{equation}
-t_{S-I} = \frac{l_{S-I}}{v_{S-I}} = \frac{4 mm}{1.745 mm/ \mu s}
+t_{S-I} = \frac{l_{S-I}}{v_{S-I}} = \frac{4 mm}{1.73 mm/ \mu s}
 \end{equation}

### 2018-05-01 Chapter-1/1-Theory.tex
s

@@
 
-Historically, neutrino interactions with the nucleus in the GeV region are divided into three categories as a function of increasing neutrino energy: quasi elastic (QE), resonant (RES), and deep inelastic (DIS) scattering.
-All current and forthcoming oscillation experiments live in the 0.1-10 GeV transition region, which encompasses the energy where the QE neutrino-nucleus interaction transitions into RES and the energy  where RES  transitions into DIS.
- Schematically, neutrino and antineutrino QE charge current scattering refers to the process $\nu_l n \rightarrow l^- p$ and $\bar\nu_l p \rightarrow l^+ n$ where a charged lepton and single nucleon are ejected in the elastic interaction, leaving the target nucleus in its ground state.
+Historically, neutrino interactions with the nucleus in the GeV region are divided into three categories into three categories whose contributions change as a function of increasing neutrino energy:: quasi elastic (QE), resonant (RES), and deep inelastic (DIS) scattering.
+All current and forthcoming oscillation experiments live in the 0.1-10 GeV transition region, which encompasses the energy where the QE neutrino-nucleus interaction transitions into RES and then into DIS.
+For scattering off free nucleons, neutrino and antineutrino QE charge current scattering refers to the process $\nu_l n \rightarrow l^- p$ and $\bar\nu_l p \rightarrow l^+ n$ where a charged lepton and single nucleon are ejected in the elastic interaction.
 %of a neutrino (or antineutrino) with a nucleon in the target material.
- Resonant scattering refers to an inelastic collision producing a nucleon excited state ($\Delta$,N*) -- the resonance, indeed -- which then quickly decays, most often to a nucleon and single-pion final state.
+ Resonant scattering refers to an inelastic collision producing a nucleon excited state ($\Delta$,N*) -- the resonance -- which then quickly decays, most often to a nucleon and single-pion final state.
 DIS refers to the head-on collision between the neutrino and a parton inside the nucleon, producing hadronization and subsequent  abundant production of mesons and nucleons.
@@
 
-As we will discuss in Chapter \ref{ch:2}, the properties of argon make it a good candidate for interacting medium in neutrino experiments; in particular the density of its interaction centers  augments the yield of neutrino interactions and allows for relatively compact detectors.
+As we will discuss in Chapter \ref{ch:2}, the properties of argon make it a good candidate for an interacting medium in neutrino experiments; in particular the density of its interaction centers  increases the yield of neutrino interactions and allows for relatively compact detectors.
 Though, the choice of a relatively heavy nuclear target comes at the cost of enhancing  nuclear effects which modify the kinematic and final state of the neutrino interaction products.
 
-Nuclear effects can potentially affect the neutrino event rates, nucleon emission, neutrino energy reconstruction, and the neutrino/antineutrino ratios, carrying deep implications for oscillation experiments.
+Nuclear effects can potentially affect neutrino event rates, final state particle emission, neutrino energy reconstruction, and the neutrino/antineutrino ratios, carrying deep implications for oscillation experiments.
 Even in the case of ``simple" QE scattering, intra-nuclear hadron rescattering and correlation effects between the target nucleons can cause the ejection of additional nucleons in the final state, modifying the final state kinematics and topology.
-In case of resonant and DIS scattering, the hadronic interactions of meson and nucleons produced in the decay of the resonance or during hadronization  complicate this picture even more.
+In the case of resonant and DIS scattering, the hadronic interactions of meson and nucleons produced in the decay of the resonance or during hadronization  complicate this picture even more.
 A large source of uncertainty in modeling  nuclear effects in neutrino interactions come from mesons interactions (and re-interactions) in the nucleus, e.g., pion re-scattering, charge exchange, and absorption.
@@
 % and of light mesons interaction with nuclei studies.
- We describe the signal signature and historic measurements of pion-nucleus cross section, as well as the implementation of this cross sections in the current version of the simulation package used by LArIAT.
+ We describe the signal signature and historic measurements of pion-nucleus cross section, as well as the implementation of these cross sections in the current version of the simulation package used by LArIAT.
 
@@
 Pions are a common product of neutrino interaction, especially in resonant scattering, DIS and coherent pion production.
-For neutrino experiments in argon, there are two main reasons why understanding pion hadronic interactions with argon is important: to model the behavior of the pion inside the nucleus struck by the neutrino and to model the behavior of the pion during its propagation inside the detector medium.
+For neutrino experiments in argon, there are two main reasons why understanding pion hadronic interactions with argon is important: to model the behavior of the pion inside the target nucleus  and to model the behavior of the pion during its propagation inside the detector medium.
 
-Assumptions on the nuclear models and on the interaction of hadrons inside the nucleus performed at the level of the neutrino event generator bridge the measurement of the products of a neutrino interaction to the reconstruction of the neutrino energy and flavor.
-Thus, understanding pion hadronic interactions with the nucleus is particularly important to model correctly resonant, DIS and coherent pion production in neutrino interactions, where the presence of pions in the nucleus is abundant.
+Assumptions on the nuclear modeling and on the interaction of hadrons inside the nucleus performed at the level of the neutrino event generator bridge the measurement of the products of a neutrino interaction to the reconstruction of the neutrino energy and flavor.
+Thus, understanding pion hadronic interactions with the nucleus is particularly important to model correctly resonant, DIS and coherent pion production in neutrino interactions.
 For example, in case of resonant scattering,
@@
 \end{equation}
-the $\Delta$ and $N^*$ and excited states will decay hadronically in matters of $\sim10^{-24}$~s inside the nucleus producing pions which will bounce within the nuclear medium.
+the $\Delta$ and $N^*$ and excited states will decay hadronically in matters of $\sim10^{-24}$~s inside the nucleus producing pions which which will have many chances to re-interact as they exit the target medium.
 The decay modes for the lower mass $\Delta$ (1232) and $N^*$(1440)  are listed in table \ref{tab:decaysdelta}.
@@
 We illustrate here the conceptual basis of the GENIE Neutrino Generator~\cite{Andreopoulos:2009rq} as an example, since GENIE is one the most popular event generators for liquid argon experiments.
-For example, the nuclear model used by GENIE for all processes is a relativistic Fermi gas (RFG) modified to incorporate short range nucleon-nucleon correlations \cite{PhysRevD.24.1400}.
+For example, the nuclear model used by GENIE for all processes is a Relativistic Fermi Gas (RFG) model modified to incorporate nucleon-nucleon correlations \cite{PhysRevD.24.1400}.
 This means that the initial momentum and binding energy of the struck nucleon is determined by assuming nucleons inside the nucleus are quasi-free, acting independently in the mean field of the nucleus.
-For A $ >$ 20 like argon, the 2-parameter Woods-Saxon shell model for density function is used.
-The GENIE module INTRANUKE \cite{Juszczak2006} is used to simulate the final-state interactions (FSI) which is the hadron re-interaction inside the nucleus.
+For A $ >$ 20 such as argon, the 2-parameter Woods-Saxon shell model for density function is used.
+The GENIE module INTRANUKE \cite{Juszczak2006} is used to simulate final-state interactions (FSI) which model  hadron re-interactions inside the nucleus.
 This module places the outgoing particles in the nucleus and propagates them using the ``hA model".
@@
 When possible, external hadron-nucleus scattering data are used to tune INTRANUKE.
-Since no data is available for Argon, GENIE uses an interpolation of data from heavier and lighter nuclei for the pion-argon cross section leading to big uncertainties in the INTRANUKE module.
+Since no data is available for Argon, GENIE uses an interpolation of data from heavier and lighter nuclei for the pion-argon cross section leading to large (10?s of \%) resultant uncertainties in the INTRANUKE module.
 
-Once the pion has left the target nucleus, the pion-argon hadronic cross section plays an important role in the pion transportation inside the argon medium: processes like pion absorption with emission of nucleons or pion charge exchange can greatly modify the topology of a neutrino interactions in the detector and lead to errors in the event classification.
+Once the pion has left the target nucleus, the pion-argon hadronic cross section also plays an important role in the pion transportation inside the argon medium: processes such as pion absorption or pion charge exchange can greatly modify the topology of a neutrino interactions in the detector and lead to significant modifications in the event classification.
 Being able to reconstruct the details of pions inside the detector is an imperative for modern liquid argon neutrino experiment to achieve the design resolution for their key physics measurements.
+%For example, if a pi+ charge exchanges into a pi0, an event which would not have been a background to a nue search now becomes one.
+So you need to know with what frequency this happens in a given nucleus for a given energy pion.
+Also, if a pion gets absorbed in the target nucleus, this energy is  transferred into nucleons, some of which can be  invisible (e.g., neutrons) to the  detector thus leading to a underestimate of  incoming neutrino energy.
 
@@
 \end{equation}
-the reaction channel is further characterizes by several exclusive channels with defined topologies,
+the reaction channel is further characterized by several exclusive channels with defined topologies,
 \begin{equation}
@@
 Since the pion mass is two orders of magnitude greater than the electron mass, the spatial wave form of the pion will overlap more with the nucleus compared to the electron case.
-After the electromagnetic formation of the pionic atom, the pion will be quickly absorbed by the nucleus, which is put in an excited state.
+After the electromagnetic formation of the pionic atom, the pion will get quickly absorbed by the nucleus, which is put in an excited state.
 The nucleus then de-excites with the emission of low energy nucleons and photons.
-Pion capture is predominant compared to pion decay, the other important process for very low energy pions.
+Pion capture is dominant compared to pion decay, the other important process for very low energy pions.
 The decay of a pion is governed by the weak force; the pion decay life time is $\tau_\pi = 2.6 \times 10^{-8}$ s and the main decay mode is  $\pi^-\rightarrow \mu^- + \bar\nu_\mu$ (BR 99.98\%).
@@
 However, data on argon are rare: the total differential hadronic cross section has never been measured before on argon.
-Simulation packages like Geant4 base their pion transportation for argon on data from lighter and heavier nuclei: the goal of LArIAT's dedicated measurement on argon is to bridge this gap in data, thus reducing the uncertainties related to pions interaction in argon in both neutrino event generators and in simulation packages of pion transportation.
+Simulation packages such as Geant4 base their pion transportation for argon on data from lighter and heavier nuclei: the goal of LArIAT's dedicated measurement on argon is to bridge this gap in data, thus reducing the uncertainties related to pion interactions in argon in both neutrino event generators and in simulation packages of pion transportation.
 
@@
 In particular, pions (and kaons) transportation is achieved through the Geant4 FTFP\_BERT physics list.
-In this physics list, Geant4 uses the Bertini cascade model  \cite{Wright:2015xia}  to simulate the products of the pion-nucleus interaction as well as the secondary hadrons re-interactions inside the target nucleus (intra-nuclear cascade).
+In this physics list, Geant4 uses the Bertini cascade model  \cite{Wright:2015xia}  to simulate the products of the pion-nucleus interaction as well as secondary hadronic re-interactions inside the target nucleus (intra-nuclear cascade).
 The target nucleus is represented as a continuous gas where the nuclear potential follows concentrical shells whose depths approximate the Woods-Saxon shape.
@@
 The work in \cite{Nutini2015}  reviews  the current status of negative and positive pion simulation in Geant4 and GENIE for $^{12}$C, $^{56}$Fe, and $^{40}$Ca.
-From that work, we report the results for $^{12}$C in Figure \ref{fig:pionC} as it allows a direct comparison between Geant4, GENIE and data.
-Geant4 predictions for $\pi^-$  on Carbon are in good agreement with data over all the spectrum, while GENIE predictions seem to show some features at around 500 MeV and 900 MeV, maybe due to higher resonances in the hA model.
-From the same work, we also report the negative pion cross section on $^{40}$Ca in Figure \ref{fig:pionCa}, since this is the nuclear medium closest argon with some available data.
+From that work, we report the results for $^{12}$C in Figure \ref{fig:pionC} as it allows a direct comparison between Geant4, GENIE and and pion re-scattering data.
+Geant4 predictions for $\pi^-$  on Carbon are in good agreement with data over  the entire spectrum spectrum, while GENIE predictions seem to show some features at around 500 MeV and 900 MeV, maybe due to higher resonances in the hA model.
+From the same work, we also report the negative pion cross section on $^{40}$Ca in Figure \ref{fig:pionCa}, since this is the nuclear medium closest to argon.
 The predictions from both Geant4 and GENIE agree with data in the high energy region; the Geant4 and GENIE predictions diverge in the resonance region, where data is not available.

### 2018-05-01 Chapter-4/4-InteractionsInLAr.tex
s

@@
 
+\begin{table}[]
+\centering
+\begin{tabular}{|l|c|c|}
+\hline
+                                                        & Run-II Negative Polarity   &  Run-II Positive Polarity  \\ \hline
+Events Reconstructed in Beamline &  158396  & 260810  \\ \hline
+Events with Plausible Trajectory     &   147468 & 240954  \\ \hline
+Candidate $\pi^-/\mu^-/e^-$            &   138481 &     N.A.   \\ \hline
+Candidate $K^+$                            &    N.A       & 2837     \\ \hline
+Events Surviving Pile Up Filter       &                 &              \\ \hline
+Events with WC2TPC Match          &                 &              \\ \hline
+Events Surviving Shower Filter      &                 &  N.A.     \\ \hline
+\end{tabular}
+\caption{Number of data events for Run-II Negative and Positive polarity }
+\label{tab:beamlineDataSelection}
+\end{table}
+
 \subsubsection{Particle Identification in the beamline}
@@
 m_{Beam} = \frac{p_{Beam}}{c}\sqrt{\biggl(\frac{TOF*c}{l}\biggr)^2 -1},
+\label{eq:mass}
 \end{equation}
@@
 \begin{itemize}
-\item \underline{$\pi, \mu, e$:} 0~MeV $<$ mass $<$ 350~MeV
+\item \underline{$\pi, \mu, e$:}  mass $<$ 350~MeV
 
@@
 \begin{figure}
-%\captionsetup{justification=raggedright}  
-\begin{minipage}[b]{.53\textwidth}  
   \centering  
-\includegraphics[width=3in]{Chapter-4/Images/massRunIIPos.png}
-\end{minipage}%  
-\begin{minipage}[b]{0.53\textwidth}  
-  \centering  
-\includegraphics[width=3in]{Chapter-4/Images/face-smile-big.png}
-\end{minipage}
-\caption{The mass plotted for a sample of Run-II events reconstructed in the beamline, negative polarity runs on the left and positive polarity runs on the right.
-The classification of the events into $\pi, \mu, e$, kaon, or proton is based on this distribution.\textcolor{red}{CHANGE PLOTS}}
+\includegraphics[width=\textwidth]{Chapter-4/Images/massRunII.png}
+\caption{Distribution of the beamline mass as calculated according to equation \ref{eq:mass} for the Run-II events reconstructed in the beamline, negative polarity runs on the left and positive polarity runs on the right.
+The classification of the events into $\pi^\pm/ \mu^\pm/e^\pm$, K$^\pm$, or (anti)proton is based on these distributions, whose selection cut are represented by the vertical colored lines.}
 \label{fig:mass}

### 2018-05-03 Chapter-4/4-InteractionsInLAr.tex
s

@@
 Events with WC2TPC Match          &                 &              \\ \hline
-Events Surviving Shower Filter      &                 &  N.A.     \\ \hline
+Events Surviving Shower Filter      &   56526    &  N.A.     \\ \hline
 \end{tabular}

### 2018-05-13 Chapter-4/4-Methodology.tex
s

--- 2018-05-01 Chapter-4/4-InteractionsInLAr.tex
@@
 \chapter{Total Hadronic Cross Section Measurement Methodology}\label{ch:Interactions}
+This chapter describes the general procedure employed to measure a total hadronic differential cross section in LArIAT.
+Albeit with small differences, both the  ($\pi^{-}$,Ar) and (K$^{+}$,Ar) total hadronic cross section measurements rely on the same procedure described in details in the following sections.
+We start by selecting the particle of interest using a combination of beamline detectors and TPC information (Section \ref{ch:ParticleSelectionMethod}).
+We then perform a handshake between the beamline information and the TPC tracking to assure the selection of the right TPC track (Section \ref{ch:WC2TPCMatchMethod}).
+Finally, we apply the ``thin slice" method and measure the ``raw" hadronic cross section (Section \ref{ch:ThinSliceMethod}).
+A series of corrections are then evaluated to obtain the ``true" cross section (Section \ref{ch:MCCorrections}).
 
-%the prediction of the total hadronic interaction cross section ($\pi^{\pm}$, Ar) for thin-target simulations from two Monte Carlo generators (Geant 4.10.1 with Bertini Cascade model~\cite{geant4, g4bert} and Genie v2.8.2 with intranuke-hA model).
-The thick target simulation used a simple stand-alone Geant4 simulation  (i.e., no other detector features were taken into account except the geometry of the thin slices in the LAr volume).
-Fig.~\ref{fig:xsplot} shows the resulting total ${\pi^-}$ cross section extracted by the sliced TPC technique; it agrees well with the Geant 4 thin-target cross section.
-The Genie thin-target cross section for $^{40}$Ar shown in this figure is significantly different than that of Geant at low kinetic energies due to the fact that Geant4 models are tuned on $^{12}$C while Genie ones are tuned on a much heavier $^{56}$Fe target.
-The extrapolated cross section predicted for $^{40}$Ar can be then very different, especially in the resonance region where the model is strongly target-dependent from one generator to another.
+At the end of this chapter, we show a sanity check of the methodology against MC truth information (Section \ref{ch:procedureTesting}).
 
-%The comparison of the Geant4 thin- and thick-target cross section results demonstrates the power of the "sliced TPC" method for the measurement of the ($\pi^{\pm}$, Ar) cross section in LArIAT TPC geometry.
-section
+\section{Event Selection}\label{ch:ParticleSelectionMethod}
+The measurement of the ($\pi^{-}$,Ar) and (K$^{+}$,Ar) total hadronic cross section in LArIAT starts by selecting the pool of pion or kaon candidates and measuring their momentum.
+This is done through the series of selections on  beamline and TPC information described in the next sections.
+The summary of the event selection in data is reported in Table \ref{tab:beamlineDataSelection}.
 
-%Having validated the ``Sliced TPC'' technique and shown that the this technique recovers the simulated cross-section, we now move to performing this measurement utilizing the complete LArIAT simulation within the LArSoft \cite{} based simulation.
-Furthermore, we move from utilizing particle level MC-truth information to performing fully automated reconstruction of the charged hadron events within the LArTPC.
-
-%Since the kaon cross section in argon has never been measured before, the Geant4 Monte Carlo tunes kaon transportation in argon by extrapolation from lighter and heavier nuclei.
-As shown in the previous section,  kaon data on carbon are available and  can be used as a metric to evaluate the Geant4 prediction performances.
-Figure \ref{fig:TrueCarbon} shows the total hadronic cross section for carbon implemented in Geant4 10.01.p3 overlaid with the Bugg and Friedman data.
-Unfortunately, the current version of Geant4 does not reproduce the data for carbon closely.
-On one hand, this evidence makes us even more wary when using the Monte Carlo in simulating the kaon-argon interactions.
-On the other, it further highlights the importance of kaon measurements.
-
-\section{How to Measure a Hadron Cross Section in LArIAT}\label{ch:methodology}
-We use both the LArIAT  beamline detectors and the LArTPC information to measure the total hadronic cross sections in argon.
-Albeit with small differences, both the  $\pi^{-}$~-~Ar and K$^{+}$~-~Ar total hadronic cross section measurements rely on the same procedure described in details in the following sections: we select the particle of interest using a combination of beamline detectors and TPC information (section \ref{ch:ParticleSelectionMethod}), we perform a handshake between the beamline information and the TPC tracking to assure we are selecting the right TPC track (section \ref{ch:WC2TPCMatchMethod}), and we apply the ``thin slice" method to get to the final result (section \ref{ch:ThinSliceMethod}).
-At the end of this chapter, we show a cross check of this method against MC truth information (section \ref{ch:procedureTesting}).
-
-\subsection{Event Selection}\label{ch:ParticleSelectionMethod}
-\subsubsection{Beamline events}\label{ch:beamlineDetectorsData}
-As will be clear in section \ref{ch:ThinSliceMethod}, beamline particle identification and momentum measurement before entering the TPC are fundamental information for the hadronic cross sections measurements in LArIAT.
-Thus, we apply selections to the LArIAT data to keep only events whose wire chamber and time of flight information is registered.
-Additionally, we perform a check of the plausibility of the trajectory inside the beamline detectors: given the position of the hits in the four wire chambers, we make sure the particle trajectory does not cross any impenetrable material such as the collimator and the magnets steel.
-
-\begin{table}[]
+\begin{table}[b]
 \centering
@@
 Events with Plausible Trajectory            &   147468 & 240954  \\ \hline
-Candidate $\pi^-/\mu^-/e^-$            &   138481 &     N.A.   \\ \hline
-Candidate $K^+$                            &    N.A       & 2837     \\ \hline
-Events Surviving Pile Up Filter       &                 &              \\ \hline
-Events with WC2TPC Match          &                 &              \\ \hline
-Events Surviving Shower Filter      &                 &  N.A.     \\ \hline
+Beamline $\pi^-/\mu^-/e^-$  Candidate  &   138481 &     N.A.   \\ \hline
+Beamline $K^+$   Candidate                 &    N.A       & 2837     \\ \hline
+Events Surviving Pile Up Filter              &   108929  &              \\ \hline
+Events with WC2TPC Match                 &    41757   &              \\ \hline
+Events Surviving Shower Filter             &    40841    &  N.A.     \\ \hline
+Available Events For Cross Section      &   40841    &       \\ \hline
 \end{tabular}
@@
 \end{table}
+
+\subsubsection{Selection of Beamline Events}\label{ch:beamlineDetectorsData}
+As shown in equation \ref{eq:enFF}, we leverage the beamline particle identification and momentum measurement before entering the TPC as in input to evaluate the kinetic energy for the hadrons used in the  cross sections measurements.
+Thus, we select the LArIAT data to keep only events whose wire chamber and time of flight information is registered (line 2 in in Table \ref{tab:beamlineDataSelection}).
+Additionally, we perform a check of the plausibility of the trajectory inside the beamline detectors: given the position of the hits in the four wire chambers, we make sure the particle's trajectory does not cross any impenetrable material such as the collimator and the magnets steel (line 3 in in Table \ref{tab:beamlineDataSelection}).
 
@@
 \end{equation}
- where $c$ is the speed of light and $l$ is the length of the particle trajectory between the time of flight paddels.
+ where $c$ is the speed of light and $l$ is the length of the particle's trajectory between the time of flight paddels.
 
@@
 \begin{itemize}
-\item \underline{$\pi, \mu, e$:}  mass $<$ 350~MeV
+\item \underline{$\pi/\mu/e$:}  mass $<$ 350~MeV
 
@@
 \end{itemize}
+
+Lines 4 and 5 in in Table \ref{tab:beamlineDataSelection} show the number of negative $\pi/\mu/e$ and positive $K$ candidates which pass the mass selection for LArIAT Run-II data.
 
@@
 
-\subsubsection{Additional Particle Identification technique}\label{ch:electrons}
-In the case of the $\pi^-$-Ar cross section, the resolution of  beamline mass spectrometer is not sufficient to select a beam of pure pions.
-In fact, muons and electrons survive the selection on the beamline mass value.
-It is important to notice that the composition of the negative polarity beam is mostly pions, as discussed in section \ref{ch:beamlineComposition}.
+\subsubsection{TPC Selection: Halo mitigation }\label{ch:pileUp}
+The secondary beam impinging on LArIAT secondary target produces a plethora of particles which propagates downstream.
+The presence of upstream and downstream collimators greatly abates the number of particles tracing down the LArIAT tertiary beamline.
+However, it is possible that more than one particle sneaks into the LArTPC during its readout time: the TPC readout is triggered by the particle  firing the beamline detectors, but particles from the beam halo might be present in the TPC at the same time.
+We call ``pile up" the additional traces in the TPC.
+We adjusted the primary beam intensity between LArIAT Run I and Run II to reduce the presence of events with high pile up particles in the data sample.
+For the cross section analyses, we remove events with more than 4 tracks in the first 14 cm upstream portion of the TPC from the sample (line 6 in in Table \ref{tab:beamlineDataSelection}).
+
+\subsubsection{TPC Selection: Shower Removal}\label{ch:electrons}
+In the case of the ($\pi^-$,Ar) cross section, the resolution of  beamline mass spectrometer is not sufficient to select a beam of pure pions.
+In fact, muons and electrons survive the selection on the beamline mass.
+It is important to notice that the composition of the negative polarity beam is mostly pions, as will be discussed in section \ref{ch:beamlineComposition}.
 Anyhow, we devise a selection on the TPC information to mitigate the presence of electrons in the sample used for the pion cross section.
 The selection relies on the different topologies of a pion and an electron event in the argon: while the former will trace a track inside the TPC active volume, the latter will tend to ``shower", i.e.
-interact with the medium, produce bremsstrahlung photons which pair convert into several short tracks.
-%We provide details of this selection in section \ref{ch:electronCuts}.
+interact with the medium, producing bremsstrahlung photons which pair convert into several short tracks.
+In order to remove the shower topology, we create a region of interest (ROI) around the TPC track corresponding to the beamline particle (more details on this in the next section).
+We look for short tracks contained in the ROI, as depicted in figure \ref{fig:showerFilt}:  if more then 5 tracks shorter than 10 cm are in the ROI, we reject the event.
+Line 8 in in Table \ref{tab:beamlineDataSelection}) shows the number of events surviving this selection.
 
-\subsubsection{Pile up mitigation }\label{ch:pileUp}
-The secondary beam impinging on LArIAT secondary target produces a plethora  of particles.
-The presence of upstream and downstream collimators greatly abates the number of particles tracing down the LArIAT beamline.
-However, more than one beamline particles, or particles produced from the beam interaction with the beamline detectors, may sneak into the LArTPC during its readout time.
-The TPC readout is triggered by the actual particle  firing the beamline detectors; we call ``pile up" the additional traces in the TPC.
-We adjusted the primary beam intensity between LArIAT Run I and Run II to minimize the presence of events with high pile up particles in the data sample.
-For the cross section analyses, we remove events with more than 4 tracks in the first 14 cm upstream portion of the TPC from the sample.
+\begin{figure}
+  \centering  
+\includegraphics[width=3in]{Chapter-4/Images/Shower.png}
+\caption{Visual rendering of the shower filter.
+The ROI is a cut cone, with a small radius of 4 cm, a big radius of 10 cm and an height of 42 cm (corresponding to 3 radiation lengths for electrons in Argon).}
+\label{fig:showerFilt}
+\end{figure}
 
-\subsection{Wire Chamber to TPC Match}\label{ch:WC2TPCMatchMethod}
-For each event passing the selection on its beamline information we need to identify the track inside the TPC corresponding to the particle which triggered the beamline detectors, a procedure we refer to as ``WC to TPC match" (WC2TPC for short).
+\section{Beamline and TPC Handshake: the Wire Chamber to TPC Match}\label{ch:WC2TPCMatchMethod}
+For each event passing the selection on its beamline information, we need to identify the track inside the TPC corresponding to the particle which triggered the beamline detectors, a procedure we refer to as ``WC to TPC match" (WC2TPC for short).
 In general, the TPC tracking algorithm will reconstruct more than one track in the event, partially due to the fact that hadrons interact in the chamber and partially because of pile up particles during the triggered TPC drift time, as shown in figure \ref{fig:kaonInteraction}.
@@
 
-\subsection{The Thin Slice Method}\label{ch:ThinSliceMethod}
-\subsubsection{Cross Sections on Thin Target}
+\section{The Thin Slice Method}\label{ch:ThinSliceMethod}
+\subsection{Cross Sections on Thin Target}
 Cross section measurements on a thin target have been the bread and butter of nuclear and particle experimentalists since the Geiger-Marsden experiments \cite{Geiger1909}.
@@
 
-\subsubsection{Not-so-Thin Target: Slicing the Argon}
+\subsection{Not-so-Thin Target: Slicing the Argon}
 The LArIAT TPC, with its 90 cm of length, is not a thin target. However, the fine-grained tracking of the LArIAT LArTPC allows us to treat the argon volume as a sequence of many adjacent thin targets.
@@
 
-\subsection{Procedure testing with truth quantities}\label{ch:procedureTesting}
+\section{Procedure testing with truth quantities}\label{ch:procedureTesting}
 The $\pi^{-}$-Ar and K$^{+}$-Ar total hadronic cross section implemented in Geant4 can be used as a tool to validate the measurement methodology.
@@
 %\end{comment}
+
+\section{Corrections to Data}\label{ch:MCCorrections}

### 2018-05-13 Chapter-5/5-Preparatory.tex
s

--- 2018-04-30 Chapter-5/5-SamplesPreparation.tex
@@
-\chapter{Samples Preparation}\label{ch:samples}
-This chapter describes the preparation of the data and Monte Carlo samples used for the cross section analyses. This entails:
+\chapter{Preparatory Work}\label{ch:samples}
+This chapter describes the preparatory work done on the the data and Monte Carlo samples used for the cross section analyses. This entails:
+
 \begin{enumerate}
-\item the beamline event selection on data,
 \item the MC production,
@@
 
-\section{LArIAT Data}
-\section{LArIAT Monte Carlo}
+\section{Construction of a Monte Carlo Simulation for LArIAT}
+\subsection{G4Beamline}\label{ch:beamlineComposition}
+\subsection{Data Driven MC}\label{sec:DDMC}
+
+\section{Tracking Studies}
+
+\section{Energy Calibration and Studies}
+
+\section{Estimate of Energy Loss before the TPC}
+
+\begin{comment}
+\section{Construction of a Monte Carlo Simulation for LArIAT}
 For the simulation of LArIAT events and their particle make up, we use a combination of two MC generators: the G4Beamline Monte Carlo and the Data Driven single particle Monte Carlo (DDMC).
@@
 As described in section \ref{sec:SignalProc}, this is a multi-step procedure.
-In LArIAT, we first correct the raw charge by the electronic noise on the considered wire \cite{technote}, then by the electron life time \cite{technote},  and then by the recombination effect using the ArgoNeut recombination values.
+In LArIAT, we first correct the raw charge by the electronic noise on the considered wire \cite{technote}, then by the electron lifetime \cite{LArIATLifeTime},  and then by the recombination using the ArgoNeut recombination values.
 Lastly, we apply overall calibration of the energy, i.e.
-we determine the ``calorimetry constants" using the procedure described in this paragraph.
+we determine the ``calorimetry constants" using the procedure described in this section.
 
-The general idea is to use a parametrization of the energy deposited per unit length (dE/dX) as a function of momentum to independently determine the calorimetry constants for the data and Monte Carlo in the LArIAT Run-II data samples.
-This is done by comparing the stopping power measured on reconstructed quantities against the Bethe-Bloch theoretical prediction for various particle species.
+We independently determine  the calorimetry constants for Data and Monte Carlo in the LArIAT Run-II Data samples using  a parametrization of the energy deposited per unit length (dE/dX) as a function of momentum.
+This is done by comparing the stopping power measured on reconstructed quantities against the Bethe-Bloch theoretical prediction for various particle species (see equation \ref{eq:BB}).
+We obtain the theoretical expectation for the dE/dX most probable value of pions ($\pi$), muons ($\mu$), kaons ($K$), and protons ($p$) in the momentum range most relevant for LArIAT (Figure \ref{fig:PDGEnergyLossArgon}) using the tables provided by the Particle Data Group \cite{Patrignani:2016xqp} for liquid argon \cite{PDG-Argon}.
 
-This is best represented by Figure \ref{fig:PDGEnergyLoss}, taken from the Particle Data Group \cite{PDG}.
-Using the tables provided by the PDG for liquid argon (\cite{PDG-Argon}), we calculate the theoretical values for the most probably value for the stopping power of pions ($\pi$), muons ($\mu$), kaons ($K$), and protons ($p$) in the momentum range most relevant for LArIAT, shown in Figure \ref{fig:PDGEnergyLossArgon}.
+The basic idea of this calibration technique is to utilize the most upstream portion of a TPC track which has a well known momentum and particle species to measure its $dE/dX$.
+Once a sample of particles dE/dX has been measured at various momenta, we then tune to calorimetry constants within the reconstruction software to align these measured values to match the theoretical ones found in Figure \ref{fig:PDGEnergyLossArgon}.
+
+In data, we start by selecting a sample of beamline positive pion candidates without any restriction on their measured momentum.
+We then apply the WC2TPC match and subtract the energy loss upstream to the TPC front face, determining the momentum at the TPC front face.
+For each surviving pion candidate,  we measure the dE/dx at each of the first 12 spacepoints associated the 3D reconstructed track, corresponding to a $\sim$ 5 cm portion.
+These dE/dX measurements are then put into a histogram that corresponds to measured momentum of the track.
+The dE/dX histograms are sampled every 50 MeV in momentum (e.g.
+150~MeV/c $< P <$ 200~MeV/c, 200~MeV/c $< P <$ 250/c~MeV, etc...).
+This process of selecting, sampling, and recording the dE/dX for various momentum bins is repeated over the entire sample of events, allowing us to collect sufficient statistic in most of the momentum bins between 150~MeV/c and 1100~MeV/c.
+Each 50 MeV/c momentum binned dE/dX histogram is now fit with a simple Landau function.
+The most probable value (MPV) and the associated error on the MPV from the fit are extracted and plotted on Figure \ref{fig:PDGEnergyLossArgon}.
+Depending on the outcome of the fit, we modify the calorimetry constants and we repeat the procedure until a qualitative agreement is achieved.
+We perform this  tuning for the collection and induction plane separately.
+As a cross check to the determined calorimetry constants using the positive pions, we plot the dE/dx vs momentum distribution all the other particle species identifiable in the beamline data ($\pi$, K , p, in both polarities) against the corresponding Beth-Bloch prediction in a similar fashion.
+On average, pions and muons only lose $\sim$10 MeV in this 5~cm section of the track and protons lose $\sim$20 MeV.
+Thus choosing 50 MeV size bins for our histograms covers the energy spread within those bins due to energy loss from ionization.
+The results of the tuning and cross check for Run-II data on the collection plane is shown in Figure \ref{fig:BBandData} positive polarity data on the left, negative polarity data on the right.
+
+In MC, we simulate the corresponding positive pion sample with the DDMC (see section \ref{sec:DDMC}) and follow the same steps as in data.
+The calorimetry tuning is explained in all his gory details in \cite{technote}.
 
@@
 \end{figure}
-
-Using the predictions in Figure \ref{fig:PDGEnergyLossArgon}, allows us to tune the calorimetry constants.
-The goal is to have the data and the Bethe-Bloch prediction agree across the broad range of momentum.
-
-The basic idea of this calibration technique is to utilize a portion of a track within the LArTPC that has a well known momentum and particle species to measure the stopping power (or  energy deposited per unit length, $dE/dX$) as recorded inside the TPC.
-Once a sample of particles $dE/dX$ has been measured at various momenta, we then tune to calorimetry constants within the reconstruction software to align these measured values to match the theoretical ones found in Figure \ref{fig:PDGEnergyLossArgon}.
-
-The calibration procedure follows the following steps:
-\begin{itemize}
-\item \textbf{Species and Momentum Determination} We first select a sample of beamline events that correspond to either a sample of $\pi, \mu, e$, kaons or protons.
-This is done by selecting based on the invariant beamline mass as shown in section \ref{}.
-The beamline particle is matched to the TPC information.
-In order to determine the particle momentum at the TPC front face, we use the momentum measured by the LArIAT spectrometer and we apply a flat correction for the energy loss between the beamline and the TPC.
-For the Monte Carlo, no such beamline identification is done and instead we generate DDMC for each particle species.
-
-\item \textbf{dE/dX sampling:} With the track within the TPC identified and the momentum for that track measured, we require the track to be of a minimum length of 10~cm long (to ensure we are away from any interaction point where the track may be broken into subsequent tracks).
-We then take the first twelve spacepoints of the track (excluding the first point to avoid edge effects near the field cage) and sample the reconstructed dE/dX for each point along the track.
-On average, this samples 5~cm of the track (shown in Section \ref{sec:Results}).
-These dE/dX measurements are then put into a histogram that corresponds to measured momentum of the track.
-The dE/dX histograms are sampled every 50 MeV in momentum (e.g.
-150~MeV $< P <$ 200~MeV, 200~MeV $< P <$ 250~MeV, etc...).
-On average, pions and muons only lose $\sim$10 MeV in this 5~cm section of the track and protons lose $\sim$20 MeV.
-Thus choosing 50 MeV size bins for our histograms covers the energy spread within those bins due to energy loss from ionization.
-This process of selecting, sampling, and recording the dE/dX for various momentum bins is now repeated over the entire sample of events, allowing us to collect sufficient statistic in most of the momentum bins between 150~MeV and 1100~MeV.
-
-\item \textbf{Fit, tune, repeat:} Each 50 MeV momentum binned dE/dX histogram is now fit with a simple Landau function.
-The most probable value (MPV) and the associated error on the MPV from the fit are extracted and plotted on Figure \ref{fig:PDGEnergyLossArgon}.
-Depending on the outcome of the fit, the calorimetry constants are either tuned up or down.
-The values are tuned for both the collection and induction plane to try to achieve the best possible match to the theoretical curve.
-The exact match is left as a qualitative exercise and is not quantitatively evaluated.
-\end{itemize}
 
@@
 The distributions for data and MC for $\sigma_\alpha$ and $\sigma_S$ are given in \ref{fig:trackingResolution}.
+
+\end{comment}

### 2018-05-13 Chapter-5/Old/5-SamplesPreparation.tex
s

--- 2018-05-13 Chapter-5/5-Preparatory.tex
@@
-\chapter{Preparatory Work}\label{ch:samples}
-This chapter describes the preparatory work done on the the data and Monte Carlo samples used for the cross section analyses. This entails:
-
+\chapter{Samples Preparation}\label{ch:samples}
+This chapter describes the preparation of the data and Monte Carlo samples used for the cross section analyses. This entails:
 \begin{enumerate}
+\item the beamline event selection on data,
 \item the MC production,
@@
 
-\section{Construction of a Monte Carlo Simulation for LArIAT}
-\subsection{G4Beamline}\label{ch:beamlineComposition}
-\subsection{Data Driven MC}\label{sec:DDMC}
-
-\section{Tracking Studies}
-
-\section{Energy Calibration and Studies}
-
-\section{Estimate of Energy Loss before the TPC}
-
-\begin{comment}
-\section{Construction of a Monte Carlo Simulation for LArIAT}
+\section{LArIAT Data}
+\section{LArIAT Monte Carlo}
 For the simulation of LArIAT events and their particle make up, we use a combination of two MC generators: the G4Beamline Monte Carlo and the Data Driven single particle Monte Carlo (DDMC).
@@
 The distributions for data and MC for $\sigma_\alpha$ and $\sigma_S$ are given in \ref{fig:trackingResolution}.
-
-\end{comment}

### 2018-05-13 Chapter-6/6-PionCrossSection.tex
s

\chapter{Negative Pion Cross Section Measurement}\label{ch:PionXS}
\section{Raw Cross Section}\label{ch:PionXSRaw}

\section{Background Subtracted Cross Section}\label{ch:PionXSBkgSub}

\section{Efficiency Corrected Cross Section}\label{ch:PionXSEffCorr}

### 2018-05-13 Chapter-7/7-KaonCrossSection.tex
s

\chapter{Positive Kaon Cross Section Measurement}\label{ch:KaonXS}
\section{Raw Cross Section}\label{ch:KaonXSRaw}

### 2018-05-13 thesis_main.tex
s

--- 2017-12-06 thesis_main.tex
@@
 
-\documentclass[letterpaper,12pt,draft]{yalephd}
+\documentclass[letterpaper,12pt]{yalephd}
 % remove draft option for final printing.
 % font size must be between 10pt-12pt.
+\usepackage{caption,setspace}
+\usepackage{amsmath} 
+\usepackage{rotating}
 
 \usepackage{geometry} % you need this for yalephd.cls to work.
-\usepackage{graphicx} % you probably want the rest of these.
+\usepackage{graphicx}
+\usepackage[graphicx]{realboxes}
+ % you probably want the rest of these.
 \usepackage{dcolumn}
 \usepackage{bm}
+\usepackage{hyperref}
+
+\usepackage{mathtools}
+\DeclarePairedDelimiter\bra{\langle}{\rvert}
+\DeclarePairedDelimiter\ket{\lvert}{\rangle}
+\DeclarePairedDelimiterX\braket[2]{\langle}{\rangle}{#1 \delimsize\vert #2}
+\usepackage{textgreek}
+
+\usepackage{multirow}
 \usepackage{amsmath}
@@
 
+\usepackage{lineno,xcolor}
+
+\linenumbers
+
 \begin{document}
@@
 
-%  \begin{dedication}
-%A mia mamma e mio babbo,\\
-%grazie per le radici e grazie per le ali.
-%    \par   %% or a blank line
-%    \vspace{2\baselineskip}
-%To my mom and dad,\\
-%thank you for the roots and thank you for the wings.
-%    \vspace{\baselineskip}
-%  \end{dedication}
+  \begin{dedication}
+A mia mamma e mio babbo,\\
+grazie per le radici e grazie per le ali.
+    \par   %% or a blank line
+    \vspace{2\baselineskip}
+To my mom and dad,\\
+thank you for the roots and thank you for the wings.
+    \vspace{\baselineskip}
+  \end{dedication}
+
 \tableofcontents
-\listoffigures % remove this if you have no figures.
-\listoftables % remove this if you have no tables.
+%\listoffigures % remove this if you have no figures.
+%\listoftables % remove this if you have no tables.
 
 \chapter{Acknowledgements} % this needs to be before \mainmatter.
-A lot of people are awesome. But not you, who are reading my thesis before it's done.
+
+{\raggedleft ``\emph{Dunque io ringrazio tutti quanti.} \par}
+{\raggedleft \emph{Specie la mia mamma che mi ha fatto cos\'i funky.}"\par}
+{\raggedleft -- Articolo 31, Tanqi Funky, 1996 -- \par}
+\vspace{0.5cm}
+
+{\raggedleft ``\emph{At last, I thank everyone.} \par}
+{\raggedleft \emph{Especially my mom who made me so funky.}"\par}
+{\raggedleft -- Articolo 31, Tanqi Funky, 1996 -- \par}
+\vspace{0.5cm}
+
+A lot of people are awesome, especially you, since you probably agreed to read this when it was a draft.
+
+%Bonnie. I don't even know where to start.
+%Flavio
+%
+
+%A big thank you to my academic sisters, a bunch of extraordinary women who have taught me so much about myself and our field.
+%From the youngest to the oldest (again academically speaking): Supraja, you worked the magic of enriching my past, thank you from the bottom of my heart;  Brooke, the depth of your knowledge, your drive and love for our field has been and will always be a constant inspiration;  Ariana, thanks for keeping me sane for the last four years, if you run for president, remember to delete  our Skype chats, love and hugs to you dudette; and last but not least Xiao, I don't think Fermilab has given me a bigger gift than your friendship.
+
+%Shany
+%Jen 
+%Jonathan
+%Andrzej
+%Anto
+%LArIAT & MicroBooNE
+%Louise
 
@@
 \mainmatter
+%\input{Chapters/0-Introduction}
+%\input{Chapter-1/1-Theory}
+%\input{Chapter-2/2-LArTPCs}
+%\input{Chapter-3/3-TheLArIATExperiment}
+\input{Chapter-4/4-Methodology}
 
-\input{0-Introduction}
-\input{1-LArTPCs}
-\input{2-TheLArIATExperiment}
-\input{3-KaonInteractionsInLAr}
-\input{4-DataCollection}
-\input{5-EnergyCalibration}
-\input{6-TrackingOptimization}
-\input{7-KaonCrossSectionAnalysis}
+\input{Chapter-5/5-Preparatory}
+\input{Chapter-6/6-PionCrossSection}
+\input{Chapter-7/7-KaonCrossSection}
+
+%\input{Chapter-7/7-BackgroundSub}
+%\input{Chapter-8/8-CrossSectionAnalysis}\input{Chapter-9/9-Uncertainty}
+
+% Only call appendix once, here.
+\appendix
+%\input{Chapters/A-KaonAna}
+\input{AppendixB-EField/B-LArIATEField}
+%\input{Chapters/C-MicroBooNECRT}
+
+% Old chapters configuration %
+%\input{Chapters/4-DataCollection}
+%\input{Chapters/4-MonteCarlo}
+%\input{Chapters/5-EnergyCalibration}
+%\input{Chapter-8/8-TrackingOptimization}
 

### 2018-05-14 Chapter-4/4-Methodology.tex
s

@@
 Beamline $K^+$   Candidate                 &    N.A       & 2837     \\ \hline
-Events Surviving Pile Up Filter              &   108929  &              \\ \hline
-Events with WC2TPC Match                 &    41757   &              \\ \hline
+Events Surviving Pile Up Filter              &   108929  & 2389       \\ \hline
+Events with WC2TPC Match                 &    41757   & 1081 \\ \hline
 Events Surviving Shower Filter             &    40841    &  N.A.     \\ \hline
-Available Events For Cross Section      &   40841    &       \\ \hline
+Available Events For Cross Section      &   40841    &   1081    \\ \hline
 \end{tabular}
@@
   \centering  
-\includegraphics[width=3in]{Chapter-4/Images/Shower.png}
+\includegraphics[width=\textwidth]{Chapter-4/Images/Shower.png}
 \caption{Visual rendering of the shower filter.
@@
 For each event passing the selection on its beamline information, we need to identify the track inside the TPC corresponding to the particle which triggered the beamline detectors, a procedure we refer to as ``WC to TPC match" (WC2TPC for short).
-In general, the TPC tracking algorithm will reconstruct more than one track in the event, partially due to the fact that hadrons interact in the chamber and partially because of pile up particles during the triggered TPC drift time, as shown in figure \ref{fig:kaonInteraction}.
-\textcolor{red}{ADD EVENT DISPLAYS}
+In general, the TPC tracking algorithm will reconstruct more than one track in the event, partially due to the fact that hadrons interact in the chamber and partially because of pile up particles during the triggered TPC readout time, as shown in figure~\ref{fig:kaonInteraction}.
 
@@
   \centering  
-\includegraphics[width=3in]{Chapter-4/Images/face-smile-big.png}
+\includegraphics[width=\textwidth]{Chapter-4/Images/KaonExample.png}
 \caption{Kaon candidate event: on the right, event display showing raw quantities; on the left, event display showing reconstructed tracks.
@@
 We attempt to uniquely match one wire chamber track to one and only one reconstructed TPC track. 
-In data, this match leverages on a geometrical selection exploiting both the position of the wire chamber and TPC tracks, and the angle between them.
-We consider only TPC tracks whose first point is in the first 2 cm upstream portion of the TPC for the match.
-We project the wire chamber track to the TPC front face where we define the $x_{FF}$ and $y_{FF}$ coordinates used for evaluating the match.
-We define $\Delta$X as the difference between the $x$ position of the most upstream point of the TPC track and $x_{FF}$.
+In order to determine if the presence of a match, we apply a geometrical selection on the relative the position of the wire chamber and TPC tracks. 
+We start by considering only TPC tracks whose first point is in the first 2 cm upstream portion of the TPC for the match.
+We project the wire chamber track to the TPC front face where we define the coordinates of the projected point as  $x_{FF}$ and $y_{FF}$.
+For each considered TPC track, we define $\Delta$X as the difference between the $x$ position of the most upstream point of the TPC track and $x_{FF}$.
 $\Delta$Y is defined analogously.
 We define the radius difference, $\Delta$R, as $ \Delta \text{R} =  \sqrt{ \Delta \text{X}^2 +  \Delta \text{Y}^2}  $.
-The angle between the incident WC track and the TPC track in the plane that contains them defines $\alpha$.
+We define  as $\alpha$ the angle between the incident WC track and the TPC track in the plane that contains them.
 If  $\Delta \text{R} < 4 $~cm, $\alpha < 8^\circ $,  a match between WC-track and TPC reconstructed track is found.
-We describe  how we determinate the best value for the radius and angular selection in sec \ref{ch:WC2TPCMatchOptimization}.
+We describe  how we determine the value for the radius and angular selection in sec \ref{ch:WC2TPCMatchOptimization}.
 In MC, we mimic the matching between the WC and the TPC track by constructing a fake WC track using truth information at wire chamber four.
@@
 
+\begin{figure}
+  \centering  
+\includegraphics[width=\textwidth]{Chapter-4/Images/Shower.png}
+\caption{Visual rendering of the wire chamber to TPC match.}
+\label{fig:showerFilt}
+\end{figure}
+
 \section{The Thin Slice Method}\label{ch:ThinSliceMethod}
+Once the pool of hadron candidate is selected and the TPC track corresponding to the beamline event is identified, we apply the thin slice method to measure the cross section, as the following sections describe.
 \subsection{Cross Sections on Thin Target}

### 2018-05-14 Chapter-4/4-Methodology.tex
s

@@
-\chapter{Total Hadronic Cross Section Measurement Methodology}\label{ch:Interactions}
+\chapter{Total Ha($\pi^{-}$,Ar)dronic Cross Section Measurement Methodology}\label{ch:Interactions}
 This chapter describes the general procedure employed to measure a total hadronic differential cross section in LArIAT.
@@
 \section{The Thin Slice Method}\label{ch:ThinSliceMethod}
-Once the pool of hadron candidate is selected and the TPC track corresponding to the beamline event is identified, we apply the thin slice method to measure the cross section, as the following sections describe.
+Once we have selected the pool of hadron candidates and we have identified the TPC track corresponding to the beamline event, we apply the thin slice method to measure the cross section, as the following sections describe.
 \subsection{Cross Sections on Thin Target}
@@
 In general, the target is not a single particle, but rather a slab of material containing many diffusion centers.
-The so-called  ``thin target" approximation assumes that the target centers are uniformly distributed in the material and that the target is thin compared to the interaction length so that no center of interaction sits in front of another.
+The so-called  ``thin target" approximation assumes that the target centers are uniformly distributed in the material and that the target is thin compared to the projectile interaction length,WC2TPC so that no center of interaction sits in front of another.
 In this approximation, the ratio between the number of particles interacting in the target $N_{Interacting}$ and number of incident particles $N_{Incident}$ determines the interaction probability $P_{Interacting}$, which is the complementary to one of the survival probability $P_{Survival}$.
@@
 \subsection{Not-so-Thin Target: Slicing the Argon}
-The LArIAT TPC, with its 90 cm of length, is not a thin target. However, the fine-grained tracking of the LArIAT LArTPC allows us to treat the argon volume as a sequence of many adjacent thin targets.
+The interaction length of pions and kaons in argon is expected to be of the order of 50 cm for pions and 100 cm for kaons.
+Thus, the LArIAT TPC, with its 90 cm of length, is not a thin target.
+However, the fine-grained tracking of the LArIAT LArTPC allows us to treat the argon volume as a sequence of many adjacent thin targets.
 
@@
 Considering each slice {\emph{j}}  a ``thin target",  we can apply the cross section calculation from Eq.~\ref{eq:thinTargetXSSolved} iteratively, evaluating the kinetic energy of the hadron as it enters each slice, $E_{j}^{kin}$.
-For each WC-to-TPC matched particle, the energy of the hadron entering the TPC is known thanks to the momentum and mass determination by the tertiary beamline,
+For each WC2TPC matched particle, the energy of the hadron entering the TPC is known thanks to the momentum and mass determination by the tertiary beamline,
 
@@
 
+\subsection{Corrections to the Raw Cross Section}\label{ch:MCCorrections}
+%%%%%%%%%%%%%%%%%%%%%%%%%%
+Equation \ref{eq:thinTargetXSSolved}  is a prescription for measuring the cross section in case of a pure beam of the hadron of interest and 100\% efficiency in the determination of the interaction point.
+For example, if LArIAT had a beam of pure pions and were 100\% efficient in determining the interaction point within the TPC, the pion cross section in each energy bin would be given by
+
+\begin{equation}
+ \sigma^{\pi^-}(E_{i})  = \frac{1}{n \delta X}\frac{N^{\pi^-}_{ \text{Interacting}} (E_{i})}{N^{\pi^-}_{ \text{Incident}}(E_{i})}.
+\label{eq:thinTargetXSSolved2}
+\end{equation}
+
+Unfortunately, this is not the case.
+In fact, the selection used to isolate pions in the LArIAT beam allows for the presence of some muons and electrons as background.
+Also, the LArIAT TPC is not 100\% efficient in determining the interaction point.
+Therefore we need to apply two corrections evaluated on the MC in order to extract the cross section from LArIAT data: the background subtraction and the efficiency correction.
+Still using the pion case as example, we estimate the pion cross section in each energy bin changing  the equation \ref{eq:thinTargetXSSolved2} into
+\begin{equation}
+ \sigma^{\pi^-}(E_{i})  =\frac{1}{n \delta X}\frac{N^{\pi^-}_{ \text{Interacting}} (E_{i})}{N^{\pi^-}_{ \text{Incident}}(E_{i})} = \frac{1}{n \delta X}\frac{ \epsilon^{inc}_i [ N^{ \text{TOT}}_{ \text{Interacting}} (E_{i}) - B_{ \text{Interacting}} (E_i)] }{   \epsilon^{int}_i [N^{ \text{TOT}}_{ \text{Incident}}(E_{i}) - B_{ \text{Incident}} (E_i)]},
+\label{eq:True}
+\end{equation}
+
+where  $N^{\text{TOT}}_{\text{Interacting}} (E_{i})$ and $N^{\text{TOT}}_{\text{Incident}} (E_{i})$ is the measured content of the interacting and incident histograms for events that pass the event selection, $B_{interacting} (E_i)$ and $B_{\text{Incident}} (E_i)$ represent the contributions from beamline background, and  $\epsilon^{int}_i$ and  $\epsilon^{inc}_i$ are the efficiency corrections for said histograms.
+
+As we will show in section \ref{sec:Correction}, the background subtraction for the interacting and incident histograms can be translated into a corresponding corrections $C^{\pi MC}_{Interacting} (E_{i})$ and $C^{\pi MC}_{Incident} (E_{i})$ and the cross section re-written as follows
+
+\begin{equation}
+   \sigma^{\pi^-}(E_{i})  = \frac{1}{n \delta X}\frac{ \epsilon^{inc}_i  \text{  } N^{\text{TOT}}_{\text{Interacting}} (E_{i}) \text{  } C^{\pi MC}_{Interacting} (E_{i}) }{   \epsilon^{int}_i \text{  }N^{\text{TOT}}_{\text{Incident}} (E_{i})\text{  } C^{\pi MC}_{Incident} (E_{i})}.
+\label{eq:C}
+\end{equation}
+
+%The following sections describe the procedures used to evaluate  the background subtraction (section \ref{sec:beamCont}) and the efficiency correction (section \ref{sec:EffCorrection}), as well as  their uncertainties.
+%The reader might be concerned about bin-by-bin migration of events in the interacting and incident plots due to the finite resolution of the energy reconstruction.
+In section \ref{sec:Energy}, we make an argument to why we expect the smearing matrix to be extremely close to diagonal, such that its calculation and relative corrections are left for an improvement of the analysis.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \section{Procedure testing with truth quantities}\label{ch:procedureTesting}
-The $\pi^{-}$-Ar and K$^{+}$-Ar total hadronic cross section implemented in Geant4 can be used as a tool to validate the measurement methodology.
+The ($\pi^{-}$,Ar) and (K$^{+}$,Ar) total hadronic cross section implemented in Geant4 can be used as a tool to validate the measurement methodology.
 We describe here a closure test done on Monte Carlo to prove that the methodology of slicing the TPC retrieves the underlying cross section distribution implemented in Geant4 within the statistical uncertainty.
@@
 
-For the validation test, we fire about 390000 pions and 140000 kaons inside the LArIAT TPC active volume using the Data Driven Monte Carlo (see section \ref{sec:DDMC}).
+For the validation test, we fire about a sample of pions and a sample of kaons inside the LArIAT TPC active volume using the Data Driven Monte Carlo (see section \ref{sec:DDMC}).
 We apply  the thin-sliced method using only true quantities to calculate the hadron kinetic energy at each slab in order to decouple reconstruction effects from issues with the methodology.
@@
 Then, we recursively subtracted it from the hadron kinetic energy at the TPC front face to evaluate the kinetic energy at each slab until the true interaction point is reached.
+Since the MC is a pure beam of the hadron of interest and truth information is used to retrieve the interaction point, no correction is applied.
 Doing so, we obtain the true interacting and incident distributions for the considered hadron and we obtain the true MC cross section as a function of the hadron true kinetic energy.
@@
 \end{minipage}
-\caption{Hadronic cross sections for $\pi^-$-Ar (left) and K$^+$-Ar (right) implemented in Geant4 10.01.p3 (solid lines) overlaid the true MC cross section as obtained with the sliced TPC method (markers).
+\caption{Hadronic cross sections for ($\pi^-$,Ar) on the left and (K$^+$,Ar) on the right as implemented in Geant4 10.01.p3 (solid lines) overlaid the true MC cross section as obtained with the sliced TPC method (markers).
 The total cross section is shown in green,  the elastic cross section in blue and the inelastic cross section in red.}
@@
 %\end{comment}
-
-\section{Corrections to Data}\label{ch:MCCorrections}

### 2018-05-14 Chapter-4/4-Methodology.tex
s

@@
 Albeit with small differences, both the  ($\pi^{-}$,Ar) and (K$^{+}$,Ar) total hadronic cross section measurements rely on the same procedure described in details in the following sections.
-We start by selecting the particle of interest using a combination of beamline detectors and TPC information (Section \ref{ch:ParticleSelectionMethod}).
+We start by selecting the particle of interest using a combination of beamline detectors and TPC information (Section \ref{ch:ParticleSelectionMeEquationthod}).
 We then perform a handshake between the beamline information and the TPC tracking to assure the selection of the right TPC track (Section \ref{ch:WC2TPCMatchMethod}).
@@
 
-Considering each slice {\emph{j}}  a ``thin target",  we can apply the cross section calculation from Eq.~\ref{eq:thinTargetXSSolved} iteratively, evaluating the kinetic energy of the hadron as it enters each slice, $E_{j}^{kin}$.
+Considering each slice {\emph{j}}  a ``thin target",  we can apply the cross section calculation from Equation~\ref{eq:thinTargetXSSolved} iteratively, evaluating the kinetic energy of the hadron as it enters each slice, $E_{j}^{kin}$.
 For each WC2TPC matched particle, the energy of the hadron entering the TPC is known thanks to the momentum and mass determination by the tertiary beamline,
@@
 Therefore we need to apply two corrections evaluated on the MC in order to extract the cross section from LArIAT data: the background subtraction and the efficiency correction.
-Still using the pion case as example, we estimate the pion cross section in each energy bin changing  the equation \ref{eq:thinTargetXSSolved2} into
+Still using the pion case as example, we estimate the pion cross section in each energy bin changing  Equation \ref{eq:thinTargetXSSolved2} into
 \begin{equation}

### 2018-05-14 Chapter-3/3-TheLArIATExperiment.tex
s

@@
 For LArIAT Run II, the High Yield sample is about three times the Picky Tracks statistics.
-For the first measurements of the LArIAT hadronic cross section, we use the Picky Tracks sample because the uncertainty on the momentum is smaller and the comparison with the beamline MC results is straightforward compared with the High Yield sample;  a possible future update and cross check of these analysis would be the use of the High Yield sample.
 

### 2018-05-14 Chapter-4/4-Methodology.tex
s

@@
-\chapter{Total Ha($\pi^{-}$,Ar)dronic Cross Section Measurement Methodology}\label{ch:Interactions}
+\chapter{Total Hadronic Cross Section Measurement Methodology}\label{ch:Interactions}
 This chapter describes the general procedure employed to measure a total hadronic differential cross section in LArIAT.
@@
 
-\subsubsection{Selection of Beamline Events}\label{ch:beamlineDetectorsData}
+\subsection{Selection of Beamline Events}\label{ch:beamlineDetectorsData}
 As shown in equation \ref{eq:enFF}, we leverage the beamline particle identification and momentum measurement before entering the TPC as in input to evaluate the kinetic energy for the hadrons used in the  cross sections measurements.
@@
 
-\subsubsection{Particle Identification in the beamline}
+\subsection{Particle Identification in the Beamline}
 In data, the main tool to establish the identity of the hadron of interest is the LArIAT tertiary beamline, in its function of mass spectrometer.
@@
 
-\subsubsection{TPC Selection: Halo mitigation }\label{ch:pileUp}
+\subsection{TPC Selection: Halo Mitigation }\label{ch:pileUp}
 The secondary beam impinging on LArIAT secondary target produces a plethora of particles which propagates downstream.
@@
 
-\subsubsection{TPC Selection: Shower Removal}\label{ch:electrons}
+\subsection{TPC Selection: Shower Removal}\label{ch:electrons}
 In the case of the ($\pi^-$,Ar) cross section, the resolution of  beamline mass spectrometer is not sufficient to select a beam of pure pions.

### 2018-05-14 Chapter-5/5-Preparatory.tex
s

@@
 \chapter{Preparatory Work}\label{ch:samples}
-This chapter describes the preparatory work done on the the data and Monte Carlo samples used for the cross section analyses. This entails:
+%{\raggedleft ``\emph{Il dolce non lo mangi mai, ma qualche volta ti rifai.} \par}
+%{\raggedleft \emph{Abbracciami}"\par}
+%{\raggedleft -- Pietro Ciampi, L'amore e' tutto qui, 1971 -- \par}
+%\vspace{0.5cm}
 
-\begin{enumerate}
-\item the MC production,
-\item the energy calibration of the detector both in data and MC,
-\item the optimization of the tracking algorithm for the total cross section analyses.
-\end{enumerate}
+%{\raggedleft ``\emph{You never eat dessert, but, sometimes, you make up for it.} \par}
+%{\raggedleft \emph{Hug me}"\par}
+%{\raggedleft -- Pietro Ciampi, L'amore e' tutto qui, 1971 -- \par}
+%\vspace{0.5cm}
 
-\section{Construction of a Monte Carlo Simulation for LArIAT}
+This chapter describes the preparatory work done on the the data and Monte Carlo samples used for the cross section analyses. 
+This entails the choice of the data set and the production of the information needed to construct the Monte Carlo Simulation~(section \ref{sec:dataSet}),  the construction and use of said Monte Carlo simulation~(section \ref{sec:MCSet}), the study and optimization of the tracking in the TPC for the cross section analyses~(section \ref{sec:TrackingStudies}),  the calibration of the calorimetry response and related energy studies~(section \ref{ch:energyCal}).
+
+\section{Cross Section Analyses Data Set}\label{sec:dataSet}
+We choose LArIAT Run-II as the data period for the  ($\pi^{-}$,Ar) and (K$^{+}$,Ar) total hadronic cross section analyses. 
+Data taking for the this period started on 03/15/2016  and ended on 07/31/2016. 
+Since we are interested in beamline and TPC information, we ask basic requirements on the operational status of the time of fight, wire chambers and TPC to form the good run list for this period, which we informally call ``lovely runs".
+
+The subset of lovely runs  chosen for the  ($\pi^{-}$,Ar) total hadronic cross section analysis includes only the -60A and -100A magnet configurations in negative polarity, even if LArIAT explored several other beamline configurations during Run-II.
+The -60A and -100A combined data set accounts for approximately 90\% of the total Run-II negative polarity runs.
+Since the production of beamline Monte Carlo depends on the wanted beamline configuration, the choice of only two beamline settings limits the need for beamline MC production.
+
+Similarly, the subset of lovely runs chosen  for the (K$^{+}$,Ar)  total hadronic cross section analysis includes only the +60A and +100A magnet configurations in positive polarity.
+It should be noted that kaons are extremely rare in the +60A sample, thus the data sample for the (K$^{+}$,Ar) cross section after the mass selection is about 90\% +100A runs, as shown in Table \ref{tab:databreakdown}.
+
+For the first measurements in LArIAT that uses both beamline and TPC information, we choose strict requirements on the reconstruction of the WC tracks, the so-called ``Picky Track" sample (see \ref{sec:MWPCfunc}).
+This choice presents two advantages:  the uncertainty on the momentum reconstruction for the ``Picky Tracks" sample is smaller compared to the ``High Yield" sample, and the comparison with the beamline MC results is straightforward.
+A possible future update and cross check of these analysis would be the use of the High Yield sample, where the statistics is about three times higher.
+
+The breakdown of beamline events as a function of the magnets settings is shown in Table \ref{tab:databreakdown}. 
+The choice of the data sets determines the production of beamline MC and serves as basis for the production of Data Driven MC, as shown in the next sections.
+
+\begin{table}[b]
+\centering
+\begin{tabular}{|l|c|c|c|}  
+\hline
+                                                              & I = 60 A          & I = 100 A   & Total     \\ \hline
+Data Events after $\pi/\mu/e$ Mass Selection     &     67068          &  71413  & 138481 \\ \hline
+Data Events after $K$ Mass Selection                &     274              &   2563   & 2837  \\ \hline
+\end{tabular}
+\caption{Number of data events which fit the $\pi/\mu/e$ or $K$ mass hypothesis as a function of magnet settings.}
+\label{tab:databreakdown}
+\end{table}
+
+\section{Construction of a Monte Carlo Simulation for LArIAT}\label{sec:MCSet}
+For the simulation of LArIAT events and their particle make up, we use a combination of two MC generators: the G4Beamline Monte Carlo and the Data Driven single particle Monte Carlo (DDMC).
+We use the G4Beamline MC to simulate the particle transportation in the beamline and calculate the particle composition of the beam just after the fourth Wire Chamber (WC4).
+In order to simulate the beamline particles after WC4 and in the TPC, we use the DDMC.
+
 \subsection{G4Beamline}\label{ch:beamlineComposition}
-\subsection{Data Driven MC}\label{sec:DDMC}
-
-\section{Tracking Studies}
-
-\section{Energy Calibration and Studies}
-
-\section{Estimate of Energy Loss before the TPC}
-
-\begin{comment}
-\section{Construction of a Monte Carlo Simulation for LArIAT}
-For the simulation of LArIAT events and their particle make up, we use a combination of two MC generators: the G4Beamline Monte Carlo and the Data Driven single particle Monte Carlo (DDMC).
-We use the G4Beamline MC to simulate the particle transportation in the beamline and calculate the particle composition of the beam just past the fourth Wire Chamber (WC4).
-In order to simulate the beam line particles after WC4 and in the TPC, we use the DDMC.
-\subsection{G4Beamline}\label{beamlineComposition}
-G4Beamline simulates the beam collision with the LArIAT secondary target, the energy deposited by the particles in the LArIAT beamline detectors and the action of the LArIAT magnets, effectively accounting for particle transportation through the beam line from the LArIAT target until ``Big Disk", a fictional, void detector located just before the cryostat.
+G4Beamline simulates the beam collision with the LArIAT secondary target, the energy deposited by the particles in the LArIAT beamline detectors and the action of the LArIAT magnets, effectively accounting for particle transportation through the beam line from the LArIAT target until ``Big Disk", a fictional, void detector located just before the LArIAT cryostat.
  At the moment of this writing, G4Beamline does not simulated the responses of the beam line detectors.
@@
 For the kaon cross section analysis the relevant beam conditions is a secondary beam energy of 64 GeV, positive polarity magnet with current of 100 A.
+
+\subsection{Data Driven MC}\label{sec:DDMC}
+\subsection{Estimate of Energy Loss before the TPC}
+
+\section{Tracking Studies}\label{sec:TrackingStudies}
+\subsection{Study of WC to TPC Match}\label{ch:WC2TPCMatchOptimization}
+
+\section{Energy Calibration and Studies}\label{ch:energyCal}
+
+\begin{comment}
+\section{Construction of a Monte Carlo Simulation for LArIAT}
+\subsection{G4Beamline}\label{beamlineComposition}
 

### 2018-05-14 Chapter-5/5-Preparatory.tex
s

@@
 \subsection{G4Beamline}\label{ch:beamlineComposition}
-G4Beamline simulates the beam collision with the LArIAT secondary target, the energy deposited by the particles in the LArIAT beamline detectors and the action of the LArIAT magnets, effectively accounting for particle transportation through the beam line from the LArIAT target until ``Big Disk", a fictional, void detector located just before the LArIAT cryostat.
+G4Beamline simulates the beam collision with the LArIAT secondary target, the energy deposited by the particles in the LArIAT beamline detectors, and the action of the LArIAT magnets, effectively accounting for particle transportation through the beam line from the LArIAT target until ``Big Disk", a fictional, void detector located just before the LArIAT cryostat.
  At the moment of this writing, G4Beamline does not simulated the responses of the beam line detectors.
@@
 In order to ease the handshake between G4Beamline and the DDMC, we ask for the beam composition just after WC4.
-Since LArIAT data are taken under different beam conditions, G4Beamline simulates separately the beam composition according to the magnets' settings and the secondary beam intensity.
+Since LArIAT data are taken under different beam conditions, we need to simulate separately the beam composition according to the magnets' settings and the secondary beam intensity with G4Beamline.
 For the pion cross section analysis the relevant beam conditions are  secondary beam energy of 64 GeV, negative polarity magnet with current of 100 A and 60 A.
@@
 
-\subsection{Data Driven MC}\label{sec:DDMC}
-\subsection{Estimate of Energy Loss before the TPC}
+\subsubsection{Beam Composition for Negative Pion Cross Section}
+Even if pions are by far the biggest beam component in negative polarity runs, the LArIAT tertiary beam is not a pure pion beam.
+While useful to discriminate between pions, kaons, and protons, the beamline detectors are not sensitive enough to  discriminate among the lighter particles in the beam: electrons, muons and pions fall under the same mass hypothesis.
+Thus, we need to assess the contamination from beamline particles other than pions in the event selections used for the pion cross section analysis and correct for its effects.
+The first step of this process is assessing the percentage of electrons and muons in the $\pi/\mu/e$ beamline candidates via the G4Beamline MC.
+The full treatment of the beamline contamination in the pion cross section calculation is described in section \ref{ch:PionXSBkgSub}.
+Since the beamline composition is a function of the magnet settings, we simulate separately events for magnet current of -60A and -100A. 
+Figure \ref{fig:BeamComposition} shows the momentum predictions from G4Beamline overlaid with data for the 60A runs (left) and for the 100A runs (right).
+The predictions for electrons, muons and pions have been staggered and their sum is area normalized to data.
+Albeit not perfect, these plots show a reasonable agreement between the momentum shapes in data and MC.
+We attribute  the difference in shape to the lack of simulation of the WC efficiency in the MC which is momentum dependent and leads to enhance the number events in the center of the momentum distribution.
 
-\section{Tracking Studies}\label{sec:TrackingStudies}
-\subsection{Study of WC to TPC Match}\label{ch:WC2TPCMatchOptimization}
+\begin{figure}
+\includegraphics[width=0.5\textwidth,height=\textheight,keepaspectratio]{Chapter-7/Images/Beam60A.png}
+\includegraphics[width=0.5\textwidth,height=\textheight,keepaspectratio]{Chapter-7/Images/Beam100A.png}
+\caption{Beam composition for the -60A runs (left) and -100A runs (right).
+The solid blue plot represents the simulated pion content, the yellow plot represents the simulated muon content and the grey plot represents the simulated electron content.
+The plots are area normalized to the number of data events, shown in red.
+}
+\label{fig:BeamComposition}
+\end{figure}
 
-\section{Energy Calibration and Studies}\label{ch:energyCal}
+Table \ref{tab:beamline} shows the beam composition per magnet setting after the mass selection according to the G4Beamline simulation.
+\begin{table}[]
+\centering
+\begin{tabular}{|l|c|c|}
+\hline
+                     & I = -60 A           & I = -100 A \\ \hline
+G4Pions       &   68.8 \%           &      87.4 \%        \\ \hline
+G4Muons     &     4.6 \%           &        3.7 \%         \\ \hline
+G4Electrons &   26.6 \%           &        8.9 \%        \\ \hline
+\end{tabular}
+\caption{Simulated beamline composition per magnet settings}
+\label{tab:beamline}
+\end{table}
 
-\begin{comment}
-\section{Construction of a Monte Carlo Simulation for LArIAT}
-\subsection{G4Beamline}\label{beamlineComposition}
+\subsubsection{Beam Composition for Positive Kaon Cross Section}
+In the positive polarity runs, the tertiary beam composition is mainly pions and protons.
+The left side of Figure \ref{fig:BeamCompositionPos} shows the  predictions for the momentum spectra for the 100A positive runs  according to  G4Beamline (solid colors) overlaid with data (black points).
+Since the LArIAT beamline detectors can discriminate between kaons and other particles, we do not rely on the G4Beamline simulation to estimate the beamline contamination in the pool of kaon candidates (as in the case of the pion cross section), but rather we use a data drive approach.
+The basic idea of this data driven approach is to estimate the bleed over from high and low mass peaks under the kaon peak by fitting the tails of the $\pi/\mu/e$ and proton mass distributions, as shown in Figure \ref{fig:BeamCompositionPos} right side.
+Since the shape of the tails is unknown, the estimate is done multiple times varying the range and shape for reasonable functions. 
+For example, to estimate the proton content under the kaon peak, we start by fitting the left tail of the proton mass distribution with a gaussian function between 650 $MeV/c^2$ and 750 $MeV/c^2$.% in a reasonable range and with a reasonable function.
+We extend the fit function under the kaon peak and integrate the between 350-650 $MeV/c^2$.
+We integrate the mass histogram in the same range and calculate the proton contamination as the ratio between the two integrals.
+We repeat this procedure for several fit shapes (gaussian, linear and exponential functions are used) and tail ranges.
+Finally, we calculate the contamination as the weighted average of single estimates, where the weights are calculated to be the $1./\chi^2$ of the tail fits.
+The procedure is repeated for lighter particles mass peak independently.
+With 12 iterations of this method we find a proton contamination of  0.2 $\pm$ 0.5 \%  and a contamination from the lighter particles of 5 $\pm$ 2 \% .
 
-\textcolor{red}{DECIDE IF YOU WANT THE BEAM COMPOSITION HERE}
-%Figure 16 shows the tertiary beam spectra for the 64 GeV and 100 A.
-
-%In Table 3, the beam composition is given in terms of percentage of different particle species per spill for positive polarity.
-The values reported are the weighted average on the two beam conditions considered.
-The weights are calculated according to the fourth column of Table 2.
+\begin{figure}
+\includegraphics[width=0.5\textwidth,height=\textheight,keepaspectratio]{Chapter-5/Images/Beam100Pos.png}
+\includegraphics[width=0.5\textwidth,height=\textheight,keepaspectratio]{Chapter-5/Images/MassPos.png}
+\caption{$Left.$ Beam composition for the +100A runs after WC4 (no mass selection applied).
+The solid blue plot represents the simulated pion content, the yellow plot represents the simulated muon content and the grey plot represents the simulated positron content, the red the proton content and the mustard the kaon content.
+The plots are area normalized to the number of data events, shown in black.$Right.$ Mass distribution for the Run-II positive runs, where the area under the kaon mass peak is highlighted in purple.
+The area under the extension of a possible fit for the proton tail is highlighted in red.
+}
+\label{fig:BeamCompositionPos}
+\end{figure}
 
 \subsection{Data Driven MC}\label{sec:DDMC}
-The Data Driven single particle Monte Carlo (DDMC) is a single particle MC gun which simulates the particle transportation from WC4 into the TPC leveraging on the beamline data information.
-The DDMC uses the data momentum and position at WC4 to derive its initial conditions: a general sketch of the DDMC workflow is shown in Figure \ref{fig:DDMCSketch}.
+The Data Driven single particle Monte Carlo (DDMC) is a single particle gun which simulates the particle transportation from WC4 into the TPC leveraging on the beamline data information.
+The DDMC uses the data momentum and position at WC4 to derive the event generation: a general sketch of the DDMC workflow is shown in Figure \ref{fig:DDMCSketch}.
 
 When producing a DDMC sample, beam line data from a particular running period and/or running condition are selected first.
+For example, data for the negative 60A runs and for the negative 100A runs inform the event generation stage of two different DDMC samples.
 Figure \ref{fig:DDMCQuantities}  schematically shows the data quantities of interest leveraged from data: the momentum ($P_x, P_y, P_z$) and position ($X, Y$) at WC4.
-For each data event, we obtain the particle position ($X, Y$) at WC4 directly from the data measurement.
-On the contrary, we calculate the components of the momentum using the beamline measurement of the momentum magnitude (see section \ref{sec:MWPCfunc}) in conjunction with the hits on WC3 and WC4 to determine the direction of the momentum vector, as described in \ref{sec:MWPCfunc}.
-The momentum and position of the selected data is sampled thousand of times through a 5-dimensional hit-or-miss sampling procedure.
-This produces MC distributions with the same momentum and position distributions as data, with the additional benefit of accounting for the correlations between the considered variables.
+For each data event, we obtain the  particle position ($X, Y$) at WC4 directly from the data measurement; we calculate the components of the momentum using the beamline measurement of the momentum magnitude in conjunction with the hits on WC3 and WC4 to determine the direction of the momentum vector, as described in section \ref{sec:MWPCfunc}.
+The momentum and position of the selected data
+
+The momentum and position of the selected data form a 5-dimensional tupla, which we sample thousand of times through a 5-dimensional hit-or-miss sampling procedure to generate the MC events.
+This produces MC $P_x, P_y, P_z, X, Y$ distributions  with the same momentum and position distributions as data, with the additional benefit of accounting for the correlations between the considered variables.
 A LArSoft simulation module then launches single particle MC from z = -100 cm (the location of the WC4) using the sampled momentum and position distributions as a template.
 As an example, the results of the DDMC generation compared to data for the pion 60A sample are shown in figure \ref{DDMCComparison}; as expected, MC and data agree within the statistical uncertainty by construction.
-Using this technique ensures the MC and data particles have very similar momentum, position and angular distributions at WC4 and allow us to us the MC sample in several occasions, for example to calibrate the energy loss upstream of the TPC or account for the WC2TPC match inefficiency.
+Using this technique ensures the MC and data particles have very similar momentum, position and angular distributions at WC4 and allow us to us the MC sample in several occasions, for example to calibrate the energy loss upstream of the TPC (see Section \ref{ch:eloss}) or to study the tracking and the calorimetric performance (sections \ref{sec:TrackingStudies} and \ref{ch:energyCal}).
 A small caveat is in order here: the DDMC is a single particle Monte Carlo, which means that the beam pile-up is not simulated.
+
 Three sample of \textcolor{red}{NUMBERS} pions, muons and electrons, as well as  a sample of \textcolor{red}{NUMBERS} kaons have been generated with the DDMC and are used for the MC cross section study.
@@
 \end{figure}
+
+\subsection{Estimate of Energy Loss before the TPC}\label{ch:eloss}
+
+\section{Tracking Studies}\label{sec:TrackingStudies}
+\subsection{Study of WC to TPC Match}\label{ch:WC2TPCMatchOptimization}
+
+\section{Energy Calibration and Studies}\label{ch:energyCal}
+
+\begin{comment}
 

### 2018-05-14 Chapter-7/7-BackgroundSub.tex
s

@@
 \section{Assessing Beamline Contamination}
-Even if pions are by far the biggest beam component in negative polarity runs, the LArIAT beam is not a pure pion beam.
+%Even if pions are by far the biggest beam component in negative polarity runs, the LArIAT beam is not a pure pion beam.
 While useful to discriminate between pions, kaons, and protons, the beamline detectors are not sensitive enough to  discriminate among the lighter particles in the beam: electrons, muons and pions fall under the same mass hypothesis.
@@
 
-\subsection{Electron and Muon contamination}
-We estimate the percentage of electrons and muons in the beam via the beamline MC. 
-Since the beamline composition is a function of the magnet settings, we simulate separately events for magnet current of -60A and -100A.
+%\subsection{Electron and Muon contamination}\label{ch:beamlineComposition}
+%We estimate the percentage of electrons and muons in the beam via the G4Beamline MC. 
+%Since the beamline composition is a function of the magnet settings, we simulate separately events for magnet current of -60A and -100A.
 
-\begin{figure}
-\includegraphics[width=0.5\textwidth,height=\textheight,keepaspectratio]{Chapter-7/Images/Beam60A.png}
-\includegraphics[width=0.5\textwidth,height=\textheight,keepaspectratio]{Chapter-7/Images/Beam100A.png}
-\caption{Beam composition for the -60A runs (left) and -100A runs (right).
-The solid blue plot represents the simulated pion content, the yellow plot represents the simulated muon content and the grey plot represents the simulated electron content.
-The plots are area normalized to the number of data events, shown in red.
+%\begin{figure}
+%\includegraphics[width=0.5\textwidth,height=\textheight,keepaspectratio]{Chapter-7/Images/Beam60A.png}
+%\includegraphics[width=0.5\textwidth,height=\textheight,keepaspectratio]{Chapter-7/Images/Beam100A.png}
+%\caption{Beam composition for the -60A runs (left) and -100A runs (right).
+The solid blue plot represents the simulated pion content, the yellow plot %represents the simulated muon content and the grey plot represents the simulated electron content.
+The plots are area normalized to the number of %data events, shown in red.
 }
-\label{fig:BeamComposition}
-\end{figure}
+%\label{fig:BeamComposition}
+%\end{figure}
 
-Table \ref{tab:beamline} shows the beam composition per magnet setting after the mass selection according to the G4Beamline simulation.
-\begin{table}[]
-\centering
-\begin{tabular}{|l|c|c|}
-\hline
-                     & I = -60 A           & I = -100 A \\ \hline
-G4Pions       &   68.8 \%           &      87.4 \%        \\ \hline
-G4Muons     &     4.6 \%           &        3.7 \%         \\ \hline
-G4Electrons &   26.6 \%           &        8.9 \%        \\ \hline
-\end{tabular}
-\caption{Simulated beamline composition per magnet settings}
-\label{tab:beamline}
-\end{table}
+%Table \ref{tab:beamline} shows the beam composition per magnet setting after the mass selection according to the G4Beamline simulation.
+%\begin{table}[]
+%\centering
+%\begin{tabular}{|l|c|c|}
+%\hline
+ %                    & I = -60 A           & I = -100 A \\ \hline
+%G4Pions       &   68.8 \%           &      87.4 \%        \\ \hline
+%G4Muons     &     4.6 \%           &        3.7 \%         \\ \hline
+%G4Electrons &   26.6 \%           &        8.9 \%        \\ \hline
+%\end{tabular}
+%\caption{Simulated beamline composition per magnet settings}
+%\label{tab:beamline}
+%\end{table}
 
-\begin{table}[]
-\centering
-\begin{tabular}{|l|c|c|c|c|c|}
-\hline
-                                                     & I = -60 A          & I = -100 A   & Total     &  w$_{60A}$ &  w$_{100A}$\\ \hline
-N Data Events after Mass Selection     &     70192          &  76056       & 146248 & 0.48 & 0.52\\ \hline
-%Data events for Cross Section          &                         &                   &              & & \\ \hline
-%Estimated e in the beamline             &   18671.1         &    6769.0  &              & & \\ \hline
-%Estimated $\mu$ in the beamline     &     3228.8         &     2814.1     &              & & \\ \hline
-%Estimated $\pi$ in the beamline       &    48292.1        &   66472.9   &              & & \\ \hline
-\end{tabular}
-\caption{Number of data events which fit the pion mass hypothesis as a function of magnet settings. The last two columns represent the faction of the data in the given magnet setting.}
-\label{tab:databreakdown}
-\end{table}
+%\begin{table}[]
+%\centering
+%\begin{tabular}{|l|c|c|c|c|c|}
+%\hline
+%                                                     & I = -60 A          & I = -100 A   & Total     &  w$_{60A}$ &  w$_{100A}$\\ \hline
+%N Data Events after Mass Selection     &     70192          &  76056       & 146248 & 0.48 & 0.52\\ \hline
+%%Data events for Cross Section          &                         &                   &              & & \\ \hline
+%%Estimated e in the beamline             &   18671.1         &    6769.0  &              & & \\ \hline
+%%Estimated $\mu$ in the beamline     &     3228.8         &     2814.1     &              & & \\ \hline
+%%Estimated $\pi$ in the beamline       &    48292.1        &   66472.9   &              & & \\ \hline
+%\end{tabular}
+%\caption{Number of data events which fit the pion mass hypothesis as a function of magnet settings. The last two columns represent the faction of the %data in the given magnet setting.}
+%\label{tab:databreakdown}
+%\end{table}
 
-We calculate the electron to pion, as well as the muon to pion ratio on the whole sample as the weighted sum of the corresponding ratio in the two current settings, 
-\begin{equation}
-\frac{N_e}{N_\pi}_{Data} = w_{60A}\frac{N_e}{N_\pi}_{60A}  + w_{100A}\frac{N_e}{N_\pi}_{100A},
-\end{equation}
-\begin{equation}
-\frac{N_\mu}{N_\pi}_{Data} = w_{60A}\frac{N_\mu}{N_\pi}_{60A}  + w_{100A}\frac{N_\mu}{N_\pi}_{100A},
-\end{equation}
-where the weights $w_{60A}$ and $w_{100A}$ are the percentage of events in the corresponding magnet configuration passing the mass selection in data, as shown in table \ref{tab:databreakdown}.
-Figure \ref{fig:BeamComposition} shows the mometum predictions from G4Beamline overlaid with data for the 60A runs (left) and for the 100A runs (right).
-The predictions for electrons, muons and pions have been staggered and their sum is area normalized to data.
+%We calculate the electron to pion, as well as the muon to pion ratio on the whole sample as the weighted sum of the corresponding ratio in the two %current settings, 
+%\begin{equation}
+%\frac{N_e}{N_\pi}_{Data} = w_{60A}\frac{N_e}{N_\pi}_{60A}  + w_{100A}\frac{N_e}{N_\pi}_{100A},
+%\end{equation}
+%\begin{equation}
+%\frac{N_\mu}{N_\pi}_{Data} = w_{60A}\frac{N_\mu}{N_\pi}_{60A}  + w_{100A}\frac{N_\mu}{N_\pi}_{100A},
+%\end{equation}
+%where the weights $w_{60A}$ and $w_{100A}$ are the percentage of events in the corresponding magnet configuration passing the mass selection in %data, as shown in table \ref{tab:databreakdown}.
+Figure \ref{fig:BeamComposition} shows the mometum predictions from G4Beamline overlaid with %data for the 60A runs (left) and for the 100A runs (right).
+The predictions for electrons, muons and pions have been staggered and their sum is area %normalized to data.
 Albeit not perfect, these plots show a reasonable agreement between the momentum shapes in data and MC.
-We attribute  the difference in shape to the lack of simulation of the WC efficiency in the MC which is momentum dependent and leads to enhance the number events in the center of the momentum distribution.
+We attribute  the %difference in shape to the lack of simulation of the WC efficiency in the MC which is momentum dependent and leads to enhance the number events in %the center of the momentum distribution.
 

### 2018-05-14 Chapter-5/5-Preparatory.tex
s

@@
 \begin{figure}
-\includegraphics[width=0.5\textwidth,height=\textheight,keepaspectratio]{Chapter-7/Images/Beam60A.png}
-\includegraphics[width=0.5\textwidth,height=\textheight,keepaspectratio]{Chapter-7/Images/Beam100A.png}
+\includegraphics[width=0.5\textwidth,height=\textheight,keepaspectratio]{Chapter-5/Images/Beam60A.png}
+\includegraphics[width=0.5\textwidth,height=\textheight,keepaspectratio]{Chapter-5/Images/Beam100A.png}
 \caption{Beam composition for the -60A runs (left) and -100A runs (right).
@@
 For each data event, we obtain the  particle position ($X, Y$) at WC4 directly from the data measurement; we calculate the components of the momentum using the beamline measurement of the momentum magnitude in conjunction with the hits on WC3 and WC4 to determine the direction of the momentum vector, as described in section \ref{sec:MWPCfunc}.
-The momentum and position of the selected data
+The momentum and position of the selected data form a 5-dimensional tupla, which we sample thousands of times through a 5-dimensional hit-or-miss sampling procedure to generate the MC events.
+This produces MC $P_x, P_y, P_z, X, Y$ distributions  with the same momentum and position distributions as data, with the additional benefit of accounting for the correlations between the considered variables.
+As an example, the results of the DDMC generation compared to data for the kaon +100A sample are shown in figure \ref{DDMCComparison} for the $P_z$, $X$ and $Y$ distributions; as expected, MC and data agree within the statistical uncertainty by construction.
+A LArSoft simulation module then launches single particle MC from z = -100 cm (the location of the WC4) using the MC generated events.
+The particles are free to decay and interact in their path from WC4 to the TPC according to the Geant4 simulation.
 
-The momentum and position of the selected data form a 5-dimensional tupla, which we sample thousand of times through a 5-dimensional hit-or-miss sampling procedure to generate the MC events.
-This produces MC $P_x, P_y, P_z, X, Y$ distributions  with the same momentum and position distributions as data, with the additional benefit of accounting for the correlations between the considered variables.
-A LArSoft simulation module then launches single particle MC from z = -100 cm (the location of the WC4) using the sampled momentum and position distributions as a template.
-As an example, the results of the DDMC generation compared to data for the pion 60A sample are shown in figure \ref{DDMCComparison}; as expected, MC and data agree within the statistical uncertainty by construction.
-Using this technique ensures the MC and data particles have very similar momentum, position and angular distributions at WC4 and allow us to us the MC sample in several occasions, for example to calibrate the energy loss upstream of the TPC (see Section \ref{ch:eloss}) or to study the tracking and the calorimetric performance (sections \ref{sec:TrackingStudies} and \ref{ch:energyCal}).
+Using the DDMC technique ensures that the MC and data particles have very similar momentum, position and angular distributions at WC4 and allows us to use the MC sample in several occasions, for example to calibrate the energy loss upstream of the TPC (see Section \ref{ch:eloss}) or to study the tracking and the calorimetric performance (sections \ref{sec:TrackingStudies} and \ref{ch:energyCal}).
 A small caveat is in order here: the DDMC is a single particle Monte Carlo, which means that the beam pile-up is not simulated.
 
-Three sample of \textcolor{red}{NUMBERS} pions, muons and electrons, as well as  a sample of \textcolor{red}{NUMBERS} kaons have been generated with the DDMC and are used for the MC cross section study.
+Six samples are the basis fo the MC used in the pion cross section measurement: three samples of  $\sim$340000 pions, muons and electrons to simulate the negative 60A runs, and three samples of $\sim$340000 pions, muons and electrons for the negative 100A runs.
+
+The MC used for the kaon cross section analysis is a sample of \textcolor{red}{NUMBERS}  kaons.
 
@@
 \centering
-\includegraphics[width=\textwidth]{Chapter-5/Images/Pz.png}
-\caption{Comparison between generated quantities and data distributions for the 60A pion sample: Z component of the momentum (top left), X position at Wire Chamber 4 (top right), Y position at Wire Chamber 4 (bottom).}
+\includegraphics[width=0.48\textwidth]{Chapter-5/Images/DDMCPz.png}
+\includegraphics[width=0.48\textwidth]{Chapter-5/Images/DDMCX.png}
+\includegraphics[width=0.48\textwidth]{Chapter-5/Images/DDMCY.png}
+\caption{Comparison between generated quantities and data distributions for the 100A kaon sample: Z component of the momentum at WC4 (top left), X position at Wire Chamber 4 (top right), Y position at Wire Chamber 4 (bottom).}
 \label{fig:DDMCComparison}
@@
 \subsection{Estimate of Energy Loss before the TPC}\label{ch:eloss}
+The beamline particles travel a path from where their  momentum is measured in the beamline until they are tracked again inside the TPC.
+In the LArIAT geometry, a particle leaving the WC4 will encounter the materials listed in Table \ref{tab:budget} before being registered again.
+The energy lost by the particle in this non-instrumented material modifies the particle's kinetic energy and directly affects the cross section measurement, as shown in equation \ref{eq:enFF}.
+
+\begin{table}[h!]
+\centering
+\begin{tabular}{|l|l|l|}
+\hline
+Material  & density {[}g/cm$^3${]} & width {[}cm{]}    \\ \hline
+Fiberglass laminate (G10)      & 1.7                             & 1.28                              \\
+Liquid Argon                           & 1.4                             & 3.20                             \\
+Stainless Steel                        & 7.7                            & 0.23                             \\
+Titanium                                  & 4.5                            & 0.04                             \\ 
+Air                                            &  1.2 $\cdot10^{-3}$  & 89.43                              \\
+Plastic Scintillator                    & 1.03                          & 1.20 (+ 1.30)                             \\ \hline
+\end{tabular}
+\caption{LArIAT material budget from WC4 to the TPC Front Face.}
+\label{tab:budget}
+\end{table}
+
+We derive an estimate of the energy loss between the beamline momentum measurement and the TPC ($E_{loss}$) from the Data Driven Monte Carlo using the pion and kaon samples separately, since this quantity is not  measurable directly on data.
+The $E_{loss}$ distribution for the 60A  and 100A pion sample is shown in figure \ref{fig:ELoss60A}, left and right respectively.
+A clear double peaked structure is visible, which is due to the particles either missing or hitting the HALO paddle: a schematic rendering of this occurrence is  shown in figure \ref{fig:Halo}.
+The kinematic at WC4 determines the trajectory of a particle and whether or not it will hit the halo paddle.
+In figure \ref{fig:PxVsXTrue} , we plot the true  $X$ component of the momentum versus the true $X$ position at WC4 for pions missing the halo paddle (left) and for pions hitting the halo paddle (right) for the 60A MC simulation runs -- analogous plots are obtained with the 100A simulation.
+These distributions can be separated drawing a line in this position-momentum space.
+We use a logistic regression  \cite{agresti2013categorical}  as a classifier to find the best separating line, shown in both plots as the red line.
+We classify as ``hitting the halo paddle" all pions whose $P_x$ and $X$ are such that $$P_x +0.02* X - 0.4 < 0 $$ and as ``missing the halo  paddle" all pions whose $P_x$ and $X$ are such that $$P_x +0.02*X - 0.4 > 0, $$ where the coefficients of the line are empirically found by the logistic regression estimation.
+Overall, this simple classifier classifies in the right category (hit or miss) about 86\% of the pion events.
+We apply the same classifier on data.
+We assign  $E_{loss} = 32 \pm 4 $~MeV for events classified as ``hitting the halo paddle"; we assign  $E_{loss} = 24 \pm 3 $~MeV for events classified as ``missing the halo paddle".
+
+%We use the separation of these two distributions to decide what the energy loss for each event on data. 
+%Thus,  we assign the value for energy loss is used in the data.
+
+\begin{figure}[hbpt]
+\centering
+\includegraphics[width=0.45\textwidth]{Chapter-5/Images/E_loss60A.png}
+\includegraphics[width=0.45\textwidth]{Chapter-5/Images/E_loss100A.png}
+\caption{True energy loss between WC4 and the TPC front face according to the MC simulation of the 60A runs (left) and of the 100A runs (right).
+The distribution for the whole data sample is shown in blue, the distribution for the pions missing the halo is shown in red, and the distribution for the pions hitting the halo is shown in green.
+}
+\label{fig:ELoss60A}
+\end{figure}
+
+\begin{figure}[hbpt]
+\centering
+\includegraphics[scale=0.5]{Chapter-5/Images/Halo.png}
+\caption{Schematic rendering of the particle path between WC4 and the TPC front face.
+The paddle with the hollow central circle represents the Halo paddle.
+We illustrate two possible trajectories: in black, a trajectory that miss the paddle and goes through the hole in the Halo, in blue a trajectory that hits the Halo paddle and goes through the scintillation material.}
+\label{fig:Halo}
+\end{figure}
+
+\begin{figure}[hbpt]
+\centering
+\includegraphics[width=\textwidth]{Chapter-5/Images/PXVsX60A.png}
+\caption{Horizontal component of the true momentum vs the horizontal position at WC4 for MC simulated pions of the 60A runs.
+The plot on the left shows the distribution for pion that miss the halo paddle and the plot on the right shows the distributions for pions that hit the halo.
+The form of the classifier is overlaid to both plots (red line).}
+\label{fig:PxVsXTrue}
+\end{figure}
 

### 2018-05-14 Chapter-8/8-CrossSectionAnalysis.tex
s

@@
 \section{Estimate of $E_{loss}$ before the TPC}
-The beamline particles travel a path from when their momentum is measured by the beamline detector, until they are tracked again inside the TPC.
-In the current LArIAT geometry, a particle leaving the fourth wire chamber will encounter the materials listed in Table \ref{tab:budget} before being registered again.
-The energy lost by the particle in this non-instrumented material modifies the particle's kinetic energy and directly affects the cross section measurement, as shown in equation \ref{eq:enFF}.
-
-\begin{table}[h!]
-\centering
-\begin{tabular}{|l|l|l|}
-\hline
-Material  & density {[}g/cm$^3${]} & width {[}cm{]}    \\ \hline
-Fiberglass laminate (G10)      & 1.7                             & 1.28                              \\
-Liquid Argon                           & 1.4                             & 3.20                             \\
-Stainless Steel                        & 7.7                            & 0.23                             \\
-Titanium                                  & 4.5                            & 0.04                             \\ 
-Air                                            &  1.2 $\cdot10^{-3}$  & 89.43                              \\
-Plastic Scintillator                    & 1.03                          & 1.20 (+ 1.30)                             \\ \hline
-\end{tabular}
-\caption{LArIAT material budget from WC4 to the TPC Front Face.}
-\label{tab:budget}
-\end{table}
-
-We derive an estimate of the energy loss between the beamline momentum measurement and the TPC ($E_{loss}$) from the Data Driven Monte Carlo using the pion and kaon samples separately, since this quantity is not  measurable directly on data.
-The $E_{loss}$ distribution for the 60A  and 100A pion sample is shown in figure \ref{fig:ELoss60A}, left and right respectively.
-A clear double peaked structure is visible, which is due to the particles either missing or hitting the HALO paddle: a schematic rendering of this occurrence is  shown in figure \ref{fig:Halo}.
-The kinematic at WC4 determines the trajectory of a particle and whether or not it will hit the halo paddle.
-In figure \ref{fig:PxVsXTrue} , we plot the true  $X$ component of the momentum versus the true $X$ position at WC4 for pions missing the halo paddle (left) and for pions hitting the halo paddle (right) for the 60A MC simulation runs -- analogous plots are obtained with the 100A simulation.
-These distributions can be separated drawing a line in this position-momentum space.
-We use a logistic regression  \cite{agresti2013categorical}  as a classifier to find the best separating line, shown in both plots as the red line.
-We classify as ``hitting the halo paddle" all pions whose $P_x$ and $X$ are such that $$P_x +0.02* X - 0.4 < 0 $$ and as ``missing the halo  paddle" all pions whose $P_x$ and $X$ are such that $$P_x +0.02*X - 0.4 > 0, $$ where the coefficients of the line are empirically found by the logistic regression estimation.
-Overall, this simple classifier classifies in the right category (hit or miss) about 86\% of the pion events.
-We apply the same classifier on data.
-We assign  $E_{loss} = 32 \pm 4 $~MeV for events classified as ``hitting the halo paddle"; we assign  $E_{loss} = 24 \pm 3 $~MeV for events classified as ``missing the halo paddle".
-
-%We use the separation of these two distributions to decide what the energy loss for each event on data. 
-%Thus,  we assign the value for energy loss is used in the data.
-
-\begin{figure}[hbpt]
-\centering
-\includegraphics[width=0.45\textwidth]{Chapter-5/Images/E_loss60A.png}
-\includegraphics[width=0.45\textwidth]{Chapter-5/Images/E_loss100A.png}
-\caption{True energy loss between WC4 and the TPC front face according to the MC simulation of the 60A runs (left) and of the 100A runs (right).
-The distribution for the whole data sample is shown in blue, the distribution for the pions missing the halo is shown in red, and the distribution for the pions hitting the halo is shown in green.
-}
-\label{fig:ELoss60A}
-\end{figure}
-
-\begin{figure}[hbpt]
-\centering
-\includegraphics[scale=0.5]{Chapter-5/Images/Halo.png}
-\caption{Schematic rendering of the particle path between WC4 and the TPC front face.
-The paddle with the hollow central circle represents the Halo paddle.
-We illustrate two possible trajectories: in black, a trajectory that miss the paddle and goes through the hole in the Halo, in blue a trajectory that hits the Halo paddle and goes through the scintillation material.}
-\label{fig:Halo}
-\end{figure}
-
-\begin{figure}[hbpt]
-\centering
-\includegraphics[width=\textwidth]{Chapter-5/Images/PXVsX60A.png}
-\caption{Horizontal component of the true momentum vs the horizontal position at WC4 for MC simulated pions of the 60A runs.
-The plot on the left shows the distribution for pion that miss the halo paddle and the plot on the right shows the distributions for pions that hit the halo.
-The form of the classifier is overlaid to both plots (red line).}
-\label{fig:PxVsXTrue}
-\end{figure}
-
 \section{Interacting and Incident Distributions}

### 2018-05-14 Chapter-5/5-Preparatory.tex
s

@@
 
-We derive an estimate of the energy loss between the beamline momentum measurement and the TPC ($E_{loss}$) from the Data Driven Monte Carlo using the pion and kaon samples separately, since this quantity is not  measurable directly on data.
+We derive an estimate of the energy loss between the beamline momentum measurement and the TPC ($E_{loss}$) from the pion DDMC sample, since this quantity is not  measurable directly on data. 
 The $E_{loss}$ distribution for the 60A  and 100A pion sample is shown in figure \ref{fig:ELoss60A}, left and right respectively.
@@
 The kinematic at WC4 determines the trajectory of a particle and whether or not it will hit the halo paddle.
-In figure \ref{fig:PxVsXTrue} , we plot the true  $X$ component of the momentum versus the true $X$ position at WC4 for pions missing the halo paddle (left) and for pions hitting the halo paddle (right) for the 60A MC simulation runs -- analogous plots are obtained with the 100A simulation.
+In figure \ref{fig:PxVsXTrue} , we plot the true  horizontal component of the momentum $P_x$ versus the true $X$ position at WC4 for pions missing the halo paddle (left) and for pions hitting the halo paddle (right) for the 60A MC simulation runs -- analogous plots are obtained with the 100A simulation.
 These distributions can be separated drawing a line in this position-momentum space.
@@
 We classify as ``hitting the halo paddle" all pions whose $P_x$ and $X$ are such that $$P_x +0.02* X - 0.4 < 0 $$ and as ``missing the halo  paddle" all pions whose $P_x$ and $X$ are such that $$P_x +0.02*X - 0.4 > 0, $$ where the coefficients of the line are empirically found by the logistic regression estimation.
-Overall, this simple classifier classifies in the right category (hit or miss) about 86\% of the pion events.
+Overall, this simple methode classifies in the right category (hit or miss) about 86\% of the pion events.
 We apply the same classifier on data.
-We assign  $E_{loss} = 32 \pm 4 $~MeV for events classified as ``hitting the halo paddle"; we assign  $E_{loss} = 24 \pm 3 $~MeV for events classified as ``missing the halo paddle".
+We assign  $E_{loss} = 32 \pm 4 $~MeV for pion events classified as ``hitting the halo paddle"; we assign  $E_{loss} = 24 \pm 3 $~MeV for pion events classified as ``missing the halo paddle".
 
@@
 \includegraphics[width=0.45\textwidth]{Chapter-5/Images/E_loss100A.png}
-\caption{True energy loss between WC4 and the TPC front face according to the MC simulation of the 60A runs (left) and of the 100A runs (right).
+\caption{True energy loss between WC4 and the TPC front face according to the MC simulation of negative pions of the 60A runs (left) and of the 100A runs (right).
 The distribution for the whole data sample is shown in blue, the distribution for the pions missing the halo is shown in red, and the distribution for the pions hitting the halo is shown in green.

### 2018-05-14 Chapter-5/5-Preparatory.tex
s

@@
 Overall, this simple methode classifies in the right category (hit or miss) about 86\% of the pion events.
+In MC, we assign  $E_{loss} = 32 \pm 4 $~MeV for pion events classified as ``hitting the halo paddle"; we assign  $E_{loss} = 24 \pm 3 $~MeV for pion events classified as ``missing the halo paddle".
 We apply the same classifier on data.
-We assign  $E_{loss} = 32 \pm 4 $~MeV for pion events classified as ``hitting the halo paddle"; we assign  $E_{loss} = 24 \pm 3 $~MeV for pion events classified as ``missing the halo paddle".
+A survey of the simulated geometry showed an excess of 3 cm of un-instrumented argon compared with the measured geometry used in data.
+Thus, we account for this difference by assigning in data $E_{loss} = 24 \pm 6 $~MeV for pion events classified as ``hitting the halo paddle" and  $E_{loss} = 17 \pm 6 $~MeV for pion events classified as ``missing the halo paddle", where the uncertainty is derived as the standard deviation of the double peaked distribution.
+
+The analogous study on the kaon MC sample resulted in an energy loss of $E_{loss} = 37 \pm 5 $~MeV for kaon events classified as ``hitting the halo paddle" and $E_{loss} = 31 \pm 4 $~MeV for kaon events classified as ``missing the halo paddle".
+
+s\begin{table}[b]
+\centering
+\begin{tabular}{|l|c|c|}  
+\hline
+                          & Hitting Halo          & Missing Halo     \\ \hline
+Pion  MC           &  $32 \pm 4 $         &    $24 \pm 3$     \\ \hline
+Pion Data          &  $25 \pm 6$          &    $17 \pm 6 $    \\ \hline
+Kaon  MC          &  $37 \pm 5 $        &     $31 \pm 4 $    \\ \hline
+Kaon Data         &  $26 \pm 6 $        &     $22 \pm 6 $    \\ \hline
+\end{tabular}
+\caption{Energy loss for pions and kaons.}
+\label{tab:Eloss}
+\end{table}
 

### 2018-05-14 Chapter-5/5-Preparatory.tex
s

--- 2018-05-14 Chapter-5/5-Preparatory.tex
@@
 Overall, this simple methode classifies in the right category (hit or miss) about 86\% of the pion events.
+In MC, we assign  $E_{loss} = 32 \pm 4 $~MeV for pion events classified as ``hitting the halo paddle"; we assign  $E_{loss} = 24 \pm 3 $~MeV for pion events classified as ``missing the halo paddle".
 We apply the same classifier on data.
-We assign  $E_{loss} = 32 \pm 4 $~MeV for pion events classified as ``hitting the halo paddle"; we assign  $E_{loss} = 24 \pm 3 $~MeV for pion events classified as ``missing the halo paddle".
+A scan of the simulated geometry showed an excess of 3 cm of un-instrumented argon compared with the surveyed detector geometry.
+We account for this difference by assigning in data $E_{loss} = 24 \pm 6 $~MeV for pion events classified as ``hitting the halo paddle" and  $E_{loss} = 17 \pm 6 $~MeV for pion events classified as ``missing the halo paddle", where the uncertainty is derived as the standard deviation of the double peaked distribution.
+
+The summary of the values for used for $E_{Loss}$ for the pion sample is listed in table \ref{tab:Eloss}  with the analogous results for the study on the kaon case.
+
+\begin{table}[b]
+\centering
+\begin{tabular}{|l|c|c|}  
+\hline
+                          &  \multicolumn{2}{c|}{E$_{loss}$ [MeV]}    \\ \hline
+                          & Hitting Halo          & Missing Halo     \\ \hline
+Pion  MC           &  $32 \pm 4 $         &    $24 \pm 3$     \\ \hline
+Pion Data          &  $25 \pm 6$          &    $17 \pm 6 $    \\ \hline
+Kaon  MC          &  $37 \pm 5 $        &     $31 \pm 4 $    \\ \hline
+Kaon Data         &  $26 \pm 6 $        &     $22 \pm 6 $    \\ \hline
+\end{tabular}
+\caption{Energy loss for pions and kaons.}
+\label{tab:Eloss}
+\end{table}
 
@@
 \section{Tracking Studies}\label{sec:TrackingStudies}
+In this section, we describe three studies.
+The first is a justification of the selection criteria for the beamline handshake with the TPC information.
+We perform this study to boost  the correct identification of the particles in the TPC associated with the beamline information, while maintaining sufficient statistics for the cross section measurement.
+The second study is an optimization of the tracking algorithm, with the scope of maximizing the identification of the hadronic interaction point inside the TPC.
+These two studies are related, since the optimization of the tracking is performed on TPC tracks which have been matched to the wire chamber track; in turn, the tracking algorithm for TPC tracks determines the number of reconstructed tracks in each event used to try the matching with the wire chamber track.
+Starting with a sensible tracking reconstruction, we perform the WC2TPC matching optimization first, then the tracking optimization.
+The WC2TPC match purity and efficiency  are then calculated again with the optimized tracking.
+The third study is an evaluation of  the angular resolution of the tracking algorithm in data and MC, which is particularly important in the context of the cross section analyses.
+
 \subsection{Study of WC to TPC Match}\label{ch:WC2TPCMatchOptimization}
+\subsection{Tracking Optimization}\label{ch:TrackOptimization}
+\subsection{Study of WC to TPC Match}\label{ch:WC2TPCMatchOptimization}
+
+\subsection{Angular Resolution}\label{sec:angleRes}
+Scope of this study is to understand and compare the tracking performances and angular resolution of the TPC tracking on data and MC. 
+We use the angular resolution of the tracking to determine  the value of smallest angle that we can reconstruct with a non-zero efficiency, effectively determining a selection on the angular distribution of the cross section measurement due to the tracking performance.
+
+We start by selecting all the WC2TPC matched tracks used for the cross section analysis.
+These tracks can contain from a minimum of 3 3D-space points to a maximum of 240  3D-space points.
+We fit a line to all the 3D-space points associated with the track.
+For each track we calculate the average distance between each point in space and the fit line as follows 
+\begin{equation} 
+\bar d = \frac{\sum^N_i d_i}{N},
+\end{equation} 
+where $N$ is the number of 3D-space points of the track and $d_i$ is the distance of the $i$-th space point to the line fit.
+Several tests to compare the goodness of fit between data and MC have been considered.
+We decided to use $\bar d$ for its straightforward interpretation.
+The $\bar d$ distribution for data and MC is shown in Figure \ref{fig:Chi2AllPts} and shows a relatively good agreement between data and MC.
+
+A visual representation of the procedure used to evaluate the angular resolution is shown in Figure \ref{fig:AngResProcedure}. 
+For each track, we order the space points according to their Z position along the positive beam direction (panel a) and we split them in two sets: the first set contains all the points belonging to the first half of the track and the second set contains all the points belonging the second half of the track.
+We remove the last four points in the first set and the first four points in the second set, so to have a gap in the middle of the original track (panel b).
+We fit the first and the second set of points with two lines  (panel c).
+We then calculate the angle between the fit of the first and second half $\alpha$ (panel d).
+The angle $\alpha$ determines the spatial resolution of the tracking.
+The distributions for data and MC for $\alpha$ are given in Figure \ref{fig:trackingResolution}.
+The mean of the data and MC angular resolution are respectively
+
+\begin{equation}
+\bar\alpha_{Data} = (5.0 \pm 4.5) \text{ deg} 
+\end{equation}
+
+\begin{equation}
+\bar\alpha_{MC} = (4.5 \pm 3.9) \text{ deg}. 
+\end{equation}
+
+Interaction angles smaller than the angle resolution are indistinguishable for the reconstruction.
+Therefore, we assess our ability to measure the cross section to be limited to interaction angles greater than 5.0 deg.
+More accurate studies of the angular resolution as a function of the kinetic energy and track length, albeit interesting, are left for an improvement of the analysis.
+
+It is beneficial to take a moment to describe the definition of interaction angle.
+In case of elastic scattering, the definition is straightforward: the interaction angle is the angle between the incoming and outgoing pion, i.e.
+
+\begin{equation}  
+\theta = \cos^{-1} \Big(\frac{\vec p _{\text{incoming}}  \cdot\vec p _{\text{outgoing}}}{|\vec p _{\text{incoming}}|  |\vec p _{\text{outgoing}}| }\Big).
+\end{equation}   
+In case of inelastic scattering,  the presence of several topologies requires a more complex definition, as shown in figure \ref{fig:scatterPic}.
+We define the scattering angle as the biggest of the angles between the incoming pion and the visible daughters, where the visible daughters are charged particles that travel more than 0.47 cm in the detector (see panel a); in case all the daughters are invisible, the angle is assigned to be 90 deg (see panel b).
+We chose this working definition of scattering angle for inelastic scattering keeping in mind how our tracking reconstruction works: the tracking will stop correctly in case of all the daughters are not visible in the detector and it is likely to stop correctly if multiple daughters form an interaction vertex.
+The only ``dangerous" case is the production of one charged daughter plus neutrals, which we can study with this working definition of scattering angle (see panel c).
+
+We can see the effects of the angular resolution on the cross section by plotting the true Geant4 cross section for interaction angles greater than a minimum interaction angle.
+Figure \ref{fig:trueWithAngles} shows the  true Geant4 cross section  for interaction angles greater than 0 deg (green), 4.5 deg (red), 5.0 deg  (blue) and 9.0 deg (yellow).
+A small $0.5 \text{ deg}$ systematic shift between the mean of the data and MC angular resolution is present.%, which we account for in the context of the MC efficiency correction to the cross section, as presented in \ref{sec:angSys}.
+
+\begin{figure}[ht]
+\begin{minipage}[t]{0.45\linewidth}
+\centering
+\includegraphics[width=\textwidth]{Chapter-5/Images/cDAvg.png}
+\caption[]{Distributions of the average distance between each 3D point in space and the fit line,  $\bar d$ for the data used in the pion cross section analysis and the pion only DDMC.
+The distributions are area normalized.
+} \label{fig:Chi2AllPts}
+\end{minipage}
+\hspace{0.5cm}
+\begin{minipage}[t]{0.45\linewidth}
+\centering
+\includegraphics[width=\textwidth]{Chapter-5/Images/cTrackingDeg.png}
+\caption[]{Distributions of angular resolution $\alpha$ for data used in the pion cross section analysis and pion only DDMC. The distributions are area normalized. } \label{fig:trackingResolution}
+\end{minipage}
+\end{figure}
+
+\begin{figure}[ht]
+\begin{minipage}[t]{0.45\linewidth}
+\centering
+\includegraphics[width=\textwidth]{Chapter-5/Images/TrackingProcedure.png}
+\caption{A visual representation of the procedure used to evaluate the angular resolution.}
+\label{fig:AngResProcedure}
+\end{minipage}
+\hspace{0.5cm}
+\begin{minipage}[t]{0.45\linewidth}
+\centering
+\includegraphics[width=\textwidth]{Chapter-5/Images/Daughters.png}
+\caption{A visual representation of the scattering angle definition in case of inelastic scattering.}
+\label{fig:scatterPic}
+\end{minipage}
+\end{figure}
+
+\begin{figure}[p]
+\centering
+\includegraphics[width=0.48\textwidth]{Chapter-5/Images/cTrueXSAngle.png}
+\caption{ True ($\pi^-{-}, Ar$) cross section for interaction angles greater than 0 deg (green), 4.5 deg (red), 5.0 deg  (blue) and 9.0 deg (yellow). }
+\label{fig:trueWithAngles}
+\end{figure}
 
@@
 \section{Tracking Studies}
-In this section, we describe three studies.
-The first is a justification of the selection criteria for the beamline handshake with the TPC information.
-We perform this study to boost  the correct identification of the particles in the TPC associated with the beamline information, while maintaining sufficient statistics for the cross section measurement.
-The second study is an optimization of the tracking algorithm, with the scope of maximizing the identification of the hadronic interaction point inside the TPC.
-These two studies are related, since the optimization of the tracking is performed on TPC tracks which have been matched to the wire chamber track; in turn, the tracking algorithm for TPC tracks determine the number of reconstructed tracks in each event used to try the matching with the wire chamber track.
-Starting with a sensible tracking reconstruction, we perform the WC2TPC matching optimization first, then the tracking optimization.
-The WC2TPC match purity and efficiency  are then calculated again with the optimized tracking.
 
@@
 We then apply the same WC to TPC matching algorithm as in data described in \ref{ch:WC2TPCMatchMethod}.
-
-%In data, we attempt to uniquely match one WC-Track to one and only one reconstructed TPC track.
-This match is done by using in the $X$ and $Y$ coordinate of the extrapolated WC-Track to the upstream most point of the reconstructed TPC Track and by using the angle between the incoming track angle and the reconstructed TPC.
-We define $\Delta$X as the difference between the $x$ position of the most upstream point of the TPC track and the $x$ position of the WC track as projected to the TPC front face.
-$\Delta$Y is defined analogously.
-We define  $\Delta$R as $ \Delta \text{R} =  \sqrt{ \Delta \text{X}^2 +  \Delta \text{Y}^2}  $.
-The angle between the incident WC Track and the TPC track in the plane that contains them defines $\alpha$.
-
-%We define a match between WC-track and TPC reconstructed track if  $\Delta \text{R} < r_{T}$, $\alpha < \alpha_{T}$ and the Z position of the first reconstructed point of the TPC track is within 2 cm from the TPC front face.
-The determination of the best $r_{T}$ and $\alpha_{T}$ is the scope of the following section.
-
-%In MC, we mimic the matching between the WC and the TPC track on Monte Carlo by constructing a fake WC track using truth information at wire chamber four.
-We then apply the same WC to TPC matching algorithm as in data.
 

### 2018-05-18 Chapter-2/2-LArTPCs.tex
s

@@
 \chapter{Liquid Argon Detectors at the Intensity Frontier}\label{ch:2}
+%{\raggedleft ``\emph{Beat-up little seagull, on a marble stair} \par}
+%{\raggedleft \emph{Tryin' to find the ocean, lookin' everywhere.}"\par}
+%{\raggedleft -- Nina Simone, Baltimore, 1978 -- \par}
+%\vspace{0.5cm}
+
 In the next few years, LArTPCs will  be the tools to answer some of the burning questions in neutrino physics today.
@@
 Surface LArTPC of the size of several meters are expected to be modestly impacted from the space charge effect, where charge build-up create anisotropy of the electric field magnitude of the order of  5\% at a drift field of 500 V/cm \cite{SpaceCharge}.
-The smallness of the LArIAT drift volume is such that  effect of  space charge on the electric field is expected to be negligible.
-\textcolor{red}{CHIEDI A FLAVIO}
+The smallness of the LArIAT drift volume and its relatively high electric field are such that the effect of  space charge is expected to be negligible.
 

### 2018-05-18 Chapter-4/4-Methodology.tex
s

@@
   \centering  
-\includegraphics[width=\textwidth]{Chapter-4/Images/Shower.png}
+\includegraphics[width=\textwidth]{Chapter-4/Images/WC2TPCMatchTracks.png}
 \caption{Visual rendering of the wire chamber to TPC match.}
@@
 
-\subsection{Not-so-Thin Target: Slicing the Argon}
+\subsection{Not-so-Thin Target: Slicing the Argon}\label{ch:XSRaw}
 The interaction length of pions and kaons in argon is expected to be of the order of 50 cm for pions and 100 cm for kaons.
@@
 The cross section as a function of kinetic energy, $\sigma_{TOT}( E^{kin})$ will then be proportional to the ratio $\frac{N_{Interacting}( E^{kin})}{N_{Incident}( E^{kin})}$ .
-
-The statistical uncertainty for each energy bin is calculated by error propagation from the statistical  uncertainty on $N_{Incident}$ and $N_{Interacting}$. 
-Since the number of incident hadrons in each energy bin is given by a simple counting, we assume that $N_{Incident}$ is distributed as a poissonian with mean and $\sigma^2$ equal to $N_{Incident}$ in each bin.
-On the other hand, $N_{Interacting}$ follows a binomial distribution: a particle in a given energy bin might or might not interact.  The square of the variance for the binomial is given by  
-\begin{equation}
-\sigma^2 = \mathcal{N}P_{Interacting}(1-P_{Interacting});
-\label{eq:binVar}
-\end{equation}
-
-since the interaction probability $P_{Interacting}$ is $\frac{ N_{Interacting}}{N_{Incident}}$ and the number of tries $\mathcal{N}$ is $N_{Incident}$, equation \ref{eq:binVar} translates into
-\begin{equation}
-\sigma^2 = N_{Incident}\frac{ N_{Interacting}}{N_{Incident}} (1-\frac{ N_{Interacting}}{N_{Incident}}) = N_{Interacting}(1-\frac{ N_{Interacting}}{N_{Incident}}).
-\end{equation}
-
-$N_{Incident}$ and $N_{Interacting}$ are not independent.
-The uncertainty on the cross section is thus calculated as 
-\begin{equation}
-\delta\sigma_{tot}(E) = \sigma_{tot}(E) \Big(\frac{\delta N_{Interacting}}{N_{Interacting}}+\frac{\delta N_{Incident}}{N_{Incident}}\Big) 
-\end{equation}
-where:
-\begin{eqnarray}
-\delta N_{Incident} = \sqrt[]{N_{Incident}} \\
-\delta N_{Interacting} = \sqrt[]{N_{Interacting}\Big(1-\frac{ N_{Interacting}}{N_{Incident}}\Big)}.
-\end{eqnarray}
 

### 2018-05-18 Chapter-5/5-Preparatory.tex
s

@@
 The WC2TPC match purity and efficiency  are then calculated again with the optimized tracking.
+
 The third study is an evaluation of  the angular resolution of the tracking algorithm in data and MC, which is particularly important in the context of the cross section analyses.
@@
 \subsection{Study of WC to TPC Match}\label{ch:WC2TPCMatchOptimization}
+
+Plots I want in this section:
+\begin{enumerate}
+\item WC2TPC MC DeltaX, DeltaY and $\alpha$
+\end{enumerate}
+
+Scope of this study is assessing the goodness of the wire chamber to TPC match on Monte Carlo and decide the selection values we will use on data.
+A word of caution is necessary here.
+With this study, we want to minimize pathologies associated with the presence of the primary hadron itself, e.g.
+the incorrect association between the beamline hadron and its decay products inside the TPC.
+Assessing the contamination from pile-up\footnote{We remind the reader that the DDMC is a single particle Monte Carlo, where the beam pile up is not simulated.}, albeit related, is beyond the scope of this study.
+
+In MC, we are able to define a correct WC2TPC match using the Geant4 truth information. We are thus able to count how many times the WC tracks is associated with the wrong TPC reconstructed track.
+
+We define a correct match if the all following conditions are met:
+\begin{itemize}
+\item[-] the length of the true primary Geant4 track in the TPC is greater than 2 cm,  
+\item[-] the length of the reconstructed track length is greater than 2 cm,
+\item[-] the Z position of the first reconstructed point is within 2 cm from the TPC front face
+\item[-] the distance between the reconstructed track and the true entering point is the minimum compared with all the other reconstructed tracks.
+\end{itemize}
+
+In order to count the wrong matches, we consider all the reconstructed tracks whose Z position of the first reconstructed point lies within 2 cm from the TPC front face.
+Events with true length in TPC $<$ 2 cm are included.
+Since hadrons are shot 100 cm upstream from the TPC front face, the following two scenarios are possible from a truth standpoint: 
+\begin{itemize}
+\item[[$Ta$]] the primary hadron decays or interact strongly before getting to the TPC,
+\item[[$Tb$]] the primary hadron enters the TPC.
+\end{itemize}
+
+As described in Section \ref{ch:WC2TPCMatchMethod}, we define a WC2TPC match according to the relative position of the WC and TPC track parametrized with $\Delta R$ and the angle between them, parametrized with $\alpha$.
+Once we choose the selection values $r_{T}$ and $\alpha_{T}$ to determine a reconstructed WC2TPC match, the following five scenarios are possible in the truth to reconstruction interplay :
+\begin{itemize}
+\item[1)] only the correct track is matched
+\item[2)] only one wrong track is matched 
+\item[3)] the correct track and one (or more) wrong tracks are matched
+\item[4)] multiple wrong tracks  matched.
+\item[5)] no reconstructed tracks are matched
+\end{itemize}
+
+Since we keep only events with one and only one match, we discard cases 3), 4) and 5) from the events used in the cross section measurement.
+For each set of $r_{T}$ and $\alpha_{T}$ selection value, we define purity and efficiency of the selection as follows:
+\begin{equation}
+\text{Efficiency} = \frac{\text{Number of events correctly matched}}{\text{ Number of events with primary in TPC}},
+\end{equation}
+
+\begin{equation}
+\text{Purity} = \frac{\text{Number of events correctly matched}}{\text{Total number of matched events}}.
+\end{equation}
+
+Figure \ref{fig:EffPurityK} shows the efficiency (left) and purity (right) for WC2TPC match as a function of the radius, $r_{T}$, and angle, $\alpha_{T}$, selection value.
+It is apparent how both efficiency and purity are fairly flat as a function of the radius selection value at a given angle.
+This is not surprising.
+Since we are studying a single particle gun Monte Carlo sample, the wrong matches can occur only for mis-tracking of the primary or for association with decay products;  decay products will tend to be produced at large angles compared to the primary, but could be fairly close to the in $x$ and $y$ projection of the primary.
+The radius cut would play a key role in removing pile up events.
+
+For LArIAT cross section measurements, we generally prefer purity over efficiency, since a sample of particles of a pure species will lead to a better measurement.
+Obviously, purity should be balanced with a sensible efficiency to avoid rejecting the whole sample.
+
+We choose $(\alpha_{T}$, $r_{T}) = (8 \text{ deg}, 4 \text{ cm} )$ and get a MC 85\% efficiency and 98\% purity for the kaon sample and a MC 95\% efficiency and 90\% purity for the pion sample.
+
+\begin{figure}[hpbt]
+\centering
+\includegraphics[width=15cm]{Chapter-5/Images/KEffPurity.png}
+\caption{Efficiency (left) and purity (right) for WC2TPC match as a function of the radius and angle selections for the kaon sample.}
+\label{fig:EffPurityK}
+\end{figure}
+
 \subsection{Tracking Optimization}\label{ch:TrackOptimization}
-\subsection{Study of WC to TPC Match}\label{ch:WC2TPCMatchOptimization}
 
@@
 We use the angular resolution of the tracking to determine  the value of smallest angle that we can reconstruct with a non-zero efficiency, effectively determining a selection on the angular distribution of the cross section measurement due to the tracking performance.
+This study is performed on the pion sample, but its results are extrapolated to the kaon case.
 
@@
 \begin{equation}
-\bar\alpha_{Data} = (5.0 \pm 4.5) \text{ deg} 
+\bar\alpha_{Data} = (5.0 \pm 4.5) \text{ deg}, 
 \end{equation}
@@
 \subsection{Selection Study for the Wire Chamber to TPC Match}\label{ch:WC2TPCMatchOptimization}
-Plots I want in this section:
-\begin{enumerate}
-\item WC2TPC MC DeltaX, DeltaY and $\alpha$
-\end{enumerate}
-
-Scope of this study is assessing the goodness of the wire chamber to TPC match on Monte Carlo and decide the selection values we will use on data.
-A word of caution is necessary here.
-With this study, we want to minimize pathologies associated with the presence of the primary hadron itself, e.g.
-the incorrect association between the beamline hadron and its decay products inside the TPC.
-Assessing the contamination from pile-up\footnote{We remind the reader that the DDMC is a single particle Monte Carlo, where the beam pile up is not simulated.}, albeit related, is beyond the scope of this study.
-
-In MC, we are able to define a correct WC2TPC match using the Geant4 truth information. We are thus able to count how many times the WC tracks is associated with the wrong TPC reconstructed track.
-
-We define a correct match if the all following conditions are met:
-\begin{itemize}
-\item[-] the length of the true primary Geant4 track in the TPC is greater than 2 cm,  
-\item[-] the length of the reconstructed track length is greater than 2 cm,
-\item[-] the Z position of the first reconstructed point is within 2 cm from the TPC front face
-\item[-] the distance between the reconstructed track and the true entering point is the minimum compared with all the other reconstructed tracks.
-\end{itemize}
-
-In order to count the wrong matches, we consider all the reconstructed tracks whose Z position of the first reconstructed point lies within 2 cm from the TPC front face.
-Events with true length in TPC $<$ 2 cm are included.
-Since hadrons are shot 100 cm upstream from the TPC front face, the following two scenarios are possible from a truth standpoint: 
-\begin{itemize}
-\item[[$Ta$]] the primary hadron decays or interact strongly before getting to the TPC,
-\item[[$Tb$]] the primary hadron enters the TPC.
-\end{itemize}
-
-Once we choose the selection cuts to determine a reconstructed wire chamber-to-TPC match $r_{T}$ and $\alpha_{T}$, the following five scenarios are possible in the truth to reconstruction interplay : 
-\begin{itemize}
-\item[1)] only the correct track is matched
-\item[2)] only one wrong track is matched 
-\item[3)] the correct track and one (or more) wrong tracks are matched
-\item[4)] multiple wrong tracks  matched.
-\item[5)] no reconstructed tracks are matched
-\end{itemize}
-
-Since we keep only events with one and only one match, we discard cases 3), 4) and 5) from the events used in the cross section measurement.
-For each set of $r_{T}$ and $\alpha_{T}$ selection value, we define purity and efficiency of the selection as follows:
-\begin{equation}
-\text{Efficiency} = \frac{\text{Number of events correctly matched}}{\text{ Number of events with primary in TPC}}
-\end{equation}
-
-\begin{equation}
-\text{Purity} = \frac{\text{Number of events correctly matched}}{\text{Total number of matched events}}.
-\end{equation}
-
-Figure \ref{fig:EffPurityK} shows the efficiency (left) and purity (right) for wire chamber-to-TPC match as a function of the radius, $r_{T}$, and angle, $\alpha_{T}$, selection value.
-It is apparent how both efficiency and purity are fairly flat as a function of the radius selection value at a given angle.
-This is not surprising.
-Since we are studying a single particle gun Monte Carlo sample, the wrong matches can occur only for mis-tracking of the primary or for association with decay products;  decay products will tend to be produced at large angles compared to the primary, but could be fairly close to the in $x$ and $y$ projection of the primary.
-The radius cut would play a key role in removing pile up events.
-
-For LArIAT cross section measurements, we generally prefer purity over efficiency, since a sample of particles of a pure species will lead to a better measurement.
-Obviously, purity should be balanced with a sensible efficiency to avoid rejecting the whole sample.
-
-We choose $(\alpha_{T}$, $r_{T}) = (8 \text{ deg}, 4 \text{ cm} )$ and get a MC 85\% efficiency and 98\% purity for the kaon sample and a MC \textcolor{red}{BOH}\% efficiency and 98\% purity for the \textcolor{red}{BOH} sample.
-
-\begin{figure}[hpbt]
-\centering
-\includegraphics[width=15cm]{Chapter-5/Images/KEffPurity.png}
-\caption{Efficiency (left) and purity (right) for wire chamber-to-TPC match as a function of the radius and angle selections.}
-\label{fig:EffPurityK}
-\end{figure}
-
 \subsection{Interaction Point Optimization}\label{ch:TrackingOptimization}
@@
 
-\subsection{Tracking spatial and angular resolution}
-Scope of this study is understanding and comparing the tracking spatial and angular resolution on data and MC.
-We start by selecting all the WC2TPC matched tracks. 
-We fit a line on all the space points of the track and calculate the $\chi^2$. The $\chi^2$ distribution for data and MC is shown in Figure \ref{fig:Chi2AllPts}.
-
-For the spatial and angular resolution study, we reject tracks with less than 14 space points.
-For each track, we order the space points according to their Z position and we split them in two sets: the first set counts all the points belonging to the first half of the track and the second set counts all the points belonging to the second half of the track.
-We remove the last 5 points in the first set and the first 5 points in the second set, so to have a gap in the middle of the original track.
-We fit the first and the second set of points with a line separately.
-We reject the event entirely if the  $\chi^2$ for the fit of either of the halves is greater than four.
-We define a track middle plane as the plane perpendicular to the original track fit, positioned in the middle of its length.
-We project the tracks on the middle plane and calculate the impact parameter, $d$, i.e.
-the distance between the projected points.
-We also calculate the angle between the original track direction and the fit of the first and second half, called $\alpha_1$ and $\alpha_2$ respectively.
-The spatial resolution of the track will be $\sigma_S = \frac{d}{\sqrt 2}$ while the angular resolution of the tracks will be  $\sigma_\alpha = \alpha_1 - \alpha_2$.
-The distributions for data and MC for $\sigma_\alpha$ and $\sigma_S$ are given in \ref{fig:trackingResolution}.
-
 \end{comment}

### 2018-05-18 Chapter-6/6-PionCrossSection.tex
s

@@
 \chapter{Negative Pion Cross Section Measurement}\label{ch:PionXS}
+%{\raggedleft ``\emph{Y ella es flama que se eleva, Y es un p\'ajaro a volar} \par}
+%{\raggedleft \emph{En la noche que se incendia, Estrella de oscuridad}"\par}
+%{\raggedleft \emph{Que busca entre la tiniebla, La dulce hoguera del beso}"\par}
+%{\raggedleft -- Lila Downs, Benediction And Dream,  -- \par}
+
 \section{Raw Cross Section}\label{ch:PionXSRaw}
+We measure the ($\pi^-$-Ar) cross section as a function of the kinetic energy in the two chosen data sets, the 60A and 100A negative runs. 
+As will be clarified in \ref{ch:PionXSCorrections},  the corrections to the raw cross section depend on the beam conditions and need to proceed independently for the two data sets.
+Thus, we present here the two measurements separately.
 
-\section{Background Subtracted Cross Section}\label{ch:PionXSBkgSub}
+As stated in section \ref{ch:XSRaw},  the raw cross section is given by the equation \label{eq:thinTargetXSSolved}
+\begin{equation}
+ \sigma_{TOT} (E_i)  = \frac{1}{n \delta X}\frac{N_{Interacting}(E_i)}{N_{Incident}(E_i)}.
+\end{equation}
 
-\section{Efficiency Corrected Cross Section}\label{ch:PionXSEffCorr}
+where $N_{Interacting}$  is the number of particles interacting in an argon slice at kinetic energy $E_i$, $N_{Incident}$ is number of particles incident  on the argon slice at   kinetic energy $E_i$,  $n$ is the density of the target centers  and $\delta X$ is the thickness of the argon slice.
+
+Figure \ref{fig:InteractingRaw} shows the interacting histogram for the 60A dataset on the left and for the 100A dataset on the right. 
+Figure \ref{fig:IncidentRaw} shows the incident histogram for the 60A dataset on the left and for the 100A dataset on the right. 
+Figure \ref{fig:XSRaw} shows the raw cross section for the 60A dataset on the left and for the 100A dataset on the right.
+On all plots the same color scheme is used: the statistical uncertainty is shown in azure, while the systematic uncertainty is shown in blue.
+The calculation of the statistical unceratinty is laid out in section \ref{ch:StatUncertaintyXSRaw}, while the systematics on this \ref{ch:SysUncertaintyXSRaw}.
+
+\begin{figure}[p]
+\centering  
+\includegraphics[width=0.48\textwidth]{Chapter-4/Images/face-smile-big.png}
+\includegraphics[width=0.48\textwidth]{Chapter-4/Images/face-smile-big.png}
+\caption{Raw number of interacting pion candidates as a function of the reconstructed kinetic energy for the 60A runs (lef) and for the 100A runs (right).
+The statistical uncertainties are shown in azure, the systematic uncertainties in blue.}
+\label{fig:InteractingRaw}
+\end{figure}
+
+\begin{figure}
+\centering  
+\includegraphics[width=0.48\textwidth]{Chapter-4/Images/face-smile-big.png}
+\includegraphics[width=0.48\textwidth]{Chapter-4/Images/face-smile-big.png}
+\caption{Raw number of incident pion candidates as a function of the reconstructed kinetic energy for the 60A runs (lef) and for the 100A runs (right).
+The statistical uncertainties are shown in azure, the systematic uncertainties in blue.}
+\label{fig:IncidentRaw}
+\end{figure}
+
+\begin{figure}
+\centering  
+\includegraphics[width=0.48\textwidth]{Chapter-4/Images/face-smile-big.png}
+\includegraphics[width=0.48\textwidth]{Chapter-4/Images/face-smile-big.png}
+\caption{Raw ($\pi^-$-Ar) total hadronic cross section for the 60A runs (lef) and for the 100A runs (right). The statistical uncertainties are shown in azure, the systematic uncertainties in blue.}
+\label{fig:XSRaw}
+\end{figure}
+
+\subsection{Statistical Uncertainty}\label{ch:StatUncertaintyXSRaw}
+The statistical uncertainty for each kinetic energy bin of the cross section plot is calculated by error propagation from the statistical uncertainty on $N_{Incident}$ and $N_{Interacting}$ correspondent bin.
+Since the number of incident hadrons in each energy bin is given by a simple counting, we assume that $N_{Incident}$ is distributed as a poissonian with mean and $\sigma^2$ equal to $N_{Incident}$ in each bin.
+On the other hand, $N_{Interacting}$ follows a binomial distribution: a particle in a given energy bin might or might not interact.  The square of the variance for the binomial is given by  
+\begin{equation}
+\sigma^2 = \mathcal{N}P_{Interacting}(1-P_{Interacting});
+\label{eq:binVar}
+\end{equation}
+
+since the interaction probability $P_{Interacting}$ is $\frac{ N_{Interacting}}{N_{Incident}}$ and the number of tries $\mathcal{N}$ is $N_{Incident}$, equation \ref{eq:binVar} translates into
+\begin{equation}
+\sigma^2 = N_{Incident}\frac{ N_{Interacting}}{N_{Incident}} (1-\frac{ N_{Interacting}}{N_{Incident}}) = N_{Interacting}(1-\frac{ N_{Interacting}}{N_{Incident}}).
+\end{equation}
+
+$N_{Incident}$ and $N_{Interacting}$ are not independent.
+The uncertainty on the cross section is thus calculated as 
+\begin{equation}
+\delta\sigma_{tot}(E) = \sigma_{tot}(E) \Big(\frac{\delta N_{Interacting}}{N_{Interacting}}+\frac{\delta N_{Incident}}{N_{Incident}}\Big) 
+\end{equation}
+where:
+\begin{eqnarray}
+\delta N_{Incident} = \sqrt[]{N_{Incident}} \\
+\delta N_{Interacting} = \sqrt[]{N_{Interacting}\Big(1-\frac{ N_{Interacting}}{N_{Incident}}\Big)}.
+\end{eqnarray}
+
+\subsection{Treatment of Systematics} \label{ch:SysUncertaintyXSRaw}
+
+\section{Corrections to the Raw Cross Section}\label{ch:PionXSCorrections}
+\subsection{Treatment of Systematics}

### 2018-05-18 Chapter-5/5-Preparatory.tex
s

@@
 
+\clearpage
+
 \section{Energy Calibration and Studies}\label{ch:energyCal} 
+\subsection{Energy Calibration}\label{ch:energyCalibration}
+\subsection{Uncertainty on Kinetic Energy}\label{ch:kinEn}
+The measured kinetic energy of a hadron candidate at each argon slab determines which bins of the interacting and incident histograms a selected event is going to fill.
+With this study, we determine the uncertainty of the kinetic energy measurement which we will propagate into the cross section measurement, as discussed in Section \ref{ch:SysUncertaintyXSRaw} for the pion cross section and in Section \ref{ch:SysUncertaintyXSRawKaon} for the kaon cross section.
+
+The kinetic energy of a hadron at the $j^{\text{th}}$ slice of argon in the TPC is given by
+
+\begin{equation}
+KE_{j} = \sqrt{p_{Beam}^2 + m_{Beam}^2} - m_{Beam}^2 - E_{Loss} - E_{\text{FF-j}},
+\end{equation}
+
+where $p_{Beam}$ is the momentum measured by the beamline detectors,  $m_{Beam}$ is the mass of the hadron as reported in the PDG,  $E_{Loss}$  is the energy loss between the beamline and the TPC, and $ E_{\text{FF-j}}$ is the energy that the hadron deposited from the TPC front face until the $j^{\text{th}}$ slice.
+The uncertainty on $KE_{j}$ is then given by 
+\begin{equation}
+\delta KE_{j} = \sqrt{\delta p_{Beam}^2 + \delta E_{Loss}^2 +  \delta  E_{\text{dep FF-j}}^2},
+\end{equation}
+
+where we have dropped the uncertainty on the mass, since it is orders of magnitude smaller than the other uncertainties.
+We assume the relative uncertainty on $p_{Beam}$ to be 2\%, and the uncertainty on the energy loss upstream to be 7~MeV, as calculated in Section \ref{ch:eloss}.
+We describe the estimate of the uncertainty on $E_{\text{FF-j}}$ in the rest of this section.
+
+The energy deposited from the TPC front face until the $j^{\text{th}}$ slice is the sum of the measured energy deposited in each previous slabs $E_{i}$, i.e.
+\begin{equation}
+E_{\text{FF-j}} = \sum_{i<j} E_{i}, 
+\end{equation}
+where $E_{i}$ is measured in each slab as  the product of the stopping power,  $dE/dX_{i}$,  and the track pitch, $Pitch_i$, for that point. 
+If we assume conservatively that the measurements of $E_{i}$ are not independent from one another, the uncertainty on $E_{\text{FF-j}}$ becomes
+\begin{equation}
+\delta E_{\text{FF-j}} = (j-1) \delta E_{i}, 
+\end{equation}
+where $\delta E_{i}$ is the uncertainty on the energy loss in one slab of argon.
+
+Figure \ref{fig:EnergyDeposited} shows the distribution of the energy deposited in each slab of argon, for the 60A negative pion dataset in black and for the pion only MC in blue.
+The distributions are fitted with a landau displayed in red for data and in teal for MC.
+The uncertainty on $E_{i}$ is given by the width of the Landau fit to the data.
+A small systematic uncertainty  is given by a 1.0\% difference between the most probable value of the landau fits in data and MC.
+
+\begin{figure}[htb]
+\centering
+\includegraphics[width=0.7\textwidth]{Chapter-5/Images/DepEnergy_Fit_v4.png}
+\caption[]{ Energy Deposited in Pion MC and 60A data.  } \label{fig:EnergyDeposited}
+\end{figure}
+
+So the uncertainty on the incident kinetic energy is given by
+
+\begin{equation}
+\delta KE^{Incident} = \sqrt{ (\delta KE_{Initial})^2 + (\delta E^{Slab}_{j})^2} = \sqrt{(12 MeV)^2 + (2 MeV)^2} = 12.1 MeV
+\end{equation}
+
+Figure \ref{fig:EnergyDepositedStacked} shows the stacked version of the Energy Deposited plots with the backgrounds stacked.
+The backgrounds are given in the ratio of 68.8\% pion, 4.6\% muon, and 26.6\% electron.
+Once they are taken in these ratios, the sum of the MC is normalized to the sum of the data.
+
+\begin{figure}[htb]
+\centering
+\includegraphics[width=0.48\textwidth]{Chapter-5/Images/DepEnergy_Stacked_v1.png}
+\includegraphics[width=0.48\textwidth]{Chapter-5/Images/DepEnergy_Stacked_v3.png}
+\caption[]{ Energy Deposited with all the MC and 60A data.  } \label{fig:EnergyDepositedStacked}
+\end{figure}
+
+The energy at the interacting point is given by
+\begin{equation}
+KE_{Interaction} = \sqrt{P_{WCtrk}^2 + m_{\pi}^2} - E_{Loss} - (\Sigma dE/dX_{i} \times Pitch)
+\end{equation}
+
+and has the exact same uncertainty as the incident kinetic energy plot. Thus these estimates can be applied to getting the uncertainty on the energy of the reconstructed cross-section.
+
+A study we did was to look at the difference between DATA/MC in the dE/dX and energy deposited.
+We basically found there is very little difference between the two and we try to quantify how much the difference is.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{dE/dX}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+Figure \ref{fig:dEdXLinearScale} shows the output of the fit of the Pion MC and the 60 Amp data.
+The MC is normalized to the data and both are fit to a Landau function.
+\footnote{The entries at dE/dX = 0 come from an uninitialized variable and can/should be taken out of these plots}
+
+\begin{figure}[htb]
+\centering
+\includegraphics[width=0.48\textwidth]{Chapter-5/Images/dEdX_Fit_v1.png}
+\includegraphics[width=0.48\textwidth]{Chapter-5/Images/dEdX_Fit_v4.png}
+\caption[]{ dE/dX for 60Amp data and data driven pion MC, both fit with a Landau  } \label{fig:dEdXLinearScale}
+\end{figure}
+
+The difference between the two MPV's, is 2.4\% between the data and the MC.
+
+Figure \ref{fig:dEdXLinearStacked} shows the stacked version of the dE/dX with the backgrounds stacked.
+The backgrounds are given in the ratio of 68.8\% pion, 4.6\% muon, and 26.6\% electron.
+Once they are taken in these ratios, the sum of the MC is normalized to the sum of the data.
+
+\begin{figure}[htb]
+\centering
+\includegraphics[width=0.48\textwidth]{Chapter-5/Images/dEdX_stacked_v1.png}
+\includegraphics[width=0.48\textwidth]{Chapter-5/Images/dEdX_stacked_v4.png}
+\caption[]{ Stacked versions of the dE/dX with the data and electron/muon/pion MC.  } \label{fig:dEdXLinearStacked}
+\end{figure}
+
+For completeness, the log scale versions of are shown in Figure \ref{fig:dEdXLogScale}.
+
+\begin{figure}[htb]
+\centering
+\includegraphics[width=0.48\textwidth]{Chapter-5/Images/dEdX_Fit_v2.png}
+\includegraphics[width=0.48\textwidth]{Chapter-5/Images/dEdX_Fit_v3.png}
+\includegraphics[width=0.48\textwidth]{Chapter-5/Images/dEdX_stacked_v2.png}
+\includegraphics[width=0.48\textwidth]{Chapter-5/Images/dEdX_stacked_v3.png}
+\caption[]{ dE/dX for 60Amp data and MC shown in log scale  } \label{fig:dEdXLogScale}
+\end{figure}
+
+Plotting scripts can be found here on lariatgpvm \begin{verbatim}
+/lariat/app/users/jasaadi/v06_34_01_PionWeek/PlottingScripts
+\end{verbatim} and the samples were put here \begin{verbatim}
+/lariat/data/users/elenag/theFinalPions/TPCDATA
+
+/lariat/data/users/elenag/theFinalPions/TPC_MC/
+\end{verbatim}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Energy Deposited}\label{sec:Energy}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 

### 2018-05-18 Chapter-6/6-PionCrossSection.tex
s

@@
 %{\raggedleft \emph{Que busca entre la tiniebla, La dulce hoguera del beso}"\par}
-%{\raggedleft -- Lila Downs, Benediction And Dream,  -- \par}
+%{\raggedleft -- Lila Downs, Benediction And Dream,  2002-- \par}
 
@@
 \centering  
-\includegraphics[width=0.48\textwidth]{Chapter-4/Images/face-smile-big.png}
-\includegraphics[width=0.48\textwidth]{Chapter-4/Images/face-smile-big.png}
+\includegraphics[width=0.48\textwidth]{Chapter-6/Images/Plots60A_StatSyst.png}
+\includegraphics[width=0.48\textwidth]{Chapter-6/Images/Plots100A_StatSyst.png}
 \caption{Raw number of interacting pion candidates as a function of the reconstructed kinetic energy for the 60A runs (lef) and for the 100A runs (right).
@@
 \centering  
-\includegraphics[width=0.48\textwidth]{Chapter-4/Images/face-smile-big.png}
-\includegraphics[width=0.48\textwidth]{Chapter-4/Images/face-smile-big.png}
+\includegraphics[width=0.48\textwidth]{Chapter-6/Images/Plots60A_StatSysc.png}
+\includegraphics[width=0.48\textwidth]{Chapter-6/Images/Plots100A_StatSysc.png}
 \caption{Raw number of incident pion candidates as a function of the reconstructed kinetic energy for the 60A runs (lef) and for the 100A runs (right).
@@
 \centering  
-\includegraphics[width=0.48\textwidth]{Chapter-4/Images/face-smile-big.png}
-\includegraphics[width=0.48\textwidth]{Chapter-4/Images/face-smile-big.png}
+\includegraphics[width=0.48\textwidth]{Chapter-6/Images/Plots60A.png}
+\includegraphics[width=0.48\textwidth]{Chapter-6/Images/Plots100A.png}
 \caption{Raw ($\pi^-$-Ar) total hadronic cross section for the 60A runs (lef) and for the 100A runs (right). The statistical uncertainties are shown in azure, the systematic uncertainties in blue.}

### 2018-05-19 Chapter-5/5-Preparatory.tex
s

@@
 This produces MC $P_x, P_y, P_z, X, Y$ distributions  with the same momentum and position distributions as data, with the additional benefit of accounting for the correlations between the considered variables.
-As an example, the results of the DDMC generation compared to data for the kaon +100A sample are shown in figure \ref{DDMCComparison} for the $P_z$, $X$ and $Y$ distributions; as expected, MC and data agree within the statistical uncertainty by construction.
+As an example, the results of the DDMC generation compared to data for the kaon +100A sample are shown in figure \ref{fig:DDMCComparison} for the $P_z$, $X$ and $Y$ distributions; as expected, MC and data agree within the statistical uncertainty by construction.
 A LArSoft simulation module then launches single particle MC from z = -100 cm (the location of the WC4) using the MC generated events.
@@
 \includegraphics[width=0.48\textwidth]{Chapter-5/Images/cTrueXSAngle.png}
-\caption{ True ($\pi^-{-}, Ar$) cross section for interaction angles greater than 0 deg (green), 4.5 deg (red), 5.0 deg  (blue) and 9.0 deg (yellow). }
+\caption{ True ($\pi^-, Ar$) cross section for interaction angles greater than 0 deg (green), 4.5 deg (red), 5.0 deg  (blue) and 9.0 deg (yellow). }
 \label{fig:trueWithAngles}
@@
 \clearpage
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\section{Energy Calibration and Studies}\label{ch:energyCal} 
+\section{Calorimetry Studies}\label{ch:energyCal} 
+The ability to measure the kinetic energy of hadrons in the TPC is fundamental for the cross section analyses. 
+Thus, we describe first how we calibrate the TPC calorimetric response (Section \ref{ch:energyCalibration}) and how we measure the kinetic energy of the hadrons in the TPC (Section \ref{ch:kinEn}).
+
 \subsection{Energy Calibration}\label{ch:energyCalibration}
-\subsection{Uncertainty on Kinetic Energy}\label{ch:kinEn}
+Scope of the energy calibration is to identify the factors which convert the charge collected (dQ) to energy deposited in the chamber (dE).
+As described in section \ref{sec:SignalProc}, this is a multi-step procedure.
+In LArIAT, we first correct the raw charge by the electronic noise on the considered wire \cite{LArIATdqdx}, then by the electron lifetime \cite{LArIATLifeTime},  and then by the recombination using the ArgoNeut recombination values.
+Lastly, we apply overall calibration of the energy, i.e.
+we determine the ``calorimetry constants" using the procedure described in this section.
+
+We independently determine  the calorimetry constants for Data and Monte Carlo in the LArIAT Run-II Data samples using  a parametrization of the stopping power (a.k.a.
+energy deposited per unit length, $dE/dX$)  as a function of momentum.
+This is done by comparing the stopping power measured on reconstructed quantities against the Bethe-Bloch theoretical prediction for various particle species (see Equation \ref{eq:BB}).
+We obtain the theoretical expectation for the $dE/dX$ most probable value of pions ($\pi$), muons ($\mu$), kaons ($K$), and protons ($p$) in the momentum range most relevant for LArIAT (Figure \ref{fig:PDGEnergyLossArgon}) using the tables provided by the Particle Data Group \cite{Patrignani:2016xqp} for liquid argon \cite{PDG-Argon}.
+
+The basic idea of this calibration technique is to utilize a sample of beamline events with known particle species and momentum to measure the $dE/dX$ of the corresponding tracks in the TPC.
+In particular, we decided to use positive pions as calibration sample and all the other particle species as cross check.
+Once the $dE/dX$ of a sample of positive pions  has been measured at various momenta, we tune to calorimetry constants within the reconstruction software to align the measured values to match the theoretical ones found in Figure \ref{fig:PDGEnergyLossArgon}.
+
+In data, we start by selecting a sample of beamline positive pion beamline candidates without any restriction on their measured momentum\footnote{it should be noted that some muon and position contamination is present in the $\pi^+$ sample}.
+We then apply the WC2TPC match and subtract the energy loss upstream to the TPC front face, determining the momentum at the TPC front face.
+For each surviving pion candidate,  we measure the $dE/dx$ at each of the first 12 spacepoints associated the 3D reconstructed track, corresponding to a $\sim$ 5 cm portion.
+These $dE/dX$ measurements are then put into a histogram that corresponds to measured momentum of the track.
+The $dE/dX$ histograms are sampled every 50 MeV/c in momentum (e.g.
+150~MeV/c $< P <$ 200~MeV/c, 200~MeV/c $< P <$ 250/c~MeV, etc...).
+This process of selecting, sampling, and recording the $dE/dX$ for various momentum bins is repeated over the entire sample of events, allowing us to collect sufficient statistic in most of the momentum bins between 150~MeV/c and 1100~MeV/c.
+On average, pions and muons only lose $\sim$10 MeV in this 5~cm section of the track and protons lose $\sim$20 MeV.
+Thus choosing 50 MeV/c size bins for our histograms covers the energy spread within those bins due to energy loss from ionization for all the particle species identifiable in the beamline.
+Each 50 MeV/c momentum binned $dE/dX$ histogram is now fit with a simple Landau function.
+The most probable value (MPV) and the associated error on the MPV from the fit are extracted and plotted against the theoretical prediction Figure \ref{fig:PDGEnergyLossArgon}.
+Depending on the outcome of the fit, we modify the calorimetry constants and we repeat the procedure until a qualitative agreement is achieved.
+We perform this  tuning for the collection and induction plane separately.
+As a cross check to the calorimetry constants determined using the positive pions, we lock the constants and  plot the $dE/dx$ versus momentum distribution of all the other particle species identifiable in the beamline data ($\pi/\mu/e$, K , p, in both polarities) against the corresponding Beth-Bloch prediction.
+The agreement between data from the other particle species and the predictions is the expected result of this cross check.
+The results of the tuning and cross check for Run-II data on the collection plane is shown in Figure \ref{fig:BBandData}  negative polarity data on top, positive polarity data on the bottom.
+
+In MC, we simulate the corresponding positive pion sample with the DDMC (see section \ref{sec:DDMC}) and follow the same steps as in data.
+More details on the calorimetry tuning can be found in \cite{LArIATCalo}.
+
+\textcolor{red}{Add agreement between data and MC for dedx for pions}
+
+\begin{figure}[htb]
+\centering
+\includegraphics[width=0.50\textwidth]{Chapter-5/Images/dEdXvsMomentumTemplate.png}
+\caption{Stopping power for pions, muons, kaons, and protons in liquid argon over the momentum range most relvant for LArIAT according to the Beth-Bloch equation.
+The solid lines represent the prediction for the mean energy $dE/dX$, while the dashed lines are the predictions for the MPV.}
+\label{fig:PDGEnergyLossArgon}
+\end{figure}
+
+\begin{figure}[htb]
+\centering
+\includegraphics[width=0.50\textwidth]{Chapter-5/Images/RunIINegTotaldEdXvsMomentum.png}
+\includegraphics[width=0.50\textwidth]{Chapter-5/Images/RunIIPosTotaldEdXvsMomentum.png}
+\caption{Stopping power versus Momentum for Run-II negative (top) and positive (bottom) polarity data.
+We achieve the agreement between the Bethe-Bloch predictions and the distribution obtained with of the positive pions (top plot, red dots) by tuning the calorimetry constants.
+Once the calorimetry constants are locked in, the agreement between the other particle species and the Bethe-Bloch predictions follows naturally.}
+\label{fig:BBandData}
+\end{figure}
+
+%\begin{figure}[htb]
+%\centering
+%\includegraphics[width=0.50\textwidth]{images/CalibrationExample.png}
+%\caption{Illustration of the calibration technique.
+Here we depict a 325 MeV wire chamber track (shown in green) which enters the TPC (taking into account the energy loss from the upstream material) and we sample the first 12 spacepoints (shown in teal) to extract the dE/dX distribution which is fit with a Landau.}
+%\label{fig:CalibrationExample}
+%\end{figure}
+
+\subsection{Kinetic Energy Measurement}\label{ch:kinEn}
 The measured kinetic energy of a hadron candidate at each argon slab determines which bins of the interacting and incident histograms a selected event is going to fill.
-With this study, we determine the uncertainty of the kinetic energy measurement which we will propagate into the cross section measurement, as discussed in Section \ref{ch:SysUncertaintyXSRaw} for the pion cross section and in Section \ref{ch:SysUncertaintyXSRawKaon} for the kaon cross section.
+In this section, we define the measurement on the kinetic energy and  determine the related uncertainty.
+We will propagate this uncertainty into the cross section measurement, as discussed in Section \ref{ch:SysUncertaintyXSRaw} for the pion cross section and in Section \ref{ch:SysUncertaintyXSRawKaon} for the kaon cross section.
 
@@
 
-The energy deposited from the TPC front face until the $j^{\text{th}}$ slice is the sum of the measured energy deposited in each previous slabs $E_{i}$, i.e.
+The energy deposited by the hadron from the TPC front face until the $j^{\text{th}}$ slice is the sum of the measured energy deposited in each previous slabs $E_{i}$, i.e.
 \begin{equation}
@@
 
-Figure \ref{fig:EnergyDeposited} shows the distribution of the energy deposited in each slab of argon, for the 60A negative pion dataset in black and for the pion only MC in blue.
+The left side of Figure \ref{fig:EnergyDeposited} shows the distribution of the energy deposited in each slab of argon, for the 60A negative pion dataset in black and for the pion only MC in blue.
+The analogous plot for the -100A negative pion data set  is show on the right side of Figure \ref{fig:EnergyDeposited}.
 The distributions are fitted with a landau displayed in red for data and in teal for MC.
@@
 \centering
-\includegraphics[width=0.7\textwidth]{Chapter-5/Images/DepEnergy_Fit_v4.png}
-\caption[]{ Energy Deposited in Pion MC and 60A data.  } \label{fig:EnergyDeposited}
+\includegraphics[width=0.48\textwidth]{Chapter-5/Images/DepEnergy_Fit_v4.png}
+\includegraphics[width=0.48\textwidth]{Chapter-5/Images/DepEnergy_Fit_v4100A.png}
+\caption[]{ Energy deposited  $E_{i}$ in a single slab of argon for the pion -60A runs (left) and -100A runs (right).
+The data is shown in black, the MC in blue.
+The distributions are fitted with a landau displayed in red for data and in teal for MC.
+} \label{fig:EnergyDeposited}
 \end{figure}
 
-So the uncertainty on the incident kinetic energy is given by
-
-\begin{equation}
-\delta KE^{Incident} = \sqrt{ (\delta KE_{Initial})^2 + (\delta E^{Slab}_{j})^2} = \sqrt{(12 MeV)^2 + (2 MeV)^2} = 12.1 MeV
-\end{equation}
+\begin{comment}
 
@@
 and has the exact same uncertainty as the incident kinetic energy plot. Thus these estimates can be applied to getting the uncertainty on the energy of the reconstructed cross-section.
-
-A study we did was to look at the difference between DATA/MC in the dE/dX and energy deposited.
-We basically found there is very little difference between the two and we try to quantify how much the difference is.
 
@@
 
-Plotting scripts can be found here on lariatgpvm \begin{verbatim}
-/lariat/app/users/jasaadi/v06_34_01_PionWeek/PlottingScripts
-\end{verbatim} and the samples were put here \begin{verbatim}
-/lariat/data/users/elenag/theFinalPions/TPCDATA
-
-/lariat/data/users/elenag/theFinalPions/TPC_MC/
-\end{verbatim}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\begin{comment}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%         Energy Calibration         %%%%%%%%%%%%%%         
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Energy Calibration}\label{ch:energyCalibration}
-Scope of the energy calibration is to identify the factors which convert the charge collected (dQ) to energy deposited in the chamber(dE).
-As described in section \ref{sec:SignalProc}, this is a multi-step procedure.
-In LArIAT, we first correct the raw charge by the electronic noise on the considered wire \cite{technote}, then by the electron lifetime \cite{LArIATLifeTime},  and then by the recombination using the ArgoNeut recombination values.
-Lastly, we apply overall calibration of the energy, i.e.
-we determine the ``calorimetry constants" using the procedure described in this section.
-
-We independently determine  the calorimetry constants for Data and Monte Carlo in the LArIAT Run-II Data samples using  a parametrization of the energy deposited per unit length (dE/dX) as a function of momentum.
-This is done by comparing the stopping power measured on reconstructed quantities against the Bethe-Bloch theoretical prediction for various particle species (see equation \ref{eq:BB}).
-We obtain the theoretical expectation for the dE/dX most probable value of pions ($\pi$), muons ($\mu$), kaons ($K$), and protons ($p$) in the momentum range most relevant for LArIAT (Figure \ref{fig:PDGEnergyLossArgon}) using the tables provided by the Particle Data Group \cite{Patrignani:2016xqp} for liquid argon \cite{PDG-Argon}.
-
-The basic idea of this calibration technique is to utilize the most upstream portion of a TPC track which has a well known momentum and particle species to measure its $dE/dX$.
-Once a sample of particles dE/dX has been measured at various momenta, we then tune to calorimetry constants within the reconstruction software to align these measured values to match the theoretical ones found in Figure \ref{fig:PDGEnergyLossArgon}.
-
-In data, we start by selecting a sample of beamline positive pion candidates without any restriction on their measured momentum.
-We then apply the WC2TPC match and subtract the energy loss upstream to the TPC front face, determining the momentum at the TPC front face.
-For each surviving pion candidate,  we measure the dE/dx at each of the first 12 spacepoints associated the 3D reconstructed track, corresponding to a $\sim$ 5 cm portion.
-These dE/dX measurements are then put into a histogram that corresponds to measured momentum of the track.
-The dE/dX histograms are sampled every 50 MeV in momentum (e.g.
-150~MeV/c $< P <$ 200~MeV/c, 200~MeV/c $< P <$ 250/c~MeV, etc...).
-This process of selecting, sampling, and recording the dE/dX for various momentum bins is repeated over the entire sample of events, allowing us to collect sufficient statistic in most of the momentum bins between 150~MeV/c and 1100~MeV/c.
-Each 50 MeV/c momentum binned dE/dX histogram is now fit with a simple Landau function.
-The most probable value (MPV) and the associated error on the MPV from the fit are extracted and plotted on Figure \ref{fig:PDGEnergyLossArgon}.
-Depending on the outcome of the fit, we modify the calorimetry constants and we repeat the procedure until a qualitative agreement is achieved.
-We perform this  tuning for the collection and induction plane separately.
-As a cross check to the determined calorimetry constants using the positive pions, we plot the dE/dx vs momentum distribution all the other particle species identifiable in the beamline data ($\pi$, K , p, in both polarities) against the corresponding Beth-Bloch prediction in a similar fashion.
-On average, pions and muons only lose $\sim$10 MeV in this 5~cm section of the track and protons lose $\sim$20 MeV.
-Thus choosing 50 MeV size bins for our histograms covers the energy spread within those bins due to energy loss from ionization.
-The results of the tuning and cross check for Run-II data on the collection plane is shown in Figure \ref{fig:BBandData} positive polarity data on the left, negative polarity data on the right.
-
-In MC, we simulate the corresponding positive pion sample with the DDMC (see section \ref{sec:DDMC}) and follow the same steps as in data.
-The calorimetry tuning is explained in all his gory details in \cite{technote}.
-
-\begin{figure}[htb]
-\centering
-%\includegraphics[width=0.50\textwidth]{images/PDGdEdX.png}
-\caption{Mean energy loss in various materials over a range of particle momenta as produced in Reference \cite{PDG}.}
-\label{fig:PDGEnergyLoss}
-\end{figure}
-
-\begin{figure}[htb]
-\centering
-%\includegraphics[width=0.50\textwidth]{images/dEdXvsMomentumTemplate}
-\caption{Mean energy loss for pions, muons, and protons in liquid argon over the momentum range most relvant for LArIAT.}
-\label{fig:PDGEnergyLossArgon}
-\end{figure}
-
-\begin{figure}[htb]
-\centering
-%\includegraphics[width=0.50\textwidth]{images/CalibrationExample.png}
-\caption{Illustration of the calibration technique.
-Here we depict a 325 MeV wire chamber track (shown in green) which enters the TPC (taking into account the energy loss from the upstream material) and we sample the first 12 spacepoints (shown in teal) to extract the dE/dX distribution which is fit with a Landau.}
-\label{fig:CalibrationExample}
-\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%         Energy Calibration         %%%%%%%%%%%%%%         
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Tracking Studies}
-
-%\section{MC sample and WC2TPC match}
-We perform the following studies on a MC sample of 191000 kaons and 359000 pions produced with the DDMC technique.
-DDMC particles are shot from the WC4 location into the TPC following the beam profile.
-We mimic the matching between the WC and the TPC track on Monte Carlo by constructing a fake WC track using truth information at wire chamber four.
-We then apply the same WC to TPC matching algorithm as in data described in \ref{ch:WC2TPCMatchMethod}.
-
-\subsection{Selection Study for the Wire Chamber to TPC Match}\label{ch:WC2TPCMatchOptimization}
-\subsection{Interaction Point Optimization}\label{ch:TrackingOptimization}
-Scheme of this subsection
-\subsubsection{Brief Explanation of the reconstruction chain}
-\subsubsection{Explanation of clustering parameters}
-\subsubsection{Figure of merit and  spanning of cluster}
-\subsubsection{Important numbers out of this optimization}
-
-Plots I want in this section:
-\begin{enumerate}
-\item Delta L, reco - true
-\item Delta L, reco - true Elastic, Delta L, reco - true Inelastic, other
-\item Length Quality cut
-\item Efficiency as a function of true KE and Angle
-\end{enumerate}
-
 \end{comment}

### 2018-05-19 Chapter-6/6-PionCrossSection.tex
s

@@
 The statistical uncertainty for each kinetic energy bin of the cross section plot is calculated by error propagation from the statistical uncertainty on $N_{Incident}$ and $N_{Interacting}$ correspondent bin.
-Since the number of incident hadrons in each energy bin is given by a simple counting, we assume that $N_{Incident}$ is distributed as a poissonian with mean and $\sigma^2$ equal to $N_{Incident}$ in each bin.
-On the other hand, $N_{Interacting}$ follows a binomial distribution: a particle in a given energy bin might or might not interact.  The square of the variance for the binomial is given by  
+Since the number of incident hadrons in each energy bin is given by a simple counting, we assume that $N_{Incident}$ is distributed as a poissonian with mean and variance equal to $N_{Incident}$ in each bin.
+On the other hand, $N_{Interacting}$ follows a binomial distribution: a particle in a given energy bin might or might not interact.  The variance for the binomial is given by  
 \begin{equation}
-\sigma^2 = \mathcal{N}P_{Interacting}(1-P_{Interacting});
+\text{\textsf{Var[$N_{Incident}$]}} = \mathcal{N}P_{Interacting}(1-P_{Interacting});
 \label{eq:binVar}
@@
 \begin{equation}
-\sigma^2 = N_{Incident}\frac{ N_{Interacting}}{N_{Incident}} (1-\frac{ N_{Interacting}}{N_{Incident}}) = N_{Interacting}(1-\frac{ N_{Interacting}}{N_{Incident}}).
+\text{\textsf{Var[$N_{Incident}$]}} = N_{Incident}\frac{ N_{Interacting}}{N_{Incident}} (1-\frac{ N_{Interacting}}{N_{Incident}}) = N_{Interacting}(1-\frac{ N_{Interacting}}{N_{Incident}}). 
 \end{equation}
@@
 $N_{Incident}$ and $N_{Interacting}$ are not independent.
-The uncertainty on the cross section is thus calculated as 
+The statistical uncertainty on the cross section is thus calculated as 
 \begin{equation}
@@
 \subsection{Treatment of Systematics} \label{ch:SysUncertaintyXSRaw}
+The only systematic effect considered in the measurement of the raw cross section come from the propagation of the uncertainty associate with the measurement of the kinetic energy at each slab.
 

### 2018-05-19 Chapter-5/5-Preparatory.tex
s

@@
 The basic idea of this calibration technique is to utilize a sample of beamline events with known particle species and momentum to measure the $dE/dX$ of the corresponding tracks in the TPC.
-In particular, we decided to use positive pions as calibration sample and all the other particle species as cross check.
-Once the $dE/dX$ of a sample of positive pions  has been measured at various momenta, we tune to calorimetry constants within the reconstruction software to align the measured values to match the theoretical ones found in Figure \ref{fig:PDGEnergyLossArgon}.
+In particular, we decided to use positive pions as calibration sample and samples from all the other particle species as cross check.
+Once the $dE/dX$ of the positive pion sample  has been measured at various momenta, we tune to calorimetry constants within the reconstruction software to align the measured values to match the theoretical ones found in Figure \ref{fig:PDGEnergyLossArgon}.
 
@@
 The most probable value (MPV) and the associated error on the MPV from the fit are extracted and plotted against the theoretical prediction Figure \ref{fig:PDGEnergyLossArgon}.
-Depending on the outcome of the fit, we modify the calorimetry constants and we repeat the procedure until a qualitative agreement is achieved.
+Depending on the outcome of the data-prediction comparison, we modify the calorimetry constants and we repeat the procedure until a qualitative agreement is achieved.
 We perform this  tuning for the collection and induction plane separately.

### 2018-05-19 Chapter-6/6-PionCrossSection.tex
s

@@
 \centering  
-\includegraphics[width=0.48\textwidth]{Chapter-6/Images/Plots60A_StatSyst.png}
-\includegraphics[width=0.48\textwidth]{Chapter-6/Images/Plots100A_StatSyst.png}
+\includegraphics[width=0.48\textwidth]{Chapter-6/Images/Plots60A_MCData_Int_StatSyst.pdf}
+\includegraphics[width=0.48\textwidth]{Chapter-6/Images/Plots100A_MCData_Int_StatSyst.pdf}
 \caption{Raw number of interacting pion candidates as a function of the reconstructed kinetic energy for the 60A runs (lef) and for the 100A runs (right).
@@
 \centering  
-\includegraphics[width=0.48\textwidth]{Chapter-6/Images/Plots60A_StatSysc.png}
-\includegraphics[width=0.48\textwidth]{Chapter-6/Images/Plots100A_StatSysc.png}
+\includegraphics[width=0.48\textwidth]{Chapter-6/Images/Plots60A_MCData_Inc_StatSyst.pdf}
+\includegraphics[width=0.48\textwidth]{Chapter-6/Images/Plots100A_MCData_Inc_StatSyst.pdf}
 \caption{Raw number of incident pion candidates as a function of the reconstructed kinetic energy for the 60A runs (lef) and for the 100A runs (right).
@@
 \centering  
-\includegraphics[width=0.48\textwidth]{Chapter-6/Images/Plots60A.png}
-\includegraphics[width=0.48\textwidth]{Chapter-6/Images/Plots100A.png}
+\includegraphics[width=0.48\textwidth]{Chapter-6/Images/Plots60A_MCData_XS_StatSyst.pdf}
+\includegraphics[width=0.48\textwidth]{Chapter-6/Images/Plots100A_MCData_XS_StatSyst.pdf}
 \caption{Raw ($\pi^-$-Ar) total hadronic cross section for the 60A runs (lef) and for the 100A runs (right). The statistical uncertainties are shown in azure, the systematic uncertainties in blue.}
@@
 \subsection{Treatment of Systematics} \label{ch:SysUncertaintyXSRaw}
-The only systematic effect considered in the measurement of the raw cross section come from the propagation of the uncertainty associate with the measurement of the kinetic energy at each slab.
+The only systematic effect considered in the measurement of the raw cross section results from the propagation of the uncertainty associate with the measurement of the kinetic energy at each slab.
 

### 2018-05-21 Chapter-4/4-Methodology.tex
s

@@
 \begin{equation}
-   \sigma^{\pi^-}(E_{i})  = \frac{1}{n \delta X}\frac{ \epsilon^{inc}_i  \text{  } N^{\text{TOT}}_{\text{Interacting}} (E_{i}) \text{  } C^{\pi MC}_{Interacting} (E_{i}) }{   \epsilon^{int}_i \text{  }N^{\text{TOT}}_{\text{Incident}} (E_{i})\text{  } C^{\pi MC}_{Incident} (E_{i})}.
+      \sigma^{\pi^-}(E_{i})  = \frac{1}{n \delta X}\frac{ \epsilon^{\text{Inc}}(E_i)  \hspace{0.2cm} C^{\pi MC}_{\text{Int}} (E_{i}) \hspace{0.2cm} N^{\text{TOT}}_{\text{Int}} (E_{i}) }{   \epsilon^{\text{Int}}(E_i) \hspace{0.2cm} C^{\pi MC}_{\text{Inc}} (E_{i}) \hspace{0.2cm}  N^{\text{TOT}}_{\text{Inc}} (E_{i})}.
 \label{eq:C}

### 2018-05-21 Chapter-6/6-PionCrossSection.tex
s

@@
 \section{Raw Cross Section}\label{ch:PionXSRaw}
-We measure the ($\pi^-$-Ar) cross section as a function of the kinetic energy in the two chosen data sets, the 60A and 100A negative runs. 
-As will be clarified in \ref{ch:PionXSCorrections},  the corrections to the raw cross section depend on the beam conditions and need to proceed independently for the two data sets.
-Thus, we present here the two measurements separately.
+We measure the ($\pi^-$-Ar) cross section as a function of the kinetic energy in the two chosen data sets, the -60A and -100A negative runs. 
+As we will clarify in Section \ref{ch:PionXSCorrections},  the corrections to the raw cross section depend on the beam conditions and need to be calculated independently for the two data sets.
+Thus, we present here the measurements on the two datasets separately.
 
@@
 \begin{equation}
- \sigma_{TOT} (E_i)  = \frac{1}{n \delta X}\frac{N_{Interacting}(E_i)}{N_{Incident}(E_i)}.
+ \sigma_{TOT} (E_i)  = \frac{1}{n \delta X}\frac{N^{\text{TOT}}_{\text{Int}}(E_i)}{N^{\text{TOT}}_{\text{Inc}}(E_i)}.
 \end{equation}
 
-where $N_{Interacting}$  is the number of particles interacting in an argon slice at kinetic energy $E_i$, $N_{Incident}$ is number of particles incident  on the argon slice at   kinetic energy $E_i$,  $n$ is the density of the target centers  and $\delta X$ is the thickness of the argon slice.
+where $N^{\text{TOT}}_{\text{Int}}$  is the number of particles interacting at kinetic energy $E_i$, $N^{\text{TOT}}_{\text{Inc}}$ is number of particles incident  on an argon slice at  kinetic energy $E_i$,  $n$ is the density of the target centers  and $\delta X$ is the thickness of the argon slice.
 
-Figure \ref{fig:InteractingRaw} shows the interacting histogram for the 60A dataset on the left and for the 100A dataset on the right. 
-Figure \ref{fig:IncidentRaw} shows the incident histogram for the 60A dataset on the left and for the 100A dataset on the right. 
-Figure \ref{fig:XSRaw} shows the raw cross section for the 60A dataset on the left and for the 100A dataset on the right.
-On all plots the same color scheme is used: the statistical uncertainty is shown in azure, while the systematic uncertainty is shown in blue.
-The calculation of the statistical unceratinty is laid out in section \ref{ch:StatUncertaintyXSRaw}, while the systematics on this \ref{ch:SysUncertaintyXSRaw}.
+Figure \ref{fig:InteractingRaw} shows the distribution of  $N^{\text{TOT}}_{\text{Int}}$  as a function of the kinetic energy for the 60A dataset on the left and for the 100A dataset on the right.
+The data central points are represented by black dots, the statistical uncertainty is shown in black, while the systematic uncertainty is shown in red.
+Data is displayed over the raw cross section obtained with a MC mixed sample of pions, muon and electrons in the percentage predicted by G4Beamline.
+The contribution from the simulated pions are shown in blue, the ones from secondaries in red, the ones from muons in yellow and the ones from electrons in gray.
+The simulated pion's and backgrounds' contributions are stacked; the sum of the integrals from each particle species is normalized to the integral of the data.
+
+Figure \ref{fig:IncidentRaw} shows the distribution of  $N^{\text{TOT}}_{\text{Inc}}$   for the 60A dataset on the left and for the 100A dataset on the right.
+Data is displayed over the MC.
+The same color scheme and normalization procedure is used for both the interacting and incident histograms.
+
+Figure \ref{fig:XSRaw} shows the raw cross section for the 60A dataset on the left and for the 100A dataset on the right, statistical uncertainty in black and systematic uncertainty in red.
+The raw data cross section is overlaid to the reconstructed MC cross section (in azure).
+
+The calculation of the statistical uncertainty for the interacting, incident and cross section measurements is laid out in Section \ref{ch:StatUncertaintyXSRaw}, while the  corresponding systematics  uncertainty on Section \ref{ch:SysUncertaintyXSRaw}.
 
@@
 \includegraphics[width=0.48\textwidth]{Chapter-6/Images/Plots100A_MCData_Int_StatSyst.pdf}
-\caption{Raw number of interacting pion candidates as a function of the reconstructed kinetic energy for the 60A runs (lef) and for the 100A runs (right).
-The statistical uncertainties are shown in azure, the systematic uncertainties in blue.}
+\caption{Raw number of interacting pion candidates as a function of the reconstructed kinetic energy for the 60A runs (left) and for the 100A runs (right).
+The statistical uncertainties are shown in black, the systematic uncertainties in red.}
 \label{fig:InteractingRaw}
@@
 \includegraphics[width=0.48\textwidth]{Chapter-6/Images/Plots100A_MCData_Inc_StatSyst.pdf}
-\caption{Raw number of incident pion candidates as a function of the reconstructed kinetic energy for the 60A runs (lef) and for the 100A runs (right).
-The statistical uncertainties are shown in azure, the systematic uncertainties in blue.}
+\caption{Raw number of incident pion candidates as a function of the reconstructed kinetic energy for the 60A runs (left) and for the 100A runs (right).
+The statistical uncertainties are shown in black, the systematic uncertainties in red.}
 \label{fig:IncidentRaw}
@@
 \includegraphics[width=0.48\textwidth]{Chapter-6/Images/Plots100A_MCData_XS_StatSyst.pdf}
-\caption{Raw ($\pi^-$-Ar) total hadronic cross section for the 60A runs (lef) and for the 100A runs (right). The statistical uncertainties are shown in azure, the systematic uncertainties in blue.}
+\caption{Raw ($\pi^-$-Ar) total hadronic cross section for the 60A runs (left) and for the 100A runs (right).
+The statistical uncertainties are shown in black, the systematic uncertainties in red.
+The raw cross section obtained with a MC mixed sample of pions, muon and electrons in the percentage predicted by G4Beamline is shown in azure.
+}
 \label{fig:XSRaw}
@@
 \subsection{Statistical Uncertainty}\label{ch:StatUncertaintyXSRaw}
-The statistical uncertainty for each kinetic energy bin of the cross section plot is calculated by error propagation from the statistical uncertainty on $N_{Incident}$ and $N_{Interacting}$ correspondent bin.
-Since the number of incident hadrons in each energy bin is given by a simple counting, we assume that $N_{Incident}$ is distributed as a poissonian with mean and variance equal to $N_{Incident}$ in each bin.
-On the other hand, $N_{Interacting}$ follows a binomial distribution: a particle in a given energy bin might or might not interact.  The variance for the binomial is given by  
+The statistical uncertainty for each kinetic energy bin of the cross section plot is calculated by error propagation from the statistical uncertainty on $N^{\text{TOT}}_{\text{Inc}}$ and $N^{\text{TOT}}_{\text{Int}}$ correspondent bin.
+Since the number of incident hadrons in each energy bin is given by a simple counting, we assume that $N^{\text{TOT}}_{\text{Inc}}$ is distributed as a poissonian with mean and variance equal to $N^{\text{TOT}}_{\text{Inc}}$ in each bin.
+On the other hand, $N^{\text{TOT}}_{\text{Int}}$ follows a binomial distribution: a particle in a given energy bin might or might not interact.  The variance for the binomial is given by  
 \begin{equation}
-\text{\textsf{Var[$N_{Incident}$]}} = \mathcal{N}P_{Interacting}(1-P_{Interacting});
+\text{\textsf{Var[}} N^{\text{TOT}}_{\text{Int}} \text{\textsf{]}}
+ = \mathcal{N}P_{Interacting}(1-P_{Interacting});
 \label{eq:binVar}
@@
 
-since the interaction probability $P_{Interacting}$ is $\frac{ N_{Interacting}}{N_{Incident}}$ and the number of tries $\mathcal{N}$ is $N_{Incident}$, equation \ref{eq:binVar} translates into
+since the interaction probability $P_{Interacting}$ is $\frac{ N^{\text{TOT}}_{\text{Int}}}{N^{\text{TOT}}_{\text{Inc}}}$ and the number of tries $\mathcal{N}$ is $N^{\text{TOT}}_{\text{Inc}}$, equation \ref{eq:binVar} translates into
 \begin{equation}
-\text{\textsf{Var[$N_{Incident}$]}} = N_{Incident}\frac{ N_{Interacting}}{N_{Incident}} (1-\frac{ N_{Interacting}}{N_{Incident}}) = N_{Interacting}(1-\frac{ N_{Interacting}}{N_{Incident}}). 
+\text{\textsf{Var[}} N^{\text{TOT}}_{\text{Int}} \text{\textsf{]}}
+= N^{\text{TOT}}_{\text{Inc}}\frac{ N^{\text{TOT}}_{\text{Int}}}{N^{\text{TOT}}_{\text{Inc}}} (1-\frac{ N^{\text{TOT}}_{\text{Int}}}{N^{\text{TOT}}_{\text{Inc}}}) = N^{\text{TOT}}_{\text{Int}}(1-\frac{ N^{\text{TOT}}_{\text{Int}}}{N^{\text{TOT}}_{\text{Inc}}}).
 \end{equation}
 
-$N_{Incident}$ and $N_{Interacting}$ are not independent.
+$N^{\text{TOT}}_{\text{Inc}}$ and $N^{\text{TOT}}_{\text{Int}}$ are not independent.
 The statistical uncertainty on the cross section is thus calculated as 
 \begin{equation}
-\delta\sigma_{tot}(E) = \sigma_{tot}(E) \Big(\frac{\delta N_{Interacting}}{N_{Interacting}}+\frac{\delta N_{Incident}}{N_{Incident}}\Big) 
+\delta\sigma_{tot}(E) = \sigma_{tot}(E) \Big(\frac{\delta N^{\text{TOT}}_{\text{Int}}}{N^{\text{TOT}}_{\text{Int}}}+\frac{\delta N^{\text{TOT}}_{\text{Inc}}}{N^{\text{TOT}}_{\text{Inc}}}\Big) 
 \end{equation}
@@
 \begin{eqnarray}
-\delta N_{Incident} = \sqrt[]{N_{Incident}} \\
-\delta N_{Interacting} = \sqrt[]{N_{Interacting}\Big(1-\frac{ N_{Interacting}}{N_{Incident}}\Big)}.
+\delta N^{\text{TOT}}_{\text{Inc}} = \sqrt[]{N^{\text{TOT}}_{\text{Inc}}} \\
+\delta N^{\text{TOT}}_{\text{Int}} = \sqrt[]{N^{\text{TOT}}_{\text{Int}}\Big(1-\frac{ N^{\text{TOT}}_{\text{Int}}}{N^{\text{TOT}}_{\text{Inc}}}\Big)}.
 \end{eqnarray}
@@
 \section{Corrections to the Raw Cross Section}\label{ch:PionXSCorrections}
-\subsection{Treatment of Systematics}
+As described in section \ref{ch:MCCorrections} as series of corrections are needed to derive the true pion cross section from the raw cross section. 
+These corrections are described in equation \ref{eq:C},
+
+\begin{equation}
+   \sigma^{\pi^-}_{TOT}(E_{i})  = \frac{1}{n \delta X}\frac{ \epsilon^{\text{Inc}}(E_i)  \hspace{0.2cm} C^{\pi MC}_{\text{Int}} (E_{i}) \hspace{0.2cm} N^{\text{TOT}}_{\text{Int}} (E_{i}) }{   \epsilon^{\text{Int}}(E_i) \hspace{0.2cm} C^{\pi MC}_{\text{Inc}} (E_{i}) \hspace{0.2cm}  N^{\text{TOT}}_{\text{Inc}} (E_{i})}.
+ \tag{\ref{eq:C}}
+\end{equation}
+
+\subsection{Background subtraction}
+
+\begin{figure}[htb]
+\centering
+\includegraphics[width=\textwidth]{Chapter-6/Images/Bkg60A_inc_int.pdf}
+\includegraphics[width=\textwidth]{Chapter-6/Images/Bkg100A_inc_int.pdf}
+\caption{.}
+\label{fig:BkgCorr}
+\end{figure}
+
+\subsubsection{Treatment of Systematics}
+
+\subsection{Efficiency Correction}
+\subsubsection{Treatment of Systematics}
+
+\subsection{Final Plots}
+
+\begin{figure}[htb]
+\centering
+\includegraphics[width=0.48\textwidth]{Chapter-6/Images/TheMoneyPlot60A.pdf}
+\includegraphics[width=0.48\textwidth]{Chapter-6/Images/TheMoneyPlot100A.pdf}
+\caption{.}
+\label{fig:FinalXSPion}
+\end{figure}
+
+\begin{figure}[htb]
+\centering
+\includegraphics[width=0.48\textwidth]{Chapter-6/Images/TheMoneyPlot.pdf}
+\caption{.}
+\label{fig:FinalXSPion}
+\end{figure}

### 2018-05-21 Chapter-6/6-PionCrossSection.tex
s

@@
 \chapter{Negative Pion Cross Section Measurement}\label{ch:PionXS}
-%{\raggedleft ``\emph{Y ella es flama que se eleva, Y es un p\'ajaro a volar} \par}
-%{\raggedleft \emph{En la noche que se incendia, Estrella de oscuridad}"\par}
-%{\raggedleft \emph{Que busca entre la tiniebla, La dulce hoguera del beso}"\par}
-%{\raggedleft -- Lila Downs, Benediction And Dream,  2002-- \par}
+%{\raggedleft ``\emph{Y ella es flama que se eleva, Y es un p\'ajaro a volar.} \par}
+%{\raggedleft \emph{En la noche que se incendia, estrella de oscuridad}\par}
+%{\raggedleft \emph{que busca entre la tiniebla, la dulce hoguera del beso.}"\par}
+%{\raggedleft -- Lila Downs, Benediction And Dream,  2002 -- \par}
+%\vspace{0.5cm}
+
+In this chapter, we show the result of the thin slice method to measure 
+the ($\pi^-$-Ar) total hadronic cross section.
+In Section \ref{ch:PionXSRaw}, we start by measuring the raw cross section, i.e.
+the cross section obtained exclusively using data reconstruction, with no additional corrections.
+In Section \ref{ch:PionXSCorrections}, we apply a series of data manipulations geared toward the correction for the background contributions and thse correction for detection inefficiency, showing the final results in Section \ref{ch:FinalPion}.
 
 \section{Raw Cross Section}\label{ch:PionXSRaw}
-We measure the ($\pi^-$-Ar) cross section as a function of the kinetic energy in the two chosen data sets, the -60A and -100A negative runs. 
-As we will clarify in Section \ref{ch:PionXSCorrections},  the corrections to the raw cross section depend on the beam conditions and need to be calculated independently for the two data sets.
-Thus, we present here the measurements on the two datasets separately.
+We measure the raw ($\pi^-$-Ar) total hadronic cross section as a function of the kinetic energy in the two chosen data sets, the -60A and -100A negative runs. 
+As we will clarify in Section \ref{ch:PionXSCorrections},  the corrections to the raw cross section depend on the beam conditions and need to be calculated independently for the two datasets.
+Thus, we present here the measurement of the raw cross section on the two datasets separately.
 
@@
 \begin{equation}
- \sigma_{TOT} (E_i)  = \frac{1}{n \delta X}\frac{N^{\text{TOT}}_{\text{Int}}(E_i)}{N^{\text{TOT}}_{\text{Inc}}(E_i)}.
+ \sigma_{TOT} (E_i)  = \frac{1}{n \delta X}\frac{N^{\text{TOT}}_{\text{Int}}(E_i)}{N^{\text{TOT}}_{\text{Inc}}(E_i)},
 \end{equation}
 
-where $N^{\text{TOT}}_{\text{Int}}$  is the number of particles interacting at kinetic energy $E_i$, $N^{\text{TOT}}_{\text{Inc}}$ is number of particles incident  on an argon slice at  kinetic energy $E_i$,  $n$ is the density of the target centers  and $\delta X$ is the thickness of the argon slice.
+where $N^{\text{TOT}}_{\text{Int}}$  is the measured number of particles interacting at kinetic energy $E_i$, $N^{\text{TOT}}_{\text{Inc}}$ is the  measured  number of particles incident  on an argon slice at  kinetic energy $E_i$,  $n$ is the density of the target centers  and $\delta X$ is the thickness of the argon slice.
+The density of the target centers and the slab thickness are $n = 0.021\cdot10^{24} \text{ cm}^{-3} $ and  $\delta X=0.47\text{ cm}$, respectively.
 
@@
 The data central points are represented by black dots, the statistical uncertainty is shown in black, while the systematic uncertainty is shown in red.
-Data is displayed over the raw cross section obtained with a MC mixed sample of pions, muon and electrons in the percentage predicted by G4Beamline.
-The contribution from the simulated pions are shown in blue, the ones from secondaries in red, the ones from muons in yellow and the ones from electrons in gray.
+Data is displayed over the $N^{\text{TOT}}_{\text{Int}}$  distribution obtained with a MC mixed sample of pions, muon and electrons (additional details on the composition will be provided in Section \ref{ch:BKGsubXS}).
+The contribution from the simulated pions is shown in blue, the one from secondaries in red, the one from muons in yellow and the ones from electrons in gray.
 The simulated pion's and backgrounds' contributions are stacked; the sum of the integrals from each particle species is normalized to the integral of the data.
@@
 Figure \ref{fig:XSRaw} shows the raw cross section for the 60A dataset on the left and for the 100A dataset on the right, statistical uncertainty in black and systematic uncertainty in red.
-The raw data cross section is overlaid to the reconstructed MC cross section (in azure).
+The raw data cross section is overlaid to the reconstructed cross section for the MC mixed sample, displayed in azure.
+Since the background contributions and the detector effects for the 60A and 100A sample are different, it is premature to compare the raw cross sections obtained from the two samples at this point.
 
-The calculation of the statistical uncertainty for the interacting, incident and cross section measurements is laid out in Section \ref{ch:StatUncertaintyXSRaw}, while the  corresponding systematics  uncertainty on Section \ref{ch:SysUncertaintyXSRaw}.
+We describe the calculation of the statistical uncertainty for the interacting, incident and cross section distributions in Section \ref{ch:StatUncertaintyXSRaw}; we describe the procedure to calculate the corresponding systematics uncertainty on Section \ref{ch:SysUncertaintyXSRaw}.
 
@@
 \caption{Raw number of incident pion candidates as a function of the reconstructed kinetic energy for the 60A runs (left) and for the 100A runs (right).
-The statistical uncertainties are shown in black, the systematic uncertainties in red.}
+The statistical uncertainty is shown in black, the systematic uncertainties in red.}
 \label{fig:IncidentRaw}
@@
 \caption{Raw ($\pi^-$-Ar) total hadronic cross section for the 60A runs (left) and for the 100A runs (right).
-The statistical uncertainties are shown in black, the systematic uncertainties in red.
+The statistical uncertainty is shown in black, the systematic uncertainties in red.
 The raw cross section obtained with a MC mixed sample of pions, muon and electrons in the percentage predicted by G4Beamline is shown in azure.
@@
 \subsection{Statistical Uncertainty}\label{ch:StatUncertaintyXSRaw}
-The statistical uncertainty for each kinetic energy bin of the cross section plot is calculated by error propagation from the statistical uncertainty on $N^{\text{TOT}}_{\text{Inc}}$ and $N^{\text{TOT}}_{\text{Int}}$ correspondent bin.
-Since the number of incident hadrons in each energy bin is given by a simple counting, we assume that $N^{\text{TOT}}_{\text{Inc}}$ is distributed as a poissonian with mean and variance equal to $N^{\text{TOT}}_{\text{Inc}}$ in each bin.
+The statistical uncertainty for a given kinetic energy bin of the cross section  is calculated by error propagation from the statistical uncertainty on $N^{\text{TOT}}_{\text{Inc}}$ and $N^{\text{TOT}}_{\text{Int}}$ correspondent bin.
+Since the number of incident particles in each energy bin is given by a simple counting, we assume that $N^{\text{TOT}}_{\text{Inc}}$ is distributed as a poissonian with mean and variance equal to $N^{\text{TOT}}_{\text{Inc}}$ in each bin.
 On the other hand, $N^{\text{TOT}}_{\text{Int}}$ follows a binomial distribution: a particle in a given energy bin might or might not interact.  The variance for the binomial is given by  
@@
 \text{\textsf{Var[}} N^{\text{TOT}}_{\text{Int}} \text{\textsf{]}}
- = \mathcal{N}P_{Interacting}(1-P_{Interacting});
+ = \mathcal{N}P_{Interacting}(1-P_{Interacting}).
 \label{eq:binVar}
@@
 
-since the interaction probability $P_{Interacting}$ is $\frac{ N^{\text{TOT}}_{\text{Int}}}{N^{\text{TOT}}_{\text{Inc}}}$ and the number of tries $\mathcal{N}$ is $N^{\text{TOT}}_{\text{Inc}}$, equation \ref{eq:binVar} translates into
+Since the interaction probability $P_{Interacting}$ is $\frac{ N^{\text{TOT}}_{\text{Int}}}{N^{\text{TOT}}_{\text{Inc}}}$ and the number of tries $\mathcal{N}$ is $N^{\text{TOT}}_{\text{Inc}}$, equation \ref{eq:binVar} translates into
 \begin{equation}
@@
 \begin{equation}
-\delta\sigma_{tot}(E) = \sigma_{tot}(E) \Big(\frac{\delta N^{\text{TOT}}_{\text{Int}}}{N^{\text{TOT}}_{\text{Int}}}+\frac{\delta N^{\text{TOT}}_{\text{Inc}}}{N^{\text{TOT}}_{\text{Inc}}}\Big) 
+\delta\sigma_{TOT}(E) = \sigma_{TOT}(E) \Big(\frac{\delta N^{\text{TOT}}_{\text{Int}}}{N^{\text{TOT}}_{\text{Int}}}+\frac{\delta N^{\text{TOT}}_{\text{Inc}}}{N^{\text{TOT}}_{\text{Inc}}}\Big) 
 \end{equation}
@@
 
-\subsection{Background subtraction}
+\subsection{Background subtraction}\label{ch:BKGsubXS}
 
@@
 
-\subsection{Final Plots}
+\section{Final Plots}\label{ch:FinalPion}
 

### 2018-05-21 Chapter-6/6-PionCrossSection.tex
s

@@
 \subsection{Treatment of Systematics} \label{ch:SysUncertaintyXSRaw}
-The only systematic effect considered in the measurement of the raw cross section results from the propagation of the uncertainty associate with the measurement of the kinetic energy at each slab.
+The only systematic effect considered in the measurement of the raw cross section results from the propagation of the uncertainty associate with the measurement of the kinetic energy at each argon slab.
+As shown in Section \ref{ch:kinEn}, the uncertainty on the kinetic energy of a pion candidate at the j$^{th}$ slab of argon  is given by
+
+\begin{eqnarray}
+\delta KE_{j} &=& \sqrt{\delta p_{Beam}^2 + \delta E_{Loss}^2 +  \delta  E_{\text{dep FF-j}}^2},\\
+&=& \sqrt{(2\% \text{ }p_{Beam})^2 +  (\sim 7 \text{ [MeV]})^2 +  (j-1)^2 (\sim0.08\text{ [MeV]})^2}.
+\end{eqnarray}
+
+We propagate this uncertainty  by varying the energy measurement $KE_{j}$ at each argon slab.
+We calculating the $N^{\text{TOT}}_{\text{Inc}}$,  $N^{\text{TOT}}_{\text{Int}}$ and cross section plots in three cases: first assigning the measured $KE_{j}$ at each kinetic energy sampling, then assigning $KE_{j} + \delta KE_{j}$, and finally assigning $KE_{j} - \delta KE_{j}$.
+The difference between the values obtained using the $KE_{j}$ sampling and the maximum and minimum values in each kinetic energy bin determines the systematic uncertainty.
+
+%We propagate this uncertainty  by calculating the $N^{\text{TOT}}_{\text{Inc}}$,  $N^{\text{TOT}}_{\text{Int}}$ and cross section plots twice: first assigning $KE_{j} + \delta KE_{j}$ at each kinetic energy sampling, then assigning $KE_{j} - \delta KE_{j}$.
+The difference between the central value and the maximum and minimum value in each kinetic energy bin gives the systematic uncertainty.
 

### 2018-05-21 AppendixB-TrackingStudies/B-Tracking.tex
s

\chapter{Tracking Studies}\label{ch:AppendixTrack}

### 2018-05-21 AppendixC-EnergyCalibration/C-EnergyCalibration.tex
s

\chapter{Energy Calibration}\label{ch:AppendixEnergy}

### 2018-05-21 Chapter-1/1-Theory.tex
s

@@
 Figure \ref{fig:TrueCarbon} shows the total hadronic cross section for carbon implemented in Geant4 10.01.p3 overlaid with the Bugg and Friedman data.
-Unfortunately, the current version of Geant4 does not reproduce the data for carbon closely.
+Unfortunately, version 10.01.p3\footnote{It should be noted that the latest Geant4 version, 10.03.p3, uses a different parametrization for the kaon cross section and retrieves a better agreement with data.}  of Geant4, which is the version used for the simulation in this work, does not reproduce the data for carbon closely.
 On one hand, this evidence makes us even more wary when using the Monte Carlo in simulating the kaon-argon interactions.

### 2018-05-21 Chapter-4/4-Methodology.tex
s

--- 2018-05-18 Chapter-4/4-Methodology.tex
@@
 \chapter{Total Hadronic Cross Section Measurement Methodology}\label{ch:Interactions}
-This chapter describes the general procedure employed to measure a total hadronic differential cross section in LArIAT.
+%{\raggedleft ``\emph{Like a lemon to the lime and the bubble to the bee}" \par}
+%{\raggedleft -- Eazy-E, Gimmie that Nutt,  1993 -- \par}
+%\vspace{0.5cm}
+
+This chapter describes the general procedure employed to measure a total hadronic differential cross 10.03.p1section in LArIAT.
 Albeit with small differences, both the  ($\pi^{-}$,Ar) and (K$^{+}$,Ar) total hadronic cross section measurements rely on the same procedure described in details in the following sections.
-We start by selecting the particle of interest using a combination of beamline detectors and TPC information (Section \ref{ch:ParticleSelectionMeEquationthod}).
+We start by selecting the particle of interest using a combination of beamline detectors and TPC information (Section \ref{ch:ParticleSelectionMethod}).
 We then perform a handshake between the beamline information and the TPC tracking to assure the selection of the right TPC track (Section \ref{ch:WC2TPCMatchMethod}).
@@
 
-At the end of this chapter, we show a sanity check of the methodology against MC truth information (Section \ref{ch:procedureTesting}).
+At the end of this chapter, we show a sanity check of the methodology by applying the thin slice method employing only MC truth information and retrieving the Geant4 tabulated cross section for pions and kaons (Section \ref{ch:procedureTesting}).
 
@@
 \hline
-                                                        & Run-II Negative Polarity   &  Run-II Positive Polarity  \\ \hline
-Events Reconstructed in Beamline        &  158396  & 260810  \\ \hline
-Events with Plausible Trajectory            &   147468 & 240954  \\ \hline
-Beamline $\pi^-/\mu^-/e^-$  Candidate  &   138481 &     N.A.   \\ \hline
-Beamline $K^+$   Candidate                 &    N.A       & 2837     \\ \hline
-Events Surviving Pile Up Filter              &   108929  & 2389       \\ \hline
-Events with WC2TPC Match                 &    41757   & 1081 \\ \hline
-Events Surviving Shower Filter             &    40841    &  N.A.     \\ \hline
-Available Events For Cross Section      &   40841    &   1081    \\ \hline
+                                                        & Run-II Neg Pol   &  Run-II Pos Pol  \\ \hline
+1. Events Reconstructed in Beamline        &  158396  & 260810  \\ \hline
+2. Events with Plausible Trajectory            &   147468 & 240954  \\ \hline
+3. Beamline $\pi^-/\mu^-/e^-$  Candidate  &   138481 &     N.A.   \\ \hline
+4. Beamline $K^+$   Candidate                 &    N.A       & 2837     \\ \hline
+5. Events Surviving Pile Up Filter              &   108929  & 2389       \\ \hline
+6. Events with WC2TPC Match                 &    41757   & 1081 \\ \hline
+7. Events Surviving Shower Filter             &    40841    &  N.A.     \\ \hline
+8. Available Events For Cross Section      &   40841    &   1081    \\ \hline
 \end{tabular}
@@
 As shown in equation \ref{eq:enFF}, we leverage the beamline particle identification and momentum measurement before entering the TPC as in input to evaluate the kinetic energy for the hadrons used in the  cross sections measurements.
-Thus, we select the LArIAT data to keep only events whose wire chamber and time of flight information is registered (line 2 in in Table \ref{tab:beamlineDataSelection}).
-Additionally, we perform a check of the plausibility of the trajectory inside the beamline detectors: given the position of the hits in the four wire chambers, we make sure the particle's trajectory does not cross any impenetrable material such as the collimator and the magnets steel (line 3 in in Table \ref{tab:beamlineDataSelection}).
+Thus, we select the LArIAT data to keep only events whose wire chamber and time of flight information is registered (line 1 in in Table \ref{tab:beamlineDataSelection}).
+Additionally, we perform a check of the plausibility of the trajectory inside the beamline detectors: given the position of the hits in the four wire chambers, we make sure the particle's trajectory does not cross any impenetrable material such as the collimator and the magnets steel (line 2 in in Table \ref{tab:beamlineDataSelection}).
 
@@
 
-Lines 4 and 5 in in Table \ref{tab:beamlineDataSelection} show the number of negative $\pi/\mu/e$ and positive $K$ candidates which pass the mass selection for LArIAT Run-II data.
+Lines 3 and 4 in in Table \ref{tab:beamlineDataSelection} show the number of negative $\pi/\mu/e$ and positive $K$ candidates which pass the mass selection for LArIAT Run-II data.
 
@@
 We adjusted the primary beam intensity between LArIAT Run I and Run II to reduce the presence of events with high pile up particles in the data sample.
-For the cross section analyses, we remove events with more than 4 tracks in the first 14 cm upstream portion of the TPC from the sample (line 6 in in Table \ref{tab:beamlineDataSelection}).
+For the cross section analyses, we remove events with more than 4 tracks in the first 14 cm upstream portion of the TPC from the sample (line 5 in in Table \ref{tab:beamlineDataSelection}).
 
@@
 It is important to notice that the composition of the negative polarity beam is mostly pions, as will be discussed in section \ref{ch:beamlineComposition}.
-Anyhow, we devise a selection on the TPC information to mitigate the presence of electrons in the sample used for the pion cross section.
+Still, we devise a selection on the TPC information to mitigate the presence of electrons in the sample used for the pion cross section.
 The selection relies on the different topologies of a pion and an electron event in the argon: while the former will trace a track inside the TPC active volume, the latter will tend to ``shower", i.e.
 interact with the medium, producing bremsstrahlung photons which pair convert into several short tracks.
-In order to remove the shower topology, we create a region of interest (ROI) around the TPC track corresponding to the beamline particle (more details on this in the next section).
+In order to remove the shower topology, we create a region of interest (ROI) around the TPC track corresponding to the beamline particle.
 We look for short tracks contained in the ROI, as depicted in figure \ref{fig:showerFilt}:  if more then 5 tracks shorter than 10 cm are in the ROI, we reject the event.
-Line 8 in in Table \ref{tab:beamlineDataSelection}) shows the number of events surviving this selection.
+Line 7 in in Table \ref{tab:beamlineDataSelection} shows the number of events surviving this selection.
 
@@
 A kink is visible in the kaon ionization, signature of a hadronic interaction: the tracking correctly stops at the kink position and two tracks are formed.
-An additional pile-up track is so present in the event (top track).}
+An additional pile-up track is so present in the event (top track in red).}
 \label{fig:kaonInteraction}
@@
 We attempt to uniquely match one wire chamber track to one and only one reconstructed TPC track. 
-In order to determine if the presence of a match, we apply a geometrical selection on the relative the position of the wire chamber and TPC tracks. 
+In order to determine if a match is present, we apply a geometrical selection on the relative the position of the wire chamber and TPC tracks. 
 We start by considering only TPC tracks whose first point is in the first 2 cm upstream portion of the TPC for the match.
@@
 We define  as $\alpha$ the angle between the incident WC track and the TPC track in the plane that contains them.
-If  $\Delta \text{R} < 4 $~cm, $\alpha < 8^\circ $,  a match between WC-track and TPC reconstructed track is found.
-We describe  how we determine the value for the radius and angular selection in sec \ref{ch:WC2TPCMatchOptimization}.
+If  $\Delta \text{R} < 4 $~cm, $\alpha < 8^\circ $,  a match between WC-track and TPC track is found.
+We describe  how we determine the value for the radius and angular selection in Section \ref{ch:WC2TPCMatchOptimization}.
+We discard events with multiple WC2TPC matches.
+We use only those TPC tracks that are matched to WC tracks in the cross section calculation.
+Line 6 in Table \ref{tab:beamlineDataSelection} shows the number of events where a unique WC2TPC match was found.
+
 In MC, we mimic the matching between the WC and the TPC track by constructing a fake WC track using truth information at wire chamber four.
 We then apply the same WC to TPC matching algorithm as in data.
-We discard events with multiple WC2TPC matches. We use only those TPC tracks that are matched to WC tracks in the cross section calculation.
 
@@
 Cross section measurements on a thin target have been the bread and butter of nuclear and particle experimentalists since the Geiger-Marsden experiments \cite{Geiger1909}.
-At their core, these types of experiments consist in shooting a beam of particles with a known flux on a thin target and recording the outgoing flux.
+At their core, this type of experiments consists in shooting a beam of particles with a known flux on a thin slab of material and recording the outgoing flux.
 
 In general, the target is not a single particle, but rather a slab of material containing many diffusion centers.
-The so-called  ``thin target" approximation assumes that the target centers are uniformly distributed in the material and that the target is thin compared to the projectile interaction length,WC2TPC so that no center of interaction sits in front of another.
-In this approximation, the ratio between the number of particles interacting in the target $N_{Interacting}$ and number of incident particles $N_{Incident}$ determines the interaction probability $P_{Interacting}$, which is the complementary to one of the survival probability $P_{Survival}$.
+The so-called  ``thin target" approximation assumes that the target centers are uniformly distributed in the material and that the target is thin compared to the projectile interaction length, so that no center of interaction sits in front of another.
+In this approximation, the ratio between the number of particles interacting in the target $N_{\text{Int}}$ and number of incident particles $N_{\text{Inc}}$ on the target determines the interaction probability $P_{Interacting}$, which is the complementary to one of the survival probability $P_{Survival}$.
 Equation \ref{eq:thinTargetXS} 
 \begin{equation}
-P_{Survival} = 1- P_{Interacting} = 1 - \frac{N_{Interacting}}{N_{Incident}} = e^{-\sigma_{TOT} n \delta X}
+P_{Survival} = 1- P_{Interacting} = 1 - \frac{N_{\text{Int}}}{N_{\text{Inc}}} = e^{-\sigma_{TOT} n \delta X}
 \label{eq:thinTargetXS}
@@
 describes the probability for a particle to survive the thin target.
-This formula relates the total cross section $\sigma_{TOT}$, the density of the target centers  $n$  and  the thickness of the target  along the incident hadron direction $\delta X$, to the interaction probability\footnote{The scattering center density in the target, {\emph{n}},  relates to the argon density $\rho$, the Avogadro number  $ N_{A} $ and the argon molar mass $m_A$ as $n=\frac{\rho N_{A} }{m_A}$.}.
-If the target is thin compared to the interaction length of the process considered, we can Taylor expand the exponential function in equation \ref{eq:thinTargetXS} and find a simple proportionality relationship between the number of incident and interacting particles, and the cross section, as shown in equation \ref{eq:thinTargetXSTaylor}:
+This formula relates  the interaction probability to the total hadronic cross section ($\sigma_{TOT}$), the density of the target centers ($n$)\footnote{The scattering center density in the target, {\emph{n}},  relates to the argon density $\rho$, the Avogadro number  $ N_{A} $ and the argon molar mass $m_A$ as $n=\frac{\rho N_{A} }{m_A}$.}    and  the thickness of the target  along the incident hadron direction ($\delta X$).
+If the target is thin compared to the interaction length of the process considered, we can Taylor expand the exponential function in equation \ref{eq:thinTargetXS} and find a simple proportionality relationship between the cross section and the number of incident and interacting particles, as shown in equation \ref{eq:thinTargetXSTaylor}:
 \begin{equation}
-1 - \frac{N_{Interacting}}{N_{Incident}} =  1 -\sigma_{TOT} n \delta X + O(\delta X^2).
+1 - \frac{N_{\text{Int}}}{N_{\text{Inc}}} =  1 -\sigma_{TOT} n \delta X + O(\delta X^2).
 \label{eq:thinTargetXSTaylor}
@@
 \begin{equation}
- \sigma_{TOT}  = \frac{1}{n \delta X}\frac{N_{Interacting}}{N_{Incident}}.
+ \sigma_{TOT}  = \frac{1}{n \delta X}\frac{N_{\text{Int}}}{N_{\text{Inc}}}.
 \label{eq:thinTargetXSSolved}
@@
 The wires are oriented at +/- $60^{\circ}$ from the vertical direction at 4 mm spacing, while the beam direction is oriented 3 degrees off the $z$ axis in the $XZ$ plane.
-The wires collect signals proportional to the energy loss of the hadron along its path in a  $\delta${\emph{X}} = 4 mm/sin($60^{\circ}$) $\approx$ 4.7~mm slab of liquid argon.
+The wires collect signals proportional to the energy loss of the hadron along its path in a  $\delta${\emph{X}} = 4 mm/(sin($60^{\circ}$)cos($3^{\circ}$)) $\approx$ 4.7~mm slab of liquid argon.
 Thus, one can think to slice the TPC into many thin targets of $\delta${\emph{X}} = 4.7~mm thickness along the direction of the incident particle, making a measurement at each wire along the path.
@@
 \end{equation}
-where $E_{loss}$ is a correction for the energy loss in the dead material between the beamline and the TPC front face.
+where $E_{loss}$ is a correction for the energy loss in the uninstrumented material between the beamline and the TPC front face.
 The  energy of the hadron at each slab is determined by subtracting the energy released by the particle in the previous slabs.
@@
 \begin{equation}
- E_{j}^{kin} =  E^{kin}_{Front Face} - \sum_{i < j} \Delta E_i,
+ E_{j}^{kin} =  E^{kin}_{Front Face} - \sum_{i < j} E_{\text{Dep},i},
 \label{eq:KEj}
 \end{equation}
-where $\Delta E_i$ is the energy deposited at each argon slice before the $j^{th}$ point as measured by the calorimetry associated with the tracking.
+where $E_{\text{Dep},i}$ is the energy deposited at each argon slice before the $j^{th}$ point as measured by the calorimetry associated with the tracking.
 
-If the particle enters a slice, it contributes to $N_{Incident}( E^{kin})$ in the energy bin corresponding to its kinetic energy in that slice.
-If it interacts in the slice, it then also contributes to $N_{Interacting}(E^{kin})$ in the appropriate energy bin.
-The cross section as a function of kinetic energy, $\sigma_{TOT}( E^{kin})$ will then be proportional to the ratio $\frac{N_{Interacting}( E^{kin})}{N_{Incident}( E^{kin})}$ .
+If the particle enters a slice, it contributes to $N_{\text{Inc}}( E^{kin})$ in the energy bin corresponding to its kinetic energy in that slice.
+If it interacts in the slice, it also contributes to $N_{\text{Int}}(E^{kin})$ in the appropriate energy bin.
+The cross section as a function of kinetic energy, $\sigma_{TOT}( E^{kin})$ will then be proportional to the ratio $\frac{N_{\text{Int}}( E^{kin})}{N_{\text{Inc}}( E^{kin})}$.
+
+Our goal is to measure the total interaction cross section, independently  from the topology of the interaction.
+Thus, we determine that a hadron interacted simply by requiring that the last point of the WC2TPC matched track lies inside the fiducial volume, whose boundaries are defined in Table \ref{tab:FidVol}.
+If the TPC track stops within the fiducial volume, its last point will be the interaction point; if the track crosses the boundaries of the fiducial volume, the track will be considered ``through going" and no interaction point will be found.
+The only slabs considered to fill the  $N_{\text{Inc}})$ and  $N_{\text{Inc}}$ plots are the slabs included the fiducial volume.
+
+\begin{table}[t]
+\centering
+\begin{tabular}{|l|r|r|}
+\hline
+& min   &  max  \\ \hline
+$X$ & 1 cm   & 46 cm  \\ \hline
+$Y$ & -15 cm   & 15  cm  \\ \hline
+$Z$ & 0 cm   & 86 cm  \\ \hline
+\end{tabular}
+\caption{Fiducial volume boundaries used to determine cross section interaction point. }
+\label{tab:FidVol}
+\end{table}
 
@@
 \begin{equation}
- \sigma^{\pi^-}(E_{i})  = \frac{1}{n \delta X}\frac{N^{\pi^-}_{ \text{Interacting}} (E_{i})}{N^{\pi^-}_{ \text{Incident}}(E_{i})}.
+ \sigma^{\pi^-}_{TOT}(E_{i})  = \frac{1}{n \delta X}\frac{N^{\pi^-}_{ \text{Int}} (E_{i})}{N^{\pi^-}_{ \text{Inc}}(E_{i})}.
 \label{eq:thinTargetXSSolved2}
@@
 Also, the LArIAT TPC is not 100\% efficient in determining the interaction point.
-Therefore we need to apply two corrections evaluated on the MC in order to extract the cross section from LArIAT data: the background subtraction and the efficiency correction.
+Therefore we need to apply two corrections evaluated on the MC in order to extract the true cross section from LArIAT data: the background subtraction and the efficiency correction.
 Still using the pion case as example, we estimate the pion cross section in each energy bin changing  Equation \ref{eq:thinTargetXSSolved2} into
 \begin{equation}
- \sigma^{\pi^-}(E_{i})  =\frac{1}{n \delta X}\frac{N^{\pi^-}_{ \text{Interacting}} (E_{i})}{N^{\pi^-}_{ \text{Incident}}(E_{i})} = \frac{1}{n \delta X}\frac{ \epsilon^{inc}_i [ N^{ \text{TOT}}_{ \text{Interacting}} (E_{i}) - B_{ \text{Interacting}} (E_i)] }{   \epsilon^{int}_i [N^{ \text{TOT}}_{ \text{Incident}}(E_{i}) - B_{ \text{Incident}} (E_i)]},
+ \sigma^{\pi^-}_{TOT}(E_{i})  =\frac{1}{n \delta X}\frac{N^{\pi^-}_{ \text{Int}} (E_{i})}{N^{\pi^-}_{ \text{Inc}}(E_{i})} = \frac{1}{n \delta X}\frac{ \epsilon^{\text{Inc}}(E_i) [ N^{ \text{TOT}}_{ \text{Int}} (E_{i}) - B_{ \text{Int}} (E_i)] }{   \epsilon^{\text{Int}}(E_i) [N^{ \text{TOT}}_{ \text{Inc}}(E_{i}) - B_{ \text{Inc}} (E_i)]},
 \label{eq:True}
@@
 
-where  $N^{\text{TOT}}_{\text{Interacting}} (E_{i})$ and $N^{\text{TOT}}_{\text{Incident}} (E_{i})$ is the measured content of the interacting and incident histograms for events that pass the event selection, $B_{interacting} (E_i)$ and $B_{\text{Incident}} (E_i)$ represent the contributions from beamline background, and  $\epsilon^{int}_i$ and  $\epsilon^{inc}_i$ are the efficiency corrections for said histograms.
+where  $N^{\text{TOT}}_{\text{Int}} (E_{i})$ and $N^{\text{TOT}}_{\text{Incident}} (E_{i})$ is the measured content of the interacting and incident histograms for events that pass the event selection, $B_{\text{Int}} (E_i)$ and $B_{\text{Inc}} (E_i)$ represent the contributions from beamline background to the interacting and incident histograms respectively, and  $\epsilon^{\text{Int}}(E_i)$ and  $\epsilon^{\text{Inc}}(E_i)$ are the efficiency corrections for said histograms.
 
-As we will show in section \ref{sec:Correction}, the background subtraction for the interacting and incident histograms can be translated into a corresponding corrections $C^{\pi MC}_{Interacting} (E_{i})$ and $C^{\pi MC}_{Incident} (E_{i})$ and the cross section re-written as follows
+As we will show in section \ref{sec:Correction}, the background subtraction for the interacting and incident histograms can be translated into a corresponding relative pion content $C^{\pi MC}_{Interacting} (E_{i})$ and $C^{\pi MC}_{Incident} (E_{i})$ and the cross section re-written as follows
 
 \begin{equation}
-   \sigma^{\pi^-}(E_{i})  = \frac{1}{n \delta X}\frac{ \epsilon^{inc}_i  \text{  } N^{\text{TOT}}_{\text{Interacting}} (E_{i}) \text{  } C^{\pi MC}_{Interacting} (E_{i}) }{   \epsilon^{int}_i \text{  }N^{\text{TOT}}_{\text{Incident}} (E_{i})\text{  } C^{\pi MC}_{Incident} (E_{i})}.
+      \sigma^{\pi^-}_{TOT}(E_{i})  = \frac{1}{n \delta X}\frac{ \epsilon^{\text{Inc}}(E_i)  \hspace{0.2cm} C^{\pi MC}_{\text{Int}} (E_{i}) \hspace{0.2cm} N^{\text{TOT}}_{\text{Int}} (E_{i}) }{   \epsilon^{\text{Int}}(E_i) \hspace{0.2cm} C^{\pi MC}_{\text{Inc}} (E_{i}) \hspace{0.2cm}  N^{\text{TOT}}_{\text{Inc}} (E_{i})}.
 \label{eq:C}
@@
 
-For pions in the considered energy range, \textcolor{red}{the Geant4 inelastic model adopted to is ``BertiniCascade", while the elastic model ``hElasticLHEP".}
-For kaons, the Geant4 inelastic model adopted to is ``BertiniCascade", while the elastic model ``hElasticLHEP".
+For pions and kaons in the considered energy range, the Geant4 inelastic model adopted is ``BertiniCascade"; the pion elastic cross sections are modeled on Chips, while the kaon elastic cross sections are modeled on Gheisha and Chips.
 
-For the validation test, we fire about a sample of pions and a sample of kaons inside the LArIAT TPC active volume using the Data Driven Monte Carlo (see section \ref{sec:DDMC}).
-We apply  the thin-sliced method using only true quantities to calculate the hadron kinetic energy at each slab in order to decouple reconstruction effects from issues with the methodology.
+For the validation test, we fire a sample of pions and a sample of kaons inside the LArIAT TPC active volume using the Data Driven Monte Carlo (see section \ref{sec:DDMC}).
+We apply  the thin-sliced method using only true quantities to calculate the hadron kinetic energy at each slab in order to decouple reconstruction effects from possible issues with the methodology.
 For each slab of 4.7 mm length along the path of the hadron, we integrate the true energy deposition as given by the Geant4 transportation model.
@@
 
-Figure \ref{fig:TrueMCXS2} shows the total hadronic cross section for argon implemented in Geant4 10.01.p3 (solid lines) overlaid with the true MC cross section as obtained with the sliced TPC method (markers) for pions on the left and kaons on the right; the total cross section is shown in green,  the elastic cross section in blue and the inelastic cross section in red.
+Figure \ref{fig:TrueMCXS2} shows the total hadronic cross section for argon implemented in Geant4 10.03.p1 (solid lines) overlaid with the true MC cross section as obtained with the sliced TPC method (markers) for pions on the left and kaons on the right; the total cross section is shown in green,  the elastic cross section in blue and the inelastic cross section in red.
 The nice agreement with the Geant4 distribution and the cross section  obtained with the sliced TPC method gives us confidence in the  validity of the methodology.
@@
 \end{minipage}
-\caption{Hadronic cross sections for ($\pi^-$,Ar) on the left and (K$^+$,Ar) on the right as implemented in Geant4 10.01.p3 (solid lines) overlaid the true MC cross section as obtained with the sliced TPC method (markers).
+\caption{Hadronic cross sections for ($\pi^-$,Ar) on the left and (K$^+$,Ar) on the right as implemented in Geant4 10.03.p1 (solid lines) overlaid the true MC cross section as obtained with the sliced TPC method (markers).
 The total cross section is shown in green,  the elastic cross section in blue and the inelastic cross section in red.}

### 2018-05-21 Chapter-5/5-Preparatory.tex
s

@@
-\chapter{Preparatory Work}\label{ch:samples}
+\chapter{Data and MC preparation for the Cross Section Measurements}\label{ch:samples}
 %{\raggedleft ``\emph{Il dolce non lo mangi mai, ma qualche volta ti rifai.} \par}
@@
 This chapter describes the preparatory work done on the the data and Monte Carlo samples used for the cross section analyses. 
-This entails the choice of the data set and the production of the information needed to construct the Monte Carlo Simulation~(section \ref{sec:dataSet}),  the construction and use of said Monte Carlo simulation~(section \ref{sec:MCSet}), the study and optimization of the tracking in the TPC for the cross section analyses~(section \ref{sec:TrackingStudies}),  the calibration of the calorimetry response and related energy studies~(section \ref{ch:energyCal}).
+This entails the choice of the datasets and the production of the information needed to construct the Monte Carlo Simulation~(section \ref{sec:dataSet}),  the construction and use of said Monte Carlo simulation~(section \ref{sec:MCSet}), the study and optimization of the tracking in the TPC for the cross section analyses~(section \ref{sec:TrackingStudies}),  the calibration of the calorimetry response and related energy studies~(section \ref{ch:energyCal}).
 
-\section{Cross Section Analyses Data Set}\label{sec:dataSet}
+\section{Cross Section Analyses Data Sets}\label{sec:dataSet}
 We choose LArIAT Run-II as the data period for the  ($\pi^{-}$,Ar) and (K$^{+}$,Ar) total hadronic cross section analyses. 
@@
 The -60A and -100A combined data set accounts for approximately 90\% of the total Run-II negative polarity runs.
-Since the production of beamline Monte Carlo depends on the wanted beamline configuration, the choice of only two beamline settings limits the need for beamline MC production.
+The  choice of the main two beamline settings limits the need for the production of many MC sets and related corrections, still maintaining a high number of events.
+%Since the production of beamline Monte Carlo depends on the wanted beamline configuration, the choice of only two beamline settings limits the need for beamline MC production.
 
@@
 
-For the first measurements in LArIAT that uses both beamline and TPC information, we choose strict requirements on the reconstruction of the WC tracks, the so-called ``Picky Track" sample (see \ref{sec:MWPCfunc}).
+For the first measurements in LArIAT that uses both beamline and TPC information, we choose strict requirements on the reconstruction of the WC tracks, the so-called ``Picky Track" sample (see Section \ref{sec:MWPCfunc}).
 This choice presents two advantages:  the uncertainty on the momentum reconstruction for the ``Picky Tracks" sample is smaller compared to the ``High Yield" sample, and the comparison with the beamline MC results is straightforward.
@@
 \section{Construction of a Monte Carlo Simulation for LArIAT}\label{sec:MCSet}
-For the simulation of LArIAT events and their particle make up, we use a combination of two MC generators: the G4Beamline Monte Carlo and the Data Driven single particle Monte Carlo (DDMC).
+For the simulation of LArIAT events and for the simulation of the datasets' particle make up, we use a combination of two MC generators: the G4Beamline Monte Carlo and the Data Driven single particle Monte Carlo (DDMC).
 We use the G4Beamline MC to simulate the particle transportation in the beamline and calculate the particle composition of the beam just after the fourth Wire Chamber (WC4).
@@
 \subsection{G4Beamline}\label{ch:beamlineComposition}
-G4Beamline simulates the beam collision with the LArIAT secondary target, the energy deposited by the particles in the LArIAT beamline detectors, and the action of the LArIAT magnets, effectively accounting for particle transportation through the beam line from the LArIAT target until ``Big Disk", a fictional, void detector located just before the LArIAT cryostat.
- At the moment of this writing, G4Beamline does not simulated the responses of the beam line detectors.
+G4Beamline simulates the beam collision with the LArIAT secondary target, the energy deposited by the particles in the LArIAT beamline detectors, and the action of the LArIAT magnets, effectively accounting for particle transportation through the beamline from the LArIAT target until ``Big Disk", a fictional, void detector located just before the LArIAT cryostat.
+ At the moment of this writing, G4Beamline does not simulated the responses of the beamline detectors.
 It is possible to interrogate the truth level information of the simulated particles in several points of the geometry.
@@
 The first step of this process is assessing the percentage of electrons and muons in the $\pi/\mu/e$ beamline candidates via the G4Beamline MC.
-The full treatment of the beamline contamination in the pion cross section calculation is described in section \ref{ch:PionXSBkgSub}.
 Since the beamline composition is a function of the magnet settings, we simulate separately events for magnet current of -60A and -100A. 
@@
 
+The estimated beam composition is used as a basis to estimate the background contamination in the  ($\pi^{-}$,Ar) cross section measurement, whose  full treatment is described in section \ref{ch:PionXSBkgSub}.
+
 \subsubsection{Beam Composition for Positive Kaon Cross Section}
@@
 Since the shape of the tails is unknown, the estimate is done multiple times varying the range and shape for reasonable functions. 
-For example, to estimate the proton content under the kaon peak, we start by fitting the left tail of the proton mass distribution with a gaussian function between 650 $MeV/c^2$ and 750 $MeV/c^2$.% in a reasonable range and with a reasonable function.
-We extend the fit function under the kaon peak and integrate the between 350-650 $MeV/c^2$.
+For example, to estimate the proton content under the kaon peak, we start by fitting the left tail of the proton mass distribution with a gaussian function between 650 $MeV/c^2$ and 750 MeV/c$^2$.% in a reasonable range and with a reasonable function.
+  We extend the fit function under the kaon peak and integrate the extended fit function between 350-650 MeV/c$^2$.
 We integrate the mass histogram in the same range and calculate the proton contamination as the ratio between the two integrals.
-We repeat this procedure for several fit shapes (gaussian, linear and exponential functions are used) and tail ranges.
+We repeat this procedure for several fit shapes (gaussian, linear and exponential functions) and tail ranges.
 Finally, we calculate the contamination as the weighted average of single estimates, where the weights are calculated to be the $1./\chi^2$ of the tail fits.
@@
 With 12 iterations of this method we find a proton contamination of  0.2 $\pm$ 0.5 \%  and a contamination from the lighter particles of 5 $\pm$ 2 \% .
+The estimate of the proton background is currently not used in the kaon cross section analysis, but it is a fundamental step to retrieve the true kaon cross section which will be implemented in the analysis next step.
 
@@
 \includegraphics[width=0.5\textwidth,height=\textheight,keepaspectratio]{Chapter-5/Images/MassPos.png}
-\caption{$Left.$ Beam composition for the +100A runs after WC4 (no mass selection applied).
-The solid blue plot represents the simulated pion content, the yellow plot represents the simulated muon content and the grey plot represents the simulated positron content, the red the proton content and the mustard the kaon content.
-The plots are area normalized to the number of data events, shown in black.$Right.$ Mass distribution for the Run-II positive runs, where the area under the kaon mass peak is highlighted in purple.
+\caption{\emph{Left:} Beam composition for the +100A runs after WC4 (no mass selection applied).
+The solid colors represent the contributions from the G4Beamline simulated particles: blue plot represents the simulated pion content, the yellow plot represents the simulated muon content and the grey plot represents the simulated positron content, the red the proton content and the mustard the kaon content.
+The plots are area normalized to the number of data events, shown in black.
+\emph{Right:} Mass distribution for the Run-II positive runs, where the area under the kaon mass peak is highlighted in purple.
 The area under the extension of a possible fit for the proton tail is highlighted in red.
@@
 
-When producing a DDMC sample, beam line data from a particular running period and/or running condition are selected first.
+When producing a DDMC sample, beamline data from a particular running period and/or running condition are selected first.
 For example, data for the negative 60A runs and for the negative 100A runs inform the event generation stage of two different DDMC samples.
@@
 The momentum and position of the selected data form a 5-dimensional tupla, which we sample thousands of times through a 5-dimensional hit-or-miss sampling procedure to generate the MC events.
-This produces MC $P_x, P_y, P_z, X, Y$ distributions  with the same momentum and position distributions as data, with the additional benefit of accounting for the correlations between the considered variables.
+This generates MC events  with the same momentum and position distributions as data, with the additional benefit of accounting for the correlations between the $P_x, P_y, P_z, X, Y$ variables.
 As an example, the results of the DDMC generation compared to data for the kaon +100A sample are shown in figure \ref{fig:DDMCComparison} for the $P_z$, $X$ and $Y$ distributions; as expected, MC and data agree within the statistical uncertainty by construction.
@@
 
-Using the DDMC technique ensures that the MC and data particles have very similar momentum, position and angular distributions at WC4 and allows us to use the MC sample in several occasions, for example to calibrate the energy loss upstream of the TPC (see Section \ref{ch:eloss}) or to study the tracking and the calorimetric performance (sections \ref{sec:TrackingStudies} and \ref{ch:energyCal}).
+Using the DDMC technique ensures that the MC and data particles have very similar momentum, position and angular distributions at WC4 and allows us to use the MC sample in several occasions: to calibrate the energy loss upstream of the TPC (see Section \ref{ch:eloss}), to estimate the background contamination to the pion cross section (see Section \ref{ch:PionXSBkgSub}), or to study the tracking and the calorimetric performance (sections \ref{sec:TrackingStudies} and \ref{ch:energyCal}).
 A small caveat is in order here: the DDMC is a single particle Monte Carlo, which means that the beam pile-up is not simulated.
@@
 
-\subsection{Estimate of Energy Loss before the TPC}\label{ch:eloss}
+\section{Estimate of Backgrounds in the Pion Cross Section}\label{ch:PionXSBkgSub}
+We use the beamline simulation and the DDMC simulation to estimate the background in the total hadronic pion cross section.
+Two categories of background exists for the negative pion cross section measurement: the one related to the pion interaction in the chamber, discussed in Section \ref{ch:CaptureAndDecay} and the one related to the beamline contamination, discussed in Section \ref{ch:PionXSBkgSub2}.
+
+\subsection{Background from Pion Capture and Decay}\label{ch:CaptureAndDecay}
+Our goal is to measure the total hadronic cross section for negative pions in argon.
+Since pion capture can be classified as an electromagnetic process and pion decay is a week process,  capture and decay represent unwanted interactions.
+We present here a study of capture and decay in Monte Carlo and the solution we adopted to mitigate their occurrence in the data sample.
+
+For this MC study, we use a sample of  MC pions generated according to the  $-6$0A beam profile with the DDMC (see Section \ref{sec:DDMC}).
+It is important to notice that capture occurs predominantly at rest, while decay may occur both in flight and at rest.
+Thus, we can highly mitigate capture and decay at rest by removing pions which would release all their energy in the TPC and stop.
+This translates into a momentum selection, where we keep only events whose WC momentum is above a certain threshold.
+Figure \ref{fig:CaptureMom} shows the true momentum distribution for the primary pions\footnote{We use here the Geant4 denomination ``primary" to indicate that the pion considered does not undergo interactions modifying its energy before getting to the TPC.
+In fact, not every pion shot from wire chamber four will arrive to the TPC as primary,  some will decay or interact before the TPC.}  that arrive to the TPC (pink), that capture (green) or decay (blue) inside the TPC, on a linear and log scale vertical axis.
+
+In order to choose the selection value for the wire chamber momentum, it is beneficial to estimate the ratio of events which capture or decay that survive the selection in MC as a function of the momentum threshold, and compare it with the survival ratio for all events.
+This is done in figure \ref{fig:survRatio}.
+We define the survival ratio simply  as the number of events surviving the true momentum selection divided by the number of events of that category.
+We calculate the survival ratio separately for the three event categories explained above: total (pink), capture (green) and decay (blue).
+Selecting pions with momentum greater than 420 MeV/c reduces the capture events by ~99\% while maintaining about 80\% of the total data sample. 
+Figure \ref{fig:evtRatio} shows the ratio of events which end their life in capture (green) or decay (blue) over the total number of events as a as a function of the true momentum at wire chamber four.
+This ratio is slightly dependent on the inelastic cross section implemented in Geant4, as we are able to register a pion capture (or decay) only if it did not interact inelastically in the TPC.
+We choose a momentum threshold of 420~MeV/c because the percentage of capture events drops below 1\% and the percentage of decays is never above 2\% for momenta greater than 420~MeV/c.
+After the momentum selection, we evaluate the contribution of capture and decay to be a negligibly small background to the cross section measurement compared to the background related to the beamline.
+
+\begin{figure}[p]
+\centering
+\includegraphics[width=15cm]{Chapter-7/Images/CDAsMomentumFunct.png}
+\caption{True momentum distribution at wire chamber 4 for every simulated pion arriving in the TPC (pink), ending its life in capture (green) or in decay (blue) in the TPC, linear vertical axis on the left, logarithmic on the right.
+}
+\label{fig:CaptureMom}
+\end{figure}
+
+\begin{figure}[p]
+\centering
+\begin{minipage}[t]{0.45\textwidth}
+\centering
+\includegraphics[width=7.5cm]{Chapter-7/Images/CDThreshold.png}
+\caption{Survival ratio as a function of selection threshold on true momentum at wire chamber four for for every simulated pion arriving in the TPC (pink), capture (green) or in decay (blue).   }
+\label{fig:survRatio}
+\end{minipage}\hfill
+\begin{minipage}[t]{0.45\textwidth}
+\centering
+\includegraphics[width=7.5cm]{Chapter-7/Images/CDRatio.png}
+\caption{Ratio between the capture (green) and decay (blue) events over the total number of events as a as a function of the true momentum at wire chamber four.}
+\label{fig:evtRatio}
+\end{minipage}
+\end{figure}
+
+\subsection{Beamline Background}\label{ch:PionXSBkgSub2}
+As shown in Section \ref{ch:beamlineComposition}, we use G4Beamline to estimate the beam composition at WC4, obtaining the composition shown in Table \ref{tab:beamline}.
+Once we have estimated the beam composition at WC4, the next step to estimate the beamline background in the cross section measurement by propagating pions, muons and electrons to the TPC and evaluate their contribution to the cross section.
+To do so, we simulate the same number of electrons, muons and pions with the DDMC and we apply the same selection filters on the three samples.
+The number of events per particle species surviving this selection is shown on table \ref{tab:MCafterCutContaminants}.
+
+\begin{table}[]
+\centering
+\begin{tabular}{| l | l | l | l | l | l | l | l | }
+\hline
+ &  \multicolumn{3}{|c|}{Magnet Current -60A} & \multicolumn{3}{|c|}{Magnet Current -100 A}\\
+
+                                                  & MC $\pi^-$   & MC  $ \mu^-$ & MC  $e^-$ & MC  $\pi^-$ & MC  $\mu^-$ & MC  $e^-$  \\
+\hline
+&  &  &  & & &\\  
+Total Initial events                     & 334500  & 334500 & 334500 &344500 &344500& 344500 \\
+After Multiplicity Rejection        & 330668  & 333420 & 198065 &326576 &344208& 201380 \\
+After WC2TPC Selection          & 218239  & 296333 & 91139  &230418 &300228& 98834  \\
+Evts After Shower Rejection     & 208063  & 288914 &  20293 &219882 &293585& 17780  \\
+&  &  &  & & &\\  
+  \hline
+&  &  &  & & &\\  
+Selection Survival Rate           &62.3\% & 86.6\% & 6.1\% & 63.8\%& 85.5\%& 5.2\%\\
+Beam Composition  @WC4      &  68.8\%   &  4.6 \%  & 26.6 \%    & 87.4 \% & 3.7 \%  & 8.9 \% \\ %
+Beam Composition  @TPC FF &  88.5\%   & 8.2\%   & 3.3 \%   & 94.0\%	& 5.3\% & 0.7\%\\
+                                                  &                      &                       &                   &                       &                        &\\  
+\hline
+\end{tabular}
+\caption{MC selection flow per particle species.}
+\label{tab:MCafterCutContaminants}
+\end{table}
+
+\begin{figure}[p]
+\centering
+%\includegraphics[width=\textwidth]{Studies/Figures/Staggered60A.png}
+\caption{Left: staggered contributions to the interacting kinetic energy distribution for electron (grey), muons (yellow) and pion (blue) in the 60A simulation sample.
+Right: staggered contributions to the incident kinetic energy distribution for electron (grey), muons (yellow) and pion (blue) in the 60A simulation sample.
+}
+\label{fig:stag60A}
+\end{figure}
+
+In order to reproduce the closest make up of the beam to data, we weight each event of a given particle species according to the estimated beam composition.
+In case of 60A runs, for example, the weights are 0.688 for pions,  0.046 for muons  and 0.266 for electrons.
+We produce accordingly the interacting and incident histograms for the events surviving the selection, staggering the contributions for each particle species, as shown in Figure  \ref{fig:stag60A}.
+From those histograms, we are able to evaluate the relative contribution of  pions and  background to each bin of the interacting and incident histograms separately and obtain the respective corrections for data.
+We take here the interacting histogram as example, noting that the derivation of the correction for the incident histogram is identical.
+The number of entries in each bin of the interacting plot (Figure \ref{fig:stag60A} left) is  $N^{\text{TOT}}_{\text{Interacting}} (E_{i})$, equal to the sum of the pions and background in that bin, namely
+
+\begin{equation}
+N^{\text{TOT}}_{\text{Interacting}} (E_{i}) =  N^\pi_{Interacting} (E_{i}) + \underbrace{ N^\mu_{Interacting} (E_{i}) + N^e_{Interacting} (E_{i}) + N^{Secondary}_{Interacting} (E_{i}) }_{B_{Interacting} (E_i)}.
+\end{equation}
+Thus, the relative contribution of pions to each bin in MC can be calculated as follows
+\begin{equation}
+C^{\pi MC}_{Interacting} (E_{i}) =  \frac{N^{\pi MC}_{Interacting}}{ N^{TOT MC}_{Interacting} (E_{i}) } =    \frac{N^{TOT MC}_{Interacting} (E_{i}) - B^ {MC}_{Interacting} (E_i)}{ N^{TOT MC}_{Interacting} (E_{i})}.
+\end{equation}
+
+In order to evaluate the pion content of each bin in data, we scale the measured bin content by the corresponding  pion contribution found in MC, as follows
+\begin{equation}
+N^{\pi Reco Data}_{Interacting} = N^{TOT Data}_{Interacting} (E_{i}) - B^{Data}_{Interacting} (E_i)  =  C^{\pi MC}_{Interacting} (E_{i}) N^{TOT Data}_{Interacting} (E_{i}).
+\end{equation}
+
+\section{Estimate of Energy Loss before the TPC}\label{ch:eloss}
 The beamline particles travel a path from where their  momentum is measured in the beamline until they are tracked again inside the TPC.
@@
 
-We derive an estimate of the energy loss between the beamline momentum measurement and the TPC ($E_{loss}$) from the pion DDMC sample, since this quantity is not  measurable directly on data. 
+We derive an estimate of the energy loss between the beamline momentum measurement and the TPC ($E_{loss}$) from the pion and kaon DDMC samples, since this quantity is not  measurable directly on data.
 The $E_{loss}$ distribution for the 60A  and 100A pion sample is shown in figure \ref{fig:ELoss60A}, left and right respectively.
@@
 We classify as ``hitting the halo paddle" all pions whose $P_x$ and $X$ are such that $$P_x +0.02* X - 0.4 < 0 $$ and as ``missing the halo  paddle" all pions whose $P_x$ and $X$ are such that $$P_x +0.02*X - 0.4 > 0, $$ where the coefficients of the line are empirically found by the logistic regression estimation.
-Overall, this simple methode classifies in the right category (hit or miss) about 86\% of the pion events.
+Overall, this simple method classifies in the right category (hit or miss) about 86\% of the pion events.
 In MC, we assign  $E_{loss} = 32 \pm 4 $~MeV for pion events classified as ``hitting the halo paddle"; we assign  $E_{loss} = 24 \pm 3 $~MeV for pion events classified as ``missing the halo paddle".
 We apply the same classifier on data.
-A scan of the simulated geometry showed an excess of 3 cm of un-instrumented argon compared with the surveyed detector geometry.
+A scan of the simulated geometry showed an excess of 3 cm of uninstrumented argon compared with the surveyed detector geometry.
 We account for this difference by assigning in data $E_{loss} = 24 \pm 6 $~MeV for pion events classified as ``hitting the halo paddle" and  $E_{loss} = 17 \pm 6 $~MeV for pion events classified as ``missing the halo paddle", where the uncertainty is derived as the standard deviation of the double peaked distribution.

### 2018-05-21 Chapter-6/6-PionCrossSection.tex
s

@@
 In Section \ref{ch:PionXSRaw}, we start by measuring the raw cross section, i.e.
-the cross section obtained exclusively using data reconstruction, with no additional corrections.
-In Section \ref{ch:PionXSCorrections}, we apply a series of data manipulations geared toward the correction for the background contributions and thse correction for detection inefficiency, showing the final results in Section \ref{ch:FinalPion}.
+the cross section obtained exclusively using data reconstruction, without any additional corrections.
+In Section \ref{ch:PionXSCorrections}, we apply a statistical subtraction of the background contributions based on simulation and a correction for detection inefficiency.
+The final results are presented in Section \ref{ch:FinalPion}.
 
@@
 \begin{eqnarray}
-\delta KE_{j} &=& \sqrt{\delta p_{Beam}^2 + \delta E_{Loss}^2 +  \delta  E_{\text{dep FF-j}}^2},\\
-&=& \sqrt{(2\% \text{ }p_{Beam})^2 +  (\sim 7 \text{ [MeV]})^2 +  (j-1)^2 (\sim0.08\text{ [MeV]})^2}.
+\delta KE_{j} &=& \sqrt{\delta p_{Beam}^2 + \delta E_{Loss}^2 +  \delta  E_{\text{dep FF-j}}^2}\\
+&=& \sqrt{(2\% \text{ }p_{Beam})^2 +  (\sim 6 \text{ [MeV]})^2 +  (j-1)^2 (\sim0.08\text{ [MeV]})^2}.
 \end{eqnarray}
@@
 We propagate this uncertainty  by varying the energy measurement $KE_{j}$ at each argon slab.
-We calculating the $N^{\text{TOT}}_{\text{Inc}}$,  $N^{\text{TOT}}_{\text{Int}}$ and cross section plots in three cases: first assigning the measured $KE_{j}$ at each kinetic energy sampling, then assigning $KE_{j} + \delta KE_{j}$, and finally assigning $KE_{j} - \delta KE_{j}$.
+We measure $N^{\text{TOT}}_{\text{Inc}}$,  $N^{\text{TOT}}_{\text{Int}}$ and the cross section  in three cases: first assigning the measured $KE_{j}$ at each kinetic energy sampling, then assigning $KE_{j} + \delta KE_{j}$, and finally assigning $KE_{j} - \delta KE_{j}$.
 The difference between the values obtained using the $KE_{j}$ sampling and the maximum and minimum values in each kinetic energy bin determines the systematic uncertainty.
@@
 As described in section \ref{ch:MCCorrections} as series of corrections are needed to derive the true pion cross section from the raw cross section. 
-These corrections are described in equation \ref{eq:C},
+The corrected cross section is given in equation \ref{eq:C},
 
@@
 
+Section \label{ch:BKGsubXS} describes the procedure employed to obtain  
+$C^{\pi MC}_{\text{Int}} (E_{i})$  and  $C^{\pi MC}_{\text{Inc}} (E_{i})$ and the propagation to the cross section measurement of the relative uncertainties.
+
+Section \label{ch:ch:EFFXS} describes the procedure employed to obtain  
+$\epsilon^{\text{Int}}(E_i)$  and $\epsilon^{\text{Inc}}(E_i)$ and the propagation to the cross section measurement of the relative uncertainties.
+
 \subsection{Background subtraction}\label{ch:BKGsubXS}
+Even if pions are by far the biggest component of the beam in negative polarity runs, the LArIAT beam is not a pure pion beam.
+While useful to discriminate  pions/muons/electrons from kaons, and protons, the beamline detectors are not sensitive enough to  discriminate among the lighter particles in the beam: electrons, muons and pions fall under the same mass hypothesis.
+Thus, we need to assess the background from beamline particles other than pions in the event selections used for the pion cross section analysis and correct for its effects.
+
+\subsubsection{Beam Composition}\label{sec:BeamAtWC4}
+We define beamline background every TPC track matched to the WC track which is not a primary pion. Potentially, there are 4 different types of beamline background:
+\begin{itemize}
+\item[]1) electrons,
+\item[]2) muons,
+\item[]3) secondaries from pion events,
+\item[]4) matched pile up events.
+\end{itemize}
+
+The first step is to estimate what percentage of events used in the cross section calculation is not a primary pion.
+The next two sections will illustrate this estimate for the electrons, muons and secondaries from pion event.
+We estimate the last type of background, the ``matched pile up" events, to be a negligible fraction, because of the definition of the WC2TPC match: we deem the probability of a single match with a halo particle in the absence of a beamline particle\footnote{ Events with multiple WC2TPC matches are always rejected.} negligibly small.
+\textcolor{red}{SHOW VTX distribution in WC2TPC match}
+
+\subsubsection{Background from Beamline Electrons and Muons}\label{stionEMu}
+\begin{figure}[b]
+%\includegraphics[width=0.5\textwidth,height=\textheight,keepaspectratio]{Studies/Figures//Beam60A.png}
+%\includegraphics[width=0.5\textwidth,height=\textheight,keepaspectratio]{Studies/Figures//Beam100A.png}
+\caption{Beam composition for the -60A runs (left) and -100A runs (right).
+The solid blue plot represents the simulated pion content, the yellow plot represents the simulated muon content and the grey plot represents the simulated electron content.
+The plots are area normalized to the number of data events, shown in red.
+}
+\label{fig:BeamComposition}
+\end{figure}
+
+We estimate the percentage of electrons and muons in the beam via the G4Beamline MC. 
+Since the beamline composition is a function of the magnet settings, we simulate separately events for magnet current of -60A and -100A. 
+Table \ref{tab:beamline} shows the beam composition per magnet setting after the mass selection according to the G4Beamline simulation.
+\begin{table}[p]
+\centering
+\begin{tabular}{|l|c|c|}
+\hline
+                     & I = -60 A           & I = -100 A \\ \hline
+G4Pions       &   68.8 \%           &      87.4 \%        \\ \hline
+G4Muons     &     4.6 \%           &        3.7 \%         \\ \hline
+G4Electrons &   26.6 \%           &        8.9 \%        \\ \hline
+\end{tabular}
+\caption{Simulated beamline composition per magnet settings}
+\label{tab:beamline}
+\end{table}
+
+Figure \ref{fig:BeamComposition} shows the momentum predictions from G4Beamline overlaid with data for the 60A runs (left) and for the 100A runs (right).
+The predictions for electrons, muons and pions have been staggered and their sum is area normalized to data, which is shown in red.
+Albeit not perfect, these plots show a reasonable agreement between the momentum shapes in data and MC.
+We attribute  the difference in shape to the lack of simulation of the WC efficiency in the MC which is momentum dependent and leads to enhance the number events in the center of the momentum distribution.
+
+Once the beam composition at WC4 is know,  we simulate the electrons, muons and pions with the DDMC and we subject the three samples to the same selection chain (WC2TPC match, shower filter, pile up filter).
+The percentage of electrons and muons surviving the selection chain weighted by the beam composition is the  electron and muon background in the pion cross section sample, as shown in Table \ref{tab:MCafterCutContaminants}.
+
+\subsubsection{Background from secondaries at TPC Front Face}
+Pions can travel the length of the LArIAT beamline and interact hadronically in the steel or in the non-instrumented argon upstream to the TPC front face.
+Or, they could decay in flight between WC4 and the TPC.
+One of the interaction products can leak into the TPC and be matched with the WC track, contributing to the pool of events used for the cross section calculation.
+We call this type of particles ``secondaries" from pion events, with a terminology inspired by Geant4.
+We estimate the number of secondaries using the DDMC pion sample.
+The percentage of secondaries is given by the number of matched WC2TPC tracks whose corresponding particle is not flagged as primary by Geant4.
+The secondary to pion ratio is 4.9\% in the 60A sample and $Y$\% in the 100A sample.
+
+\subsection{Background Contribution to the Cross Section}\label{sec:Correction}
+
+Figures \ref{fig:CorrectionsBeam} show $C^{\pi MC}_{Interacting} (E_{i})$ and $C^{\pi MC}_{Incident} (E_{i})$ as a function of the kinetic energy for the 60A runs and their systematic uncertainty.
+We take a 100\% systematic uncertainty on the muon and electron content: we calculate the extreme values of  $C^{\pi MC}_{Interacting} (E_{i})$ and $C^{\pi MC}_{Incident} (E_{i})$ in each bin changing the beam composition for the configurations listed in Table \ref{tab:beamlineSys}.
+
+\begin{table}[p]
+\centering
+\begin{tabular}{| l | l | l | l | l | l | l | l | }
+\hline
+ &  \multicolumn{3}{|c|}{Magnet Current -60A} & \multicolumn{3}{|c|}{Magnet Current -100 A}\\
+
+                                                  & MC $\pi^-$   & MC  $ \mu^-$ & MC  $e^-$ & MC  $\pi^-$ & MC  $\mu^-$ & MC  $e^-$  \\
+\hline
+Composition 2x muons          & 64.2	\%&9.2 \%&	26.6 \%&	83.7 \%&	7.4	\%&8.9 \% \\
+Composition 0.5x muons       &71.1	\%&2.3 \%&	26.6 \%&	89.2 \%&	1.9	\%&8.9 \% \\
+Composition 2x Electrons      &42.2	\%&4.6 \%&	53.2 \%&	78.5	\%&  3.7	\%&17.8 \%\\
+Composition 0.5x Electrons   &82.1	\%&4.6 \%&	13.3 \%&	91.9 \%&	3.7	\%&4.4 \% \\
+\hline
+\end{tabular}
+\caption{Beam composition variation for the study of systematics due to beam contamination.}
+\label{tab:beamlineSys}
+\end{table}
+
+\begin{figure}[p]
+\centering
+%\includegraphics[width=\textwidth]{Studies/Figures/c60A_Multiple.png}
+\caption{Left: relative pion content for interacting histogram a function of kinetic energy for the 60A runs, statistics uncertainty in azure and systematic uncertainty in blue.
+Right: relative pion content  for incident histogram a function of kinetic energy for the 60A runs, statistics uncertainty in azure and systematic uncertainty in blue.}
+\label{fig:CorrectionsBeam}
+\end{figure}
 
@@
 
-\subsection{Efficiency Correction}
+\subsection{Efficiency Correction}\label{ch:EFFXS}
+The interaction point for a track used in the total hadronic cross section analysis is defined to be the last point of the WC2TPC matched track which lies inside the fiducial volume.
+This definition is independent from the topology of the interaction.
+If the TPC track stops within the fiducial volume, its last point will be the interaction point, no matter what the products of the interaction look like; if the track crosses the boundaries of the fiducial volume, the track will be considered ``through going" and no interaction point will be found.
+Given this definition, it is evident that we rely on the tracking algorithm to discern where the interaction occurred in the TPC  and correctly stop the tracking.
+The tracking algorithm has an intrinsic angle resolution as shown in section \ref{sec:angleRes}, which limits its efficiency, especially in the case of elastic scattering occurring a low angles.
+Thus, we need to apply an efficiency correction to data in order to retrieve the true cross section.
+The efficiency correction is evaluated separately for the interacting and incident histograms, namely $\epsilon^{int}_i$ and  $\epsilon^{inc}_i$, and propagated to the cross section as shown in  equation \ref{eq:True}.
+
+\subsubsection{Efficiency Correction: Procedure}\label{sec:EffCorrection}
+We describe here the procedure to calculate the efficiency correction taking the interacting histogram as example and noting that the procedure is identical for the  incident histogram.
+
+We derive the correction on a set of pure pion MC, calculating its value bin by bin as the ratio between the true bin content and the correspondent reconstructed bin content.
+The correction is then applied to the relevant bin in data.
+In formulae, the efficiency correction is calculated to be
+
+\begin{equation}
+\epsilon^{\text{int}}_i  =  \frac{N^{\text{ $\pi$ Reco MC}}_{\text{Interacting}} (E_{i})}{ N^{\text{ $\pi$ True MC}}_{\text{Interacting}} (E_{i})  },
+\end{equation}
+
+where $N^{\text{ $\pi$ True MC}}_{\text{Interacting}} (E_{i}) $ is the content of the $i$-th bin in for the true interacting histogram, and $N^{\text{ $\pi$ Reco MC}}_{\text{Interacting}} (E_{i}) $ is the content of the $i$-th bin in for the reconstructed interacting histogram.
+The correction is applied to data as follows
+
+\begin{equation}
+N^{\text{ $\pi$ True Data}}_{\text{Interacting}} (E_{i})  =  \frac{N^{\text{$\pi$ Reco Data}}_{\text{Interacting}} (E_{i})}{\epsilon^{\text{int}}_i} = N^{\text{$\pi$ Reco Data}}_{\text{Interacting}} (E_{i}) \frac{N^{\text{ $\pi$ True MC}}_{\text{Interacting}} (E_{i})}{ N^{\text{ $\pi$ Reco MC}}_{\text{Interacting}} (E_{i})}.
+\end{equation}
+
+where $N^{\text{$\pi$ Reco Data}}_{\text{Interacting}} (E_{i})$ is the background subtracted bin content of the $i$-th bin in for the reconstructed interacting histogram for data, i.e. 
+\begin{equation}
+N^{\text{$\pi$ Reco Data}}_{\text{Interacting}} (E_{i}) =  N^{\text{TOT Data}}_{\text{Interacting}} (E_{i}) - B^{\text{Data}}_{\text{Interacting}} (E_i)  =  C^{\text{$\pi$ MC}}_{\text{Interacting}} (E_{i}) N^{\text{TOT Data}}_{\text{Interacting}} (E_{i}).
+\end{equation}
+
+Figures \ref{fig:EffCorr} show $\epsilon^{\text{int}}_i(E_{i})$ and $\epsilon^{\text{int}}_i (E_{i})$ as a function of the kinetic energy for the 60A runs and their systematic uncertainty.
+
+%############################# Systematics
+
+In section \ref{sec:angleRes}, we estimated the angular resolution for data and MC to be $\bar\alpha_{Data} = (5.0 \pm 4.5) \text{ deg}$  and 
+$\bar\alpha_{MC} = (4.5 \pm 3.9) \text{ deg}$, respectively.
+Interaction angles smaller than the angular resolution are indistinguishable for the reconstruction.
+Thus, we claim we are able to  measure the cross section for interaction angles greater than 5.0 deg.
+Geant4 simulates interactions at all angles, as shown in figure \ref{trueScatteringAngle}.
+In order to calculate the efficiency correction,  we select events which have an interaction angle greater than a given $\alpha_res$ to construct the true interacting and incident histograms (the denominator of the efficiency correction).
+
+The systematics on the efficiency correction is estimated by varying the value of $\alpha_{res}$ to be
+
 \subsubsection{Treatment of Systematics}
 
-\section{Final Plots}\label{ch:FinalPion}
+\clearpage
+\section{Results}\label{ch:FinalPion}
+Figure \ref{fig:FinalXSPion} show the measurement of the ($\pi^-$-Ar) total hadronic cross section for  scattering angles greater than 5$^\circ$, as the result of the background subtraction and efficiency correction to the raw cross section.
+The top left plot is the measurement obtained on the 60A data, statistical uncertainty in black and systematic uncertainty in red.
+The top right plot is the measurement obtained on the 100A data, statistical uncertainty in black and systematic uncertainty in blue.
+The bottom plot shows the two measurements overlaid.
+In all three plot, the Geant4 prediction for the total hadronic cross section for angle scattering greater than 5$^\circ$ is displayed in green.
 
@@
 \includegraphics[width=0.48\textwidth]{Chapter-6/Images/TheMoneyPlot100A.pdf}
-\caption{.}
+\includegraphics[width=0.48\textwidth]{Chapter-6/Images/TheMoneyPlot.pdf}
+\caption{ \emph{Top Left:} ($\pi^-$-Ar) total hadronic cross section for  scattering angles greater than 5$^\circ$ measured in the 60A sample, statistical uncertainty in black and systematic uncertainty in red.
+The Geant4 prediction for the total hadronic cross section for angle scattering greater than 5$^\circ$ is displayed in green.
+\\
+\emph{Top Right:} ($\pi^-$-Ar) total hadronic cross section for  scattering angles greater than 5$^\circ$ measured in the 100A sample, statistical uncertainty in black and systematic uncertainty in blue.
+The Geant4 prediction for the total hadronic cross section for angle scattering greater than 5$^\circ$ is displayed in green.\\
+\emph{Bottom:} ($\pi^-$-Ar) total hadronic cross section measurements in the 60A and 100A samples overlaid with the  Geant4 prediction (green).}
 \label{fig:FinalXSPion}
 \end{figure}
-
-\begin{figure}[htb]
-\centering
-\includegraphics[width=0.48\textwidth]{Chapter-6/Images/TheMoneyPlot.pdf}
-\caption{.}
-\label{fig:FinalXSPion}
-\end{figure}

### 2018-05-21 Chapter-5/5-Preparatory.tex
s

@@
 \section{Estimate of Backgrounds in the Pion Cross Section}\label{ch:PionXSBkgSub}
+
 We use the beamline simulation and the DDMC simulation to estimate the background in the total hadronic pion cross section.
@@
 
+\clearpage
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Beamline Background}\label{ch:PionXSBkgSub2}
+We define beamline background every TPC track matched to the WC track which is not a primary pion. Potentially, there are 4 different types of beamline background:
+\begin{itemize}
+\item[]1) electrons,
+\item[]2) muons,
+\item[]3) secondaries from pion events,
+\item[]4) matched pile up events.
+\end{itemize}
+
+The first step is to estimate what percentage of events used in the cross section calculation is not a primary pion.
+We start by noting that the last type of background, the ``matched pile up" events, is a negligible fraction, because of the definition of the WC2TPC match: we deem the probability of a single match with a halo particle in the absence of a beamline particle\footnote{ Events with multiple WC2TPC matches are always rejected.} negligibly small.
+%\textcolor{red}{SHOW VTX distribution in WC2TPC match}
+
 As shown in Section \ref{ch:beamlineComposition}, we use G4Beamline to estimate the beam composition at WC4, obtaining the composition shown in Table \ref{tab:beamline}.
-Once we have estimated the beam composition at WC4, the next step to estimate the beamline background in the cross section measurement by propagating pions, muons and electrons to the TPC and evaluate their contribution to the cross section.
+The next step to estimate the beamline background in the cross section measurement is propagating pions, muons and electrons to the TPC and evaluate their contribution to the cross section.
 To do so, we simulate the same number of electrons, muons and pions with the DDMC and we apply the same selection filters on the three samples.
 The number of events per particle species surviving this selection is shown on table \ref{tab:MCafterCutContaminants}.
+
+Pions can travel the length of the LArIAT beamline and interact hadronically in the steel or in the non-instrumented argon upstream to the TPC front face.
+Or, they could decay in flight between WC4 and the TPC.
+One of the interaction products can leak into the TPC and be matched with the WC track, contributing to the pool of events used for the cross section calculation.
+We call this type of particles ``secondaries" from pion events, with a terminology inspired by Geant4.
+We estimate the number of secondaries using the DDMC pion sample.
+The percentage of secondaries is given by the number of matched WC2TPC tracks whose corresponding particle is not flagged as primary by Geant4.
+The secondary to pion ratio is 4.9\% in the 60A sample and 4.3\% in the 100A sample.
 
@@
 \centering
-%\includegraphics[width=\textwidth]{Studies/Figures/Staggered60A.png}
+\includegraphics[width=\textwidth]{Chapter-5/Images/Background60A.pdf}
 \caption{Left: staggered contributions to the interacting kinetic energy distribution for electron (grey), muons (yellow) and pion (blue) in the 60A simulation sample.

### 2018-05-21 Chapter-6/6-PionCrossSection.tex
s

@@
 \subsection{Background subtraction}\label{ch:BKGsubXS}
-Even if pions are by far the biggest component of the beam in negative polarity runs, the LArIAT beam is not a pure pion beam.
+%Even if pions are by far the biggest component of the beam in negative polarity runs, the LArIAT beam is not a pure pion beam.
 While useful to discriminate  pions/muons/electrons from kaons, and protons, the beamline detectors are not sensitive enough to  discriminate among the lighter particles in the beam: electrons, muons and pions fall under the same mass hypothesis.
@@
 
-\subsubsection{Beam Composition}\label{sec:BeamAtWC4}
-We define beamline background every TPC track matched to the WC track which is not a primary pion. Potentially, there are 4 different types of beamline background:
-\begin{itemize}
-\item[]1) electrons,
-\item[]2) muons,
-\item[]3) secondaries from pion events,
-\item[]4) matched pile up events.
-\end{itemize}
+%\subsubsection{Beam Composition}\label{sec:BeamAtWC4}
+%We define beamline background every TPC track matched to the WC track which is not a primary pion. Potentially, there are 4 different types of beamline background:
+%\begin{itemize}
+%\item[]1) electrons,
+%\item[]2) muons,
+%\item[]3) secondaries from pion events,
+%\item[]4) matched pile up events.
+%\end{itemize}
 
-The first step is to estimate what percentage of events used in the cross section calculation is not a primary pion.
+%The first step is to estimate what percentage of events used in the cross section calculation is not a primary pion.
 The next two sections will illustrate this estimate for the electrons, muons and secondaries from pion event.
-We estimate the last type of background, the ``matched pile up" events, to be a negligible fraction, because of the definition of the WC2TPC match: we deem the probability of a single match with a halo particle in the absence of a beamline particle\footnote{ Events with multiple WC2TPC matches are always rejected.} negligibly small.
+%We estimate the last type of background, the ``matched pile up" events, to be a negligible fraction, because of the definition of the WC2TPC match: we deem the probability of a single match with a halo particle in the absence of a beamline particle\footnote{ Events with multiple WC2TPC matches are always rejected.} negligibly small.
 \textcolor{red}{SHOW VTX distribution in WC2TPC match}
 
-\subsubsection{Background from Beamline Electrons and Muons}\label{stionEMu}
-\begin{figure}[b]
+%\subsubsection{Background from Beamline Electrons and Muons}\label{stionEMu}
+%\begin{figure}[b]
 %\includegraphics[width=0.5\textwidth,height=\textheight,keepaspectratio]{Studies/Figures//Beam60A.png}
 %\includegraphics[width=0.5\textwidth,height=\textheight,keepaspectratio]{Studies/Figures//Beam100A.png}
-\caption{Beam composition for the -60A runs (left) and -100A runs (right).
+%\caption{Beam composition for the -60A runs (left) and -100A runs (right).
 The solid blue plot represents the simulated pion content, the yellow plot represents the simulated muon content and the grey plot represents the simulated electron content.
@@
 }
-\label{fig:BeamComposition}
-\end{figure}
+%\label{fig:BeamComposition}
+%\end{figure}
 
-We estimate the percentage of electrons and muons in the beam via the G4Beamline MC. 
-Since the beamline composition is a function of the magnet settings, we simulate separately events for magnet current of -60A and -100A. 
-Table \ref{tab:beamline} shows the beam composition per magnet setting after the mass selection according to the G4Beamline simulation.
-\begin{table}[p]
-\centering
-\begin{tabular}{|l|c|c|}
-\hline
-                     & I = -60 A           & I = -100 A \\ \hline
-G4Pions       &   68.8 \%           &      87.4 \%        \\ \hline
-G4Muons     &     4.6 \%           &        3.7 \%         \\ \hline
-G4Electrons &   26.6 \%           &        8.9 \%        \\ \hline
-\end{tabular}
-\caption{Simulated beamline composition per magnet settings}
-\label{tab:beamline}
-\end{table}
+%We estimate the percentage of electrons and muons in the beam via the G4Beamline MC. 
+%Since the beamline composition is a function of the magnet settings, we simulate separately events for magnet current of -60A and -100A. 
+%Table \ref{tab:beamline} shows the beam composition per magnet setting after the mass selection according to the G4Beamline simulation.
+%\begin{table}[p]
+%\centering
+%\begin{tabular}{|l|c|c|}
+%\hline
+ %                   & I = -60 A           & I = -100 A \\ \hline
+%G4Pions       &   68.8 \%           &      87.4 \%        \\ \hline
+%G4Muons     &     4.6 \%           &        3.7 \%         \\ \hline
+%G4Electrons &   26.6 \%           &        8.9 \%        \\ \hline
+%\end{tabular}
+%\caption{Simulated beamline composition per magnet settings}
+%\label{tab:beamline}
+%\end{table}
 
-Figure \ref{fig:BeamComposition} shows the momentum predictions from G4Beamline overlaid with data for the 60A runs (left) and for the 100A runs (right).
+%Figure \ref{fig:BeamComposition} shows the momentum predictions from G4Beamline overlaid with data for the 60A runs (left) and for the 100A runs (right).
 The predictions for electrons, muons and pions have been staggered and their sum is area normalized to data, which is shown in red.
@@
 
-Once the beam composition at WC4 is know,  we simulate the electrons, muons and pions with the DDMC and we subject the three samples to the same selection chain (WC2TPC match, shower filter, pile up filter).
+%Once the beam composition at WC4 is know,  we simulate the electrons, muons and pions with the DDMC and we subject the three samples to the same selection chain (WC2TPC match, shower filter, pile up filter).
 The percentage of electrons and muons surviving the selection chain weighted by the beam composition is the  electron and muon background in the pion cross section sample, as shown in Table \ref{tab:MCafterCutContaminants}.
 
-\subsubsection{Background from secondaries at TPC Front Face}
-Pions can travel the length of the LArIAT beamline and interact hadronically in the steel or in the non-instrumented argon upstream to the TPC front face.
+%\subsubsection{Background from secondaries at TPC Front Face}
+%Pions can travel the length of the LArIAT beamline and interact hadronically in the steel or in the non-instrumented argon upstream to the TPC front face.
 Or, they could decay in flight between WC4 and the TPC.

### 2018-05-21 Chapter-4/4-Methodology.tex
s

@@
 
-As we will show in section \ref{sec:Correction}, the background subtraction for the interacting and incident histograms can be translated into a corresponding relative pion content $C^{\pi MC}_{Interacting} (E_{i})$ and $C^{\pi MC}_{Incident} (E_{i})$ and the cross section re-written as follows
+As we will show in section \ref{ch:PionXSBkgSub}, the background subtraction for the interacting and incident histograms can be translated into a corresponding relative pion content $C^{\pi MC}_{Interacting} (E_{i})$ and $C^{\pi MC}_{Incident} (E_{i})$ and the cross section re-written as follows
 

### 2018-05-21 Chapter-5/5-Preparatory.tex
s

@@
 
-\begin{figure}[p]
+\begin{figure}[]
 \centering
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Beamline Background}\label{ch:PionXSBkgSub2}
+\subsection{Contributions from the Beamline Background}\label{ch:PionXSBkgSub2}
 We define beamline background every TPC track matched to the WC track which is not a primary pion. Potentially, there are 4 different types of beamline background:
@@
 
-The first step is to estimate what percentage of events used in the cross section calculation is not a primary pion.
+The first step to quantify the effect of the beamline background on the pion cross section is to estimate what percentage of events used in the cross section calculation is not a primary pion.
 We start by noting that the last type of background, the ``matched pile up" events, is a negligible fraction, because of the definition of the WC2TPC match: we deem the probability of a single match with a halo particle in the absence of a beamline particle\footnote{ Events with multiple WC2TPC matches are always rejected.} negligibly small.
 %\textcolor{red}{SHOW VTX distribution in WC2TPC match}
+As shown in Section \ref{ch:beamlineComposition}, we use G4Beamline to estimate the percentage of pions, muons and electrons at WC4, obtaining the composition shown in Table \ref{tab:beamline}.
+The next step is to simulate those pions, muons and electrons from WC4 to the TPC with the DDMC and evaluate their contribution to the cross section.
+To do so, we start by simulating the same number of electrons, muons and pions with the DDMC and we apply the same selection filters on the three samples.
+The number of events per particle species surviving this selection is shown on table \ref{tab:MCafterCutContaminants}.
+In order to reproduce the closest make up of the beam to data, we weight each event of a given particle species according to the estimated beam composition.
+In case of 60A runs, for example, the weights are 0.688 for pions,  0.046 for muons  and 0.266 for electrons.
 
-As shown in Section \ref{ch:beamlineComposition}, we use G4Beamline to estimate the beam composition at WC4, obtaining the composition shown in Table \ref{tab:beamline}.
-The next step to estimate the beamline background in the cross section measurement is propagating pions, muons and electrons to the TPC and evaluate their contribution to the cross section.
-To do so, we simulate the same number of electrons, muons and pions with the DDMC and we apply the same selection filters on the three samples.
-The number of events per particle species surviving this selection is shown on table \ref{tab:MCafterCutContaminants}.
-
-Pions can travel the length of the LArIAT beamline and interact hadronically in the steel or in the non-instrumented argon upstream to the TPC front face.
+It should be noted that pions may  interact hadronically in the steel or in the non-instrumented argon upstream to the TPC front face while travelling the length of between WC4 and the TPC.
 Or, they could decay in flight between WC4 and the TPC.
@@
 
-In order to reproduce the closest make up of the beam to data, we weight each event of a given particle species according to the estimated beam composition.
-In case of 60A runs, for example, the weights are 0.688 for pions,  0.046 for muons  and 0.266 for electrons.
-We produce accordingly the interacting and incident histograms for the events surviving the selection, staggering the contributions for each particle species, as shown in Figure  \ref{fig:stag60A}.
-From those histograms, we are able to evaluate the relative contribution of  pions and  background to each bin of the interacting and incident histograms separately and obtain the respective corrections for data.
+We evaluate the beamline background contribution to the cross section by producing the interacting and incident histograms for the events surviving the selection, staggering the contributions for each particle species, as shown in Figure  \ref{fig:stag60A}.
+From those histograms, we are able to evaluate the contribution of  pions and  beamline backgrounds to each bin of the interacting and incident histograms separately and obtain the relative pion content.
+The relative pion content in each bin for the interacting and incident histograms represents the correction applied to data.
 We take here the interacting histogram as example, noting that the derivation of the correction for the incident histogram is identical.
-The number of entries in each bin of the interacting plot (Figure \ref{fig:stag60A} left) is  $N^{\text{TOT}}_{\text{Interacting}} (E_{i})$, equal to the sum of the pions and background in that bin, namely
+The number of entries in each bin of the interacting plot (Figure \ref{fig:stag60A} left) is  $N^{\text{TOT}}_{\text{Int}} (E_{i})$, equal to the sum of the pions and beamline backgrounds in that bin, namely
 
 \begin{equation}
-N^{\text{TOT}}_{\text{Interacting}} (E_{i}) =  N^\pi_{Interacting} (E_{i}) + \underbrace{ N^\mu_{Interacting} (E_{i}) + N^e_{Interacting} (E_{i}) + N^{Secondary}_{Interacting} (E_{i}) }_{B_{Interacting} (E_i)}.
+N^{\text{TOT}}_{\text{Int}} (E_{i}) =  N^\pi_{\text{Int}} (E_{i}) + \underbrace{ N^\mu_{\text{Int}} (E_{i}) + N^e_{\text{Int}} (E_{i}) + N^{Secondary}_{\text{Int}} (E_{i}) }_{B_{\text{Int}} (E_i)}.
 \end{equation}
-Thus, the relative contribution of pions to each bin in MC can be calculated as follows
+Thus, the relative pion content to each bin in MC can be calculated as follows
 \begin{equation}
-C^{\pi MC}_{Interacting} (E_{i}) =  \frac{N^{\pi MC}_{Interacting}}{ N^{TOT MC}_{Interacting} (E_{i}) } =    \frac{N^{TOT MC}_{Interacting} (E_{i}) - B^ {MC}_{Interacting} (E_i)}{ N^{TOT MC}_{Interacting} (E_{i})}.
+C^{\pi MC}_{\text{Int}} (E_{i}) =  \frac{N^{\pi MC}_{\text{Int}}}{ N^{TOT MC}_{\text{Int}} (E_{i}) } =    \frac{N^{TOT MC}_{\text{Int}} (E_{i}) - B^ {MC}_{\text{Int}} (E_i)}{ N^{TOT MC}_{\text{Int}} (E_{i})}.
 \end{equation}
 
-In order to evaluate the pion content of each bin in data, we scale the measured bin content by the corresponding  pion contribution found in MC, as follows
+In order to evaluate the pion content of each bin in data, we scale the measured bin by the corresponding relative pion content found in MC, as follows
 \begin{equation}
-N^{\pi Reco Data}_{Interacting} = N^{TOT Data}_{Interacting} (E_{i}) - B^{Data}_{Interacting} (E_i)  =  C^{\pi MC}_{Interacting} (E_{i}) N^{TOT Data}_{Interacting} (E_{i}).
+N^{\pi Reco Data}_{\text{Int}} = N^{TOT Data}_{\text{Int}} (E_{i}) - B^{Data}_{\text{Int}} (E_i)  =  C^{\pi MC}_{\text{Int}} (E_{i}) N^{TOT Data}_{\text{Int}} (E_{i}).
 \end{equation}
+
+The pion content is evaluated separately in the interacting and incident histograms. Their ratio determines a correction to the measured raw cross section. 
+For example, the measured raw  cross section of a sample with enhanced muons content will tend to be lower than the raw cross section of a muon free sample.
+This is because most of the muons will cross the TPC without stopping, thus contributing almost exclusively to the incident histogram, forcing the pion content to be lower in the incident histogram than in the interacting; thus, the correction will tend to enhance the cross section.
 

### 2018-05-21 Chapter-6/6-PionCrossSection.tex
s

@@
 
-Section \label{ch:BKGsubXS} describes the procedure employed to obtain  
-$C^{\pi MC}_{\text{Int}} (E_{i})$  and  $C^{\pi MC}_{\text{Inc}} (E_{i})$ and the propagation to the cross section measurement of the relative uncertainties.
+Section \ref{ch:BKGsubXSPuppa} describes the evaluation of pion content in the interacting and incident histograms, ($C^{\pi MC}_{\text{Int}} (E_{i})$  and  $C^{\pi MC}_{\text{Inc}} (E_{i})$) and the propagation to the cross section measurement of the relative systematic uncertainties.
 
-Section \label{ch:ch:EFFXS} describes the procedure employed to obtain  
-$\epsilon^{\text{Int}}(E_i)$  and $\epsilon^{\text{Inc}}(E_i)$ and the propagation to the cross section measurement of the relative uncertainties.
+Section \ref{ch:EFFXS} describes the procedure employed to obtain  the efficiency corrections $\epsilon^{\text{Int}}(E_i)$  and $\epsilon^{\text{Inc}}(E_i)$ and the propagation to the cross section measurement of the relative uncertainties.
 
-\subsection{Background subtraction}\label{ch:BKGsubXS}
-%Even if pions are by far the biggest component of the beam in negative polarity runs, the LArIAT beam is not a pure pion beam.
-While useful to discriminate  pions/muons/electrons from kaons, and protons, the beamline detectors are not sensitive enough to  discriminate among the lighter particles in the beam: electrons, muons and pions fall under the same mass hypothesis.
-Thus, we need to assess the background from beamline particles other than pions in the event selections used for the pion cross section analysis and correct for its effects.
-
-%\subsubsection{Beam Composition}\label{sec:BeamAtWC4}
-%We define beamline background every TPC track matched to the WC track which is not a primary pion. Potentially, there are 4 different types of beamline background:
-%\begin{itemize}
-%\item[]1) electrons,
-%\item[]2) muons,
-%\item[]3) secondaries from pion events,
-%\item[]4) matched pile up events.
-%\end{itemize}
-
-%The first step is to estimate what percentage of events used in the cross section calculation is not a primary pion.
-The next two sections will illustrate this estimate for the electrons, muons and secondaries from pion event.
-%We estimate the last type of background, the ``matched pile up" events, to be a negligible fraction, because of the definition of the WC2TPC match: we deem the probability of a single match with a halo particle in the absence of a beamline particle\footnote{ Events with multiple WC2TPC matches are always rejected.} negligibly small.
-\textcolor{red}{SHOW VTX distribution in WC2TPC match}
-
-%\subsubsection{Background from Beamline Electrons and Muons}\label{stionEMu}
-%\begin{figure}[b]
-%\includegraphics[width=0.5\textwidth,height=\textheight,keepaspectratio]{Studies/Figures//Beam60A.png}
-%\includegraphics[width=0.5\textwidth,height=\textheight,keepaspectratio]{Studies/Figures//Beam100A.png}
-%\caption{Beam composition for the -60A runs (left) and -100A runs (right).
-The solid blue plot represents the simulated pion content, the yellow plot represents the simulated muon content and the grey plot represents the simulated electron content.
-The plots are area normalized to the number of data events, shown in red.
-}
-%\label{fig:BeamComposition}
-%\end{figure}
-
-%We estimate the percentage of electrons and muons in the beam via the G4Beamline MC. 
-%Since the beamline composition is a function of the magnet settings, we simulate separately events for magnet current of -60A and -100A. 
-%Table \ref{tab:beamline} shows the beam composition per magnet setting after the mass selection according to the G4Beamline simulation.
-%\begin{table}[p]
-%\centering
-%\begin{tabular}{|l|c|c|}
-%\hline
- %                   & I = -60 A           & I = -100 A \\ \hline
-%G4Pions       &   68.8 \%           &      87.4 \%        \\ \hline
-%G4Muons     &     4.6 \%           &        3.7 \%         \\ \hline
-%G4Electrons &   26.6 \%           &        8.9 \%        \\ \hline
-%\end{tabular}
-%\caption{Simulated beamline composition per magnet settings}
-%\label{tab:beamline}
-%\end{table}
-
-%Figure \ref{fig:BeamComposition} shows the momentum predictions from G4Beamline overlaid with data for the 60A runs (left) and for the 100A runs (right).
-The predictions for electrons, muons and pions have been staggered and their sum is area normalized to data, which is shown in red.
-Albeit not perfect, these plots show a reasonable agreement between the momentum shapes in data and MC.
-We attribute  the difference in shape to the lack of simulation of the WC efficiency in the MC which is momentum dependent and leads to enhance the number events in the center of the momentum distribution.
-
-%Once the beam composition at WC4 is know,  we simulate the electrons, muons and pions with the DDMC and we subject the three samples to the same selection chain (WC2TPC match, shower filter, pile up filter).
-The percentage of electrons and muons surviving the selection chain weighted by the beam composition is the  electron and muon background in the pion cross section sample, as shown in Table \ref{tab:MCafterCutContaminants}.
-
-%\subsubsection{Background from secondaries at TPC Front Face}
-%Pions can travel the length of the LArIAT beamline and interact hadronically in the steel or in the non-instrumented argon upstream to the TPC front face.
-Or, they could decay in flight between WC4 and the TPC.
-One of the interaction products can leak into the TPC and be matched with the WC track, contributing to the pool of events used for the cross section calculation.
-We call this type of particles ``secondaries" from pion events, with a terminology inspired by Geant4.
-We estimate the number of secondaries using the DDMC pion sample.
-The percentage of secondaries is given by the number of matched WC2TPC tracks whose corresponding particle is not flagged as primary by Geant4.
-The secondary to pion ratio is 4.9\% in the 60A sample and $Y$\% in the 100A sample.
-
-\subsection{Background Contribution to the Cross Section}\label{sec:Correction}
-
-Figures \ref{fig:CorrectionsBeam} show $C^{\pi MC}_{Interacting} (E_{i})$ and $C^{\pi MC}_{Incident} (E_{i})$ as a function of the kinetic energy for the 60A runs and their systematic uncertainty.
-We take a 100\% systematic uncertainty on the muon and electron content: we calculate the extreme values of  $C^{\pi MC}_{Interacting} (E_{i})$ and $C^{\pi MC}_{Incident} (E_{i})$ in each bin changing the beam composition for the configurations listed in Table \ref{tab:beamlineSys}.
-
-\begin{table}[p]
-\centering
-\begin{tabular}{| l | l | l | l | l | l | l | l | }
-\hline
- &  \multicolumn{3}{|c|}{Magnet Current -60A} & \multicolumn{3}{|c|}{Magnet Current -100 A}\\
-
-                                                  & MC $\pi^-$   & MC  $ \mu^-$ & MC  $e^-$ & MC  $\pi^-$ & MC  $\mu^-$ & MC  $e^-$  \\
-\hline
-Composition 2x muons          & 64.2	\%&9.2 \%&	26.6 \%&	83.7 \%&	7.4	\%&8.9 \% \\
-Composition 0.5x muons       &71.1	\%&2.3 \%&	26.6 \%&	89.2 \%&	1.9	\%&8.9 \% \\
-Composition 2x Electrons      &42.2	\%&4.6 \%&	53.2 \%&	78.5	\%&  3.7	\%&17.8 \%\\
-Composition 0.5x Electrons   &82.1	\%&4.6 \%&	13.3 \%&	91.9 \%&	3.7	\%&4.4 \% \\
-\hline
-\end{tabular}
-\caption{Beam composition variation for the study of systematics due to beam contamination.}
-\label{tab:beamlineSys}
-\end{table}
-
-\begin{figure}[p]
-\centering
-%\includegraphics[width=\textwidth]{Studies/Figures/c60A_Multiple.png}
-\caption{Left: relative pion content for interacting histogram a function of kinetic energy for the 60A runs, statistics uncertainty in azure and systematic uncertainty in blue.
-Right: relative pion content  for incident histogram a function of kinetic energy for the 60A runs, statistics uncertainty in azure and systematic uncertainty in blue.}
-\label{fig:CorrectionsBeam}
-\end{figure}
-
-\begin{figure}[htb]
-\centering
-\includegraphics[width=\textwidth]{Chapter-6/Images/Bkg60A_inc_int.pdf}
-\includegraphics[width=\textwidth]{Chapter-6/Images/Bkg100A_inc_int.pdf}
-\caption{.}
-\label{fig:BkgCorr}
-\end{figure}
-
-\subsubsection{Treatment of Systematics}
+\subsection{Background subtraction}\label{ch:BKGsubXSPuppa}
+We use the procedure described in \ref{ch:PionXSBkgSub2} to evaluate the pion content in the interacting $C^{\pi MC}_{\text{Int}} (E_{i})$  and incident $C^{\pi MC}_{\text{Inc}} (E_{i})$) histograms.
 
@@
 \end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{comment}
+\subsection{Background subtraction}\label{ch:BKGsubXS}
+%Even if pions are by far the biggest component of the beam in negative polarity runs, the LArIAT beam is not a pure pion beam.
+While useful to discriminate  pions/muons/electrons from kaons, and protons, the beamline detectors are not sensitive enough to  discriminate among the lighter particles in the beam: electrons, muons and pions fall under the same mass hypothesis.
+Thus, we need to assess the background from beamline particles other than pions in the event selections used for the pion cross section analysis and correct for its effects.
+
+%\subsubsection{Beam Composition}\label{sec:BeamAtWC4}
+%We define beamline background every TPC track matched to the WC track which is not a primary pion. Potentially, there are 4 different types of beamline background:
+%\begin{itemize}
+%\item[]1) electrons,
+%\item[]2) muons,
+%\item[]3) secondaries from pion events,
+%\item[]4) matched pile up events.
+%\end{itemize}
+
+%The first step is to estimate what percentage of events used in the cross section calculation is not a primary pion.
+The next two sections will illustrate this estimate for the electrons, muons and secondaries from pion event.
+%We estimate the last type of background, the ``matched pile up" events, to be a negligible fraction, because of the definition of the WC2TPC match: we deem the probability of a single match with a halo particle in the absence of a beamline particle\footnote{ Events with multiple WC2TPC matches are always rejected.} negligibly small.
+\textcolor{red}{SHOW VTX distribution in WC2TPC match}
+
+%\subsubsection{Background from Beamline Electrons and Muons}\label{stionEMu}
+%\begin{figure}[b]
+%\includegraphics[width=0.5\textwidth,height=\textheight,keepaspectratio]{Studies/Figures//Beam60A.png}
+%\includegraphics[width=0.5\textwidth,height=\textheight,keepaspectratio]{Studies/Figures//Beam100A.png}
+%\caption{Beam composition for the -60A runs (left) and -100A runs (right).
+The solid blue plot represents the simulated pion content, the yellow plot represents the simulated muon content and the grey plot represents the simulated electron content.
+The plots are area normalized to the number of data events, shown in red.
+}
+%\label{fig:BeamComposition}
+%\end{figure}
+
+%We estimate the percentage of electrons and muons in the beam via the G4Beamline MC. 
+%Since the beamline composition is a function of the magnet settings, we simulate separately events for magnet current of -60A and -100A. 
+%Table \ref{tab:beamline} shows the beam composition per magnet setting after the mass selection according to the G4Beamline simulation.
+%\begin{table}[p]
+%\centering
+%\begin{tabular}{|l|c|c|}
+%\hline
+ %                   & I = -60 A           & I = -100 A \\ \hline
+%G4Pions       &   68.8 \%           &      87.4 \%        \\ \hline
+%G4Muons     &     4.6 \%           &        3.7 \%         \\ \hline
+%G4Electrons &   26.6 \%           &        8.9 \%        \\ \hline
+%\end{tabular}
+%\caption{Simulated beamline composition per magnet settings}
+%\label{tab:beamline}
+%\end{table}
+
+%Figure \ref{fig:BeamComposition} shows the momentum predictions from G4Beamline overlaid with data for the 60A runs (left) and for the 100A runs (right).
+The predictions for electrons, muons and pions have been staggered and their sum is area normalized to data, which is shown in red.
+Albeit not perfect, these plots show a reasonable agreement between the momentum shapes in data and MC.
+We attribute  the difference in shape to the lack of simulation of the WC efficiency in the MC which is momentum dependent and leads to enhance the number events in the center of the momentum distribution.
+
+%Once the beam composition at WC4 is know,  we simulate the electrons, muons and pions with the DDMC and we subject the three samples to the same selection chain (WC2TPC match, shower filter, pile up filter).
+The percentage of electrons and muons surviving the selection chain weighted by the beam composition is the  electron and muon background in the pion cross section sample, as shown in Table \ref{tab:MCafterCutContaminants}.
+
+%\subsubsection{Background from secondaries at TPC Front Face}
+%Pions can travel the length of the LArIAT beamline and interact hadronically in the steel or in the non-instrumented argon upstream to the TPC front face.
+Or, they could decay in flight between WC4 and the TPC.
+One of the interaction products can leak into the TPC and be matched with the WC track, contributing to the pool of events used for the cross section calculation.
+We call this type of particles ``secondaries" from pion events, with a terminology inspired by Geant4.
+%We estimate the number of secondaries using the DDMC pion sample.
+The percentage of secondaries is given by the number of matched WC2TPC tracks whose corresponding particle is not flagged as primary by Geant4.
+The secondary to pion ratio is 4.9\% in the 60A sample and $Y$\% in the 100A sample.
+
+\subsection{Background Contribution to the Cross Section}\label{sec:Correction}
+
+Figures \ref{fig:CorrectionsBeam} show $C^{\pi MC}_{Interacting} (E_{i})$ and $C^{\pi MC}_{Incident} (E_{i})$ as a function of the kinetic energy for the 60A runs and their systematic uncertainty.
+We take a 100\% systematic uncertainty on the muon and electron content: we calculate the extreme values of  $C^{\pi MC}_{Interacting} (E_{i})$ and $C^{\pi MC}_{Incident} (E_{i})$ in each bin changing the beam composition for the configurations listed in Table \ref{tab:beamlineSys}.
+
+\begin{table}[p]
+\centering
+\begin{tabular}{| l | l | l | l | l | l | l | l | }
+\hline
+ &  \multicolumn{3}{|c|}{Magnet Current -60A} & \multicolumn{3}{|c|}{Magnet Current -100 A}\\
+
+                                                  & MC $\pi^-$   & MC  $ \mu^-$ & MC  $e^-$ & MC  $\pi^-$ & MC  $\mu^-$ & MC  $e^-$  \\
+\hline
+Composition 2x muons          & 64.2	\%&9.2 \%&	26.6 \%&	83.7 \%&	7.4	\%&8.9 \% \\
+Composition 0.5x muons       &71.1	\%&2.3 \%&	26.6 \%&	89.2 \%&	1.9	\%&8.9 \% \\
+Composition 2x Electrons      &42.2	\%&4.6 \%&	53.2 \%&	78.5	\%&  3.7	\%&17.8 \%\\
+Composition 0.5x Electrons   &82.1	\%&4.6 \%&	13.3 \%&	91.9 \%&	3.7	\%&4.4 \% \\
+\hline
+\end{tabular}
+\caption{Beam composition variation for the study of systematics due to beam contamination.}
+\label{tab:beamlineSys}
+\end{table}
+
+\begin{figure}[p]
+\centering
+%\includegraphics[width=\textwidth]{Studies/Figures/c60A_Multiple.png}
+\caption{Left: relative pion content for interacting histogram a function of kinetic energy for the 60A runs, statistics uncertainty in azure and systematic uncertainty in blue.
+Right: relative pion content  for incident histogram a function of kinetic energy for the 60A runs, statistics uncertainty in azure and systematic uncertainty in blue.}
+\label{fig:CorrectionsBeam}
+\end{figure}
+
+\begin{figure}[htb]
+\centering
+\includegraphics[width=\textwidth]{Chapter-6/Images/Bkg60A_inc_int.pdf}
+\includegraphics[width=\textwidth]{Chapter-6/Images/Bkg100A_inc_int.pdf}
+\caption{.}
+\label{fig:BkgCorr}
+\end{figure}
+
+\subsubsection{Treatment of Systematics}
+
+\end{comment}

### 2018-05-22 Chapter-6/6-PionCrossSection.tex
s

@@
 \section{Corrections to the Raw Cross Section}\label{ch:PionXSCorrections}
-As described in section \ref{ch:MCCorrections} as series of corrections are needed to derive the true pion cross section from the raw cross section. 
-The corrected cross section is given in equation \ref{eq:C},
+As described in section \ref{ch:MCCorrections}, we need to apply a background correction and an efficiency correction in order to derive the true pion cross section from the raw cross section.
+The true cross section is given in equation \ref{eq:C},
 
@@
 \subsection{Background subtraction}\label{ch:BKGsubXSPuppa}
-We use the procedure described in \ref{ch:PionXSBkgSub2} to evaluate the pion content in the interacting $C^{\pi MC}_{\text{Int}} (E_{i})$  and incident $C^{\pi MC}_{\text{Inc}} (E_{i})$) histograms.
+We use the procedure described in \ref{ch:PionXSBkgSub2} to evaluate the relative pion content in the interacting histogram $C^{\pi MC}_{\text{Int}} (E_{i})$  and the relative pion content in the incident $C^{\pi MC}_{\text{Inc}} (E_{i})$.
+We start by evaluating the relative pion content assuming the beamline composition simulated by G4Beamline, whose pion, muon and electron percentages per beam condition are reported again in the first line of Table \ref{tab:beamlineSys}.
+The left side of Figure \ref{fig:CorrectionsBeam} shows the  MC estimated  relative pion content for the interacting histogram as function of kinetic energy for the 60A runs (top) and 100A runs (bottom).
+The right side of the same figure shows the  MC estimated  relative pion content for the incident histogram as function of kinetic energy for the 60A runs (top) and 100A runs (bottom).
+In Figure \ref{fig:CorrectionsBeam} the central curves displayed in light blue are obtained using the beamline composition as predicted by G4Beamline: these are the correction curves for the relative pion content applied to data.
+
+So, the question now becomes: how well do we know the beamline composition?
+In absence of additional data constraints,  we take a 100\% systematic uncertainty on the electron content, reported in lines 3 and 4 of Table \ref{tab:beamlineSys}.
+The effect of doubling or halving the electron percentage in the beam on the pion relative content is displayed in red in Figure \ref{fig:CorrectionsBeam}.
+We reserve a slightly different treatment for the muon content.
+Since G4Beamline tracks only particles which cross all the wire chambers, pion events that decay in flight from WC1 to WC4 are not recorded by G4Beamline.
+Pion decays in the beamline could be trigger the beamline detectors in data, if the produced muon proceeds in the beamline.
+Thus, we take the G4Beamline prediction for muons as a lower bound in the composition: the effect of doubling the muon content (line 2 in Table \ref{tab:beamlineSys}) is shown in blue on Figure \ref{fig:CorrectionsBeam}.
+A future study of data from additional beamline detectors such as the Aerogel Chernkov detectors \cite{detectorPaper} or the muon range stack (see Section \ref{sec:MuRS}) has the potential of a narrowing the systematics uncertainty coming from the beamline compositon.
+
+We propagate the uncertainty on the beamline composition  as a systematic uncertainty to the cross section by varying the beam composition for all the cases listed in Table \ref{tab:beamlineSys} and evaluating variation of obtained  data cross sections in each bin.
+This systematic uncertainty is summed in quadrature with the statistical uncertainty and the systematic uncertainty related to the kinetic energy measurement.
+
+\begin{table}[p]
+\centering
+\begin{tabular}{| l | l | l | l | l | l | l | l | }
+\hline
+ &  \multicolumn{3}{|c|}{Magnet Current -60A} & \multicolumn{3}{|c|}{Magnet Current -100 A}\\
+
+                                                  & MC $\pi^-$   & MC  $ \mu^-$ & MC  $e^-$ & MC  $\pi^-$ & MC  $\mu^-$ & MC  $e^-$  \\
+\hline
+Expected Composition          & 68.8	\%&4.6 \%&	26.6 \%&	87.4 \%&	3.7	\%&8.9 \% \\
+Composition 2x Muons          & 64.2	\%&9.2 \%&	26.6 \%&	83.7 \%&	7.4	\%&8.9 \% \\
+Composition 2x Electrons      &42.2	\%&4.6 \%&	53.2 \%&	78.5	\%&  3.7	\%&17.8 \%\\
+Composition 0.5x Electrons   &82.1	\%&4.6 \%&	13.3 \%&	91.9 \%&	3.7	\%&4.4 \% \\
+\hline
+\end{tabular}
+\caption{Beam composition variation for the study of systematics due to beam contamination.}
+\label{tab:beamlineSys}
+\end{table}
+
+\begin{figure}[p]
+\centering
+\includegraphics[width=\textwidth]{Chapter-6/Images/Bkg60A_inc_int.pdf}
+\includegraphics[width=\textwidth]{Chapter-6/Images/Bkg100A_inc_int.pdf}
+\caption{\emph{Left:} MC estimated relative pion content for interacting histogram a function of kinetic energy for the 60A runs (top) and 100A runs (bottom), statistics uncertainty in azure and systematic uncertainty in blue.
+\emph{Right:}  MC estimated relative pion content  for incident histogram a function of kinetic energy for the 60A runs (top)  and 100A (bottom), statistics uncertainty in azure and systematic uncertainty in blue.}
+\label{fig:CorrectionsBeam}
+\end{figure}
 
@@
 
-Figures \ref{fig:CorrectionsBeam} show $C^{\pi MC}_{Interacting} (E_{i})$ and $C^{\pi MC}_{Incident} (E_{i})$ as a function of the kinetic energy for the 60A runs and their systematic uncertainty.
-We take a 100\% systematic uncertainty on the muon and electron content: we calculate the extreme values of  $C^{\pi MC}_{Interacting} (E_{i})$ and $C^{\pi MC}_{Incident} (E_{i})$ in each bin changing the beam composition for the configurations listed in Table \ref{tab:beamlineSys}.
-
-\begin{table}[p]
-\centering
-\begin{tabular}{| l | l | l | l | l | l | l | l | }
-\hline
- &  \multicolumn{3}{|c|}{Magnet Current -60A} & \multicolumn{3}{|c|}{Magnet Current -100 A}\\
-
-                                                  & MC $\pi^-$   & MC  $ \mu^-$ & MC  $e^-$ & MC  $\pi^-$ & MC  $\mu^-$ & MC  $e^-$  \\
-\hline
-Composition 2x muons          & 64.2	\%&9.2 \%&	26.6 \%&	83.7 \%&	7.4	\%&8.9 \% \\
-Composition 0.5x muons       &71.1	\%&2.3 \%&	26.6 \%&	89.2 \%&	1.9	\%&8.9 \% \\
-Composition 2x Electrons      &42.2	\%&4.6 \%&	53.2 \%&	78.5	\%&  3.7	\%&17.8 \%\\
-Composition 0.5x Electrons   &82.1	\%&4.6 \%&	13.3 \%&	91.9 \%&	3.7	\%&4.4 \% \\
-\hline
-\end{tabular}
-\caption{Beam composition variation for the study of systematics due to beam contamination.}
-\label{tab:beamlineSys}
-\end{table}
-
-\begin{figure}[p]
-\centering
-%\includegraphics[width=\textwidth]{Studies/Figures/c60A_Multiple.png}
-\caption{Left: relative pion content for interacting histogram a function of kinetic energy for the 60A runs, statistics uncertainty in azure and systematic uncertainty in blue.
-Right: relative pion content  for incident histogram a function of kinetic energy for the 60A runs, statistics uncertainty in azure and systematic uncertainty in blue.}
-\label{fig:CorrectionsBeam}
-\end{figure}
-
-\begin{figure}[htb]
-\centering
-\includegraphics[width=\textwidth]{Chapter-6/Images/Bkg60A_inc_int.pdf}
-\includegraphics[width=\textwidth]{Chapter-6/Images/Bkg100A_inc_int.pdf}
-\caption{.}
-\label{fig:BkgCorr}
-\end{figure}
-
 \subsubsection{Treatment of Systematics}

### 2018-05-22 Chapter-6/6-PionCrossSection.tex
s

@@
 Thus, we need to apply an efficiency correction to data in order to retrieve the true cross section.
-The efficiency correction is evaluated separately for the interacting and incident histograms, namely $\epsilon^{int}_i$ and  $\epsilon^{inc}_i$, and propagated to the cross section as shown in  equation \ref{eq:True}.
+The efficiency correction is evaluated separately for the interacting and incident histograms, namely $\epsilon^{\text{int}}_i$ and  $\epsilon^{\text{inc}}_i$, and propagated to the cross section as shown in  equation \ref{eq:C}.
 
@@
 \begin{equation}
-\epsilon^{\text{int}}_i  =  \frac{N^{\text{ $\pi$ Reco MC}}_{\text{Interacting}} (E_{i})}{ N^{\text{ $\pi$ True MC}}_{\text{Interacting}} (E_{i})  },
+ \epsilon^{\text{Int}}(E_i)  =  \frac{N^{\text{ $\pi$ Reco MC}}_{\text{Interacting}} (E_{i})}{ N^{\text{ $\pi$ True MC}}_{\text{Interacting}} (E_{i})  },
 \end{equation}
 
-where $N^{\text{ $\pi$ True MC}}_{\text{Interacting}} (E_{i}) $ is the content of the $i$-th bin in for the true interacting histogram, and $N^{\text{ $\pi$ Reco MC}}_{\text{Interacting}} (E_{i}) $ is the content of the $i$-th bin in for the reconstructed interacting histogram.
+where $N^{\text{ $\pi$ True MC}}_{\text{Int}} (E_{i}) $ is the content of the $i$-th bin in for the true interacting histogram, and $N^{\text{ $\pi$ Reco MC}}_{\text{Int}} (E_{i}) $ is the content of the $i$-th bin in for the reconstructed interacting histogram.
 The correction is applied to data as follows
@@
 \begin{equation}
-N^{\text{ $\pi$ True Data}}_{\text{Interacting}} (E_{i})  =  \frac{N^{\text{$\pi$ Reco Data}}_{\text{Interacting}} (E_{i})}{\epsilon^{\text{int}}_i} = N^{\text{$\pi$ Reco Data}}_{\text{Interacting}} (E_{i}) \frac{N^{\text{ $\pi$ True MC}}_{\text{Interacting}} (E_{i})}{ N^{\text{ $\pi$ Reco MC}}_{\text{Interacting}} (E_{i})}.
+N^{\text{ $\pi$ True Data}}_{\text{Int}} (E_{i})  =  \frac{N^{\text{$\pi$ Reco Data}}_{\text{Int}} (E_{i})}{\epsilon^{\text{Int}} (E_{i}) } = N^{\text{$\pi$ Reco Data}}_{\text{Int}} (E_{i}) \frac{N^{\text{ $\pi$ True MC}}_{\text{Int}} (E_{i})}{ N^{\text{ $\pi$ Reco MC}}_{\text{Int}} (E_{i})}.
 \end{equation}
 
-where $N^{\text{$\pi$ Reco Data}}_{\text{Interacting}} (E_{i})$ is the background subtracted bin content of the $i$-th bin in for the reconstructed interacting histogram for data, i.e. 
+where $N^{\text{$\pi$ Reco Data}}_{\text{Int}} (E_{i})$ is the background subtracted bin content of the $i$-th bin in for the reconstructed interacting histogram for data, i.e. 
 \begin{equation}
-N^{\text{$\pi$ Reco Data}}_{\text{Interacting}} (E_{i}) =  N^{\text{TOT Data}}_{\text{Interacting}} (E_{i}) - B^{\text{Data}}_{\text{Interacting}} (E_i)  =  C^{\text{$\pi$ MC}}_{\text{Interacting}} (E_{i}) N^{\text{TOT Data}}_{\text{Interacting}} (E_{i}).
+N^{\text{$\pi$ Reco Data}}_{\text{Int}} (E_{i}) =  N^{\text{TOT Data}}_{\text{Int}} (E_{i}) - B^{\text{Data}}_{\text{Int}} (E_i)  =  C^{\text{$\pi$ MC}}_{\text{Int}} (E_{i}) N^{\text{TOT Data}}_{\text{Int}} (E_{i}).
 \end{equation}
-
-Figures \ref{fig:EffCorr} show $\epsilon^{\text{int}}_i(E_{i})$ and $\epsilon^{\text{int}}_i (E_{i})$ as a function of the kinetic energy for the 60A runs and their systematic uncertainty.
-
-%############################# Systematics
 
@@
 $\bar\alpha_{MC} = (4.5 \pm 3.9) \text{ deg}$, respectively.
-Interaction angles smaller than the angular resolution are indistinguishable for the reconstruction.
+Most interaction angles smaller than the angular resolution will thus be indistinguishable  for the reconstruction.
 Thus, we claim we are able to  measure the cross section for interaction angles greater than 5.0 deg.
-Geant4 simulates interactions at all angles, as shown in figure \ref{trueScatteringAngle}.
-In order to calculate the efficiency correction,  we select events which have an interaction angle greater than a given $\alpha_res$ to construct the true interacting and incident histograms (the denominator of the efficiency correction).
+Geant4 simulates interactions at all angles, as shown in figure \ref{fig:trueScatteringAngle}.
+In order to calculate the efficiency correction,  we select events which have an interaction angle greater than a given $\alpha_{res}$ to construct the true interacting and incident histograms (the denominator of the efficiency correction).
+The systematics on the efficiency correction is estimated by varying the value of $\alpha_{res}$ between 0 deg and 4.5 deg and propagating the uncertainty on the cross section.
 
-The systematics on the efficiency correction is estimated by varying the value of $\alpha_{res}$ to be
+Figure \ref{fig:EffCorr60A} shows $\epsilon^{\text{Int}}(E_{i})$ in the left side and $ \epsilon^{\text{Inc}}(E_i)$ on the right as a function of the kinetic energy for the 60A runs and their systematic uncertainty.
+Similarly, figure \ref{fig:EffCorr100A} shows $\epsilon^{\text{Int}}(E_{i})$ in the left side and $ \epsilon^{\text{Inc}}(E_i)$ on the right as a function of the kinetic energy for the 100A runs and their systematic uncertainty.
 
-\subsubsection{Treatment of Systematics}
+\begin{figure}[p]
+\centering
+\includegraphics[width=\textwidth]{Chapter-6/Images/60AEffCorr.pdf}
+\caption{\emph{Left:} Efficiency correction on the 60A interacting histogram, statistical uncertainty in blue, systematic uncertainty in red.
+\emph{Right:}  Efficiency correction on the 60A incident histogram, statistical uncertainty in blue, systematic uncertainty in red.}
+\label{fig:EffCorr60A}
+\end{figure}
+
+\begin{figure}[p]
+\centering
+\includegraphics[width=\textwidth]{Chapter-6/Images/100AEffCorr.pdf}
+\caption{\emph{Left:} Efficiency correction on the 100A interacting histogram, statistical uncertainty in blue, systematic uncertainty in red.
+\emph{Right:}  Efficiency correction on the 100A incident histogram, statistical uncertainty in blue, systematic uncertainty in red.}
+\label{fig:EffCorr100A}
+\end{figure}
+
+\begin{figure}[p]
+\centering
+\includegraphics[width=0.48\textwidth]{Chapter-5/Images//cAngleTrue.png}
+\caption{Distribution of the true scattering angle for a pion elastic scattering off the argon nucleus as simulated by Geant4.}
+\label{fig:trueScatteringAngle}
+\end{figure}
 
@@
 In all three plot, the Geant4 prediction for the total hadronic cross section for angle scattering greater than 5$^\circ$ is displayed in green.
+
+The systematic uncertainty on the cross section is the sum in quadrature of the statistical uncertainty, the systematic uncertainty related to the kinetic energy measurement, the systematic uncertainty related to the beam composition and the systematic uncertainty related to the efficiency correction.
 

### 2018-05-22 Chapter-2/2-LArTPCs.tex
s

@@
 \chapter{Liquid Argon Detectors at the Intensity Frontier}\label{ch:2}
-%{\raggedleft ``\emph{Beat-up little seagull, on a marble stair} \par}
-%{\raggedleft \emph{Tryin' to find the ocean, lookin' everywhere.}"\par}
-%{\raggedleft -- Nina Simone, Baltimore, 1978 -- \par}
-%\vspace{0.5cm}
+{\raggedleft ``\emph{Beat-up little seagull, on a marble stair} \par}
+{\raggedleft \emph{Tryin' to find the ocean, lookin' everywhere.}"\par}
+{\raggedleft -- Nina Simone, Baltimore, 1978 -- \par}
+\vspace{0.5cm}
 

### 2018-05-22 Chapter-3/3-TheLArIATExperiment.tex
s

@@
 \chapter{LArIAT: Liquid Argon In A Testbeam}\label{sec:experimentDescription}
+{\raggedleft ``\emph{But hey we need to be somewhat foolish,} \par}
+{\raggedleft \emph{feebleminded, wrong and senseless.}"\par}
+{\raggedleft -- Agnes Obel, Avenue, 2010 -- \par}
+\vspace{0.5cm}
+
 In this chapter, we describe the LArIAT experimental setup.

### 2018-05-22 Chapter-4/4-Methodology.tex
s

@@
 \chapter{Total Hadronic Cross Section Measurement Methodology}\label{ch:Interactions}
-%{\raggedleft ``\emph{Like a lemon to the lime and the bubble to the bee}" \par}
-%{\raggedleft -- Eazy-E, Gimmie that Nutt,  1993 -- \par}
-%\vspace{0.5cm}
+{\raggedleft ``\emph{Like a lemon to the lime and the bubble to the bee}" \par}
+{\raggedleft -- Eazy-E, Gimmie that Nutt,  1993 -- \par}
+\vspace{0.5cm}
 

### 2018-05-22 Chapter-5/5-Preparatory.tex
s

@@
 \chapter{Data and MC preparation for the Cross Section Measurements}\label{ch:samples}
-%{\raggedleft ``\emph{Il dolce non lo mangi mai, ma qualche volta ti rifai.} \par}
-%{\raggedleft \emph{Abbracciami}"\par}
-%{\raggedleft -- Pietro Ciampi, L'amore e' tutto qui, 1971 -- \par}
-%\vspace{0.5cm}
-
-%{\raggedleft ``\emph{You never eat dessert, but, sometimes, you make up for it.} \par}
-%{\raggedleft \emph{Hug me}"\par}
-%{\raggedleft -- Pietro Ciampi, L'amore e' tutto qui, 1971 -- \par}
-%\vspace{0.5cm}
+{\raggedleft ``\emph{Il dolce non lo mangi mai, ma qualche volta ti rifai.} \par}
+{\raggedleft \emph{Abbracciami}"\par}
+{\raggedleft -- Pietro Ciampi, L'amore e' tutto qui, 1971 -- \par}
+\vspace{0.5cm}
 

### 2018-05-22 Chapter-6/6-PionCrossSection.tex
s

@@
 \chapter{Negative Pion Cross Section Measurement}\label{ch:PionXS}
-%{\raggedleft ``\emph{Y ella es flama que se eleva, Y es un p\'ajaro a volar.} \par}
-%{\raggedleft \emph{En la noche que se incendia, estrella de oscuridad}\par}
-%{\raggedleft \emph{que busca entre la tiniebla, la dulce hoguera del beso.}"\par}
-%{\raggedleft -- Lila Downs, Benediction And Dream,  2002 -- \par}
-%\vspace{0.5cm}
+{\raggedleft ``\emph{Y ella es flama que se eleva, Y es un p\'ajaro a volar.} \par}
+{\raggedleft \emph{En la noche que se incendia, estrella de oscuridad}\par}
+{\raggedleft \emph{que busca entre la tiniebla, la dulce hoguera del beso.}"\par}
+{\raggedleft -- Lila Downs, Benediction And Dream,  2002 -- \par}
+\vspace{0.5cm}
 

### 2018-05-22 AppendixB-TrackingStudies/B-Tracking.tex
s

@@
 \chapter{Tracking Studies}\label{ch:AppendixTrack}
+In this section, we describe three studies.
+The first is a justification of the selection criteria for the beamline handshake with the TPC information.
+We perform this study to boost  the correct identification of the particles in the TPC associated with the beamline information, while maintaining sufficient statistics for the cross section measurement.
+The second study is an optimization of the tracking algorithm, with the scope of maximizing the identification of the hadronic interaction point inside the TPC.
+These two studies are related, since the optimization of the tracking is performed on TPC tracks which have been matched to the wire chamber track; in turn, the tracking algorithm for TPC tracks determines the number of reconstructed tracks in each event used to try the matching with the wire chamber track.
+Starting with a sensible tracking reconstruction, we perform the WC2TPC matching optimization first, then the tracking optimization.
+The WC2TPC match purity and efficiency  are then calculated again with the optimized tracking.
+
+\subsection{Study of WC to TPC Match}\label{ch:WC2TPCMatchOptimization}
+
+Plots I want in this section:
+\begin{enumerate}
+\item WC2TPC MC DeltaX, DeltaY and $\alpha$
+\end{enumerate}
+
+Scope of this study is assessing the goodness of the wire chamber to TPC match on Monte Carlo and decide the selection values we will use on data.
+A word of caution is necessary here.
+With this study, we want to minimize pathologies associated with the presence of the primary hadron itself, e.g.
+the incorrect association between the beamline hadron and its decay products inside the TPC.
+Assessing the contamination from pile-up\footnote{We remind the reader that the DDMC is a single particle Monte Carlo, where the beam pile up is not simulated.}, albeit related, is beyond the scope of this study.
+
+In MC, we are able to define a correct WC2TPC match using the Geant4 truth information. We are thus able to count how many times the WC tracks is associated with the wrong TPC reconstructed track.
+
+We define a correct match if the all following conditions are met:
+\begin{itemize}
+\item[-] the length of the true primary Geant4 track in the TPC is greater than 2 cm,  
+\item[-] the length of the reconstructed track length is greater than 2 cm,
+\item[-] the Z position of the first reconstructed point is within 2 cm from the TPC front face
+\item[-] the distance between the reconstructed track and the true entering point is the minimum compared with all the other reconstructed tracks.
+\end{itemize}
+
+In order to count the wrong matches, we consider all the reconstructed tracks whose Z position of the first reconstructed point lies within 2 cm from the TPC front face.
+Events with true length in TPC $<$ 2 cm are included.
+Since hadrons are shot 100 cm upstream from the TPC front face, the following two scenarios are possible from a truth standpoint: 
+\begin{itemize}
+\item[[$Ta$]] the primary hadron decays or interact strongly before getting to the TPC,
+\item[[$Tb$]] the primary hadron enters the TPC.
+\end{itemize}
+
+As described in Section \ref{ch:WC2TPCMatchMethod}, we define a WC2TPC match according to the relative position of the WC and TPC track parametrized with $\Delta R$ and the angle between them, parametrized with $\alpha$.
+Once we choose the selection values $r_{T}$ and $\alpha_{T}$ to determine a reconstructed WC2TPC match, the following five scenarios are possible in the truth to reconstruction interplay :
+\begin{itemize}
+\item[1)] only the correct track is matched
+\item[2)] only one wrong track is matched 
+\item[3)] the correct track and one (or more) wrong tracks are matched
+\item[4)] multiple wrong tracks  matched.
+\item[5)] no reconstructed tracks are matched
+\end{itemize}
+
+Since we keep only events with one and only one match, we discard cases 3), 4) and 5) from the events used in the cross section measurement.
+For each set of $r_{T}$ and $\alpha_{T}$ selection value, we define purity and efficiency of the selection as follows:
+\begin{equation}
+\text{Efficiency} = \frac{\text{Number of events correctly matched}}{\text{ Number of events with primary in TPC}},
+\end{equation}
+
+\begin{equation}
+\text{Purity} = \frac{\text{Number of events correctly matched}}{\text{Total number of matched events}}.
+\end{equation}
+
+Figure \ref{fig:EffPurityK} shows the efficiency (left) and purity (right) for WC2TPC match as a function of the radius, $r_{T}$, and angle, $\alpha_{T}$, selection value.
+It is apparent how both efficiency and purity are fairly flat as a function of the radius selection value at a given angle.
+This is not surprising.
+Since we are studying a single particle gun Monte Carlo sample, the wrong matches can occur only for mis-tracking of the primary or for association with decay products;  decay products will tend to be produced at large angles compared to the primary, but could be fairly close to the in $x$ and $y$ projection of the primary.
+The radius cut would play a key role in removing pile up events.
+
+For LArIAT cross section measurements, we generally prefer purity over efficiency, since a sample of particles of a pure species will lead to a better measurement.
+Obviously, purity should be balanced with a sensible efficiency to avoid rejecting the whole sample.
+
+We choose $(\alpha_{T}$, $r_{T}) = (8 \text{ deg}, 4 \text{ cm} )$ and get a MC 85\% efficiency and 98\% purity for the kaon sample and a MC 95\% efficiency and 90\% purity for the pion sample.
+
+\begin{figure}[hpbt]
+\centering
+\includegraphics[width=15cm]{Chapter-5/Images/KEffPurity.png}
+\caption{Efficiency (left) and purity (right) for WC2TPC match as a function of the radius and angle selections for the kaon sample.}
+\label{fig:EffPurityK}
+\end{figure}
+
+\subsection{Tracking Optimization}\label{ch:TrackOptimization}

### 2018-05-22 AppendixC-EnergyCalibration/C-EnergyCalibration.tex
s

@@
-\chapter{Energy Calibration}\label{ch:AppendixEnergy}
+\chapter{Energy Calibration}\label{ch:energyCalibration}
+
+Scope of the energy calibration is to identify the factors which convert the charge collected (dQ) to energy deposited in the chamber (dE).
+As described in section \ref{sec:SignalProc}, this is a multi-step procedure.
+In LArIAT, we first correct the raw charge by the electronic noise on the considered wire \cite{LArIATdqdx}, then by the electron lifetime \cite{LArIATLifeTime},  and then by the recombination using the ArgoNeut recombination values.
+Lastly, we apply overall calibration of the energy, i.e.
+we determine the ``calorimetry constants" using the procedure described in this section.
+
+We independently determine  the calorimetry constants for Data and Monte Carlo in the LArIAT Run-II Data samples using  a parametrization of the stopping power (a.k.a.
+energy deposited per unit length, $dE/dX$)  as a function of momentum.
+This is done by comparing the stopping power measured on reconstructed quantities against the Bethe-Bloch theoretical prediction for various particle species (see Equation \ref{eq:BB}).
+We obtain the theoretical expectation for the $dE/dX$ most probable value of pions ($\pi$), muons ($\mu$), kaons ($K$), and protons ($p$) in the momentum range most relevant for LArIAT (Figure \ref{fig:PDGEnergyLossArgon}) using the tables provided by the Particle Data Group \cite{Patrignani:2016xqp} for liquid argon \cite{PDG-Argon}.
+
+The basic idea of this calibration technique is to utilize a sample of beamline events with known particle species and momentum to measure the $dE/dX$ of the corresponding tracks in the TPC.
+In particular, we decided to use positive pions as calibration sample and samples from all the other particle species as cross check.
+Once the $dE/dX$ of the positive pion sample  has been measured at various momenta, we tune to calorimetry constants within the reconstruction software to align the measured values to match the theoretical ones found in Figure \ref{fig:PDGEnergyLossArgon}.
+
+In data, we start by selecting a sample of beamline positive pion beamline candidates without any restriction on their measured momentum\footnote{it should be noted that some muon and position contamination is present in the $\pi^+$ sample}.
+We then apply the WC2TPC match and subtract the energy loss upstream to the TPC front face, determining the momentum at the TPC front face.
+For each surviving pion candidate,  we measure the $dE/dx$ at each of the first 12 spacepoints associated the 3D reconstructed track, corresponding to a $\sim$ 5 cm portion.
+These $dE/dX$ measurements are then put into a histogram that corresponds to measured momentum of the track.
+The $dE/dX$ histograms are sampled every 50 MeV/c in momentum (e.g.
+150~MeV/c $< P <$ 200~MeV/c, 200~MeV/c $< P <$ 250/c~MeV, etc...).
+This process of selecting, sampling, and recording the $dE/dX$ for various momentum bins is repeated over the entire sample of events, allowing us to collect sufficient statistic in most of the momentum bins between 150~MeV/c and 1100~MeV/c.
+On average, pions and muons only lose $\sim$10 MeV in this 5~cm section of the track and protons lose $\sim$20 MeV.
+Thus choosing 50 MeV/c size bins for our histograms covers the energy spread within those bins due to energy loss from ionization for all the particle species identifiable in the beamline.
+Each 50 MeV/c momentum binned $dE/dX$ histogram is now fit with a simple Landau function.
+The most probable value (MPV) and the associated error on the MPV from the fit are extracted and plotted against the theoretical prediction Figure \ref{fig:PDGEnergyLossArgon}.
+Depending on the outcome of the data-prediction comparison, we modify the calorimetry constants and we repeat the procedure until a qualitative agreement is achieved.
+We perform this  tuning for the collection and induction plane separately.
+As a cross check to the calorimetry constants determined using the positive pions, we lock the constants and  plot the $dE/dx$ versus momentum distribution of all the other particle species identifiable in the beamline data ($\pi/\mu/e$, K , p, in both polarities) against the corresponding Beth-Bloch prediction.
+The agreement between data from the other particle species and the predictions is the expected result of this cross check.
+The results of the tuning and cross check for Run-II data on the collection plane is shown in Figure \ref{fig:BBandData}  negative polarity data on top, positive polarity data on the bottom.
+
+In MC, we simulate the corresponding positive pion sample with the DDMC (see section \ref{sec:DDMC}) and follow the same steps as in data.
+More details on the calorimetry tuning can be found in \cite{LArIATCalo}.
+
+\textcolor{red}{Add agreement between data and MC for dedx for pions}
+
+\begin{figure}[htb]
+\centering
+\includegraphics[width=0.50\textwidth]{Chapter-5/Images/dEdXvsMomentumTemplate.png}
+\caption{Stopping power for pions, muons, kaons, and protons in liquid argon over the momentum range most relvant for LArIAT according to the Beth-Bloch equation.
+The solid lines represent the prediction for the mean energy $dE/dX$, while the dashed lines are the predictions for the MPV.}
+\label{fig:PDGEnergyLossArgon}
+\end{figure}
+
+\begin{figure}[htb]
+\centering
+\includegraphics[width=0.50\textwidth]{Chapter-5/Images/RunIINegTotaldEdXvsMomentum.png}
+\includegraphics[width=0.50\textwidth]{Chapter-5/Images/RunIIPosTotaldEdXvsMomentum.png}
+\caption{Stopping power versus Momentum for Run-II negative (top) and positive (bottom) polarity data.
+We achieve the agreement between the Bethe-Bloch predictions and the distribution obtained with of the positive pions (top plot, red dots) by tuning the calorimetry constants.
+Once the calorimetry constants are locked in, the agreement between the other particle species and the Bethe-Bloch predictions follows naturally.}
+\label{fig:BBandData}
+\end{figure}
+
+%\begin{figure}[htb]
+%\centering
+%\includegraphics[width=0.50\textwidth]{images/CalibrationExample.png}
+%\caption{Illustration of the calibration technique.
+Here we depict a 325 MeV wire chamber track (shown in green) which enters the TPC (taking into account the energy loss from the upstream material) and we sample the first 12 spacepoints (shown in teal) to extract the dE/dX distribution which is fit with a Landau.}
+%\label{fig:CalibrationExample}
+%\end{figure}

### 2018-05-22 Chapter-4/4-Methodology.tex
s

@@
 
-This chapter describes the general procedure employed to measure a total hadronic differential cross 10.03.p1section in LArIAT.
-Albeit with small differences, both the  ($\pi^{-}$,Ar) and (K$^{+}$,Ar) total hadronic cross section measurements rely on the same procedure described in details in the following sections.
+This chapter describes the general procedure employed to measure a total hadronic differential cross section in LArIAT.
+Albeit with small differences, both the  ($\pi^{-}$,Ar) and (K$^{+}$,Ar) total hadronic cross section measurements rely on the same procedure.
 We start by selecting the particle of interest using a combination of beamline detectors and TPC information (Section \ref{ch:ParticleSelectionMethod}).
-We then perform a handshake between the beamline information and the TPC tracking to assure the selection of the right TPC track (Section \ref{ch:WC2TPCMatchMethod}).
+We then perform a handshake between the beamline information and the TPC tracking to assure the selection of the correct TPC track (Section \ref{ch:WC2TPCMatchMethod}).
 Finally, we apply the ``thin slice" method and measure the ``raw" hadronic cross section (Section \ref{ch:ThinSliceMethod}).
@@
 \subsection{Selection of Beamline Events}\label{ch:beamlineDetectorsData}
-As shown in equation \ref{eq:enFF}, we leverage the beamline particle identification and momentum measurement before entering the TPC as in input to evaluate the kinetic energy for the hadrons used in the  cross sections measurements.
+As shown in equation \ref{eq:enFF}, we leverage the beamline particle identification and momentum measurement before entering the TPC as an input to evaluate the kinetic energy for the hadrons used in the  cross sections measurements.
 Thus, we select the LArIAT data to keep only events whose wire chamber and time of flight information is registered (line 1 in in Table \ref{tab:beamlineDataSelection}).
@@
 \begin{itemize}
-\item \underline{$\pi/\mu/e$:}  mass $<$ 350~MeV
+\item \underline{$\pi/\mu/e$:}  mass $<$ 350~MeV/c$^2$
 
-\item \underline{kaon:} 350~MeV $<$ mass $<$ 650~MeV
+\item \underline{kaon:} 350~MeV $<$ mass $<$ 650~MeV/c$^2$
 
-\item \underline{proton:} 650~MeV $<$ mass $<$ 3000~MeV.
+\item \underline{proton:} 650~MeV $<$ mass $<$ 3000~MeV/c$^2$.
 
@@
 
-In general, the target is not a single particle, but rather a slab of material containing many diffusion centers.
+In general even in the case of thin target, the target is not a single particle, but rather a slab of material containing many diffusion centers.
 The so-called  ``thin target" approximation assumes that the target centers are uniformly distributed in the material and that the target is thin compared to the projectile interaction length, so that no center of interaction sits in front of another.
-In this approximation, the ratio between the number of particles interacting in the target $N_{\text{Int}}$ and number of incident particles $N_{\text{Inc}}$ on the target determines the interaction probability $P_{Interacting}$, which is the complementary to one of the survival probability $P_{Survival}$.
+In this approximation, the ratio between the number of particles interacting in the target $N_{\text{Int}}$ and the number of incident particles $N_{\text{Inc}}$ on the target determines the interaction probability $P_{Interacting}$, which is the complementary to one of the survival probability $P_{Survival}$.
 Equation \ref{eq:thinTargetXS} 
@@
 If the TPC track stops within the fiducial volume, its last point will be the interaction point; if the track crosses the boundaries of the fiducial volume, the track will be considered ``through going" and no interaction point will be found.
-The only slabs considered to fill the  $N_{\text{Inc}})$ and  $N_{\text{Inc}}$ plots are the slabs included the fiducial volume.
+The only points of the hadronic candidate track considered to fill the  $N_{\text{Inc}})$ and  $N_{\text{Inc}}$ plots are the ones contained in the fiducial volume.
 
@@
 Unfortunately, this is not the case.
-In fact, the selection used to isolate pions in the LArIAT beam allows for the presence of some muons and electrons as background.
+In fact, the selection used to isolate pions in the LArIAT beam allows for the presence of some muons and electrons as background, while the kaon selection allows for a small percentage of protons (see Section \ref{ch:beamlineComposition}).
 Also, the LArIAT TPC is not 100\% efficient in determining the interaction point.
@@
 
-where  $N^{\text{TOT}}_{\text{Int}} (E_{i})$ and $N^{\text{TOT}}_{\text{Incident}} (E_{i})$ is the measured content of the interacting and incident histograms for events that pass the event selection, $B_{\text{Int}} (E_i)$ and $B_{\text{Inc}} (E_i)$ represent the contributions from beamline background to the interacting and incident histograms respectively, and  $\epsilon^{\text{Int}}(E_i)$ and  $\epsilon^{\text{Inc}}(E_i)$ are the efficiency corrections for said histograms.
+where  $N^{\text{TOT}}_{\text{Int}} (E_{i})$ and $N^{\text{TOT}}_{\text{Incident}} (E_{i})$ is the measured content of the interacting and incident histograms for events that pass the event selection, $B_{\text{Int}} (E_i)$ and $B_{\text{Inc}} (E_i)$ represent the contributions from the background to the interacting and incident histograms respectively, and  $\epsilon^{\text{Int}}(E_i)$ and  $\epsilon^{\text{Inc}}(E_i)$ are the efficiency corrections for said histograms.
 
@@
 
-For pions and kaons in the considered energy range, the Geant4 inelastic model adopted is ``BertiniCascade"; the pion elastic cross sections are modeled on Chips, while the kaon elastic cross sections are modeled on Gheisha and Chips.
+For pions and kaons in the considered energy range, the Geant4 inelastic model adopted is ``BertiniCascade"; the pion elastic cross sections are tabulated from on Chips, while the kaon elastic cross sections are tabulated on Gheisha and Chips.
 
@@
 Since the MC is a pure beam of the hadron of interest and truth information is used to retrieve the interaction point, no correction is applied.
-Doing so, we obtain the true interacting and incident distributions for the considered hadron and we obtain the true MC cross section as a function of the hadron true kinetic energy.
+Doing so, we obtain the true interacting and incident distributions for the considered hadron, from which we derive  the true MC cross section as a function of the hadron true kinetic energy.
 

### 2018-05-22 Chapter-5/5-Preparatory.tex
s

@@
 
-This chapter describes the preparatory work done on the the data and Monte Carlo samples used for the cross section analyses. 
-This entails the choice of the datasets and the production of the information needed to construct the Monte Carlo Simulation~(section \ref{sec:dataSet}),  the construction and use of said Monte Carlo simulation~(section \ref{sec:MCSet}), the study and optimization of the tracking in the TPC for the cross section analyses~(section \ref{sec:TrackingStudies}),  the calibration of the calorimetry response and related energy studies~(section \ref{ch:energyCal}).
+This chapter describes the  work done on the  data and Monte Carlo samples in preparation for the cross section analyses. 
+This entails the choice of the datasets and the production of the information needed to construct the Monte Carlo Simulation~(Section  \ref{sec:dataSet}),  the construction and use of said Monte Carlo simulation~(section \ref{sec:MCSet}), the study of backgrounds for the pion cross section~(Section  \ref{ch:PionXSBkgSub}), the study of the energy loss between WC4 and TPC (Section \ref{ch:eloss}), the study of the tracking in the TPC~(Section  \ref{sec:TrackingStudies}), and study of  the calorimetry response ~(Section  \ref{ch:energyCal}).
 
@@
 The -60A and -100A combined data set accounts for approximately 90\% of the total Run-II negative polarity runs.
-The  choice of the main two beamline settings limits the need for the production of many MC sets and related corrections, still maintaining a high number of events.
+The  choice of the main two beamline settings limits the need for the production of many different MC sets and related corrections, still maintaining a high number of events.
 %Since the production of beamline Monte Carlo depends on the wanted beamline configuration, the choice of only two beamline settings limits the need for beamline MC production.
@@
 Albeit not perfect, these plots show a reasonable agreement between the momentum shapes in data and MC.
-We attribute  the difference in shape to the lack of simulation of the WC efficiency in the MC which is momentum dependent and leads to enhance the number events in the center of the momentum distribution.
+We attribute the difference in shape to a two approximations performed in the MC.
+Firstly, G4Beamline lacks the simulation of the WC efficiency which is momentum dependent and leads to enhance the number events in the center of the momentum distribution.
+Secondly, G4Beamline stop tracking pions and their products if they decay in after WC1; in data, pion decays in flight can still create a tigger if the produced muon travels thought the beamline detectors.
+In the pion cross section analysis, these differences between data and G4Beamline are accounted for as a systematic uncertainty related to the beam composition (see Section \ref{ch:BKGsubXSPuppa}).
+%We attribute  the difference in shape to the lack of simulation of the WC efficiency in the MC which is momentum dependent and leads to enhance the number events in the center of the momentum distribution.
 
@@
 Since the shape of the tails is unknown, the estimate is done multiple times varying the range and shape for reasonable functions. 
-For example, to estimate the proton content under the kaon peak, we start by fitting the left tail of the proton mass distribution with a gaussian function between 650 $MeV/c^2$ and 750 MeV/c$^2$.% in a reasonable range and with a reasonable function.
+For example, to estimate the proton content under the kaon peak, we start by fitting the left tail of the proton mass distribution with a gaussian function between 650 $MeV/c^2$ and 750 MeV/c$^2$.
+% in a reasonable range and with a reasonable function.
  We extend the fit function under the kaon peak and integrate the extended fit function between 350-650 MeV/c$^2$.
@@
 We repeat this procedure for several fit shapes (gaussian, linear and exponential functions) and tail ranges.
-Finally, we calculate the contamination as the weighted average of single estimates, where the weights are calculated to be the $1./\chi^2$ of the tail fits.
+Finally, we calculate the contamination as the weighted average of single estimates, where the weights are calculated to be the $1./|1-\chi^2|$ of the tail fits.
 The procedure is repeated for lighter particles mass peak independently.
-With 12 iterations of this method we find a proton contamination of  0.2 $\pm$ 0.5 \%  and a contamination from the lighter particles of 5 $\pm$ 2 \% .
-The estimate of the proton background is currently not used in the kaon cross section analysis, but it is a fundamental step to retrieve the true kaon cross section which will be implemented in the analysis next step.
+With 12 iterations of this method we find a proton contamination of  5.0 $\pm$ 2.0 \%  and a contamination from the lighter particles of 0.2 $\pm$ 0.5 \% .
+The estimate of the proton background is currently not used in the kaon cross section analysis, but it is a fundamental step to retrieve the true kaon cross section which will be implemented in the further development of the analysis.
 
@@
 The momentum and position of the selected data form a 5-dimensional tupla, which we sample thousands of times through a 5-dimensional hit-or-miss sampling procedure to generate the MC events.
-This generates MC events  with the same momentum and position distributions as data, with the additional benefit of accounting for the correlations between the $P_x, P_y, P_z, X, Y$ variables.
+This sampling generates MC events  with the same momentum and position distributions as data, with the additional benefit of accounting for the correlations between the $P_x, P_y, P_z, X, Y$ variables.
 As an example, the results of the DDMC generation compared to data for the kaon +100A sample are shown in figure \ref{fig:DDMCComparison} for the $P_z$, $X$ and $Y$ distributions; as expected, MC and data agree within the statistical uncertainty by construction.
-A LArSoft simulation module then launches single particle MC from z = -100 cm (the location of the WC4) using the MC generated events.
+A LArSoft simulation module then launches single particle MC from z = -100 cm (the location of the WC4) using the generated events.
 The particles are free to decay and interact in their path from WC4 to the TPC according to the Geant4 simulation.
 
-Using the DDMC technique ensures that the MC and data particles have very similar momentum, position and angular distributions at WC4 and allows us to use the MC sample in several occasions: to calibrate the energy loss upstream of the TPC (see Section \ref{ch:eloss}), to estimate the background contamination to the pion cross section (see Section \ref{ch:PionXSBkgSub}), or to study the tracking and the calorimetric performance (sections \ref{sec:TrackingStudies} and \ref{ch:energyCal}).
+Using the DDMC technique ensures that the MC and data particles have very similar momentum, position and angular distributions at WC4 and allows us to use the MC sample in several occasions: to estimate the background contamination to the pion cross section (see Section \ref{ch:PionXSBkgSub}), to calibrate the energy loss upstream of the TPC (see Section \ref{ch:eloss}),  or to study the tracking and the calorimetric performance (sections \ref{sec:TrackingStudies} and \ref{ch:energyCal}).
 A small caveat is in order here: the DDMC is a single particle Monte Carlo, which means that the beam pile-up is not simulated.
 
-Six samples are the basis fo the MC used in the pion cross section measurement: three samples of  $\sim$340000 pions, muons and electrons to simulate the negative 60A runs, and three samples of $\sim$340000 pions, muons and electrons for the negative 100A runs.
-
-The MC used for the kaon cross section analysis is a sample of \textcolor{red}{NUMBERS}  kaons.
+We generate six samples for the pion cross section measurement: three samples of  $\sim$330000 pions, muons and electrons to simulate the negative 60A runs, and three samples of $\sim$340000 pions, muons and electrons for the negative 100A runs.
+We generate a sample of  195000  kaons for the kaon cross section analysis.
 
@@
 Figure \ref{fig:CaptureMom} shows the true momentum distribution for the primary pions\footnote{We use here the Geant4 denomination ``primary" to indicate that the pion considered does not undergo interactions modifying its energy before getting to the TPC.
-In fact, not every pion shot from wire chamber four will arrive to the TPC as primary,  some will decay or interact before the TPC.}  that arrive to the TPC (pink), that capture (green) or decay (blue) inside the TPC, on a linear and log scale vertical axis.
+In fact, not every pion shot from wire chamber four will arrive to the TPC as primary,  some will decay or interact before the TPC.}  that arrive to the TPC (pink), that capture (green) or decay (blue) inside the TPC, on a linear scale (left) and on a log scale (right) vertical axis.
 
@@
 \section{Tracking Studies}\label{sec:TrackingStudies}
-In this section, we describe three studies.
-The first is a justification of the selection criteria for the beamline handshake with the TPC information.
-We perform this study to boost  the correct identification of the particles in the TPC associated with the beamline information, while maintaining sufficient statistics for the cross section measurement.
-The second study is an optimization of the tracking algorithm, with the scope of maximizing the identification of the hadronic interaction point inside the TPC.
-These two studies are related, since the optimization of the tracking is performed on TPC tracks which have been matched to the wire chamber track; in turn, the tracking algorithm for TPC tracks determines the number of reconstructed tracks in each event used to try the matching with the wire chamber track.
-Starting with a sensible tracking reconstruction, we perform the WC2TPC matching optimization first, then the tracking optimization.
-The WC2TPC match purity and efficiency  are then calculated again with the optimized tracking.
 
 The third study is an evaluation of  the angular resolution of the tracking algorithm in data and MC, which is particularly important in the context of the cross section analyses.
-
-\subsection{Study of WC to TPC Match}\label{ch:WC2TPCMatchOptimization}
-
-Plots I want in this section:
-\begin{enumerate}
-\item WC2TPC MC DeltaX, DeltaY and $\alpha$
-\end{enumerate}
-
-Scope of this study is assessing the goodness of the wire chamber to TPC match on Monte Carlo and decide the selection values we will use on data.
-A word of caution is necessary here.
-With this study, we want to minimize pathologies associated with the presence of the primary hadron itself, e.g.
-the incorrect association between the beamline hadron and its decay products inside the TPC.
-Assessing the contamination from pile-up\footnote{We remind the reader that the DDMC is a single particle Monte Carlo, where the beam pile up is not simulated.}, albeit related, is beyond the scope of this study.
-
-In MC, we are able to define a correct WC2TPC match using the Geant4 truth information. We are thus able to count how many times the WC tracks is associated with the wrong TPC reconstructed track.
-
-We define a correct match if the all following conditions are met:
-\begin{itemize}
-\item[-] the length of the true primary Geant4 track in the TPC is greater than 2 cm,  
-\item[-] the length of the reconstructed track length is greater than 2 cm,
-\item[-] the Z position of the first reconstructed point is within 2 cm from the TPC front face
-\item[-] the distance between the reconstructed track and the true entering point is the minimum compared with all the other reconstructed tracks.
-\end{itemize}
-
-In order to count the wrong matches, we consider all the reconstructed tracks whose Z position of the first reconstructed point lies within 2 cm from the TPC front face.
-Events with true length in TPC $<$ 2 cm are included.
-Since hadrons are shot 100 cm upstream from the TPC front face, the following two scenarios are possible from a truth standpoint: 
-\begin{itemize}
-\item[[$Ta$]] the primary hadron decays or interact strongly before getting to the TPC,
-\item[[$Tb$]] the primary hadron enters the TPC.
-\end{itemize}
-
-As described in Section \ref{ch:WC2TPCMatchMethod}, we define a WC2TPC match according to the relative position of the WC and TPC track parametrized with $\Delta R$ and the angle between them, parametrized with $\alpha$.
-Once we choose the selection values $r_{T}$ and $\alpha_{T}$ to determine a reconstructed WC2TPC match, the following five scenarios are possible in the truth to reconstruction interplay :
-\begin{itemize}
-\item[1)] only the correct track is matched
-\item[2)] only one wrong track is matched 
-\item[3)] the correct track and one (or more) wrong tracks are matched
-\item[4)] multiple wrong tracks  matched.
-\item[5)] no reconstructed tracks are matched
-\end{itemize}
-
-Since we keep only events with one and only one match, we discard cases 3), 4) and 5) from the events used in the cross section measurement.
-For each set of $r_{T}$ and $\alpha_{T}$ selection value, we define purity and efficiency of the selection as follows:
-\begin{equation}
-\text{Efficiency} = \frac{\text{Number of events correctly matched}}{\text{ Number of events with primary in TPC}},
-\end{equation}
-
-\begin{equation}
-\text{Purity} = \frac{\text{Number of events correctly matched}}{\text{Total number of matched events}}.
-\end{equation}
-
-Figure \ref{fig:EffPurityK} shows the efficiency (left) and purity (right) for WC2TPC match as a function of the radius, $r_{T}$, and angle, $\alpha_{T}$, selection value.
-It is apparent how both efficiency and purity are fairly flat as a function of the radius selection value at a given angle.
-This is not surprising.
-Since we are studying a single particle gun Monte Carlo sample, the wrong matches can occur only for mis-tracking of the primary or for association with decay products;  decay products will tend to be produced at large angles compared to the primary, but could be fairly close to the in $x$ and $y$ projection of the primary.
-The radius cut would play a key role in removing pile up events.
-
-For LArIAT cross section measurements, we generally prefer purity over efficiency, since a sample of particles of a pure species will lead to a better measurement.
-Obviously, purity should be balanced with a sensible efficiency to avoid rejecting the whole sample.
-
-We choose $(\alpha_{T}$, $r_{T}) = (8 \text{ deg}, 4 \text{ cm} )$ and get a MC 85\% efficiency and 98\% purity for the kaon sample and a MC 95\% efficiency and 90\% purity for the pion sample.
-
-\begin{figure}[hpbt]
-\centering
-\includegraphics[width=15cm]{Chapter-5/Images/KEffPurity.png}
-\caption{Efficiency (left) and purity (right) for WC2TPC match as a function of the radius and angle selections for the kaon sample.}
-\label{fig:EffPurityK}
-\end{figure}
-
-\subsection{Tracking Optimization}\label{ch:TrackOptimization}
 
@@
 
-\subsection{Energy Calibration}\label{ch:energyCalibration}
-Scope of the energy calibration is to identify the factors which convert the charge collected (dQ) to energy deposited in the chamber (dE).
-As described in section \ref{sec:SignalProc}, this is a multi-step procedure.
-In LArIAT, we first correct the raw charge by the electronic noise on the considered wire \cite{LArIATdqdx}, then by the electron lifetime \cite{LArIATLifeTime},  and then by the recombination using the ArgoNeut recombination values.
-Lastly, we apply overall calibration of the energy, i.e.
-we determine the ``calorimetry constants" using the procedure described in this section.
-
-We independently determine  the calorimetry constants for Data and Monte Carlo in the LArIAT Run-II Data samples using  a parametrization of the stopping power (a.k.a.
-energy deposited per unit length, $dE/dX$)  as a function of momentum.
-This is done by comparing the stopping power measured on reconstructed quantities against the Bethe-Bloch theoretical prediction for various particle species (see Equation \ref{eq:BB}).
-We obtain the theoretical expectation for the $dE/dX$ most probable value of pions ($\pi$), muons ($\mu$), kaons ($K$), and protons ($p$) in the momentum range most relevant for LArIAT (Figure \ref{fig:PDGEnergyLossArgon}) using the tables provided by the Particle Data Group \cite{Patrignani:2016xqp} for liquid argon \cite{PDG-Argon}.
-
-The basic idea of this calibration technique is to utilize a sample of beamline events with known particle species and momentum to measure the $dE/dX$ of the corresponding tracks in the TPC.
-In particular, we decided to use positive pions as calibration sample and samples from all the other particle species as cross check.
-Once the $dE/dX$ of the positive pion sample  has been measured at various momenta, we tune to calorimetry constants within the reconstruction software to align the measured values to match the theoretical ones found in Figure \ref{fig:PDGEnergyLossArgon}.
-
-In data, we start by selecting a sample of beamline positive pion beamline candidates without any restriction on their measured momentum\footnote{it should be noted that some muon and position contamination is present in the $\pi^+$ sample}.
-We then apply the WC2TPC match and subtract the energy loss upstream to the TPC front face, determining the momentum at the TPC front face.
-For each surviving pion candidate,  we measure the $dE/dx$ at each of the first 12 spacepoints associated the 3D reconstructed track, corresponding to a $\sim$ 5 cm portion.
-These $dE/dX$ measurements are then put into a histogram that corresponds to measured momentum of the track.
-The $dE/dX$ histograms are sampled every 50 MeV/c in momentum (e.g.
-150~MeV/c $< P <$ 200~MeV/c, 200~MeV/c $< P <$ 250/c~MeV, etc...).
-This process of selecting, sampling, and recording the $dE/dX$ for various momentum bins is repeated over the entire sample of events, allowing us to collect sufficient statistic in most of the momentum bins between 150~MeV/c and 1100~MeV/c.
-On average, pions and muons only lose $\sim$10 MeV in this 5~cm section of the track and protons lose $\sim$20 MeV.
-Thus choosing 50 MeV/c size bins for our histograms covers the energy spread within those bins due to energy loss from ionization for all the particle species identifiable in the beamline.
-Each 50 MeV/c momentum binned $dE/dX$ histogram is now fit with a simple Landau function.
-The most probable value (MPV) and the associated error on the MPV from the fit are extracted and plotted against the theoretical prediction Figure \ref{fig:PDGEnergyLossArgon}.
-Depending on the outcome of the data-prediction comparison, we modify the calorimetry constants and we repeat the procedure until a qualitative agreement is achieved.
-We perform this  tuning for the collection and induction plane separately.
-As a cross check to the calorimetry constants determined using the positive pions, we lock the constants and  plot the $dE/dx$ versus momentum distribution of all the other particle species identifiable in the beamline data ($\pi/\mu/e$, K , p, in both polarities) against the corresponding Beth-Bloch prediction.
-The agreement between data from the other particle species and the predictions is the expected result of this cross check.
-The results of the tuning and cross check for Run-II data on the collection plane is shown in Figure \ref{fig:BBandData}  negative polarity data on top, positive polarity data on the bottom.
-
-In MC, we simulate the corresponding positive pion sample with the DDMC (see section \ref{sec:DDMC}) and follow the same steps as in data.
-More details on the calorimetry tuning can be found in \cite{LArIATCalo}.
-
-\textcolor{red}{Add agreement between data and MC for dedx for pions}
-
-\begin{figure}[htb]
-\centering
-\includegraphics[width=0.50\textwidth]{Chapter-5/Images/dEdXvsMomentumTemplate.png}
-\caption{Stopping power for pions, muons, kaons, and protons in liquid argon over the momentum range most relvant for LArIAT according to the Beth-Bloch equation.
-The solid lines represent the prediction for the mean energy $dE/dX$, while the dashed lines are the predictions for the MPV.}
-\label{fig:PDGEnergyLossArgon}
-\end{figure}
-
-\begin{figure}[htb]
-\centering
-\includegraphics[width=0.50\textwidth]{Chapter-5/Images/RunIINegTotaldEdXvsMomentum.png}
-\includegraphics[width=0.50\textwidth]{Chapter-5/Images/RunIIPosTotaldEdXvsMomentum.png}
-\caption{Stopping power versus Momentum for Run-II negative (top) and positive (bottom) polarity data.
-We achieve the agreement between the Bethe-Bloch predictions and the distribution obtained with of the positive pions (top plot, red dots) by tuning the calorimetry constants.
-Once the calorimetry constants are locked in, the agreement between the other particle species and the Bethe-Bloch predictions follows naturally.}
-\label{fig:BBandData}
-\end{figure}
-
-%\begin{figure}[htb]
-%\centering
-%\includegraphics[width=0.50\textwidth]{images/CalibrationExample.png}
-%\caption{Illustration of the calibration technique.
-Here we depict a 325 MeV wire chamber track (shown in green) which enters the TPC (taking into account the energy loss from the upstream material) and we sample the first 12 spacepoints (shown in teal) to extract the dE/dX distribution which is fit with a Landau.}
-%\label{fig:CalibrationExample}
-%\end{figure}
-
 \subsection{Kinetic Energy Measurement}\label{ch:kinEn}
@@
 In this section, we define the measurement on the kinetic energy and  determine the related uncertainty.
-We will propagate this uncertainty into the cross section measurement, as discussed in Section \ref{ch:SysUncertaintyXSRaw} for the pion cross section and in Section \ref{ch:SysUncertaintyXSRawKaon} for the kaon cross section.
+We will propagate this uncertainty into the cross section measurement, as discussed in Section \ref{ch:SysUncertaintyXSRaw} for the pion cross section and in Section \ref{ch:KaonXSRawUnc} for the kaon cross section.
 

### 2018-05-22 Chapter-4/4-Methodology.tex
s

@@
 {\raggedleft ``\emph{Like a lemon to the lime and the bubble to the bee}" \par}
-{\raggedleft -- Eazy-E, Gimmie that Nutt,  1993 -- \par}
+{\raggedleft -- Eazy-E, Gimmie that *,  1993 -- \par}
 \vspace{0.5cm}

### 2018-05-22 Chapter-5/5-Preparatory.tex
s

@@
 
-In order to choose the selection value for the wire chamber momentum, it is beneficial to estimate the ratio of events which capture or decay that survive the selection in MC as a function of the momentum threshold, and compare it with the survival ratio for all events.
+In order to choose the selection value for the wire chamber momentum, it is beneficial to estimate the ratio of events which capture or decay that survive the selection in MC as a function of the momentum threshold, and compare it with the survival ratio for all the 60A events.
 This is done in figure \ref{fig:survRatio}.
@@
 We calculate the survival ratio separately for the three event categories explained above: total (pink), capture (green) and decay (blue).
-Selecting pions with momentum greater than 420 MeV/c reduces the capture events by ~99\% while maintaining about 80\% of the total data sample. 
+Selecting pions with momentum greater than 420 MeV/c reduces the capture events by ~99\% while maintaining about 80\% of the 60A data sample and almost the entire 100A sample. 
 Figure \ref{fig:evtRatio} shows the ratio of events which end their life in capture (green) or decay (blue) over the total number of events as a as a function of the true momentum at wire chamber four.
@@
 We choose a momentum threshold of 420~MeV/c because the percentage of capture events drops below 1\% and the percentage of decays is never above 2\% for momenta greater than 420~MeV/c.
-After the momentum selection, we evaluate the contribution of capture and decay to be a negligibly small background to the cross section measurement compared to the background related to the beamline.
+After the momentum selection, we evaluate the contribution of capture and decay to be a negligibly small background to the cross section measurement compared to the background related to the beamline which we will address in the next section.
 
@@
 The next step is to simulate those pions, muons and electrons from WC4 to the TPC with the DDMC and evaluate their contribution to the cross section.
-To do so, we start by simulating the same number of electrons, muons and pions with the DDMC and we apply the same selection filters on the three samples.
+To do so, we start by simulating the same number of electrons, muons and pions with the DDMC and we apply the same selection chain (i.e.
+track multiplicity rejection, WC2TPC acceptance and shower rejection) on the three samples.
 The number of events per particle species surviving this selection is shown on table \ref{tab:MCafterCutContaminants}.
@@
 The $E_{loss}$ distribution for the 60A  and 100A pion sample is shown in figure \ref{fig:ELoss60A}, left and right respectively.
+The $E_{loss}$ distribution for the  whole kaon sample is shown in figure \ref{fig:ELossKaons}.
 A clear double peaked structure is visible, which is due to the particles either missing or hitting the HALO paddle: a schematic rendering of this occurrence is  shown in figure \ref{fig:Halo}.
 The kinematic at WC4 determines the trajectory of a particle and whether or not it will hit the halo paddle.
-In figure \ref{fig:PxVsXTrue} , we plot the true  horizontal component of the momentum $P_x$ versus the true $X$ position at WC4 for pions missing the halo paddle (left) and for pions hitting the halo paddle (right) for the 60A MC simulation runs -- analogous plots are obtained with the 100A simulation.
+In figure \ref{fig:PxVsXTrue} , we plot the true  horizontal component of the momentum $P_x$ versus the true $X$ position at WC4 for pions missing the halo paddle (left) and for pions hitting the halo paddle (right) for the -60A MC simulation runs -- analogous plots are obtained with the -100A pion simulation and with the kaon simulation.
 These distributions can be separated drawing a line in this position-momentum space.
@@
 We apply the same classifier on data.
+
 A scan of the simulated geometry showed an excess of 3 cm of uninstrumented argon compared with the surveyed detector geometry.
@@
 Pion Data          &  $25 \pm 6$          &    $17 \pm 6 $    \\ \hline
-Kaon  MC          &  $37 \pm 5 $        &     $31 \pm 4 $    \\ \hline
-Kaon Data         &  $26 \pm 6 $        &     $22 \pm 6 $    \\ \hline
+Kaon  MC          &  $38 \pm 6 $        &     $31 \pm 5 $    \\ \hline
+Kaon Data         &  $26 \pm 7 $        &     $22 \pm 7 $    \\ \hline
 \end{tabular}
@@
 \label{fig:ELoss60A}
+\end{figure}
+
+\begin{figure}[hbpt]
+\centering
+\includegraphics[width=0.45\textwidth]{Chapter-5/Images/ELossKaons.pdf}
+\caption{True energy loss between WC4 and the TPC front face according to the MC simulation of positive kaons in the 60A and 100A combined sample.
+The distribution for the whole data sample is shown in blue, the distribution for the kaons missing the halo is shown in red, and the distribution for the kaons hitting the halo is shown in green.
+}
+\label{fig:ELossKaons}
 \end{figure}
@@
 \section{Tracking Studies}\label{sec:TrackingStudies}
-
-The third study is an evaluation of  the angular resolution of the tracking algorithm in data and MC, which is particularly important in the context of the cross section analyses.
+The tracking of hadrons in the TPC determines both the beamline to TPC handshake and the identification of the interaction point within the TPC.
+Thus, it plays a fundamental role in the cross section measurements.
+We performed several studies geared towards the optimization of the package for tracking in the TPC.
+In particular, we studied a suitable set of parameters for the WC2TPC match and we optimized the clustering algorithm to maximize the efficiency of finding the interaction point on MC.
+Given the technical nature of these studies, we report them in Appendix \ref{ch:AppendixTrack}.
+We report here the evaluation of  the angular resolution of the tracking algorithm in data and MC, due to its implication on the physics measurement.
 

### 2018-05-22 Chapter-5/5-Preparatory.tex
g

@@
 \centering
+\includegraphics[width=\textwidth]{Chapter-5/Images/TrackingProcedure.png}
+\caption{A visual representation of the procedure used to evaluate the angular resolution.}
+\label{fig:AngResProcedure}
+\end{minipage}
+\hspace{0.5cm}
+\begin{minipage}[t]{0.45\linewidth}
+\centering
+\includegraphics[width=\textwidth]{Chapter-5/Images/Daughters.png}
+\caption{A visual representation of the scattering angle definition in case of inelastic scattering.}
+\label{fig:scatterPic}
+\end{minipage}
+\end{figure}
+
+\begin{figure}[ht]
+\begin{minipage}[t]{0.45\linewidth}
+\centering
 \includegraphics[width=\textwidth]{Chapter-5/Images/cDAvg.png}
@@
 \centering
-\includegraphics[width=\textwidth]{Chapter-5/Images/TrackingProcedure.png}
-\caption{A visual representation of the procedure used to evaluate the angular resolution.}
-\label{fig:AngResProcedure}
+\includegraphics[width=\textwidth]{Chapter-5/Images/DavgKaons.pdf}
+\caption[]{Distributions of the average distance between each 3D point in space and the fit line,  $\bar d$ for the data used in the kaon cross section analysis and the kaon only DDMC.
+The distributions are area normalized.
+} \label{fig:Chi2AllPts}
 \end{minipage}
@@
 \centering
-\includegraphics[width=\textwidth]{Chapter-5/Images/Daughters.png}
-\caption{A visual representation of the scattering angle definition in case of inelastic scattering.}
-\label{fig:scatterPic}
+\includegraphics[width=\textwidth]{Chapter-5/Images/KaonAngRes.png}
+\caption[]{Distributions of angular resolution $\alpha$ for data used in the kaon cross section analysis and kaon only DDMC. The distributions are area normalized. } \label{fig:trackingResolutionK}
 \end{minipage}

### 2018-05-23 Chapter-5/5-Preparatory.tex
s

@@
 We use the angular resolution of the tracking to determine  the value of smallest angle that we can reconstruct with a non-zero efficiency, effectively determining a selection on the angular distribution of the cross section measurement due to the tracking performance.
-This study is performed on the pion sample, but its results are extrapolated to the kaon case.
+%\textcolor{red}{This study is performed on the pion sample, but its results are extrapolated to the kaon case.}
 
@@
 We decided to use $\bar d$ for its straightforward interpretation.
-The $\bar d$ distribution for data and MC is shown in Figure \ref{fig:Chi2AllPts} and shows a relatively good agreement between data and MC.
+The $\bar d$ distribution for data and MC is shown in Figure \ref{fig:Chi2AllPts} for pions and in Figure \ref{fig:Chi2AllPtsK} for kaons and shows a relatively good agreement between data and MC.
 
@@
 The angle $\alpha$ determines the spatial resolution of the tracking.
-The distributions for data and MC for $\alpha$ are given in Figure \ref{fig:trackingResolution}.
-The mean of the data and MC angular resolution are respectively
+The distributions for data and MC for $\alpha$ are given in Figure \ref{fig:trackingResolution} for pions and in Figure \ref{fig:trackingResolutionK} for kaons.
+The mean of the data and MC angular resolution are reported in Table {tab:AngRes} for pions and kaons in data and MC.
 
-\begin{equation}
-\bar\alpha_{Data} = (5.0 \pm 4.5) \text{ deg}, 
-\end{equation}
-
-\begin{equation}
-\bar\alpha_{MC} = (4.5 \pm 3.9) \text{ deg}. 
-\end{equation}
+\begin{table}[b]
+\centering
+\begin{tabular}{|l|c|c|}  
+\hline
+                      &       Data          & MC     \\ \hline
+Pions             &   $\bar\alpha_{Data} = (5.0 \pm 4.5) \text{ deg}$          & $\bar\alpha_{MC} = (4.5 \pm 3.9) \text{ deg}$ \\ \hline
+Kaons            &   $\bar\alpha_{Data} = (4.3 \pm 3.7) \text{ deg}$          & $\bar\alpha_{MC} = (4.4 \pm 3.6) \text{ deg}$  \\ \hline
+\end{tabular}
+\caption{Angular resolution for Pion and Kaon tracking in both data and MC.}
+\label{tab:AngRes}
+\end{table}
 
@@
 It is beneficial to take a moment to describe the definition of interaction angle.
-In case of elastic scattering, the definition is straightforward: the interaction angle is the angle between the incoming and outgoing pion, i.e.
+In case of elastic scattering, the definition is straightforward: the interaction angle is the angle between the incoming and outgoing hadron, i.e.
 
@@
 In case of inelastic scattering,  the presence of several topologies requires a more complex definition, as shown in figure \ref{fig:scatterPic}.
-We define the scattering angle as the biggest of the angles between the incoming pion and the visible daughters, where the visible daughters are charged particles that travel more than 0.47 cm in the detector (see panel a); in case all the daughters are invisible, the angle is assigned to be 90 deg (see panel b).
-We chose this working definition of scattering angle for inelastic scattering keeping in mind how our tracking reconstruction works: the tracking will stop correctly in case of all the daughters are not visible in the detector and it is likely to stop correctly if multiple daughters form an interaction vertex.
+We define the scattering angle as the biggest of the angles between the incoming hadron and the visible daughters, where the visible daughters are charged particles that travel more than 0.47 cm in the detector (see panel a); in case all the daughters are invisible, the angle is assigned to be 90 deg (see panel b).
+We chose this working definition of scattering angle for inelastic scattering keeping in mind how our tracking reconstruction works: the tracking will stop correctly non of the daughters are is visible in the detector and it is likely to stop correctly if multiple daughters form an interaction vertex.
 The only ``dangerous" case is the production of one charged daughter plus neutrals, which we can study with this working definition of scattering angle (see panel c).
@@
 The distributions are area normalized.
-} \label{fig:Chi2AllPts}
+} \label{fig:Chi2AllPtsK}
 \end{minipage}
@@
 \section{Calorimetry Studies}\label{ch:energyCal} 
-The ability to measure the kinetic energy of hadrons in the TPC is fundamental for the cross section analyses. 
-Thus, we describe first how we calibrate the TPC calorimetric response (Section \ref{ch:energyCalibration}) and how we measure the kinetic energy of the hadrons in the TPC (Section \ref{ch:kinEn}).
+The measured kinetic energy of a hadron candidate at each argon slab determines which bins of the interacting and incident histograms a selected event is going to fill.
+Thus, the energy measurement provided by the LArTPC is fundamental for the cross section analysis.
+In Appendix  \ref{ch:energyCalibration}, we describe how we calibrate the TPC calorimetric response.
+In the following section, we describe how we measure the kinetic energy of the hadrons in the TPC.
 
 \subsection{Kinetic Energy Measurement}\label{ch:kinEn}
-The measured kinetic energy of a hadron candidate at each argon slab determines which bins of the interacting and incident histograms a selected event is going to fill.
 In this section, we define the measurement on the kinetic energy and  determine the related uncertainty.
@@
 } \label{fig:EnergyDeposited}
+\end{figure}
+
+\begin{figure}[htb]
+\centering
+\includegraphics[width=0.48\textwidth]{Chapter-5/Images/KaonEdep.pdf}
+\caption[]{ Energy deposited  $E_{i}$ in a single slab of argon for the kaons of the +60A runs and +100A runs.
+The data is shown in black, the MC in blue.
+The distributions are fitted with a landau displayed in red for data and in teal for MC.
+} \label{fig:EnergyDepositedK}
 \end{figure}

### 2018-05-23 Chapter-6/6-PionCrossSection.tex
s

@@
 
-As stated in section \ref{ch:XSRaw},  the raw cross section is given by the equation \label{eq:thinTargetXSSolved}
+As stated in section \ref{ch:XSRaw},  the raw cross section is given by the equation \ref{eq:thinTargetXSSolved}
 \begin{equation}
@@
 \delta KE_{j} &=& \sqrt{\delta p_{Beam}^2 + \delta E_{Loss}^2 +  \delta  E_{\text{dep FF-j}}^2}\\
-&=& \sqrt{(2\% \text{ }p_{Beam})^2 +  (\sim 6 \text{ [MeV]})^2 +  (j-1)^2 (\sim0.08\text{ [MeV]})^2}.
+&=& \sqrt{(2\% \text{ }p_{Beam})^2 +  ( 6 \text{ [MeV]})^2 +  (j-1)^2 (\sim0.08\text{ [MeV]})^2}.
 \end{eqnarray}

### 2018-05-23 Chapter-7/7-KaonCrossSection.tex
s

@@
 \chapter{Positive Kaon Cross Section Measurement}\label{ch:KaonXS}
+
+In this chapter, we show the result of the thin slice method to measure  the ($K^+$-Ar) total hadronic cross section.
+In Section \ref{ch:KaonXSRaw}, we start by measuring the raw cross section.
+In Section \ref{ch:KaonXSCorrections}, we apply a statistical subtraction of the background contributions based on simulation and a correction for detection inefficiency.
+The final results are presented in Section \ref{ch:FinalKaon}.
+
 \section{Raw Cross Section}\label{ch:KaonXSRaw}
+We measure the raw ($K^+$-Ar) total hadronic cross section as a function of the kinetic energy in the combined +60A and +100A dataset.
+
+Similar to the pion case,  the raw cross section is given by the equation \ref{eq:thinTargetXSSolved}
+\begin{equation}
+ \sigma_{TOT} (E_i)  = \frac{1}{n \delta X}\frac{N^{\text{TOT}}_{\text{Int}}(E_i)}{N^{\text{TOT}}_{\text{Inc}}(E_i)},
+\end{equation}
+
+where $N^{\text{TOT}}_{\text{Int}}$  is the measured number of particles interacting at kinetic energy $E_i$, $N^{\text{TOT}}_{\text{Inc}}$ is the  measured  number of particles incident  on an argon slice at  kinetic energy $E_i$,  $n$ is the density of the target centers  and $\delta X$ is the thickness of the argon slice.
+The density of the target centers and the slab thickness are $n = 0.021\cdot10^{24} \text{ cm}^{-3} $ and  $\delta X=0.47\text{ cm}$, respectively.
+
+As in the case of pions, kaons might decay or interact between WC4 and the TPC front face.
+Some of the interaction products may be wrongly matched to the WC track, forming the ``secondary" particle's background in the kaon sample.
+We estimate the effect of the contamination of secondaries thorugh the DDMC kaon sample.
+Figure \ref{fig:InteractingRawK} shows the distribution of  $N^{\text{TOT}}_{\text{Int}}$  as a function of the kinetic energy.
+The data central points are represented by black dots, the statistical uncertainty is shown in black, while the systematic uncertainty is shown in red.
+Data is displayed over the $N^{\text{TOT}}_{\text{Int}}$  distribution obtained with a DDMC  sample of kaons shot from WC4.
+The contribution from the simulated kaons is shown in pink, the one from secondaries in red.
+The simulated kaon's and secondaries' contributions are stacked; the sum of their integrals is normalized to the integral of the data.
+
+Figure \ref{fig:IncidentRawK} shows the distribution of  $N^{\text{TOT}}_{\text{Inc}}$.
+Data is displayed over the MC.
+The same color scheme and normalization procedure is used for both the interacting and incident histograms.
+
+Figure \ref{fig:XSRawK} shows the raw cross section, statistical uncertainty in black and systematic uncertainty in red.
+The raw data cross section is overlaid to the reconstructed cross section for the MC mixed sample, displayed in azure.
+We calculate the statistical uncertainty for the interacting, incident and cross section distributions in a similar fashion to the pion case as described in Section \ref{ch:StatUncertaintyXSRaw}.
+
+As in the pion case, the only systematic effect considered in the measurement of the raw cross section results from the propagation of the uncertainty associate with the measurement of the kinetic energy at each argon slab.
+For kaons, the uncertainty on the kinetic energy of a candidate at the j$^{th}$ slab of argon  is given by
+
+\begin{eqnarray}
+\delta KE_{j} &=& \sqrt{\delta p_{Beam}^2 + \delta E_{Loss}^2 +  \delta  E_{\text{dep FF-j}}^2}\\
+&=& \sqrt{(2\% \text{ }p_{Beam})^2 +  (7 \text{ [MeV]})^2 +  (j-1)^2 (\sim0.18\text{ [MeV]})^2}.
+\end{eqnarray}
+
+We propagate this uncertainty  by varying the energy measurement $KE_{j}$ at each argon slab.
+We measure $N^{\text{TOT}}_{\text{Inc}}$,  $N^{\text{TOT}}_{\text{Int}}$ and the cross section  in three cases: first assigning the measured $KE_{j}$ at each kinetic energy sampling, then assigning $KE_{j} + \delta KE_{j}$, and finally assigning $KE_{j} - \delta KE_{j}$.
+The difference between the values obtained using the $KE_{j}$ sampling and the maximum and minimum values in each kinetic energy bin determines the systematic uncertainty.
+
+\begin{figure}[]
+\centering
+\begin{minipage}[t]{0.45\textwidth}
+\centering
+\includegraphics[width=\textwidth]{Chapter-7/Images/Plots_MCData_Int_StatSystK.pdf}
+\caption{Raw number of interacting kaon candidates as a function of the reconstructed kinetic energy. The statistical uncertainties are shown in black, the systematic uncertainties in red.}
+\label{fig:InteractingRawK}
+\end{minipage}\hfill
+\begin{minipage}[t]{0.45\textwidth}
+\centering
+\includegraphics[width=\textwidth]{Chapter-7/Images/Plots_MCData_Inc_StatSystK.pdf}
+\caption{Raw number of incident kaon candidates as a function of the reconstructed kinetic energy. The statistical uncertainty is shown in black, the systematic uncertainties in red.}
+\label{fig:IncidentRawK}
+\end{minipage}
+\end{figure}
+
+\begin{figure}
+\centering  
+\includegraphics[width=0.48\textwidth]{Chapter-7/Images/Plots_MCData_XS_StatSystK.pdf}
+\caption{Raw ($K^+$-Ar) total hadronic cross section.
+The statistical uncertainty is shown in black, the systematic uncertainties in red.
+The raw cross section obtained with a MC sample of kaons is shown in azure.
+For the MC cross section,  we include the contributions from secondaries.
+}
+\label{fig:XSRaw}
+\end{figure}
+
+\begin{comment}
+
+\section{Corrections to the Raw Cross Section}\label{ch:KaonXSCorrections}
+As described in section \ref{ch:MCCorrections}, we need to apply a background correction and an efficiency correction in order to derive the true Kaon cross section from the raw cross section.
+The true cross section is given in equation \ref{eq:C},
+
+\begin{equation}
+   \sigma^{\pi^-}_{TOT}(E_{i})  = \frac{1}{n \delta X}\frac{ \epsilon^{\text{Inc}}(E_i)  \hspace{0.2cm} C^{\pi MC}_{\text{Int}} (E_{i}) \hspace{0.2cm} N^{\text{TOT}}_{\text{Int}} (E_{i}) }{   \epsilon^{\text{Int}}(E_i) \hspace{0.2cm} C^{\pi MC}_{\text{Inc}} (E_{i}) \hspace{0.2cm}  N^{\text{TOT}}_{\text{Inc}} (E_{i})}.
+ \tag{\ref{eq:C}}
+\end{equation}
+
+Section \ref{ch:BKGsubXSPuppa} describes the evaluation of Kaon content in the interacting and incident histograms, ($C^{\pi MC}_{\text{Int}} (E_{i})$  and  $C^{\pi MC}_{\text{Inc}} (E_{i})$) and the propagation to the cross section measurement of the relative systematic uncertainties.
+
+Section \ref{ch:EFFXS} describes the procedure employed to obtain  the efficiency corrections $\epsilon^{\text{Int}}(E_i)$  and $\epsilon^{\text{Inc}}(E_i)$ and the propagation to the cross section measurement of the relative uncertainties.
+
+\subsection{Background subtraction}\label{ch:BKGsubXSPuppa}
+We use the procedure described in \ref{ch:KaonXSBkgSub2} to evaluate the relative Kaon content in the interacting histogram $C^{\pi MC}_{\text{Int}} (E_{i})$  and the relative Kaon content in the incident $C^{\pi MC}_{\text{Inc}} (E_{i})$.
+We start by evaluating the relative Kaon content assuming the beamline composition simulated by G4Beamline, whose Kaon, muon and electron percentages per beam condition are reported again in the first line of Table \ref{tab:beamlineSys}.
+The left side of Figure \ref{fig:CorrectionsBeam} shows the  MC estimated  relative Kaon content for the interacting histogram as function of kinetic energy for the 60A runs (top) and 100A runs (bottom).
+The right side of the same figure shows the  MC estimated  relative Kaon content for the incident histogram as function of kinetic energy for the 60A runs (top) and 100A runs (bottom).
+In Figure \ref{fig:CorrectionsBeam} the central curves displayed in light blue are obtained using the beamline composition as predicted by G4Beamline: these are the correction curves for the relative Kaon content applied to data.
+
+So, the question now becomes: how well do we know the beamline composition?
+In absence of additional data constraints,  we take a 100\% systematic uncertainty on the electron content, reported in lines 3 and 4 of Table \ref{tab:beamlineSys}.
+The effect of doubling or halving the electron percentage in the beam on the Kaon relative content is displayed in red in Figure \ref{fig:CorrectionsBeam}.
+We reserve a slightly different treatment for the muon content.
+Since G4Beamline tracks only particles which cross all the wire chambers, Kaon events that decay in flight from WC1 to WC4 are not recorded by G4Beamline.
+Kaon decays in the beamline could be trigger the beamline detectors in data, if the produced muon proceeds in the beamline.
+Thus, we take the G4Beamline prediction for muons as a lower bound in the composition: the effect of doubling the muon content (line 2 in Table \ref{tab:beamlineSys}) is shown in blue on Figure \ref{fig:CorrectionsBeam}.
+A future study of data from additional beamline detectors such as the Aerogel Chernkov detectors \cite{detectorPaper} or the muon range stack (see Section \ref{sec:MuRS}) has the potential of a narrowing the systematics uncertainty coming from the beamline compositon.
+
+We propagate the uncertainty on the beamline composition  as a systematic uncertainty to the cross section by varying the beam composition for all the cases listed in Table \ref{tab:beamlineSys} and evaluating variation of obtained  data cross sections in each bin.
+This systematic uncertainty is summed in quadrature with the statistical uncertainty and the systematic uncertainty related to the kinetic energy measurement.
+
+\begin{table}[p]
+\centering
+\begin{tabular}{| l | l | l | l | l | l | l | l | }
+\hline
+ &  \multicolumn{3}{|c|}{Magnet Current -60A} & \multicolumn{3}{|c|}{Magnet Current -100 A}\\
+
+                                                  & MC $\pi^-$   & MC  $ \mu^-$ & MC  $e^-$ & MC  $\pi^-$ & MC  $\mu^-$ & MC  $e^-$  \\
+\hline
+Expected Composition          & 68.8	\%&4.6 \%&	26.6 \%&	87.4 \%&	3.7	\%&8.9 \% \\
+Composition 2x Muons          & 64.2	\%&9.2 \%&	26.6 \%&	83.7 \%&	7.4	\%&8.9 \% \\
+Composition 2x Electrons      &42.2	\%&4.6 \%&	53.2 \%&	78.5	\%&  3.7	\%&17.8 \%\\
+Composition 0.5x Electrons   &82.1	\%&4.6 \%&	13.3 \%&	91.9 \%&	3.7	\%&4.4 \% \\
+\hline
+\end{tabular}
+\caption{Beam composition variation for the study of systematics due to beam contamination.}
+\label{tab:beamlineSys}
+\end{table}
+
+\begin{figure}[p]
+\centering
+\includegraphics[width=\textwidth]{Chapter-6/Images/Bkg60A_inc_int.pdf}
+\includegraphics[width=\textwidth]{Chapter-6/Images/Bkg100A_inc_int.pdf}
+\caption{\emph{Left:} MC estimated relative Kaon content for interacting histogram a function of kinetic energy for the 60A runs (top) and 100A runs (bottom), statistics uncertainty in azure and systematic uncertainty in blue.
+\emph{Right:}  MC estimated relative Kaon content  for incident histogram a function of kinetic energy for the 60A runs (top)  and 100A (bottom), statistics uncertainty in azure and systematic uncertainty in blue.}
+\label{fig:CorrectionsBeam}
+\end{figure}
+
+\subsection{Efficiency Correction}\label{ch:EFFXS}
+The interaction point for a track used in the total hadronic cross section analysis is defined to be the last point of the WC2TPC matched track which lies inside the fiducial volume.
+This definition is independent from the topology of the interaction.
+If the TPC track stops within the fiducial volume, its last point will be the interaction point, no matter what the products of the interaction look like; if the track crosses the boundaries of the fiducial volume, the track will be considered ``through going" and no interaction point will be found.
+Given this definition, it is evident that we rely on the tracking algorithm to discern where the interaction occurred in the TPC  and correctly stop the tracking.
+The tracking algorithm has an intrinsic angle resolution as shown in section \ref{sec:angleRes}, which limits its efficiency, especially in the case of elastic scattering occurring a low angles.
+Thus, we need to apply an efficiency correction to data in order to retrieve the true cross section.
+The efficiency correction is evaluated separately for the interacting and incident histograms, namely $\epsilon^{\text{int}}_i$ and  $\epsilon^{\text{inc}}_i$, and propagated to the cross section as shown in  equation \ref{eq:C}.
+
+\subsubsection{Efficiency Correction: Procedure}\label{sec:EffCorrection}
+We describe here the procedure to calculate the efficiency correction taking the interacting histogram as example and noting that the procedure is identical for the  incident histogram.
+
+We derive the correction on a set of pure Kaon MC, calculating its value bin by bin as the ratio between the true bin content and the correspondent reconstructed bin content.
+The correction is then applied to the relevant bin in data.
+In formulae, the efficiency correction is calculated to be
+
+\begin{equation}
+ \epsilon^{\text{Int}}(E_i)  =  \frac{N^{\text{ $\pi$ Reco MC}}_{\text{Interacting}} (E_{i})}{ N^{\text{ $\pi$ True MC}}_{\text{Interacting}} (E_{i})  },
+\end{equation}
+
+where $N^{\text{ $\pi$ True MC}}_{\text{Int}} (E_{i}) $ is the content of the $i$-th bin in for the true interacting histogram, and $N^{\text{ $\pi$ Reco MC}}_{\text{Int}} (E_{i}) $ is the content of the $i$-th bin in for the reconstructed interacting histogram.
+The correction is applied to data as follows
+
+\begin{equation}
+N^{\text{ $\pi$ True Data}}_{\text{Int}} (E_{i})  =  \frac{N^{\text{$\pi$ Reco Data}}_{\text{Int}} (E_{i})}{\epsilon^{\text{Int}} (E_{i}) } = N^{\text{$\pi$ Reco Data}}_{\text{Int}} (E_{i}) \frac{N^{\text{ $\pi$ True MC}}_{\text{Int}} (E_{i})}{ N^{\text{ $\pi$ Reco MC}}_{\text{Int}} (E_{i})}.
+\end{equation}
+
+where $N^{\text{$\pi$ Reco Data}}_{\text{Int}} (E_{i})$ is the background subtracted bin content of the $i$-th bin in for the reconstructed interacting histogram for data, i.e. 
+\begin{equation}
+N^{\text{$\pi$ Reco Data}}_{\text{Int}} (E_{i}) =  N^{\text{TOT Data}}_{\text{Int}} (E_{i}) - B^{\text{Data}}_{\text{Int}} (E_i)  =  C^{\text{$\pi$ MC}}_{\text{Int}} (E_{i}) N^{\text{TOT Data}}_{\text{Int}} (E_{i}).
+\end{equation}
+
+In section \ref{sec:angleRes}, we estimated the angular resolution for data and MC to be $\bar\alpha_{Data} = (5.0 \pm 4.5) \text{ deg}$  and 
+$\bar\alpha_{MC} = (4.5 \pm 3.9) \text{ deg}$, respectively.
+Most interaction angles smaller than the angular resolution will thus be indistinguishable  for the reconstruction.
+Thus, we claim we are able to  measure the cross section for interaction angles greater than 5.0 deg.
+Geant4 simulates interactions at all angles, as shown in figure \ref{fig:trueScatteringAngle}.
+In order to calculate the efficiency correction,  we select events which have an interaction angle greater than a given $\alpha_{res}$ to construct the true interacting and incident histograms (the denominator of the efficiency correction).
+The systematics on the efficiency correction is estimated by varying the value of $\alpha_{res}$ between 0 deg and 4.5 deg and propagating the uncertainty on the cross section.
+
+Figure \ref{fig:EffCorr60A} shows $\epsilon^{\text{Int}}(E_{i})$ in the left side and $ \epsilon^{\text{Inc}}(E_i)$ on the right as a function of the kinetic energy for the 60A runs and their systematic uncertainty.
+Similarly, figure \ref{fig:EffCorr100A} shows $\epsilon^{\text{Int}}(E_{i})$ in the left side and $ \epsilon^{\text{Inc}}(E_i)$ on the right as a function of the kinetic energy for the 100A runs and their systematic uncertainty.
+
+\begin{figure}[p]
+\centering
+\includegraphics[width=\textwidth]{Chapter-6/Images/60AEffCorr.pdf}
+\caption{\emph{Left:} Efficiency correction on the 60A interacting histogram, statistical uncertainty in blue, systematic uncertainty in red.
+\emph{Right:}  Efficiency correction on the 60A incident histogram, statistical uncertainty in blue, systematic uncertainty in red.}
+\label{fig:EffCorr60A}
+\end{figure}
+
+\begin{figure}[p]
+\centering
+\includegraphics[width=\textwidth]{Chapter-6/Images/100AEffCorr.pdf}
+\caption{\emph{Left:} Efficiency correction on the 100A interacting histogram, statistical uncertainty in blue, systematic uncertainty in red.
+\emph{Right:}  Efficiency correction on the 100A incident histogram, statistical uncertainty in blue, systematic uncertainty in red.}
+\label{fig:EffCorr100A}
+\end{figure}
+
+\begin{figure}[p]
+\centering
+\includegraphics[width=0.48\textwidth]{Chapter-5/Images//cAngleTrue.png}
+\caption{Distribution of the true scattering angle for a Kaon elastic scattering off the argon nucleus as simulated by Geant4.}
+\label{fig:trueScatteringAngle}
+\end{figure}
+
+\clearpage
+\section{Results}\label{ch:FinalKaon}
+Figure \ref{fig:FinalXSKaon} show the measurement of the ($\pi^-$-Ar) total hadronic cross section for  scattering angles greater than 5$^\circ$, as the result of the background subtraction and efficiency correction to the raw cross section.
+The top left plot is the measurement obtained on the 60A data, statistical uncertainty in black and systematic uncertainty in red.
+The top right plot is the measurement obtained on the 100A data, statistical uncertainty in black and systematic uncertainty in blue.
+The bottom plot shows the two measurements overlaid.
+In all three plot, the Geant4 prediction for the total hadronic cross section for angle scattering greater than 5$^\circ$ is displayed in green.
+
+The systematic uncertainty on the cross section is the sum in quadrature of the statistical uncertainty, the systematic uncertainty related to the kinetic energy measurement, the systematic uncertainty related to the beam composition and the systematic uncertainty related to the efficiency correction.
+
+\begin{figure}[htb]
+\centering
+\includegraphics[width=0.48\textwidth]{Chapter-6/Images/TheMoneyPlot60A.pdf}
+\includegraphics[width=0.48\textwidth]{Chapter-6/Images/TheMoneyPlot100A.pdf}
+\includegraphics[width=0.48\textwidth]{Chapter-6/Images/TheMoneyPlot.pdf}
+\caption{ \emph{Top Left:} ($\pi^-$-Ar) total hadronic cross section for  scattering angles greater than 5$^\circ$ measured in the 60A sample, statistical uncertainty in black and systematic uncertainty in red.
+The Geant4 prediction for the total hadronic cross section for angle scattering greater than 5$^\circ$ is displayed in green.
+\\
+\emph{Top Right:} ($\pi^-$-Ar) total hadronic cross section for  scattering angles greater than 5$^\circ$ measured in the 100A sample, statistical uncertainty in black and systematic uncertainty in blue.
+The Geant4 prediction for the total hadronic cross section for angle scattering greater than 5$^\circ$ is displayed in green.\\
+\emph{Bottom:} ($\pi^-$-Ar) total hadronic cross section measurements in the 60A and 100A samples overlaid with the  Geant4 prediction (green).}
+\label{fig:FinalXSKaon}
+\end{figure}
+
+\subsection{Background subtraction}\label{ch:BKGsubXS}
+
+\subsection{Background Contribution to the Cross Section}\label{sec:Correction}
+
+\subsubsection{Treatment of Systematics}
+
+\end{comment}

### 2018-05-24 Chapter-3/3-TheLArIATExperiment.tex
s

--- 2018-05-14 Chapter-3/3-TheLArIATExperiment.tex
@@
 \chapter{LArIAT: Liquid Argon In A Testbeam}\label{sec:experimentDescription}
+{\raggedleft ``\emph{But, hey we need to be somewhat foolish...} \par}
+{\raggedleft -- Agnes Obel, Avenue, 2010 -- \par}
+\vspace{0.5cm}
+
 In this chapter, we describe the LArIAT experimental setup.

### 2018-05-24 Chapter-6/6-PionCrossSection.tex
s

@@
 \includegraphics[width=\textwidth]{Chapter-6/Images/Bkg100A_inc_int.pdf}
-\caption{\emph{Left:} MC estimated relative pion content for interacting histogram a function of kinetic energy for the 60A runs (top) and 100A runs (bottom), statistics uncertainty in azure and systematic uncertainty in blue.
-\emph{Right:}  MC estimated relative pion content  for incident histogram a function of kinetic energy for the 60A runs (top)  and 100A (bottom), statistics uncertainty in azure and systematic uncertainty in blue.}
+\caption{\emph{Left:} MC estimated relative pion content for interacting histogram a function of kinetic energy for the 60A runs (top) and 100A runs (bottom), predicted background content in azure and muon and electron content variation in blue and red.
+\emph{Right:}  MC estimated relative pion content  for incident histogram a function of kinetic energy for the 60A runs (top)  and 100A (bottom), predicted background content in azure and muon and electron content variation in blue and red}
 \label{fig:CorrectionsBeam}

### 2018-05-24 Chapter-7/7-KaonCrossSection.tex
s

--- 2018-05-13 Chapter-7/7-KaonCrossSection.tex
@@
 \chapter{Positive Kaon Cross Section Measurement}\label{ch:KaonXS}
+
+In this chapter, we show the result of the thin slice method to measure  the ($K^+$-Ar) total hadronic cross section.
+In Section \ref{ch:KaonXSRaw}, we start by measuring the raw cross section.
+In Section \ref{ch:KaonXSCorrections}, we apply a statistical subtraction of the background contributions based on simulation and a correction for detection inefficiency.
+The final results are presented in Section \ref{ch:FinalKaon}.
+
 \section{Raw Cross Section}\label{ch:KaonXSRaw}
+We measure the raw ($K^+$-Ar) total hadronic cross section as a function of the kinetic energy in the combined +60A and +100A dataset.
+
+Similar to the pion case,  the raw cross section is given by the equation \ref{eq:thinTargetXSSolved}
+\begin{equation}
+ \sigma_{TOT} (E_i)  = \frac{1}{n \delta X}\frac{N^{\text{TOT}}_{\text{Int}}(E_i)}{N^{\text{TOT}}_{\text{Inc}}(E_i)},
+\end{equation}
+
+where $N^{\text{TOT}}_{\text{Int}}$  is the measured number of particles interacting at kinetic energy $E_i$, $N^{\text{TOT}}_{\text{Inc}}$ is the  measured  number of particles incident  on an argon slice at  kinetic energy $E_i$,  $n$ is the density of the target centers  and $\delta X$ is the thickness of the argon slice.
+The density of the target centers and the slab thickness are $n = 0.021\cdot10^{24} \text{ cm}^{-3} $ and  $\delta X=0.47\text{ cm}$, respectively.
+
+As in the case of pions, kaons might decay or interact between WC4 and the TPC front face.
+Some of the interaction products may be wrongly matched to the WC track, forming the ``secondary" particle's background in the kaon sample.
+We estimate the effect of the contamination of secondaries through  the DDMC kaon sample.
+Figure \ref{fig:InteractingRawK} shows the distribution of  $N^{\text{TOT}}_{\text{Int}}$  as a function of the kinetic energy.
+The data central points are represented by black dots, the statistical uncertainty is shown in black, while the systematic uncertainty is shown in red.
+Data is displayed over the $N^{\text{TOT}}_{\text{Int}}$  distribution obtained with a DDMC  sample of kaons shot from WC4.
+The contribution from the simulated kaons which interact hadronically is shown in pink, the contributions from kaon decay is shown in orange and the one from secondaries in red. 
+The simulated kaon's and secondaries' contributions are stacked; the sum of their integrals is normalized to the integral of the data.
+
+Figure \ref{fig:IncidentRawK} shows the distribution of  $N^{\text{TOT}}_{\text{Inc}}$.
+Data is displayed over the MC.
+For the  $N^{\text{TOT}}_{\text{Inc}}$ distribution we do not make a distinction between kaons that decay or interact hadronically because any kaon independently from its final interaction contributes to the flux of incident particles at given kinetic energy.
+The same normalization procedure is used for both the interacting and incident histograms.
+
+Figure \ref{fig:XSRawK} shows the raw cross section, statistical uncertainty in black and systematic uncertainty in red.
+The raw data cross section is overlaid to the reconstructed cross section for the MC mixed sample, displayed in azure.
+We calculate the statistical uncertainty for the interacting, incident and cross section distributions in a similar fashion to the pion case as described in Section \ref{ch:StatUncertaintyXSRaw}.
+
+As in the pion case, the only systematic effect considered in the measurement of the raw cross section results from the propagation of the uncertainty associate with the measurement of the kinetic energy at each argon slab.
+For kaons, the uncertainty on the kinetic energy of a candidate at the j$^{th}$ slab of argon  is given by
+
+\begin{eqnarray}
+\delta KE_{j} &=& \sqrt{\delta p_{Beam}^2 + \delta E_{Loss}^2 +  \delta  E_{\text{dep FF-j}}^2}\\
+&=& \sqrt{(2\% \text{ }p_{Beam})^2 +  (7 \text{ [MeV]})^2 +  (j-1)^2 (\sim0.18\text{ [MeV]})^2}.
+\end{eqnarray}
+
+We propagate this uncertainty  by varying the energy measurement $KE_{j}$ at each argon slab.
+We measure $N^{\text{TOT}}_{\text{Inc}}$,  $N^{\text{TOT}}_{\text{Int}}$ and the cross section  in three cases: first assigning the measured $KE_{j}$ at each kinetic energy sampling, then assigning $KE_{j} + \delta KE_{j}$, and finally assigning $KE_{j} - \delta KE_{j}$.
+The difference between the values obtained using the $KE_{j}$ sampling and the maximum and minimum values in each kinetic energy bin determines the systematic uncertainty.
+
+\begin{figure}[]
+\centering
+\begin{minipage}[t]{0.45\textwidth}
+\centering
+\includegraphics[width=\textwidth]{Chapter-7/Images/Plots_MCData_Int_StatSystK_WithDK.pdf}
+\caption{Raw number of interacting kaon candidates as a function of the reconstructed kinetic energy. The statistical uncertainties are shown in black, the systematic uncertainties in red.}
+\label{fig:InteractingRawK}
+\end{minipage}\hfill
+\begin{minipage}[t]{0.45\textwidth}
+\centering
+\includegraphics[width=\textwidth]{Chapter-7/Images/Plots_MCData_Inc_StatSystK_WithDK.pdf}
+\caption{Raw number of incident kaon candidates as a function of the reconstructed kinetic energy. The statistical uncertainty is shown in black, the systematic uncertainties in red.}
+\label{fig:IncidentRawK}
+\end{minipage}
+\end{figure}
+
+\begin{figure}
+\centering  
+\includegraphics[width=0.48\textwidth]{Chapter-7/Images/Plots_MCData_XS_StatSystK_WithDK.pdf}
+\caption{Raw ($K^+$-Ar) total hadronic cross section.
+The statistical uncertainty is shown in black, the systematic uncertainties in red.
+The raw cross section obtained with a MC sample of kaons is shown in azure.
+For the MC cross section,  we include the contributions from secondaries.
+}
+\label{fig:XSRawK}
+\end{figure}
+
+\section{Corrections to the Raw Cross Section}\label{ch:KaonXSCorrections}
+As described in section \ref{ch:MCCorrections}, we need to apply a background correction and an efficiency correction in order to derive the true Kaon cross section from the raw cross section.
+The true cross section is given in equation \ref{eq:C},
+
+\begin{equation}
+   \sigma^{K^+}_{TOT}(E_{i})  = \frac{1}{n \delta X}\frac{ \epsilon^{\text{Inc}}(E_i)  \hspace{0.2cm} C^{K MC}_{\text{Int}} (E_{i}) \hspace{0.2cm} N^{\text{TOT}}_{\text{Int}} (E_{i}) }{   \epsilon^{\text{Int}}(E_i) \hspace{0.2cm} C^{K MC}_{\text{Inc}} (E_{i}) \hspace{0.2cm}  N^{\text{TOT}}_{\text{Inc}} (E_{i})}.
+ \tag{\ref{eq:C}}
+\end{equation}
+
+Currently, the only background considered for the kaon hadronic cross section comes from the presence of secondaries.
+A further development of the analysis will need to account for the presence of a small proton contamination.
+Figure \ref{fig:CorrectionsBeamK} shows the relative kaon content for the interacting and incident histograms.
+\begin{figure}[p]
+\centering
+\includegraphics[width=\textwidth]{Chapter-7/Images/KaonBkgSub_WithDK.pdf}
+\caption{\emph{Left:} MC estimated relative kaon content for kaons interacting hadronically as function of kinetic energy.
+\emph{Right:}  MC estimated relative kaon content  for incident histogram a function of kinetic energy.}
+\label{fig:CorrectionsBeamK}
+\end{figure}
+
+As described in \ref{ch:EFFXS} for the pion case, we derive the correction on a set of pure kaon MC, calculating its value bin by bin as the ratio between the true bin content and the correspondent reconstructed bin content.
+The correction is then applied to the relevant bin in data.
+The efficiency correction is evaluated separately for the interacting and incident histograms, namely $\epsilon^{\text{int}}_i$ and  $\epsilon^{\text{inc}}_i$, and propagated to the cross section as shown in  equation \ref{eq:C}.
+
+In section \ref{sec:angleRes}, we estimated the angular resolution for data and MC to be $\bar\alpha_{Data} = (4.3 \pm 3.7) \text{ deg}$  and  $\bar\alpha_{MC} = (4.4 \pm 3.6) \text{ deg}$, respectively.
+Most interaction angles smaller than the angular resolution will thus be indistinguishable  for the reconstruction.
+Thus, we claim we are able to  measure the cross section for interaction angles greater than 4.5 deg.
+Geant4 simulates interactions at all angles: in order to calculate the efficiency correction,  we select events which have an interaction angle greater than a $\alpha_{res}$ to construct the true interacting and incident histograms (the denominator of the efficiency correction).
+The systematics on the efficiency correction is estimated by varying the value of $\alpha_{res}$ between 0 deg and 4.5 deg and propagating the uncertainty on the cross section.
+
+Figure \ref{fig:EffCorrK} shows $\epsilon^{\text{Int}}(E_{i})$ in the left side and $ \epsilon^{\text{Inc}}(E_i)$ on the right as a function of the kinetic energy for the kaon sample and their systematic uncertainty.
+
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{Chapter-7/Images/EffCorrK.pdf}
+\caption{\emph{Left:} Efficiency correction on the interacting histogram, statistical uncertainty in blue, systematic uncertainty in red.
+\emph{Right:}  Efficiency correction on the incident histogram, statistical uncertainty in blue, systematic uncertainty in red.}
+\label{fig:EffCorrK}
+\end{figure}
+
+\section{Results}\label{ch:FinalKaon}
+\section{Results}\label{ch:FinalKaon}
+Figure \ref{fig:FinalXSKaon} show the measurement of the ($K^+$-Ar) total hadronic cross section for  scattering angles greater than 5$^\circ$, as the result of the background subtraction and efficiency correction to the raw cross section.
+The plot shows the measurement obtained on the full dataset, statistical uncertainty in black and systematic uncertainty in red.
+The Geant4 prediction for the total hadronic cross section for angle scattering greater than 5$^\circ$ is displayed in green.
+
+The systematic uncertainty on the cross section is the sum in quadrature of the statistical uncertainty, the systematic uncertainty related to the kinetic energy measurement, the systematic uncertainty related to the beam composition and the systematic uncertainty related to the efficiency correction.
+
+\begin{figure}[htb]
+\centering
+\includegraphics[width=0.48\textwidth]{Chapter-7/Images/TheMoneyPlotK.pdf}
+\caption{ ($K^+$-Ar) total hadronic cross section for  scattering angles greater than 5$^\circ$ measured in the 60A sample, statistical uncertainty in black and systematic uncertainty in red.
+The Geant4 prediction for the total hadronic cross section for angle scattering greater than 5$^\circ$ is displayed in green.
+}
+\label{fig:FinalXSKaon}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{comment}
+
+\clearpage
+\section{Results}\label{ch:FinalKaon}
+
+\end{comment}

### 2018-05-24 Chapter-1/1-Theory.tex
s

@@
 
-\section{The Standard Model}
+\begin{figure}[h!]
+\includegraphics[scale=0.2,right]{Chapter-1/Images/Bach.png}
+\end{figure}
+
+{\raggedleft -- J. S. Bach, Suite No. 1 in G major, BWV 1007, 1720 ca. -- \par}
+\vspace{0.5cm}
+
+In this chapter, we set the ($\pi^-$- Ar) and ($K^+$- Ar) total hadronic cross section measurements into the greater theoretical and phenomenological framework.
+We start by briefly describing the Standard Model (Section \ref{ch:SM}), with particular attention to neutrinos and neutrino interactions (Section \ref{ch:nus} ).
+We then describe some of the open questions in neutrino physics today and Beyond Standard Model theories (\ref{ch:BMS}) setting the stage for the measurements reported in this work (Section \ref{ch:MotHadXS}).
+
+\section{The Standard Model}\label{ch:SM}
 The Standard Model (SM) of particle physics is the most accurate theoretical description of the subatomic world and, in general, one of the most precisely tested theories in the history of physics.
@@
 
-The Standard Model is a gauge theory based on the local group of symmetry
+The Standard Model is a gauge theory based on the local symmetry group
 \begin{equation}
-G_{SM} = SU(3)_C  \otimes SU(2)_T \otimes U(1)_Y
+G_{SM} = SU(3)_C  \otimes SU(2)_L \otimes U(1)_Y
 \label{eq:SMGroup}
@@
 
-where the subscripts indicate the conserved charges: the strong charge, or color C, the weak isospin T (or rather its third component T3) and the hypercharge Y.
-These quantities can be related to the electric charge Q through the Gell-Mann-Nishijima relation:
+where the subscripts C indicates the conserved strong charge (color),   and the subscripts Y indicates the conserved hypercharge. 
+If we indicated with T the weak isospin T and with T3  its third component, hypercharge can be related to the electric charge Q through the Gell-Mann-Nishijima relation:
 \begin{equation}
@@
 
-In the quantum field framework, the elementary particles correspond to the irreducible representations of the G$_{SM}$ symmetry group.
+In the quantum field framework, the SM fields correspond to the irreducible representations of the G$_{SM}$ symmetry group.
 In particular, the particles are divided in two categories, fermions and bosons, according to their spin-statistics.
@@
 \end{tabular}
-\caption{SM elementary fermions. The subscripts L and R indicate respectively the negative helicity (left-handed) and the positive helicity (right-handed).}
+\caption{SM elementary fermionic fields. The subscripts L and R indicate respectively the negative chirality (left-handed) and the positive chirality (right-handed).}
 \label{tab:SMParticles}
@@
 
-Quarks can interact via all three the fundamental forces; they are triplets of SU(3)$_C$, that is they can exist in three different colors: C = R, G, B.
-If one chooses a base where $u$, $c$ and $t$ quarks are simultaneously eigenstates of both the strong and the weak interactions, the remaining eigenstates are usually written as $d$, $s$ and $b$ for the strong interaction and $d'$, $s'$ and $b'$ for the weak interaction, because the latter ones are the result of a Cabibbo rotation on the first ones.
+Quarks can interact via all three the fundamental forces; they are triplets of SU(3)$_C$, that is they can exist in three different colors.
+If one chooses a base where $u$, $c$ and $t$ quarks are simultaneously eigenstates of both the strong and the weak interactions, the remaining eigenstates are usually written as $d$, $s$ and $b$ for the strong interaction and $d'$, $s'$ and $b'$ for the weak interaction, because the latter ones are the result of a CKM rotation on the first ones.
 Charged leptons interact via the weak and the electromagnetic forces, while neutrinos only interact via the weak force. 
@@
 A gauge theory by itself cannot provide a description of massive particles, but it is experimentally well know that most of the elementary particles have non-zero masses.
-The introduction of massive fields in the Standard Model lagrangian would make the theory non-renormalizable, and - so far - mathematically impossible to handle.
+The introduction of massive fields in the Standard Model lagrangian would make the theory not gauge invariant, resulting ill-defined.
 This problem is solved in the SM by the introduction of a scalar iso-doublet $\Phi(x)$, the Higgs field, which gives mass to W$^\pm$ and Z$^0$ gauge bosons through the electroweak symmetry breaking mechanism and to the fermions through Yukawa coupling \cite{Higgs1964,Higgs19642}.
@@
 
-\section{Neutrinos:  tiny cracks in the Standard Model}
+\section{Neutrinos:  tiny cracks in the Standard Model}\label{ch:nus}
+To our current knowledge, neutrinos are the most abundant fermion in the Universe.
+And yet, they are maybe the most mysterious particle in the SM: they generate theoretical puzzles and experimental challenges.
+In this section, we treat neutrinos within and beyond the SM and describe the make up of their interaction with matter.
+
 \subsection{Neutrinos in the Standard Model}
-Neutrino were introduced in the SM as left-handed massless Weyl spinors.
-The Dirac equation of motion
+Neutrino can be introduced in the SM as left-handed massless Weyl spinors.
+The Dirac equation of motion for a free field
 \begin{equation}
@@
 Pauli initially rejected the description of a physical particle through a single Weyl spinor because of its implication of parity violation.
-In fact, since the spatial inversion operator throws $\psi_R \leftrightarrow \psi_L$, parity is conserved only if the both the chiral components exist at the same time.
+In fact, since the spatial inversion operator throws $\psi_R \leftrightarrow \psi_L$, parity is conserved only if both chiral components exist at the same time.
 For the neutrino introduction in the SM, experiments came in help of the theoretical description.
@@
 
-The symmetry group $SU(2)_T \otimes U(1)_Y$ is the only group relevant for neutrino interactions.
-The SM electroweak lagrangian is the most general renormalizable lagrangian invariant under the local symmetry group $SU(2)_T \otimes U(1)_Y$.
+The symmetry group $SU(2)_L \otimes U(1)_Y$ is the only group relevant for neutrino interactions.
+The SM electroweak lagrangian is the most general renormalizable lagrangian invariant under the local symmetry group $SU(2)_L \otimes U(1)_Y$.
 The lagrangian couples the weak isotopic spin doublets and singlets described in Table \ref{tab:SMParticles} with the gauge bosons  $A^{\mu}_{a}$ ($a$ $=$ 1,2,3) and $B^{\mu}$, and Higgs doublet $\Phi(x)$:
@@
 This phenomenon, called ``neutrino oscillations",  was predicted first by Bruno Pontecorvo in 1957 ~\cite{Pontecorvo:1967fh}.
-Neutrino oscillations are possible only if the neutrino flavor eigenstate are not identical to the mass eigenstates, thus resulting in the first evidence of physics beyond the Standard Model.
+Neutrino oscillations are possible only if the neutrino flavor eigenstate are not identical to the mass eigenstates.
+Thus, the observation of neutrino oscillation results in the first evidence of physics beyond the Standard Model.
 A minimal extension of the SM introduces three mass eigenstates, $\ket{\nu_i}$ ($i = 1,2, 3$), whose mass $m_i$ is well defined.
-The unitary Pontecorvo-Maki-Nakagawa-Sakata matrix transforms the spinor wave functions ($\psi$) of each component  between the flavor and mass bases as follows
+The unitary Pontecorvo-Maki-Nakagawa-Sakata matrix transforms the mass base into the flavor base as follows
 
 \begin{equation} 
-\sum_\alpha \psi_\alpha \ket{\nu_\alpha} =  \sum_i \psi_i \ket{\nu_i}, \rightarrow \psi_\alpha =  U_{PMNS} \psi_i, 
+ \ket{\nu_\alpha} =   U_{PMNS} \ket{\nu_i},
 \end{equation}
@@
 $\theta_{23}$            & 45$^\circ$                      & 9.0\%   & Super Kamiokande, MINOS,\\
-$\Delta m^2_{23}$    & 2.5 $10^{-3}$ eV$^2$   & 1.8\% &  No$\nu$a, MACRO \\
+|$\Delta m^2_{32}$|    & 2.5 $10^{-3}$ eV$^2$   & 1.8\% &  No$\nu$a, MACRO \\
 \hline
 $\theta_{12}$            & 34$^\circ$                     & 5.8\% &  SNO,Gallex,\\
-$\Delta m^2_{12}$    & 7.4 $10^{-5}$ eV$^2$   & 2.8\% &  SAGE,  KamLAND \\
+|$\Delta m^2_{12}$|    & 7.4 $10^{-5}$ eV$^2$   & 2.8\% &  SAGE,  KamLAND \\
 \hline
 $\theta_{13}$            &   9$^\circ$                    & 4.7\%  & DAYA  Bay, \\
-$\Delta m^2_{13}$    & 2.5 $10^{-3}$ eV$^2$   & 1.8\% &  RENO\\
+|$\Delta m^2_{32}$|    & 2.5 $10^{-3}$ eV$^2$   & 1.8\% &  RENO\\
 \hline
@@
 Historically, neutrino interactions with the nucleus in the GeV region are divided into three categories into three categories whose contributions change as a function of increasing neutrino energy:: quasi elastic (QE), resonant (RES), and deep inelastic (DIS) scattering.
-All current and forthcoming oscillation experiments live in the 0.1-10 GeV transition region, which encompasses the energy where the QE neutrino-nucleus interaction transitions into RES and then into DIS.
+All current and forthcoming oscillation experiments on neutrino beams live in the 0.1-10 GeV transition region, which encompasses the energy where the QE neutrino-nucleus interaction transitions into RES and then into DIS.
 For scattering off free nucleons, neutrino and antineutrino QE charge current scattering refers to the process $\nu_l n \rightarrow l^- p$ and $\bar\nu_l p \rightarrow l^+ n$ where a charged lepton and single nucleon are ejected in the elastic interaction.
@@
 
-\section{Beyond the Standard Model}
+\section{Beyond the Standard Model}\label{ch:BMS}
 The discovery of neutrino oscillation and its implication of non-zero neutrino mass mark  the beginning of a new, exciting era in neutrino physics: the era of physics Beyond the Standard Model (BSM) at the intensity frontier.
@@
 \ref{eq:SMGroup}.
-\textcolor{red}{Drawing a parallel with the electroweak symmetry breaking mechanism, where the $ SU(2)_T \otimes U(1)_Y$ is recovered from $U(1)_{EM}$, an interesting line of simplification for the SM group structure would be to devise a similar mechanism where $SU(3)_C  \otimes SU(2)_T \otimes U(1)_Y$ is recovered from an hypothetical larger group.
+\textcolor{red}{Drawing a parallel with the electroweak symmetry breaking mechanism, where the $ SU(2)_L \otimes U(1)_Y$ is recovered from $U(1)_{EM}$, an interesting line of simplification for the SM group structure would be to devise a similar mechanism where $SU(3)_C  \otimes SU(2)_L \otimes U(1)_Y$ is recovered from an hypothetical larger group.
 IS THIS CORRECT?} Just as the electroweak unification becomes evident at energies higher than the Higgs VEV,  a direct manifestation of Grand Unification Theories (GUTs) would occur at even higher energies.
@@
 
-\textbf{SU(5).} The smallest simple group containing $SU(3)_C  \otimes SU(2)_T \otimes U(1)_Y$ is SU(5), as shown first by Georgi and Glashow in~\cite{PhysRevLett.32.438}.
+\textbf{SU(5).} The smallest simple group containing $SU(3)_C  \otimes SU(2)_L \otimes U(1)_Y$ is SU(5), as shown first by Georgi and Glashow in~\cite{PhysRevLett.32.438}.
 Quarks and leptons in this group fit the $\bar 5$ and 10 representations.
@@
 
-%The $SU(5)$ group is the smallest simple group containing $SU(3)_C  \otimes SU(2)_T \otimes U(1)_Y$. Georgi and Glashow proposed the ``minimal SU(5)"
+%The $SU(5)$ group is the smallest simple group containing $SU(3)_C  \otimes SU(2)_L \otimes U(1)_Y$. Georgi and Glashow proposed the ``minimal SU(5)"
 
@@
 
-\section{Motivations for Hadronic Cross Sections in Argon}
+\section{Motivations for Hadronic Cross Sections in Argon} \label{ch:MotHadXS} 
 Critical challenges await  the next decade of high energy physics at the intensity frontier.

### 2018-05-24 Chapter-2/2-LArTPCs.tex
s

@@
 \chapter{Liquid Argon Detectors at the Intensity Frontier}\label{ch:2}
-{\raggedleft ``\emph{Beat-up little seagull, on a marble stair} \par}
-{\raggedleft \emph{Tryin' to find the ocean, lookin' everywhere.}"\par}
-{\raggedleft -- Nina Simone, Baltimore, 1978 -- \par}
+{\raggedleft ``\emph{Don't you know, honey,} \par}
+{\raggedleft \emph{Ain't nobody ever gonna love you, the way I try to do?}"\par}
+{\raggedleft -- Janis Joplin, Cry Baby, 1971 -- \par}
 \vspace{0.5cm}
@@
 In the next few years, LArTPCs will  be the tools to answer some of the burning questions in neutrino physics today.
-This section illustrates the operational principles of this detector technology, as well as the scope of the key detectors in the US liquid argon program -- SBN, DUNE and LArIAT.
+This chapter illustrates the operational principles of this detector technology, as well as the scope of the key detectors in the US liquid argon program -- SBN, DUNE and LArIAT.
 
 \section{The Liquid Argon Time Projection Chamber Technology}
+In this section, we outline an extremely brief history of Time Projection Chambers as particle detectors, focusing on their incarnation as Argon detectors for neutrino physics.
+We further describe the working principles of Liquid Argon Time Projection Chambers, leading  to the description of the event reconstruction in LArTPC.
+
 \subsection{TPCs, Neutrinos \& Argon}
@@
 \section{The Intensity Frontier Program}
-\subsection{Prospects for LArTPCs in Neutrino Physics: SBN and DUNE}
-The ArgoNeut  experiment \cite{ArgoNeuT-det} initiated the US LArTPC neutrino program.
-Following the success of this small TPC on the NuMI beam, a wide program of LArTPCs on neutrino beams has flourished.
+This section highlights the role of Liquid Argon Time Projection Chambers at the Intensity frontier.
+In particular, we show the prospects for the exploration of neutrino physics (Section \ref{ch:NuPhysLAr}) and GUT models (Section \ref{ch:GutsLAr}) in current and forthcoming LAr experiments.
+In Section \label{ch:LArIATIntro}, we introduce LArIAT and its role in the Intensity Frontier panorama.
+
+\subsection{Prospects for LArTPCs in Neutrino Physics: SBN and DUNE}\label{ch:NuPhysLAr}
+The ArgoNeut  experiment \cite{ArgoNeuT-det} together the LAr R\&D experiments TallBo and the Yale TPC  initiated the US LArTPC neutrino program.
+Following the success of the ArgoNeut small TPC on the NuMI beam, a wide program of LArTPCs on neutrino beams has flourished.
 The construction of  LArTPCs as near and far detectors at different baseline allows for the exploration of some of the fundamental questions in neutrino physics today illustrated in section \ref{ch:questions}.
@@
 
-\subsection{Prospects for LArTPCs in GUT Physics: DUNE}
+\subsection{Prospects for LArTPCs in GUT Physics: DUNE}\label{ch:GutsLAr}
 The experimental exploration of a manifestation of Grand Unified Theory is possible in DUNE thanks to its sheer mass.
@@
 %\subsection{Nucleon Decay Detection in LAr}
-\subsection{Enabling the next generation of discoveries: LArIAT}
+\subsection{Enabling the next generation of discoveries: LArIAT}\label{ch:LArIATIntro}
 LArIAT, a small LArTPC in a test beam,  is designed to perform an extensive physics campaign centered on charged particle cross section measurements while characterizing the detector performance for future LArTPCs.

### 2018-05-24 Chapter-3/3-TheLArIATExperiment.tex
s

@@
 \chapter{LArIAT: Liquid Argon In A Testbeam}\label{sec:experimentDescription}
-{\raggedleft ``\emph{But, hey we need to be somewhat foolish...} \par}
+{\raggedleft ``\emph{But, hey we need to be somewhat foolish...}" \par}
 {\raggedleft -- Agnes Obel, Avenue, 2010 -- \par}
@@
 For LArIAT Run II, the High Yield sample is about three times the Picky Tracks statistics.
+We assume an uncertainty of 2\% for four-point WC track, momentum uncertainty as reported for the same beamline in \cite{MinervaTestbeam}.
 
 %We use Picky Tracks to calibrate the momentum measurement for the High Yield sample, in particular to obtain a momentum correction for  tracks missing information from the central WCs;  this correction adds an uncertainty of approximately 2\% to the momentum calculation of the High Yield sample.
-
-\textcolor{red}{Four point track momentum uncertainty}
 
@@
 \section{In the Cryostat}
+The heart of the LArIAT experiment lives in the LArIAT cryostat.
+In this section, we describe the cryogenic system and the argon purity (Section \ref{ch:Cryo}), the LArIAT TPC (Section \ref{sec:TPCCharge}) and light collection system (\ref{sec:TPCLight}).
 \subsection{Cryogenics and Argon Purity}\label{ch:Cryo}

### 2018-05-24 Chapter-7/7-KaonCrossSection.tex
s

@@
 \chapter{Positive Kaon Cross Section Measurement}\label{ch:KaonXS}
+{\raggedleft ``\emph{Beat-up little seagull, on a marble stair} \par}
+{\raggedleft \emph{Tryin' to find the ocean, lookin' everywhere.}"\par}
+{\raggedleft -- Nina Simone, Baltimore, 1978 -- \par}
+\vspace{0.5cm}
 
@@
 \section{Results}\label{ch:FinalKaon}
-\section{Results}\label{ch:FinalKaon}
 Figure \ref{fig:FinalXSKaon} show the measurement of the ($K^+$-Ar) total hadronic cross section for  scattering angles greater than 5$^\circ$, as the result of the background subtraction and efficiency correction to the raw cross section.
@@
 
-The systematic uncertainty on the cross section is the sum in quadrature of the statistical uncertainty, the systematic uncertainty related to the kinetic energy measurement, the systematic uncertainty related to the beam composition and the systematic uncertainty related to the efficiency correction.
+The systematic uncertainty on the cross section is the sum in quadrature of the statistical uncertainty, the systematic uncertainty related to the kinetic energy measurement and the systematic uncertainty related to the efficiency correction.
 

### 2018-05-24 Chapter-0/0-Introduction.tex
s

--- 2018-01-30 Chapter-0/0-Introduction.tex
@@
-\chapter{Introduction}
+\chapter{Introduction}\label{ch:-1}
 
-\section{The Standard Model}
-The Standard Model (SM) of particle physics the most accurate theoretical description of the subatomic world and, more generically, one of the most precisely tested theories in the history of physics.
-The SM describes the strong, electromagnetic and weak interactions among  elementary particles in the framework of quantum field theory, accounting for the unification of electromagnetic and weak interactions for energies above the  vacuum expectation value of the Higgs field.
-The SM does not describe gravity or general relativity.
+This thesis work concerns the first measurement of the  ($\pi^-$-Ar)  total hadronic cross section  in the 100-1000 MeV  kinetic energy range and the first measurement of the ($K^+$-Ar) total hadronic cross section  in the 100-650 MeV  kinetic energy range.
+We performed these measurements at the LArIAT experiment,  a small (0.25 ton)  Liquid Argon Time Projection Chamber (LArTPC) on a beam of charge particles at the Fermilab Test Beam Facility.
+Albeit particle and nuclear physics have a long history of hadronic cross section measurements, the work outlined in this thesis presents a new methodology -- the ``thin slice method" -- for cross section measurements possible only thanks to the excellent tracking and calorimetry capabilities of the LArTPC technology.
+The LArTPC technology allows to see unprecedented details of particle interactions in argon
 
-The Standard Model is a gauge theory based on the local group of symmetry
-\begin{equation}
-G_{SM} = SU(3)_C  \otimes SU(2)_T \otimes U(1)_Y
-\end{equation}
+A renewed interest for precision measurements of hadronic cross sections in argon arises from the current panorama of experimental neutrino physics.
 
-where the subscripts indicate the conserved charges: the strong charge, or color C, the weak isospin T (or rather its third component T3) and the hypercharge Y.
-These quantities can be related to the electric charge Q through the Gell-Mann-Nishijima relation:
-\begin{equation}
-Q = \frac{Y}{2} + T_3.
-\end{equation}
+The discovery of the Higgs boson in 2012 marked the triumph of the Standard Model of Particle Physics; exploring what lays beyond is the real challenge in our field today. 
+Since their formulation in 1930, neutrinos have been a source of surprises (and Nobel Prizes) for particle physicists, tiny cracks in our understanding of Nature.
+In particular, the discovery of neutrino oscillation represents the first evidence of physics Beyond the Standard Model (BSM).
+From a theoretical point of view, the field is developing new theories to account for the small but non-zero mass of neutrinos, while trying to remain consistent with the rest of the Standard Model.
+From an experimental point of view, we are developing technologies and huge collaborations to probe these theories.
+Even today, neutrinos might hold the key to the next generation of discoveries in particle physics.
 
-In the quantum field framework, the elementary particles correspond to the irreducible representations of the G$_{SM}$ symmetry group.
-In particular, the particles are divided in two categories, fermions and bosons, according to their spin-statistics.
-Described by the Fermi-Dirac statistics, Fermions have half-integer spin and are sometimes called ``matter-particles".
-Bosons or ``force carriers" have integer spin, follow the Bose-Einstein statistics and mediate the interaction between fermions.
-The fundamental fermions and their quantum numbers are listed in Tab \ref{tab:SMParticles}.
+According to the latest Particle Physics Project Prioritization Panel  \cite{P5}, the US particle physics panorama is directing a substantial effort towards the exploration of the Intensity Frontier.
+In particular, the near future will see the development of a Short Baseline Neutrino Program (SBN) and long baseline neutrino program with  DUNE as far detector.
+Both these program are based on the Liquid Argon Time Projection Chamber (LArTPC) detector technology.
+The US liquid argon program has the potential to answer many of the fundamental open questions in particle physics today, such as: is there a fourth generation neutrino?
+is CP violated in the lepton sector?
+are there any additional symmetries?
+And, can we find an indication of Grand Unified Theories?
 
-\begin{table}[]
-\centering
-\begin{tabular}{|cccc|c|c|c|}\hline
-Generation               & I                   & II                  & III                 & T                       & Y                   & Q                   \\\hline
+The SBN program at Fermilab is tasked with conclusively addressing the existence of a fourth neutrino generation in the  $\Delta m^2= \Delta m^2_{14} \sim [0.1 - 10]$ eV$^2$ parameter space.
+The SBN program entails three surface LArTPCs positioned on the Booster Neutrino Beam at different distances from the neutrino production in oder to fully exploit  the L/E dependence of the oscillation pattern:  SBND (100 m from the decay pipe), MicroBooNE (450 m), and ICARUS (600 m).
+SBN will also perform an extensive
+program of neutrino cross section measurements, fundamental to abate systematics in the oscillation analyses in both SBN and DUNE.
 
-\multirow{6}{*}{Leptons}                          &                     &                     &                     &                         &                     &                     \\
-& $\begin{pmatrix}\ \nu_e\\ e \end{pmatrix}_L$ & $\begin{pmatrix}\ \nu_\mu\\ \mu \end{pmatrix}_L$ & $\begin{pmatrix}\ \nu_\tau\\ \tau \end{pmatrix}_L$ & $\begin{matrix}\ 1/2\\ -1/2 \end{matrix}$ & $\begin{matrix}\ -1\\ -1 \end{matrix}$ & $\begin{matrix}\ 0\\ -1 \end{matrix}$ \\
-                         &                     &                     &                     &                         &                     &                     \\
-                         & $e_R$         & $\mu_R$      & $\tau_R$                  & 0                      & -2                  & 1    \\
-                         &                     &                     &                     &                         &                     &                     \\\hline
-\multirow{7}{*}{Quarks} 
-                         &                     &                     &                     &                         &                     &                     \\
-& $\begin{pmatrix}\ u\\ d' \end{pmatrix}_L$ & $\begin{pmatrix}\ c\\ s' \end{pmatrix}_L$ & $\begin{pmatrix}\ t\\ b' \end{pmatrix}_L$ & $\begin{matrix}\ 1/2\\ -1/2 \end{matrix}$ & $\begin{matrix}\ 1/3\\ 1/3 \end{matrix}$ & $\begin{matrix}\ 2/3\\ -1/3 \end{matrix}$ \\
-                         &                     &                     &                     &                         &                     &                     \\
-& $\begin{matrix}\ u_R\\ d'_R \end{matrix}$ & $\begin{matrix}\ c_R\\ s'_R \end{matrix}$ & $\begin{matrix}\ t_R\\ b'_R \end{matrix}$ & $\begin{matrix}\ 0\\ 0 \end{matrix}$ & $\begin{matrix}\ 4/3\\ -2/3 \end{matrix}$ & $\begin{matrix}\ 2/3\\ -1/3 \end{matrix}$ \\
-                         &                     &                     &                     &                         &                     &                    \\\hline
-\end{tabular}
-\caption{SM elementary fermions. The subscripts L and R indicate respectively the negative helicity (left-handed) and the positive helicity (right-handed).}
-\label{tab:SMParticles}
-\end{table}
+The long baseline neutrino program and DUNE in particular has a wide physics program.
+For what it concerns neutrino physics, DUNE has the capability of solving the mass hierarchy and octant problem,  and discovering CP violation in the neutrino sector.
+Besides its neutrino program, DUNE can open an experimental window on Grand Unified Theories (GUTs).
+GUTs could potentially answer fundamental questions such as the existence of non-zero neutrino masses and matter-antimatter asymmetry, explaining some ``accidents" in the Standard Models, such as the exact cancellation of the  proton and the electron charge.
+Probing GUTs at the unification energy scale is impossible by any foreseeable collider experiment.
+An indirect proof is thus needed such as baryon number violation, which is predicted by almost every GUT in the form of proton decay, bounded nucleon decay or $n-\bar n$ oscillations on long time-scales.
+Historically, the dominant technology used in these searches has been water Cherenkov detectors, with Super-Kamiokande setting all the current experimental limits on the decay lifetimes at the order of $\sim 10 ^{34}$ years.
+The DUNE far detector and its non-accelerator physics program is a interesting new actor on this stage.
+LArTPCs can in fact complement nucleon decay searches in modes where water Cherenkov detectors are less sensitive, especially $p\rightarrow K^+\bar{\nu}$  \cite{Adams:2013qkq}.
 
-Quarks can interact via all three the fundamental forces; they are triplets of SU(3)$_C$, that is they can exist in three different colors: C = R, G, B.
-If one chooses a base where $u$, $c$ and $t$ quarks are simultaneously eigenstates of both the strong and the weak interactions, the remaining eigenstates are usually written as $d$, $s$ and $b$ for the strong interaction and $d'$, $s'$ and $b'$ for the weak interaction, because the latter ones are the result of a Cabibbo rotation on the first ones.
-Charged leptons interact via the weak and the electromagnetic forces, while neutrinos only interact via the weak force. 
-The gauge group univocally determines the number of gauge bosons that carry the interaction; the gauge bosons correspond to the generators of the group: eight gluons (g) for the strong interaction, one photon ($\gamma$) and three bosons (W$^\pm$, Z$^0$) for the electroweak interaction.
-A gauge theory by itself can not provide a description of massive particles, but it is experimentally well know that most of the elementary particles have non-zero masses.
-The introduction of massive fields in the Standard Model lagrangian would make the theory non-renormalizable, and - so far - mathematically impossible to handle.
-This problem is solved in the Standard Model by the introduction of a scalar iso-doublet $\Phi(x)$, the Higgs field, which gives mass to W$^\pm$ and Z$^0$ gauge bosons through the electroweak symmetry breaking and to the fermions through Yukawa coupling \cite{Higgs1964,Higgs19642}.
+\textcolor{red}{Brief history of nu detectors}
 
-\section{Neutrinos in the Standard Model}
-The neutrino was introduced in the SM as a left-handed massless Weyl spinor.
-The Dirac equation of motion
-\begin{equation}
-(i\gamma^ \mu \partial_\mu - m) \psi = 0
-\end{equation}
-for a fermionic field 
-\begin{equation}
- \psi =  \psi_L +  \psi_R
-\end{equation}
-is equivalent to the equaitons
-\begin{equation}
-i\gamma^ \mu \partial_\mu  \psi_L = m \psi_R
-\end{equation}
-\begin{equation}
-i\gamma^ \mu \partial_\mu  \psi_R = m \psi_L
-\end{equation}
+\begin{itemize}
+%\item Challenges Beyond the SM
+%\item Many in text generation of neutrino detector
+\item Brief history of nu detectors
+\item Calibration : LArIAT
+\item The measurements
+\end{itemize}
 
-for the chiral fields $\psi_R$ and $\psi_L$, whose evolution in space and time is coupled through the mass $m$.
-If the fermion is massless, the chiral fields decouple and the fermion can be described by a single Weyl spinor with two independent components.
-Pauli initially rejected the description of a physical particle through a single Wyle spinor because of its implication of parity violation.
-In fact, since the spatial inversion operator throws $\psi_R \leftrightarrow \psi_L$, parity is conserved only if the both the chiral components exist at the same time.
-\textcolor{red}{ADD CITATIONS}.
-For the neutrino introduction in the SM, experiments came in help of the theoretical description:  the constraint of parity conservation weakened after Wu's experiment \textcolor{red}{ADD CITATIONS AND DATES}, there was no experimental indication for massive neutrinos and neutrinos likely interacted only via the left-handed component.
+LArTPCs provide excellent electron/photon separation \cite{Acciarri:2016sli} lacking in Cherenkov detectors which can be leveraged to abate the photon background from neutral current interactions  in $\nu_e$ searches.
+LArTPCs share superb tracking capability with bubble chamber detectors, with several additional benefits.
+They are electronically read out and self triggered detectors; they provide full 3D-imaging with millimeter resolution, precise calorimetric reconstruction and excellent particle identification.
+Plus, female physicists are actually writing code and running experiments, not (only) staring at images \textcolor{red}{-- Too much?
+--} \cite{2006physics4152G}.
 
-The symmetry group $SU(2)_T \otimes U(1)_Y$ is the only group relevant for neutrino interactions.
-The SM electroweak lagrangian is the most general renormalizable lagrangian invariant under the local symmetry group $SU(2)_T \otimes U(1)_Y$.
-The lagrangian couples the weak isotopic spin doublets and singlets described in \ref{tab:SMParticles} with the gauge bosons  $A^{\mu}_{a}$ ($a$ $=$ 1,2,3) and $B^{\mu}$, and Higgs doublet $\Phi(x)$:
+The complexity of the LArTPC technology for neutrino detection is due to several reasons.
+Argon is a fairly heavy element, which means that nuclear effects play an important role in the looks of the interaction topology.
+Also, data on charged particle interaction in argon is scarce;  no measurement of hadronic cross sections for pions or kaons is available for argon.
+Secondly,
 
-\begin{eqnarray}
-\lefteqn{\mathcal{L} = i\sum_{\alpha=e,\mu,\tau} \bar{L}'_{\alpha L}  \slashed D L'_{\alpha L} + 
- i\sum_{\alpha=1,2,3} \bar{Q}'_{\alpha L}  \slashed D Q'_{\alpha L} {}}
- \nonumber\\
- & & {} + i\sum_{\alpha=e,\mu,\tau} \bar{l}'_{\alpha R}  \slashed D l'_{\alpha R} + i\sum_{\alpha=d,s,b} \bar{q}'^D_{\alpha R}  \slashed D q'^D_{\alpha R} + i\sum_{\alpha=u,c,t} \bar{q}'^U_{\alpha R}  \slashed D q'^U_{\alpha R}
- \nonumber\\
- & & {} -\frac{1}{4}A_{\mu \nu}A^{\mu \nu} - \frac{1}{4}B_{\mu \nu}B^{\mu \nu}
- \nonumber\\
- & & {} +(D_{\rho}\Phi)^\dagger(D^{\rho}\Phi) - \mu^2\Phi^\dagger\Phi - \lambda(\Phi^\dagger\Phi)^2 
- \nonumber\\
- & & {} -\sum_{\alpha,\beta=e,\mu,\tau} \Big(Y'^l_{\alpha\beta}\bar{L}'_{\alpha L}  \Phi l'_{\beta R} + Y'^{l*}_{\alpha\beta}\bar{l}'_{\beta R}  \Phi^\dagger L'_{\alpha L}\Big)
-  \nonumber\\
- & & {} -\sum_{\alpha=1,2,3} \sum_{\beta=d,s,b} \Big(Y'^D_{\alpha\beta}\bar{Q}'_{\alpha L}  \Phi q'^D_{\beta R} + Y'^{D*}_{\alpha\beta}\bar{q}'^D_{\beta R}  \Phi^\dagger Q'_{\alpha L}\Big)
-  \nonumber\\
- & & {} -\sum_{\alpha=1,2,3} \sum_{\beta=u,c,t} \Big(Y'^U_{\alpha\beta}\bar{Q}'_{\alpha L}   \widetilde{\Phi} q'^U_{\beta R} + Y'^{U*}_{\alpha\beta}\bar{q}'^U_{\beta R} \widetilde{\Phi}^\dagger Q'_{\alpha L}\Big).
-\end{eqnarray}
+LArIAT \cite{Cavanna:2014iqa}, a small (0.25 ton) LArTPC on a beam of charge particles at the Fermilab Test Beam Facility, is performing precise cross section measurements of charged particles in argon.
+The LArIAT suite of cross section measurements is a critical set of results which will be employed by both the short- and long- baseline neutrino programs.
+This work is a measurement of the negative pion argon and the positive kaon-argon total hadronic differential cross sections.
 
-The first two lines of the lagrangian summarize the kinetic terms for the fermionic fields and their coupling to the gauge bosons $A^{\mu\nu}_a$, $B^{\mu\nu}$ \footnote{In gauge theories the ordinary derivative $\partial_\mu$  is substitued with èthe covariant derivative $D_\mu$.
-Here $D_\mu = \partial_\mu + igA_\mu \cdot I + ig'B_\mu\frac{Y}{2}$, where I and Y are the SU(2)$_L$ and U(1)$_Y$ generators, respectively.}.
-The third line describes the kinetic terms and the self-coupling terms of the gauge bosons.
-The forth line is the Higgs lagrangian, which results in the spontaneous symmetry breaking.
-The last three lines describe the Yukawa coupling between fermions and the Higgs field, origin of the fermion's mass.
+The ($\pi^-$-Ar) hadronic cross section is a fundamental input for neutrino detectors in liquid argon. In fact, reconstructing neutrino interactions at the GeV scale where pion production is abundant.
 
-The coupling between left-handed and right-handed field generates the mass term for fermions.
-The SM assumes only left-handed components for neutrinos, thus implying zero neutrino mass.
-Since any linear combination of massless fields results in a massless field, the flavor eigenstates are identical to the mass eigenstates in the SM.
+The   total hadronic differential cross sections is of particular interest for proton decay searches in DUNE, whose most interesting proton decay channel is $p\rightarrow K^+ \bar \nu$.
 
-\section{Beyond the Standard Model}
-The discovery of neutrino oscillation and its implication of non-zero neutrino mass marks  the beginning of a new, exciting era in neutrino physics: the era of physics Beyond the Standard Model (BSM) in the neutrino sector.
-We are currently searching for new, deeper theories that can accommodate neutrinos with non-zero mass, while remaining consistent with the rest of the Standard Model.
-%More excitingly, we need to probe these theories experimentally.
-\subsection{Neutrino Oscillations}
-\subsection{Proton Decay}
+The cross section analyses exploit the totality of LArIAT's experimental handles; they rely on beam line detector information as well as both calorimetry and tracking in the TPC.
+These analyses are LArIAT's first physics results.
+
+The  ($K^+$-Ar) total hadronic differential cross section in LArIAT is particularly relevant for a high identification efficiency in the context of proton decay searches in DUNE in the  $p\rightarrow K^+\bar{\nu}$  channel.
+In fact, the kaon-argon cross section affects the kaon topology by modifying the kaon tracking and energy reconstruction, impacting the basis for kaon identification in a LArTPC.
+
+A series of additional studies and calibrations were necessary to perform the cross section analyses.
+Appendix \ref{ch:AppendixB} shows a measurement of the LArIAT LArTPC electric field using cosmic data.
+Appendix \ref{ch:AppendixB} shows an optimization of the tracking algorithms geared towards maximizing the efficiency of finding the hadronic interaction point.
+Appendix \ref{ch:AppendixB} shows the calorimetry calibration of the LArIAT LArTPC, which is a pivotal measurement to enable any physics analysis with TPC data.
+
+In order to measure the kaon-argon differential cross section, several steps were necessary.
+The analysis starts by identifying a sample of kaons in the beam line and assessing the beam line contaminations.
+It proceeds with tracking the kaon candidates in the TPC and measuring their calorimetry.
+Then,  the hadronic interaction point is identified and distinguished from weak interaction cases.
+The last step of the analysis, which is the energy unfolding, is now ongoing.
+With  this measurement, I acquired extensive experience in tracking hadrons and measuring their energy inside the TPC.
+This entailed both an in-depth study of the tracking algorithms performances on data and a dive into the GEANT4 transportation models.
+Since no previous kaon measurement has been performed on argon, simulation packages such as GEANT4 currently interpolate the kaon transportation on (scarce) data from lighter and heavier nuclei.
+This is suboptimal for proton decay searches.
+In fact, an extremely high signal detection efficiency, measured by testing identification algorithms on simulation, is extremely difficult to obtain if the simulation is not reliable.
+By adding the experimental data points for kaon-argon cross section to the simulation, my thesis analysis will cover this gap of knowledge.
+
+Chapters:\ref{ch:TheTheory}
+Chapter \ref{ch:2}
+Chapter \ref{sec:experimentDescription}

### 2018-05-24 Chapter-2/2-LArTPCs.tex
s

@@
 {\raggedleft \emph{Ain't nobody ever gonna love you, the way I try to do?}"\par}
-{\raggedleft -- Janis Joplin, Cry Baby, 1971 -- \par}
+{\raggedleft -- Janis Joplin,  1971 -- \par}%Cry Baby,
 \vspace{0.5cm}

### 2018-05-24 Chapter-3/3-TheLArIATExperiment.tex
s

@@
 {\raggedleft ``\emph{But, hey we need to be somewhat foolish...}" \par}
-{\raggedleft -- Agnes Obel, Avenue, 2010 -- \par}
+{\raggedleft -- Agnes Obel, 2010 -- \par} %, Avenue,
 \vspace{0.5cm}

### 2018-05-24 Chapter-4/4-Methodology.tex
s

@@
 {\raggedleft ``\emph{Like a lemon to the lime and the bubble to the bee}" \par}
-{\raggedleft -- Eazy-E, Gimmie that *,  1993 -- \par}
+{\raggedleft -- Eazy-E,   1993 -- \par}%Gimmie that *,
 \vspace{0.5cm}

### 2018-05-24 Chapter-5/5-Preparatory.tex
s

@@
 {\raggedleft \emph{Abbracciami}"\par}
-{\raggedleft -- Pietro Ciampi, L'amore e' tutto qui, 1971 -- \par}
+{\raggedleft -- Pietro Ciampi,  1971 -- \par}%L'amore e' tutto qui,
 \vspace{0.5cm}

### 2018-05-24 Chapter-6/6-PionCrossSection.tex
s

@@
 {\raggedleft \emph{que busca entre la tiniebla, la dulce hoguera del beso.}"\par}
-{\raggedleft -- Lila Downs, Benediction And Dream,  2002 -- \par}
+{\raggedleft -- Lila Downs,   2002 -- \par}%Benediction And Dream,
 \vspace{0.5cm}

### 2018-05-24 Chapter-7/7-KaonCrossSection.tex
s

@@
 {\raggedleft \emph{Tryin' to find the ocean, lookin' everywhere.}"\par}
-{\raggedleft -- Nina Simone, Baltimore, 1978 -- \par}
+{\raggedleft -- Nina Simone, 1978 -- \par}% Baltimore,
 \vspace{0.5cm}

### 2018-05-25 Chapter-0/0-Introduction.tex
s

@@
-\chapter{Introduction}\label{ch:-1}
+\chapter*{Introduction}\label{ch:-1}
 
 This thesis work concerns the first measurement of the  ($\pi^-$-Ar)  total hadronic cross section  in the 100-1000 MeV  kinetic energy range and the first measurement of the ($K^+$-Ar) total hadronic cross section  in the 100-650 MeV  kinetic energy range.
-We performed these measurements at the LArIAT experiment,  a small (0.25 ton)  Liquid Argon Time Projection Chamber (LArTPC) on a beam of charge particles at the Fermilab Test Beam Facility.
-Albeit particle and nuclear physics have a long history of hadronic cross section measurements, the work outlined in this thesis presents a new methodology -- the ``thin slice method" -- for cross section measurements possible only thanks to the excellent tracking and calorimetry capabilities of the LArTPC technology.
-The LArTPC technology allows to see unprecedented details of particle interactions in argon
-
-A renewed interest for precision measurements of hadronic cross sections in argon arises from the current panorama of experimental neutrino physics.
+We performed these measurements at the LArIAT experiment,  a small (0.25 ton)  Liquid Argon Time Projection Chamber (LArTPC) on a beam of charged particles at the Fermilab Test Beam Facility.
+Albeit particle and nuclear physics have a long history of hadronic cross section measurements, the work outlined in this thesis presents a new methodology -- the ``thin slice method" -- for cross section measurements, possible only thanks to the detection capabilities of the LArTPC technology.
+The LArTPC technology allows to see unprecedented details of particle interactions in argon given its full 3D-imaging with millimeter resolution and precise calorimetric reconstruction.
+A renewed interest for precision measurements of hadronic cross sections, particularly in argon, arises from the current  panorama of experimental  particle physics at the intensity frontier.
 
@@
 
-According to the latest Particle Physics Project Prioritization Panel  \cite{P5}, the US particle physics panorama is directing a substantial effort towards the exploration of the Intensity Frontier.
+According to the latest Particle Physics Project Prioritization Panel  \cite{P5}, the US particle physics panorama is directing a substantial effort towards the exploration of the intensity frontier.
 In particular, the near future will see the development of a Short Baseline Neutrino Program (SBN) and long baseline neutrino program with  DUNE as far detector.
-Both these program are based on the Liquid Argon Time Projection Chamber (LArTPC) detector technology.
+Both these programs are based on the Liquid Argon Time Projection Chamber (LArTPC) detector technology.
 The US liquid argon program has the potential to answer many of the fundamental open questions in particle physics today, such as: is there a fourth generation neutrino?
@@
 
-The long baseline neutrino program and DUNE in particular has a wide physics program.
-For what it concerns neutrino physics, DUNE has the capability of solving the mass hierarchy and octant problem,  and discovering CP violation in the neutrino sector.
+DUNE has a vast neutrino and non-accelerator physics reach.
+For what it concerns neutrino physics, oscillation analyses in DUNE have the capability of solving the mass hierarchy and octant problem,  and discovering CP violation in the neutrino sector.
 Besides its neutrino program, DUNE can open an experimental window on Grand Unified Theories (GUTs).
 GUTs could potentially answer fundamental questions such as the existence of non-zero neutrino masses and matter-antimatter asymmetry, explaining some ``accidents" in the Standard Models, such as the exact cancellation of the  proton and the electron charge.
-Probing GUTs at the unification energy scale is impossible by any foreseeable collider experiment.
-An indirect proof is thus needed such as baryon number violation, which is predicted by almost every GUT in the form of proton decay, bounded nucleon decay or $n-\bar n$ oscillations on long time-scales.
+Directly probing GUTs at the unification energy scale is impossible by any foreseeable collider experiment.
+We then need an indirect proof such as baryon number violation, which is predicted by almost every GUT in the form of proton decay, bounded nucleon decay or $n-\bar n$ oscillations on long time-scales.
 Historically, the dominant technology used in these searches has been water Cherenkov detectors, with Super-Kamiokande setting all the current experimental limits on the decay lifetimes at the order of $\sim 10 ^{34}$ years.
@@
 
-LArIAT \cite{Cavanna:2014iqa}, a small (0.25 ton) LArTPC on a beam of charge particles at the Fermilab Test Beam Facility, is performing precise cross section measurements of charged particles in argon.
+LArIAT \cite{Cavanna:2014iqa}, a small (0.25 ton) LArTPC on a beam of charged particles at the Fermilab Test Beam Facility, is performing precise cross section measurements of charged particles in argon.
 The LArIAT suite of cross section measurements is a critical set of results which will be employed by both the short- and long- baseline neutrino programs.
@@
 In fact, an extremely high signal detection efficiency, measured by testing identification algorithms on simulation, is extremely difficult to obtain if the simulation is not reliable.
-By adding the experimental data points for kaon-argon cross section to the simulation, my thesis analysis will cover this gap of knowledge.
+By adding the experimental data points for kaon-argon cross section to the simulation, my thesis analysis will cover this gap of knowledge.\\
 
-Chapters:\ref{ch:TheTheory}
-Chapter \ref{ch:2}
-Chapter \ref{sec:experimentDescription}
+This body of work is divided in 8 chapters.
+We provide a description of the theoretical framework for the measurements in  Chapter \ref{ch:TheTheory}. Chapter \ref{ch:2} outlines the LArTPC detector technology, while
+Chapter \ref{sec:experimentDescription} describes LArIAT experimental setup.
+We present the event selection for both the pion and kaon analyses, as well as the ``thin-slice method" in Chapter \ref{ch:Interactions}.
+Chapter \ref{ch:samples}  describes the work done on the data and Monte Carlo samples in preparation of the cross section analyses.
+Chapter \ref{ch:PionXS} shows  the results for the ($\pi^-$-Ar) total hadronic cross section measurement.
+Chapter \ref{ch:KaonXS} shows  the results for the ($K^+$-Ar) total hadronic cross section measurement.
+We draw the final remarks on this work in Chapter \ref{ch:Conclusions}

### 2018-05-25 thesis_main.tex
s

@@
 \usepackage{hyperref}
+\usepackage[export]{adjustbox}
 
@@
 
-\setcounter{chapter}{-1}
+%\setcounter{chapter}{-1}
 
@@
 % Need to define title before the abstract.
-\title{Measurement of total hadronic differential cross sections in the LArIAT experiment}
+\title{Measurement of ($\pi^-$-Ar) and ($K^+$-Ar) total hadronic cross sections in the LArIAT experiment}
 \author{Elena Gramellini}
@@
 \maketitle
-\makecopyright{2017}
+\makecopyright{2018}
 
@@
 
-%Bonnie. I don't even know where to start.
-%Flavio
-%
+I jumped the pond a little less than 5 years ago because I had a taste of what being a physicists in the US might mean and wanted more. 
+I came to this vast (and at times inhospitable) land thinking I would learn a lot of physics (which I did) and trying to keep doing research for as long as I could (which I'm fortunate enough to say I'll still do for a while).
+Little did I know that getting my PhD would be first and foremost a journey to self discovery, the beginning of a path to asymptotically tend to the person and the scientist I want to be.
+A journey impossible without the help of many people, which I will now proceed to thank, hoping that the acknowledgements won't be longer than the thesis itself.
 
-%A big thank you to my academic sisters, a bunch of extraordinary women who have taught me so much about myself and our field.
-%From the youngest to the oldest (again academically speaking): Supraja, you worked the magic of enriching my past, thank you from the bottom of my heart;  Brooke, the depth of your knowledge, your drive and love for our field has been and will always be a constant inspiration;  Ariana, thanks for keeping me sane for the last four years, if you run for president, remember to delete  our Skype chats, love and hugs to you dudette; and last but not least Xiao, I don't think Fermilab has given me a bigger gift than your friendship.
+Mom, you are and will always be the hero of my story; I know this has been challenging for you, thanks for being there for me at every hour of the day and the night, for being always honest with me and for somehow handling this whole US-Italy thing gracefully.
+I can't believe you still haven't met Bonnie.
+Ah, and don't worry, I'll always translate things for you.
 
-%Shany
-%Jen 
-%Jonathan
-%Andrzej
-%Anto
+Bonnie Fleming, my advisor, I don't even know where to start in thanking you.
+Thanks for listening to me before I proved I was worth be listened to; thanks for always believing I could make it, even at the bottom of my doubts; thanks for the laughters, thanks for leading by example, thanks for  letting me part of the most wonderful research group ;
+
+%Flavio Cavanna 
+%Ornella Palamara, I once told Andrzej I think of you as my spirit animal: fierce, independent and direct; he replied I still need to eat a lot of paninis to get there and I can't agree more.
+
+%A big thank you to my academic sisters, from the youngest to the oldest (again academically speaking): Supraja Balasubramanian, you worked the magic of enriching my past, thank you from the bottom of my heart;  Brooke Morrison, the depth of your knowledge, your drive and love for our field has been and will always be a constant inspiration;  Ariana  Hackenburg, thanks for keeping me sane for the last four years, if you run for president -- as you should -- remember to delete our Skype chats, love and hugs to you dudette; and last but not least Xiao Luo, Fermilab has given me lots of extraordinary gifts, but no one  bigger than your friendship.
+
+%Jen Raaf
+%Jonathan Asaadi
+%Andrzej M. Szelc
+%Roberto Acciarri
 %LArIAT & MicroBooNE
-%Louise
+%To all the members of the NPG, thanks for all the laughters and all the wise insights. You're the smartest bunch of idiots I've ever met, I love y'all.
+%Shany %Louise
+
+%To the extraordinary people who stayed close, even with an ocean in between: la Tasso, Marci, la Bea, la Kittirn, Alvin, la Viola e Brogna, Sighi and all the people from the ``La Chiappia Nuova" group, thanks for making me feel like I never left.
+%Finally, Anto, cause people might think I'm nice and easy, but you know I'm not. And you're still here.
 
@@
 \mainmatter
-%\input{Chapters/0-Introduction}
-%\input{Chapter-1/1-Theory}
-%\input{Chapter-2/2-LArTPCs}
-%\input{Chapter-3/3-TheLArIATExperiment}
+\input{Chapter-0/0-Introduction}
+\input{Chapter-1/1-Theory}
+\input{Chapter-2/2-LArTPCs}
+\input{Chapter-3/3-TheLArIATExperiment}
 \input{Chapter-4/4-Methodology}
-
 \input{Chapter-5/5-Preparatory}
@@
 \input{Chapter-7/7-KaonCrossSection}
+\input{Chapter-8/8-Conclusions}
 
-%\input{Chapter-7/7-BackgroundSub}
-%\input{Chapter-8/8-CrossSectionAnalysis}\input{Chapter-9/9-Uncertainty}
+%
+%\input{Chapter-9/9-Uncertainty}
 
@@
 \appendix
-%\input{Chapters/A-KaonAna}
-\input{AppendixB-EField/B-LArIATEField}
-%\input{Chapters/C-MicroBooNECRT}
+%%\input{Chapters/A-KaonAna}
+\input{AppendixB-TrackingStudies/B-Tracking}
+\input{AppendixC-EnergyCalibration/C-EnergyCalibration}
+%\input{AppendixB-EField/B-LArIATEField}
+%%\input{Chapters/C-MicroBooNECRT}
 

### 2018-05-25 Chapter-0/0-Introduction.tex
s

@@
-\chapter*{Introduction}\label{ch:-1}
+`\chapter{Introduction}\label{ch:-1}
 
@@
 We performed these measurements at the LArIAT experiment,  a small (0.25 ton)  Liquid Argon Time Projection Chamber (LArTPC) on a beam of charged particles at the Fermilab Test Beam Facility.
-Albeit particle and nuclear physics have a long history of hadronic cross section measurements, the work outlined in this thesis presents a new methodology -- the ``thin slice method" -- for cross section measurements, possible only thanks to the detection capabilities of the LArTPC technology.
-The LArTPC technology allows to see unprecedented details of particle interactions in argon given its full 3D-imaging with millimeter resolution and precise calorimetric reconstruction.
+Albeit particle and nuclear physics have a long history of hadronic cross section measurements, the work outlined in this thesis presents a new methodology -- the ``thin slice method" -- for cross section measurements in argon, possible only thanks to the detection capabilities of the LArTPC technology.
+The combination of fine-grained tracking and excellent calorimetric information provided by the LArTPC technology  allows to see unprecedented details of particle interactions in argon and, in LArIAT, to measure the kinetic energy of a hadron at each step of the tracking.
 A renewed interest for precision measurements of hadronic cross sections, particularly in argon, arises from the current  panorama of experimental  particle physics at the intensity frontier.
@@
 From an experimental point of view, we are developing technologies and huge collaborations to probe these theories.
-Even today, neutrinos might hold the key to the next generation of discoveries in particle physics.
+As we enter the era of precision measurements of neutrino interaction, neutrinos might hold the key to the next generation of discoveries in particle physics.
 
-According to the latest Particle Physics Project Prioritization Panel  \cite{P5}, the US particle physics panorama is directing a substantial effort towards the exploration of the intensity frontier.
-In particular, the near future will see the development of a Short Baseline Neutrino Program (SBN) and long baseline neutrino program with  DUNE as far detector.
-Both these programs are based on the Liquid Argon Time Projection Chamber (LArTPC) detector technology.
+Experimentally, precision measurements can be achieved only if the detector technology is able to resolve the fine details of a statistically relevant number of interactions. 
+With ``fine details" here we mean the ability to distinguish the many products of the neutrino interaction, such as protons, pions, muons and electrons, and to measure their energy.
+Historically,  bubble chamber neutrino detectors were the first revolution in neutrino detection: for example, the spatial resolution of Gargamelle allowed the discovery of neutrino neutral current interaction.
+Despite the high precision of bubble chambers images, this technology is hard to scale to massive size, making statistical analyses on neutrino interactions almost impossible to perform.
+To make up for the small neutrino interaction cross section, neutrino experiments moved to very large size, at the expenses of spatial precision.
+This is the case for the detectors which discovered neutrino oscillation:  both Super-Kamiokande and SNO are massive Cherenkov detectors.
+With LArTPCs, the field is gaining again bubble-chamber like precision but at massive scales.
+Following the recommendations of  the latest Particle Physics Project Prioritization Panel  \cite{P5}, the US particle physics panorama is directing a substantial effort towards the exploration of the intensity frontier through the construction of massive LArTPCs.
+In particular, the near future will see the development of a Short Baseline Neutrino Program (SBN) and long baseline neutrino program  (DUNE), both based on the LArTPC detector technology.
 The US liquid argon program has the potential to answer many of the fundamental open questions in particle physics today, such as: is there a fourth generation neutrino?
@@
 are there any additional symmetries?
-And, can we find an indication of Grand Unified Theories?
+and, can we find an indication of Grand Unified Theories?
 
@@
 We then need an indirect proof such as baryon number violation, which is predicted by almost every GUT in the form of proton decay, bounded nucleon decay or $n-\bar n$ oscillations on long time-scales.
-Historically, the dominant technology used in these searches has been water Cherenkov detectors, with Super-Kamiokande setting all the current experimental limits on the decay lifetimes at the order of $\sim 10 ^{34}$ years.
+Historically, the main technology used in these searches has been water Cherenkov detectors, with Super-Kamiokande setting all the current experimental limits on the decay lifetimes at the order of $\sim 10 ^{34}$ years.
 The DUNE far detector and its non-accelerator physics program is a interesting new actor on this stage.
@@
 
-\textcolor{red}{Brief history of nu detectors}
-
-\begin{itemize}
-%\item Challenges Beyond the SM
-%\item Many in text generation of neutrino detector
-\item Brief history of nu detectors
-\item Calibration : LArIAT
-\item The measurements
-\end{itemize}
-
+Such a diverse physics program speaks to the versatility of the LArTPC technology.
 LArTPCs provide excellent electron/photon separation \cite{Acciarri:2016sli} lacking in Cherenkov detectors which can be leveraged to abate the photon background from neutral current interactions  in $\nu_e$ searches.
-LArTPCs share superb tracking capability with bubble chamber detectors, with several additional benefits.
+LArTPCs also share superb tracking capability with bubble chamber detectors, with several additional benefits.
 They are electronically read out and self triggered detectors; they provide full 3D-imaging with millimeter resolution, precise calorimetric reconstruction and excellent particle identification.
-Plus, female physicists are actually writing code and running experiments, not (only) staring at images \textcolor{red}{-- Too much?
+%Plus, female physicists are actually writing code and running experiments, not (only) staring at images \textcolor{red}{-- Too much?
 --} \cite{2006physics4152G}.
 
+The amount of information a LArTPC can provide makes these detectors rather complex: a series of dedicated measurements is necessary to obtain meaningful physics results from a LArTPC.
 The complexity of the LArTPC technology for neutrino detection is due to several reasons.
 Argon is a fairly heavy element, which means that nuclear effects play an important role in the looks of the interaction topology.
-Also, data on charged particle interaction in argon is scarce;  no measurement of hadronic cross sections for pions or kaons is available for argon.
-Secondly,
+For example, pions are one of the main products of neutrino interactions; yet,  since data on charged particle interaction in argon is scarce, neutrino event generators have big uncertainties in the re-scattering simulation of hadrons in argon.
+%No measurement of hadronic cross sections for pions or kaons is available for argon (yet).
+The amount of details in an LArTPC event is easily  parsed by human eye, but can make automatic event reconstruction rather challenging.
+Thus, reconstruction algorithms in LArTPC need to be tune to recognize the different topologies of the neutrino interaction products in argon.
+This is particularly true for pions, since they are an abundant product of the neutrino interactionsl the occurrence of a pion interaction in argon can modify the topology of the neutrino event, causing a misidentification of the neutrino interaction.
 
-LArIAT \cite{Cavanna:2014iqa}, a small (0.25 ton) LArTPC on a beam of charged particles at the Fermilab Test Beam Facility, is performing precise cross section measurements of charged particles in argon.
-The LArIAT suite of cross section measurements is a critical set of results which will be employed by both the short- and long- baseline neutrino programs.
-This work is a measurement of the negative pion argon and the positive kaon-argon total hadronic differential cross sections.
-
-The ($\pi^-$-Ar) hadronic cross section is a fundamental input for neutrino detectors in liquid argon. In fact, reconstructing neutrino interactions at the GeV scale where pion production is abundant.
-
-The   total hadronic differential cross sections is of particular interest for proton decay searches in DUNE, whose most interesting proton decay channel is $p\rightarrow K^+ \bar \nu$.
-
+The LArIAT \cite{Cavanna:2014iqa} experiment is performing precise cross section measurements of charged particles in argon to bridge this gap of knowledge. 
+The LArIAT LArTPC sits on a beam of charged particles at the Fermilab Test Beam Facility; the beam provides charge particles of the type and energy range relevant for neutrino interaction of both SBN and DUNE.
+The ($\pi^-$-Ar) hadronic cross section is a fundamental input for neutrino detectors in liquid argon, as pion interactions can modify the topology and energy reconstruction of neutrino events in the GeV range, where pion production is abundant.
+The  ($K^+$-Ar) total hadronic differential cross section in LArIAT is particularly relevant for a high identification efficiency in the context of proton decay searches in DUNE in the  $p\rightarrow K^+\bar{\nu}$  channel.
+In fact, the kaon-argon cross section affects the kaon topology by modifying the kaon tracking and energy reconstruction, impacting the basis for kaon identification in a LArTPC.
 The cross section analyses exploit the totality of LArIAT's experimental handles; they rely on beam line detector information as well as both calorimetry and tracking in the TPC.
 These analyses are LArIAT's first physics results.
-
-The  ($K^+$-Ar) total hadronic differential cross section in LArIAT is particularly relevant for a high identification efficiency in the context of proton decay searches in DUNE in the  $p\rightarrow K^+\bar{\nu}$  channel.
-In fact, the kaon-argon cross section affects the kaon topology by modifying the kaon tracking and energy reconstruction, impacting the basis for kaon identification in a LArTPC.
-
-A series of additional studies and calibrations were necessary to perform the cross section analyses.
-Appendix \ref{ch:AppendixB} shows a measurement of the LArIAT LArTPC electric field using cosmic data.
-Appendix \ref{ch:AppendixB} shows an optimization of the tracking algorithms geared towards maximizing the efficiency of finding the hadronic interaction point.
-Appendix \ref{ch:AppendixB} shows the calorimetry calibration of the LArIAT LArTPC, which is a pivotal measurement to enable any physics analysis with TPC data.
-
-In order to measure the kaon-argon differential cross section, several steps were necessary.
-The analysis starts by identifying a sample of kaons in the beam line and assessing the beam line contaminations.
-It proceeds with tracking the kaon candidates in the TPC and measuring their calorimetry.
-Then,  the hadronic interaction point is identified and distinguished from weak interaction cases.
-The last step of the analysis, which is the energy unfolding, is now ongoing.
-With  this measurement, I acquired extensive experience in tracking hadrons and measuring their energy inside the TPC.
-This entailed both an in-depth study of the tracking algorithms performances on data and a dive into the GEANT4 transportation models.
-Since no previous kaon measurement has been performed on argon, simulation packages such as GEANT4 currently interpolate the kaon transportation on (scarce) data from lighter and heavier nuclei.
-This is suboptimal for proton decay searches.
-In fact, an extremely high signal detection efficiency, measured by testing identification algorithms on simulation, is extremely difficult to obtain if the simulation is not reliable.
-By adding the experimental data points for kaon-argon cross section to the simulation, my thesis analysis will cover this gap of knowledge.\\
+In order to measure total hadronic  argon cross sections, several steps are necessary.
+The analysis starts by identifying a sample of the hadron of interest in the beam line and assessing the beam line contaminations.
+It proceeds with tracking the hadron candidates in the TPC and measuring their calorimetry at each point in the tracking: the fine sampling of an hadron in the TPC forms the set of ``incident" hadrons.
+Then, the hadronic interaction point is identified and the raw cross section is calculated.
+Two correction\\
 
@@
 We draw the final remarks on this work in Chapter \ref{ch:Conclusions}
+
+A series of additional studies and calibrations were necessary to perform the cross section analyses.
+Appendix \ref{ch:AppendixB} shows a measurement of the LArIAT LArTPC electric field using cosmic data.
+Appendix \ref{ch:AppendixB} shows an optimization of the tracking algorithms geared towards maximizing the efficiency of finding the hadronic interaction point.
+Appendix \ref{ch:AppendixB} shows the calorimetry calibration of the LArIAT LArTPC, which is a pivotal measurement to enable any physics analysis with TPC data.

### 2018-05-25 thesis_main.tex
s

@@
 
-A lot of people are awesome, especially you, since you probably agreed to read this when it was a draft.
+%A lot of people are awesome, especially you, since you probably agreed to read this when it was a draft.
 
-I jumped the pond a little less than 5 years ago because I had a taste of what being a physicists in the US might mean and wanted more. 
-I came to this vast (and at times inhospitable) land thinking I would learn a lot of physics (which I did) and trying to keep doing research for as long as I could (which I'm fortunate enough to say I'll still do for a while).
+%I jumped the pond a little less than 5 years ago because I had a taste of what being a physicists in the US might be like and wanted more. 
+%I came to this vast and at times inhospitable land thinking I would learn a lot of physics (which I did) and trying to keep doing research for as long as I could (which I'm fortunate enough to say I'll still do for a while).
 Little did I know that getting my PhD would be first and foremost a journey to self discovery, the beginning of a path to asymptotically tend to the person and the scientist I want to be.
@@
 
-Mom, you are and will always be the hero of my story; I know this has been challenging for you, thanks for being there for me at every hour of the day and the night, for being always honest with me and for somehow handling this whole US-Italy thing gracefully.
+%Mom, you are and will always be the hero of my story; I know this has been challenging for you, thanks for being there for me at every hour of the day and the night, for being always honest with me and for somehow handling this whole US-Italy thing gracefully.
 I can't believe you still haven't met Bonnie.
@@
 
-Bonnie Fleming, my advisor, I don't even know where to start in thanking you.
-Thanks for listening to me before I proved I was worth be listened to; thanks for always believing I could make it, even at the bottom of my doubts; thanks for the laughters, thanks for leading by example, thanks for  letting me part of the most wonderful research group ;
+%Bonnie Fleming, my advisor, I don't even know where to start in thanking you.
+Thanks for listening to me before I proved I was worth be listened to; thanks for always believing I could make it, even at the bottom of my doubts; thanks for all the laughters, for all the freedom, for all the great advices; thanks for leading by example; thanks for being always supportive of all my crazy side projects; thanks for choosing each and every person in the most wonderful research group ever and thanks for letting me be part of it.
+Thanks for showing me that even Bonnie Freaking Fleming is human, has her downs and still she strives.
 
@@
 % Starts proper arabic numbering of pages and chapters.
+\input{Chapter-0/0-Introduction}
 \mainmatter
-\input{Chapter-0/0-Introduction}
 \input{Chapter-1/1-Theory}

### 2018-05-25 Chapter-0/0-Introduction.tex
s

@@
 
-Experimentally, precision measurements can be achieved only if the detector technology is able to resolve the fine details of a statistically relevant number of interactions. 
+Experimentally, precision measurements can be achieved only if the detector technology is able to resolve the fine details of a neutrino interaction and to record a statistically relevant number of neutrinos.
 With ``fine details" here we mean the ability to distinguish the many products of the neutrino interaction, such as protons, pions, muons and electrons, and to measure their energy.
@@
 Argon is a fairly heavy element, which means that nuclear effects play an important role in the looks of the interaction topology.
-For example, pions are one of the main products of neutrino interactions; yet,  since data on charged particle interaction in argon is scarce, neutrino event generators have big uncertainties in the re-scattering simulation of hadrons in argon.
+For example, pions are one of the main products of neutrino interactions; yet,  since data on charged particle interaction in argon is scarce, neutrino event generators have big uncertainties in the re-scattering simulation of pion in argon.
 %No measurement of hadronic cross sections for pions or kaons is available for argon (yet).
@@
 Thus, reconstruction algorithms in LArTPC need to be tune to recognize the different topologies of the neutrino interaction products in argon.
-This is particularly true for pions, since they are an abundant product of the neutrino interactionsl the occurrence of a pion interaction in argon can modify the topology of the neutrino event, causing a misidentification of the neutrino interaction.
+This is particularly true for pions, since they are an abundant product of the neutrino interactions: the occurrence of a pion interaction in argon can modify the topology of the neutrino event, causing a misidentification of the neutrino interaction.
 
 The LArIAT \cite{Cavanna:2014iqa} experiment is performing precise cross section measurements of charged particles in argon to bridge this gap of knowledge. 
-The LArIAT LArTPC sits on a beam of charged particles at the Fermilab Test Beam Facility; the beam provides charge particles of the type and energy range relevant for neutrino interaction of both SBN and DUNE.
+The LArIAT LArTPC sits on a beam of charged particles at the Fermilab Test Beam Facility which provides charge particles of the type and energy range relevant for neutrino interaction of both SBN and DUNE.
 The ($\pi^-$-Ar) hadronic cross section is a fundamental input for neutrino detectors in liquid argon, as pion interactions can modify the topology and energy reconstruction of neutrino events in the GeV range, where pion production is abundant.
@@
 In fact, the kaon-argon cross section affects the kaon topology by modifying the kaon tracking and energy reconstruction, impacting the basis for kaon identification in a LArTPC.
+
 The cross section analyses exploit the totality of LArIAT's experimental handles; they rely on beam line detector information as well as both calorimetry and tracking in the TPC.
@@
 In order to measure total hadronic  argon cross sections, several steps are necessary.
-The analysis starts by identifying a sample of the hadron of interest in the beam line and assessing the beam line contaminations.
-It proceeds with tracking the hadron candidates in the TPC and measuring their calorimetry at each point in the tracking: the fine sampling of an hadron in the TPC forms the set of ``incident" hadrons.
+The analyses start by identifying a sample of the hadron of interest in the beam line and assessing the beam line contaminations.
+It proceeds with tracking the hadron candidates in the TPC and measuring their kinetic energy at each point in the tracking: the fine sampling of an hadron in the TPC forms the set of ``incident" hadrons.
 Then, the hadronic interaction point is identified and the raw cross section is calculated.
-Two correction\\
+Two corrections are then applied to the raw cross section -- a background subtractions and a correction for detector effects -- to obtain the true cross section measurement.\\
 
-This body of work is divided in 8 chapters.
+This body of work is organized in 8 chapters.
 We provide a description of the theoretical framework for the measurements in  Chapter \ref{ch:TheTheory}. Chapter \ref{ch:2} outlines the LArTPC detector technology, while
 Chapter \ref{sec:experimentDescription} describes LArIAT experimental setup.
-We present the event selection for both the pion and kaon analyses, as well as the ``thin-slice method" in Chapter \ref{ch:Interactions}.
+We present the event selection for both the pion and kaon analyses, as well as the ``thin slice method" in Chapter \ref{ch:Interactions}.
 Chapter \ref{ch:samples}  describes the work done on the data and Monte Carlo samples in preparation of the cross section analyses.
@@
 Appendix \ref{ch:AppendixB} shows a measurement of the LArIAT LArTPC electric field using cosmic data.
-Appendix \ref{ch:AppendixB} shows an optimization of the tracking algorithms geared towards maximizing the efficiency of finding the hadronic interaction point.
-Appendix \ref{ch:AppendixB} shows the calorimetry calibration of the LArIAT LArTPC, which is a pivotal measurement to enable any physics analysis with TPC data.
+Appendix \ref{ch:AppendixTrack} shows an optimization of the tracking algorithms geared towards maximizing the efficiency of finding the hadronic interaction point.
+Appendix \ref{ch:energyCalibration} shows the calorimetry calibration of the LArIAT LArTPC, which is a pivotal measurement to enable any physics analysis with TPC data.

### 2018-05-25 Chapter-8/8-Conclusions.tex
s

\chapter{Conclusions}\label{ch:Conclusions}
In the era of neutrino precision measurements, of huge liquid argon detectors and of massive amount of information from LArTPCs, a renew interest for an ancient measurement arises: the measurement of hadronic interactions with matter. With this work, we presented the first ever ($\pi^-$-Ar) and ($K^+$-Ar) total hadronic cross section measurements as a function of the hadron kinetic energy. These analyses are the first physics analyses developed by the LArIAT experiment.
Both the analysis follow a similar workflow and  they rely on beam line detector information as well as both calorimetry and tracking in the TPC.

In order to measure ($\pi^-$-Ar) total hadronic  argon cross sections, we start by selecting pion beamline candidates through a series of selections on the beamline and TPC information apt to maximize the number of pions in the selection over the number of muons and electrons. We use the LArIAT beamline MC to estimate the beam composition of the selected beamline candidare and we propagate the particle species to the LArAIT TPC constructing a properly weighted sample with the DDMC.

 The analyses start by identifying a sample of the hadron of interest in the beam line and assessing the beam line contaminations. It proceeds with tracking the hadron candidates in the TPC and measuring their kinetic energy at each point in the tracking: the fine sampling of an hadron in the TPC forms the set of ``incident" hadrons.  Then, the hadronic interaction point is identified and the raw cross section is calculated. Two corrections are then applied to the raw cross section -- a background subtractions and a correction for detector effects -- to obtain the true cross section measurement.\\

These analyses' work flow will serve as a basis for the future cross section measurements of pions and kaons in the exclusive channels.

### 2018-05-28 AppendixB-TrackingStudies/B-Tracking.tex
s

@@
-\chapter{Tracking Studies}\label{ch:AppendixTrack}
-In this section, we describe three studies.
+\chapter{Additional Tracking Studies for LArIAT Cross Section Analyses}\label{ch:AppendixTrack}
+In this section, we describe two studies.
 The first is a justification of the selection criteria for the beamline handshake with the TPC information.
@@
 
-\subsection{Study of WC to TPC Match}\label{ch:WC2TPCMatchOptimization}
+\subsection{Study of WC to TPC Match}
 
-Plots I want in this section:
-\begin{enumerate}
-\item WC2TPC MC DeltaX, DeltaY and $\alpha$
-\end{enumerate}
+%Plots I want in this section:
+%\begin{enumerate}
+%\item WC2TPC MC DeltaX, DeltaY and $\alpha$
+%\end{enumerate}
 
-Scope of this study is assessing the goodness of the wire chamber to TPC match on Monte Carlo and decide the selection values we will use on data.
+Scope of this study is assessing the performances of the WC2TPC match on Monte Carlo (see Section \ref{ch:WC2TPCMatchMethod} ) and decide the selection values we will use on data.
 A word of caution is necessary here.
@@
 \subsection{Tracking Optimization}\label{ch:TrackOptimization}
+We perform an optimization of the clustering algorithm (see Section \ref{sec:SignalProc}) with the scope of maximizing the efficiency of finding the interaction point for the total hadronic cross section measurements.
+We define as the interaction point the most downstream point of a WC2TPC matched TPC tracks within the TPC fiducial volume.
+Since all the WC2TPC tracks are by definition beam particles, tracks travel from upstream to downstream in the TPC; thus, identifying the interaction point means to stop the tracking correctly.
+
+TrajCluster is the package used to cluster hits in LArIAT; this package counts more than 20 tunable parameters.
+A standard method to develop clustering algorithms and checking their performances is to ``hand scan", which means recognizing the effect of parameters tuning by looking at a series of data event displays.
+Albeit we recognize the importance of hand scanning as a great diagnosis tool, we developed a fully automated optimization package which compares MC reconstructed information to MC truth.
+
+We start by defining a figure of merit in order to discern what makes a parameter configuration better than an other.
+We chose the percentage of events whose reconstructed and true length differ less than 2 cm.
+We then identify the parameters in TrajCluster that are most important to  correctly stop  the tracking and an appropriate range of values for each of them.
+We chose to optimize the parameters that leverage on the angle between consecutive groups of hits, the number of hits use in the cluster fit and the average hit charge to stop the tracking.
+We define a configuration space with all  possible combination of values for the chosen parameters and we perform reconstruction one combination at a time: the combination with the highest figure of merit determines the optimized tracking reconstruction.
+
+We chose construct the combination space using a total of 5 parameters, 3 values each and two iterations of the method (for a total of 486 combinations).
+We run the combinations on a sample of 100000 pion events.
+After the optimization, the most upstream point of the tracking is correctly identified 99.5\% of the times, the most downstream point is correctly identified 62.5\% of the times, the tracking stops short about 15\% of the times and misses the interaction point 22.5\% of the times.
+Hand scanning confirmed that the missed interaction points happen in the vast majority of cases for very shallow angles, as shown in the event display in Figure, or in the case of angles visible only in one projection plane.
+We also noticed that the premature stopping of the tracks is often related to the presence of delta rays parallel to the track.
+We see room of improvement, such as the delta ray removal and a forced track breaking in case of a kink present in a single plane, for a future analysis.
+\textcolor{red}{ADD evd}
+
+The procedure behind this optimization package is virtually applicable to any LArSoft module where it is possible to define figure of merit.

### 2018-05-28 Chapter-8/8-Conclusions.tex
s

@@
 \chapter{Conclusions}\label{ch:Conclusions}
-In the era of neutrino precision measurements, of huge liquid argon detectors and of massive amount of information from LArTPCs, a renew interest for an ancient measurement arises: the measurement of hadronic interactions with matter.
+In the era of neutrino precision measurements, of huge liquid argon detectors and of massive amount of information from LArTPCs, a renewed interest for an ancient measurement arises: the measurement of hadronic interactions with matter.
 With this work, we presented the first ever ($\pi^-$-Ar) and ($K^+$-Ar) total hadronic cross section measurements as a function of the hadron kinetic energy.
@@
 In order to measure ($\pi^-$-Ar) total hadronic  argon cross sections, we start by selecting pion beamline candidates through a series of selections on the beamline and TPC information apt to maximize the number of pions in the selection over the number of muons and electrons.
-We use the LArIAT beamline MC to estimate the beam composition of the selected beamline candidare and we propagate the particle species to the LArAIT TPC constructing a properly weighted sample with the DDMC.
+We use the LArIAT beamline MC to estimate the beam composition of the selected beamline candidates and we propagate them to the LArAIT TPC constructing a properly weighted sample with the DDMC.
+We apply the thin slice method on the pion candidates and obtain the raw cross section measurement.
+From the simulated sample, we obtain two corrections accounting for the beamline background contamination and for detector effects.
+Finally, we apply the corrections to data and measure the true cross section.
 
- The analyses start by identifying a sample of the hadron of interest in the beam line and assessing the beam line contaminations.
-It proceeds with tracking the hadron candidates in the TPC and measuring their kinetic energy at each point in the tracking: the fine sampling of an hadron in the TPC forms the set of ``incident" hadrons.
-Then, the hadronic interaction point is identified and the raw cross section is calculated.
-Two corrections are then applied to the raw cross section -- a background subtractions and a correction for detector effects -- to obtain the true cross section measurement.\\
+In order to measure ($K^+$-Ar) total hadronic  argon cross sections, we follow a similar procedure, i.e.
+we apply the thin slice method on kaon candidates identified in the beamline to obtain the raw cross section.
+We apply a background correction and a correction for detector effects to the raw cross section.
+The background correction accounts for the presence of secondary particles in both the interacting and incident histograms and for the presence of decay events in the interacting plot.
 
-These analyses' work flow will serve as a basis for the future cross section measurements of pions and kaons in the exclusive channels.
+The final results for the ($\pi^-$-Ar) and ($K^+$-Ar) total hadronic cross section are shown side by side in figure \ref{fig:finalfinal}.
+
+\begin{figure}[htb]
+\centering
+\includegraphics[width=0.48\textwidth]{Chapter-6/Images/TheMoneyPlot.pdf}
+\includegraphics[width=0.48\textwidth]{Chapter-7/Images/TheMoneyPlotK.pdf}
+\caption{\emph{Left:} ($\pi^-$-Ar) total hadronic cross section measurements in the 60A and 100A samples overlaid with the  Geant4 prediction (green).
+\emph{Right:} ($K^+$-Ar) total hadronic cross section for  scattering angles greater than 5$^\circ$ measured in the 60A sample, statistical uncertainty in black and systematic uncertainty in red.
+The Geant4 prediction for the total hadronic cross section for angle scattering greater than 5$^\circ$ is displayed in green.
+}
+\label{fig:finalfinal}
+\end{figure}
+
+These analyses' will serve as a basis for the future cross section measurements of pions and kaons for the exclusive channels in LArIAT.

### 2018-05-29 thesis_main.tex
s

@@
 \begin{abstract}
-Abstract goes here. Limit 750 words.
+%Abstract goes here. Limit 750 words.
+
+The Liquid Argon Time Projection Chamber (LArTPC) represents one of the
+most advanced experimental technologies for physics at the Intensity
+Frontier due to its full 3D-imaging, excellent particle identification
+and precise calorimetric energy reconstruction. By deploying a LArTPC in
+a dedicated calibration test beam line at Fermilab, LArIAT (Liquid
+Argon In A Testbeam) aims to experimentally calibrate this technology
+in a controlled environment and to provide physics results key to the
+neutrino oscillation physics and proton decay searches of the Short
+Baseline Neutrino and Long Baseline Neutrino programs.
+
+LArIAT's physics program entails a vast set of topics with a
+particular focus on the study of nuclear effects such as pion and
+kaon characteristic interaction modes.
+
+This thesis presents two world's first measurements: the measurement of ($\pi^-$-Ar) total hadronic cross section in the 100-1000 MeV kinetic energy range and the measurement of the ($K^+$-Ar) total hadronic cross section in the 100-600 MeV kinetic energy range.
+The analyses devised for these measurements use both the core elements of LArIAT: beamline and TPC.
+The first step in each analysis is the development of an event selection based on  beamline and TPC information geared towards the identification of the hadron of interest.
+We then proceed to match the beamline candidate to a suitable TPC track.
+Finally,  we apply the ``thin slice method" technique and measure the cross section, correcting for backgrounds and detector effects.
+The thin slice technique is a new method to measure hadron-argon cross sections possible only thanks to the combination of the tracking and calorimetry capability of the LArTPC technology.
+Albeit never on argon, the hadronic cross section of pions has been measured before on several different elements in thin target experiments, leading to solid predictions in the argon case.
+Even through the use of a different technique, our measurement of the  ($\pi^-$-Ar) total hadronic cross section is in general agreement with the prediction by the Geant4 Bertini Cascade model which are based on data from thin target experiments.
+On the contrary, cross section measurements for kaons are extremely rare, thus more difficult to model.
+Not surprisingly, our measurement of the  ($K^+$-Ar) total hadronic cross section is mostly in disagreement with the Geant4 prediction.
+
+This thesis also reports two  ancillary  detector physics measurements necessary for the cross section analyses: the measurements of the LArIAT electric field and calorimetry constants.
+We developed a technique to measure the LArIAT electric field using cathode to anode piercing tracks with cosmic data.
+We applied a new technique for the measurement of the calorimetry calibration constants based on the particles' momentum measurement.
+
+The ($\pi^-$-Ar) and the ($K^+$-Ar) total hadronic cross measurements are the first physics results of the LArIAT experiment and will be the basis for the future LArIAT measurements of pion and kaon cross sections in the exclusive channels.
+
 \end{abstract}
@@
 
-%A lot of people are awesome, especially you, since you probably agreed to read this when it was a draft.
+A lot of people are awesome, especially you, since you probably agreed to read this when it was a draft.
 
 %I jumped the pond a little less than 5 years ago because I had a taste of what being a physicists in the US might be like and wanted more. 
-%I came to this vast and at times inhospitable land thinking I would learn a lot of physics (which I did) and trying to keep doing research for as long as I could (which I'm fortunate enough to say I'll still do for a while).
+%I came to this vast and at times inhospitable land thinking I would learn a lot of physics (which I did) and hoping I'd keep doing research for as long as I could (which I'm fortunate enough to say I'll still do for a while).
 Little did I know that getting my PhD would be first and foremost a journey to self discovery, the beginning of a path to asymptotically tend to the person and the scientist I want to be.
@@
 
-%Flavio Cavanna 
-%Ornella Palamara, I once told Andrzej I think of you as my spirit animal: fierce, independent and direct; he replied I still need to eat a lot of paninis to get there and I can't agree more.
+%Flavio Cavanna, my co-advisor, working with you has been one of the greatest pleasures of my academic life; the extent of your knowledge never ceases to amaze me and the freshness of your enthusiasm is always contagious.
+Thanks for all the time spent in your office banging our heads on these cross section analyses.
+
+%Ornella Palamara, after a particularly heated exchange at a wine and cheese, I once told Andrzej I think of you as my spirit animal: fierce, knowledgeable and direct; he replied I still need to walk many miles to get there and I can't agree more.
 
@@
 
-%Jen Raaf
-%Jonathan Asaadi
-%Andrzej M. Szelc
-%Roberto Acciarri
-%LArIAT & MicroBooNE
+%Jen Raaf and Jonathan Asaadi, these analyses would have not seen the light of day without you.
+Thanks for your many inputs and for letting your door open to talk about physics or my latest paranoia.
+Thanks for the patience and understanding.
+And thanks for leading that small experiment with a big heart called LArIAT and all its crazy ponies.
+Speaking of which, thanks to all my collaborators, especially Irene Nutini, who started the path for the pion analysis, Greg Pulliam, who suffered with me at the end of said path, Dan the Man, Jason, Johnny, Will Fo and Mr Bojangles.
+A special token of gratitude goes to Roberto Acciarri and Andrzej M.
+Szelc, you are so much more to me than collaborators and previous car owners, you know you'll always have a home in Chicago.
+
 %To all the members of the NPG, thanks for all the laughters and all the wise insights. You're the smartest bunch of idiots I've ever met, I love y'all.
-%Shany %Louise
 
-%To the extraordinary people who stayed close, even with an ocean in between: la Tasso, Marci, la Bea, la Kittirn, Alvin, la Viola e Brogna, Sighi and all the people from the ``La Chiappia Nuova" group, thanks for making me feel like I never left.
-%Finally, Anto, cause people might think I'm nice and easy, but you know I'm not. And you're still here.
+%Thanks to the MicroBooNE collaboration Sam kazu and kirby
+
+%Shany Danieli and Louise Laage, you're the piece of heart I left in New Haven.
+Shany, I don't know what made you think I was the cool kid in the group, but I'm so glad you stuck with me.
+I'm not sure I would have survived my first semester at Yale without you: you shine in your smartness and lift other people with you -- it's a rare quality in our field.
+Louise, my dear, thanks for always finding the time to listening, for your attuned spirit and for the magnolias.
+
+%Thanks to all the people which helped me make Chicago and Fermilab my new home, in particular David Caratelli,  Mateus Carneiro, Pilar Coloma, Ciaran Hughes, Gordan Krnjaic, Seyda Ipek, Kiel Howe and Rob Fine.
+
+%To the extraordinary people who stayed close, even with an ocean in between: la Tasso, Marci, la Bea, la Kittirn, Alvin, la Viola e Brogna, Sighi, Eleonora and all the people from the ``La Chiappia Nuova" group, thanks for making me feel like I never left.
+
+%Finally, thanks to Anto, my husband, my topo and my adventure companion, cause people might think I'm nice and easy, but you know I'm not...
+And still, you followed me across half the world and you're still here, so that I could never feel alone.
 
@@
 \appendix
-%%\input{Chapters/A-KaonAna}
+\input{AppendixB-EField/B-LArIATEField}
 \input{AppendixB-TrackingStudies/B-Tracking}
 \input{AppendixC-EnergyCalibration/C-EnergyCalibration}
-%\input{AppendixB-EField/B-LArIATEField}
+
 %%\input{Chapters/C-MicroBooNECRT}

### 2018-05-29 Chapter-4/4-Methodology.tex
s

@@
 
-This chapter describes the general procedure employed to measure a total hadronic differential cross section in LArIAT.
+This chapter describes the general procedure employed to measure  total hadronic interaction cross sections on Ar in LArIAT.
 Albeit with small differences, both the  ($\pi^{-}$,Ar) and (K$^{+}$,Ar) total hadronic cross section measurements rely on the same procedure.
 We start by selecting the particle of interest using a combination of beamline detectors and TPC information (Section \ref{ch:ParticleSelectionMethod}).
-We then perform a handshake between the beamline information and the TPC tracking to assure the selection of the correct TPC track (Section \ref{ch:WC2TPCMatchMethod}).
-Finally, we apply the ``thin slice" method and measure the ``raw" hadronic cross section (Section \ref{ch:ThinSliceMethod}).
-A series of corrections are then evaluated to obtain the ``true" cross section (Section \ref{ch:MCCorrections}).
+We then perform a handshake between the beamline information and the TPC tracking to assure the selection of the correct TPC track (Section \ref{ch:WC2TPCMatchMethod}) associated to the corresponding beam particle.
+We then apply the ``thin slice" method to measure the ``raw" hadronic cross section (Section \ref{ch:ThinSliceMethod}).
+A series of corrections are then evaluated and applied to obtain the final cross section (Section \ref{ch:MCCorrections}).
 
-At the end of this chapter, we show a sanity check of the methodology by applying the thin slice method employing only MC truth information and retrieving the Geant4 tabulated cross section for pions and kaons (Section \ref{ch:procedureTesting}).
+At the end of this chapter, we show a sanity check of the methodology by applying the thin slice method employing only MC truth information and retrieving the expected MC cross section for pions and kaons (Section \ref{ch:procedureTesting}).
 
 \section{Event Selection}\label{ch:ParticleSelectionMethod}
-The measurement of the ($\pi^{-}$,Ar) and (K$^{+}$,Ar) total hadronic cross section in LArIAT starts by selecting the pool of pion or kaon candidates and measuring their momentum.
+The measurement of the ($\pi^{-}$,Ar) and (K$^{+}$,Ar) total hadronic cross section in LArIAT starts by selecting the pool of pion or kaon candidates and measuring their momentum before they enter the LAr volume.
 This is done through the series of selections on  beamline and TPC information described in the next sections.
@@
 \subsection{Selection of Beamline Events}\label{ch:beamlineDetectorsData}
-As shown in equation \ref{eq:enFF}, we leverage the beamline particle identification and momentum measurement before entering the TPC as an input to evaluate the kinetic energy for the hadrons used in the  cross sections measurements.
-Thus, we select the LArIAT data to keep only events whose wire chamber and time of flight information is registered (line 1 in in Table \ref{tab:beamlineDataSelection}).
+We leverage the beamline particle identification and momentum measurement before entering the TPC as an input to evaluate the kinetic energy for the hadrons used in the  cross sections measurements.
+To this end, we select the LArIAT data to keep only events whose wire chamber and time of flight information is registered (line 1 in in Table \ref{tab:beamlineDataSelection}).
 Additionally, we perform a check of the plausibility of the trajectory inside the beamline detectors: given the position of the hits in the four wire chambers, we make sure the particle's trajectory does not cross any impenetrable material such as the collimator and the magnets steel (line 2 in in Table \ref{tab:beamlineDataSelection}).
@@
 The presence of upstream and downstream collimators greatly abates the number of particles tracing down the LArIAT tertiary beamline.
-However, it is possible that more than one particle sneaks into the LArTPC during its readout time: the TPC readout is triggered by the particle  firing the beamline detectors, but particles from the beam halo might be present in the TPC at the same time.
+However, it is possible that more than one particle sneaks into the LArTPC during its readout time: the TPC readout is triggered by the particle  firing the series of beamline detectors along our tertiary beamline, but particles from the beam halo might also be present in the TPC at the same time.
 We call ``pile up" the additional traces in the TPC.
@@
 In the case of the ($\pi^-$,Ar) cross section, the resolution of  beamline mass spectrometer is not sufficient to select a beam of pure pions.
-In fact, muons and electrons survive the selection on the beamline mass.
+In fact,  muons which are close in mass to the pions and relativistic electrons survive the selection on the beamline mass.
 It is important to notice that the composition of the negative polarity beam is mostly pions, as will be discussed in section \ref{ch:beamlineComposition}.
 Still, we devise a selection on the TPC information to mitigate the presence of electrons in the sample used for the pion cross section.
-The selection relies on the different topologies of a pion and an electron event in the argon: while the former will trace a track inside the TPC active volume, the latter will tend to ``shower", i.e.
+The selection relies on the different topologies of a pion and an electron event when propagating in liquid argon: while the former will trace a track inside the TPC active volume, the latter will tend to ``shower", i.e.
 interact with the medium, producing bremsstrahlung photons which pair convert into several short tracks.
@@
 For each event passing the selection on its beamline information, we need to identify the track inside the TPC corresponding to the particle which triggered the beamline detectors, a procedure we refer to as ``WC to TPC match" (WC2TPC for short).
-In general, the TPC tracking algorithm will reconstruct more than one track in the event, partially due to the fact that hadrons interact in the chamber and partially because of pile up particles during the triggered TPC readout time, as shown in figure~\ref{fig:kaonInteraction}.
+In general, the TPC tracking algorithm can reconstruct more than one track in the event, partially due to the fact that hadrons interact in the chamber and partially because of pile up particles during the triggered TPC readout time, as shown in figure~\ref{fig:kaonInteraction}.
 
@@
 
-We attempt to uniquely match one wire chamber track to one and only one reconstructed TPC track. 
-In order to determine if a match is present, we apply a geometrical selection on the relative the position of the wire chamber and TPC tracks. 
+We attempt to uniquely match one wire chamber track (see Section \ref{sec:MWPCfunc}) to one and only one reconstructed TPC track. 
+In order to determine if a match is present, we apply a geometrical selection on the relative position of the wire chamber and TPC tracks. 
 We start by considering only TPC tracks whose first point is in the first 2 cm upstream portion of the TPC for the match.
@@
 
-In MC, we mimic the matching between the WC and the TPC track by constructing a fake WC track using truth information at wire chamber four.
+In MC, we mimic the matching between the WC and the TPC track by constructing an artificial WC track using truth information at wire chamber four.
 We then apply the same WC to TPC matching algorithm as in data.
@@
 \section{The Thin Slice Method}\label{ch:ThinSliceMethod}
-Once we have selected the pool of hadron candidates and we have identified the TPC track corresponding to the beamline event, we apply the thin slice method to measure the cross section, as the following sections describe.
+
+Once we have selected the 40841 beamline pion candidates  and the 1081 beamline kaon candidates, and we have identified the TPC corresponding track, we apply the thin slice method to measure the cross section, as the following sections describe.
 \subsection{Cross Sections on Thin Target}
@@
 The so-called  ``thin target" approximation assumes that the target centers are uniformly distributed in the material and that the target is thin compared to the projectile interaction length, so that no center of interaction sits in front of another.
-In this approximation, the ratio between the number of particles interacting in the target $N_{\text{Int}}$ and the number of incident particles $N_{\text{Inc}}$ on the target determines the interaction probability $P_{Interacting}$, which is the complementary to one of the survival probability $P_{Survival}$.
+In this approximation, the ratio between the number of particles interacting in the target $N_{\text{Int}}$ and the number of incident particles $N_{\text{Inc}}$ on the target estimates the interaction probability $P_{Interacting}$, which is the complementary to one of the survival probability $P_{Survival}$.
 Equation \ref{eq:thinTargetXS} 
 \begin{equation}
-P_{Survival} = 1- P_{Interacting} = 1 - \frac{N_{\text{Int}}}{N_{\text{Inc}}} = e^{-\sigma_{TOT} n \delta X}
+P_{Survival} = 1- P_{Interacting} = 1 - \frac{N_{\text{Int}}}{N_{\text{Inc}}} = e^{-\sigma_{TOT}\text{ } n \text{ }\delta X}
 \label{eq:thinTargetXS}
@@
 \begin{equation}
-1 - \frac{N_{\text{Int}}}{N_{\text{Inc}}} =  1 -\sigma_{TOT} n \delta X + O(\delta X^2).
+1 - \frac{N_{\text{Int}}}{N_{\text{Inc}}} =  1 -\sigma_{TOT} \text{  }n \text{  }\delta X + O(\delta X^2).
 \label{eq:thinTargetXSTaylor}
@@
 \begin{equation}
- \sigma_{TOT}  = \frac{1}{n \delta X}\frac{N_{\text{Int}}}{N_{\text{Inc}}}.
+ \sigma_{TOT}  = \frac{1}{n \text{ }\delta X}\frac{N_{\text{Int}}}{N_{\text{Inc}}}.
 \label{eq:thinTargetXSSolved}
@@
 
-\subsection{Not-so-Thin Target: Slicing the Argon}\label{ch:XSRaw}
-The interaction length of pions and kaons in argon is expected to be of the order of 50 cm for pions and 100 cm for kaons.
+\subsection{Not-so-Thin Target: Slicing the Liquid Argon Volume}\label{ch:XSRaw}
+The interaction length of pions and kaons in liquid argon is expected to be of the order of 50 cm for pions and 100 cm for kaons.
 Thus, the LArIAT TPC, with its 90 cm of length, is not a thin target.
@@
 
-As described in Chapter \ref{sec:experimentDescription}, LArIAT wire planes consist of 240 wires each.
-The wires are oriented at +/- $60^{\circ}$ from the vertical direction at 4 mm spacing, while the beam direction is oriented 3 degrees off the $z$ axis in the $XZ$ plane.
-The wires collect signals proportional to the energy loss of the hadron along its path in a  $\delta${\emph{X}} = 4 mm/(sin($60^{\circ}$)cos($3^{\circ}$)) $\approx$ 4.7~mm slab of liquid argon.
+As described in Chapter \ref{sec:experimentDescription}, LArIAT induction and collection planes consist of 240 wires each at 4 mm spacing.
+The wires are oriented at +/- $60^{\circ}$ from the vertical direction, while the beam direction is oriented 3 degrees off the $z$ axis in the $XZ$ plane.
+The collection wires collect signals proportional to the energy deposited by the hadron along its path in a  $\delta${\emph{X}} = 4 mm/(sin($60^{\circ}$)cos($3^{\circ}$)) $\approx$ 4.7~mm slab of liquid argon.
 Thus, one can think to slice the TPC into many thin targets of $\delta${\emph{X}} = 4.7~mm thickness along the direction of the incident particle, making a measurement at each wire along the path.
@@
 \end{equation}
-where $E_{loss}$ is a correction for the energy loss in the uninstrumented material between the beamline and the TPC front face.
-The  energy of the hadron at each slab is determined by subtracting the energy released by the particle in the previous slabs.
-For example, at the $j^{th}$ point of a track, the kinetic energy will be
+where $E_{loss}$ is a correction for the kinetic energy loss in the uninstrumented material between the beamline and the TPC front face.
+While propagating through the target,  the kinetic energy of the hadron at each slab is determined by subtracting the energy deposited by the particle in the previous slabs.
+For example, at the $j^{th}$ slab of a track, the kinetic energy will be
 
@@
 
-If the particle enters a slice, it contributes to $N_{\text{Inc}}( E^{kin})$ in the energy bin corresponding to its kinetic energy in that slice.
-If it interacts in the slice, it also contributes to $N_{\text{Int}}(E^{kin})$ in the appropriate energy bin.
-The cross section as a function of kinetic energy, $\sigma_{TOT}( E^{kin})$ will then be proportional to the ratio $\frac{N_{\text{Int}}( E^{kin})}{N_{\text{Inc}}( E^{kin})}$.
+If the particle enters a slice, it contributes to the $N_{\text{Inc}}( E^{kin})$ distribution in the energy bin corresponding to its kinetic energy in that slice.
+While into the slice, a hadron may or may not interact.
+If it interacts in the slice, it  contributes also to the $N_{\text{Int}}(E^{kin})$ distribution in the appropriate energy bin; this occurrence corresponds to the end of the hadron tracking.
+If the hadron does not interact, it will enter the next slice and the interaction evaluation starts again.
+The process is applied to all the hadrons in the sample; the cross section as a function of kinetic energy, $\sigma_{TOT}( E^{kin})$ is then evaluated to be proportional to the ratio $\frac{N_{\text{Int}}( E^{kin})}{N_{\text{Inc}}( E^{kin})}$ -- bin by bin ratio.
 
 Our goal is to measure the total interaction cross section, independently  from the topology of the interaction.
-Thus, we determine that a hadron interacted simply by requiring that the last point of the WC2TPC matched track lies inside the fiducial volume, whose boundaries are defined in Table \ref{tab:FidVol}.
-If the TPC track stops within the fiducial volume, its last point will be the interaction point; if the track crosses the boundaries of the fiducial volume, the track will be considered ``through going" and no interaction point will be found.
-The only points of the hadronic candidate track considered to fill the  $N_{\text{Inc}})$ and  $N_{\text{Inc}}$ plots are the ones contained in the fiducial volume.
+Thus, we determine that a hadron interacted simply by requiring that the last point of the WC2TPC matched track lies in a slice within the fiducial volume, whose boundaries are defined in Table \ref{tab:FidVol}.
+If the TPC track ends within the fiducial volume, its last point will be the interaction point; if the track crosses the boundaries of the fiducial volume, the track will be considered ``through going" and no interaction point will be found.
+The only points of the hadronic candidate track considered to fill the  $N_{\text{Int}}$ and  $N_{\text{Inc}}$ distributions are the ones contained in the fiducial volume.
+\textcolor{red}{ADD DECAY}
 

### 2018-05-29 thesis_main.tex
s

@@
 
-This thesis presents two world's first measurements: the measurement of ($\pi^-$-Ar) total hadronic cross section in the 100-1000 MeV kinetic energy range and the measurement of the ($K^+$-Ar) total hadronic cross section in the 100-600 MeV kinetic energy range.
+This thesis presents two world's first measurements: the measurement of ($\pi^-$-Ar) total hadronic cross section in the 100-1050 MeV kinetic energy range and the measurement of the ($K^+$-Ar) total hadronic cross section in the 100-650 MeV kinetic energy range.
 The analyses devised for these measurements use both the core elements of LArIAT: beamline and TPC.
@@
 We then proceed to match the beamline candidate to a suitable TPC track.
-Finally,  we apply the ``thin slice method" technique and measure the cross section, correcting for backgrounds and detector effects.
+Finally,  we apply the ``thin slice method" technique and measure the cross section, correcting for background and detector effects.
 The thin slice technique is a new method to measure hadron-argon cross sections possible only thanks to the combination of the tracking and calorimetry capability of the LArTPC technology.
 Albeit never on argon, the hadronic cross section of pions has been measured before on several different elements in thin target experiments, leading to solid predictions in the argon case.
-Even through the use of a different technique, our measurement of the  ($\pi^-$-Ar) total hadronic cross section is in general agreement with the prediction by the Geant4 Bertini Cascade model which are based on data from thin target experiments.
+Through the use of a different technique, our measurement of the  ($\pi^-$-Ar) total hadronic cross section is in general agreement with the prediction by the Geant4 Bertini Cascade model which are based on data from thin target experiments.
 On the contrary, cross section measurements for kaons are extremely rare, thus more difficult to model.
@@
 This thesis also reports two  ancillary  detector physics measurements necessary for the cross section analyses: the measurements of the LArIAT electric field and calorimetry constants.
-We developed a technique to measure the LArIAT electric field using cathode to anode piercing tracks with cosmic data.
+We developed a technique to measure the LArIAT electric field using cathode-anode piercing tracks with cosmic data.
 We applied a new technique for the measurement of the calorimetry calibration constants based on the particles' momentum measurement.
@@
 \mainmatter
-\input{Chapter-1/1-Theory}
-\input{Chapter-2/2-LArTPCs}
-\input{Chapter-3/3-TheLArIATExperiment}
+%\input{Chapter-1/1-Theory}
+%\input{Chapter-2/2-LArTPCs}
+%\input{Chapter-3/3-TheLArIATExperiment}
 \input{Chapter-4/4-Methodology}
@@
 
-%
-%\input{Chapter-9/9-Uncertainty}
-
 % Only call appendix once, here.
@@
 \input{AppendixC-EnergyCalibration/C-EnergyCalibration}
-
-%%\input{Chapters/C-MicroBooNECRT}
-
-% Old chapters configuration %
-%\input{Chapters/4-DataCollection}
-%\input{Chapters/4-MonteCarlo}
-%\input{Chapters/5-EnergyCalibration}
-%\input{Chapter-8/8-TrackingOptimization}
 

### 2018-05-29 Chapter-4/4-Methodology.tex
s

@@
 The only points of the hadronic candidate track considered to fill the  $N_{\text{Int}}$ and  $N_{\text{Inc}}$ distributions are the ones contained in the fiducial volume.
-\textcolor{red}{ADD DECAY}
+
+ A notable background pertinent only to the $N_{\text{Int}}$  distribution are cases in which the hadrons decays inside the TPC.
+In those cases in fact, the tracking ends inside the TPC but the interaction is not hadronic.
+The handling of decay background is treated in a slightly different way for the pion and kaon section, details can be found in sections \ref{ch:PionXSBkgSub} and \ref{ch:KaonXSRaw} respectively.
 
@@
 Equation \ref{eq:thinTargetXSSolved}  is a prescription for measuring the cross section in case of a pure beam of the hadron of interest and 100\% efficiency in the determination of the interaction point.
-For example, if LArIAT had a beam of pure pions and were 100\% efficient in determining the interaction point within the TPC, the pion cross section in each energy bin would be given by
+For example, if LArIAT had a beam of pure pions and were 100\% efficient in determining the interaction point within the TPC, the pion cross section as a function of  kinetic energy (estimated at the central value of the energy bin $E_i$) would be given by
 
@@
 Unfortunately, this is not the case.
-In fact, the selection used to isolate pions in the LArIAT beam allows for the presence of some muons and electrons as background, while the kaon selection allows for a small percentage of protons (see Section \ref{ch:beamlineComposition}).
-Also, the LArIAT TPC is not 100\% efficient in determining the interaction point.
-Therefore we need to apply two corrections evaluated on the MC in order to extract the true cross section from LArIAT data: the background subtraction and the efficiency correction.
+In fact, the selection used to isolate pions in the LArIAT beam allows for the presence of some muons and electrons as background, while the kaon selection allows for a small contamination of protons (see Section \ref{ch:beamlineComposition}).
+Also, the LArIAT TPC tracking algorithm is not 100\% efficient in determining the interaction point.
+Therefore we need to apply two corrections evaluated on the MC in order to extract the final cross section from LArIAT data: i) a background subtraction and ii) a correction for reconstruction effects.
 Still using the pion case as example, we estimate the pion cross section in each energy bin changing  Equation \ref{eq:thinTargetXSSolved2} into
 \begin{equation}
- \sigma^{\pi^-}_{TOT}(E_{i})  =\frac{1}{n \delta X}\frac{N^{\pi^-}_{ \text{Int}} (E_{i})}{N^{\pi^-}_{ \text{Inc}}(E_{i})} = \frac{1}{n \delta X}\frac{ \epsilon^{\text{Inc}}(E_i) [ N^{ \text{TOT}}_{ \text{Int}} (E_{i}) - B_{ \text{Int}} (E_i)] }{   \epsilon^{\text{Int}}(E_i) [N^{ \text{TOT}}_{ \text{Inc}}(E_{i}) - B_{ \text{Inc}} (E_i)]},
+ \sigma^{\pi^-}_{TOT}(E_{i})  =\frac{1}{n\text{ } \delta X}\frac{N^{\pi^-}_{ \text{Int}} (E_{i})}{N^{\pi^-}_{ \text{Inc}}(E_{i})} = \frac{1}{n \text{ }\delta X}\frac{ \epsilon^{\text{Inc}}(E_i) [ N^{ \text{TOT}}_{ \text{Int}} (E_{i}) - B_{ \text{Int}} (E_i)] }{   \epsilon^{\text{Int}}(E_i) [N^{ \text{TOT}}_{ \text{Inc}}(E_{i}) - B_{ \text{Inc}} (E_i)]},
 \label{eq:True}
@@
 
-where  $N^{\text{TOT}}_{\text{Int}} (E_{i})$ and $N^{\text{TOT}}_{\text{Incident}} (E_{i})$ is the measured content of the interacting and incident histograms for events that pass the event selection, $B_{\text{Int}} (E_i)$ and $B_{\text{Inc}} (E_i)$ represent the contributions from the background to the interacting and incident histograms respectively, and  $\epsilon^{\text{Int}}(E_i)$ and  $\epsilon^{\text{Inc}}(E_i)$ are the efficiency corrections for said histograms.
+where  $N^{\text{TOT}}_{\text{Int}} (E_{i})$ and $N^{\text{TOT}}_{\text{Incident}} (E_{i})$ is the measured content of the interacting and incident histograms for events that pass the event selection, $B_{\text{Int}} (E_i)$ and $B_{\text{Inc}} (E_i)$ represent the contributions from the background to the interacting and incident histograms respectively, and  $\epsilon^{\text{Int}}(E_i)$ and  $\epsilon^{\text{Inc}}(E_i)$ are the corrections for reconstruction effects.
 
-As we will show in section \ref{ch:PionXSBkgSub}, the background subtraction for the interacting and incident histograms can be translated into a corresponding relative pion content $C^{\pi MC}_{Interacting} (E_{i})$ and $C^{\pi MC}_{Incident} (E_{i})$ and the cross section re-written as follows
+As we will show in section \ref{ch:PionXSBkgSub}, the background subtraction for the interacting and incident histograms can be translated into corresponding relative pion content factors $C^{\pi MC}_{\text{Int}} (E_{i})$ and $C^{\pi MC}_{\text{Inc}} (E_{i})$ and the cross section re-written as follows
 
 \begin{equation}
-      \sigma^{\pi^-}_{TOT}(E_{i})  = \frac{1}{n \delta X}\frac{ \epsilon^{\text{Inc}}(E_i)  \hspace{0.2cm} C^{\pi MC}_{\text{Int}} (E_{i}) \hspace{0.2cm} N^{\text{TOT}}_{\text{Int}} (E_{i}) }{   \epsilon^{\text{Int}}(E_i) \hspace{0.2cm} C^{\pi MC}_{\text{Inc}} (E_{i}) \hspace{0.2cm}  N^{\text{TOT}}_{\text{Inc}} (E_{i})}.
+      \sigma^{\pi^-}_{TOT}(E_{i})  = \frac{1}{n\text{ } \delta X}\frac{ \epsilon^{\text{Inc}}(E_i)  \hspace{0.2cm} C^{\pi MC}_{\text{Int}} (E_{i}) \hspace{0.2cm} N^{\text{TOT}}_{\text{Int}} (E_{i}) }{   \epsilon^{\text{Int}}(E_i) \hspace{0.2cm} C^{\pi MC}_{\text{Inc}} (E_{i}) \hspace{0.2cm}  N^{\text{TOT}}_{\text{Inc}} (E_{i})}.
 \label{eq:C}
@@
 
-\section{Procedure testing with truth quantities}\label{ch:procedureTesting}
+\section{Procedure testing with MC truth quantities}\label{ch:procedureTesting}
 The ($\pi^{-}$,Ar) and (K$^{+}$,Ar) total hadronic cross section implemented in Geant4 can be used as a tool to validate the measurement methodology.
-We describe here a closure test done on Monte Carlo to prove that the methodology of slicing the TPC retrieves the underlying cross section distribution implemented in Geant4 within the statistical uncertainty.
-%under the working assumption of perfect reconstruction.
+We describe here a closure test done on Monte Carlo to prove that the methodology of slicing the TPC retrieves the underlying cross section distribution implemented in Geant4 within the MC statistical uncertainty.
 
-For pions and kaons in the considered energy range, the Geant4 inelastic model adopted is ``BertiniCascade"; the pion elastic cross sections are tabulated from on Chips, while the kaon elastic cross sections are tabulated on Gheisha and Chips.
+For pions and kaons in the considered energy range, the Geant4 inelastic model adopted is ``BertiniCascade"; the pion elastic cross sections are tabulated from Chips, while the kaon elastic cross sections are tabulated on Gheisha and Chips.
 
-For the validation test, we fire a sample of pions and a sample of kaons inside the LArIAT TPC active volume using the Data Driven Monte Carlo (see section \ref{sec:DDMC}).
+For the validation test, we fire a sample of pions and a sample of kaons inside the LArIAT TPC active volume using the Data Driven Monte Carlo, a procedure described in Section \ref{sec:DDMC}.
 We apply the thin-sliced method using only true quantities to calculate the hadron kinetic energy at each slab in order to decouple reconstruction effects from possible issues with the methodology.
-For each slab of 4.7 mm length along the path of the hadron, we integrate the true energy deposition as given by the Geant4 transportation model.
+For each slab of 4.7 mm length along the path of the hadron, we integrate the true energy deposition as given by the Geant4 transport model.
 Then, we recursively subtracted it from the hadron kinetic energy at the TPC front face to evaluate the kinetic energy at each slab until the true interaction point is reached.
-Since the MC is a pure beam of the hadron of interest and truth information is used to retrieve the interaction point, no correction is applied.
-Doing so, we obtain the true interacting and incident distributions for the considered hadron, from which we derive  the true MC cross section as a function of the hadron true kinetic energy.
+Since the MC is a pure beam of the hadron of interest and truth information is used to retrieve the interaction point, no background correction or reconstruction effects correction is applied.
+Doing so, we obtain the true interacting and incident distributions for the considered hadron, whose ratio leads to  the true MC cross section as a function of the hadron kinetic energy.
 
-Figure \ref{fig:TrueMCXS2} shows the total hadronic cross section for argon implemented in Geant4 10.03.p1 (solid lines) overlaid with the true MC cross section as obtained with the sliced TPC method (markers) for pions on the left and kaons on the right; the total cross section is shown in green,  the elastic cross section in blue and the inelastic cross section in red.
+Figure \ref{fig:TrueMCXS2} shows the total hadronic cross section for argon implemented in Geant4 10.03.p1 (solid lines) overlaid with the true MC cross section as obtained with the sliced TPC method (markers) for pions on the left and kaons on the right; the total cross section is shown in green.
+For completeness, we also report the contributions from  the elastic cross section (in blue) and the inelastic cross section (in red), available at the MC level.
 The nice agreement with the Geant4 distribution and the cross section  obtained with the sliced TPC method gives us confidence in the  validity of the methodology.

### 2018-05-30 Chapter-1/1-Theory.tex
s

@@
 
-\chapter{The theoretical framework}
+\chapter{The theoretical framework}\label{ch:TheTheory}
 
@@
 
-{\raggedleft -- J. S. Bach, Suite No. 1 in G major, BWV 1007, 1720 ca. -- \par}
+{\raggedleft -- J. S. Bach, 1720 ca. -- \par} %, Suite No. 1 in G major, BWV 1007
 \vspace{0.5cm}
@@
 
-\subsubsection{$\pi^{-}$Ar Hadronic Interaction: Signal Signatures}
+\subsubsection{$\pi^{-}$Ar Hadronic Interaction: Signal Signatures}\label{ch:signature}
 Strong hadronic interaction models \cite{9780198520085,9780471779957} predict the pion interaction processes with argon in the [100 - 1200] MeV energy range. 

### 2018-05-30 Chapter-5/5-Preparatory.tex
s

@@
 Data taking for the this period started on 03/15/2016  and ended on 07/31/2016. 
-Since we are interested in beamline and TPC information, we ask basic requirements on the operational status of the time of fight, wire chambers and TPC to form the good run list for this period, which we informally call ``lovely runs".
+Since we are interested in beamline and TPC information, we ask basic requirements on the operational status of the time of flight counters, wire chambers and TPC to form the good run list for this period, which we informally call ``lovely runs".
 
@@
 The -60A and -100A combined data set accounts for approximately 90\% of the total Run-II negative polarity runs.
-The  choice of the main two beamline settings limits the need for the production of many different MC sets and related corrections, still maintaining a high number of events.
+The  choice of these two main beamline settings limits the need for the production of many different MC sets and related corrections, still maintaining a high number of events.
 %Since the production of beamline Monte Carlo depends on the wanted beamline configuration, the choice of only two beamline settings limits the need for beamline MC production.
@@
 
-For the first measurements in LArIAT that uses both beamline and TPC information, we choose strict requirements on the reconstruction of the WC tracks, the so-called ``Picky Track" sample (see Section \ref{sec:MWPCfunc}).
+For these first measurements that make use of both the LArIAT beamline and TPC information, we choose strict requirements on the reconstruction of the WC tracks, the so-called ``Picky Track" sample (see Section \ref{sec:MWPCfunc}), where we require a single hit in each and every wire chamber detector to reconstruct the WC track.
 This choice presents two advantages:  the uncertainty on the momentum reconstruction for the ``Picky Tracks" sample is smaller compared to the ``High Yield" sample, and the comparison with the beamline MC results is straightforward.
@@
 For the simulation of LArIAT events and for the simulation of the datasets' particle make up, we use a combination of two MC generators: the G4Beamline Monte Carlo and the Data Driven single particle Monte Carlo (DDMC).
-We use the G4Beamline MC to simulate the particle transportation in the beamline and calculate the particle composition of the beam just after the fourth Wire Chamber (WC4).
-In order to simulate the beamline particles after WC4 and in the TPC, we use the DDMC.
+We use the G4Beamline MC to simulate the particle transport in the LArIAT tertiary beamline and calculate the particle composition of the beam just after the last Wire Chamber (WC4).
+We use the DDMC  to simulate the particles after WC4 along the beamline, close to the beam window in the LAr cryostat and in the TPC.
 
 \subsection{G4Beamline}\label{ch:beamlineComposition}
-G4Beamline simulates the beam collision with the LArIAT secondary target, the energy deposited by the particles in the LArIAT beamline detectors, and the action of the LArIAT magnets, effectively accounting for particle transportation through the beamline from the LArIAT target until ``Big Disk", a fictional, void detector located just before the LArIAT cryostat.
+G4Beamline simulates the beam collision at the LArIAT secondary target, the energy deposited by the particles in the LArIAT beamline detectors, and the action of the LArIAT magnets, effectively accounting for particle transport through the beamline from the LArIAT target until ``Big Disk", a fictional, void detector located just before the LArIAT cryostat.
  At the moment of this writing, G4Beamline does not simulated the responses of the beamline detectors.
@@
 In order to ease the handshake between G4Beamline and the DDMC, we ask for the beam composition just after WC4.
-Since LArIAT data are taken under different beam conditions, we need to simulate separately the beam composition according to the magnets' settings and the secondary beam intensity with G4Beamline.
-For the pion cross section analysis the relevant beam conditions are  secondary beam energy of 64 GeV, negative polarity magnet with current of 100 A and 60 A.
-For the kaon cross section analysis the relevant beam conditions is a secondary beam energy of 64 GeV, positive polarity magnet with current of 100 A.
+Since LArIAT data are taken under different beam conditions, we need to simulate separately the beam composition according to the magnets' settings and the secondary pion beam intensity with G4Beamline.
+For the pion cross section analysis the relevant beam conditions are  secondary pion beam energy of 64 GeV, negative polarity magnet with current of 100 A and 60 A.
+For the kaon cross section analysis the relevant beam conditions is a secondary  pion beam energy of 64 GeV, positive polarity magnet with current of 100 A.
 
@@
 Even if pions are by far the biggest beam component in negative polarity runs, the LArIAT tertiary beam is not a pure pion beam.
-While useful to discriminate between pions, kaons, and protons, the beamline detectors are not sensitive enough to  discriminate among the lighter particles in the beam: electrons, muons and pions fall under the same mass hypothesis.
-Thus, we need to assess the contamination from beamline particles other than pions in the event selections used for the pion cross section analysis and correct for its effects.
-The first step of this process is assessing the percentage of electrons and muons in the $\pi/\mu/e$ beamline candidates via the G4Beamline MC.
+While useful to discriminate between pions, kaons, and (anti)protons, the beamline detectors are not sensitive enough to  discriminate among the lighter particles in the beam: electrons, muons and pions fall under the same mass hypothesis.
+Thus, we need to assess the contamination from beamline particles other than pions in the event selections used for the pion cross section analysis and correct for this background.
+The first step of this process is assessing the percentage of electrons and muons in the $\pi/\mu/e$ beamline candidates via the G4Beamline MC, as we deem the percentage of kaons and antiprotons negligible after the mass selection (< 1 \textperthousand).
 Since the beamline composition is a function of the magnet settings, we simulate separately events for magnet current of -60A and -100A. 
@@
 The predictions for electrons, muons and pions have been staggered and their sum is area normalized to data.
-Albeit not perfect, these plots show a reasonable agreement between the momentum shapes in data and MC.
-We attribute the difference in shape to a two approximations performed in the MC.
+Albeit not perfect, these plots show a reasonable agreement in momentum shapes between data and MC.
+We attribute the difference in shape (longer tail in data) to a two approximations performed in the MC.
 Firstly, G4Beamline lacks the simulation of the WC efficiency which is momentum dependent and leads to enhance the number events in the center of the momentum distribution.
-Secondly, G4Beamline stop tracking pions and their products if they decay in after WC1; in data, pion decays in flight can still create a tigger if the produced muon travels thought the beamline detectors.
-In the pion cross section analysis, these differences between data and G4Beamline are accounted for as a systematic uncertainty related to the beam composition (see Section \ref{ch:BKGsubXSPuppa}).
+Secondly, G4Beamline stop tracking pions and their products if they decay after WC1; in data, pion decays in flight can still create a tigger if the produced muon travels thought the beamline detectors.
+In the pion cross section analysis, these differences between data and the G4Beamline prediction are accounted for as a systematic uncertainty related to the beam composition (see Section \ref{ch:BKGsubXSPuppa}).
 %We attribute  the difference in shape to the lack of simulation of the WC efficiency in the MC which is momentum dependent and leads to enhance the number events in the center of the momentum distribution.
@@
 \subsection{Data Driven MC}\label{sec:DDMC}
-The Data Driven single particle Monte Carlo (DDMC) is a single particle gun which simulates the particle transportation from WC4 into the TPC leveraging on the beamline data information.
+The Data Driven single particle Monte Carlo (DDMC) is a single particle gun which simulates the particle transport from WC4 into the TPC leveraging on the beamline data information.
 The DDMC uses the data momentum and position at WC4 to derive the event generation: a general sketch of the DDMC workflow is shown in Figure \ref{fig:DDMCSketch}.
@@
 For each data event, we obtain the  particle position ($X, Y$) at WC4 directly from the data measurement; we calculate the components of the momentum using the beamline measurement of the momentum magnitude in conjunction with the hits on WC3 and WC4 to determine the direction of the momentum vector, as described in section \ref{sec:MWPCfunc}.
-The momentum and position of the selected data form a 5-dimensional tupla, which we sample thousands of times through a 5-dimensional hit-or-miss sampling procedure to generate the MC events.
+The momentum and position of the selected data events form a 5-dimensional series of tuples.
+The DDMC event generator samples from the joint distribution of these five quantities using a 5-dimensional hit-or-miss sampling procedure.
 This sampling generates MC events  with the same momentum and position distributions as data, with the additional benefit of accounting for the correlations between the $P_x, P_y, P_z, X, Y$ variables.
@@
 
-Using the DDMC technique ensures that the MC and data particles have very similar momentum, position and angular distributions at WC4 and allows us to use the MC sample in several occasions: to estimate the background contamination to the pion cross section (see Section \ref{ch:PionXSBkgSub}), to calibrate the energy loss upstream of the TPC (see Section \ref{ch:eloss}),  or to study the tracking and the calorimetric performance (sections \ref{sec:TrackingStudies} and \ref{ch:energyCal}).
+Using the DDMC technique ensures that the MC and data particles have very similar momentum, position and angular distributions at WC4 and allows us to use the MC sample in several occasions: to estimate the background contamination to the pion cross section (see Section \ref{ch:PionXSBkgSub}), to calibrate the energy loss upstream of the TPC (see Section \ref{ch:eloss}),  and to study the tracking and the calorimetric performance  in the LArTPC (sections \ref{sec:TrackingStudies} and \ref{ch:energyCal}).
 A small caveat is in order here: the DDMC is a single particle Monte Carlo, which means that the beam pile-up is not simulated.
@@
 We use the beamline simulation and the DDMC simulation to estimate the background in the total hadronic pion cross section.
-Two categories of background exists for the negative pion cross section measurement: the one related to the pion interaction in the chamber, discussed in Section \ref{ch:CaptureAndDecay} and the one related to the beamline contamination, discussed in Section \ref{ch:PionXSBkgSub2}.
+Two categories of background exists for the negative pion cross section measurement: the background related to the pion alternative processes to hadronic interaction inside the LArTPC, discussed in Section \ref{ch:CaptureAndDecay} and the background related to the beamline contamination, discussed in Section \ref{ch:PionXSBkgSub2}.
 
@@
 Our goal is to measure the total hadronic cross section for negative pions in argon.
-Since pion capture can be classified as an electromagnetic process and pion decay is a week process,  capture and decay represent unwanted interactions.
-We present here a study of capture and decay in Monte Carlo and the solution we adopted to mitigate their occurrence in the data sample.
+Since pion capture can be classified as an electromagnetic process and pion decay is a weak process (as discussed in Section \ref{ch:signature}),  capture and decay represent alternative processes to the hadronic interaction, resulting therefore into a background for the cross section analysis.
+We present here a study of capture and decay in Monte Carlo and the solution we adopted to remove their occurrence in the data sample.
 
@@
 It is important to notice that capture occurs predominantly at rest, while decay may occur both in flight and at rest.
-Thus, we can highly mitigate capture and decay at rest by removing pions which would release all their energy in the TPC and stop.
-This translates into a momentum selection, where we keep only events whose WC momentum is above a certain threshold.
+Thus, we can highly mitigate capture and decay at rest by removing pions whose kinetic energy at the TPC front face is low enough to be completely released by ionization in the TPC, eventually bringing the pion to stop within the fiducial volume.
+This translates into a beam momentum selection, where we keep only events whose WC momentum is above a certain threshold.
 Figure \ref{fig:CaptureMom} shows the true momentum distribution for the primary pions\footnote{We use here the Geant4 denomination ``primary" to indicate that the pion considered does not undergo interactions modifying its energy before getting to the TPC.
@@
 
-In order to choose the selection value for the wire chamber momentum, it is beneficial to estimate the ratio of events which capture or decay that survive the selection in MC as a function of the momentum threshold, and compare it with the survival ratio for all the 60A events.
+In order to choose the selection value for the wire chamber momentum, it is beneficial to estimate the fraction of MC capture and decay events that survive the momentum selection as a function of the momentum threshold, and compare it with the survival fraction for all the 60A events.
 This is done in figure \ref{fig:survRatio}.
@@
 We calculate the survival ratio separately for the three event categories explained above: total (pink), capture (green) and decay (blue).
-Selecting pions with momentum greater than 420 MeV/c reduces the capture events by ~99\% while maintaining about 80\% of the 60A data sample and almost the entire 100A sample. 
-Figure \ref{fig:evtRatio} shows the ratio of events which end their life in capture (green) or decay (blue) over the total number of events as a as a function of the true momentum at wire chamber four.
+Selecting pions with momentum greater than 420 MeV/c removes $\sim$99\% the capture events while maintaining about 80\% of the 60A data sample and almost the entire 100A sample. 
+Figure \ref{fig:evtRatio} shows the ratio of events which end their life in capture (green) or decay (blue) over the total number of events as a as a function of the true MC momentum at WC4.
 This ratio is slightly dependent on the inelastic cross section implemented in Geant4, as we are able to register a pion capture (or decay) only if it did not interact inelastically in the TPC.
@@
 \includegraphics[width=15cm]{Chapter-7/Images/CDAsMomentumFunct.png}
-\caption{True momentum distribution at wire chamber 4 for every simulated pion arriving in the TPC (pink), ending its life in capture (green) or in decay (blue) in the TPC, linear vertical axis on the left, logarithmic on the right.
+\caption{MC momentum distribution at wire chamber 4 for every simulated pion arriving in the TPC (pink), ending its life in capture (green) or in decay (blue) in the TPC, linear vertical axis on the left, logarithmic on the right.
 }
@@
 \includegraphics[width=7.5cm]{Chapter-7/Images/CDThreshold.png}
-\caption{Survival ratio as a function of selection threshold on true momentum at wire chamber four for for every simulated pion arriving in the TPC (pink), capture (green) or in decay (blue).   }
+\caption{Survival ratio as a function of selection threshold on MC momentum at wire chamber four for for every simulated pion arriving in the TPC (pink), capture (green) or in decay (blue).   }
 \label{fig:survRatio}
@@
 \includegraphics[width=7.5cm]{Chapter-7/Images/CDRatio.png}
-\caption{Ratio between the capture (green) and decay (blue) events over the total number of events as a as a function of the true momentum at wire chamber four.}
+\caption{Ratio between the capture (green) and decay (blue) events over the total number of events as a as a function of the MC momentum at wire chamber four.}
 \label{fig:evtRatio}
@@
 \item[]2) muons,
-\item[]3) secondaries from pion events,
+\item[]3) secondaries from pion interactions in beamline elements,
 \item[]4) matched pile up events.
@@
 The next step is to simulate those pions, muons and electrons from WC4 to the TPC with the DDMC and evaluate their contribution to the cross section.
-To do so, we start by simulating the same number of electrons, muons and pions with the DDMC and we apply the same selection chain (i.e.
-track multiplicity rejection, WC2TPC acceptance and shower rejection) on the three samples.
+To do so, we start by generating the same number of electrons, muons and pions with the DDMC and we apply the same selection chain applied in data (i.e.
+track multiplicity rejection, WC2TPC match and shower rejection) on the three samples.
 The number of events per particle species surviving this selection is shown on table \ref{tab:MCafterCutContaminants}.
-In order to reproduce the closest make up of the beam to data, we weight each event of a given particle species according to the estimated beam composition.
+In order to reproduce the data beamline composition, we weight each event of a given particle species according to the estimated content for that species as found with the G4Beamline simulation.
 In case of 60A runs, for example, the weights are 0.688 for pions,  0.046 for muons  and 0.266 for electrons.
 
-It should be noted that pions may  interact hadronically in the steel or in the non-instrumented argon upstream to the TPC front face while travelling the length of between WC4 and the TPC.
+It should be noted that pions may  interact hadronically in the steel (cryostat wall) or in the non-instrumented argon upstream to the TPC front face while traveling the length of between WC4 and the TPC.
 Or, they could decay in flight between WC4 and the TPC.
-One of the interaction products can leak into the TPC and be matched with the WC track, contributing to the pool of events used for the cross section calculation.
+One of the interaction or decay products can leak into the TPC and be matched with the WC track, contributing to the pool of events used for the cross section calculation.
 We call this type of particles ``secondaries" from pion events, with a terminology inspired by Geant4.
@@
 The percentage of secondaries is given by the number of matched WC2TPC tracks whose corresponding particle is not flagged as primary by Geant4.
-The secondary to pion ratio is 4.9\% in the 60A sample and 4.3\% in the 100A sample.
+The secondary particles to primary pion ratio is 4.9\% in the 60A sample and 4.3\% in the 100A sample.
 
@@
 &  &  &  & & &\\  
-Selection Survival Rate           &62.3\% & 86.6\% & 6.1\% & 63.8\%& 85.5\%& 5.2\%\\
+Selection Survival Rate           &62.2\% & 86.4\% & 6.1\% & 63.8\%& 85.2\%& 5.2\%\\
 Beam Composition  @WC4      &  68.8\%   &  4.6 \%  & 26.6 \%    & 87.4 \% & 3.7 \%  & 8.9 \% \\ %
-Beam Composition  @TPC FF &  88.5\%   & 8.2\%   & 3.3 \%   & 94.0\%	& 5.3\% & 0.7\%\\
+Expected Composition in XS sample &  88.5\%   & 8.2\%   & 3.3 \%   & 94.0\%	& 5.3\% & 0.7\%\\
                                                   &                      &                       &                   &                       &                        &\\  
@@
 
-We evaluate the beamline background contribution to the cross section by producing the interacting and incident histograms for the events surviving the selection, staggering the contributions for each particle species, as shown in Figure  \ref{fig:stag60A}.
+\begin{figure}[p]
+\centering
+\includegraphics[width=\textwidth]{Chapter-5/Images/Background100A.pdf}
+\caption{Left: staggered contributions to the interacting kinetic energy distribution for electron (grey), muons (yellow) and pion (blue) in the 100A simulation sample.
+Right: staggered contributions to the incident kinetic energy distribution for electron (grey), muons (yellow) and pion (blue) in the 100A simulation sample.
+}
+\label{fig:stag100A}
+\end{figure}
+
+We evaluate the beamline background contribution to the cross section by producing the interacting and incident histograms for the signal and background events surviving the selection, staggering the contributions for each particle species, as shown in Figure  \ref{fig:stag60A} for the -60A case and  in Figure \ref{fig:stag100A} for the -100A case.
 From those histograms, we are able to evaluate the contribution of  pions and  beamline backgrounds to each bin of the interacting and incident histograms separately and obtain the relative pion content.
@@
 We take here the interacting histogram as example, noting that the derivation of the correction for the incident histogram is identical.
-The number of entries in each bin of the interacting plot (Figure \ref{fig:stag60A} left) is  $N^{\text{TOT}}_{\text{Int}} (E_{i})$, equal to the sum of the pions and beamline backgrounds in that bin, namely
+The number of entries in each bin of the interacting histogram (Figure \ref{fig:stag60A} left) is  $N^{\text{TOT}}_{\text{Int}} (E_{i})$, equal to the sum of the pions and beamline backgrounds flagged as interacting in that bin, namely
 
@@
 This is because most of the muons will cross the TPC without stopping, thus contributing almost exclusively to the incident histogram, forcing the pion content to be lower in the incident histogram than in the interacting; thus, the correction will tend to enhance the cross section.
+We present the estimation of $C^{\pi MC}_{\text{Int}} (E_{i})$ and $C^{\pi MC}_{\text{Inc}} (E_{i})$ and their associated systematic uncertainty in chapter discussing the pion cross section measurement (Figure \ref{fig:CorrectionsBeam}).
 
@@
 The $E_{loss}$ distribution for the  whole kaon sample is shown in figure \ref{fig:ELossKaons}.
-A clear double peaked structure is visible, which is due to the particles either missing or hitting the HALO paddle: a schematic rendering of this occurrence is  shown in figure \ref{fig:Halo}.
-The kinematic at WC4 determines the trajectory of a particle and whether or not it will hit the halo paddle.
+A clear double peaked structure is visible.
+After a long investigation on the origin of this unexpected shape, we determined that its origin is due to the particles either missing or hitting the HALO paddle (see Section \ref{sec:MuRS} for the HALO paddle description): a schematic rendering of this occurrence is  shown in figure \ref{fig:Halo}.
+The kinematic at WC4 determines the trajectory of a particle and indicates whether or not it will hit the halo paddle, whose positioning and central hole size were not sufficiently well defined when the beamline instrumentation was installed.
 In figure \ref{fig:PxVsXTrue} , we plot the true  horizontal component of the momentum $P_x$ versus the true $X$ position at WC4 for pions missing the halo paddle (left) and for pions hitting the halo paddle (right) for the -60A MC simulation runs -- analogous plots are obtained with the -100A pion simulation and with the kaon simulation.
@@
 In MC, we assign  $E_{loss} = 32 \pm 4 $~MeV for pion events classified as ``hitting the halo paddle"; we assign  $E_{loss} = 24 \pm 3 $~MeV for pion events classified as ``missing the halo paddle".
+These values are the average and width of the two landau distributions underneath the double peaked distribution.
 We apply the same classifier on data.
 
-A scan of the simulated geometry showed an excess of 3 cm of uninstrumented argon compared with the surveyed detector geometry.
+A late scan of the simulated geometry showed an excess of 3 cm of uninstrumented argon compared with the surveyed detector geometry.
+This excess has an effect on the $E_{loss}$: MC particles traverse more uninstrumented material  (i.e.
+loose more energy in the argon before the TPC front face) than data particles.
 We account for this difference by assigning in data $E_{loss} = 24 \pm 6 $~MeV for pion events classified as ``hitting the halo paddle" and  $E_{loss} = 17 \pm 6 $~MeV for pion events classified as ``missing the halo paddle", where the uncertainty is derived as the standard deviation of the double peaked distribution.
@@
 Thus, it plays a fundamental role in the cross section measurements.
-We performed several studies geared towards the optimization of the package for tracking in the TPC.
+We performed several studies geared towards the optimization of the LArSoft software package for tracking in the TPC.
 In particular, we studied a suitable set of parameters for the WC2TPC match and we optimized the clustering algorithm to maximize the efficiency of finding the interaction point on MC.
 Given the technical nature of these studies, we report them in Appendix \ref{ch:AppendixTrack}.
-We report here the evaluation of  the angular resolution of the tracking algorithm in data and MC, due to its implication on the physics measurement.
+We only report here the evaluation of  the angular resolution of the tracking algorithm in data and MC, due to its important implications on the physics measurement.
 
 \subsection{Angular Resolution}\label{sec:angleRes}
-Scope of this study is to understand and compare the tracking performances and angular resolution of the TPC tracking on data and MC. 
-We use the angular resolution of the tracking to determine  the value of smallest angle that we can reconstruct with a non-zero efficiency, effectively determining a selection on the angular distribution of the cross section measurement due to the tracking performance.
+Scope of this study is to understand and compare the tracking performances and the angular resolution of the TPC tracking on data and MC. 
+We use the angular resolution of the tracking to determine  the value of smallest angle that the tracking algorithms allow to reconstruct with a non-zero efficiency, effectively determining a selection on the distribution of the scattering angle of hadronic interaction entering the cross section measurement.
 %\textcolor{red}{This study is performed on the pion sample, but its results are extrapolated to the kaon case.}
@@
 These tracks can contain from a minimum of 3 3D-space points to a maximum of 240  3D-space points.
-We fit a line to all the 3D-space points associated with the track.
+We fit a straight line to all the 3D-space points associated with the track.
 For each track we calculate the average distance between each point in space and the fit line as follows 
@@
 We then calculate the angle between the fit of the first and second half $\alpha$ (panel d).
-The angle $\alpha$ determines the spatial resolution of the tracking.
+The angle $\alpha$ determines the angular resolution of the tracking.
 The distributions for data and MC for $\alpha$ are given in Figure \ref{fig:trackingResolution} for pions and in Figure \ref{fig:trackingResolutionK} for kaons.
-The mean of the data and MC angular resolution are reported in Table {tab:AngRes} for pions and kaons in data and MC.
+The mean of the data and MC angular resolution are reported in Table \ref{tab:AngRes} for pions and kaons in data and MC.
 
@@
 
-Interaction angles smaller than the angle resolution are indistinguishable for the reconstruction.
+Interaction angles smaller than the angular resolution are indistinguishable for the reconstruction.
 Therefore, we assess our ability to measure the cross section to be limited to interaction angles greater than 5.0 deg.
@@
 
+As we discussed in Section \ref{ch:signature}, several different interaction topologies are included as signal in the total hadronic cross section.
+The ability to detect a minimum interaction angle  and to stop the tracking accordingly mainly effects two interaction channels: the pion elastic interaction (see Table \ref{tab:pionFinal}, second line) and the pion inelastic interaction in case of neutral particle emission (see Table \ref{tab:pionFinal}, fourth line); the overall effect of this limitation is to reduced the cross section measurement to the measurement of the cross section relative to interaction angles greater than a $\sim$5.0 deg .
 It is beneficial to take a moment to describe the definition of interaction angle.
@@
 \end{equation}   
-In case of inelastic scattering,  the presence of several topologies requires a more complex definition, as shown in figure \ref{fig:scatterPic}.
-We define the scattering angle as the biggest of the angles between the incoming hadron and the visible daughters, where the visible daughters are charged particles that travel more than 0.47 cm in the detector (see panel a); in case all the daughters are invisible, the angle is assigned to be 90 deg (see panel b).
-We chose this working definition of scattering angle for inelastic scattering keeping in mind how our tracking reconstruction works: the tracking will stop correctly non of the daughters are is visible in the detector and it is likely to stop correctly if multiple daughters form an interaction vertex.
+In case of the reaction channel,  the presence of several topologies requires a more complex definition, as shown in figure \ref{fig:scatterPic}.
+We define the scattering angle as the biggest of the angles between the incoming hadron and the visible daughters, where the visible daughters are charged particles that travel more than the average pitch length ($\delta X = 47$ mm) in the detector (see panel a); in case all the daughters are invisible, the angle is assigned to be 90 deg (see panel b).
+We chose this working definition of scattering angle for inelastic scattering keeping in mind how our tracking reconstruction works: the tracking will stop correctly if none of the daughters is visible in the detector and it is likely to stop correctly if multiple visible daughters form an interaction vertex.
 The only ``dangerous" case is the production of one charged daughter plus neutrals, which we can study with this working definition of scattering angle (see panel c).
 
-We can see the effects of the angular resolution on the cross section by plotting the true Geant4 cross section for interaction angles greater than a minimum interaction angle.
-Figure \ref{fig:trueWithAngles} shows the  true Geant4 cross section  for interaction angles greater than 0 deg (green), 4.5 deg (red), 5.0 deg  (blue) and 9.0 deg (yellow).
-A small $0.5 \text{ deg}$ systematic shift between the mean of the data and MC angular resolution is present.%, which we account for in the context of the MC efficiency correction to the cross section, as presented in \ref{sec:angSys}.
+Once we fix the scattering angle definition, we can study the effects of the angular resolution on the cross section by plotting the true Geant4 total hadronic cross section for interaction angles greater than a minimum interaction angle.
+The left side of Figure \ref{fig:trueWithAngles} shows the  true Geant4 cross section  for interaction angles greater than 0 deg (green), 4.5 deg (red) corresponding to the MC angular resolution, 5.0 deg  (blue) corresponding to the data angular resolution, and 9.0 deg (yellow).
+A small $0.5 \text{ deg}$ systematic shift between the mean of the data and MC angular resolution is present, which has a negligible impact on the cross section.
+The right side of Figure \ref{fig:trueWithAngles} shows the ratio between the  true cross section for interaction angles greater than 5 deg and the true interaction cross section for all angles;  the cross section for angles greater than 5$^\circ$ accounts for more than 80\% than the total cross section in every energy bin.
+%, which we account for in the context of the MC efficiency correction to the cross section, as presented in \ref{sec:angSys}.
 
@@
 \includegraphics[width=0.48\textwidth]{Chapter-5/Images/cTrueXSAngle.png}
-\caption{ True ($\pi^-, Ar$) cross section for interaction angles greater than 0 deg (green), 4.5 deg (red), 5.0 deg  (blue) and 9.0 deg (yellow). }
+\includegraphics[width=0.48\textwidth]{Chapter-5/Images/RatioXSs.pdf}
+\caption{\emph{Left:} True ($\pi^-, Ar$) cross section for interaction angles greater than 0 deg (green), 4.5 deg (red), 5.0 deg  (blue) and 9.0 deg (yellow).
+\emph{Right:} Ratio between the true cross section for interaction angles greater than 5 deg and the true interaction cross section for all angles.}
 \label{fig:trueWithAngles}

### 2018-05-30 Chapter-5/5-Preparatory.tex
s

@@
 In Appendix  \ref{ch:energyCalibration}, we describe how we calibrate the TPC calorimetric response.
-In the following section, we describe how we measure the kinetic energy of the hadrons in the TPC.
+In this section, we describe how we measure the kinetic energy of the hadrons in the TPC.
 
@@
 \end{equation}
-where $E_{i}$ is measured in each slab as  the product of the stopping power,  $dE/dX_{i}$,  and the track pitch, $Pitch_i$, for that point. 
-If we assume conservatively that the measurements of $E_{i}$ are not independent from one another, the uncertainty on $E_{\text{FF-j}}$ becomes
+where $E_{i}$ is measured in each slab as  the product of the stopping power,  $dE/dX_{i}$,  and the track pitch for that point.
+Since the measurements of the energy deposited in each slab rely on the same global calorimetric procedure and tracking algorithms, we assume conservatively that that the measurements of $E_{i}$ are not independent from one another; thus, the uncertainty on $E_{\text{FF-j}}$ becomes
 \begin{equation}
-\delta E_{\text{FF-j}} = (j-1) \delta E_{i}, 
+\delta E_{\text{FF-j}} =  \sum_{i<j}  \delta E_{i} = (j-1) \delta E_{i}, 
 \end{equation}

### 2018-05-30 Chapter-6/6-PionCrossSection.tex
s

@@
 the cross section obtained exclusively using data reconstruction, without any additional corrections.
-In Section \ref{ch:PionXSCorrections}, we apply a statistical subtraction of the background contributions based on simulation and a correction for detection inefficiency.
+In Section \ref{ch:PionXSCorrections}, we apply the statistical subtraction of the background contributions based on simulation and the correction for reconstruction effects.
 The final results are presented in Section \ref{ch:FinalPion}.
@@
 We measure the raw ($\pi^-$-Ar) total hadronic cross section as a function of the kinetic energy in the two chosen data sets, the -60A and -100A negative runs. 
-As we will clarify in Section \ref{ch:PionXSCorrections},  the corrections to the raw cross section depend on the beam conditions and need to be calculated independently for the two datasets.
+As we will clarify in Section \ref{ch:PionXSCorrections},  the corrections to the raw cross section depend on the beam settings and need to be calculated independently for the two datasets.
 Thus, we present here the measurement of the raw cross section on the two datasets separately.
@@
 \begin{equation}
- \sigma_{TOT} (E_i)  = \frac{1}{n \delta X}\frac{N^{\text{TOT}}_{\text{Int}}(E_i)}{N^{\text{TOT}}_{\text{Inc}}(E_i)},
+ \sigma_{TOT} (E_i)  = \frac{1}{n\text{ } \delta X}\frac{N^{\text{TOT}}_{\text{Int}}(E_i)}{N^{\text{TOT}}_{\text{Inc}}(E_i)},  \tag{\ref{eq:thinTargetXSSolved}}
 \end{equation}
@@
 The data central points are represented by black dots, the statistical uncertainty is shown in black, while the systematic uncertainty is shown in red.
-Data is displayed over the $N^{\text{TOT}}_{\text{Int}}$  distribution obtained with a MC mixed sample of pions, muon and electrons (additional details on the composition will be provided in Section \ref{ch:BKGsubXS}).
+Data is displayed over the $N^{\text{TOT}}_{\text{Int}}$  distribution obtained with a MC sample of pions, muon and electrons weighted by the beam composition (additional details on the composition will be provided in Section \ref{ch:BKGsubXS}).
 The contribution from the simulated pions is shown in blue, the one from secondaries in red, the one from muons in yellow and the ones from electrons in gray.
@@
 Figure \ref{fig:XSRaw} shows the raw cross section for the 60A dataset on the left and for the 100A dataset on the right, statistical uncertainty in black and systematic uncertainty in red.
-The raw data cross section is overlaid to the reconstructed cross section for the MC mixed sample, displayed in azure.
+The raw data cross section is overlaid to the reconstructed cross section for the MC sample, displayed in azure.
 Since the background contributions and the detector effects for the 60A and 100A sample are different, it is premature to compare the raw cross sections obtained from the two samples at this point.
@@
 
-Since the interaction probability $P_{Interacting}$ is $\frac{ N^{\text{TOT}}_{\text{Int}}}{N^{\text{TOT}}_{\text{Inc}}}$ and the number of tries $\mathcal{N}$ is $N^{\text{TOT}}_{\text{Inc}}$, equation \ref{eq:binVar} translates into
+Since the interaction probability $P_{Interacting}$ is estimated by $\frac{ N^{\text{TOT}}_{\text{Int}}}{N^{\text{TOT}}_{\text{Inc}}}$ and the number of tries $\mathcal{N}$ is $N^{\text{TOT}}_{\text{Inc}}$, equation \ref{eq:binVar} translates into
 \begin{equation}
@@
 $N^{\text{TOT}}_{\text{Inc}}$ and $N^{\text{TOT}}_{\text{Int}}$ are not independent.
-The statistical uncertainty on the cross section is thus calculated as 
+In fact, the population of a given bin for the interacting histogram always implies at least the population of the same bin in the incident histogram  (and possibly other incidents bins at higher energies).
+Thus, we conservatively calculate the statistical uncertainty on the cross section as
 \begin{equation}
@@
 \section{Corrections to the Raw Cross Section}\label{ch:PionXSCorrections}
-As described in section \ref{ch:MCCorrections}, we need to apply a background correction and an efficiency correction in order to derive the true pion cross section from the raw cross section.
-The true cross section is given in equation \ref{eq:C},
+As described in section \ref{ch:MCCorrections}, we need to apply a background correction and an efficiency correction in order to derive the pion cross section from the raw cross section.
+The  cross section is given in equation \ref{eq:C},
 
 \begin{equation}
-   \sigma^{\pi^-}_{TOT}(E_{i})  = \frac{1}{n \delta X}\frac{ \epsilon^{\text{Inc}}(E_i)  \hspace{0.2cm} C^{\pi MC}_{\text{Int}} (E_{i}) \hspace{0.2cm} N^{\text{TOT}}_{\text{Int}} (E_{i}) }{   \epsilon^{\text{Int}}(E_i) \hspace{0.2cm} C^{\pi MC}_{\text{Inc}} (E_{i}) \hspace{0.2cm}  N^{\text{TOT}}_{\text{Inc}} (E_{i})}.
+   \sigma^{\pi^-}_{TOT}(E_{i})  = \frac{1}{n \text{ } \delta X}\frac{ \epsilon^{\text{Inc}}(E_i)  \hspace{0.2cm} C^{\pi MC}_{\text{Int}} (E_{i}) \hspace{0.2cm} N^{\text{TOT}}_{\text{Int}} (E_{i}) }{   \epsilon^{\text{Int}}(E_i) \hspace{0.2cm} C^{\pi MC}_{\text{Inc}} (E_{i}) \hspace{0.2cm}  N^{\text{TOT}}_{\text{Inc}} (E_{i})}.
  \tag{\ref{eq:C}}
@@
 We use the procedure described in \ref{ch:PionXSBkgSub2} to evaluate the relative pion content in the interacting histogram $C^{\pi MC}_{\text{Int}} (E_{i})$  and the relative pion content in the incident $C^{\pi MC}_{\text{Inc}} (E_{i})$.
-We start by evaluating the relative pion content assuming the beamline composition simulated by G4Beamline, whose pion, muon and electron percentages per beam condition are reported again in the first line of Table \ref{tab:beamlineSys}.
+We start by evaluating the relative pion content assuming the beamline composition simulated by G4Beamline, whose pion, muon and electron percentages per beam setting are reported again in the first line of Table \ref{tab:beamlineSys}.
 The left side of Figure \ref{fig:CorrectionsBeam} shows the  MC estimated  relative pion content for the interacting histogram as function of kinetic energy for the 60A runs (top) and 100A runs (bottom).
 The right side of the same figure shows the  MC estimated  relative pion content for the incident histogram as function of kinetic energy for the 60A runs (top) and 100A runs (bottom).
-In Figure \ref{fig:CorrectionsBeam} the central curves displayed in light blue are obtained using the beamline composition as predicted by G4Beamline: these are the correction curves for the relative pion content applied to data.
+In Figure \ref{fig:CorrectionsBeam} the central curves displayed in light blue are obtained using the beamline composition as predicted by G4Beamline: these are the correction curves for the relative pion content applied to data in equation \ref{eq:C}.
 
@@
 Since G4Beamline tracks only particles which cross all the wire chambers, pion events that decay in flight from WC1 to WC4 are not recorded by G4Beamline.
-Pion decays in the beamline could be trigger the beamline detectors in data, if the produced muon proceeds in the beamline.
+Pion decays in the beamline could  trigger the beamline detectors in data, if the produced muon propagates forward along the beamline.
 Thus, we take the G4Beamline prediction for muons as a lower bound in the composition: the effect of doubling the muon content (line 2 in Table \ref{tab:beamlineSys}) is shown in blue on Figure \ref{fig:CorrectionsBeam}.
@@
 
-\subsection{Efficiency Correction}\label{ch:EFFXS}
-The interaction point for a track used in the total hadronic cross section analysis is defined to be the last point of the WC2TPC matched track which lies inside the fiducial volume.
-This definition is independent from the topology of the interaction.
-If the TPC track stops within the fiducial volume, its last point will be the interaction point, no matter what the products of the interaction look like; if the track crosses the boundaries of the fiducial volume, the track will be considered ``through going" and no interaction point will be found.
-Given this definition, it is evident that we rely on the tracking algorithm to discern where the interaction occurred in the TPC  and correctly stop the tracking.
+\subsection{Correction for Reconstruction Effects}\label{ch:EFFXS}
+The interaction point for a particle in the selected sample for the total hadronic cross section analysis is the last point of its track that lies inside the LArTPC fiducial volume.
+This definition holds regardless the type of the interaction, i.e.
+if the TPC track ends within the fiducial volume, its last point will be the interaction point, no matter what the products of the interaction look like; on the other hand, if the track crosses the boundaries of the fiducial volume, the particle will be considered ``through going" and no interaction point will be found.
+Given this definition, it is evident that we rely on the tracking algorithm to discern where the interaction occurred in the TPC  and correctly end the tracking.
 The tracking algorithm has an intrinsic angle resolution as shown in section \ref{sec:angleRes}, which limits its efficiency, especially in the case of elastic scattering occurring a low angles.
-Thus, we need to apply an efficiency correction to data in order to retrieve the true cross section.
-The efficiency correction is evaluated separately for the interacting and incident histograms, namely $\epsilon^{\text{int}}_i$ and  $\epsilon^{\text{inc}}_i$, and propagated to the cross section as shown in  equation \ref{eq:C}.
+\textcolor{red}{Plus, there are instance where INSERT HERE THE STUFF ABOUT THE MIGRATION}
+Thus, we need to apply a correction accounting for mis-reconstruction of the interaction point in order to retrieve the true cross section.
+This correction is evaluated separately for the interacting and incident histograms bin by bin, namely $\epsilon^{\text{int}}(E_i)$ and  $\epsilon^{\text{inc}}(E_i)$, and applied in the cross section formula as shown in  equation \ref{eq:C}.
 
-\subsubsection{Efficiency Correction: Procedure}\label{sec:EffCorrection}
-We describe here the procedure to calculate the efficiency correction taking the interacting histogram as example and noting that the procedure is identical for the  incident histogram.
+\subsubsection{Reconstruction Effects Correction: Procedure}\label{sec:EffCorrection}
+We describe here the procedure to calculate the mis-reconstruction correction taking the interacting distribution as example and noting that the procedure is identical for the incident distribution.
 
-We derive the correction on a set of pure pion MC, calculating its value bin by bin as the ratio between the true bin content and the correspondent reconstructed bin content.
-The correction is then applied to the relevant bin in data.
-In formulae, the efficiency correction is calculated to be
+In section \ref{sec:angleRes}, we estimated the angular resolution for data and MC to be $\bar\alpha_{Data} = (5.0 \pm 4.5) \text{ deg}$  and 
+$\bar\alpha_{MC} = (4.5 \pm 3.9) \text{ deg}$, respectively.
+Most interaction angles smaller than the angular resolution will thus be indistinguishable  for the reconstruction.
+Thus, we claim we are able to  measure the cross section for interaction angles greater than 5.0 deg.
+Geant4 simulates interactions at all angles, as shown in figure \ref{fig:trueScatteringAngle}.
+In order to calculate the correction for reconstruction effects,  we select events which have an interaction angle greater than a given $\alpha_{res}$ to construct the true interacting and incident histograms (the denominator of the correction).
+
+We derive the correction $\epsilon^{\text{int}}(E_i)$ on a set of pure pion MC, calculating its value bin by bin as the ratio between the true bin content and the correspondent reconstructed bin content.
+The true interacting distribution is obtained applying the thin slice method on true MC energy deposition up to the MC flagged true interaction point for interaction angles greater than 5$^\circ$.
+The reconstructed MC interacting distribution is obtained treating the MC events through the same reconstruction process as data: the interaction point is given by the end of the tracking and its energy is given by the reconstructed calorimetric information.
+The correction is then applied to in data bin by bin.
+In formulae, the correction is calculated to be
 
@@
 
-where $N^{\text{ $\pi$ True MC}}_{\text{Int}} (E_{i}) $ is the content of the $i$-th bin in for the true interacting histogram, and $N^{\text{ $\pi$ Reco MC}}_{\text{Int}} (E_{i}) $ is the content of the $i$-th bin in for the reconstructed interacting histogram.
+where $N^{\text{ $\pi$ True MC}}_{\text{Int}} (E_{i}) $ is the content of the $i$-th bin in the true interacting histogram, and $N^{\text{ $\pi$ Reco MC}}_{\text{Int}} (E_{i}) $ is the content of the $i$-th bin in the reconstructed interacting histogram.
 The correction is applied to data as follows
@@
 \begin{equation}
-N^{\text{ $\pi$ True Data}}_{\text{Int}} (E_{i})  =  \frac{N^{\text{$\pi$ Reco Data}}_{\text{Int}} (E_{i})}{\epsilon^{\text{Int}} (E_{i}) } = N^{\text{$\pi$ Reco Data}}_{\text{Int}} (E_{i}) \frac{N^{\text{ $\pi$ True MC}}_{\text{Int}} (E_{i})}{ N^{\text{ $\pi$ Reco MC}}_{\text{Int}} (E_{i})}.
+N^{\text{ $\pi$  Data}}_{\text{Int}} (E_{i})  =  \frac{N^{\text{$\pi$ Reco Data}}_{\text{Int}} (E_{i})}{\epsilon^{\text{Int}} (E_{i}) } = N^{\text{$\pi$ Reco Data}}_{\text{Int}} (E_{i}) \frac{N^{\text{ $\pi$ True MC}}_{\text{Int}} (E_{i})}{ N^{\text{ $\pi$ Reco MC}}_{\text{Int}} (E_{i})}.
 \end{equation}
@@
 
-In section \ref{sec:angleRes}, we estimated the angular resolution for data and MC to be $\bar\alpha_{Data} = (5.0 \pm 4.5) \text{ deg}$  and 
-$\bar\alpha_{MC} = (4.5 \pm 3.9) \text{ deg}$, respectively.
-Most interaction angles smaller than the angular resolution will thus be indistinguishable  for the reconstruction.
-Thus, we claim we are able to  measure the cross section for interaction angles greater than 5.0 deg.
-Geant4 simulates interactions at all angles, as shown in figure \ref{fig:trueScatteringAngle}.
-In order to calculate the efficiency correction,  we select events which have an interaction angle greater than a given $\alpha_{res}$ to construct the true interacting and incident histograms (the denominator of the efficiency correction).
-The systematics on the efficiency correction is estimated by varying the value of $\alpha_{res}$ between 0 deg and 4.5 deg and propagating the uncertainty on the cross section.
+ The systematics on this correction is estimated by varying the value of $\alpha_{res}$ between 0 deg and 4.5 deg and propagating the uncertainty on the cross section.
 
@@
 \includegraphics[width=\textwidth]{Chapter-6/Images/60AEffCorr.pdf}
-\caption{\emph{Left:} Efficiency correction on the 60A interacting histogram, statistical uncertainty in blue, systematic uncertainty in red.
-\emph{Right:}  Efficiency correction on the 60A incident histogram, statistical uncertainty in blue, systematic uncertainty in red.}
+\caption{\emph{Left:} Reconstruction effects correction on the 60A interacting histogram, statistical uncertainty in blue, systematic uncertainty in red.
+\emph{Right:}  Reconstruction effects correction on the 60A incident histogram, statistical uncertainty in blue, systematic uncertainty in red.}
 \label{fig:EffCorr60A}
@@
 \includegraphics[width=\textwidth]{Chapter-6/Images/100AEffCorr.pdf}
-\caption{\emph{Left:} Efficiency correction on the 100A interacting histogram, statistical uncertainty in blue, systematic uncertainty in red.
-\emph{Right:}  Efficiency correction on the 100A incident histogram, statistical uncertainty in blue, systematic uncertainty in red.}
+\caption{\emph{Left:} Reconstruction effects correction on the 100A interacting histogram, statistical uncertainty in blue, systematic uncertainty in red.
+\emph{Right:}  Reconstruction effects correction on the 100A incident histogram, statistical uncertainty in blue, systematic uncertainty in red.}
 \label{fig:EffCorr100A}
@@
 \section{Results}\label{ch:FinalPion}
-Figure \ref{fig:FinalXSPion} show the measurement of the ($\pi^-$-Ar) total hadronic cross section for  scattering angles greater than 5$^\circ$, as the result of the background subtraction and efficiency correction to the raw cross section.
+Figure \ref{fig:FinalXSPion} show the measurement of the ($\pi^-$-Ar) total hadronic cross section for  scattering angles greater than 5$^\circ$, as the result of the background subtraction and reconstruction effects correction to the raw cross section.
 The top left plot is the measurement obtained on the 60A data, statistical uncertainty in black and systematic uncertainty in red.
@@
 
-The systematic uncertainty on the cross section is the sum in quadrature of the statistical uncertainty, the systematic uncertainty related to the kinetic energy measurement, the systematic uncertainty related to the beam composition and the systematic uncertainty related to the efficiency correction.
+The systematic uncertainty on the cross section is the sum in quadrature of the statistical uncertainty, the systematic uncertainty related to the kinetic energy measurement, the systematic uncertainty related to the beam composition and the systematic uncertainty related to the reconstruction effects correction.
 

### 2018-05-30 Chapter-7/7-KaonCrossSection.tex
s

@@
 In Section \ref{ch:KaonXSRaw}, we start by measuring the raw cross section.
-In Section \ref{ch:KaonXSCorrections}, we apply a statistical subtraction of the background contributions based on simulation and a correction for detection inefficiency.
+In Section \ref{ch:KaonXSCorrections}, we apply a statistical subtraction of the background contributions based on simulation and a correction for reconstruction effects.
 The final results are presented in Section \ref{ch:FinalKaon}.
@@
 \begin{equation}
- \sigma_{TOT} (E_i)  = \frac{1}{n \delta X}\frac{N^{\text{TOT}}_{\text{Int}}(E_i)}{N^{\text{TOT}}_{\text{Inc}}(E_i)},
+ \sigma_{TOT} (E_i)  = \frac{1}{n \delta X}\frac{N^{\text{TOT}}_{\text{Int}}(E_i)}{N^{\text{TOT}}_{\text{Inc}}(E_i)}, \tag{\ref{eq:thinTargetXSSolved}}
 \end{equation}
@@
 Data is displayed over the MC.
-For the  $N^{\text{TOT}}_{\text{Inc}}$ distribution we do not make a distinction between kaons that decay or interact hadronically because any kaon independently from its final interaction contributes to the flux of incident particles at given kinetic energy.
+For the  $N^{\text{TOT}}_{\text{Inc}}$ distribution we do not make a distinction between kaons that decay or interact hadronically because any kaon contributes to the flux of incident particles at given kinetic energy independently from its final interaction.
 The same normalization procedure is used for both the interacting and incident histograms.
@@
 \section{Corrections to the Raw Cross Section}\label{ch:KaonXSCorrections}
-As described in section \ref{ch:MCCorrections}, we need to apply a background correction and an efficiency correction in order to derive the true Kaon cross section from the raw cross section.
-The true cross section is given in equation \ref{eq:C},
+As described in section \ref{ch:MCCorrections}, we need to apply a background correction and a correction for reconstruction effects in order to derive the  kaon cross section from the raw cross section.
+The  cross section is given in equation \ref{eq:C},
 
@@
 
-Currently, the only background considered for the kaon hadronic cross section comes from the presence of secondaries.
+Currently, the only background considered for the kaon hadronic cross section comes from the presence of secondaries and decay interaction points.
 A further development of the analysis will need to account for the presence of a small proton contamination.
-Figure \ref{fig:CorrectionsBeamK} shows the relative kaon content for the interacting and incident histograms.
+Figure \ref{fig:CorrectionsBeamK} shows the relative kaon content suitable for the hadronic cross section for the interacting and incident histograms.
 \begin{figure}[p]
@@
 The correction is then applied to the relevant bin in data.
-The efficiency correction is evaluated separately for the interacting and incident histograms, namely $\epsilon^{\text{int}}_i$ and  $\epsilon^{\text{inc}}_i$, and propagated to the cross section as shown in  equation \ref{eq:C}.
+The correction for recontruction effects is evaluated separately for the interacting and incident histograms, namely $\epsilon^{\text{int}}(E_i)$ and  $\epsilon^{\text{inc}}(E_i)$, and propagated to the cross section as shown in  equation \ref{eq:C}.
 
@@
 Thus, we claim we are able to  measure the cross section for interaction angles greater than 4.5 deg.
-Geant4 simulates interactions at all angles: in order to calculate the efficiency correction,  we select events which have an interaction angle greater than a $\alpha_{res}$ to construct the true interacting and incident histograms (the denominator of the efficiency correction).
-The systematics on the efficiency correction is estimated by varying the value of $\alpha_{res}$ between 0 deg and 4.5 deg and propagating the uncertainty on the cross section.
+Geant4 simulates interactions at all angles: in order to calculate the correction for reconstruction effects,  we select events which have an interaction angle greater than a $\alpha_{res}$ to construct the true interacting and incident histograms (the denominator of the correction).
+The systematics on this correction is estimated by varying the value of $\alpha_{res}$ between 0 deg and 4.5 deg and propagating the uncertainty on the cross section.
 
@@
 \includegraphics[width=\textwidth]{Chapter-7/Images/EffCorrK.pdf}
-\caption{\emph{Left:} Efficiency correction on the interacting histogram, statistical uncertainty in blue, systematic uncertainty in red.
-\emph{Right:}  Efficiency correction on the incident histogram, statistical uncertainty in blue, systematic uncertainty in red.}
+\caption{\emph{Left:} Reconstruction  effects correction on the interacting histogram, statistical uncertainty in blue, systematic uncertainty in red.
+\emph{Right:}  Reconstruction  effects  correction on the incident histogram, statistical uncertainty in blue, systematic uncertainty in red.}
 \label{fig:EffCorrK}
@@
 \section{Results}\label{ch:FinalKaon}
-Figure \ref{fig:FinalXSKaon} show the measurement of the ($K^+$-Ar) total hadronic cross section for  scattering angles greater than 5$^\circ$, as the result of the background subtraction and efficiency correction to the raw cross section.
+Figure \ref{fig:FinalXSKaon} show the measurement of the ($K^+$-Ar) total hadronic cross section for  scattering angles greater than 5$^\circ$, as the result of the background subtraction and reconstruction  effects  correction to the raw cross section.
 The plot shows the measurement obtained on the full dataset, statistical uncertainty in black and systematic uncertainty in red.
@@
 
-The systematic uncertainty on the cross section is the sum in quadrature of the statistical uncertainty, the systematic uncertainty related to the kinetic energy measurement and the systematic uncertainty related to the efficiency correction.
+The systematic uncertainty on the cross section is the sum in quadrature of the statistical uncertainty, the systematic uncertainty related to the kinetic energy measurement and the systematic uncertainty related to the reconstruction  effects  correction.
 

### 2018-05-30 thesis_main.tex
s

@@
 \usepackage{rotating}
+%\usepackage{newtxtext,newtxmath,amsmath}
+\usepackage{ textcomp }
 
@@
 
-A lot of people are awesome, especially you, since you probably agreed to read this when it was a draft.
+%A lot of people are awesome, especially you, since you probably agreed to read this when it was a draft.
 
@@
 
-%Mom, you are and will always be the hero of my story; I know this has been challenging for you, thanks for being there for me at every hour of the day and the night, for being always honest with me and for somehow handling this whole US-Italy thing gracefully.
+Mom, you are and will always be the hero of my story; I know this has been challenging for you, thanks for being there for me at every hour of the day and the night, for being always honest with me and for somehow handling this whole US-Italy thing gracefully.
 I can't believe you still haven't met Bonnie.
@@
 
-%Bonnie Fleming, my advisor, I don't even know where to start in thanking you.
+Bonnie Fleming, my advisor, I don't even know where to start in thanking you.
 Thanks for listening to me before I proved I was worth be listened to; thanks for always believing I could make it, even at the bottom of my doubts; thanks for all the laughters, for all the freedom, for all the great advices; thanks for leading by example; thanks for being always supportive of all my crazy side projects; thanks for choosing each and every person in the most wonderful research group ever and thanks for letting me be part of it.
@@
 
-%Flavio Cavanna, my co-advisor, working with you has been one of the greatest pleasures of my academic life; the extent of your knowledge never ceases to amaze me and the freshness of your enthusiasm is always contagious.
-Thanks for all the time spent in your office banging our heads on these cross section analyses.
+Flavio Cavanna, my co-advisor, working with you has been one of the greatest pleasures of my academic life; the extent of your knowledge never ceases to amaze me and the freshness of your enthusiasm is always contagious.
+Thanks for all the time spent in your office banging our heads on these cross section analyses and thanks for convincing me to join LArIAT in the first place.
 
-%Ornella Palamara, after a particularly heated exchange at a wine and cheese, I once told Andrzej I think of you as my spirit animal: fierce, knowledgeable and direct; he replied I still need to walk many miles to get there and I can't agree more.
+Ornella Palamara, after a particularly heated exchange at a wine and cheese, I once told Andrzej I think of you as my spirit animal: fierce, knowledgeable and direct; he replied I still need to walk many miles to get there and I can't agree more.
 
-%A big thank you to my academic sisters, from the youngest to the oldest (again academically speaking): Supraja Balasubramanian, you worked the magic of enriching my past, thank you from the bottom of my heart;  Brooke Morrison, the depth of your knowledge, your drive and love for our field has been and will always be a constant inspiration;  Ariana  Hackenburg, thanks for keeping me sane for the last four years, if you run for president -- as you should -- remember to delete our Skype chats, love and hugs to you dudette; and last but not least Xiao Luo, Fermilab has given me lots of extraordinary gifts, but no one  bigger than your friendship.
+A big thank you to my academic sisters, from the youngest to the oldest (again academically speaking): Supraja Balasubramanian, you worked the magic of enriching my past, thank you from the bottom of my heart;  Brooke Morrison, the depth of your knowledge, your drive and love for our field has been and will always be a constant inspiration;  Ariana  Hackenburg, thanks for keeping me sane for the last four years, if you run for president -- as you should -- remember to delete our Skype chats, love and hugs to you dudette; and last but not least Xiao Luo, Fermilab has given me lots of extraordinary gifts, but no one  bigger than your friendship.
 
-%Jen Raaf and Jonathan Asaadi, these analyses would have not seen the light of day without you.
-Thanks for your many inputs and for letting your door open to talk about physics or my latest paranoia.
+Jen Raaf and Jonathan Asaadi, these analyses would have not seen the light of day without you.
+Thanks for your many inputs and for letting your door open to talk about physics and/or my latest paranoia.
 Thanks for the patience and understanding.
@@
 
-%To all the members of the NPG, thanks for all the laughters and all the wise insights. You're the smartest bunch of idiots I've ever met, I love y'all.
+To all the members of the NPG, thanks for all the laughters and all the wise insights. You're the smartest bunch of idiots I've ever met, I love y'all.
 
-%Thanks to the MicroBooNE collaboration Sam kazu and kirby
+Thanks to the MicroBooNE collaboration Sam kazu and kirby
 
-%Shany Danieli and Louise Laage, you're the piece of heart I left in New Haven.
+Shany Danieli and Louise Laage, you're the piece of heart I left in New Haven.
 Shany, I don't know what made you think I was the cool kid in the group, but I'm so glad you stuck with me.
 I'm not sure I would have survived my first semester at Yale without you: you shine in your smartness and lift other people with you -- it's a rare quality in our field.
-Louise, my dear, thanks for always finding the time to listening, for your attuned spirit and for the magnolias.
+Louise, my dear, thanks for always finding the time to listen, for your attuned spirit and for the magnolias.
 
-%Thanks to all the people which helped me make Chicago and Fermilab my new home, in particular David Caratelli,  Mateus Carneiro, Pilar Coloma, Ciaran Hughes, Gordan Krnjaic, Seyda Ipek, Kiel Howe and Rob Fine.
+Thanks to all the people which helped me make Chicago and Fermilab my new home, in particular David Caratelli,  Mateus Carneiro, Pilar Coloma, Ciaran Hughes, Gordan Krnjaic, Seyda Ipek, Kiel Howe and Rob Fine.
 
-%To the extraordinary people who stayed close, even with an ocean in between: la Tasso, Marci, la Bea, la Kittirn, Alvin, la Viola e Brogna, Sighi, Eleonora and all the people from the ``La Chiappia Nuova" group, thanks for making me feel like I never left.
+To the extraordinary people who stayed close, even with an ocean in between: la Tasso, Marci, la Bea, la Kittirn, Alvin, la Viola e Brogna, Sighi, Eleonora and all the people from the ``La Chiappia Nuova" group, thanks for making me feel like I never left.
 
-%Finally, thanks to Anto, my husband, my topo and my adventure companion, cause people might think I'm nice and easy, but you know I'm not...
+Finally, thanks to Anto, my husband, my topo and my adventure companion, cause people might think I'm nice and easy, but you know I'm not...
 And still, you followed me across half the world and you're still here, so that I could never feel alone.
@@
 \mainmatter
-%\input{Chapter-1/1-Theory}
-%\input{Chapter-2/2-LArTPCs}
-%\input{Chapter-3/3-TheLArIATExperiment}
+\input{Chapter-1/1-Theory}
+\input{Chapter-2/2-LArTPCs}
+\input{Chapter-3/3-TheLArIATExperiment}
 \input{Chapter-4/4-Methodology}

### 2018-05-30 Chapter-4/4-Methodology.tex
s

@@
 We look for short tracks contained in the ROI, as depicted in figure \ref{fig:showerFilt}:  if more then 5 tracks shorter than 10 cm are in the ROI, we reject the event.
-Line 7 in in Table \ref{tab:beamlineDataSelection} shows the number of events surviving this selection.
+Line 7 in  Table \ref{tab:beamlineDataSelection} shows the number of events surviving this selection; that table also shows that this selection is applied after the wire chamber to TPC match (discussed in the next section).
+The WC2TPC match already lower the presence of electrons in the sample, which is further reduced by the shower filter.
 
@@
 Thus, the LArIAT TPC, with its 90 cm of length, is not a thin target.
-However, the fine-grained tracking of the LArIAT LArTPC allows us to treat the argon volume as a sequence of many adjacent thin targets.
+However, the granularity of the LArIAT LArTPC detector allows us to treat the argon volume as a sequence of many adjacent thin targets.
 
@@
 The collection wires collect signals proportional to the energy deposited by the hadron along its path in a  $\delta${\emph{X}} = 4 mm/(sin($60^{\circ}$)cos($3^{\circ}$)) $\approx$ 4.7~mm slab of liquid argon.
-Thus, one can think to slice the TPC into many thin targets of $\delta${\emph{X}} = 4.7~mm thickness along the direction of the incident particle, making a measurement at each wire along the path.
+Thus, one can think to slice the TPC into many thin targets of $\delta${\emph{X}} = 4.7~mm thickness along the direction of the incident particle, making a measurement at each wire along the path, as sketched in Figure \ref{fig:TPCGran}.
+
+\begin{figure}
+  \centering  
+\includegraphics[width=0.8\textwidth]{Chapter-4/Images/LArSlice.png}
+\caption{Representation of sliced LAr Volume.}
+\label{fig:TPCGran}
+\end{figure}
 

### 2018-05-30 Chapter-6/6-PionCrossSection.tex
s

@@
 \section{Results}\label{ch:FinalPion}
-Figure \ref{fig:FinalXSPion} show the measurement of the ($\pi^-$-Ar) total hadronic cross section for  scattering angles greater than 5$^\circ$, as the result of the background subtraction and reconstruction effects correction to the raw cross section.
-The top left plot is the measurement obtained on the 60A data, statistical uncertainty in black and systematic uncertainty in red.
-The top right plot is the measurement obtained on the 100A data, statistical uncertainty in black and systematic uncertainty in blue.
+Figure \ref{fig:FinalXSPion} shows the measurement of the ($\pi^-$-Ar) total hadronic cross section for  scattering angles greater than 5$^\circ$, as the result of the background subtraction and reconstruction effects correction to the raw cross section.
+The top left plot is the measurement obtained on the 60A data, statistical and systematic uncertainty shown in red.
+The top right plot is the measurement obtained on the 100A data, statistical uncertainty and systematic uncertainty shown in blue.
 The bottom plot shows the two measurements overlaid.
@@
 The systematic uncertainty on the cross section is the sum in quadrature of the statistical uncertainty, the systematic uncertainty related to the kinetic energy measurement, the systematic uncertainty related to the beam composition and the systematic uncertainty related to the reconstruction effects correction.
+
+The cross section measurements in the two datasets agree within the systematic uncertainty, even if the cross section measured in the 60A data is lower than the one measured in the 100A data in all overlapping bins except for the 200-250 MeV bin.
 
@@
 \label{fig:FinalXSPion}
+\end{figure}
+
+The top panel of Figure \ref{fig:FinalXSResPion} shows the cross section obtained combining the two datasets.
+To combine the datasets, we merge the corrected interacting histogram for the 60A data with the interacting histogram for the 100A data; we apply the same merging with the incident histograms.
+We then calculate the cross section central point and statistical uncertainty using the merged interacting and incident histograms.
+The systematic uncertainty is calculated as the sum in quadrature of the original systematic uncertainties on the cross section measurements on the two separated data sets.
+
+The bottom panel of figure  \ref{fig:FinalXSResPion} shows the relative difference between the measured cross section and the Geant4 predicted cross section.
+A couple of considerations are in order here.
+On one hand, we note the general agreement between the measured and predicted cross sections.
+We expected the pion cross section in argon to be well modeled in Geant4, given the vast set of measurements available for pion interaction -- albeit not on argon.
+Thus, this agreement augments our confidence in the experimental methodology used.
+On the other hand, this agreement is not perfect.
+As a matter of fact, its imperfections are rather suggestive: we see a slight shift of the delta resonance at higher energies and a small shape difference, with a steeper drop between the delta region and the flatter part at energies greater than 400 MeV.
+Since we see room for improvement of the systematic uncertainties, we look forward to see if this difference in shape will be confirmed by further analyses.
+
+\begin{figure}[htb]
+\centering
+\includegraphics[width=\textwidth]{Chapter-6/Images/TheRealMoneyPlot.pdf}
+\caption{ \emph{Top:} ($\pi^-$-Ar) total hadronic cross section for  scattering angles greater than 5$^\circ$ measured in the combined sample, statistical uncertainty and systematic uncertainty in black.
+The Geant4 prediction for the total hadronic cross section for angle scattering greater than 5$^\circ$ is displayed in green.
+\emph{Bottom:} Relative difference between the measured cross section and the Geant4 prediction.
+}
+\label{fig:FinalXSResPion}
 \end{figure}

### 2018-05-30 thesis_main.tex
s

@@
 
-\linenumbers
+%\linenumbers
 
@@
 % Need to define title before the abstract.
-\title{Measurement of ($\pi^-$-Ar) and ($K^+$-Ar) total hadronic cross sections in the LArIAT experiment}
+\title{Measurement of ($\pi^-$-Ar) and ($K^+$-Ar) total hadronic cross sections at the LArIAT experiment}
 \author{Elena Gramellini}
@@
 neutrino oscillation physics and proton decay searches of the Short
-Baseline Neutrino and Long Baseline Neutrino programs.
+Baseline Neutrino (SBND, MicroBooNE, ICARUS) and Long Baseline Neutrino programs (DUNE).
 
@@
 Finally,  we apply the ``thin slice method" technique and measure the cross section, correcting for background and detector effects.
-The thin slice technique is a new method to measure hadron-argon cross sections possible only thanks to the combination of the tracking and calorimetry capability of the LArTPC technology.
-Albeit never on argon, the hadronic cross section of pions has been measured before on several different elements in thin target experiments, leading to solid predictions in the argon case.
+The thin slice technique is a new method to measure hadron-argon cross sections possible only due to the combination of the tracking and calorimetry capability of the LArTPC technology.
+Albeit never on argon, the hadronic cross section of pions has been measured before on several different elements in thin target experiments, leading to solid predictions for measurements on argon.
 Through the use of a different technique, our measurement of the  ($\pi^-$-Ar) total hadronic cross section is in general agreement with the prediction by the Geant4 Bertini Cascade model which are based on data from thin target experiments.
 On the contrary, cross section measurements for kaons are extremely rare, thus more difficult to model.
-Not surprisingly, our measurement of the  ($K^+$-Ar) total hadronic cross section is mostly in disagreement with the Geant4 prediction.
+Our measurement of the  ($K^+$-Ar) total hadronic cross section is mostly in disagreement with the Geant4 prediction and provides new experimental data to properly tune the Geant4 models.
 

### 2018-05-31 Chapter-0/0-Introduction.tex
s

@@
-`\chapter{Introduction}\label{ch:-1}
-
-This thesis work concerns the first measurement of the  ($\pi^-$-Ar)  total hadronic cross section  in the 100-1000 MeV  kinetic energy range and the first measurement of the ($K^+$-Ar) total hadronic cross section  in the 100-650 MeV  kinetic energy range.
-We performed these measurements at the LArIAT experiment,  a small (0.25 ton)  Liquid Argon Time Projection Chamber (LArTPC) on a beam of charged particles at the Fermilab Test Beam Facility.
-Albeit particle and nuclear physics have a long history of hadronic cross section measurements, the work outlined in this thesis presents a new methodology -- the ``thin slice method" -- for cross section measurements in argon, possible only thanks to the detection capabilities of the LArTPC technology.
-The combination of fine-grained tracking and excellent calorimetric information provided by the LArTPC technology  allows to see unprecedented details of particle interactions in argon and, in LArIAT, to measure the kinetic energy of a hadron at each step of the tracking.
-A renewed interest for precision measurements of hadronic cross sections, particularly in argon, arises from the current  panorama of experimental  particle physics at the intensity frontier.
-
+\chapter{Introduction}\label{ch:-1}
 The discovery of the Higgs boson in 2012 marked the triumph of the Standard Model of Particle Physics; exploring what lays beyond is the real challenge in our field today. 
@@
 From an experimental point of view, we are developing technologies and huge collaborations to probe these theories.
-As we enter the era of precision measurements of neutrino interaction, neutrinos might hold the key to the next generation of discoveries in particle physics.
+As we enter the era of high statistics, precision measurements of neutrino interactions, neutrinos might hold the key to the next generation of discoveries in particle physics.
+
+This thesis work describes the first measurement of the  ($\pi^-$-Ar)  total hadronic cross section  in the 100-1000 MeV  kinetic energy range and the first measurement of the ($K^+$-Ar) total hadronic cross section  in the 100-650 MeV  kinetic energy range.
+These measurements were performed with the LArIAT experiment,  a small (0.25 ton)  Liquid Argon Time Projection Chamber (LArTPC) in a beam of charged particles at the Fermilab Test Beam Facility.
+Particle and nuclear physics have a long history of hadronic cross section measurements; what makes these measurements unique is both the target (argon) and the methodology used -- the ``thin slice method" -- which takes advantage of the detection capabilities of the LArTPC technology.
+The combination of fine-grained tracking and excellent calorimetric information provided by the LArTPC technology  enables to the measurement of unprecedented details of particle interactions in argon and, in LArIAT, to measure the kinetic energy of a hadron at each step along the particle traces.
+A renewed interest for precision measurements of hadronic cross sections, particularly in argon, arises from the current  panorama of experimental  particle physics at the intensity frontier, in particular neutrino physics in LArTPCs.
 
@@
 Following the recommendations of  the latest Particle Physics Project Prioritization Panel  \cite{P5}, the US particle physics panorama is directing a substantial effort towards the exploration of the intensity frontier through the construction of massive LArTPCs.
-In particular, the near future will see the development of a Short Baseline Neutrino Program (SBN) and long baseline neutrino program  (DUNE), both based on the LArTPC detector technology.
+In particular, we are seeing the development of a Short Baseline Neutrino Program (SBN) and long baseline neutrino program  (DUNE), both based on the LArTPC detector technology.
 The US liquid argon program has the potential to answer many of the fundamental open questions in particle physics today, such as: is there a fourth generation neutrino?
@@
 The SBN program at Fermilab is tasked with conclusively addressing the existence of a fourth neutrino generation in the  $\Delta m^2= \Delta m^2_{14} \sim [0.1 - 10]$ eV$^2$ parameter space.
-The SBN program entails three surface LArTPCs positioned on the Booster Neutrino Beam at different distances from the neutrino production in oder to fully exploit  the L/E dependence of the oscillation pattern:  SBND (100 m from the decay pipe), MicroBooNE (450 m), and ICARUS (600 m).
+The SBN program entails three surface LArTPCs positioned on the Booster Neutrino Beam at different distances from the neutrino production in oder to fully exploit  the L/E dependence of the oscillation pattern:  SBND (110 m from the decay pipe), MicroBooNE (470 m), and ICARUS (600 m).
 SBN will also perform an extensive
@@
 The complexity of the LArTPC technology for neutrino detection is due to several reasons.
-Argon is a fairly heavy element, which means that nuclear effects play an important role in the looks of the interaction topology.
+Argon is a fairly heavy element, which means that nuclear effects play an important role in the interaction topology.
 For example, pions are one of the main products of neutrino interactions; yet,  since data on charged particle interaction in argon is scarce, neutrino event generators have big uncertainties in the re-scattering simulation of pion in argon.
 %No measurement of hadronic cross sections for pions or kaons is available for argon (yet).
-The amount of details in an LArTPC event is easily  parsed by human eye, but can make automatic event reconstruction rather challenging.
-Thus, reconstruction algorithms in LArTPC need to be tune to recognize the different topologies of the neutrino interaction products in argon.
-This is particularly true for pions, since they are an abundant product of the neutrino interactions: the occurrence of a pion interaction in argon can modify the topology of the neutrino event, causing a misidentification of the neutrino interaction.
+The amount of details in an LArTPC event can easily  parsed by human eye, but can make automatic event reconstruction rather challenging.
+Thus, reconstruction algorithms in LArTPC need to be tuned to recognize the different topologies of the neutrino interaction products in argon.
+This is particularly true for pions, since they are an abundant product of the neutrino interactions: the occurrence of a pion interaction in argon can modify the topology of the neutrino event, causing a misidentification of the neutrino interaction channel.
 
-The LArIAT \cite{Cavanna:2014iqa} experiment is performing precise cross section measurements of charged particles in argon to bridge this gap of knowledge. 
+The LArIAT \cite{Cavanna:2014iqa} experiment is performing precise cross section measurements of charged particles in argon to address this gap of knowledge. 
 The LArIAT LArTPC sits on a beam of charged particles at the Fermilab Test Beam Facility which provides charge particles of the type and energy range relevant for neutrino interaction of both SBN and DUNE.
@@
 Then, the hadronic interaction point is identified and the raw cross section is calculated.
-Two corrections are then applied to the raw cross section -- a background subtractions and a correction for detector effects -- to obtain the true cross section measurement.\\
+Two corrections are then applied to the raw cross section -- a background subtractions and a correction for detector effects -- to obtain the cross section measurement, presented here.\\
 

### 2018-05-31 Chapter-6/6-PionCrossSection.tex
s

@@
 On one hand, we note the general agreement between the measured and predicted cross sections.
-We expected the pion cross section in argon to be well modeled in Geant4, given the vast set of measurements available for pion interaction -- albeit not on argon.
-Thus, this agreement augments our confidence in the experimental methodology used.
+We expected the pion cross section in argon to be well modeled in Geant4, given the vast set of measurements available for pion interaction on different targets -- even if no data is available for argon.
+Thus, this agreement confirms our confidence in the experimental methodology used.
 On the other hand, this agreement is not perfect.
 As a matter of fact, its imperfections are rather suggestive: we see a slight shift of the delta resonance at higher energies and a small shape difference, with a steeper drop between the delta region and the flatter part at energies greater than 400 MeV.
-Since we see room for improvement of the systematic uncertainties, we look forward to see if this difference in shape will be confirmed by further analyses.
+As we showed in Figure \ref{fig:pi1}, the shape of the delta resonance changes as a function of the nuclear target; thus it is plausible that the Geant4 model under estimates the cross section for argon in the delta resonance region and overestimates it in the higher energy region.
+Since we see room for improvement of the systematic uncertainties, we look forward to seeing if this difference in shape will be confirmed by further analyses, leading to a re-tuning of the model.
 

### 2018-05-31 Chapter-7/7-KaonCrossSection.tex
s

@@
 Figure \ref{fig:FinalXSKaon} show the measurement of the ($K^+$-Ar) total hadronic cross section for  scattering angles greater than 5$^\circ$, as the result of the background subtraction and reconstruction  effects  correction to the raw cross section.
-The plot shows the measurement obtained on the full dataset, statistical uncertainty in black and systematic uncertainty in red.
+The plot shows the measurement obtained on the full dataset, statistical and systematic uncertainty in black.
 The Geant4 prediction for the total hadronic cross section for angle scattering greater than 5$^\circ$ is displayed in green.
@@
 
+The measured cross section is mostly in disagreement with the predicted cross section; in particular it is significantly higher -- on the order of 20\%-30\% higher -- in the bins lower than 400 MeV.
+The 100 MeV to 400 MeV range is the region of interest for the expected kaons coming from a proton decay event.
+
 \begin{figure}[htb]
 \centering
-\includegraphics[width=0.48\textwidth]{Chapter-7/Images/TheMoneyPlotK.pdf}
-\caption{ ($K^+$-Ar) total hadronic cross section for  scattering angles greater than 5$^\circ$ measured in the 60A sample, statistical uncertainty in black and systematic uncertainty in red.
+\includegraphics[width=\textwidth]{Chapter-7/Images/TheMoneyPlotK.pdf}
+\caption{ \emph{Top:} ($K^+$-Ar) total hadronic cross section for  scattering angles greater than 5$^\circ$ measured in the combined +60A and +100A data samples, statistical uncertainty and systematic uncertainty in black.
 The Geant4 prediction for the total hadronic cross section for angle scattering greater than 5$^\circ$ is displayed in green.
-}
+\emph{Bottom:} Relative difference between the measured cross section and the Geant4 prediction.}
 \label{fig:FinalXSKaon}

### 2018-05-31 Chapter-8/8-Conclusions.tex
s

@@
 \chapter{Conclusions}\label{ch:Conclusions}
-In the era of neutrino precision measurements, of huge liquid argon detectors and of massive amount of information from LArTPCs, a renewed interest for an ancient measurement arises: the measurement of hadronic interactions with matter.
+In the era of neutrino precision physics, of huge liquid argon detectors and of massive amount of information from LArTPCs, a renewed interest in an ancient measurement arises: the measurement of hadronic interactions with matter.
 With this work, we presented the first ever ($\pi^-$-Ar) and ($K^+$-Ar) total hadronic cross section measurements as a function of the hadron kinetic energy.
-These analyses are the first physics analyses developed by the LArIAT experiment.
-Both the analysis follow a similar workflow and  they rely on beam line detector information as well as both calorimetry and tracking in the TPC.
+Both the analysis follow a similar workflow and  they rely on beam line detector information as well as both calorimetry and tracking in the LArTPC.
 
-In order to measure ($\pi^-$-Ar) total hadronic  argon cross sections, we start by selecting pion beamline candidates through a series of selections on the beamline and TPC information apt to maximize the number of pions in the selection over the number of muons and electrons.
+For the ($\pi^-$-Ar) total hadronic cross section measurement, we start by identifying pion beamline candidates through a series of selections on the beamline and TPC information which maximize the number of pions over other particle species (muons and electrons).
 We use the LArIAT beamline MC to estimate the beam composition of the selected beamline candidates and we propagate them to the LArAIT TPC constructing a properly weighted sample with the DDMC.
-We apply the thin slice method on the pion candidates and obtain the raw cross section measurement.
-From the simulated sample, we obtain two corrections accounting for the beamline background contamination and for detector effects.
-Finally, we apply the corrections to data and measure the true cross section.
+We apply the thin slice method on the pion candidates and obtain the raw cross section.
+From the simulated sample, we obtain two corrections accounting for the beamline background contamination and for reconstruction effects.
+Finally, we apply the corrections to data and measure the cross section.
 
 In order to measure ($K^+$-Ar) total hadronic  argon cross sections, we follow a similar procedure, i.e.
-we apply the thin slice method on kaon candidates identified in the beamline to obtain the raw cross section.
-We apply a background correction and a correction for detector effects to the raw cross section.
+we apply the thin slice method on kaon candidates identified in the beamline and obtain the raw cross section.
+We then apply a MC derived background correction and a correction for detector effects to obtain the cross section.
 The background correction accounts for the presence of secondary particles in both the interacting and incident histograms and for the presence of decay events in the interacting plot.
 
-The final results for the ($\pi^-$-Ar) and ($K^+$-Ar) total hadronic cross section are shown side by side in figure \ref{fig:finalfinal}.
+The final results for the ($\pi^-$-Ar) and ($K^+$-Ar) total hadronic cross section are shown side by side in the top part of Figure \ref{fig:finalfinal}.
+The bottom part of the same figure shows the relative deviation from the cross section predicted by Geant4.
+In the pion case, the measured cross section mostly agrees with the Geant4 prediction.
+A hint of a difference in shape is present: Geant4 seems to underestimate the cross section in the delta resonance region and overestimate it for energies greater than 400 MeV.
+Since the   ($\pi^-$-Ar) has never been measured before on Argon and since previous measurements on different nuclei show that the shape of the delta resonance peak varies as a function of the target nucleus, this hint to a shape disagreement is an interesting feature of this measurement which LArIAT will explore with a future development of the analysis.
+In the kaon case, the measured cross section is mostly higher than the Geant4 prediction.
+Contrary to the pion case, previous data on kaons are scarce even on different nuclei; thus, we do not expect the Geant4 prediction to be as solid in the kaon case as it is in the pion case.
+The nice agreement between the measured cross section and the prediction of the pion case gives us confidence in the methodology, thus confidence in the kaon measurement.
+With the kaon cross section measurement, we hope to inform the next generation of Geant4 model tuning, thus providing a more reliable tool in the simulation for proton decay searches.
+
+The ($\pi^-$-Ar) and ($K^+$-Ar) total hadronic cross section analyses are the first physics analyses developed by the LArIAT experiment and  will serve as a basis for the future cross section measurements of pions and kaons for the exclusive channels.
 
@@
 \centering
-\includegraphics[width=0.48\textwidth]{Chapter-6/Images/TheMoneyPlot.pdf}
+\includegraphics[width=0.48\textwidth]{Chapter-6/Images/TheRealMoneyPlot.pdf}
 \includegraphics[width=0.48\textwidth]{Chapter-7/Images/TheMoneyPlotK.pdf}
-\caption{\emph{Left:} ($\pi^-$-Ar) total hadronic cross section measurements in the 60A and 100A samples overlaid with the  Geant4 prediction (green).
-\emph{Right:} ($K^+$-Ar) total hadronic cross section for  scattering angles greater than 5$^\circ$ measured in the 60A sample, statistical uncertainty in black and systematic uncertainty in red.
-The Geant4 prediction for the total hadronic cross section for angle scattering greater than 5$^\circ$ is displayed in green.
+\caption{\emph{Left:} ($\pi^-$-Ar) total hadronic cross section measurements in the combined sample overlaid with the  Geant4 prediction (green).
+\emph{Right:} ($K^+$-Ar) total hadronic cross section for  scattering angles greater than 5$^\circ$ measured.
+The respective Geant4 prediction for the total hadronic cross section for angle scattering greater than 5$^\circ$ is displayed in green.
 }
@@
 \end{figure}
-
-These analyses' will serve as a basis for the future cross section measurements of pions and kaons for the exclusive channels in LArIAT.

### 2018-05-31 thesis_main.tex
s

@@
 
-{\raggedleft ``\emph{Dunque io ringrazio tutti quanti.} \par}
-{\raggedleft \emph{Specie la mia mamma che mi ha fatto cos\'i funky.}"\par}
-{\raggedleft -- Articolo 31, Tanqi Funky, 1996 -- \par}
-\vspace{0.5cm}
+\noindent ``\emph{Dunque io ringrazio tutti quanti.} \hfill ``\emph{At last, I thank everyone.}
 
-{\raggedleft ``\emph{At last, I thank everyone.} \par}
-{\raggedleft \emph{Especially my mom who made me so funky.}"\par}
+\noindent \emph{Specie la mia mamma} \hfill \emph{Especially my mom}
+
+\noindent \emph{che mi ha fatto cos\'i funky.}" \hfill \emph{who made me so funky.}"
+
 {\raggedleft -- Articolo 31, Tanqi Funky, 1996 -- \par}
@@
 
-%I jumped the pond a little less than 5 years ago because I had a taste of what being a physicists in the US might be like and wanted more. 
-%I came to this vast and at times inhospitable land thinking I would learn a lot of physics (which I did) and hoping I'd keep doing research for as long as I could (which I'm fortunate enough to say I'll still do for a while).
+I jumped the pond a little less than 5 years ago because I had a taste of what being a physicists in the US might be like and wanted more.
+I came to this vast and sometimes inhospitable land thinking I'd learn a lot of physics (which I did) and hoping I'd keep doing research for as long as I could (which I'm fortunate enough to say I'll still do for a while).
 Little did I know that getting my PhD would be first and foremost a journey to self discovery, the beginning of a path to asymptotically tend to the person and the scientist I want to be.
-A journey impossible without the help of many people, which I will now proceed to thank, hoping that the acknowledgements won't be longer than the thesis itself.
+A journey impossible without the help of many people, which I will now proceed to thank; let's hope that the acknowledgements won't be longer than the thesis.
 
@@
 
-A big thank you to my academic sisters, from the youngest to the oldest (again academically speaking): Supraja Balasubramanian, you worked the magic of enriching my past, thank you from the bottom of my heart;  Brooke Morrison, the depth of your knowledge, your drive and love for our field has been and will always be a constant inspiration;  Ariana  Hackenburg, thanks for keeping me sane for the last four years, if you run for president -- as you should -- remember to delete our Skype chats, love and hugs to you dudette; and last but not least Xiao Luo, Fermilab has given me lots of extraordinary gifts, but no one  bigger than your friendship.
+A big thank you to my academic sisters, from the youngest to the oldest (again academically speaking): Supraja Balasubramanian, you worked the magic of enriching my past, thank you from the bottom of my heart;  Brooke Morrison, the depth of your knowledge, your drive and love for our field has been a constant inspiration;  Ariana  Hackenburg, thanks for keeping me sane for the last four years, if you run for president -- as you should -- remember to delete our Skype chats, love and hugs to you dudette; and last but not least Xiao Luo, Fermilab has given me lots of extraordinary gifts, but no one  bigger than your friendship.
 
@@
 
+Thanks to the MicroBooNE experiment which has been an incredible learning experience.
+In particular, I want to thank  Sam Zeller for the countless pieces of advice,  my CRT troopers Martin, David, Rui, Igor and Linda  (if I know something about hardware it's because of you), and Kazu and  Kirby (if I know something about software it's because of you).
+
 To all the members of the NPG, thanks for all the laughters and all the wise insights. You're the smartest bunch of idiots I've ever met, I love y'all.
-
-Thanks to the MicroBooNE collaboration Sam kazu and kirby
 
@@
 
-Thanks to all the people which helped me make Chicago and Fermilab my new home, in particular David Caratelli,  Mateus Carneiro, Pilar Coloma, Ciaran Hughes, Gordan Krnjaic, Seyda Ipek, Kiel Howe and Rob Fine.
+Thanks to all the people which helped me make Chicago and Fermilab my new home, in particular David Caratelli,  Mateus Carneiro, Pilar Coloma, Ciaran Hughes, Gordan Krnjaic, Seyda Ipek, Kiel Howe, Adi Ashkenazi and Rob Fine.
 
-To the extraordinary people who stayed close, even with an ocean in between: la Tasso, Marci, la Bea, la Kittirn, Alvin, la Viola e Brogna, Sighi, Eleonora and all the people from the ``La Chiappia Nuova" group, thanks for making me feel like I never left.
+To the wonderful people who stayed close even with an ocean in between: la Tasso, Marci, la Bea, la Kittirn, Alvin, la Viola e Brogna, Sighi, la Secchi and the people from the ``La Chiappia Nuova" group, thanks for making me feel like I never left.
 
 Finally, thanks to Anto, my husband, my topo and my adventure companion, cause people might think I'm nice and easy, but you know I'm not...
-And still, you followed me across half the world and you're still here, so that I could never feel alone.
+And still, you followed me across half the world and you're still here, so that I can never feel alone.
 

### 2018-05-31 Chapter-1/1-Theory.tex
s

@@
 In fact, since the spatial inversion operator throws $\psi_R \leftrightarrow \psi_L$, parity is conserved only if both chiral components exist at the same time.
-For the neutrino introduction in the SM, experiments came in help of the theoretical description.
+For the neutrino introduction in the SM, experiments were critical to a more accurate theoretical description.
 The constraint of parity conservation weakened after Wu's experiment in 1957 \cite{PhysRev.105.1413}.
@@
 
-where $c$ e $s$ stand respectively for cosine and sine of the corresponding  mixing angles ($\theta_{12}$, $\theta_{23}$ and $\theta_{13}$), $\delta$ is the Dirac CP violation phase, $\alpha_1$ and $\alpha_{2}$  are the eventual Majorana CP violation phases.
+where $c$ and $s$ stand respectively for cosine and sine of the corresponding  mixing angles ($\theta_{12}$, $\theta_{23}$ and $\theta_{13}$), $\delta$ is the Dirac CP violation phase, $\alpha_1$ and $\alpha_{2}$  are the eventual Majorana CP violation phases.
 Experimental results on neutrino oscillations are generally reported in terms of the mixing angles and of the squared mass splitting $\Delta m^2_{ab} = m^2_{a} - m^2_{b}$, where $a$ and $b$ represent the mass eigenstates.
-A summary of the current status of experimental results, albeit partial, is given in table \ref{tab:nuosc}.
+A summary of the current status of experimental results is given in table \ref{tab:nuosc}.
 
@@
 
-Historically, neutrino interactions with the nucleus in the GeV region are divided into three categories into three categories whose contributions change as a function of increasing neutrino energy:: quasi elastic (QE), resonant (RES), and deep inelastic (DIS) scattering.
-All current and forthcoming oscillation experiments on neutrino beams live in the 0.1-10 GeV transition region, which encompasses the energy where the QE neutrino-nucleus interaction transitions into RES and then into DIS.
+Historically, neutrino interactions with the nucleus in the GeV region are divided into three categories  whose contributions change as a function of increasing neutrino energy:: quasi elastic (QE), resonant (RES), and deep inelastic (DIS) scattering.
+All current and near future oscillation experiments on neutrino beams live in the 0.1-10 GeV transition region, which encompasses the energy where the QE neutrino-nucleus interaction transitions into RES and then into DIS.
 For scattering off free nucleons, neutrino and antineutrino QE charge current scattering refers to the process $\nu_l n \rightarrow l^- p$ and $\bar\nu_l p \rightarrow l^+ n$ where a charged lepton and single nucleon are ejected in the elastic interaction.
@@
 
-A renewed interest for neutrino cross section measurements surged in recent years, along with a lively discussion on the data reporting; the historical method of reporting the neutrino cross section as a function of the neutrino energy or momentum transferred shakes under the weight of its dependency on the chosen nuclear model.
+There has been a renewed interest in neutrino cross section measurements in recent years, along with a lively discussion on the data reporting; the historical method of reporting the neutrino cross section as a function of the neutrino energy or momentum transferred shakes under the weight of its dependency on the chosen nuclear model.
 On one hand, correcting for nuclear effects in neutrino interaction can introduce unwanted sources of uncertainty and model dependency especially due to the mis-modeling of the meson interactions.
@@
 In SUSY, every fundamental particle in the SM has a ``superpartner", identical in each quantum number except for the spin-statistics: the fermion supersymmetric partners are bosons and vice versa.
-Collider experiments (mainly LHC)  constrain the mass of the supersymmetric partners to be very heavy\cite{LHC}.
+Collider experiments (mainly LHC)  constrain the mass of the supersymmetric partners to be very heavy.
 The SU(5), SU(10) groups with a SUSY twist are the basic groups for SUSY GUTs.
@@
 \begin{itemize}
-\item[-] the assessment of the existence of right-handed sterile neutrinos via the study of accelerator neutrinos on a short baseline (SBN), 
+\item[-] the assessment of the existence of sterile neutrinos via the study of accelerator neutrinos on a short baseline (SBN), 
 \item[-] the determination of the sign of $\Delta m^2_{13}$ (or $\Delta m^2_{23}$), i.e., the neutrino mass hierarchy via the study of accelerator neutrinos on a long baseline (DUNE),
@@
 When possible, external hadron-nucleus scattering data are used to tune INTRANUKE.
-Since no data is available for Argon, GENIE uses an interpolation of data from heavier and lighter nuclei for the pion-argon cross section leading to large (10?s of \%) resultant uncertainties in the INTRANUKE module.
+Since no data is available for Argon, GENIE uses an interpolation of data from heavier and lighter nuclei for the pion-argon cross section leading to resultant uncertainties in the INTRANUKE module of the order of 10-20\%.
 
@@
 
-LArIAT tiny active volume makes it impossible for the experiment to place competitive limits on nucleon decay searches.
+LArIAT's tiny active volume makes it impossible for the experiment to place competitive limits on nucleon decay searches.
 However,  LArIAT provides excellent data to characterize kaons in liquid argon for the ``LAr golden mode", $p \rightarrow K^+ \bar{\nu}$.

### 2018-05-31 Chapter-2/2-LArTPCs.tex
s

@@
 Historically, the active medium in ionization detectors has been in the gaseous form.
-Carlo Rubbia first proposed the use of a Liquid Argon TPC for a neutrino experiment, ICARUS \cite{Rubbia:1977zz}, in 1977.
+Willis and Radeka first proposed the use of liquid-argon ionization chambers as total-absorption detectors in their pioneer work of 1974 \cite{WILLIS1974221}
+Carlo Rubbia imported this concept to neutrino detection with the ICARUS experiment \cite{Rubbia:1977zz}, in 1977.
 Using nobles elements in the liquid form for neutrino detectors is advantageous for several reasons.
@@
 LArTPCs are some times referred as to ``electronic" bubble-chambers, for the similarity in the tracking and energy resolution which is coupled with an electronic readout of the imaging information in LArTPCs.
-Compared to these historic detectors however, LArTPC bestow tridimensional tracking and a self triggering mechanism provided by the scintillation light in the liquid argon.
+Compared to these historic detectors however, LArTPC bestow tridimensional tracking, calorimetry and a self triggering mechanism provided by the scintillation light in the liquid argon.
 An event display of a $\nu_\mu$ CC interaction candidate in the MicroBooNE detector is shown in picture \ref{fig:NuEvd} to display the level of spatial details these detectors are capable of; the color scale of the image is proportional to the energy deposited, hinting to these calorimetry capabilities of the detectors.
@@
 Argon ionization and scintillation are the processes leveraged to detect particles in the LArTPC active volume.
-When a ionizing radiation traverses the argon active volume it leaves a trail of ionization electrons along its trajectory and it excites the argon producing scintillation light -- details on the production and detection of ionization charge and scintillation light are provided in \ref{sec:light} and \ref{sec:light} respectively.
+When a ionizing radiation traverses the argon active volume it leaves a trail of ionization electrons along its trajectory and it excites the argon producing scintillation light -- details on the production and detection of ionization charge and scintillation light are provided in \ref{sec:light}.
 The optical detector sees the argon scintillation light in matters of nanoseconds.
@@
 Thus, argon  is constantly  filtered in the cryogenic system, which reduce the oxygen and water contamination to less than 100 parts per trillion.
-The filtration system depends on the size and drift distance of the experiment and, for experiments on several meters scale, it includes an argon recirculation system.
+The filtration system depends on the size and drift distance of the experiment and, for experiments on several meters scale, it includes an argon recirculation system \cite{MicroBooNE-det}.
 
@@
 The scattering length has been measured to be 66 cm \cite{Ishida1997} , shorter than the theoretical prediction of $\sim$~90 cm \cite{Teague1968}; this value is short enough to be relevant for the current size of LArTPCs detectors.
-In fact, Rayleigh scattering worsen the resolution on $t_0$, the start time for charge drifting, and  alters the light directionality, complicating the matching between light and charge coming from the same object in case of multiple charged particles in the detector.
+In fact, Rayleigh scattering worsens the resolution on $t_0$, the start time for charge drifting, and  alters the light directionality, complicating the matching between light and charge coming from the same object in case of multiple charged particles in the detector.
 
@@
 The next steps in the event reconstruction chain will then decide if rejecting hits with poor fits.
-It is important to notice how the height and width of the hit depend on the topology of the event: for example, a particle running  parallel to the wire planes will leave a series of sharp hits on many consecutive wires, while a particle traveling towards the planes will leave a long, wide hit on very few wires.
+It is important to notice how the height and width of the hit depend on the topology of the event: for example, a particle running  parallel to the wire planes will leave a series of narrow individual hits, one on each consecutive wire, while a particle traveling towards the planes will leave long, wide hits on very few wires.
 The height of the hits and their integral is proportional to the charge collected on the wire, so it depends on the particle type.\\
@@
 \textbf{3D Tracking.} The 3D tracking set of algorithms uses clusters close in time on the induction and collection planes as starting point to form a 3D track.
-Firstly, it construct a tentative 3D trajectory using the edges of the clusters.
-Then, it  projected back the tentative trajectory on to the planes and adjusts the parameters of the 3D track fit such that they minimize the distance between the fit projections and the track hits in all wire planes simultaneously.
-Tridimensional tracking can use multiple clusters in one plane, but it can never break them in smaller groups of hits.
+Firstly, it constructs a tentative 3D trajectory using the edges of the clusters.
+Then, it  projects back the tentative trajectory on to the planes and adjusts the parameters of the 3D track fit such that they minimize the distance between the fit projections and the track hits in all wire planes simultaneously.
+Tridimensional tracking can use multiple clusters in one plane, but it can never break them into smaller groups of hits.
 This algorithm was first developed for the ICARUS collaboration\cite{Antonello2013}.
@@
 Lastly an overall calibration of the energy, explained in detail in section \ref{ch:energyCalibration}, is applied and the calorimetric information for the given track is assigned.
-Even if calorimetry is done in 2D, it benefits from the 3D tracking information; typical information available after the calorimetric reconstruction are the total energy deposited by the particle and its stopping power $dE/dx$ at each ``track pitch", i.e.
+Even if calorimetry is done in 2D, it benefits from the 3D tracking information; typical objects available after the calorimetric reconstruction are the total energy deposited by the particle and its stopping power $dE/dx$ at each ``track pitch", i.e.
 at each 2D projection on the wire plane of the 3D trajectory.
@@
 In particular, we show the prospects for the exploration of neutrino physics (Section \ref{ch:NuPhysLAr}) and GUT models (Section \ref{ch:GutsLAr}) in current and forthcoming LAr experiments.
-In Section \label{ch:LArIATIntro}, we introduce LArIAT and its role in the Intensity Frontier panorama.
+In Section \ref{ch:LArIATIntro}, we introduce LArIAT and its role in the Intensity Frontier panorama.
 
 \subsection{Prospects for LArTPCs in Neutrino Physics: SBN and DUNE}\label{ch:NuPhysLAr}
-The ArgoNeut  experiment \cite{ArgoNeuT-det} together the LAr R\&D experiments TallBo and the Yale TPC  initiated the US LArTPC neutrino program.
+The ArgoNeut  experiment \cite{ArgoNeuT-det} together with the LAr R\&D experiments TallBo and the Yale TPC  initiated the US LArTPC neutrino program.
 Following the success of the ArgoNeut small TPC on the NuMI beam, a wide program of LArTPCs on neutrino beams has flourished.
@@
 
-The Short-Baseline Neutrino (SBN) \cite{Antonello:2015lea} program at Fermilab  is tasked with conclusively assess the nature of the ``LSND and MiniBooNE anomalies" \cite{Aguilar:2001ty, Athanassopoulos:1997pv, Aguilar-Arevalo:2013pmq}, resolving the mystery of  sterile neutrinos at the eV$^2$ scale.
-The SBN program entails three surface LArTPCs positioned on the Booster Neutrino Beam at different distances from the neutrino production in oder to fully exploit  the L/E dependence of the oscillation pattern:  SBND (100 m from the decay pipe), MicroBooNE (450 m), and ICARUS (600 m).
+The Short-Baseline Neutrino (SBN) \cite{Antonello:2015lea} program at Fermilab  is tasked with conclusively addressing the nature of the ``LSND and MiniBooNE anomalies" \cite{Aguilar:2001ty, Athanassopoulos:1997pv, Aguilar-Arevalo:2013pmq}, resolving the mystery of  sterile neutrinos at the eV$^2$ scale.
+The SBN program is comprised of three surface LArTPCs positioned on the Booster Neutrino Beam at different distances from the neutrino production in order to fully exploit  the L/E dependence of the oscillation pattern:  SBND (110 m from the decay pipe), MicroBooNE (470 m), and ICARUS (600 m).
 Within the oscillation context, the choice of the LArTPC technology for the SBN detectors changes the set of systematics with respect to LSND and MiniBooNE, whose detection techniques were both based on Cherenkov light.
-In particular, LArTPCs provide excellent electron/photon separation \cite{Acciarri:2016sli} lacking in Cherenkov detectors which can be leveraged to abate the photon background from neutral current interactions  in $\nu_e$ searches.
+In particular, LArTPCs provide excellent electron/photon separation \cite{Acciarri:2016sli} lacking in Cherenkov detectors which can be leveraged to reduce the photon background from neutral current interactions  in $\nu_e$ searches.
 MicroBooNE\cite{MicroBooNE-det}, the first detector of the SBN program to be fully operational, started its first neutrino run in October 2015.
-MicroBooNE is a 89 ton active volume LArTPC, single drift chamber with TPC dimensions of 2.6 m (drift) x 2.3 m (heigh) x 10.4 m (depth).
-MicroBooNE is positioned at a very similar L/E on the Booster neutrino beam as MiniBooNE has the scope to directly cross check the MiniBooNE oscillation measurement.
-In case MicroBooNE confirms the presence of the ``low energy excess" anomaly, SBND and ICARUS will provide the full measurement of the oscillation parameters.
+MicroBooNE is a 85 ton active volume LArTPC, single drift chamber with TPC dimensions of 2.6 m (drift) x 2.3 m (heigh) x 10.4 m (depth).
+MicroBooNE is positioned at a the same energy and very similar baseline on the Booster neutrino beam as MiniBooNE; MicroBooNE has the scope to directly cross check the MiniBooNE oscillation measurement.
+In case MicroBooNE confirms the presence of the ``low energy excess" (LEE) anomaly, SBND and ICARUS will confirm the LEE is due to oscillations (length dependent).
+In case MicroBooNE does not confirm the LEE, the SBN program will fully address the entire allowed parameter space.
 SBND and ICARUS are both dual drift chambers, whose active volume is respectively 112 ton and 600 ton.
@@
 Shooting neutrinos from Fermilab for 800 miles to the SURF laboratory in South Dakota, DUNE is tasked with preforming conclusive measurements of CP violation in the lepton sector,  the neutrino mass ordering and the $\theta_{23}$ octant.
-The DUNE far detector will count four 10 kton LArTPCs, roughly of dimensions of  19 m (horizontally) x 18 m (vertically) x 66 m (depth).
+The DUNE far detector is comprised of four 10 kton fiducial volume LArTPCs, roughly of dimensions of  19 m (horizontally) x 18 m (vertically) x 66 m (depth).
 

### 2018-05-31 Chapter-3/3-TheLArIATExperiment.tex
s

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-The instrumentation of  LArIAT tertiary beam and the TPC components have changed several times during the three years of LArIAT data taking.
+The instrumentation of the LArIAT tertiary beam and the TPC components have changed several times during the three years of LArIAT data taking.
 The following paragraphs describe the components operational during ``Run II", the data taking period relevant to the hadron cross section measurements considered in this thesis.
@@
 For the Run II data taking period, the current settings explored were 60A (B $\sim$0.21 T) and 100A (B $\sim$0.35 T) in both polarities.
-Albeit advantageous to enrich the tertiary beam composition with high mass particles such as kaons, we never pushed the magnets current over 100 A, not to incur in overheating.
+Albeit advantageous to enrich the tertiary beam composition with high mass particles such as kaons, we never pushed the magnets current over 100 A, so as not to incur in overheating.
 During operation, we operated an air and water cooling system on the magnets and we remotely monitored the magnet temperatures.
@@
 Each instrumented slot is equipped with four scintillation bars each, positioned horizontally in the direction orthogonal to the beam.
-Each scintillator bar measures  \textcolor{red}{?
-x ?
-x 2}~cm and it is read out by one PMT.
+Each scintillator is read out by one PMT.
 
@@
 It is worth mentioning here the presence of an additional scintillation paddle between WC4 and the downstream paddle of the TOF system, called halo.
-The halo is a 39x38x1 cm$^3$ paddle with a 6.5 cm radius hole in the center, whose original function was to reject  beam particles slightly offset from the beamline center.
+The halo is a 39 x 38 x 1 cm$^3$ paddle with a 6.5 cm radius hole in the center, whose original function was to reject  beam particles slightly offset from the beamline center.
 Data from this paddle turned out to be unusable, so our data events include both particle going through the halo scintillation material or through the halo hole.
@@
 As in any other LArTPC, argon purity is a crucial parameter for LArIAT.
-Indeed, the presence of contaminants affects both the basic working principles of a LArTPC, as shown in section \ref{sec:LArTPCWorkingPrinciple}: electronegative contaminants such as oxygen and water decrease the number of ionization electrons collected on the wires after drifting through the volume.
+Indeed, the presence of contaminants affects both the basic working principles of a LArTPC, as shown in Section \ref{sec:LArTPCWorkingPrinciple}: electronegative contaminants such as oxygen and water decrease the number of ionization electrons collected on the wires after drifting through the volume.
 In addition, contaminants such as Nitrogen decrease the light yield from scintillation light, especially in its slow component.
@@
 LArIAT  performs the amplification stage directly in cold with  amplifiers %developed by the Brookhaven National Lab (BNL) and
-mounted on the TPC frame inside the liquid argon. The BNL ASICs adopted in LAriAT are designated as LArASIC, version 4-star and are the same used by the MicroBooNE experiment \cite{Acciarri2017}.
+mounted on the TPC frame inside the liquid argon. The BNL ASICs adopted in LArIAT are designated as LArASIC, version 4-star and are the same used by the MicroBooNE experiment \cite{Acciarri2017}.
 %The ASIC signals for each wire are then driven out of the vessel  to DAQ boards that act as waveform recorders.
@@
 
-Fundamental inputs to the trigger card come from the TOF (see section~\ref{sec:TOF}) and the wire chambers (see section~\ref{sec:MWPC}), as activity in these systems points to the presence of a charged particle in tertiary beam line.
+Fundamental inputs to the trigger card come from the TOF (see Section~\ref{sec:TOF}) and the wire chambers (see Section~\ref{sec:MWPC}), as activity in these systems points to the presence of a charged particle in tertiary beam line.
 In particular, the discriminated pulses from the TOF PMTs form a NIM logic pulse for the trigger logic.
@@
 LArIAT is a complex ensemble of systems which needed to be monitored simultaneously during data taking.
-We performed the monitoring of the systems operations with a slow control system, a DAQ monitoring system and a low level data quality monitoring described in the following sections.
+We performed the monitoring of the systems operations with a slow control system, a DAQ monitoring system and a low level data quality monitoring system described in the following sections.
 
@@
 Its simple GUI allowed us to change the operating parameters and to graph the trends of several variables of interest for all of the tertiary beam detectors.
-Among the most important quantities monitored by Synoptic there are the level of argon in both the inner vessel and the external dewar, the operating voltages of cathode and wire planes, of the PMTs and SiPMs, and of the four wire chambers, as well as the magnet temperatures.
+Among the most important quantities monitored by Synoptic are the level of argon in both the inner vessel and the external dewar, the operating voltages of cathode and wire planes, of the PMTs and SiPMs, and of the four wire chambers, as well as the magnet temperatures.
 Figure \ref{fig:synoptics} shows an example of the monitoring system.

### 2018-05-31 Chapter-4/4-Methodology.tex
s

@@
 The secondary beam impinging on LArIAT secondary target produces a plethora of particles which propagates downstream.
-The presence of upstream and downstream collimators greatly abates the number of particles tracing down the LArIAT tertiary beamline.
+The presence of upstream and downstream collimators greatly abates the number of particles traveling down the LArIAT tertiary beamline.
 However, it is possible that more than one particle sneaks into the LArTPC during its readout time: the TPC readout is triggered by the particle  firing the series of beamline detectors along our tertiary beamline, but particles from the beam halo might also be present in the TPC at the same time.
@@
 In order to remove the shower topology, we create a region of interest (ROI) around the TPC track corresponding to the beamline particle.
-We look for short tracks contained in the ROI, as depicted in figure \ref{fig:showerFilt}:  if more then 5 tracks shorter than 10 cm are in the ROI, we reject the event.
+We look for short tracks contained in the ROI, as depicted in Figure \ref{fig:showerFilt}:  if more then 5 tracks shorter than 10 cm are in the ROI, we reject the event.
 Line 7 in  Table \ref{tab:beamlineDataSelection} shows the number of events surviving this selection; that table also shows that this selection is applied after the wire chamber to TPC match (discussed in the next section).
@@
 For each event passing the selection on its beamline information, we need to identify the track inside the TPC corresponding to the particle which triggered the beamline detectors, a procedure we refer to as ``WC to TPC match" (WC2TPC for short).
-In general, the TPC tracking algorithm can reconstruct more than one track in the event, partially due to the fact that hadrons interact in the chamber and partially because of pile up particles during the triggered TPC readout time, as shown in figure~\ref{fig:kaonInteraction}.
+In general, the TPC tracking algorithm can reconstruct more than one track in the event, partially due to the fact that hadrons interact in the chamber and partially because of pile up particles during the triggered TPC readout time, as shown in Figure~\ref{fig:kaonInteraction}.
 
@@
 If  $\Delta \text{R} < 4 $~cm, $\alpha < 8^\circ $,  a match between WC-track and TPC track is found.
-We describe  how we determine the value for the radius and angular selection in Section \ref{ch:WC2TPCMatchOptimization}.
+We describe  how we determine the value for the radius and angular selection in Appendix \ref{ch:AppendixTrack}.
 We discard events with multiple WC2TPC matches.
@@
 
-As we will show in section \ref{ch:PionXSBkgSub}, the background subtraction for the interacting and incident histograms can be translated into corresponding relative pion content factors $C^{\pi MC}_{\text{Int}} (E_{i})$ and $C^{\pi MC}_{\text{Inc}} (E_{i})$ and the cross section re-written as follows
+As we will show in Section \ref{ch:PionXSBkgSub}, the background subtraction for the interacting and incident histograms can be translated into corresponding relative pion content factors $C^{\pi MC}_{\text{Int}} (E_{i})$ and $C^{\pi MC}_{\text{Inc}} (E_{i})$ and the cross section re-written as follows
 

### 2018-05-31 Chapter-5/5-Preparatory.tex
s

@@
 This chapter describes the  work done on the  data and Monte Carlo samples in preparation for the cross section analyses. 
-This entails the choice of the datasets and the production of the information needed to construct the Monte Carlo Simulation~(Section  \ref{sec:dataSet}),  the construction and use of said Monte Carlo simulation~(section \ref{sec:MCSet}), the study of backgrounds for the pion cross section~(Section  \ref{ch:PionXSBkgSub}), the study of the energy loss between WC4 and TPC (Section \ref{ch:eloss}), the study of the tracking in the TPC~(Section  \ref{sec:TrackingStudies}), and study of  the calorimetry response ~(Section  \ref{ch:energyCal}).
+This entails the choice of the datasets and the production of the information needed to construct the Monte Carlo Simulation~(Section  \ref{sec:dataSet}),  the construction and use of said Monte Carlo simulation~(Section \ref{sec:MCSet}), the study of backgrounds for the pion cross section~(Section  \ref{ch:PionXSBkgSub}), the study of the energy loss between WC4 and TPC (Section \ref{ch:eloss}), the study of the tracking in the TPC~(Section  \ref{sec:TrackingStudies}), and study of  the calorimetry response ~(Section  \ref{ch:energyCal}).
 
@@
 
-The estimated beam composition is used as a basis to estimate the background contamination in the  ($\pi^{-}$,Ar) cross section measurement, whose  full treatment is described in section \ref{ch:PionXSBkgSub}.
+The estimated beam composition is used as a basis to estimate the background contamination in the  ($\pi^{-}$,Ar) cross section measurement, whose  full treatment is described in Section \ref{ch:PionXSBkgSub}.
 
@@
 Figure \ref{fig:DDMCQuantities}  schematically shows the data quantities of interest leveraged from data: the momentum ($P_x, P_y, P_z$) and position ($X, Y$) at WC4.
-For each data event, we obtain the  particle position ($X, Y$) at WC4 directly from the data measurement; we calculate the components of the momentum using the beamline measurement of the momentum magnitude in conjunction with the hits on WC3 and WC4 to determine the direction of the momentum vector, as described in section \ref{sec:MWPCfunc}.
+For each data event, we obtain the  particle position ($X, Y$) at WC4 directly from the data measurement; we calculate the components of the momentum using the beamline measurement of the momentum magnitude in conjunction with the hits on WC3 and WC4 to determine the direction of the momentum vector, as described in Section \ref{sec:MWPCfunc}.
 The momentum and position of the selected data events form a 5-dimensional series of tuples.
@@
 
-Using the DDMC technique ensures that the MC and data particles have very similar momentum, position and angular distributions at WC4 and allows us to use the MC sample in several occasions: to estimate the background contamination to the pion cross section (see Section \ref{ch:PionXSBkgSub}), to calibrate the energy loss upstream of the TPC (see Section \ref{ch:eloss}),  and to study the tracking and the calorimetric performance  in the LArTPC (sections \ref{sec:TrackingStudies} and \ref{ch:energyCal}).
+Using the DDMC technique ensures that the MC and data particles have very similar momentum, position and angular distributions at WC4 and allows us to use the MC sample in several occasions: to estimate the background contamination to the pion cross section (see Section \ref{ch:PionXSBkgSub}), to calibrate the energy loss upstream of the TPC (see Section \ref{ch:eloss}),  and to study the tracking and the calorimetric performance  in the LArTPC (Sections \ref{sec:TrackingStudies} and \ref{ch:energyCal}).
 A small caveat is in order here: the DDMC is a single particle Monte Carlo, which means that the beam pile-up is not simulated.
@@
 In order to choose the selection value for the wire chamber momentum, it is beneficial to estimate the fraction of MC capture and decay events that survive the momentum selection as a function of the momentum threshold, and compare it with the survival fraction for all the 60A events.
-This is done in figure \ref{fig:survRatio}.
+This is done in Figure \ref{fig:survRatio}.
 We define the survival ratio simply  as the number of events surviving the true momentum selection divided by the number of events of that category.
@@
 The kinematic at WC4 determines the trajectory of a particle and indicates whether or not it will hit the halo paddle, whose positioning and central hole size were not sufficiently well defined when the beamline instrumentation was installed.
-In figure \ref{fig:PxVsXTrue} , we plot the true  horizontal component of the momentum $P_x$ versus the true $X$ position at WC4 for pions missing the halo paddle (left) and for pions hitting the halo paddle (right) for the -60A MC simulation runs -- analogous plots are obtained with the -100A pion simulation and with the kaon simulation.
+In Figure \ref{fig:PxVsXTrue} , we plot the true  horizontal component of the momentum $P_x$ versus the true $X$ position at WC4 for pions missing the halo paddle (left) and for pions hitting the halo paddle (right) for the -60A MC simulation runs -- analogous plots are obtained with the -100A pion simulation and with the kaon simulation.
 These distributions can be separated drawing a line in this position-momentum space.
@@
 In MC, we assign  $E_{loss} = 32 \pm 4 $~MeV for pion events classified as ``hitting the halo paddle"; we assign  $E_{loss} = 24 \pm 3 $~MeV for pion events classified as ``missing the halo paddle".
-These values are the average and width of the two landau distributions underneath the double peaked distribution.
+These values are the average and width of the two Landau distributions underneath the double peaked distribution.
 We apply the same classifier on data.
@@
 
-The summary of the values for used for $E_{Loss}$ for the pion sample is listed in table \ref{tab:Eloss}  with the analogous results for the study on the kaon case.
+The summary of the values for used for $E_{Loss}$ for the pion sample is listed in Table \ref{tab:Eloss}  with the analogous results for the study on the kaon case.
 
@@
 \subsection{Angular Resolution}\label{sec:angleRes}
-Scope of this study is to understand and compare the tracking performances and the angular resolution of the TPC tracking on data and MC. 
+The scope of this study is to understand and compare the tracking performances and the angular resolution of the TPC tracking on data and MC. 
 We use the angular resolution of the tracking to determine  the value of smallest angle that the tracking algorithms allow to reconstruct with a non-zero efficiency, effectively determining a selection on the distribution of the scattering angle of hadronic interaction entering the cross section measurement.
@@
 In this section, we define the measurement on the kinetic energy and  determine the related uncertainty.
-We will propagate this uncertainty into the cross section measurement, as discussed in Section \ref{ch:SysUncertaintyXSRaw} for the pion cross section and in Section \ref{ch:KaonXSRawUnc} for the kaon cross section.
+We will propagate this uncertainty into the cross section measurement, as discussed in Section \ref{ch:SysUncertaintyXSRaw} for the pion cross section and in Section \ref{ch:KaonXSRaw} for the kaon cross section.
 
@@
 The analogous plot for the -100A negative pion data set  is show on the right side of Figure \ref{fig:EnergyDeposited}.
-The distributions are fitted with a landau displayed in red for data and in teal for MC.
+The distributions are fitted with a Landau displayed in red for data and in teal for MC.
 The uncertainty on $E_{i}$ is given by the width of the Landau fit to the data.
-A small systematic uncertainty  is given by a 1.0\% difference between the most probable value of the landau fits in data and MC.
+A small systematic uncertainty  is given by a 1.0\% difference between the most probable value of the Landau fits in data and MC.
 
@@
 The data is shown in black, the MC in blue.
-The distributions are fitted with a landau displayed in red for data and in teal for MC.
+The distributions are fitted with a Landau displayed in red for data and in teal for MC.
 } \label{fig:EnergyDeposited}
@@
 The data is shown in black, the MC in blue.
-The distributions are fitted with a landau displayed in red for data and in teal for MC.
+The distributions are fitted with a Landau displayed in red for data and in teal for MC.
 } \label{fig:EnergyDepositedK}

### 2018-05-31 Chapter-6/6-PionCrossSection.tex
s

@@
 
-As stated in section \ref{ch:XSRaw},  the raw cross section is given by the equation \ref{eq:thinTargetXSSolved}
+As stated in Section \ref{ch:XSRaw},  the raw cross section is given by the Equation \ref{eq:thinTargetXSSolved}
 \begin{equation}
@@
 The data central points are represented by black dots, the statistical uncertainty is shown in black, while the systematic uncertainty is shown in red.
-Data is displayed over the $N^{\text{TOT}}_{\text{Int}}$  distribution obtained with a MC sample of pions, muon and electrons weighted by the beam composition (additional details on the composition will be provided in Section \ref{ch:BKGsubXS}).
+Data is displayed over the $N^{\text{TOT}}_{\text{Int}}$  distribution obtained with a MC sample of pions, muon and electrons weighted by the beam composition (additional details on the composition will be provided in Section \ref{ch:BKGsubXSPuppa}).
 The contribution from the simulated pions is shown in blue, the one from secondaries in red, the one from muons in yellow and the ones from electrons in gray.
@@
 
-Since the interaction probability $P_{Interacting}$ is estimated by $\frac{ N^{\text{TOT}}_{\text{Int}}}{N^{\text{TOT}}_{\text{Inc}}}$ and the number of tries $\mathcal{N}$ is $N^{\text{TOT}}_{\text{Inc}}$, equation \ref{eq:binVar} translates into
+Since the interaction probability $P_{Interacting}$ is estimated by $\frac{ N^{\text{TOT}}_{\text{Int}}}{N^{\text{TOT}}_{\text{Inc}}}$ and the number of tries $\mathcal{N}$ is $N^{\text{TOT}}_{\text{Inc}}$, Equation \ref{eq:binVar} translates into
 \begin{equation}
@@
 \section{Corrections to the Raw Cross Section}\label{ch:PionXSCorrections}
-As described in section \ref{ch:MCCorrections}, we need to apply a background correction and an efficiency correction in order to derive the pion cross section from the raw cross section.
-The  cross section is given in equation \ref{eq:C},
+As described in Section \ref{ch:MCCorrections}, we need to apply a background correction and an efficiency correction in order to derive the pion cross section from the raw cross section.
+The  cross section is given in Equation \ref{eq:C},
 
@@
 The right side of the same figure shows the  MC estimated  relative pion content for the incident histogram as function of kinetic energy for the 60A runs (top) and 100A runs (bottom).
-In Figure \ref{fig:CorrectionsBeam} the central curves displayed in light blue are obtained using the beamline composition as predicted by G4Beamline: these are the correction curves for the relative pion content applied to data in equation \ref{eq:C}.
+In Figure \ref{fig:CorrectionsBeam} the central curves displayed in light blue are obtained using the beamline composition as predicted by G4Beamline: these are the correction curves for the relative pion content applied to data in Equation \ref{eq:C}.
 
@@
 Given this definition, it is evident that we rely on the tracking algorithm to discern where the interaction occurred in the TPC  and correctly end the tracking.
-The tracking algorithm has an intrinsic angle resolution as shown in section \ref{sec:angleRes}, which limits its efficiency, especially in the case of elastic scattering occurring a low angles.
+The tracking algorithm has an intrinsic angle resolution as shown in Section \ref{sec:angleRes}, which limits its efficiency, especially in the case of elastic scattering occurring a low angles.
 \textcolor{red}{Plus, there are instance where INSERT HERE THE STUFF ABOUT THE MIGRATION}
 Thus, we need to apply a correction accounting for mis-reconstruction of the interaction point in order to retrieve the true cross section.
-This correction is evaluated separately for the interacting and incident histograms bin by bin, namely $\epsilon^{\text{int}}(E_i)$ and  $\epsilon^{\text{inc}}(E_i)$, and applied in the cross section formula as shown in  equation \ref{eq:C}.
+This correction is evaluated separately for the interacting and incident histograms bin by bin, namely $\epsilon^{\text{int}}(E_i)$ and  $\epsilon^{\text{inc}}(E_i)$, and applied in the cross section formula as shown in  Equation \ref{eq:C}.
 
@@
 
-In section \ref{sec:angleRes}, we estimated the angular resolution for data and MC to be $\bar\alpha_{Data} = (5.0 \pm 4.5) \text{ deg}$  and 
+In Section \ref{sec:angleRes}, we estimated the angular resolution for data and MC to be $\bar\alpha_{Data} = (5.0 \pm 4.5) \text{ deg}$  and 
 $\bar\alpha_{MC} = (4.5 \pm 3.9) \text{ deg}$, respectively.
@@
 Figure \ref{fig:EffCorr60A} shows $\epsilon^{\text{Int}}(E_{i})$ in the left side and $ \epsilon^{\text{Inc}}(E_i)$ on the right as a function of the kinetic energy for the 60A runs and their systematic uncertainty.
-Similarly, figure \ref{fig:EffCorr100A} shows $\epsilon^{\text{Int}}(E_{i})$ in the left side and $ \epsilon^{\text{Inc}}(E_i)$ on the right as a function of the kinetic energy for the 100A runs and their systematic uncertainty.
+Similarly, Figure \ref{fig:EffCorr100A} shows $\epsilon^{\text{Int}}(E_{i})$ in the left side and $ \epsilon^{\text{Inc}}(E_i)$ on the right as a function of the kinetic energy for the 100A runs and their systematic uncertainty.
 

### 2018-05-31 Chapter-7/7-KaonCrossSection.tex
s

@@
 
-Similar to the pion case,  the raw cross section is given by the equation \ref{eq:thinTargetXSSolved}
+Similar to the pion case,  the raw cross section is given by the Equation \ref{eq:thinTargetXSSolved}
 \begin{equation}
@@
 \section{Corrections to the Raw Cross Section}\label{ch:KaonXSCorrections}
-As described in section \ref{ch:MCCorrections}, we need to apply a background correction and a correction for reconstruction effects in order to derive the  kaon cross section from the raw cross section.
-The  cross section is given in equation \ref{eq:C},
+As described in Section \ref{ch:MCCorrections}, we need to apply a background correction and a correction for reconstruction effects in order to derive the  kaon cross section from the raw cross section.
+The  cross section is given in Equation \ref{eq:C},
 
@@
 The correction is then applied to the relevant bin in data.
-The correction for recontruction effects is evaluated separately for the interacting and incident histograms, namely $\epsilon^{\text{int}}(E_i)$ and  $\epsilon^{\text{inc}}(E_i)$, and propagated to the cross section as shown in  equation \ref{eq:C}.
+The correction for recontruction effects is evaluated separately for the interacting and incident histograms, namely $\epsilon^{\text{int}}(E_i)$ and  $\epsilon^{\text{inc}}(E_i)$, and propagated to the cross section as shown in  Equation \ref{eq:C}.
 
-In section \ref{sec:angleRes}, we estimated the angular resolution for data and MC to be $\bar\alpha_{Data} = (4.3 \pm 3.7) \text{ deg}$  and  $\bar\alpha_{MC} = (4.4 \pm 3.6) \text{ deg}$, respectively.
+In Section \ref{sec:angleRes}, we estimated the angular resolution for data and MC to be $\bar\alpha_{Data} = (4.3 \pm 3.7) \text{ deg}$  and  $\bar\alpha_{MC} = (4.4 \pm 3.6) \text{ deg}$, respectively.
 Most interaction angles smaller than the angular resolution will thus be indistinguishable  for the reconstruction.

### 2018-05-31 Chapter-3/3-TheLArIATExperiment.tex
s

@@
 The TDC time resolution is 1.18~ns/bin and can accept 2 edges per 9~ns.
-The maximum event rate acceptable by the chamber system is 1 MHz: this rate is not a limiting factor considering that \textcolor{red}{the rate of the tertiary particle beam at the first wire chamber is estimated to be less than 15 kHz}.
+The maximum event rate acceptable by the chamber system is 1 MHz.%: this rate is not a limiting factor considering that the rate of the tertiary particle beam at the first wire chamber is estimated to be less than 15 kHz}.
 A full spill of data occurring once per supercycle is stored on the TDC board memory at once and read out by a specially designed controller.

### 2018-05-31 Chapter-6/6-PionCrossSection.tex
s

@@
 
-The bottom panel of figure  \ref{fig:FinalXSResPion} shows the relative difference between the measured cross section and the Geant4 predicted cross section.
+The bottom panel of figure  \ref{fig:FinalXSResPion} shows the relative deviation between the measured cross section and the Geant4 predicted cross section.
 A couple of considerations are in order here.
 On one hand, we note the general agreement between the measured and predicted cross sections.
-We expected the pion cross section in argon to be well modeled in Geant4, given the vast set of measurements available for pion interaction on different targets -- even if no data is available for argon.
+We expected the pion cross section in argon to be well modeled in Geant4, given the vast set of measurements available for pion interaction on different targets -- even if no previous data is available on argon.
 Thus, this agreement confirms our confidence in the experimental methodology used.

### 2018-05-31 Chapter-7/7-KaonCrossSection.tex
s

@@
 The measured cross section is mostly in disagreement with the predicted cross section; in particular it is significantly higher -- on the order of 20\%-30\% higher -- in the bins lower than 400 MeV.
-The 100 MeV to 400 MeV range is the region of interest for the expected kaons coming from a proton decay event.
+Such difference is unlikely to be accounted for the small proton contamination expected in the data sample.
+It is worth noticing that the 100 MeV to 400 MeV energy range is the region of interest for the expected kaons coming from a proton decay event.
+With this result, we will inform a tuning of the Geant4 model for kaon cross section.
 
@@
 The Geant4 prediction for the total hadronic cross section for angle scattering greater than 5$^\circ$ is displayed in green.
-\emph{Bottom:} Relative difference between the measured cross section and the Geant4 prediction.}
+\emph{Bottom:} Relative deviation between the measured cross section and the Geant4 prediction.}
 \label{fig:FinalXSKaon}

### 2018-05-31 Chapter-8/8-Conclusions.tex
s

@@
 With this work, we presented the first ever ($\pi^-$-Ar) and ($K^+$-Ar) total hadronic cross section measurements as a function of the hadron kinetic energy.
-Both the analysis follow a similar workflow and  they rely on beam line detector information as well as both calorimetry and tracking in the LArTPC.
+Both the analyses follow a similar workflow and  they rely on beam line detector information as well as both calorimetry and tracking in the LArTPC.
 
 For the ($\pi^-$-Ar) total hadronic cross section measurement, we start by identifying pion beamline candidates through a series of selections on the beamline and TPC information which maximize the number of pions over other particle species (muons and electrons).
-We use the LArIAT beamline MC to estimate the beam composition of the selected beamline candidates and we propagate them to the LArAIT TPC constructing a properly weighted sample with the DDMC.
+We use the LArIAT beamline MC to estimate the beam composition of the selected beamline candidates and we propagate them to the LArTPC constructing a properly weighted sample with the DDMC.
 We apply the thin slice method on the pion candidates and obtain the raw cross section.
@@
 we apply the thin slice method on kaon candidates identified in the beamline and obtain the raw cross section.
-We then apply a MC derived background correction and a correction for detector effects to obtain the cross section.
+We then apply a MC derived background correction and a correction for detector effects to measure the cross section.
 The background correction accounts for the presence of secondary particles in both the interacting and incident histograms and for the presence of decay events in the interacting plot.
@@
 A hint of a difference in shape is present: Geant4 seems to underestimate the cross section in the delta resonance region and overestimate it for energies greater than 400 MeV.
-Since the   ($\pi^-$-Ar) has never been measured before on Argon and since previous measurements on different nuclei show that the shape of the delta resonance peak varies as a function of the target nucleus, this hint to a shape disagreement is an interesting feature of this measurement which LArIAT will explore with a future development of the analysis.
+Since the   ($\pi^-$-Ar) has never been measured before on argon and since previous measurements on different nuclei show that the shape of the delta resonance peak varies as a function of the target nucleus, this hint is an interesting feature of the measurement which LArIAT will explore with a future development of the analysis.
 In the kaon case, the measured cross section is mostly higher than the Geant4 prediction.
-Contrary to the pion case, previous data on kaons are scarce even on different nuclei; thus, we do not expect the Geant4 prediction to be as solid in the kaon case as it is in the pion case.
+Contrary to the pion case, previous measurements of kaon interaction on any nuclei are scarce; thus, we do not expect the Geant4 prediction to be as solid in the kaon case as it is in the pion case -- as illustrated  by Figure \ref{fig:TrueCarbon} for kaon cross section on carbon.
 The nice agreement between the measured cross section and the prediction of the pion case gives us confidence in the methodology, thus confidence in the kaon measurement.
@@
 
-The ($\pi^-$-Ar) and ($K^+$-Ar) total hadronic cross section analyses are the first physics analyses developed by the LArIAT experiment and  will serve as a basis for the future cross section measurements of pions and kaons for the exclusive channels.
+The ($\pi^-$-Ar) and ($K^+$-Ar) total hadronic cross section analyses are the first physics analyses developed by the LArIAT experiment and  will serve as a basis for the future cross section measurements of pion and kaon cross sections in the exclusive channels.
 
@@
 \includegraphics[width=0.48\textwidth]{Chapter-7/Images/TheMoneyPlotK.pdf}
-\caption{\emph{Left:} ($\pi^-$-Ar) total hadronic cross section measurements in the combined sample overlaid with the  Geant4 prediction (green).
+\caption{\emph{Left:} ($\pi^-$-Ar) total hadronic cross section measurement for  scattering angles greater than 5$^\circ$ measured in the combined sample overlaid with the  Geant4 prediction (green).
 \emph{Right:} ($K^+$-Ar) total hadronic cross section for  scattering angles greater than 5$^\circ$ measured.

### 2018-05-31 thesis_main.tex
s

@@
 
-\noindent ``\emph{Dunque io ringrazio tutti quanti.} \hfill ``\emph{At last, I thank everyone.}
+\noindent ``\emph{Dunque io ringrazio tutti quanti.} \hfill ``\emph{So, I thank everyone.}
 
@@
 Little did I know that getting my PhD would be first and foremost a journey to self discovery, the beginning of a path to asymptotically tend to the person and the scientist I want to be.
-A journey impossible without the help of many people, which I will now proceed to thank; let's hope that the acknowledgements won't be longer than the thesis.
+A journey impossible without the help of many people, which I will now proceed to thank.
+Let's hope that the acknowledgements won't be longer than the thesis itself.
 
-Mom, you are and will always be the hero of my story; I know this has been challenging for you, thanks for being there for me at every hour of the day and the night, for being always honest with me and for somehow handling this whole US-Italy thing gracefully.
+Mom, you are and will always be the hero of my story; I know this has been challenging for you, thanks for being there for me at every hour of the day and the night, for being always honest and for somehow handling this whole US-Italy thing gracefully.
 I can't believe you still haven't met Bonnie.
@@
 Thanks for listening to me before I proved I was worth be listened to; thanks for always believing I could make it, even at the bottom of my doubts; thanks for all the laughters, for all the freedom, for all the great advices; thanks for leading by example; thanks for being always supportive of all my crazy side projects; thanks for choosing each and every person in the most wonderful research group ever and thanks for letting me be part of it.
-Thanks for showing me that even Bonnie Freaking Fleming is human, has her downs and still she strives.
+Thanks for showing me that even Bonnie %Freaking
+Fleming is human, has her downs and still she thrives.
 
@@
 A special token of gratitude goes to Roberto Acciarri and Andrzej M.
-Szelc, you are so much more to me than collaborators and previous car owners, you know you'll always have a home in Chicago.
+Szelc, you are so much more to me than collaborators and previous car owners, you know you'll always have a home in Chicago (or wherever this job will bring me).
 
 Thanks to the MicroBooNE experiment which has been an incredible learning experience.
-In particular, I want to thank  Sam Zeller for the countless pieces of advice,  my CRT troopers Martin, David, Rui, Igor and Linda  (if I know something about hardware it's because of you), and Kazu and  Kirby (if I know something about software it's because of you).
+I want to thank  Sam Zeller for the countless pieces of advice,  my CRT troopers Martin, David, Rui, Igor and Linda  -- if I know something about hardware it's because of you -- and Kazu and  Kirby, who I shamelessly bothered many times for software help, sushi and coffee.
 
@@
 
-Shany Danieli and Louise Laage, you're the piece of heart I left in New Haven.
+Shany Danieli, Louise Laage and Flor Orosz Hunziker, you're the piece of heart I left in New Haven.
 Shany, I don't know what made you think I was the cool kid in the group, but I'm so glad you stuck with me.
-I'm not sure I would have survived my first semester at Yale without you: you shine in your smartness and lift other people with you -- it's a rare quality in our field.
+I'm not sure I would have survived my first semester at Yale without you.
+Your intelligence is matched only by your kindness: you have the rare gift to shine and make others shine with you.
 Louise, my dear, thanks for always finding the time to listen, for your attuned spirit and for the magnolias.
+Flor, thanks for welcoming me to the new world with love and giving me the right definition of ``America".
 
@@
 Finally, thanks to Anto, my husband, my topo and my adventure companion, cause people might think I'm nice and easy, but you know I'm not...
-And still, you followed me across half the world and you're still here, so that I can never feel alone.
+And still, you followed me across half the world and you're here, so that I can never feel alone.
 

### 2018-05-31 Chapter-4/4-Methodology.tex
s

@@
 We look for short tracks contained in the ROI, as depicted in Figure \ref{fig:showerFilt}:  if more then 5 tracks shorter than 10 cm are in the ROI, we reject the event.
-Line 7 in  Table \ref{tab:beamlineDataSelection} shows the number of events surviving this selection; that table also shows that this selection is applied after the wire chamber to TPC match (discussed in the next section).
-The WC2TPC match already lower the presence of electrons in the sample, which is further reduced by the shower filter.
+Line 7 in  Table \ref{tab:beamlineDataSelection} shows the number of events surviving this selection; that table also shows that this selection is applied after the beamline event is matched to TPC particle (discussed in the next section).
+This match already lowers the presence of electrons in the sample, which is further reduced by the shower filter.
 
@@
 \begin{equation}
- \sigma^{\pi^-}_{TOT}(E_{i})  = \frac{1}{n \delta X}\frac{N^{\pi^-}_{ \text{Int}} (E_{i})}{N^{\pi^-}_{ \text{Inc}}(E_{i})}.
+ \sigma^{\pi^-}_{TOT}(E_{i})  = \frac{1}{n\text{ } \delta X}\frac{N^{\pi^-}_{ \text{Int}} (E_{i})}{N^{\pi^-}_{ \text{Inc}}(E_{i})}.
 \label{eq:thinTargetXSSolved2}
@@
 In fact, the selection used to isolate pions in the LArIAT beam allows for the presence of some muons and electrons as background, while the kaon selection allows for a small contamination of protons (see Section \ref{ch:beamlineComposition}).
-Also, the LArIAT TPC tracking algorithm is not 100\% efficient in determining the interaction point.
-Therefore we need to apply two corrections evaluated on the MC in order to extract the final cross section from LArIAT data: i) a background subtraction and ii) a correction for reconstruction effects.
+Also, the LArTPC tracking algorithm is not 100\% efficient in determining the interaction point.
+This inefficiency occurs in two fashions: i) the tracking algorithm does not stop at the interaction point and continues adding hits from a particle past it (this happens especially in the case of shallow elastic scattering), ii) the tracking stops prematurely.
+These two cases have different consequences on the population of the interacting and incident distributions.
+In the first case, the interacting histogram will be underpopulated and the incident histogram might be overpopulated.
+In case of premature end of tracking, the interacting histogram will be overpopulated at energies greater than the eventual interaction, while the incident histogram will be underpopulated.
+Given the importance of tracking for the cross section measurements,  we report an optimization to maximize the identification of the interaction point in Appendix \ref{ch:AppendixTrack}.
+
+Therefore, we apply two corrections evaluated on MC in order to extract the final cross section from LArIAT data: i) a background subtraction and ii) a correction for reconstruction effects. 
 Still using the pion case as example, we estimate the pion cross section in each energy bin changing  Equation \ref{eq:thinTargetXSSolved2} into

### 2018-05-31 Chapter-5/5-Preparatory.tex
s

@@
 Given the technical nature of these studies, we report them in Appendix \ref{ch:AppendixTrack}.
-We only report here the evaluation of  the angular resolution of the tracking algorithm in data and MC, due to its important implications on the physics measurement.
+We only report here the evaluation of  the angular resolution of the tracking algorithm and a study of the track pitch in data and MC, due to their important implications on the physics measurement.
 
@@
 \label{fig:trueWithAngles}
+\end{figure}
+
+\subsection{Track Pitch}\label{sec:trackPitch}
+The granularity of the LArTPC detector leads to the fine-sampling of a hadron inside the argon active volume.
+This detector granularity translates into the ``track pitch" quantity, which is the distance between two consecutive 3D points of the tracking.
+The track pitch determines the width of the slab $\delta X$ in the cross section Equation \ref{eq:thinTargetXSSolved2}.
+Figure \ref{fig:Pitch} shows the pitch distribution in data (black) and MC (red) for the pion combined sample on the left and the kaon sample on the right.
+The MC and data distributions agree rather well.
+Thus, we estimate the slab width to be $\delta X = 0.47 \pm 0.03 $ cm.
+
+\begin{figure}[p]
+\centering
+\includegraphics[width=0.48\textwidth]{Chapter-5/Images/PitchPi.pdf}
+\includegraphics[width=0.48\textwidth]{Chapter-5/Images/PitchK.pdf}
+\caption{\emph{Left:} Pitch distribution for the pion combined sample 60A and 100A, data displayed in black, MC in red. 
+\emph{Right:} Pitch distribution for the kaon sample, data displayed in black, MC in red.}
+\label{fig:Pitch}
 \end{figure}

### 2018-05-31 Chapter-6/6-PionCrossSection.tex
s

@@
 \begin{equation}
-\delta\sigma_{TOT}(E) = \sigma_{TOT}(E) \Big(\frac{\delta N^{\text{TOT}}_{\text{Int}}}{N^{\text{TOT}}_{\text{Int}}}+\frac{\delta N^{\text{TOT}}_{\text{Inc}}}{N^{\text{TOT}}_{\text{Inc}}}\Big) 
+\frac{\delta\sigma_{TOT}(E)}{\sigma_{TOT}(E)} =  \Big(\frac{\delta N^{\text{TOT}}_{\text{Int}}}{N^{\text{TOT}}_{\text{Int}}}+\frac{\delta N^{\text{TOT}}_{\text{Inc}}}{N^{\text{TOT}}_{\text{Inc}}}\Big)
 \end{equation}
@@
 \subsection{Treatment of Systematics} \label{ch:SysUncertaintyXSRaw}
-The only systematic effect considered in the measurement of the raw cross section results from the propagation of the uncertainty associate with the measurement of the kinetic energy at each argon slab.
+The systematic effects considered in the measurement of the raw cross section results from the uncertainty on the track pitch  ($\delta X = 0.47 \pm 0.03 $ cm, shown in \ref{sec:trackPitch})  and from the propagation of the uncertainty associate with the measured kinetic energy at each argon slab.
 As shown in Section \ref{ch:kinEn}, the uncertainty on the kinetic energy of a pion candidate at the j$^{th}$ slab of argon  is given by
@@
 The tracking algorithm has an intrinsic angle resolution as shown in Section \ref{sec:angleRes}, which limits its efficiency, especially in the case of elastic scattering occurring a low angles.
-\textcolor{red}{Plus, there are instance where INSERT HERE THE STUFF ABOUT THE MIGRATION}
-Thus, we need to apply a correction accounting for mis-reconstruction of the interaction point in order to retrieve the true cross section.
+Also, as mentioned in Section \ref{ch:MCCorrections}, the tracking algorithm can prematurely stop, mis-identifying an interaction point.
+Thus, we need to apply a correction accounting for mis-reconstruction of the interaction point in order to retrieve the final cross section.
 This correction is evaluated separately for the interacting and incident histograms bin by bin, namely $\epsilon^{\text{int}}(E_i)$ and  $\epsilon^{\text{inc}}(E_i)$, and applied in the cross section formula as shown in  Equation \ref{eq:C}.
@@
 Geant4 simulates interactions at all angles, as shown in figure \ref{fig:trueScatteringAngle}.
-In order to calculate the correction for reconstruction effects,  we select events which have an interaction angle greater than a given $\alpha_{res}$ to construct the true interacting and incident histograms (the denominator of the correction).
+In order to calculate the correction for reconstruction effects,  we select events which have an interaction angle greater than a given $\alpha_{res}$ to construct the true MC interacting and incident histograms (the denominator of the correction).
 
@@
 
- The systematics on this correction is estimated by varying the value of $\alpha_{res}$ between 0 deg and 4.5 deg and propagating the uncertainty on the cross section.
+ The systematics on this correction is estimated by varying the value of $\alpha_{res}$ between 0 deg and 5.0 deg and propagating the uncertainty on the cross section.
 
@@
 \section{Results}\label{ch:FinalPion}
-Figure \ref{fig:FinalXSPion} shows the measurement of the ($\pi^-$-Ar) total hadronic cross section for  scattering angles greater than 5$^\circ$, as the result of the background subtraction and reconstruction effects correction to the raw cross section.
+Figure \ref{fig:FinalXSPion} shows the measurement of the ($\pi^-$-Ar) total hadronic cross section,
+
+\begin{equation}
+ \sigma^{\pi^-}_{TOT}(E_{i})  =\frac{1}{n\text{ } \delta X}\frac{N^{\pi^-}_{ \text{Int}} (E_{i})}{N^{\pi^-}_{ \text{Inc}}(E_{i})} = \frac{1}{n \text{ }\delta X}\frac{ \epsilon^{\text{Inc}}(E_i) [ N^{ \text{TOT}}_{ \text{Int}} (E_{i}) - B_{ \text{Int}} (E_i)] }{   \epsilon^{\text{Int}}(E_i) [N^{ \text{TOT}}_{ \text{Inc}}(E_{i}) - B_{ \text{Inc}} (E_i)]},
+\end{equation}
+
+for  scattering angles greater than 5$^\circ$, as the result of the background subtraction and reconstruction effects correction to the raw cross section.
 The top left plot is the measurement obtained on the 60A data, statistical and systematic uncertainty shown in red.
@@
 
-The systematic uncertainty on the cross section is the sum in quadrature of the statistical uncertainty, the systematic uncertainty related to the kinetic energy measurement, the systematic uncertainty related to the beam composition and the systematic uncertainty related to the reconstruction effects correction.
+The systematic uncertainty on the cross section is the sum in quadrature of the statistical uncertainty, the systematic uncertainty calculated for the raw cross section, the systematic uncertainty related to the beam composition and the systematic uncertainty related to the reconstruction effects correction.
 

### 2018-05-31 Chapter-7/7-KaonCrossSection.tex
s

@@
 where $N^{\text{TOT}}_{\text{Int}}$  is the measured number of particles interacting at kinetic energy $E_i$, $N^{\text{TOT}}_{\text{Inc}}$ is the  measured  number of particles incident  on an argon slice at  kinetic energy $E_i$,  $n$ is the density of the target centers  and $\delta X$ is the thickness of the argon slice.
-The density of the target centers and the slab thickness are $n = 0.021\cdot10^{24} \text{ cm}^{-3} $ and  $\delta X=0.47\text{ cm}$, respectively.
+The density of the target centers and the slab thickness are $n = 0.021\cdot10^{24} \text{ cm}^{-3} $ and  $\delta X=0.47 \pm 0.03\text{ cm}$, respectively.
 
@@
 
-As in the pion case, the only systematic effect considered in the measurement of the raw cross section results from the propagation of the uncertainty associate with the measurement of the kinetic energy at each argon slab.
+As in the pion case, the systematic effects considered in the measurement of the raw cross section results from the uncertainty on the pitch ($\delta X = 0.47 \pm 0.03$ cm) and from the propagation of the uncertainty associate with the measurement of the kinetic energy at each argon slab.
 For kaons, the uncertainty on the kinetic energy of a candidate at the j$^{th}$ slab of argon  is given by
@@
 \section{Results}\label{ch:FinalKaon}
-Figure \ref{fig:FinalXSKaon} show the measurement of the ($K^+$-Ar) total hadronic cross section for  scattering angles greater than 5$^\circ$, as the result of the background subtraction and reconstruction  effects  correction to the raw cross section.
+Figure \ref{fig:FinalXSKaon} show the measurement of the ($K^+$-Ar) total hadronic cross section,
+\begin{equation}
+ \sigma^{K^+}_{TOT}(E_{i})  =\frac{1}{n\text{ } \delta X}\frac{N^{K^+}_{ \text{Int}} (E_{i})}{N^{K^+}_{ \text{Inc}}(E_{i})} = \frac{1}{n \text{ }\delta X}\frac{ \epsilon^{\text{Inc}}(E_i) [ N^{ \text{TOT}}_{ \text{Int}} (E_{i}) - B_{ \text{Int}} (E_i)] }{   \epsilon^{\text{Int}}(E_i) [N^{ \text{TOT}}_{ \text{Inc}}(E_{i}) - B_{ \text{Inc}} (E_i)]},
+\end{equation}
+
+for  scattering angles greater than 5$^\circ$, as the result of the background subtraction and reconstruction  effects  correction to the raw cross section.
 The plot shows the measurement obtained on the full dataset, statistical and systematic uncertainty in black.
@@
 
-The systematic uncertainty on the cross section is the sum in quadrature of the statistical uncertainty, the systematic uncertainty related to the kinetic energy measurement and the systematic uncertainty related to the reconstruction  effects  correction.
+The systematic uncertainty on the cross section is the sum in quadrature of the statistical uncertainty, the systematic uncertainty calculated for the raw cross section and the systematic uncertainty related to the reconstruction  effects  correction.
 

### 2018-05-31 AppendixC-EnergyCalibration/C-EnergyCalibration.tex
s

@@
 
-\textcolor{red}{Add agreement between data and MC for dedx for pions}
+After the calibration is done separately on data and MC, we can compare the resulting dE/dX distributions; this is done for the set of pion beamline candidates and pion MC on the left side of Figure \ref{fig:dedx}.
+On the right side of the same figure, we report the data-MC comparison for the kaon sample used in the cross section.
+The distributions are fitted with simple Landau functions.
+As expected, the Landau MPV for data  is consistent with the MC MPV for both pions and kaons.
+For both the pion and the kaon case, the width of the Landau is a bit wider in data than in MC; this difference might be due to an underestimate of the electronic noise in the MC simulation.
 
@@
 \centering
-\includegraphics[width=0.50\textwidth]{Chapter-5/Images/RunIINegTotaldEdXvsMomentum.png}
-\includegraphics[width=0.50\textwidth]{Chapter-5/Images/RunIIPosTotaldEdXvsMomentum.png}
+\includegraphics[width=0.48\textwidth]{Chapter-5/Images/RunIINegTotaldEdXvsMomentum.png}
+\includegraphics[width=0.48\textwidth]{Chapter-5/Images/RunIIPosTotaldEdXvsMomentum.png}
 \caption{Stopping power versus Momentum for Run-II negative (top) and positive (bottom) polarity data.
@@
 \label{fig:BBandData}
+\end{figure}
+
+\begin{figure}[htb]
+\centering
+\includegraphics[width=0.48\textwidth]{AppendixC-EnergyCalibration/dEdXPions.pdf}
+\includegraphics[width=0.48\textwidth]{AppendixC-EnergyCalibration/dEdXKaons.pdf}
+\caption{\emph{Left:} dE/dx distribution for $\pi^-/\mu^-/e^-$ data (black) and Pion MC (blue).
+The Landau fit for data is shown in red, the one for MC in teal.\emph{Right:} dE/dx distribution for $K^+$ data (black) and Kaon MC (blue).
+The Landau fit for data is shown in red, the one for MC in teal.
+All the distributions are area normalized.}
+\label{fig:dedx}
 \end{figure}

### 2018-05-31 Chapter-3/3-TheLArIATExperiment.tex
s

@@
 The other two drift regions are those between the shield plane and the induction plane (S-I), and between the induction plane and the collection plane (I-C).
-The electric field in these regions is chosen to satisfy the charge transparency condition and allow for 100$\%$ transmission of the drifting electrons through the shield and the induction planes.
-
+The electric field in these regions is chosen to satisfy the charge transparency condition  and allow for 100$\%$ transmission of the drifting electrons through the shield and the induction planes; for example, the transparency condition between for the shield plane reads
+\begin{equation}
+\frac{E_{S-I}}{E_{C-S}} > \frac{1+\rho}{1-\rho},
+\end{equation}
+where $E_{C-S}$ is the electric field in the region between the cathode and the shield plane, $E_{S-I}$ is the electric field in the region between the shield and the induction plane, and $\rho = \frac{2\pi \text{ }r}{D}$ is a factor accounting for the radius of the wires $r$ (152 $\mu m$) and the gap distance $D$ (4 mm).
+An analogous condition can be written for the induction plane.
 \begin{figure}[htb]

### 2018-06-01 Chapter-1/1-Theory.tex
s

@@
 A summary of the pion final states in order of pion multiplicity for the reaction channel is given in table \ref{tab:pionFinal}.
+Figure \ref{fig:PionsEvd} shows examples of the topologies due to the pion-argon hadronic interaction as they appear in the  LArIAT data: elastic and inelastic scattering, pion absorption with emission of protons, charge exchange and production of pions.
+
+\begin{figure}[hbpt]
+\centering
+\includegraphics[width=\textwidth]{Chapter-1/Images/PionsEvd.pdf}
+\caption{Pion-Argon interaction topologies due to elastic scattering, inelastic scattering, pion absorption with emission of protons, charge exchange and production of pions as seen in LArIAT data (induction plane only).}
+\label{fig:PionsEvd}
+\end{figure}
+
 Pion capture and pion decay at rest dominate the cross section under 100 MeV.
@@
 
+Figure \ref{fig:KaonsEvd} shows examples of the topologies due to the kaon-argon hadronic interaction as they appear in the  LArIAT data: elastic and inelastic scattering.
+
+\begin{figure}[hbpt]
+\centering
+\includegraphics[width=0.7\textwidth]{Chapter-1/Images/KaonsEvd.pdf}
+\caption{Kaon-Argon interaction topologies due to elastic scattering and inelastic scattering as seen in LArIAT data (collection plane only).}
+\label{fig:KaonsEvd}
+\end{figure}
+
 For the LArIAT cross section analysis, the kaons considered span a momentum inside the TPC from 100 MeV/c to 800 MeV/c.

### 2018-06-01 Chapter-7/7-KaonCrossSection.tex
s

@@
 
-The measured cross section is mostly in disagreement with the predicted cross section; in particular it is significantly higher -- on the order of 20\%-30\% higher -- in the bins lower than 400 MeV.
+The measured cross section is mostly in tension with the predicted cross section; in particular it is significantly higher -- on the order of 20\%-30\% higher -- in the bins lower than 400 MeV.
 Such difference is unlikely to be accounted for the small proton contamination expected in the data sample.

### 2018-06-01 thesis_main.tex
s

@@
 This thesis presents two world's first measurements: the measurement of ($\pi^-$-Ar) total hadronic cross section in the 100-1050 MeV kinetic energy range and the measurement of the ($K^+$-Ar) total hadronic cross section in the 100-650 MeV kinetic energy range.
-The analyses devised for these measurements use both the core elements of LArIAT: beamline and TPC.
+The analyses devised for these measurements use both the core elements of LArIAT: beamline and LArTPC.
 The first step in each analysis is the development of an event selection based on  beamline and TPC information geared towards the identification of the hadron of interest.
 We then proceed to match the beamline candidate to a suitable TPC track.
-Finally,  we apply the ``thin slice method" technique and measure the cross section, correcting for background and detector effects.
+Finally,  we apply the ``thin slice method" technique and measure the cross section, correcting for background contamination and detector effects.
 The thin slice technique is a new method to measure hadron-argon cross sections possible only due to the combination of the tracking and calorimetry capability of the LArTPC technology.
-Albeit never on argon, the hadronic cross section of pions has been measured before on several different elements in thin target experiments, leading to solid predictions for measurements on argon.
-Through the use of a different technique, our measurement of the  ($\pi^-$-Ar) total hadronic cross section is in general agreement with the prediction by the Geant4 Bertini Cascade model which are based on data from thin target experiments.
-On the contrary, cross section measurements for kaons are extremely rare, thus more difficult to model.
-Our measurement of the  ($K^+$-Ar) total hadronic cross section is mostly in disagreement with the Geant4 prediction and provides new experimental data to properly tune the Geant4 models.
+Albeit never on argon, the hadronic cross section of pions has been extensively measured before on lighter and heavier different elements in thin target experiments, leading to solid predictions for measurements on argon.
+Through the use of a different technique, our measurement of the  ($\pi^-$-Ar) total hadronic cross section is in general agreement with the predictions by the Geant4 Bertini Cascade model which are based on data from thin target experiments.
+On the contrary, cross section measurements for kaons are extremely scarce, thus more difficult to model.
+Our measurement of the  ($K^+$-Ar) total hadronic cross section is mostly in tension with the Geant4 prediction over the explored range of energies and provides new experimental data to properly compare existing interaction models.
 
@@
 The ($\pi^-$-Ar) and the ($K^+$-Ar) total hadronic cross measurements are the first physics results of the LArIAT experiment and will be the basis for the future LArIAT measurements of pion and kaon cross sections in the exclusive channels.
+The outcome of these measurements will ultimately enable to quantify and reduce the systematic associated with the hadronic interaction models in neutrino-argon interactions.
 

### 2018-06-01 Chapter-0/0-Introduction.tex
s

@@
 Particle and nuclear physics have a long history of hadronic cross section measurements; what makes these measurements unique is both the target (argon) and the methodology used -- the ``thin slice method" -- which takes advantage of the detection capabilities of the LArTPC technology.
-The combination of fine-grained tracking and excellent calorimetric information provided by the LArTPC technology  enables to the measurement of unprecedented details of particle interactions in argon and, in LArIAT, to measure the kinetic energy of a hadron at each step along the particle traces.
+The combination of fine-grained tracking and excellent calorimetric information provided by the LArTPC technology  enables the measurement of unprecedented details of particle interactions in argon and, in LArIAT, to measure the kinetic energy of a hadron at each step along the particle traces.
 A renewed interest for precision measurements of hadronic cross sections, particularly in argon, arises from the current  panorama of experimental  particle physics at the intensity frontier, in particular neutrino physics in LArTPCs.
@@
 With ``fine details" here we mean the ability to distinguish the many products of the neutrino interaction, such as protons, pions, muons and electrons, and to measure their energy.
-Historically,  bubble chamber neutrino detectors were the first revolution in neutrino detection: for example, the spatial resolution of Gargamelle allowed the discovery of neutrino neutral current interaction.
+Historically,  bubble chamber neutrino detectors were the first revolution in neutrino detection: for example, the spatial resolution of Gargamelle allowed the discovery of neutrino neutral current interaction\cite{HASERT1973138}.
 Despite the high precision of bubble chambers images, this technology is hard to scale to massive size, making statistical analyses on neutrino interactions almost impossible to perform.
@@
 %No measurement of hadronic cross sections for pions or kaons is available for argon (yet).
-The amount of details in an LArTPC event can easily  parsed by human eye, but can make automatic event reconstruction rather challenging.
+The amount of details in a LArTPC event can easily be parsed by human eye, but can make automatic event reconstruction rather challenging.
 Thus, reconstruction algorithms in LArTPC need to be tuned to recognize the different topologies of the neutrino interaction products in argon.
-This is particularly true for pions, since they are an abundant product of the neutrino interactions: the occurrence of a pion interaction in argon can modify the topology of the neutrino event, causing a misidentification of the neutrino interaction channel.
+This is particularly true for pions, since they are copiously produced of the neutrino interactions: the occurrence of a pion interaction in argon can modify the topology of the neutrino event, causing a misidentification of the neutrino interaction channel.
 
@@
 The ($\pi^-$-Ar) hadronic cross section is a fundamental input for neutrino detectors in liquid argon, as pion interactions can modify the topology and energy reconstruction of neutrino events in the GeV range, where pion production is abundant.
-The  ($K^+$-Ar) total hadronic differential cross section in LArIAT is particularly relevant for a high identification efficiency in the context of proton decay searches in DUNE in the  $p\rightarrow K^+\bar{\nu}$  channel.
+The  ($K^+$-Ar) total hadronic cross section in LArIAT is particularly relevant for a high identification efficiency in the context of proton decay searches in DUNE in the  $p\rightarrow K^+\bar{\nu}$  channel.
 In fact, the kaon-argon cross section affects the kaon topology by modifying the kaon tracking and energy reconstruction, impacting the basis for kaon identification in a LArTPC.
@@
 These analyses are LArIAT's first physics results.
-In order to measure total hadronic  argon cross sections, several steps are necessary.
+In order to measure total hadronic cross sections on argon, several steps are necessary.
 The analyses start by identifying a sample of the hadron of interest in the beam line and assessing the beam line contaminations.
 It proceeds with tracking the hadron candidates in the TPC and measuring their kinetic energy at each point in the tracking: the fine sampling of an hadron in the TPC forms the set of ``incident" hadrons.
-Then, the hadronic interaction point is identified and the raw cross section is calculated.
+Then, the hadronic interaction point is identified and the raw cross section is calculated via the ``thin slice method".
 Two corrections are then applied to the raw cross section -- a background subtractions and a correction for detector effects -- to obtain the cross section measurement, presented here.\\
@@
 Chapter \ref{sec:experimentDescription} describes LArIAT experimental setup.
-We present the event selection for both the pion and kaon analyses, as well as the ``thin slice method" in Chapter \ref{ch:Interactions}.
+We present the event selection for both the pion and kaon analyses, as well as the thin slice method in Chapter \ref{ch:Interactions}.
 Chapter \ref{ch:samples}  describes the work done on the data and Monte Carlo samples in preparation of the cross section analyses.

### 2018-06-01 Chapter-1/1-Theory.tex
s

@@
 The gauge group univocally determines the number of gauge bosons that carry the interaction; the gauge bosons correspond to the generators of the group: eight gluons (g) for the strong interaction, one photon ($\gamma$) and three bosons (W$^\pm$, Z$^0$) for the electroweak interaction.
-A gauge theory by itself cannot provide a description of massive particles, but it is experimentally well know that most of the elementary particles have non-zero masses.
+A gauge theory by itself cannot provide a description of massive particles, but it is experimentally well known that most of the elementary particles have non-zero masses.
 The introduction of massive fields in the Standard Model lagrangian would make the theory not gauge invariant, resulting ill-defined.
@@
 \end{equation}
-is equivalent to the equaitons
+is equivalent to the equations
 \begin{equation}
@@
 In fact, since the spatial inversion operator throws $\psi_R \leftrightarrow \psi_L$, parity is conserved only if both chiral components exist at the same time.
-For the neutrino introduction in the SM, experiments were critical to a more accurate theoretical description.
+For the neutrino introduction in the SM, experiments were critical to the formulation of an accurate theoretical description.
 The constraint of parity conservation weakened after Wu's experiment in 1957 \cite{PhysRev.105.1413}.
@@
 
-The first two lines of the lagrangian summarize the kinetic terms for the fermionic fields and their coupling to the gauge bosons $A^{\mu\nu}_a$, $B^{\mu\nu}$ \footnote{In gauge theories the ordinary derivative $\partial_\mu$  is substitued with the covariant derivative $D_\mu$.
+The first two lines of the lagrangian summarize the kinetic terms for the fermionic fields and their coupling to the gauge bosons $A^{\mu\nu}_a$, $B^{\mu\nu}$ \footnote{In gauge theories the ordinary derivative $\partial_\mu$  is substituted with the covariant derivative $D_\mu$.
 Here $D_\mu = \partial_\mu + igA_\mu \cdot I + ig'B_\mu\frac{Y}{2}$, where I and Y are the SU(2)$_L$ and U(1)$_Y$ generators, respectively.}.
 The third line describes the kinetic terms and the self-coupling terms of the gauge bosons.
-The forth line is the Higgs lagrangian, which results in the spontaneous symmetry breaking.
+The fourth line is the Higgs lagrangian, which results in the spontaneous symmetry breaking.
 The last three lines describe the Yukawa coupling between fermions and the Higgs field, origin of the fermions' mass.
@@
 \subsection{Neutrino Oscillations}
-The determination of the flavor of a neutrino dynamically arises from the corresponding charged lepton associated in a change current interaction; for example, a $\nu_e$ is a neutrino which produces an $e^-$, a $\bar\nu_\mu$ is a neutrino which produces a $\mu^+$, $etc$.
+The determination of the flavor of a neutrino dynamically arises from the corresponding charged lepton associated in a charged current interaction; for example, a $\nu_e$ is a neutrino which produces an $e^-$, a $\bar\nu_\mu$ is a neutrino which produces a $\mu^+$, $etc$.
 The neutrino flavor eigenstates $\ket{\nu_\alpha}$,  with $\alpha = e,\mu,\tau$, are orthogonal to each other and form a base for the weak interaction matrix.
@@
 \hline
-$\theta_{23}$            & 45$^\circ$                      & 9.0\%   & Super Kamiokande, MINOS,\\
-|$\Delta m^2_{32}$|    & 2.5 $10^{-3}$ eV$^2$   & 1.8\% &  No$\nu$a, MACRO \\
+$\theta_{23}$            & $\sim$45$^\circ$                      & 9.0\%   & Super Kamiokande, MINOS,\\
+|$\Delta m^2_{32}$|    & 2.5 $10^{-3}$ eV$^2$   & 1.8\% &  NO$\nu$a, MACRO, Super Kamiokande\\
 \hline
@@
 \hline
-$\theta_{13}$            &   9$^\circ$                    & 4.7\%  & DAYA  Bay, \\
+$\theta_{13}$            &   9$^\circ$                    & 4.7\%  & Daya  Bay, \\
 |$\Delta m^2_{32}$|    & 2.5 $10^{-3}$ eV$^2$   & 1.8\% &  RENO\\
@@
 All neutrino experiments involving the detection of single neutrinos are concerned with neutrino interactions (and neutrino cross sections) on nuclei. 
-Given the invisible nature of the neutrino, characterizing the products of its interaction is the only method to a) assess the neutrino presence, b) detect its flavor in case of a charge current interaction and c) eventually reconstruct its energy.
+Given the invisible nature of the neutrino, characterizing the products of its interaction is the only method to a) assess the neutrino presence, b) detect its flavor in case of a charged current interaction and c) eventually reconstruct its energy.
 
-Historically, neutrino interactions with the nucleus in the GeV region are divided into three categories  whose contributions change as a function of increasing neutrino energy:: quasi elastic (QE), resonant (RES), and deep inelastic (DIS) scattering.
+Historically, neutrino interactions with the nucleus in the GeV region are divided into three categories  whose contributions change as a function of increasing neutrino energy: quasi-elastic (QE), resonant (RES), and deep inelastic (DIS) scattering.
 All current and near future oscillation experiments on neutrino beams live in the 0.1-10 GeV transition region, which encompasses the energy where the QE neutrino-nucleus interaction transitions into RES and then into DIS.
-For scattering off free nucleons, neutrino and antineutrino QE charge current scattering refers to the process $\nu_l n \rightarrow l^- p$ and $\bar\nu_l p \rightarrow l^+ n$ where a charged lepton and single nucleon are ejected in the elastic interaction.
+For scattering off free nucleons, neutrino and antineutrino QE charged current scattering refers to the process $\nu_l n \rightarrow l^- p$ and $\bar\nu_l p \rightarrow l^+ n$ where a charged lepton and single nucleon are ejected in the elastic interaction.
 %of a neutrino (or antineutrino) with a nucleon in the target material.
@@
 DIS refers to the head-on collision between the neutrino and a parton inside the nucleon, producing hadronization and subsequent  abundant production of mesons and nucleons.
-In addition to such interactions between the neutrino and a single component of the nucleus, neutrinos can also interact with the nucleus as a whole, albeit more rarely, a well documented process called coherent meson production scattering \cite{Vilain1993}; the signature of such process is the production of a distinctly forward-scattered single meson final state, most often a pion.
+In addition to such interactions between the neutrino and a single component of the nucleus, neutrinos can also interact with the nucleus as a whole, albeit more rarely, a well documented process called coherent meson production \cite{Vilain1993}; the signature of such process is the production of a distinctly forward-scattered single meson final state, most often a pion.
 This simple picture of neutrino interactions works rather well for scattering
@@
 
-Nuclear effects can potentially affect neutrino event rates, final state particle emission, neutrino energy reconstruction, and the neutrino/antineutrino ratios, carrying deep implications for oscillation experiments.
+Nuclear effects can potentially affect measured neutrino event rates, final state particle emission, neutrino energy reconstruction, and the neutrino/antineutrino ratios, carrying deep implications for oscillation experiments.
 Even in the case of ``simple" QE scattering, intra-nuclear hadron rescattering and correlation effects between the target nucleons can cause the ejection of additional nucleons in the final state, modifying the final state kinematics and topology.
 In the case of resonant and DIS scattering, the hadronic interactions of meson and nucleons produced in the decay of the resonance or during hadronization  complicate this picture even more.
-A large source of uncertainty in modeling  nuclear effects in neutrino interactions come from mesons interactions (and re-interactions) in the nucleus, e.g., pion re-scattering, charge exchange, and absorption.
+A large source of uncertainty in modeling  nuclear effects in neutrino interactions comes from meson interactions (and re-interactions) in the nucleus, e.g., pion re-scattering, charge exchange, and absorption.
 
 There has been a renewed interest in neutrino cross section measurements in recent years, along with a lively discussion on the data reporting; the historical method of reporting the neutrino cross section as a function of the neutrino energy or momentum transferred shakes under the weight of its dependency on the chosen nuclear model.
-On one hand, correcting for nuclear effects in neutrino interaction can introduce unwanted sources of uncertainty and model dependency especially due to the mis-modeling of the meson interactions.
+On one hand, correcting for nuclear effects in neutrino interactions can introduce unwanted sources of uncertainty and model dependency, especially due to the mis-modeling of the meson interactions.
 On the other, avoiding this correction makes a comparison between neutrino interactions on different target nuclei extremely difficult.
@@
 Data on neutrino scattering off many different nuclei are available for both charged current (CC) and neutral current (NC) channels, as summarized in \cite{RevModPhys.84.1307}.
-A summary of the results on QE, resonant and DIS scattering for neutrinos and antineutrinos from accelerators on different target is reported in Figure \ref{fig:nuXS}, where the (NUANCE) \cite{Casper2002} event generator is used as comparison with the theory.
+A summary of the results of QE, resonant and DIS scattering for neutrinos and antineutrinos from accelerators on different target is reported in Figure \ref{fig:nuXS}, where the (NUANCE) \cite{Casper2002} event generator prediction is shown in comparison with experimental data.
 \begin{figure}[hbpt]
@@
 \subsection{Open Questions in Neutrino Physics}\label{ch:questions}
-On one hand, the last three decades of experiments in neutrino oscillations brought spectacular advancements in the understanding of the oscillations pattern, measuring the neutrino mixing angles and mass splitting with a precision of less than 10\%.
+On one hand, the last three decades of experiments in neutrino oscillations brought spectacular advancements in the understanding of the oscillations pattern, measuring the neutrino mixing angles and mass splitting with a precision better than 10\%.
 On the other, they opened the field for a series of questions needing experimental answers.
@@
 The  LSND and MiniBooNE $\nu_e$ and $\bar\nu_e$ appearance results, known as the ``LSND and MiniBooNE anomalies" \cite{Aguilar:2001ty, Athanassopoulos:1997pv, Aguilar-Arevalo:2013pmq}, may be interpreted under the assumption of a new right-handed neutrino.
-The additional neutrino needs to be ``sterile", i.e needs not to couple with the electroweak force carriers, in order to meet the constraint imposed by  the measurement of the width of the Z boson~\cite{2006257}.
+The additional neutrino needs to be ``sterile" i.e, needs not to couple with the electroweak force carriers, in order to meet the constraint imposed by  the measurement of the width of the Z boson~\cite{2006257}.
 The new sterile neutrino would mainly be composed of a heavy neutrino $\nu_4$ with mass $m_4$ such that  $m_1, m_2, m_3 \ll m_4$ and  $\Delta m^2= \Delta m^2_{14} \sim [0.1 - 10]$ eV$^2$.
@@
 \textbf{CP Violation In Lepton Sector.} The measurement of non-zero value for the oscillation parameter $\theta_{13}$ allows the exploration of low-energy CP violation in the lepton sector at neutrino long baseline oscillation experiments, enabling the possibility to measure the Dirac CP-violating phase $\delta$.
-Exciting theoretical results tie $\delta$ directly to the generation of the baryon asymmetry of the Universe at the Grand Unified Theory scale \textcolor{red}{a couple of cit would be nice}.
+Exciting theoretical results tie $\delta$ directly to the generation of the baryon asymmetry of the Universe at the Grand Unified Theory scale.
 According to the theoretical model described in \cite{PASCOLI20071}, for example, leptogenesis can be achieved if $|\sin\theta_{13} \sin \delta| > 0.11$, i.e.
@@
 the distance between the neutrino production and detection points, and $\Delta_{ab}$ a factor proportional to the sign and magnitude of the mass splitting.
-From these equations, it is clear how the relative  large value of $\theta_{13}$ is a happy accident necessary not to completely suppress the sensitivity to CP violation.
-The equations also show how the sensitivity to $\delta$ is tied to the measurement of the least precisely measured mixing angle,  $\theta_{23}$ (via the $\sin2\theta_{23}$ term) and to an other unknown quantity, the neutrino ``mass hierarchy" (via the $\Delta_{ab}$ terms).
+From these equations, it is clear how the relatively  large value of $\theta_{13}$ is a happy accident necessary not to completely suppress the sensitivity to CP violation.
+The equations also show how the sensitivity to $\delta$ is tied to the measurement of the least precisely measured mixing angle,  $\theta_{23}$ (via the $\sin2\theta_{23}$ term) and to another unknown quantity, the neutrino ``mass hierarchy" (via the $\Delta_{ab}$ terms).
 The precise determination of $\theta_{23}$ is often referred as to ``the octant problem".
@@
 The smallness of neutrino masses can only be pinned on a tiny Yukawa coupling which is not justified by the theory.\\
-In 1937, Majorana demonstrated that the introduction of a two components spinor is sufficient to describe a massive fermion \cite{Majorana1937}.
-The Dirac equations of motion for the chiral fields (equations \ref{eq:15} and \ref{eq:16}) hold true in the case of two components spinor under the assumption that the chiral components $\psi_R$ and $\psi_L$ are correlated through the charge conjugation matrix $\mathcal{C}$, $\psi_R = \mathcal{C}\bar\psi_L$.
+In 1937, Majorana demonstrated that the introduction of a two-component spinor is sufficient to describe a massive fermion \cite{Majorana1937}.
+The Dirac equations of motion for the chiral fields (equations \ref{eq:15} and \ref{eq:16}) hold true in the case of a two-component spinor under the assumption that the chiral components $\psi_R$ and $\psi_L$ are correlated through the charge conjugation matrix $\mathcal{C}$, $\psi_R = \mathcal{C}\bar\psi_L$.
 Therefore the theory is applicable only to neutral fermions.
@@
  \end{equation}
- where $g$ is the coupling coefficient, $v$ the Higgs VEV and $\mathcal{M}$ a constant with the dimension of the mass proportional to the scale of new physics.
+ where $g$ is the coupling coefficient, $v$ the Higgs VEV, and $\mathcal{M}$ a constant with the dimension of the mass proportional to the scale of new physics.
 The  $\mathcal{L}_5$ term would introduce a non-renormalizable term in the lagrangian, since it has dimensions of energy to the fifth power.
-This is not allowed in the SM lagrangian; however, the existence of such terms is plausible if the we consider the SM as an effective theory at low energy, manifestation of the symmetry breaking of a more general theory at higher energy, e.g.
+This is not allowed in the SM lagrangian; however, the existence of such terms is plausible if we consider the SM as an effective theory at low energy, manifestation of the symmetry breaking of a more general theory at higher energy, e.g.
 a Grand Unified Theory (GUT), and not the definitive theory.
-The mass term in eq \ref{eq:majoMass} implies the neutrino mass to be  $m = \frac{g v^2}{\mathcal{M}}$.
-The coupling coefficient can be of the order of any other fermion's coupling coefficient, since the smallness of neutrino masses is achieved by the big value of the new physics mass scale alone.
+The mass term in Equation \ref{eq:majoMass} implies the neutrino mass to be  $m = \frac{g v^2}{\mathcal{M}}$.
+The coupling coefficient can be of the order of any other fermion's coupling coefficient, since the smallness of neutrino masses is achieved by the large value of the new physics mass scale alone.
 This vanilla formulation is the conceptual basis for many flavors of \emph{seesaw mechanism} ~\cite{Yanagida1980}, which we will not discuss here in any detail.
@@
 Thus, it is impossible to discern the neutrino nature through kinematic effects such as neutrino oscillations.
-Neutrinoless double beta decay searches are the most promising way to understand the nature of the neutrino and are therefore subject of great theoretical and experimental interest.
+Neutrinoless double beta decay ($0\nu\beta\beta$) searches are the most promising way to understand the nature of the neutrino and are therefore a subject of great theoretical and experimental interest.
 Observation of the lepton number violating process  $0\nu\beta\beta$  would imply neutrinos have a Majorana component.
@@
 
-The SM does not include a suitable dark matter candidate and a mechanisms that accounts for the baryon asymmetry of the universe.
+The SM does not include a suitable dark matter candidate and a mechanism that accounts for the baryon asymmetry of the universe.
 Additionally, up to a total of 25 parameters remain seemingly arbitrary and need to be fitted to data: 3 gauge couplings, 9 charged fermion masses, 3 mixing angles and one CP phase in the CKM matrix, the Higgs mass and quartic coupling, $\theta_{QCD}$, 3 neutrino mixing angles, 1 Dirac phase and, eventually,  2 Majorana phases.
 
-From a group theory perspective, the SM has a rather complex group structure, where a gauge group is formed with the direct product of other three groups as shown in eq.
-\ref{eq:SMGroup}.
-\textcolor{red}{Drawing a parallel with the electroweak symmetry breaking mechanism, where the $ SU(2)_L \otimes U(1)_Y$ is recovered from $U(1)_{EM}$, an interesting line of simplification for the SM group structure would be to devise a similar mechanism where $SU(3)_C  \otimes SU(2)_L \otimes U(1)_Y$ is recovered from an hypothetical larger group.
-IS THIS CORRECT?} Just as the electroweak unification becomes evident at energies higher than the Higgs VEV,  a direct manifestation of Grand Unification Theories (GUTs) would occur at even higher energies.
+From a group theory perspective, the SM has a rather complex group structure, where a gauge group is formed with the direct product of other three groups as shown in Equation \ref{eq:SMGroup}.
+Drawing a parallel with the electroweak symmetry breaking mechanism, where the $ SU(2)_L \otimes U(1)_Y$ is recovered from $U(1)_{EM}$, an interesting line of simplification for the SM group structure would be to devise a similar mechanism where $SU(3)_C  \otimes SU(2)_L \otimes U(1)_Y$ is recovered from an hypothetical larger group.
+Just as the electroweak unification becomes evident at energies higher than the Higgs VEV,  a direct manifestation of Grand Unification Theories (GUTs) would occur at even higher energies.
 
-As the smallness of neutrino masses suggests the existence of a higher mass scale, an other, even stronger, hint to Grand Unification comes from the slope of running of the coupling constants.
-The coupling constants for the electromagnetic, weak and strong interactions in the SM vary as a function of the interaction energy as shown in figure \ref{fig:constants};  they do not exactly meet under the current experimental constraints, but their trend is interesting enough to push for the construction of theories where perfect unification is achieved through the addition of new particles.
+As the smallness of neutrino masses suggests the existence of a higher mass scale, another, even stronger, hint for Grand Unification comes from the slope of running of the coupling constants.
+The coupling constants for the electromagnetic, weak and strong interactions in the SM vary as a function of the interaction energy as shown in Figure \ref{fig:constants};  they do not exactly meet under the current experimental constraints, but their trend is interesting enough to push for the construction of theories where perfect unification is achieved through the addition of new particles.
 
@@
 The new super heavy bosons are colored and form a weak doublet.
-Their are the mediator of the interaction that turns quarks into leptons, leading to predict the existence of processes that violate baryon number, such as $p \rightarrow \pi^0  + e^+$  (see fig~ \ref{fig:MandatoryFeynmannDiagrams}, right).
-The prediction for proton decay lifetime, $\tau_p \sim \frac{M^4_X}{m^5_p}\sim10^{30\pm1.5}$~years, is unfortunately experimentally disproved by IMB and Super-Kamiokande \cite{PhysRevD.42.2974,PhysRevD.95.012004}.
+They are the mediator of the interaction that turns quarks into leptons, leading to predict the existence of processes that violate baryon number, such as $p \rightarrow \pi^0  + e^+$  (see Figure~ \ref{fig:MandatoryFeynmannDiagrams}, right).
+The prediction for proton decay lifetime, $\tau_p \sim \frac{M^4_X}{m^5_p}\sim10^{30\pm1.5}$~years, is unfortunately experimentally disproven by IMB and Super-Kamiokande \cite{PhysRevD.42.2974,PhysRevD.95.012004}.
 
@@
 $u$ and $d$) need to decay into second and third generation particles (such as $s$).
-Predictions on the proton decay lifetime depend on the chosen SUSY model; again, some of the predictions are not excluded by the experiments \cite{PhysRevD.55.6986, PhysRevD.67.075007,Lopez:1991eca}.
+Predictions of the proton decay lifetime depend on the chosen SUSY model; again, some of the predictions are not excluded by the experiments \cite{PhysRevD.55.6986, PhysRevD.67.075007,Lopez:1991eca}.
 
@@
 \item[-] the determination of the sign of $\Delta m^2_{13}$ (or $\Delta m^2_{23}$), i.e., the neutrino mass hierarchy via the study of accelerator neutrinos on a long baseline (DUNE),
-\item[-] the determination of the octant, i.e.  whether $\theta_{23}$ is maximal, via the study of accelerator neutrinos on a long baseline (DUNE),
-\item[-] the determination the status of CP symmetry in the lepton sector, via the study of accelerator neutrinos on a long baseline (DUNE),
+\item[-] the determination of the octant, i.e.,  whether $\theta_{23}$ is maximal, via the study of accelerator neutrinos on a long baseline (DUNE),
+\item[-] the determination of the status of CP symmetry in the lepton sector, via the study of accelerator neutrinos on a long baseline (DUNE),
 \item[-] the search for observables predicted by GUTs, such as proton decay via the study of non accelerator physics in massive underground detectors (DUNE).
@@
 % and of light mesons interaction with nuclei studies.
- We describe the signal signature and historic measurements of pion-nucleus cross section, as well as the implementation of these cross sections in the current version of the simulation package used by LArIAT.
+ We describe the signal signature and historic measurements of pion-nucleus cross sections, as well as the implementation of these cross sections in the current version of the simulation package used by LArIAT.
 
@@
 As outlined in \ref{ch:NuInt}, neutrino experiments use the products of neutrino interactions to identify the energy and flavor of the incoming neutrino.
-Pions are a common product of neutrino interaction, especially in resonant scattering, DIS and coherent pion production.
+Pions are a common product of neutrino interactions, especially in resonant scattering, DIS and coherent pion production.
 For neutrino experiments in argon, there are two main reasons why understanding pion hadronic interactions with argon is important: to model the behavior of the pion inside the target nucleus  and to model the behavior of the pion during its propagation inside the detector medium.
@@
 \end{equation}
-the $\Delta$ and $N^*$ and excited states will decay hadronically in matters of $\sim10^{-24}$~s inside the nucleus producing pions which which will have many chances to re-interact as they exit the target medium.
+the $\Delta$ and $N^*$ and excited states will decay hadronically in matters of $\sim10^{-24}$~s inside the nucleus, producing pions which which will have many chances to re-interact as they exit the target medium.
 The decay modes for the lower mass $\Delta$ (1232) and $N^*$(1440)  are listed in table \ref{tab:decaysdelta}.
-\begin{table}[b]
+\begin{table}[h]
 \centering
@@
 This module places the outgoing particles in the nucleus and propagates them using the ``hA model".
-In the INRANUKE hA model, hadrons can undergo at most one FSI per event.
+In the INTRANUKE hA model, hadrons can undergo at most one FSI per event.
 When possible, external hadron-nucleus scattering data are used to tune INTRANUKE.
-Since no data is available for Argon, GENIE uses an interpolation of data from heavier and lighter nuclei for the pion-argon cross section leading to resultant uncertainties in the INTRANUKE module of the order of 10-20\%.
+Since no data are available for argon, GENIE uses an interpolation of data from heavier and lighter nuclei for the pion-argon cross section leading to resultant uncertainties in the INTRANUKE module of the order of 10-20\%.
 
-Once the pion has left the target nucleus, the pion-argon hadronic cross section also plays an important role in the pion transportation inside the argon medium: processes such as pion absorption or pion charge exchange can greatly modify the topology of a neutrino interactions in the detector and lead to significant modifications in the event classification.
-Being able to reconstruct the details of pions inside the detector is an imperative for modern liquid argon neutrino experiment to achieve the design resolution for their key physics measurements.
+Once the pion has left the target nucleus, the pion-argon hadronic cross section also plays an important role in the pion transportation inside the argon medium: processes such as pion absorption or pion charge exchange can greatly modify the topology of a neutrino interaction	 in the detector and lead to significant modifications in the event classification.
+Being able to reconstruct the details of pions inside the detector is an imperative for modern liquid argon neutrino experiments to achieve the design resolution for their key physics measurements.
 %For example, if a pi+ charge exchanges into a pi0, an event which would not have been a background to a nue search now becomes one.
@@
                       &              &                     &           \\
- 2+ & \textbf{Pion Production}, $\sigma_{\pi prod}$    &           $\pi^- + N \rightarrow \ge2 \pi $ + nucleons            &  Possible if pion K.E $\ge$ 500 Mev/c       \\
+ 2+ & \textbf{Pion Production}, $\sigma_{\pi prod}$    &           $\pi^- + N \rightarrow \ge2 \pi $ + nucleons            &  Possible if pion K.E $\ge$ 500 MeV       \\
                     &              &                     &           \\
@@
 %}
-\caption{Summary of negative pion hadronic interactions of the reaction channel as a function of the pion multiplicity in the final state in the energy range [100-1200] MeV. }
+\caption{Summary of negative pion hadronic interactions of the reaction channel as a function of the pion multiplicity in the final state for the energy range [100-1200] MeV. }
 \label{tab:pionFinal}
@@
 Experimental results on several nuclei as reported in \cite{PhysRevC.14.635} are shown in Figure \ref{fig:pi1}; it is interesting to notice here how the shape of the $\Delta$ resonance becomes less pronounced as a function of the mass number of the target nucleus.
-Pion interactions with heavier nuclei also shift the peak of the resonance at lower energy; this effect is due to kinematic considerations and to the difference in propagation of the $\Delta$ inside the nucleus.
-Multiple scattering effect modify the resonance width, which is larger than the natural-decay width.
+Pion interactions with heavier nuclei also shift the peak of the resonance to lower energy; this effect is due to kinematic considerations and to the difference in propagation of the $\Delta$ inside the nucleus.
+Multiple scattering effects modify the resonance width, which is larger than the natural-decay width.
 As an example of a fairly well studied target, Figure \ref{fig:pi2} reports the negative pion cross section on Carbon for the elastic and reaction\footnote{ This paper calls  ``inelastic interaction" what we refer as to ``reaction channel".} channels, and their sum \cite{BINON1970168}.
@@
 LArIAT uses Geant4 as the default simulation package.
-In particular, pions (and kaons) transportation is achieved through the Geant4 FTFP\_BERT physics list.
+In particular, pion (and kaon) transportation is achieved through the Geant4 FTFP\_BERT physics list.
 In this physics list, Geant4 uses the Bertini cascade model  \cite{Wright:2015xia}  to simulate the products of the pion-nucleus interaction as well as secondary hadronic re-interactions inside the target nucleus (intra-nuclear cascade).
 The target nucleus is represented as a continuous gas where the nuclear potential follows concentrical shells whose depths approximate the Woods-Saxon shape.
-The CERN-HERA compilations \cite{Compilation1,Compilation2} of hadron-nucleon interaction data is the data base used for the decision making process after the cascade is invoked.
+The CERN-HERA compilations \cite{Compilation1,Compilation2} of hadron-nucleon interaction data is the database used for the decision making process after the cascade is invoked.
 The cross section model determines if the pion interacts, the eventual type of interaction and the interaction multiplicity.
-For  hadron projectiles with energy less than 20 GeV, Geant4 reports  the uncertainty on the cross section model to be about the size of the error bars on the data used, or about 10\%, increasing to 20-30\% in energy regions where data is sparse.
+For  hadron projectiles with energy less than 20 GeV, Geant4 reports  the uncertainty on the cross section model to be about the size of the error bars on the data used, or about 10\%, increasing to 20-30\% in energy regions where data are sparse.
 
@@
 The work in \cite{Nutini2015}  reviews  the current status of negative and positive pion simulation in Geant4 and GENIE for $^{12}$C, $^{56}$Fe, and $^{40}$Ca.
-From that work, we report the results for $^{12}$C in Figure \ref{fig:pionC} as it allows a direct comparison between Geant4, GENIE and and pion re-scattering data.
-Geant4 predictions for $\pi^-$  on Carbon are in good agreement with data over  the entire spectrum spectrum, while GENIE predictions seem to show some features at around 500 MeV and 900 MeV, maybe due to higher resonances in the hA model.
+From that work, we report the results for $^{12}$C in Figure \ref{fig:pionC} as it allows a direct comparison between Geant4, GENIE and pion re-scattering data.
+Geant4 predictions for $\pi^-$  on Carbon are in good agreement with data over  the entire  spectrum, while GENIE predictions seem to show some features at around 500 MeV and 900 MeV, maybe due to higher resonances in the hA model.
 From the same work, we also report the negative pion cross section on $^{40}$Ca in Figure \ref{fig:pionCa}, since this is the nuclear medium closest to argon.
-The predictions from both Geant4 and GENIE agree with data in the high energy region; the Geant4 and GENIE predictions diverge in the resonance region, where data is not available.
+The predictions from both Geant4 and GENIE agree with data in the high energy region; the Geant4 and GENIE predictions diverge in the resonance region, where data are not available.
 These few examples highlight how cross section data for the specific nucleus considered in the neutrino experiments is fundamental to inform the Monte Carlo simulation.
 
-For the LArIAT simulation of the MC sample used in the $\pi^-$ argon total hadronic cross section measurement we use the Geant4 Bertini Cascade model, whose  predictions for the total, elastic and reaction hadronic cross sections are show in Figure \ref{fig:PionTrueArgon}.
+For the LArIAT simulation of the MC sample used in the $\pi^-$ argon total hadronic cross section measurement we use the Geant4 Bertini Cascade model, whose  predictions for the total, elastic and reaction hadronic cross sections are shown in Figure \ref{fig:PionTrueArgon}.
 
@@
 	   	        \caption{ Total, elastic and reaction cross section for $\pi^-$ on $^{12}$C.
-Comparison between results from Geant4 simulation (Bertini cascade model), Genie simulation (hA model), and experimental data \cite{Clough1974, PhysRevC.23.2173, ALLARDYCE19731,osti_4488432}.}
+Comparison between results from Geant4 simulation (Bertini cascade model), GENIE simulation (hA model), and experimental data \cite{Clough1974, PhysRevC.23.2173, ALLARDYCE19731,osti_4488432}.}
         \label{fig:pionC}
@@
         \caption{Total, elastic and reaction cross section for $\pi^-$ on $^{40}$Ca.
-Comparison between results from Geant4 simulation (Bertini cascade model),  Genie simulation (hA model), and experimental data \cite{ALLARDYCE19731}.}
+Comparison between results from Geant4 simulation (Bertini cascade model),  GENIE simulation (hA model), and experimental data \cite{ALLARDYCE19731}.}
         \label{fig:pionCa}
@@
 \includegraphics[width=3in]{Chapter-1/Images/cPionsG4.png}
-\caption{Total, elastic and reaction hadronic cross section for $\pi^-$-argon implemented in Geant4  10.01.p3.}
+\caption{Total, elastic and reaction hadronic cross section for $\pi^-$-argon implemented in Geant4  10.01.p3 using FTBF\_BERT physics list.}
 \label{fig:PionTrueArgon}
@@
 % and of light mesons interaction with nuclei studies.
- We then describe the signal signature and historical measurements of kaon-nucleus cross section, as well as the implementation of this cross sections in the current version of the simulation package used by LArIAT.
+ We then describe the signal signature and historical measurements of kaon-nucleus cross section, as well as the implementation of this cross section in the current version of the simulation package used by LArIAT.
 

### 2018-06-01 Chapter-3/3-TheLArIATExperiment.tex
s

@@
 
-\subsection{LArTPC: Light Collection System}\label{sec:TPCLight} 
+\subsection{LArTPC: Light Collection}\label{sec:TPCLight} 
 The collection of scintillation photons is the second mechanism of particle detection in argon other than the ionization electrons.

### 2018-06-01 Chapter-5/5-Preparatory.tex
s

@@
 Beam Composition  @WC4      &  68.8\%   &  4.6 \%  & 26.6 \%    & 87.4 \% & 3.7 \%  & 8.9 \% \\ %
-Expected Composition in XS sample &  88.5\%   & 8.2\%   & 3.3 \%   & 94.0\%	& 5.3\% & 0.7\%\\
+Expected Composition&  88.5\%   & 8.2\%   & 3.3 \%   & 94.0\%	& 5.3\% & 0.7\%\\
+                                    in XS sample               &                      &                       &                   &                       &                        &\\  
                                                   &                      &                       &                   &                       &                        &\\  

### 2018-06-25 AppendixB-EField/B-LArIATEField.tex
s

@@
 
-Before getting into the details of the measurement procedures, it is important to explicit the relationship between some  quantities in play.
-The electric field and the drift velocity ($v_{drift}$) are related as follows
+Before getting into the details of the measurement procedures, it is important to clarify the relationship between the  quantities in play.
+The electric field ($E_{field}$) and the drift velocity ($v_{drift}$) are related as follows
 \begin{equation} v_{drift} = \mu(E_{field},T) E_{field}, \label{eq:vd}
@@
 where $\Delta x$ is the distance between the edges of the drift region.
-Table \ref{tab:Efields} reports the values of the electric field, drift velocity, and drift times for the smaller drift volumes.
+Table \ref{tab:Efields} reports the values of the electric field, drift velocity, and drift times for the TPC's small drift volumes.
 
@@
 \includegraphics[width=3in]{AppendixB-EField/Images/CircuitLArIAT.png}
-\caption{LArIAT HV simple schematics.}
+\caption{LArIAT HV simple scheme.}
 \label{fig:circuit}
@@
 \end{figure}
-We devise an independent method to measure the drift time (and consequently drift velocity and electric field) using TPC cathode to anode piercing tracks.
+We devised an independent method to measure the drift time (and consequently drift velocity and electric field) using TPC cathode to anode piercing tracks.
 We use this method as a cross check to the single line method.
@@
 \begin{itemize}
-\item[0.] Select cosmic ray events with only 1 reconstructed track 
-\item[1.] Reduce the events to the one containing tracks that cross both anode and cathode
-\item[2.] Identify the first and last hit of the track
-\item[3.] Measure the time difference between these two hits ($\Delta t$).
+\item[1.] Select cosmic ray events with only 1 reconstructed track 
+\item[2.] Reduce the events to the ones containing tracks that cross both anode and cathode
+\item[3.] Identify the first and last hit of the track
+\item[4.] Measure the time difference between these two hits ($\Delta t$).
 \end{itemize}
@@
 We choose cosmic events to allow for a high number of anode to cathode piercing tracks (ACP tracks), rejecting  beam  events where the particles travel almost perpendicularly to drift direction.
-We select events with only one reconstructed track to  maximize the chance of selecting a single crossing muon (no-michel electron).
-We utilize ACP tracks because their hits span the full drift length of the TPC, see figure \ref{fig:Scheme}, allowing us to define where the first and last hit of the tracks are located in space regardless of our assumption of the electric field.
+We select events with only one reconstructed track to  maximize the chance of selecting a single crossing muon (no michel electron).
+We utilize ACP tracks because their hits span the full drift length of the TPC, see Figure \ref{fig:Scheme}, allowing us to define where the first and last hit of the tracks are located in space regardless of our assumption of the electric field.
 %The definition of the last hit is easy: it is the hit closest to the cathode.
@@
 For each track passing our selection, we loop through the associated hits to retrieve the timing information.
-The analysis is performed separately on hits on the collection plane and induction plane, but lead to consistent results.
+The analysis is performed separately on hits on the collection plane and induction plane, but leads to consistent results.
 As an example of the time difference, figures \ref{fig:Run2PosColFit} and \ref{fig:Run2PosIndFit} represent the difference in time between the last and first hit of the selected tracks for Run-II Positive Polarity sample on the collection and induction plane respectively.

### 2018-06-25 AppendixB-TrackingStudies/B-Tracking.tex
s

@@
 
-Scope of this study is assessing the performances of the WC2TPC match on Monte Carlo (see Section \ref{ch:WC2TPCMatchMethod} ) and decide the selection values we will use on data.
+Scope of this study is assessing the performances of the WC2TPC match on Monte Carlo (see Section \ref{ch:WC2TPCMatchMethod}) and decide the selection values we will use on data.
 A word of caution is necessary here.
@@
 We chose the percentage of events whose reconstructed and true length differ less than 2 cm.
+In order to identify the true interaction point, no selection is performed on the scattering angle: even hadronic interaction flagged by Geant4 with very shallow angles (< 5$^\circ$) are kept to perform the optimization.
 We then identify the parameters in TrajCluster that are most important to  correctly stop  the tracking and an appropriate range of values for each of them.
@@
 After the optimization, the most upstream point of the tracking is correctly identified 99.5\% of the times, the most downstream point is correctly identified 62.5\% of the times, the tracking stops short about 15\% of the times and misses the interaction point 22.5\% of the times.
-Hand scanning confirmed that the missed interaction points happen in the vast majority of cases for very shallow angles, as shown in the event display in Figure, or in the case of angles visible only in one projection plane.
+Hand scanning confirmed that the missed interaction points happen in the vast majority of cases for very shallow angles, as shown in the event display in Figure \ref{evd:kink}, or in the case of angles visible only in one projection plane.
 We also noticed that the premature stopping of the tracks is often related to the presence of delta rays parallel to the track.
 We see room of improvement, such as the delta ray removal and a forced track breaking in case of a kink present in a single plane, for a future analysis.
-\textcolor{red}{ADD evd}
 
 The procedure behind this optimization package is virtually applicable to any LArSoft module where it is possible to define figure of merit.
+
+\begin{figure}[hpbt]
+\centering
+\includegraphics[width=15cm]{AppendixB-TrackingStudies/Kink.png}
+\caption{Example of a shallow angle hadronic interaction ``missed" by the TrajCluster.}
+\label{evd:kink}
+\end{figure}

### 2018-06-25 AppendixC-EnergyCalibration/C-EnergyCalibration.tex
s

@@
 
-Scope of the energy calibration is to identify the factors which convert the charge collected (dQ) to energy deposited in the chamber (dE).
-As described in section \ref{sec:SignalProc}, this is a multi-step procedure.
+The scope of the energy calibration is to identify the factors which convert the charge collected (dQ) to the energy deposited in the chamber (dE).
+As described in Section \ref{sec:SignalProc}, this is a multi-step procedure.
 In LArIAT, we first correct the raw charge by the electronic noise on the considered wire \cite{LArIATdqdx}, then by the electron lifetime \cite{LArIATLifeTime},  and then by the recombination using the ArgoNeut recombination values.
-Lastly, we apply overall calibration of the energy, i.e.
-we determine the ``calorimetry constants" using the procedure described in this section.
+Lastly, we apply overall calibration of the energy, i.e., we determine the ``calorimetry constants" using the procedure described in this section.
 
@@
 The basic idea of this calibration technique is to utilize a sample of beamline events with known particle species and momentum to measure the $dE/dX$ of the corresponding tracks in the TPC.
-In particular, we decided to use positive pions as calibration sample and samples from all the other particle species as cross check.
+In particular, we decided to use positive pions as the calibration sample and samples from all the other particle species as a cross check.
 Once the $dE/dX$ of the positive pion sample  has been measured at various momenta, we tune to calorimetry constants within the reconstruction software to align the measured values to match the theoretical ones found in Figure \ref{fig:PDGEnergyLossArgon}.
+This is a new technique for calorimetry calibration in liquid argon developed by a joint ArgoNeut and  LArIAT effort.
 
-In data, we start by selecting a sample of beamline positive pion beamline candidates without any restriction on their measured momentum\footnote{it should be noted that some muon and position contamination is present in the $\pi^+$ sample}.
+In data, we start by selecting a sample of beamline positive pion beamline candidates without any restriction on their measured momentum\footnote{It should be noted that some muon and position contamination is present in the $\pi^+$ sample.}.
 We then apply the WC2TPC match and subtract the energy loss upstream to the TPC front face, determining the momentum at the TPC front face.
@@
 We perform this  tuning for the collection and induction plane separately.
-As a cross check to the calorimetry constants determined using the positive pions, we lock the constants and  plot the $dE/dx$ versus momentum distribution of all the other particle species identifiable in the beamline data ($\pi/\mu/e$, K , p, in both polarities) against the corresponding Beth-Bloch prediction.
+As a cross check to the calorimetry constants determined using the positive pions, we lock the constants and  plot the $dE/dx$ versus momentum distribution of all the other particle species identifiable in the beamline data ($\pi/\mu/e$, K , p, in both polarities) against the corresponding Bethe-Bloch prediction.
 The agreement between data from the other particle species and the predictions is the expected result of this cross check.
@@
 
-After the calibration is done separately on data and MC, we can compare the resulting dE/dX distributions; this is done for the set of pion beamline candidates and pion MC on the left side of Figure \ref{fig:dedx}.
-On the right side of the same figure, we report the data-MC comparison for the kaon sample used in the cross section.
+After the calibration is done separately on data and MC, we can compare the resulting dE/dX distributions; this is done for the set of pion beamline candidates and pion MC used in the cross section analysis on the left side of Figure \ref{fig:dedx}.
+On the right side of the same figure, we report the data-MC comparison for the kaon sample used in the cross section analysis.
 The distributions are fitted with simple Landau functions.
@@
 \caption{\emph{Left:} dE/dx distribution for $\pi^-/\mu^-/e^-$ data (black) and Pion MC (blue).
-The Landau fit for data is shown in red, the one for MC in teal.\emph{Right:} dE/dx distribution for $K^+$ data (black) and Kaon MC (blue).
+The Landau fit for data is shown in red, the one for MC in teal.
+\emph{Right:} dE/dx distribution for $K^+$ data (black) and Kaon MC (blue).
 The Landau fit for data is shown in red, the one for MC in teal.

### 2018-06-25 Chapter-0/0-Introduction.tex
s

@@
 To make up for the small neutrino interaction cross section, neutrino experiments moved to very large size, at the expenses of spatial precision.
-This is the case for the detectors which discovered neutrino oscillation:  both Super-Kamiokande and SNO are massive Cherenkov detectors.
+This is the case for the detectors which discovered neutrino oscillation:  both Super-Kamiokande and SNO are massive Cherenkov detectors \cite{PDGOsc}.
 With LArTPCs, the field is gaining again bubble-chamber like precision but at massive scales.
@@
 DUNE has a vast neutrino and non-accelerator physics reach.
-For what it concerns neutrino physics, oscillation analyses in DUNE have the capability of solving the mass hierarchy and octant problem,  and discovering CP violation in the neutrino sector.
+Within neutrino physics, oscillation analyses in DUNE have the capability of solving the mass hierarchy and octant problem,  and discovering CP violation in the neutrino sector.
 Besides its neutrino program, DUNE can open an experimental window on Grand Unified Theories (GUTs).
@@
 Argon is a fairly heavy element, which means that nuclear effects play an important role in the interaction topology.
-For example, pions are one of the main products of neutrino interactions; yet,  since data on charged particle interaction in argon is scarce, neutrino event generators have big uncertainties in the re-scattering simulation of pion in argon.
+For example, pions are one of the main products of neutrino interactions; yet,  since data on charged particle interaction in argon is scarce, neutrino event generators have big uncertainties in the re-scattering simulation of pions in argon.
 %No measurement of hadronic cross sections for pions or kaons is available for argon (yet).

### 2018-06-25 Chapter-1/1-Theory.tex
s

@@
 \subsection{Neutrinos in the Standard Model}
-Neutrino can be introduced in the SM as left-handed massless Weyl spinors.
+Neutrinos can be introduced in the SM as left-handed massless Weyl spinors.
 The Dirac equation of motion for a free field
@@
 Experimental results on neutrino oscillations are generally reported in terms of the mixing angles and of the squared mass splitting $\Delta m^2_{ab} = m^2_{a} - m^2_{b}$, where $a$ and $b$ represent the mass eigenstates.
-A summary of the current status of experimental results is given in table \ref{tab:nuosc}.
+A summary of the current status of experimental results extracted from \cite{PDGOsc} is given in table \ref{tab:nuosc}.
 
@@
 \centering
-\caption{Summary of experimental results on neutrino oscillation parameters. \textcolor{red}{ADD CITATIONS}}
+\caption{Summary of experimental results on neutrino oscillation parameters extracted from \cite{PDGOsc}.}
 \label{tab:nuosc}
@@
 $\theta_{23}$            & $\sim$45$^\circ$                      & 9.0\%   & Super Kamiokande, MINOS,\\
-|$\Delta m^2_{32}$|    & 2.5 $10^{-3}$ eV$^2$   & 1.8\% &  NO$\nu$a, MACRO, Super Kamiokande\\
+$|\Delta m^2_{32}|$    & 2.5 $10^{-3}$ eV$^2$   & 1.8\% &  NO$\nu$a, MACRO, Super Kamiokande\\
 \hline
 $\theta_{12}$            & 34$^\circ$                     & 5.8\% &  SNO,Gallex,\\
-|$\Delta m^2_{12}$|    & 7.4 $10^{-5}$ eV$^2$   & 2.8\% &  SAGE,  KamLAND \\
+$|\Delta m^2_{12}|$    & 7.4 $10^{-5}$ eV$^2$   & 2.8\% &  SAGE,  KamLAND \\
 \hline
 $\theta_{13}$            &   9$^\circ$                    & 4.7\%  & Daya  Bay, \\
-|$\Delta m^2_{32}$|    & 2.5 $10^{-3}$ eV$^2$   & 1.8\% &  RENO\\
+$|\Delta m^2_{32}|$    & 2.5 $10^{-3}$ eV$^2$   & 1.8\% &  RENO\\
 \hline
@@
 \centering
-\includegraphics[width=0.7\textwidth]{Chapter-1/Images/KaonsEvd.pdf}
+\includegraphics[width=\textwidth]{Chapter-1/Images/KaonsEvd.pdf}
 \caption{Kaon-Argon interaction topologies due to elastic scattering and inelastic scattering as seen in LArIAT data (collection plane only).}

### 2018-06-25 Chapter-2/2-LArTPCs.tex
s

@@
 \section{The Liquid Argon Time Projection Chamber Technology}
-In this section, we outline an extremely brief history of Time Projection Chambers as particle detectors, focusing on their incarnation as Argon detectors for neutrino physics.
+In this section, we outline an extremely brief history of Time Projection Chambers as particle detectors, focusing on their incarnation as argon detectors for neutrino physics.
 We further describe the working principles of Liquid Argon Time Projection Chambers, leading  to the description of the event reconstruction in LArTPC.
@@
 where $\lambda = 1/N_{e}\sigma_i$, with $N_{e}$ being the electron density of $\sigma_i$ the ionization cross-section per electron.
-About 66\% of the ionizing collisions in Argon produce only a single electron/ion pair \cite{0034-4885-73-11-116201}; in the other cases, the transferred kinetic energy is enough for the primary electron to liberate one or more secondary electrons, which usually stay close to the original pair.
+About 66\% of the ionizing collisions in argon produce only a single electron/ion pair \cite{0034-4885-73-11-116201}; in the other cases, the transferred kinetic energy is enough for the primary electron to liberate one or more secondary electrons, which usually stay close to the original pair.
 Occasionally, electrons can receive enough energy to be ejected with high energy, forming a so-called ``$\delta$-ray": a detectable  short track off the particle trajectory, as shown in figure \ref{fig:delta}.
@@
 \includegraphics[width=\textwidth]{Chapter-2/Images/Delta.png}
-\caption{Events display for a LArIAT pion absorption candidate on the induction plane, with highlighted delta ray.}
+\caption{Event display for a LArIAT pion absorption candidate on the induction plane, with highlighted delta ray.}
 \label{fig:delta}
@@
 The information resulting from this process on a single spike form a single reconstructed ``hit", see Figure \ref{fig:SignalProc} panel c).
-The next steps in the event reconstruction chain will then decide if rejecting hits with poor fits.
+The next steps in the event reconstruction chain will then decide which hits to use according to their goodness of fit.
 It is important to notice how the height and width of the hit depend on the topology of the event: for example, a particle running  parallel to the wire planes will leave a series of narrow individual hits, one on each consecutive wire, while a particle traveling towards the planes will leave long, wide hits on very few wires.
@@
 The development of different approaches to accomplish this task is an extremely hot topic in LArTPC event reconstruction which spans from more traditional approaches such as line-clustering  \cite{Barker2011} to the use of machine learning tools \cite{1748-0221-12-03-P03011}.
-Generally speaking, the scope of hit clustering and event reconstruction to provide shower-like or track like-objects with an associated energy reconstruction.
+Generally speaking, the scope of hit clustering and event reconstruction is to provide shower-like or track like-objects with an associated energy reconstruction.
 This is because different particles have different topology in the detector -- electrons and photon create electromagnetic showers,  resulting in shower-like topologies, while muons and hadrons  leave track-like signals.
-For the scope of these thesis, we will describe only LArIAT's approach to track reconstruction even if we recognize the breath of LArTPC event reconstruction is much wider.
+For the scope of this thesis, we will describe only LArIAT's approach to track reconstruction even if we recognize the breath of LArTPC event reconstruction is much wider.
 We are interested in the reconstruction of pions and kaons in the active volume, whose topology is track-like.\\
@@
 Even if calorimetry is done in 2D, it benefits from the 3D tracking information; typical objects available after the calorimetric reconstruction are the total energy deposited by the particle and its stopping power $dE/dx$ at each ``track pitch", i.e.
-at each 2D projection on the wire plane of the 3D trajectory.
+at each segment between two 3D points projected on the wire plane.
 

### 2018-06-25 Chapter-3/3-TheLArIATExperiment.tex
s

@@
 { Magnetic field over current as a function of the current, for one NDB magnet (excitation curve).
-The data was collected using two hall probes (blue and green).
+The data were collected using two hall probes (blue and green).
 We fit the readings with a cubic function (black) to average of measurements (red) given in the legend \cite{detectorPaper}.}
@@
 For the Run II data taking period, the current settings explored were 60A (B $\sim$0.21 T) and 100A (B $\sim$0.35 T) in both polarities.
-Albeit advantageous to enrich the tertiary beam composition with high mass particles such as kaons, we never pushed the magnets current over 100 A, so as not to incur in overheating.
+Albeit advantageous to enrich the tertiary beam composition with high mass particles such as kaons, we never pushed the magnets current over 100 A, so to not overheat them.
 During operation, we operated an air and water cooling system on the magnets and we remotely monitored the magnet temperatures.
@@
 \caption{
-{One of the four Multi Wire Proportional Chambers (WC) used in the LArIAT tertiary beamline and related read-out electronics.}
+{One of the four Multi-Wire Proportional Chambers (WC) used in the LArIAT tertiary beamline and related read-out electronics.}
 }
@@
 The TDC time resolution is 1.18~ns/bin and can accept 2 edges per 9~ns.
-The maximum event rate acceptable by the chamber system is 1 MHz.%: this rate is not a limiting factor considering that the rate of the tertiary particle beam at the first wire chamber is estimated to be less than 15 kHz}.
+The maximum event rate acceptable by the chamber system is 1 MHz. 
 A full spill of data occurring once per supercycle is stored on the TDC board memory at once and read out by a specially designed controller.
@@
 Cosmic ray muons whose average energy is in the few GeV range crossing both anode and cathode populate the events triggered this way.
-This particularly useful sample of tracks is associated can be used for many tasks; for example, we use anode-cathode piercing tracks to cross check the TPC electric field on data (see Appendix \ref{ch:AppendixB}), to calibrate the charge response of the TPC wires for the full TPC volume and to measure the electron lifetime in the chamber \cite{LArIATLifeTime}.
+This particularly useful sample of tracks can be used for many tasks; for example, we use anode-cathode piercing tracks to cross check the TPC electric field on data (see Appendix \ref{ch:AppendixB}), to calibrate the charge response of the TPC wires for the full TPC volume and to measure the electron lifetime in the chamber \cite{LArIATLifeTime}.
 

### 2018-06-25 Chapter-4/4-Methodology.tex
s

@@
 
-This chapter describes the general procedure employed to measure  total hadronic interaction cross sections on Ar in LArIAT.
+This chapter describes the general procedure employed to measure  total hadronic interaction cross sections on argon in LArIAT.
 Albeit with small differences, both the  ($\pi^{-}$,Ar) and (K$^{+}$,Ar) total hadronic cross section measurements rely on the same procedure.
@@
 If the particle enters a slice, it contributes to the $N_{\text{Inc}}( E^{kin})$ distribution in the energy bin corresponding to its kinetic energy in that slice.
-While into the slice, a hadron may or may not interact.
+Within the slice, the hadron may or may not interact.
 If it interacts in the slice, it  contributes also to the $N_{\text{Int}}(E^{kin})$ distribution in the appropriate energy bin; this occurrence corresponds to the end of the hadron tracking.

### 2018-06-25 Chapter-5/5-Preparatory.tex
s

@@
 G4Beamline simulates the beam collision at the LArIAT secondary target, the energy deposited by the particles in the LArIAT beamline detectors, and the action of the LArIAT magnets, effectively accounting for particle transport through the beamline from the LArIAT target until ``Big Disk", a fictional, void detector located just before the LArIAT cryostat.
- At the moment of this writing, G4Beamline does not simulated the responses of the beamline detectors.
+ At the moment of this writing, G4Beamline does not simulate the responses of the beamline detectors.
 It is possible to interrogate the truth level information of the simulated particles in several points of the geometry.
@@
 Thus, we need to assess the contamination from beamline particles other than pions in the event selections used for the pion cross section analysis and correct for this background.
-The first step of this process is assessing the percentage of electrons and muons in the $\pi/\mu/e$ beamline candidates via the G4Beamline MC, as we deem the percentage of kaons and antiprotons negligible after the mass selection (< 1 \textperthousand).
+The first step of this process is assessing the percentage of electrons and muons in the $\pi/\mu/e$ beamline candidates via the G4Beamline MC, as we deem the percentage of kaons and antiprotons negligible after the mass selection ($<$ 1 \textperthousand).
 Since the beamline composition is a function of the magnet settings, we simulate separately events for magnet current of -60A and -100A. 
@@
 
-\subsubsection{Beam Composition for Positive Kaon Cross Section}
+\subsubsection{Beam Composition for Positive Kaon Cross Section}\label{ch:protons}
 In the positive polarity runs, the tertiary beam composition is mainly pions and protons.
@@
 The procedure is repeated for lighter particles mass peak independently.
-With 12 iterations of this method we find a proton contamination of  5.0 $\pm$ 2.0 \%  and a contamination from the lighter particles of 0.2 $\pm$ 0.5 \% .
+With 12 iterations of this method we find a proton contamination of  5.0 $\pm$ 2.0 \%  and a contamination from the lighter particles of 0.2 $\pm$ 0.5 \%.
 The estimate of the proton background is currently not used in the kaon cross section analysis, but it is a fundamental step to retrieve the true kaon cross section which will be implemented in the further development of the analysis.
@@
 \subsection{Contributions from the Beamline Background}\label{ch:PionXSBkgSub2}
-We define beamline background every TPC track matched to the WC track which is not a primary pion. Potentially, there are 4 different types of beamline background:
+We define ``beamline background" as every TPC track matched to the WC track which is not a primary pion. Potentially, there are 4 different types of beamline background:
 \begin{itemize}
@@
 track multiplicity rejection, WC2TPC match and shower rejection) on the three samples.
-The number of events per particle species surviving this selection is shown on table \ref{tab:MCafterCutContaminants}.
+The number of events per particle species surviving this selection is shown on Table \ref{tab:MCafterCutContaminants}.
 In order to reproduce the data beamline composition, we weight each event of a given particle species according to the estimated content for that species as found with the G4Beamline simulation.
@@
 One of the interaction or decay products can leak into the TPC and be matched with the WC track, contributing to the pool of events used for the cross section calculation.
-We call this type of particles ``secondaries" from pion events, with a terminology inspired by Geant4.
+We call these occurrences ``secondaries" from pion events, with a terminology inspired by Geant4.
 We estimate the number of secondaries using the DDMC pion sample.
@@
 In Figure \ref{fig:PxVsXTrue} , we plot the true  horizontal component of the momentum $P_x$ versus the true $X$ position at WC4 for pions missing the halo paddle (left) and for pions hitting the halo paddle (right) for the -60A MC simulation runs -- analogous plots are obtained with the -100A pion simulation and with the kaon simulation.
-These distributions can be separated drawing a line in this position-momentum space.
+These distributions can be separated by a line in this position-momentum space.
 We use a logistic regression  \cite{agresti2013categorical}  as a classifier to find the best separating line, shown in both plots as the red line.
@@
 The paddle with the hollow central circle represents the Halo paddle.
-We illustrate two possible trajectories: in black, a trajectory that miss the paddle and goes through the hole in the Halo, in blue a trajectory that hits the Halo paddle and goes through the scintillation material.}
+We illustrate two possible trajectories: in black, a trajectory that misses the paddle and goes through the hole in the Halo, in blue a trajectory that hits the Halo paddle and goes through the scintillation material.}
 \label{fig:Halo}
@@
 \end{equation}   
-In case of the reaction channel,  the presence of several topologies requires a more complex definition, as shown in figure \ref{fig:scatterPic}.
+In case of the reaction channel,  the presence of several topologies requires a more complex definition, as shown in Figure \ref{fig:scatterPic}.
 We define the scattering angle as the biggest of the angles between the incoming hadron and the visible daughters, where the visible daughters are charged particles that travel more than the average pitch length ($\delta X = 47$ mm) in the detector (see panel a); in case all the daughters are invisible, the angle is assigned to be 90 deg (see panel b).
-We chose this working definition of scattering angle for inelastic scattering keeping in mind how our tracking reconstruction works: the tracking will stop correctly if none of the daughters is visible in the detector and it is likely to stop correctly if multiple visible daughters form an interaction vertex.
+We chose this working definition of scattering angle for inelastic scattering keeping in mind how our tracking reconstruction works: the tracking will stop correctly if none of the daughters are visible in the detector and it is likely to stop correctly if multiple visible daughters form an interaction vertex.
 The only ``dangerous" case is the production of one charged daughter plus neutrals, which we can study with this working definition of scattering angle (see panel c).
@@
 The left side of Figure \ref{fig:trueWithAngles} shows the  true Geant4 cross section  for interaction angles greater than 0 deg (green), 4.5 deg (red) corresponding to the MC angular resolution, 5.0 deg  (blue) corresponding to the data angular resolution, and 9.0 deg (yellow).
-A small $0.5 \text{ deg}$ systematic shift between the mean of the data and MC angular resolution is present, which has a negligible impact on the cross section.
+As reported in Table \ref{tab:AngRes}, we find a $0.5 \text{ deg}$ difference between the mean of the data and MC angular resolution, which has a negligible impact on the cross section.
 The right side of Figure \ref{fig:trueWithAngles} shows the ratio between the  true cross section for interaction angles greater than 5 deg and the true interaction cross section for all angles;  the cross section for angles greater than 5$^\circ$ accounts for more than 80\% than the total cross section in every energy bin.
@@
 where $E_{i}$ is measured in each slab as  the product of the stopping power,  $dE/dX_{i}$,  and the track pitch for that point.
-Since the measurements of the energy deposited in each slab rely on the same global calorimetric procedure and tracking algorithms, we assume conservatively that that the measurements of $E_{i}$ are not independent from one another; thus, the uncertainty on $E_{\text{FF-j}}$ becomes
+Since the measurements of the energy deposited in each slab rely on the same global calorimetric procedure and tracking algorithms, we assume conservatively that the measurements of $E_{i}$ are not independent from one another; thus, the uncertainty on $E_{\text{FF-j}}$ becomes
 \begin{equation}

### 2018-06-25 Chapter-6/6-PionCrossSection.tex
s

@@
 
-In this chapter, we show the result of the thin slice method to measure 
+In this chapter, we show the results of the thin slice method to measure 
 the ($\pi^-$-Ar) total hadronic cross section.
 In Section \ref{ch:PionXSRaw}, we start by measuring the raw cross section, i.e.
-the cross section obtained exclusively using data reconstruction, without any additional corrections.
+the cross section obtained exclusively using data reconstruction, without any additional correction.
 In Section \ref{ch:PionXSCorrections}, we apply the statistical subtraction of the background contributions based on simulation and the correction for reconstruction effects.
@@
 As we will clarify in Section \ref{ch:PionXSCorrections},  the corrections to the raw cross section depend on the beam settings and need to be calculated independently for the two datasets.
-Thus, we present here the measurement of the raw cross section on the two datasets separately.
+Thus, we present here the measurements of the raw cross section on the two datasets separately.
 
@@
 Figure \ref{fig:IncidentRaw} shows the distribution of  $N^{\text{TOT}}_{\text{Inc}}$   for the 60A dataset on the left and for the 100A dataset on the right.
-Data is displayed over the MC.
+Data is displayed over MC.
 The same color scheme and normalization procedure is used for both the interacting and incident histograms.
@@
 $N^{\text{TOT}}_{\text{Inc}}$ and $N^{\text{TOT}}_{\text{Int}}$ are not independent.
-In fact, the population of a given bin for the interacting histogram always implies at least the population of the same bin in the incident histogram  (and possibly other incidents bins at higher energies).
+In fact, populating a given bin for the interacting histogram always implies at least populating the same bin in the incident histogram  (and possibly other incidents bins at higher energies).
 Thus, we conservatively calculate the statistical uncertainty on the cross section as
@@
 \subsection{Treatment of Systematics} \label{ch:SysUncertaintyXSRaw}
-The systematic effects considered in the measurement of the raw cross section results from the uncertainty on the track pitch  ($\delta X = 0.47 \pm 0.03 $ cm, shown in \ref{sec:trackPitch})  and from the propagation of the uncertainty associate with the measured kinetic energy at each argon slab.
+The systematic effects considered in the measurement of the raw cross section result from the uncertainty on the track pitch  ($\delta X = 0.47 \pm 0.03 $ cm, shown in \ref{sec:trackPitch})  and from the propagation of the uncertainty associate with the measured kinetic energy at each argon slab.
 As shown in Section \ref{ch:kinEn}, the uncertainty on the kinetic energy of a pion candidate at the j$^{th}$ slab of argon  is given by
@@
 
-We propagate the uncertainty on the beamline composition  as a systematic uncertainty to the cross section by varying the beam composition for all the cases listed in Table \ref{tab:beamlineSys} and evaluating variation of obtained  data cross sections in each bin.
-This systematic uncertainty is summed in quadrature with the statistical uncertainty and the systematic uncertainty related to the kinetic energy measurement.
+We propagate the uncertainty on the beamline composition  as a systematic uncertainty to the cross section by varying the beam composition for all the cases listed in Table \ref{tab:beamlineSys} and evaluating the variation of the obtained  data cross sections in each bin.
+This systematic uncertainty is summed in quadrature with the statistical uncertainty and the systematic uncertainty calculated for the raw cross section.
 
@@
 This definition holds regardless the type of the interaction, i.e.
-if the TPC track ends within the fiducial volume, its last point will be the interaction point, no matter what the products of the interaction look like; on the other hand, if the track crosses the boundaries of the fiducial volume, the particle will be considered ``through going" and no interaction point will be found.
+if the TPC track ends within the fiducial volume, its last point will be the interaction point, no matter what the products of the interaction look like; conversely, if the track crosses the boundaries of the fiducial volume, the particle will be considered ``through going" and no interaction point will be found.
 Given this definition, it is evident that we rely on the tracking algorithm to discern where the interaction occurred in the TPC  and correctly end the tracking.
@@
 The bottom plot shows the two measurements overlaid.
-In all three plot, the Geant4 prediction for the total hadronic cross section for angle scattering greater than 5$^\circ$ is displayed in green.
+In all three plot, the FTFP\_BERT Geant4 prediction for the total hadronic cross section for angle scattering greater than 5$^\circ$ is displayed in green.
 
@@
 The cross section measurements in the two datasets agree within the systematic uncertainty, even if the cross section measured in the 60A data is lower than the one measured in the 100A data in all overlapping bins except for the 200-250 MeV bin.
+With the exception of the highest energy bins, the systematic uncertainties are the dominant uncertainties for both the 60A and 100A datasets.
 
@@
 \includegraphics[width=0.48\textwidth]{Chapter-6/Images/TheMoneyPlot.pdf}
-\caption{ \emph{Top Left:} ($\pi^-$-Ar) total hadronic cross section for  scattering angles greater than 5$^\circ$ measured in the 60A sample, statistical uncertainty in black and systematic uncertainty in red.
-The Geant4 prediction for the total hadronic cross section for angle scattering greater than 5$^\circ$ is displayed in green.
+\caption{ \emph{Top Left:} ($\pi^-$-Ar) total hadronic cross section for  scattering angles greater than 5$^\circ$ measured in the 60A sample, statistical uncertainty and systematic uncertainty in red.
+The FTFP\_BERT  Geant4 prediction for the total hadronic cross section for angle scattering greater than 5$^\circ$ is displayed in green.
 \\
-\emph{Top Right:} ($\pi^-$-Ar) total hadronic cross section for  scattering angles greater than 5$^\circ$ measured in the 100A sample, statistical uncertainty in black and systematic uncertainty in blue.
-The Geant4 prediction for the total hadronic cross section for angle scattering greater than 5$^\circ$ is displayed in green.\\
-\emph{Bottom:} ($\pi^-$-Ar) total hadronic cross section measurements in the 60A and 100A samples overlaid with the  Geant4 prediction (green).}
+\emph{Top Right:} ($\pi^-$-Ar) total hadronic cross section for  scattering angles greater than 5$^\circ$ measured in the 100A sample, statistical uncertainty and systematic uncertainty in blue.
+The FTFP\_BERT  Geant4 prediction for the total hadronic cross section for angle scattering greater than 5$^\circ$ is displayed in green.\\
+\emph{Bottom:} ($\pi^-$-Ar) total hadronic cross section measurements in the 60A and 100A samples overlaid with the  FTFP\_BERT Geant4 prediction (green).}
 \label{fig:FinalXSPion}
@@
 The top panel of Figure \ref{fig:FinalXSResPion} shows the cross section obtained combining the two datasets.
-To combine the datasets, we merge the corrected interacting histogram for the 60A data with the interacting histogram for the 100A data; we apply the same merging with the incident histograms.
+To combine the datasets, we merge the corrected interacting histogram for the 60A data with the corrected interacting histogram for the 100A data; we apply the same merging with the incident histograms.
 We then calculate the cross section central point and statistical uncertainty using the merged interacting and incident histograms.
 The systematic uncertainty is calculated as the sum in quadrature of the original systematic uncertainties on the cross section measurements on the two separated data sets.
+Again, with the exception of the highest energy bins (energy greater than 800 MeV), the systematic uncertainties are the dominant uncertainties; the lion's share of the systematic uncertainty results from our uncertainty on the beam composition, which we plan to constrain with different beamline data in a future iteration of the analysis.
 
-The bottom panel of figure  \ref{fig:FinalXSResPion} shows the relative deviation between the measured cross section and the Geant4 predicted cross section.
+The bottom panel of Figure  \ref{fig:FinalXSResPion} shows the relative deviation between the measured cross section and the FTFP\_BERT  Geant4 predicted cross section.
 A couple of considerations are in order here.
@@
 We expected the pion cross section in argon to be well modeled in Geant4, given the vast set of measurements available for pion interaction on different targets -- even if no previous data is available on argon.
-Thus, this agreement confirms our confidence in the experimental methodology used.
+Thus, the data-MC agreement is expected and confirms the validity of the experimental methodology used.
 On the other hand, this agreement is not perfect.

### 2018-06-25 Chapter-7/7-KaonCrossSection.tex
s

@@
 
-In this chapter, we show the result of the thin slice method to measure  the ($K^+$-Ar) total hadronic cross section.
+In this chapter, we show the results of the thin slice method to measure  the ($K^+$-Ar) total hadronic cross section.
 In Section \ref{ch:KaonXSRaw}, we start by measuring the raw cross section.
@@
 Figure \ref{fig:IncidentRawK} shows the distribution of  $N^{\text{TOT}}_{\text{Inc}}$.
-Data is displayed over the MC.
+Data is displayed over MC.
 For the  $N^{\text{TOT}}_{\text{Inc}}$ distribution we do not make a distinction between kaons that decay or interact hadronically because any kaon contributes to the flux of incident particles at given kinetic energy independently from its final interaction.
@@
 Figure \ref{fig:XSRawK} shows the raw cross section, statistical uncertainty in black and systematic uncertainty in red.
-The raw data cross section is overlaid to the reconstructed cross section for the MC mixed sample, displayed in azure.
+The raw data cross section is overlaid to the reconstructed cross section for the MC  sample, displayed in azure.
 We calculate the statistical uncertainty for the interacting, incident and cross section distributions in a similar fashion to the pion case as described in Section \ref{ch:StatUncertaintyXSRaw}.
@@
 Currently, the only background considered for the kaon hadronic cross section comes from the presence of secondaries and decay interaction points.
-A further development of the analysis will need to account for the presence of a small proton contamination.
+A further development of the analysis will need to account for the presence of a small proton contamination (estimated to be $\sim 5\%$ in Section \ref{ch:protons}).
 Figure \ref{fig:CorrectionsBeamK} shows the relative kaon content suitable for the hadronic cross section for the interacting and incident histograms.
@@
 The systematic uncertainty on the cross section is the sum in quadrature of the statistical uncertainty, the systematic uncertainty calculated for the raw cross section and the systematic uncertainty related to the reconstruction  effects  correction.
+In every bin, the statistical uncertainty is the dominant uncertainty; a future iteration of the analysis using the ``High Yield" sample, which is expected to boost the statistics by a factor of 3, will reduce the statistical uncertainty.
 

### 2018-06-25 Drafting.tex
s

@@
-\subsubsection{Inputs to the Trigger System}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% yale_thesis.tex
+% Elena Gramellini
+% Started: 2017/12/05
+%
+% A bare, sample template for a Yale PhD thesis using yalephd.cls
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-Two primary inputs to the trigger card are from the time of flight (see Sec.~\ref{sec:TOF}) and the wire chamber (see Sec.~\ref{sec:MWPC}) systems in the tertiary beamline.
-Coincidence of activity in both of these systems strongly suggests that a charged particle has made the journey down the tertiary beam line from the copper target to the LArTPC in the cryostat, and that we will have a measure of its momentum and its velocity.
-PMT pulses from the time of flight system each pass through a
-%FIXME WHAT MODEL linear fanout?
-linear fan-out, one output of which is threshold discriminated by a
-%FIXME WHAT MODEL
-discriminator to produce a NIM logic pulse for use in trigger logic.
-On each of the upstream or downstream TOF paddles, we form a coincidence to within 20~ns of pulses from all the PMTs observing that same block of scintillator.
-The upstream paddle coincidence signal is delayed 20~ns to allow any approximately lightspeed particles to travel 6.5~m to the downstream paddle.
-The same upstream paddle coincidence signal is also widened to 100~ns to allow for slow-traveling (high-mass) particles.
-Coincidences of these upstream and downstream TOF trigger input signals can be made in the V1495 trigger card.
+\documentclass[letterpaper,12pt]{yalephd}
+% remove draft option for final printing.
+% font size must be between 10pt-12pt.
+\usepackage{caption,setspace}
+\usepackage{amsmath} 
+\usepackage{rotating}
+%\usepackage{newtxtext,newtxmath,amsmath}
+\usepackage{ textcomp }
 
-Each wire chamber is read out by four multi-hit TDCs (time to digital converters), with sixty-four wires per TDC and two TDCs per plane (one horizontal and one vertical).
-The TDCs each provide a logical ``fast'' OR of their inputs, indicating that one or more of their sixty-four wires went over the settable threshold.
-Using NIM logic units, the OR of the horizontal wires and the OR of the vertical wires are input to a coincidence unit for each wire chamber, providing a single logical pulse for each of the four wire chambers, indicating that at least one horizontal wire and one vertical wire saw significant signal.
-The single culminating logical pulses from each of the four wire chambers make up the first four inputs to the V1495 trigger card.
-Within the user-programmable FPGA, the V1495 looks for the coincidence within 20~ns of at least three of these four specially-treated logical inputs.
+\usepackage{geometry} % you need this for yalephd.cls to work.
+\usepackage{graphicx}
+\usepackage[graphicx]{realboxes}
+ % you probably want the rest of these.
+\usepackage{dcolumn}
+\usepackage{bm}
+\usepackage{hyperref}
+\usepackage[export]{adjustbox}
 
-Another primary input to the trigger card is from the cosmic towers (see Section~\ref{sec:CosmicRayPaddle}).
-To capture cosmic ray events in which a minimally ionizing cosmic ray muon crossed the TPC along the body diagonal, NIM modules form the logical coincidences from the two cosmic towers, one upper and one lower paddle assembly, in each combination.
-The OR of these is provided as an input to the V1495.
+\usepackage{mathtools}
+\DeclarePairedDelimiter\bra{\langle}{\rvert}
+\DeclarePairedDelimiter\ket{\lvert}{\rangle}
+\DeclarePairedDelimiterX\braket[2]{\langle}{\rangle}{#1 \delimsize\vert #2}
+\usepackage{textgreek}
 
-Three important logic pulses are derived from the timing of the beam.
-These include a pulse in a brief window before the beam, a pulse indicating that the beam is on, and a pulse which defines the beam-free period which may be used for collecting cosmic-ray events.
-An adjustable pulser is a fourth trigger input which does not depend on any particular activity in the experiment hall,  useful for collecting background events with zero bias.
+\usepackage{multirow}
+\usepackage{amsmath}
+\usepackage{amsfonts}
+\usepackage{amssymb}
+\usepackage{appendix}
+\usepackage{comment}
+\usepackage{cite}
+\usepackage{color}
+\bibliographystyle{plain}
+\usepackage{slashed}
 
-The PMTs observing liquid argon scintillation light (see Section \ref{sec:PhotonSystem}) produce pulses which form the foundation of several interesting trigger inputs.
-Thresholding a copy of each PMT pulse (after amplification), and requiring a coincidence of pulses within $\sim$20~ns, creates simple trigger inputs indicating ionizing radiation was produced in the TPC.
-This scintillation logic pulse is used to initiate a gate which spans the length of the TPC drift time, creating a logic signal which is remains ``on'' while significant drift charge may still be present in the TPC.
-In addition, requiring a delayed coincidence of two subsequent scintillation logic pulses, separated by a variable length of time ranging from 300~ns to 7~$\mu$s, is used to create a trigger input to select events where a cosmic muon stops and decays to a Michel electron in the TPC.
-A few different versions of this light-based trigger were implemented throughout the course of LArIAT's run time to allow reconstruction and calorimetric studies of Michel electrons.
-Figure~\ref{michel_logic} shows a schematic diagram of the logic comprising the Michel electron trigger.
+%\setcounter{chapter}{-1}
 
-\begin{figure}
-\includegraphics[width=\textwidth]{figures/trigger_michellogic.png}
-\caption{\label{michel_logic}A schematic diagram of the trigger logic used to select Michel electron events during the cosmic readout window of the LArIAT supercycle.
-The two PMT signals refer to the Hamamatsu (``HMM'') and ETL PMTs described in Section~\ref{sec:PhotonSystem}.
-For some data-taking periods in Run-II, un-amplified pulses were discriminated at 180 mV to act as a veto on events that may saturate the dynamic range of the V1751 digitizer.
-The discriminator thresholds used on the amplified (x10) PMT signal copies (\emph{ThA}, \emph{ThB}) as well as the Gate Delay period, were adjusted between run periods while experimenting with different configurations.}
-\end{figure}
+\newenvironment{dedication}
+  {%\clearpage           % we want a new page          %% I commented this
+   \thispagestyle{empty}% no header and footer
+   \vspace*{\stretch{1}}% some space at the top
+   \itshape             % the text is in italics
+   \raggedleft          % flush to the right margin
+  }
+  {\par % end the paragraph
+   \vspace{\stretch{3}} % space at bottom is three times that at the top
+   \clearpage           % finish off the page
+  }
 
-Further trigger inputs come from the beam line instrumentation behind the LArTPC cryostat, the PMTs of the Punch-Through scintillator paddles and those of the scintillator paddles instrumenting the Muon Range Stack.
-The PMT pulses of all four of the broad-faced Punch Through paddles are discriminated to form logic pulses.
-A single logic pulse is formed from these, indicating activity in at least two overlapping paddles at the rear of cryostat, before the steel block of the range stack.
-PMT pulses from the Muon Range Stack are amplified and threshold discriminated.
-These MuRS paddle pulses are then combined as in the Punch Through, creating single-bit indicators for each of the four instrumented layers that at least one pair of overlapping scintillator paddles sent signals within a 20~ns coincidence window.
+\usepackage{lineno,xcolor}
 
-\subsubsection{Trigger Decision and Issuance}
+%\linenumbers
+\begin{document}
+\begin{equation}
+E^{\text{kin}}_{\text{Front Face}} = \sqrt{p_{Beam}^2 + m_{Beam}^2} - m_{Beam}^2 - E_{Loss} 
+\end{equation}
 
-The V1495 may be configured to have up to sixteen trigger patterns and sixteen veto patterns, based on the trigger input signals.
-A trigger pattern is defined as the AND of one or more defined inputs, and may include the NOT of the AND of further inputs.
-Veto patterns are independently defined in the same way, but they have a very different effect.
-When any of the trigger patterns fire, a ``fast trigger'' signal is issued and an adjustable countdown is initiated.
-If the countdown completes without a veto pattern firing, the ``slow trigger'' signal is also issued and on a distinct hardware channel.
-Otherwise, if a veto pattern fires during the countdown, the slow trigger signal is vetoed.
+\begin{document}
 
-The fast trigger signal prompts readout of all the `short' data buffers, which include the V1751 modules, the V1495 itself, and the MWPC controller.
-The V1751 buffers typically contain digitized PMT signals from the time of flight and cryogenic light collection detectors.
-Readout of the TPC wire signals, which are much longer and more numerous, is only prompted at the issuance of the slow trigger.
+\begin{equation}
+E^\text{kin}_j = E^{\text{kin}}_{\text{Front Face}} - \sum_{j<i} E_{\text{dep i}} 
+\end{equation}
+
+\begin{equation}
+E_{i} = \sqrt{p_{Beam}^2 + m_{Beam}^2} - m_{Beam}^2 - E_{Loss} - E_{\text{dep FF-i}}
+\end{equation}
+
+\begin{equation}
+\delta E_{i} = \sqrt{\delta p_{Beam}^2 + \delta E_{Loss}^2 +  \delta  E_{\text{dep FF-i}}^2}
+\end{equation}
+
+\begin{equation}
+E_{\text{dep FF-i}} = \sum_{j<i} E_{\text{dep i}}  \Rightarrow \delta E_{\text{dep FF-i}} = (i -1) \delta E_{\text{dep i}}
+\end{equation}
+
+\begin{equation}
+      \sigma^{\pi^-}_{TOT}(E_{i})  = \frac{1}{n\text{ } \delta X}\frac{ \epsilon^{\text{Inc}}(E_i)  \hspace{0.2cm} C^{\pi MC}_{\text{Int}} (E_{i}) \hspace{0.2cm} N^{\text{TOT}}_{\text{Int}} (E_{i}) }{   \epsilon^{\text{Int}}(E_i) \hspace{0.2cm} C^{\pi MC}_{\text{Inc}} (E_{i}) \hspace{0.2cm}  N^{\text{TOT}}_{\text{Inc}} (E_{i})}.
+\label{eq:C}
+\end{equation}
+
+\begin{equation}
+      \sigma^{K^+}_{TOT}(E_{i})  = \frac{1}{n\text{ } \delta X}\frac{ \epsilon^{\text{Inc}}(E_i)  \hspace{0.2cm} C^{K MC}_{\text{Int}} (E_{i}) \hspace{0.2cm} N^{\text{TOT}}_{\text{Int}} (E_{i}) }{   \epsilon^{\text{Int}}(E_i) \hspace{0.2cm} C^{K MC}_{\text{Inc}} (E_{i}) \hspace{0.2cm}  N^{\text{TOT}}_{\text{Inc}} (E_{i})}.
+\label{eq:C}
+\end{equation}
+
+\end{document}

### 2018-06-25 thesis_main.tex
s

@@
 \advisor{Bonnie T. Fleming}
-\date{Date you'll receive your degree} % usually not \today.
+\date{08/15/2018} % usually not \today.
 
@@
 
-This thesis presents two world's first measurements: the measurement of ($\pi^-$-Ar) total hadronic cross section in the 100-1050 MeV kinetic energy range and the measurement of the ($K^+$-Ar) total hadronic cross section in the 100-650 MeV kinetic energy range.
+This thesis presents two measurements performed for the first time: the measurement of ($\pi^-$-Ar) total hadronic cross section in the 100-1050 MeV kinetic energy range and the measurement of the ($K^+$-Ar) total hadronic cross section in the 100-650 MeV kinetic energy range.
 The analyses devised for these measurements use both the core elements of LArIAT: beamline and LArTPC.
@@
 Thanks to the MicroBooNE experiment which has been an incredible learning experience.
-I want to thank  Sam Zeller for the countless pieces of advice,  my CRT troopers Martin, David, Rui, Igor and Linda  -- if I know something about hardware it's because of you -- and Kazu and  Kirby, who I shamelessly bothered many times for software help, sushi and coffee.
+I want to thank  Sam Zeller for the countless pieces of advice, Jessica Esquivel, whose passion and resilience I deeply admire,  my CRT troopers Martin, David, Rui, Igor and Linda  -- if I know something about hardware it's because of you -- and Kazu and  Kirby, who I shamelessly bothered many times for software help, sushi and coffee.
 
@@
 Louise, my dear, thanks for always finding the time to listen, for your attuned spirit and for the magnolias.
-Flor, thanks for welcoming me to the new world with love and giving me the right definition of ``America".
+Flor, thanks for welcoming me to the new world with love and providing me with the correct definition of ``America".
 

### 2018-07-19 AppendixB-EField/B-LArIATEField.tex
s

@@
 \centering
-\includegraphics[width=3in]{AppendixB-EField/Images/CircuitLArIAT.png}
+\includegraphics[width=\textwidth]{AppendixB-EField/Images/CircuitLArIAT.png}
 \caption{LArIAT HV simple scheme.}
@@
 \centering
-\includegraphics[width=3in]{AppendixB-EField/Images/glassman_current_20160525-30.png}
+\includegraphics[width=\textwidth]{AppendixB-EField/Images/glassman_current_20160525-30.png}
 \caption{Current reading from the Glassman between May 25th and May 30th, 2016 (typical Run-II conditions).}
@@
 \centering
-\includegraphics[width=3in]{AppendixB-EField/Images/RunIIPosCol.png}
+\includegraphics[width=\textwidth]{AppendixB-EField/Images/RunIIPosCol.png}
 \caption{Collection plane $\Delta$t fit for Run II positive polarity ACP data  selected tracks.}
@@
 \centering
-\includegraphics[width=3in]{AppendixB-EField/Images/RunIIPosInd.png}
+\includegraphics[width=\textwidth]{AppendixB-EField/Images/RunIIPosInd.png}
 \caption{Induction plane $\Delta$t fit for Run II positive polarity ACP data  selected tracks.}
@@
     %\resizebox{0.45\textwidth}{!}{%
-    \begin{tabular}{|c|c|c|c|}
+    \begin{tabular}{|l|c|c|c|}
       \multicolumn{4}{c}{\textbf{Delta t$_{drift}$, drift v and E field with ACP tracks}} \\
       \hline \hline
-       Data Period  & $\Delta t_{Drift}$ [$\mu s$] & Drift velocity [mm/$\mu$s] & E field [V/cm] \\
+       Data Period  & $\Delta t_{Drift}$ [$\mu s$] & $v_{Drift}$ [mm/$\mu$s] & $E_{\text{field}}$ [V/cm] \\
        \hline

### 2018-07-19 Chapter-1/1-Theory.tex
s

@@
 where the subscripts C indicates the conserved strong charge (color),   and the subscripts Y indicates the conserved hypercharge. 
-If we indicated with T the weak isospin T and with T3  its third component, hypercharge can be related to the electric charge Q through the Gell-Mann-Nishijima relation:
+If we indicated with T the weak isospin T and with T$_3$  its third component, hypercharge can be related to the electric charge Q through the Gell-Mann-Nishijima relation:
 \begin{equation}
@@
 
-where $c$ and $s$ stand respectively for cosine and sine of the corresponding  mixing angles ($\theta_{12}$, $\theta_{23}$ and $\theta_{13}$), $\delta$ is the Dirac CP violation phase, $\alpha_1$ and $\alpha_{2}$  are the eventual Majorana CP violation phases.
+where $c$ and $s$ stand respectively for cosine and sine of the corresponding  mixing angles
+($\theta_{12}$, $\theta_{23}$ and $\theta_{13}$), $\delta$ is the Dirac CP violation phase,
+$\alpha_1$ and $\alpha_{2}$  are the eventual Majorana CP violation phases.
 Experimental results on neutrino oscillations are generally reported in terms of the mixing angles and of the squared mass splitting $\Delta m^2_{ab} = m^2_{a} - m^2_{b}$, where $a$ and $b$ represent the mass eigenstates.

### 2018-07-19 Chapter-5/5-Preparatory.tex
s

@@
 &  &  &  & & &\\  
-Total Initial events                     & 334500  & 334500 & 334500 &344500 &344500& 344500 \\
-After Multiplicity Rejection        & 330668  & 333420 & 198065 &326576 &344208& 201380 \\
+Total Initial Events                     & 334500  & 334500 & 334500 &344500 &344500& 344500 \\
+Pass Multiplicity Filter        & 330668  & 333420 & 198065 &326576 &344208& 201380 \\
 After WC2TPC Selection          & 218239  & 296333 & 91139  &230418 &300228& 98834  \\
-Evts After Shower Rejection     & 208063  & 288914 &  20293 &219882 &293585& 17780  \\
+Pass Shower Filter     & 208063  & 288914 &  20293 &219882 &293585& 17780  \\
 &  &  &  & & &\\  

### 2018-07-19 Chapter-6/6-PionCrossSection.tex
s

@@
 Expected Composition   & 68.8	\%&4.6 \%&	26.6 \%&	87.4 \%&	3.7	\%&8.9 \% \\
-Composition 2x Muons          & 64.2	\%&9.2 \%&	26.6 \%&	83.7 \%&	7.4	\%&8.9 \% \\
-Composition 2x Electrons      &42.2	\%&4.6 \%&	53.2 \%&	78.5	\%&  3.7	\%&17.8 \%\\
-Composition 0.5x Electrons   &82.1	\%&4.6 \%&	13.3 \%&	91.9 \%&	3.7	\%&4.4 \% \\
+Composition 2 x $\mu$  & 64.2	\%&9.2 \%&	26.6 \%&	83.7 \%&	7.4	\%&8.9 \% \\
+Composition 2 x $e$      &42.2	\%&4.6 \%&	53.2 \%&	78.5	\%&  3.7	\%&17.8 \%\\
+Composition 0.5 x $e$   &82.1	\%&4.6 \%&	13.3 \%&	91.9 \%&	3.7	\%&4.4 \% \\
 \hline

### 2018-08-13 AppendixB-TrackingStudies/B-Tracking.tex
s

@@
 
-\subsection{Study of WC to TPC Match}
+\section{Study of WC to TPC Match}
 
@@
 
-\subsection{Tracking Optimization}\label{ch:TrackOptimization}
+\section{Tracking Optimization}\label{ch:TrackOptimization}
 We perform an optimization of the clustering algorithm (see Section \ref{sec:SignalProc}) with the scope of maximizing the efficiency of finding the interaction point for the total hadronic cross section measurements.

### 2018-08-13 Chapter-1/1-Theory.tex
s

@@
 
-The key elements of a neutrino event generators for resonance and DIS events are the nuclear model and the hadron treatment (both production and transportation).
+The key elements of  neutrino event generators for resonance and DIS events are the nuclear model and the hadron treatment (both production and transportation).
 We illustrate here the conceptual basis of the GENIE Neutrino Generator~\cite{Andreopoulos:2009rq} as an example, since GENIE is one the most popular event generators for liquid argon experiments.
@@
 
+%\begin{figure}
+%\captionsetup{justification=raggedright}  
+%	\begin{minipage}[t]{.5\textwidth}  
+%	  \centering  
+%	   \includegraphics[width=\textwidth]{Chapter-1/Images/IreneCarbon.png}
+%	   	        \caption{ Total, elastic and reaction cross section for $\pi^-$ on $^{12}$C.
+Comparison between results from Geant4 simulation (Bertini cascade model), GENIE simulation (hA model), and experimental data \cite{Clough1974, PhysRevC.23.2173, ALLARDYCE19731,osti_4488432}.}
+%        \label{fig:pionC}
+%	\end{minipage}%  
+%	\begin{minipage}[t]{0.5\textwidth}  
+%	  \centering  
+%	\includegraphics[width=\textwidth]{Chapter-1/Images/IreneCalcium.png}
+%        \caption{Total, elastic and reaction cross section for $\pi^-$ on $^{40}$Ca.
+Comparison between results from Geant4 simulation (Bertini cascade model),  GENIE simulation (hA model), and experimental data \cite{ALLARDYCE19731}.}
+%        \label{fig:pionCa}
+%	\end{minipage}
+%	\par
+%\end{figure}
+
 \begin{figure}
-\captionsetup{justification=raggedright}  
-	\begin{minipage}[t]{.53\textwidth}  
 \centering
-	   \includegraphics[width=3in]{Chapter-1/Images/IreneCarbon.png}
+	   \includegraphics[width=0.5\textwidth]{Chapter-1/Images/IreneCarbon.png}
 	   	        \caption{ Total, elastic and reaction cross section for $\pi^-$ on $^{12}$C.
@@
         \label{fig:pionC}
-	\end{minipage}%  
-	\begin{minipage}[t]{0.53\textwidth}  
+\end{figure}
+
+\begin{figure}
 \centering
-	\includegraphics[width=3in]{Chapter-1/Images/IreneCalcium.png}
+	\includegraphics[width=0.5\textwidth]{Chapter-1/Images/IreneCalcium.png}
         \caption{Total, elastic and reaction cross section for $\pi^-$ on $^{40}$Ca.
@@
         \label{fig:pionCa}
-	\end{minipage}
-	\par
 \end{figure}
@@
 	  \centering  
-	   \includegraphics[width=3in]{Chapter-1/Images/Friedmann.png}
+	   \includegraphics[width=2.8in]{Chapter-1/Images/Friedmann.png}
 	   	        \caption{Ratios between experimental and calculated cross sections as from \cite{Friedman:1997eq}. Top: Total cross sections. \\Bottom: reaction cross sections.}
@@
 	  \centering  
-	\includegraphics[width=3in]{Chapter-1/Images/Bugg.png}
+	\includegraphics[width=2.8in]{Chapter-1/Images/Bugg.png}
         \caption{Total K$^+$  and K$^-$ cross sections on carbon as from \cite{PhysRev.168.1466}.}

### 2018-08-13 Chapter-2/2-LArTPCs.tex
s

@@
 EOS/HISS and ALICE ), dark matter (ArDM), rare decays and capture (e.g.
-TRIUMP, MuCap),  neutrino detectors and nucleon decay (ICARUS, SBN, DUNE), and neutrino less double beta decay (Next).
+TRIUMP, MuCap),  neutrino detectors and nucleon decay (ICARUS, SBN, DUNE), and neutrino less double beta decay (Next, EXO-200\/nEXO).
 A nice review of the history of TPCs and working principles is provided in \cite{0034-4885-73-11-116201}.
@@
 Thus, at the passage of charged particles, liquids generally produce more ionization electrons than gases for the same deposited energy, forcing the particles to deposit more energy in a shorter range.
-The downside of using noble liquid elements in experiments is that they require expensive cryogenic systems to cool the gas until it transitions to its the liquid form.
+The downside of using noble liquid elements in experiments is that they require expensive cryogenic systems to cool the gas until it transitions to its liquid form.
 The properties of liquid argon in comparison liquid xenon -- a popular choice for dark matter and neutrinoless double beta decay detectors -- are summarized in table \ref{tab:properties}.
@@
 To the bare bones, a LArTPC is a bulk of liquid argon sandwiched in a flat capacitor, equipped with a light collection system, as the cartoon in \ref{fig:cartoon} shows.
-A uniform electric field of the order of 500 V/cm is maintained constant between the faces of the capacitor.
+A uniform electric field of the order of 500 V/cm is maintained constant between the conductive faces and field shaping rings are used to avoid fringing fields.
 The anode is sensitive to ionization charge and it is usually made of two or more planes segmented into several hundreds parallel sense wires a few millimeters apart; different geometries for the anode segmentation are under study \cite{1748-0221-8-07-P07002}.
@@
 In fact, since the constant electric field implies that the drift velocity is also constant, the position of the original ionization is simply given by the multiplication of the drift velocity by the drift time, where the ``drift time" is the difference between $t_0$ and the charge arrival time on the wire planes.
-The spacial resolution on this dimension is limited by the time resolution of the electronics or by longitudinal diffusion of the electrons.
+The spatial resolution on this dimension is limited by the time resolution of the electronics or by longitudinal diffusion of the electrons.
 The spatial information on the different wire planes maps a bi-dimensional projection of the interaction pattern in the plane perpendicular to the drift direction.

### 2018-08-13 Chapter-4/4-Methodology.tex
s

@@
 \end{equation}
- where $c$ is the speed of light and $l$ is the length of the particle's trajectory between the time of flight paddels.
+ where $c$ is the speed of light and $l$ is the length of the particle's trajectory between the time of flight paddles.
 
@@
 $\Delta$Y is defined analogously.
-We define the radius difference, $\Delta$R, as $ \Delta \text{R} =  \sqrt{ \Delta \text{X}^2 +  \Delta \text{Y}^2}  $.
+We define the radius difference, $\Delta$R, as $ \Delta \text{R} =  \sqrt{ \Delta \text{X}^2 +  \Delta \text{Y}^2}  $, see Figure \ref{fig:Match}.
 We define  as $\alpha$ the angle between the incident WC track and the TPC track in the plane that contains them.
@@
 \caption{Visual rendering of the wire chamber to TPC match.}
-\label{fig:showerFilt}
+\label{fig:Match}
 \end{figure}

### 2018-08-13 Chapter-5/5-Preparatory.tex
s

@@
 Firstly, G4Beamline lacks the simulation of the WC efficiency which is momentum dependent and leads to enhance the number events in the center of the momentum distribution.
-Secondly, G4Beamline stop tracking pions and their products if they decay after WC1; in data, pion decays in flight can still create a tigger if the produced muon travels thought the beamline detectors.
+Secondly, G4Beamline stop tracking pions and their products if they decay after WC1; in data, pion decays in flight can still create a tigger if the produced muon travels through the beamline detectors.
 In the pion cross section analysis, these differences between data and the G4Beamline prediction are accounted for as a systematic uncertainty related to the beam composition (see Section \ref{ch:BKGsubXSPuppa}).

### 2018-08-13 thesis_main.tex
s

@@
 \advisor{Bonnie T. Fleming}
-\date{08/15/2018} % usually not \today.
+\date{December 2018} % usually not \today.
 
@@
 \tableofcontents
-%\listoffigures % remove this if you have no figures.
-%\listoftables % remove this if you have no tables.
+\listoffigures % remove this if you have no figures.
+\listoftables % remove this if you have no tables.
 
@@
 
-I jumped the pond a little less than 5 years ago because I had a taste of what being a physicists in the US might be like and wanted more.
+I jumped the pond a little less than 5 years ago because I had a taste of what being a physicist in the US might be like and wanted more.
 I came to this vast and sometimes inhospitable land thinking I'd learn a lot of physics (which I did) and hoping I'd keep doing research for as long as I could (which I'm fortunate enough to say I'll still do for a while).
@@
 %\end{table}
-
+\backmatter
 % Any chapters such as End Notes go after this.
-\backmatter
+%\clearpage
+\addcontentsline{toc}{chapter}{Bibliography}
 
 \bibliography{bib.bib}
-%\bibliographystyle{plain}
+\bibliographystyle{plain}
 % for your own sake, use a bibtex file, so all of the numbering of references will be done

### 2018-08-13 Chapter-3/3-TheLArIATExperiment.tex
s

@@
 The piping was cleaned to remove oil and grease before being welded into the system.
-LArIAT uses a filtration system designed for the Liquid Argon Purity Demonstrator (LAPD)~\cite{LAPD}: half of a 77~liter filter contains a 4A molecular sieve (Sigma-Aldrich~\cite{sigma-aldrich}) able to remove mainly water, while the other half contains BASF~CU-0226~S, a highly dispersed copper oxide impregnated on a high surface area alumina, apt to remove mainly oxygen~\cite{basf}.
+LArIAT uses a filtration system designed for the Liquid Argon Purity Demonstrator (LAPD)~\cite{LAPD}: half of a 77~liter filter contains a 4A molecular sieve able to remove mainly water, while the other half contains BASF~CU-0226~S, a highly dispersed copper oxide impregnated on a high surface area alumina, apt to remove mainly oxygen.
 A single pass of argon in the filter is sufficient to achieve the necessary purity, unless the filter is saturated.

### 2018-08-13 Chapter-1/1-Theory.tex
the unbound copy

@@
 The introduction of massive fields in the Standard Model lagrangian would make the theory not gauge invariant, resulting ill-defined.
-This problem is solved in the SM by the introduction of a scalar iso-doublet $\Phi(x)$, the Higgs field, which gives mass to W$^\pm$ and Z$^0$ gauge bosons through the electroweak symmetry breaking mechanism and to the fermions through Yukawa coupling \cite{Higgs1964,Higgs19642}.
+This problem is solved in the SM by the introduction of a scalar iso-doublet $\Phi(x)$, the Higgs field, which gives mass to W$^\pm$ and Z$^0$ gauge bosons through the electroweak symmetry breaking mechanism and to the fermions through Yukawa coupling \cite{Higgs1964}.
 The discovery of the Higgs boson in 2012 by the LHC experiments \cite{201230,Aad2012} marked the ultimate confirmation  of a long history of successful predictions by the SM.
@@
 The SM electroweak lagrangian is the most general renormalizable lagrangian invariant under the local symmetry group $SU(2)_L \otimes U(1)_Y$.
-The lagrangian couples the weak isotopic spin doublets and singlets described in Table \ref{tab:SMParticles} with the gauge bosons  $A^{\mu}_{a}$ ($a$ $=$ 1,2,3) and $B^{\mu}$, and Higgs doublet $\Phi(x)$:
+The lagrangian couples the weak isotopic spin doublets and singlets ($L_{\alpha}$, $Q_{\alpha}$,$l_{\alpha}$, $q_{\alpha}$) described in Table \ref{tab:SMParticles}  with the gauge bosons  $A^{\mu}_{a}$ ($a$ $=$ 1,2,3) and $B^{\mu}$, and Higgs doublet $\Phi(x)$:
 

### 2018-08-13 Chapter-3/3-TheLArIATExperiment.tex
the unbound copy

@@
 Indeed, the presence of contaminants affects both the basic working principles of a LArTPC, as shown in Section \ref{sec:LArTPCWorkingPrinciple}: electronegative contaminants such as oxygen and water decrease the number of ionization electrons collected on the wires after drifting through the volume.
-In addition, contaminants such as Nitrogen decrease the light yield from scintillation light, especially in its slow component.
+In addition,
+contaminants such as Nitrogen decrease the light yield from scintillation light, especially in its slow component.
 In LArIAT, contaminations should not exceed the level of 0.2 parts per billion (ppb).
@@
 The piping was cleaned to remove oil and grease before being welded into the system.
-LArIAT uses a filtration system designed for the Liquid Argon Purity Demonstrator (LAPD)~\cite{LAPD}: half of a 77~liter filter contains a 4A molecular sieve able to remove mainly water, while the other half contains BASF~CU-0226~S, a highly dispersed copper oxide impregnated on a high surface area alumina, apt to remove mainly oxygen.
+LArIAT uses a filtration system designed for the Liquid Argon Purity Demonstrator (LAPD)~\cite{1748-0221-9-07-P07005}: half of a 77~liter filter contains a 4A molecular sieve able to remove mainly water, while the other half contains BASF~CU-0226~S, a highly dispersed copper oxide impregnated on a high surface area alumina, apt to remove mainly oxygen.
 A single pass of argon in the filter is sufficient to achieve the necessary purity, unless the filter is saturated.

### 2018-12-30 Chapter-6/6-PionCrossSection.tex
s
half of lariat prel

@@
 To combine the datasets, we merge the corrected interacting histogram for the 60A data with the corrected interacting histogram for the 100A data; we apply the same merging with the incident histograms.
-We then calculate the cross section central point and statistical uncertainty using the merged interacting and incident histograms.
-The systematic uncertainty is calculated as the sum in quadrature of the original systematic uncertainties on the cross section measurements on the two separated data sets.
+We then calculate the cross section central point, statistical uncertainty and systematic uncertainty relative to the kinetic energy using the merged interacting and incident histograms.
+The systematic uncertainty relative to the corrections is calculated as the sum in quadrature of the original systematic uncertainties on the cross section measurements on the two separated data sets.
 Again, with the exception of the highest energy bins (energy greater than 800 MeV), the systematic uncertainties are the dominant uncertainties; the lion's share of the systematic uncertainty results from our uncertainty on the beam composition, which we plan to constrain with different beamline data in a future iteration of the analysis.

### 2018-12-30 Drafting.tex
s
half of lariat prel

@@
 \begin{document}
+
+\begin{align}
+C^{\text{$\pi$ MC}}_{\text{Int}} (E_{i}) &= \frac{N^{\text{$\pi$ MC}}_{\text{Int}} (E_{i})}{N^{\text{TOT MC}}_{\text{Int}} (E_{i})}\\
+&= \frac{ N^{\text{TOT MC}}_{\text{Int}} (E_{i}) - B^{\text{TOT MC}}_{\text{Int}} (E_{i}) }{N^{\text{TOT MC}}_{\text{Int}} (E_{i})}
+\end{align}
+
+\begin{align}
+C^{\text{$\pi$ MC}}_{\text{Inc}} (E_{i}) &= \frac{N^{\text{$\pi$ MC}}_{\text{Inc}} (E_{i})}{N^{\text{TOT MC}}_{\text{Inc}} (E_{i})}\\
+&= \frac{ N^{\text{TOT MC}}_{\text{Inc}} (E_{i}) - B^{\text{TOT MC}}_{\text{Inc}} (E_{i}) }{N^{\text{TOT MC}}_{\text{Inc}} (E_{i})}
+\end{align}
+
 \begin{equation}
-E^{\text{kin}}_{\text{Front Face}} = \sqrt{p_{Beam}^2 + m_{Beam}^2} - m_{Beam}^2 - E_{Loss} 
+E^{\text{kin}}_{\text{Front Face}} = \sqrt{p_{Beam}^2 + m_{Beam}^2} - m_{Beam} - E_{Loss} 
 \end{equation}
 
-\begin{document}
+\begin{equation}
+U^\text{Int} _{i, j} (E_i) 
+\end{equation}
+
+\begin{equation}
+U^\text{Inc} _{i, j} (E_i) 
+\end{equation}
 
@@
 \begin{equation}
-E_{i} = \sqrt{p_{Beam}^2 + m_{Beam}^2} - m_{Beam}^2 - E_{Loss} - E_{\text{dep FF-i}}
+E_{i} = \sqrt{p_{Beam}^2 + m_{Beam}^2} - m_{Beam} - E_{Loss} - E_{\text{dep FF-j}}
 \end{equation}
@@
 \begin{equation}
-\delta E_{i} = \sqrt{\delta p_{Beam}^2 + \delta E_{Loss}^2 +  \delta  E_{\text{dep FF-i}}^2}
+E_{\text{slice j}} = \sqrt{p_{Beam}^2 + m_{Beam}^2} - m_{Beam} - E_{Loss} - E_{\text{dep FF-j}}
 \end{equation}
@@
 \begin{equation}
-E_{\text{dep FF-i}} = \sum_{j<i} E_{\text{dep i}}  \Rightarrow \delta E_{\text{dep FF-i}} = (i -1) \delta E_{\text{dep i}}
+\delta E_{\text{slice j}} = \sqrt{\delta p_{Beam}^2 + \delta E_{Loss}^2 +  \delta  E_{\text{dep FF-j}}^2}
+\end{equation}
+
+\begin{equation}
+E_{\text{dep FF-j}} = \sum_{j<s} E_{\text{dep s}}  \Rightarrow \delta E_{\text{dep FF-j}} = (j -1) \delta E_{\text{dep s}}
 \end{equation}
@@
 
+\begin{equation}
+P_\text{Int} = \frac{N_\text{Int}}{N_\text{Inc}} = 1 - e^{-\sigma_\text{TOT} \text{ n }\delta X} 
+\end{equation}
+
+\begin{equation}
+\sigma_\text{TOT}(E)  \sim \frac{1}{\text{ n }\delta X} \frac{N_\text{Int}}{N_\text{Inc}}
+\end{equation}
+
+\begin{equation}
+\sigma_\text{TOT}(E_i)  = \frac{1}{\text{ n }\delta X} \frac{N^{\pi^-}_\text{Int}(E_i)}{N^{\pi^-}_\text{Inc}(E_i)}
+\end{equation}
+
+\begin{equation}
+\mathcal{L}(\mu_0; \sigma^2_0; \Delta\theta_0, \Delta\theta_1) = \prod^1_{i=0} f_X(\Delta\theta_i, \mu_0, \sigma^2_0 ) \Rightarrow 
+\end{equation}
+
+\begin{equation}\log{\mathcal{L}} = -\frac{1}{2}\log(2\pi) - \log{\sigma_0} - \frac{1}{2}\frac{(\Delta\theta_0 - \mu_0)^2}{\sigma^2_0} + \text{same for $\Delta\theta_1$}
+\end{equation}
+
+\begin{table}[]
+\centering
+\begin{tabular}{|l|c|c|}
+\hline
+                     & Low E Beam          & High E Beam\\ \hline
+G4Pions       &   70.9 \%          &       82.3 \%         \\ \hline
+G4Muons     &   14.6 \%           &      13.5 \%         \\ \hline
+G4Electrons &   14.5 \%           &       4.2 \%         \\ \hline
+\end{tabular}
+\caption{Simulated beamline composition per magnet settings}
+\label{tab:beamline}
+\end{table}
+
+\begin{table}[b]
+\centering
+\begin{tabular}{|l|c|}
+\hline
+                                                        & Run-II Neg Pol     \\ \hline
+1. Events Reconstructed in Beamline        &  158396    \\ \hline
+2. Events with Plausible Trajectory            &   147468   \\ \hline
+3. Beamline $\pi^-/\mu^-/e^-$  Candidate  &   138481    \\ \hline
+4. Events Surviving Pile Up Filter              &   108929        \\ \hline
+5. Events with WC2TPC Match                 &    41757   \\ \hline
+6. Events Surviving Shower Filter             &    40841       \\ \hline
+7. Available Events For Cross Section      &   40841       \\ \hline
+\end{tabular}
+\caption{Number of data events for Run-II Negative and Positive polarity }
+\label{tab:beamlineDataSelection}
+\end{table}
+
 \end{document}

### 2018-12-30 thesis_main.tex
s
half of lariat prel

@@
 \usepackage{slashed}  
+\usepackage{url}
+\usepackage{hyperref}
 
@@
 % Need to define title before the abstract.
-\title{Measurement of ($\pi^-$-Ar) and ($K^+$-Ar) total hadronic cross sections at the LArIAT experiment}
+\title{Measurement of the Negative Pion and Positive Kaon Total Hadronic Cross Sections on Argon at the LArIAT Experiment}
 \author{Elena Gramellini}
@@
 
-This thesis presents two measurements performed for the first time: the measurement of ($\pi^-$-Ar) total hadronic cross section in the 100-1050 MeV kinetic energy range and the measurement of the ($K^+$-Ar) total hadronic cross section in the 100-650 MeV kinetic energy range.
+This thesis presents two measurements performed for the first time: the measurement of the negative pion total hadronic cross section on argon in the 100-1050 MeV kinetic energy range and the measurement of the positive kaon total hadronic cross section on argon in the 100-650 MeV kinetic energy range.
 The analyses devised for these measurements use both the core elements of LArIAT: beamline and LArTPC.
@@
 Albeit never on argon, the hadronic cross section of pions has been extensively measured before on lighter and heavier different elements in thin target experiments, leading to solid predictions for measurements on argon.
-Through the use of a different technique, our measurement of the  ($\pi^-$-Ar) total hadronic cross section is in general agreement with the predictions by the Geant4 Bertini Cascade model which are based on data from thin target experiments.
+Through the use of a different technique, our measurement of the negative pion total hadronic cross section on argon is in general agreement with the predictions by the Geant4 Bertini Cascade model which are based on data from thin target experiments.
 On the contrary, cross section measurements for kaons are extremely scarce, thus more difficult to model.
-Our measurement of the  ($K^+$-Ar) total hadronic cross section is mostly in tension with the Geant4 prediction over the explored range of energies and provides new experimental data to properly compare existing interaction models.
+Our measurement of the  positive kaon total hadronic cross section on argon is mostly in tension with the Geant4 prediction over the explored range of energies and provides new experimental data to properly compare existing interaction models.
 
@@
 
-The ($\pi^-$-Ar) and the ($K^+$-Ar) total hadronic cross measurements are the first physics results of the LArIAT experiment and will be the basis for the future LArIAT measurements of pion and kaon cross sections in the exclusive channels.
+The negative pion and the positive kaon total hadronic cross measurements are the first physics results of the LArIAT experiment and will be the basis for the future LArIAT measurements of pion and kaon cross sections in the exclusive channels.
 The outcome of these measurements will ultimately enable to quantify and reduce the systematic associated with the hadronic interaction models in neutrino-argon interactions.
